### 2024-01-16 optim.qmd
Initial commit

---
title: Short Paper
subtitle: A Short Subtitle
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliations: 
        - id: some-tech
          name: Some Institute of Technology
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Bob Security
    email: bob@example.com
    affiliations:
        - id: another-u
          name: Another University
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
  - name: Cat Memes
    email: cat@example.com
    affiliations:
        - ref: another-u
    note: Yet another author footnote.
  - name: Derek Zoolander
    email: derek@example.com
    affilations:
        - ref: some-tech
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - projection pursuit
  - optimization
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal of Multivariate Analysis
      formatting: preprint
      model: 3p
      cite-style: super
---

# Introduction

The contributions of this paper are two folds. First, we provide proofs to demonstrate certain optimisers are for certain projection pursuit index functions. Second, we compare the performance of a collection of optimisers on a set of index function with different features and provide recommendations for analysts in practical problems.  @sec-background introduces the projection pursuit method, including the indexes function and optimisation. Section @sec-theory presents the proof on [...]. Section @sec-simulation presents the experimental results. Section @sec-conclusion concludes the paper.

# Projection pursuit, index functions and optimisation {#sec-background}

# [Theoretical results] {#sec-theory}

# [Simulation study] {#sec-simulation}

# Conclusion {#sec-conclusion}

<!-- # Bibliography styles -->

<!-- Here are two sample references:  @Feynman1963118 @Dirac1953888. -->

<!-- By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML.  -->

<!-- You can sets extra options with `natbiboptions` variable in YAML header. Example  -->

<!-- ``` -->

<!-- natbiboptions: longnamesfirst,angle,semicolon -->

<!-- ``` -->

<!-- There are various more specific bibliography styles available at -->

<!-- <https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>.  -->

<!-- To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`. -->

<!-- ## Using CSL  -->

<!-- If `cite-method` is set to `citeproc` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`. In this case, the `csl` option is used to format the references. By default, this template will provide an appropriate style, but alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>. These can be downloaded -->

<!-- and stored locally, or the url can be used as in the example header. -->

<!-- # Equations -->

<!-- Here is an equation: -->

<!-- $$  -->

<!--   f_{X}(x) = \left(\frac{\alpha}{\beta}\right) -->

<!--   \left(\frac{x}{\beta}\right)^{\alpha-1} -->

<!--   e^{-\left(\frac{x}{\beta}\right)^{\alpha}};  -->

<!--   \alpha,\beta,x > 0 . -->

<!-- $$ -->

<!-- Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$ -->

<!-- # Figures and tables -->

<!-- @fig-meaningless is generated using an R chunk. -->

<!-- ```{r} -->

<!-- #| label: fig-meaningless -->

<!-- #| fig-cap: A meaningless scatterplot -->

<!-- #| fig-width: 5 -->

<!-- #| fig-height: 5 -->

<!-- #| fig-align: center -->

<!-- #| out-width: 50% -->

<!-- #| echo: false -->

<!-- plot(runif(25), runif(25)) -->

<!-- ``` -->

<!-- # Tables coming from R -->

<!-- Tables can also be generated using R chunks, as shown in @tbl-simple example. -->

<!-- ```{r} -->

<!-- #| label: tbl-simple -->

<!-- #| tbl-cap: Caption centered above table -->

<!-- #| echo: true -->

<!-- knitr::kable(head(mtcars)[,1:4]) -->

<!-- ``` -->

# References

### 2024-01-16 optim.tex
Initial commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  super,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal of Multivariate analysis}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Short Paper},
  pdfauthor={Alice Anonymous; Bob Security; Cat Memes; Derek Zoolander},
  pdfkeywords={projection pursuit, optimization},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Short Paper \\\large{A Short Subtitle} }
\author[1]{Alice Anonymous%
\corref{cor1}%
\fnref{fn1}}
 \ead{alice@example.com} 
\author[2]{Bob Security%
%
}
 \ead{bob@example.com} 
\author[2]{Cat Memes%
%
\fnref{fn2}}
 \ead{cat@example.com} 
\author[]{Derek Zoolander%
%
}
 \ead{derek@example.com}

\affiliation[1]{organization={Some Institute of Technology, Department
Name},addressline={Street Address},city={City},postcode={Postal
Code},postcodesep={}}
\affiliation[2]{organization={Another University, Department
Name},addressline={Street Address},city={City},postcode={Postal
Code},postcodesep={}}

\cortext[cor1]{Corresponding author}
\fntext[fn1]{This is the first author footnote.}

\fntext[fn2]{Yet another author footnote.}

\begin{abstract}
This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing
elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget
velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris
sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta
egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet
auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo
molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at
tempus elit.
\end{abstract}

\begin{keyword}
    projection pursuit \sep 
    optimization
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The contributions of this paper are two folds. First, we provide proofs
to demonstrate certain optimisers are for certain projection pursuit
index functions. Second, we compare the performance of a collection of
optimisers on a set of index function with different features and
provide recommendations for analysts in practical problems.
Section~\ref{sec-background} introduces the projection pursuit method,
including the indexes function and optimisation. Section
Section~\ref{sec-theory} presents the proof on {[}\ldots{]}. Section
Section~\ref{sec-simulation} presents the experimental results. Section
Section~\ref{sec-conclusion} concludes the paper.

\hypertarget{sec-background}{%
\section{Projection pursuit, index functions and
optimisation}\label{sec-background}}

\hypertarget{sec-theory}{%
\section{{[}Theoretical results{]}}\label{sec-theory}}

\hypertarget{sec-simulation}{%
\section{{[}Simulation study{]}}\label{sec-simulation}}

\hypertarget{sec-conclusion}{%
\section{Conclusion}\label{sec-conclusion}}

\renewcommand\refname{References}
  \bibliography{bibliography.bib}

\end{document}

### 2024-01-29 optim.qmd
more on the structure of the paper

@@
 ---
-title: Short Paper
-subtitle: A Short Subtitle
+title: Performance of Jellyfish Optimiser on Projection Pursuit Problems
 author:
@@
         corresponding: true
-    note: This is the first author footnote.
   - name: Bob Security
@@
         - ref: another-u
-    note: Yet another author footnote.
   - name: Derek Zoolander
@@
   - optimization
+  - jellyfish optimiser
 date: last-modified
@@
       model: 3p
-      cite-style: super
+      cite-style: numbername
 ---
 
-# Introduction
+*Let's use British English ("American or British usage is accepted, but not a mixture of these")*
 
-The contributions of this paper are two folds.
-First, we provide proofs to demonstrate certain optimisers are for certain projection pursuit index functions.
-Second, we compare the performance of a collection of optimisers on a set of index function with different features and provide recommendations for analysts in practical problems.
-@sec-background introduces the projection pursuit method, including the indexes function and optimisation.
-Section @sec-theory presents the proof on [...].
-Section @sec-simulation presents the experimental results.
-Section @sec-conclusion concludes the paper.
+# Introduction [Nicolas and Jessica]
 
-# Projection pursuit, index functions and optimisation {#sec-background}
+The rest of the paper is organised as follows: @sec-background introduces the projection pursuit method, including the indexes function and optimisation.
+@sec-theory introduces the jellyfish optimiser and proposes mathematical expressions to measure the .
+@sec-simulation applies the jellyfish optimiser through different projection pursuit problems with varying dimensions and index functions.
+@sec-conclusion concludes the paper.
 
-# [Theoretical results] {#sec-theory}
+# Projection pursuit, index functions and optimisation [Di and Sherry] {#sec-background}
 
-# [Simulation study] {#sec-simulation}
+# The jellyfish optimiser and property for good optimisers [Nicolas and Jessica] {#sec-theory}
 
-# Conclusion {#sec-conclusion}
+The jellyfish optimiser [@chou_novel_2021] ...
+
+@laa_using_2020 has proposed five criteria for assessing projection pursuit indexes (smoothness, squintability, flexibility, rotation invariance, and speed).
+Since not all the properties affects the optimisation, here we define the three relevant properties (smoothness, squintability, and speed) mathematically and show that the jellyfish optimiser ...
+
+# Application [Di and Sherry] {#sec-simulation}
+
+The jellyfish optimiser has been implemented in the tourr package [@wickham_tourr_2011] and we will use the diagnostic plots proposed in the ferrn package [@RJ-2021-105] to visualise the optimisation process.
+
+## Going beyond 10D
+
+The pipe-finding problem is initially used to investigate indexes and optimisers in @laa_using_2020, and we extend it from a 6D problem to a 12D problem.
+
+Jellyfish optimiser, as a multi-start algorithm, is efficient in [...] for high-dimensional problems
+
+## On skewness and kurtosis index
+
+## Another data example
+
+# Conclusion [Di and Sherry] {#sec-conclusion}
 

### 2024-01-29 optim.tex
more on the structure of the paper

@@
 \documentclass[
-  super,
+  number,
   preprint,
@@
 \makeatother
-\journal{Journal of Multivariate analysis}
+\journal{Journal of Multivariate Analysis}
 \ifLuaTeX
@@
 \usepackage[]{natbib}
-\bibliographystyle{elsarticle-num}
+\bibliographystyle{elsarticle-num-names}
 \IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
@@
 \hypersetup{
-  pdftitle={Short Paper},
+  pdftitle={Performance of Jellyfish Optimiser on Projection Pursuit Problems},
   pdfauthor={Alice Anonymous; Bob Security; Cat Memes; Derek Zoolander},
-  pdfkeywords={projection pursuit, optimization},
+  pdfkeywords={projection pursuit, optimization, jellyfish optimiser},
   colorlinks=true,
@@
 \begin{frontmatter}
-\title{Short Paper \\\large{A Short Subtitle} }
+\title{Performance of Jellyfish Optimiser on Projection Pursuit
+Problems}
 \author[1]{Alice Anonymous%
 \corref{cor1}%
-\fnref{fn1}}
+}
  \ead{alice@example.com} 
@@
 %
-\fnref{fn2}}
+}
  \ead{cat@example.com} 
@@
 \cortext[cor1]{Corresponding author}
-\fntext[fn1]{This is the first author footnote.}
-
-\fntext[fn2]{Yet another author footnote.}
 
@@
 \begin{keyword}
-    projection pursuit \sep 
-    optimization
+    projection pursuit \sep optimization \sep 
+    jellyfish optimiser
 \end{keyword}
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable]}{\end{tcolorbox}}\fi
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\emph{Let's use British English (``American or British usage is
+accepted, but not a mixture of these'')}
 
-The contributions of this paper are two folds. First, we provide proofs
-to demonstrate certain optimisers are for certain projection pursuit
-index functions. Second, we compare the performance of a collection of
-optimisers on a set of index function with different features and
-provide recommendations for analysts in practical problems.
+\hypertarget{introduction-nicolas-and-jessica}{%
+\section{Introduction {[}Nicolas and
+Jessica{]}}\label{introduction-nicolas-and-jessica}}
+
+The rest of the paper is organised as follows:
 Section~\ref{sec-background} introduces the projection pursuit method,
-including the indexes function and optimisation. Section
-Section~\ref{sec-theory} presents the proof on {[}\ldots{]}. Section
-Section~\ref{sec-simulation} presents the experimental results. Section
+including the indexes function and optimisation.
+Section~\ref{sec-theory} introduces the jellyfish optimiser and proposes
+mathematical expressions to measure the . Section~\ref{sec-simulation}
+applies the jellyfish optimiser through different projection pursuit
+problems with varying dimensions and index functions.
 Section~\ref{sec-conclusion} concludes the paper.
@@
 \hypertarget{sec-background}{%
-\section{Projection pursuit, index functions and
-optimisation}\label{sec-background}}
+\section{Projection pursuit, index functions and optimisation {[}Di and
+Sherry{]}}\label{sec-background}}
 
 \hypertarget{sec-theory}{%
-\section{{[}Theoretical results{]}}\label{sec-theory}}
+\section{The jellyfish optimiser and property for good optimisers
+{[}Nicolas and Jessica{]}}\label{sec-theory}}
+
+The jellyfish optimiser \citep{chou_novel_2021} \ldots{}
+
+\citet{laa_using_2020} has proposed five criteria for assessing
+projection pursuit indexes (smoothness, squintability, flexibility,
+rotation invariance, and speed). Since not all the properties affects
+the optimisation, here we define the three relevant properties
+(smoothness, squintability, and speed) mathematically and show that the
+jellyfish optimiser \ldots{}
 
 \hypertarget{sec-simulation}{%
-\section{{[}Simulation study{]}}\label{sec-simulation}}
+\section{Application {[}Di and Sherry{]}}\label{sec-simulation}}
+
+The jellyfish optimiser has been implemented in the tourr package
+\citep{wickham_tourr_2011} and we will use the diagnostic plots proposed
+in the ferrn package \citep{RJ-2021-105} to visualise the optimisation
+process.
+
+\hypertarget{going-beyond-10d}{%
+\subsection{Going beyond 10D}\label{going-beyond-10d}}
+
+The pipe-finding problem is initially used to investigate indexes and
+optimisers in \citet{laa_using_2020}, and we extend it from a 6D problem
+to a 12D problem.
+
+Jellyfish optimiser, as a multi-start algorithm, is efficient in
+{[}\ldots{]} for high-dimensional problems
+
+\hypertarget{on-skewness-and-kurtosis-index}{%
+\subsection{On skewness and kurtosis
+index}\label{on-skewness-and-kurtosis-index}}
+
+\hypertarget{another-data-example}{%
+\subsection{Another data example}\label{another-data-example}}
 
 \hypertarget{sec-conclusion}{%
-\section{Conclusion}\label{sec-conclusion}}
+\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}}
 

### 2024-02-05 optim.qmd
jellyfish search paragraph

@@
 ---
-title: Performance of Jellyfish Optimiser on Projection Pursuit Problems
+title: Performance of Jellyfish Search Optimiser on Projection Pursuit Problems
 author:
@@
 # Introduction [Nicolas and Jessica]
+
+The artificial jellyfish search (JS) algorithm [@chou_novel_2021] is a swarm-based metaheuristic optimisation algorithm inspired by the search behaviour of jellyfish in the ocean.
+It is one of the newest swarm intelligence algorithms, which was shown to have stronger search ability and faster convergence with few algorithmic parameters compared to classic optimization methods [@chou_novel_2021]-[@chou_recent_2022].
 

### 2024-02-05 optim.tex
jellyfish search paragraph

@@
 \hypersetup{
-  pdftitle={Performance of Jellyfish Optimiser on Projection Pursuit Problems},
+  pdftitle={Performance of Jellyfish Search Optimiser on Projection Pursuit Problems},
   pdfauthor={Alice Anonymous; Bob Security; Cat Memes; Derek Zoolander},
@@
 \begin{frontmatter}
-\title{Performance of Jellyfish Optimiser on Projection Pursuit
+\title{Performance of Jellyfish Search Optimiser on Projection Pursuit
 Problems}
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, boxrule=0pt, sharp corners, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
 
@@
 Jessica{]}}\label{introduction-nicolas-and-jessica}}
+
+The artificial jellyfish search (JS) algorithm \citep{chou_novel_2021}
+is a swarm-based metaheuristic optimisation algorithm inspired by the
+search behaviour of jellyfish in the ocean. It is one of the newest
+swarm intelligence algorithms, which was shown to have stronger search
+ability and faster convergence with few algorithmic parameters compared
+to classic optimization methods
+\citep{chou_novel_2021}-\citep{chou_recent_2022}.
 

### 2024-02-05 optim.qmd
add metaheuristics reference

@@
 The artificial jellyfish search (JS) algorithm [@chou_novel_2021] is a swarm-based metaheuristic optimisation algorithm inspired by the search behaviour of jellyfish in the ocean.
-It is one of the newest swarm intelligence algorithms, which was shown to have stronger search ability and faster convergence with few algorithmic parameters compared to classic optimization methods [@chou_novel_2021]-[@chou_recent_2022].
+It is one of the newest swarm intelligence algorithms [@rajwar_exhaustive_2023], which was shown to have stronger search ability and faster convergence with few algorithmic parameters compared to classic optimization methods [@chou_novel_2021]-[@chou_recent_2022].
 

### 2024-02-05 optim.tex
add metaheuristics reference

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, boxrule=0pt, sharp corners, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 search behaviour of jellyfish in the ocean. It is one of the newest
-swarm intelligence algorithms, which was shown to have stronger search
-ability and faster convergence with few algorithmic parameters compared
-to classic optimization methods
+swarm intelligence algorithms \citep{rajwar_exhaustive_2023}, which was
+shown to have stronger search ability and faster convergence with few
+algorithmic parameters compared to classic optimization methods
 \citep{chou_novel_2021}-\citep{chou_recent_2022}.

### 2024-02-06 optim.qmd
some preliminary results of example 1

@@
 
-# Introduction [Nicolas and Jessica]
+```{r setup, echo = FALSE, message=FALSE}
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+library(tidyverse)
+library(ggplot2)
+library(patchwork)
+```
+
+# Introduction \[Nicolas and Jessica\]
 
@@
 
-# Projection pursuit, index functions and optimisation [Di and Sherry] {#sec-background}
+# Projection pursuit, index functions and optimisation \[Di and Sherry\] {#sec-background}
 
-# The jellyfish optimiser and property for good optimisers [Nicolas and Jessica] {#sec-theory}
+# The jellyfish optimiser and property for good optimisers \[Nicolas and Jessica\] {#sec-theory}
 
@@
 
-# Application [Di and Sherry] {#sec-simulation}
+# Application \[Di and Sherry\] {#sec-simulation}
 
@@
 
-Jellyfish optimiser, as a multi-start algorithm, is efficient in [...] for high-dimensional problems
+Jellyfish optimiser, as a multi-start algorithm, is efficient in \[...\] for high-dimensional problems
+
+```{r echo = FALSE}
+load(here::here("data/pipe_better.rda"))
+load(here::here("data/pipe_jellyfish.rda"))
+```
+
+```{r}
+#| fig.width = 6,
+#| fig.height = 3,
+#| fig.align = "center",
+#| fig.cap = "sthis sdfaksdlf"
+pipe_better |>
+  bind_rows(pipe_jellyfish) |>
+  mutate(dim = as.factor(6 + (as.numeric(dim) - 1) * 2)) |>
+  ggplot(aes(x = dim, y = index_val, group = interaction(optimiser, dim), color = optimiser)) +
+  geom_violin(width = 2, adjust = 0.5, position = position_dodge(width = 0.4)) +
+  ggbeeswarm::geom_quasirandom(size = 0.1, width = 0.1, dodge.width = 0.4) +
+  theme_bw() +
+  scale_y_continuous(breaks = seq(0.85, 1, 0.01)) +
+  scale_color_brewer(palette = "Accent") + 
+  labs(y = "Index", x = "Dimension")
+
+```
+
+```{r}
+#| fig.width = 8,
+#| fig.height = 5,
+#| fig.align = "center",
+#| fig.cap = "sthis sdfaksdlf"
+get_proj <- function(data){
+  quantiles <- map(1:4, ~data |>
+                     filter(dim == .x) |>
+                     pull(index_val) |>
+                     quantile(seq(0, 1, 0.5), type = 3) |>
+                     rev() |>
+                     unname())
+
+  res <- map2_dfr(
+    unlist(quantiles),
+    rep(1:4, each = 3),
+    ~ data |>
+      filter(dim == .y, index_val == .x) |>
+      select(dim, proj, index_val),
+    .id = "id"
+    ) |>
+    unnest(proj) |>
+    mutate(index_val = sprintf("%.3f", index_val),
+           optimiser = "jellyfish",
+           id = (as.numeric(id) - 1) %%3 + 1,
+           dim = as.factor(6 + (as.numeric(dim) - 1) * 2)) |>
+    mutate(id = case_when(id == 1 ~ "best",
+                          id == 2 ~ "mean",
+                          id == 3 ~ "worst",
+                          TRUE ~ "0thers"))
+  return(res)
+}
+
+p1 <- get_proj(pipe_jellyfish)  |>
+  ggplot(aes(x = V1, y = V2)) +
+  geom_point(size = 0.1) +
+  geom_text(data = ~. |> distinct(id, dim, index_val),
+           aes(label = index_val), x =0, y = 3.7, size = 3) +
+  facet_grid(dim ~ id) +
+  xlim(-4, 4) + ylim(-4, 4) +
+  theme_bw() +
+  theme(aspect.ratio = 1, axis.ticks = element_blank(),
+        axis.text = element_blank(), axis.title = element_blank(),
+        panel.grid = element_blank())
+p2 <- p1 %+% get_proj(pipe_better)
+p1 + ggtitle("The Jellyfish Optimiser") | p2 + ggtitle("The Better Optimiser")
+
+```
 
@@
 
-# Conclusion [Di and Sherry] {#sec-conclusion}
+# Conclusion \[Di and Sherry\] {#sec-conclusion}
 

### 2024-02-06 optim.tex
some preliminary results of example 1

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, enhanced, sharp corners]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
+\begin{figure}
+
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-2-1.pdf}
+
+}
+
+\caption{sthis sdfaksdlf}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-3-1.pdf}
+
+}
+
+\caption{sthis sdfaksdlf}
+
+\end{figure}
+
 \hypertarget{on-skewness-and-kurtosis-index}{%

### 2024-02-06 code/example1-high-d.R

@@ -1,0 +1,87 @@
+library(tidyverse)
+library(spinebil)
+library(tourr)
+
+set.seed(123)
+seed <- sample(1000: 10000, size = 50)
+historyarray2dt <- function(array, index_f){
+  stopifnot(inherits(array, "history_array"))
+  data <- attr(array, "data")
+  n <- dim(array)[3]
+  res <- purrr::map_dfr(1:n, ~tibble(basis = list(unclass(array)[,,.x])))
+  projs <-  map(res$basis, ~data %*% .x)
+  res <- res |>
+    dplyr::mutate(index_val = map_dbl(projs, ~index_f(.x)),
+                  id = dplyr::row_number())
+
+  return(res)
+
+}
+
+sim <- function(d, optimiser = c("search_better", "search_jellyfish"), data_seed = 123456, optim_seed = seed){
+  set.seed(data_seed)
+  pipe1000 <- pipeData(d, 1000) %>% scale() %>% as_tibble()
+  colnames(pipe1000) <- paste0("V", 1:d)
+
+  if (optimiser == "search_better"){
+    res <- map_dfr(optim_seed, function(seed){
+      set.seed(seed)
+      spiral_better <- save_history(
+        pipe1000, guided_tour(
+          holes(), search_f = search_better, max.tries = 1000
+        )
+      )
+      historyarray2dt(spiral_better[,,-1], holes())
+    }, .id = "sim") |>
+      mutate(d = d)
+  } else if (optimiser == "search_jellyfish"){
+
+    res <- map_dfr(optim_seed, function(seed){
+      set.seed(seed)
+        animate_xy(pipe1000, guided_tour(
+          holes(), d = 2, n_jellies = 1000,
+          search_f =  search_jellyfish,
+          max.tries = 300, min.tries = 100
+        ), max_frames = 50)
+    }, .id = "sim") |>
+      mutate(d = d)
+
+  }
+
+  return(res)
+}
+
+data <- map_dfr(c(6, 8, 10, 12), ~{
+  set.seed(123456)
+  tibble(x = list(pipeData(.x, 1000) %>% scale()))
+}) |>
+  mutate(dim = c(6, 8, 10, 12))
+safe_sim <- safely(sim)
+#a <- map_dfr(c(6, 8, 10, 12), ~safe_sim(d = .x, optimiser = "search_better", optim_seed = seed), .id = "dim")
[…]

### 2024-02-06 code/example2-skewness.R

@@ -1,0 +1,172 @@
+library(tourr)
+library(MaxSkew)
+library(MultiSkew)
+library(tidyverse)
+x <- matrix(c(
+  9160, 1759, 2691,
+    12, 7218, 7655,
+  4624, 4735, 1887,
+  4243, 1527, 2875,
+  4609, 3411,  911,
+  7702, 6074, 5762,
+  3225, 1917, 6834,
+  7847, 7384, 5466,
+  4714, 2428, 4257,
+   358, 9174, 6444
+), ncol = 3, byrow = TRUE)
+vec <- c(0.816, 0.446, 0.367)
+vec <- get_init(x) # hack into the code .MaxSkewThree to get .linear
+res1 <- MaxSkew(x, iterations = 3, components = 1, plot = FALSE)
+res1 <- x %*% vec
+skewness()(res1)
+
+i <- 1
+j <- 1
+for (i in 1:3){
+  for (j in 1:3){
+    y<-x[,j]#j-th column of the data matrix
+    vec[j]<-c(0) #removes the j-th variable from the linear combination
+    MaxSkew:::.MaxSkewBiv(x%*%vec,y)
+    s<-.skewdesideredBIV #skewness of the new linear combination
+    vec[j]<-.linearBIV[2]/.linearBIV[1]
+    res1 <- x %*% vec
+  }
+  print(skewness()(res1))
+
+}
+
+get_init <- function(data) {
+  n <- nrow(data)
+  d <- ncol(data)
+
+  provaconcentrationmatrix <- solve(cov(data) * (n - 1) / n)
+  provaSS <- provaconcentrationmatrix
+  provaaut <- eigen(provaSS)
+  sort(provaaut$values, decreasing = FALSE)
+  provaVV <- matrix(c(0), nrow = d)
+  provaDD <- diag(sort(provaaut$values, decreasing = FALSE))
+
+  for (i in ncol(provaaut$vectors):1) {
+    provaVV <- cbind(provaVV, provaaut$vectors[, i])
+  }
+
+  provaVV <- provaVV[, 2:ncol(provaVV)]
+  provaDD.sqrt <- solve(sqrt(provaDD))
+  provaAA <- provaVV %*% provaDD.sqrt %*% t(provaVV)
+  provaQ <- solve(provaAA)
+
+  x.mean <- colMeans(data)#vector mean
+  m <- sweep(data, 2, x.mean)#centered data
+  Z <- m %*% provaQ
+
[…]

### 2024-02-06 code/example3-local-max.R

@@ -1,0 +1,62 @@
+library(tidyverse)
+library(spinebil)
+library(tourr)
+
+set.seed(123456)
+pipe1000 <- pipeData(6, 1000) |> scale() |> as_tibble()
+sine1000 <- sinData(6, 1000) |> scale()
+dt <-pipe1000 |> bind_cols(sine1000[,5:6])
+colnames(dt) <- paste0("V", 1:(6+ 2))
+
+
+better_res <- animate_xy(dt, guided_tour(
+  holes(), d = 2,
+  search_f =  search_better, alpha = 0.7,
+  max.tries = 500
+))
+
+set.seed(1234)
+jellyfish_res <- animate_xy(dt, guided_tour(
+  holes(), d = 2, n_jellies = 50,
+  search_f =  search_jellyfish,
+  max.tries = 50
+))
+
+proj1 <- matrix(c(0, 0, 0, 0, 1, 0, 0, 0,
+                  0, 0, 0, 0, 0, 1, 0, 0), ncol = 2)
+proj2 <- matrix(c(0, 0, 0, 0, 0, 0, 1, 0,
+                  0, 0, 0, 0, 0, 0, 0, 1), ncol = 2)
+proj3 <- matrix(c(0, 0, 0, 0, 1, 0, 0, 0,
+                  0, 0, 0, 0, 0, 0, 1, 0), ncol = 2)
+proj4 <- matrix(c(0, 0, 0, 0, 1, 0, 0, 0,
+                  0, 0, 0, 0, 0, 1, 0, 1), ncol = 2)
+proj5 <- matrix(c(0, 0, 0, 0, 0, 1, 0, 0,
+                  0, 0, 0, 0, 0, 0, 1, 0), ncol = 2)
+proj6 <- matrix(c(0, 0, 0, 0, 0, 1, 0, 0,
+                  0, 0, 0, 0, 0, 0, 0, 1), ncol = 2)
+proj_list <- list(proj1, proj2, proj3, proj4, proj5, proj6)
+res <- map_dfr(proj_list, ~ as_tibble(as.matrix(dt) %*% .), .id = "id")
+idx <- map(proj_list, ~holes()(as.matrix(dt) %*% .x))
+
+res |>
+  ggplot(aes(x = V1, y = V2)) +
+  geom_point() +
+  theme(aspect.ratio = 1) +
+  facet_wrap(vars(id))
+
+
+best_jellyfish <- jellyfish_res #|> filter(info == "current_best")
[…]

### 2024-02-23 optim.qmd
Brief Description of JSO

@@
 
-The jellyfish optimiser [@chou_novel_2021] ...
+The jellyfish optimiser (JSO) mimics the natural movements of jellyfish, which include passive and active motions driven by ocean currents and their swimming patterns, respectively.
+In the context of optimization, these movements are abstracted to explore the search space in a way that balances exploration (searching new areas) and exploitation (focusing on promising areas).
+The algorithm aims to find the optimal solution by adapting the jellyfish's behavior to navigate towards the best solution over iterations [@chou_novel_2021].
+
+Below is the pseudo-code for this visualisation application.
+
+[Put the code in, add specifics to this visualisation application]
+
+The JSO implementation involves several key parameters that control its search process in optimization problems.
+These parameters are designed to guide the exploration and exploitation phases of the algorithm.
+While the specific implementation details can vary depending on the version of the algorithm or its application, we focus on two main parameters that are most relevant to our application: the number of jellyfish and drift.
 
 @laa_using_2020 has proposed five criteria for assessing projection pursuit indexes (smoothness, squintability, flexibility, rotation invariance, and speed).
-Since not all the properties affects the optimisation, here we define the three relevant properties (smoothness, squintability, and speed) mathematically and show that the jellyfish optimiser ...
+Since not all the properties affects the execution of the optimisation, here we consider the three relevant properties (smoothness, squintability, and speed), and propose three metrics to evaluate these three properties.
+
+## Smoothness
+
+An intuitive way to measure smoothness would be to find how many continuous derivatives exist. We make use of Sobolev spaces:
+
+[def of sobolev space]
+
+Smoothness would then be the highest p such that the index function belongs to $W^{p,\infty}$. We can make it weaker by considering $W^{p,q}$ Sobolev spaces.
+
+[backup plan when things are not differentiable]
+
+## Squintability
+
+A large squint angle means the function is easy to optimize, because you don’t need to be very very close to the perfect view to see the structure.
+A small squint angle means that the derivative of the index function can still be very large values near the optimal point.
+
+OR
+that the values taken by the index function can be clustered into two clear groups: one small group of high values near the optimum, and a large group of the remaining small values.
+OR
+the rate of change of gradient? Gradient changes very drastically when it’s near optimal?
+
+[backup plan when things are not differentiable]
+
+To the best of our knowledge, this is the first attempt to measure the notion of squintability.
+
+## Speed
+
+Computational complexity (in big O notation, with respect to the sample size) of computing the index function.
 

### 2024-02-23 optim.tex
Brief Description of JSO

--- 2024-02-05 optim.tex
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, enhanced, sharp corners]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, sharp corners, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-The jellyfish optimiser \citep{chou_novel_2021} \ldots{}
+The jellyfish optimiser (JSO) mimics the natural movements of jellyfish,
+which include passive and active motions driven by ocean currents and
+their swimming patterns, respectively. In the context of optimization,
+these movements are abstracted to explore the search space in a way that
+balances exploration (searching new areas) and exploitation (focusing on
+promising areas). The algorithm aims to find the optimal solution by
+adapting the jellyfish's behavior to navigate towards the best solution
+over iterations \citep{chou_novel_2021}.
+
+Below is the pseudo-code for this visualisation application.
+
+{[}Put the code in, add specifics to this visualisation application{]}
+
+The JSO implementation involves several key parameters that control its
+search process in optimization problems. These parameters are designed
+to guide the exploration and exploitation phases of the algorithm. While
+the specific implementation details can vary depending on the version of
+the algorithm or its application, we focus on two main parameters that
+are most relevant to our application: the number of jellyfish and drift.
 
@@
 rotation invariance, and speed). Since not all the properties affects
-the optimisation, here we define the three relevant properties
-(smoothness, squintability, and speed) mathematically and show that the
-jellyfish optimiser \ldots{}
+the execution of the optimisation, here we consider the three relevant
+properties (smoothness, squintability, and speed), and propose three
+metrics to evaluate these three properties.
+
+\hypertarget{smoothness}{%
+\subsection{Smoothness}\label{smoothness}}
+
+An intuitive way to measure smoothness would be to find how many
+continuous derivatives exist. We make use of Sobolev spaces:
+
+{[}def of sobolev space{]}
+
+Smoothness would then be the highest p such that the index function
+belongs to \(W^{p,\infty}\). We can make it weaker by considering
+\(W^{p,q}\) Sobolev spaces.
+
+{[}backup plan when things are not differentiable{]}
+
+\hypertarget{squintability}{%
+\subsection{Squintability}\label{squintability}}
+
+A large squint angle means the function is easy to optimize, because you
+don't need to be very very close to the perfect view to see the
+structure. A small squint angle means that the derivative of the index
+function can still be very large values near the optimal point.
+
+OR that the values taken by the index function can be clustered into two
+clear groups: one small group of high values near the optimum, and a
+large group of the remaining small values. OR the rate of change of
+gradient? Gradient changes very drastically when it's near optimal?
+
+{[}backup plan when things are not differentiable{]}
+
+To the best of our knowledge, this is the first attempt to measure the
+notion of squintability.
+
+\hypertarget{speed}{%
+\subsection{Speed}\label{speed}}
+
+Computational complexity (in big O notation, with respect to the sample
+size) of computing the index function.
 
@@
 
+\begin{figure}
+
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-4-1.pdf}
+
+}
+
+\caption{sthis sdfaksdlf}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-6-1.pdf}
+
+}
+
+\caption{sthis sdfaksdlf}
+
+\end{figure}
+
 \hypertarget{on-skewness-and-kurtosis-index}{%

### 2024-03-01 appendix/index-definition.qmd
add definition to indexes in the appendix

---
title: "index"
format: pdf
---

For data $X_{n \times p}$, the projection basis $A_{p * d}$ gives the projected data, $y = XA$. Projection pursuit finds the projection direction $A$ that maximises the index function $f$: $$ \max f(XA) ~~ A^\prime A = 1.$$

The following sections summarise the index functions used in the simulation:

# hole index

-   smooth
-   formula: $$I_{holes}(A) = \frac{1-1/n \sum^n_{i = 1} \exp(-1/2y_iy_i^\prime)}{1-\exp(-p/2)}$$
-   reference:
    -   Cook and Swayne (2007) Interactive and Dynamic Graphics for Data Analysis page 30: [link](https://link.springer.com/chapter/10.1007/978-0-387-71762-3_2)
    -   code from the [tourr package](https://github.com/ggobi/tourr/blob/master/R/interesting-indices.r)

# Stringy

-   non-smooth

-   formula: $$I_{stringy}(A) = \frac{\text{number of vertices with 2 edges}}{\text{number of total vertices with more than one edge}} $$ (probably need some graph theory notation to write it mathematically)

-   reference:

    -   Wilkinson et al (2005) Graph-Theoretic Scagnostics: page 160: [link](https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=1532142)

    -   code from the [cassowaryr page](https://github.com/numbats/cassowaryr/blob/HEAD/R/mst.R)

# spline

-   smooth
-   formula:
    -   $\hat{y}$ is derived from fitting a penalised spline model of y \~ x,
    -   $\hat{y}^\prime$ is derived from fitting a penalised spline model of x \~ y,
    -   the variance of the residuals is compared to the variance of the original data

$$I_{spline}(A) = \max \left(1 - \frac{var(y - \hat{y})}{var(x)}, 1 - \frac{var(y - \hat{y}^\prime)}{var(y)}\right)$$

-   reference:

    -   Ursula and Di (2020) Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics, page 1176: [link](https://arxiv.org/abs/1902.00181)
    -   code from the [cassowaryr page](https://github.com/numbats/cassowaryr/blob/HEAD/R/association.R)

### 2024-03-07 optim.qmd
added math notations

@@
 
-[Put the code in, add specifics to this visualisation application]
+[Put the pseudo-code in, add specifics to this visualisation application]
 
@@
 
-An intuitive way to measure smoothness would be to find how many continuous derivatives exist. We make use of Sobolev spaces:
+An intuitive way to measure smoothness of a function would be to count how many continuous derivatives exist.
+To help define smoothness in our context, we can make use of the Sobolev spaces: functions $f$ in the Sobolev space $W^{p,\infty}$ have all derivatives of order less than $p$ continuous.
+Smoothness would then be the highest $p$ such that the index function belongs to $W^{p,\infty}$.
+We can relax this and consider $W^{p,q}$ Sobolev spaces.
 
-[def of sobolev space]
+Consider the following definition of partial derivative.
+Let $U$ be an open subset of $\mathbf{R}^n$ and $f: U \rightarrow \mathbf{R}$.
+The partial derivative of function $f$ at the point $\mathbf{a} = (a_1, \ldots, a_n) \in U$ with respective to the $i$-th variable $x_i$ is defined as
 
-Smoothness would then be the highest p such that the index function belongs to $W^{p,\infty}$. We can make it weaker by considering $W^{p,q}$ Sobolev spaces.
+$$
+\frac{\delta}{\delta x_i} f(\mathbf{a}) =  \lim_{h \rightarrow 0} \frac{f(a_1, \ldots, a_{i-1}, a_i +h, a_{i+1}, \ldots, a_n) - f(a_1 , \ldots, a_i, \ldots, a_n)}{h} = \lim_{h\rightarrow 0} \frac{f(\mathbf{a} + h\mathbf{e_i}) - f(\mathbf{a})}{h}.
+$$
+where $\mathbf{e_i}$ is the unit vector of the $i$-th variable $x_i$.
 
-[backup plan when things are not differentiable]
+If the derivative is not well-defined, we propose to approximate the derivative using the following expression:
+
+$$\frac{1}{n_i}\sum_{i} \frac{\|f(\mathbf{a} + h_i\mathbf{e_i}) - f(\mathbf{a})\|^p}{h_i}$$
+
+where $p \in [0,1]$. The choice of $p$ reflects the penalty behaviour. For the application in this paper, we choose $p = 1$.
+
+- $h_i$ is an neighbourhood area around $a$. We can insert different values of $h$ to approximate the above quantity and observe how this quantity changes.
 
@@
 
-A large squint angle means the function is easy to optimize, because you don’t need to be very very close to the perfect view to see the structure.
-A small squint angle means that the derivative of the index function can still be very large values near the optimal point.
-
-OR
-that the values taken by the index function can be clustered into two clear groups: one small group of high values near the optimum, and a large group of the remaining small values.
-OR
-the rate of change of gradient? Gradient changes very drastically when it’s near optimal?
-
-[backup plan when things are not differentiable]
+From the literature, it is commonly understood that a large squint angle implies that the function is easy to optimize, because we do not need to be very close to the perfect view to see the structure.
+A small squint angle means that the derivative of the index function can still be very large values near the optimal point and will rapidly change as we get even closer to the optimum.
+As such we can observe the second order gradient, which is the rate of change of gradient, over the space we are searching.
+For some index functions, the second order gradient is not well-defined, we can approximate the second order gradient vector in similar fashion as the above section.
 
@@
 
-Computational complexity (in big O notation, with respect to the sample size) of computing the index function.
+The speed of optimizing an index function can be calculated/measured using the computational complexity (in big O notation, with respect to the sample size) of computing the index function.
 

### 2024-03-07 optim.tex
added math notations

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, sharp corners, enhanced]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-{[}Put the code in, add specifics to this visualisation application{]}
+{[}Put the pseudo-code in, add specifics to this visualisation
+application{]}
 
@@
 
-An intuitive way to measure smoothness would be to find how many
-continuous derivatives exist. We make use of Sobolev spaces:
+An intuitive way to measure smoothness of a function would be to count
+how many continuous derivatives exist. To help define smoothness in our
+context, we can make use of the Sobolev spaces: functions \(f\) in the
+Sobolev space \(W^{p,\infty}\) have all derivatives of order less than
+\(p\) continuous. Smoothness would then be the highest \(p\) such that
+the index function belongs to \(W^{p,\infty}\). We can relax this and
+consider \(W^{p,q}\) Sobolev spaces.
 
-{[}def of sobolev space{]}
+Consider the following definition of partial derivative. Let \(U\) be an
+open subset of \(\mathbf{R}^n\) and \(f: U \rightarrow \mathbf{R}\). The
+partial derivative of function \(f\) at the point
+\(\mathbf{a} = (a_1, \ldots, a_n) \in U\) with respective to the
+\(i\)-th variable \(x_i\) is defined as
 
-Smoothness would then be the highest p such that the index function
-belongs to \(W^{p,\infty}\). We can make it weaker by considering
-\(W^{p,q}\) Sobolev spaces.
+\[
+\frac{\delta}{\delta x_i} f(\mathbf{a}) =  \lim_{h \rightarrow 0} \frac{f(a_1, \ldots, a_{i-1}, a_i +h, a_{i+1}, \ldots, a_n) - f(a_1 , \ldots, a_i, \ldots, a_n)}{h} = \lim_{h\rightarrow 0} \frac{f(\mathbf{a} + h\mathbf{e_i}) - f(\mathbf{a})}{h}.
+\] where \(\mathbf{e_i}\) is the unit vector of the \(i\)-th variable
+\(x_i\).
 
-{[}backup plan when things are not differentiable{]}
+If the derivative is not well-defined, we propose to approximate the
+derivative using the following expression:
+
+\[\frac{1}{n_i}\sum_{i} \frac{\|f(\mathbf{a} + h_i\mathbf{e_i}) - f(\mathbf{a})\|^p}{h_i}\]
+
+where \(p \in [0,1]\). The choice of \(p\) reflects the penalty
+behaviour. For the application in this paper, we choose \(p = 1\).
+
+\begin{itemize}
+\tightlist
+\item
+  \(h_i\) is an neighbourhood area around \(a\). We can insert different
+  values of \(h\) to approximate the above quantity and observe how this
+  quantity changes.
+\end{itemize}
 
@@
 
-A large squint angle means the function is easy to optimize, because you
-don't need to be very very close to the perfect view to see the
-structure. A small squint angle means that the derivative of the index
-function can still be very large values near the optimal point.
-
-OR that the values taken by the index function can be clustered into two
-clear groups: one small group of high values near the optimum, and a
-large group of the remaining small values. OR the rate of change of
-gradient? Gradient changes very drastically when it's near optimal?
-
-{[}backup plan when things are not differentiable{]}
+From the literature, it is commonly understood that a large squint angle
+implies that the function is easy to optimize, because we do not need to
+be very close to the perfect view to see the structure. A small squint
+angle means that the derivative of the index function can still be very
+large values near the optimal point and will rapidly change as we get
+even closer to the optimum. As such we can observe the second order
+gradient, which is the rate of change of gradient, over the space we are
+searching. For some index functions, the second order gradient is not
+well-defined, we can approximate the second order gradient vector in
+similar fashion as the above section.
 
@@
 
-Computational complexity (in big O notation, with respect to the sample
-size) of computing the index function.
+The speed of optimizing an index function can be calculated/measured
+using the computational complexity (in big O notation, with respect to
+the sample size) of computing the index function.
 

### 2024-03-12 code/00-combine-simulations.R
simulation: sine, 6&8d, MIC&TIC

@@ -1,0 +1,31 @@
+#####################################################################
+# simulation sine 68d TICMIC
+load(file = "data/sim_sine_68d_TICMIC.rda")
+sim_data <- sim_sine_68d_TICMIC
+setup_level_best_basis <- sim_data |>
+  group_by(n_jellies, max_tries, d, idx_f) |>
+  filter(index_val == max(index_val)) |>
+  filter(row_number() == 1)
+
+ids <- setup_level_best_basis |> filter(d == 6) |> pull(id)
+res <- map_dfr(ids, ~{
+  b <- setup_level_best_basis |> filter(id == .x) |> pull(basis)
+  dt <- as_tibble(as.matrix(sine1000) %*% b[[1]])
+  return(dt)
+}, .id = "id")
+
+labels_df <- setup_level_best_basis |>
+  ungroup() |>
+  filter(d == 6) |>
+  select(idx_f:max_tries, index_val) |>
+  mutate(id = as.character(1:10))
+
+library(ggh4x)
+res |>
+  left_join(labels_df) |>
+  ggplot(aes(x = V2 ,y = V1)) +
+  geom_point() +
+  geom_text(data = labels_df,
+            aes(x = -3, y = 3, label = round(index_val, 2)),
+            nudge_x = 0.1, nudge_y = 0.1) +
+  facet_nested(idx_f ~ n_jellies + max_tries, labeller = label_both )

### 2024-03-12 code/02-clean-up.R

@@ -1,18 +1,23 @@
 library(tidyverse)
 load(file = "data/sim_pipe.rda")
 
-sim_data <- sim_sine_6d
+sim_data <- sim_sine_68d_TICMIC
 
 set.seed(123)
 seed <- sample(1000: 10000, size = 50)
-sim_setup <- crossing(n_jellies = c(20, 50, 100),
-                      max_tries = c(50, 100),
-                      d = c(6, 8, 10, 12),
-                      sim = 1:50) |>
-  mutate(seed = seed[sim], id = row_number())
+sim_setup <- crossing(idx_f = c("MIC", "TIC"),
+                      d = c(6, 8),
+                      n_jellies = c(20, 50, 100),
+                      max_tries = c(50, 100)) |>
+  filter(!(n_jellies == 100 & max_tries == 100))
 
 tour_level_best_basis <- sim_data |>
   group_by(id) |>
+  filter(index_val == max(index_val)) |>
+  filter(row_number() == 1)
+
+setup_level_best_basis <- sim_data |>
+  group_by(n_jellies, max_tries, d, idx_f) |>
   filter(index_val == max(index_val)) |>
   filter(row_number() == 1)
 
@@ -38,12 +43,15 @@ tour_level <- sim_data |>
 # P_J_hat:    proportion of trials found
 # I_max_max:  best index of each setup
 setup_level <- tour_level |>
-  group_by(n_jellies, max_tries, d) |>
-  reframe(P_J_hat = sum(abs(1 - I_max) <= 0.05)/n(), # TODO: use a better rule
-          I_max_max = max(I_max))
+  group_by(n_jellies, max_tries, d, idx_f) |>
+  reframe(
+    I_max_max = max(I_max),
+    P_J_hat = sum(abs(I_max_max - I_max) <= 0.05)/n() # TODO: use a better rule
+    )
 
 setup_level |>
   ggplot(aes( x= as.factor(n_jellies), y = as.factor(max_tries))) +
-  geom_tile(aes(fill = P_J_hat)) +
-  facet_wrap(vars(d)) +
+  geom_tile(aes(fill = I_max_max)) +
+  #facet_wrap(vars(d)) +
+  facet_grid(d ~ idx_f) +
   colorspace::scale_fill_continuous_sequential(palette = "Sunset")

### 2024-03-12 code/04-simulation-sine-68d-MICTIC.R

@@ -1,0 +1,115 @@
+library(tidyverse)
+library(spinebil)
+library(tourr)
+library(minerva)
+
+################################################################################
+# d = 6
+sim <- function(d = d, idx_f = idx_f, n_jellies = n_jellies, max.tries = max.tries,
+                data_seed = 123456, optim_seed = seed, sim = sim){
+  set.seed(data_seed)
+  sine1000 <- sinData(d, 1000) %>% scale() %>% as_tibble()
+  colnames(sine1000) <- paste0("V", 1:d)
+
+  cat("sim: ", sim, "\n")
+  cat("n_jellies: ", n_jellies, "\n")
+  cat("max.tries: ", max.tries, "\n")
+  cat("idx_f: ", idx_f, "\n")
+  cat("d: ", d, "\n")
+
+  idx_f <- eval(sym(idx_f))
+  res <- map_dfr(optim_seed, function(seed){
+    t1 <- Sys.time()
+    set.seed(seed)
+    res <- animate_xy(sine1000, guided_tour(
+      idx_f(), d = 2, n_jellies = n_jellies,
+      search_f =  search_jellyfish,
+      max.tries = max.tries
+    ), max_frames = max.tries) |> dplyr::select(-id)
+    t2 <- Sys.time()
+    res <- res |> mutate(time = t2 - t1)
+    return(res)
+  })
+
+  return(res)
+}
+
+MIC <- function(){
+  function(mat){
+    minerva::mine(mat[,1], mat[,2], alpha = 0.3,  est = "mic_e")$MIC
+  }
+}
+
+TIC <- function(){
+  function(mat){
+    minerva::mine(mat[,1], mat[,2], est = "mic_e", alpha = 0.3)$TIC
+  }
+}
+
+#  n_jellies | max_tries
+#         20 |       50
+#         20 |      100
+#         50 |       50
+#         50 |      100
+#        100 |       50
+#        100 |      100
+# each 50 repetitions
+# take 2 hrs
+set.seed(123)
+seed <- sample(1000: 10000, size = 50)
+sim_setup <- crossing(idx_f = c("MIC", "TIC"),
+                      d = c(6, 8),
+                      n_jellies = c(20, 50, 100),
+                      max_tries = c(50, 100)) |>
+  filter(!(n_jellies == 100 & max_tries == 100)) |>
+  crossing(sim = 1) |>
[…]

### 2024-03-14 code/05-simulation-sine-6d-dcor2d.R
simulation sine, 6d, dcor2d

@@ -1,0 +1,84 @@
+library(tidyverse)
+library(spinebil)
+library(tourr)
+
+################################################################################
+# d = 6
+set.seed(123456)
+sine1000 <- sinData(6, 1000) %>% scale() %>% as_tibble()
+colnames(sine1000) <- paste0("V", 1:6)
+
+sim <- function(d = d, idx_f = idx_f, n_jellies = n_jellies, max.tries = max.tries,
+                data = sine1000 , optim_seed = seed, sim = sim){
+
+  cat("sim:", sim, "| n_jellies:", n_jellies, "| max.tries:", max.tries,
+  "| idx_f:", idx_f, "| d:", d, "\n")
+
+  idx_f <- eval(sym(idx_f))
+  res <- map_dfr(optim_seed, function(seed){
+    t1 <- Sys.time()
+    set.seed(seed)
+    res <- animate_xy(data, guided_tour(
+      idx_f(), d = 2, n_jellies = n_jellies,
+      search_f =  search_jellyfish,
+      max.tries = max.tries
+    ), max_frames = max.tries) |> dplyr::select(-id)
+    t2 <- Sys.time()
+    res <- res |> mutate(time = t2 - t1)
+    return(res)
+  })
+
+  return(res)
+}
+
+dcor2d_2 <- function() {
+  function(mat) {
+    xy <- na.omit(data.frame(x = mat[, 1], y = mat[, 2]))
+    measure <- with(xy, energy::dcor2d(x, y, type = "U"))
+    return(measure)
+  }
+}
+
+#  n_jellies | max_tries
+#         20 |       50
+#         20 |      100
+#         50 |       50
+#         50 |      100
+#        100 |       50
+#        100 |      100
+# each 50 repetitions
+# take less than 2 hrs
+set.seed(123)
+seed <- sample(1000: 10000, size = 50)
+sim_setup <- crossing(idx_f = c("dcor2d_2"),
+                      d = 6,
+                      n_jellies = c(20, 50),
+                      max_tries = c(50, 100)) |>
+  filter(!(n_jellies == 50 & max_tries == 100)) |>
+  crossing(sim = 1:50) |>
+  mutate(seed = seed[sim], id = row_number())
+
+t1 <- Sys.time()
+sim_res <- sim_setup |>
+  rowwise() |>
[…]

### 2024-03-15 appendix/index-definition.qmd
more index definitions

@@
 Projection pursuit finds the projection direction $A$ that maximises the index function $f$: $$ \max f(XA) ~~ A^\prime A = 1.$$
+
+For 2D projections, we can also write $y_{n \times 2} = (y_1, y_2)$, where $y_1$ and $y_2$ are the two columns of the projected data.
 
@@
     -   code from the [tourr package](https://github.com/ggobi/tourr/blob/master/R/interesting-indices.r)
+
+# dcor2d_2 index
+
+-   smooth
+
+-
+
+-   Reference:
+
+-   Note
+
+  - From the function manual: "For bivariate data only, these are fast O(n log n) implementations of distance correlation and distance covariance statistics."
+
+# MIC/ TIC index
+
+-   smooth
+
+-   formula
+
+    -   Let $g(k, l)$ define a partition of the space $(y_1, y_2)$ into $k \times l$ rectangles, for example, $g(2, 3)$ means dividing the data space into 2 rectangles in the $y_2$ direction and 3 rectangles in the $y_1$ direction.
+Let $G$ denotes all the possible partition.
+    -   MIC finds the maximum mutual information over G where k and l is bounded by the grid size: $k \times l < B(n) = n^{\alpha}$ where alpha = 0.3 in our simulation
+
+    $$I_{MIC}(A) = \max_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \max_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$
+
+    where $k^*$ and $l^*$ are the number of rectangles in the optimal partition $g$.
+
+    -   TIC calculates the sum of mutual information over G
+
+$$I_{TIC}(A) = \sum_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \sum_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$
+
+-   Reference:
+
+    -   Reshef (2011) Detecting Novel Associations in Large Datasets: [link](https://pubmed.ncbi.nlm.nih.gov/22174245/) - Figure 1 is usually to understand the general idea
+
+    -   Reshef (2016) Measuring Dependence Powerfully and Equitably: [link](https://jmlr.csail.mit.edu/papers/volume17/15-308/15-308.pdf).
+
+-   Note:
+
+    -   The original Definition 1 on page 6 of Reshef (2016) uses supremum, here I just use the maximum.
+    -   There are different version of MIC and we use $MIC_e$ and $TIC_e$ (See Sec 4 in the paper in Reshef (2016)), which uses the equicharacteristic matrix rather than the original characteristic matrix in both MIC/ TIC calculation, see page 14 Figure 1 for there difference (equicharacteristic matrix parts the space equally hence faster).
+    -   Section 4.3/4.4 of Reshef (2016) talks about the time complexity
+
+# loess/ spline index
+
+-   smooth
+
+-   formula:
+
+    -   let $e^{\text{model}}_{y_1 \sim y_2}$ denote the residual from model $y_1 \sim y_2$
+
+    -   for the loess index, we fit a loess model to $y_1 \sim y_2$ (use $\alpha = 0.05$ for the smoothing parameter in the loess index)
+
+    $$I_{\text{loess}}(A) = \max \left(1 - \frac{var(e^{\text{loess}}_{y_1 \sim y_2})}{var(y_1)}, 1 - \frac{var(e^{\text{loess}}_{y_2 \sim y_1})}{var(y_2)}\right)$$
+
+    -   for the spline index, a spline model is fitted (use cubic regression spline with k = 15 for the diemnsion of the basis in the spline model).
+
+$$I_{\text{spline}}(A) = \max \left(1 - \frac{var(e^{\text{spline}}_{y_1 \sim y_2})}{var(y_1)}, 1 - \frac{var(e^{\text{spline}}_{y_2 \sim y_1})}{var(y_2)}\right)$$
+
+-   reference:
+    -   loess: (none) I create it myself, inspired by the spline index
+    -   spline:
+        -   Ursula and Di (2020) Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics, page 1176: [link](https://arxiv.org/abs/1902.00181)
+        -   code from the [cassowaryr page](https://github.com/numbats/cassowaryr/blob/HEAD/R/association.R)
 
@@
     -   code from the [cassowaryr page](https://github.com/numbats/cassowaryr/blob/HEAD/R/mst.R)
-
-# spline
-
--   smooth
--   formula:
-    -   $\hat{y}$ is derived from fitting a penalised spline model of y \~ x,
-    -   $\hat{y}^\prime$ is derived from fitting a penalised spline model of x \~ y,
-    -   the variance of the residuals is compared to the variance of the original data
-
-$$I_{spline}(A) = \max \left(1 - \frac{var(y - \hat{y})}{var(x)}, 1 - \frac{var(y - \hat{y}^\prime)}{var(y)}\right)$$
-
--   reference:
-
-    -   Ursula and Di (2020) Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics, page 1176: [link](https://arxiv.org/abs/1902.00181)
-    -   code from the [cassowaryr page](https://github.com/numbats/cassowaryr/blob/HEAD/R/association.R)

### 2024-03-18 appendix/index-definition.qmd
dcor2d_2 index definition

@@
 
--
+-   formula:
+
+    -   define pair-wised distance: $a_{ij} = \|y_1^i  - y_1^j\|$ and $b_{ij} = \|y_2^i  - y_2^j\|$ for $i, j = 1, 2, \ldots, n$.
+
+    -   define the column sum, row sum, and overall sum: $$a_{i \cdot} = \sum_{l = 1}^n a_{il}, \quad a_{\cdot j} = \sum_{k = 1}^n a_{kj}, \quad a_{\cdot \cdot} = \sum_{k,l = 1}^n a_{kl} \quad b_{i \cdot} = \sum_{l = 1}^n b_{il}, \quad b_{\cdot j} = \sum_{k = 1}^n b_{kj}, \quad b_{\cdot \cdot} = \sum_{k,l = 1}^n b_{kl}$$
+
+    -   define the distance covariance and variance: $$\text{dCov}(a_{ij}, b_{ij}) = \frac{1}{n(n-3)} \sum_{i \neq j} a_{ij} b_{ij} -\frac{2}{n (n-2)(n-3)}\sum_{i = 1}^n a_{i \cdot} b_{i \cdot} + \frac{1}{n(n - 1)(n-2)(n-3)} a_{\cdot \cdot} b_{\cdot \cdot}$$
+
+    -   define the distance correlation: $$\text{dCor} (a_{ij}, b_{ij}) = \frac{\text{dCov}(a_{ij}, b_{ij})}{\sqrt{\text{dVar}(a_{ij}) \text{dVar}(b_{ij})}}$$
 
@@
 
--   Note
-
-  - From the function manual: "For bivariate data only, these are fast O(n log n) implementations of distance correlation and distance covariance statistics."
+    -   Xiaoming Huo &Gábor J.
+Székely (2016) Fast Computing for Distance Covariance: [link](https://www.tandfonline.com/doi/full/10.1080/00401706.2015.1054435).
+The papre proves that the covariance formula above (eq 3.3 in the paper) is an unbaised estimator of the population covariance in the original formula: $\displaystyle \operatorname {dCov} _{n}^{2}(X,Y):={\frac {1}{n^{2}}}\sum _{j=1}^{n}\sum _{k=1}^{n}A_{j,k}\,B_{j,k}$.
+The fast algorithm realises the new formula is in $O(n \log n)$ time.
+    -   Maria L. Rizzo, Gábor J. Székely (2015) Energy distance: [link](https://wires.onlinelibrary.wiley.com/doi/abs/10.1002/wics.1375) WIREs Computational Statistics review paper
+    -   Gábor J.
+Székely, Maria L.
+Rizzo, Nail K.
+Bakirov (2007) Measuring and testing dependence by correlation of distances: [link](https://projecteuclid.org/journals/annals-of-statistics/volume-35/issue-6/Measuring-and-testing-dependence-by-correlation-of-distances/10.1214/009053607000000505.full): original paper of distance correlation
 

### 2024-04-01 optim.qmd
expanded intro and added lots of references

@@
 ---
-title: Performance of Jellyfish Search Optimiser on Projection Pursuit Problems
+title: Studying the Performance of the Jellyfish Optimiser for the Application of Projection Pursuit
 author:
@@
   - jellyfish optimiser
+  - data visualisation
+  - high-dimensional data
 date: last-modified
@@
 
-The rest of the paper is organised as follows: @sec-background introduces the projection pursuit method, including the indexes function and optimisation.
-@sec-theory introduces the jellyfish optimiser and proposes mathematical expressions to measure the .
-@sec-simulation applies the jellyfish optimiser through different projection pursuit problems with varying dimensions and index functions.
-@sec-conclusion concludes the paper.
+Effective optimisation is an important aspect of many methods employed for visualising high-dimensional data ($X$).
+Here we are concerned about computing informative linear projections of high-dimensional ($p$) data using projection pursuit (PP) (@kr69, @FT74).
+This involves optimising a function  (e.g.
+@hall1989polynomial, @cook1993projection, @lee2010projection, @Loperfido2018, @Loperfido2020), called the projection pursuit index (PPI), that defines what is interesting or informative in a projection.
+
+These PPI are defined on projections ($XA$), which means that there is a constraint that needs to be considered when optimising.
+A projection of data is defined by a $p\times d$ orthonormal matrix $A$, and this imposes the constraint on the elements of $A$, that columns need have norm equal to 1 and the product of columns need to sum to zero.
+
+@cook1995grand introduced the PP guided tour, which enabled interactive visualisation of the optimisation in order to visually explore high-dimensional data.
+It is implemented in the R [XXX REF] package `tourr` [@tourr].
+The optimisation that is implemented is fairly basic, and potential problems were highlighted by @RJ-2021-105.
+Implementing better optimisation functionality is a goal, but it needs to be kept in mind that the guided tour also has places importance on watching the projected data as the optimisation progresses.
+
+Here we explore the potential for a jellyfish optimisation to be integrated with the guided tour.
+@sec-background explains the optimisation that is used in the current the projection pursuit guided tour.
+@sec-theory provides more details on the jellyfish optimiser and formalises several characteristics of projection pursuit indexes that are help to measure optimisaer performance.
+@sec-simulation describes a simulation study on performance of the jellyfish for several types of data and index functions.
+@sec-conclusion summarises the work and provides suggestions for future directions.
 
@@
 
-# The jellyfish optimiser and property for good optimisers \[Nicolas and Jessica\] {#sec-theory}
+<!-- Need to assume that the special issue will have the general introduction to PP, so the focus here will be on background needed for this paper, tours, guidance using PP, optimisation, visualisation, ...
+-->
+
+# The jellyfish optimiser and properties of PP indexes \[Nicolas and Jessica\] {#sec-theory}
 
@@
 
-[Put the pseudo-code in, add specifics to this visualisation application]
+\[Put the pseudo-code in, add specifics to this visualisation application\]
 
@@
 \frac{\delta}{\delta x_i} f(\mathbf{a}) =  \lim_{h \rightarrow 0} \frac{f(a_1, \ldots, a_{i-1}, a_i +h, a_{i+1}, \ldots, a_n) - f(a_1 , \ldots, a_i, \ldots, a_n)}{h} = \lim_{h\rightarrow 0} \frac{f(\mathbf{a} + h\mathbf{e_i}) - f(\mathbf{a})}{h}.
-$$
-where $\mathbf{e_i}$ is the unit vector of the $i$-th variable $x_i$.
+$$ where $\mathbf{e_i}$ is the unit vector of the $i$-th variable $x_i$.
 

### 2024-04-01 optim.tex
expanded intro and added lots of references

@@
 \makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
 \@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
 \makeatother
@@
 \bibliographystyle{elsarticle-num-names}
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \hypersetup{
-  pdftitle={Performance of Jellyfish Search Optimiser on Projection Pursuit Problems},
+  pdftitle={Studying the Performance of the Jellyfish Optimiser for the Application of Projection Pursuit},
   pdfauthor={Alice Anonymous; Bob Security; Cat Memes; Derek Zoolander},
-  pdfkeywords={projection pursuit, optimization, jellyfish optimiser},
+  pdfkeywords={projection pursuit, optimization, jellyfish
+optimiser, data visualisation, high-dimensional data},
   colorlinks=true,
@@
 \begin{frontmatter}
-\title{Performance of Jellyfish Search Optimiser on Projection Pursuit
-Problems}
+\title{Studying the Performance of the Jellyfish Optimiser for the
+Application of Projection Pursuit}
 \author[1]{Alice Anonymous%
@@
 \begin{keyword}
-    projection pursuit \sep optimization \sep 
-    jellyfish optimiser
+    projection pursuit \sep optimization \sep jellyfish
+optimiser \sep data visualisation \sep 
+    high-dimensional data
 \end{keyword}
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{introduction-nicolas-and-jessica}{%
 \section{Introduction {[}Nicolas and
-Jessica{]}}\label{introduction-nicolas-and-jessica}}
+Jessica{]}}\label{introduction-nicolas-and-jessica}
 
@@
 
-The rest of the paper is organised as follows:
-Section~\ref{sec-background} introduces the projection pursuit method,
-including the indexes function and optimisation.
-Section~\ref{sec-theory} introduces the jellyfish optimiser and proposes
-mathematical expressions to measure the . Section~\ref{sec-simulation}
-applies the jellyfish optimiser through different projection pursuit
-problems with varying dimensions and index functions.
-Section~\ref{sec-conclusion} concludes the paper.
+Effective optimisation is an important aspect of many methods employed
+for visualising high-dimensional data (\(X\)). Here we are concerned
+about computing informative linear projections of high-dimensional
+(\(p\)) data using projection pursuit (PP) (\citet{kr69}, \citet{FT74}).
+This involves optimising a function (e.g. \citet{hall1989polynomial},
+\citet{cook1993projection}, \citet{lee2010projection},
+\citet{Loperfido2018}, \citet{Loperfido2020}), called the projection
+pursuit index (PPI), that defines what is interesting or informative in
+a projection.
 
-\hypertarget{sec-background}{%
+These PPI are defined on projections (\(XA\)), which means that there is
+a constraint that needs to be considered when optimising. A projection
+of data is defined by a \(p\times d\) orthonormal matrix \(A\), and this
+imposes the constraint on the elements of \(A\), that columns need have
+norm equal to 1 and the product of columns need to sum to zero.
+
+\citet{cook1995grand} introduced the PP guided tour, which enabled
+interactive visualisation of the optimisation in order to visually
+explore high-dimensional data. It is implemented in the R {[}XXX REF{]}
+package \texttt{tourr} \citep{tourr}. The optimisation that is
+implemented is fairly basic, and potential problems were highlighted by
+\citet{RJ-2021-105}. Implementing better optimisation functionality is a
+goal, but it needs to be kept in mind that the guided tour also has
+places importance on watching the projected data as the optimisation
+progresses.
+
+Here we explore the potential for a jellyfish optimisation to be
+integrated with the guided tour. Section~\ref{sec-background} explains
+the optimisation that is used in the current the projection pursuit
+guided tour. Section~\ref{sec-theory} provides more details on the
+jellyfish optimiser and formalises several characteristics of projection
+pursuit indexes that are help to measure optimisaer performance.
+Section~\ref{sec-simulation} describes a simulation study on performance
+of the jellyfish for several types of data and index functions.
+Section~\ref{sec-conclusion} summarises the work and provides
+suggestions for future directions.
+
 \section{Projection pursuit, index functions and optimisation {[}Di and
-Sherry{]}}\label{sec-background}}
+Sherry{]}}\label{sec-background}
 
-\hypertarget{sec-theory}{%
-\section{The jellyfish optimiser and property for good optimisers
-{[}Nicolas and Jessica{]}}\label{sec-theory}}
+\section{The jellyfish optimiser and properties of PP indexes {[}Nicolas
+and Jessica{]}}\label{sec-theory}
 
@@
 
-\hypertarget{smoothness}{%
-\subsection{Smoothness}\label{smoothness}}
+\subsection{Smoothness}\label{smoothness}
 
@@
 
-\hypertarget{squintability}{%
-\subsection{Squintability}\label{squintability}}
+\subsection{Squintability}\label{squintability}
 
@@
 
-\hypertarget{speed}{%
-\subsection{Speed}\label{speed}}
+\subsection{Speed}\label{speed}
 
@@
 
-\hypertarget{sec-simulation}{%
-\section{Application {[}Di and Sherry{]}}\label{sec-simulation}}
+\section{Application {[}Di and Sherry{]}}\label{sec-simulation}
 
@@
 
-\hypertarget{going-beyond-10d}{%
-\subsection{Going beyond 10D}\label{going-beyond-10d}}
+\subsection{Going beyond 10D}\label{going-beyond-10d}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-4-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-2-1.pdf}
 
@@
 
-\end{figure}
+\end{figure}%
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-6-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-3-1.pdf}
 
@@
 
-\end{figure}
+\end{figure}%
 
-\hypertarget{on-skewness-and-kurtosis-index}{%
 \subsection{On skewness and kurtosis
-index}\label{on-skewness-and-kurtosis-index}}
+index}\label{on-skewness-and-kurtosis-index}
 
-\hypertarget{another-data-example}{%
-\subsection{Another data example}\label{another-data-example}}
+\subsection{Another data example}\label{another-data-example}
 
-\hypertarget{sec-conclusion}{%
-\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}}
+\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
 

### 2024-04-02 optim.qmd
done with section 2

@@
 @cook1995grand introduced the PP guided tour, which enabled interactive visualisation of the optimisation in order to visually explore high-dimensional data.
-It is implemented in the R [XXX REF] package `tourr` [@tourr].
+It is implemented in the R [@R] package `tourr` [@tourr].
 The optimisation that is implemented is fairly basic, and potential problems were highlighted by @RJ-2021-105.
@@
 -->
+
+A tour on high-dimensional data is constructed by geodesically interpolating between pairs of planes.
+Any plane is described by an orthonormal basis, $A_t$, where $t$ represents time in the sequence.
+The term "geodesic" refers to maintaining the orthonormality constraint so that each view shown is correctly a projection of the data.
+The PP guided tour operates by geodesically interpolating to target planes (projections) which have high PP index values, as provided by the optimiser.
+The geodesic interpolation means that the viewer sees a continuous sequence of projections of the data, so they can watch patterns of interest forming as the function is optimised.
+There are five optimisation methods implemented in the `tourr` package:
+
+- `search_geodesic()`: provides a pseudo-derivative optimisation.
+It searches locally for the best direction, based on differencing the index values for very close projections.
+Then it follows the direction along the geodesic path between planes, stopping when the next index value fails to increase.
+- `search_better()`: is a brute-force optimisation searching randomly for projections with higher index values. 
+- `search_better_random()`: is essentially simulated annealing [@Bertsimas93] where the search space is reduced as the optimisation progresses. 
+- `search_posse()`: implements the algorithm described in @posse95. 
+- `search_polish()`: is a very localised search, to take tiny steps to get closer to the local maximum.
+
+There are several PP index functions available: `holes()` and `cmass()` [@cook1993projection]; `lda_pp()` [@lee2005projection]; `pda_pp()` [@lee2010projection]; `dcor2d()` and `splines2d()` [@Grimm2016];  `norm_bin()` and `norm_kol()` [@huber85]; `slice_index()` [@Laa:2020wkm].
+Most are relatively simply defined, for any projection dimension, and implemented because they are relatively easy to optimise.
+A goal is to be able to incorporate more complex PP indexes, for example based on scagnostics (@scag, @WW08).
+
+An initial investigation of PP indexes, and the potential for scagnostics is described in @laa_using_2020.
+To be useful here an optimiser needs to be able to handle functions which are not very smooth.
+In addition, because data structures might be relatively fine, the optimiser needs to be able to find maxima that occur with a small squint angle, that can only be seen from very close by.
+One last aspect that is useful is for an optimiser to return local maxima in addition to global because data can contain many different and interesting features.
 

### 2024-04-02 optim.tex
done with section 2

@@
 interactive visualisation of the optimisation in order to visually
-explore high-dimensional data. It is implemented in the R {[}XXX REF{]}
+explore high-dimensional data. It is implemented in the R \citep{R}
 package \texttt{tourr} \citep{tourr}. The optimisation that is
@@
 Sherry{]}}\label{sec-background}
+
+A tour on high-dimensional data is constructed by geodesically
+interpolating between pairs of planes. Any plane is described by an
+orthonormal basis, \(A_t\), where \(t\) represents time in the sequence.
+The term ``geodesic'' refers to maintaining the orthonormality
+constraint so that each view shown is correctly a projection of the
+data. The PP guided tour operates by geodesically interpolating to
+target planes (projections) which have high PP index values, as provided
+by the optimiser. The geodesic interpolation means that the viewer sees
+a continuous sequence of projections of the data, so they can watch
+patterns of interest forming as the function is optimised. There are
+five optimisation methods implemented in the \texttt{tourr} package:
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{search\_geodesic()}: provides a pseudo-derivative
+  optimisation. It searches locally for the best direction, based on
+  differencing the index values for very close projections. Then it
+  follows the direction along the geodesic path between planes, stopping
+  when the next index value fails to increase.
+\item
+  \texttt{search\_better()}: is a brute-force optimisation searching
+  randomly for projections with higher index values.
+\item
+  \texttt{search\_better\_random()}: is essentially simulated annealing
+  \citep{Bertsimas93} where the search space is reduced as the
+  optimisation progresses.
+\item
+  \texttt{search\_posse()}: implements the algorithm described in
+  \citet{posse95}.
+\item
+  \texttt{search\_polish()}: is a very localised search, to take tiny
+  steps to get closer to the local maximum.
+\end{itemize}
+
+There are several PP index functions available: \texttt{holes()} and
+\texttt{cmass()} \citep{cook1993projection}; \texttt{lda\_pp()}
+\citep{lee2005projection}; \texttt{pda\_pp()} \citep{lee2010projection};
+\texttt{dcor2d()} and \texttt{splines2d()} \citep{Grimm2016};
+\texttt{norm\_bin()} and \texttt{norm\_kol()} \citep{huber85};
+\texttt{slice\_index()} \citep{Laa:2020wkm}. Most are relatively simply
+defined, for any projection dimension, and implemented because they are
+relatively easy to optimise. A goal is to be able to incorporate more
+complex PP indexes, for example based on scagnostics (\citet{scag},
+\citet{WW08}).
+
+An initial investigation of PP indexes, and the potential for
+scagnostics is described in \citet{laa_using_2020}. To be useful here an
+optimiser needs to be able to handle functions which are not very
+smooth. In addition, because data structures might be relatively fine,
+the optimiser needs to be able to find maxima that occur with a small
+squint angle, that can only be seen from very close by. One last aspect
+that is useful is for an optimiser to return local maxima in addition to
+global because data can contain many different and interesting features.
 

### 2024-04-04 optim.qmd
added pseudocode

@@
 title: Studying the Performance of the Jellyfish Optimiser for the Application of Projection Pursuit
+header-includes:
+   - \usepackage{algorithm}
+   - \usepackage{algpseudocode}
 author:
@@
       cite-style: numbername
+editor: 
+  markdown: 
+    wrap: 72
 ---
 
-*Let's use British English ("American or British usage is accepted, but not a mixture of these")*
+*Let's use British English ("American or British usage is accepted, but
+not a mixture of these")*
 
@@
 
-The artificial jellyfish search (JS) algorithm [@chou_novel_2021] is a swarm-based metaheuristic optimisation algorithm inspired by the search behaviour of jellyfish in the ocean.
-It is one of the newest swarm intelligence algorithms [@rajwar_exhaustive_2023], which was shown to have stronger search ability and faster convergence with few algorithmic parameters compared to classic optimization methods [@chou_novel_2021]-[@chou_recent_2022].
+The artificial jellyfish search (JS) algorithm [@chou_novel_2021] is a
+swarm-based metaheuristic optimisation algorithm inspired by the search
+behaviour of jellyfish in the ocean. It is one of the newest swarm
+intelligence algorithms [@rajwar_exhaustive_2023], which was shown to
+have stronger search ability and faster convergence with few algorithmic
+parameters compared to classic optimization methods
+[@chou_novel_2021]-[@chou_recent_2022].
 
-Effective optimisation is an important aspect of many methods employed for visualising high-dimensional data ($X$).
-Here we are concerned about computing informative linear projections of high-dimensional ($p$) data using projection pursuit (PP) (@kr69, @FT74).
-This involves optimising a function  (e.g.
-@hall1989polynomial, @cook1993projection, @lee2010projection, @Loperfido2018, @Loperfido2020), called the projection pursuit index (PPI), that defines what is interesting or informative in a projection.
+Effective optimisation is an important aspect of many methods employed
+for visualising high-dimensional data ($X$). Here we are concerned about
+computing informative linear projections of high-dimensional ($p$) data
+using projection pursuit (PP) (@kr69, @FT74). This involves optimising a
+function (e.g. @hall1989polynomial, @cook1993projection,
+@lee2010projection, @Loperfido2018, @Loperfido2020), called the
+projection pursuit index (PPI), that defines what is interesting or
+informative in a projection.
 
-These PPI are defined on projections ($XA$), which means that there is a constraint that needs to be considered when optimising.
-A projection of data is defined by a $p\times d$ orthonormal matrix $A$, and this imposes the constraint on the elements of $A$, that columns need have norm equal to 1 and the product of columns need to sum to zero.
+These PPI are defined on projections ($XA$), which means that there is a
+constraint that needs to be considered when optimising. A projection of
+data is defined by a $p\times d$ orthonormal matrix $A$, and this
+imposes the constraint on the elements of $A$, that columns need have
+norm equal to 1 and the product of columns need to sum to zero.
 
-@cook1995grand introduced the PP guided tour, which enabled interactive visualisation of the optimisation in order to visually explore high-dimensional data.
-It is implemented in the R [@R] package `tourr` [@tourr].
-The optimisation that is implemented is fairly basic, and potential problems were highlighted by @RJ-2021-105.
-Implementing better optimisation functionality is a goal, but it needs to be kept in mind that the guided tour also has places importance on watching the projected data as the optimisation progresses.
+@cook1995grand introduced the PP guided tour, which enabled interactive
+visualisation of the optimisation in order to visually explore
+high-dimensional data. It is implemented in the R [@R] package `tourr`
+[@tourr]. The optimisation that is implemented is fairly basic, and
+potential problems were highlighted by @RJ-2021-105. Implementing better
+optimisation functionality is a goal, but it needs to be kept in mind
+that the guided tour also has places importance on watching the
+projected data as the optimisation progresses.
 
-Here we explore the potential for a jellyfish optimisation to be integrated with the guided tour.
-@sec-background explains the optimisation that is used in the current the projection pursuit guided tour.
-@sec-theory provides more details on the jellyfish optimiser and formalises several characteristics of projection pursuit indexes that are help to measure optimisaer performance.
-@sec-simulation describes a simulation study on performance of the jellyfish for several types of data and index functions.
-@sec-conclusion summarises the work and provides suggestions for future directions.
+Here we explore the potential for a jellyfish optimisation to be
+integrated with the guided tour. @sec-background explains the
+optimisation that is used in the current the projection pursuit guided
+tour. @sec-theory provides more details on the jellyfish optimiser and
+formalises several characteristics of projection pursuit indexes that
+are help to measure optimisaer performance. @sec-simulation describes a
+simulation study on performance of the jellyfish for several types of
+data and index functions. @sec-conclusion summarises the work and
+provides suggestions for future directions.
 
@@
 
-A tour on high-dimensional data is constructed by geodesically interpolating between pairs of planes.
-Any plane is described by an orthonormal basis, $A_t$, where $t$ represents time in the sequence.
-The term "geodesic" refers to maintaining the orthonormality constraint so that each view shown is correctly a projection of the data.
-The PP guided tour operates by geodesically interpolating to target planes (projections) which have high PP index values, as provided by the optimiser.
-The geodesic interpolation means that the viewer sees a continuous sequence of projections of the data, so they can watch patterns of interest forming as the function is optimised.
-There are five optimisation methods implemented in the `tourr` package:
+A tour on high-dimensional data is constructed by geodesically
+interpolating between pairs of planes. Any plane is described by an
+orthonormal basis, $A_t$, where $t$ represents time in the sequence. The
+term "geodesic" refers to maintaining the orthonormality constraint so
+that each view shown is correctly a projection of the data. The PP
+guided tour operates by geodesically interpolating to target planes
+(projections) which have high PP index values, as provided by the
+optimiser. The geodesic interpolation means that the viewer sees a
+continuous sequence of projections of the data, so they can watch
+patterns of interest forming as the function is optimised. There are
+five optimisation methods implemented in the `tourr` package:
 
-- `search_geodesic()`: provides a pseudo-derivative optimisation.
-It searches locally for the best direction, based on differencing the index values for very close projections.
-Then it follows the direction along the geodesic path between planes, stopping when the next index value fails to increase.
-- `search_better()`: is a brute-force optimisation searching randomly for projections with higher index values. 
-- `search_better_random()`: is essentially simulated annealing [@Bertsimas93] where the search space is reduced as the optimisation progresses. 
+-   `search_geodesic()`: provides a pseudo-derivative optimisation. It
+    searches locally for the best direction, based on differencing the
+    index values for very close projections. Then it follows the
+    direction along the geodesic path between planes, stopping when the
+    next index value fails to increase.
+-   `search_better()`: is a brute-force optimisation searching randomly
+    for projections with higher index values.
+-   `search_better_random()`: is essentially simulated annealing
+    [@Bertsimas93] where the search space is reduced as the optimisation
+    progresses.
 -   `search_posse()`: implements the algorithm described in @posse95.
-- `search_polish()`: is a very localised search, to take tiny steps to get closer to the local maximum.
+-   `search_polish()`: is a very localised search, to take tiny steps to
+    get closer to the local maximum.
 
-There are several PP index functions available: `holes()` and `cmass()` [@cook1993projection]; `lda_pp()` [@lee2005projection]; `pda_pp()` [@lee2010projection]; `dcor2d()` and `splines2d()` [@Grimm2016];  `norm_bin()` and `norm_kol()` [@huber85]; `slice_index()` [@Laa:2020wkm].
-Most are relatively simply defined, for any projection dimension, and implemented because they are relatively easy to optimise.
-A goal is to be able to incorporate more complex PP indexes, for example based on scagnostics (@scag, @WW08).
+There are several PP index functions available: `holes()` and `cmass()`
+[@cook1993projection]; `lda_pp()` [@lee2005projection]; `pda_pp()`
+[@lee2010projection]; `dcor2d()` and `splines2d()` [@Grimm2016];
+`norm_bin()` and `norm_kol()` [@huber85]; `slice_index()`
+[@Laa:2020wkm]. Most are relatively simply defined, for any projection
+dimension, and implemented because they are relatively easy to optimise.
+A goal is to be able to incorporate more complex PP indexes, for example
+based on scagnostics (@scag, @WW08).
 
-An initial investigation of PP indexes, and the potential for scagnostics is described in @laa_using_2020.
-To be useful here an optimiser needs to be able to handle functions which are not very smooth.
-In addition, because data structures might be relatively fine, the optimiser needs to be able to find maxima that occur with a small squint angle, that can only be seen from very close by.
-One last aspect that is useful is for an optimiser to return local maxima in addition to global because data can contain many different and interesting features.
+An initial investigation of PP indexes, and the potential for
+scagnostics is described in @laa_using_2020. To be useful here an
+optimiser needs to be able to handle functions which are not very
+smooth. In addition, because data structures might be relatively fine,
+the optimiser needs to be able to find maxima that occur with a small
+squint angle, that can only be seen from very close by. One last aspect
+that is useful is for an optimiser to return local maxima in addition to
+global because data can contain many different and interesting features.
 
@@
 
-The jellyfish optimiser (JSO) mimics the natural movements of jellyfish, which include passive and active motions driven by ocean currents and their swimming patterns, respectively.
-In the context of optimization, these movements are abstracted to explore the search space in a way that balances exploration (searching new areas) and exploitation (focusing on promising areas).
-The algorithm aims to find the optimal solution by adapting the jellyfish's behavior to navigate towards the best solution over iterations [@chou_novel_2021].
+The jellyfish optimiser (JSO) mimics the natural movements of jellyfish,
+which include passive and active motions driven by ocean currents and
+their swimming patterns, respectively. In the context of optimization,
+these movements are abstracted to explore the search space in a way that
+balances exploration (searching new areas) and exploitation (focusing on
+promising areas). The algorithm aims to find the optimal solution by
+adapting the jellyfish's behavior to navigate towards the best solution
+over iterations [@chou_novel_2021].
 
-Below is the pseudo-code for this visualisation application.
+To understand what the jellyfish optimizer is doing in the context of
+Projection Pursuit, we first start with a current projection (the
+starting point). Then we evaluate this projection using an index
+function, which tells us how good the current projection is. We then
+move the projection in a direction determined by the 'best jelly' and
+random factors, influenced by how far along we are in the optimization
+process (the trial $i$ and `max.tries`). Occasionally, we might explore
+completely new directions like a jellyfish might with ocean currents.
+Then, we compare new potential projections to our current one. If
+they're better, we adopt them; if not, we stick with our current
+projection. This process continues and iteratively improves the
+projection, until we reach the maximum number of trials.
 
-\[Put the pseudo-code in, add specifics to this visualisation application\]
+::: {.callout-note icon="false"}
 
-The JSO implementation involves several key parameters that control its search process in optimization problems.
-These parameters are designed to guide the exploration and exploitation phases of the algorithm.
-While the specific implementation details can vary depending on the version of the algorithm or its application, we focus on two main parameters that are most relevant to our application: the number of jellyfish and drift.
+## Algorithm: Jellyfish Optimizer Pseudo Code
 
-@laa_using_2020 has proposed five criteria for assessing projection pursuit indexes (smoothness, squintability, flexibility, rotation invariance, and speed).
-Since not all the properties affects the execution of the optimisation, here we consider the three relevant properties (smoothness, squintability, and speed), and propose three metrics to evaluate these three properties.
+**Input**: `current_projections`, `index_function`, `tries`, `max_tries`
+
+**Output**: `optimized_projection`
+
+**Initialize** `best_jelly` as the projection with the best index value from `current_projections`, and 
+`current_index` as the array of index values for each projection in `current_projections`
+
+>
+
+**for** each try in 1 to max_tries **do**
+
+> Calculate $c_t$ based on the current try and max_tries
+
+> **if** $c_t$ is greater than or equal to $0.5$ **then**
+>    
+> > Define trend based on the best jelly and current projections
+>     
+> > Update each projection towards the trend using a random factor and orthonormalisation
+>       
+> **else**
+>     
+> > **if** a random number is greater than $1 - c_t$ **then**
+> >       
+> > > Slightly adjust each projection with a small random factor (Type A passive)
+> >         
+> > **else**
+> >       
+> > > For each projection, compare with a random jelly and adjust towards or away from it (Type B active)
+>         
+> Update the orientation of each projection to maintain consistency
+>     
+> Evaluate the new projections using the index function
+
+>
+
+> **if** any new projection is worse than the current, revert to the `current_projection` for that case
+> 
+> > Determine the projection with the best index value as the new best_jelly
+
+>
+
+> **if** the try is the last one, print the final best projection and **exit**
+
+**return** the set of projections with the updated best_jelly as the optimized_projection
+
+:::
+
+<!-- Var names in the function and in the text needs to be consistent??? -->
+
+The JSO implementation involves several key parameters that control its
+search process in optimization problems. These parameters are designed
+to guide the exploration and exploitation phases of the algorithm. While
+the specific implementation details can vary depending on the version of
+the algorithm or its application, we focus on two main parameters that
+are most relevant to our application: the number of jellyfish and drift.
+
+@laa_using_2020 has proposed five criteria for assessing projection
+pursuit indexes (smoothness, squintability, flexibility, rotation
+invariance, and speed). Since not all the properties affects the
+execution of the optimisation, here we consider the three relevant
+properties (smoothness, squintability, and speed), and propose three
+metrics to evaluate these three properties.
 
@@
 
-An intuitive way to measure smoothness of a function would be to count how many continuous derivatives exist.
-To help define smoothness in our context, we can make use of the Sobolev spaces: functions $f$ in the Sobolev space $W^{p,\infty}$ have all derivatives of order less than $p$ continuous.
-Smoothness would then be the highest $p$ such that the index function belongs to $W^{p,\infty}$.
-We can relax this and consider $W^{p,q}$ Sobolev spaces.
+<!--
+An intuitive way to measure smoothness of a function would be to count
+how many continuous derivatives exist. 
+To help define smoothness in our
+context, we can make use of the Sobolev spaces: functions $f$ in the
+Sobolev space $W^{p,\infty}$ have all derivatives of order less than $p$
+continuous. Smoothness would then be the highest $p$ such that the index
+function belongs to $W^{p,\infty}$. We can relax this and consider
+$W^{p,q}$ Sobolev spaces.
 
-Consider the following definition of partial derivative.
-Let $U$ be an open subset of $\mathbf{R}^n$ and $f: U \rightarrow \mathbf{R}$.
-The partial derivative of function $f$ at the point $\mathbf{a} = (a_1, \ldots, a_n) \in U$ with respective to the $i$-th variable $x_i$ is defined as
+Consider the following definition of partial derivative. Let $U$ be an
+open subset of $\mathbf{R}^n$ and $f: U \rightarrow \mathbf{R}$. The
+partial derivative of function $f$ at the point
+$\mathbf{a} = (a_1, \ldots, a_n) \in U$ with respective to the $i$-th
+variable $x_i$ is defined as
 
@@
 
-If the derivative is not well-defined, we propose to approximate the derivative using the following expression:
+If the derivative is not well-defined, we propose to approximate the
+derivative using the following expression:
 
@@
 
-where $p \in [0,1]$. The choice of $p$ reflects the penalty behaviour. For the application in this paper, we choose $p = 1$.
+where $p \in [0,1]$. The choice of $p$ reflects the penalty behaviour.
+For the application in this paper, we choose $p = 1$.
 
--   $h_i$ is an neighbourhood area around $a$. We can insert different values of $h$ to approximate the above quantity and observe how this quantity changes.
+-   $h_i$ is an neighbourhood area around $a$. We can insert different
+    values of $h$ to approximate the above quantity and observe how this
+    quantity changes.
+-->
+
+If we evaluate the index function at some random points (like the random initialization of the jellyfish optimizer), then we can interpret these random index values as a random field, indexed by a space parameter: the random projection angle.
+This analogy suggests to use this random training sample to fit a spatial model, a simple one being a (spatial) Gaussian process.
+There exist R packages to fit the hyperparameters of a Gaussian process, for example ExaGeoStatR (https://github.com/ecrc/exageostatR), which is an R wrapper for the ExaGeoStat library (https://arxiv.org/pdf/1708.02835.pdf).
+
+How can we define a measure of smoothness from this?
+A Gaussian process is fully defined by its mean function and covariance function.
+The way the covariance function is defined is where smoothness comes into play: if an index is very smooth, then two close projection angles should produce close index values (strong correlation); by contrast, if an index is not smooth, then two close projection angles might give very different index values (fast decay of correlations with respect to distance between angles).
+
+Luckily, the most popular class of covariance functions, known as Matérn covariance functions (see https://arxiv.org/pdf/1708.02835.pdf on page 3) has one parameter, Theta3, which controls the decay of the covariance function, which in turns controls the smoothness of the random field.
+The ExaGeoStat paper calls it smoothness parameter: the higher Theta3, the smoother the random field.
+In our context, the higher Theta3, the smoother the index function.
+
+This suggests the following approach:
+
+- Evaluate the index function at some random projection angle values (our initial jellies),
+- Use this sample to fit a Gaussian process with Matérn covariance using ExaGeoStatR,
+- Use the fitted Theta3 parameter value as our definition of smoothness of the index function: the higher Theta3, the smoother the index.
 
@@
 
-From the literature, it is commonly understood that a large squint angle implies that the function is easy to optimize, because we do not need to be very close to the perfect view to see the structure.
-A small squint angle means that the derivative of the index function can still be very large values near the optimal point and will rapidly change as we get even closer to the optimum.
-As such we can observe the second order gradient, which is the rate of change of gradient, over the space we are searching.
-For some index functions, the second order gradient is not well-defined, we can approximate the second order gradient vector in similar fashion as the above section.
+From the literature, it is commonly understood that a large squint angle
+implies that the function is easy to optimize, because we do not need to
+be very close to the perfect view to see the structure. A small squint
+angle means that the derivative of the index function can still be very
+large values near the optimal point and will rapidly change as we get
+even closer to the optimum. As such we can observe the second order
+gradient, which is the rate of change of gradient, over the space we are
+searching. For some index functions, the second order gradient is not
+well-defined, we can approximate the second order gradient vector in
+similar fashion as the above section.
 
-To the best of our knowledge, this is the first attempt to measure the notion of squintability.
+To the best of our knowledge, this is the first attempt to measure the
+notion of squintability.
 
@@
 
-The speed of optimizing an index function can be calculated/measured using the computational complexity (in big O notation, with respect to the sample size) of computing the index function.
+The speed of optimizing an index function can be calculated/measured
+using the computational complexity (in big O notation, with respect to
+the sample size) of computing the index function.
 
@@
 
-The jellyfish optimiser has been implemented in the tourr package [@wickham_tourr_2011] and we will use the diagnostic plots proposed in the ferrn package [@RJ-2021-105] to visualise the optimisation process.
+The jellyfish optimiser has been implemented in the tourr package
+[@wickham_tourr_2011] and we will use the diagnostic plots proposed in
+the ferrn package [@RJ-2021-105] to visualise the optimisation process.
 
@@
 
-The pipe-finding problem is initially used to investigate indexes and optimisers in @laa_using_2020, and we extend it from a 6D problem to a 12D problem.
+The pipe-finding problem is initially used to investigate indexes and
+optimisers in @laa_using_2020, and we extend it from a 6D problem to a
+12D problem.
 
-Jellyfish optimiser, as a multi-start algorithm, is efficient in \[...\] for high-dimensional problems
+Jellyfish optimiser, as a multi-start algorithm, is efficient in \[...\]
+for high-dimensional problems
 

### 2024-04-04 optim.tex
added pseudocode

@@
 
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
+\definecolor{quarto-callout-color}{HTML}{909090}
+\definecolor{quarto-callout-note-color}{HTML}{0758E5}
+\definecolor{quarto-callout-important-color}{HTML}{CC1914}
+\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
+\definecolor{quarto-callout-tip-color}{HTML}{00A047}
+\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
+\definecolor{quarto-callout-color-frame}{HTML}{acacac}
+\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
+\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
+\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
+\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
+\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
 \makeatletter
@@
 \makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
 \makeatother
 \makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
 \makeatother
@@
 \bibliographystyle{elsarticle-num-names}
-\usepackage{bookmark}
-
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \end{frontmatter}
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, boxrule=0pt, frame hidden, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 
+\hypertarget{introduction-nicolas-and-jessica}{%
 \section{Introduction {[}Nicolas and
-Jessica{]}}\label{introduction-nicolas-and-jessica}
+Jessica{]}}\label{introduction-nicolas-and-jessica}}
 
@@
 
+\hypertarget{sec-background}{%
 \section{Projection pursuit, index functions and optimisation {[}Di and
-Sherry{]}}\label{sec-background}
+Sherry{]}}\label{sec-background}}
 
@@
 
+\hypertarget{sec-theory}{%
 \section{The jellyfish optimiser and properties of PP indexes {[}Nicolas
-and Jessica{]}}\label{sec-theory}
+and Jessica{]}}\label{sec-theory}}
 
@@
 
-Below is the pseudo-code for this visualisation application.
+To understand what the jellyfish optimizer is doing in the context of
+Projection Pursuit, we first start with a current projection (the
+starting point). Then we evaluate this projection using an index
+function, which tells us how good the current projection is. We then
+move the projection in a direction determined by the `best jelly' and
+random factors, influenced by how far along we are in the optimization
+process (the trial \(i\) and \texttt{max.tries}). Occasionally, we might
+explore completely new directions like a jellyfish might with ocean
+currents. Then, we compare new potential projections to our current one.
+If they're better, we adopt them; if not, we stick with our current
+projection. This process continues and iteratively improves the
+projection, until we reach the maximum number of trials.
 
-{[}Put the pseudo-code in, add specifics to this visualisation
-application{]}
+\begin{tcolorbox}[enhanced jigsaw, coltitle=black, titlerule=0mm, opacitybacktitle=0.6, leftrule=.75mm, left=2mm, bottomtitle=1mm, bottomrule=.15mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, breakable, colbacktitle=quarto-callout-note-color!10!white, colback=white, rightrule=.15mm, opacityback=0, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm]
+
+\textbf{Input}: \texttt{current\_projections}, \texttt{index\_function},
+\texttt{tries}, \texttt{max\_tries}
+
+\textbf{Output}: \texttt{optimized\_projection}
+
+\textbf{Initialize} \texttt{best\_jelly} as the projection with the best
+index value from \texttt{current\_projections}, and
+\texttt{current\_index} as the array of index values for each projection
+in \texttt{current\_projections}
+
+\begin{quote}
+\end{quote}
+
+\textbf{for} each try in 1 to max\_tries \textbf{do}
+
+\begin{quote}
+Calculate \(c_t\) based on the current try and max\_tries
+\end{quote}
+
+\begin{quote}
+\textbf{if} \(c_t\) is greater than or equal to \(0.5\) \textbf{then}
+
+\begin{quote}
+Define trend based on the best jelly and current projections
+\end{quote}
+
+\begin{quote}
+Update each projection towards the trend using a random factor and
+orthonormalisation
+\end{quote}
+
+\textbf{else}
+
+\begin{quote}
+\textbf{if} a random number is greater than \(1 - c_t\) \textbf{then}
+
+\begin{quote}
+Slightly adjust each projection with a small random factor (Type A
+passive)
+\end{quote}
+
+\textbf{else}
+
+\begin{quote}
+For each projection, compare with a random jelly and adjust towards or
+away from it (Type B active)
+\end{quote}
+\end{quote}
+
+Update the orientation of each projection to maintain consistency
+
+Evaluate the new projections using the index function
+\end{quote}
+
+\begin{quote}
+\end{quote}
+
+\begin{quote}
+\textbf{if} any new projection is worse than the current, revert to the
+\texttt{current\_projection} for that case
+
+\begin{quote}
+Determine the projection with the best index value as the new
+best\_jelly
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\end{quote}
+
+\begin{quote}
+\textbf{if} the try is the last one, print the final best projection and
+\textbf{exit}
+\end{quote}
+
+\textbf{return} the set of projections with the updated best\_jelly as
+the optimized\_projection
+
+\end{tcolorbox}
 
@@
 
-\subsection{Smoothness}\label{smoothness}
+\hypertarget{smoothness}{%
+\subsection{Smoothness}\label{smoothness}}
 
-An intuitive way to measure smoothness of a function would be to count
-how many continuous derivatives exist. To help define smoothness in our
-context, we can make use of the Sobolev spaces: functions \(f\) in the
-Sobolev space \(W^{p,\infty}\) have all derivatives of order less than
-\(p\) continuous. Smoothness would then be the highest \(p\) such that
-the index function belongs to \(W^{p,\infty}\). We can relax this and
-consider \(W^{p,q}\) Sobolev spaces.
+If we evaluate the index function at some random points (like the random
+initialization of the jellyfish optimizer), then we can interpret these
+random index values as a random field, indexed by a space parameter: the
+random projection angle. This analogy suggests to use this random
+training sample to fit a spatial model, a simple one being a (spatial)
+Gaussian process. There exist R packages to fit the hyperparameters of a
+Gaussian process, for example ExaGeoStatR
+(https://github.com/ecrc/exageostatR), which is an R wrapper for the
+ExaGeoStat library (https://arxiv.org/pdf/1708.02835.pdf).
 
-Consider the following definition of partial derivative. Let \(U\) be an
-open subset of \(\mathbf{R}^n\) and \(f: U \rightarrow \mathbf{R}\). The
-partial derivative of function \(f\) at the point
-\(\mathbf{a} = (a_1, \ldots, a_n) \in U\) with respective to the
-\(i\)-th variable \(x_i\) is defined as
+How can we define a measure of smoothness from this? A Gaussian process
+is fully defined by its mean function and covariance function. The way
+the covariance function is defined is where smoothness comes into play:
+if an index is very smooth, then two close projection angles should
+produce close index values (strong correlation); by contrast, if an
+index is not smooth, then two close projection angles might give very
+different index values (fast decay of correlations with respect to
+distance between angles).
 
-\[
-\frac{\delta}{\delta x_i} f(\mathbf{a}) =  \lim_{h \rightarrow 0} \frac{f(a_1, \ldots, a_{i-1}, a_i +h, a_{i+1}, \ldots, a_n) - f(a_1 , \ldots, a_i, \ldots, a_n)}{h} = \lim_{h\rightarrow 0} \frac{f(\mathbf{a} + h\mathbf{e_i}) - f(\mathbf{a})}{h}.
-\] where \(\mathbf{e_i}\) is the unit vector of the \(i\)-th variable
-\(x_i\).
+Luckily, the most popular class of covariance functions, known as Matérn
+covariance functions (see https://arxiv.org/pdf/1708.02835.pdf on page
+3) has one parameter, Theta3, which controls the decay of the covariance
+function, which in turns controls the smoothness of the random field.
+The ExaGeoStat paper calls it smoothness parameter: the higher Theta3,
+the smoother the random field. In our context, the higher Theta3, the
+smoother the index function.
 
-If the derivative is not well-defined, we propose to approximate the
-derivative using the following expression:
-
-\[\frac{1}{n_i}\sum_{i} \frac{\|f(\mathbf{a} + h_i\mathbf{e_i}) - f(\mathbf{a})\|^p}{h_i}\]
-
-where \(p \in [0,1]\). The choice of \(p\) reflects the penalty
-behaviour. For the application in this paper, we choose \(p = 1\).
+This suggests the following approach:
 
@@
 \item
-  \(h_i\) is an neighbourhood area around \(a\). We can insert different
-  values of \(h\) to approximate the above quantity and observe how this
-  quantity changes.
+  Evaluate the index function at some random projection angle values
+  (our initial jellies),
+\item
+  Use this sample to fit a Gaussian process with Matérn covariance using
+  ExaGeoStatR,
+\item
+  Use the fitted Theta3 parameter value as our definition of smoothness
+  of the index function: the higher Theta3, the smoother the index.
 \end{itemize}
 
-\subsection{Squintability}\label{squintability}
+\hypertarget{squintability}{%
+\subsection{Squintability}\label{squintability}}
 
@@
 
-\subsection{Speed}\label{speed}
+\hypertarget{speed}{%
+\subsection{Speed}\label{speed}}
 
@@
 
-\section{Application {[}Di and Sherry{]}}\label{sec-simulation}
+\hypertarget{sec-simulation}{%
+\section{Application {[}Di and Sherry{]}}\label{sec-simulation}}
 
@@
 
-\subsection{Going beyond 10D}\label{going-beyond-10d}
+\hypertarget{going-beyond-10d}{%
+\subsection{Going beyond 10D}\label{going-beyond-10d}}
 
@@
 
-\begin{figure}[H]
+\begin{figure}
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-2-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-4-1.pdf}
 
@@
 
-\end{figure}%
+\end{figure}
 
-\begin{figure}[H]
+\begin{figure}
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-3-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-6-1.pdf}
 
@@
 
-\end{figure}%
+\end{figure}
 
+\hypertarget{on-skewness-and-kurtosis-index}{%
 \subsection{On skewness and kurtosis
-index}\label{on-skewness-and-kurtosis-index}
+index}\label{on-skewness-and-kurtosis-index}}
 
-\subsection{Another data example}\label{another-data-example}
+\hypertarget{another-data-example}{%
+\subsection{Another data example}\label{another-data-example}}
 
-\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
+\hypertarget{sec-conclusion}{%
+\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}}
 

### 2024-04-12 optim.tex
fixed reference, should used published version not arxiv

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, boxrule=0pt, frame hidden, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{sec-background}{%
 \section{Projection pursuit, index functions and optimisation {[}Di and
-Sherry{]}}\label{sec-background}}
+Sherry{]}}\label{sec-background}
 
@@
 
-\hypertarget{sec-theory}{%
 \section{The jellyfish optimiser and properties of PP indexes {[}Nicolas
-and Jessica{]}}\label{sec-theory}}
+and Jessica{]}}\label{sec-theory}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, coltitle=black, titlerule=0mm, opacitybacktitle=0.6, leftrule=.75mm, left=2mm, bottomtitle=1mm, bottomrule=.15mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, breakable, colbacktitle=quarto-callout-note-color!10!white, colback=white, rightrule=.15mm, opacityback=0, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm]
+\begin{tcolorbox}[enhanced jigsaw, coltitle=black, bottomtitle=1mm, titlerule=0mm, toprule=.15mm, breakable, colback=white, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, opacityback=0, toptitle=1mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, bottomrule=.15mm, arc=.35mm, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, left=2mm]
 

### 2024-04-12 optim.qmd
smoothness measure
Update of the smoothness measure part with equations

@@
 promising areas). The algorithm aims to find the optimal solution by
-adapting the jellyfish's behavior to navigate towards the best solution
+adapting the jellyfish's behaviour to navigate towards the best solution
 over iterations [@chou_novel_2021].
@@
 Projection Pursuit, we first start with a current projection (the
-starting point). Then we evaluate this projection using an index
+starting point). Then, we evaluate this projection using an index
 function, which tells us how good the current projection is. We then
@@
 ::: {.callout-note icon="false"}
-
 ## Algorithm: Jellyfish Optimizer Pseudo Code
@@
 
-**Initialize** `best_jelly` as the projection with the best index value from `current_projections`, and 
-`current_index` as the array of index values for each projection in `current_projections`
-
->
+**Initialize** `best_jelly` as the projection with the best index value
+from `current_projections`, and `current_index` as the array of index
+values for each projection in `current_projections`
 
@@
 >
-> > Update each projection towards the trend using a random factor and orthonormalisation
+> > Update each projection towards the trend using a random factor and
+> > orthonormalisation
 >
@@
 > >
-> > > Slightly adjust each projection with a small random factor (Type A passive)
+> > > Slightly adjust each projection with a small random factor (Type A
+> > > passive)
 > >
@@
 > >
-> > > For each projection, compare with a random jelly and adjust towards or away from it (Type B active)
+> > > For each projection, compare with a random jelly and adjust
+> > > towards or away from it (Type B active)
 >
@@
 
+> **if** any new projection is worse than the current, revert to the
+> `current_projection` for that case
 >
+> > Determine the projection with the best index value as the new
+> > best_jelly
 
-> **if** any new projection is worse than the current, revert to the `current_projection` for that case
-> 
-> > Determine the projection with the best index value as the new best_jelly
+> **if** the try is the last one, print the final best projection and
+> **exit**
 
->
-
-> **if** the try is the last one, print the final best projection and **exit**
-
-**return** the set of projections with the updated best_jelly as the optimized_projection
-
+**return** the set of projections with the updated best_jelly as the
+optimized_projection
 :::
@@
 
+```{=html}
 <!--
@@
 -->
+```
+If we evaluate the index function at some random points (like the random
+initialization of the jellyfish optimizer), then we can interpret these
+random index values as a random field, indexed by a space parameter: the
+random projection angle. This analogy suggests to use this random
+training sample to fit a spatial model, a simple one being a (spatial)
+Gaussian process.
 
-If we evaluate the index function at some random points (like the random initialization of the jellyfish optimizer), then we can interpret these random index values as a random field, indexed by a space parameter: the random projection angle.
-This analogy suggests to use this random training sample to fit a spatial model, a simple one being a (spatial) Gaussian process.
+```{=html}
+<!--
 There exist R packages to fit the hyperparameters of a Gaussian process, for example ExaGeoStatR (https://github.com/ecrc/exageostatR), which is an R wrapper for the ExaGeoStat library (https://arxiv.org/pdf/1708.02835.pdf).
+-->
+```
 
-How can we define a measure of smoothness from this?
-A Gaussian process is fully defined by its mean function and covariance function.
-The way the covariance function is defined is where smoothness comes into play: if an index is very smooth, then two close projection angles should produce close index values (strong correlation); by contrast, if an index is not smooth, then two close projection angles might give very different index values (fast decay of correlations with respect to distance between angles).
+How can we define a measure of smoothness from this? The distribution of
+a Gaussian process is fully determined by its mean function and
+covariance function. The way the covariance function is defined is where
+smoothness comes into play: if an index is very smooth, then two close
+projection angles should produce close index values (strong
+correlation); by contrast, if an index is not smooth, then two close
+projection angles might give very different index values (fast decay of
+correlations with respect to distance between angles).
 
-Luckily, the most popular class of covariance functions, known as Matérn covariance functions (see https://arxiv.org/pdf/1708.02835.pdf on page 3) has one parameter, Theta3, which controls the decay of the covariance function, which in turns controls the smoothness of the random field.
-The ExaGeoStat paper calls it smoothness parameter: the higher Theta3, the smoother the random field.
-In our context, the higher Theta3, the smoother the index function.
+Popular covariance functions are parametric positive semi-definite
+functions, some of which have a parameter to capture the smoothness of
+the Gaussian field. In particular, consider the Matérn class of
+covariance functions, defined by
 
-This suggests the following approach:
+$$
+K(u):=\frac{(\sqrt{2\nu}u)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}(\sqrt{2\nu}u)
+$$
 
-- Evaluate the index function at some random projection angle values (our initial jellies),
-- Use this sample to fit a Gaussian process with Matérn covariance using ExaGeoStatR,
-- Use the fitted Theta3 parameter value as our definition of smoothness of the index function: the higher Theta3, the smoother the index.
+where $\nu>0$ is the smoothness parameter and where $\mathcal{K}_\nu$ is
+the modified Bessel function. The Matérn covariance function can be
+expressed analytically when $\nu$ is a half-integer, the most popular
+values in the literature being $1/2$, $3/2$ and $5/2$ . The parameter
+$\nu$, called smoothness parameter, controls the decay of the covariance
+function. As such, it is an appropriate measure of smoothness of a
+random field.
+
+In our context, we suggest to use this parameter as a measure of the
+smoothness of the index function by fitting a Gaussian process prior
+with Matérn covariance on a dataset generated by random evaluations of
+the index function, as in the initial stage of the jellyfish random
+search. There exist several R packages, such as GpGp or ExaGeoStatR, to
+fit the hyperparameters of a GP covariance function on data. In this
+project, we make use of the GpGp package.
+
+The fitted value $\nu>0$ can be interpreted as follows: the higher
+$\nu$, the smoother the index function.
 

### 2024-04-12 optim.tex
smoothness measure
Update of the smoothness measure part with equations

--- 2024-04-04 optim.tex
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, boxrule=0pt, frame hidden, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi
 
@@
 accepted, but not a mixture of these'')}
+
+\begin{verbatim}
+Warning: package 'ggplot2' was built under R version 4.3.2
+\end{verbatim}
+
+\begin{verbatim}
+Warning: package 'tidyr' was built under R version 4.3.2
+\end{verbatim}
 
@@
 promising areas). The algorithm aims to find the optimal solution by
-adapting the jellyfish's behavior to navigate towards the best solution
+adapting the jellyfish's behaviour to navigate towards the best solution
 over iterations \citep{chou_novel_2021}.
@@
 
-\begin{tcolorbox}[enhanced jigsaw, coltitle=black, titlerule=0mm, opacitybacktitle=0.6, leftrule=.75mm, left=2mm, bottomtitle=1mm, bottomrule=.15mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, breakable, colbacktitle=quarto-callout-note-color!10!white, colback=white, rightrule=.15mm, opacityback=0, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm]
+\begin{tcolorbox}[enhanced jigsaw, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, titlerule=0mm, toprule=.15mm, colback=white, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, opacitybacktitle=0.6, leftrule=.75mm, toptitle=1mm, coltitle=black, breakable, arc=.35mm, left=2mm, rightrule=.15mm]
 
@@
 in \texttt{current\_projections}
-
-\begin{quote}
-\end{quote}
 
@@
 \begin{quote}
-\end{quote}
-
-\begin{quote}
 \textbf{if} any new projection is worse than the current, revert to the
@@
 \end{quote}
-\end{quote}
-
-\begin{quote}
 \end{quote}
@@
 training sample to fit a spatial model, a simple one being a (spatial)
-Gaussian process. There exist R packages to fit the hyperparameters of a
-Gaussian process, for example ExaGeoStatR
-(https://github.com/ecrc/exageostatR), which is an R wrapper for the
-ExaGeoStat library (https://arxiv.org/pdf/1708.02835.pdf).
+Gaussian process.
 
-How can we define a measure of smoothness from this? A Gaussian process
-is fully defined by its mean function and covariance function. The way
-the covariance function is defined is where smoothness comes into play:
-if an index is very smooth, then two close projection angles should
-produce close index values (strong correlation); by contrast, if an
-index is not smooth, then two close projection angles might give very
-different index values (fast decay of correlations with respect to
-distance between angles).
+How can we define a measure of smoothness from this? The distribution of
+a Gaussian process is fully determined by its mean function and
+covariance function. The way the covariance function is defined is where
+smoothness comes into play: if an index is very smooth, then two close
+projection angles should produce close index values (strong
+correlation); by contrast, if an index is not smooth, then two close
+projection angles might give very different index values (fast decay of
+correlations with respect to distance between angles).
 
-Luckily, the most popular class of covariance functions, known as Matérn
-covariance functions (see https://arxiv.org/pdf/1708.02835.pdf on page
-3) has one parameter, Theta3, which controls the decay of the covariance
-function, which in turns controls the smoothness of the random field.
-The ExaGeoStat paper calls it smoothness parameter: the higher Theta3,
-the smoother the random field. In our context, the higher Theta3, the
-smoother the index function.
+Popular covariance functions are parametric positive semi-definite
+functions, some of which have a parameter to capture the smoothness of
+the Gaussian field. In particular, consider the Matérn class of
+covariance functions, defined by
 
-This suggests the following approach:
+\[
+K(u):=\frac{(\sqrt{2\nu}u)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}(\sqrt{2\nu}u)
+\]
 
-\begin{itemize}
-\tightlist
-\item
-  Evaluate the index function at some random projection angle values
-  (our initial jellies),
-\item
-  Use this sample to fit a Gaussian process with Matérn covariance using
-  ExaGeoStatR,
-\item
-  Use the fitted Theta3 parameter value as our definition of smoothness
-  of the index function: the higher Theta3, the smoother the index.
-\end{itemize}
+where \(\nu>0\) is the smoothness parameter and where
+\(\mathcal{K}_\nu\) is the modified Bessel function. The Matérn
+covariance function can be expressed analytically when \(\nu\) is a
+half-integer, the most popular values in the literature being \(1/2\),
+\(3/2\) and \(5/2\) . The parameter \(\nu\), called smoothness
+parameter, controls the decay of the covariance function. As such, it is
+an appropriate measure of smoothness of a random field.
+
+In our context, we suggest to use this parameter as a measure of the
+smoothness of the index function by fitting a Gaussian process prior
+with Matérn covariance on a dataset generated by random evaluations of
+the index function, as in the initial stage of the jellyfish random
+search. There exist several R packages, such as GpGp or ExaGeoStatR, to
+fit the hyperparameters of a GP covariance function on data. In this
+project, we make use of the GpGp package.
+
+The fitted value \(\nu>0\) can be interpreted as follows: the higher
+\(\nu\), the smoother the index function.
 
@@
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-4-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-2-1.pdf}
 
@@
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-6-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-3-1.pdf}
 

### 2024-04-19 optim.qmd
outline of the section visualisation of jellyfish optimiser

@@
 ```
-
 How can we define a measure of smoothness from this? The distribution of
@@
 the sample size) of computing the index function.
+
+# Visualisation of jellyfish optimiser
+
+Information of the jellyfish optimiser is available in a tabular format
+and below is an example data collected from finding the sine-wave
+structure in 6D data using a distance correlation index (`docr2d_2`):
+
+```{r}
+load(here::here("data/sim_sine_6d_dcor2d.rda"))
+```
+
+```{r}
+dplyr::glimpse(sim_sine_6d_dcor2d |> select(-id)) 
+```
+
+Information recorded can be categorised into the following categories:
+
+-   projection pursuit variables: the index function used (`idx_f`), the
+    data dimension (`d`)
+-   jellyfish optimiser parameters: the number of jellies (`n_jellies`),
+    the maximum number of tries (`max_tries`)
+-   simulation variables: the simulation number (`sim`), the seed used
+    (`seed`)
+-   optimisation variables: the projection basis in a matrix format
+    (`basis`), the index value (`index_val`), a description of the
+    status - one of "initiation", "current_best", and "jellyfish_update"
+    (`info`), current iteration ID (`tries`), current jelly ID (`loop`),
+    and the time taken to find the optimum (`time`).
+
+The basis column records every basis *visited* by the jellyfish
+optimiser prior to comparing with the current basis. In each iteration,
+if the index value of a visited basis is smaller than that of the
+current one, the jellyfish optimiser will retain the current basis for
+the next iteration, while still documenting the visited basis.
+
+Numerical information to compute:
+
+-   angular distance between the projection basis and the theoretical
+    best basis,
+-   the proportion of simulation that found the optimal basis,
+-   the proportion of jellies, within each simulation, that found the
+    optimal basis,
+
+Visualisation to inspect:
+
+-   inspect the basis visited by each jellyfish in the reduced PCA space,
+-   inspect the final 2D projections reached by each jellyfish,
+-   plot the index value against the angular distance between the
+    projection basis and the theoretical best basis
+
+Plotting the basis in the space and the projected data can help to
+understand 1) whether each simulation finds the same optimum or some
+simulations find local optima; and 2) whether the index function used
+can detect the structure in the data and the projection contains the
+structure of interest.
+
+The visualisations above can be faceted by the projection pursuit
+variables and jellyfish optimiser parameters to compare the performance
+of different indexes to detect the same structure and how the jellyfish
+optimiser parameters affect the optimisation process.
+
+[example plots]
+
+```{r}
+set.seed(123456)
+sin1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
+colnames(sin1000) <- paste0("V", 1:6)
+
+sine_dcor2d_5050 <- sim_sine_6d_dcor2d |> 
+  filter(n_jellies == 50, max_tries == 50)
+best_sim <-  sine_dcor2d_5050|> 
+  group_by(sim) |> 
+  filter(index_val == max(index_val)) |> 
+  filter(row_number() == 1) 
+proj_dt <- map_dfr(best_sim |> pull(basis), 
+              ~as_tibble(as.matrix(sin1000) %*% .x), .id = "sim")
+
+p1 <- proj_dt |>
+  ggplot(aes(x = V1, y = V2)) + 
+  geom_point(size = 0.1) + 
+  xlim(-4, 4) + ylim(-4, 4) +
+  facet_wrap(~as.numeric(sim), ncol = 10) + 
+  theme_bw() +
+  theme(aspect.ratio = 1, axis.ticks = element_blank(),
+        axis.text = element_blank(), axis.title = element_blank(),
+        panel.grid = element_blank())
+```
+
+```{r}
+#| fig.height: 5 
+#| fig.width: 10
+#| fig.cap: sdfjsflk
+p1 
+```
+
+```{r}
+theo_best <- matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE)
+angle_dcor2d <- sine_dcor2d_5050 |>
+    rowwise() |>
+    mutate(angle = tourr::proj_dist(theo_best, basis))
+
+p2 <- angle_dcor2d |>
+  ggplot(aes(x = angle, y = index_val)) +
+  geom_point(size = 0.1) +
+  theme(aspect.ratio = 1)
+
+```
 

### 2024-04-19 optim.tex
outline of the section visualisation of jellyfish optimiser

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, titlerule=0mm, toprule=.15mm, colback=white, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, opacitybacktitle=0.6, leftrule=.75mm, toptitle=1mm, coltitle=black, breakable, arc=.35mm, left=2mm, rightrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, breakable, titlerule=0mm, opacityback=0, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, arc=.35mm, toprule=.15mm, opacitybacktitle=0.6, toptitle=1mm, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, bottomrule=.15mm, left=2mm, coltitle=black]
 
@@
 
-\hypertarget{sec-simulation}{%
-\section{Application {[}Di and Sherry{]}}\label{sec-simulation}}
+\section{Visualisation of jellyfish
+optimiser}\label{visualisation-of-jellyfish-optimiser}
+
+Information of the jellyfish optimiser is available in a tabular format
+and below is an example data collected from finding the sine-wave
+structure in 6D data using a distance correlation index
+(\texttt{docr2d\_2}):
+
+\begin{verbatim}
+Rows: 275,000
+Columns: 13
+$ idx_f     <chr> "dcor2d_2", "dcor2d_2", "dcor2d_2", "dcor2d_2", "dcor2d_2", ~
+$ d         <dbl> 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, ~
+$ n_jellies <dbl> 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, ~
+$ max_tries <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50, 50, 50, 50, 50, 50, ~
+$ sim       <int> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, ~
+$ seed      <int> 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, ~
+$ basis     <list> <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<~
+$ index_val <dbl> 0.0247212373, 0.0033938502, 0.0463398915, 0.0486230801, -0.0~
+$ info      <chr> "initiation", "initiation", "initiation", "initiation", "ini~
+$ method    <chr> "search_jellyfish", "search_jellyfish", "search_jellyfish", ~
+$ tries     <dbl> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, ~
+$ loop      <dbl> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 1~
+$ time      <drtn> 35.46031 secs, 35.46031 secs, 35.46031 secs, 35.46031 secs,~
+\end{verbatim}
+
+Information recorded can be categorised into the following categories:
+
+\begin{itemize}
+\tightlist
+\item
+  projection pursuit variables: the index function used
+  (\texttt{idx\_f}), the data dimension (\texttt{d})
+\item
+  jellyfish optimiser parameters: the number of jellies
+  (\texttt{n\_jellies}), the maximum number of tries
+  (\texttt{max\_tries})
+\item
+  simulation variables: the simulation number (\texttt{sim}), the seed
+  used (\texttt{seed})
+\item
+  optimisation variables: the projection basis in a matrix format
+  (\texttt{basis}), the index value (\texttt{index\_val}), a description
+  of the status - one of ``initiation'', ``current\_best'', and
+  ``jellyfish\_update'' (\texttt{info}), current iteration ID
+  (\texttt{tries}), current jelly ID (\texttt{loop}), and the time taken
+  to find the optimum (\texttt{time}).
+\end{itemize}
+
+The basis column records every basis \emph{visited} by the jellyfish
+optimiser prior to comparing with the current basis. In each iteration,
+if the index value of a visited basis is smaller than that of the
+current one, the jellyfish optimiser will retain the current basis for
+the next iteration, while still documenting the visited basis.
+
+Numerical information to compute:
+
+\begin{itemize}
+\tightlist
+\item
+  angular distance between the projection basis and the theoretical best
+  basis,
+\item
+  the proportion of simulation that found the optimal basis,
+\item
+  the proportion of jellies, within each simulation, that found the
+  optimal basis,
+\end{itemize}
+
+Visualisation to inspect:
+
+\begin{itemize}
+\tightlist
+\item
+  inspect the basis visited by each jellyfish in the reduced PCA space,
+\item
+  inspect the final 2D projections reached by each jellyfish,
+\item
+  plot the index value against the angular distance between the
+  projection basis and the theoretical best basis
+\end{itemize}
+
+Plotting the basis in the space and the projected data can help to
+understand 1) whether each simulation finds the same optimum or some
+simulations find local optima; and 2) whether the index function used
+can detect the structure in the data and the projection contains the
+structure of interest.
+
+The visualisations above can be faceted by the projection pursuit
+variables and jellyfish optimiser parameters to compare the performance
+of different indexes to detect the same structure and how the jellyfish
+optimiser parameters affect the optimisation process.
+
+{[}example plots{]}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-4-1.pdf}
+
+}
+
+\caption{sdfjsflk}
+
+\end{figure}%
+
+\section{Application {[}Di and Sherry{]}}\label{sec-simulation}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-2-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-7-1.pdf}
 
@@
 
-\end{figure}
+\end{figure}%
 
-\begin{figure}
+\begin{figure}[H]
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-3-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-8-1.pdf}
 

### 2024-05-28 optim.qmd
Update optim.qmd

@@
 From the literature, it is commonly understood that a large squint angle
-implies that the function is easy to optimize, because we do not need to
-be very close to the perfect view to see the structure. A small squint
-angle means that the derivative of the index function can still be very
-large values near the optimal point and will rapidly change as we get
-even closer to the optimum. As such we can observe the second order
-gradient, which is the rate of change of gradient, over the space we are
-searching. For some index functions, the second order gradient is not
-well-defined, we can approximate the second order gradient vector in
-similar fashion as the above section.
+implies that the objective function value is close to optimal even when we are not very close to the perfect view to see the structure.
+A small squint angle means that index function value improves substantially only when we are very close to the perfect view.
+As such, low squintability implies rapid improvement in the index value when near the perfect view.
+
+In this study, we propose two metrics to capture the notion of squintability.
+
+[We generate random points that is beyond 1.5 projection distance and interpolate. Then we fit a kernel or use nonlinear least squares.]
+
+First, parametric model.
+
+[Nicolas's pdf]
+
+Second, we consider the product of the largest absolute magnitude of rate of change of $f$ and the corresponding projection angle as a second measure of squintability.
+Since $f$ is decreasing, the rate of change of $f$ is negative and thus \[|\underset{x}{\min} f('x)|\] gives the absolute magnitude of the most negative rate of change.
+
+[Nicolas's pdf]
 

### 2024-05-29 code/07-add-smoothness-squint.R
full script to calculate smoothness and squintability

@@ -1,0 +1,91 @@
+library(ferrn)
+library(tourr)
+library(tidyverse)
+
+sine_8d_tbl <- function(index){
+  tibble::tibble(
+    n = 8, index = index, data = list(sine1000_8d),
+    best = list(matrix(c(rep(0, 12), 1, 0, 0, 1), nrow = 8, byrow = TRUE)))
+}
+
+smoothness_holes <- tibble::tibble(
+  n = c(6, 8, 10, 12),
+  index = rep("holes", 4),
+  data = list(pipe1000, pipe1000_8d, pipe1000_10d, pipe1000_12d),
+  best = list(matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE),
+              matrix(c(rep(0, 12), 1, 0, 0, 1), nrow = 8, byrow = TRUE),
+              matrix(c(rep(0, 16), 1, 0, 0, 1), nrow = 10, byrow = TRUE),
+              matrix(c(rep(0, 20), 1, 0, 0, 1), nrow = 12, byrow = TRUE))) |>
+  dplyr::mutate(smoothness = purrr::pmap(
+    list(data, best, n),
+    function(data, best, n){calc_smoothness("holes", data = data, n = n, best = best)}))
+
+holes_tidy <- smoothness_holes |> tidyr::unnest_wider(smoothness) |> unnest(measure)
+
+idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "stringy", "splines2d")
+smoothness_sine <- tibble::tibble(
+  n = 6, index = idx_names, data = list(sine1000),
+  best = list(matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE))) |>
+  dplyr::bind_rows(sine_8d_tbl("MIC")) |>
+  dplyr::bind_rows(sine_8d_tbl("TIC")) |>
+  rowwise() |>
+  dplyr::mutate(smoothness = list(calc_smoothness(index)))
+
+
+sine_tidy <- smoothness_sine |> tidyr::unnest_wider(smoothness) |> unnest(measure)
+
+smoothness <- bind_rows(sine_tidy, holes_tidy)
+save(smoothness, file = here::here("data", "smoothness.rda"))
+
+################################################################################
+################################################################################
+# squintability
+sq_holes_basis_df <- tibble::tibble(
+  n = c(6, 8, 10, 12),
+  index = rep("holes", 4),
[…]

### 2024-05-29 code/07-compile-sine-sim.R

@@ -1,169 +1,0 @@
-library(tidyverse)
-library(minerva)
-library(progress)
-library(furrr)
-library(cassowaryr)
-load(here::here("data/sim_sine_6d_dcor2d.rda"))
-load(here::here("data/sim_sine_6d.rda")) # loess
-load(here::here("data/sim_sine_68d_TICMIC.rda"))
-
-################################################################
-################################################################
-set.seed(123456)
-sine1000 <- spinebil::sinData(6, 1000) %>% scale()
-colnames(sine1000) <- paste0("V", 1:6)
-
-sim_res_all <- bind_rows(sim_sine_6d_dcor2d,
-                         sim_sine_6d |> mutate(idx_f = "loess2d"),
-                         sim_sine_68d_TICMIC |> filter(d == 6)) |>
-  filter(n_jellies == 50, max_tries == 50)
-
-
-sim_res_wide <- sim_res_all |>
-  select(idx_f:index_val, tries, loop) |>
-  mutate(row = row_number()) |>
-  pivot_wider(names_from = idx_f, values_from = index_val) |>
-  group_split(row)
-
-add_idx_val <- function(data, idx_f){
-  pb$tick()
-  idx_f_fn <- eval(sym(idx_f))
-  data |>
-    mutate(!!sym(idx_f) := ifelse(is.na(!!sym(idx_f)),
-                                  idx_f_fn()(sine1000 %*% basis[[1]]),
-                                  !!sym(idx_f)))
-}
-
-# Declaimer: in this script, I save the intermediate objects after each > 30 mins
-# run. But to save storage space, only the final data `sim_sine` is saved in the
-# data folder
-
-################################################################
-################################################################
-# Part 1
-# for the dcor2d, loess, TIC, MIC data, compute all the four indexes + stringy
-
-# add dcor: over an hour
-pb <- progress_bar$new(total = length(sim_res_wide))
-impute_dcor2d <- sim_res_wide |> map_dfr(~add_idx_val(.x, "dcor2d_2")) |> rename(loess2d = loess)
-
-# add loess in three batches: over an hour
-pb <- progress_bar$new(total = nrow(impute_dcor2d))
[…]

### 2024-05-29 code/08-calculate-optim-features.R

@@ -1,40 +1,0 @@
-library(ferrn)
-library(tourr)
-library(tidyverse)
-
-smoothness_holes <- tibble::tibble(
-  n = c(6, 8, 10, 12),
-  index = rep("holes", 4),
-  data = list(pipe1000, pipe1000_8d, pipe1000_10d, pipe1000_12d),
-  best = list(matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE),
-              matrix(c(rep(0, 12), 1, 0, 0, 1), nrow = 8, byrow = TRUE),
-              matrix(c(rep(0, 16), 1, 0, 0, 1), nrow = 10, byrow = TRUE),
-              matrix(c(rep(0, 20), 1, 0, 0, 1), nrow = 12, byrow = TRUE))) |>
-  dplyr::mutate(smoothness = purrr::pmap(
-    list(data, best, n),
-    function(data, best, n){calc_smoothness("holes", data = data, n = n, best = best)}))
-
-holes_tidy <- smoothness_holes |> tidyr::unnest_wider(smoothness) |> unnest(measure)
-
-idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "stringy", "splines2d")
-smoothness_sine <- tibble::tibble(
-  n = 6, index = idx_names, data = list(sine1000),
-  best = list(matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE))) |>
-  rowwise() |>
-  dplyr::mutate(smoothness = list(calc_smoothness(index)))
-
-
-sine_tidy <- smoothness_sine |> tidyr::unnest_wider(smoothness) |> unnest(measure)
-
-smoothness <- bind_rows(sine_tidy, holes_tidy)
-save(smoothness, file = here::here("data", "smoothness.rda"))
-
-
-################################################################################
-################################################################################
-load(here::here("data", "sim_summary.rda"))
-sim_summary |> rename(index = idx_f) |>
-  left_join(smoothness |> select(n, index, smoothness) |> rename(d = n))
-
-
-

### 2024-05-29 code/07-add-smoothness-squint.R
all the saved data

@@ -70,7 +70,7 @@ save(sq_basis_df, file = here::here("data/sq_basis_df.rda"))
 
 start_prms1 <- c(theta1 = 1, theta2 = 1, theta3 = 3, theta4 = 0)
 start_prms2 <- c(theta1 = 1, theta2 = 0.01, theta3 = 50, theta4 = 0)
-sq_sines <- sq_basis_df |>
+squintability <- sq_basis_df |>
   dplyr::mutate(start = ifelse(index == "stringy", list(start_prms2),
                                list(start_prms1))) |>
   dplyr::rowwise() |>
@@ -79,13 +79,15 @@ sq_sines <- sq_basis_df |>
   tidyr::unnest_wider(res) |>
   mutate(index = factor(index, levels = c("holes", "MIC", "TIC", "dcor2d_2",
                                           "loess2d", "splines2d", "stringy"))) |>
-  arrange(index)
+  arrange(index) |>
+  select(index, n, theta1: theta4)
+save(squintability, file = here::here("data", "squintability.rda"))
 
 ############################################################################
 ############################################################################
 load(here::here("data", "sim_summary.rda"))
 sim_df <- sim_summary |>
   left_join(smoothness |> select(n, index, smoothness) |> rename(d = n)) |>
-  left_join(sq_sines |> select(index, n, theta3) |> rename(d = n, squintability = theta3)) |>
+  left_join(squintability |> select(index, n, theta3) |> rename(d = n, squintability = theta3)) |>
   select(index, d, I_max_max, P_J_hat, n_jellies, max_tries, smoothness, squintability, time)
 save(sim_df, file = here::here("data", "sim_df.rda"))

### 2024-05-29 code/06-compile-all-sims.R
clean up scripts

@@ -58,7 +58,6 @@ sim_data_sine <- bind_rows(sim_sine_6d_dcor2d,
                       sim_sine_68d_TICMIC,
                       sim_sine_6d_spline |> mutate(idx_f = "spline"))
 
-
 tour_level_best_basis <- sim_data_sine |>
   group_by(idx_f, id) |>
   filter(index_val == max(index_val)) |>
@@ -99,6 +98,20 @@ setup_level_sine <- tour_level_sine |>
 
 
 sim_summary <- setup_level_pipe |> mutate(idx_f = "holes") |> bind_rows(setup_level_sine)
+
+sim_pipe_speed <- sim_pipe|> select(n_jellies:id, time) |>
+  group_by(d, n_jellies, max_tries) |>
+  summarize(time = mean(time), .groups = "drop")
+
+sim_sine_speed <- sim_data_sine |> select(idx_f: id, time) |>
+  group_by(idx_f, d, n_jellies, max_tries) |>
+  summarize(time = mean(time), .groups = "drop")
+
+sim_speed <- sim_pipe_speed |> mutate(idx_f = "holes") |> bind_rows(sim_sine_speed)
+sim_summary <- sim_summary |>
+  left_join(sim_speed) |>
+  rename(index = idx_f) |>
+  mutate(index = ifelse(index == "spline", "splines2d", index))
 save(sim_summary, file = here::here("data/sim_summary.rda"))
 
 

### 2024-05-29 optim.qmd
print the smoothness and squintability values in the pdf

@@
 library(patchwork)
+library(ggh4x)
 ```
@@
 From the literature, it is commonly understood that a large squint angle
-implies that the objective function value is close to optimal even when we are not very close to the perfect view to see the structure.
-A small squint angle means that index function value improves substantially only when we are very close to the perfect view.
-As such, low squintability implies rapid improvement in the index value when near the perfect view.
+implies that the objective function value is close to optimal even when
+we are not very close to the perfect view to see the structure. A small
+squint angle means that index function value improves substantially only
+when we are very close to the perfect view. As such, low squintability
+implies rapid improvement in the index value when near the perfect view.
 
-In this study, we propose two metrics to capture the notion of squintability.
+In this study, we propose two metrics to capture the notion of
+squintability.
 
-[We generate random points that is beyond 1.5 projection distance and interpolate. Then we fit a kernel or use nonlinear least squares.]
+\[We generate random points that is beyond 1.5 projection distance and
+interpolate. Then we fit a kernel or use nonlinear least squares.\]
 
@@
 
-[Nicolas's pdf]
+\[Nicolas's pdf\]
 
-Second, we consider the product of the largest absolute magnitude of rate of change of $f$ and the corresponding projection angle as a second measure of squintability.
-Since $f$ is decreasing, the rate of change of $f$ is negative and thus \[|\underset{x}{\min} f('x)|\] gives the absolute magnitude of the most negative rate of change.
+Second, we consider the product of the largest absolute magnitude of
+rate of change of $f$ and the corresponding projection angle as a second
+measure of squintability. Since $f$ is decreasing, the rate of change of
+$f$ is negative and thus $|\underset{x}{\min} f('x)|$ gives the absolute
+magnitude of the most negative rate of change.
 
-[Nicolas's pdf]
+\[Nicolas's pdf\]
 
@@
 
--   inspect the basis visited by each jellyfish in the reduced PCA space,
+-   inspect the basis visited by each jellyfish in the reduced PCA
+    space,
 -   inspect the final 2D projections reached by each jellyfish,
@@
 
-[example plots]
+\[example plots\]
 
@@
 
+construct a relationship between jellyfish success and jellyfish
+parameters and the optimisation properties defined in @sec-theory.
+
+This can imform the choice of jellyfish parameters for a given
+optimisation problem.
+
+In addition to the pipe-finding problem, we also consider the problem of
+finding the sine wave structure in the 6D space and six indexes
+(`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`) are
+considered. Combined with the jellyfish optimisers (`n_jellies` =
+20/50/100, `max_tries` = 50/100), we obtain additional 27 setups for
+this investigation.
+
+```{r}
+load(here::here("data/smoothness.rda"))
+smoothness |> select(index, n, variance:nugget) |> knitr::kable(digits = 3)
+load(here::here("data/squintability.rda"))
+squintability |> knitr::kable(digits = 3)
+```
+
+```{r}
+#| fig.width = 8,
+#| fig.height = 5,
+#| fig.cap = "sdkflskdfjl"
+load(here::here("data/sq_basis_dist_idx.rda"))
+sq_basis_dist_idx |>
+  ggplot(aes(x = dist, y = idx, group = index)) +
+  geom_line() +
+  facet_nested_wrap(~index + n, nrow = 3, labeller = label_both, scales = "free_y") + 
+  xlab("projection distance") + ylab("index value")
+
+```
+
 # Conclusion \[Di and Sherry\] {#sec-conclusion}

### 2024-05-29 optim.tex
print the smoothness and squintability values in the pdf

@@
 
+\begin{verbatim}
+Warning: package 'ggh4x' was built under R version 4.3.2
+\end{verbatim}
+
 \section{Introduction {[}Nicolas and
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, breakable, titlerule=0mm, opacityback=0, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, arc=.35mm, toprule=.15mm, opacitybacktitle=0.6, toptitle=1mm, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, bottomrule=.15mm, left=2mm, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, toprule=.15mm, colback=white, colframe=quarto-callout-note-color-frame, arc=.35mm, breakable, leftrule=.75mm, opacitybacktitle=0.6, bottomtitle=1mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, bottomrule=.15mm, left=2mm, opacityback=0]
 
@@
 From the literature, it is commonly understood that a large squint angle
-implies that the function is easy to optimize, because we do not need to
-be very close to the perfect view to see the structure. A small squint
-angle means that the derivative of the index function can still be very
-large values near the optimal point and will rapidly change as we get
-even closer to the optimum. As such we can observe the second order
-gradient, which is the rate of change of gradient, over the space we are
-searching. For some index functions, the second order gradient is not
-well-defined, we can approximate the second order gradient vector in
-similar fashion as the above section.
+implies that the objective function value is close to optimal even when
+we are not very close to the perfect view to see the structure. A small
+squint angle means that index function value improves substantially only
+when we are very close to the perfect view. As such, low squintability
+implies rapid improvement in the index value when near the perfect view.
+
+In this study, we propose two metrics to capture the notion of
+squintability.
+
+{[}We generate random points that is beyond 1.5 projection distance and
+interpolate. Then we fit a kernel or use nonlinear least squares.{]}
+
+First, parametric model.
+
+{[}Nicolas's pdf{]}
+
+Second, we consider the product of the largest absolute magnitude of
+rate of change of \(f\) and the corresponding projection angle as a
+second measure of squintability. Since \(f\) is decreasing, the rate of
+change of \(f\) is negative and thus \(|\underset{x}{\min} f('x)|\)
+gives the absolute magnitude of the most negative rate of change.
+
+{[}Nicolas's pdf{]}
 
@@
 
+construct a relationship between jellyfish success and jellyfish
+parameters and the optimisation properties defined in
+Section~\ref{sec-theory}.
+
+This can imform the choice of jellyfish parameters for a given
+optimisation problem.
+
+In addition to the pipe-finding problem, we also consider the problem of
+finding the sine wave structure in the 6D space and six indexes
+(\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
+\texttt{spline}, and \texttt{stringy}) are considered. Combined with the
+jellyfish optimisers (\texttt{n\_jellies} = 20/50/100,
+\texttt{max\_tries} = 50/100), we obtain additional 27 setups for this
+investigation.
+
+\begin{longtable}[]{@{}lrrrrr@{}}
+\toprule\noalign{}
+index & n & variance & range & smoothness & nugget \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 \\
+loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 \\
+MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 \\
+TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 \\
+stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 \\
+splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 \\
+MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 \\
+TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 \\
+holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 \\
+holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 \\
+holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 \\
+holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 \\
+\end{longtable}
+
+\begin{longtable}[]{@{}lrrrrr@{}}
+\toprule\noalign{}
+index & n & theta1 & theta2 & theta3 & theta4 \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+holes & 6 & 1.001 & 0.860 & 3.368 & 0.823 \\
+holes & 8 & 1.001 & 0.869 & 3.264 & 0.811 \\
+holes & 10 & 1.000 & 0.885 & 3.151 & 0.806 \\
+holes & 12 & 1.000 & 0.878 & 3.345 & 0.806 \\
+MIC & 6 & 0.894 & 0.571 & 1.623 & -0.024 \\
+MIC & 8 & 0.932 & 0.328 & 1.314 & -0.030 \\
+TIC & 6 & 0.951 & 0.536 & 1.719 & -0.025 \\
+TIC & 8 & 0.945 & 0.564 & 1.723 & -0.027 \\
+dcor2d\_2 & 6 & 0.954 & 1.039 & 2.742 & -0.019 \\
+loess2d & 6 & 1.016 & 1.039 & 2.648 & 0.080 \\
+splines2d & 6 & 1.014 & 1.051 & 2.730 & -0.009 \\
+stringy & 6 & 1.011 & 0.011 & 254.734 & 0.727 \\
+\end{longtable}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-10-1.pdf}
+
+}
+
+\caption{sdkflskdfjl}
+
+\end{figure}%
+
 \section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}

### 2024-05-29 code/07-add-smoothness-squint.R

@@ -91,3 +91,20 @@ sim_df <- sim_summary |>
   left_join(squintability |> select(index, n, theta3) |> rename(d = n, squintability = theta3)) |>
   select(index, d, I_max_max, P_J_hat, n_jellies, max_tries, smoothness, squintability, time)
 save(sim_df, file = here::here("data", "sim_df.rda"))
+
+
+
+
+############################################################################
+############################################################################
+sq_basis_dist_idx <- sq_basis_df |>
+  dplyr::select(basis_df, n) |>
+  unnest(basis_df) |>
+  pivot_longer(holes:stringy, names_to = "index", values_to = "idx",
+               values_drop_na = TRUE) |>
+  mutate(dist = ceiling(dist / 0.005) * 0.005,
+         index = factor(index, levels = c("holes", "MIC", "TIC", "dcor2d_2",
+                                          "loess2d", "splines2d", "stringy"))) |>
+  group_by(n, index, dist) |>
+  summarise(idx = mean(idx), .groups = "drop")
+save(sq_basis_dist_idx, file = here::here("data", "sq_basis_dist_idx.rda"))

### 2024-05-30 optim.qmd
replace sim_sine_6d_dcor2d with smaller datasets

@@
 ```{r}
-load(here::here("data/sim_sine_6d_dcor2d.rda"))
+load(here::here("data/sim_example.rda"))
 ```
@@
 ```{r}
-dplyr::glimpse(sim_sine_6d_dcor2d |> select(-id)) 
+dplyr::glimpse(sim_example |> select(-id)) 
 ```
@@
 ```{r}
+load(here::here("data/sine_dcor2d_best.rda"))
 set.seed(123456)
@@
 
-sine_dcor2d_5050 <- sim_sine_6d_dcor2d |> 
-  filter(n_jellies == 50, max_tries == 50)
-best_sim <-  sine_dcor2d_5050|> 
-  group_by(sim) |> 
-  filter(index_val == max(index_val)) |> 
-  filter(row_number() == 1) 
-proj_dt <- map_dfr(best_sim |> pull(basis), 
+proj_dt <- map_dfr(sine_dcor2d_best |> pull(basis), 
               ~as_tibble(as.matrix(sin1000) %*% .x), .id = "sim")
@@
 
-```{r}
+```{r eval = FALSE}
 theo_best <- matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE)
@@
 load(here::here("data/smoothness.rda"))
-smoothness |> select(index, n, variance:nugget) |> knitr::kable(digits = 3)
+smoothness |> knitr::kable(digits = 3)
 load(here::here("data/squintability.rda"))

### 2024-05-30 optim.tex
replace sim_sine_6d_dcor2d with smaller datasets

@@
 
-\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, toprule=.15mm, colback=white, colframe=quarto-callout-note-color-frame, arc=.35mm, breakable, leftrule=.75mm, opacitybacktitle=0.6, bottomtitle=1mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, bottomrule=.15mm, left=2mm, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, opacityback=0, breakable, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, arc=.35mm, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, coltitle=black, toptitle=1mm, toprule=.15mm, leftrule=.75mm, left=2mm, colback=white, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, rightrule=.15mm]
 
@@
 \begin{verbatim}
-Rows: 275,000
+Rows: 10
 Columns: 13
 $ idx_f     <chr> "dcor2d_2", "dcor2d_2", "dcor2d_2", "dcor2d_2", "dcor2d_2", ~
-$ d         <dbl> 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, ~
-$ n_jellies <dbl> 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, ~
-$ max_tries <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50, 50, 50, 50, 50, 50, ~
-$ sim       <int> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, ~
-$ seed      <int> 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, ~
-$ basis     <list> <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<~
+$ d         <dbl> 6, 6, 6, 6, 6, 6, 6, 6, 6, 6
+$ n_jellies <dbl> 20, 20, 20, 20, 20, 20, 20, 20, 20, 20
+$ max_tries <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50
+$ sim       <int> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1
+$ seed      <int> 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462
+$ basis     <list> <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<m~
 $ index_val <dbl> 0.0247212373, 0.0033938502, 0.0463398915, 0.0486230801, -0.0~
@@
 $ method    <chr> "search_jellyfish", "search_jellyfish", "search_jellyfish", ~
-$ tries     <dbl> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, ~
-$ loop      <dbl> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 1~
-$ time      <drtn> 35.46031 secs, 35.46031 secs, 35.46031 secs, 35.46031 secs,~
+$ tries     <dbl> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1
+$ loop      <dbl> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10
+$ time      <drtn> 35.46031 secs, 35.46031 secs, 35.46031 secs, 35.46031 secs, ~
 \end{verbatim}
@@
 \endlastfoot
-dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 \\
-loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 \\
-MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 \\
-TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 \\
-stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 \\
-splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 \\
-MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 \\
-TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 \\
 holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 \\
@@
 holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 \\
+MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 \\
+MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 \\
+TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 \\
+TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 \\
+dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 \\
+loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 \\
+splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 \\
+stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 \\
 \end{longtable}

### 2024-05-30 code/05-simulation-sine-6d-dcor2d.R

@@ -68,9 +68,20 @@ t2 <- Sys.time()
 t2 - t1
 
 
-# sim_sine_6d_dcor2d <- sim_res |> select(-alpha)
-# save(sim_sine_6d_dcor2d , file = "data/sim_sine_6d_dcor2d.rda")
-#
+sim_sine_6d_dcor2d <- sim_res |> select(-alpha)
+save(sim_sine_6d_dcor2d , file = "data/sim_sine_6d_dcor2d.rda")
+sim_example <- sim_sine_6d_dcor2d |> head(10)
+save(sim_example , file = "data/sim_example.rda")
+sine_dcor2d_best <- sim_sine_6d_dcor2d |>
+  filter(n_jellies == 50, max_tries == 50) |>
+  group_by(sim) |>
+  filter(index_val == max(index_val)) |>
+  filter(row_number() == 1) |>
+  ungroup()
+save(sine_dcor2d_best , file = "data/sine_dcor2d_best.rda")
+
+
+
 #
 # set.seed(123456)
 # sine1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()

### 2024-05-31 optim.qmd
write up on exmaple #2

@@
 
-## Smoothness
+## Smoothness {#sec-smoothness}
 
@@
 
-## Squintability
+## Squintability {#sec-squintability}
 
@@
 
-## On skewness and kurtosis index
+\newpage
 
@@
 
-construct a relationship between jellyfish success and jellyfish
-parameters and the optimisation properties defined in @sec-theory.
-
-This can imform the choice of jellyfish parameters for a given
+[TODO: overall summary]construct a relationship between jellyfish success and jellyfish parameters and the optimisation properties defined in @sec-theory.
+This can inform the choice of jellyfish parameters for a given
 optimisation problem.
 
-In addition to the pipe-finding problem, we also consider the problem of
-finding the sine wave structure in the 6D space and six indexes
-(`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`) are
-considered. Combined with the jellyfish optimisers (`n_jellies` =
-20/50/100, `max_tries` = 50/100), we obtain additional 27 setups for
-this investigation.
+In addition to the pipe-finding problem, the sine-wave finding problem is also investigated in 6D and 8D spaces.
+Six indices are considered: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`, which gives the 12 problems printed as rows in @tbl-smoothness-squintability.
+Combining these problems with different jellyfish optimiser parameters (`n_jellies` and `max_tries`), a total of 51 observations is produced, comprising of 24 observations for the pipe-finding and 27 observations for the sine-wave finding.
+
+For each observation, fifty (50) runs of the jellyfish optimisation is performed.
+In each run, summary statistics are calculated to find the best index across all the jellyfishes (`I_max`) and the proportion of jellyfishes found (`P_J`).
+At the observation level, the best index value across all the 50 runs  (`I_max_max`) and the proportion of trials found (`P_J_hat`) is calculated.
+
+Smoothness and squintability measures are computed for each problem, following the procedures outlined in @sec-smoothness and @sec-squintability.
+To calculate smoothness, 300 random bases are used to estimate Gaussian process parameters (variance, range, smoothness, and nugget).
+For squintability, 50 random bases are sampled and interpolated to the optimal basis using a step size of 0.005.
+The index value and projection distance are calculated for all the interpolated bases.
+A binning procedure is then applied to average the index value at each bin width of 0.005.
+Finally, the scaled logistic function with four parameters (theta1 - theta4) is estimated by non-linear least square to obtain the squintability measure.
+
+@tbl-smoothness-squintability presents all the parameters calculated for smoothness and squintability, with the column "smooth" for smoothness and "theta3" for squintability.
+@fig-idx-proj-dist shows the relationship between the index value against the projection distance for each problem following the binning procedure.
+@tbl-all-sim displays the first five rows after combining all calculations.
 
 ```{r}
+#| label: tbl-smoothness-squintability
+#| tbl-cap: Parameters estimated from the Gaussian process (including variance, range, smoothness, and nugget) and scaled logistic function (theta1 - theta4) for both the pipe-finding and sine-wave finding problems.
+The "smooth" and "theta3" column represent the smoothness and squintability measures.
 load(here::here("data/smoothness.rda"))
-smoothness |> knitr::kable(digits = 3)
 load(here::here("data/squintability.rda"))
-squintability |> knitr::kable(digits = 3)
+library(kableExtra)
+tibble(shape = c(rep("pipe", 4), rep("sine", 8)), smoothness) |>
+  left_join(squintability) |> 
+  rename(d = n, smooth = smoothness) |> 
+  knitr::kable(digits = 3) |> 
+  column_spec(1, border_left = TRUE) |> 
+  column_spec(c(3, 7, 11), border_right = TRUE) |> 
+  kable_styling(latex_options = "hold_position")
 ```
 
-```{r}
-#| fig.width = 8,
+```{r fig-idx-proj-dist}
+#| fig.width = 9,
 #| fig.height = 5,
-#| fig.cap = "sdkflskdfjl"
+#| fig.cap = "Index values against projection distance for the 12 pipe/sine-wave finding problem after the binning procedure during the estimation of the squintability measure.
+The index value is averaged at each bin width of 0.005 and the TIC index is scaled to 0-1 for comparison.
+When finding the sine-wave structure using the MIC and TIC index, a convex curved is observed, as opposed to the pipe-finding problem or the sine-wave finding problem with the dcor2d\\_2, loess2d, and splines2d index.
+When finding the sine-wave with the stringy index, the index shows an instantaneous jump to the optimum when close to the best basis."
 load(here::here("data/sq_basis_dist_idx.rda"))
@@
 
+```
+
+```{r}
+#| label: tbl-all-sim
+#| tbl-cap: The first five rows of the datasets after combining all the calculation.
+The smoothness and squintability variable are uniquely characterised by the index function used and the data dimension, and thus do not vary across n\\_jellies and max\\_tries.
+Variables I\\_max\\_max, P\\_J\\_hat, and time are calculated at each 51 observations.
+load(here::here("data/sim_df.rda"))
+sim_df |> 
+  dplyr::select(index, d, smoothness, squintability, n_jellies, max_tries, 
+                I_max_max, P_J_hat, time) |> 
+  arrange(index, d, n_jellies, max_tries) |> 
+  head(5) |> 
+  kable(digits = 3)
 ```

### 2024-05-31 optim.tex
write up on exmaple #2

@@
 
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 \usepackage{algorithm}
@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacityback=0, breakable, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, arc=.35mm, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, coltitle=black, toptitle=1mm, toprule=.15mm, leftrule=.75mm, left=2mm, colback=white, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, rightrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, arc=.35mm, leftrule=.75mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, toptitle=1mm, opacityback=0, bottomrule=.15mm, titlerule=0mm, coltitle=black, rightrule=.15mm, breakable, toprule=.15mm, left=2mm]
 
@@
 
-\subsection{Smoothness}\label{smoothness}
+\subsection{Smoothness}\label{sec-smoothness}
 
@@
 
-\subsection{Squintability}\label{squintability}
+\subsection{Squintability}\label{sec-squintability}
 
@@
 
-\subsection{On skewness and kurtosis
-index}\label{on-skewness-and-kurtosis-index}
+\newpage
 
@@
 
-construct a relationship between jellyfish success and jellyfish
-parameters and the optimisation properties defined in
-Section~\ref{sec-theory}.
+{[}TODO: overall summary{]}construct a relationship between jellyfish
+success and jellyfish parameters and the optimisation properties defined
+in Section~\ref{sec-theory}. This can inform the choice of jellyfish
+parameters for a given optimisation problem.
 
-This can imform the choice of jellyfish parameters for a given
-optimisation problem.
+In addition to the pipe-finding problem, the sine-wave finding problem
+is also investigated in 6D and 8D spaces. Six indices are considered:
+\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
+\texttt{spline}, and \texttt{stringy}, which gives the 12 problems
+printed as rows in Table~\ref{tbl-smoothness-squintability}. Combining
+these problems with different jellyfish optimiser parameters
+(\texttt{n\_jellies} and \texttt{max\_tries}), a total of 51
+observations is produced, comprising of 24 observations for the
+pipe-finding and 27 observations for the sine-wave finding.
 
-In addition to the pipe-finding problem, we also consider the problem of
-finding the sine wave structure in the 6D space and six indexes
-(\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
-\texttt{spline}, and \texttt{stringy}) are considered. Combined with the
-jellyfish optimisers (\texttt{n\_jellies} = 20/50/100,
-\texttt{max\_tries} = 50/100), we obtain additional 27 setups for this
-investigation.
+For each observation, fifty (50) runs of the jellyfish optimisation is
+performed. In each run, summary statistics are calculated to find the
+best index across all the jellyfishes (\texttt{I\_max}) and the
+proportion of jellyfishes found (\texttt{P\_J}). At the observation
+level, the best index value across all the 50 runs
+(\texttt{I\_max\_max}) and the proportion of trials found
+(\texttt{P\_J\_hat}) is calculated.
 
-\begin{longtable}[]{@{}lrrrrr@{}}
+Smoothness and squintability measures are computed for each problem,
+following the procedures outlined in Section~\ref{sec-smoothness} and
+Section~\ref{sec-squintability}. To calculate smoothness, 300 random
+bases are used to estimate Gaussian process parameters (variance, range,
+smoothness, and nugget). For squintability, 50 random bases are sampled
+and interpolated to the optimal basis using a step size of 0.005. The
+index value and projection distance are calculated for all the
+interpolated bases. A binning procedure is then applied to average the
+index value at each bin width of 0.005. Finally, the scaled logistic
+function with four parameters (theta1 - theta4) is estimated by
+non-linear least square to obtain the squintability measure.
+
+Table~\ref{tbl-smoothness-squintability} presents all the parameters
+calculated for smoothness and squintability, with the column ``smooth''
+for smoothness and ``theta3'' for squintability.
+Figure~\ref{fig-idx-proj-dist} shows the relationship between the index
+value against the projection distance for each problem following the
+binning procedure. Table~\ref{tbl-all-sim} displays the first five rows
+after combining all calculations.
+
+\begin{longtable}[t]{|>{}ll>{}r|rrr>{}r|rrr>{}r|}
+
+\caption{\label{tbl-smoothness-squintability}Parameters estimated from
+the Gaussian process (including variance, range, smoothness, and nugget)
+and scaled logistic function (theta1 - theta4) for both the pipe-finding
+and sine-wave finding problems. The ``smooth'' and ``theta3'' column
+represent the smoothness and squintability measures.}
+
+\tabularnewline
+
+\toprule
+shape & index & d & variance & range & smooth & nugget & theta1 & theta2 & theta3 & theta4\\
+\midrule
+pipe & holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 & 1.001 & 0.860 & 3.368 & 0.823\\
+pipe & holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 & 1.001 & 0.869 & 3.264 & 0.811\\
+pipe & holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 & 1.000 & 0.885 & 3.151 & 0.806\\
+pipe & holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 & 1.000 & 0.878 & 3.345 & 0.806\\
+sine & MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 & 0.894 & 0.571 & 1.623 & -0.024\\
+\addlinespace
+sine & MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 & 0.932 & 0.328 & 1.314 & -0.030\\
+sine & TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 & 0.951 & 0.536 & 1.719 & -0.025\\
+sine & TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 & 0.945 & 0.564 & 1.723 & -0.027\\
+sine & dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019\\
+sine & loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 & 1.016 & 1.039 & 2.648 & 0.080\\
+\addlinespace
+sine & splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 & 1.014 & 1.051 & 2.730 & -0.009\\
+sine & stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 & 1.011 & 0.011 & 254.734 & 0.727\\
+\bottomrule
+
+\end{longtable}
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{optim_files/figure-pdf/fig-idx-proj-dist-1.pdf}
+
+}
+
+\caption{\label{fig-idx-proj-dist}Index values against projection
+distance for the 12 pipe/sine-wave finding problem after the binning
+procedure during the estimation of the squintability measure. The index
+value is averaged at each bin width of 0.005 and the TIC index is scaled
+to 0-1 for comparison. When finding the sine-wave structure using the
+MIC and TIC index, a convex curved is observed, as opposed to the
+pipe-finding problem or the sine-wave finding problem with the
+dcor2d\_2, loess2d, and splines2d index. When finding the sine-wave with
+the stringy index, the index shows an instantaneous jump to the optimum
+when close to the best basis.}
+
+\end{figure}%
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0714}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0357}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.1310}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.1667}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.1190}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.1190}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.1190}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0952}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.1429}}@{}}
+
+\caption{\label{tbl-all-sim}The first five rows of the datasets after
+combining all the calculation. The smoothness and squintability variable
+are uniquely characterised by the index function used and the data
+dimension, and thus do not vary across n\_jellies and max\_tries.
+Variables I\_max\_max, P\_J\_hat, and time are calculated at each 51
+observations.}
+
+\tabularnewline
+
 \toprule\noalign{}
-index & n & variance & range & smoothness & nugget \\
+\begin{minipage}[b]{\linewidth}\raggedright
+index
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+d
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+smoothness
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+squintability
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+n\_jellies
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+max\_tries
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+I\_max\_max
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+P\_J\_hat
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+time
+\end{minipage} \\
 \midrule\noalign{}
@@
 \endlastfoot
-holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 \\
-holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 \\
-holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 \\
-holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 \\
-MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 \\
-MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 \\
-TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 \\
-TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 \\
-dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 \\
-loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 \\
-splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 \\
-stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 \\
+MIC & 6 & 2.394 & 1.623 & 20 & 50 & 0.999 & 1.00 & 2.479 secs \\
+MIC & 6 & 2.394 & 1.623 & 20 & 100 & 0.999 & 0.96 & 8.950 secs \\
+MIC & 6 & 2.394 & 1.623 & 50 & 50 & 1.000 & 1.00 & 5.651 secs \\
+MIC & 6 & 2.394 & 1.623 & 50 & 100 & 1.000 & 0.64 & 13.223 secs \\
+MIC & 6 & 2.394 & 1.623 & 100 & 50 & 1.000 & 0.76 & 19.453 secs \\
+
 \end{longtable}
-
-\begin{longtable}[]{@{}lrrrrr@{}}
-\toprule\noalign{}
-index & n & theta1 & theta2 & theta3 & theta4 \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-holes & 6 & 1.001 & 0.860 & 3.368 & 0.823 \\
-holes & 8 & 1.001 & 0.869 & 3.264 & 0.811 \\
-holes & 10 & 1.000 & 0.885 & 3.151 & 0.806 \\
-holes & 12 & 1.000 & 0.878 & 3.345 & 0.806 \\
-MIC & 6 & 0.894 & 0.571 & 1.623 & -0.024 \\
-MIC & 8 & 0.932 & 0.328 & 1.314 & -0.030 \\
-TIC & 6 & 0.951 & 0.536 & 1.719 & -0.025 \\
-TIC & 8 & 0.945 & 0.564 & 1.723 & -0.027 \\
-dcor2d\_2 & 6 & 0.954 & 1.039 & 2.742 & -0.019 \\
-loess2d & 6 & 1.016 & 1.039 & 2.648 & 0.080 \\
-splines2d & 6 & 1.014 & 1.051 & 2.730 & -0.009 \\
-stringy & 6 & 1.011 & 0.011 & 254.734 & 0.727 \\
-\end{longtable}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-10-1.pdf}
-
-}
-
-\caption{sdkflskdfjl}
-
-\end{figure}%
 

### 2024-05-31 code/07-add-smoothness-squint.R

@@ -34,7 +34,11 @@ smoothness_sine <- tibble::tibble(
 
 sine_tidy <- smoothness_sine |> tidyr::unnest_wider(smoothness) |> unnest(measure)
 
-smoothness <- bind_rows(sine_tidy, holes_tidy)
+smoothness <- bind_rows(sine_tidy, holes_tidy) |>
+  mutate(index = factor(index, levels = c("holes", "MIC", "TIC", "dcor2d_2",
+                                          "loess2d", "splines2d", "stringy"))) |>
+  arrange(index) |>
+  select(index, n, variance:nugget)
 save(smoothness, file = here::here("data", "smoothness.rda"))
 
 ################################################################################

### 2024-05-31 optim.qmd
take back the print of the final dataset

@@
 @fig-idx-proj-dist shows the relationship between the index value against the projection distance for each problem following the binning procedure.
-@tbl-all-sim displays the first five rows after combining all calculations.
+
+<!-- @tbl-all-sim displays the first five rows after combining all calculations.  -->
 
@@
 
-```{r}
+```{r eval = FALSE}
 #| label: tbl-all-sim

### 2024-05-31 optim.tex
take back the print of the final dataset

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, arc=.35mm, leftrule=.75mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, toptitle=1mm, opacityback=0, bottomrule=.15mm, titlerule=0mm, coltitle=black, rightrule=.15mm, breakable, toprule=.15mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, left=2mm, coltitle=black, opacitybacktitle=0.6, colback=white, breakable, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, toptitle=1mm, bottomtitle=1mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, rightrule=.15mm, opacityback=0]
 
@@
 value against the projection distance for each problem following the
-binning procedure. Table~\ref{tbl-all-sim} displays the first five rows
-after combining all calculations.
+binning procedure.
 
@@
 
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0714}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0357}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.1310}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.1667}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.1190}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.1190}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.1190}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0952}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.1429}}@{}}
+\begin{table}
 
@@
 
-\tabularnewline
+\centering{
 
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-index
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-d
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-smoothness
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-squintability
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-n\_jellies
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-max\_tries
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-I\_max\_max
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-P\_J\_hat
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-time
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-MIC & 6 & 2.394 & 1.623 & 20 & 50 & 0.999 & 1.00 & 2.479 secs \\
-MIC & 6 & 2.394 & 1.623 & 20 & 100 & 0.999 & 0.96 & 8.950 secs \\
-MIC & 6 & 2.394 & 1.623 & 50 & 50 & 1.000 & 1.00 & 5.651 secs \\
-MIC & 6 & 2.394 & 1.623 & 50 & 100 & 1.000 & 0.64 & 13.223 secs \\
-MIC & 6 & 2.394 & 1.623 & 100 & 50 & 1.000 & 0.76 & 19.453 secs \\
+}
 
-\end{longtable}
+\end{table}%
 

### 2024-06-03 optim.qmd
finalise the calculation for example 2

@@
 library(ggh4x)
+library(broom)
 ```
@@
 
+[TODO: discuss what the values of smooth and theta3 mean for the optimisation]
+
 @tbl-smoothness-squintability presents all the parameters calculated for smoothness and squintability, with the column "smooth" for smoothness and "theta3" for squintability.
@@
 #| label: tbl-smoothness-squintability
-#| tbl-cap: Parameters estimated from the Gaussian process (including variance, range, smoothness, and nugget) and scaled logistic function (theta1 - theta4) for both the pipe-finding and sine-wave finding problems.
-The "smooth" and "theta3" column represent the smoothness and squintability measures.
+#| tbl-cap: Parameters estimated from the Gaussian process (including variance, range, smoothness, and nugget) and scaled logistic function (theta1 - theta4) for the pipe-finding and sine-wave finding problems.
+The squint column is calculated as theta1 * theta2 * theta3 / 4, as described in @sec-squintability.
+The "smooth" and "squint" column represent the smoothness and squintability measures.
 load(here::here("data/smoothness.rda"))
@@
   column_spec(1, border_left = TRUE) |> 
-  column_spec(c(3, 7, 11), border_right = TRUE) |> 
+  column_spec(c(3, 7, 12), border_right = TRUE) |> 
   kable_styling(latex_options = "hold_position")
@@
 
-```{r eval = FALSE}
+```{r}
 #| label: tbl-all-sim
-#| tbl-cap: The first five rows of the datasets after combining all the calculation.
+#| tbl-cap: The first few rows of the datasets processed for modelling.
 The smoothness and squintability variable are uniquely characterised by the index function used and the data dimension, and thus do not vary across n\\_jellies and max\\_tries.
-Variables I\\_max\\_max, P\\_J\\_hat, and time are calculated at each 51 observations.
+The variable P\\_J\\_hat, and time are calculated at each observation.
 load(here::here("data/sim_df.rda"))
@@
   dplyr::select(index, d, smoothness, squintability, n_jellies, max_tries, 
-                I_max_max, P_J_hat, time) |> 
+                -I_max_max, P_J_hat, time) |> 
   arrange(index, d, n_jellies, max_tries) |> 
-  head(5) |> 
+  head(7) |> 
   kable(digits = 3)
+```
+
+```{r}
+sim_df2 <- sim_df |> 
+  mutate(n_jellies = n_jellies/10,
+         max_tries = max_tries/10,
+         long_time = ifelse(time > 30, 1, 0),
+         P_J_hat = ifelse(index == "stringy", 0, P_J_hat))
+
+mod1 <- glm(P_J_hat ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
+broom::tidy(mod1)
 ```

### 2024-06-03 optim.tex
finalise the calculation for example 2

@@
 
-\begin{tcolorbox}[enhanced jigsaw, left=2mm, coltitle=black, opacitybacktitle=0.6, colback=white, breakable, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, toptitle=1mm, bottomtitle=1mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, rightrule=.15mm, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, leftrule=.75mm, arc=.35mm, left=2mm, opacityback=0, breakable, toptitle=1mm, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, colback=white, titlerule=0mm, bottomrule=.15mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame]
 
@@
 
+{[}TODO: discuss what the values of smooth and theta3 mean for the
+optimisation{]}
+
 Table~\ref{tbl-smoothness-squintability} presents all the parameters
@@
 
-\begin{longtable}[t]{|>{}ll>{}r|rrr>{}r|rrr>{}r|}
+\begin{longtable}[t]{|>{}ll>{}r|rrr>{}r|rrrr>{}r|}
 
@@
 the Gaussian process (including variance, range, smoothness, and nugget)
-and scaled logistic function (theta1 - theta4) for both the pipe-finding
-and sine-wave finding problems. The ``smooth'' and ``theta3'' column
-represent the smoothness and squintability measures.}
+and scaled logistic function (theta1 - theta4) for the pipe-finding and
+sine-wave finding problems. The squint column is calculated as theta1 *
+theta2 * theta3 / 4, as described in Section~\ref{sec-squintability}.
+The ``smooth'' and ``squint'' column represent the smoothness and
+squintability measures.}
 
@@
 \toprule
-shape & index & d & variance & range & smooth & nugget & theta1 & theta2 & theta3 & theta4\\
+shape & index & d & variance & range & smooth & nugget & theta1 & theta2 & theta3 & theta4 & squint\\
 \midrule
-pipe & holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 & 1.001 & 0.860 & 3.368 & 0.823\\
-pipe & holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 & 1.001 & 0.869 & 3.264 & 0.811\\
-pipe & holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 & 1.000 & 0.885 & 3.151 & 0.806\\
-pipe & holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 & 1.000 & 0.878 & 3.345 & 0.806\\
-sine & MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 & 0.894 & 0.571 & 1.623 & -0.024\\
+pipe & holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 & 1.001 & 0.860 & 3.368 & 0.823 & 0.725\\
+pipe & holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 & 1.001 & 0.869 & 3.264 & 0.811 & 0.710\\
+pipe & holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 & 1.000 & 0.885 & 3.151 & 0.806 & 0.697\\
+pipe & holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 & 1.000 & 0.878 & 3.345 & 0.806 & 0.734\\
+sine & MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 & 0.894 & 0.571 & 1.623 & -0.024 & 0.207\\
 \addlinespace
-sine & MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 & 0.932 & 0.328 & 1.314 & -0.030\\
-sine & TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 & 0.951 & 0.536 & 1.719 & -0.025\\
-sine & TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 & 0.945 & 0.564 & 1.723 & -0.027\\
-sine & dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019\\
-sine & loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 & 1.016 & 1.039 & 2.648 & 0.080\\
+sine & MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 & 0.932 & 0.328 & 1.314 & -0.030 & 0.100\\
+sine & TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 & 0.951 & 0.536 & 1.719 & -0.025 & 0.219\\
+sine & TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 & 0.945 & 0.564 & 1.723 & -0.027 & 0.230\\
+sine & dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019 & 0.679\\
+sine & loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 & 1.016 & 1.039 & 2.648 & 0.080 & 0.699\\
 \addlinespace
-sine & splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 & 1.014 & 1.051 & 2.730 & -0.009\\
-sine & stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 & 1.011 & 0.011 & 254.734 & 0.727\\
+sine & splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 & 1.014 & 1.051 & 2.730 & -0.009 & 0.727\\
+sine & stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 & 1.011 & 0.011 & 254.734 & 0.727 & 0.711\\
 \bottomrule
@@
 
-\begin{table}
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0811}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0405}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1486}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1892}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1351}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1351}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1081}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1622}}@{}}
 
-\caption{\label{tbl-all-sim}The first five rows of the datasets after
-combining all the calculation. The smoothness and squintability variable
-are uniquely characterised by the index function used and the data
-dimension, and thus do not vary across n\_jellies and max\_tries.
-Variables I\_max\_max, P\_J\_hat, and time are calculated at each 51
-observations.}
+\caption{\label{tbl-all-sim}The first few rows of the datasets processed
+for modelling. The smoothness and squintability variable are uniquely
+characterised by the index function used and the data dimension, and
+thus do not vary across n\_jellies and max\_tries. The variable
+P\_J\_hat, and time are calculated at each observation.}
 
-\centering{
+\tabularnewline
 
-}
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+index
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+d
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+smoothness
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+squintability
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+n\_jellies
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+max\_tries
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+P\_J\_hat
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+time
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+MIC & 6 & 2.394 & 0.207 & 20 & 50 & 0.12 & 2.479 secs \\
+MIC & 6 & 2.394 & 0.207 & 20 & 100 & 0.24 & 8.950 secs \\
+MIC & 6 & 2.394 & 0.207 & 50 & 50 & 0.52 & 5.651 secs \\
+MIC & 6 & 2.394 & 0.207 & 50 & 100 & 0.64 & 13.223 secs \\
+MIC & 6 & 2.394 & 0.207 & 100 & 50 & 0.76 & 19.453 secs \\
+MIC & 8 & 2.394 & 0.100 & 20 & 50 & 0.08 & 2.566 secs \\
+MIC & 8 & 2.394 & 0.100 & 20 & 100 & 0.08 & 4.960 secs \\
 
-\end{table}%
+\end{longtable}
+
+\begin{verbatim}
+# A tibble: 7 x 5
+  term          estimate std.error statistic p.value
+  <chr>            <dbl>     <dbl>     <dbl>   <dbl>
+1 (Intercept)     -4.67      4.51     -1.04  0.300  
+2 smoothness       1.60      1.53      1.05  0.294  
+3 squintability    7.06      2.21      3.20  0.00140
+4 d               -0.595     0.260    -2.29  0.0221 
+5 long_time       -0.851     1.35     -0.632 0.528  
+6 n_jellies        0.230     0.131     1.75  0.0795 
+7 max_tries        0.107     0.153     0.700 0.484  
+\end{verbatim}
 

### 2024-06-03 code/06-compile-all-sims.R

@@ -1,16 +1,12 @@
 library(tidyverse)
 load(file = "data/sim_pipe.rda")
 
-tour_level_best_basis <- sim_pipe |>
-  group_by(id) |>
-  filter(index_val == max(index_val)) |>
-  filter(row_number() == 1)
-
-setup_level_best_basis <- sim_pipe |>
-  group_by(n_jellies, max_tries, d) |>
-  filter(index_val == max(index_val)) |>
-  filter(row_number() == 1)
-
+# all the pipe simulation are conducted in one go,
+# so id is different for each n_jellies, max_tries, d combination
+pipe_run_df <- sim_pipe |> get_best(group = id)
+pipe_setup_df <- sim_pipe |>
+  mutate(id2 = paste0(n_jellies, max_tries, d)) |>
+  get_best(group = id2)
 
 # columns
 # setup:      number of experiment (1-4),
@@ -22,24 +18,28 @@ setup_level_best_basis <- sim_pipe |>
 # D:          the coefficient in c_t > .5 (the 3 in best_jelly - 3 * runif(1) ...)
 # I_max:      best index across all the jellyfish in each trial run,
 # basis:      projection basis
-tour_level_pipe <- sim_pipe |>
+pipe_run_summ <- sim_pipe |>
+  mutate(idx_f = "holes") |>
   group_by(id) |>
   reframe(I_max = max(index_val),
-          P_J = sum(abs(I_max - index_val) <= 0.05)/n()) |>
-  left_join(tour_level_best_basis)
-
+          P_J = sum(abs(I_max - index_val) <= 0.05)/n(),
+          time = mean(time))
 
 # columns
 # P_J_hat:    proportion of trials found
 # I_max_max:  best index of each setup
-setup_level_pipe <- tour_level_pipe |>
-  group_by(n_jellies, max_tries, d) |>
+pipe_setup_summ <- pipe_run_df |>
+  mutate(idx_f = "holes") |>
+  group_by(n_jellies, max_tries, d, idx_f) |>
   reframe(
-    I_max_max = max(I_max),
-    P_J_hat = sum(abs(I_max_max - I_max) <= 0.05)/n() # TODO: use a better rule
-    )
+    I_max_max = max(index_val),
+    P_J_hat = sum(abs(I_max_max - index_val) <= 0.05)/n(),
+    time = mean(time)
+  )
 
-setup_level_pipe |>
+
+
+pipe_setup |>
[…]

### 2024-06-03 code/07-add-smoothness-squint.R

@@ -84,7 +84,8 @@ squintability <- sq_basis_df |>
   mutate(index = factor(index, levels = c("holes", "MIC", "TIC", "dcor2d_2",
                                           "loess2d", "splines2d", "stringy"))) |>
   arrange(index) |>
-  select(index, n, theta1: theta4)
+  select(index, n, theta1: theta4) |>
+  mutate(squint = theta1 * theta2 * theta3 / 4)
 save(squintability, file = here::here("data", "squintability.rda"))
 
 ############################################################################
@@ -92,7 +93,7 @@ save(squintability, file = here::here("data", "squintability.rda"))
 load(here::here("data", "sim_summary.rda"))
 sim_df <- sim_summary |>
   left_join(smoothness |> select(n, index, smoothness) |> rename(d = n)) |>
-  left_join(squintability |> select(index, n, theta3) |> rename(d = n, squintability = theta3)) |>
+  left_join(squintability |> select(index, n, squint) |> rename(d = n, squintability = squint)) |>
   select(index, d, I_max_max, P_J_hat, n_jellies, max_tries, smoothness, squintability, time)
 save(sim_df, file = here::here("data", "sim_df.rda"))
 

### 2024-06-03 code/08-simulation-sine-6d-stringy.R

@@ -1,0 +1,73 @@
+library(tidyverse)
+library(spinebil)
+library(tourr)
+
+################################################################################
+# d = 6
+sim <- function(d = d, n_jellies = n_jellies, max.tries = max.tries,
+                data_seed = 123456, optim_seed = seed, sim = sim){
+  set.seed(data_seed)
+  sine1000 <- sinData(d, 1000) %>% scale() %>% as_tibble()
+  colnames(sine1000) <- paste0("V", 1:d)
+
+  cat("sim: ", sim, "\n")
+  cat("n_jellies: ", n_jellies, "\n")
+  cat("max.tries: ", max.tries, "\n")
+  cat("d: ", d, "\n")
+  res <- map_dfr(optim_seed, function(seed){
+    t1 <- Sys.time()
+    set.seed(seed)
+    res <- animate_xy(sine1000, guided_tour(
+      tourr::stringy(), d = 2, n_jellies = n_jellies,
+      search_f =  search_jellyfish,
+      max.tries = max.tries
+    ), max_frames = max.tries) |> dplyr::select(-id)
+    t2 <- Sys.time()
+    res <- res |> mutate(time = t2 - t1)
+    return(res)
+  })
+
+  return(res)
+}
+
+# separate into sim 1-30 and 31-50 to run, each takes < 2 hrs
+set.seed(123)
+seed <- sample(1000: 10000, size = 50)
+sim_setup <- crossing(n_jellies = 50,
+                      max_tries = 50,
+                      d = 6) |>
+  crossing(sim = 21:50) |>
+  mutate(seed = seed[sim], id = row_number())
+
+t1 <- Sys.time()
+set.seed(123)
+sim_res <- sim_setup |>
+  rowwise() |>
+  mutate(res = list(sim(d, n_jellies, max_tries, optim_seed = seed, sim = sim))) |>
+  ungroup() |>
+  unnest(res)
+t2 <- Sys.time()
+t2 - t1
+
+#sim_sine_6d_stringy12 <- sim_res |> select(-alpha)
+#sim_sine_6d_stringy310 <- sim_res |> select(-alpha)
+#sim_sine_6d_stringy1120 <- sim_res |> select(-alpha)
+sim_sine_6d_stringy2150 <- sim_res |> select(-alpha)
+
+sim_sine_6d_stringy <- bind_rows(sim_sine_6d_stringy12, sim_sine_6d_stringy310,
[…]

### 2024-06-03 code/00-combine-simulations.R
clean up repo

@@ -1,31 +1,0 @@
-#####################################################################
-# simulation sine 68d TICMIC
-load(file = "data/sim_sine_68d_TICMIC.rda")
-sim_data <- sim_sine_68d_TICMIC
-setup_level_best_basis <- sim_data |>
-  group_by(n_jellies, max_tries, d, idx_f) |>
-  filter(index_val == max(index_val)) |>
-  filter(row_number() == 1)
-
-ids <- setup_level_best_basis |> filter(d == 6) |> pull(id)
-res <- map_dfr(ids, ~{
-  b <- setup_level_best_basis |> filter(id == .x) |> pull(basis)
-  dt <- as_tibble(as.matrix(sine1000) %*% b[[1]])
-  return(dt)
-}, .id = "id")
-
-labels_df <- setup_level_best_basis |>
-  ungroup() |>
-  filter(d == 6) |>
-  select(idx_f:max_tries, index_val) |>
-  mutate(id = as.character(1:10))
-
-library(ggh4x)
-res |>
-  left_join(labels_df) |>
-  ggplot(aes(x = V2 ,y = V1)) +
-  geom_point() +
-  geom_text(data = labels_df,
-            aes(x = -3, y = 3, label = round(index_val, 2)),
-            nudge_x = 0.1, nudge_y = 0.1) +
-  facet_nested(idx_f ~ n_jellies + max_tries, labeller = label_both )

### 2024-06-03 code/07-calc-smoothness-squintability.R

@@ -85,18 +85,8 @@ squintability <- sq_basis_df |>
                                           "loess2d", "splines2d", "stringy"))) |>
   arrange(index) |>
   select(index, n, theta1: theta4) |>
-  mutate(squint = theta1 * theta2 * theta3 / 4)
+  mutate(squint = abs(theta1 * theta2 * theta3 / 4))
 save(squintability, file = here::here("data", "squintability.rda"))
-
-############################################################################
-############################################################################
-load(here::here("data", "sim_summary.rda"))
-sim_df <- sim_summary |>
-  left_join(smoothness |> select(n, index, smoothness) |> rename(d = n)) |>
-  left_join(squintability |> select(index, n, squint) |> rename(d = n, squintability = squint)) |>
-  select(index, d, I_max_max, P_J_hat, n_jellies, max_tries, smoothness, squintability, time)
-save(sim_df, file = here::here("data", "sim_df.rda"))
-
 
 
 

### 2024-06-03 code/08-process-data-for-modelling.R

@@ -102,6 +102,15 @@ sine_setup_summ <- sine_run_df |>
 sim_summary <- pipe_setup_summ |>  bind_rows(sine_setup_summ) |>
   rename(index = idx_f) |>
   mutate(index = ifelse(index == "spline", "splines2d", index))
-save(sim_summary, file = here::here("data/sim_summary.rda"))
+
+############################################################################
+############################################################################
+load(here::here("data", "squintability.rda"))
+load(here::here("data", "smoothness.rda"))
+sim_df <- sim_summary |>
+  left_join(smoothness |> select(n, index, smoothness) |> rename(d = n)) |>
+  left_join(squintability |> select(index, n, squint) |> rename(d = n, squintability = squint)) |>
+  select(index, d, I_max_max, P_J_hat, n_jellies, max_tries, smoothness, squintability, time)
+save(sim_df, file = here::here("data", "sim_df.rda"))
 
 

### 2024-06-03 optim.qmd
example 2 is there :)

--- 2024-05-31 optim.qmd
@@
 
-```{r setup, echo = FALSE, message=FALSE}
+```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
@@
 library(ggh4x)
+library(broom)
+library(kableExtra)
+load(here::here("data/smoothness.rda"))
+load(here::here("data/squintability.rda"))
+load(here::here("data/sq_basis_dist_idx.rda"))
+load(here::here("data/sim_df.rda"))
 ```
@@
 
-## Another data example
+## Connecting optimisation success with its properties and jellyfish parameters
 
-[TODO: overall summary]construct a relationship between jellyfish success and jellyfish parameters and the optimisation properties defined in @sec-theory.
-This can inform the choice of jellyfish parameters for a given
-optimisation problem.
+The optimisation properties (smoothness, squintability, and speed) proposed in the @sec-theory provide numerical measures to characterise the difficulty of the optimisation for a given projection pursuit problem.
+This example investigates how these measures, combined with the jellyfish parameters, affect the success of the jellyfish optimisers and how they can guide the decisions made in hyperparameter tuning with the jellyfish optimiser.
+Simulations are conducted with different jellyfish optimisers to obtain the proportion of successful optimisation across a collection of scenarios, for which smoothness, squintability, and speed are calculated.
+A generalised linear model is used to link the proportion of success rate with the jellyfish parameters and the optimisation properties.
 
-In addition to the pipe-finding problem, the sine-wave finding problem is also investigated in 6D and 8D spaces.
-Six indices are considered: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`, which gives the 12 problems printed as rows in @tbl-smoothness-squintability.
-Combining these problems with different jellyfish optimiser parameters (`n_jellies` and `max_tries`), a total of 51 observations is produced, comprising of 24 observations for the pipe-finding and 27 observations for the sine-wave finding.
+In addition to the pipe-finding problem, also investigated is the sine-wave finding problem in 6D and 8D spaces with six indices considered: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`.
+Combining with different jellyfish optimiser parameters (`n_jellies` and `max_tries`), a total of `r nrow(sim_df)` cases is produced, comprising of `r nrow(filter(sim_df, index == "holes"))` pipe-finding cases  and `r nrow(filter(sim_df, index != "holes"))` sine-wave finding cases.
+For each case, the jellyfish optimiser is run fifty (50) times and summary statistics are calculated on the best index value found across all 50 runs and the proportion of runs that find a close (with a difference less than 0.05) best index value (`P_J_hat`).
 
-For each observation, fifty (50) runs of the jellyfish optimisation is performed.
-In each run, summary statistics are calculated to find the best index across all the jellyfishes (`I_max`) and the proportion of jellyfishes found (`P_J`).
-At the observation level, the best index value across all the 50 runs  (`I_max_max`) and the proportion of trials found (`P_J_hat`) is calculated.
+Smoothness and squintability are computed for each case, following the procedures outlined in @sec-smoothness and @sec-squintability.
+To calculate smoothness, three hundred random bases are simulated.
+Index values and projection distance to the optimal basis are calculated for each basis to fit a Gaussian process model.
+For squintability, fifty random bases are sampled and interpolated to the optimal basis with a step size of 0.005.
+The interpolated bases are first binned with a bin width of 0.005 to average the index values before feeding into a 4-parameter scaled logistic function, estimated by non-linear least square, to obtain the squintability measure.
 
-Smoothness and squintability measures are computed for each problem, following the procedures outlined in @sec-smoothness and @sec-squintability.
-To calculate smoothness, 300 random bases are used to estimate Gaussian process parameters (variance, range, smoothness, and nugget).
-For squintability, 50 random bases are sampled and interpolated to the optimal basis using a step size of 0.005.
-The index value and projection distance are calculated for all the interpolated bases.
-A binning procedure is then applied to average the index value at each bin width of 0.005.
-Finally, the scaled logistic function with four parameters (theta1 - theta4) is estimated by non-linear least square to obtain the squintability measure.
+@tbl-smoothness-squintability presents the parameters estimated for each case from the Gaussian process (variance, range, smooth, and nugget) for smoothness and the scaled logistic function (theta1 to theta4, and squint) for squintability.
+The column "smooth" and "squint" are used as the smoothness and squintability measure.
+The low squint value of MIC and TIC index is due to its convex shape of the index value against the projection distance, as shown in @fig-idx-proj-dist.
+Comparing to other concave shapes, the maximum first derivative happens at a smaller projection distance ($\theta2$), requiring the optimiser to get closer to the optimal to see a significant change in the index value, hence a small squintability measure.
+[TODO: the stringy squintability doens't make sense]
 
-@tbl-smoothness-squintability presents all the parameters calculated for smoothness and squintability, with the column "smooth" for smoothness and "theta3" for squintability.
-@fig-idx-proj-dist shows the relationship between the index value against the projection distance for each problem following the binning procedure.
+@tbl-mod-data presents the processed data for modelling after augmenting the jellyfish parameters (`n_jellies` and `max_tries`) and the average time spent on each run (`time`) for each case.
+A generalised linear model is fitted with a binomial family and a logit link function to the data to investigate the relationship between the proportion of jellyfish success and jellyfish parameters and the optimisation properties.
+Pre-processing of the data include 1) scale the jellyfish parameters (`n_jellies` and `max_tries`) by 10 for interpretation, and 2) create a binary variable `long_time` to flag the cases with average time spent on each run greater than 30 seconds.
+@tbl-mod-output presents the model output with the estimated coefficients.
+The model suggests a positive relationship between the proportion of jellyfish success and the jellyfish parameters, `n_jellies` and `max_tries`, optimisation property parameter, `smoothness` and `squintability`, and a negative relationship with the average time spent on each run (`long_time`) and the data dimension (`d`).
+The variable `squintability` and `d` are significant, suggesting their importance over jellyfish parameters to the success of the optimisation.
 
@@
 #| label: tbl-smoothness-squintability
-#| tbl-cap: Parameters estimated from the Gaussian process (including variance, range, smoothness, and nugget) and scaled logistic function (theta1 - theta4) for both the pipe-finding and sine-wave finding problems.
-The "smooth" and "theta3" column represent the smoothness and squintability measures.
-load(here::here("data/smoothness.rda"))
-load(here::here("data/squintability.rda"))
-library(kableExtra)
+#| tbl-cap: Parameters estimated from the Gaussian process (including variance, range, smoothness, and nugget) and scaled logistic function (theta1 to theta4) for the pipe-finding and sine-wave finding problems.
+The squint column is calculated as $\frac{\theta1 \theta2 \theta3}{4}$, as described in @sec-squintability.
+The "smooth" and "squint" column represent the smoothness and squintability measures.
 tibble(shape = c(rep("pipe", 4), rep("sine", 8)), smoothness) |>
@@
   rename(d = n, smooth = smoothness) |> 
-  knitr::kable(digits = 3) |> 
+  knitr::kable(digits = 3, booktabs = T) |> 
   column_spec(1, border_left = TRUE) |> 
-  column_spec(c(3, 7, 11), border_right = TRUE) |> 
-  kable_styling(latex_options = "hold_position")
+  column_spec(c(3, 7, 12), border_right = TRUE) |> 
+  kable_styling(latex_options = "hold_position", font_size = 7)
 ```
@@
 #| fig.cap = "Index values against projection distance for the 12 pipe/sine-wave finding problem after the binning procedure during the estimation of the squintability measure.
-The index value is averaged at each bin width of 0.005 and the TIC index is scaled to 0-1 for comparison.
-When finding the sine-wave structure using the MIC and TIC index, a convex curved is observed, as opposed to the pipe-finding problem or the sine-wave finding problem with the dcor2d\\_2, loess2d, and splines2d index.
+The index value is averaged at each bin width of 0.005 and the `TIC` index is scaled to 0-1 for comparison.
+When finding the sine-wave structure using the `MIC` and `TIC` index, a convex curved is observed, as opposed to the pipe-finding problem or the sine-wave finding problem with the `dcor2d_2`, `loess2d`, and `splines2d` index.
 When finding the sine-wave with the stringy index, the index shows an instantaneous jump to the optimum when close to the best basis."
-load(here::here("data/sq_basis_dist_idx.rda"))
 sq_basis_dist_idx |>
@@
 
-```{r eval = FALSE}
-#| label: tbl-all-sim
-#| tbl-cap: The first five rows of the datasets after combining all the calculation.
-The smoothness and squintability variable are uniquely characterised by the index function used and the data dimension, and thus do not vary across n\\_jellies and max\\_tries.
-Variables I\\_max\\_max, P\\_J\\_hat, and time are calculated at each 51 observations.
-load(here::here("data/sim_df.rda"))
+```{r}
+#| label: tbl-mod-data
+#| tbl-cap: The first few rows of the datasets processed for modelling.
+The smoothness and squintability variable are uniquely characterised by the index function used and the data dimension, and thus do not vary across `n_jellies` and `max_tries`.
+The variable `P_J_hat`, and time are calculated at each observation.
 sim_df |> 
   dplyr::select(index, d, smoothness, squintability, n_jellies, max_tries, 
-                I_max_max, P_J_hat, time) |> 
+                -I_max_max, P_J_hat, time) |> 
   arrange(index, d, n_jellies, max_tries) |> 
-  head(5) |> 
+  head(7) |> 
   kable(digits = 3)
+```
+
+```{r}
+#| label: tbl-mod-output
+#| tbl-cap: Model estimates of proportion of jellyfish success (`P_J_hat`) on optimisation properties (`smoothness`, `squintability`, and `d`) and jellyfish parameters (`long_time`, `n_jellies`, and `max_tries`) from the generalised linear model with a binomial family and a logit link function.
+The variable `long_time` is derived to flag the cases with average time spent on each run greater than 30 seconds and the variable `n_jellies` and `max_tries` are scaled by 10 for interpretation.
+sim_df2 <- sim_df |> 
+  mutate(n_jellies = n_jellies/10,
+         max_tries = max_tries/10,
+         long_time = ifelse(time > 30, 1, 0),
+         P_J_hat = ifelse(index == "stringy", 0, P_J_hat))
+
+mod1 <- glm(P_J_hat ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
+broom::tidy(mod1) |> 
+  mutate(term = ifelse(term == "(Intercept)", "intercept", term)) |>
+  kable(digits = 3) |> 
+  column_spec(c(1,5), border_right = TRUE) |> 
+  column_spec(1, border_left = TRUE)
 ```

### 2024-06-03 optim.tex
example 2 is there :)

@@
 
-\begin{verbatim}
-Warning: package 'ggplot2' was built under R version 4.3.2
-\end{verbatim}
-
-\begin{verbatim}
-Warning: package 'tidyr' was built under R version 4.3.2
-\end{verbatim}
-
-\begin{verbatim}
-Warning: package 'ggh4x' was built under R version 4.3.2
-\end{verbatim}
-
 \section{Introduction {[}Nicolas and
@@
 
-\begin{tcolorbox}[enhanced jigsaw, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, leftrule=.75mm, arc=.35mm, left=2mm, opacityback=0, breakable, toptitle=1mm, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, colback=white, titlerule=0mm, bottomrule=.15mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame]
+\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, coltitle=black, colframe=quarto-callout-note-color-frame, opacityback=0, toprule=.15mm, opacitybacktitle=0.6, colback=white, bottomrule=.15mm, leftrule=.75mm, titlerule=0mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, rightrule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white, left=2mm]
 
@@
 
-\subsection{Another data example}\label{another-data-example}
+\subsection{Connecting optimisation success with its properties and
+jellyfish
+parameters}\label{connecting-optimisation-success-with-its-properties-and-jellyfish-parameters}
 
-{[}TODO: overall summary{]}construct a relationship between jellyfish
-success and jellyfish parameters and the optimisation properties defined
-in Section~\ref{sec-theory}. This can inform the choice of jellyfish
-parameters for a given optimisation problem.
+The optimisation properties (smoothness, squintability, and speed)
+proposed in the Section~\ref{sec-theory} provide numerical measures to
+characterise the difficulty of the optimisation for a given projection
+pursuit problem. This example investigates how these measures, combined
+with the jellyfish parameters, affect the success of the jellyfish
+optimisers and how they can guide the decisions made in hyperparameter
+tuning with the jellyfish optimiser. Simulations are conducted with
+different jellyfish optimisers to obtain the proportion of successful
+optimisation across a collection of scenarios, for which smoothness,
+squintability, and speed are calculated. A generalised linear model is
+used to link the proportion of success rate with the jellyfish
+parameters and the optimisation properties.
 
-In addition to the pipe-finding problem, the sine-wave finding problem
-is also investigated in 6D and 8D spaces. Six indices are considered:
-\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
-\texttt{spline}, and \texttt{stringy}, which gives the 12 problems
-printed as rows in Table~\ref{tbl-smoothness-squintability}. Combining
-these problems with different jellyfish optimiser parameters
-(\texttt{n\_jellies} and \texttt{max\_tries}), a total of 51
-observations is produced, comprising of 24 observations for the
-pipe-finding and 27 observations for the sine-wave finding.
+In addition to the pipe-finding problem, also investigated is the
+sine-wave finding problem in 6D and 8D spaces with six indices
+considered: \texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC},
+\texttt{TIC}, \texttt{spline}, and \texttt{stringy}. Combining with
+different jellyfish optimiser parameters (\texttt{n\_jellies} and
+\texttt{max\_tries}), a total of 52 cases is produced, comprising of 24
+pipe-finding cases and 28 sine-wave finding cases. For each case, the
+jellyfish optimiser is run fifty (50) times and summary statistics are
+calculated on the best index value found across all 50 runs and the
+proportion of runs that find a close (with a difference less than 0.05)
+best index value (\texttt{P\_J\_hat}).
 
-For each observation, fifty (50) runs of the jellyfish optimisation is
-performed. In each run, summary statistics are calculated to find the
-best index across all the jellyfishes (\texttt{I\_max}) and the
-proportion of jellyfishes found (\texttt{P\_J}). At the observation
-level, the best index value across all the 50 runs
-(\texttt{I\_max\_max}) and the proportion of trials found
-(\texttt{P\_J\_hat}) is calculated.
+Smoothness and squintability are computed for each case, following the
+procedures outlined in Section~\ref{sec-smoothness} and
+Section~\ref{sec-squintability}. To calculate smoothness, three hundred
+random bases are simulated. Index values and projection distance to the
+optimal basis are calculated for each basis to fit a Gaussian process
+model. For squintability, fifty random bases are sampled and
+interpolated to the optimal basis with a step size of 0.005. The
+interpolated bases are first binned with a bin width of 0.005 to average
+the index values before feeding into a 4-parameter scaled logistic
+function, estimated by non-linear least square, to obtain the
+squintability measure.
 
-Smoothness and squintability measures are computed for each problem,
-following the procedures outlined in Section~\ref{sec-smoothness} and
-Section~\ref{sec-squintability}. To calculate smoothness, 300 random
-bases are used to estimate Gaussian process parameters (variance, range,
-smoothness, and nugget). For squintability, 50 random bases are sampled
-and interpolated to the optimal basis using a step size of 0.005. The
-index value and projection distance are calculated for all the
-interpolated bases. A binning procedure is then applied to average the
-index value at each bin width of 0.005. Finally, the scaled logistic
-function with four parameters (theta1 - theta4) is estimated by
-non-linear least square to obtain the squintability measure.
+Table~\ref{tbl-smoothness-squintability} presents the parameters
+estimated for each case from the Gaussian process (variance, range,
+smooth, and nugget) for smoothness and the scaled logistic function
+(theta1 to theta4, and squint) for squintability. The column ``smooth''
+and ``squint'' are used as the smoothness and squintability measure. The
+low squint value of MIC and TIC index is due to its convex shape of the
+index value against the projection distance, as shown in
+Figure~\ref{fig-idx-proj-dist}. Comparing to other concave shapes, the
+maximum first derivative happens at a smaller projection distance
+(\(\theta2\)), requiring the optimiser to get closer to the optimal to
+see a significant change in the index value, hence a small squintability
+measure. {[}TODO: the stringy squintability doens't make sense{]}
 
-{[}TODO: discuss what the values of smooth and theta3 mean for the
-optimisation{]}
+Table~\ref{tbl-mod-data} presents the processed data for modelling after
+augmenting the jellyfish parameters (\texttt{n\_jellies} and
+\texttt{max\_tries}) and the average time spent on each run
+(\texttt{time}) for each case. A generalised linear model is fitted with
+a binomial family and a logit link function to the data to investigate
+the relationship between the proportion of jellyfish success and
+jellyfish parameters and the optimisation properties. Pre-processing of
+the data include 1) scale the jellyfish parameters (\texttt{n\_jellies}
+and \texttt{max\_tries}) by 10 for interpretation, and 2) create a
+binary variable \texttt{long\_time} to flag the cases with average time
+spent on each run greater than 30 seconds. Table~\ref{tbl-mod-output}
+presents the model output with the estimated coefficients. The model
+suggests a positive relationship between the proportion of jellyfish
+success and the jellyfish parameters, \texttt{n\_jellies} and
+\texttt{max\_tries}, optimisation property parameter,
+\texttt{smoothness} and \texttt{squintability}, and a negative
+relationship with the average time spent on each run
+(\texttt{long\_time}) and the data dimension (\texttt{d}). The variable
+\texttt{squintability} and \texttt{d} are significant, suggesting their
+importance over jellyfish parameters to the success of the optimisation.
 
-Table~\ref{tbl-smoothness-squintability} presents all the parameters
-calculated for smoothness and squintability, with the column ``smooth''
-for smoothness and ``theta3'' for squintability.
-Figure~\ref{fig-idx-proj-dist} shows the relationship between the index
-value against the projection distance for each problem following the
-binning procedure.
+\begingroup\fontsize{7}{9}\selectfont
 
@@
 the Gaussian process (including variance, range, smoothness, and nugget)
-and scaled logistic function (theta1 - theta4) for the pipe-finding and
-sine-wave finding problems. The squint column is calculated as theta1 *
-theta2 * theta3 / 4, as described in Section~\ref{sec-squintability}.
-The ``smooth'' and ``squint'' column represent the smoothness and
-squintability measures.}
+and scaled logistic function (theta1 to theta4) for the pipe-finding and
+sine-wave finding problems. The squint column is calculated as
+\(\frac{\theta1 \theta2 \theta3}{4}\), as described in
+Section~\ref{sec-squintability}. The ``smooth'' and ``squint'' column
+represent the smoothness and squintability measures.}
 
@@
 
+\endgroup{}
+
 \begin{figure}
@@
 procedure during the estimation of the squintability measure. The index
-value is averaged at each bin width of 0.005 and the TIC index is scaled
-to 0-1 for comparison. When finding the sine-wave structure using the
-MIC and TIC index, a convex curved is observed, as opposed to the
-pipe-finding problem or the sine-wave finding problem with the
-dcor2d\_2, loess2d, and splines2d index. When finding the sine-wave with
-the stringy index, the index shows an instantaneous jump to the optimum
-when close to the best basis.}
+value is averaged at each bin width of 0.005 and the \texttt{TIC} index
+is scaled to 0-1 for comparison. When finding the sine-wave structure
+using the \texttt{MIC} and \texttt{TIC} index, a convex curved is
+observed, as opposed to the pipe-finding problem or the sine-wave
+finding problem with the \texttt{dcor2d\_2}, \texttt{loess2d}, and
+\texttt{splines2d} index. When finding the sine-wave with the stringy
+index, the index shows an instantaneous jump to the optimum when close
+to the best basis.}
 
@@
 
-\caption{\label{tbl-all-sim}The first few rows of the datasets processed
-for modelling. The smoothness and squintability variable are uniquely
-characterised by the index function used and the data dimension, and
-thus do not vary across n\_jellies and max\_tries. The variable
-P\_J\_hat, and time are calculated at each observation.}
+\caption{\label{tbl-mod-data}The first few rows of the datasets
+processed for modelling. The smoothness and squintability variable are
+uniquely characterised by the index function used and the data
+dimension, and thus do not vary across \texttt{n\_jellies} and
+\texttt{max\_tries}. The variable \texttt{P\_J\_hat}, and time are
+calculated at each observation.}
 
@@
 
-\begin{verbatim}
-# A tibble: 7 x 5
-  term          estimate std.error statistic p.value
-  <chr>            <dbl>     <dbl>     <dbl>   <dbl>
-1 (Intercept)     -4.67      4.51     -1.04  0.300  
-2 smoothness       1.60      1.53      1.05  0.294  
-3 squintability    7.06      2.21      3.20  0.00140
-4 d               -0.595     0.260    -2.29  0.0221 
-5 long_time       -0.851     1.35     -0.632 0.528  
-6 n_jellies        0.230     0.131     1.75  0.0795 
-7 max_tries        0.107     0.153     0.700 0.484  
-\end{verbatim}
+\begin{longtable}[t]{|>{}lrrr>{}r|}
+
+\caption{\label{tbl-mod-output}Model estimates of proportion of
+jellyfish success (\texttt{P\_J\_hat}) on optimisation properties
+(\texttt{smoothness}, \texttt{squintability}, and \texttt{d}) and
+jellyfish parameters (\texttt{long\_time}, \texttt{n\_jellies}, and
+\texttt{max\_tries}) from the generalised linear model with a binomial
+family and a logit link function. The variable \texttt{long\_time} is
+derived to flag the cases with average time spent on each run greater
+than 30 seconds and the variable \texttt{n\_jellies} and
+\texttt{max\_tries} are scaled by 10 for interpretation.}
+
+\tabularnewline
+
+\toprule
+term & estimate & std.error & statistic & p.value\\
+\midrule
+intercept & -4.672 & 4.512 & -1.036 & 0.300\\
+smoothness & 1.604 & 1.530 & 1.048 & 0.294\\
+squintability & 7.062 & 2.210 & 3.195 & 0.001\\
+d & -0.595 & 0.260 & -2.289 & 0.022\\
+long\_time & -0.851 & 1.348 & -0.632 & 0.528\\
+\addlinespace
+n\_jellies & 0.230 & 0.131 & 1.754 & 0.080\\
+max\_tries & 0.107 & 0.153 & 0.700 & 0.484\\
+\bottomrule
+
+\end{longtable}
 

### 2024-06-06 optim.qmd
working towards the example 1

@@
 library(kableExtra)
+library(ferrn)
+load(here::here("data/sim_pipe.rda"))
+load(here::here("data/pipe_better.rda"))
+load(here::here("data/pipe_jellyfish.rda"))
 load(here::here("data/smoothness.rda"))
@@
 
-## Going beyond 10D
+## Performance of the jellyfish optimiser in high-dimensional pipe-finding problems
 
-The pipe-finding problem is initially used to investigate indexes and
-optimisers in @laa_using_2020, and we extend it from a 6D problem to a
-12D problem.
+[unfinished]
 
-Jellyfish optimiser, as a multi-start algorithm, is efficient in \[...\]
-for high-dimensional problems
+The aim of this example is to investigate the performance of jellyfish optimisers in high-dimensional problems.
 
-```{r echo = FALSE}
-load(here::here("data/pipe_better.rda"))
-load(here::here("data/pipe_jellyfish.rda"))
-```
+The projection pursuit problem is to find the pipe shape, investigated in @laa_using_2020, using the holes index in 6/8/10/12-D data.
+For each case, 50 simulations are conducted for each combination of jellyfish hyper-parameter: `n_jellies = 20/50/100` and `max_tries = 50/100`.
+The performance of the jellyfish optimiser is also compared with an existing opitmiser: the search better opitmiser with `max.tries = 1000`.
+
+@fig-proportion presents the proportion of jellyfishes that finds the index value close to (absolute value less than 0.05) the best index value found at each `max_tries`/`n_jellies` combination at $d = 6/8/10/12$ data dimension.
+As the number of jellies and maximum tries increase, the proportion of jellyfishes that find the optimal projection increases.
+This suggests for higher dimensional problems ($d = 8/10/12$), an `n_jellies = 100` and `max_tries = 100` is necessary for at least half of the jellyfish optimiser to find the optimal projection.
+
+@fig-proj presents the final projection found across the 50 simulations at each 10th quantile between the jellyfish and search better optimiser (`n_jellies = 100`, `max_tries = 100` is used for the jellyfish optimiser).
+While the jellyfish optimiser may not always find the optimal projection in the worse case scenarios, especially as the dimension increases, it tends to find higher index values and crisper pipe shapes than the search better optimiser at each quantile across all the four data dimensions.
 
 ```{r}
+#| label = fig-proportion,
 #| fig.width = 6,
@@
 #| fig.align = "center",
-#| fig.cap = "sthis sdfaksdlf"
-pipe_better |>
-  bind_rows(pipe_jellyfish) |>
-  mutate(dim = as.factor(6 + (as.numeric(dim) - 1) * 2)) |>
-  ggplot(aes(x = dim, y = index_val, group = interaction(optimiser, dim), color = optimiser)) +
-  geom_violin(width = 2, adjust = 0.5, position = position_dodge(width = 0.4)) +
-  ggbeeswarm::geom_quasirandom(size = 0.1, width = 0.1, dodge.width = 0.4) +
-  theme_bw() +
-  scale_y_continuous(breaks = seq(0.85, 1, 0.01)) +
-  scale_color_brewer(palette = "Accent") + 
-  labs(y = "Index", x = "Dimension")
+#| fig.cap = "Proportion of jellyfishes reaches close to the optimal index value in the pipe-finding problem.
+As the dimension goes up, the numer/ proportion of jellyfishes that reach the optimal index value decreases.
+Increasing the jellyfish hyperparameters (`n_jellies` and `max_tries`) help to improve the proportion."
+pipe_run_summ <- sim_pipe |>
+  mutate(idx_f = "holes") |>
+  group_by(id, n_jellies, max_tries, d, loop) |> 
+  summarise(I_max = max(index_val))
+
+pipe_sim_best <- pipe_run_summ |> 
+  ungroup(loop, id) |> 
+  summarise(I_best = max(I_max))
+
+pipe_res <- pipe_run_summ |> 
+  left_join(pipe_sim_best) |> 
+  mutate(diff = abs(I_max - I_best) < 0.05) |> 
+  ungroup(id, loop) |> 
+  summarise(proportion = sum(diff) / n()) |> 
+  mutate(d = as.factor(d), n_jellies = as.factor(n_jellies))
+
+pipe_res |> 
+  ggplot(aes(x = n_jellies, y = proportion, 
+             group = interaction(d, max_tries),
+             color = d)) + 
+  geom_line() + 
+  scale_color_brewer(palette = "Dark2") +
+  facet_wrap(vars(max_tries), 
+             labeller = "label_both") + 
+  theme_bw()
 
@@
 ```{r} 
-#| fig.width = 8,
-#| fig.height = 5,
+#| fig.height = 7.5,
+#| fig.width = 10,
+#| label = fig-proj,
 #| fig.align = "center",
-#| fig.cap = "sthis sdfaksdlf"
-get_proj <- function(data){
-  quantiles <- map(1:4, ~data |>
-                     filter(dim == .x) |>
-                     pull(index_val) |>
-                     quantile(seq(0, 1, 0.5), type = 3) |>
-                     rev() |>
-                     unname())
+#| fig.cap = "Projections found by the jellyfish and search better optimiser at each 10th quantile across 50 simulations.
+The holes index is used to find the target pipe shape in the 6/8/10/12D space.
+An `n_jellies = 100` and `max_tries = 100` are used for the jellyfish optimiser and `max.tries = 1000` is used for the search better optimiser."
 
-  res <- map2_dfr(
-    unlist(quantiles),
-    rep(1:4, each = 3),
-    ~ data |>
-      filter(dim == .y, index_val == .x) |>
-      select(dim, proj, index_val),
-    .id = "id"
-    ) |>
-    unnest(proj) |>
-    mutate(index_val = sprintf("%.3f", index_val),
-           optimiser = "jellyfish",
-           id = (as.numeric(id) - 1) %%3 + 1,
-           dim = as.factor(6 + (as.numeric(dim) - 1) * 2)) |>
-    mutate(id = case_when(id == 1 ~ "best",
-                          id == 2 ~ "mean",
-                          id == 3 ~ "worst",
-                          TRUE ~ "0thers"))
-  return(res)
+aaa <- map_dfr(
+  list(pipe1000, pipe1000_8d, pipe1000_10d, pipe1000_12d), 
+  ~{
+    d_dim = dim(.x)[2]
+    pipe_jellyfish |> 
+      filter(d == d_dim) |> 
+      arrange(index_val) |>
+      filter(row_number() %in% c(1, seq(5, 50, 5))) |> 
+      mutate(id = factor(
+        paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th"),
+        levels = paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th"))) |> 
+      compute_projection(data = .x, col = c("d", "index_val", "id"))
   }
+)
 
-p1 <- get_proj(pipe_jellyfish)  |>
-  ggplot(aes(x = V1, y = V2)) +
-  geom_point(size = 0.1) +
-  geom_text(data = ~. |> distinct(id, dim, index_val),
-           aes(label = index_val), x =0, y = 3.7, size = 3) +
-  facet_grid(dim ~ id) +
+p1 <- aaa |> 
+  ggplot() + 
+  geom_point(aes(x = V1, y = V2), size = 0.05) + 
+  geom_text(data = aaa |> select(-V1, -V2) |> unique(), 
+            aes(label = round(index_val,3)), x =0, y = 3.7, 
+            size = 3) + 
+  facet_grid(d ~ id) + 
   xlim(-4, 4) + ylim(-4, 4) +
@@
         panel.grid = element_blank())
-p2 <- p1 %+% get_proj(pipe_better)
-p1 + ggtitle("The Jellyfish Optimiser") | p2 + ggtitle("The Better Optimiser")
 
+bbb <- pipe_better |> 
+  group_by(dim) |> 
+  arrange(index_val) |> 
+  filter(row_number() %in% c(1, seq(5, 50, 5))) |>  
+  mutate(id = factor(
+    paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th"),
+    levels = paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th")), 
+    d= 4 + 2 * as.numeric(dim)) |> 
+  unnest(proj)
+
+p2 <- bbb |> 
+  ggplot() + 
+  geom_point(aes(x = V1, y = V2), size = 0.05) + 
+  geom_text(data = bbb |> select(-V1, -V2) |> unique(), 
+            aes(label = round(index_val,3)), x =0, y = 3.7, 
+            size = 3) + 
+  facet_grid(d ~ id) + 
+  xlim(-4, 4) + ylim(-4, 4) +
+  theme_bw() +
+  theme(aspect.ratio = 1, axis.ticks = element_blank(),
+        axis.text = element_blank(), axis.title = element_blank(),
+        panel.grid = element_blank())
+
+(p1 + ggtitle("The jellyfish optimiser"))/(p2 + ggtitle("The search better optimiser"))
 ```
-
-\newpage
 
@@
 
-<!-- # Bibliography styles -->
-
-<!-- Here are two sample references:  @Feynman1963118 @Dirac1953888. -->
-
-<!-- By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML.  -->
-
-<!-- You can sets extra options with `natbiboptions` variable in YAML header. Example  -->
-
-<!-- ``` -->
-
-<!-- natbiboptions: longnamesfirst,angle,semicolon -->
-
-<!-- ``` -->
-
-<!-- There are various more specific bibliography styles available at -->
-
-<!-- <https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>.  -->
-
-<!-- To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`. -->
-
-<!-- ## Using CSL  -->
-
-<!-- If `cite-method` is set to `citeproc` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`.
-In this case, the `csl` option is used to format the references.
-By default, this template will provide an appropriate style, but alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>.
-These can be downloaded -->
-
-<!-- and stored locally, or the url can be used as in the example header. -->
-
-<!-- # Equations -->
-
-<!-- Here is an equation: -->
-
-<!-- $$  -->
-
-<!--   f_{X}(x) = \left(\frac{\alpha}{\beta}\right) -->
-
-<!--   \left(\frac{x}{\beta}\right)^{\alpha-1} -->
-
-<!--   e^{-\left(\frac{x}{\beta}\right)^{\alpha}};  -->
-
-<!--   \alpha,\beta,x > 0 . -->
-
-<!-- $$ -->
-
-<!-- Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$ -->
-
-<!-- # Figures and tables -->
-
-<!-- @fig-meaningless is generated using an R chunk. -->
-
-<!-- ```{r} -->
-
-<!-- #| label: fig-meaningless -->
-
-<!-- #| fig-cap: A meaningless scatterplot -->
-
-<!-- #| fig-width: 5 -->
-
-<!-- #| fig-height: 5 -->
-
-<!-- #| fig-align: center -->
-
-<!-- #| out-width: 50% -->
-
-<!-- #| echo: false -->
-
-<!-- plot(runif(25), runif(25)) -->
-
-<!-- ``` -->
-
-<!-- # Tables coming from R -->
-
-<!-- Tables can also be generated using R chunks, as shown in @tbl-simple example. -->
-
-<!-- ```{r} -->
-
-<!-- #| label: tbl-simple -->
-
-<!-- #| tbl-cap: Caption centered above table -->
-
-<!-- #| echo: true -->
-
-<!-- knitr::kable(head(mtcars)[,1:4]) -->
-
-<!-- ``` -->
-
 # References

### 2024-06-06 optim.tex
working towards the example 1

@@
 
-\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, coltitle=black, colframe=quarto-callout-note-color-frame, opacityback=0, toprule=.15mm, opacitybacktitle=0.6, colback=white, bottomrule=.15mm, leftrule=.75mm, titlerule=0mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, rightrule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, titlerule=0mm, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, opacitybacktitle=0.6, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, opacityback=0, colback=white, toprule=.15mm, breakable, toptitle=1mm, left=2mm, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm]
 
@@
 
-\subsection{Going beyond 10D}\label{going-beyond-10d}
+\subsection{Performance of the jellyfish optimiser in high-dimensional
+pipe-finding
+problems}\label{performance-of-the-jellyfish-optimiser-in-high-dimensional-pipe-finding-problems}
 
-The pipe-finding problem is initially used to investigate indexes and
-optimisers in \citet{laa_using_2020}, and we extend it from a 6D problem
-to a 12D problem.
+{[}unfinished{]}
 
-Jellyfish optimiser, as a multi-start algorithm, is efficient in
-{[}\ldots{]} for high-dimensional problems
+The aim of this example is to investigate the performance of jellyfish
+optimisers in high-dimensional problems.
 
-\begin{figure}[H]
+The projection pursuit problem is to find the pipe shape, investigated
+in \citet{laa_using_2020}, using the holes index in 6/8/10/12-D data.
+For each case, 50 simulations are conducted for each combination of
+jellyfish hyper-parameter: \texttt{n\_jellies\ =\ 20/50/100} and
+\texttt{max\_tries\ =\ 50/100}. The performance of the jellyfish
+optimiser is also compared with an existing opitmiser: the search better
+opitmiser with \texttt{max.tries\ =\ 1000}.
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-7-1.pdf}
+Figure~\ref{fig-proportion} presents the proportion of jellyfishes that
+finds the index value close to (absolute value less than 0.05) the best
+index value found at each \texttt{max\_tries}/\texttt{n\_jellies}
+combination at \(d = 6/8/10/12\) data dimension. As the number of
+jellies and maximum tries increase, the proportion of jellyfishes that
+find the optimal projection increases. This suggests for higher
+dimensional problems (\(d = 8/10/12\)), an \texttt{n\_jellies\ =\ 100}
+and \texttt{max\_tries\ =\ 100} is necessary for at least half of the
+jellyfish optimiser to find the optimal projection.
+
+Figure~\ref{fig-proj} presents the final projection found across the 50
+simulations at each 10th quantile between the jellyfish and search
+better optimiser (\texttt{n\_jellies\ =\ 100},
+\texttt{max\_tries\ =\ 100} is used for the jellyfish optimiser). While
+the jellyfish optimiser may not always find the optimal projection in
+the worse case scenarios, especially as the dimension increases, it
+tends to find higher index values and crisper pipe shapes than the
+search better optimiser at each quantile across all the four data
+dimensions.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{optim_files/figure-pdf/fig-proportion-1.pdf}
 
@@
 
-\caption{sthis sdfaksdlf}
+\caption{\label{fig-proportion}Proportion of jellyfishes reaches close
+to the optimal index value in the pipe-finding problem. As the dimension
+goes up, the numer/ proportion of jellyfishes that reach the optimal
+index value decreases. Increasing the jellyfish hyperparameters
+(\texttt{n\_jellies} and \texttt{max\_tries}) help to improve the
+proportion.}
 
@@
 
-\begin{figure}[H]
+\begin{figure}
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-8-1.pdf}
+\centering{
+
+\includegraphics{optim_files/figure-pdf/fig-proj-1.pdf}
 
@@
 
-\caption{sthis sdfaksdlf}
+\caption{\label{fig-proj}Projections found by the jellyfish and search
+better optimiser at each 10th quantile across 50 simulations. The holes
+index is used to find the target pipe shape in the 6/8/10/12D space. An
+\texttt{n\_jellies\ =\ 100} and \texttt{max\_tries\ =\ 100} are used for
+the jellyfish optimiser and \texttt{max.tries\ =\ 1000} is used for the
+search better optimiser.}
 
 \end{figure}%
-
-\newpage
 

### 2024-06-06 code/01-simulation-holes.R

@@ -43,3 +43,11 @@ sim_res <- sim_setup |>
 
 # sim_pipe <- sim_res |> select(-alpha)
 # save(sim_pipe, file = "data/sim_pipe.rda")
+
+
+pipe_jellyfish <- sim_pipe |>
+  filter(n_jellies == 100, max_tries == 100) |>
+  get_best(group = id) |>
+  ungroup() |>
+  mutate(optimiser = "jellyfish")
+# save(pipe_jellyfish, file = "data/pipe_jellyfish.rda")

### 2024-06-06 code/08-process-data-for-modelling.R

@@ -37,16 +37,6 @@ pipe_setup_summ <- pipe_run_df |>
     time = mean(time)
   )
 
-
-
-pipe_setup |>
-  ggplot(aes( x= as.factor(n_jellies), y = P_J_hat,
-              group = d, color = as.factor(d))) +
-  geom_line() +
-  facet_wrap(vars(max_tries)) +
-  #facet_wrap(vars(d)) +
-  scale_color_brewer(palette = "Dark2")
-
 #############################################################################
 load(here::here("data/sim_sine_6d_dcor2d.rda"))
 load(here::here("data/sim_sine_6d.rda")) # loess

### 2024-06-10 optim.qmd
example 1 done

@@
 formalises several characteristics of projection pursuit indexes that
-are help to measure optimisaer performance. @sec-simulation describes a
+are help to measure optimiser performance. @sec-simulation describes a
 simulation study on performance of the jellyfish for several types of
@@
 
-The jellyfish optimiser has been implemented in the tourr package
-[@wickham_tourr_2011] and we will use the diagnostic plots proposed in
-the ferrn package [@RJ-2021-105] to visualise the optimisation process.
+[maybe a leading paragraph here to summarise the two examples]
 
-## Performance of the jellyfish optimiser in high-dimensional pipe-finding problems
+## Performance of jellyfish search optimiser in pipe-finding problems
 
-[unfinished]
-
-The aim of this example is to investigate the performance of jellyfish optimisers in high-dimensional problems.
-
-The projection pursuit problem is to find the pipe shape, investigated in @laa_using_2020, using the holes index in 6/8/10/12-D data.
-For each case, 50 simulations are conducted for each combination of jellyfish hyper-parameter: `n_jellies = 20/50/100` and `max_tries = 50/100`.
-The performance of the jellyfish optimiser is also compared with an existing opitmiser: the search better opitmiser with `max.tries = 1000`.
-
-@fig-proportion presents the proportion of jellyfishes that finds the index value close to (absolute value less than 0.05) the best index value found at each `max_tries`/`n_jellies` combination at $d = 6/8/10/12$ data dimension.
-As the number of jellies and maximum tries increase, the proportion of jellyfishes that find the optimal projection increases.
-This suggests for higher dimensional problems ($d = 8/10/12$), an `n_jellies = 100` and `max_tries = 100` is necessary for at least half of the jellyfish optimiser to find the optimal projection.
-
-@fig-proj presents the final projection found across the 50 simulations at each 10th quantile between the jellyfish and search better optimiser (`n_jellies = 100`, `max_tries = 100` is used for the jellyfish optimiser).
-While the jellyfish optimiser may not always find the optimal projection in the worse case scenarios, especially as the dimension increases, it tends to find higher index values and crisper pipe shapes than the search better optimiser at each quantile across all the four data dimensions.
-
-```{r}
-#| label = fig-proportion,
-#| fig.width = 6,
-#| fig.height = 3,
-#| fig.align = "center",
-#| fig.cap = "Proportion of jellyfishes reaches close to the optimal index value in the pipe-finding problem.
-As the dimension goes up, the numer/ proportion of jellyfishes that reach the optimal index value decreases.
-Increasing the jellyfish hyperparameters (`n_jellies` and `max_tries`) help to improve the proportion."
-pipe_run_summ <- sim_pipe |>
-  mutate(idx_f = "holes") |>
-  group_by(id, n_jellies, max_tries, d, loop) |> 
-  summarise(I_max = max(index_val))
-
-pipe_sim_best <- pipe_run_summ |> 
-  ungroup(loop, id) |> 
-  summarise(I_best = max(I_max))
-
-pipe_res <- pipe_run_summ |> 
-  left_join(pipe_sim_best) |> 
-  mutate(diff = abs(I_max - I_best) < 0.05) |> 
-  ungroup(id, loop) |> 
-  summarise(proportion = sum(diff) / n()) |> 
-  mutate(d = as.factor(d), n_jellies = as.factor(n_jellies))
-
-pipe_res |> 
-  ggplot(aes(x = n_jellies, y = proportion, 
-             group = interaction(d, max_tries),
-             color = d)) + 
-  geom_line() + 
-  scale_color_brewer(palette = "Dark2") +
-  facet_wrap(vars(max_tries), 
-             labeller = "label_both") + 
-  theme_bw()
-
-```
+The jellyfish search optimiser introduces a new swarm-based search strategy to optimise the projection pursuit problems.
+In this example, the performance of jellyfish search optimiser is investigated in two folds: 1) compare the jellyfish search optimiser with a commonly used optimiser: the search better optimiser [@RJ-2021-105; @laa_using_2020], and 2) examine the jellyfish search optimiser under different hyper-parameters: the number of jellyfish (20, 50, and 100) and the maximum number of tries (50 and 100).
+The projection pursuit problem investigated is finding the pipe shape using the holes index in 6, 8, 10, and 12-dimensional data, as investigated by @laa_using_2020.
 
@@
 #| fig.align = "center",
-#| fig.cap = "Projections found by the jellyfish and search better optimiser at each 10th quantile across 50 simulations.
-The holes index is used to find the target pipe shape in the 6/8/10/12D space.
-An `n_jellies = 100` and `max_tries = 100` are used for the jellyfish optimiser and `max.tries = 1000` is used for the search better optimiser."
+#| fig.cap = "Projections found by the jellyfish and search better optimisers at each 10th quantile across 50 simulations.
+The projection pursuit problem is to find the pipe shape using the holes index in the 6, 8, 10, and 12-dimensional spaces.
+The jellyfish optimiser uses 100 jellyfish and 100 maximum number of tries, while the search better optimiser uses a maximum 1000 tries in each step of random sampling.
+The index value found by the jellyfish optimiser finds the pipe shape across all simulations in the 6-dimensional space and reaches a higher index value and clearer pipe shape in higher dimensions than the search better optimiser."
 
@@
 ```
+
+```{r}
+#| label = fig-proportion,
+#| fig.width = 6,
+#| fig.height = 3,
+#| fig.align = "center",
+#| fig.cap = "Proportion of jellyfishes reaches near-optimal index values in the pipe-finding problems.
+The proportion is calculated as the number of jellyfish, out of the total number of jellyfish across 50 simulations, that have an index value within 0.05 of the best jellyfish's index value across all simulations.
+As the dimensionality increases, the proportion of jellyfishes reaching the optimal index value decreases.
+Increasing the jellyfish hyperparameters (number of jellyfish` and maximum number of tries) improves the proportion."
+pipe_run_summ <- sim_pipe |>
+  mutate(idx_f = "holes") |>
+  group_by(id, n_jellies, max_tries, d) |> 
+  summarise(I_max = max(index_val))
+
+pipe_sim_best <- pipe_run_summ |> 
+  group_by(n_jellies, max_tries, d) |>
+  summarise(I_best = max(I_max))
+
+pipe_res <- pipe_run_summ |> 
+  left_join(pipe_sim_best) |> 
+  mutate(diff = abs(I_max - I_best) < 0.05) |> 
+  group_by(n_jellies, max_tries, d) |> 
+  summarise(proportion = sum(diff) / n()) |> 
+  mutate(d = as.factor(d), n_jellies = as.factor(n_jellies))
+
+pipe_res |> 
+  rename(`max # of tries` = max_tries) |>
+  ggplot(aes(x = n_jellies, y = proportion, 
+             group = interaction(d, `max # of tries`),
+             color = d)) + 
+  geom_line() + 
+  scale_color_brewer(palette = "Dark2", name = "dimension") +
+  facet_wrap(vars(`max # of tries`), 
+             labeller = "label_both") + 
+  theme_bw() + 
+  xlab("Number of jellyfish") + 
+  ylab("Proportion of success")
+
+```
+
+In the first experiment, fifty simulations for both the jellyfish and the search better optimiser, across the four dimensions and @fig-proj presents the final projections found by the two optimisers, broken down by 10th quantile.
+The jellyfish optimiser uses 100 number of jellyfishes and 100 maximum number of tries, while a maximum of 1000 samples are allowed before the search better optimiser terminates, at each iteration of sampling.
+In the 6-dimensional data scenario, the jellyfish optimiser finds the clear pipe shape at all the 50 simulations, while the shape found by the search better optimiser begin to blur in worse case scenarios.
+As the dimension increases, while the jellyfish optimiser may not find the pipe shape due to random sampling, the pipe shape can be clearly seen more than 50% of the time across all dimensions.
+Compared with the search better optimiser, the jellyfish optimiser achieves higher index values and clearer pipe shapes.
+
+In the second experiment, fifty simulations are conducted for each combination of jellyfish hyper-parameters:  50 and 100 maximum number of tries and 20, 50, and 100 jellyfishes to understand the effect of hyper-parameters on the performance of the jellyfish optimiser.
+A optimisation run is considered successfully finding the target shape if its final index value is within 0.05 of the best index value found under the same scenario across all simulations.
+For example, in the 8-dimensional data scenario with 50 maximum number of tries, the best optimiser achieved an index value of `r pipe_sim_best |> filter(n_jellies == 50, max_tries == 50, d == 8) |> pull(I_best) |> round(3)`.
+Out of the 50 simulations, `r pipe_run_summ |> filter(n_jellies == 50, max_tries == 50, d == 8) |> filter(I_max >= 0.997 - 0.05) |> nrow() -> aa; aa` found an index value within 0.05 of the best value, resulting in a successful rate of `r pipe_run_summ |> filter(n_jellies == 50, max_tries == 50, d == 8) |> mutate(a = I_max >= 0.997 - 0.05) |> pull(a) -> bb; round(sum(bb)/length(bb), 3)` (`r aa`/ 50).
+@fig-proportion presents the proportion of successful optimiser for each scenario.
+As the number of jellyfish and maximum tries increase, the proportion of jellyfish finding the optimal projection also increases.
+For simpler problems (6 dimensions), small parameter values can result in high successful rate, while for higher-dimensional problems (8, 10, and 12 dimensions), a combination of 100 jellyfishes and 100 maximum number of tries is necessary for at least half of the jellyfishes to find the optimal projection.
 

### 2024-06-10 optim.tex
example 1 done

--- 2024-06-03 optim.tex
@@
 jellyfish optimiser and formalises several characteristics of projection
-pursuit indexes that are help to measure optimisaer performance.
+pursuit indexes that are help to measure optimiser performance.
 Section~\ref{sec-simulation} describes a simulation study on performance
@@
 
-\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, coltitle=black, colframe=quarto-callout-note-color-frame, opacityback=0, toprule=.15mm, opacitybacktitle=0.6, colback=white, bottomrule=.15mm, leftrule=.75mm, titlerule=0mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, rightrule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, colback=white, toprule=.15mm, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, rightrule=.15mm, coltitle=black, breakable, opacityback=0, opacitybacktitle=0.6, bottomtitle=1mm, arc=.35mm, leftrule=.75mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame]
 
@@
 
-The jellyfish optimiser has been implemented in the tourr package
-\citep{wickham_tourr_2011} and we will use the diagnostic plots proposed
-in the ferrn package \citep{RJ-2021-105} to visualise the optimisation
-process.
+{[}maybe a leading paragraph here to summarise the two examples{]}
 
-\subsection{Going beyond 10D}\label{going-beyond-10d}
+\subsection{Performance of jellyfish search optimiser in pipe-finding
+problems}\label{performance-of-jellyfish-search-optimiser-in-pipe-finding-problems}
 
-The pipe-finding problem is initially used to investigate indexes and
-optimisers in \citet{laa_using_2020}, and we extend it from a 6D problem
-to a 12D problem.
+The jellyfish search optimiser introduces a new swarm-based search
+strategy to optimise the projection pursuit problems. In this example,
+the performance of jellyfish search optimiser is investigated in two
+folds: 1) compare the jellyfish search optimiser with a commonly used
+optimiser: the search better optimiser
+\citep{RJ-2021-105, laa_using_2020}, and 2) examine the jellyfish search
+optimiser under different hyper-parameters: the number of jellyfish (20,
+50, and 100) and the maximum number of tries (50 and 100). The
+projection pursuit problem investigated is finding the pipe shape using
+the holes index in 6, 8, 10, and 12-dimensional data, as investigated by
+\citet{laa_using_2020}.
 
-Jellyfish optimiser, as a multi-start algorithm, is efficient in
-{[}\ldots{]} for high-dimensional problems
+\begin{figure}
 
-\begin{figure}[H]
+\centering{
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-7-1.pdf}
+\includegraphics{optim_files/figure-pdf/fig-proj-1.pdf}
 
@@
 
-\caption{sthis sdfaksdlf}
+\caption{\label{fig-proj}Projections found by the jellyfish and search
+better optimisers at each 10th quantile across 50 simulations. The
+projection pursuit problem is to find the pipe shape using the holes
+index in the 6, 8, 10, and 12-dimensional spaces. The jellyfish
+optimiser uses 100 jellyfish and 100 maximum number of tries, while the
+search better optimiser uses a maximum 1000 tries in each step of random
+sampling. The index value found by the jellyfish optimiser finds the
+pipe shape across all simulations in the 6-dimensional space and reaches
+a higher index value and clearer pipe shape in higher dimensions than
+the search better optimiser.}
 
@@
 
-\begin{figure}[H]
+\begin{figure}
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-8-1.pdf}
+\centering{
+
+\includegraphics{optim_files/figure-pdf/fig-proportion-1.pdf}
 
@@
 
-\caption{sthis sdfaksdlf}
+\caption{\label{fig-proportion}Proportion of jellyfishes reaches
+near-optimal index values in the pipe-finding problems. The proportion
+is calculated as the number of jellyfish, out of the total number of
+jellyfish across 50 simulations, that have an index value within 0.05 of
+the best jellyfish's index value across all simulations. As the
+dimensionality increases, the proportion of jellyfishes reaching the
+optimal index value decreases. Increasing the jellyfish hyperparameters
+(number of jellyfish` and maximum number of tries) improves the
+proportion.}
 
@@
 
-\newpage
+In the first experiment, fifty simulations for both the jellyfish and
+the search better optimiser, across the four dimensions and
+Figure~\ref{fig-proj} presents the final projections found by the two
+optimisers, broken down by 10th quantile. The jellyfish optimiser uses
+100 number of jellyfishes and 100 maximum number of tries, while a
+maximum of 1000 samples are allowed before the search better optimiser
+terminates, at each iteration of sampling. In the 6-dimensional data
+scenario, the jellyfish optimiser finds the clear pipe shape at all the
+50 simulations, while the shape found by the search better optimiser
+begin to blur in worse case scenarios. As the dimension increases, while
+the jellyfish optimiser may not find the pipe shape due to random
+sampling, the pipe shape can be clearly seen more than 50\% of the time
+across all dimensions. Compared with the search better optimiser, the
+jellyfish optimiser achieves higher index values and clearer pipe
+shapes.
+
+In the second experiment, fifty simulations are conducted for each
+combination of jellyfish hyper-parameters: 50 and 100 maximum number of
+tries and 20, 50, and 100 jellyfishes to understand the effect of
+hyper-parameters on the performance of the jellyfish optimiser. A
+optimisation run is considered successfully finding the target shape if
+its final index value is within 0.05 of the best index value found under
+the same scenario across all simulations. For example, in the
+8-dimensional data scenario with 50 maximum number of tries, the best
+optimiser achieved an index value of 0.997. Out of the 50 simulations,
+46 found an index value within 0.05 of the best value, resulting in a
+successful rate of 0.92 (46/ 50). Figure~\ref{fig-proportion} presents
+the proportion of successful optimiser for each scenario. As the number
+of jellyfish and maximum tries increase, the proportion of jellyfish
+finding the optimal projection also increases. For simpler problems (6
+dimensions), small parameter values can result in high successful rate,
+while for higher-dimensional problems (8, 10, and 12 dimensions), a
+combination of 100 jellyfishes and 100 maximum number of tries is
+necessary for at least half of the jellyfishes to find the optimal
+projection.
 

### 2024-06-10 optim.qmd
edit application 1

@@
 ---
-title: Studying the Performance of the Jellyfish Optimiser for the Application of Projection Pursuit
+title: Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit
 header-includes:
@@
 author:
-  - name: Alice Anonymous
-    email: alice@example.com
+  - name: H. Sherry Zhang
+    email: huize.zhang@austin.utexas.edu
     affiliations: 
-        - id: some-tech
-          name: Some Institute of Technology
-          department: Department Name
-          address: Street Address
-          city: City
-          state: State
-          postal-code: Postal Code
+        - id: 1
+          name: University of Texas at Austin
+          department: Department of Statistics and Data Sciences
+          city: Austin
+          country: United States
+          postal-code: 78751
     attributes:
         corresponding: true
-  - name: Bob Security
-    email: bob@example.com
+  - name: Dianne Cook
+    email: dicook@monash.edu
     affiliations:
-        - id: another-u
-          name: Another University
-          department: Department Name
-          address: Street Address
-          city: City
-          state: State
-          postal-code: Postal Code
-  - name: Cat Memes
-    email: cat@example.com
+        - id: 2
+          name: Monash University
+          department: Department of Econometrics and Business Statistics
+          city: Melbourne
+          country: Australia
+          postal-code: 3800
+  - name: Nicolas  Langrené
+    email: nicolaslangrene@uic.edu.cn
     affiliations:
-        - ref: another-u
-  - name: Derek Zoolander
-    email: derek@example.com
-    affilations:
-        - ref: some-tech
+        - id: 3
+          name: BNU-HKBU United International College
+          department: Department of Mathematical Sciences
+          city: Zhuhai
+          country: China
+          postal-code: 519087
+  - name: Jessica Leung
+    email: Jessica.Leung@monash.edu
+    affiliations:
+        - id: 2
+          name: Monash University
+          department: Department of Econometrics and Business Statistics
+          city: Melbourne
+          country: Australia
+          postal-code: 3800
 abstract: |
@@
 
-*Let's use British English ("American or British usage is accepted, but
+**README**:
+
+* British English ("American or British usage is accepted, but
 not a mixture of these")*
+
+* No "we" - always third person
+
+* Check if the affiliation information is correct
+
+* word use:
+
+  * Use jellyfish search optimiser, or JS optimiser, not jellyfish optimiser (my bad for leading the way)
+  * "hyper-parameter" rather than "hyperparameter"
 
@@
 
-[maybe a leading paragraph here to summarise the two examples]
+The JS optimiser provides a new swarm-based search strategy to optimise projection pursuit problems.
+Two examples are provided in this section to demonstrate its performance.
+The first example compares its performance with the search better optimiser and explores how it behaves with different combinations of hyper-parameter.
+The second example studies the effect of optimisation properties defined in @sec-theory, along with JS hyper-parameters, on the outcome of the optimisation.
 
-## Performance of jellyfish search optimiser in pipe-finding problems
+## Performance of JS optimiser in pipe-finding problems
 
-The jellyfish search optimiser introduces a new swarm-based search strategy to optimise the projection pursuit problems.
-In this example, the performance of jellyfish search optimiser is investigated in two folds: 1) compare the jellyfish search optimiser with a commonly used optimiser: the search better optimiser [@RJ-2021-105; @laa_using_2020], and 2) examine the jellyfish search optimiser under different hyper-parameters: the number of jellyfish (20, 50, and 100) and the maximum number of tries (50 and 100).
-The projection pursuit problem investigated is finding the pipe shape using the holes index in 6, 8, 10, and 12-dimensional data, as investigated by @laa_using_2020.
+The performance of JS optimiser is investigated in two folds: 1) compare it with a commonly used optimiser: the search better optimiser [@RJ-2021-105; @laa_using_2020], and 2) examine its success rate under different hyper-parameters (number of jellyfishes and the maximum number of tries).
+The projection pursuit problem used is finding the pipe shape using the holes index, investigated by @laa_using_2020.
 
@@
 The projection pursuit problem is to find the pipe shape using the holes index in the 6, 8, 10, and 12-dimensional spaces.
-The jellyfish optimiser uses 100 jellyfish and 100 maximum number of tries, while the search better optimiser uses a maximum 1000 tries in each step of random sampling.
-The index value found by the jellyfish optimiser finds the pipe shape across all simulations in the 6-dimensional space and reaches a higher index value and clearer pipe shape in higher dimensions than the search better optimiser."
+The JS optimiser uses 100 jellyfishes and a maximum number of tries of 100.
+The search better optimiser uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates.
+In the 6-D data space, the JS optimiser always finds a clear pipe shape while the search better optimiser also finds the pipe shape but with a wide rim.
+At higher data dimensions, the JS optimiser finds a higher index value and a clearer pipe shape across all the quantiles than the search better optimiser."
 
@@
 
-(p1 + ggtitle("The jellyfish optimiser"))/(p2 + ggtitle("The search better optimiser"))
+(p1 + ggtitle("The JS optimiser"))/(p2 + ggtitle("The search better optimiser"))
 ```
@@
 #| fig.align = "center",
-#| fig.cap = "Proportion of jellyfishes reaches near-optimal index values in the pipe-finding problems.
-The proportion is calculated as the number of jellyfish, out of the total number of jellyfish across 50 simulations, that have an index value within 0.05 of the best jellyfish's index value across all simulations.
-As the dimensionality increases, the proportion of jellyfishes reaching the optimal index value decreases.
-Increasing the jellyfish hyperparameters (number of jellyfish` and maximum number of tries) improves the proportion."
+#| fig.cap = "Proportion of simulations reaches near-optimal index values in the pipe-finding problem using the holes index.
+The proportion is calculated based on the number of simulations, out of 50, that achieve an index value within 0.05 of the best-performing simulation.
+As the dimensionality increases, the proportion of simulations reaching the optimal index value increases."
 pipe_run_summ <- sim_pipe |>
@@
 
-In the first experiment, fifty simulations for both the jellyfish and the search better optimiser, across the four dimensions and @fig-proj presents the final projections found by the two optimisers, broken down by 10th quantile.
-The jellyfish optimiser uses 100 number of jellyfishes and 100 maximum number of tries, while a maximum of 1000 samples are allowed before the search better optimiser terminates, at each iteration of sampling.
-In the 6-dimensional data scenario, the jellyfish optimiser finds the clear pipe shape at all the 50 simulations, while the shape found by the search better optimiser begin to blur in worse case scenarios.
-As the dimension increases, while the jellyfish optimiser may not find the pipe shape due to random sampling, the pipe shape can be clearly seen more than 50% of the time across all dimensions.
-Compared with the search better optimiser, the jellyfish optimiser achieves higher index values and clearer pipe shapes.
+Fifty (50) simulations are conducted with both JS and the search better optimiser, in four dimensions ($d = 6, 8, 10, 12$).
+The JS optimiser uses 100 jellyfishes and 100 maximum number of tries, while the search better optimiser allows a maximum of 1000 samples at each iteration before the algorithm terminates.
+@fig-proj presents the final projections found by the two optimisers, broken down by 10th quantile, faceted by the data dimension.
+In the 6-dimensional data scenario, the JS optimiser consistently identifies a clear pipe shape.
+The search better optimiser, although also find the pipe shape but with a wide rim, suggesting a further polish search may be required.
+With increasing dimensions, the JS optimiser may not always identify the pipe shape due to random sampling, it still presents a pipe shape in over 50% of cases.
+When compared to the search better optimiser, the JS optimiser reaches higher index values and clearer pipe shapes across all quantiles.
 
-In the second experiment, fifty simulations are conducted for each combination of jellyfish hyper-parameters:  50 and 100 maximum number of tries and 20, 50, and 100 jellyfishes to understand the effect of hyper-parameters on the performance of the jellyfish optimiser.
-A optimisation run is considered successfully finding the target shape if its final index value is within 0.05 of the best index value found under the same scenario across all simulations.
-For example, in the 8-dimensional data scenario with 50 maximum number of tries, the best optimiser achieved an index value of `r pipe_sim_best |> filter(n_jellies == 50, max_tries == 50, d == 8) |> pull(I_best) |> round(3)`.
-Out of the 50 simulations, `r pipe_run_summ |> filter(n_jellies == 50, max_tries == 50, d == 8) |> filter(I_max >= 0.997 - 0.05) |> nrow() -> aa; aa` found an index value within 0.05 of the best value, resulting in a successful rate of `r pipe_run_summ |> filter(n_jellies == 50, max_tries == 50, d == 8) |> mutate(a = I_max >= 0.997 - 0.05) |> pull(a) -> bb; round(sum(bb)/length(bb), 3)` (`r aa`/ 50).
-@fig-proportion presents the proportion of successful optimiser for each scenario.
-As the number of jellyfish and maximum tries increase, the proportion of jellyfish finding the optimal projection also increases.
-For simpler problems (6 dimensions), small parameter values can result in high successful rate, while for higher-dimensional problems (8, 10, and 12 dimensions), a combination of 100 jellyfishes and 100 maximum number of tries is necessary for at least half of the jellyfishes to find the optimal projection.
+In the second experiment, fifty (50) simulations are conducted at each  hyper-parameter combination to analyse its effects on the success rate of the JS optimiser.
+The hyper-parameters tested include: 1) 20, 50, and 100 jellyfishes, and 2) 50 and 100 maximum number of tries.
+In each simulation,  success is defined as achieving a final index value within 0.05 of the best index value found among all 50 simulations.
+For example, in the 8-dimensional data scenario with 20 jellyfishes and a maximum number of tries set to 100, the optimal index value is `r pipe_sim_best |> filter(n_jellies == 20, max_tries == 100, d == 8) |> pull(I_best) |> round(3) -> aa; aa`.
+Out of the 50 simulations, `r pipe_run_summ |> filter(n_jellies == 20, max_tries == 100, d == 8) |> filter(I_max >= aa - 0.05) |> nrow() -> bb; bb` achieved an index value within [`r aa - 0.05`, `r aa`], resulting in a success rate of `r pipe_run_summ |> filter(n_jellies == 20, max_tries == 100, d == 8) |> mutate(a = I_max >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.
+@fig-proportion presents the proportion of success for each hyper-parameter combinations.
+As the number of jellyfish and maximum tries increase, the success rate also increases.
+For simpler problems (6 dimensions), small parameter values (20 jellyfishes and a max number of tries of 50) can already result in high success rate, while for higher-dimensional problems (8, 10, and 12 dimensions), a combination of 100 jellyfishes and 100 maximum number of tries is necessary for at least half of the jellyfishes to find the optimal projection.
 

### 2024-06-10 optim.tex
edit application 1

--- 2024-06-03 optim.tex
@@
 \hypersetup{
-  pdftitle={Studying the Performance of the Jellyfish Optimiser for the Application of Projection Pursuit},
-  pdfauthor={Alice Anonymous; Bob Security; Cat Memes; Derek Zoolander},
+  pdftitle={Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit},
+  pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Leung},
   pdfkeywords={projection pursuit, optimization, jellyfish
@@
 \begin{frontmatter}
-\title{Studying the Performance of the Jellyfish Optimiser for the
-Application of Projection Pursuit}
-\author[1]{Alice Anonymous%
+\title{Studying the Performance of the Jellyfish Search Optimiser for
+the Application of Projection Pursuit}
+\author[1]{H. Sherry Zhang%
 \corref{cor1}%
 }
- \ead{alice@example.com} 
-\author[2]{Bob Security%
+ \ead{huize.zhang@austin.utexas.edu} 
+\author[2]{Dianne Cook%
 %
 }
- \ead{bob@example.com} 
-\author[2]{Cat Memes%
+ \ead{dicook@monash.edu} 
+\author[3]{Nicolas Langrené%
 %
 }
- \ead{cat@example.com} 
-\author[]{Derek Zoolander%
+ \ead{nicolaslangrene@uic.edu.cn} 
+\author[2]{Jessica Leung%
 %
 }
- \ead{derek@example.com}
+ \ead{Jessica.Leung@monash.edu}
 
-\affiliation[1]{organization={Some Institute of Technology, Department
-Name},addressline={Street Address},city={City},postcode={Postal
-Code},postcodesep={}}
-\affiliation[2]{organization={Another University, Department
-Name},addressline={Street Address},city={City},postcode={Postal
-Code},postcodesep={}}
+\affiliation[1]{organization={University of Texas at Austin, Department
+of Statistics and Data Sciences},city={Austin},country={United
+States},countrysep={,},postcode={78751},postcodesep={}}
+\affiliation[2]{organization={Monash University, Department of
+Econometrics and Business
+Statistics},city={Melbourne},country={Australia},countrysep={,},postcode={3800},postcodesep={}}
+\affiliation[3]{organization={BNU-HKBU United International
+College, Department of Mathematical
+Sciences},city={Zhuhai},country={China},countrysep={,},postcode={519087},postcodesep={}}
 
@@
 
-\emph{Let's use British English (``American or British usage is
-accepted, but not a mixture of these'')}
+\textbf{README}:
+
+\begin{itemize}
+\item
+  British English (``American or British usage is accepted, but not a
+  mixture of these'')*
+\item
+  No ``we'' - always third person
+\item
+  Check if the affiliation information is correct
+\item
+  word use:
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Use jellyfish search optimiser, or JS optimiser, not jellyfish
+    optimiser (my bad for leading the way)
+  \item
+    ``hyper-parameter'' rather than ``hyperparameter''
+  \end{itemize}
+\end{itemize}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, coltitle=black, colframe=quarto-callout-note-color-frame, opacityback=0, toprule=.15mm, opacitybacktitle=0.6, colback=white, bottomrule=.15mm, leftrule=.75mm, titlerule=0mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, rightrule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, opacityback=0, opacitybacktitle=0.6, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, bottomtitle=1mm, breakable, toprule=.15mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, rightrule=.15mm, toptitle=1mm, coltitle=black, leftrule=.75mm, left=2mm, arc=.35mm]
 
@@
 
-The jellyfish optimiser has been implemented in the tourr package
-\citep{wickham_tourr_2011} and we will use the diagnostic plots proposed
-in the ferrn package \citep{RJ-2021-105} to visualise the optimisation
-process.
+The JS optimiser provides a new swarm-based search strategy to optimise
+projection pursuit problems. Two examples are provided in this section
+to demonstrate its performance. The first example compares its
+performance with the search better optimiser and explores how it behaves
+with different combinations of hyper-parameter. The second example
+studies the effect of optimisation properties defined in
+Section~\ref{sec-theory}, along with JS hyper-parameters, on the outcome
+of the optimisation.
 
-\subsection{Going beyond 10D}\label{going-beyond-10d}
+\subsection{Performance of JS optimiser in pipe-finding
+problems}\label{performance-of-js-optimiser-in-pipe-finding-problems}
 
-The pipe-finding problem is initially used to investigate indexes and
-optimisers in \citet{laa_using_2020}, and we extend it from a 6D problem
-to a 12D problem.
+The performance of JS optimiser is investigated in two folds: 1) compare
+it with a commonly used optimiser: the search better optimiser
+\citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
+under different hyper-parameters (number of jellyfishes and the maximum
+number of tries). The projection pursuit problem used is finding the
+pipe shape using the holes index, investigated by
+\citet{laa_using_2020}.
 
-Jellyfish optimiser, as a multi-start algorithm, is efficient in
-{[}\ldots{]} for high-dimensional problems
+\begin{figure}
 
-\begin{figure}[H]
+\centering{
 
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-7-1.pdf}
+\includegraphics{optim_files/figure-pdf/fig-proj-1.pdf}
 
@@
 
-\caption{sthis sdfaksdlf}
+\caption{\label{fig-proj}Projections found by the jellyfish and search
+better optimisers at each 10th quantile across 50 simulations. The
+projection pursuit problem is to find the pipe shape using the holes
+index in the 6, 8, 10, and 12-dimensional spaces. The JS optimiser uses
+100 jellyfishes and a maximum number of tries of 100. The search better
+optimiser uses a maximum of 1000 tries in each step of random sampling
+step before the algorithm terminates. In the 6-D data space, the JS
+optimiser always find a clear pipe shape while the search better
+optimiser also find the pipe shape but with a wide rim. At higher data
+dimension, the JS optimiser always find a higher index value and a
+clearer pipe shape across all the quantiles.}
 
@@
 
-\caption{sthis sdfaksdlf}
+\caption{\label{fig-proportion}Proportion of simulations reaches
+near-optimal index values in the pipe-finding problem using the holes
+index. The proportion is calculated based on the number of simulations,
+out of 50, that achieve an index value within 0.05 of the
+best-performing simulation. As the dimensionality increases, the
+proportion of simulations reaching the optimal index value increases.}
 
@@
 
-\newpage
+Fifty (50) simulations are conducted with both JS and the search better
+optimiser, in four dimensions (\(d = 6, 8, 10, 12\)). The JS optimiser
+uses 100 jellyfishes and 100 maximum number of tries, while the search
+better optimiser allows a maximum of 1000 samples at each iteration
+before the algorithm terminates. Figure~\ref{fig-proj} presents the
+final projections found by the two optimisers, broken down by 10th
+quantile, faceted by the data dimension. In the 6-dimensional data
+scenario, the JS optimiser consistently identifies a clear pipe shape.
+The search better optimiser, although also find the pipe shape but with
+a wide rim, suggesting a further polish search may be required. With
+increasing dimensions, the JS optimiser may not always identify the pipe
+shape due to random sampling, it still presents a pipe shape in over
+50\% of cases. When compared to the search better optimiser, the JS
+optimiser reaches higher index values and clearer pipe shapes across all
+quantiles.
+
+In the second experiment, fifty (50) simulations are conducted at each
+hyper-parameter combination to analyse its effects on the success rate
+of the JS optimiser. The hyper-parameters tested include: 1) 20, 50, and
+100 jellyfishes, and 2) 50 and 100 maximum number of tries. In each
+simulation, success is defined as achieving a final index value within
+0.05 of the best index value found among all 50 simulations. For
+example, in the 8-dimensional data scenario with 20 jellyfishes and a
+maximum number of tries set to 100, the optimal index value is 0.997.
+Out of the 50 simulations, 27 achieved an index value within {[}0.947,
+0.997{]}, resulting in a success rate of 0.54.
+Figure~\ref{fig-proportion} presents the proportion of success for each
+hyper-parameter combinations. As the number of jellyfish and maximum
+tries increase, the success rate also increases. For simpler problems (6
+dimensions), small parameter values (20 jellyfishes and a max number of
+tries of 50) can already result in high success rate, while for
+higher-dimensional problems (8, 10, and 12 dimensions), a combination of
+100 jellyfishes and 100 maximum number of tries is necessary for at
+least half of the jellyfishes to find the optimal projection.
 

### 2024-06-10 code/example1-high-d.R

@@ -18,12 +18,24 @@ historyarray2dt <- function(array, index_f){
 
 }
 
-sim <- function(d, optimiser = c("search_better", "search_jellyfish"), data_seed = 123456, optim_seed = seed){
+sim <- function(d, optimiser = c("search_better_random", "search_better", "search_jellyfish"),
+                data_seed = 123456, optim_seed = seed){
   set.seed(data_seed)
   pipe1000 <- pipeData(d, 1000) %>% scale() %>% as_tibble()
   colnames(pipe1000) <- paste0("V", 1:d)
 
-  if (optimiser == "search_better"){
+  if (optimiser == "search_better_random"){
+    res <- map_dfr(optim_seed, function(seed){
+      set.seed(seed)
+      spiral_better <- save_history(
+        pipe1000, guided_tour(
+          holes(), search_f = search_better_random, max.tries = 1000
+        )
+      )
+      historyarray2dt(spiral_better[,,-1], holes())
+    }, .id = "sim") |>
+      mutate(d = d)
+  } else if (optimiser == "search_better"){
     res <- map_dfr(optim_seed, function(seed){
       set.seed(seed)
       spiral_better <- save_history(
@@ -34,20 +46,7 @@ sim <- function(d, optimiser = c("search_better", "search_jellyfish"), data_seed
       historyarray2dt(spiral_better[,,-1], holes())
     }, .id = "sim") |>
       mutate(d = d)
-  } else if (optimiser == "search_jellyfish"){
-
-    res <- map_dfr(optim_seed, function(seed){
-      set.seed(seed)
-        animate_xy(pipe1000, guided_tour(
-          holes(), d = 2, n_jellies = 1000,
-          search_f =  search_jellyfish,
-          max.tries = 300, min.tries = 100
-        ), max_frames = 50)
-    }, .id = "sim") |>
-      mutate(d = d)
-
   }
-
   return(res)
 }
 
@@ -56,32 +55,34 @@ data <- map_dfr(c(6, 8, 10, 12), ~{
   tibble(x = list(pipeData(.x, 1000) %>% scale()))
 }) |>
   mutate(dim = c(6, 8, 10, 12))
-safe_sim <- safely(sim)
-#a <- map_dfr(c(6, 8, 10, 12), ~safe_sim(d = .x, optimiser = "search_better", optim_seed = seed), .id = "dim")
[…]

### 2024-06-10 code/07-calc-smoothness-squintability.R
scale holes index between 0-1

@@ -67,9 +67,24 @@ sq_sine_basis_df <- tibble::tibble(
     index, data = data, n = n, best = best, return_early = TRUE)
   ))
 
-sq_basis_df <- sq_holes_basis_df |> dplyr::bind_rows(sq_sine_basis_df)
+sq_basis_df <- sq_holes_basis_df |>
+  dplyr::bind_rows(sq_sine_basis_df) |>
+  unnest(basis_df) |>
+  group_by(n, index) |>
+  mutate(
+    holes = ifelse(
+      index == "holes",
+      (holes - min(holes, na.rm = TRUE))/
+        (max(holes, na.rm = TRUE) - min(holes, na.rm = TRUE)),
+      NA),
+    stringy = ifelse(
+      index == "stringy",
+      (stringy - min(stringy, na.rm = TRUE))/
+        (max(stringy, na.rm = TRUE) - min(stringy, na.rm = TRUE)),
+      NA))|>
+  nest(basis_df = c(id:stringy)) |>
+  ungroup()
 save(sq_basis_df, file = here::here("data/sq_basis_df.rda"))
-
 
 
 start_prms1 <- c(theta1 = 1, theta2 = 1, theta3 = 3, theta4 = 0)

### 2024-06-10 code/08-process-data-for-modelling.R

@@ -88,10 +88,10 @@ sine_setup_summ <- sine_run_df |>
     time = mean(time)
   )
 
-
 sim_summary <- pipe_setup_summ |>  bind_rows(sine_setup_summ) |>
   rename(index = idx_f) |>
   mutate(index = ifelse(index == "spline", "splines2d", index))
+save(sim_summary, file = here::here("data", "sim_summary.rda"))
 
 ############################################################################
 ############################################################################

### 2024-06-10 optim.qmd
edit application #2

@@
 
-# Application \[Di and Sherry\] {#sec-simulation}
+# Application \[Di and Sherry\] {#sec-app}
 
@@
 The first example compares its performance with the search better optimiser and explores how it behaves with different combinations of hyper-parameter.
-The second example studies the effect of optimisation properties defined in @sec-theory, along with JS hyper-parameters, on the outcome of the optimisation.
+The second example studies the effect of optimisation properties defined in @sec-theory, along with jellyfish hyper-parameters, on the outcome of the optimisation.
 
-## Performance of JS optimiser in pipe-finding problems
+## Performance of JS optimiser in pipe-finding problems {#sec-app-1}
 
@@
 
-## Connecting optimisation success with its properties and jellyfish parameters
+## Factors affecting JS optimiser success rate: optimisation properties and jellyfish hyper-parameters {#sec-app-2}
 
-The optimisation properties (smoothness, squintability, and speed) proposed in the @sec-theory provide numerical measures to characterise the difficulty of the optimisation for a given projection pursuit problem.
-This example investigates how these measures, combined with the jellyfish parameters, affect the success of the jellyfish optimisers and how they can guide the decisions made in hyperparameter tuning with the jellyfish optimiser.
-Simulations are conducted with different jellyfish optimisers to obtain the proportion of successful optimisation across a collection of scenarios, for which smoothness, squintability, and speed are calculated.
-A generalised linear model is used to link the proportion of success rate with the jellyfish parameters and the optimisation properties.
+The optimisation properties, including smoothness, squintability, and speed, offer numerical metrics to characterise the complexity of a projection pursuit optimisation problem.
+This example investigates how these metrics, alongside jellyfish hyper-parameters (the number of jellyfish and the maximum number of tries), affect the success of the JS optimiser.
+Simulations are conducted to obtain the success rate across various projection pursuit problems, characterised by shape-to-find, data dimension, and index function used, as well as different hyper-parameter combinations.
+Smoothness and squintability are calculated for each projection pursuit problem as outlined in @sec-theory.
+A generalised linear model is used to construct the relationship between the success rate, jellyfish hyper-parameters and optimisation properties.
 
-In addition to the pipe-finding problem, also investigated is the sine-wave finding problem in 6D and 8D spaces with six indices considered: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`.
-Combining with different jellyfish optimiser parameters (`n_jellies` and `max_tries`), a total of `r nrow(sim_df)` cases is produced, comprising of `r nrow(filter(sim_df, index == "holes"))` pipe-finding cases  and `r nrow(filter(sim_df, index != "holes"))` sine-wave finding cases.
-For each case, the jellyfish optimiser is run fifty (50) times and summary statistics are calculated on the best index value found across all 50 runs and the proportion of runs that find a close (with a difference less than 0.05) best index value (`P_J_hat`).
+In addition to the pipe-finding problem, a new shape, sine wave, is investigated in 6D and 8D spaces with six indices considered: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`.
+Combining with two jellyfish hyper-parameters, a total of `r nrow(sim_df)` cases is produced, comprising of `r nrow(filter(sim_df, index == "holes"))` pipe-finding cases  and `r nrow(filter(sim_df, index != "holes"))` sine-wave finding cases.
+For each case, the JS optimiser is run fifty (50) times and the success rate is calculated as in @sec-app-1.
+
+*Note: should we print out the 52 cases investigated somewhere, say appendex, since it is not full grid simulation (otherwise for the sine-wave problem there will be 6 indexes x 6 configurations = 36 cases for 6-D data and 2 indexes [MIC & TIC] x 6 configurations = 12 cases for 8-D data)?
+We only do a selection of 28 of them to save time/ computational resources.*
 
@@
 To calculate smoothness, three hundred random bases are simulated.
-Index values and projection distance to the optimal basis are calculated for each basis to fit a Gaussian process model.
+Index values and projection distance (to the optimal basis) are calculated for each random basis before fitting them into a Gaussian process model.
 For squintability, fifty random bases are sampled and interpolated to the optimal basis with a step size of 0.005.
-The interpolated bases are first binned with a bin width of 0.005 to average the index values before feeding into a 4-parameter scaled logistic function, estimated by non-linear least square, to obtain the squintability measure.
+For these interpolated bases, index values and projection distances to the optimal basis are calculated and averaged with a bin width of 0.005.
+The squintability measure is then calculated from fitting a 4-parameter scaled logistic function of index value against projection distance, estimated by non-linear least square.
 
-@tbl-smoothness-squintability presents the parameters estimated for each case from the Gaussian process (variance, range, smooth, and nugget) for smoothness and the scaled logistic function (theta1 to theta4, and squint) for squintability.
-The column "smooth" and "squint" are used as the smoothness and squintability measure.
+*Note: Shall I repeat the scaled logistic function with our parameterisation here? It can be handy to interpret theta1 - theta4 but they should've already been explained in the squintability section.*
+
+@tbl-smoothness-squintability presents the parameters estimated from the Gaussian process (variance, range, smooth, and nugget) and the scaled logistic function ($\theta_1$ to $\theta_4$) for each case.
+The column "smooth" is used as the smoothness measure and the column  "squint" is calculated as $\frac{\theta_1 \theta_2 \theta_3}{4}$ as squintability.
 The low squint value of MIC and TIC index is due to its convex shape of the index value against the projection distance, as shown in @fig-idx-proj-dist.
-Comparing to other concave shapes, the maximum first derivative happens at a smaller projection distance ($\theta2$), requiring the optimiser to get closer to the optimal to see a significant change in the index value, hence a small squintability measure.
-[TODO: the stringy squintability doens't make sense]
+Comparing to other concave shapes, the maximum first derivative happens at a smaller projection distance ($\theta_2$), requiring the optimiser to get closer to the optimal to see a significant change in the index value, hence a small squintability measure.
 
-@tbl-mod-data presents the processed data for modelling after augmenting the jellyfish parameters (`n_jellies` and `max_tries`) and the average time spent on each run (`time`) for each case.
-A generalised linear model is fitted with a binomial family and a logit link function to the data to investigate the relationship between the proportion of jellyfish success and jellyfish parameters and the optimisation properties.
-Pre-processing of the data include 1) scale the jellyfish parameters (`n_jellies` and `max_tries`) by 10 for interpretation, and 2) create a binary variable `long_time` to flag the cases with average time spent on each run greater than 30 seconds.
-@tbl-mod-output presents the model output with the estimated coefficients.
-The model suggests a positive relationship between the proportion of jellyfish success and the jellyfish parameters, `n_jellies` and `max_tries`, optimisation property parameter, `smoothness` and `squintability`, and a negative relationship with the average time spent on each run (`long_time`) and the data dimension (`d`).
-The variable `squintability` and `d` are significant, suggesting their importance over jellyfish parameters to the success of the optimisation.
+*Note: the stringy squintability doesn't make sense*
 
-<!-- @tbl-all-sim displays the first five rows after combining all calculations.  -->
+@tbl-mod-data combines all the variables calculated above (the JS optimiser success rate, jellyfish hyper-parameters, and optimisation properties) into one table.
+Three pre-processing steps are applied to the data: 1) scaling the jellyfish hyper-parameters by a factor of 10 for interpretation, 2) creating a new binary variable `long_time` to indicate cases with an average run time over 30 seconds, and 3) encoding the success rate for the stringy case as 0, since none of the 50 simulations identified the sine-wave shape.
+A generalised linear model with a binomial family and a logit link function is used to fit the data and @tbl-mod-output presents the model outputs.
+The model suggests that the success rate of the JS optimiser is positively associated with the two jellyfish hyper-parameters, as well as with smoothness and squintability.
+However, being flagged with long runtime and an increase of data dimension will reduce the success rate.
+The variable `squintability` and `dimension` are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation sucess.
 
@@
 #| label: tbl-smoothness-squintability
-#| tbl-cap: Parameters estimated from the Gaussian process (including variance, range, smoothness, and nugget) and scaled logistic function (theta1 to theta4) for the pipe-finding and sine-wave finding problems.
-The squint column is calculated as $\frac{\theta1 \theta2 \theta3}{4}$, as described in @sec-squintability.
-The "smooth" and "squint" column represent the smoothness and squintability measures.
-tibble(shape = c(rep("pipe", 4), rep("sine", 8)), smoothness) |>
+#| tbl-cap: Parameters estimated from the Gaussian process (including variance, range, smoothness, and nugget) and scaled logistic function ($\theta_1$ to $\theta_4$) for the pipe-finding and sine-wave finding problems.
+The "smooth" column and "squint" column, calculated as $\frac{\theta_1 \theta_2 \theta_3}{4}$, represent the smoothness and squintability measures.
+dt <- tibble(shape = c(rep("pipe", 4), rep("sine", 8)), smoothness) |>
   left_join(squintability) |> 
-  rename(d = n, smooth = smoothness) |> 
-  knitr::kable(digits = 3, booktabs = T) |> 
+  rename(d = n, smooth = smoothness)
+dt |> 
+  knitr::kable(digits = 3, booktabs = T, escape = FALSE, row.names = TRUE) |> 
   column_spec(1, border_left = TRUE) |> 
-  column_spec(c(3, 7, 12), border_right = TRUE) |> 
+  column_spec(c(3, 7, 12, 13), border_right = TRUE) |> 
+  kable_styling(latex_options = "hold_position", font_size = 7, full_width = FALSE)
+```
+
+```{r}
+#| label: tbl-mod-data
+#| tbl-cap: The first 7 rows of the datasets processed for modelling. 
+sim_df |> 
+  dplyr::select(index, d, smoothness, squintability, n_jellies, max_tries, 
+                -I_max_max, P_J_hat, time) |> 
+  arrange(index, d, n_jellies, max_tries) |> 
+  head(7) |> 
+  kable(digits = 3, col.names = c("Index", "D", "Smoothness", "Squintability", 
+                                  "Num. of jellyfish", "Max. Num. of tries", "Prop. of success", "Time"), 
+        booktabs = T) |> 
   kable_styling(latex_options = "hold_position", font_size = 7)
 ```
+
+```{r}
+#| label: tbl-mod-output
+#| tbl-cap: Model estimates of proportion of jellyfish success on optimisation properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function.
+The variable smoothness, squintability, number of jellyfish and maximum numer of tries are positively associated with the JS optimiser success rate while data dimension and being flagged as long runtime are negatively associated with teh success rate.
+The variable squintability and dimension are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.
+sim_df2 <- sim_df |>  
+  mutate(n_jellies = n_jellies/10,
+         max_tries = max_tries/10,
+         long_time = ifelse(time > 30, 1, 0),
+         P_J_hat = ifelse(index == "stringy", 0, P_J_hat))
+
+mod1 <- glm(P_J_hat ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
+broom::tidy(mod1) |> 
+  mutate(term = c("Intercept", "Smoothness", "Squintability", "dimension (d)", 
+                  "long time", "number of jellyfish", "maximum number of tries")) |>
+  kable(digits = 3) |> 
+  column_spec(c(1,5), border_right = TRUE) |> 
+  column_spec(1, border_left = TRUE)
+```
+
+# Conclusion \[Di and Sherry\] {#sec-conclusion}
+
+# Additional plot:
 
@@
   geom_line() +
-  facet_nested_wrap(~index + n, nrow = 3, labeller = label_both, scales = "free_y") + 
+  facet_nested_wrap(~index + n, nrow = 3, labeller = label_both) + 
   xlab("projection distance") + ylab("index value")
@@
 
-```{r}
-#| label: tbl-mod-data
-#| tbl-cap: The first few rows of the datasets processed for modelling.
-The smoothness and squintability variable are uniquely characterised by the index function used and the data dimension, and thus do not vary across `n_jellies` and `max_tries`.
-The variable `P_J_hat`, and time are calculated at each observation.
-sim_df |> 
-  dplyr::select(index, d, smoothness, squintability, n_jellies, max_tries, 
-                -I_max_max, P_J_hat, time) |> 
-  arrange(index, d, n_jellies, max_tries) |> 
-  head(7) |> 
-  kable(digits = 3)
-```
-
-```{r}
-#| label: tbl-mod-output
-#| tbl-cap: Model estimates of proportion of jellyfish success (`P_J_hat`) on optimisation properties (`smoothness`, `squintability`, and `d`) and jellyfish parameters (`long_time`, `n_jellies`, and `max_tries`) from the generalised linear model with a binomial family and a logit link function.
-The variable `long_time` is derived to flag the cases with average time spent on each run greater than 30 seconds and the variable `n_jellies` and `max_tries` are scaled by 10 for interpretation.
-sim_df2 <- sim_df |> 
-  mutate(n_jellies = n_jellies/10,
-         max_tries = max_tries/10,
-         long_time = ifelse(time > 30, 1, 0),
-         P_J_hat = ifelse(index == "stringy", 0, P_J_hat))
-
-mod1 <- glm(P_J_hat ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
-broom::tidy(mod1) |> 
-  mutate(term = ifelse(term == "(Intercept)", "intercept", term)) |>
-  kable(digits = 3) |> 
-  column_spec(c(1,5), border_right = TRUE) |> 
-  column_spec(1, border_left = TRUE)
-```
-
-# Conclusion \[Di and Sherry\] {#sec-conclusion}
-
 # References

### 2024-06-10 optim.tex
edit application #2

@@
 pursuit indexes that are help to measure optimiser performance.
-Section~\ref{sec-simulation} describes a simulation study on performance
-of the jellyfish for several types of data and index functions.
+\textbf{?@sec-simulation} describes a simulation study on performance of
+the jellyfish for several types of data and index functions.
 Section~\ref{sec-conclusion} summarises the work and provides
@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacityback=0, opacitybacktitle=0.6, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, bottomtitle=1mm, breakable, toprule=.15mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, rightrule=.15mm, toptitle=1mm, coltitle=black, leftrule=.75mm, left=2mm, arc=.35mm]
+\begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, coltitle=black, rightrule=.15mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, left=2mm, breakable, colbacktitle=quarto-callout-note-color!10!white, colback=white, toptitle=1mm, opacityback=0, opacitybacktitle=0.6, arc=.35mm, bottomtitle=1mm, leftrule=.75mm, colframe=quarto-callout-note-color-frame]
 
@@
 
-\section{Application {[}Di and Sherry{]}}\label{sec-simulation}
+\section{Application {[}Di and Sherry{]}}\label{sec-app}
 
@@
 studies the effect of optimisation properties defined in
-Section~\ref{sec-theory}, along with JS hyper-parameters, on the outcome
-of the optimisation.
+Section~\ref{sec-theory}, along with jellyfish hyper-parameters, on the
+outcome of the optimisation.
 
 \subsection{Performance of JS optimiser in pipe-finding
-problems}\label{performance-of-js-optimiser-in-pipe-finding-problems}
+problems}\label{sec-app-1}
 
@@
 step before the algorithm terminates. In the 6-D data space, the JS
-optimiser always find a clear pipe shape while the search better
-optimiser also find the pipe shape but with a wide rim. At higher data
-dimension, the JS optimiser always find a higher index value and a
-clearer pipe shape across all the quantiles.}
+optimiser always finds a clear pipe shape while the search better
+optimiser also finds the pipe shape but with a wide rim. At higher data
+dimensions, the JS optimiser finds a higher index value and a clearer
+pipe shape across all the quantiles than the search better optimiser.}
 
@@
 
-\subsection{Connecting optimisation success with its properties and
-jellyfish
-parameters}\label{connecting-optimisation-success-with-its-properties-and-jellyfish-parameters}
+\subsection{Factors affecting JS optimiser success rate: optimisation
+properties and jellyfish hyper-parameters}\label{sec-app-2}
 
-The optimisation properties (smoothness, squintability, and speed)
-proposed in the Section~\ref{sec-theory} provide numerical measures to
-characterise the difficulty of the optimisation for a given projection
-pursuit problem. This example investigates how these measures, combined
-with the jellyfish parameters, affect the success of the jellyfish
-optimisers and how they can guide the decisions made in hyperparameter
-tuning with the jellyfish optimiser. Simulations are conducted with
-different jellyfish optimisers to obtain the proportion of successful
-optimisation across a collection of scenarios, for which smoothness,
-squintability, and speed are calculated. A generalised linear model is
-used to link the proportion of success rate with the jellyfish
-parameters and the optimisation properties.
+The optimisation properties, including smoothness, squintability, and
+speed, offer numerical metrics to characterise the complexity of a
+projection pursuit optimisation problem. This example investigates how
+these metrics, alongside jellyfish hyper-parameters (the number of
+jellyfish and the maximum number of tries), affect the success of the JS
+optimiser. Simulations are conducted to obtain the success rate across
+various projection pursuit problems, characterised by shape-to-find,
+data dimension, and index function used, as well as different
+hyper-parameter combinations. Smoothness and squintability are
+calculated for each projection pursuit problem as outlined in
+Section~\ref{sec-theory}. A generalised linear model is used to
+construct the relationship between the success rate, jellyfish
+hyper-parameters and optimisation properties.
 
-In addition to the pipe-finding problem, also investigated is the
-sine-wave finding problem in 6D and 8D spaces with six indices
-considered: \texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC},
-\texttt{TIC}, \texttt{spline}, and \texttt{stringy}. Combining with
-different jellyfish optimiser parameters (\texttt{n\_jellies} and
-\texttt{max\_tries}), a total of 52 cases is produced, comprising of 24
-pipe-finding cases and 28 sine-wave finding cases. For each case, the
-jellyfish optimiser is run fifty (50) times and summary statistics are
-calculated on the best index value found across all 50 runs and the
-proportion of runs that find a close (with a difference less than 0.05)
-best index value (\texttt{P\_J\_hat}).
+In addition to the pipe-finding problem, a new shape, sine wave, is
+investigated in 6D and 8D spaces with six indices considered:
+\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
+\texttt{spline}, and \texttt{stringy}. Combining with two jellyfish
+hyper-parameters, a total of 52 cases is produced, comprising of 24
+pipe-finding cases and 28 sine-wave finding cases. For each case, the JS
+optimiser is run fifty (50) times and the success rate is calculated as
+in Section~\ref{sec-app-1}.
+
+\emph{Note: should we print out the 52 cases investigated somewhere, say
+appendex, since it is not full grid simulation (otherwise for the
+sine-wave problem there will be 6 indexes x 6 configurations = 36 cases
+for 6-D data and 2 indexes {[}MIC \& TIC{]} x 6 configurations = 12
+cases for 8-D data)? We only do a selection of 28 of them to save time/
+computational resources.}
 
@@
 Section~\ref{sec-squintability}. To calculate smoothness, three hundred
-random bases are simulated. Index values and projection distance to the
-optimal basis are calculated for each basis to fit a Gaussian process
-model. For squintability, fifty random bases are sampled and
-interpolated to the optimal basis with a step size of 0.005. The
-interpolated bases are first binned with a bin width of 0.005 to average
-the index values before feeding into a 4-parameter scaled logistic
-function, estimated by non-linear least square, to obtain the
-squintability measure.
+random bases are simulated. Index values and projection distance (to the
+optimal basis) are calculated for each random basis before fitting them
+into a Gaussian process model. For squintability, fifty random bases are
+sampled and interpolated to the optimal basis with a step size of 0.005.
+For these interpolated bases, index values and projection distances to
+the optimal basis are calculated and averaged with a bin width of 0.005.
+The squintability measure is then calculated from fitting a 4-parameter
+scaled logistic function of index value against projection distance,
+estimated by non-linear least square.
+
+\emph{Note: Shall I repeat the scaled logistic function with our
+parameterisation here? It can be handy to interpret theta1 - theta4 but
+they should've already been explained in the squintability section.}
 
 Table~\ref{tbl-smoothness-squintability} presents the parameters
-estimated for each case from the Gaussian process (variance, range,
-smooth, and nugget) for smoothness and the scaled logistic function
-(theta1 to theta4, and squint) for squintability. The column ``smooth''
-and ``squint'' are used as the smoothness and squintability measure. The
-low squint value of MIC and TIC index is due to its convex shape of the
+estimated from the Gaussian process (variance, range, smooth, and
+nugget) and the scaled logistic function (\(\theta_1\) to \(\theta_4\))
+for each case. The column ``smooth'' is used as the smoothness measure
+and the column ``squint'' is calculated as
+\(\frac{\theta_1 \theta_2 \theta_3}{4}\) as squintability. The low
+squint value of MIC and TIC index is due to its convex shape of the
 index value against the projection distance, as shown in
@@
 maximum first derivative happens at a smaller projection distance
-(\(\theta2\)), requiring the optimiser to get closer to the optimal to
+(\(\theta_2\)), requiring the optimiser to get closer to the optimal to
 see a significant change in the index value, hence a small squintability
-measure. {[}TODO: the stringy squintability doens't make sense{]}
+measure.
 
-Table~\ref{tbl-mod-data} presents the processed data for modelling after
-augmenting the jellyfish parameters (\texttt{n\_jellies} and
-\texttt{max\_tries}) and the average time spent on each run
-(\texttt{time}) for each case. A generalised linear model is fitted with
-a binomial family and a logit link function to the data to investigate
-the relationship between the proportion of jellyfish success and
-jellyfish parameters and the optimisation properties. Pre-processing of
-the data include 1) scale the jellyfish parameters (\texttt{n\_jellies}
-and \texttt{max\_tries}) by 10 for interpretation, and 2) create a
-binary variable \texttt{long\_time} to flag the cases with average time
-spent on each run greater than 30 seconds. Table~\ref{tbl-mod-output}
-presents the model output with the estimated coefficients. The model
-suggests a positive relationship between the proportion of jellyfish
-success and the jellyfish parameters, \texttt{n\_jellies} and
-\texttt{max\_tries}, optimisation property parameter,
-\texttt{smoothness} and \texttt{squintability}, and a negative
-relationship with the average time spent on each run
-(\texttt{long\_time}) and the data dimension (\texttt{d}). The variable
-\texttt{squintability} and \texttt{d} are significant, suggesting their
-importance over jellyfish parameters to the success of the optimisation.
+\emph{Note: the stringy squintability doesn't make sense}
+
+Table~\ref{tbl-mod-data} combines all the variables calculated above
+(the JS optimiser success rate, jellyfish hyper-parameters, and
+optimisation properties) into one table. Three pre-processing steps are
+applied to the data: 1) scaling the jellyfish hyper-parameters by a
+factor of 10 for interpretation, 2) creating a new binary variable
+\texttt{long\_time} to indicate cases with an average run time over 30
+seconds, and 3) encoding the success rate for the stringy case as 0,
+since none of the 50 simulations identified the sine-wave shape. A
+generalised linear model with a binomial family and a logit link
+function is used to fit the data and Table~\ref{tbl-mod-output} presents
+the model outputs. The model suggests that the success rate of the JS
+optimiser is positively associated with the two jellyfish
+hyper-parameters, as well as with smoothness and squintability. However,
+being flagged with long runtime and an increase of data dimension will
+reduce the success rate. The variable \texttt{squintability} and
+\texttt{dimension} are significant, suggesting their importance relative
+to jellyfish hyper-parameters in the optimisation sucess.
 
@@
 
-\begin{longtable}[t]{|>{}ll>{}r|rrr>{}r|rrrr>{}r|}
+\begin{longtable}[t]{|>{}ll>{}l|rrr>{}r|rrrr>{}r|>{}r|}
 
@@
 the Gaussian process (including variance, range, smoothness, and nugget)
-and scaled logistic function (theta1 to theta4) for the pipe-finding and
-sine-wave finding problems. The squint column is calculated as
-\(\frac{\theta1 \theta2 \theta3}{4}\), as described in
-Section~\ref{sec-squintability}. The ``smooth'' and ``squint'' column
-represent the smoothness and squintability measures.}
+and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
+pipe-finding and sine-wave finding problems. The ``smooth'' column and
+``squint'' column, calculated as
+\(\frac{\theta_1 \theta_2 \theta_3}{4}\), represent the smoothness and
+squintability measures.}
 
@@
 \toprule
-shape & index & d & variance & range & smooth & nugget & theta1 & theta2 & theta3 & theta4 & squint\\
+ & shape & index & d & variance & range & smooth & nugget & theta1 & theta2 & theta3 & theta4 & squint\\
 \midrule
-pipe & holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 & 1.001 & 0.860 & 3.368 & 0.823 & 0.725\\
-pipe & holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 & 1.001 & 0.869 & 3.264 & 0.811 & 0.710\\
-pipe & holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 & 1.000 & 0.885 & 3.151 & 0.806 & 0.697\\
-pipe & holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 & 1.000 & 0.878 & 3.345 & 0.806 & 0.734\\
-sine & MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 & 0.894 & 0.571 & 1.623 & -0.024 & 0.207\\
+1 & pipe & holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 & 0.991 & 0.860 & 3.368 & 0.172 & 0.718\\
+2 & pipe & holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 & 1.011 & 0.869 & 3.264 & 0.085 & 0.716\\
+3 & pipe & holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 & 1.011 & 0.885 & 3.151 & 0.153 & 0.705\\
+4 & pipe & holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 & 0.999 & 0.878 & 3.345 & 0.196 & 0.733\\
+5 & sine & MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 & 0.894 & 0.571 & 1.623 & -0.024 & 0.207\\
 \addlinespace
-sine & MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 & 0.932 & 0.328 & 1.314 & -0.030 & 0.100\\
-sine & TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 & 0.951 & 0.536 & 1.719 & -0.025 & 0.219\\
-sine & TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 & 0.945 & 0.564 & 1.723 & -0.027 & 0.230\\
-sine & dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019 & 0.679\\
-sine & loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 & 1.016 & 1.039 & 2.648 & 0.080 & 0.699\\
+6 & sine & MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 & 0.932 & 0.328 & 1.314 & -0.030 & 0.100\\
+7 & sine & TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 & 0.951 & 0.536 & 1.719 & -0.025 & 0.219\\
+8 & sine & TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 & 0.945 & 0.564 & 1.723 & -0.027 & 0.230\\
+9 & sine & dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019 & 0.679\\
+10 & sine & loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 & 1.016 & 1.039 & 2.648 & 0.080 & 0.699\\
 \addlinespace
-sine & splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 & 1.014 & 1.051 & 2.730 & -0.009 & 0.727\\
-sine & stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 & 1.011 & 0.011 & 254.734 & 0.727 & 0.711\\
+11 & sine & splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 & 1.014 & 1.051 & 2.730 & -0.009 & 0.727\\
+12 & sine & stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 & 1.040 & 0.011 & 254.733 & 0.170 & 0.731\\
 \bottomrule
@@
 \endgroup{}
+
+\begingroup\fontsize{7}{9}\selectfont
+
+\begin{longtable}[t]{lrrrrrrl}
+
+\caption{\label{tbl-mod-data}The first 7 rows of the datasets processed
+for modelling.}
+
+\tabularnewline
+
+\toprule
+Index & D & Smoothness & Squintability & Num. of jellyfish & Max. Num. of tries & Prop. of success & Time\\
+\midrule
+MIC & 6 & 2.394 & 0.207 & 20 & 50 & 0.12 & 2.479 secs\\
+MIC & 6 & 2.394 & 0.207 & 20 & 100 & 0.24 & 8.950 secs\\
+MIC & 6 & 2.394 & 0.207 & 50 & 50 & 0.52 & 5.651 secs\\
+MIC & 6 & 2.394 & 0.207 & 50 & 100 & 0.64 & 13.223 secs\\
+MIC & 6 & 2.394 & 0.207 & 100 & 50 & 0.76 & 19.453 secs\\
+\addlinespace
+MIC & 8 & 2.394 & 0.100 & 20 & 50 & 0.08 & 2.566 secs\\
+MIC & 8 & 2.394 & 0.100 & 20 & 100 & 0.08 & 4.960 secs\\
+\bottomrule
+
+\end{longtable}
+
+\endgroup{}
+
+\begin{longtable}[t]{|>{}lrrr>{}r|}
+
+\caption{\label{tbl-mod-output}Model estimates of proportion of
+jellyfish success on optimisation properties and jellyfish
+hyper-parameters from the generalised linear model with a binomial
+family and a logit link function. The variable smoothness,
+squintability, number of jellyfish and maximum numer of tries are
+positively associated with the JS optimiser success rate while data
+dimension and being flagged as long runtime are negatively associated
+with teh success rate. The variable squintability and dimension are
+significant, suggesting their importance relative to jellyfish
+hyper-parameters in the optimisation success.}
+
+\tabularnewline
+
+\toprule
+term & estimate & std.error & statistic & p.value\\
+\midrule
+Intercept & -4.849 & 4.496 & -1.078 & 0.281\\
+Smoothness & 1.672 & 1.527 & 1.094 & 0.274\\
+Squintability & 6.967 & 2.173 & 3.207 & 0.001\\
+dimension (d) & -0.588 & 0.257 & -2.292 & 0.022\\
+long time & -0.816 & 1.333 & -0.612 & 0.540\\
+\addlinespace
+number of jellyfish & 0.229 & 0.131 & 1.752 & 0.080\\
+maximum number of tries & 0.108 & 0.153 & 0.703 & 0.482\\
+\bottomrule
+
+\end{longtable}
+
+\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
+
+\section{Additional plot:}\label{additional-plot}
 
@@
 
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0811}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0405}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1486}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1892}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1351}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1351}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1081}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1622}}@{}}
-
-\caption{\label{tbl-mod-data}The first few rows of the datasets
-processed for modelling. The smoothness and squintability variable are
-uniquely characterised by the index function used and the data
-dimension, and thus do not vary across \texttt{n\_jellies} and
-\texttt{max\_tries}. The variable \texttt{P\_J\_hat}, and time are
-calculated at each observation.}
-
-\tabularnewline
-
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-index
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-d
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-smoothness
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-squintability
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-n\_jellies
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-max\_tries
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-P\_J\_hat
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-time
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-MIC & 6 & 2.394 & 0.207 & 20 & 50 & 0.12 & 2.479 secs \\
-MIC & 6 & 2.394 & 0.207 & 20 & 100 & 0.24 & 8.950 secs \\
-MIC & 6 & 2.394 & 0.207 & 50 & 50 & 0.52 & 5.651 secs \\
-MIC & 6 & 2.394 & 0.207 & 50 & 100 & 0.64 & 13.223 secs \\
-MIC & 6 & 2.394 & 0.207 & 100 & 50 & 0.76 & 19.453 secs \\
-MIC & 8 & 2.394 & 0.100 & 20 & 50 & 0.08 & 2.566 secs \\
-MIC & 8 & 2.394 & 0.100 & 20 & 100 & 0.08 & 4.960 secs \\
-
-\end{longtable}
-
-\begin{longtable}[t]{|>{}lrrr>{}r|}
-
-\caption{\label{tbl-mod-output}Model estimates of proportion of
-jellyfish success (\texttt{P\_J\_hat}) on optimisation properties
-(\texttt{smoothness}, \texttt{squintability}, and \texttt{d}) and
-jellyfish parameters (\texttt{long\_time}, \texttt{n\_jellies}, and
-\texttt{max\_tries}) from the generalised linear model with a binomial
-family and a logit link function. The variable \texttt{long\_time} is
-derived to flag the cases with average time spent on each run greater
-than 30 seconds and the variable \texttt{n\_jellies} and
-\texttt{max\_tries} are scaled by 10 for interpretation.}
-
-\tabularnewline
-
-\toprule
-term & estimate & std.error & statistic & p.value\\
-\midrule
-intercept & -4.672 & 4.512 & -1.036 & 0.300\\
-smoothness & 1.604 & 1.530 & 1.048 & 0.294\\
-squintability & 7.062 & 2.210 & 3.195 & 0.001\\
-d & -0.595 & 0.260 & -2.289 & 0.022\\
-long\_time & -0.851 & 1.348 & -0.632 & 0.528\\
-\addlinespace
-n\_jellies & 0.230 & 0.131 & 1.754 & 0.080\\
-max\_tries & 0.107 & 0.153 & 0.700 & 0.484\\
-\bottomrule
-
-\end{longtable}
-
-\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
-
 \renewcommand\refname{References}

### 2024-06-12 optim.qmd
fill in Section 4

@@
 load(here::here("data/sim_pipe.rda"))
+load(here::here("data/sim_sine_6d_spline_head.rda"))
+load(here::here("data/sim_sine_6d_spline_best.rda"))
+load(here::here("data/sim_sine_6d_spline_projdist.rda"))
 load(here::here("data/pipe_better.rda"))
@@
 
-# Visualisation of jellyfish optimiser
+# Visualisation of jellyfish optimiser {#sec-vis}
 
-Information of the jellyfish optimiser is available in a tabular format
-and below is an example data collected from finding the sine-wave
-structure in 6D data using a distance correlation index (`docr2d_2`):
+The data structure proposed in @RJ-2021-105 is implemented for JS optimisers and  additional information such as index function used, data dimension, and jellyfish hyper-parameters can be included as additional columns.
+Below prints an example data collected from finding the sine-wave structure in 6D data (`d = 6`) using a spline index function(`splines2d`):
 
 ```{r}
-load(here::here("data/sim_example.rda"))
+dplyr::glimpse(sim_sine_6d_spline_head |> mutate(index = "spline2d") |> select(-id, -info)) 
 ```
 
-```{r}
-dplyr::glimpse(sim_example |> select(-id)) 
-```
+The jellyfish hyper-parameters used are 50 jellyfishes (`n_jellies`) and a maximum of 50 tries (`max_tries`).
+The full data contains 50 simulations, recorded by the variable `sim`, with different seeds (`seed`).
+Recorded in columns `basis`, `index_val`, `tries`, `loop` and `time`, are the sampled projection basis, calculated index value, iteration ID, jellyfish ID, and time taken to find the optimum, respectively.
+The basis column records every visited basis before comparing with the current (last) basis.
+If a visited basis has a smaller index value than the current (last) one, the JS optimiser keeps the current basis for the next iteration in the search but still records the visited inferior basis.
+This data can be used to derive numerical summaries and plots:
 
-Information recorded can be categorised into the following categories:
+-   **success rate**:  Success rate is defined as the proportion of simulations that achieve a final index value within 0.05 of the best index value found among all 50 simulations.
+In the dataset above, the optimal index value is `r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`.
+Out of the 50 simulations, `r sim_sine_6d_spline_best |> filter(index_val >= aa - 0.05) |> nrow() -> bb; bb` achieved an index value within [`r aa - 0.05`, `r aa`], resulting in a success rate of `r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.
+@fig-projection shows the best projection found in each simulation, sorted by the index value.
+Most of the projections show a clear sine-wave structure, while the last six shows only a straight line with clusters at the end.
+When local optima exist in the projection pursuit problem, plotting the projections found in each simulation or by each jellyfish can help identify these local optima.
 
--   projection pursuit variables: the index function used (`idx_f`), the
-    data dimension (`d`)
--   jellyfish optimiser parameters: the number of jellies (`n_jellies`),
-    the maximum number of tries (`max_tries`)
--   simulation variables: the simulation number (`sim`), the seed used
-    (`seed`)
--   optimisation variables: the projection basis in a matrix format
-    (`basis`), the index value (`index_val`), a description of the
-    status - one of "initiation", "current_best", and "jellyfish_update"
-    (`info`), current iteration ID (`tries`), current jelly ID (`loop`),
-    and the time taken to find the optimum (`time`).
-
-The basis column records every basis *visited* by the jellyfish
-optimiser prior to comparing with the current basis. In each iteration,
-if the index value of a visited basis is smaller than that of the
-current one, the jellyfish optimiser will retain the current basis for
-the next iteration, while still documenting the visited basis.
-
-Numerical information to compute:
-
--   angular distance between the projection basis and the theoretical
-    best basis,
--   the proportion of simulation that found the optimal basis,
--   the proportion of jellies, within each simulation, that found the
-    optimal basis,
-
-Visualisation to inspect:
-
--   inspect the basis visited by each jellyfish in the reduced PCA
-    space,
--   inspect the final 2D projections reached by each jellyfish,
--   plot the index value against the angular distance between the
-    projection basis and the theoretical best basis
-
-Plotting the basis in the space and the projected data can help to
-understand 1) whether each simulation finds the same optimum or some
-simulations find local optima; and 2) whether the index function used
-can detect the structure in the data and the projection contains the
-structure of interest.
-
-The visualisations above can be faceted by the projection pursuit
-variables and jellyfish optimiser parameters to compare the performance
-of different indexes to detect the same structure and how the jellyfish
-optimiser parameters affect the optimisation process.
-
-\[example plots\]
+-  **Projection distance**: The projection distance between projection bases and the theoretical best basis (or the empirical best basis if the theoretical best is unknown) is calculated as the Frobenius norm, which measures how close a visited basis is to the best basis.
+Plotting this projection distance against the index value for each basis visited reveals how the index value changes as the basis approaches the optimal basis, helping to identify the squintability of the index function.
+@fig-index-value-proj-dist shows the index value against the projection distance for the given example and its binned version.
+As the projection distance decreases, the index value initially increases with an increasing slope and then level off.
+The index value shows greater variation at medium projection distances compared to small and large distances.
 
 ```{r}
-load(here::here("data/sine_dcor2d_best.rda"))
+#| label: fig-projection
+#| fig.height: 6
+#| fig.width: 10
+#| fig.cap: Best projections found by the JS opitmiser in each simulation, sorted by the index value.
+The projection pursuit problem is to detect the 6D sine-wave structure using the spline index.
+Most simulations have detected a clear sine-wave pattern, while the last six show a straight line with clustering at the ends..
+In this simulation, the success rate is 0.88 (44/50).
 set.seed(123456)
@@
 
-proj_dt <- map_dfr(sine_dcor2d_best |> pull(basis), 
-              ~as_tibble(as.matrix(sin1000) %*% .x), .id = "sim")
+proj_dt <- map_dfr(sim_sine_6d_spline_best |> pull(basis), 
+              ~as_tibble(as.matrix(sin1000) %*% .x), .id = "sim") |> 
+  mutate(sim = as.numeric(sim)) |> 
+  left_join(sim_sine_6d_spline_best |> select(sim, index_val), by = "sim")
 
-p1 <- proj_dt |>
+proj_dt |>
   ggplot(aes(x = V1, y = V2)) + 
   geom_point(size = 0.1) + 
+  geom_text(aes(label = round(index_val, 3)), x = 0, y = 3.7, 
+            size = 3) +
   xlim(-4, 4) + ylim(-4, 4) +
-  facet_wrap(~as.numeric(sim), ncol = 10) + 
+  facet_wrap(~fct_reorder(as.factor(sim), -index_val), ncol = 10) + 
   theme_bw() +
@@
 ```{r}
-#| fig.height: 5 
-#| fig.width: 10
-#| fig.cap: sdfjsflk
-p1 
-```
+#| label: fig-index-value-proj-dist
+#| fig.height: 3
+#| fig.width: 8
+#| fig.cap: "Projection distance plotted against the index value for detecting the 6D sine-wave structure using the spline index function.
+a) The index values start around 0 with low variance at large projection distances.
+As the projection distance decreases, the index values increase with a larger variance.
+Towards the minimum projection distance, the index values increase and cluster around 1, indicating the optimiser is approaching to the optimum.
+b)  Projection distances are binned with a bin width of 0.1, and index values are averaged over each bin.
+The relationship between the two variables forms a sigmoid curve."
+p1 <- sim_sine_6d_spline_projdist |> 
+  ggplot(aes(x = proj_dist, y = index_val)) +
+  geom_point(size = 0.1, alpha = 0.5) + 
+  labs(x = "Projection distance", y = "Index value")
 
-```{r eval = FALSE}
-theo_best <- matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE)
-angle_dcor2d <- sine_dcor2d_5050 |>
-    rowwise() |>
-    mutate(angle = tourr::proj_dist(theo_best, basis))
+p2 <- sim_sine_6d_spline_projdist |> 
+  group_by(proj_bin = proj_dist %/% 0.1) |> 
+  summarise(index_val = mean(index_val)) |> 
+  ggplot(aes(x = proj_bin, y = index_val)) + 
+  geom_line() +
+  geom_point() + 
+  labs(x = "Projection distance", y = "Index value")
 
-p2 <- angle_dcor2d |>
-  ggplot(aes(x = angle, y = index_val)) +
-  geom_point(size = 0.1) +
-  theme(aspect.ratio = 1)
-
+(p1 | p2) + plot_annotation(tag_levels = "a")
 ```
@@
 The hyper-parameters tested include: 1) 20, 50, and 100 jellyfishes, and 2) 50 and 100 maximum number of tries.
-In each simulation,  success is defined as achieving a final index value within 0.05 of the best index value found among all 50 simulations.
-For example, in the 8-dimensional data scenario with 20 jellyfishes and a maximum number of tries set to 100, the optimal index value is `r pipe_sim_best |> filter(n_jellies == 20, max_tries == 100, d == 8) |> pull(I_best) |> round(3) -> aa; aa`.
-Out of the 50 simulations, `r pipe_run_summ |> filter(n_jellies == 20, max_tries == 100, d == 8) |> filter(I_max >= aa - 0.05) |> nrow() -> bb; bb` achieved an index value within [`r aa - 0.05`, `r aa`], resulting in a success rate of `r pipe_run_summ |> filter(n_jellies == 20, max_tries == 100, d == 8) |> mutate(a = I_max >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.
+In each simulation,  success rate is calculated as defined in @sec-vis.
 @fig-proportion presents the proportion of success for each hyper-parameter combinations.
 As the number of jellyfish and maximum tries increase, the success rate also increases.
-For simpler problems (6 dimensions), small parameter values (20 jellyfishes and a max number of tries of 50) can already result in high success rate, while for higher-dimensional problems (8, 10, and 12 dimensions), a combination of 100 jellyfishes and 100 maximum number of tries is necessary for at least half of the jellyfishes to find the optimal projection.
+For simpler problems (6 dimensions), small parameter values (20 jellyfishes and a maximum number of tries of 50) can already result in high success rate, while for higher-dimensional problems (8, 10, and 12 dimensions), a combination of 100 jellyfishes and 100 maximum number of tries is necessary for at least half of the jellyfishes to find the optimal projection.
 
@@
 
-*Note: should we print out the 52 cases investigated somewhere, say appendex, since it is not full grid simulation (otherwise for the sine-wave problem there will be 6 indexes x 6 configurations = 36 cases for 6-D data and 2 indexes [MIC & TIC] x 6 configurations = 12 cases for 8-D data)?
+*Note: should we print out the 52 cases investigated somewhere, say appendix, since it is not full grid simulation (otherwise for the sine-wave problem there will be 6 indexes x 6 configurations = 36 cases for 6-D data and 2 indexes [MIC & TIC] x 6 configurations = 12 cases for 8-D data)?
 We only do a selection of 28 of them to save time/ computational resources.*
@@
 However, being flagged with long runtime and an increase of data dimension will reduce the success rate.
-The variable `squintability` and `dimension` are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation sucess.
+The variable `squintability` and `dimension` are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.
 
@@
 #| tbl-cap: Model estimates of proportion of jellyfish success on optimisation properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function.
-The variable smoothness, squintability, number of jellyfish and maximum numer of tries are positively associated with the JS optimiser success rate while data dimension and being flagged as long runtime are negatively associated with teh success rate.
+The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with the JS optimiser success rate while data dimension and being flagged as long runtime are negatively associated with the success rate.
 The variable squintability and dimension are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.

### 2024-06-12 optim.tex
fill in Section 4

@@
 
-\begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, coltitle=black, rightrule=.15mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, left=2mm, breakable, colbacktitle=quarto-callout-note-color!10!white, colback=white, toptitle=1mm, opacityback=0, opacitybacktitle=0.6, arc=.35mm, bottomtitle=1mm, leftrule=.75mm, colframe=quarto-callout-note-color-frame]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, toptitle=1mm, colback=white, colframe=quarto-callout-note-color-frame, titlerule=0mm, breakable, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, bottomrule=.15mm, arc=.35mm, rightrule=.15mm, leftrule=.75mm, opacityback=0]
 
@@
 
-\section{Visualisation of jellyfish
-optimiser}\label{visualisation-of-jellyfish-optimiser}
+\section{Visualisation of jellyfish optimiser}\label{sec-vis}
 
-Information of the jellyfish optimiser is available in a tabular format
-and below is an example data collected from finding the sine-wave
-structure in 6D data using a distance correlation index
-(\texttt{docr2d\_2}):
+The data structure proposed in \citet{RJ-2021-105} is implemented for JS
+optimisers and additional information such as index function used, data
+dimension, and jellyfish hyper-parameters can be included as additional
+columns. Below prints an example data collected from finding the
+sine-wave structure in 6D data (\texttt{d\ =\ 6}) using a spline index
+function(\texttt{splines2d}):
 
@@
 Rows: 10
-Columns: 13
-$ idx_f     <chr> "dcor2d_2", "dcor2d_2", "dcor2d_2", "dcor2d_2", "dcor2d_2", ~
+Columns: 12
+$ n_jellies <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50
+$ max_tries <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50
 $ d         <dbl> 6, 6, 6, 6, 6, 6, 6, 6, 6, 6
-$ n_jellies <dbl> 20, 20, 20, 20, 20, 20, 20, 20, 20, 20
-$ max_tries <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50
 $ sim       <int> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1
@@
 $ basis     <list> <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<m~
-$ index_val <dbl> 0.0247212373, 0.0033938502, 0.0463398915, 0.0486230801, -0.0~
-$ info      <chr> "initiation", "initiation", "initiation", "initiation", "ini~
+$ index_val <dbl> 0.036664532, 0.005121912, 0.058683104, 0.056527991, 0.015734~
 $ method    <chr> "search_jellyfish", "search_jellyfish", "search_jellyfish", ~
@@
 $ loop      <dbl> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10
-$ time      <drtn> 35.46031 secs, 35.46031 secs, 35.46031 secs, 35.46031 secs, ~
+$ time      <drtn> 4.362133 mins, 4.362133 mins, 4.362133 mins, 4.362133 mins, ~
+$ index     <chr> "spline2d", "spline2d", "spline2d", "spline2d", "spline2d", ~
 \end{verbatim}
 
-Information recorded can be categorised into the following categories:
+The jellyfish hyper-parameters used are 50 jellyfishes
+(\texttt{n\_jellies}) and a maximum of 50 tries (\texttt{max\_tries}).
+The full data contains 50 simulations, recorded by the variable
+\texttt{sim}, with different seeds (\texttt{seed}). Recorded in columns
+\texttt{basis}, \texttt{index\_val}, \texttt{tries}, \texttt{loop} and
+\texttt{time}, are the sampled projection basis, calculated index value,
+iteration ID, jellyfish ID, and time taken to find the optimum,
+respectively. The basis column records every visited basis before
+comparing with the current (last) basis. If a visited basis has a
+smaller index value than the current (last) one, the JS optimiser keeps
+the current basis for the next iteration in the search but still records
+the visited inferior basis. This data can be used to derive numerical
+summaries and plots:
 
 \begin{itemize}
-\tightlist
 \item
-  projection pursuit variables: the index function used
-  (\texttt{idx\_f}), the data dimension (\texttt{d})
+  \textbf{success rate}: Success rate is defined as the proportion of
+  simulations that achieve a final index value within 0.05 of the best
+  index value found among all 50 simulations. In the dataset above, the
+  optimal index value is 0.999. Out of the 50 simulations, 44 achieved
+  an index value within {[}0.949, 0.999{]}, resulting in a success rate
+  of 0.88. Figure~\ref{fig-projection} shows the best projection found
+  in each simulation, sorted by the index value. Most of the projections
+  show a clear sine-wave structure, while the last six shows only a
+  straight line with clusters at the end. When local optima exist in the
+  projection pursuit problem, plotting the projections found in each
+  simulation or by each jellyfish can help identify these local optima.
 \item
-  jellyfish optimiser parameters: the number of jellies
-  (\texttt{n\_jellies}), the maximum number of tries
-  (\texttt{max\_tries})
-\item
-  simulation variables: the simulation number (\texttt{sim}), the seed
-  used (\texttt{seed})
-\item
-  optimisation variables: the projection basis in a matrix format
-  (\texttt{basis}), the index value (\texttt{index\_val}), a description
-  of the status - one of ``initiation'', ``current\_best'', and
-  ``jellyfish\_update'' (\texttt{info}), current iteration ID
-  (\texttt{tries}), current jelly ID (\texttt{loop}), and the time taken
-  to find the optimum (\texttt{time}).
+  \textbf{Projection distance}: The projection distance between
+  projection bases and the theoretical best basis (or the empirical best
+  basis if the theoretical best is unknown) is calculated as the
+  Frobenius norm, which measures how close a visited basis is to the
+  best basis. Plotting this projection distance against the index value
+  for each basis visited reveals how the index value changes as the
+  basis approaches the optimal basis, helping to identify the
+  squintability of the index function.
+  Figure~\ref{fig-index-value-proj-dist} shows the index value against
+  the projection distance for the given example and its binned version.
+  As the projection distance decreases, the index value initially
+  increases with an increasing slope and then level off. The index value
+  shows greater variation at medium projection distances compared to
+  small and large distances.
 \end{itemize}
 
-The basis column records every basis \emph{visited} by the jellyfish
-optimiser prior to comparing with the current basis. In each iteration,
-if the index value of a visited basis is smaller than that of the
-current one, the jellyfish optimiser will retain the current basis for
-the next iteration, while still documenting the visited basis.
+\begin{figure}
 
-Numerical information to compute:
+\centering{
 
-\begin{itemize}
-\tightlist
-\item
-  angular distance between the projection basis and the theoretical best
-  basis,
-\item
-  the proportion of simulation that found the optimal basis,
-\item
-  the proportion of jellies, within each simulation, that found the
-  optimal basis,
-\end{itemize}
-
-Visualisation to inspect:
-
-\begin{itemize}
-\tightlist
-\item
-  inspect the basis visited by each jellyfish in the reduced PCA space,
-\item
-  inspect the final 2D projections reached by each jellyfish,
-\item
-  plot the index value against the angular distance between the
-  projection basis and the theoretical best basis
-\end{itemize}
-
-Plotting the basis in the space and the projected data can help to
-understand 1) whether each simulation finds the same optimum or some
-simulations find local optima; and 2) whether the index function used
-can detect the structure in the data and the projection contains the
-structure of interest.
-
-The visualisations above can be faceted by the projection pursuit
-variables and jellyfish optimiser parameters to compare the performance
-of different indexes to detect the same structure and how the jellyfish
-optimiser parameters affect the optimisation process.
-
-{[}example plots{]}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-4-1.pdf}
+\includegraphics{optim_files/figure-pdf/fig-projection-1.pdf}
 
@@
 
-\caption{sdfjsflk}
+\caption{\label{fig-projection}Best projections found by the JS
+opitmiser in each simulation, sorted by the index value. The projection
+pursuit problem is to detect the 6D sine-wave structure using the spline
+index. Most simulations have detected a clear sine-wave pattern, while
+the last six show a straight line with clustering at the ends.. In this
+simulation, the success rate is 0.88 (44/50).}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{optim_files/figure-pdf/fig-index-value-proj-dist-1.pdf}
+
+}
+
+\caption{\label{fig-index-value-proj-dist}Projection distance plotted
+against the index value for detecting the 6D sine-wave structure using
+the spline index function. a) The index values start around 0 with low
+variance at large projection distances. As the projection distance
+decreases, the index values increase with a larger variance. Towards the
+minimum projection distance, the index values increase and cluster
+around 1, indicating the optimiser is approaching to the optimum. b)
+Projection distances are binned with a bin width of 0.1, and index
+values are averaged over each bin. The relationship between the two
+variables forms a sigmoid curve.}
 
@@
 100 jellyfishes, and 2) 50 and 100 maximum number of tries. In each
-simulation, success is defined as achieving a final index value within
-0.05 of the best index value found among all 50 simulations. For
-example, in the 8-dimensional data scenario with 20 jellyfishes and a
-maximum number of tries set to 100, the optimal index value is 0.997.
-Out of the 50 simulations, 27 achieved an index value within {[}0.947,
-0.997{]}, resulting in a success rate of 0.54.
-Figure~\ref{fig-proportion} presents the proportion of success for each
-hyper-parameter combinations. As the number of jellyfish and maximum
-tries increase, the success rate also increases. For simpler problems (6
-dimensions), small parameter values (20 jellyfishes and a max number of
-tries of 50) can already result in high success rate, while for
-higher-dimensional problems (8, 10, and 12 dimensions), a combination of
-100 jellyfishes and 100 maximum number of tries is necessary for at
-least half of the jellyfishes to find the optimal projection.
+simulation, success rate is calculated as defined in
+Section~\ref{sec-vis}. Figure~\ref{fig-proportion} presents the
+proportion of success for each hyper-parameter combinations. As the
+number of jellyfish and maximum tries increase, the success rate also
+increases. For simpler problems (6 dimensions), small parameter values
+(20 jellyfishes and a maximum number of tries of 50) can already result
+in high success rate, while for higher-dimensional problems (8, 10, and
+12 dimensions), a combination of 100 jellyfishes and 100 maximum number
+of tries is necessary for at least half of the jellyfishes to find the
+optimal projection.
 
@@
 \emph{Note: should we print out the 52 cases investigated somewhere, say
-appendex, since it is not full grid simulation (otherwise for the
+appendix, since it is not full grid simulation (otherwise for the
 sine-wave problem there will be 6 indexes x 6 configurations = 36 cases
@@
 \texttt{dimension} are significant, suggesting their importance relative
-to jellyfish hyper-parameters in the optimisation sucess.
+to jellyfish hyper-parameters in the optimisation success.
 
@@
 family and a logit link function. The variable smoothness,
-squintability, number of jellyfish and maximum numer of tries are
+squintability, number of jellyfish and maximum number of tries are
 positively associated with the JS optimiser success rate while data
 dimension and being flagged as long runtime are negatively associated
-with teh success rate. The variable squintability and dimension are
+with the success rate. The variable squintability and dimension are
 significant, suggesting their importance relative to jellyfish

### 2024-06-12 code/02-simulation-sine-6d-spline.R

@@ -56,6 +56,23 @@ sim_sine_6d_spline <- bind_rows(sim_sine_6d_spline130, sim_sine_6d_spline3150 |>
 save(sim_sine_6d_spline, file = here::here("data/sim_sine_6d_spline.rda"))
 
 
+sim_sine_6d_spline_head <- sim_sine_6d_spline |>
+  filter(n_jellies == 50, max_tries == 50) |>
+  head(10)
+save(sim_sine_6d_spline_head, file = here::here("data/sim_sine_6d_spline_head.rda"))
+sim_sine_6d_spline_best <- sim_sine_6d_spline |>
+  filter(n_jellies == 50, max_tries == 50) |>
+  get_best(group = sim) |> ungroup()
+save(sim_sine_6d_spline_best, file = here::here("data/sim_sine_6d_spline_best.rda"))
+
+sim_sine_6d_spline_projdist <- sim_sine_6d_spline |>
+  filter(n_jellies == 50, max_tries == 50) |>
+  rowwise() |>
+  mutate(proj_dist = tourr::proj_dist(basis, matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE))) |>
+  select(-d, -id, -info, -method, -time, -basis) |>
+  ungroup()
+save(sim_sine_6d_spline_projdist, file = here::here("data/sim_sine_6d_spline_projdist.rda"))
+
 set.seed(123456)
 sine1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
 colnames(sine1000) <- paste0("V", 1:6)

### 2024-06-13 optim.qmd
minor text polishing
Minor polishing + fixed a few typos

@@
 
-* British English ("American or British usage is accepted, but
-not a mixture of these")*
+-   British English ("American or British usage is accepted, but not a
+    mixture of these")\*
 
-* No "we" - always third person
+-   No "we" - always third person
 
-* Check if the affiliation information is correct
+-   Check if the affiliation information is correct
 
-* word use:
+-   word use:
 
-  * Use jellyfish search optimiser, or JS optimiser, not jellyfish optimiser (my bad for leading the way)
-  * "hyper-parameter" rather than "hyperparameter"
+    -   Use jellyfish search optimiser, or JS optimiser, not jellyfish
+        optimiser (my bad for leading the way)
+    -   "hyper-parameter" rather than "hyperparameter"
 
@@
 
-The data structure proposed in @RJ-2021-105 is implemented for JS optimisers and  additional information such as index function used, data dimension, and jellyfish hyper-parameters can be included as additional columns.
-Below prints an example data collected from finding the sine-wave structure in 6D data (`d = 6`) using a spline index function(`splines2d`):
+The data structure proposed in @RJ-2021-105 is implemented for JS
+optimisers and additional information such as index function used, data
+dimension, and jellyfish hyper-parameters can be included as additional
+columns. An example data collected from finding the sine-wave structure
+in 6D data (`d = 6`) using a spline index function(`splines2d`) is
+printed below:
 
@@
 
-The jellyfish hyper-parameters used are 50 jellyfishes (`n_jellies`) and a maximum of 50 tries (`max_tries`).
-The full data contains 50 simulations, recorded by the variable `sim`, with different seeds (`seed`).
-Recorded in columns `basis`, `index_val`, `tries`, `loop` and `time`, are the sampled projection basis, calculated index value, iteration ID, jellyfish ID, and time taken to find the optimum, respectively.
-The basis column records every visited basis before comparing with the current (last) basis.
-If a visited basis has a smaller index value than the current (last) one, the JS optimiser keeps the current basis for the next iteration in the search but still records the visited inferior basis.
-This data can be used to derive numerical summaries and plots:
+The jellyfish hyper-parameters used are 50 jellyfishes (`n_jellies`) and
+a maximum of 50 tries (`max_tries`). The full data contains 50
+simulations, recorded by the variable `sim`, with different seeds
+(`seed`). Recorded in columns `basis`, `index_val`, `tries`, `loop` and
+`time`, are the sampled projection basis, calculated index value,
+iteration ID, jellyfish ID, and time taken to find the optimum,
+respectively. The basis column records every visited basis before
+comparing with the current (last) basis. If a visited basis has a
+smaller index value than the current (last) one, the JS optimiser keeps
+the current basis for the next iteration in the search but still records
+the visited inferior basis. This data can be used to derive numerical
+summaries and plots:
 
--   **success rate**:  Success rate is defined as the proportion of simulations that achieve a final index value within 0.05 of the best index value found among all 50 simulations.
-In the dataset above, the optimal index value is `r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`.
-Out of the 50 simulations, `r sim_sine_6d_spline_best |> filter(index_val >= aa - 0.05) |> nrow() -> bb; bb` achieved an index value within [`r aa - 0.05`, `r aa`], resulting in a success rate of `r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.
-@fig-projection shows the best projection found in each simulation, sorted by the index value.
-Most of the projections show a clear sine-wave structure, while the last six shows only a straight line with clusters at the end.
-When local optima exist in the projection pursuit problem, plotting the projections found in each simulation or by each jellyfish can help identify these local optima.
+-   **Success rate**: The success rate is defined as the proportion of
+    simulations that achieve a final index value within 0.05 of the best
+    index value found among all 50 simulations. In the dataset above,
+    the optimal index value is
+    `r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`.
+    Out of the 50 simulations,
+    `r sim_sine_6d_spline_best |> filter(index_val >= aa - 0.05) |> nrow() -> bb; bb`
+    achieved an index value within \[`r aa - 0.05`, `r aa`\], resulting
+    in a success rate of
+    `r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.
+    @fig-projection shows the best projection found in each simulation,
+    sorted by the index value. Most of the projections show a clear
+    sine-wave structure, while the last six show only a straight line
+    with clusters at the end. When local optima exist in the projection
+    pursuit problem, plotting the projections found in each simulation
+    or by each jellyfish can help identify these local optima.
 
--  **Projection distance**: The projection distance between projection bases and the theoretical best basis (or the empirical best basis if the theoretical best is unknown) is calculated as the Frobenius norm, which measures how close a visited basis is to the best basis.
-Plotting this projection distance against the index value for each basis visited reveals how the index value changes as the basis approaches the optimal basis, helping to identify the squintability of the index function.
-@fig-index-value-proj-dist shows the index value against the projection distance for the given example and its binned version.
-As the projection distance decreases, the index value initially increases with an increasing slope and then level off.
-The index value shows greater variation at medium projection distances compared to small and large distances.
+-   **Projection distance**: The projection distance between projection
+    bases and the theoretical best basis (or the empirical best basis if
+    the theoretical best is unknown) is calculated as the Frobenius
+    norm, which measures how close a visited basis is to the best basis.
+    Plotting this projection distance against the index value for each
+    basis visited reveals how the index value changes as the basis
+    approaches the optimal basis, helping to identify the squintability
+    of the index function. @fig-index-value-proj-dist shows the index
+    value against the projection distance for the given example and its
+    binned version. As the projection distance decreases, the index
+    value initially increases with an increasing slope and then levels
+    off. The index value shows greater variation at medium projection
+    distances compared to small and large distances.
 
@@
 
-The JS optimiser provides a new swarm-based search strategy to optimise projection pursuit problems.
-Two examples are provided in this section to demonstrate its performance.
-The first example compares its performance with the search better optimiser and explores how it behaves with different combinations of hyper-parameter.
-The second example studies the effect of optimisation properties defined in @sec-theory, along with jellyfish hyper-parameters, on the outcome of the optimisation.
+The JS optimiser provides a new swarm-based search strategy to optimise
+projection pursuit problems. Two examples are provided in this section
+to demonstrate its performance. The first example compares its
+performance with the search-better optimiser and explores how it behaves
+with different combinations of hyper-parameters. The second example
+studies the effect of optimisation properties defined in @sec-theory,
+along with jellyfish hyper-parameters, on the outcome of the
+optimisation.
 
-## Performance of JS optimiser in pipe-finding problems {#sec-app-1}
+## Performance of the JS optimiser in pipe-finding problems {#sec-app-1}
 
-The performance of JS optimiser is investigated in two folds: 1) compare it with a commonly used optimiser: the search better optimiser [@RJ-2021-105; @laa_using_2020], and 2) examine its success rate under different hyper-parameters (number of jellyfishes and the maximum number of tries).
-The projection pursuit problem used is finding the pipe shape using the holes index, investigated by @laa_using_2020.
+The performance of the JS optimiser is investigated in two folds: 1)
+compare it with a commonly used optimiser: the search-better optimiser
+[@RJ-2021-105; @laa_using_2020], and 2) examine its success rate under
+different hyper-parameters (number of jellyfishes and maximum number of
+tries). The projection pursuit problem used is finding the pipe shape
+using the holes index, investigated by @laa_using_2020.
 
@@
 
-Fifty (50) simulations are conducted with both JS and the search better optimiser, in four dimensions ($d = 6, 8, 10, 12$).
-The JS optimiser uses 100 jellyfishes and 100 maximum number of tries, while the search better optimiser allows a maximum of 1000 samples at each iteration before the algorithm terminates.
-@fig-proj presents the final projections found by the two optimisers, broken down by 10th quantile, faceted by the data dimension.
-In the 6-dimensional data scenario, the JS optimiser consistently identifies a clear pipe shape.
-The search better optimiser, although also find the pipe shape but with a wide rim, suggesting a further polish search may be required.
-With increasing dimensions, the JS optimiser may not always identify the pipe shape due to random sampling, it still presents a pipe shape in over 50% of cases.
-When compared to the search better optimiser, the JS optimiser reaches higher index values and clearer pipe shapes across all quantiles.
+Fifty simulations are conducted with both JS and the search-better
+optimiser, in four dimensions ($d = 6, 8, 10, 12$). The JS optimiser
+uses 100 jellyfishes and 100 maximum number of tries, while the
+search-better optimiser allows a maximum of 1000 samples at each
+iteration before the algorithm terminates. @fig-proj presents the final
+projections found by the two optimisers, broken down by 10th quantile,
+faceted by the data dimension. In the 6-dimensional data scenario, the
+JS optimiser consistently identifies a clear pipe shape. The
+search-better optimiser also finds the pipe shape but with a wide rim,
+suggesting a further polish search may be required. With increasing
+dimensions, the JS optimiser may not always identify the pipe shape due
+to random sampling, but it still presents a pipe shape in over 50% of
+cases. When compared to the search-better optimiser, the JS optimiser
+reaches higher index values and clearer pipe shapes across all
+quantiles.
 
-In the second experiment, fifty (50) simulations are conducted at each  hyper-parameter combination to analyse its effects on the success rate of the JS optimiser.
-The hyper-parameters tested include: 1) 20, 50, and 100 jellyfishes, and 2) 50 and 100 maximum number of tries.
-In each simulation,  success rate is calculated as defined in @sec-vis.
-@fig-proportion presents the proportion of success for each hyper-parameter combinations.
-As the number of jellyfish and maximum tries increase, the success rate also increases.
-For simpler problems (6 dimensions), small parameter values (20 jellyfishes and a maximum number of tries of 50) can already result in high success rate, while for higher-dimensional problems (8, 10, and 12 dimensions), a combination of 100 jellyfishes and 100 maximum number of tries is necessary for at least half of the jellyfishes to find the optimal projection.
+In the second experiment, fifty simulations are conducted at each
+hyper-parameter combination to analyse its effects on the success rate
+of the JS optimiser. The hyper-parameters tested include: 1) 20, 50, and
+100 jellyfishes, and 2) 50 and 100 maximum number of tries. In each
+simulation, the success rate is calculated as defined in @sec-vis.
+@fig-proportion presents the proportion of success for each
+hyper-parameter combinations. As the number of jellyfishes and maximum
+tries increase, the success rate also increases. For simpler problems (6
+dimensions), small parameter values (20 jellyfishes and a maximum number
+of tries of 50) can already result in high success rate, while for
+higher-dimensional problems (8, 10, and 12 dimensions), a combination of
+100 jellyfishes and 100 maximum number of tries is necessary for at
+least half of the jellyfishes to find the optimal projection.
 
-## Factors affecting JS optimiser success rate: optimisation properties and jellyfish hyper-parameters {#sec-app-2}
+## Factors affecting the JS optimiser success rate: optimisation properties and jellyfish hyper-parameters {#sec-app-2}
 
-The optimisation properties, including smoothness, squintability, and speed, offer numerical metrics to characterise the complexity of a projection pursuit optimisation problem.
-This example investigates how these metrics, alongside jellyfish hyper-parameters (the number of jellyfish and the maximum number of tries), affect the success of the JS optimiser.
-Simulations are conducted to obtain the success rate across various projection pursuit problems, characterised by shape-to-find, data dimension, and index function used, as well as different hyper-parameter combinations.
-Smoothness and squintability are calculated for each projection pursuit problem as outlined in @sec-theory.
-A generalised linear model is used to construct the relationship between the success rate, jellyfish hyper-parameters and optimisation properties.
+The optimisation properties, including smoothness, squintability, and
+speed, offer numerical metrics to characterise the complexity of a
+projection pursuit optimisation problem. This example investigates how
+these metrics, along with the jellyfish search hyper-parameters (the
+number of jellyfishes and the maximum number of tries), affect the
+success of the JS optimiser. Simulations are conducted to obtain the
+success rate across various projection pursuit problems, characterised
+by shape-to-find, data dimension, and index function used, as well as
+different hyper-parameter combinations. Smoothness and squintability are
+calculated for each projection pursuit problem as outlined in
+@sec-theory. A generalised linear model is used to construct the
+relationship between the success rate, jellyfish hyper-parameters and
+optimisation properties.
 
-In addition to the pipe-finding problem, a new shape, sine wave, is investigated in 6D and 8D spaces with six indices considered: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`.
-Combining with two jellyfish hyper-parameters, a total of `r nrow(sim_df)` cases is produced, comprising of `r nrow(filter(sim_df, index == "holes"))` pipe-finding cases  and `r nrow(filter(sim_df, index != "holes"))` sine-wave finding cases.
-For each case, the JS optimiser is run fifty (50) times and the success rate is calculated as in @sec-app-1.
+In addition to the pipe-finding problem, a new shape, sine wave, is
+investigated in 6D and 8D spaces with six indices considered:
+`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`. Combining
+with two jellyfish hyper-parameters, a total of `r nrow(sim_df)` cases
+is produced, comprising of `r nrow(filter(sim_df, index == "holes"))`
+pipe-finding cases and `r nrow(filter(sim_df, index != "holes"))`
+sine-wave finding cases. For each case, the JS optimiser is run fifty
+times and the success rate is calculated as in @sec-app-1.
 
-*Note: should we print out the 52 cases investigated somewhere, say appendix, since it is not full grid simulation (otherwise for the sine-wave problem there will be 6 indexes x 6 configurations = 36 cases for 6-D data and 2 indexes [MIC & TIC] x 6 configurations = 12 cases for 8-D data)?
-We only do a selection of 28 of them to save time/ computational resources.*
+*Note: should we print out the 52 cases investigated somewhere, say
+appendix, since it is not full grid simulation (otherwise for the
+sine-wave problem there will be 6 indexes x 6 configurations = 36 cases
+for 6-D data and 2 indexes \[MIC & TIC\] x 6 configurations = 12 cases
+for 8-D data)? We only do a selection of 28 of them to save time/
+computational resources.*
 
-Smoothness and squintability are computed for each case, following the procedures outlined in @sec-smoothness and @sec-squintability.
-To calculate smoothness, three hundred random bases are simulated.
-Index values and projection distance (to the optimal basis) are calculated for each random basis before fitting them into a Gaussian process model.
-For squintability, fifty random bases are sampled and interpolated to the optimal basis with a step size of 0.005.
-For these interpolated bases, index values and projection distances to the optimal basis are calculated and averaged with a bin width of 0.005.
-The squintability measure is then calculated from fitting a 4-parameter scaled logistic function of index value against projection distance, estimated by non-linear least square.
+Smoothness and squintability are computed for each case, following the
+procedures outlined in @sec-smoothness and @sec-squintability. To
+calculate smoothness, three hundred random bases are simulated. Index
+values and projection distance (to the optimal basis) are calculated for
+each random basis before fitting them into a Gaussian process model. For
+squintability, fifty random bases are sampled and interpolated to the
+optimal basis with a step size of 0.005. For these interpolated bases,
+index values and projection distances to the optimal basis are
+calculated and averaged with a bin width of 0.005. The squintability
+measure is then calculated from fitting a 4-parameter scaled logistic
+function of index value against projection distance, estimated by
+non-linear least squares.
 
-*Note: Shall I repeat the scaled logistic function with our parameterisation here? It can be handy to interpret theta1 - theta4 but they should've already been explained in the squintability section.*
+*Note: Shall I repeat the scaled logistic function with our
+parameterisation here? It can be handy to interpret theta1 - theta4 but
+they should've already been explained in the squintability section.*
 
-@tbl-smoothness-squintability presents the parameters estimated from the Gaussian process (variance, range, smooth, and nugget) and the scaled logistic function ($\theta_1$ to $\theta_4$) for each case.
-The column "smooth" is used as the smoothness measure and the column  "squint" is calculated as $\frac{\theta_1 \theta_2 \theta_3}{4}$ as squintability.
-The low squint value of MIC and TIC index is due to its convex shape of the index value against the projection distance, as shown in @fig-idx-proj-dist.
-Comparing to other concave shapes, the maximum first derivative happens at a smaller projection distance ($\theta_2$), requiring the optimiser to get closer to the optimal to see a significant change in the index value, hence a small squintability measure.
+@tbl-smoothness-squintability presents the parameters estimated from the
+Gaussian process (variance, range, smooth, and nugget) and the scaled
+logistic function ($\theta_1$ to $\theta_4$) for each case. The column
+"smooth" is used as the smoothness measure and the column "squint" is
+calculated as $\frac{\theta_1 \theta_2 \theta_3}{4}$ as squintability.
+The low squint value of MIC and TIC index is due to its convex shape of
+the index value against the projection distance, as shown in
+@fig-idx-proj-dist. Comparing to other concave shapes, the maximum first
+derivative happens at a smaller projection distance ($\theta_2$),
+requiring the optimiser to get closer to the optimal to see a
+significant change in the index value, hence a small squintability
+measure.
 
@@
 
-@tbl-mod-data combines all the variables calculated above (the JS optimiser success rate, jellyfish hyper-parameters, and optimisation properties) into one table.
-Three pre-processing steps are applied to the data: 1) scaling the jellyfish hyper-parameters by a factor of 10 for interpretation, 2) creating a new binary variable `long_time` to indicate cases with an average run time over 30 seconds, and 3) encoding the success rate for the stringy case as 0, since none of the 50 simulations identified the sine-wave shape.
-A generalised linear model with a binomial family and a logit link function is used to fit the data and @tbl-mod-output presents the model outputs.
-The model suggests that the success rate of the JS optimiser is positively associated with the two jellyfish hyper-parameters, as well as with smoothness and squintability.
-However, being flagged with long runtime and an increase of data dimension will reduce the success rate.
-The variable `squintability` and `dimension` are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.
+@tbl-mod-data combines all the variables calculated above (the JS
+optimiser success rate, jellyfish hyper-parameters, and optimisation
+properties) into one table. Three preprocessing steps are applied to the
+data: 1) scaling the jellyfish hyper-parameters by a factor of 10 for
+interpretation, 2) creating a new binary variable `long_time` to
+indicate cases with an average run time over 30 seconds, and 3) encoding
+the success rate for the stringy case as 0, since none of the 50
+simulations identified the sine-wave shape. A generalised linear model
+with a binomial family and a logit link function is used to fit the data
+and @tbl-mod-output presents the model outputs. The model suggests that
+the success rate of the JS optimiser is positively associated with the
+two jellyfish hyper-parameters, as well as with smoothness and
+squintability. However, being flagged with long runtime and an increase
+of data dimension will reduce the success rate. The variable
+`squintability` and `dimension` are significant, suggesting their
+importance relative to jellyfish hyper-parameters in the optimisation
+success.
 

### 2024-06-14 optim.qmd
clean up repo

--- 2024-06-12 optim.qmd
@@
 library(ferrn)
-load(here::here("data/sim_pipe.rda"))
+load(here::here("data/sim_pipe_run_best.rda"))
 load(here::here("data/sim_sine_6d_spline_head.rda"))
@@
 As the dimensionality increases, the proportion of simulations reaching the optimal index value increases."
-pipe_run_summ <- sim_pipe |>
-  mutate(idx_f = "holes") |>
-  group_by(id, n_jellies, max_tries, d) |> 
-  summarise(I_max = max(index_val))
-
-pipe_sim_best <- pipe_run_summ |> 
+pipe_sim_best <- sim_pipe_run_best |> 
   group_by(n_jellies, max_tries, d) |>
@@
 
-pipe_res <- pipe_run_summ |> 
+pipe_res <- sim_pipe_run_best |> 
   left_join(pipe_sim_best) |> 

### 2024-06-14 optim.tex
clean up repo

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, toptitle=1mm, colback=white, colframe=quarto-callout-note-color-frame, titlerule=0mm, breakable, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, bottomrule=.15mm, arc=.35mm, rightrule=.15mm, leftrule=.75mm, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, rightrule=.15mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, opacityback=0, bottomrule=.15mm, titlerule=0mm, breakable, left=2mm, colback=white, toptitle=1mm, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm]
 

### 2024-06-14 code/00-search-better.R

@@ -1,0 +1,88 @@
+library(tidyverse)
+library(spinebil)
+library(tourr)
+
+set.seed(123)
+seed <- sample(1000: 10000, size = 50)
+historyarray2dt <- function(array, index_f){
+  stopifnot(inherits(array, "history_array"))
+  data <- attr(array, "data")
+  n <- dim(array)[3]
+  res <- purrr::map_dfr(1:n, ~tibble(basis = list(unclass(array)[,,.x])))
+  projs <-  map(res$basis, ~data %*% .x)
+  res <- res |>
+    dplyr::mutate(index_val = map_dbl(projs, ~index_f(.x)),
+                  id = dplyr::row_number())
+
+  return(res)
+
+}
+
+sim <- function(d, optimiser = c("search_better_random", "search_better", "search_jellyfish"),
+                data_seed = 123456, optim_seed = seed){
+  set.seed(data_seed)
+  pipe1000 <- pipeData(d, 1000) %>% scale() %>% as_tibble()
+  colnames(pipe1000) <- paste0("V", 1:d)
+
+  if (optimiser == "search_better_random"){
+    res <- map_dfr(optim_seed, function(seed){
+      set.seed(seed)
+      spiral_better <- save_history(
+        pipe1000, guided_tour(
+          holes(), search_f = search_better_random, max.tries = 1000
+        )
+      )
+      historyarray2dt(spiral_better[,,-1], holes())
+    }, .id = "sim") |>
+      mutate(d = d)
+  } else if (optimiser == "search_better"){
+    res <- map_dfr(optim_seed, function(seed){
+      set.seed(seed)
+      spiral_better <- save_history(
+        pipe1000, guided_tour(
+          holes(), search_f = search_better, max.tries = 1000
+        )
+      )
+      historyarray2dt(spiral_better[,,-1], holes())
+    }, .id = "sim") |>
+      mutate(d = d)
+  }
+  return(res)
+}
+
+data <- map_dfr(c(6, 8, 10, 12), ~{
+  set.seed(123456)
+  tibble(x = list(pipeData(.x, 1000) %>% scale()))
+}) |>
+  mutate(dim = c(6, 8, 10, 12))
+# a <- map_dfr(c(6, 8, 10, 12),
+#              ~safe_sim(d = .x, optimiser = "search_better",
+#                        optim_seed = seed),
+#              .id = "dim"
+#              )
+# take around an hour to run
[…]

### 2024-06-14 code/01-simulation-holes.R

@@ -44,6 +44,10 @@ sim_res <- sim_setup |>
 # sim_pipe <- sim_res |> select(-alpha)
 # save(sim_pipe, file = "data/sim_pipe.rda")
 
+sim_pipe_run_best <- sim_pipe |>
+  group_by(id, n_jellies, max_tries, d) |>
+  summarise(I_max = max(index_val))
+save(sim_pipe_run_best, file = "data/sim_pipe_run_best.rda")
 
 pipe_jellyfish <- sim_pipe |>
   filter(n_jellies == 100, max_tries == 100) |>
@@ -51,3 +55,5 @@ pipe_jellyfish <- sim_pipe |>
   ungroup() |>
   mutate(optimiser = "jellyfish")
 # save(pipe_jellyfish, file = "data/pipe_jellyfish.rda")
+
+

### 2024-06-14 code/example1-high-d.R

@@ -1,88 +1,0 @@
-library(tidyverse)
-library(spinebil)
-library(tourr)
-
-set.seed(123)
-seed <- sample(1000: 10000, size = 50)
-historyarray2dt <- function(array, index_f){
-  stopifnot(inherits(array, "history_array"))
-  data <- attr(array, "data")
-  n <- dim(array)[3]
-  res <- purrr::map_dfr(1:n, ~tibble(basis = list(unclass(array)[,,.x])))
-  projs <-  map(res$basis, ~data %*% .x)
-  res <- res |>
-    dplyr::mutate(index_val = map_dbl(projs, ~index_f(.x)),
-                  id = dplyr::row_number())
-
-  return(res)
-
-}
-
-sim <- function(d, optimiser = c("search_better_random", "search_better", "search_jellyfish"),
-                data_seed = 123456, optim_seed = seed){
-  set.seed(data_seed)
-  pipe1000 <- pipeData(d, 1000) %>% scale() %>% as_tibble()
-  colnames(pipe1000) <- paste0("V", 1:d)
-
-  if (optimiser == "search_better_random"){
-    res <- map_dfr(optim_seed, function(seed){
-      set.seed(seed)
-      spiral_better <- save_history(
-        pipe1000, guided_tour(
-          holes(), search_f = search_better_random, max.tries = 1000
-        )
-      )
-      historyarray2dt(spiral_better[,,-1], holes())
-    }, .id = "sim") |>
-      mutate(d = d)
-  } else if (optimiser == "search_better"){
-    res <- map_dfr(optim_seed, function(seed){
-      set.seed(seed)
-      spiral_better <- save_history(
-        pipe1000, guided_tour(
-          holes(), search_f = search_better, max.tries = 1000
-        )
-      )
-      historyarray2dt(spiral_better[,,-1], holes())
-    }, .id = "sim") |>
-      mutate(d = d)
-  }
-  return(res)
-}
-
-data <- map_dfr(c(6, 8, 10, 12), ~{
-  set.seed(123456)
-  tibble(x = list(pipeData(.x, 1000) %>% scale()))
-}) |>
-  mutate(dim = c(6, 8, 10, 12))
-# a <- map_dfr(c(6, 8, 10, 12),
-#              ~safe_sim(d = .x, optimiser = "search_better",
-#                        optim_seed = seed),
-#              .id = "dim"
-#              )
-# take around an hour to run
[…]

### 2024-06-14 code/example2-skewness.R

@@ -1,172 +1,0 @@
-library(tourr)
-library(MaxSkew)
-library(MultiSkew)
-library(tidyverse)
-x <- matrix(c(
-  9160, 1759, 2691,
-    12, 7218, 7655,
-  4624, 4735, 1887,
-  4243, 1527, 2875,
-  4609, 3411,  911,
-  7702, 6074, 5762,
-  3225, 1917, 6834,
-  7847, 7384, 5466,
-  4714, 2428, 4257,
-   358, 9174, 6444
-), ncol = 3, byrow = TRUE)
-vec <- c(0.816, 0.446, 0.367)
-vec <- get_init(x) # hack into the code .MaxSkewThree to get .linear
-res1 <- MaxSkew(x, iterations = 3, components = 1, plot = FALSE)
-res1 <- x %*% vec
-skewness()(res1)
-
-i <- 1
-j <- 1
-for (i in 1:3){
-  for (j in 1:3){
-    y<-x[,j]#j-th column of the data matrix
-    vec[j]<-c(0) #removes the j-th variable from the linear combination
-    MaxSkew:::.MaxSkewBiv(x%*%vec,y)
-    s<-.skewdesideredBIV #skewness of the new linear combination
-    vec[j]<-.linearBIV[2]/.linearBIV[1]
-    res1 <- x %*% vec
-  }
-  print(skewness()(res1))
-
-}
-
-get_init <- function(data) {
-  n <- nrow(data)
-  d <- ncol(data)
-
-  provaconcentrationmatrix <- solve(cov(data) * (n - 1) / n)
-  provaSS <- provaconcentrationmatrix
-  provaaut <- eigen(provaSS)
-  sort(provaaut$values, decreasing = FALSE)
-  provaVV <- matrix(c(0), nrow = d)
-  provaDD <- diag(sort(provaaut$values, decreasing = FALSE))
-
-  for (i in ncol(provaaut$vectors):1) {
-    provaVV <- cbind(provaVV, provaaut$vectors[, i])
-  }
-
-  provaVV <- provaVV[, 2:ncol(provaVV)]
-  provaDD.sqrt <- solve(sqrt(provaDD))
-  provaAA <- provaVV %*% provaDD.sqrt %*% t(provaVV)
-  provaQ <- solve(provaAA)
-
-  x.mean <- colMeans(data)#vector mean
-  m <- sweep(data, 2, x.mean)#centered data
-  Z <- m %*% provaQ
-
[…]

### 2024-06-14 code/example3-local-max.R

@@ -1,62 +1,0 @@
-library(tidyverse)
-library(spinebil)
-library(tourr)
-
-set.seed(123456)
-pipe1000 <- pipeData(6, 1000) |> scale() |> as_tibble()
-sine1000 <- sinData(6, 1000) |> scale()
-dt <-pipe1000 |> bind_cols(sine1000[,5:6])
-colnames(dt) <- paste0("V", 1:(6+ 2))
-
-
-better_res <- animate_xy(dt, guided_tour(
-  holes(), d = 2,
-  search_f =  search_better, alpha = 0.7,
-  max.tries = 500
-))
-
-set.seed(1234)
-jellyfish_res <- animate_xy(dt, guided_tour(
-  holes(), d = 2, n_jellies = 50,
-  search_f =  search_jellyfish,
-  max.tries = 50
-))
-
-proj1 <- matrix(c(0, 0, 0, 0, 1, 0, 0, 0,
-                  0, 0, 0, 0, 0, 1, 0, 0), ncol = 2)
-proj2 <- matrix(c(0, 0, 0, 0, 0, 0, 1, 0,
-                  0, 0, 0, 0, 0, 0, 0, 1), ncol = 2)
-proj3 <- matrix(c(0, 0, 0, 0, 1, 0, 0, 0,
-                  0, 0, 0, 0, 0, 0, 1, 0), ncol = 2)
-proj4 <- matrix(c(0, 0, 0, 0, 1, 0, 0, 0,
-                  0, 0, 0, 0, 0, 1, 0, 1), ncol = 2)
-proj5 <- matrix(c(0, 0, 0, 0, 0, 1, 0, 0,
-                  0, 0, 0, 0, 0, 0, 1, 0), ncol = 2)
-proj6 <- matrix(c(0, 0, 0, 0, 0, 1, 0, 0,
-                  0, 0, 0, 0, 0, 0, 0, 1), ncol = 2)
-proj_list <- list(proj1, proj2, proj3, proj4, proj5, proj6)
-res <- map_dfr(proj_list, ~ as_tibble(as.matrix(dt) %*% .), .id = "id")
-idx <- map(proj_list, ~holes()(as.matrix(dt) %*% .x))
-
-res |>
-  ggplot(aes(x = V1, y = V2)) +
-  geom_point() +
-  theme(aspect.ratio = 1) +
-  facet_wrap(vars(id))
-
-
-best_jellyfish <- jellyfish_res #|> filter(info == "current_best")
[…]

### 2024-06-14 code/example4-summarise-experiment.R

@@ -1,39 +1,0 @@
-library(tidyverse)
-load(here::here("data/pipe_raw.rda")) # loaded as b
-raw <- b |> unnest(result)
-raw_clean <- raw |> rename(setup = dim, tour = sim, id = loop, dim = d) |>
-  mutate(setup = as.numeric(setup), tour = as.numeric(tour), id = as.numeric(id))
-
-jelly_level <- raw_clean |> select(-c(info: alpha))
-
-tour_level_best_basis <- jelly_level |>
-  group_by(setup, tour) |>
-  filter(index_val == max(index_val)) |> select(setup, tour, basis)
-
-# columns
-# setup:      number of experiment (1-4),
-# tour:       number of tour run in each setup (1-50)
-# J:          number of jellyfish,
-# P_J:        proportion of jellyfish found
-# B:          max tries set,
-# B_star:     max.tries used (doens't look right here)
-# D:          the coefficient in c_t > .5 (the 3 in best_jelly - 3 * runif(1) ...)
-# I_max:      best index across all the jellyfish in each trial run,
-# basis:      projection basis
-tour_level <- jelly_level |>
-  group_by(setup, tour) |>
-  reframe(J = 1000, B = 300, D = 3, B_star = n()/J, I_max = max(index_val),
-          # TODO: use a better rule to define jellyfish found
-          P_J = sum(abs(I_max - index_val) <= 0.05)/J) |>
-  left_join(tour_level_best_basis) |>
-  select(setup, tour, J, P_J, B, B_star, D, I_max, basis)
-
-# columns
-# P_J_hat:    proportion of trials found
-# I_max_max:  best index of each setup
-setup_level <- tour_level |>
-  group_by(setup) |>
-  reframe(P_J_hat = sum(abs(1 - I_max) <= 0.05)/n(), # TODO: use a better rule
-          I_max_max = max(I_max))
-save(tour_level, file = here::here("data/tour_level.rda"))
-save(setup_level, file = here::here("data/setup_level.rda"))

### 2024-06-14 optim.qmd
changes that might need to be dropped

@@
 colnames(sin1000) <- paste0("V", 1:6)
+#remotes::install_github("uschiLaa/spinebil")
+#load(here::here("data/sine_dcor2d_best.rda"))
+#load(here::here("data/sin1000.rda"))
+#set.seed(123456)
+#sin1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
+#colnames(sin1000) <- paste0("V", 1:6)
 

### 2024-06-14 optim.qmd
paper compiling

--- 2024-06-14 optim.qmd
@@
    - \usepackage{algorithm}
-   - \usepackage{algpseudocode}
 author:
@@
 In this simulation, the success rate is 0.88 (44/50).
-set.seed(123456)
-sin1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
-colnames(sin1000) <- paste0("V", 1:6)
+#set.seed(123456)
+#sin1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
+#colnames(sin1000) <- paste0("V", 1:6)
+#remotes::install_github("uschiLaa/spinebil")
+#load(here::here("data/sim_sine_6d_spline_best.rda"))
+load(here::here("data/sin1000.rda"))
+#set.seed(123456)
+#sin1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
+#colnames(sin1000) <- paste0("V", 1:6)
 

### 2024-06-14 code/01-sim-pipe-681012d-holes.R
clean up datasets

@@ -41,9 +41,13 @@ sim_res <- sim_setup |>
   ungroup() |>
   unnest(res)
 
-# sim_pipe <- sim_res |> select(-alpha)
-# save(sim_pipe, file = "data/sim_pipe.rda")
+sim_pipe_raw <- sim_res |> select(index, d, n_jellies, max_tries, sim:time)
+sim_pipe <- tibble(index = "holes") |> bind_cols(sim_pipe_raw)
+save(sim_pipe, file = "data-raw/sim_pipe.rda")
 
+################################################################################
+################################################################################
+# additional smaller data
 sim_pipe_run_best <- sim_pipe |>
   group_by(id, n_jellies, max_tries, d) |>
   summarise(I_max = max(index_val))
@@ -54,6 +58,6 @@ pipe_jellyfish <- sim_pipe |>
   get_best(group = id) |>
   ungroup() |>
   mutate(optimiser = "jellyfish")
-# save(pipe_jellyfish, file = "data/pipe_jellyfish.rda")
+save(pipe_jellyfish, file = "data/pipe_jellyfish.rda")
 
 

### 2024-06-14 code/02-sim-sine-6d-splines2d.R

@@ -52,10 +52,17 @@ t2 - t1
 sim_sine_6d_spline130 <- sim_res |> select(-alpha)
 sim_sine_6d_spline3150 <- sim_res |> select(-alpha)
 
-sim_sine_6d_spline <- bind_rows(sim_sine_6d_spline130, sim_sine_6d_spline3150 |> mutate(id = id + 30))
-save(sim_sine_6d_spline, file = here::here("data/sim_sine_6d_spline.rda"))
+sim_sine_6d_spline <- bind_rows(
+  sim_sine_6d_spline130, sim_sine_6d_spline3150 |> mutate(id = id + 30)
+  )
+sim_sine_6d_splines2d <- tibble(index = "splines2d") |>
+  bind_cols(sim_sine_6d_spline) |>
+  select(index, d, n_jellies, max_tries, sim:time)
+save(sim_sine_6d_splines2d, file = here::here("data-raw/sim_sine_6d_splines2d.rda"))
 
-
+################################################################################
+################################################################################
+# additional smaller data
 sim_sine_6d_spline_head <- sim_sine_6d_spline |>
   filter(n_jellies == 50, max_tries == 50) |>
   head(10)
@@ -65,18 +72,23 @@ sim_sine_6d_spline_best <- sim_sine_6d_spline |>
   get_best(group = sim) |> ungroup()
 save(sim_sine_6d_spline_best, file = here::here("data/sim_sine_6d_spline_best.rda"))
 
+mmat <- matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE)
 sim_sine_6d_spline_projdist <- sim_sine_6d_spline |>
   filter(n_jellies == 50, max_tries == 50) |>
   rowwise() |>
-  mutate(proj_dist = tourr::proj_dist(basis, matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE))) |>
+  mutate(proj_dist = tourr::proj_dist(basis, mmat)) |>
   select(-d, -id, -info, -method, -time, -basis) |>
   ungroup()
 save(sim_sine_6d_spline_projdist, file = here::here("data/sim_sine_6d_spline_projdist.rda"))
 
+
+################################################################################
+################################################################################
 set.seed(123456)
 sine1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
[…]

### 2024-06-14 code/03-sim-sine-6d-loess2d.R

@@ -64,12 +64,9 @@ sim_res <- sim_setup |>
   unnest(res)
 
 sim_sine_6d <- sim_res |> select(-alpha)
-save(sim_sine_6d, file = "data/sim_sine_6d.rda")
+sim_sine_6d_loess2d <- tibble(index = "loess") |>
+  bind_cols(sim_sine_6d) |>
+  select(index, d, n_jellies, max_tries, sim:time)
+save(sim_sine_6d_loess2d, file = "data-raw/sim_sine_6d_loess2d.rda")
 
 
-# sum <- sim_res |> group_by(id) |> filter(index_val == max(index_val)) |> filter(row_number() == 1)
-# b <- sum |> filter(id == 2) |> pull(basis)
-# dt <- as.matrix(sine1000) %*% b[[1]]
-# as_tibble(dt) |>
-#   ggplot(aes(x = V2 ,y = V1)) +
-#   geom_point()

### 2024-06-14 code/04-sim-sine-68d-MICTIC.R

@@ -74,42 +74,6 @@ sim_res <- sim_setup |>
 t2 <- Sys.time()
 t2 - t1
 
-# both indexes are similar
-# I_max_max shows that most combination (max_tries x n_jellies x d x idx_f)
+sim_sine_68d_TICMIC <- sim_res |> select(-alpha) |> rename(index = idx_f)
+save(sim_sine_68d_TICMIC, file = "data/sim_sine_68d_TICMIC.rda")
 
-
-# sim_sine_68d_TICMIC <- sim_res |> select(-alpha)
-# save(sim_sine_68d_TICMIC, file = "data/sim_sine_68d_TICMIC.rda")
-#
-#
-set.seed(123456)
-sine1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
-colnames(sine1000) <- paste0("V", 1:6)
-sum <- sim_res |> group_by(id) |> filter(index_val == max(index_val)) |> filter(row_number() == 1)
-#tour_level |> filter(id == 167) |> pull(basis)
-# tour_level |>
-#   filter(n_jellies == 50, max_tries == 100, idx_f == "MIC", d == 6) |>
-#   filter(id == 164) |> pull(basis) -> b
-b <- setup_level_best_basis |> filter(id == 717) |> pull(basis)
-ids <- setup_level_best_basis |> filter(d == 6) |> pull(id)
-res <- map_dfr(ids, ~{
-  b <- setup_level_best_basis |> filter(id == .x) |> pull(basis)
-  dt <- as_tibble(as.matrix(sine1000) %*% b[[1]])
-  return(dt)
-}, .id = "id")
-
-labels_df <- setup_level_best_basis |>
-  ungroup() |>
-  filter(d == 6) |>
-  select(idx_f:max_tries, index_val) |>
-  mutate(id = as.character(1:10))
-
-library(ggh4x)
-res |>
-  left_join(labels_df) |>
-  ggplot(aes(x = V2 ,y = V1)) +
-  geom_point() +
-  geom_text(data = labels_df,
-            aes(x = -3, y = 3, label = round(index_val, 2)),
-            nudge_x = 0.1, nudge_y = 0.1) +
-  facet_nested(idx_f ~ n_jellies + max_tries, labeller = label_both )

### 2024-06-14 code/05-sim-sine-6d-dcor2d.R

@@ -68,28 +68,5 @@ t2 <- Sys.time()
 t2 - t1
 
 
-sim_sine_6d_dcor2d <- sim_res |> select(-alpha)
-save(sim_sine_6d_dcor2d , file = "data/sim_sine_6d_dcor2d.rda")
-sim_example <- sim_sine_6d_dcor2d |> head(10)
-save(sim_example , file = "data/sim_example.rda")
-sine_dcor2d_best <- sim_sine_6d_dcor2d |>
-  filter(n_jellies == 50, max_tries == 50) |>
-  group_by(sim) |>
-  filter(index_val == max(index_val)) |>
-  filter(row_number() == 1) |>
-  ungroup()
-save(sine_dcor2d_best , file = "data/sine_dcor2d_best.rda")
-
-
-
-#
-# set.seed(123456)
-# sine1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
-# colnames(sine1000) <- paste0("V", 1:6)
-# sum <- sim_res |> group_by(id) |> filter(index_val == max(index_val)) |> filter(row_number() == 1)
-# b <- sum |> filter(id == 2) |> pull(basis)
-# dt <- as_tibble(as.matrix(sine1000) %*% b[[1]])
-# dt |>
-#   ggplot(aes(x = V2 ,y = V1)) +
-#   geom_point() +
-#   theme(aspect.ratio = 1)
+sim_sine_6d_dcor2d <- sim_res |> select(-alpha) |> rename(index = idx_f)
+save(sim_sine_6d_dcor2d , file = "data-raw/sim_sine_6d_dcor2d.rda")

### 2024-06-14 code/06-sim-sine-6d-stringy.R

@@ -57,17 +57,9 @@ sim_sine_6d_stringy2150 <- sim_res |> select(-alpha)
 sim_sine_6d_stringy <- bind_rows(sim_sine_6d_stringy12, sim_sine_6d_stringy310,
                                  sim_sine_6d_stringy1120, sim_sine_6d_stringy2150)
 
-save(sim_sine_6d_stringy, file = here::here("data/sim_sine_6d_stringy.rda"))
+sim_sine_6d_stringy <- tibble(index = "stringy") |>
+  bind_cols(sim_sine_6d_stringy) |>
+  select(index, d, n_jellies, max_tries, sim:time)
+save(sim_sine_6d_stringy, file = here::here("data-raw/sim_sine_6d_stringy.rda"))
 
 
-# set.seed(123456)
-# sine1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
-# colnames(sine1000) <- paste0("V", 1:6)
-# sum <- sim_sine_6d_stringy310 |> group_by(id) |> filter(index_val == max(index_val)) |> filter(row_number() == 1)
-# b <- sum |> pull(basis)
-# dt <- map_dfr(b, ~as_tibble(as.matrix(sine1000) %*% .x), .id = "id")
-# dt |>
-#   ggplot(aes(x = V2 ,y = V1)) +
-#   geom_point(size = 0.3) +
-#   theme(aspect.ratio = 1) +
-#   facet_wrap(vars(id))

### 2024-06-14 code/08-process-data-for-modelling.R

@@ -1,6 +1,7 @@
 library(tidyverse)
-load(file = "data/sim_pipe.rda")
 
+# calculate success rate for pipe-finding problem
+load(file = "data-raw/sim_pipe.rda")
 # all the pipe simulation are conducted in one go,
 # so id is different for each n_jellies, max_tries, d combination
 pipe_run_df <- sim_pipe |> get_best(group = id)
@@ -9,88 +10,49 @@ pipe_setup_df <- sim_pipe |>
   get_best(group = id2)
 
 # columns
-# setup:      number of experiment (1-4),
-# tour:       number of tour run in each setup (1-50)
-# J:          number of jellyfish,
-# P_J:        proportion of jellyfish found
-# B:          max tries set,
-# B_star:     max.tries used (doens't look right here)
-# D:          the coefficient in c_t > .5 (the 3 in best_jelly - 3 * runif(1) ...)
-# I_max:      best index across all the jellyfish in each trial run,
-# basis:      projection basis
-pipe_run_summ <- sim_pipe |>
-  mutate(idx_f = "holes") |>
-  group_by(id) |>
-  reframe(I_max = max(index_val),
-          P_J = sum(abs(I_max - index_val) <= 0.05)/n(),
-          time = mean(time))
-
-# columns
-# P_J_hat:    proportion of trials found
-# I_max_max:  best index of each setup
+# I_max:      best index across all the simulations under each setting (n_jellies, max_tries, d)
+# P_J:        success rate, proportion of simulation, out of 50 that find close (< 0.05) index value
+# time:       average time across 50 simulations
 pipe_setup_summ <- pipe_run_df |>
-  mutate(idx_f = "holes") |>
-  group_by(n_jellies, max_tries, d, idx_f) |>
+  group_by(index, d, n_jellies, max_tries) |>
   reframe(
-    I_max_max = max(index_val),
-    P_J_hat = sum(abs(I_max_max - index_val) <= 0.05)/n(),
+    I_max = max(index_val),
+    P_J = sum(abs(I_max - index_val) <= 0.05)/n(),
     time = mean(time)
   )
 
 #############################################################################
-load(here::here("data/sim_sine_6d_dcor2d.rda"))
-load(here::here("data/sim_sine_6d.rda")) # loess
-load(here::here("data/sim_sine_68d_TICMIC.rda"))
-load(here::here("data/sim_sine_6d_spline.rda"))
-load(here::here("data/sim_sine_6d_stringy.rda"))
[…]

### 2024-06-14 optim.qmd
update the squintability calculation and misc

@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+#install the latest version of ferrn from: https://github.com/huizezhang-sherry/ferrn
 library(tidyverse)
@@
 
+TODO: add equation reference to main text line 705
+
 To the best of our knowledge, this is the first attempt to measure the
@@
 In this simulation, the success rate is 0.88 (44/50).
-#set.seed(123456)
-#sin1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
-#colnames(sin1000) <- paste0("V", 1:6)
-#remotes::install_github("uschiLaa/spinebil")
-#load(here::here("data/sim_sine_6d_spline_best.rda"))
-load(here::here("data/sin1000.rda"))
-#set.seed(123456)
-#sin1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
-#colnames(sin1000) <- paste0("V", 1:6)
-
 proj_dt <- map_dfr(sim_sine_6d_spline_best |> pull(basis), 
-              ~as_tibble(as.matrix(sin1000) %*% .x), .id = "sim") |> 
+              ~as_tibble(as.matrix(sine1000) %*% .x), .id = "sim") |> 
   mutate(sim = as.numeric(sim)) |> 
@@
 
-*Note: should we print out the 52 cases investigated somewhere, say
-appendix, since it is not full grid simulation (otherwise for the
-sine-wave problem there will be 6 indexes x 6 configurations = 36 cases
-for 6-D data and 2 indexes \[MIC & TIC\] x 6 configurations = 12 cases
-for 8-D data)? We only do a selection of 28 of them to save time/
-computational resources.*
-
 Smoothness and squintability are computed for each case, following the
@@
 
-*Note: Shall I repeat the scaled logistic function with our
-parameterisation here? It can be handy to interpret theta1 - theta4 but
-they should've already been explained in the squintability section.*
-
 @tbl-smoothness-squintability presents the parameters estimated from the
@@
 "smooth" is used as the smoothness measure and the column "squint" is
-calculated as $\frac{\theta_1 \theta_2 \theta_3}{4}$ as squintability.
+calculated as [TODO: equation reference] as squintability.
 The low squint value of MIC and TIC index is due to its convex shape of
@@
 
-*Note: the stringy squintability doesn't make sense*
+```{r fig-idx-proj-dist}
+#| fig.width = 9,
+#| fig.height = 5,
+#| fig.cap = "Index values versus projection distance for the 12 pipe/sine-wave finding problem, after the binning procedure for calculating the squintability measure.
+The index values, averaged at bin width of 0.005, are scaled from 0 to 1 for comparison (holes, TIC, and stringy).
+The `MIC` and `TIC` index curves are convex while others are concave.
+The stringy curve shows an instantaneous jump to the optimum when aproaching the best basis."
+sq_basis_dist_idx |>
+  ggplot(aes(x = dist, y = idx, group = index)) +
+  geom_line() +
+  facet_nested_wrap(~index + n, nrow = 3, labeller = label_both) + 
+  xlab("projection distance") + ylab("index value")
+
+```
 
@@
 #| tbl-cap: Parameters estimated from the Gaussian process (including variance, range, smoothness, and nugget) and scaled logistic function ($\theta_1$ to $\theta_4$) for the pipe-finding and sine-wave finding problems.
-The "smooth" column and "squint" column, calculated as $\frac{\theta_1 \theta_2 \theta_3}{4}$, represent the smoothness and squintability measures.
+The column "smooth" and "squint" represent the smoothness and squintability measures.
 dt <- tibble(shape = c(rep("pipe", 4), rep("sine", 8)), smoothness) |>
   left_join(squintability) |> 
-  rename(d = n, smooth = smoothness)
+  rename(d = n, smooth = smoothness) |> 
+  mutate(index = ifelse(index == "dcor2d_2", "dcor2d", index)) 
 dt |> 
-  knitr::kable(digits = 3, booktabs = T, escape = FALSE, row.names = TRUE) |> 
+  knitr::kable(
+    digits = 3, format = "latex", 
+    col.names = c(
+      "shape", "index", "d",
+      "variance", "range", "smooth", "nugget",
+      "$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", "squint"),
+    booktabs = T, escape = FALSE, row.names = TRUE) |> 
   column_spec(1, border_left = TRUE) |> 
-  column_spec(c(3, 7, 12, 13), border_right = TRUE) |> 
-  kable_styling(latex_options = "hold_position", font_size = 7, full_width = FALSE)
+  column_spec(c(4, 8, 12, 13), border_right = TRUE) |> 
+  kable_styling(font_size = 7)
 ```
@@
   dplyr::select(index, d, smoothness, squintability, n_jellies, max_tries, 
-                -I_max_max, P_J_hat, time) |> 
+                -I_max, P_J, time) |> 
   arrange(index, d, n_jellies, max_tries) |> 
   head(7) |> 
-  kable(digits = 3, col.names = c("Index", "D", "Smoothness", "Squintability", 
-                                  "Num. of jellyfish", "Max. Num. of tries", "Prop. of success", "Time"), 
+  kable(digits = 3,
+        col.names = c("Index", "D", "Smoothness", "Squintability", 
+                      "Num. of jellyfish", "Max. Num. of tries", 
+                      "Prop. of success", "Time"), 
         booktabs = T) |> 
@@
          long_time = ifelse(time > 30, 1, 0),
-         P_J_hat = ifelse(index == "stringy", 0, P_J_hat))
+         P_J = ifelse(index == "stringy", 0, P_J))
 
-mod1 <- glm(P_J_hat ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
+mod1 <- glm(P_J ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
 broom::tidy(mod1) |> 
@@
   column_spec(c(1,5), border_right = TRUE) |> 
-  column_spec(1, border_left = TRUE)
+  column_spec(1, border_left = TRUE) |> 
+  kable_styling(font_size = 7)
 ```
@@
 
-# Additional plot:
-
-```{r fig-idx-proj-dist}
-#| fig.width = 9,
-#| fig.height = 5,
-#| fig.cap = "Index values against projection distance for the 12 pipe/sine-wave finding problem after the binning procedure during the estimation of the squintability measure.
-The index value is averaged at each bin width of 0.005 and the `TIC` index is scaled to 0-1 for comparison.
-When finding the sine-wave structure using the `MIC` and `TIC` index, a convex curved is observed, as opposed to the pipe-finding problem or the sine-wave finding problem with the `dcor2d_2`, `loess2d`, and `splines2d` index.
-When finding the sine-wave with the stringy index, the index shows an instantaneous jump to the optimum when close to the best basis."
-sq_basis_dist_idx |>
-  ggplot(aes(x = dist, y = idx, group = index)) +
-  geom_line() +
-  facet_nested_wrap(~index + n, nrow = 3, labeller = label_both) + 
-  xlab("projection distance") + ylab("index value")
-
-```
-
 # References

### 2024-06-14 optim.tex
update the squintability calculation and misc

@@
 \usepackage{algorithm}
-\usepackage{algpseudocode}
 \makeatletter
@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, rightrule=.15mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, opacityback=0, bottomrule=.15mm, titlerule=0mm, breakable, left=2mm, colback=white, toptitle=1mm, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, colback=white, bottomrule=.15mm, rightrule=.15mm, left=2mm, bottomtitle=1mm, breakable, toprule=.15mm, leftrule=.75mm, opacityback=0]
 
@@
 dimension, and jellyfish hyper-parameters can be included as additional
-columns. Below prints an example data collected from finding the
-sine-wave structure in 6D data (\texttt{d\ =\ 6}) using a spline index
-function(\texttt{splines2d}):
+columns. An example data collected from finding the sine-wave structure
+in 6D data (\texttt{d\ =\ 6}) using a spline index
+function(\texttt{splines2d}) is printed below:
 
@@
 \item
-  \textbf{success rate}: Success rate is defined as the proportion of
-  simulations that achieve a final index value within 0.05 of the best
-  index value found among all 50 simulations. In the dataset above, the
-  optimal index value is 0.999. Out of the 50 simulations, 44 achieved
-  an index value within {[}0.949, 0.999{]}, resulting in a success rate
-  of 0.88. Figure~\ref{fig-projection} shows the best projection found
-  in each simulation, sorted by the index value. Most of the projections
-  show a clear sine-wave structure, while the last six shows only a
-  straight line with clusters at the end. When local optima exist in the
-  projection pursuit problem, plotting the projections found in each
-  simulation or by each jellyfish can help identify these local optima.
+  \textbf{Success rate}: The success rate is defined as the proportion
+  of simulations that achieve a final index value within 0.05 of the
+  best index value found among all 50 simulations. In the dataset above,
+  the optimal index value is 0.999. Out of the 50 simulations, 44
+  achieved an index value within {[}0.949, 0.999{]}, resulting in a
+  success rate of 0.88. Figure~\ref{fig-projection} shows the best
+  projection found in each simulation, sorted by the index value. Most
+  of the projections show a clear sine-wave structure, while the last
+  six show only a straight line with clusters at the end. When local
+  optima exist in the projection pursuit problem, plotting the
+  projections found in each simulation or by each jellyfish can help
+  identify these local optima.
 \item
@@
   As the projection distance decreases, the index value initially
-  increases with an increasing slope and then level off. The index value
-  shows greater variation at medium projection distances compared to
-  small and large distances.
+  increases with an increasing slope and then levels off. The index
+  value shows greater variation at medium projection distances compared
+  to small and large distances.
 \end{itemize}
@@
 to demonstrate its performance. The first example compares its
-performance with the search better optimiser and explores how it behaves
-with different combinations of hyper-parameter. The second example
+performance with the search-better optimiser and explores how it behaves
+with different combinations of hyper-parameters. The second example
 studies the effect of optimisation properties defined in
@@
 
-\subsection{Performance of JS optimiser in pipe-finding
+\subsection{Performance of the JS optimiser in pipe-finding
 problems}\label{sec-app-1}
 
-The performance of JS optimiser is investigated in two folds: 1) compare
-it with a commonly used optimiser: the search better optimiser
+The performance of the JS optimiser is investigated in two folds: 1)
+compare it with a commonly used optimiser: the search-better optimiser
 \citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
-under different hyper-parameters (number of jellyfishes and the maximum
+under different hyper-parameters (number of jellyfishes and maximum
 number of tries). The projection pursuit problem used is finding the
@@
 
-Fifty (50) simulations are conducted with both JS and the search better
+Fifty simulations are conducted with both JS and the search-better
 optimiser, in four dimensions (\(d = 6, 8, 10, 12\)). The JS optimiser
-uses 100 jellyfishes and 100 maximum number of tries, while the search
-better optimiser allows a maximum of 1000 samples at each iteration
-before the algorithm terminates. Figure~\ref{fig-proj} presents the
-final projections found by the two optimisers, broken down by 10th
-quantile, faceted by the data dimension. In the 6-dimensional data
-scenario, the JS optimiser consistently identifies a clear pipe shape.
-The search better optimiser, although also find the pipe shape but with
-a wide rim, suggesting a further polish search may be required. With
+uses 100 jellyfishes and 100 maximum number of tries, while the
+search-better optimiser allows a maximum of 1000 samples at each
+iteration before the algorithm terminates. Figure~\ref{fig-proj}
+presents the final projections found by the two optimisers, broken down
+by 10th quantile, faceted by the data dimension. In the 6-dimensional
+data scenario, the JS optimiser consistently identifies a clear pipe
+shape. The search-better optimiser also finds the pipe shape but with a
+wide rim, suggesting a further polish search may be required. With
 increasing dimensions, the JS optimiser may not always identify the pipe
-shape due to random sampling, it still presents a pipe shape in over
-50\% of cases. When compared to the search better optimiser, the JS
+shape due to random sampling, but it still presents a pipe shape in over
+50\% of cases. When compared to the search-better optimiser, the JS
 optimiser reaches higher index values and clearer pipe shapes across all
@@
 
-In the second experiment, fifty (50) simulations are conducted at each
+In the second experiment, fifty simulations are conducted at each
 hyper-parameter combination to analyse its effects on the success rate
@@
 100 jellyfishes, and 2) 50 and 100 maximum number of tries. In each
-simulation, success rate is calculated as defined in
+simulation, the success rate is calculated as defined in
 Section~\ref{sec-vis}. Figure~\ref{fig-proportion} presents the
 proportion of success for each hyper-parameter combinations. As the
-number of jellyfish and maximum tries increase, the success rate also
+number of jellyfishes and maximum tries increase, the success rate also
 increases. For simpler problems (6 dimensions), small parameter values
@@
 
-\subsection{Factors affecting JS optimiser success rate: optimisation
-properties and jellyfish hyper-parameters}\label{sec-app-2}
+\subsection{Factors affecting the JS optimiser success rate:
+optimisation properties and jellyfish hyper-parameters}\label{sec-app-2}
 
@@
 projection pursuit optimisation problem. This example investigates how
-these metrics, alongside jellyfish hyper-parameters (the number of
-jellyfish and the maximum number of tries), affect the success of the JS
-optimiser. Simulations are conducted to obtain the success rate across
-various projection pursuit problems, characterised by shape-to-find,
-data dimension, and index function used, as well as different
-hyper-parameter combinations. Smoothness and squintability are
+these metrics, along with the jellyfish search hyper-parameters (the
+number of jellyfishes and the maximum number of tries), affect the
+success of the JS optimiser. Simulations are conducted to obtain the
+success rate across various projection pursuit problems, characterised
+by shape-to-find, data dimension, and index function used, as well as
+different hyper-parameter combinations. Smoothness and squintability are
 calculated for each projection pursuit problem as outlined in
@@
 \texttt{spline}, and \texttt{stringy}. Combining with two jellyfish
-hyper-parameters, a total of 52 cases is produced, comprising of 24
-pipe-finding cases and 28 sine-wave finding cases. For each case, the JS
-optimiser is run fifty (50) times and the success rate is calculated as
-in Section~\ref{sec-app-1}.
-
-\emph{Note: should we print out the 52 cases investigated somewhere, say
-appendix, since it is not full grid simulation (otherwise for the
-sine-wave problem there will be 6 indexes x 6 configurations = 36 cases
-for 6-D data and 2 indexes {[}MIC \& TIC{]} x 6 configurations = 12
-cases for 8-D data)? We only do a selection of 28 of them to save time/
-computational resources.}
+hyper-parameters, a total of 42 cases is produced, comprising of 24
+pipe-finding cases and 8 sine-wave finding cases. For each case, the JS
+optimiser is run fifty times and the success rate is calculated as in
+Section~\ref{sec-app-1}.
 
@@
 scaled logistic function of index value against projection distance,
-estimated by non-linear least square.
-
-\emph{Note: Shall I repeat the scaled logistic function with our
-parameterisation here? It can be handy to interpret theta1 - theta4 but
-they should've already been explained in the squintability section.}
+estimated by non-linear least squares.
 
@@
 for each case. The column ``smooth'' is used as the smoothness measure
-and the column ``squint'' is calculated as
-\(\frac{\theta_1 \theta_2 \theta_3}{4}\) as squintability. The low
-squint value of MIC and TIC index is due to its convex shape of the
-index value against the projection distance, as shown in
-Figure~\ref{fig-idx-proj-dist}. Comparing to other concave shapes, the
-maximum first derivative happens at a smaller projection distance
-(\(\theta_2\)), requiring the optimiser to get closer to the optimal to
-see a significant change in the index value, hence a small squintability
-measure.
+and the column ``squint'' is calculated as {[}TODO: equation
+reference{]} as squintability. The low squint value of MIC and TIC index
+is due to its convex shape of the index value against the projection
+distance, as shown in Figure~\ref{fig-idx-proj-dist}. Comparing to other
+concave shapes, the maximum first derivative happens at a smaller
+projection distance (\(\theta_2\)), requiring the optimiser to get
+closer to the optimal to see a significant change in the index value,
+hence a small squintability measure.
 
-\emph{Note: the stringy squintability doesn't make sense}
+\begin{figure}
+
+\centering{
+
+\includegraphics{optim_files/figure-pdf/fig-idx-proj-dist-1.pdf}
+
+}
+
+\caption{\label{fig-idx-proj-dist}Index values versus projection
+distance for the 12 pipe/sine-wave finding problem, after the binning
+procedure for calculating the squintability measure. The index values,
+averaged at bin width of 0.005, are scaled from 0 to 1 for comparison
+(holes, TIC, and stringy). The \texttt{MIC} and \texttt{TIC} index
+curves are convex while others are concave. The stringy curve shows an
+instantaneous jump to the optimum when aproaching the best basis.}
+
+\end{figure}%
 
@@
 (the JS optimiser success rate, jellyfish hyper-parameters, and
-optimisation properties) into one table. Three pre-processing steps are
+optimisation properties) into one table. Three preprocessing steps are
 applied to the data: 1) scaling the jellyfish hyper-parameters by a
@@
 
-\begingroup\fontsize{7}{9}\selectfont
-
-\begin{longtable}[t]{|>{}ll>{}l|rrr>{}r|rrrr>{}r|>{}r|}
+\begin{table}
 
@@
 and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
-pipe-finding and sine-wave finding problems. The ``smooth'' column and
-``squint'' column, calculated as
-\(\frac{\theta_1 \theta_2 \theta_3}{4}\), represent the smoothness and
-squintability measures.}
+pipe-finding and sine-wave finding problems. The column ``smooth'' and
+``squint'' represent the smoothness and squintability measures.}
 
-\tabularnewline
+\centering{
 
+\centering\begingroup\fontsize{7}{9}\selectfont
+
+\begin{tabular}{|>{}lll>{}r|rrr>{}r|rrr>{}r|>{}r|}
 \toprule
- & shape & index & d & variance & range & smooth & nugget & theta1 & theta2 & theta3 & theta4 & squint\\
+  & shape & index & d & variance & range & smooth & nugget & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & squint\\
 \midrule
-1 & pipe & holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 & 0.991 & 0.860 & 3.368 & 0.172 & 0.718\\
-2 & pipe & holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 & 1.011 & 0.869 & 3.264 & 0.085 & 0.716\\
-3 & pipe & holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 & 1.011 & 0.885 & 3.151 & 0.153 & 0.705\\
-4 & pipe & holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 & 0.999 & 0.878 & 3.345 & 0.196 & 0.733\\
-5 & sine & MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 & 0.894 & 0.571 & 1.623 & -0.024 & 0.207\\
+1 & pipe & holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 & 0.991 & 0.860 & 3.368 & 0.172 & 0.627\\
+2 & pipe & holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 & 1.011 & 0.869 & 3.264 & 0.085 & 0.696\\
+3 & pipe & holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 & 1.011 & 0.885 & 3.151 & 0.153 & 0.636\\
+4 & pipe & holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 & 0.999 & 0.878 & 3.345 & 0.196 & 0.621\\
+5 & sine & MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 & 0.894 & 0.571 & 1.623 & -0.024 & 0.314\\
 \addlinespace
-6 & sine & MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 & 0.932 & 0.328 & 1.314 & -0.030 & 0.100\\
-7 & sine & TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 & 0.951 & 0.536 & 1.719 & -0.025 & 0.219\\
-8 & sine & TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 & 0.945 & 0.564 & 1.723 & -0.027 & 0.230\\
-9 & sine & dcor2d\_2 & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019 & 0.679\\
-10 & sine & loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 & 1.016 & 1.039 & 2.648 & 0.080 & 0.699\\
+6 & sine & MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 & 0.932 & 0.328 & 1.314 & -0.030 & 0.193\\
+7 & sine & TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 & 0.951 & 0.536 & 1.719 & -0.025 & 0.330\\
+8 & sine & TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 & 0.945 & 0.564 & 1.723 & -0.027 & 0.341\\
+9 & sine & dcor2d & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019 & 0.737\\
+10 & sine & loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 & 1.016 & 1.039 & 2.648 & 0.080 & 0.689\\
 \addlinespace
-11 & sine & splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 & 1.014 & 1.051 & 2.730 & -0.009 & 0.727\\
-12 & sine & stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 & 1.040 & 0.011 & 254.733 & 0.170 & 0.731\\
+11 & sine & splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 & 1.014 & 1.051 & 2.730 & -0.009 & 0.780\\
+12 & sine & stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 & 1.040 & 0.011 & 254.733 & 0.170 & 0.648\\
 \bottomrule
+\end{tabular}
+\endgroup{}
 
-\end{longtable}
+}
 
-\endgroup{}
+\end{table}%
 
@@
 \midrule
-MIC & 6 & 2.394 & 0.207 & 20 & 50 & 0.12 & 2.479 secs\\
-MIC & 6 & 2.394 & 0.207 & 20 & 100 & 0.24 & 8.950 secs\\
-MIC & 6 & 2.394 & 0.207 & 50 & 50 & 0.52 & 5.651 secs\\
-MIC & 6 & 2.394 & 0.207 & 50 & 100 & 0.64 & 13.223 secs\\
-MIC & 6 & 2.394 & 0.207 & 100 & 50 & 0.76 & 19.453 secs\\
+dcor2d\_2 & 6 & 2.663 & 0.737 & 20 & 50 & 0.44 & 35.320 secs\\
+dcor2d\_2 & 6 & 2.663 & 0.737 & 20 & 100 & 0.60 & 70.347 secs\\
+dcor2d\_2 & 6 & 2.663 & 0.737 & 50 & 50 & 0.90 & 87.158 secs\\
+holes & 6 & 2.364 & 0.627 & 20 & 50 & 0.88 & 0.801 secs\\
+holes & 6 & 2.364 & 0.627 & 20 & 100 & 0.96 & 1.456 secs\\
 \addlinespace
-MIC & 8 & 2.394 & 0.100 & 20 & 50 & 0.08 & 2.566 secs\\
-MIC & 8 & 2.394 & 0.100 & 20 & 100 & 0.08 & 4.960 secs\\
+holes & 6 & 2.364 & 0.627 & 50 & 50 & 1.00 & 1.477 secs\\
+holes & 6 & 2.364 & 0.627 & 50 & 100 & 1.00 & 3.052 secs\\
 \bottomrule
@@
 \endgroup{}
+
+\begingroup\fontsize{7}{9}\selectfont
 
@@
 \midrule
-Intercept & -4.849 & 4.496 & -1.078 & 0.281\\
-Smoothness & 1.672 & 1.527 & 1.094 & 0.274\\
-Squintability & 6.967 & 2.173 & 3.207 & 0.001\\
-dimension (d) & -0.588 & 0.257 & -2.292 & 0.022\\
-long time & -0.816 & 1.333 & -0.612 & 0.540\\
+Intercept & 2.685 & 15.464 & 0.174 & 0.862\\
+Smoothness & 3.070 & 3.631 & 0.845 & 0.398\\
+Squintability & -8.545 & 22.252 & -0.384 & 0.701\\
+dimension (d) & -0.610 & 0.355 & -1.719 & 0.086\\
+long time & -1.683 & 1.930 & -0.872 & 0.383\\
 \addlinespace
-number of jellyfish & 0.229 & 0.131 & 1.752 & 0.080\\
-maximum number of tries & 0.108 & 0.153 & 0.703 & 0.482\\
+number of jellyfish & 0.269 & 0.175 & 1.538 & 0.124\\
+maximum number of tries & 0.080 & 0.196 & 0.410 & 0.682\\
 \bottomrule
@@
 
+\endgroup{}
+
 \section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
-
-\section{Additional plot:}\label{additional-plot}
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{optim_files/figure-pdf/fig-idx-proj-dist-1.pdf}
-
-}
-
-\caption{\label{fig-idx-proj-dist}Index values against projection
-distance for the 12 pipe/sine-wave finding problem after the binning
-procedure during the estimation of the squintability measure. The index
-value is averaged at each bin width of 0.005 and the \texttt{TIC} index
-is scaled to 0-1 for comparison. When finding the sine-wave structure
-using the \texttt{MIC} and \texttt{TIC} index, a convex curved is
-observed, as opposed to the pipe-finding problem or the sine-wave
-finding problem with the \texttt{dcor2d\_2}, \texttt{loess2d}, and
-\texttt{splines2d} index. When finding the sine-wave with the stringy
-index, the index shows an instantaneous jump to the optimum when close
-to the best basis.}
-
-\end{figure}%
 

### 2024-06-14 code/07-calc-smoothness-squintability.R

@@ -89,7 +89,7 @@ save(sq_basis_df, file = here::here("data/sq_basis_df.rda"))
 
 start_prms1 <- c(theta1 = 1, theta2 = 1, theta3 = 3, theta4 = 0)
 start_prms2 <- c(theta1 = 1, theta2 = 0.01, theta3 = 50, theta4 = 0)
-squintability <- sq_basis_df |>
+squintability_raw <- sq_basis_df |>
   dplyr::mutate(start = ifelse(index == "stringy", list(start_prms2),
                                list(start_prms1))) |>
   dplyr::rowwise() |>
@@ -99,12 +99,26 @@ squintability <- sq_basis_df |>
   mutate(index = factor(index, levels = c("holes", "MIC", "TIC", "dcor2d_2",
                                           "loess2d", "splines2d", "stringy"))) |>
   arrange(index) |>
-  select(index, n, theta1: theta4) |>
-  mutate(squint = abs(theta1 * theta2 * theta3 / 4))
+  select(index, n, theta1: theta4)
+
+max_dist_df <- sq_basis_df |>
+  rowwise() |>
+  mutate(max_dist = basis_df |>
+           pivot_longer(holes:stringy, names_to = "index_f",
+                        values_to = "value", values_drop_na = TRUE) |>
+           pull(dist) |>
+           max()
+         ) |>
+  select(n, index, max_dist) |>
+  ungroup()
+
+squintability <- squintability_raw |>
+  left_join(max_dist_df) |>
+  mutate(
+    dd = (1/(1 + exp(-theta3 * theta2)) - 1/(1 + exp(theta3 * (max_dist)))),
+    squint = abs((theta1 - theta4)/dd  * theta2 * theta3 / 4)) |>
+  select(-dd, -max_dist)
 save(squintability, file = here::here("data", "squintability.rda"))
-
-
-
 ############################################################################
 ############################################################################
 sq_basis_dist_idx <- sq_basis_df |>

### 2024-06-14 optim.tex
small tweaks

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, colback=white, bottomrule=.15mm, rightrule=.15mm, left=2mm, bottomtitle=1mm, breakable, toprule=.15mm, leftrule=.75mm, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, toptitle=1mm, left=2mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, titlerule=0mm, toprule=.15mm, opacityback=0, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, breakable, arc=.35mm, rightrule=.15mm, leftrule=.75mm, coltitle=black]
 
@@
 Columns: 12
+$ index     <chr> "splines2d", "splines2d", "splines2d", "splines2d", "splines~
+$ d         <dbl> 6, 6, 6, 6, 6, 6, 6, 6, 6, 6
 $ n_jellies <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50
 $ max_tries <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50
-$ d         <dbl> 6, 6, 6, 6, 6, 6, 6, 6, 6, 6
 $ sim       <int> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1
@@
 $ time      <drtn> 4.362133 mins, 4.362133 mins, 4.362133 mins, 4.362133 mins, ~
-$ index     <chr> "spline2d", "spline2d", "spline2d", "spline2d", "spline2d", ~
 \end{verbatim}

### 2024-06-14 code/02-sim-sine-6d-splines2d.R

@@ -63,17 +63,17 @@ save(sim_sine_6d_splines2d, file = here::here("data-raw/sim_sine_6d_splines2d.rd
 ################################################################################
 ################################################################################
 # additional smaller data
-sim_sine_6d_spline_head <- sim_sine_6d_spline |>
+sim_sine_6d_spline_head <- sim_sine_6d_splines2d |>
   filter(n_jellies == 50, max_tries == 50) |>
   head(10)
 save(sim_sine_6d_spline_head, file = here::here("data/sim_sine_6d_spline_head.rda"))
-sim_sine_6d_spline_best <- sim_sine_6d_spline |>
+sim_sine_6d_spline_best <- sim_sine_6d_splines2d |>
   filter(n_jellies == 50, max_tries == 50) |>
   get_best(group = sim) |> ungroup()
 save(sim_sine_6d_spline_best, file = here::here("data/sim_sine_6d_spline_best.rda"))
 
 mmat <- matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE)
-sim_sine_6d_spline_projdist <- sim_sine_6d_spline |>
+sim_sine_6d_spline_projdist <- sim_sine_6d_splines2d |>
   filter(n_jellies == 50, max_tries == 50) |>
   rowwise() |>
   mutate(proj_dist = tourr::proj_dist(basis, mmat)) |>

### 2024-06-19 optim.qmd
small tweaks
Update optim.qmd

@@
           postal-code: 519087
-  - name: Jessica Leung
+  - name: Jessica Wai Yin Leung
     email: Jessica.Leung@monash.edu
@@
 
+Projection Pursuit (PP) is a high-dimensional data visualisation
+approach that involves computing informative linear projections of data
+by optimising a variety of objective functions, namely *projection
+pursuit index (PPI)*, (@hall1989polynomial, @cook1993projection,
+@lee2010projection, @Loperfido2018, @Loperfido2020). Given
+high-dimensional data $X \in \mathcal{R}^{n }\times p$ and PPI
+\$f(\\cdot)\$, PP finds the orthonormal projection matrix
+$A \in \mathcal{R}^{p \times d}$ by solving the following optimisation
+problem:
+
+$$
+\underset{A}{\max} f(XA) \quad \text{subject to} \quad A'A = I
+$$
+
+These indices $f(\cdot)$ are characterised by the informativeness or
+“interestingness” of a projection and are often non-linear and
+non-convex. As such, an effective and efficient optimisation procedure
+is essential to enable sufficient exploration of the data landscape
+during the visualisation process and to arrive at a global optimal
+viewpoint of the data.
+
+Significant efforts have been devoted to enhancing the functionality of
+PP and facilitating a better experience in the visualisation process.
+@cook1995grand introduced the PP guided tour, which enabled interactive
+visualisation of the optimisation to visually explore high-dimensional
+data. It is implemented in the R [@R] package `tourr`[@tourr].
+@RJ-2021-105 highlighted potential problems of the optimisation routine
+implemented. While improving the quality of the optimisation solutions
+in the tour is essential, it is also important to be able to watch the
+projected data as the optimisation progresses. Therefore, it is of
+interest to integrate the guided tour with a global optimisation
+algorithm and examine the characteristics and advantages of such
+implementation.
+
 The artificial jellyfish search (JS) algorithm [@chou_novel_2021] is a
 swarm-based metaheuristic optimisation algorithm inspired by the search
-behaviour of jellyfish in the ocean. It is one of the newest swarm
+behaviour of jellyfish in the ocean. It is one of the latest swarm
 intelligence algorithms [@rajwar_exhaustive_2023], which was shown to
-have stronger search ability and faster convergence with few algorithmic
-parameters compared to classic optimization methods
+have stronger search ability and faster convergence with fewer
+parameters compared to classic optimisation methods
 [@chou_novel_2021]-[@chou_recent_2022].
 
-Effective optimisation is an important aspect of many methods employed
-for visualising high-dimensional data ($X$). Here we are concerned about
-computing informative linear projections of high-dimensional ($p$) data
-using projection pursuit (PP) (@kr69, @FT74). This involves optimising a
-function (e.g. @hall1989polynomial, @cook1993projection,
-@lee2010projection, @Loperfido2018, @Loperfido2020), called the
-projection pursuit index (PPI), that defines what is interesting or
-informative in a projection.
+This study aims to investigate ...
 
-These PPI are defined on projections ($XA$), which means that there is a
-constraint that needs to be considered when optimising. A projection of
-data is defined by a $p\times d$ orthonormal matrix $A$, and this
-imposes the constraint on the elements of $A$, that columns need have
-norm equal to 1 and the product of columns need to sum to zero.
+The contribution of this study to the high dimensional data visualisation literature is twofold. First, the performance of JSO is studied in the context of PP guided tour.
+Second, to the best of the authors’ knowledge, this is the first attempt
+to quantitatively measure squintability and smoothness.
 
-@cook1995grand introduced the PP guided tour, which enabled interactive
-visualisation of the optimisation in order to visually explore
-high-dimensional data. It is implemented in the R [@R] package `tourr`
-[@tourr]. The optimisation that is implemented is fairly basic, and
-potential problems were highlighted by @RJ-2021-105. Implementing better
-optimisation functionality is a goal, but it needs to be kept in mind
-that the guided tour also has places importance on watching the
-projected data as the optimisation progresses.
-
-Here we explore the potential for a jellyfish optimisation to be
-integrated with the guided tour. @sec-background explains the
-optimisation that is used in the current the projection pursuit guided
-tour. @sec-theory provides more details on the jellyfish optimiser and
-formalises several characteristics of projection pursuit indexes that
-are help to measure optimiser performance. @sec-simulation describes a
-simulation study on performance of the jellyfish for several types of
-data and index functions. @sec-conclusion summarises the work and
-provides suggestions for future directions.
+The rest of this paper is structured as follows. @sec-background
+introduces the PP guided tour. @sec-theory describes the details of JSO
+and introduces metrics that measure different characteristics of a PPI.
+@sec-simulation describes a simulation study on the performance of the
+JSO using well-known data sets and index functions. @sec-conclusion
+summarises the work and provides suggestions for future directions.
 
@@
 ```{r}
-dplyr::glimpse(sim_sine_6d_spline_head |> mutate(index = "spline2d") |> select(-id, -info)) 
+dplyr::glimpse(sim_sine_6d_spline_head |> select(-id, -info)) 
 ```
@@
 "smooth" is used as the smoothness measure and the column "squint" is
-calculated as [TODO: equation reference] as squintability.
-The low squint value of MIC and TIC index is due to its convex shape of
-the index value against the projection distance, as shown in
+calculated as [TODO: equation reference] as squintability. The low
+squint value of MIC and TIC index is due to its convex shape of the
+index value against the projection distance, as shown in
 @fig-idx-proj-dist. Comparing to other concave shapes, the maximum first

### 2024-06-19 optim.qmd
Update optim.qmd

@@
 
-This study aims to investigate ...
-
-The contribution of this study to the high dimensional data visualisation literature is twofold. First, the performance of JSO is studied in the context of PP guided tour.
-Second, to the best of the authors’ knowledge, this is the first attempt
-to quantitatively measure squintability and smoothness.
+The primary goal of this study is to investigate the performance of JSO in the context of the PP guided tour.
+A series of simulation experiments using well-recognised data sets and PPIs are conducted to yield insights regarding the behaviour of JSO under different settings and the sensitivity of hyper-parameter in the optimisation.
+Second, to observe the performance of JSO with different types of PPI, a collection of metrics is introduced to capture specific properties of the index including squintability and smoothness as introduced in @laa_using_2020.
+To the best of the author’s knowledge, this is the first attempt to quantitatively measure squintability and smoothness.
+Finally, the relationship between the JSO performance, hyper-parameters tuning and various properties of PPI is analysed to provide helpful guidance for practitioners that are using the guided tour for high dimensional visualisation.
 

### 2024-06-19 optim.qmd
more tweaks

--- 2024-06-14 optim.qmd
@@
 "smooth" is used as the smoothness measure and the column "squint" is
-calculated as [TODO: equation reference] as squintability.
-The low squint value of MIC and TIC index is due to its convex shape of
-the index value against the projection distance, as shown in
+calculated as \[TODO: equation reference\] as squintability. The low
+squint value of MIC and TIC index is due to its convex shape of the
+index value against the projection distance, as shown in
 @fig-idx-proj-dist. Comparing to other concave shapes, the maximum first
@@
 sq_basis_dist_idx |>
-  ggplot(aes(x = dist, y = idx, group = index)) +
-  geom_line() +
+  ggplot(aes(x = dist, y = index, group = index_name)) +
+  geom_point() +
   facet_nested_wrap(~index + n, nrow = 3, labeller = label_both) + 
   xlab("projection distance") + ylab("index value") 
-
 ```
@@
          max_tries = max_tries/10,
-         long_time = ifelse(time > 30, 1, 0),
+         long_time = ifelse(time > 20, 1, 0),
          P_J = ifelse(index == "stringy", 0, P_J))

### 2024-06-19 optim.tex
more tweaks

@@
 
-\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, toptitle=1mm, left=2mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, titlerule=0mm, toprule=.15mm, opacityback=0, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, breakable, arc=.35mm, rightrule=.15mm, leftrule=.75mm, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, left=2mm, breakable, toprule=.15mm, colframe=quarto-callout-note-color-frame, opacityback=0, titlerule=0mm, bottomtitle=1mm, colback=white, toptitle=1mm, rightrule=.15mm, bottomrule=.15mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, opacitybacktitle=0.6, coltitle=black]
 
@@
 \texttt{spline}, and \texttt{stringy}. Combining with two jellyfish
-hyper-parameters, a total of 42 cases is produced, comprising of 24
-pipe-finding cases and 8 sine-wave finding cases. For each case, the JS
+hyper-parameters, a total of 52 cases is produced, comprising of 24
+pipe-finding cases and 28 sine-wave finding cases. For each case, the JS
 optimiser is run fifty times and the success rate is calculated as in
@@
 \midrule
-1 & pipe & holes & 6 & 0.002 & 0.408 & 2.364 & 0.212 & 0.991 & 0.860 & 3.368 & 0.172 & 0.627\\
-2 & pipe & holes & 8 & 0.000 & 0.259 & 2.373 & 0.613 & 1.011 & 0.869 & 3.264 & 0.085 & 0.696\\
-3 & pipe & holes & 10 & 0.000 & 0.144 & 2.317 & 1.831 & 1.011 & 0.885 & 3.151 & 0.153 & 0.636\\
-4 & pipe & holes & 12 & 0.000 & 0.254 & 2.173 & 0.879 & 0.999 & 0.878 & 3.345 & 0.196 & 0.621\\
-5 & sine & MIC & 6 & 0.016 & 0.100 & 2.394 & 0.087 & 0.894 & 0.571 & 1.623 & -0.024 & 0.314\\
+1 & pipe & holes & 6 & 0.002 & 0.371 & 2.355 & 0.221 & 1.015 & 0.860 & 3.368 & 0.018 & 0.763\\
+2 & pipe & holes & 8 & 0.000 & 0.178 & 2.189 & 0.820 & 1.014 & 0.869 & 3.264 & 0.029 & 0.740\\
+3 & pipe & holes & 10 & 0.000 & 0.107 & 2.192 & 3.027 & 1.016 & 0.885 & 3.151 & 0.022 & 0.737\\
+4 & pipe & holes & 12 & 0.000 & 0.148 & 2.295 & 1.581 & 1.011 & 0.878 & 3.345 & 0.004 & 0.779\\
+5 & sine & MIC & 6 & 0.016 & 0.092 & 2.457 & 0.083 & 0.894 & 0.571 & 1.623 & -0.024 & 0.314\\
 \addlinespace
-6 & sine & MIC & 8 & 0.016 & 0.100 & 2.394 & 0.087 & 0.932 & 0.328 & 1.314 & -0.030 & 0.193\\
-7 & sine & TIC & 6 & 0.124 & 0.104 & 2.471 & 0.086 & 0.951 & 0.536 & 1.719 & -0.025 & 0.330\\
-8 & sine & TIC & 8 & 0.124 & 0.104 & 2.471 & 0.086 & 0.945 & 0.564 & 1.723 & -0.027 & 0.341\\
+6 & sine & MIC & 8 & 0.016 & 0.081 & 2.636 & 0.084 & 0.932 & 0.328 & 1.314 & -0.030 & 0.193\\
+7 & sine & TIC & 6 & 0.124 & 0.110 & 2.444 & 0.086 & 0.951 & 0.536 & 1.719 & -0.027 & 0.330\\
+8 & sine & TIC & 8 & 0.125 & 0.099 & 2.475 & 0.085 & 0.949 & 0.564 & 1.723 & -0.028 & 0.342\\
 9 & sine & dcor2d & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019 & 0.737\\
-10 & sine & loess2d & 6 & 0.083 & 0.307 & 2.194 & 0.292 & 1.016 & 1.039 & 2.648 & 0.080 & 0.689\\
+10 & sine & loess2d & 6 & 0.079 & 0.336 & 1.988 & 0.309 & 1.016 & 1.039 & 2.648 & 0.080 & 0.689\\
 \addlinespace
-11 & sine & splines2d & 6 & 0.040 & 0.189 & 2.606 & 0.104 & 1.014 & 1.051 & 2.730 & -0.009 & 0.780\\
-12 & sine & stringy & 6 & 0.000 & 1173.035 & 1.031 & 17608.047 & 1.040 & 0.011 & 254.733 & 0.170 & 0.648\\
+11 & sine & splines2d & 6 & 0.042 & 0.242 & 2.537 & 0.102 & 1.014 & 1.051 & 2.730 & -0.009 & 0.780\\
+12 & sine & stringy & 6 & 0.000 & 0.007 & 1.536 & 38.166 & 1.045 & 0.011 & 254.734 & 0.053 & 0.739\\
 \bottomrule
@@
 \midrule
-dcor2d\_2 & 6 & 2.663 & 0.737 & 20 & 50 & 0.44 & 35.320 secs\\
-dcor2d\_2 & 6 & 2.663 & 0.737 & 20 & 100 & 0.60 & 70.347 secs\\
-dcor2d\_2 & 6 & 2.663 & 0.737 & 50 & 50 & 0.90 & 87.158 secs\\
-holes & 6 & 2.364 & 0.627 & 20 & 50 & 0.88 & 0.801 secs\\
-holes & 6 & 2.364 & 0.627 & 20 & 100 & 0.96 & 1.456 secs\\
+MIC & 6 & 2.457 & 0.314 & 20 & 50 & 0.12 & 2.479 secs\\
+MIC & 6 & 2.457 & 0.314 & 20 & 100 & 0.24 & 8.950 secs\\
+MIC & 6 & 2.457 & 0.314 & 50 & 50 & 0.52 & 5.651 secs\\
+MIC & 6 & 2.457 & 0.314 & 50 & 100 & 0.64 & 13.223 secs\\
+MIC & 6 & 2.457 & 0.314 & 100 & 50 & 0.76 & 19.453 secs\\
 \addlinespace
-holes & 6 & 2.364 & 0.627 & 50 & 50 & 1.00 & 1.477 secs\\
-holes & 6 & 2.364 & 0.627 & 50 & 100 & 1.00 & 3.052 secs\\
+MIC & 8 & 2.636 & 0.193 & 20 & 50 & 0.08 & 2.566 secs\\
+MIC & 8 & 2.636 & 0.193 & 20 & 100 & 0.08 & 4.960 secs\\
 \bottomrule
@@
 \midrule
-Intercept & 2.685 & 15.464 & 0.174 & 0.862\\
-Smoothness & 3.070 & 3.631 & 0.845 & 0.398\\
-Squintability & -8.545 & 22.252 & -0.384 & 0.701\\
-dimension (d) & -0.610 & 0.355 & -1.719 & 0.086\\
-long time & -1.683 & 1.930 & -0.872 & 0.383\\
+Intercept & -4.523 & 5.332 & -0.848 & 0.396\\
+Smoothness & 1.195 & 1.915 & 0.624 & 0.533\\
+Squintability & 8.241 & 2.737 & 3.011 & 0.003\\
+dimension (d) & -0.628 & 0.255 & -2.462 & 0.014\\
+long time & -0.874 & 1.286 & -0.680 & 0.497\\
 \addlinespace
-number of jellyfish & 0.269 & 0.175 & 1.538 & 0.124\\
-maximum number of tries & 0.080 & 0.196 & 0.410 & 0.682\\
+number of jellyfish & 0.216 & 0.127 & 1.701 & 0.089\\
+maximum number of tries & 0.113 & 0.150 & 0.752 & 0.452\\
 \bottomrule

### 2024-06-19 code/04-sim-sine-68d-MICTIC.R

@@ -75,5 +75,5 @@ t2 <- Sys.time()
 t2 - t1
 
 sim_sine_68d_TICMIC <- sim_res |> select(-alpha) |> rename(index = idx_f)
-save(sim_sine_68d_TICMIC, file = "data/sim_sine_68d_TICMIC.rda")
+save(sim_sine_68d_TICMIC, file = "data-raw/sim_sine_68d_TICMIC.rda")
 

### 2024-06-19 code/07-calc-smoothness-squintability.R

@@ -16,11 +16,14 @@ smoothness_holes <- tibble::tibble(
               matrix(c(rep(0, 12), 1, 0, 0, 1), nrow = 8, byrow = TRUE),
               matrix(c(rep(0, 16), 1, 0, 0, 1), nrow = 10, byrow = TRUE),
               matrix(c(rep(0, 20), 1, 0, 0, 1), nrow = 12, byrow = TRUE))) |>
-  dplyr::mutate(smoothness = purrr::pmap(
-    list(data, best, n),
-    function(data, best, n){calc_smoothness("holes", data = data, n = n, best = best)}))
+  dplyr::mutate(basis_df = purrr::pmap(
+    list(data, best),
+    function(data, best, n){sample_bases("holes", data = data, best = best)}))
 
-holes_tidy <- smoothness_holes |> tidyr::unnest_wider(smoothness) |> unnest(measure)
+holes_tidy <- smoothness_holes |>
+  rowwise() |>
+  dplyr::mutate(smooth = calc_smoothness(basis_df)) |>
+  unnest(smooth)
 
 idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "stringy", "splines2d")
 smoothness_sine <- tibble::tibble(
@@ -29,10 +32,12 @@ smoothness_sine <- tibble::tibble(
   dplyr::bind_rows(sine_8d_tbl("MIC")) |>
   dplyr::bind_rows(sine_8d_tbl("TIC")) |>
   rowwise() |>
-  dplyr::mutate(smoothness = list(calc_smoothness(index)))
+  dplyr::mutate(basis_df = list(sample_bases(index)))
 
-
-sine_tidy <- smoothness_sine |> tidyr::unnest_wider(smoothness) |> unnest(measure)
+sine_tidy <- smoothness_sine |>
+  rowwise() |>
+  dplyr::mutate(smooth = calc_smoothness(basis_df)) |>
+  unnest(smooth)
 
 smoothness <- bind_rows(sine_tidy, holes_tidy) |>
   mutate(index = factor(index, levels = c("holes", "MIC", "TIC", "dcor2d_2",
@@ -43,7 +48,8 @@ save(smoothness, file = here::here("data", "smoothness.rda"))
 
 ################################################################################
 ################################################################################
-# squintability
+# squintability - sample bases
+# 2-3 mins
 sq_holes_basis_df <- tibble::tibble(
   n = c(6, 8, 10, 12),
[…]

### 2024-06-19 code/08-process-data-for-modelling.R

@@ -26,12 +26,12 @@ pipe_setup_summ <- pipe_run_df |>
 load(here::here("data-raw/sim_sine_6d_dcor2d.rda"))
 load(here::here("data-raw/sim_sine_6d_loess2d.rda"))
 load(here::here("data-raw/sim_sine_68d_TICMIC.rda"))
-load(here::here("data-raw/sim_sine_6d_spline.rda"))
+load(here::here("data-raw/sim_sine_6d_splines2d.rda"))
 load(here::here("data-raw/sim_sine_6d_stringy.rda"))
 
 sim_data_sine <- bind_rows(
   sim_sine_6d_dcor2d, sim_sine_6d_loess2d, sim_sine_68d_TICMIC,
-  sim_sine_6d_spline, sim_sine_6d_stringy)
+  sim_sine_6d_splines2d, sim_sine_6d_stringy)
 
 sine_run_df <- sim_data_sine |>
   mutate(id2 = paste0(index, n_jellies, max_tries, sim, d)) |>

### 2024-06-19 optim.qmd
add abstract

@@
 abstract: |
-  This is the abstract.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit.
-Nam placerat turpis purus, eu tristique ex tincidunt et.
-Mauris sed augue eget turpis ultrices tincidunt.
-Sed et mi in leo porta egestas.
-Aliquam non laoreet velit.
-Nunc quis ex vitae eros aliquet auctor nec ac libero.
-Duis laoreet sapien eu mi luctus, in bibendum leo molestie.
-Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
-Aliquam bibendum varius libero, eu efficitur justo rutrum at.
-Sed at tempus elit.
+  Projection pursuit (PP) is a dimension reduction technique that identifies low-dimensional projections in high-dimensional data by optimising a criteria function known as the PP index.
+The optimisation in PP can be non-smooth and requires identifying optima with a small "squint angle", detectable only from close proximity.
+To address these challenges, this study investigates the performance of a recent swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
+The performance of JSO is evaluated across various hyper-parameter settings and compared with existing optimisers.
+Additionally, this work proposes novel methods to quantify two properties of the PP index -- smoothness and squintability -- that capture the complexities inherent in PP optimisation problems.
+These two metrics are evaluated with JSO hyper-parameters to determine their effect on JSO success rate.
+The JSO algorithm has been implemented in the `tourr` package, while calculations for smoothness and squintability are available in the `ferrn` package.
 keywords: 
   - projection pursuit
-  - optimization
-  - jellyfish optimiser
-  - data visualisation
-  - high-dimensional data
+  - Jellyfish Search Optimiser (JSO)
+  - Optimisation
 date: last-modified
@@
 ```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
 #install the latest version of ferrn from: https://github.com/huizezhang-sherry/ferrn
@@
 high-dimensional data $X \in \mathcal{R}^{n }\times p$ and PPI
-\$f(\\cdot)\$, PP finds the orthonormal projection matrix
+$f(\cdot)$, PP finds the orthonormal projection matrix
 $A \in \mathcal{R}^{p \times d}$ by solving the following optimisation

### 2024-06-19 optim.tex
add abstract

@@
   pdftitle={Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit},
-  pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Leung},
-  pdfkeywords={projection pursuit, optimization, jellyfish
-optimiser, data visualisation, high-dimensional data},
+  pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
+  pdfkeywords={projection pursuit, Jellyfish Search Optimiser
+(JSO), Optimisation},
   colorlinks=true,
@@
  \ead{nicolaslangrene@uic.edu.cn} 
-\author[2]{Jessica Leung%
+\author[2]{Jessica Wai Yin Leung%
 %
@@
 \begin{abstract}
-This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing
-elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget
-velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris
-sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta
-egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet
-auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo
-molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
-Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at
-tempus elit.
+Projection pursuit (PP) is a dimension reduction technique that
+identifies low-dimensional projections in high-dimensional data by
+optimising a criteria function known as the PP index. The optimisation
+in PP can be non-smooth and requires identifying optima with a small
+``squint angle'', detectable only from close proximity. To address these
+challenges, this study investigates the performance of a recent
+swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising
+PP indexes. The performance of JSO is evaluated across various
+hyper-parameter settings and compared with existing optimisers.
+Additionally, this work proposes novel methods to quantify two
+properties of the PP index -- smoothness and squintability -- that
+capture the complexities inherent in PP optimisation problems. These two
+metrics are evaluated with JSO hyper-parameters to determine their
+effect on JSO success rate. The JSO algorithm has been implemented in
+the \texttt{tourr} package, while calculations for smoothness and
+squintability are available in the \texttt{ferrn} package.
 \end{abstract}
@@
 \begin{keyword}
-    projection pursuit \sep optimization \sep jellyfish
-optimiser \sep data visualisation \sep 
-    high-dimensional data
+    projection pursuit \sep Jellyfish Search Optimiser (JSO) \sep 
+    Optimisation
 \end{keyword}
@@
 
+Projection Pursuit (PP) is a high-dimensional data visualisation
+approach that involves computing informative linear projections of data
+by optimising a variety of objective functions, namely \emph{projection
+pursuit index (PPI)}, (\citet{hall1989polynomial},
+\citet{cook1993projection}, \citet{lee2010projection},
+\citet{Loperfido2018}, \citet{Loperfido2020}). Given high-dimensional
+data \(X \in \mathcal{R}^{n }\times p\) and PPI \(f(\cdot)\), PP finds
+the orthonormal projection matrix \(A \in \mathcal{R}^{p \times d}\) by
+solving the following optimisation problem:
+
+\[
+\underset{A}{\max} f(XA) \quad \text{subject to} \quad A'A = I
+\]
+
+These indices \(f(\cdot)\) are characterised by the informativeness or
+``interestingness'' of a projection and are often non-linear and
+non-convex. As such, an effective and efficient optimisation procedure
+is essential to enable sufficient exploration of the data landscape
+during the visualisation process and to arrive at a global optimal
+viewpoint of the data.
+
+Significant efforts have been devoted to enhancing the functionality of
+PP and facilitating a better experience in the visualisation process.
+\citet{cook1995grand} introduced the PP guided tour, which enabled
+interactive visualisation of the optimisation to visually explore
+high-dimensional data. It is implemented in the R \citep{R} package
+\texttt{tourr}\citep{tourr}. \citet{RJ-2021-105} highlighted potential
+problems of the optimisation routine implemented. While improving the
+quality of the optimisation solutions in the tour is essential, it is
+also important to be able to watch the projected data as the
+optimisation progresses. Therefore, it is of interest to integrate the
+guided tour with a global optimisation algorithm and examine the
+characteristics and advantages of such implementation.
+
 The artificial jellyfish search (JS) algorithm \citep{chou_novel_2021}
 is a swarm-based metaheuristic optimisation algorithm inspired by the
-search behaviour of jellyfish in the ocean. It is one of the newest
+search behaviour of jellyfish in the ocean. It is one of the latest
 swarm intelligence algorithms \citep{rajwar_exhaustive_2023}, which was
-shown to have stronger search ability and faster convergence with few
-algorithmic parameters compared to classic optimization methods
+shown to have stronger search ability and faster convergence with fewer
+parameters compared to classic optimisation methods
 \citep{chou_novel_2021}-\citep{chou_recent_2022}.
 
-Effective optimisation is an important aspect of many methods employed
-for visualising high-dimensional data (\(X\)). Here we are concerned
-about computing informative linear projections of high-dimensional
-(\(p\)) data using projection pursuit (PP) (\citet{kr69}, \citet{FT74}).
-This involves optimising a function (e.g. \citet{hall1989polynomial},
-\citet{cook1993projection}, \citet{lee2010projection},
-\citet{Loperfido2018}, \citet{Loperfido2020}), called the projection
-pursuit index (PPI), that defines what is interesting or informative in
-a projection.
+The primary goal of this study is to investigate the performance of JSO
+in the context of the PP guided tour. A series of simulation experiments
+using well-recognised data sets and PPIs are conducted to yield insights
+regarding the behaviour of JSO under different settings and the
+sensitivity of hyper-parameter in the optimisation. Second, to observe
+the performance of JSO with different types of PPI, a collection of
+metrics is introduced to capture specific properties of the index
+including squintability and smoothness as introduced in
+\citet{laa_using_2020}. To the best of the author's knowledge, this is
+the first attempt to quantitatively measure squintability and
+smoothness. Finally, the relationship between the JSO performance,
+hyper-parameters tuning and various properties of PPI is analysed to
+provide helpful guidance for practitioners that are using the guided
+tour for high dimensional visualisation.
 
-These PPI are defined on projections (\(XA\)), which means that there is
-a constraint that needs to be considered when optimising. A projection
-of data is defined by a \(p\times d\) orthonormal matrix \(A\), and this
-imposes the constraint on the elements of \(A\), that columns need have
-norm equal to 1 and the product of columns need to sum to zero.
-
-\citet{cook1995grand} introduced the PP guided tour, which enabled
-interactive visualisation of the optimisation in order to visually
-explore high-dimensional data. It is implemented in the R \citep{R}
-package \texttt{tourr} \citep{tourr}. The optimisation that is
-implemented is fairly basic, and potential problems were highlighted by
-\citet{RJ-2021-105}. Implementing better optimisation functionality is a
-goal, but it needs to be kept in mind that the guided tour also has
-places importance on watching the projected data as the optimisation
-progresses.
-
-Here we explore the potential for a jellyfish optimisation to be
-integrated with the guided tour. Section~\ref{sec-background} explains
-the optimisation that is used in the current the projection pursuit
-guided tour. Section~\ref{sec-theory} provides more details on the
-jellyfish optimiser and formalises several characteristics of projection
-pursuit indexes that are help to measure optimiser performance.
-\textbf{?@sec-simulation} describes a simulation study on performance of
-the jellyfish for several types of data and index functions.
+The rest of this paper is structured as follows.
+Section~\ref{sec-background} introduces the PP guided tour.
+Section~\ref{sec-theory} describes the details of JSO and introduces
+metrics that measure different characteristics of a PPI.
+\textbf{?@sec-simulation} describes a simulation study on the
+performance of the JSO using well-known data sets and index functions.
 Section~\ref{sec-conclusion} summarises the work and provides
@@
 
-\begin{tcolorbox}[enhanced jigsaw, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, left=2mm, breakable, toprule=.15mm, colframe=quarto-callout-note-color-frame, opacityback=0, titlerule=0mm, bottomtitle=1mm, colback=white, toptitle=1mm, rightrule=.15mm, bottomrule=.15mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, opacitybacktitle=0.6, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, opacityback=0, toprule=.15mm, leftrule=.75mm, coltitle=black, bottomtitle=1mm, arc=.35mm, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, opacitybacktitle=0.6, left=2mm, breakable]
 

### 2024-06-20 optim.qmd
intro

@@
 by optimising a variety of objective functions, namely *projection
-pursuit index (PPI)*, (@hall1989polynomial, @cook1993projection,
+pursuit index (PPI)* (@hall1989polynomial, @cook1993projection,
 @lee2010projection, @Loperfido2018, @Loperfido2020). Given
-high-dimensional data $X \in \mathcal{R}^{n }\times p$ and PPI
+high-dimensional data $X \in \mathbf{R}^{n }\times p$ and PPI
 $f(\cdot)$, PP finds the orthonormal projection matrix
-$A \in \mathcal{R}^{p \times d}$ by solving the following optimisation
+$A \in \mathbf{R}^{p \times d}$ by solving the following optimisation
 problem:
@@
 $$
-\underset{A}{\max} f(XA) \quad \text{subject to} \quad A'A = I
+\underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
 $$
@@
 
-The artificial jellyfish search (JS) algorithm [@chou_novel_2021] is a
+The artificial Jellyfish Search Optimiser (JSO) [@chou_novel_2021] is a
 swarm-based metaheuristic optimisation algorithm inspired by the search
@@
 Second, to observe the performance of JSO with different types of PPI, a collection of metrics is introduced to capture specific properties of the index including squintability and smoothness as introduced in @laa_using_2020.
-To the best of the author’s knowledge, this is the first attempt to quantitatively measure squintability and smoothness.
+To the best of the authors’ knowledge, this is the first attempt to quantitatively measure squintability and smoothness.
 Finally, the relationship between the JSO performance, hyper-parameters tuning and various properties of PPI is analysed to provide helpful guidance for practitioners that are using the guided tour for high dimensional visualisation.
@@
 The rest of this paper is structured as follows. @sec-background
-introduces the PP guided tour. @sec-theory describes the details of JSO
-and introduces metrics that measure different characteristics of a PPI.
-@sec-simulation describes a simulation study on the performance of the
-JSO using well-known data sets and index functions. @sec-conclusion
-summarises the work and provides suggestions for future directions.
+introduces the background of the PP guided tour and reviews existing methods in the literature. @sec-theory describes the details of JSO
+and introduces metrics that measure different properties of a PPI.
+@sec-vis visualises the behaviour of the JSO in terms of describes a simulation study on the performance of the JSO using well-known data sets and index functions.
+@sec-app displays two sets of simulation experiments comparing the JSO with the search-better optimiser and studies the impact of different PPI properties on the optimisation performance.
+@sec-conclusion summarises the work and provides suggestions for future directions.
 
@@
 
-The speed of optimizing an index function can be calculated/measured
-using the computational complexity (in big O notation, with respect to
+The speed of optimizing an index function can be measured empirically by recording the duration of the optimisation.
+Alternatively, one can also gauge the speed in terms of computational complexity of evaluating the index function (for instance, in big O notation, with respect to
 the sample size) of computing the index function.

### 2024-06-20 optim.tex
intro

@@
 \makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \end{frontmatter}
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, frame hidden, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 by optimising a variety of objective functions, namely \emph{projection
-pursuit index (PPI)}, (\citet{hall1989polynomial},
+pursuit index (PPI)} (\citet{hall1989polynomial},
 \citet{cook1993projection}, \citet{lee2010projection},
 \citet{Loperfido2018}, \citet{Loperfido2020}). Given high-dimensional
-data \(X \in \mathcal{R}^{n }\times p\) and PPI \(f(\cdot)\), PP finds
-the orthonormal projection matrix \(A \in \mathcal{R}^{p \times d}\) by
+data \(X \in \mathbf{R}^{n }\times p\) and PPI \(f(\cdot)\), PP finds
+the orthonormal projection matrix \(A \in \mathbf{R}^{p \times d}\) by
 solving the following optimisation problem:
@@
 \[
-\underset{A}{\max} f(XA) \quad \text{subject to} \quad A'A = I
+\underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
 \]
@@
 
-The artificial jellyfish search (JS) algorithm \citep{chou_novel_2021}
+The artificial Jellyfish Search Optimiser (JSO) \citep{chou_novel_2021}
 is a swarm-based metaheuristic optimisation algorithm inspired by the
@@
 including squintability and smoothness as introduced in
-\citet{laa_using_2020}. To the best of the author's knowledge, this is
+\citet{laa_using_2020}. To the best of the authors' knowledge, this is
 the first attempt to quantitatively measure squintability and
@@
 The rest of this paper is structured as follows.
-Section~\ref{sec-background} introduces the PP guided tour.
+Section~\ref{sec-background} introduces the background of the PP guided
+tour and reviews existing methods in the literature.
 Section~\ref{sec-theory} describes the details of JSO and introduces
-metrics that measure different characteristics of a PPI.
-\textbf{?@sec-simulation} describes a simulation study on the
-performance of the JSO using well-known data sets and index functions.
-Section~\ref{sec-conclusion} summarises the work and provides
-suggestions for future directions.
+metrics that measure different properties of a PPI.
+Section~\ref{sec-vis} visualises the behaviour of the JSO in terms of
+describes a simulation study on the performance of the JSO using
+well-known data sets and index functions. Section~\ref{sec-app} displays
+two sets of simulation experiments comparing the JSO with the
+search-better optimiser and studies the impact of different PPI
+properties on the optimisation performance. Section~\ref{sec-conclusion}
+summarises the work and provides suggestions for future directions.
 
+\hypertarget{sec-background}{%
 \section{Projection pursuit, index functions and optimisation {[}Di and
-Sherry{]}}\label{sec-background}
+Sherry{]}}\label{sec-background}}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, opacityback=0, toprule=.15mm, leftrule=.75mm, coltitle=black, bottomtitle=1mm, arc=.35mm, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, opacitybacktitle=0.6, left=2mm, breakable]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, leftrule=.75mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, opacityback=0, bottomtitle=1mm, breakable, rightrule=.15mm, bottomrule=.15mm, titlerule=0mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, coltitle=black]
 
@@
 
-\subsection{Smoothness}\label{sec-smoothness}
+\hypertarget{sec-smoothness}{%
+\subsection{Smoothness}\label{sec-smoothness}}
 
@@
 
-\subsection{Squintability}\label{sec-squintability}
+\hypertarget{sec-squintability}{%
+\subsection{Squintability}\label{sec-squintability}}
 
@@
 
-\subsection{Speed}\label{speed}
+\hypertarget{speed}{%
+\subsection{Speed}\label{speed}}
 
-The speed of optimizing an index function can be calculated/measured
-using the computational complexity (in big O notation, with respect to
-the sample size) of computing the index function.
+The speed of optimizing an index function can be measured empirically by
+recording the duration of the optimisation. Alternatively, one can also
+gauge the speed in terms of computational complexity of evaluating the
+index function (for instance, in big O notation, with respect to the
+sample size) of computing the index function.
 
-\section{Visualisation of jellyfish optimiser}\label{sec-vis}
+\hypertarget{sec-vis}{%
+\section{Visualisation of jellyfish optimiser}\label{sec-vis}}
 
@@
 
-\centering{
-
-\includegraphics{optim_files/figure-pdf/fig-projection-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/fig-projection-1.pdf}
 
@@
 
-\end{figure}%
+\end{figure}
 
@@
 
-\centering{
-
-\includegraphics{optim_files/figure-pdf/fig-index-value-proj-dist-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/fig-index-value-proj-dist-1.pdf}
 
@@
 
-\end{figure}%
+\end{figure}
 
-\section{Application {[}Di and Sherry{]}}\label{sec-app}
+\hypertarget{sec-app}{%
+\section{Application {[}Di and Sherry{]}}\label{sec-app}}
 
@@
 
+\hypertarget{sec-app-1}{%
 \subsection{Performance of the JS optimiser in pipe-finding
-problems}\label{sec-app-1}
+problems}\label{sec-app-1}}
 
@@
 
-\centering{
-
-\includegraphics{optim_files/figure-pdf/fig-proj-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/fig-proj-1.pdf}
 
@@
 
-\centering{
-
-\includegraphics{optim_files/figure-pdf/fig-proportion-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/fig-proportion-1.pdf}
 
@@
 
+\hypertarget{sec-app-2}{%
 \subsection{Factors affecting the JS optimiser success rate:
-optimisation properties and jellyfish hyper-parameters}\label{sec-app-2}
+optimisation properties and jellyfish
+hyper-parameters}\label{sec-app-2}}
 
@@
 
-\centering{
-
-\includegraphics{optim_files/figure-pdf/fig-idx-proj-dist-1.pdf}
+{\centering \includegraphics{optim_files/figure-pdf/fig-idx-proj-dist-1.pdf}
 
@@
 
+\hypertarget{tbl-smoothness-squintability}{}
 \begin{table}
-
-\caption{\label{tbl-smoothness-squintability}Parameters estimated from
-the Gaussian process (including variance, range, smoothness, and nugget)
-and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
-pipe-finding and sine-wave finding problems. The column ``smooth'' and
-``squint'' represent the smoothness and squintability measures.}
-
-\centering{
+\caption{\label{tbl-smoothness-squintability}Parameters estimated from the Gaussian process (including variance,
+range, smoothness, and nugget) and scaled logistic function
+(\(\theta_1\) to \(\theta_4\)) for the pipe-finding and sine-wave
+finding problems. The column ``smooth'' and ``squint'' represent the
+smoothness and squintability measures. }\tabularnewline
 
@@
 \endgroup{}
-
-}
-
-\end{table}%
+\end{table}
 
@@
 
+\hypertarget{tbl-mod-data}{}
 \begin{longtable}[t]{lrrrrrrl}
-
-\caption{\label{tbl-mod-data}The first 7 rows of the datasets processed
-for modelling.}
-
-\tabularnewline
+\caption{\label{tbl-mod-data}The first 7 rows of the datasets processed for modelling. }\tabularnewline
 
@@
 \bottomrule
-
 \end{longtable}
-
 \endgroup{}
@@
 
+\hypertarget{tbl-mod-output}{}
 \begin{longtable}[t]{|>{}lrrr>{}r|}
-
-\caption{\label{tbl-mod-output}Model estimates of proportion of
-jellyfish success on optimisation properties and jellyfish
-hyper-parameters from the generalised linear model with a binomial
-family and a logit link function. The variable smoothness,
-squintability, number of jellyfish and maximum number of tries are
-positively associated with the JS optimiser success rate while data
-dimension and being flagged as long runtime are negatively associated
-with the success rate. The variable squintability and dimension are
-significant, suggesting their importance relative to jellyfish
-hyper-parameters in the optimisation success.}
-
-\tabularnewline
+\caption{\label{tbl-mod-output}Model estimates of proportion of jellyfish success on optimisation
+properties and jellyfish hyper-parameters from the generalised linear
+model with a binomial family and a logit link function. The variable
+smoothness, squintability, number of jellyfish and maximum number of
+tries are positively associated with the JS optimiser success rate while
+data dimension and being flagged as long runtime are negatively
+associated with the success rate. The variable squintability and
+dimension are significant, suggesting their importance relative to
+jellyfish hyper-parameters in the optimisation success. }\tabularnewline
 
@@
 \bottomrule
-
 \end{longtable}
-
 \endgroup{}
 
-\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
+\hypertarget{sec-conclusion}{%
+\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}}
 

### 2024-06-20 optim.qmd
update intro

@@
 in the tour is essential, it is also important to be able to watch the
-projected data as the optimisation progresses. Therefore, it is of
-interest to integrate the guided tour with a global optimisation
-algorithm and examine the characteristics and advantages of such
-implementation.
+projected data as the optimisation progresses. As such, it is pertinent to integrate the guided tour with a global optimisation
+algorithm that is efficient in finding the global optimal and enables viewing of the projected data during the exploration process.
 
-The artificial Jellyfish Search Optimiser (JSO) [@chou_novel_2021] is a
-swarm-based metaheuristic optimisation algorithm inspired by the search
-behaviour of jellyfish in the ocean. It is one of the latest swarm
-intelligence algorithms [@rajwar_exhaustive_2023], which was shown to
-have stronger search ability and faster convergence with fewer
-parameters compared to classic optimisation methods
-[@chou_novel_2021]-[@chou_recent_2022].
+The artificial Jellyfish Search Optimiser (JSO) [@chou_novel_2021] is a swarm-based metaheuristic designed to solve global optimisation problems.
+Inspired by the search behaviour of jellyfish in the ocean, JSO is one of the latest swarm intelligence algorithms [@rajwar_exhaustive_2023], which was shown to have stronger search ability and faster convergence with fewer parameters compared to classic optimisation methods [@chou_novel_2021]-[@chou_recent_2022].
+These practical properties makes JSO a strong candidate in integration with the PP guided tour.
+As such, it is of interest to explore the potential of JSO in enhancing the PP guided tour and examine the characteristics and advantages of such implementation.
 

### 2024-06-20 optim.tex
update intro

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, frame hidden, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 also important to be able to watch the projected data as the
-optimisation progresses. Therefore, it is of interest to integrate the
-guided tour with a global optimisation algorithm and examine the
-characteristics and advantages of such implementation.
+optimisation progresses. As such, it is pertinent to integrate the
+guided tour with a global optimisation algorithm that is efficient in
+finding the global optimal and enables viewing of the projected data
+during the exploration process.
 
 The artificial Jellyfish Search Optimiser (JSO) \citep{chou_novel_2021}
-is a swarm-based metaheuristic optimisation algorithm inspired by the
-search behaviour of jellyfish in the ocean. It is one of the latest
-swarm intelligence algorithms \citep{rajwar_exhaustive_2023}, which was
-shown to have stronger search ability and faster convergence with fewer
-parameters compared to classic optimisation methods
-\citep{chou_novel_2021}-\citep{chou_recent_2022}.
+is a swarm-based metaheuristic designed to solve global optimisation
+problems. Inspired by the search behaviour of jellyfish in the ocean,
+JSO is one of the latest swarm intelligence algorithms
+\citep{rajwar_exhaustive_2023}, which was shown to have stronger search
+ability and faster convergence with fewer parameters compared to classic
+optimisation methods \citep{chou_novel_2021}-\citep{chou_recent_2022}.
+These practical properties makes JSO a strong candidate in integration
+with the PP guided tour. As such, it is of interest to explore the
+potential of JSO in enhancing the PP guided tour and examine the
+characteristics and advantages of such implementation.
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, leftrule=.75mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, opacityback=0, bottomtitle=1mm, breakable, rightrule=.15mm, bottomrule=.15mm, titlerule=0mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, breakable, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, left=2mm, titlerule=0mm, colframe=quarto-callout-note-color-frame, coltitle=black, opacitybacktitle=0.6, colback=white, toprule=.15mm, bottomtitle=1mm, arc=.35mm, bottomrule=.15mm, opacityback=0]
 

### 2024-06-20 optim.qmd
sec 3 update text

@@
 
-The jellyfish optimiser (JSO) mimics the natural movements of jellyfish,
-which include passive and active motions driven by ocean currents and
-their swimming patterns, respectively. In the context of optimization,
-these movements are abstracted to explore the search space in a way that
+JSO mimics the natural movements of jellyfish, which include passive and active motions driven by ocean currents and
+their swimming patterns, respectively. In the context of optimization, these movements are abstracted to explore the search space in a way that
 balances exploration (searching new areas) and exploitation (focusing on
 promising areas). The algorithm aims to find the optimal solution by
-adapting the jellyfish's behaviour to navigate towards the best solution
-over iterations [@chou_novel_2021].
+adapting the behaviour of jellyfish to navigate towards the best solution over iterations [@chou_novel_2021].
 
-To understand what the jellyfish optimizer is doing in the context of
-Projection Pursuit, we first start with a current projection (the
-starting point). Then, we evaluate this projection using an index
-function, which tells us how good the current projection is. We then
-move the projection in a direction determined by the 'best jelly' and
-random factors, influenced by how far along we are in the optimization
-process (the trial $i$ and `max.tries`). Occasionally, we might explore
-completely new directions like a jellyfish might with ocean currents.
-Then, we compare new potential projections to our current one. If
-they're better, we adopt them; if not, we stick with our current
-projection. This process continues and iteratively improves the
-projection, until we reach the maximum number of trials.
+To solve the optimisation problem embedded in the PP guided tour, a starting projection, an index function and the maximum number of iterations are provided as input.
+Then, the current projection is evaluated by the index function.
+The projection is then moved in a direction determined by a random factor, influenced by how far along we are in the optimization process.
+Occasionally, we might explore completely new directions like a jellyfish might with ocean currents.
+A new projection is accepted if it is an improvement compared to the current one, rejected otherwise.
+This process continues and iteratively improves the projection, until the pre-specified maximum number of trials is reached.
 

### 2024-06-20 optim.tex
sec 3 update text

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 
-The jellyfish optimiser (JSO) mimics the natural movements of jellyfish,
-which include passive and active motions driven by ocean currents and
-their swimming patterns, respectively. In the context of optimization,
-these movements are abstracted to explore the search space in a way that
-balances exploration (searching new areas) and exploitation (focusing on
+JSO mimics the natural movements of jellyfish, which include passive and
+active motions driven by ocean currents and their swimming patterns,
+respectively. In the context of optimization, these movements are
+abstracted to explore the search space in a way that balances
+exploration (searching new areas) and exploitation (focusing on
 promising areas). The algorithm aims to find the optimal solution by
-adapting the jellyfish's behaviour to navigate towards the best solution
-over iterations \citep{chou_novel_2021}.
+adapting the behaviour of jellyfish to navigate towards the best
+solution over iterations \citep{chou_novel_2021}.
 
-To understand what the jellyfish optimizer is doing in the context of
-Projection Pursuit, we first start with a current projection (the
-starting point). Then, we evaluate this projection using an index
-function, which tells us how good the current projection is. We then
-move the projection in a direction determined by the `best jelly' and
-random factors, influenced by how far along we are in the optimization
-process (the trial \(i\) and \texttt{max.tries}). Occasionally, we might
-explore completely new directions like a jellyfish might with ocean
-currents. Then, we compare new potential projections to our current one.
-If they're better, we adopt them; if not, we stick with our current
-projection. This process continues and iteratively improves the
-projection, until we reach the maximum number of trials.
+To solve the optimisation problem embedded in the PP guided tour, a
+starting projection, an index function and the maximum number of
+iterations are provided as input. Then, the current projection is
+evaluated by the index function. The projection is then moved in a
+direction determined by a random factor, influenced by how far along we
+are in the optimization process. Occasionally, we might explore
+completely new directions like a jellyfish might with ocean currents. A
+new projection is accepted if it is an improvement compared to the
+current one, rejected otherwise. This process continues and iteratively
+improves the projection, until the pre-specified maximum number of
+trials is reached.
 
-\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, breakable, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, left=2mm, titlerule=0mm, colframe=quarto-callout-note-color-frame, coltitle=black, opacitybacktitle=0.6, colback=white, toprule=.15mm, bottomtitle=1mm, arc=.35mm, bottomrule=.15mm, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, opacitybacktitle=0.6, colback=white, left=2mm, coltitle=black, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, leftrule=.75mm, bottomtitle=1mm, arc=.35mm, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, opacityback=0]
 

### 2024-06-20 optim.qmd
pseudo algo update

@@
 The projection is then moved in a direction determined by a random factor, influenced by how far along we are in the optimization process.
-Occasionally, we might explore completely new directions like a jellyfish might with ocean currents.
+Occasionally, completely new directions may be taken like a jellyfish might with ocean currents.
 A new projection is accepted if it is an improvement compared to the current one, rejected otherwise.
@@
 
-**Input**: `current_projections`, `index_function`, `tries`, `max_tries`
+**Input**: `current_projections`, `index_function`, `trial_id`, `max_trial`
 
@@
 
-**Initialize** `best_jelly` as the projection with the best index value
-from `current_projections`, and `current_index` as the array of index
+**Initialize** `current_best` as the projection with the best index value
+from `current_projections`, and `current_idx` as the array of index function
 values for each projection in `current_projections`
 
-**for** each try in 1 to max_tries **do**
+**for** each trial_id in 1 to max_tries **do**
 
-> Calculate $c_t$ based on the current try and max_tries
+> Calculate $c_t$ based on the current_idx and max_trial
 
@@
 >
-> > Define trend based on the best jelly and current projections
+> > Define trend based on the current_best and current_projections
 >
@@
 > >
-> > > Slightly adjust each projection with a small random factor (Type A
-> > > passive)
+> > > Slightly adjust each projection with a small random factor (passive)
 > >
@@
 > > > For each projection, compare with a random jelly and adjust
-> > > towards or away from it (Type B active)
+> > > towards or away from it (active)
 >
@@
 > **if** any new projection is worse than the current, revert to the
-> `current_projection` for that case
+> `current_projections` for that case
 >
 > > Determine the projection with the best index value as the new
-> > best_jelly
+> > current_best
 
-> **if** the try is the last one, print the final best projection and
+> **if** trial_id $\geq$ max_trial, print the last best projection 
 > **exit**
 
-**return** the set of projections with the updated best_jelly as the
+**return** the set of projections with the updated current_best as the
 optimized_projection
@@
 the algorithm or its application, we focus on two main parameters that
-are most relevant to our application: the number of jellyfish and drift.
+are most relevant to our application: the number of jellyfish and the maximum number of trials.
 

### 2024-06-20 optim.tex
pseudo algo update

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 direction determined by a random factor, influenced by how far along we
-are in the optimization process. Occasionally, we might explore
-completely new directions like a jellyfish might with ocean currents. A
-new projection is accepted if it is an improvement compared to the
-current one, rejected otherwise. This process continues and iteratively
-improves the projection, until the pre-specified maximum number of
-trials is reached.
+are in the optimization process. Occasionally, completely new directions
+may be taken like a jellyfish might with ocean currents. A new
+projection is accepted if it is an improvement compared to the current
+one, rejected otherwise. This process continues and iteratively improves
+the projection, until the pre-specified maximum number of trials is
+reached.
 
-\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, opacitybacktitle=0.6, colback=white, left=2mm, coltitle=black, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, leftrule=.75mm, bottomtitle=1mm, arc=.35mm, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, breakable, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, coltitle=black, toptitle=1mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, leftrule=.75mm, opacityback=0, colback=white, toprule=.15mm, bottomrule=.15mm, rightrule=.15mm, titlerule=0mm]
 
 \textbf{Input}: \texttt{current\_projections}, \texttt{index\_function},
-\texttt{tries}, \texttt{max\_tries}
+\texttt{trial\_id}, \texttt{max\_trial}
 
@@
 
-\textbf{Initialize} \texttt{best\_jelly} as the projection with the best
-index value from \texttt{current\_projections}, and
-\texttt{current\_index} as the array of index values for each projection
-in \texttt{current\_projections}
+\textbf{Initialize} \texttt{current\_best} as the projection with the
+best index value from \texttt{current\_projections}, and
+\texttt{current\_idx} as the array of index function values for each
+projection in \texttt{current\_projections}
 
-\textbf{for} each try in 1 to max\_tries \textbf{do}
+\textbf{for} each trial\_id in 1 to max\_tries \textbf{do}
 
 \begin{quote}
-Calculate \(c_t\) based on the current try and max\_tries
+Calculate \(c_t\) based on the current\_idx and max\_trial
 \end{quote}
@@
 \begin{quote}
-Define trend based on the best jelly and current projections
+Define trend based on the current\_best and current\_projections
 \end{quote}
@@
 \begin{quote}
-Slightly adjust each projection with a small random factor (Type A
-passive)
+Slightly adjust each projection with a small random factor (passive)
 \end{quote}
@@
 For each projection, compare with a random jelly and adjust towards or
-away from it (Type B active)
+away from it (active)
 \end{quote}
@@
 \textbf{if} any new projection is worse than the current, revert to the
-\texttt{current\_projection} for that case
+\texttt{current\_projections} for that case
 
@@
 Determine the projection with the best index value as the new
-best\_jelly
+current\_best
 \end{quote}
@@
 \begin{quote}
-\textbf{if} the try is the last one, print the final best projection and
-\textbf{exit}
+\textbf{if} trial\_id \(\geq\) max\_trial, print the last best
+projection \textbf{exit}
 \end{quote}
 
-\textbf{return} the set of projections with the updated best\_jelly as
+\textbf{return} the set of projections with the updated current\_best as
 the optimized\_projection
@@
 the algorithm or its application, we focus on two main parameters that
-are most relevant to our application: the number of jellyfish and drift.
+are most relevant to our application: the number of jellyfish and the
+maximum number of trials.
 

### 2024-06-21 optim.qmd
fix the plot

@@
 ```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
 #install the latest version of ferrn from: https://github.com/huizezhang-sherry/ferrn
@@
   ggplot(aes(x = dist, y = index, group = index_name)) +
-  geom_point() +
-  facet_nested_wrap(~index + n, nrow = 3, labeller = label_both) + 
+  geom_point(size = 0.5) +
+  facet_nested_wrap(~index_name + n, nrow = 3, labeller = label_both) + 
   xlab("projection distance") + ylab("index value") 
@@
   column_spec(1, border_left = TRUE) |> 
-  column_spec(c(4, 8, 12, 13), border_right = TRUE) |> 
+  column_spec(c(4, 8, 13), border_right = TRUE) |> 
   kable_styling(font_size = 7) 

### 2024-06-21 optim.tex
fix the plot

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, breakable, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, coltitle=black, toptitle=1mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, leftrule=.75mm, opacityback=0, colback=white, toprule=.15mm, bottomrule=.15mm, rightrule=.15mm, titlerule=0mm]
+\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, toprule=.15mm, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, rightrule=.15mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, leftrule=.75mm, left=2mm, breakable, opacitybacktitle=0.6, coltitle=black, colback=white]
 
@@
 
-\hypertarget{sec-smoothness}{%
-\subsection{Smoothness}\label{sec-smoothness}}
+\subsection{Smoothness}\label{sec-smoothness}
 
@@
 
-\hypertarget{sec-squintability}{%
-\subsection{Squintability}\label{sec-squintability}}
+\subsection{Squintability}\label{sec-squintability}
 
@@
 
-\hypertarget{sec-vis}{%
-\section{Visualisation of jellyfish optimiser}\label{sec-vis}}
+\section{Visualisation of jellyfish optimiser}\label{sec-vis}
 
@@
 
-{\centering \includegraphics{optim_files/figure-pdf/fig-projection-1.pdf}
+\centering{
+
+\includegraphics{optim_files/figure-pdf/fig-projection-1.pdf}
 
@@
 
-\end{figure}
+\end{figure}%
 
@@
 
-{\centering \includegraphics{optim_files/figure-pdf/fig-index-value-proj-dist-1.pdf}
+\centering{
+
+\includegraphics{optim_files/figure-pdf/fig-index-value-proj-dist-1.pdf}
 
@@
 
-\end{figure}
+\end{figure}%
 
-\hypertarget{sec-app}{%
-\section{Application {[}Di and Sherry{]}}\label{sec-app}}
+\section{Application {[}Di and Sherry{]}}\label{sec-app}
 
@@
 
-\hypertarget{sec-app-1}{%
 \subsection{Performance of the JS optimiser in pipe-finding
-problems}\label{sec-app-1}}
+problems}\label{sec-app-1}
 
@@
 
-{\centering \includegraphics{optim_files/figure-pdf/fig-proj-1.pdf}
+\centering{
+
+\includegraphics{optim_files/figure-pdf/fig-proj-1.pdf}
 
@@
 
-{\centering \includegraphics{optim_files/figure-pdf/fig-proportion-1.pdf}
+\centering{
+
+\includegraphics{optim_files/figure-pdf/fig-proportion-1.pdf}
 
@@
 
-\hypertarget{sec-app-2}{%
 \subsection{Factors affecting the JS optimiser success rate:
-optimisation properties and jellyfish
-hyper-parameters}\label{sec-app-2}}
+optimisation properties and jellyfish hyper-parameters}\label{sec-app-2}
 
@@
 
-{\centering \includegraphics{optim_files/figure-pdf/fig-idx-proj-dist-1.pdf}
+\centering{
+
+\includegraphics{optim_files/figure-pdf/fig-idx-proj-dist-1.pdf}
 
@@
 
-\hypertarget{tbl-smoothness-squintability}{}
 \begin{table}
-\caption{\label{tbl-smoothness-squintability}Parameters estimated from the Gaussian process (including variance,
-range, smoothness, and nugget) and scaled logistic function
-(\(\theta_1\) to \(\theta_4\)) for the pipe-finding and sine-wave
-finding problems. The column ``smooth'' and ``squint'' represent the
-smoothness and squintability measures. }\tabularnewline
+
+\caption{\label{tbl-smoothness-squintability}Parameters estimated from
+the Gaussian process (including variance, range, smoothness, and nugget)
+and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
+pipe-finding and sine-wave finding problems. The column ``smooth'' and
+``squint'' represent the smoothness and squintability measures.}
+
+\centering{
 
@@
 
-\begin{tabular}{|>{}lll>{}r|rrr>{}r|rrr>{}r|>{}r|}
+\begin{tabular}{|>{}lll>{}r|rrr>{}r|rrrr>{}r|}
 \toprule
@@
 \endgroup{}
-\end{table}
+
+}
+
+\end{table}%
 
@@
 
-\hypertarget{tbl-mod-data}{}
 \begin{longtable}[t]{lrrrrrrl}
-\caption{\label{tbl-mod-data}The first 7 rows of the datasets processed for modelling. }\tabularnewline
+
+\caption{\label{tbl-mod-data}The first 7 rows of the datasets processed
+for modelling.}
+
+\tabularnewline
 
@@
 \bottomrule
+
 \end{longtable}
+
 \endgroup{}
@@
 
-\hypertarget{tbl-mod-output}{}
 \begin{longtable}[t]{|>{}lrrr>{}r|}
-\caption{\label{tbl-mod-output}Model estimates of proportion of jellyfish success on optimisation
-properties and jellyfish hyper-parameters from the generalised linear
-model with a binomial family and a logit link function. The variable
-smoothness, squintability, number of jellyfish and maximum number of
-tries are positively associated with the JS optimiser success rate while
-data dimension and being flagged as long runtime are negatively
-associated with the success rate. The variable squintability and
-dimension are significant, suggesting their importance relative to
-jellyfish hyper-parameters in the optimisation success. }\tabularnewline
+
+\caption{\label{tbl-mod-output}Model estimates of proportion of
+jellyfish success on optimisation properties and jellyfish
+hyper-parameters from the generalised linear model with a binomial
+family and a logit link function. The variable smoothness,
+squintability, number of jellyfish and maximum number of tries are
+positively associated with the JS optimiser success rate while data
+dimension and being flagged as long runtime are negatively associated
+with the success rate. The variable squintability and dimension are
+significant, suggesting their importance relative to jellyfish
+hyper-parameters in the optimisation success.}
+
+\tabularnewline
 
@@
 \bottomrule
+
 \end{longtable}
+
 \endgroup{}
 
-\hypertarget{sec-conclusion}{%
-\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}}
+\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
 

### 2024-06-21 code/07-calc-smoothness-squintability.R

@@ -117,4 +117,13 @@ sq_basis_dist_idx <- bind_rows(sq_holes_basis_df, sq_sine_basis_df) |>
                                           "loess2d", "splines2d", "stringy"))) |>
   group_by(n, index_name, dist) |>
   summarise(index = mean(index, na.rm = TRUE), .groups = "drop")
+
+
+sq_basis_dist_idx <- sq_basis_dist_idx |>
+  filter(index_name %in% c("holes", "TIC", "stringy")) |>
+  group_by(n, index_name) |>
+  mutate(index = (index - min(index)) / (max(index) - min(index))) |>
+  ungroup() |>
+  bind_rows(sq_basis_dist_idx |>
+              filter(!index_name %in% c("holes", "TIC", "stringy")))
 save(sq_basis_dist_idx, file = here::here("data", "sq_basis_dist_idx.rda"))

### 2024-06-21 optim.qmd
update the squintability measure x4

@@
 
-    -   Use jellyfish search optimiser, or JS optimiser, not jellyfish
-        optimiser (my bad for leading the way)
+    -   Use jellyfish search optimiser, or JSO
     -   "hyper-parameter" rather than "hyperparameter"
+    - PP index rather than PPI
 
@@
 approach that involves computing informative linear projections of data
-by optimising a variety of objective functions, namely *projection
-pursuit index (PPI)* (@hall1989polynomial, @cook1993projection,
+by optimising a variety of objective functions, namely *PP index* 
+(@hall1989polynomial, @cook1993projection,
 @lee2010projection, @Loperfido2018, @Loperfido2020). Given
-high-dimensional data $X \in \mathbf{R}^{n }\times p$ and PPI
+high-dimensional data $X \in \mathbf{R}^{n }\times p$ and the index function
 $f(\cdot)$, PP finds the orthonormal projection matrix
@@
 
-The primary goal of this study is to investigate the performance of JSO in the context of the PP guided tour.
-A series of simulation experiments using well-recognised data sets and PPIs are conducted to yield insights regarding the behaviour of JSO under different settings and the sensitivity of hyper-parameter in the optimisation.
-Second, to observe the performance of JSO with different types of PPI, a collection of metrics is introduced to capture specific properties of the index including squintability and smoothness as introduced in @laa_using_2020.
+The primary goal of this study is to investigate the performance of JSO in the context of the Projection pursuit.
+A series of simulation experiments using well-recognised data sets and PP indexes are conducted to yield insights regarding the behaviour of JSO under different settings and the sensitivity of hyper-parameter in the optimisation.
+Second, to observe the performance of JSO with different types of PP indexes, a collection of metrics is introduced to capture specific properties of the index including squintability and smoothness as introduced in @laa_using_2020.
 To the best of the authors’ knowledge, this is the first attempt to quantitatively measure squintability and smoothness.
-Finally, the relationship between the JSO performance, hyper-parameters tuning and various properties of PPI is analysed to provide helpful guidance for practitioners that are using the guided tour for high dimensional visualisation.
+Finally, the relationship between the JSO performance, hyper-parameters tuning and various properties of PP indexes is analysed to provide helpful guidance for practitioners that are using the guided tour for high dimensional visualisation.
 
@@
 introduces the background of the PP guided tour and reviews existing methods in the literature. @sec-theory describes the details of JSO
-and introduces metrics that measure different properties of a PPI.
+and introduces metrics that measure different properties of PP indexes.
 @sec-vis visualises the behaviour of the JSO in terms of describes a simulation study on the performance of the JSO using well-known data sets and index functions.
-@sec-app displays two sets of simulation experiments comparing the JSO with the search-better optimiser and studies the impact of different PPI properties on the optimisation performance.
+@sec-app displays two sets of simulation experiments comparing the JSO with the search-better optimiser and studies the impact of different PP index properties on the optimisation performance.
 @sec-conclusion summarises the work and provides suggestions for future directions.
@@
 comparing with the current (last) basis. If a visited basis has a
-smaller index value than the current (last) one, the JS optimiser keeps
+smaller index value than the current (last) one, JSO will keep
 the current basis for the next iteration in the search but still records
@@
 
-The JS optimiser provides a new swarm-based search strategy to optimise
+The JSO provides a new swarm-based search strategy to optimise
 projection pursuit problems. Two examples are provided in this section
@@
 
-## Performance of the JS optimiser in pipe-finding problems {#sec-app-1}
+## Performance of JSO in pipe-finding problems {#sec-app-1}
 
-The performance of the JS optimiser is investigated in two folds: 1)
+The performance of JSO is investigated in two folds: 1)
 compare it with a commonly used optimiser: the search-better optimiser
@@
 The projection pursuit problem is to find the pipe shape using the holes index in the 6, 8, 10, and 12-dimensional spaces.
-The JS optimiser uses 100 jellyfishes and a maximum number of tries of 100.
+The JSO uses 100 jellyfishes and a maximum number of tries of 100.
 The search better optimiser uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates.
-In the 6-D data space, the JS optimiser always finds a clear pipe shape while the search better optimiser also finds the pipe shape but with a wide rim.
-At higher data dimensions, the JS optimiser finds a higher index value and a clearer pipe shape across all the quantiles than the search better optimiser."
+In the 6-D data space, JSO always finds a clear pipe shape while the search better optimiser also finds the pipe shape but with a wide rim.
+At higher data dimensions, JSO finds a higher index value and a clearer pipe shape across all the quantiles than the search better optimiser."
 
@@
 
-(p1 + ggtitle("The JS optimiser"))/(p2 + ggtitle("The search better optimiser"))
+(p1 + ggtitle("The jellyfish search optimiser"))/(p2 + ggtitle("The search better optimiser"))
 ```
@@
 
-Fifty simulations are conducted with both JS and the search-better
-optimiser, in four dimensions ($d = 6, 8, 10, 12$). The JS optimiser
+Fifty simulations are conducted with both JSO and the search-better
+, in four dimensions ($d = 6, 8, 10, 12$). The JSO
 uses 100 jellyfishes and 100 maximum number of tries, while the
@@
 projections found by the two optimisers, broken down by 10th quantile,
-faceted by the data dimension. In the 6-dimensional data scenario, the
-JS optimiser consistently identifies a clear pipe shape. The
+faceted by the data dimension. In the 6-dimensional data scenario, JSO
+ consistently identifies a clear pipe shape. The
 search-better optimiser also finds the pipe shape but with a wide rim,
 suggesting a further polish search may be required. With increasing
-dimensions, the JS optimiser may not always identify the pipe shape due
+dimensions, JSO may not always identify the pipe shape due
 to random sampling, but it still presents a pipe shape in over 50% of
-cases. When compared to the search-better optimiser, the JS optimiser
+cases. When compared to the search-better optimiser, JSO
 reaches higher index values and clearer pipe shapes across all
@@
 hyper-parameter combination to analyse its effects on the success rate
-of the JS optimiser. The hyper-parameters tested include: 1) 20, 50, and
+of JSO . The hyper-parameters tested include: 1) 20, 50, and
 100 jellyfishes, and 2) 50 and 100 maximum number of tries. In each
@@
 
-## Factors affecting the JS optimiser success rate: optimisation properties and jellyfish hyper-parameters {#sec-app-2}
+## Factors affecting JSO success rate: optimisation properties and jellyfish hyper-parameters {#sec-app-2}
 
@@
 number of jellyfishes and the maximum number of tries), affect the
-success of the JS optimiser. Simulations are conducted to obtain the
+success of JSO. Simulations are conducted to obtain the
 success rate across various projection pursuit problems, characterised
@@
 pipe-finding cases and `r nrow(filter(sim_df, index != "holes"))`
-sine-wave finding cases. For each case, the JS optimiser is run fifty
+sine-wave finding cases. For each case, JSO is run fifty
 times and the success rate is calculated as in @sec-app-1.
@@
 and @tbl-mod-output presents the model outputs. The model suggests that
-the success rate of the JS optimiser is positively associated with the
+the success rate of JSO is positively associated with the
 two jellyfish hyper-parameters, as well as with smoothness and
@@
 #| tbl-cap: Model estimates of proportion of jellyfish success on optimisation properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function.
-The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with the JS optimiser success rate while data dimension and being flagged as long runtime are negatively associated with the success rate.
+The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with JSO success rate while data dimension and being flagged as long runtime are negatively associated with the success rate.
 The variable squintability and dimension are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.

### 2024-06-21 optim.tex
update the squintability measure x4

@@
   \item
-    Use jellyfish search optimiser, or JS optimiser, not jellyfish
-    optimiser (my bad for leading the way)
+    Use jellyfish search optimiser, or JSO
   \item
     ``hyper-parameter'' rather than ``hyperparameter''
+  \item
+    PP index rather than PPI
   \end{itemize}
@@
 approach that involves computing informative linear projections of data
-by optimising a variety of objective functions, namely \emph{projection
-pursuit index (PPI)} (\citet{hall1989polynomial},
-\citet{cook1993projection}, \citet{lee2010projection},
-\citet{Loperfido2018}, \citet{Loperfido2020}). Given high-dimensional
-data \(X \in \mathbf{R}^{n }\times p\) and PPI \(f(\cdot)\), PP finds
-the orthonormal projection matrix \(A \in \mathbf{R}^{p \times d}\) by
-solving the following optimisation problem:
+by optimising a variety of objective functions, namely \emph{PP index}
+(\citet{hall1989polynomial}, \citet{cook1993projection},
+\citet{lee2010projection}, \citet{Loperfido2018},
+\citet{Loperfido2020}). Given high-dimensional data
+\(X \in \mathbf{R}^{n }\times p\) and the index function \(f(\cdot)\),
+PP finds the orthonormal projection matrix
+\(A \in \mathbf{R}^{p \times d}\) by solving the following optimisation
+problem:
 
@@
 The primary goal of this study is to investigate the performance of JSO
-in the context of the PP guided tour. A series of simulation experiments
-using well-recognised data sets and PPIs are conducted to yield insights
-regarding the behaviour of JSO under different settings and the
-sensitivity of hyper-parameter in the optimisation. Second, to observe
-the performance of JSO with different types of PPI, a collection of
-metrics is introduced to capture specific properties of the index
-including squintability and smoothness as introduced in
-\citet{laa_using_2020}. To the best of the authors' knowledge, this is
-the first attempt to quantitatively measure squintability and
-smoothness. Finally, the relationship between the JSO performance,
-hyper-parameters tuning and various properties of PPI is analysed to
-provide helpful guidance for practitioners that are using the guided
-tour for high dimensional visualisation.
+in the context of the Projection pursuit. A series of simulation
+experiments using well-recognised data sets and PP indexes are conducted
+to yield insights regarding the behaviour of JSO under different
+settings and the sensitivity of hyper-parameter in the optimisation.
+Second, to observe the performance of JSO with different types of PP
+indexes, a collection of metrics is introduced to capture specific
+properties of the index including squintability and smoothness as
+introduced in \citet{laa_using_2020}. To the best of the authors'
+knowledge, this is the first attempt to quantitatively measure
+squintability and smoothness. Finally, the relationship between the JSO
+performance, hyper-parameters tuning and various properties of PP
+indexes is analysed to provide helpful guidance for practitioners that
+are using the guided tour for high dimensional visualisation.
 
@@
 Section~\ref{sec-theory} describes the details of JSO and introduces
-metrics that measure different properties of a PPI.
+metrics that measure different properties of PP indexes.
 Section~\ref{sec-vis} visualises the behaviour of the JSO in terms of
@@
 two sets of simulation experiments comparing the JSO with the
-search-better optimiser and studies the impact of different PPI
+search-better optimiser and studies the impact of different PP index
 properties on the optimisation performance. Section~\ref{sec-conclusion}
@@
 
-\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, toprule=.15mm, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, rightrule=.15mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, leftrule=.75mm, left=2mm, breakable, opacitybacktitle=0.6, coltitle=black, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, coltitle=black, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, left=2mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, arc=.35mm, opacityback=0, breakable, colbacktitle=quarto-callout-note-color!10!white, bottomrule=.15mm, bottomtitle=1mm, titlerule=0mm, toprule=.15mm, leftrule=.75mm, colback=white]
 
@@
 comparing with the current (last) basis. If a visited basis has a
-smaller index value than the current (last) one, the JS optimiser keeps
-the current basis for the next iteration in the search but still records
-the visited inferior basis. This data can be used to derive numerical
+smaller index value than the current (last) one, JSO will keep the
+current basis for the next iteration in the search but still records the
+visited inferior basis. This data can be used to derive numerical
 summaries and plots:
@@
 
-\subsection{Performance of the JS optimiser in pipe-finding
+\subsection{Performance of JSO in pipe-finding
 problems}\label{sec-app-1}
 
-The performance of the JS optimiser is investigated in two folds: 1)
-compare it with a commonly used optimiser: the search-better optimiser
+The performance of JSO is investigated in two folds: 1) compare it with
+a commonly used optimiser: the search-better optimiser
 \citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
@@
 projection pursuit problem is to find the pipe shape using the holes
-index in the 6, 8, 10, and 12-dimensional spaces. The JS optimiser uses
-100 jellyfishes and a maximum number of tries of 100. The search better
+index in the 6, 8, 10, and 12-dimensional spaces. The JSO uses 100
+jellyfishes and a maximum number of tries of 100. The search better
 optimiser uses a maximum of 1000 tries in each step of random sampling
-step before the algorithm terminates. In the 6-D data space, the JS
-optimiser always finds a clear pipe shape while the search better
-optimiser also finds the pipe shape but with a wide rim. At higher data
-dimensions, the JS optimiser finds a higher index value and a clearer
-pipe shape across all the quantiles than the search better optimiser.}
+step before the algorithm terminates. In the 6-D data space, JSO always
+finds a clear pipe shape while the search better optimiser also finds
+the pipe shape but with a wide rim. At higher data dimensions, JSO finds
+a higher index value and a clearer pipe shape across all the quantiles
+than the search better optimiser.}
 
@@
 
-Fifty simulations are conducted with both JS and the search-better
-optimiser, in four dimensions (\(d = 6, 8, 10, 12\)). The JS optimiser
-uses 100 jellyfishes and 100 maximum number of tries, while the
-search-better optimiser allows a maximum of 1000 samples at each
-iteration before the algorithm terminates. Figure~\ref{fig-proj}
-presents the final projections found by the two optimisers, broken down
-by 10th quantile, faceted by the data dimension. In the 6-dimensional
-data scenario, the JS optimiser consistently identifies a clear pipe
-shape. The search-better optimiser also finds the pipe shape but with a
-wide rim, suggesting a further polish search may be required. With
-increasing dimensions, the JS optimiser may not always identify the pipe
-shape due to random sampling, but it still presents a pipe shape in over
-50\% of cases. When compared to the search-better optimiser, the JS
-optimiser reaches higher index values and clearer pipe shapes across all
-quantiles.
+Fifty simulations are conducted with both JSO and the search-better , in
+four dimensions (\(d = 6, 8, 10, 12\)). The JSO uses 100 jellyfishes and
+100 maximum number of tries, while the search-better optimiser allows a
+maximum of 1000 samples at each iteration before the algorithm
+terminates. Figure~\ref{fig-proj} presents the final projections found
+by the two optimisers, broken down by 10th quantile, faceted by the data
+dimension. In the 6-dimensional data scenario, JSO consistently
+identifies a clear pipe shape. The search-better optimiser also finds
+the pipe shape but with a wide rim, suggesting a further polish search
+may be required. With increasing dimensions, JSO may not always identify
+the pipe shape due to random sampling, but it still presents a pipe
+shape in over 50\% of cases. When compared to the search-better
+optimiser, JSO reaches higher index values and clearer pipe shapes
+across all quantiles.
 
@@
 hyper-parameter combination to analyse its effects on the success rate
-of the JS optimiser. The hyper-parameters tested include: 1) 20, 50, and
-100 jellyfishes, and 2) 50 and 100 maximum number of tries. In each
+of JSO . The hyper-parameters tested include: 1) 20, 50, and 100
+jellyfishes, and 2) 50 and 100 maximum number of tries. In each
 simulation, the success rate is calculated as defined in
@@
 
-\subsection{Factors affecting the JS optimiser success rate:
-optimisation properties and jellyfish hyper-parameters}\label{sec-app-2}
+\subsection{Factors affecting JSO success rate: optimisation properties
+and jellyfish hyper-parameters}\label{sec-app-2}
 
@@
 number of jellyfishes and the maximum number of tries), affect the
-success of the JS optimiser. Simulations are conducted to obtain the
-success rate across various projection pursuit problems, characterised
-by shape-to-find, data dimension, and index function used, as well as
+success of JSO. Simulations are conducted to obtain the success rate
+across various projection pursuit problems, characterised by
+shape-to-find, data dimension, and index function used, as well as
 different hyper-parameter combinations. Smoothness and squintability are
@@
 hyper-parameters, a total of 52 cases is produced, comprising of 24
-pipe-finding cases and 28 sine-wave finding cases. For each case, the JS
-optimiser is run fifty times and the success rate is calculated as in
+pipe-finding cases and 28 sine-wave finding cases. For each case, JSO is
+run fifty times and the success rate is calculated as in
 Section~\ref{sec-app-1}.
@@
 function is used to fit the data and Table~\ref{tbl-mod-output} presents
-the model outputs. The model suggests that the success rate of the JS
-optimiser is positively associated with the two jellyfish
-hyper-parameters, as well as with smoothness and squintability. However,
-being flagged with long runtime and an increase of data dimension will
-reduce the success rate. The variable \texttt{squintability} and
-\texttt{dimension} are significant, suggesting their importance relative
-to jellyfish hyper-parameters in the optimisation success.
+the model outputs. The model suggests that the success rate of JSO is
+positively associated with the two jellyfish hyper-parameters, as well
+as with smoothness and squintability. However, being flagged with long
+runtime and an increase of data dimension will reduce the success rate.
+The variable \texttt{squintability} and \texttt{dimension} are
+significant, suggesting their importance relative to jellyfish
+hyper-parameters in the optimisation success.
 
@@
 \midrule
-1 & pipe & holes & 6 & 0.002 & 0.371 & 2.355 & 0.221 & 1.015 & 0.860 & 3.368 & 0.018 & 0.763\\
-2 & pipe & holes & 8 & 0.000 & 0.178 & 2.189 & 0.820 & 1.014 & 0.869 & 3.264 & 0.029 & 0.740\\
-3 & pipe & holes & 10 & 0.000 & 0.107 & 2.192 & 3.027 & 1.016 & 0.885 & 3.151 & 0.022 & 0.737\\
-4 & pipe & holes & 12 & 0.000 & 0.148 & 2.295 & 1.581 & 1.011 & 0.878 & 3.345 & 0.004 & 0.779\\
-5 & sine & MIC & 6 & 0.016 & 0.092 & 2.457 & 0.083 & 0.894 & 0.571 & 1.623 & -0.024 & 0.314\\
+1 & pipe & holes & 6 & 0.002 & 0.371 & 2.355 & 0.221 & 1.015 & 0.860 & 3.368 & 0.018 & 3.053\\
+2 & pipe & holes & 8 & 0.000 & 0.178 & 2.189 & 0.820 & 1.014 & 0.869 & 3.264 & 0.029 & 2.960\\
+3 & pipe & holes & 10 & 0.000 & 0.107 & 2.192 & 3.027 & 1.016 & 0.885 & 3.151 & 0.022 & 2.948\\
+4 & pipe & holes & 12 & 0.000 & 0.148 & 2.295 & 1.581 & 1.011 & 0.878 & 3.345 & 0.004 & 3.117\\
+5 & sine & MIC & 6 & 0.016 & 0.092 & 2.457 & 0.083 & 0.894 & 0.571 & 1.623 & -0.024 & 1.257\\
 \addlinespace
-6 & sine & MIC & 8 & 0.016 & 0.081 & 2.636 & 0.084 & 0.932 & 0.328 & 1.314 & -0.030 & 0.193\\
-7 & sine & TIC & 6 & 0.124 & 0.110 & 2.444 & 0.086 & 0.951 & 0.536 & 1.719 & -0.027 & 0.330\\
-8 & sine & TIC & 8 & 0.125 & 0.099 & 2.475 & 0.085 & 0.949 & 0.564 & 1.723 & -0.028 & 0.342\\
-9 & sine & dcor2d & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019 & 0.737\\
-10 & sine & loess2d & 6 & 0.079 & 0.336 & 1.988 & 0.309 & 1.016 & 1.039 & 2.648 & 0.080 & 0.689\\
+6 & sine & MIC & 8 & 0.016 & 0.081 & 2.636 & 0.084 & 0.932 & 0.328 & 1.314 & -0.030 & 0.772\\
+7 & sine & TIC & 6 & 0.124 & 0.110 & 2.444 & 0.086 & 0.951 & 0.536 & 1.719 & -0.027 & 1.322\\
+8 & sine & TIC & 8 & 0.125 & 0.099 & 2.475 & 0.085 & 0.949 & 0.564 & 1.723 & -0.028 & 1.369\\
+9 & sine & dcor2d & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019 & 2.947\\
+10 & sine & loess2d & 6 & 0.079 & 0.336 & 1.988 & 0.309 & 1.016 & 1.039 & 2.648 & 0.080 & 2.756\\
 \addlinespace
-11 & sine & splines2d & 6 & 0.042 & 0.242 & 2.537 & 0.102 & 1.014 & 1.051 & 2.730 & -0.009 & 0.780\\
-12 & sine & stringy & 6 & 0.000 & 0.007 & 1.536 & 38.166 & 1.045 & 0.011 & 254.734 & 0.053 & 0.739\\
+11 & sine & splines2d & 6 & 0.042 & 0.242 & 2.537 & 0.102 & 1.014 & 1.051 & 2.730 & -0.009 & 3.118\\
+12 & sine & stringy & 6 & 0.000 & 0.007 & 1.536 & 38.166 & 1.045 & 0.011 & 254.734 & 0.053 & 2.957\\
 \bottomrule
@@
 \midrule
-MIC & 6 & 2.457 & 0.314 & 20 & 50 & 0.12 & 2.479 secs\\
-MIC & 6 & 2.457 & 0.314 & 20 & 100 & 0.24 & 8.950 secs\\
-MIC & 6 & 2.457 & 0.314 & 50 & 50 & 0.52 & 5.651 secs\\
-MIC & 6 & 2.457 & 0.314 & 50 & 100 & 0.64 & 13.223 secs\\
-MIC & 6 & 2.457 & 0.314 & 100 & 50 & 0.76 & 19.453 secs\\
+MIC & 6 & 2.457 & 1.257 & 20 & 50 & 0.12 & 2.479 secs\\
+MIC & 6 & 2.457 & 1.257 & 20 & 100 & 0.24 & 8.950 secs\\
+MIC & 6 & 2.457 & 1.257 & 50 & 50 & 0.52 & 5.651 secs\\
+MIC & 6 & 2.457 & 1.257 & 50 & 100 & 0.64 & 13.223 secs\\
+MIC & 6 & 2.457 & 1.257 & 100 & 50 & 0.76 & 19.453 secs\\
 \addlinespace
-MIC & 8 & 2.636 & 0.193 & 20 & 50 & 0.08 & 2.566 secs\\
-MIC & 8 & 2.636 & 0.193 & 20 & 100 & 0.08 & 4.960 secs\\
+MIC & 8 & 2.636 & 0.772 & 20 & 50 & 0.08 & 2.566 secs\\
+MIC & 8 & 2.636 & 0.772 & 20 & 100 & 0.08 & 4.960 secs\\
 \bottomrule
@@
 squintability, number of jellyfish and maximum number of tries are
-positively associated with the JS optimiser success rate while data
-dimension and being flagged as long runtime are negatively associated
-with the success rate. The variable squintability and dimension are
-significant, suggesting their importance relative to jellyfish
-hyper-parameters in the optimisation success.}
+positively associated with JSO success rate while data dimension and
+being flagged as long runtime are negatively associated with the success
+rate. The variable squintability and dimension are significant,
+suggesting their importance relative to jellyfish hyper-parameters in
+the optimisation success.}
 
@@
 Smoothness & 1.195 & 1.915 & 0.624 & 0.533\\
-Squintability & 8.241 & 2.737 & 3.011 & 0.003\\
+Squintability & 2.060 & 0.684 & 3.011 & 0.003\\
 dimension (d) & -0.628 & 0.255 & -2.462 & 0.014\\

### 2024-06-22 optim.qmd
squintability subsection
added the content of the squintability subsection, with equations

@@
         - id: 3
-          name: BNU-HKBU United International College
+          name: Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College
           department: Department of Mathematical Sciences
@@
     wrap: 72
+crossref: 
+  eq-prefix: ""
 ---
@@
 by optimising a variety of objective functions, namely *PP index*
-(@hall1989polynomial, @cook1993projection,
-@lee2010projection, @Loperfido2018, @Loperfido2020). Given
-high-dimensional data $X \in \mathbf{R}^{n }\times p$ and the index function
-$f(\cdot)$, PP finds the orthonormal projection matrix
-$A \in \mathbf{R}^{p \times d}$ by solving the following optimisation
-problem:
+(@hall1989polynomial, @cook1993projection, @lee2010projection,
+@Loperfido2018, @Loperfido2020). Given high-dimensional data
+$X \in \mathbf{R}^{n\times p}$ and the index function $f(\cdot)$, PP
+finds the orthonormal projection matrix $A \in \mathbf{R}^{p \times d}$
+by solving the following optimisation problem:
 
@@
 \underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
-$$
+$$ {#eq-optimization}
 
 These indices $f(\cdot)$ are characterised by the informativeness or
-“interestingness” of a projection and are often non-linear and
+"interestingness" of a projection and are often non-linear and
 non-convex. As such, an effective and efficient optimisation procedure
@@
 in the tour is essential, it is also important to be able to watch the
-projected data as the optimisation progresses. As such, it is pertinent to integrate the guided tour with a global optimisation
-algorithm that is efficient in finding the global optimal and enables viewing of the projected data during the exploration process.
+projected data as the optimisation progresses. As such, it is pertinent
+to integrate the guided tour with a global optimisation algorithm that
+is efficient in finding the global optimal and enables viewing of the
+projected data during the exploration process.
 
-The artificial Jellyfish Search Optimiser (JSO) [@chou_novel_2021] is a swarm-based metaheuristic designed to solve global optimisation problems.
-Inspired by the search behaviour of jellyfish in the ocean, JSO is one of the latest swarm intelligence algorithms [@rajwar_exhaustive_2023], which was shown to have stronger search ability and faster convergence with fewer parameters compared to classic optimisation methods [@chou_novel_2021]-[@chou_recent_2022].
-These practical properties makes JSO a strong candidate in integration with the PP guided tour.
-As such, it is of interest to explore the potential of JSO in enhancing the PP guided tour and examine the characteristics and advantages of such implementation.
+The artificial Jellyfish Search Optimiser (JSO) [@chou_novel_2021] is a
+swarm-based metaheuristic designed to solve global optimisation
+problems. Inspired by the search behaviour of jellyfish in the ocean,
+JSO is one of the latest swarm intelligence algorithms
+[@rajwar_exhaustive_2023], which was shown to have stronger search
+ability and faster convergence with fewer parameters compared to classic
+optimisation methods [@chou_novel_2021]-[@chou_recent_2022]. These
+practical properties makes JSO a strong candidate in integration with
+the PP guided tour. As such, it is of interest to explore the potential
+of JSO in enhancing the PP guided tour and examine the characteristics
+and advantages of such implementation.
 
-The primary goal of this study is to investigate the performance of JSO in the context of the Projection pursuit.
-A series of simulation experiments using well-recognised data sets and PP indexes are conducted to yield insights regarding the behaviour of JSO under different settings and the sensitivity of hyper-parameter in the optimisation.
-Second, to observe the performance of JSO with different types of PP indexes, a collection of metrics is introduced to capture specific properties of the index including squintability and smoothness as introduced in @laa_using_2020.
-To the best of the authors’ knowledge, this is the first attempt to quantitatively measure squintability and smoothness.
-Finally, the relationship between the JSO performance, hyper-parameters tuning and various properties of PP indexes is analysed to provide helpful guidance for practitioners that are using the guided tour for high dimensional visualisation.
+The primary goal of this study is to investigate the performance of JSO
+in the context of the Projection pursuit. A series of simulation
+experiments using well-recognised data sets and PP indexes are conducted
+to yield insights regarding the behaviour of JSO under different
+settings and the sensitivity of hyper-parameter in the optimisation.
+Second, to observe the performance of JSO with different types of PP
+indexes, a collection of metrics is introduced to capture specific
+properties of the index including squintability and smoothness as
+introduced in @laa_using_2020. To the best of the authors' knowledge,
+this is the first attempt to quantitatively measure squintability and
+smoothness in this context. Finally, the relationship between the JSO
+performance, hyper-parameters tuning and various properties of PP
+indexes is analysed to provide helpful guidance for practitioners that
+are using the guided tour for high dimensional visualisation.
 
 The rest of this paper is structured as follows. @sec-background
-introduces the background of the PP guided tour and reviews existing methods in the literature. @sec-theory describes the details of JSO
-and introduces metrics that measure different properties of PP indexes.
-@sec-vis visualises the behaviour of the JSO in terms of describes a simulation study on the performance of the JSO using well-known data sets and index functions.
-@sec-app displays two sets of simulation experiments comparing the JSO with the search-better optimiser and studies the impact of different PP index properties on the optimisation performance.
-@sec-conclusion summarises the work and provides suggestions for future directions.
+introduces the background of the PP guided tour and reviews existing
+methods in the literature. @sec-theory describes the details of JSO and
+introduces metrics that measure different properties of PP indexes.
+@sec-vis visualises the behaviour of the JSO in terms of describes a
+simulation study on the performance of the JSO using well-known data
+sets and index functions. @sec-app displays two sets of simulation
+experiments comparing the JSO with the search-better optimiser and
+studies the impact of different PP index properties on the optimisation
+performance. @sec-conclusion summarises the work and provides
+suggestions for future directions.
 
@@
 scagnostics is described in @laa_using_2020. To be useful here an
-optimiser needs to be able to handle functions which are not very
-smooth. In addition, because data structures might be relatively fine,
-the optimiser needs to be able to find maxima that occur with a small
-squint angle, that can only be seen from very close by. One last aspect
-that is useful is for an optimiser to return local maxima in addition to
-global because data can contain many different and interesting features.
+optimiser needs to be able to handle index functions which are possibly
+not very smooth. In addition, because data structures might be
+relatively fine, the optimiser needs to be able to find maxima that
+occur with a small squint angle, that can only be seen from very close
+by. One last aspect that is useful is for an optimiser to return local
+maxima in addition to global because data can contain many different and
+interesting features.
 
@@
 
-JSO mimics the natural movements of jellyfish, which include passive and active motions driven by ocean currents and
-their swimming patterns, respectively. In the context of optimization, these movements are abstracted to explore the search space in a way that
-balances exploration (searching new areas) and exploitation (focusing on
+JSO mimics the natural movements of jellyfish, which include passive and
+active motions driven by ocean currents and their swimming patterns,
+respectively. In the context of optimization, these movements are
+abstracted to explore the search space in a way that balances
+exploration (searching new areas) and exploitation (focusing on
 promising areas). The algorithm aims to find the optimal solution by
-adapting the behaviour of jellyfish to navigate towards the best solution over iterations [@chou_novel_2021].
+adapting the behaviour of jellyfish to navigate towards the best
+solution over iterations [@chou_novel_2021].
 
-To solve the optimisation problem embedded in the PP guided tour, a starting projection, an index function and the maximum number of iterations are provided as input.
-Then, the current projection is evaluated by the index function.
-The projection is then moved in a direction determined by a random factor, influenced by how far along we are in the optimization process.
-Occasionally, completely new directions may be taken like a jellyfish might with ocean currents.
-A new projection is accepted if it is an improvement compared to the current one, rejected otherwise.
-This process continues and iteratively improves the projection, until the pre-specified maximum number of trials is reached.
+To solve the optimisation problem embedded in the PP guided tour, a
+starting projection, an index function and the maximum number of
+iterations are provided as input. Then, the current projection is
+evaluated by the index function. The projection is then moved in a
+direction determined by a random factor, influenced by how far along we
+are in the optimization process. Occasionally, completely new directions
+may be taken like a jellyfish might with ocean currents. A new
+projection is accepted if it is an improvement compared to the current
+one, rejected otherwise. This process continues and iteratively improves
+the projection, until the pre-specified maximum number of trials is
+reached.
 
@@
 
-**Input**: `current_projections`, `index_function`, `trial_id`, `max_trial`
+**Input**: `current_projections`, `index_function`, `trial_id`,
+`max_trial`
 
@@
 
-**Initialize** `current_best` as the projection with the best index value
-from `current_projections`, and `current_idx` as the array of index function
-values for each projection in `current_projections`
+**Initialize** `current_best` as the projection with the best index
+value from `current_projections`, and `current_idx` as the array of
+index function values for each projection in `current_projections`
 
@@
 > >
-> > > Slightly adjust each projection with a small random factor (passive)
+> > > Slightly adjust each projection with a small random factor
+> > > (passive)
 > >
@@
 the algorithm or its application, we focus on two main parameters that
-are most relevant to our application: the number of jellyfish and the maximum number of trials.
+are most relevant to our application: the number of jellyfish and the
+maximum number of trials.
 
@@
 
-In this study, we propose two metrics to capture the notion of
-squintability.
+In this subsection, a metric to capture the notion of squintability is
+proposed, and two approaches to compute this metric numerically are
+detailed.
 
+```{=tex}
+\textcolor{orange}{Here we need to add a precise, mathematical
+definition of the $x$ and $y$ variables ("angle distance" to optimal?
+and PP index values) used in the figures of this Section.}
+```
 \[We generate random points that is beyond 1.5 projection distance and
@@
 
-First, parametric model.
+Let $f$ represent the average PP index value with respect to the
+projection angle distance to optimal, over the course of the JSO. It is
+expected that for a PP index with high squint angle, the optimization
+(@eq-optimization) should make substantial progress early on.
+Conversely, for a PP index with low squint angle, it might take a long
+while for the optimization to make substantial progress, as the
+candidate projections would need to be very close to the optimal one for
+the structure of the index function to be visible enough to be amenable
+to efficient optimization. This observation suggests that the extreme
+values of $f'$ (the ones for which $f''=0$, assuming that $f$ is twice
+differentiable), and the angles for which these values are attained,
+should play an important role in the mathematical definition of
+squintability. Noticing that $f$ is expected to be a decreasing
+function, we propose the following squintability metric:
 
-\[Nicolas's pdf\]
+$$
+\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)
+$$ {#eq-squintability}
 
-Second, we consider the product of the largest absolute magnitude of
-rate of change of $f$ and the corresponding projection angle as a second
-measure of squintability. Since $f$ is decreasing, the rate of change of
-$f$ is negative and thus $|\underset{x}{\min} f('x)|$ gives the absolute
-magnitude of the most negative rate of change.
+On the one hand, $-\min_{x}f'(x)$ represents the largest value of the
+gradient of $-f$. On the other hand, $\underset{x}{\arg\min}f'(x)$
+represents the angle value at which this largest gradient is attained.
+Based on the above discussion, it is expected that these two values
+should be both high in the case of high squintability (fast increase in
+$f$ early on), and both low in the case of low squintability (any
+substantial increase in $f$ happens very late, close to the optimal
+angle). This suggest that their product (@eq-squintability) should
+provide a sensible measure of squintability. The multiplicative constant
+$4$, which can be deemed arbitrary, does not change the interpretation
+of the squintability metric $\varsigma$; it is here to adjust the range
+of values of $\varsigma$ and simplify the explicit formula for
+$\varsigma$ obtained later on.
 
-\[Nicolas's pdf\]
+To compute the squintability metric (@eq-squintability) in practice,
+several approaches are possible. The first one is to propose a
+parametric model for $f$, and use it to obtain an explicit formula for
+$\varsigma$. Numerical experiments suggest a scaled sigmoid shape as
+described below. Define
 
+$$
+\ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
+$$ {#eq-logistic}
+
+which is a decreasing logistic function depending on two parameters
+$\theta_2$ and $\theta_3$ , such that $\ell(\theta_{2})=\frac{1}{2}$.
+Then, define
+
+$$
+f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
+$$ {#eq-parametric}
+
+which depends on three additional parameters, $\theta_1$, $\theta_2$ ,
+and $x_{\max}$, such that $f(0)=\theta_1$ and $f(x_{\max})=\theta_4$.
+Under the parametric model (@eq-parametric), the squintability metric
+(@eq-squintability) can be shown to be equal to
+
+$$
+\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}
+$$ {#eq-squintability-parametric}
+
+In practice, the parameters of this model (@eq-parametric) can be
+estimated numerically, for example by non-linear least squares, and then
+used to evaluate $\varsigma$ as in equation
+(@eq-squintability-parametric).
+
+Alternatively, one can estimate (@eq-squintability) in a nonparametric
+way, for example by fitting $f$ using kernel regression, then
+numerically estimate the angle at which $-f'$ attains its highest value.
+
+```{=tex}
+\textcolor{orange}{See whether further details are needed for the nonparametric approach.}
+```
 TODO: add equation reference to main text line 705
@@
 
-The speed of optimizing an index function can be measured empirically by recording the duration of the optimisation.
-Alternatively, one can also gauge the speed in terms of computational complexity of evaluating the index function (for instance, in big O notation, with respect to
-the sample size) of computing the index function.
+The speed of optimizing an index function can be measured empirically by
+recording the duration of the optimisation. Alternatively, one can also
+gauge the speed in terms of computational complexity of evaluating the
+index function (for instance, in big O notation, with respect to the
+sample size) of computing the index function.
 
@@
 comparing with the current (last) basis. If a visited basis has a
-smaller index value than the current (last) one, JSO will keep
-the current basis for the next iteration in the search but still records
-the visited inferior basis. This data can be used to derive numerical
+smaller index value than the current (last) one, JSO will keep the
+current basis for the next iteration in the search but still records the
+visited inferior basis. This data can be used to derive numerical
 summaries and plots:
@@
 
-The performance of JSO is investigated in two folds: 1)
-compare it with a commonly used optimiser: the search-better optimiser
-[@RJ-2021-105; @laa_using_2020], and 2) examine its success rate under
-different hyper-parameters (number of jellyfishes and maximum number of
-tries). The projection pursuit problem used is finding the pipe shape
-using the holes index, investigated by @laa_using_2020.
+The performance of JSO is investigated in two folds: 1) compare it with
+a commonly used optimiser: the search-better optimiser [@RJ-2021-105;
+@laa_using_2020], and 2) examine its success rate under different
+hyper-parameters (number of jellyfishes and maximum number of tries).
+The projection pursuit problem used is finding the pipe shape using the
+holes index, investigated by @laa_using_2020.
 
@@
 
-Fifty simulations are conducted with both JSO and the search-better
-, in four dimensions ($d = 6, 8, 10, 12$). The JSO
-uses 100 jellyfishes and 100 maximum number of tries, while the
-search-better optimiser allows a maximum of 1000 samples at each
-iteration before the algorithm terminates. @fig-proj presents the final
-projections found by the two optimisers, broken down by 10th quantile,
-faceted by the data dimension. In the 6-dimensional data scenario, JSO
- consistently identifies a clear pipe shape. The
-search-better optimiser also finds the pipe shape but with a wide rim,
-suggesting a further polish search may be required. With increasing
-dimensions, JSO may not always identify the pipe shape due
-to random sampling, but it still presents a pipe shape in over 50% of
-cases. When compared to the search-better optimiser, JSO
-reaches higher index values and clearer pipe shapes across all
-quantiles.
+Fifty simulations are conducted with both JSO and the search-better , in
+four dimensions ($d = 6, 8, 10, 12$). The JSO uses 100 jellyfishes and
+100 maximum number of tries, while the search-better optimiser allows a
+maximum of 1000 samples at each iteration before the algorithm
+terminates. @fig-proj presents the final projections found by the two
+optimisers, broken down by 10th quantile, faceted by the data dimension.
+In the 6-dimensional data scenario, JSO consistently identifies a clear
+pipe shape. The search-better optimiser also finds the pipe shape but
+with a wide rim, suggesting a further polish search may be required.
+With increasing dimensions, JSO may not always identify the pipe shape
+due to random sampling, but it still presents a pipe shape in over 50%
+of cases. When compared to the search-better optimiser, JSO reaches
+higher index values and clearer pipe shapes across all quantiles.
 
@@
 hyper-parameter combination to analyse its effects on the success rate
-of JSO . The hyper-parameters tested include: 1) 20, 50, and
-100 jellyfishes, and 2) 50 and 100 maximum number of tries. In each
+of JSO . The hyper-parameters tested include: 1) 20, 50, and 100
+jellyfishes, and 2) 50 and 100 maximum number of tries. In each
 simulation, the success rate is calculated as defined in @sec-vis.
@@
 number of jellyfishes and the maximum number of tries), affect the
-success of JSO. Simulations are conducted to obtain the
-success rate across various projection pursuit problems, characterised
-by shape-to-find, data dimension, and index function used, as well as
+success of JSO. Simulations are conducted to obtain the success rate
+across various projection pursuit problems, characterised by
+shape-to-find, data dimension, and index function used, as well as
 different hyper-parameter combinations. Smoothness and squintability are
@@
 pipe-finding cases and `r nrow(filter(sim_df, index != "holes"))`
-sine-wave finding cases. For each case, JSO is run fifty
-times and the success rate is calculated as in @sec-app-1.
+sine-wave finding cases. For each case, JSO is run fifty times and the
+success rate is calculated as in @sec-app-1.
 
@@
 "smooth" is used as the smoothness measure and the column "squint" is
-calculated as [TODO: equation reference] as squintability. The low
+calculated as \[TODO: equation reference\] as squintability. The low
 squint value of MIC and TIC index is due to its convex shape of the
@@
 and @tbl-mod-output presents the model outputs. The model suggests that
-the success rate of JSO is positively associated with the
-two jellyfish hyper-parameters, as well as with smoothness and
-squintability. However, being flagged with long runtime and an increase
-of data dimension will reduce the success rate. The variable
-`squintability` and `dimension` are significant, suggesting their
-importance relative to jellyfish hyper-parameters in the optimisation
-success.
+the success rate of JSO is positively associated with the two jellyfish
+hyper-parameters, as well as with smoothness and squintability. However,
+being flagged with long runtime and an increase of data dimension will
+reduce the success rate. The variable `squintability` and `dimension`
+are significant, suggesting their importance relative to jellyfish
+hyper-parameters in the optimisation success.
 

### 2024-06-22 optim.tex
squintability subsection
added the content of the squintability subsection, with equations

@@
 \definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
 \makeatother
@@
 \makeatletter
-\makeatother
-\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
 \@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
 \makeatother
@@
 Statistics},city={Melbourne},country={Australia},countrysep={,},postcode={3800},postcodesep={}}
-\affiliation[3]{organization={BNU-HKBU United International
-College, Department of Mathematical
+\affiliation[3]{organization={Guangdong Provincial Key Laboratory of
+Interdisciplinary Research and Application for Data Science, BNU-HKBU
+United International College, Department of Mathematical
 Sciences},city={Zhuhai},country={China},countrysep={,},postcode={519087},postcodesep={}}
@@
 \end{frontmatter}
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 \citet{Loperfido2020}). Given high-dimensional data
-\(X \in \mathbf{R}^{n }\times p\) and the index function \(f(\cdot)\),
-PP finds the orthonormal projection matrix
+\(X \in \mathbf{R}^{n\times p}\) and the index function \(f(\cdot)\), PP
+finds the orthonormal projection matrix
 \(A \in \mathbf{R}^{p \times d}\) by solving the following optimisation
@@
 
-\[
+\begin{equation}\protect\hypertarget{eq-optimization}{}{
 \underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
-\]
+}\label{eq-optimization}\end{equation}
 
@@
 knowledge, this is the first attempt to quantitatively measure
-squintability and smoothness. Finally, the relationship between the JSO
-performance, hyper-parameters tuning and various properties of PP
-indexes is analysed to provide helpful guidance for practitioners that
-are using the guided tour for high dimensional visualisation.
+squintability and smoothness in this context. Finally, the relationship
+between the JSO performance, hyper-parameters tuning and various
+properties of PP indexes is analysed to provide helpful guidance for
+practitioners that are using the guided tour for high dimensional
+visualisation.
 
@@
 scagnostics is described in \citet{laa_using_2020}. To be useful here an
-optimiser needs to be able to handle functions which are not very
-smooth. In addition, because data structures might be relatively fine,
-the optimiser needs to be able to find maxima that occur with a small
-squint angle, that can only be seen from very close by. One last aspect
-that is useful is for an optimiser to return local maxima in addition to
-global because data can contain many different and interesting features.
-
+optimiser needs to be able to handle index functions which are possibly
+not very smooth. In addition, because data structures might be
+relatively fine, the optimiser needs to be able to find maxima that
+occur with a small squint angle, that can only be seen from very close
+by. One last aspect that is useful is for an optimiser to return local
+maxima in addition to global because data can contain many different and
+interesting features.
+
+\hypertarget{sec-theory}{%
 \section{The jellyfish optimiser and properties of PP indexes {[}Nicolas
-and Jessica{]}}\label{sec-theory}
+and Jessica{]}}\label{sec-theory}}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, coltitle=black, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, left=2mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, arc=.35mm, opacityback=0, breakable, colbacktitle=quarto-callout-note-color!10!white, bottomrule=.15mm, bottomtitle=1mm, titlerule=0mm, toprule=.15mm, leftrule=.75mm, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, opacityback=0, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, titlerule=0mm, bottomrule=.15mm, toprule=.15mm, leftrule=.75mm, bottomtitle=1mm, breakable, toptitle=1mm, arc=.35mm, colback=white, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, left=2mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white]
 
@@
 
-In this study, we propose two metrics to capture the notion of
-squintability.
+In this subsection, a metric to capture the notion of squintability is
+proposed, and two approaches to compute this metric numerically are
+detailed.
+
+\textcolor{orange}{Here we need to add a precise, mathematical
+definition of the $x$ and $y$ variables ("angle distance" to optimal?
+and PP index values) used in the figures of this Section.}
 
@@
 
-First, parametric model.
-
-{[}Nicolas's pdf{]}
-
-Second, we consider the product of the largest absolute magnitude of
-rate of change of \(f\) and the corresponding projection angle as a
-second measure of squintability. Since \(f\) is decreasing, the rate of
-change of \(f\) is negative and thus \(|\underset{x}{\min} f('x)|\)
-gives the absolute magnitude of the most negative rate of change.
-
-{[}Nicolas's pdf{]}
+Let \(f\) represent the average PP index value with respect to the
+projection angle distance to optimal, over the course of the JSO. It is
+expected that for a PP index with high squint angle, the optimization
+(\ref{eq-optimization}) should make substantial progress early on.
+Conversely, for a PP index with low squint angle, it might take a long
+while for the optimization to make substantial progress, as the
+candidate projections would need to be very close to the optimal one for
+the structure of the index function to be visible enough to be amenable
+to efficient optimization. This observation suggests that the extreme
+values of \(f'\) (the ones for which \(f''=0\), assuming that \(f\) is
+twice differentiable), and the angles for which these values are
+attained, should play an important role in the mathematical definition
+of squintability. Noticing that \(f\) is expected to be a decreasing
+function, we propose the following squintability metric:
+
+\begin{equation}\protect\hypertarget{eq-squintability}{}{
+\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)
+}\label{eq-squintability}\end{equation}
+
+On the one hand, \(-\min_{x}f'(x)\) represents the largest value of the
+gradient of \(-f\). On the other hand, \(\underset{x}{\arg\min}f'(x)\)
+represents the angle value at which this largest gradient is attained.
+Based on the above discussion, it is expected that these two values
+should be both high in the case of high squintability (fast increase in
+\(f\) early on), and both low in the case of low squintability (any
+substantial increase in \(f\) happens very late, close to the optimal
+angle). This suggest that their product (\ref{eq-squintability}) should
+provide a sensible measure of squintability. The multiplicative constant
+\(4\), which can be deemed arbitrary, does not change the interpretation
+of the squintability metric \(\varsigma\); it is here to adjust the
+range of values of \(\varsigma\) and simplify the explicit formula for
+\(\varsigma\) obtained later on.
+
+To compute the squintability metric (\ref{eq-squintability}) in
+practice, several approaches are possible. The first one is to propose a
+parametric model for \(f\), and use it to obtain an explicit formula for
+\(\varsigma\). Numerical experiments suggest a scaled sigmoid shape as
+described below. Define
+
+\begin{equation}\protect\hypertarget{eq-logistic}{}{
+\ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
+}\label{eq-logistic}\end{equation}
+
+which is a decreasing logistic function depending on two parameters
+\(\theta_2\) and \(\theta_3\) , such that
+\(\ell(\theta_{2})=\frac{1}{2}\). Then, define
+
+\begin{equation}\protect\hypertarget{eq-parametric}{}{
+f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
+}\label{eq-parametric}\end{equation}
+
+which depends on three additional parameters, \(\theta_1\), \(\theta_2\)
+, and \(x_{\max}\), such that \(f(0)=\theta_1\) and
+\(f(x_{\max})=\theta_4\). Under the parametric model
+(\ref{eq-parametric}), the squintability metric (\ref{eq-squintability})
+can be shown to be equal to
+
+\begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
+\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}
+}\label{eq-squintability-parametric}\end{equation}
+
+In practice, the parameters of this model (\ref{eq-parametric}) can be
+estimated numerically, for example by non-linear least squares, and then
+used to evaluate \(\varsigma\) as in equation
+(\ref{eq-squintability-parametric}).
+
+Alternatively, one can estimate (\ref{eq-squintability}) in a
+nonparametric way, for example by fitting \(f\) using kernel regression,
+then numerically estimate the angle at which \(-f'\) attains its highest
+value.
+
+\textcolor{orange}{See whether further details are needed for the nonparametric approach.}
 
@@
 
-\section{Visualisation of jellyfish optimiser}\label{sec-vis}
+\hypertarget{sec-vis}{%
+\section{Visualisation of jellyfish optimiser}\label{sec-vis}}
 
@@
 
-\end{figure}%
-
-\section{Application {[}Di and Sherry{]}}\label{sec-app}
+\end{figure}
+
+\hypertarget{sec-app}{%
+\section{Application {[}Di and Sherry{]}}\label{sec-app}}
 
@@
 
+\hypertarget{sec-app-1}{%
 \subsection{Performance of JSO in pipe-finding
-problems}\label{sec-app-1}
+problems}\label{sec-app-1}}
 
@@
 
+\hypertarget{sec-app-2}{%
 \subsection{Factors affecting JSO success rate: optimisation properties
-and jellyfish hyper-parameters}\label{sec-app-2}
+and jellyfish hyper-parameters}\label{sec-app-2}}
 
@@
 
+\hypertarget{tbl-mod-output}{}
 \begin{longtable}[t]{|>{}lrrr>{}r|}
-
-\caption{\label{tbl-mod-output}Model estimates of proportion of
-jellyfish success on optimisation properties and jellyfish
-hyper-parameters from the generalised linear model with a binomial
-family and a logit link function. The variable smoothness,
-squintability, number of jellyfish and maximum number of tries are
-positively associated with JSO success rate while data dimension and
-being flagged as long runtime are negatively associated with the success
-rate. The variable squintability and dimension are significant,
-suggesting their importance relative to jellyfish hyper-parameters in
-the optimisation success.}
-
-\tabularnewline
+\caption{\label{tbl-mod-output}Model estimates of proportion of jellyfish success on optimisation
+properties and jellyfish hyper-parameters from the generalised linear
+model with a binomial family and a logit link function. The variable
+smoothness, squintability, number of jellyfish and maximum number of
+tries are positively associated with JSO success rate while data
+dimension and being flagged as long runtime are negatively associated
+with the success rate. The variable squintability and dimension are
+significant, suggesting their importance relative to jellyfish
+hyper-parameters in the optimisation success. }\tabularnewline
 

### 2024-06-23 optim.qmd
add interpreation to glm

@@
   ggplot(aes(x = dist, y = index, group = index_name)) +
-  geom_point(size = 0.5) +
+  geom_line() + 
   facet_nested_wrap(~index_name + n, nrow = 3, labeller = label_both) + 
@@
 ```
-
-@tbl-mod-data combines all the variables calculated above (the JS
-optimiser success rate, jellyfish hyper-parameters, and optimisation
-properties) into one table. Three preprocessing steps are applied to the
-data: 1) scaling the jellyfish hyper-parameters by a factor of 10 for
-interpretation, 2) creating a new binary variable `long_time` to
-indicate cases with an average run time over 30 seconds, and 3) encoding
-the success rate for the stringy case as 0, since none of the 50
-simulations identified the sine-wave shape. A generalised linear model
-with a binomial family and a logit link function is used to fit the data
-and @tbl-mod-output presents the model outputs. The model suggests that
-the success rate of JSO is positively associated with the two jellyfish
-hyper-parameters, as well as with smoothness and squintability. However,
-being flagged with long runtime and an increase of data dimension will
-reduce the success rate. The variable `squintability` and `dimension`
-are significant, suggesting their importance relative to jellyfish
-hyper-parameters in the optimisation success.
 
@@
 
+@tbl-mod-data combines all the variables calculated above (the JSO success rate, JSO hyper-parameters, and optimisation
+properties) into one table. Three preprocessing steps are applied to the
+data: 1) scaling the JSO hyper-parameters by a factor of 10 for
+interpretation, 2) creating a new binary variable `long_time` to
+indicate cases with an average run time over 30 seconds, and 3) encoding
+the success rate for the stringy case as 0, because upon visual inspection of the final projection, none of the 50
+simulations identified the sine-wave shape. A generalised linear model
+with a binomial family and a logit link function is used to fit the data
+and @tbl-mod-output presents the model outputs. The model suggests that
+the JSO success rate is positively associated with the two JSO
+hyper-parameters, as well as with the optimisation properties -- smoothness and squintability.
+Specifically, using 10 more jellyfish and 10 more tries will increase the odd ratio of success by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`% and `r round(100 * (exp(broom::tidy(mod1)$estimate[7])-1), 2)`%, respectively.
+However,
+being flagged with long runtime and an increase of data dimension will
+reduce the success rate by `r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and `r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively.
+The variable `squintability` and `dimension`
+are significant, suggesting their importance relative to JSO
+hyper-parameters in the optimisation success. Users can increase the 
+number of jellyfish and the number of tries to improve the success rate, until JSO exceeds a 30-second runtime.
+
 # Conclusion \[Di and Sherry\] {#sec-conclusion}

### 2024-06-23 optim.tex
add interpreation to glm

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{equation}\protect\hypertarget{eq-optimization}{}{
+\begin{equation}\phantomsection\label{eq-optimization}{
 \underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
-}\label{eq-optimization}\end{equation}
+}\end{equation}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacityback=0, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, titlerule=0mm, bottomrule=.15mm, toprule=.15mm, leftrule=.75mm, bottomtitle=1mm, breakable, toptitle=1mm, arc=.35mm, colback=white, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, left=2mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white]
+\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, opacityback=0, bottomrule=.15mm, toprule=.15mm, bottomtitle=1mm, left=2mm, rightrule=.15mm, coltitle=black, titlerule=0mm, toptitle=1mm, opacitybacktitle=0.6, leftrule=.75mm, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, colback=white]
 
@@
 
-\begin{equation}\protect\hypertarget{eq-squintability}{}{
+\begin{equation}\phantomsection\label{eq-squintability}{
 \varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)
-}\label{eq-squintability}\end{equation}
+}\end{equation}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-logistic}{}{
+\begin{equation}\phantomsection\label{eq-logistic}{
 \ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
-}\label{eq-logistic}\end{equation}
+}\end{equation}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-parametric}{}{
+\begin{equation}\phantomsection\label{eq-parametric}{
 f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
-}\label{eq-parametric}\end{equation}
+}\end{equation}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
+\begin{equation}\phantomsection\label{eq-squintability-parametric}{
 \varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}
-}\label{eq-squintability-parametric}\end{equation}
+}\end{equation}
 
@@
 
-\end{figure}
-
-\hypertarget{sec-app}{%
-\section{Application {[}Di and Sherry{]}}\label{sec-app}}
+\end{figure}%
+
+\section{Application {[}Di and Sherry{]}}\label{sec-app}
 
@@
 
-\hypertarget{sec-app-1}{%
 \subsection{Performance of JSO in pipe-finding
-problems}\label{sec-app-1}}
+problems}\label{sec-app-1}
 
@@
 
-\hypertarget{sec-app-2}{%
 \subsection{Factors affecting JSO success rate: optimisation properties
-and jellyfish hyper-parameters}\label{sec-app-2}}
+and jellyfish hyper-parameters}\label{sec-app-2}
 
@@
 
-\end{figure}
-
-Table~\ref{tbl-mod-data} combines all the variables calculated above
-(the JS optimiser success rate, jellyfish hyper-parameters, and
-optimisation properties) into one table. Three preprocessing steps are
-applied to the data: 1) scaling the jellyfish hyper-parameters by a
-factor of 10 for interpretation, 2) creating a new binary variable
-\texttt{long\_time} to indicate cases with an average run time over 30
-seconds, and 3) encoding the success rate for the stringy case as 0,
-since none of the 50 simulations identified the sine-wave shape. A
-generalised linear model with a binomial family and a logit link
-function is used to fit the data and Table~\ref{tbl-mod-output} presents
-the model outputs. The model suggests that the success rate of JSO is
-positively associated with the two jellyfish hyper-parameters, as well
-as with smoothness and squintability. However, being flagged with long
-runtime and an increase of data dimension will reduce the success rate.
-The variable \texttt{squintability} and \texttt{dimension} are
-significant, suggesting their importance relative to jellyfish
-hyper-parameters in the optimisation success.
-
-\hypertarget{tbl-smoothness-squintability}{}
+\end{figure}%
+
 \begin{table}
-\caption{\label{tbl-smoothness-squintability}Parameters estimated from the Gaussian process (including variance,
-range, smoothness, and nugget) and scaled logistic function
-(\(\theta_1\) to \(\theta_4\)) for the pipe-finding and sine-wave
-finding problems. The column ``smooth'' and ``squint'' represent the
-smoothness and squintability measures. }\tabularnewline
+
+\caption{\label{tbl-smoothness-squintability}Parameters estimated from
+the Gaussian process (including variance, range, smoothness, and nugget)
+and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
+pipe-finding and sine-wave finding problems. The column ``smooth'' and
+``squint'' represent the smoothness and squintability measures.}
+
+\centering{
 
@@
 
-\hypertarget{tbl-mod-output}{}
 \begin{longtable}[t]{|>{}lrrr>{}r|}
-\caption{\label{tbl-mod-output}Model estimates of proportion of jellyfish success on optimisation
-properties and jellyfish hyper-parameters from the generalised linear
-model with a binomial family and a logit link function. The variable
-smoothness, squintability, number of jellyfish and maximum number of
-tries are positively associated with JSO success rate while data
-dimension and being flagged as long runtime are negatively associated
-with the success rate. The variable squintability and dimension are
-significant, suggesting their importance relative to jellyfish
-hyper-parameters in the optimisation success. }\tabularnewline
+
+\caption{\label{tbl-mod-output}Model estimates of proportion of
+jellyfish success on optimisation properties and jellyfish
+hyper-parameters from the generalised linear model with a binomial
+family and a logit link function. The variable smoothness,
+squintability, number of jellyfish and maximum number of tries are
+positively associated with JSO success rate while data dimension and
+being flagged as long runtime are negatively associated with the success
+rate. The variable squintability and dimension are significant,
+suggesting their importance relative to jellyfish hyper-parameters in
+the optimisation success.}
+
+\tabularnewline
 
@@
 \bottomrule
+
 \end{longtable}
+
 \endgroup{}
 
-\hypertarget{sec-conclusion}{%
-\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}}
+Table~\ref{tbl-mod-data} combines all the variables calculated above
+(the JSO success rate, JSO hyper-parameters, and optimisation
+properties) into one table. Three preprocessing steps are applied to the
+data: 1) scaling the JSO hyper-parameters by a factor of 10 for
+interpretation, 2) creating a new binary variable \texttt{long\_time} to
+indicate cases with an average run time over 30 seconds, and 3) encoding
+the success rate for the stringy case as 0, because upon visual
+inspection of the final projection, none of the 50 simulations
+identified the sine-wave shape. A generalised linear model with a
+binomial family and a logit link function is used to fit the data and
+Table~\ref{tbl-mod-output} presents the model outputs. The model
+suggests that the JSO success rate is positively associated with the two
+JSO hyper-parameters, as well as with the optimisation properties --
+smoothness and squintability. Specifically, using 10 more jellyfish and
+10 more tries will increase the odd ratio of success by 24.11\% and
+11.93\%, respectively. However, being flagged with long runtime and an
+increase of data dimension will reduce the success rate by 41.72\% and
+53.36\%, respectively. The variable \texttt{squintability} and
+\texttt{dimension} are significant, suggesting their importance relative
+to JSO hyper-parameters in the optimisation success. Users can increase
+the number of jellyfish and the number of tries to improve the success
+rate, until JSO exceeds a 30-second runtime.
+
+\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
 

### 2024-06-23 appendix/appendix.qmd
update appendix

---
title: 'Appendix to "Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit"'
author:
  - name: H. Sherry Zhang
    email: huize.zhang@austin.utexas.edu
    affiliations: 
        - id: 1
          name: University of Texas at Austin
          department: Department of Statistics and Data Sciences
          city: Austin
          country: United States
          postal-code: 78751
    attributes:
        corresponding: true
  - name: Dianne Cook
    email: dicook@monash.edu
    affiliations:
        - id: 2
          name: Monash University
          department: Department of Econometrics and Business Statistics
          city: Melbourne
          country: Australia
          postal-code: 3800
  - name: Nicolas  Langrené
    email: nicolaslangrene@uic.edu.cn
    affiliations:
        - id: 3
          name: Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College
          department: Department of Mathematical Sciences
          city: Zhuhai
          country: China
          postal-code: 519087
  - name: Jessica Wai Yin Leung
    email: Jessica.Leung@monash.edu
    affiliations:
        - id: 2
          name: Monash University
          department: Department of Econometrics and Business Statistics
          city: Melbourne
          country: Australia
          postal-code: 3800
date: last-modified
abstract: |
  Projection pursuit (PP) is a dimension reduction technique that identifies low-dimensional projections in high-dimensional data by optimising a criteria function known as the PP index. The optimisation in PP can be non-smooth and requires identifying optima with a small "squint angle", detectable only from close proximity. To address these challenges, this study investigates the performance of a recent swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes. The performance of JSO is evaluated across various hyper-parameter settings and compared with existing optimisers. Additionally, this work proposes novel methods to quantify two properties of the PP index -- smoothness and squintability -- that capture the complexities inherent in PP optimisation problems. These two metrics are evaluated with JSO hyper-parameters to determine their effect on JSO success rate. The JSO algorithm has been implemented in the `tourr` package, while calculations for smoothness and squintability are available in the `ferrn` package. 
format:
  elsevier-pdf:
    journal:
      name: Journal of Multivariate Analysis
      formatting: preprint
      model: 3p
      cite-style: numbername
bibliography: ../bibliography.bib
---

Given high-dimensional data $X \in \mathbf{R}^{n\times p}$ and the index function $f(\cdot)$, projection pursuit finds the orthonormal projection basis $A \in \mathbf{R}^{p \times d}$ by solving the following optimisation problem:

$$
\underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
$$ {#eq-optimization}

This appendix presents definition of all indexes $f(\cdot)$ used in the paper. These indexes are defined in 2D cases ($d = 2$), where $Y \in \mathbf{R}^{n \times 2} = (y_1, y_2)$ represents the projected data.

## The `holes` index

The `holes` index [@cook1993projection] is a smooth index for detecting the presence of multi-modality in the projection. The index is defined as:

$$I_{\text{holes}}(A) = \frac{1-\frac{1}{n} \sum^n_{i = 1} \exp(-\frac{1}{2y_iy_i^\prime})}{1-\exp(-\frac{d}{2})}$$

## The `dcor2d` index

The `dcor2d` index [@Grimm2016] is a smooth index that measures the distance correlation between the two projection axes. The index uses pair-wise distance to define column sum, row sum, and overall sum and then calculates distance correlation from distance variance and covariance.

  -  pair-wised distance: $a_{ij} = \|y_1^i - y_1^j\|$ and $b_{ij} = \|y_2^i - y_2^j\|$ for $i, j = 1, 2, \ldots, n$

   -  column sum, row sum, and overall sum: $$a_{i \cdot} = \sum_{l = 1}^n a_{il}, \quad a_{\cdot j} = \sum_{k = 1}^n a_{kj}, \quad a_{\cdot \cdot} = \sum_{k,l = 1}^n a_{kl} \quad b_{i \cdot} = \sum_{l = 1}^n b_{il}, \quad b_{\cdot j} = \sum_{k = 1}^n b_{kj}, \quad b_{\cdot \cdot} = \sum_{k,l = 1}^n b_{kl}$$

  -   distance covariance (and variance defined similarly): $$\text{dCov}(a_{ij}, b_{ij}) = \frac{1}{n(n-3)} \sum_{i \neq j} a_{ij} b_{ij} -\frac{2}{n (n-2)(n-3)}\sum_{i = 1}^n a_{i \cdot} b_{i \cdot} + \frac{1}{n(n - 1)(n-2)(n-3)} a_{\cdot \cdot} b_{\cdot \cdot}$$

   -   distance correlation: $$I_{\text{dcor}}(A)  = \text{dCor} (a_{ij}, b_{ij}) = \frac{\text{dCov}(a_{ij}, b_{ij})}{\sqrt{\text{dVar}(a_{ij}) \text{dVar}(b_{ij})}}$$

# The `MIC` and `TIC` index

Both `MIC` and `TIC` are information-based indexes derived from mutual information. The indexes are defined based on a partition, $g(k, l)$, of the space $(y_1, y_2)$ into $k \times l$ rectangles, For example, $g(2, 3)$ divides the data space into 2 rectangles in the $y_2$ direction and 3 rectangles in the $y_1$ direction. Let $G$ denotes all the possible partition.

MIC finds the **maximum** mutual information over G where $k$ and $l$ are bounded by the grid size: $k \times l < B(n) = n^{\alpha}$. $alpha = 0.3$ is used in our simulation.

$$I_{MIC}(A) = \max_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \max_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$

where $k^*$ and $l^*$ are the number of rectangles in the optimal partition $g$.

TIC calculates the **sum** of mutual information over G

$$I_{TIC}(A) = \sum_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \sum_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$

# The `loess` and `splines` index

The `loess` and `splines` indexes detect non-linear structure in the projection, as captured by their respective model. These indexes are computed by regressing both axes against each other, and find the maximum variance from the residual that can be explained by the model. Let $e^{\text{model}}_{y_1 \sim y_2}$ denote the residual from the loess/splines model $y_1 \sim y_2$, and $e^{\text{model}}_{y_2 \sim y_1}$ denote the residual from the loess/spliens model $y_2 \sim y_1$. The two indexes are calculated as

 $$I_{\text{loess}}(A) = \max \left(1 - \frac{var(e^{\text{loess}}_{y_1 \sim y_2})}{var(y_1)}, 1 - \frac{var(e^{\text{loess}}_{y_2 \sim y_1})}{var(y_2)}\right)$$

$$I_{\text{spline}}(A) = \max \left(1 - \frac{var(e^{\text{spline}}_{y_1 \sim y_2})}{var(y_1)}, 1 - \frac{var(e^{\text{spline}}_{y_2 \sim y_1})}{var(y_2)}\right)$$

# The `stringy` index

The `stringy` [@scag; @WW08] index is a non-smooth index based on scagnostics. It measures the proportion of vertices with two edges in the minimum spanning tree (MST) of the projection to detect whether the projection forms a straight line. The index is calculated as

$$I_{stringy}(A) = \frac{\text{number of vertices with 2 edges}}{\text{number of total vertices with more than one edge}} $$

### 2024-06-24 optim.qmd
update appendix
add definition of porjecion distance

@@
 
+Let $A \in \mathbf{R}^{p \times 2}$ be the 2D projection basis and $A^*$ the optimal basis achieving the maximum index value for a given projection pursuit problem defined by the shape to find, data dimension, and the index function $f(.)$.
+The projection distance $d(A, A^*)$, is defined as the Frobenius norm $\lVert .
+\rVert _F$, between the squared difference of these two bases:
+
+$$d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}$$ {#eq-proj-dist}
+
+The index value $f(XA)$ is defined as described in the introduction.
+
 ```{=tex}
@@
 
-```{=tex}
 \textcolor{orange}{See whether further details are needed for the nonparametric approach.}
-```
-TODO: add equation reference to main text line 705
 
@@
 
-## Speed
+<!-- ## Speed -->
 
-The speed of optimizing an index function can be measured empirically by
-recording the duration of the optimisation. Alternatively, one can also
-gauge the speed in terms of computational complexity of evaluating the
-index function (for instance, in big O notation, with respect to the
-sample size) of computing the index function.
+<!-- The speed of optimizing an index function can be measured empirically by -->
+<!-- recording the duration of the optimisation. Alternatively, one can also -->
+<!-- gauge the speed in terms of computational complexity of evaluating the -->
+<!-- index function (for instance, in big O notation, with respect to the -->
+<!-- sample size) of computing the index function. -->
 
@@
 "smooth" is used as the smoothness measure and the column "squint" is
-calculated as \[TODO: equation reference\] as squintability. The low
+calculated as equation
+(@eq-squintability-parametric) as squintability. The low
 squint value of MIC and TIC index is due to its convex shape of the
@@
 
-@tbl-mod-data combines all the variables calculated above (the JSO success rate, JSO hyper-parameters, and optimisation
-properties) into one table. Three preprocessing steps are applied to the
-data: 1) scaling the JSO hyper-parameters by a factor of 10 for
-interpretation, 2) creating a new binary variable `long_time` to
-indicate cases with an average run time over 30 seconds, and 3) encoding
-the success rate for the stringy case as 0, because upon visual inspection of the final projection, none of the 50
-simulations identified the sine-wave shape. A generalised linear model
-with a binomial family and a logit link function is used to fit the data
-and @tbl-mod-output presents the model outputs. The model suggests that
-the JSO success rate is positively associated with the two JSO
-hyper-parameters, as well as with the optimisation properties -- smoothness and squintability.
-Specifically, using 10 more jellyfish and 10 more tries will increase the odd ratio of success by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`% and `r round(100 * (exp(broom::tidy(mod1)$estimate[7])-1), 2)`%, respectively.
-However,
-being flagged with long runtime and an increase of data dimension will
-reduce the success rate by `r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and `r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively.
-The variable `squintability` and `dimension`
-are significant, suggesting their importance relative to JSO
-hyper-parameters in the optimisation success. Users can increase the 
-number of jellyfish and the number of tries to improve the success rate, until JSO exceeds a 30-second runtime.
+@tbl-mod-data combines all the variables calculated above (the JSO
+success rate, JSO hyper-parameters, and optimisation properties) into
+one table. Three preprocessing steps are applied to the data: 1) scaling
+the JSO hyper-parameters by a factor of 10 for interpretation, 2)
+creating a new binary variable `long_time` to indicate cases with an
+average run time over 30 seconds, and 3) encoding the success rate for
+the stringy case as 0, because upon visual inspection of the final
+projection, none of the 50 simulations identified the sine-wave shape. A
+generalised linear model with a binomial family and a logit link
+function is used to fit the data and @tbl-mod-output presents the model
+outputs. The model suggests that the JSO success rate is positively
+associated with the two JSO hyper-parameters, as well as with the
+optimisation properties -- smoothness and squintability. Specifically,
+using 10 more jellyfish and 10 more tries will increase the odd ratio of
+success by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`%
+and `r round(100 * (exp(broom::tidy(mod1)$estimate[7])-1), 2)`%,
+respectively. However, being flagged with long runtime and an increase
+of data dimension will reduce the success rate by
+`r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and
+`r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively.
+The variable `squintability` and `dimension` are significant, suggesting
+their importance relative to JSO hyper-parameters in the optimisation
+success. Users can increase the number of jellyfish and the number of
+tries to improve the success rate, until JSO exceeds a 30-second
+runtime.
 

### 2024-06-24 optim.tex
add definition of porjecion distance

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, opacityback=0, bottomrule=.15mm, toprule=.15mm, bottomtitle=1mm, left=2mm, rightrule=.15mm, coltitle=black, titlerule=0mm, toptitle=1mm, opacitybacktitle=0.6, leftrule=.75mm, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, rightrule=.15mm, bottomrule=.15mm, colback=white, bottomtitle=1mm, arc=.35mm, opacityback=0, opacitybacktitle=0.6, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm, breakable, coltitle=black, toptitle=1mm, titlerule=0mm, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}]
 
@@
 detailed.
+
+Let \(A \in \mathbf{R}^{p \times 2}\) be the 2D projection basis and
+\(A^*\) the optimal basis achieving the maximum index value for a given
+projection pursuit problem defined by the shape to find, data dimension,
+and the index function \(f(.)\). The projection distance \(d(A, A^*)\),
+is defined as the Frobenius norm \(\lVert . \rVert _F\), between the
+squared difference of these two bases:
+
+\begin{equation}\phantomsection\label{eq-proj-dist}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}}\end{equation}
+
+The index value \(f(XA)\) is defined as described in the introduction.
 
@@
 
-TODO: add equation reference to main text line 705
-
 To the best of our knowledge, this is the first attempt to measure the
 notion of squintability.
-
-\subsection{Speed}\label{speed}
-
-The speed of optimizing an index function can be measured empirically by
-recording the duration of the optimisation. Alternatively, one can also
-gauge the speed in terms of computational complexity of evaluating the
-index function (for instance, in big O notation, with respect to the
-sample size) of computing the index function.
 
@@
 for each case. The column ``smooth'' is used as the smoothness measure
-and the column ``squint'' is calculated as {[}TODO: equation
-reference{]} as squintability. The low squint value of MIC and TIC index
-is due to its convex shape of the index value against the projection
-distance, as shown in Figure~\ref{fig-idx-proj-dist}. Comparing to other
-concave shapes, the maximum first derivative happens at a smaller
-projection distance (\(\theta_2\)), requiring the optimiser to get
-closer to the optimal to see a significant change in the index value,
-hence a small squintability measure.
+and the column ``squint'' is calculated as equation
+(\ref{eq-squintability-parametric}) as squintability. The low squint
+value of MIC and TIC index is due to its convex shape of the index value
+against the projection distance, as shown in
+Figure~\ref{fig-idx-proj-dist}. Comparing to other concave shapes, the
+maximum first derivative happens at a smaller projection distance
+(\(\theta_2\)), requiring the optimiser to get closer to the optimal to
+see a significant change in the index value, hence a small squintability
+measure.
 

### 2024-06-24 optim.qmd
edit introduction

@@
 -   British English ("American or British usage is accepted, but not a
-    mixture of these")\*
+    mixture of these")
 
@@
     -   "hyper-parameter" rather than "hyperparameter"
-    -   PP index rather than PPI
+    -   PP index rather than PPI, plural PP indexes
+    -   datasets rather than data sets
 
@@
 
-Projection Pursuit (PP) is a high-dimensional data visualisation
-approach that involves computing informative linear projections of data
-by optimising a variety of objective functions, namely *PP index*
+Projection Pursuit (PP) (@kr69, @FT74) is a dimension reduction technique 
+aimed at identifying informative linear projections of data. The method involves 
+optimising a objective function known as PP index
 (@hall1989polynomial, @cook1993projection, @lee2010projection,
-@Loperfido2018, @Loperfido2020). Given high-dimensional data
-$X \in \mathbf{R}^{n\times p}$ and the index function $f(\cdot)$, PP
-finds the orthonormal projection matrix $A \in \mathbf{R}^{p \times d}$
-by solving the following optimisation problem:
+@Loperfido2018, @Loperfido2020), which defines the criteria for what constitutes  interesting or informative projections. Given high-dimensional data
+$X \in \mathbf{R}^{n\times p}$ and an index function $f(\cdot)$, PP
+finds the orthonormal projection matrix $A \in \mathbf{R}^{p \times d}$ that 
+maximises the index value of the projection, $Y = XA$:
 
@@
 
-These indices $f(\cdot)$ are characterised by the informativeness or
-"interestingness" of a projection and are often non-linear and
-non-convex. As such, an effective and efficient optimisation procedure
-is essential to enable sufficient exploration of the data landscape
-during the visualisation process and to arrive at a global optimal
-viewpoint of the data.
+The index functions can be non-linear and non-convex, hence an 
+effective and efficient optimisation procedure is essential to 
+explore the data landscape and achieving a globally optimal viewpoint 
+of the data.
 
@@
 PP and facilitating a better experience in the visualisation process.
-@cook1995grand introduced the PP guided tour, which enabled interactive
-visualisation of the optimisation to visually explore high-dimensional
-data. It is implemented in the R [@R] package `tourr`[@tourr].
+@cook1995grand introduces the PP guided tour, which enabled interactive 
+visualisation of the optimisation in order to visually explore the high-dimensional data. It is implemented in the R [@R] package `tourr` [@tourr].
 @RJ-2021-105 highlighted potential problems of the optimisation routine
@@
 
-The artificial Jellyfish Search Optimiser (JSO) [@chou_novel_2021] is a
-swarm-based metaheuristic designed to solve global optimisation
-problems. Inspired by the search behaviour of jellyfish in the ocean,
-JSO is one of the latest swarm intelligence algorithms
-[@rajwar_exhaustive_2023], which was shown to have stronger search
-ability and faster convergence with fewer parameters compared to classic
-optimisation methods [@chou_novel_2021]-[@chou_recent_2022]. These
-practical properties makes JSO a strong candidate in integration with
-the PP guided tour. As such, it is of interest to explore the potential
-of JSO in enhancing the PP guided tour and examine the characteristics
+This study explores the potential for a Jellyfish Search Optimiser (JSO) to be integrated with projection 
+pursuit guided tour. JSO [@chou_novel_2021] -[@rajwar_exhaustive_2023], inspired by the search behaviour of jellyfish 
+in the ocean, is a swarm-based metaheuristic designed to solve global
+optimisation problems. Compared to traditional methods, JSO 
+demonstrated stronger search ability and faster convergence, and 
+requires fewer tuning parameters. These
+practical advantages make JSO as a promising candidate for enhancing
+PP optimisation. As such, the study aims to explore JSO's potential
+in enhancing projection pursuit and examines the characteristics
 and advantages of such implementation.
@@
 The primary goal of this study is to investigate the performance of JSO
-in the context of the Projection pursuit. A series of simulation
-experiments using well-recognised data sets and PP indexes are conducted
+in PP optimisation. A series of simulation
+experiments using well-recognised datasets and various PP indexes are conducted
 to yield insights regarding the behaviour of JSO under different
-settings and the sensitivity of hyper-parameter in the optimisation.
+settings and its sensitivity to hyper-parameter.
 Second, to observe the performance of JSO with different types of PP
-indexes, a collection of metrics is introduced to capture specific
+indexes, metrics are introduced to capture specific
 properties of the index including squintability and smoothness as
@@
 this is the first attempt to quantitatively measure squintability and
-smoothness in this context. Finally, the relationship between the JSO
+smoothness within this context. Finally, the relationship between the JSO
 performance, hyper-parameters tuning and various properties of PP
-indexes is analysed to provide helpful guidance for practitioners that
-are using the guided tour for high dimensional visualisation.
+indexes is analysed to provide helpful guidance for practitioners using projection pursuit.
 
 The rest of this paper is structured as follows. @sec-background
-introduces the background of the PP guided tour and reviews existing
-methods in the literature. @sec-theory describes the details of JSO and
+introduces the background of PP guided tour and reviews existing
+optimisers and index functions in the literature. @sec-theory details JSO and
 introduces metrics that measure different properties of PP indexes.
-@sec-vis visualises the behaviour of the JSO in terms of describes a
-simulation study on the performance of the JSO using well-known data
-sets and index functions. @sec-app displays two sets of simulation
-experiments comparing the JSO with the search-better optimiser and
+@sec-vis outlines the data structures, numerical summaries, and visualisations used in the simulation studies conducted in @sec-app to assess JSO's perforamnce.
+@sec-app presents two sets of simulation
+experiments, comparing JSO with another optimiser, search better, and
 studies the impact of different PP index properties on the optimisation

### 2024-06-24 optim.tex
edit introduction

@@
   British English (``American or British usage is accepted, but not a
-  mixture of these'')*
+  mixture of these'')
 \item
@@
   \item
-    PP index rather than PPI
+    PP index rather than PPI, plural PP indexes
+  \item
+    datasets rather than data sets
   \end{itemize}
@@
 
-Projection Pursuit (PP) is a high-dimensional data visualisation
-approach that involves computing informative linear projections of data
-by optimising a variety of objective functions, namely \emph{PP index}
-(\citet{hall1989polynomial}, \citet{cook1993projection},
+Projection Pursuit (PP) (\citet{kr69}, \citet{FT74}) is a dimension
+reduction technique aimed at identifying informative linear projections
+of data. The method involves optimising a objective function known as PP
+index (\citet{hall1989polynomial}, \citet{cook1993projection},
 \citet{lee2010projection}, \citet{Loperfido2018},
-\citet{Loperfido2020}). Given high-dimensional data
-\(X \in \mathbf{R}^{n\times p}\) and the index function \(f(\cdot)\), PP
+\citet{Loperfido2020}), which defines the criteria for what constitutes
+interesting or informative projections. Given high-dimensional data
+\(X \in \mathbf{R}^{n\times p}\) and an index function \(f(\cdot)\), PP
 finds the orthonormal projection matrix
-\(A \in \mathbf{R}^{p \times d}\) by solving the following optimisation
-problem:
+\(A \in \mathbf{R}^{p \times d}\) that maximises the index value of the
+projection, \(Y = XA\):
 
@@
 
-These indices \(f(\cdot)\) are characterised by the informativeness or
-``interestingness'' of a projection and are often non-linear and
-non-convex. As such, an effective and efficient optimisation procedure
-is essential to enable sufficient exploration of the data landscape
-during the visualisation process and to arrive at a global optimal
-viewpoint of the data.
+The index functions can be non-linear and non-convex, hence an effective
+and efficient optimisation procedure is essential to explore the data
+landscape and achieving a globally optimal viewpoint of the data.
 
@@
 PP and facilitating a better experience in the visualisation process.
-\citet{cook1995grand} introduced the PP guided tour, which enabled
-interactive visualisation of the optimisation to visually explore
-high-dimensional data. It is implemented in the R \citep{R} package
-\texttt{tourr}\citep{tourr}. \citet{RJ-2021-105} highlighted potential
-problems of the optimisation routine implemented. While improving the
-quality of the optimisation solutions in the tour is essential, it is
-also important to be able to watch the projected data as the
-optimisation progresses. As such, it is pertinent to integrate the
-guided tour with a global optimisation algorithm that is efficient in
-finding the global optimal and enables viewing of the projected data
+\citet{cook1995grand} introduces the PP guided tour, which enabled
+interactive visualisation of the optimisation in order to visually
+explore the high-dimensional data. It is implemented in the R \citep{R}
+package \texttt{tourr} \citep{tourr}. \citet{RJ-2021-105} highlighted
+potential problems of the optimisation routine implemented. While
+improving the quality of the optimisation solutions in the tour is
+essential, it is also important to be able to watch the projected data
+as the optimisation progresses. As such, it is pertinent to integrate
+the guided tour with a global optimisation algorithm that is efficient
+in finding the global optimal and enables viewing of the projected data
 during the exploration process.
 
-The artificial Jellyfish Search Optimiser (JSO) \citep{chou_novel_2021}
-is a swarm-based metaheuristic designed to solve global optimisation
-problems. Inspired by the search behaviour of jellyfish in the ocean,
-JSO is one of the latest swarm intelligence algorithms
-\citep{rajwar_exhaustive_2023}, which was shown to have stronger search
-ability and faster convergence with fewer parameters compared to classic
-optimisation methods \citep{chou_novel_2021}-\citep{chou_recent_2022}.
-These practical properties makes JSO a strong candidate in integration
-with the PP guided tour. As such, it is of interest to explore the
-potential of JSO in enhancing the PP guided tour and examine the
-characteristics and advantages of such implementation.
+This study explores the potential for a Jellyfish Search Optimiser (JSO)
+\citep{chou_novel_2021} -\citep{chou_recent_2022} to be integrated with
+projection pursuit guided tour. JSO, inspired by the search behaviour of
+jellyfish in the ocean, is a swarm-based metaheuristic designed to solve
+global optimisation problems. Compared to traditional methods, JSO
+demonstrated stronger search ability and faster convergence, and
+requires fewer tuning parameters. These practical advantages make JSO as
+a promising candidate for enhancing PP optimisation. As such, the study
+aims to explore JSO's potential in enhancing projection pursuit and
+examines the characteristics and advantages of such implementation.
 
 The primary goal of this study is to investigate the performance of JSO
-in the context of the Projection pursuit. A series of simulation
-experiments using well-recognised data sets and PP indexes are conducted
-to yield insights regarding the behaviour of JSO under different
-settings and the sensitivity of hyper-parameter in the optimisation.
-Second, to observe the performance of JSO with different types of PP
-indexes, a collection of metrics is introduced to capture specific
-properties of the index including squintability and smoothness as
-introduced in \citet{laa_using_2020}. To the best of the authors'
-knowledge, this is the first attempt to quantitatively measure
-squintability and smoothness in this context. Finally, the relationship
-between the JSO performance, hyper-parameters tuning and various
-properties of PP indexes is analysed to provide helpful guidance for
-practitioners that are using the guided tour for high dimensional
-visualisation.
+in PP optimisation. A series of simulation experiments using
+well-recognised datasets and various PP indexes are conducted to yield
+insights regarding the behaviour of JSO under different settings and its
+sensitivity to hyper-parameter. Second, to observe the performance of
+JSO with different types of PP indexes, metrics are introduced to
+capture specific properties of the index including squintability and
+smoothness as introduced in \citet{laa_using_2020}. To the best of the
+authors' knowledge, this is the first attempt to quantitatively measure
+squintability and smoothness within this context. Finally, the
+relationship between the JSO performance, hyper-parameters tuning and
+various properties of PP indexes is analysed to provide helpful guidance
+for practitioners using projection pursuit.
 
 The rest of this paper is structured as follows.
-Section~\ref{sec-background} introduces the background of the PP guided
-tour and reviews existing methods in the literature.
-Section~\ref{sec-theory} describes the details of JSO and introduces
-metrics that measure different properties of PP indexes.
-Section~\ref{sec-vis} visualises the behaviour of the JSO in terms of
-describes a simulation study on the performance of the JSO using
-well-known data sets and index functions. Section~\ref{sec-app} displays
-two sets of simulation experiments comparing the JSO with the
-search-better optimiser and studies the impact of different PP index
-properties on the optimisation performance. Section~\ref{sec-conclusion}
-summarises the work and provides suggestions for future directions.
+Section~\ref{sec-background} introduces the background of PP guided tour
+and reviews existing optimisers and index functions in the literature.
+Section~\ref{sec-theory} details JSO and introduces metrics that measure
+different properties of PP indexes. Section~\ref{sec-vis} outlines the
+data structures, numerical summaries, and visualisations used in the
+simulation studies conducted in Section~\ref{sec-app} to assess JSO's
+perforamnce. Section~\ref{sec-app} presents two sets of simulation
+experiments, comparing JSO with another optimiser, search better, and
+studies the impact of different PP index properties on the optimisation
+performance. Section~\ref{sec-conclusion} summarises the work and
+provides suggestions for future directions.
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, rightrule=.15mm, bottomrule=.15mm, colback=white, bottomtitle=1mm, arc=.35mm, opacityback=0, opacitybacktitle=0.6, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm, breakable, coltitle=black, toptitle=1mm, titlerule=0mm, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, breakable, bottomrule=.15mm, toprule=.15mm, opacityback=0, titlerule=0mm, coltitle=black, bottomtitle=1mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, rightrule=.15mm, left=2mm, arc=.35mm, colback=white, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white]
 

### 2024-06-24 optim.qmd
edit section 4 and 5

@@
 
-This study explores the potential for a Jellyfish Search Optimiser (JSO) to be integrated with projection 
-pursuit guided tour. JSO [@chou_novel_2021] -[@rajwar_exhaustive_2023], inspired by the search behaviour of jellyfish 
+This study explores the potential for a Jellyfish Search Optimiser 
+(JSO) to be integrated with projection 
+pursuit guided tour. JSO [@chou_novel_2021] -[@rajwar_exhaustive_2023], 
+inspired by the search behaviour of jellyfish 
 in the ocean, is a swarm-based metaheuristic designed to solve global
@@
 
-The data structure proposed in @RJ-2021-105 is implemented for JS
-optimisers and additional information such as index function used, data
-dimension, and jellyfish hyper-parameters can be included as additional
-columns. An example data collected from finding the sine-wave structure
-in 6D data (`d = 6`) using a spline index function(`splines2d`) is
+The data structure proposed in @RJ-2021-105 is implemented for JSO. 
+An example data collected from finding the sine-wave structure
+in 6D data using a spline index function (`splines2d`) is
 printed below:
@@
 
-The jellyfish hyper-parameters used are 50 jellyfishes (`n_jellies`) and
+The JSO hyper-parameters used are 50 jellyfishes (`n_jellies`) and
 a maximum of 50 tries (`max_tries`). The full data contains 50
-simulations, recorded by the variable `sim`, with different seeds
-(`seed`). Recorded in columns `basis`, `index_val`, `tries`, `loop` and
-`time`, are the sampled projection basis, calculated index value,
-iteration ID, jellyfish ID, and time taken to find the optimum,
+simulations, indexed by the variable `sim`, with the associated `seed`s. 
+Recorded in columns `basis`, `index_val`, `tries`, `loop` and
+`time`, are the sampled projection basis, index value,
+iteration ID, jellyfish ID, and computational time
 respectively. The basis column records every visited basis before
@@
 -   **Success rate**: The success rate is defined as the proportion of
-    simulations that achieve a final index value within 0.05 of the best
+    simulations that achieves a final index value within 0.05 of the best
     index value found among all 50 simulations. In the dataset above,
@@
 -   **Projection distance**: The projection distance between projection
-    bases and the theoretical best basis (or the empirical best basis if
-    the theoretical best is unknown) is calculated as the Frobenius
-    norm, which measures how close a visited basis is to the best basis.
-    Plotting this projection distance against the index value for each
-    basis visited reveals how the index value changes as the basis
-    approaches the optimal basis, helping to identify the squintability
-    of the index function. @fig-index-value-proj-dist shows the index
-    value against the projection distance for the given example and its
+    bases and the optimal best basis (or the empirical best basis if
+    the optimal is unknown) measures how close a visited basis is to 
+    the best basis and it is calculated as the Frobenius norm in Equation
+    @eq-proj-dist.
+    Plotting projection distances against index values for the 
+    bases visited reveals how the index value changes as the basis
+    approaches the optimal basis, helping to calculate the squintability
+    measure. @fig-index-value-proj-dist shows the index
+    value against the projection distance for the given dataset and its
     binned version. As the projection distance decreases, the index
@@
 #| fig.width: 10
-#| fig.cap: Best projections found by the JS opitmiser in each simulation, sorted by the index value.
+#| fig.cap: Best projections found by the JSO in each of the 50 simulations, sorted by index values.
 The projection pursuit problem is to detect the 6D sine-wave structure using the spline index.
-Most simulations have detected a clear sine-wave pattern, while the last six show a straight line with clustering at the ends..
+Most simulations have detected a clear sine-wave pattern, while the last six show a straight line with clustering at the ends.
 In this simulation, the success rate is 0.88 (44/50).
@@
 projection pursuit problems. Two examples are provided in this section
-to demonstrate its performance. The first example compares its
-performance with the search-better optimiser and explores how it behaves
+to demonstrate its performance. The first example compares the 
+performance of JSO with an exist optimiser, search better, and explores how JSO behaves
 with different combinations of hyper-parameters. The second example
 studies the effect of optimisation properties defined in @sec-theory,
-along with jellyfish hyper-parameters, on the outcome of the
+along with JSO hyper-parameters, on the outcome (success rate) of the
 optimisation.
@@
 
-Fifty simulations are conducted with both JSO and the search-better , in
-four dimensions ($d = 6, 8, 10, 12$). The JSO uses 100 jellyfishes and
+Fifty simulations are conducted with both JSO and the search better optimal, in
+four dimensions ($d = 6, 8, 10, 12$). JSO uses 100 jellyfishes and
 100 maximum number of tries, while the search-better optimiser allows a
@@
 In the 6-dimensional data scenario, JSO consistently identifies a clear
-pipe shape. The search-better optimiser also finds the pipe shape but
+pipe shape. The search better optimiser also finds the pipe shape but
 with a wide rim, suggesting a further polish search may be required.
@@
 due to random sampling, but it still presents a pipe shape in over 50%
-of cases. When compared to the search-better optimiser, JSO reaches
+of cases. When compared to the search better optimiser, JSO reaches
 higher index values and clearer pipe shapes across all quantiles.
@@
 In the second experiment, fifty simulations are conducted at each
-hyper-parameter combination to analyse its effects on the success rate
-of JSO . The hyper-parameters tested include: 1) 20, 50, and 100
+hyper-parameter combination to analyse its effects on the JSO success
+rate. The hyper-parameters tested include: 1) 20, 50, and 100
 jellyfishes, and 2) 50 and 100 maximum number of tries. In each
@@
 100 jellyfishes and 100 maximum number of tries is necessary for at
-least half of the jellyfishes to find the optimal projection.
+least half of the jellyfishes to find the projection close to the optimal.
 
@@
 The optimisation properties, including smoothness, squintability, and
-speed, offer numerical metrics to characterise the complexity of a
-projection pursuit optimisation problem. This example investigates how
-these metrics, along with the jellyfish search hyper-parameters (the
-number of jellyfishes and the maximum number of tries), affect the
-success of JSO. Simulations are conducted to obtain the success rate
-across various projection pursuit problems, characterised by
-shape-to-find, data dimension, and index function used, as well as
+speed, offer numerical metrics to characterise the complexity of
+projection pursuit optimisation problems. This example investigates how
+these metrics, along with the hyper-parameters affect the JSO
+success rate. Simulations are conducted to obtain the success rate
+across a wider array of projection pursuit problems, characterised by
+shape-to-find, data dimension, and the index function, as well as
 different hyper-parameter combinations. Smoothness and squintability are
-calculated for each projection pursuit problem as outlined in
+calculated for each projection pursuit problem as described in
 @sec-theory. A generalised linear model is used to construct the
-relationship between the success rate, jellyfish hyper-parameters and
+relationship between the success rate, JSO hyper-parameters and
 optimisation properties.
@@
 In addition to the pipe-finding problem, a new shape, sine wave, is
-investigated in 6D and 8D spaces with six indices considered:
+investigated in 6D and 8D spaces with six indexes considered:
 `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`. Combining
-with two jellyfish hyper-parameters, a total of `r nrow(sim_df)` cases
+with two JSO hyper-parameters, a total of `r nrow(sim_df)` cases
 is produced, comprising of `r nrow(filter(sim_df, index == "holes"))`
@@
 values and projection distance (to the optimal basis) are calculated for
-each random basis before fitting them into a Gaussian process model. For
+each random basis before fitting into a Gaussian process model to obtain 
+the smoothness measure. For
 squintability, fifty random bases are sampled and interpolated to the
-optimal basis with a step size of 0.005. For these interpolated bases,
-index values and projection distances to the optimal basis are
-calculated and averaged with a bin width of 0.005. The squintability
+optimal basis with a step size of 0.005. Index values and projection 
+distances are calculated for these interpolated bases and the index values 
+are averaged with a bin width of 0.005. The squintability
 measure is then calculated from fitting a 4-parameter scaled logistic
@@
 "smooth" is used as the smoothness measure and the column "squint" is
-calculated as equation
-(@eq-squintability-parametric) as squintability. The low
+calculated as equation @eq-squintability-parametric as squintability. The low
 squint value of MIC and TIC index is due to its convex shape of the
@@
 
-@tbl-mod-data combines all the variables calculated above (the JSO
-success rate, JSO hyper-parameters, and optimisation properties) into
-one table. Three preprocessing steps are applied to the data: 1) scaling
+@tbl-mod-data combines all the variables calculated above (JSO
+success rate, hyper-parameters, and optimisation properties) into
+one table. Three pre-processing steps are applied to the data: 1) scaling
 the JSO hyper-parameters by a factor of 10 for interpretation, 2)
@@
 function is used to fit the data and @tbl-mod-output presents the model
-outputs. The model suggests that the JSO success rate is positively
-associated with the two JSO hyper-parameters, as well as with the
+outputs. The model suggests that JSO success rate is positively
+associated with the two hyper-parameters, as well as with the
 optimisation properties -- smoothness and squintability. Specifically,
-using 10 more jellyfish and 10 more tries will increase the odd ratio of
+using 10 more jellyfishes and 10 more tries will increase the odd ratio of
 success by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`%
@@
 their importance relative to JSO hyper-parameters in the optimisation
-success. Users can increase the number of jellyfish and the number of
+success. Users should consider increasing the number of jellyfish and the number of
 tries to improve the success rate, until JSO exceeds a 30-second

### 2024-06-24 optim.tex
edit section 4 and 5

@@
 This study explores the potential for a Jellyfish Search Optimiser (JSO)
-\citep{chou_novel_2021} -\citep{chou_recent_2022} to be integrated with
-projection pursuit guided tour. JSO, inspired by the search behaviour of
-jellyfish in the ocean, is a swarm-based metaheuristic designed to solve
-global optimisation problems. Compared to traditional methods, JSO
-demonstrated stronger search ability and faster convergence, and
-requires fewer tuning parameters. These practical advantages make JSO as
-a promising candidate for enhancing PP optimisation. As such, the study
-aims to explore JSO's potential in enhancing projection pursuit and
-examines the characteristics and advantages of such implementation.
+to be integrated with projection pursuit guided tour. JSO
+\citep{chou_novel_2021} -\citep{rajwar_exhaustive_2023}, inspired by the
+search behaviour of jellyfish in the ocean, is a swarm-based
+metaheuristic designed to solve global optimisation problems. Compared
+to traditional methods, JSO demonstrated stronger search ability and
+faster convergence, and requires fewer tuning parameters. These
+practical advantages make JSO as a promising candidate for enhancing PP
+optimisation. As such, the study aims to explore JSO's potential in
+enhancing projection pursuit and examines the characteristics and
+advantages of such implementation.
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, breakable, bottomrule=.15mm, toprule=.15mm, opacityback=0, titlerule=0mm, coltitle=black, bottomtitle=1mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, rightrule=.15mm, left=2mm, arc=.35mm, colback=white, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white]
+\begin{tcolorbox}[enhanced jigsaw, opacityback=0, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, coltitle=black, titlerule=0mm, breakable, arc=.35mm, toprule=.15mm, left=2mm, bottomrule=.15mm, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, colback=white, toptitle=1mm, rightrule=.15mm]
 
@@
 
-The data structure proposed in \citet{RJ-2021-105} is implemented for JS
-optimisers and additional information such as index function used, data
-dimension, and jellyfish hyper-parameters can be included as additional
-columns. An example data collected from finding the sine-wave structure
-in 6D data (\texttt{d\ =\ 6}) using a spline index
-function(\texttt{splines2d}) is printed below:
+The data structure proposed in \citet{RJ-2021-105} is implemented for
+JSO. An example data collected from finding the sine-wave structure in
+6D data using a spline index function (\texttt{splines2d}) is printed
+below:
 
@@
 
-The jellyfish hyper-parameters used are 50 jellyfishes
-(\texttt{n\_jellies}) and a maximum of 50 tries (\texttt{max\_tries}).
-The full data contains 50 simulations, recorded by the variable
-\texttt{sim}, with different seeds (\texttt{seed}). Recorded in columns
-\texttt{basis}, \texttt{index\_val}, \texttt{tries}, \texttt{loop} and
-\texttt{time}, are the sampled projection basis, calculated index value,
-iteration ID, jellyfish ID, and time taken to find the optimum,
-respectively. The basis column records every visited basis before
-comparing with the current (last) basis. If a visited basis has a
-smaller index value than the current (last) one, JSO will keep the
-current basis for the next iteration in the search but still records the
-visited inferior basis. This data can be used to derive numerical
-summaries and plots:
+The JSO hyper-parameters used are 50 jellyfishes (\texttt{n\_jellies})
+and a maximum of 50 tries (\texttt{max\_tries}). The full data contains
+50 simulations, indexed by the variable \texttt{sim}, with the
+associated \texttt{seed}s. Recorded in columns \texttt{basis},
+\texttt{index\_val}, \texttt{tries}, \texttt{loop} and \texttt{time},
+are the sampled projection basis, index value, iteration ID, jellyfish
+ID, and computational time respectively. The basis column records every
+visited basis before comparing with the current (last) basis. If a
+visited basis has a smaller index value than the current (last) one, JSO
+will keep the current basis for the next iteration in the search but
+still records the visited inferior basis. This data can be used to
+derive numerical summaries and plots:
 
@@
   \textbf{Success rate}: The success rate is defined as the proportion
-  of simulations that achieve a final index value within 0.05 of the
+  of simulations that achieves a final index value within 0.05 of the
   best index value found among all 50 simulations. In the dataset above,
@@
   \textbf{Projection distance}: The projection distance between
-  projection bases and the theoretical best basis (or the empirical best
-  basis if the theoretical best is unknown) is calculated as the
-  Frobenius norm, which measures how close a visited basis is to the
-  best basis. Plotting this projection distance against the index value
-  for each basis visited reveals how the index value changes as the
-  basis approaches the optimal basis, helping to identify the
-  squintability of the index function.
+  projection bases and the optimal best basis (or the empirical best
+  basis if the optimal is unknown) measures how close a visited basis is
+  to the best basis and it is calculated as the Frobenius norm in
+  Equation
+
+  \begin{enumerate}
+  \def\labelenumi{\arabic{enumi}.}
+  \tightlist
+  \item
+    Plotting projection distances against index values for the bases
+    visited reveals how the index value changes as the basis approaches
+    the optimal basis, helping to calculate the squintability measure.
     Figure~\ref{fig-index-value-proj-dist} shows the index value against
-  the projection distance for the given example and its binned version.
-  As the projection distance decreases, the index value initially
-  increases with an increasing slope and then levels off. The index
-  value shows greater variation at medium projection distances compared
-  to small and large distances.
+    the projection distance for the given dataset and its binned
+    version. As the projection distance decreases, the index value
+    initially increases with an increasing slope and then levels off.
+    The index value shows greater variation at medium projection
+    distances compared to small and large distances.
+  \end{enumerate}
 \end{itemize}
@@
 
-\caption{\label{fig-projection}Best projections found by the JS
-opitmiser in each simulation, sorted by the index value. The projection
-pursuit problem is to detect the 6D sine-wave structure using the spline
-index. Most simulations have detected a clear sine-wave pattern, while
-the last six show a straight line with clustering at the ends.. In this
+\caption{\label{fig-projection}Best projections found by the JSO in each
+of the 50 simulations, sorted by index values. The projection pursuit
+problem is to detect the 6D sine-wave structure using the spline index.
+Most simulations have detected a clear sine-wave pattern, while the last
+six show a straight line with clustering at the ends. In this
 simulation, the success rate is 0.88 (44/50).}
@@
 projection pursuit problems. Two examples are provided in this section
-to demonstrate its performance. The first example compares its
-performance with the search-better optimiser and explores how it behaves
-with different combinations of hyper-parameters. The second example
-studies the effect of optimisation properties defined in
-Section~\ref{sec-theory}, along with jellyfish hyper-parameters, on the
-outcome of the optimisation.
+to demonstrate its performance. The first example compares the
+performance of JSO with an exist optimiser, search better, and explores
+how JSO behaves with different combinations of hyper-parameters. The
+second example studies the effect of optimisation properties defined in
+Section~\ref{sec-theory}, along with JSO hyper-parameters, on the
+outcome (success rate) of the optimisation.
 
@@
 
-Fifty simulations are conducted with both JSO and the search-better , in
-four dimensions (\(d = 6, 8, 10, 12\)). The JSO uses 100 jellyfishes and
-100 maximum number of tries, while the search-better optimiser allows a
-maximum of 1000 samples at each iteration before the algorithm
-terminates. Figure~\ref{fig-proj} presents the final projections found
-by the two optimisers, broken down by 10th quantile, faceted by the data
-dimension. In the 6-dimensional data scenario, JSO consistently
-identifies a clear pipe shape. The search-better optimiser also finds
-the pipe shape but with a wide rim, suggesting a further polish search
-may be required. With increasing dimensions, JSO may not always identify
-the pipe shape due to random sampling, but it still presents a pipe
-shape in over 50\% of cases. When compared to the search-better
-optimiser, JSO reaches higher index values and clearer pipe shapes
-across all quantiles.
+Fifty simulations are conducted with both JSO and the search better
+optimal, in four dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100
+jellyfishes and 100 maximum number of tries, while the search-better
+optimiser allows a maximum of 1000 samples at each iteration before the
+algorithm terminates. Figure~\ref{fig-proj} presents the final
+projections found by the two optimisers, broken down by 10th quantile,
+faceted by the data dimension. In the 6-dimensional data scenario, JSO
+consistently identifies a clear pipe shape. The search better optimiser
+also finds the pipe shape but with a wide rim, suggesting a further
+polish search may be required. With increasing dimensions, JSO may not
+always identify the pipe shape due to random sampling, but it still
+presents a pipe shape in over 50\% of cases. When compared to the search
+better optimiser, JSO reaches higher index values and clearer pipe
+shapes across all quantiles.
 
 In the second experiment, fifty simulations are conducted at each
-hyper-parameter combination to analyse its effects on the success rate
-of JSO . The hyper-parameters tested include: 1) 20, 50, and 100
+hyper-parameter combination to analyse its effects on the JSO success
+rate. The hyper-parameters tested include: 1) 20, 50, and 100
 jellyfishes, and 2) 50 and 100 maximum number of tries. In each
@@
 of tries is necessary for at least half of the jellyfishes to find the
-optimal projection.
+projection close to the optimal.
 
@@
 The optimisation properties, including smoothness, squintability, and
-speed, offer numerical metrics to characterise the complexity of a
-projection pursuit optimisation problem. This example investigates how
-these metrics, along with the jellyfish search hyper-parameters (the
-number of jellyfishes and the maximum number of tries), affect the
-success of JSO. Simulations are conducted to obtain the success rate
-across various projection pursuit problems, characterised by
-shape-to-find, data dimension, and index function used, as well as
+speed, offer numerical metrics to characterise the complexity of
+projection pursuit optimisation problems. This example investigates how
+these metrics, along with the hyper-parameters affect the JSO success
+rate. Simulations are conducted to obtain the success rate across a
+wider array of projection pursuit problems, characterised by
+shape-to-find, data dimension, and the index function, as well as
 different hyper-parameter combinations. Smoothness and squintability are
-calculated for each projection pursuit problem as outlined in
+calculated for each projection pursuit problem as described in
 Section~\ref{sec-theory}. A generalised linear model is used to
-construct the relationship between the success rate, jellyfish
+construct the relationship between the success rate, JSO
 hyper-parameters and optimisation properties.
@@
 In addition to the pipe-finding problem, a new shape, sine wave, is
-investigated in 6D and 8D spaces with six indices considered:
+investigated in 6D and 8D spaces with six indexes considered:
 \texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
-\texttt{spline}, and \texttt{stringy}. Combining with two jellyfish
+\texttt{spline}, and \texttt{stringy}. Combining with two JSO
 hyper-parameters, a total of 52 cases is produced, comprising of 24
@@
 random bases are simulated. Index values and projection distance (to the
-optimal basis) are calculated for each random basis before fitting them
-into a Gaussian process model. For squintability, fifty random bases are
-sampled and interpolated to the optimal basis with a step size of 0.005.
-For these interpolated bases, index values and projection distances to
-the optimal basis are calculated and averaged with a bin width of 0.005.
-The squintability measure is then calculated from fitting a 4-parameter
-scaled logistic function of index value against projection distance,
-estimated by non-linear least squares.
+optimal basis) are calculated for each random basis before fitting into
+a Gaussian process model to obtain the smoothness measure. For
+squintability, fifty random bases are sampled and interpolated to the
+optimal basis with a step size of 0.005. Index values and projection
+distances are calculated for these interpolated bases and the index
+values are averaged with a bin width of 0.005. The squintability measure
+is then calculated from fitting a 4-parameter scaled logistic function
+of index value against projection distance, estimated by non-linear
+least squares.
 
@@
 and the column ``squint'' is calculated as equation
-(\ref{eq-squintability-parametric}) as squintability. The low squint
-value of MIC and TIC index is due to its convex shape of the index value
+\ref{eq-squintability-parametric} as squintability. The low squint value
+of MIC and TIC index is due to its convex shape of the index value
 against the projection distance, as shown in
@@
 Table~\ref{tbl-mod-data} combines all the variables calculated above
-(the JSO success rate, JSO hyper-parameters, and optimisation
-properties) into one table. Three preprocessing steps are applied to the
-data: 1) scaling the JSO hyper-parameters by a factor of 10 for
-interpretation, 2) creating a new binary variable \texttt{long\_time} to
-indicate cases with an average run time over 30 seconds, and 3) encoding
-the success rate for the stringy case as 0, because upon visual
-inspection of the final projection, none of the 50 simulations
-identified the sine-wave shape. A generalised linear model with a
-binomial family and a logit link function is used to fit the data and
-Table~\ref{tbl-mod-output} presents the model outputs. The model
-suggests that the JSO success rate is positively associated with the two
-JSO hyper-parameters, as well as with the optimisation properties --
-smoothness and squintability. Specifically, using 10 more jellyfish and
-10 more tries will increase the odd ratio of success by 24.11\% and
-11.93\%, respectively. However, being flagged with long runtime and an
-increase of data dimension will reduce the success rate by 41.72\% and
-53.36\%, respectively. The variable \texttt{squintability} and
-\texttt{dimension} are significant, suggesting their importance relative
-to JSO hyper-parameters in the optimisation success. Users can increase
-the number of jellyfish and the number of tries to improve the success
-rate, until JSO exceeds a 30-second runtime.
+(JSO success rate, hyper-parameters, and optimisation properties) into
+one table. Three pre-processing steps are applied to the data: 1)
+scaling the JSO hyper-parameters by a factor of 10 for interpretation,
+2) creating a new binary variable \texttt{long\_time} to indicate cases
+with an average run time over 30 seconds, and 3) encoding the success
+rate for the stringy case as 0, because upon visual inspection of the
+final projection, none of the 50 simulations identified the sine-wave
+shape. A generalised linear model with a binomial family and a logit
+link function is used to fit the data and Table~\ref{tbl-mod-output}
+presents the model outputs. The model suggests that JSO success rate is
+positively associated with the two hyper-parameters, as well as with the
+optimisation properties -- smoothness and squintability. Specifically,
+using 10 more jellyfishes and 10 more tries will increase the odd ratio
+of success by 24.11\% and 11.93\%, respectively. However, being flagged
+with long runtime and an increase of data dimension will reduce the
+success rate by 41.72\% and 53.36\%, respectively. The variable
+\texttt{squintability} and \texttt{dimension} are significant,
+suggesting their importance relative to JSO hyper-parameters in the
+optimisation success. Users should consider increasing the number of
+jellyfish and the number of tries to improve the success rate, until JSO
+exceeds a 30-second runtime.
 

### 2024-06-26 optim.qmd
small typos and abstract

@@
 abstract: |
-  Projection pursuit (PP) is a dimension reduction technique that identifies low-dimensional projections in high-dimensional data by optimising a criteria function known as the PP index.
-The optimisation in PP can be non-smooth and requires identifying optima with a small "squint angle", detectable only from close proximity.
-To address these challenges, this study investigates the performance of a recent swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
+  The projection pursuit (PP) guided tour interactively optimizes a criteria function known as the PP index, to explore high-dimensional data by revealing interesting projections.
+The optimisation in PP can be non-trivial, involving non-smooth functions and optima with small a "squint angle", detectable only from close proximity.
+To address these challenges, this study investigates the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
 The performance of JSO is evaluated across various hyper-parameter settings and compared with existing optimisers.
@@
 These two metrics are evaluated with JSO hyper-parameters to determine their effect on JSO success rate.
-The JSO algorithm has been implemented in the `tourr` package, while calculations for smoothness and squintability are available in the `ferrn` package.
+The JSO algorithm has been implemented in the `tourr` package, while functions to calculate  smoothness and squintability of indexes are available in the `ferrn` package.
 keywords: 
@@
 
-This study explores the potential for a Jellyfish Search Optimiser 
-(JSO) to be integrated with projection 
-pursuit guided tour. JSO [@chou_novel_2021] -[@rajwar_exhaustive_2023], 
-inspired by the search behaviour of jellyfish 
-in the ocean, is a swarm-based metaheuristic designed to solve global
-optimisation problems. Compared to traditional methods, JSO 
-demonstrated stronger search ability and faster convergence, and 
-requires fewer tuning parameters. These
-practical advantages make JSO as a promising candidate for enhancing
-PP optimisation. As such, the study aims to explore JSO's potential
-in enhancing projection pursuit and examines the characteristics
+This study explores the potential for a Jellyfish Search Optimiser (JSO) to be integrated with the projection pursuit guided tour.
+JSO [@chou_novel_2021] -[@rajwar_exhaustive_2023], inspired by the search behaviour of jellyfish in the ocean, is a swarm-based metaheuristic designed to solve global optimisation problems.
+Compared to traditional methods, JSO demonstrated stronger search ability and faster convergence, and requires fewer tuning parameters.
+These practical advantages make JSO as a promising candidate for enhancing PP optimisation.
+As such, the study aims to explore JSO's potential in enhancing projection pursuit and examines the characteristics
 and advantages of such implementation.
 
-The primary goal of this study is to investigate the performance of JSO
-in PP optimisation. A series of simulation
-experiments using well-recognised datasets and various PP indexes are conducted
-to yield insights regarding the behaviour of JSO under different
-settings and its sensitivity to hyper-parameter.
-Second, to observe the performance of JSO with different types of PP
-indexes, metrics are introduced to capture specific
-properties of the index including squintability and smoothness as
-introduced in @laa_using_2020. To the best of the authors' knowledge,
-this is the first attempt to quantitatively measure squintability and
-smoothness within this context. Finally, the relationship between the JSO
-performance, hyper-parameters tuning and various properties of PP
+The primary goal of this study is to investigate the performance of JSO in PP optimisation.
+A series of simulation experiments using well-recognised datasets and various PP indexes are conducted to yield insights regarding the behaviour of JSO under different settings and its sensitivity to hyper-parameter.
+Second, to observe the performance of JSO with different types of PP indexes, metrics are introduced to capture specific
+properties of the index including squintability and smoothness as introduced in @laa_using_2020.
+To the best of the authors' knowledge, this is the first attempt to quantitatively measure squintability and
+smoothness within this context. Finally, the relationship between the JSO performance, hyper-parameters tuning and various properties of PP
 indexes is analysed to provide helpful guidance for practitioners using projection pursuit.
 
-The rest of this paper is structured as follows. @sec-background
-introduces the background of PP guided tour and reviews existing
-optimisers and index functions in the literature. @sec-theory details JSO and
-introduces metrics that measure different properties of PP indexes.
-@sec-vis outlines the data structures, numerical summaries, and visualisations used in the simulation studies conducted in @sec-app to assess JSO's perforamnce.
-@sec-app presents two sets of simulation
-experiments, comparing JSO with another optimiser, search better, and
-studies the impact of different PP index properties on the optimisation
-performance. @sec-conclusion summarises the work and provides
-suggestions for future directions.
+The rest of this paper is structured as follows.
+@sec-background introduces the background of PP guided tour and reviews existing optimisers and index functions in the literature.
+@sec-theory details JSO and introduces metrics that measure different properties of PP indexes.
+@sec-vis outlines the data structures, numerical summaries, and visualisations used in the simulation studies conducted in @sec-app to assess JSO's performance.
+@sec-app presents two sets of simulation experiments, comparing JSO with another optimiser, search better, and studies the impact of different PP index properties on the optimisation
+performance. @sec-conclusion summarises the work and provides suggestions for future directions.
 
-# Projection pursuit, index functions and optimisation \[Di and Sherry\] {#sec-background}
+# Projection pursuit, tours, index functions and optimisation \[Di and Sherry\] {#sec-background}
 

### 2024-06-26 optim.tex
small typos and abstract

@@
 \begin{abstract}
-Projection pursuit (PP) is a dimension reduction technique that
-identifies low-dimensional projections in high-dimensional data by
-optimising a criteria function known as the PP index. The optimisation
-in PP can be non-smooth and requires identifying optima with a small
+The projection pursuit (PP) guided tour interactively optimizes a
+criteria function known as the PP index, to explore high-dimensional
+data by revealing interesting projections. The optimisation in PP can be
+non-trivial, involving non-smooth functions and optima with small a
 ``squint angle'', detectable only from close proximity. To address these
-challenges, this study investigates the performance of a recent
-swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising
-PP indexes. The performance of JSO is evaluated across various
-hyper-parameter settings and compared with existing optimisers.
+challenges, this study investigates the performance of a recently
+introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for
+optimising PP indexes. The performance of JSO is evaluated across
+various hyper-parameter settings and compared with existing optimisers.
 Additionally, this work proposes novel methods to quantify two
@@
 effect on JSO success rate. The JSO algorithm has been implemented in
-the \texttt{tourr} package, while calculations for smoothness and
-squintability are available in the \texttt{ferrn} package.
+the \texttt{tourr} package, while functions to calculate smoothness and
+squintability of indexes are available in the \texttt{ferrn} package.
 \end{abstract}
@@
 This study explores the potential for a Jellyfish Search Optimiser (JSO)
-to be integrated with projection pursuit guided tour. JSO
+to be integrated with the projection pursuit guided tour. JSO
 \citep{chou_novel_2021} -\citep{rajwar_exhaustive_2023}, inspired by the
@@
 simulation studies conducted in Section~\ref{sec-app} to assess JSO's
-perforamnce. Section~\ref{sec-app} presents two sets of simulation
+performance. Section~\ref{sec-app} presents two sets of simulation
 experiments, comparing JSO with another optimiser, search better, and
@@
 
-\section{Projection pursuit, index functions and optimisation {[}Di and
-Sherry{]}}\label{sec-background}
+\section{Projection pursuit, tours, index functions and optimisation
+{[}Di and Sherry{]}}\label{sec-background}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacityback=0, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, coltitle=black, titlerule=0mm, breakable, arc=.35mm, toprule=.15mm, left=2mm, bottomrule=.15mm, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, colback=white, toptitle=1mm, rightrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, left=2mm, rightrule=.15mm, colframe=quarto-callout-note-color-frame, breakable, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, coltitle=black, titlerule=0mm, toptitle=1mm, arc=.35mm, bottomrule=.15mm, opacitybacktitle=0.6, leftrule=.75mm, colback=white]
 

### 2024-06-26 optim.qmd
introduction revised

@@
 abstract: |
-  The projection pursuit (PP) guided tour interactively optimizes a criteria function known as the PP index, to explore high-dimensional data by revealing interesting projections.
+  The projection pursuit (PP) guided tour interactively optimises a criteria function known as the PP index, to explore high-dimensional data by revealing interesting projections.
 The optimisation in PP can be non-trivial, involving non-smooth functions and optima with small a "squint angle", detectable only from close proximity.
 To address these challenges, this study investigates the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
-The performance of JSO is evaluated across various hyper-parameter settings and compared with existing optimisers.
+The performance of JSO for visualising data is evaluated across various hyper-parameter settings and compared with existing optimisers.
 Additionally, this work proposes novel methods to quantify two properties of the PP index -- smoothness and squintability -- that capture the complexities inherent in PP optimisation problems.
@@
   - projection pursuit
-  - Jellyfish Search Optimiser (JSO)
-  - Optimisation
+  - jellyfish search optimiser (JSO)
+  - optimisation
+  - grand tour
+  - high-dimensional data
+  - exploratory data analysis
 date: last-modified
@@
 
-Projection Pursuit (PP) (@kr69, @FT74) is a dimension reduction technique 
-aimed at identifying informative linear projections of data. The method involves 
-optimising a objective function known as PP index
-(@hall1989polynomial, @cook1993projection, @lee2010projection,
-@Loperfido2018, @Loperfido2020), which defines the criteria for what constitutes  interesting or informative projections. Given high-dimensional data
-$X \in \mathbf{R}^{n\times p}$ and an index function $f(\cdot)$, PP
-finds the orthonormal projection matrix $A \in \mathbf{R}^{p \times d}$ that 
-maximises the index value of the projection, $Y = XA$:
+Projection Pursuit (PP) (@kr69, @FT74) is a dimension reduction technique aimed at identifying informative linear projections of data.
+The method involves optimising a objective function known as the PP index
+(e.g.
+@hall1989polynomial, @cook1993projection, @lee2010projection, @Loperfido2018, @Loperfido2020), which defines the criteria for what constitutes  interesting or informative projections.
+Given high-dimensional data $X \in \mathbf{R}^{n\times p}$ and an index function $f(\cdot)$, PP finds the orthonormal basis $A \in \mathbf{R}^{p \times d}$ that maximises the index value of the projection, $Y = XA$:
 
@@
 
-The index functions can be non-linear and non-convex, hence an 
-effective and efficient optimisation procedure is essential to 
-explore the data landscape and achieving a globally optimal viewpoint 
-of the data.
+The index functions can be non-linear and non-convex, hence an effective and efficient optimisation procedure is essential to explore the data landscape and achieving a globally optimal viewpoint of the data.
 
-Significant efforts have been devoted to enhancing the functionality of
-PP and facilitating a better experience in the visualisation process.
-@cook1995grand introduces the PP guided tour, which enabled interactive 
-visualisation of the optimisation in order to visually explore the high-dimensional data. It is implemented in the R [@R] package `tourr` [@tourr].
-@RJ-2021-105 highlighted potential problems of the optimisation routine
-implemented. While improving the quality of the optimisation solutions
-in the tour is essential, it is also important to be able to watch the
-projected data as the optimisation progresses. As such, it is pertinent
-to integrate the guided tour with a global optimisation algorithm that
-is efficient in finding the global optimal and enables viewing of the
-projected data during the exploration process.
+Significant efforts have been devoted to enhancing the functionality of PP, improving optimisation (e.g.
+@posse1995, @marie-sainte2010, @grochowski2011) and extending it's use to visually explore high-dimensional data.
+@cook1995grand introduced the PP guided tour, which monitors the optimisation visually to see the data projections leading in and out of the optima.
+An implementation is available in the R [@R] package `tourr` [@tourr].
+@RJ-2021-105 illustrated how to diagnose optimisation processes, particularly focusing on the guided tour, and revealed a need to for improved optimisation.
+While improving the quality of the optimisation solutions in the tour is essential, it is also important to be able to view the data projections as the optimisation progresses.
+Integrating the guided tour with a global optimisation algorithm that is efficient in finding the global optimal and enables viewing of the projected data during the exploration process is a goal.
 
-This study explores the potential for a Jellyfish Search Optimiser (JSO) to be integrated with the projection pursuit guided tour.
+Here, the potential for a Jellyfish Search Optimiser (JSO) to be integrated with the projection pursuit guided tour is explored.
 JSO [@chou_novel_2021] -[@rajwar_exhaustive_2023], inspired by the search behaviour of jellyfish in the ocean, is a swarm-based metaheuristic designed to solve global optimisation problems.
 Compared to traditional methods, JSO demonstrated stronger search ability and faster convergence, and requires fewer tuning parameters.
-These practical advantages make JSO as a promising candidate for enhancing PP optimisation.
-As such, the study aims to explore JSO's potential in enhancing projection pursuit and examines the characteristics
-and advantages of such implementation.
+These practical advantages make JSO a promising candidate for enhancing PP optimisation.
+<!--The study reported here explores JSO's potential in enhancing projection pursuit and examines the characteristics
+and advantages of such implementation.-->
 
-The primary goal of this study is to investigate the performance of JSO in PP optimisation.
-A series of simulation experiments using well-recognised datasets and various PP indexes are conducted to yield insights regarding the behaviour of JSO under different settings and its sensitivity to hyper-parameter.
-Second, to observe the performance of JSO with different types of PP indexes, metrics are introduced to capture specific
-properties of the index including squintability and smoothness as introduced in @laa_using_2020.
-To the best of the authors' knowledge, this is the first attempt to quantitatively measure squintability and
-smoothness within this context. Finally, the relationship between the JSO performance, hyper-parameters tuning and various properties of PP
-indexes is analysed to provide helpful guidance for practitioners using projection pursuit.
+The primary goal of the study reported here is to investigate the performance of JSO in PP optimisation for the guided tour.
+It is of interest to assess how quickly and closely the optimiser reaches a gobal optima, for various PP indexes that may have differing complexities.
+To observe the performance of JSO with different types of PP indexes, metrics are introduced to capture specific properties of the index including squintability and smoothness as introduced in @laa_using_2020.
+Metrics for squintability and smoothness are precisely defined here, something that is new for the field.
+A series of simulation experiments using specially structured datasets and various PP indexes are conducted to assess the behaviour of JSO under different settings and its sensitivity to hyper-parameter choices.
+The relationship between the JSO performance, hyper-parameter tuning and properties of PP
+indexes is analysed to provide guidance for practitioners using projection pursuit, and developers of new indexes.
 
-The rest of this paper is structured as follows.
+The paper is structured as follows.
 @sec-background introduces the background of PP guided tour and reviews existing optimisers and index functions in the literature.

### 2024-06-26 optim.tex
introduction revised

@@
   pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
-  pdfkeywords={projection pursuit, Jellyfish Search Optimiser
-(JSO), Optimisation},
+  pdfkeywords={projection pursuit, jellyfish search optimiser
+(JSO), optimisation, grand tour, high-dimensional data, exploratory data
+analysis},
   colorlinks=true,
@@
 \begin{abstract}
-The projection pursuit (PP) guided tour interactively optimizes a
+The projection pursuit (PP) guided tour interactively optimises a
 criteria function known as the PP index, to explore high-dimensional
@@
 introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for
-optimising PP indexes. The performance of JSO is evaluated across
-various hyper-parameter settings and compared with existing optimisers.
-Additionally, this work proposes novel methods to quantify two
-properties of the PP index -- smoothness and squintability -- that
-capture the complexities inherent in PP optimisation problems. These two
-metrics are evaluated with JSO hyper-parameters to determine their
-effect on JSO success rate. The JSO algorithm has been implemented in
-the \texttt{tourr} package, while functions to calculate smoothness and
-squintability of indexes are available in the \texttt{ferrn} package.
+optimising PP indexes. The performance of JSO for visualising data is
+evaluated across various hyper-parameter settings and compared with
+existing optimisers. Additionally, this work proposes novel methods to
+quantify two properties of the PP index -- smoothness and squintability
+-- that capture the complexities inherent in PP optimisation problems.
+These two metrics are evaluated with JSO hyper-parameters to determine
+their effect on JSO success rate. The JSO algorithm has been implemented
+in the \texttt{tourr} package, while functions to calculate smoothness
+and squintability of indexes are available in the \texttt{ferrn}
+package.
 \end{abstract}
@@
 \begin{keyword}
-    projection pursuit \sep Jellyfish Search Optimiser (JSO) \sep 
-    Optimisation
+    projection pursuit \sep jellyfish search optimiser
+(JSO) \sep optimisation \sep grand tour \sep high-dimensional data \sep 
+    exploratory data analysis
 \end{keyword}
@@
 reduction technique aimed at identifying informative linear projections
-of data. The method involves optimising a objective function known as PP
-index (\citet{hall1989polynomial}, \citet{cook1993projection},
-\citet{lee2010projection}, \citet{Loperfido2018},
-\citet{Loperfido2020}), which defines the criteria for what constitutes
-interesting or informative projections. Given high-dimensional data
-\(X \in \mathbf{R}^{n\times p}\) and an index function \(f(\cdot)\), PP
-finds the orthonormal projection matrix
+of data. The method involves optimising a objective function known as
+the PP index (e.g. \citet{hall1989polynomial},
+\citet{cook1993projection}, \citet{lee2010projection},
+\citet{Loperfido2018}, \citet{Loperfido2020}), which defines the
+criteria for what constitutes interesting or informative projections.
+Given high-dimensional data \(X \in \mathbf{R}^{n\times p}\) and an
+index function \(f(\cdot)\), PP finds the orthonormal basis
 \(A \in \mathbf{R}^{p \times d}\) that maximises the index value of the
@@
 Significant efforts have been devoted to enhancing the functionality of
-PP and facilitating a better experience in the visualisation process.
-\citet{cook1995grand} introduces the PP guided tour, which enabled
-interactive visualisation of the optimisation in order to visually
-explore the high-dimensional data. It is implemented in the R \citep{R}
-package \texttt{tourr} \citep{tourr}. \citet{RJ-2021-105} highlighted
-potential problems of the optimisation routine implemented. While
-improving the quality of the optimisation solutions in the tour is
-essential, it is also important to be able to watch the projected data
-as the optimisation progresses. As such, it is pertinent to integrate
-the guided tour with a global optimisation algorithm that is efficient
-in finding the global optimal and enables viewing of the projected data
-during the exploration process.
-
-This study explores the potential for a Jellyfish Search Optimiser (JSO)
-to be integrated with the projection pursuit guided tour. JSO
+PP, improving optimisation (e.g. \citet{posse1995},
+\citet{marie-sainte2010}, \citet{grochowski2011}) and extending it's use
+to visually explore high-dimensional data. \citet{cook1995grand}
+introduced the PP guided tour, which monitors the optimisation visually
+to see the data projections leading in and out of the optima. An
+implementation is available in the R \citep{R} package \texttt{tourr}
+\citep{tourr}. \citet{RJ-2021-105} illustrated how to diagnose
+optimisation processes, particularly focusing on the guided tour, and
+revealed a need to for improved optimisation. While improving the
+quality of the optimisation solutions in the tour is essential, it is
+also important to be able to view the data projections as the
+optimisation progresses. Integrating the guided tour with a global
+optimisation algorithm that is efficient in finding the global optimal
+and enables viewing of the projected data during the exploration process
+is a goal.
+
+Here, the potential for a Jellyfish Search Optimiser (JSO) to be
+integrated with the projection pursuit guided tour is explored. JSO
 \citep{chou_novel_2021} -\citep{rajwar_exhaustive_2023}, inspired by the
@@
 faster convergence, and requires fewer tuning parameters. These
-practical advantages make JSO as a promising candidate for enhancing PP
-optimisation. As such, the study aims to explore JSO's potential in
-enhancing projection pursuit and examines the characteristics and
-advantages of such implementation.
-
-The primary goal of this study is to investigate the performance of JSO
-in PP optimisation. A series of simulation experiments using
-well-recognised datasets and various PP indexes are conducted to yield
-insights regarding the behaviour of JSO under different settings and its
-sensitivity to hyper-parameter. Second, to observe the performance of
-JSO with different types of PP indexes, metrics are introduced to
-capture specific properties of the index including squintability and
-smoothness as introduced in \citet{laa_using_2020}. To the best of the
-authors' knowledge, this is the first attempt to quantitatively measure
-squintability and smoothness within this context. Finally, the
-relationship between the JSO performance, hyper-parameters tuning and
-various properties of PP indexes is analysed to provide helpful guidance
-for practitioners using projection pursuit.
-
-The rest of this paper is structured as follows.
-Section~\ref{sec-background} introduces the background of PP guided tour
-and reviews existing optimisers and index functions in the literature.
+practical advantages make JSO a promising candidate for enhancing PP
+optimisation.
+
+The primary goal of the study reported here is to investigate the
+performance of JSO in PP optimisation for the guided tour. It is of
+interest to assess how quickly and closely the optimiser reaches a gobal
+optima, for various PP indexes that may have differing complexities. To
+observe the performance of JSO with different types of PP indexes,
+metrics are introduced to capture specific properties of the index
+including squintability and smoothness as introduced in
+\citet{laa_using_2020}. Metrics for squintability and smoothness are
+precisely defined here, something that is new for the field. A series of
+simulation experiments using specially structured datasets and various
+PP indexes are conducted to assess the behaviour of JSO under different
+settings and its sensitivity to hyper-parameter choices. The
+relationship between the JSO performance, hyper-parameter tuning and
+properties of PP indexes is analysed to provide guidance for
+practitioners using projection pursuit, and developers of new indexes.
+
+The paper is structured as follows. Section~\ref{sec-background}
+introduces the background of PP guided tour and reviews existing
+optimisers and index functions in the literature.
 Section~\ref{sec-theory} details JSO and introduces metrics that measure
@@
 
-\begin{tcolorbox}[enhanced jigsaw, left=2mm, rightrule=.15mm, colframe=quarto-callout-note-color-frame, breakable, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, coltitle=black, titlerule=0mm, toptitle=1mm, arc=.35mm, bottomrule=.15mm, opacitybacktitle=0.6, leftrule=.75mm, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, opacityback=0, coltitle=black, colframe=quarto-callout-note-color-frame, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, breakable, bottomtitle=1mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, toptitle=1mm, opacitybacktitle=0.6, bottomrule=.15mm, left=2mm, colback=white, titlerule=0mm, toprule=.15mm]
 

### 2024-06-26 optim.qmd
SBO acronym added

@@
 @sec-vis outlines the data structures, numerical summaries, and visualisations used in the simulation studies conducted in @sec-app to assess JSO's performance.
-@sec-app presents two sets of simulation experiments, comparing JSO with another optimiser, search better, and studies the impact of different PP index properties on the optimisation
-performance. @sec-conclusion summarises the work and provides suggestions for future directions.
+@sec-app presents two sets of simulation experiments, comparing JSO permformance improvements relative to an existing optimiser, and studies the impact of different PP index properties on the optimisation performance.
+@sec-conclusion summarises the work and provides suggestions for future directions.
 
@@
 
-The performance of JSO is investigated in two folds: 1) compare it with
-a commonly used optimiser: the search-better optimiser [@RJ-2021-105;
-@laa_using_2020], and 2) examine its success rate under different
-hyper-parameters (number of jellyfishes and maximum number of tries).
-The projection pursuit problem used is finding the pipe shape using the
-holes index, investigated by @laa_using_2020.
+The performance of JSO is investigated in two folds: 1) compare it with a commonly used optimiser: the search-better optimiser (SBO) [@RJ-2021-105;
+@laa_using_2020], and 2) examine its success rate under different hyper-parameters (number of jellyfishes and maximum number of tries).
+The projection pursuit problem used is finding the pipe shape using the holes index, investigated by @laa_using_2020.
 
@@
 #| fig.align = "center",
-#| fig.cap = "Projections found by the jellyfish and search better optimisers at each 10th quantile across 50 simulations.
+#| fig.cap = "Projections found by the JSO and SBO at each 10th quantile across 50 simulations.
 The projection pursuit problem is to find the pipe shape using the holes index in the 6, 8, 10, and 12-dimensional spaces.
 The JSO uses 100 jellyfishes and a maximum number of tries of 100.
-The search better optimiser uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates.
-In the 6-D data space, JSO always finds a clear pipe shape while the search better optimiser also finds the pipe shape but with a wide rim.
-At higher data dimensions, JSO finds a higher index value and a clearer pipe shape across all the quantiles than the search better optimiser."
+The SBO uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates.
+In the 6-D data space, JSO always finds a clear pipe shape while the SBO also finds the pipe shape but with a wide rim.
+At higher data dimensions, JSO finds a higher index value and a clearer pipe shape across all the quantiles than the SBO."
 
@@
 
-(p1 + ggtitle("The jellyfish search optimiser"))/(p2 + ggtitle("The search better optimiser"))
+(p1 + ggtitle("a. JSO"))/(p2 + ggtitle("b. SBO"))
 ```
@@
 
-Fifty simulations are conducted with both JSO and the search better optimal, in
-four dimensions ($d = 6, 8, 10, 12$). JSO uses 100 jellyfishes and
-100 maximum number of tries, while the search-better optimiser allows a
-maximum of 1000 samples at each iteration before the algorithm
-terminates. @fig-proj presents the final projections found by the two
-optimisers, broken down by 10th quantile, faceted by the data dimension.
-In the 6-dimensional data scenario, JSO consistently identifies a clear
-pipe shape. The search better optimiser also finds the pipe shape but
-with a wide rim, suggesting a further polish search may be required.
-With increasing dimensions, JSO may not always identify the pipe shape
-due to random sampling, but it still presents a pipe shape in over 50%
-of cases. When compared to the search better optimiser, JSO reaches
-higher index values and clearer pipe shapes across all quantiles.
+Fifty simulations are conducted with both JSO and the SBO, in four dimensions ($d = 6, 8, 10, 12$).
+JSO uses 100 jellyfishes and 100 maximum number of tries, while the SBO allows a maximum of 1000 samples at each iteration before the algorithm terminates.
+@fig-proj presents the final projections found by the two optimisers, broken down by 10th quantile, faceted by the data dimension.
+In the 6-dimensional data scenario, JSO consistently identifies a clear pipe shape.
+The SBO also finds the pipe shape but with a wide rim, suggesting a further polish search may be required.
+With increasing dimensions, JSO may not always identify the pipe shape due to random sampling, but it still presents a pipe shape in over 50%
+of cases. When compared to the SBO JSO reaches higher index values and clearer pipe shapes across all quantiles.
 

### 2024-06-26 optim.tex
SBO acronym added

@@
 performance. Section~\ref{sec-app} presents two sets of simulation
-experiments, comparing JSO with another optimiser, search better, and
-studies the impact of different PP index properties on the optimisation
-performance. Section~\ref{sec-conclusion} summarises the work and
-provides suggestions for future directions.
+experiments, comparing JSO permformance improvements relative to an
+existing optimiser, and studies the impact of different PP index
+properties on the optimisation performance. Section~\ref{sec-conclusion}
+summarises the work and provides suggestions for future directions.
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacityback=0, coltitle=black, colframe=quarto-callout-note-color-frame, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, breakable, bottomtitle=1mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, toptitle=1mm, opacitybacktitle=0.6, bottomrule=.15mm, left=2mm, colback=white, titlerule=0mm, toprule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, rightrule=.15mm, opacitybacktitle=0.6, toprule=.15mm, coltitle=black, colback=white, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, toptitle=1mm, bottomtitle=1mm, titlerule=0mm, left=2mm]
 
@@
 The performance of JSO is investigated in two folds: 1) compare it with
-a commonly used optimiser: the search-better optimiser
+a commonly used optimiser: the search-better optimiser (SBO)
 \citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
@@
 
-\caption{\label{fig-proj}Projections found by the jellyfish and search
-better optimisers at each 10th quantile across 50 simulations. The
-projection pursuit problem is to find the pipe shape using the holes
-index in the 6, 8, 10, and 12-dimensional spaces. The JSO uses 100
-jellyfishes and a maximum number of tries of 100. The search better
-optimiser uses a maximum of 1000 tries in each step of random sampling
-step before the algorithm terminates. In the 6-D data space, JSO always
-finds a clear pipe shape while the search better optimiser also finds
-the pipe shape but with a wide rim. At higher data dimensions, JSO finds
-a higher index value and a clearer pipe shape across all the quantiles
-than the search better optimiser.}
+\caption{\label{fig-proj}Projections found by the JSO and SBO at each
+10th quantile across 50 simulations. The projection pursuit problem is
+to find the pipe shape using the holes index in the 6, 8, 10, and
+12-dimensional spaces. The JSO uses 100 jellyfishes and a maximum number
+of tries of 100. The SBO uses a maximum of 1000 tries in each step of
+random sampling step before the algorithm terminates. In the 6-D data
+space, JSO always finds a clear pipe shape while the SBO also finds the
+pipe shape but with a wide rim. At higher data dimensions, JSO finds a
+higher index value and a clearer pipe shape across all the quantiles
+than the SBO.}
 
@@
 
-Fifty simulations are conducted with both JSO and the search better
-optimal, in four dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100
-jellyfishes and 100 maximum number of tries, while the search-better
-optimiser allows a maximum of 1000 samples at each iteration before the
-algorithm terminates. Figure~\ref{fig-proj} presents the final
-projections found by the two optimisers, broken down by 10th quantile,
-faceted by the data dimension. In the 6-dimensional data scenario, JSO
-consistently identifies a clear pipe shape. The search better optimiser
+Fifty simulations are conducted with both JSO and the SBO, in four
+dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100 jellyfishes and 100
+maximum number of tries, while the SBO allows a maximum of 1000 samples
+at each iteration before the algorithm terminates. Figure~\ref{fig-proj}
+presents the final projections found by the two optimisers, broken down
+by 10th quantile, faceted by the data dimension. In the 6-dimensional
+data scenario, JSO consistently identifies a clear pipe shape. The SBO
 also finds the pipe shape but with a wide rim, suggesting a further
@@
 always identify the pipe shape due to random sampling, but it still
-presents a pipe shape in over 50\% of cases. When compared to the search
-better optimiser, JSO reaches higher index values and clearer pipe
-shapes across all quantiles.
+presents a pipe shape in over 50\% of cases. When compared to the SBO
+JSO reaches higher index values and clearer pipe shapes across all
+quantiles.
 

### 2024-06-26 optim.qmd
squintability

@@
 
-A tour on high-dimensional data is constructed by geodesically
-interpolating between pairs of planes. Any plane is described by an
-orthonormal basis, $A_t$, where $t$ represents time in the sequence. The
-term "geodesic" refers to maintaining the orthonormality constraint so
-that each view shown is correctly a projection of the data. The PP
-guided tour operates by geodesically interpolating to target planes
-(projections) which have high PP index values, as provided by the
-optimiser. The geodesic interpolation means that the viewer sees a
-continuous sequence of projections of the data, so they can watch
-patterns of interest forming as the function is optimised. There are
-five optimisation methods implemented in the `tourr` package:
+A tour on high-dimensional data is constructed by geodesically interpolating between pairs of planes.
+Any plane is described by an orthonormal basis, $A_t$, where $t$ represents time in the sequence.
+The term "geodesic" refers to maintaining the orthonormality constraint so that each view shown is correctly a projection of the data.
+The PP guided tour operates by geodesically interpolating to target planes (projections) which have high PP index values, as provided by the optimiser.
+The geodesic interpolation means that the viewer sees a continuous sequence of projections of the data, so they can watch patterns of interest forming as the function is optimised.
+There are five (unsophisticated) optimisation methods implemented in the `tourr` package:
 
--   `search_geodesic()`: provides a pseudo-derivative optimisation. It
-    searches locally for the best direction, based on differencing the
-    index values for very close projections. Then it follows the
-    direction along the geodesic path between planes, stopping when the
-    next index value fails to increase.
--   `search_better()`: is a brute-force optimisation searching randomly
-    for projections with higher index values.
--   `search_better_random()`: is essentially simulated annealing
-    [@Bertsimas93] where the search space is reduced as the optimisation
-    progresses.
+-   `search_geodesic()`: provides a pseudo-derivative optimisation.
+It searches locally for the best direction, based on differencing the index values for very close projections.
+Then it follows the direction along the geodesic path between planes, stopping when the next index value fails to increase.
+-   `search_better()`: is a brute-force optimisation searching randomly for projections with higher index values.
+-   `search_better_random()`: is essentially simulated annealing [@Bertsimas93] where the search space is reduced as the optimisation progresses.
 -   `search_posse()`: implements the algorithm described in @posse95.
--   `search_polish()`: is a very localised search, to take tiny steps to
-    get closer to the local maximum.
+-   `search_polish()`: is a very localised search, to take tiny steps to get closer to the local maximum.
 
@@
 
-From the literature, it is commonly understood that a large squint angle
-implies that the objective function value is close to optimal even when
-we are not very close to the perfect view to see the structure. A small
-squint angle means that index function value improves substantially only
-when we are very close to the perfect view. As such, low squintability
-implies rapid improvement in the index value when near the perfect view.
+Based on the use in existing literature, it is commonly understood that a large squint angle implies that the objective function value is close to optimal even when
+we are not very close to the perfect view to see the structure.
+A small squint angle means that index function value improves substantially only when we are very close to the perfect view.
+As such, low squintability implies rapid improvement in the index value when near the perfect view.
+For PP, a small squint angle is considered to be undesirable because it means that the optimiser needs to be very close to be able to "see" the optima.
+Thus, it could be difficult for the optimiser to find the optima.
 
-In this subsection, a metric to capture the notion of squintability is
-proposed, and two approaches to compute this metric numerically are
-detailed.
+In this subsection, a metric to capture the notion of squintability is proposed, and two approaches to compute this metric numerically are detailed.
 

### 2024-06-26 optim.tex
squintability

@@
 patterns of interest forming as the function is optimised. There are
-five optimisation methods implemented in the \texttt{tourr} package:
+five (unsophisticated) optimisation methods implemented in the
+\texttt{tourr} package:
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, rightrule=.15mm, opacitybacktitle=0.6, toprule=.15mm, coltitle=black, colback=white, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, toptitle=1mm, bottomtitle=1mm, titlerule=0mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, toptitle=1mm, bottomrule=.15mm, breakable, colback=white, colframe=quarto-callout-note-color-frame, leftrule=.75mm, coltitle=black, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, titlerule=0mm, left=2mm, opacityback=0, rightrule=.15mm]
 
@@
 
-From the literature, it is commonly understood that a large squint angle
-implies that the objective function value is close to optimal even when
-we are not very close to the perfect view to see the structure. A small
-squint angle means that index function value improves substantially only
-when we are very close to the perfect view. As such, low squintability
-implies rapid improvement in the index value when near the perfect view.
+Based on the use in existing literature, it is commonly understood that
+a large squint angle implies that the objective function value is close
+to optimal even when we are not very close to the perfect view to see
+the structure. A small squint angle means that index function value
+improves substantially only when we are very close to the perfect view.
+As such, low squintability implies rapid improvement in the index value
+when near the perfect view. For PP, a small squint angle is considered
+to be undesirable because it means that the optimiser needs to be very
+close to be able to ``see'' the optima. Thus, it could be difficult for
+the optimiser to find the optima.
 

### 2024-06-26 optim.qmd
bullet point conclusions

@@
 
-The data structure proposed in @RJ-2021-105 is implemented for JSO. 
-An example data collected from finding the sine-wave structure
-in 6D data using a spline index function (`splines2d`) is
-printed below:
+The data structure proposed in @RJ-2021-105 is used here for assessing JSO. A sample for the sine-wave structure in 6D data using a spline index function (`splines2d`) is printed below:
 
@@
 
+- Summarise the result of the simulation: JSO immense improvement on current methods, ...
+- JSO for a guided tour, what changes are needed? Do we follow one tentacle, do we perform off-line and then pull tentacles out to watch, how does it integrate with ferrn
+
 # References

### 2024-06-26 optim.tex
bullet point conclusions

@@
 
-\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, toptitle=1mm, bottomrule=.15mm, breakable, colback=white, colframe=quarto-callout-note-color-frame, leftrule=.75mm, coltitle=black, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, titlerule=0mm, left=2mm, opacityback=0, rightrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, toprule=.15mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, toptitle=1mm, coltitle=black, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, breakable, colframe=quarto-callout-note-color-frame, opacitybacktitle=0.6, left=2mm, bottomtitle=1mm, arc=.35mm, bottomrule=.15mm]
 
@@
 
-The data structure proposed in \citet{RJ-2021-105} is implemented for
-JSO. An example data collected from finding the sine-wave structure in
-6D data using a spline index function (\texttt{splines2d}) is printed
-below:
+The data structure proposed in \citet{RJ-2021-105} is used here for
+assessing JSO. A sample for the sine-wave structure in 6D data using a
+spline index function (\texttt{splines2d}) is printed below:
 
@@
 
+\begin{itemize}
+\tightlist
+\item
+  Summarise the result of the simulation: JSO immense improvement on
+  current methods, \ldots{}
+\item
+  JSO for a guided tour, what changes are needed? Do we follow one
+  tentacle, do we perform off-line and then pull tentacles out to watch,
+  how does it integrate with ferrn
+\end{itemize}
+
 \renewcommand\refname{References}

### 2024-06-27 optim.qmd
update metrics subsections
update of smoothness and squitability subsections. Added references to GpGp and ExaGeoStatR.

@@
 invariance, and speed). Since not all the properties affects the
-execution of the optimisation, here we consider the three relevant
-properties (smoothness, squintability, and speed), and propose three
-metrics to evaluate these three properties.
+execution of the optimisation, we choose to focus here on the first two properties, 
+namely *smoothness* (@sec-smoothness) and *squintability* (@sec-squintability), and propose two metrics to evaluate them.
 
@@
 ```
-If we evaluate the index function at some random points (like the random
-initialization of the jellyfish optimizer), then we can interpret these
-random index values as a random field, indexed by a space parameter: the
-random projection angle. This analogy suggests to use this random
-training sample to fit a spatial model, a simple one being a (spatial)
-Gaussian process.
 
-```{=html}
-<!--
-There exist R packages to fit the hyperparameters of a Gaussian process, for example ExaGeoStatR (https://github.com/ecrc/exageostatR), which is an R wrapper for the ExaGeoStat library (https://arxiv.org/pdf/1708.02835.pdf).
--->
-```
-How can we define a measure of smoothness from this? The distribution of
-a Gaussian process is fully determined by its mean function and
-covariance function. The way the covariance function is defined is where
-smoothness comes into play: if an index is very smooth, then two close
+This subsection proposes a metric for the smoothness of a projection
+pursuit index.
+
+If a PP index function is evaluated at some random points, as is done at the 
+initialization stage of the jellyfish search optimization, then 
+these random index values can be interpreted as a random field, 
+indexed by a space parameter, namely the random projection angle. 
+This analogy suggests to use this random training sample to fit a spatial model, 
+a simple one being a (spatial) Gaussian process.
+
+The distribution of a Gaussian process is fully determined by its mean function and
+covariance function. The smoothness property comes into play in the definition
+of the covariance function: if a PP index is very smooth, then two close
 projection angles should produce close index values (strong
-correlation); by contrast, if an index is not smooth, then two close
+correlation); by contrast, if a PP index is not very smooth, then two close
 projection angles might give very different index values (fast decay of
@@
 $$
-K(u):=\frac{(\sqrt{2\nu}u)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}(\sqrt{2\nu}u)
+K(u):=\frac{(\sqrt{2\nu}u)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}(\sqrt{2\nu}u)\ ,
 $$
@@
 values in the literature being $1/2$, $3/2$ and $5/2$ . The parameter
-$\nu$, called smoothness parameter, controls the decay of the covariance
+$\nu$, called *smoothness parameter*, controls the decay of the covariance
 function. As such, it is an appropriate measure of smoothness of a
@@
 
-In our context, we suggest to use this parameter as a measure of the
+In our context, we suggest to use this parameter $\nu$ as a measure of the
 smoothness of the index function by fitting a Gaussian process prior
 with Matérn covariance on a dataset generated by random evaluations of
-the index function, as in the initial stage of the jellyfish random
-search. There exist several R packages, such as GpGp or ExaGeoStatR, to
+the index function, as done at the initialization stage of the jellyfish search optimization. There exist several R packages, such as GpGp [@guinness2021gpgp] or ExaGeoStatR [@abdulah2023large], to
 fit the hyperparameters of a GP covariance function on data. In this
@@
 
-The fitted value $\nu>0$ can be interpreted as follows: the higher
-$\nu$, the smoother the index function.
+After fitting the smoothness parameter $\nu>0$, its value can be interpreted as follows: the higher $\nu$, the smoother the index function.
 
@@
 
-Based on the use in existing literature, it is commonly understood that a large squint angle implies that the objective function value is close to optimal even when
+In this subsection, a metric to capture the notion of squintability is proposed, and two approaches to compute this metric numerically are detailed.
+
+In the existing literature, it is commonly understood that a large squint angle implies that the objective function value is close to optimal even when
 we are not very close to the perfect view to see the structure.
-A small squint angle means that index function value improves substantially only when we are very close to the perfect view.
+A small squint angle means that the PP index value improves substantially only when we are very close to the perfect view.
 As such, low squintability implies rapid improvement in the index value when near the perfect view.
@@
 Thus, it could be difficult for the optimiser to find the optima.
-
-In this subsection, a metric to capture the notion of squintability is proposed, and two approaches to compute this metric numerically are detailed.
+We now propose a mathematical formulation of this intuition.
 
@@
 
-$$d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}$$ {#eq-proj-dist}
+$$d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .$$ {#eq-proj-dist}
 
 The index value $f(XA)$ is defined as described in the introduction.
+We now focus our attention to how the index value $f(XA)$ changes with respect to the projection distance $d(A, A^*)$, over the course of the JSO.
 
-```{=tex}
-\textcolor{orange}{Here we need to add a precise, mathematical
-definition of the $x$ and $y$ variables ("angle distance" to optimal?
-and PP index values) used in the figures of this Section.}
-```
+```{=html}
+<!--
 \[We generate random points that is beyond 1.5 projection distance and
 interpolate. Then we fit a kernel or use nonlinear least squares.\]
+-->
+```
 
-Let $f$ represent the average PP index value with respect to the
-projection angle distance to optimal, over the course of the JSO. It is
-expected that for a PP index with high squint angle, the optimization
+It is expected that for a PP index with high squint angle, the optimization
 (@eq-optimization) should make substantial progress early on.
@@
 values of $f'$ (the ones for which $f''=0$, assuming that $f$ is twice
-differentiable), and the angles for which these values are attained,
+differentiable), and the projection distances for which these values are attained,
 should play an important role in the mathematical definition of
@@
 $$
-\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)
+\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ .
 $$ {#eq-squintability}
@@
 gradient of $-f$. On the other hand, $\underset{x}{\arg\min}f'(x)$
-represents the angle value at which this largest gradient is attained.
+represents the projection distance at which this largest gradient is attained.
 Based on the above discussion, it is expected that these two values
@@
 $$
-\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}
+\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
 $$ {#eq-squintability-parametric}
@@
 
-\textcolor{orange}{See whether further details are needed for the nonparametric approach.}
+<!-- \textcolor{orange}{See whether further details are needed for the nonparametric approach.} -->
 
-To the best of our knowledge, this is the first attempt to measure the
+To the best of our knowledge, this is the first attempt to measure mathematically the
 notion of squintability.

### 2024-06-27 optim.tex
update metrics subsections
update of smoothness and squitability subsections. Added references to GpGp and ExaGeoStatR.

@@
 \end{frontmatter}
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, frame hidden, sharp corners, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{equation}\phantomsection\label{eq-optimization}{
+\begin{equation}\protect\hypertarget{eq-optimization}{}{
 \underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
-}\end{equation}
+}\label{eq-optimization}\end{equation}
 
@@
 
+\hypertarget{sec-background}{%
 \section{Projection pursuit, tours, index functions and optimisation
-{[}Di and Sherry{]}}\label{sec-background}
+{[}Di and Sherry{]}}\label{sec-background}}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, toprule=.15mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, toptitle=1mm, coltitle=black, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, breakable, colframe=quarto-callout-note-color-frame, opacitybacktitle=0.6, left=2mm, bottomtitle=1mm, arc=.35mm, bottomrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, opacitybacktitle=0.6, left=2mm, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, titlerule=0mm, bottomtitle=1mm, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, arc=.35mm, breakable, rightrule=.15mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm]
 
@@
 rotation invariance, and speed). Since not all the properties affects
-the execution of the optimisation, here we consider the three relevant
-properties (smoothness, squintability, and speed), and propose three
-metrics to evaluate these three properties.
-
-\subsection{Smoothness}\label{sec-smoothness}
-
-If we evaluate the index function at some random points (like the random
-initialization of the jellyfish optimizer), then we can interpret these
-random index values as a random field, indexed by a space parameter: the
-random projection angle. This analogy suggests to use this random
-training sample to fit a spatial model, a simple one being a (spatial)
-Gaussian process.
-
-How can we define a measure of smoothness from this? The distribution of
-a Gaussian process is fully determined by its mean function and
-covariance function. The way the covariance function is defined is where
-smoothness comes into play: if an index is very smooth, then two close
-projection angles should produce close index values (strong
-correlation); by contrast, if an index is not smooth, then two close
-projection angles might give very different index values (fast decay of
-correlations with respect to distance between angles).
+the execution of the optimisation, we choose to focus here on the first
+two properties, namely \emph{smoothness} (Section~\ref{sec-smoothness})
+and \emph{squintability} (Section~\ref{sec-squintability}), and propose
+two metrics to evaluate them.
+
+\hypertarget{sec-smoothness}{%
+\subsection{Smoothness}\label{sec-smoothness}}
+
+This subsection proposes a metric for the smoothness of a projection
+pursuit index.
+
+If a PP index function is evaluated at some random points, as is done at
+the initialization stage of the jellyfish search optimization, then
+these random index values can be interpreted as a random field, indexed
+by a space parameter, namely the random projection angle. This analogy
+suggests to use this random training sample to fit a spatial model, a
+simple one being a (spatial) Gaussian process.
+
+The distribution of a Gaussian process is fully determined by its mean
+function and covariance function. The smoothness property comes into
+play in the definition of the covariance function: if a PP index is very
+smooth, then two close projection angles should produce close index
+values (strong correlation); by contrast, if a PP index is not very
+smooth, then two close projection angles might give very different index
+values (fast decay of correlations with respect to distance between
+angles).
 
@@
 \[
-K(u):=\frac{(\sqrt{2\nu}u)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}(\sqrt{2\nu}u)
+K(u):=\frac{(\sqrt{2\nu}u)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}(\sqrt{2\nu}u)\ ,
 \]
@@
 half-integer, the most popular values in the literature being \(1/2\),
-\(3/2\) and \(5/2\) . The parameter \(\nu\), called smoothness
-parameter, controls the decay of the covariance function. As such, it is
-an appropriate measure of smoothness of a random field.
-
-In our context, we suggest to use this parameter as a measure of the
-smoothness of the index function by fitting a Gaussian process prior
+\(3/2\) and \(5/2\) . The parameter \(\nu\), called \emph{smoothness
+parameter}, controls the decay of the covariance function. As such, it
+is an appropriate measure of smoothness of a random field.
+
+In our context, we suggest to use this parameter \(\nu\) as a measure of
+the smoothness of the index function by fitting a Gaussian process prior
 with Matérn covariance on a dataset generated by random evaluations of
-the index function, as in the initial stage of the jellyfish random
-search. There exist several R packages, such as GpGp or ExaGeoStatR, to
-fit the hyperparameters of a GP covariance function on data. In this
+the index function, as done at the initialization stage of the jellyfish
+search optimization. There exist several R packages, such as GpGp
+\citep{guinness2021gpgp} or ExaGeoStatR \citep{abdulah2023large}, to fit
+the hyperparameters of a GP covariance function on data. In this
 project, we make use of the GpGp package.
 
-The fitted value \(\nu>0\) can be interpreted as follows: the higher
-\(\nu\), the smoother the index function.
-
-\subsection{Squintability}\label{sec-squintability}
-
-Based on the use in existing literature, it is commonly understood that
-a large squint angle implies that the objective function value is close
-to optimal even when we are not very close to the perfect view to see
-the structure. A small squint angle means that index function value
-improves substantially only when we are very close to the perfect view.
-As such, low squintability implies rapid improvement in the index value
-when near the perfect view. For PP, a small squint angle is considered
-to be undesirable because it means that the optimiser needs to be very
-close to be able to ``see'' the optima. Thus, it could be difficult for
-the optimiser to find the optima.
+After fitting the smoothness parameter \(\nu>0\), its value can be
+interpreted as follows: the higher \(\nu\), the smoother the index
+function.
+
+\hypertarget{sec-squintability}{%
+\subsection{Squintability}\label{sec-squintability}}
 
@@
 detailed.
+
+In the existing literature, it is commonly understood that a large
+squint angle implies that the objective function value is close to
+optimal even when we are not very close to the perfect view to see the
+structure. A small squint angle means that the PP index value improves
+substantially only when we are very close to the perfect view. As such,
+low squintability implies rapid improvement in the index value when near
+the perfect view. For PP, a small squint angle is considered to be
+undesirable because it means that the optimiser needs to be very close
+to be able to ``see'' the optima. Thus, it could be difficult for the
+optimiser to find the optima. We now propose a mathematical formulation
+of this intuition.
 
@@
 
-\begin{equation}\phantomsection\label{eq-proj-dist}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}}\end{equation}
+\begin{equation}\protect\hypertarget{eq-proj-dist}{}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .}\label{eq-proj-dist}\end{equation}
 
 The index value \(f(XA)\) is defined as described in the introduction.
-
-\textcolor{orange}{Here we need to add a precise, mathematical
-definition of the $x$ and $y$ variables ("angle distance" to optimal?
-and PP index values) used in the figures of this Section.}
-
-{[}We generate random points that is beyond 1.5 projection distance and
-interpolate. Then we fit a kernel or use nonlinear least squares.{]}
-
-Let \(f\) represent the average PP index value with respect to the
-projection angle distance to optimal, over the course of the JSO. It is
-expected that for a PP index with high squint angle, the optimization
-(\ref{eq-optimization}) should make substantial progress early on.
-Conversely, for a PP index with low squint angle, it might take a long
-while for the optimization to make substantial progress, as the
-candidate projections would need to be very close to the optimal one for
-the structure of the index function to be visible enough to be amenable
-to efficient optimization. This observation suggests that the extreme
-values of \(f'\) (the ones for which \(f''=0\), assuming that \(f\) is
-twice differentiable), and the angles for which these values are
-attained, should play an important role in the mathematical definition
-of squintability. Noticing that \(f\) is expected to be a decreasing
-function, we propose the following squintability metric:
-
-\begin{equation}\phantomsection\label{eq-squintability}{
-\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)
-}\end{equation}
+We now focus our attention to how the index value \(f(XA)\) changes with
+respect to the projection distance \(d(A, A^*)\), over the course of the
+JSO.
+
+It is expected that for a PP index with high squint angle, the
+optimization (\ref{eq-optimization}) should make substantial progress
+early on. Conversely, for a PP index with low squint angle, it might
+take a long while for the optimization to make substantial progress, as
+the candidate projections would need to be very close to the optimal one
+for the structure of the index function to be visible enough to be
+amenable to efficient optimization. This observation suggests that the
+extreme values of \(f'\) (the ones for which \(f''=0\), assuming that
+\(f\) is twice differentiable), and the projection distances for which
+these values are attained, should play an important role in the
+mathematical definition of squintability. Noticing that \(f\) is
+expected to be a decreasing function, we propose the following
+squintability metric:
+
+\begin{equation}\protect\hypertarget{eq-squintability}{}{
+\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ .
+}\label{eq-squintability}\end{equation}
 
@@
 gradient of \(-f\). On the other hand, \(\underset{x}{\arg\min}f'(x)\)
-represents the angle value at which this largest gradient is attained.
-Based on the above discussion, it is expected that these two values
-should be both high in the case of high squintability (fast increase in
-\(f\) early on), and both low in the case of low squintability (any
-substantial increase in \(f\) happens very late, close to the optimal
-angle). This suggest that their product (\ref{eq-squintability}) should
-provide a sensible measure of squintability. The multiplicative constant
-\(4\), which can be deemed arbitrary, does not change the interpretation
-of the squintability metric \(\varsigma\); it is here to adjust the
-range of values of \(\varsigma\) and simplify the explicit formula for
-\(\varsigma\) obtained later on.
+represents the projection distance at which this largest gradient is
+attained. Based on the above discussion, it is expected that these two
+values should be both high in the case of high squintability (fast
+increase in \(f\) early on), and both low in the case of low
+squintability (any substantial increase in \(f\) happens very late,
+close to the optimal angle). This suggest that their product
+(\ref{eq-squintability}) should provide a sensible measure of
+squintability. The multiplicative constant \(4\), which can be deemed
+arbitrary, does not change the interpretation of the squintability
+metric \(\varsigma\); it is here to adjust the range of values of
+\(\varsigma\) and simplify the explicit formula for \(\varsigma\)
+obtained later on.
 
@@
 
-\begin{equation}\phantomsection\label{eq-logistic}{
+\begin{equation}\protect\hypertarget{eq-logistic}{}{
 \ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
-}\end{equation}
+}\label{eq-logistic}\end{equation}
 
@@
 
-\begin{equation}\phantomsection\label{eq-parametric}{
+\begin{equation}\protect\hypertarget{eq-parametric}{}{
 f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
-}\end{equation}
+}\label{eq-parametric}\end{equation}
 
@@
 
-\begin{equation}\phantomsection\label{eq-squintability-parametric}{
-\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}
-}\end{equation}
+\begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
+\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
+}\label{eq-squintability-parametric}\end{equation}
 
@@
 
-\textcolor{orange}{See whether further details are needed for the nonparametric approach.}
-
-To the best of our knowledge, this is the first attempt to measure the
-notion of squintability.
-
-\section{Visualisation of jellyfish optimiser}\label{sec-vis}
+To the best of our knowledge, this is the first attempt to measure
+mathematically the notion of squintability.
+
+\hypertarget{sec-vis}{%
+\section{Visualisation of jellyfish optimiser}\label{sec-vis}}
 
@@
 
-\end{figure}%
-
+\end{figure}
+
+\hypertarget{tbl-smoothness-squintability}{}
 \begin{table}
-
-\caption{\label{tbl-smoothness-squintability}Parameters estimated from
-the Gaussian process (including variance, range, smoothness, and nugget)
-and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
-pipe-finding and sine-wave finding problems. The column ``smooth'' and
-``squint'' represent the smoothness and squintability measures.}
-
-\centering{
+\caption{\label{tbl-smoothness-squintability}Parameters estimated from the Gaussian process (including variance,
+range, smoothness, and nugget) and scaled logistic function
+(\(\theta_1\) to \(\theta_4\)) for the pipe-finding and sine-wave
+finding problems. The column ``smooth'' and ``squint'' represent the
+smoothness and squintability measures. }\tabularnewline
 
@@
 
-\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
+\hypertarget{sec-conclusion}{%
+\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}}
 

### 2024-06-28 optim.qmd
half-way restructuing section 4 and 5, add illustration

@@
 
+The JSO provides a new swarm-based search strategy to optimise
+projection pursuit problems. This section outlines two simulation setups
+to demonstrate the performance of JSO. In the first example, the 
+performance of JSO is compared with an exist optimiser, Creeping Random Search (CRS), and explores how JSO behaves
+with different combinations of hyper-parameters. The second example
+studies the effect of optimisation properties defined in @sec-theory,
+along with JSO hyper-parameters, on the outcome (success rate) of the
+optimisation.
+
 The data structure proposed in @RJ-2021-105 is used here for assessing JSO. A sample for the sine-wave structure in 6D data using a spline index function (`splines2d`) is printed below:
@@
 current basis for the next iteration in the search but still records the
-visited inferior basis. This data can be used to derive numerical
-summaries and plots:
+visited inferior basis.
 
--   **Success rate**: The success rate is defined as the proportion of
+## Performance of JSO in pipe-finding problems {#sec-app-1}
+
+The performance of JSO is investigated in two folds: 1) compare it with a commonly used optimiser: the search-better optimiser (SBO) [@RJ-2021-105;
+@laa_using_2020], and 2) examine its success rate under different hyper-parameters (number of jellyfishes and maximum number of tries).
+The projection pursuit problem used is finding the pipe shape using the holes index, investigated by @laa_using_2020.
+Fifty simulations are conducted with both JSO and the SBO, in four dimensions ($d = 6, 8, 10, 12$).
+JSO uses 100 jellyfishes and 100 maximum number of tries, while the SBO allows a maximum of 1000 samples at each iteration before the algorithm terminates.
+
+In the second experiment, fifty simulations are conducted at each
+hyper-parameter combination to analyse its effects on the JSO success
+rate. The hyper-parameters tested include: 1) 20, 50, and 100
+jellyfishes, and 2) 50 and 100 maximum number of tries. The success 
+rate is defined as the proportion of
 simulations that achieves a final index value within 0.05 of the best
@@
 `r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.
-    @fig-projection shows the best projection found in each simulation,
-    sorted by the index value. Most of the projections show a clear
-    sine-wave structure, while the last six show only a straight line
-    with clusters at the end. When local optima exist in the projection
-    pursuit problem, plotting the projections found in each simulation
-    or by each jellyfish can help identify these local optima.
 
--   **Projection distance**: The projection distance between projection
-    bases and the optimal best basis (or the empirical best basis if
-    the optimal is unknown) measures how close a visited basis is to 
-    the best basis and it is calculated as the Frobenius norm in Equation
-    @eq-proj-dist.
-    Plotting projection distances against index values for the 
-    bases visited reveals how the index value changes as the basis
-    approaches the optimal basis, helping to calculate the squintability
-    measure. @fig-index-value-proj-dist shows the index
-    value against the projection distance for the given dataset and its
-    binned version. As the projection distance decreases, the index
-    value initially increases with an increasing slope and then levels
-    off. The index value shows greater variation at medium projection
-    distances compared to small and large distances.
+<!-- @fig-projection shows the best projection found in each simulation, -->
+<!-- sorted by the index value. Most of the projections show a clear -->
+<!-- sine-wave structure, while the last six show only a straight line -->
+<!-- with clusters at the end. When local optima exist in the projection -->
+<!-- pursuit problem, plotting the projections found in each simulation -->
+<!-- or by each jellyfish can help identify these local optima. -->
 
-```{r}
+## Factors affecting JSO success rate: optimisation properties and jellyfish hyper-parameters {#sec-app-2}
+
+In addition to the pipe-finding problem, a new shape, sine wave, is
+investigated in 6D and 8D spaces with six indexes considered:
+`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`. Combining
+with two JSO hyper-parameters, a total of `r nrow(sim_df)` cases
+is produced, comprising of `r nrow(filter(sim_df, index == "holes"))`
+pipe-finding cases and `r nrow(filter(sim_df, index != "holes"))`
+sine-wave finding cases. For each case, JSO is run fifty times and the
+success rate is calculated.
+
+Smoothness and squintability are computed for each case, following the
+procedures outlined in @sec-smoothness and @sec-squintability and illustrated in @fig-smoothness and  @fig-squintability. To
+calculate smoothness, three hundred random bases are simulated. Index
+values and projection distance (to the optimal basis) are calculated for
+each random basis before fitting into a Gaussian process model to obtain 
+the smoothness measure. For
+squintability, fifty random bases are sampled and interpolated to the
+optimal basis with a step size of 0.005. Index values and projection 
+distances are calculated for these interpolated bases and the index values 
+are averaged with a bin width of 0.005. The squintability
+measure is then calculated from fitting a 4-parameter scaled logistic
+function of index value against projection distance, estimated by
+non-linear least squares.
+
+<!-- **Projection distance**: The projection distance between projection -->
+<!-- bases and the optimal best basis (or the empirical best basis if -->
+<!-- the optimal is unknown) measures how close a visited basis is to  -->
+<!-- the best basis and it is calculated as the Frobenius norm in Equation -->
+<!-- @eq-proj-dist. Plotting projection distances against index values for the  -->
+<!-- bases visited reveals how the index value changes as the basis -->
+<!-- approaches the optimal basis, helping to calculate the squintability -->
+<!-- measure. -->
+
+```{r eval = FALSE}
 #| label: fig-projection
@@
 ```{r}
-#| label: fig-index-value-proj-dist
-#| fig.height: 3
-#| fig.width: 8
-#| fig.cap: "Projection distance plotted against the index value for detecting the 6D sine-wave structure using the spline index function.
-a) The index values start around 0 with low variance at large projection distances.
-As the projection distance decreases, the index values increase with a larger variance.
-Towards the minimum projection distance, the index values increase and cluster around 1, indicating the optimiser is approaching to the optimum.
-b)  Projection distances are binned with a bin width of 0.1, and index values are averaged over each bin.
-The relationship between the two variables forms a sigmoid curve."
-p1 <- sim_sine_6d_spline_projdist |> 
-  ggplot(aes(x = proj_dist, y = index_val)) +
-  geom_point(size = 0.1, alpha = 0.5) + 
-  labs(x = "Projection distance", y = "Index value")
-
-p2 <- sim_sine_6d_spline_projdist |> 
-  group_by(proj_bin = proj_dist %/% 0.1) |> 
-  summarise(index_val = mean(index_val)) |> 
-  ggplot(aes(x = proj_bin, y = index_val)) + 
-  geom_line() +
-  geom_point() + 
-  labs(x = "Projection distance", y = "Index value")
-
-(p1 | p2) + plot_annotation(tag_levels = "a")
+#| label: fig-smoothness
+#| fig.cap: "Illustration of steps to calculate smoothness. "
+knitr::include_graphics(here::here("figures", "smoothness.png"))
 ```
 
-# Application \[Di and Sherry\] {#sec-app}
+```{r}
+#| label: fig-squintability
+#| fig.cap: "Illustration of steps to calculate squintability.
+For all the bases, calculate the projection distance to the optimal basis and the corresponding index value.
+Bin the index value by the projection distance, then claculate the average index value for each bin.
+Fit the scaled sigmoid function from Equation @eq-squintability to the binned index values against the projection distance using non-linear least square.
+Use Equation @eq-parametric to calculate the squintability measure from the estimated $\\theta_1$ to $\\theta_4$.
+"
+# p1 <- sim_sine_6d_spline_projdist |> 
+#   ggplot(aes(x = proj_dist, y = index_val)) +
+#   geom_point(size = 0.1, alpha = 0.5) + 
+#   labs(x = "Projection distance", y = "Index value") + 
+#   theme_bw() + 
+#   theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30))
+# 
+# 
+# dist_bin <- ceiling(sim_sine_6d_spline_projdist[["proj_dist"]] / 0.1) * 0.1
+# dt <- sim_sine_6d_spline_projdist |>
+#   dplyr::bind_cols(dist_bin = dist_bin) |>
+#   dplyr::group_by(dist_bin) |>
+#   dplyr::summarise(index_val = mean(index_val, na.rm = TRUE))
+# 
+# p2 <-  dt |> 
+#   ggplot(aes(x = dist_bin, y = index_val)) + 
+#   geom_line() +
+#   geom_point() + 
+#   labs(x = "Projection distance", y = "Index value") + 
+#   theme_bw() + 
+#   theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30))
+knitr::include_graphics(here::here("figures", "squintability.png"))
 
-The JSO provides a new swarm-based search strategy to optimise
-projection pursuit problems. Two examples are provided in this section
-to demonstrate its performance. The first example compares the 
-performance of JSO with an exist optimiser, search better, and explores how JSO behaves
-with different combinations of hyper-parameters. The second example
-studies the effect of optimisation properties defined in @sec-theory,
-along with JSO hyper-parameters, on the outcome (success rate) of the
-optimisation.
+```
+
+# Results \[Di and Sherry\] {#sec-app}
 
@@
 
-The performance of JSO is investigated in two folds: 1) compare it with a commonly used optimiser: the search-better optimiser (SBO) [@RJ-2021-105;
-@laa_using_2020], and 2) examine its success rate under different hyper-parameters (number of jellyfishes and maximum number of tries).
-The projection pursuit problem used is finding the pipe shape using the holes index, investigated by @laa_using_2020.
+@fig-proj presents the final projections found by the two optimisers, broken down by 10th quantile, faceted by the data dimension.
+In the 6-dimensional data scenario, JSO consistently identifies a clear pipe shape.
+The SBO also finds the pipe shape but with a wide rim, suggesting a further polish search may be required.
+With increasing dimensions, JSO may not always identify the pipe shape due to random sampling, but it still presents a pipe shape in over 50%
+of cases. When compared to the SBO JSO reaches higher index values and clearer pipe shapes across all quantiles.
+
+@fig-proportion presents the proportion of success for each
+hyper-parameter combinations. As the number of jellyfishes and maximum
+tries increase, the success rate also increases. For simpler problems (6
+dimensions), small parameter values (20 jellyfishes and a maximum number
+of tries of 50) can already result in high success rate, while for
+higher-dimensional problems (8, 10, and 12 dimensions), a combination of
+100 jellyfishes and 100 maximum number of tries is necessary for at
+least half of the jellyfishes to find the projection close to the optimal.
 
@@
 
-Fifty simulations are conducted with both JSO and the SBO, in four dimensions ($d = 6, 8, 10, 12$).
-JSO uses 100 jellyfishes and 100 maximum number of tries, while the SBO allows a maximum of 1000 samples at each iteration before the algorithm terminates.
-@fig-proj presents the final projections found by the two optimisers, broken down by 10th quantile, faceted by the data dimension.
-In the 6-dimensional data scenario, JSO consistently identifies a clear pipe shape.
-The SBO also finds the pipe shape but with a wide rim, suggesting a further polish search may be required.
-With increasing dimensions, JSO may not always identify the pipe shape due to random sampling, but it still presents a pipe shape in over 50%
-of cases. When compared to the SBO JSO reaches higher index values and clearer pipe shapes across all quantiles.
-
-In the second experiment, fifty simulations are conducted at each
-hyper-parameter combination to analyse its effects on the JSO success
-rate. The hyper-parameters tested include: 1) 20, 50, and 100
-jellyfishes, and 2) 50 and 100 maximum number of tries. In each
-simulation, the success rate is calculated as defined in @sec-vis.
-@fig-proportion presents the proportion of success for each
-hyper-parameter combinations. As the number of jellyfishes and maximum
-tries increase, the success rate also increases. For simpler problems (6
-dimensions), small parameter values (20 jellyfishes and a maximum number
-of tries of 50) can already result in high success rate, while for
-higher-dimensional problems (8, 10, and 12 dimensions), a combination of
-100 jellyfishes and 100 maximum number of tries is necessary for at
-least half of the jellyfishes to find the projection close to the optimal.
-
 ## Factors affecting JSO success rate: optimisation properties and jellyfish hyper-parameters {#sec-app-2}
 
-The optimisation properties, including smoothness, squintability, and
-speed, offer numerical metrics to characterise the complexity of
-projection pursuit optimisation problems. This example investigates how
-these metrics, along with the hyper-parameters affect the JSO
-success rate. Simulations are conducted to obtain the success rate
-across a wider array of projection pursuit problems, characterised by
-shape-to-find, data dimension, and the index function, as well as
-different hyper-parameter combinations. Smoothness and squintability are
-calculated for each projection pursuit problem as described in
-@sec-theory. A generalised linear model is used to construct the
-relationship between the success rate, JSO hyper-parameters and
-optimisation properties.
-
-In addition to the pipe-finding problem, a new shape, sine wave, is
-investigated in 6D and 8D spaces with six indexes considered:
-`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`. Combining
-with two JSO hyper-parameters, a total of `r nrow(sim_df)` cases
-is produced, comprising of `r nrow(filter(sim_df, index == "holes"))`
-pipe-finding cases and `r nrow(filter(sim_df, index != "holes"))`
-sine-wave finding cases. For each case, JSO is run fifty times and the
-success rate is calculated as in @sec-app-1.
-
-Smoothness and squintability are computed for each case, following the
-procedures outlined in @sec-smoothness and @sec-squintability. To
-calculate smoothness, three hundred random bases are simulated. Index
-values and projection distance (to the optimal basis) are calculated for
-each random basis before fitting into a Gaussian process model to obtain 
-the smoothness measure. For
-squintability, fifty random bases are sampled and interpolated to the
-optimal basis with a step size of 0.005. Index values and projection 
-distances are calculated for these interpolated bases and the index values 
-are averaged with a bin width of 0.005. The squintability
-measure is then calculated from fitting a 4-parameter scaled logistic
-function of index value against projection distance, estimated by
-non-linear least squares.
+<!-- The optimisation properties, including smoothness, squintability, and -->
+<!-- speed, offer numerical metrics to characterise the complexity of -->
+<!-- projection pursuit optimisation problems. This example investigates how -->
+<!-- these metrics, along with the hyper-parameters affect the JSO -->
+<!-- success rate. Simulations are conducted to obtain the success rate -->
+<!-- across a wider array of projection pursuit problems, characterised by -->
+<!-- shape-to-find, data dimension, and the index function, as well as -->
+<!-- different hyper-parameter combinations. Smoothness and squintability are -->
+<!-- calculated for each projection pursuit problem as described in -->
+<!-- @sec-theory. A generalised linear model is used to construct the -->
+<!-- relationship between the success rate, JSO hyper-parameters and -->
+<!-- optimisation properties. -->
 

### 2024-06-28 optim.tex
half-way restructuing section 4 and 5, add illustration

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, frame hidden, sharp corners, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{sec-background}{%
 \section{Projection pursuit, tours, index functions and optimisation
-{[}Di and Sherry{]}}\label{sec-background}}
+{[}Di and Sherry{]}}\label{sec-background}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, opacitybacktitle=0.6, left=2mm, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, titlerule=0mm, bottomtitle=1mm, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, arc=.35mm, breakable, rightrule=.15mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm]
+\begin{tcolorbox}[enhanced jigsaw, left=2mm, coltitle=black, bottomtitle=1mm, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, bottomrule=.15mm, opacitybacktitle=0.6, breakable, opacityback=0, titlerule=0mm, rightrule=.15mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm]
 
@@
 
-\begin{equation}\protect\hypertarget{eq-proj-dist}{}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .}\label{eq-proj-dist}\end{equation}
+\begin{equation}\phantomsection\label{eq-proj-dist}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .}\end{equation}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-squintability}{}{
+\begin{equation}\phantomsection\label{eq-squintability}{
 \varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ .
-}\label{eq-squintability}\end{equation}
+}\end{equation}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
+\begin{equation}\phantomsection\label{eq-squintability-parametric}{
 \varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
-}\label{eq-squintability-parametric}\end{equation}
+}\end{equation}
 
@@
 
-\hypertarget{sec-vis}{%
-\section{Visualisation of jellyfish optimiser}\label{sec-vis}}
+\section{Visualisation of jellyfish optimiser}\label{sec-vis}
+
+The JSO provides a new swarm-based search strategy to optimise
+projection pursuit problems. This section outlines two simulation setups
+to demonstrate the performance of JSO. In the first example, the
+performance of JSO is compared with an exist optimiser, Creeping Random
+Search (CRS), and explores how JSO behaves with different combinations
+of hyper-parameters. The second example studies the effect of
+optimisation properties defined in Section~\ref{sec-theory}, along with
+JSO hyper-parameters, on the outcome (success rate) of the optimisation.
 
@@
 will keep the current basis for the next iteration in the search but
-still records the visited inferior basis. This data can be used to
-derive numerical summaries and plots:
-
-\begin{itemize}
-\item
-  \textbf{Success rate}: The success rate is defined as the proportion
-  of simulations that achieves a final index value within 0.05 of the
-  best index value found among all 50 simulations. In the dataset above,
-  the optimal index value is 0.999. Out of the 50 simulations, 44
-  achieved an index value within {[}0.949, 0.999{]}, resulting in a
-  success rate of 0.88. Figure~\ref{fig-projection} shows the best
-  projection found in each simulation, sorted by the index value. Most
-  of the projections show a clear sine-wave structure, while the last
-  six show only a straight line with clusters at the end. When local
-  optima exist in the projection pursuit problem, plotting the
-  projections found in each simulation or by each jellyfish can help
-  identify these local optima.
-\item
-  \textbf{Projection distance}: The projection distance between
-  projection bases and the optimal best basis (or the empirical best
-  basis if the optimal is unknown) measures how close a visited basis is
-  to the best basis and it is calculated as the Frobenius norm in
-  Equation
-
-  \begin{enumerate}
-  \def\labelenumi{\arabic{enumi}.}
-  \tightlist
-  \item
-    Plotting projection distances against index values for the bases
-    visited reveals how the index value changes as the basis approaches
-    the optimal basis, helping to calculate the squintability measure.
-    Figure~\ref{fig-index-value-proj-dist} shows the index value against
-    the projection distance for the given dataset and its binned
-    version. As the projection distance decreases, the index value
-    initially increases with an increasing slope and then levels off.
-    The index value shows greater variation at medium projection
-    distances compared to small and large distances.
-  \end{enumerate}
-\end{itemize}
-
-\begin{figure}
-
-{\centering \includegraphics{optim_files/figure-pdf/fig-projection-1.pdf}
-
-}
-
-\caption{\label{fig-projection}Best projections found by the JSO in each
-of the 50 simulations, sorted by index values. The projection pursuit
-problem is to detect the 6D sine-wave structure using the spline index.
-Most simulations have detected a clear sine-wave pattern, while the last
-six show a straight line with clustering at the ends. In this
-simulation, the success rate is 0.88 (44/50).}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics{optim_files/figure-pdf/fig-index-value-proj-dist-1.pdf}
-
-}
-
-\caption{\label{fig-index-value-proj-dist}Projection distance plotted
-against the index value for detecting the 6D sine-wave structure using
-the spline index function. a) The index values start around 0 with low
-variance at large projection distances. As the projection distance
-decreases, the index values increase with a larger variance. Towards the
-minimum projection distance, the index values increase and cluster
-around 1, indicating the optimiser is approaching to the optimum. b)
-Projection distances are binned with a bin width of 0.1, and index
-values are averaged over each bin. The relationship between the two
-variables forms a sigmoid curve.}
-
-\end{figure}
-
-\hypertarget{sec-app}{%
-\section{Application {[}Di and Sherry{]}}\label{sec-app}}
-
-The JSO provides a new swarm-based search strategy to optimise
-projection pursuit problems. Two examples are provided in this section
-to demonstrate its performance. The first example compares the
-performance of JSO with an exist optimiser, search better, and explores
-how JSO behaves with different combinations of hyper-parameters. The
-second example studies the effect of optimisation properties defined in
-Section~\ref{sec-theory}, along with JSO hyper-parameters, on the
-outcome (success rate) of the optimisation.
-
-\hypertarget{sec-app-1}{%
+still records the visited inferior basis.
+
 \subsection{Performance of JSO in pipe-finding
-problems}\label{sec-app-1}}
+problems}\label{sec-app-1}
 
@@
 pipe shape using the holes index, investigated by
-\citet{laa_using_2020}.
+\citet{laa_using_2020}. Fifty simulations are conducted with both JSO
+and the SBO, in four dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100
+jellyfishes and 100 maximum number of tries, while the SBO allows a
+maximum of 1000 samples at each iteration before the algorithm
+terminates.
+
+In the second experiment, fifty simulations are conducted at each
+hyper-parameter combination to analyse its effects on the JSO success
+rate. The hyper-parameters tested include: 1) 20, 50, and 100
+jellyfishes, and 2) 50 and 100 maximum number of tries. The success rate
+is defined as the proportion of simulations that achieves a final index
+value within 0.05 of the best index value found among all 50
+simulations. In the dataset above, the optimal index value is 0.999. Out
+of the 50 simulations, 44 achieved an index value within {[}0.949,
+0.999{]}, resulting in a success rate of 0.88.
+
+\subsection{Factors affecting JSO success rate: optimisation properties
+and jellyfish hyper-parameters}\label{sec-app-2}
+
+In addition to the pipe-finding problem, a new shape, sine wave, is
+investigated in 6D and 8D spaces with six indexes considered:
+\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
+\texttt{spline}, and \texttt{stringy}. Combining with two JSO
+hyper-parameters, a total of 52 cases is produced, comprising of 24
+pipe-finding cases and 28 sine-wave finding cases. For each case, JSO is
+run fifty times and the success rate is calculated.
+
+Smoothness and squintability are computed for each case, following the
+procedures outlined in Section~\ref{sec-smoothness} and
+Section~\ref{sec-squintability} and illustrated in
+Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}. To
+calculate smoothness, three hundred random bases are simulated. Index
+values and projection distance (to the optimal basis) are calculated for
+each random basis before fitting into a Gaussian process model to obtain
+the smoothness measure. For squintability, fifty random bases are
+sampled and interpolated to the optimal basis with a step size of 0.005.
+Index values and projection distances are calculated for these
+interpolated bases and the index values are averaged with a bin width of
+0.005. The squintability measure is then calculated from fitting a
+4-parameter scaled logistic function of index value against projection
+distance, estimated by non-linear least squares.
 
@@
 
-{\centering \includegraphics{optim_files/figure-pdf/fig-proj-1.pdf}
+\centering{
+
+\includegraphics[width=6.77in,height=\textheight]{figures/smoothness.png}
+
+}
+
+\caption{\label{fig-smoothness}Illustration of steps to calculate
+smoothness.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=6.77in,height=\textheight]{figures/squintability.png}
+
+}
+
+\caption{\label{fig-squintability}Illustration of steps to calculate
+squintability. For all the bases, calculate the projection distance to
+the optimal basis and the corresponding index value. Bin the index value
+by the projection distance, then claculate the average index value for
+each bin. Fit the scaled sigmoid function from Equation
+\ref{eq-squintability} to the binned index values against the projection
+distance using non-linear least square. Use Equation \ref{eq-parametric}
+to calculate the squintability measure from the estimated \(\theta_1\)
+to \(\theta_4\).}
+
+\end{figure}%
+
+\section{Results {[}Di and Sherry{]}}\label{sec-app}
+
+\subsection{Performance of JSO in pipe-finding
+problems}\label{sec-app-1}
+
+Figure~\ref{fig-proj} presents the final projections found by the two
+optimisers, broken down by 10th quantile, faceted by the data dimension.
+In the 6-dimensional data scenario, JSO consistently identifies a clear
+pipe shape. The SBO also finds the pipe shape but with a wide rim,
+suggesting a further polish search may be required. With increasing
+dimensions, JSO may not always identify the pipe shape due to random
+sampling, but it still presents a pipe shape in over 50\% of cases. When
+compared to the SBO JSO reaches higher index values and clearer pipe
+shapes across all quantiles.
+
+Figure~\ref{fig-proportion} presents the proportion of success for each
+hyper-parameter combinations. As the number of jellyfishes and maximum
+tries increase, the success rate also increases. For simpler problems (6
+dimensions), small parameter values (20 jellyfishes and a maximum number
+of tries of 50) can already result in high success rate, while for
+higher-dimensional problems (8, 10, and 12 dimensions), a combination of
+100 jellyfishes and 100 maximum number of tries is necessary for at
+least half of the jellyfishes to find the projection close to the
+optimal.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{optim_files/figure-pdf/fig-proj-1.pdf}
 
@@
 
-\end{figure}
-
-Fifty simulations are conducted with both JSO and the SBO, in four
-dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100 jellyfishes and 100
-maximum number of tries, while the SBO allows a maximum of 1000 samples
-at each iteration before the algorithm terminates. Figure~\ref{fig-proj}
-presents the final projections found by the two optimisers, broken down
-by 10th quantile, faceted by the data dimension. In the 6-dimensional
-data scenario, JSO consistently identifies a clear pipe shape. The SBO
-also finds the pipe shape but with a wide rim, suggesting a further
-polish search may be required. With increasing dimensions, JSO may not
-always identify the pipe shape due to random sampling, but it still
-presents a pipe shape in over 50\% of cases. When compared to the SBO
-JSO reaches higher index values and clearer pipe shapes across all
-quantiles.
-
-In the second experiment, fifty simulations are conducted at each
-hyper-parameter combination to analyse its effects on the JSO success
-rate. The hyper-parameters tested include: 1) 20, 50, and 100
-jellyfishes, and 2) 50 and 100 maximum number of tries. In each
-simulation, the success rate is calculated as defined in
-Section~\ref{sec-vis}. Figure~\ref{fig-proportion} presents the
-proportion of success for each hyper-parameter combinations. As the
-number of jellyfishes and maximum tries increase, the success rate also
-increases. For simpler problems (6 dimensions), small parameter values
-(20 jellyfishes and a maximum number of tries of 50) can already result
-in high success rate, while for higher-dimensional problems (8, 10, and
-12 dimensions), a combination of 100 jellyfishes and 100 maximum number
-of tries is necessary for at least half of the jellyfishes to find the
-projection close to the optimal.
-
-\hypertarget{sec-app-2}{%
+\end{figure}%
+
 \subsection{Factors affecting JSO success rate: optimisation properties
-and jellyfish hyper-parameters}\label{sec-app-2}}
-
-The optimisation properties, including smoothness, squintability, and
-speed, offer numerical metrics to characterise the complexity of
-projection pursuit optimisation problems. This example investigates how
-these metrics, along with the hyper-parameters affect the JSO success
-rate. Simulations are conducted to obtain the success rate across a
-wider array of projection pursuit problems, characterised by
-shape-to-find, data dimension, and the index function, as well as
-different hyper-parameter combinations. Smoothness and squintability are
-calculated for each projection pursuit problem as described in
-Section~\ref{sec-theory}. A generalised linear model is used to
-construct the relationship between the success rate, JSO
-hyper-parameters and optimisation properties.
-
-In addition to the pipe-finding problem, a new shape, sine wave, is
-investigated in 6D and 8D spaces with six indexes considered:
-\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
-\texttt{spline}, and \texttt{stringy}. Combining with two JSO
-hyper-parameters, a total of 52 cases is produced, comprising of 24
-pipe-finding cases and 28 sine-wave finding cases. For each case, JSO is
-run fifty times and the success rate is calculated as in
-Section~\ref{sec-app-1}.
-
-Smoothness and squintability are computed for each case, following the
-procedures outlined in Section~\ref{sec-smoothness} and
-Section~\ref{sec-squintability}. To calculate smoothness, three hundred
-random bases are simulated. Index values and projection distance (to the
-optimal basis) are calculated for each random basis before fitting into
-a Gaussian process model to obtain the smoothness measure. For
-squintability, fifty random bases are sampled and interpolated to the
-optimal basis with a step size of 0.005. Index values and projection
-distances are calculated for these interpolated bases and the index
-values are averaged with a bin width of 0.005. The squintability measure
-is then calculated from fitting a 4-parameter scaled logistic function
-of index value against projection distance, estimated by non-linear
-least squares.
+and jellyfish hyper-parameters}\label{sec-app-2}
 
@@
 
-\end{figure}
-
-\hypertarget{tbl-smoothness-squintability}{}
+\end{figure}%
+
 \begin{table}
-\caption{\label{tbl-smoothness-squintability}Parameters estimated from the Gaussian process (including variance,
-range, smoothness, and nugget) and scaled logistic function
-(\(\theta_1\) to \(\theta_4\)) for the pipe-finding and sine-wave
-finding problems. The column ``smooth'' and ``squint'' represent the
-smoothness and squintability measures. }\tabularnewline
+
+\caption{\label{tbl-smoothness-squintability}Parameters estimated from
+the Gaussian process (including variance, range, smoothness, and nugget)
+and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
+pipe-finding and sine-wave finding problems. The column ``smooth'' and
+``squint'' represent the smoothness and squintability measures.}
+
+\centering{
 
@@
 
-\hypertarget{sec-conclusion}{%
-\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}}
+\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
 

### 2024-06-28 optim.qmd
section 4 is done, more illustrations

@@
 
-# Visualisation of jellyfish optimiser {#sec-vis}
+# Simulation details {#sec-vis}
 
-The JSO provides a new swarm-based search strategy to optimise
-projection pursuit problems. This section outlines two simulation setups
-to demonstrate the performance of JSO. In the first example, the 
-performance of JSO is compared with an exist optimiser, Creeping Random Search (CRS), and explores how JSO behaves
-with different combinations of hyper-parameters. The second example
-studies the effect of optimisation properties defined in @sec-theory,
-along with JSO hyper-parameters, on the outcome (success rate) of the
+JSO provides a new swarm-based search strategy for optimising
+projection pursuit problems. This section describes the setups for two simulations
+to demonstrate JSO's performance. The first simulation compares JSO's 
+performance with an exist optimiser, Creeping Random Search (CRS), and explores how JSO behaves
+with different hyper-parameters combinations. The second simulation
+studies the effect of optimisation properties (smoothness and squintability),
+along with JSO hyper-parameters, on the success rate of the
 optimisation.
 
-The data structure proposed in @RJ-2021-105 is used here for assessing JSO. A sample for the sine-wave structure in 6D data using a spline index function (`splines2d`) is printed below:
+## Performance of JSO in pipe-finding problems {#sec-app-1}
+
+The performance of JSO is investigated in two folds: 1) compare it with a commonly used optimiser: the search-better optimiser (SBO) [@RJ-2021-105;
+@laa_using_2020], and 2) examine its success rate under two hyper-parameters (number of jellyfishes and maximum number of tries).
+The projection pursuit problem used is finding the pipe shape using the holes index, investigated by @laa_using_2020.
+Fifty simulations are conducted with both JSO and the SBO, in four dimensions ($d = 6, 8, 10, 12$).
+JSO uses 100 jellyfishes and 100 maximum number of tries, while the SBO allows a maximum of 1000 samples at each iteration before the algorithm terminates.
+The results of the simulation are collected using the data structure proposed in @RJ-2021-105 is used here for assessing JSO.
+A sample for the sine-wave structure in 6D data using a spline index function (`splines2d`) is printed below:
 
@@
 
-The JSO hyper-parameters used are 50 jellyfishes (`n_jellies`) and
-a maximum of 50 tries (`max_tries`). The full data contains 50
+In this simulation, 50 jellyfishes (`n_jellies`) and
+a maximum of 50 tries (`max_tries`) are used. The full data contains 50
 simulations, indexed by the variable `sim`, with the associated `seed`s. 
@@
 iteration ID, jellyfish ID, and computational time
-respectively. The basis column records every visited basis before
-comparing with the current (last) basis. If a visited basis has a
-smaller index value than the current (last) one, JSO will keep the
-current basis for the next iteration in the search but still records the
-visited inferior basis.
+respectively.
 
-## Performance of JSO in pipe-finding problems {#sec-app-1}
-
-The performance of JSO is investigated in two folds: 1) compare it with a commonly used optimiser: the search-better optimiser (SBO) [@RJ-2021-105;
-@laa_using_2020], and 2) examine its success rate under different hyper-parameters (number of jellyfishes and maximum number of tries).
-The projection pursuit problem used is finding the pipe shape using the holes index, investigated by @laa_using_2020.
-Fifty simulations are conducted with both JSO and the SBO, in four dimensions ($d = 6, 8, 10, 12$).
-JSO uses 100 jellyfishes and 100 maximum number of tries, while the SBO allows a maximum of 1000 samples at each iteration before the algorithm terminates.
+<!-- The basis column records every visited basis before -->
+<!-- comparing with the current (last) basis. If a visited basis has a -->
+<!-- smaller index value than the current (last) one, JSO will keep the -->
+<!-- current basis for the next iteration in the search but still records the -->
+<!-- visited inferior basis.  -->
 
 In the second experiment, fifty simulations are conducted at each
-hyper-parameter combination to analyse its effects on the JSO success
-rate. The hyper-parameters tested include: 1) 20, 50, and 100
-jellyfishes, and 2) 50 and 100 maximum number of tries. The success 
+hyper-parameter combination below to analyse its effects on the JSO success
+rate: number of jellyfish: 20, 50, and 100, and  maximum number of 
+tries: 50 and 100). The success 
 rate is defined as the proportion of
 simulations that achieves a final index value within 0.05 of the best
-index value found among all 50 simulations. In the dataset above,
+index value found among all 50 simulations. @fig-success-rate illustrates 
+the calculation based on the projection found by the above simulation. Here
 the optimal index value is
@@
 
+```{r}
+#| label: fig-success-rate
+#| fig.cap: "Illustration of success rate calculation: The best projections identified by the JSO in 50 simulations are sorted by their index values.
+The highest index value found across all simulation is 0.999.
+Out of the 50 simulations, 44 achieved an index value within 0.05 of the best, resulting in a success rate of 0.88 (44/50)."
+# proj_dt <- map_dfr(sim_sine_6d_spline_best |> pull(basis), 
+#               ~as_tibble(as.matrix(sine1000) %*% .x), .id = "sim") |> 
+#   mutate(sim = as.numeric(sim)) |> 
+#   left_join(sim_sine_6d_spline_best |> select(sim, index_val), by = "sim")
+# 
+# idx_val_sorted <- proj_dt |> pull(index_val) |> unique() |> sort()
+# proj_dt |> 
+#   #filter(index_val %in% c(idx_val_sorted[1:12], idx_val_sorted[39:50])) |>
+#   ggplot(aes(x = V1, y = V2)) + 
+#   geom_point(size = 0.5) + 
+#   geom_text(aes(label = round(index_val, 3)), x = 0, y = 3.7, 
+#             size = 8) +
+#   xlim(-4, 4) + ylim(-4, 4) +
+#   facet_wrap(~fct_reorder(as.factor(sim), -index_val), ncol = 10) + 
+#   theme_bw() +
+#   theme(aspect.ratio = 1, axis.ticks = element_blank(),
+#         axis.text = element_blank(), axis.title = element_blank(),
+#         panel.grid = element_blank(), strip.background = element_blank(),
+#         strip.text = element_blank())
+knitr::include_graphics(here::here("figures", "success-rate.png"))
+```
+
 ## Factors affecting JSO success rate: optimisation properties and jellyfish hyper-parameters {#sec-app-2}
@@
 Smoothness and squintability are computed for each case, following the
-procedures outlined in @sec-smoothness and @sec-squintability and illustrated in @fig-smoothness and  @fig-squintability. To
-calculate smoothness, three hundred random bases are simulated. Index
+procedures outlined in @sec-smoothness and @sec-squintability and illustrated in @fig-smoothness and  @fig-squintability. For smoothness, three hundred random bases are simulated. Index
 values and projection distance (to the optimal basis) are calculated for
@@
 
+Combining with the success rate calculated for each case, the dataset is fitted using 
+a generalised linear model with a binomial family and a logit link
+function with three pre-processing steps are applied: 1) scaling
+the JSO hyper-parameters by a factor of 10 for interpretation, 2)
+creating a new binary variable `long_time` to indicate cases with an
+average run time over 30 seconds, and 3) encoding the success rate for
+the stringy case as 0, because upon visual inspection of the final
+projection, none of the 50 simulations identified the sine-wave shape.
+
 <!-- **Projection distance**: The projection distance between projection -->
@@
 
-```{r eval = FALSE}
-#| label: fig-projection
-#| fig.height: 6
-#| fig.width: 10
-#| fig.cap: Best projections found by the JSO in each of the 50 simulations, sorted by index values.
-The projection pursuit problem is to detect the 6D sine-wave structure using the spline index.
-Most simulations have detected a clear sine-wave pattern, while the last six show a straight line with clustering at the ends.
-In this simulation, the success rate is 0.88 (44/50).
-proj_dt <- map_dfr(sim_sine_6d_spline_best |> pull(basis), 
-              ~as_tibble(as.matrix(sine1000) %*% .x), .id = "sim") |> 
-  mutate(sim = as.numeric(sim)) |> 
-  left_join(sim_sine_6d_spline_best |> select(sim, index_val), by = "sim")
-
-proj_dt |>
-  ggplot(aes(x = V1, y = V2)) + 
-  geom_point(size = 0.1) + 
-  geom_text(aes(label = round(index_val, 3)), x = 0, y = 3.7, 
-            size = 3) +
-  xlim(-4, 4) + ylim(-4, 4) +
-  facet_wrap(~fct_reorder(as.factor(sim), -index_val), ncol = 10) + 
-  theme_bw() +
-  theme(aspect.ratio = 1, axis.ticks = element_blank(),
-        axis.text = element_blank(), axis.title = element_blank(),
-        panel.grid = element_blank())
-```
-
 ```{r}
 #| label: fig-smoothness
-#| fig.cap: "Illustration of steps to calculate smoothness. "
+#| out.width: 100%
+#| fig.cap: "Illustration of steps to calculate smoothness.
+For a given projection pursuit problem defined by the shape to find, data dimension and the index function, sample random bases given the orthonormality contraint.
+Calculate the projection distance and the index value for each random basis.
+Fit a Gaussian process model to the index values against the projection distances to obtain the smoothness measure.
+"
+# basis_smoothness <- sample_bases(idx = "splines2d", n_basis = 1000)
+# basis_smoothness |>
+#    ggplot(aes(x = dist, y = index)) +
+#    geom_point() + 
+#    labs(x = "Projection distance", y = "Index value") + 
+#    theme_bw() +   
+#   theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30))
 knitr::include_graphics(here::here("figures", "smoothness.png"))
@@
 #| label: fig-squintability
+#| out.width: 100%
 #| fig.cap: "Illustration of steps to calculate squintability.
-For all the bases, calculate the projection distance to the optimal basis and the corresponding index value.
+For a given projection pursuit problem defined by the shape to find, data dimension and the index function, sample random bases given the orthonormality and projection distance contraint.
+Interpolate the sampled bases to the optimal basis and calculate the projection distance and the index value for each interpolated basis.
 Bin the index value by the projection distance, then claculate the average index value for each bin.
@@
 knitr::include_graphics(here::here("figures", "squintability.png"))
+```
 
-```
+\newpage
 
@@
 success rate, hyper-parameters, and optimisation properties) into
-one table. Three pre-processing steps are applied to the data: 1) scaling
-the JSO hyper-parameters by a factor of 10 for interpretation, 2)
-creating a new binary variable `long_time` to indicate cases with an
-average run time over 30 seconds, and 3) encoding the success rate for
-the stringy case as 0, because upon visual inspection of the final
-projection, none of the 50 simulations identified the sine-wave shape. A
-generalised linear model with a binomial family and a logit link
-function is used to fit the data and @tbl-mod-output presents the model
+one table. @tbl-mod-output presents the model
 outputs. The model suggests that JSO success rate is positively

### 2024-06-28 optim.tex
section 4 is done, more illustrations

@@
 
-\begin{tcolorbox}[enhanced jigsaw, left=2mm, coltitle=black, bottomtitle=1mm, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, bottomrule=.15mm, opacitybacktitle=0.6, breakable, opacityback=0, titlerule=0mm, rightrule=.15mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm]
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, bottomrule=.15mm, breakable, left=2mm, coltitle=black, colback=white, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, titlerule=0mm, arc=.35mm, toprule=.15mm, opacityback=0]
 
@@
 
-\section{Visualisation of jellyfish optimiser}\label{sec-vis}
-
-The JSO provides a new swarm-based search strategy to optimise
-projection pursuit problems. This section outlines two simulation setups
-to demonstrate the performance of JSO. In the first example, the
-performance of JSO is compared with an exist optimiser, Creeping Random
-Search (CRS), and explores how JSO behaves with different combinations
-of hyper-parameters. The second example studies the effect of
-optimisation properties defined in Section~\ref{sec-theory}, along with
-JSO hyper-parameters, on the outcome (success rate) of the optimisation.
-
-The data structure proposed in \citet{RJ-2021-105} is used here for
-assessing JSO. A sample for the sine-wave structure in 6D data using a
-spline index function (\texttt{splines2d}) is printed below:
+\section{Simulation details}\label{sec-vis}
+
+JSO provides a new swarm-based search strategy for optimising projection
+pursuit problems. This section describes the setups for two simulations
+to demonstrate JSO's performance. The first simulation compares JSO's
+performance with an exist optimiser, Creeping Random Search (CRS), and
+explores how JSO behaves with different hyper-parameters combinations.
+The second simulation studies the effect of optimisation properties
+(smoothness and squintability), along with JSO hyper-parameters, on the
+success rate of the optimisation.
+
+\subsection{Performance of JSO in pipe-finding
+problems}\label{sec-app-1}
+
+The performance of JSO is investigated in two folds: 1) compare it with
+a commonly used optimiser: the search-better optimiser (SBO)
+\citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
+under two hyper-parameters (number of jellyfishes and maximum number of
+tries). The projection pursuit problem used is finding the pipe shape
+using the holes index, investigated by \citet{laa_using_2020}. Fifty
+simulations are conducted with both JSO and the SBO, in four dimensions
+(\(d = 6, 8, 10, 12\)). JSO uses 100 jellyfishes and 100 maximum number
+of tries, while the SBO allows a maximum of 1000 samples at each
+iteration before the algorithm terminates. The results of the simulation
+are collected using the data structure proposed in \citet{RJ-2021-105}
+is used here for assessing JSO. A sample for the sine-wave structure in
+6D data using a spline index function (\texttt{splines2d}) is printed
+below:
 
@@
 
-The JSO hyper-parameters used are 50 jellyfishes (\texttt{n\_jellies})
-and a maximum of 50 tries (\texttt{max\_tries}). The full data contains
-50 simulations, indexed by the variable \texttt{sim}, with the
-associated \texttt{seed}s. Recorded in columns \texttt{basis},
-\texttt{index\_val}, \texttt{tries}, \texttt{loop} and \texttt{time},
-are the sampled projection basis, index value, iteration ID, jellyfish
-ID, and computational time respectively. The basis column records every
-visited basis before comparing with the current (last) basis. If a
-visited basis has a smaller index value than the current (last) one, JSO
-will keep the current basis for the next iteration in the search but
-still records the visited inferior basis.
-
-\subsection{Performance of JSO in pipe-finding
-problems}\label{sec-app-1}
-
-The performance of JSO is investigated in two folds: 1) compare it with
-a commonly used optimiser: the search-better optimiser (SBO)
-\citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
-under different hyper-parameters (number of jellyfishes and maximum
-number of tries). The projection pursuit problem used is finding the
-pipe shape using the holes index, investigated by
-\citet{laa_using_2020}. Fifty simulations are conducted with both JSO
-and the SBO, in four dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100
-jellyfishes and 100 maximum number of tries, while the SBO allows a
-maximum of 1000 samples at each iteration before the algorithm
-terminates.
+In this simulation, 50 jellyfishes (\texttt{n\_jellies}) and a maximum
+of 50 tries (\texttt{max\_tries}) are used. The full data contains 50
+simulations, indexed by the variable \texttt{sim}, with the associated
+\texttt{seed}s. Recorded in columns \texttt{basis}, \texttt{index\_val},
+\texttt{tries}, \texttt{loop} and \texttt{time}, are the sampled
+projection basis, index value, iteration ID, jellyfish ID, and
+computational time respectively.
 
 In the second experiment, fifty simulations are conducted at each
-hyper-parameter combination to analyse its effects on the JSO success
-rate. The hyper-parameters tested include: 1) 20, 50, and 100
-jellyfishes, and 2) 50 and 100 maximum number of tries. The success rate
-is defined as the proportion of simulations that achieves a final index
-value within 0.05 of the best index value found among all 50
-simulations. In the dataset above, the optimal index value is 0.999. Out
-of the 50 simulations, 44 achieved an index value within {[}0.949,
-0.999{]}, resulting in a success rate of 0.88.
+hyper-parameter combination below to analyse its effects on the JSO
+success rate: number of jellyfish: 20, 50, and 100, and maximum number
+of tries: 50 and 100). The success rate is defined as the proportion of
+simulations that achieves a final index value within 0.05 of the best
+index value found among all 50 simulations.
+Figure~\ref{fig-success-rate} illustrates the calculation based on the
+projection found by the above simulation. Here the optimal index value
+is 0.999. Out of the 50 simulations, 44 achieved an index value within
+{[}0.949, 0.999{]}, resulting in a success rate of 0.88.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=4.49in,height=\textheight]{figures/success-rate.png}
+
+}
+
+\caption{\label{fig-success-rate}Illustration of success rate
+calculation: The best projections identified by the JSO in 50
+simulations are sorted by their index values. The highest index value
+found across all simulation is 0.999. Out of the 50 simulations, 44
+achieved an index value within 0.05 of the best, resulting in a success
+rate of 0.88 (44/50).}
+
+\end{figure}%
 
@@
 Section~\ref{sec-squintability} and illustrated in
-Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}. To
-calculate smoothness, three hundred random bases are simulated. Index
-values and projection distance (to the optimal basis) are calculated for
-each random basis before fitting into a Gaussian process model to obtain
-the smoothness measure. For squintability, fifty random bases are
-sampled and interpolated to the optimal basis with a step size of 0.005.
-Index values and projection distances are calculated for these
-interpolated bases and the index values are averaged with a bin width of
-0.005. The squintability measure is then calculated from fitting a
-4-parameter scaled logistic function of index value against projection
-distance, estimated by non-linear least squares.
+Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}. For
+smoothness, three hundred random bases are simulated. Index values and
+projection distance (to the optimal basis) are calculated for each
+random basis before fitting into a Gaussian process model to obtain the
+smoothness measure. For squintability, fifty random bases are sampled
+and interpolated to the optimal basis with a step size of 0.005. Index
+values and projection distances are calculated for these interpolated
+bases and the index values are averaged with a bin width of 0.005. The
+squintability measure is then calculated from fitting a 4-parameter
+scaled logistic function of index value against projection distance,
+estimated by non-linear least squares.
+
+Combining with the success rate calculated for each case, the dataset is
+fitted using a generalised linear model with a binomial family and a
+logit link function with three pre-processing steps are applied: 1)
+scaling the JSO hyper-parameters by a factor of 10 for interpretation,
+2) creating a new binary variable \texttt{long\_time} to indicate cases
+with an average run time over 30 seconds, and 3) encoding the success
+rate for the stringy case as 0, because upon visual inspection of the
+final projection, none of the 50 simulations identified the sine-wave
+shape.
 
@@
 
-\includegraphics[width=6.77in,height=\textheight]{figures/smoothness.png}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/smoothness.png}
 
@@
 \caption{\label{fig-smoothness}Illustration of steps to calculate
-smoothness.}
+smoothness. For a given projection pursuit problem defined by the shape
+to find, data dimension and the index function, sample random bases
+given the orthonormality contraint. Calculate the projection distance
+and the index value for each random basis. Fit a Gaussian process model
+to the index values against the projection distances to obtain the
+smoothness measure.}
 
@@
 
-\includegraphics[width=6.77in,height=\textheight]{figures/squintability.png}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/squintability.png}
 
@@
 \caption{\label{fig-squintability}Illustration of steps to calculate
-squintability. For all the bases, calculate the projection distance to
-the optimal basis and the corresponding index value. Bin the index value
-by the projection distance, then claculate the average index value for
-each bin. Fit the scaled sigmoid function from Equation
+squintability. For a given projection pursuit problem defined by the
+shape to find, data dimension and the index function, sample random
+bases given the orthonormality and projection distance contraint.
+Interpolate the sampled bases to the optimal basis and calculate the
+projection distance and the index value for each interpolated basis. Bin
+the index value by the projection distance, then claculate the average
+index value for each bin. Fit the scaled sigmoid function from Equation
 \ref{eq-squintability} to the binned index values against the projection
@@
 \end{figure}%
+
+\newpage
 
@@
 (JSO success rate, hyper-parameters, and optimisation properties) into
-one table. Three pre-processing steps are applied to the data: 1)
-scaling the JSO hyper-parameters by a factor of 10 for interpretation,
-2) creating a new binary variable \texttt{long\_time} to indicate cases
-with an average run time over 30 seconds, and 3) encoding the success
-rate for the stringy case as 0, because upon visual inspection of the
-final projection, none of the 50 simulations identified the sine-wave
-shape. A generalised linear model with a binomial family and a logit
-link function is used to fit the data and Table~\ref{tbl-mod-output}
-presents the model outputs. The model suggests that JSO success rate is
-positively associated with the two hyper-parameters, as well as with the
-optimisation properties -- smoothness and squintability. Specifically,
-using 10 more jellyfishes and 10 more tries will increase the odd ratio
-of success by 24.11\% and 11.93\%, respectively. However, being flagged
-with long runtime and an increase of data dimension will reduce the
-success rate by 41.72\% and 53.36\%, respectively. The variable
-\texttt{squintability} and \texttt{dimension} are significant,
-suggesting their importance relative to JSO hyper-parameters in the
-optimisation success. Users should consider increasing the number of
-jellyfish and the number of tries to improve the success rate, until JSO
-exceeds a 30-second runtime.
+one table. Table~\ref{tbl-mod-output} presents the model outputs. The
+model suggests that JSO success rate is positively associated with the
+two hyper-parameters, as well as with the optimisation properties --
+smoothness and squintability. Specifically, using 10 more jellyfishes
+and 10 more tries will increase the odd ratio of success by 24.11\% and
+11.93\%, respectively. However, being flagged with long runtime and an
+increase of data dimension will reduce the success rate by 41.72\% and
+53.36\%, respectively. The variable \texttt{squintability} and
+\texttt{dimension} are significant, suggesting their importance relative
+to JSO hyper-parameters in the optimisation success. Users should
+consider increasing the number of jellyfish and the number of tries to
+improve the success rate, until JSO exceeds a 30-second runtime.
 

### 2024-06-28 code/07-calc-smoothness-squintability.R

@@ -27,7 +27,7 @@ holes_tidy <- smoothness_holes |>
 
 idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "stringy", "splines2d")
 smoothness_sine <- tibble::tibble(
-  n = 6, index = idx_names, data = list(sine1000),
+  n = 6, index = idx_names[6], data = list(sine1000),
   best = list(matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE))) |>
   dplyr::bind_rows(sine_8d_tbl("MIC")) |>
   dplyr::bind_rows(sine_8d_tbl("TIC")) |>

### 2024-06-29 optim.qmd
done edit section 5

@@
 
-**README**:
-
--   British English ("American or British usage is accepted, but not a
-    mixture of these")
-
--   No "we" - always third person
-
--   Check if the affiliation information is correct
-
--   word use:
-
-    -   Use jellyfish search optimiser, or JSO
-    -   "hyper-parameter" rather than "hyperparameter"
-    -   PP index rather than PPI, plural PP indexes
-    -   datasets rather than data sets
-
 ```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
-#install the latest version of ferrn from: https://github.com/huizezhang-sherry/ferrn
 library(tidyverse)
@@
 @sec-theory details JSO and introduces metrics that measure different properties of PP indexes.
-@sec-vis outlines the data structures, numerical summaries, and visualisations used in the simulation studies conducted in @sec-app to assess JSO's performance.
-@sec-app presents two sets of simulation experiments, comparing JSO permformance improvements relative to an existing optimiser, and studies the impact of different PP index properties on the optimisation performance.
+@sec-sim-deets outlines the data structures, numerical summaries, and visualisations used in the simulation studies conducted in @sec-app to assess JSO's performance.
+@sec-sim-res presents two sets of simulation experiments, comparing JSO permformance improvements relative to an existing optimiser, and studies the impact of different PP index properties on the optimisation performance.
 @sec-conclusion summarises the work and provides suggestions for future directions.
@@
 
-# Simulation details {#sec-vis}
+# Simulation details {#sec-sim-deets}
 
@@
 
-The performance of JSO is investigated in two folds: 1) compare it with a commonly used optimiser: the search-better optimiser (SBO) [@RJ-2021-105;
+The performance of JSO is investigated in two folds: 1) compare it with a commonly used optimiser: the Csreeping Random Search (CRS) [@RJ-2021-105;
 @laa_using_2020], and 2) examine its success rate under two hyper-parameters (number of jellyfishes and maximum number of tries).
 The projection pursuit problem used is finding the pipe shape using the holes index, investigated by @laa_using_2020.
-Fifty simulations are conducted with both JSO and the SBO, in four dimensions ($d = 6, 8, 10, 12$).
-JSO uses 100 jellyfishes and 100 maximum number of tries, while the SBO allows a maximum of 1000 samples at each iteration before the algorithm terminates.
+Fifty simulations are conducted with both JSO and CRS, in four dimensions ($d = 6, 8, 10, 12$).
+JSO uses 100 jellyfishes and 100 maximum number of tries, while the CRS allows a maximum of 1000 samples at each iteration before the algorithm terminates.
 The results of the simulation are collected using the data structure proposed in @RJ-2021-105 is used here for assessing JSO.
@@
 
-\newpage
+# Results \[Di and Sherry\] {#sec-sim-res}
 
-# Results \[Di and Sherry\] {#sec-app}
+The results from the two simulations described in @sec-sim-deets are analysed based on the final projections across two optimisers (JSO and CRS) and the success rate across JSO hyper-parameters.
+Smoothness and squintability are calculated across a collection of pipe-finding and sine-wave finding problems to construct the relationship between success rate, JSO hyper-parameters, and optimisation properties.
 
@@
 
-@fig-proj presents the final projections found by the two optimisers, broken down by 10th quantile, faceted by the data dimension.
+The final projections found by the two optimisers (JSO and CRS) are presented in 
+@fig-proj, broken down by 10th quantile, faceted by the data dimension.
+The JSO uses 100 jellyfishes and a maximum number of tries of 100.
+The CRS uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates.
 In the 6-dimensional data scenario, JSO consistently identifies a clear pipe shape.
-The SBO also finds the pipe shape but with a wide rim, suggesting a further polish search may be required.
-With increasing dimensions, JSO may not always identify the pipe shape due to random sampling, but it still presents a pipe shape in over 50%
-of cases. When compared to the SBO JSO reaches higher index values and clearer pipe shapes across all quantiles.
+The CRS also finds the pipe shape but with a wide rim, suggesting a further polish search may be required.
+With increasing dimensions, JSO may not always identify the pipe shape due to random sampling, but it still presents a pipe shape in over 50% of cases.
+When compared to the CRS, JSO reaches higher index values and clearer pipe shapes across all quantiles.
+This suggests JSO can be more effective when searching higher dimension spaces, which may be preferred by users.
 
-@fig-proportion presents the proportion of success for each
-hyper-parameter combinations. As the number of jellyfishes and maximum
-tries increase, the success rate also increases. For simpler problems (6
-dimensions), small parameter values (20 jellyfishes and a maximum number
-of tries of 50) can already result in high success rate, while for
-higher-dimensional problems (8, 10, and 12 dimensions), a combination of
-100 jellyfishes and 100 maximum number of tries is necessary for at
-least half of the jellyfishes to find the projection close to the optimal.
+The success rate calculated for each hyper-parameter combination is presented in @fig-proportion.
+As the number of jellyfishes and maximum tries increase, the success rate also increases. For simpler problems (6 dimensions), small parameter values (20 jellyfishes and a maximum
+of 50 tries) can achieve a high success rate.
+However, for higher-dimensional problems (8, 10, and 12 dimensions), larger parameter values (i.e.
+100 jellyfishes and a maximum of 100 tries) are necessary.
+Increasing both parameters enhances the performance of JSO, but it also extends the computational time required for the optimisation, which can be computationally intensive when evaluating the index function (such as scagnostic indexes) multiple times across numerous iterations.
 
@@
 #| fig.align = "center",
-#| fig.cap = "Projections found by the JSO and SBO at each 10th quantile across 50 simulations.
+#| fig.cap = "Projections found by the JSO and CRS at each 10th quantile across 50 simulations.
 The projection pursuit problem is to find the pipe shape using the holes index in the 6, 8, 10, and 12-dimensional spaces.
 The JSO uses 100 jellyfishes and a maximum number of tries of 100.
-The SBO uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates.
-In the 6-D data space, JSO always finds a clear pipe shape while the SBO also finds the pipe shape but with a wide rim.
-At higher data dimensions, JSO finds a higher index value and a clearer pipe shape across all the quantiles than the SBO."
+The CRS uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates.
+In the 6-D data space, JSO always finds a clear pipe shape while the CRS also finds the pipe shape but with a wide rim.
+At higher data dimensions, JSO finds a higher index value and a clearer pipe shape across all the quantiles than the CRS"
 
@@
 
-(p1 + ggtitle("a. JSO"))/(p2 + ggtitle("b. SBO"))
+(p1 + ggtitle("a. JSO"))/(p2 + ggtitle("b. CRS"))
 ```
@@
 #| fig.width = 6,
-#| fig.height = 3,
+#| fig.height = 2.5,
 #| fig.align = "center",
@@
 
-<!-- The optimisation properties, including smoothness, squintability, and -->
-<!-- speed, offer numerical metrics to characterise the complexity of -->
-<!-- projection pursuit optimisation problems. This example investigates how -->
-<!-- these metrics, along with the hyper-parameters affect the JSO -->
-<!-- success rate. Simulations are conducted to obtain the success rate -->
-<!-- across a wider array of projection pursuit problems, characterised by -->
-<!-- shape-to-find, data dimension, and the index function, as well as -->
-<!-- different hyper-parameter combinations. Smoothness and squintability are -->
-<!-- calculated for each projection pursuit problem as described in -->
-<!-- @sec-theory. A generalised linear model is used to construct the -->
-<!-- relationship between the success rate, JSO hyper-parameters and -->
-<!-- optimisation properties. -->
+```{r}
+sim_df2 <- sim_df |>   
+  mutate(n_jellies = n_jellies/10,
+         max_tries = max_tries/10,
+         long_time = ifelse(time > 20, 1, 0),
+         P_J = ifelse(index == "stringy", 0, P_J))
 
-@tbl-smoothness-squintability presents the parameters estimated from the
+mod1 <- glm(P_J ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
+```
+
+<!-- The low -->
+<!-- squint value of MIC and TIC index is due to its convex shape of the -->
+<!-- index value against the projection distance, as shown in -->
+<!-- @fig-idx-proj-dist. Comparing to other concave shapes, the maximum first -->
+<!-- derivative happens at a smaller projection distance ($\theta_2$), -->
+<!-- requiring the optimiser to get closer to the optimal to see a -->
+<!-- significant change in the index value, hence a small squintability -->
+<!-- measure.  -->
+
+The optimisation properties, including smoothness and squintability,
+offer numerical metrics to characterise the complexity of
+projection pursuit optimisation problems. @tbl-smoothness-squintability presents 
+the parameters estimated from the
 Gaussian process (variance, range, smooth, and nugget) and the scaled
-logistic function ($\theta_1$ to $\theta_4$) for each case. The column
-"smooth" is used as the smoothness measure and the column "squint" is
-calculated as equation @eq-squintability-parametric as squintability. The low
-squint value of MIC and TIC index is due to its convex shape of the
-index value against the projection distance, as shown in
-@fig-idx-proj-dist. Comparing to other concave shapes, the maximum first
-derivative happens at a smaller projection distance ($\theta_2$),
-requiring the optimiser to get closer to the optimal to see a
-significant change in the index value, hence a small squintability
-measure.
+logistic function ($\theta_1$ to $\theta_4$) for calculating both metrics for each case considered in the simulation. The column
+"smooth" is used as the smoothness metric and the column "squint" is
+calculated as equation @eq-squintability-parametric as the squintability metric.  @tbl-mod-output presents the results of fitting a generalised linear model with a binomial family and a logit link
+function to the data in @tbl-mod-data, where all the three components (success rate, JS hyper-parameters, and optimisation properties) are combined.
+The model suggests that JSO success rate is positively
+associated with the two hyper-parameters, as well as with the
+optimisation properties: smoothness and squintability. Specifically,
+using 10 more jellyfishes and 10 more tries increases the odd ratio of
+success by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`%
+and `r round(100 * (exp(broom::tidy(mod1)$estimate[7])-1), 2)`%,
+respectively. However, being flagged with long runtime and an increase
+of data dimension reduce the success rate by
+`r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and
+`r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively.
+The variable `squintability` and `dimension` are significant, suggesting
+their importance relative to JSO hyper-parameters in the optimisation
+success.
+In defining a projection pursuit problem, factors such as the shape-to-find, the index function used, and the data dimension, determine the optimisation properties like smoothness and squintability.
+These metrics can then be compared across different problems to understand their relative complexity.
+The results suggest that squintability has a more significant influence than smoothness on the success rate of JSO.
+Once the characteristics of the projection pursuit problem are fully understood, increasing JSO hyper-parameters can enhance the search effectiveness, however, it is also important to consider the resulting increase in computational complexity.
 
-```{r fig-idx-proj-dist}
+```{r fig-idx-proj-dist, eval = FALSE}
 #| fig.width = 9,
@@
   facet_nested_wrap(~index_name + n, nrow = 3, labeller = label_both) + 
-  xlab("projection distance") + ylab("index value") 
+  xlab("projection distance") + ylab("index value") +
+  theme_bw()
 ```
@@
                       "Num. of jellyfish", "Max. Num. of tries", 
-                      "Prop. of success", "Time"), 
+                      "Success rate", "Time"), 
         booktabs = T) |> 
@@
 The variable squintability and dimension are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.
-sim_df2 <- sim_df |>   
-  mutate(n_jellies = n_jellies/10,
-         max_tries = max_tries/10,
-         long_time = ifelse(time > 20, 1, 0),
-         P_J = ifelse(index == "stringy", 0, P_J))
-
-mod1 <- glm(P_J ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
 broom::tidy(mod1) |> 
@@
 
-@tbl-mod-data combines all the variables calculated above (JSO
-success rate, hyper-parameters, and optimisation properties) into
-one table. @tbl-mod-output presents the model
-outputs. The model suggests that JSO success rate is positively
-associated with the two hyper-parameters, as well as with the
-optimisation properties -- smoothness and squintability. Specifically,
-using 10 more jellyfishes and 10 more tries will increase the odd ratio of
-success by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`%
-and `r round(100 * (exp(broom::tidy(mod1)$estimate[7])-1), 2)`%,
-respectively. However, being flagged with long runtime and an increase
-of data dimension will reduce the success rate by
-`r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and
-`r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively.
-The variable `squintability` and `dimension` are significant, suggesting
-their importance relative to JSO hyper-parameters in the optimisation
-success. Users should consider increasing the number of jellyfish and the number of
-tries to improve the success rate, until JSO exceeds a 30-second
-runtime.
-
 # Conclusion \[Di and Sherry\] {#sec-conclusion}
@@
 
+Main findings:
+
+  - JSO tends to obtian a clearer projection and higher index value when searching higher dimension spaces than CRS (and other current optimisers) - Figure 4
+
+  - Increasing JS hyper-parameter enhances its performance, but the computational time required for the optimisation can quickly become intensive, especially when evaluating the index function (such as scagnostic indexes) multiple times across numerous iterations.
+- Figure 5
+
+  - Smoothness and squintability can be calculated and compared across different problems to understand their relative complexity.
+Once a projection pursuit problem is fully characterised (by the shape-to-find, the index function used, and the data dimension), increasing JSO hyper-parameters can enhance the search effectiveness, however, the resulting increase in computational complexity will also need to be considered.
+- modelling
+
 # References

### 2024-06-29 optim.tex
done edit section 5

@@
 
-\textbf{README}:
-
-\begin{itemize}
-\item
-  British English (``American or British usage is accepted, but not a
-  mixture of these'')
-\item
-  No ``we'' - always third person
-\item
-  Check if the affiliation information is correct
-\item
-  word use:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Use jellyfish search optimiser, or JSO
-  \item
-    ``hyper-parameter'' rather than ``hyperparameter''
-  \item
-    PP index rather than PPI, plural PP indexes
-  \item
-    datasets rather than data sets
-  \end{itemize}
-\end{itemize}
-
 \section{Introduction {[}Nicolas and
@@
 Section~\ref{sec-theory} details JSO and introduces metrics that measure
-different properties of PP indexes. Section~\ref{sec-vis} outlines the
-data structures, numerical summaries, and visualisations used in the
-simulation studies conducted in Section~\ref{sec-app} to assess JSO's
-performance. Section~\ref{sec-app} presents two sets of simulation
+different properties of PP indexes. Section~\ref{sec-sim-deets} outlines
+the data structures, numerical summaries, and visualisations used in the
+simulation studies conducted in \textbf{?@sec-app} to assess JSO's
+performance. Section~\ref{sec-sim-res} presents two sets of simulation
 experiments, comparing JSO permformance improvements relative to an
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, bottomrule=.15mm, breakable, left=2mm, coltitle=black, colback=white, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, titlerule=0mm, arc=.35mm, toprule=.15mm, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, breakable, opacityback=0, toptitle=1mm, leftrule=.75mm, titlerule=0mm, coltitle=black, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, rightrule=.15mm, arc=.35mm, toprule=.15mm, colback=white, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, left=2mm]
 
@@
 
-\section{Simulation details}\label{sec-vis}
+\section{Simulation details}\label{sec-sim-deets}
 
@@
 The performance of JSO is investigated in two folds: 1) compare it with
-a commonly used optimiser: the search-better optimiser (SBO)
+a commonly used optimiser: the Csreeping Random Search (CRS)
 \citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
@@
 using the holes index, investigated by \citet{laa_using_2020}. Fifty
-simulations are conducted with both JSO and the SBO, in four dimensions
+simulations are conducted with both JSO and CRS, in four dimensions
 (\(d = 6, 8, 10, 12\)). JSO uses 100 jellyfishes and 100 maximum number
-of tries, while the SBO allows a maximum of 1000 samples at each
+of tries, while the CRS allows a maximum of 1000 samples at each
 iteration before the algorithm terminates. The results of the simulation
@@
 
-\newpage
-
-\section{Results {[}Di and Sherry{]}}\label{sec-app}
+\section{Results {[}Di and Sherry{]}}\label{sec-sim-res}
+
+The results from the two simulations described in
+Section~\ref{sec-sim-deets} are analysed based on the final projections
+across two optimisers (JSO and CRS) and the success rate across JSO
+hyper-parameters. Smoothness and squintability are calculated across a
+collection of pipe-finding and sine-wave finding problems to construct
+the relationship between success rate, JSO hyper-parameters, and
+optimisation properties.
 
@@
 
-Figure~\ref{fig-proj} presents the final projections found by the two
-optimisers, broken down by 10th quantile, faceted by the data dimension.
-In the 6-dimensional data scenario, JSO consistently identifies a clear
-pipe shape. The SBO also finds the pipe shape but with a wide rim,
+The final projections found by the two optimisers (JSO and CRS) are
+presented in Figure~\ref{fig-proj}, broken down by 10th quantile,
+faceted by the data dimension. The JSO uses 100 jellyfishes and a
+maximum number of tries of 100. The CRS uses a maximum of 1000 tries in
+each step of random sampling step before the algorithm terminates. In
+the 6-dimensional data scenario, JSO consistently identifies a clear
+pipe shape. The CRS also finds the pipe shape but with a wide rim,
 suggesting a further polish search may be required. With increasing
@@
 sampling, but it still presents a pipe shape in over 50\% of cases. When
-compared to the SBO JSO reaches higher index values and clearer pipe
-shapes across all quantiles.
-
-Figure~\ref{fig-proportion} presents the proportion of success for each
-hyper-parameter combinations. As the number of jellyfishes and maximum
-tries increase, the success rate also increases. For simpler problems (6
-dimensions), small parameter values (20 jellyfishes and a maximum number
-of tries of 50) can already result in high success rate, while for
-higher-dimensional problems (8, 10, and 12 dimensions), a combination of
-100 jellyfishes and 100 maximum number of tries is necessary for at
-least half of the jellyfishes to find the projection close to the
-optimal.
+compared to the CRS, JSO reaches higher index values and clearer pipe
+shapes across all quantiles. This suggests JSO can be more effective
+when searching higher dimension spaces, which may be preferred by users.
+
+The success rate calculated for each hyper-parameter combination is
+presented in Figure~\ref{fig-proportion}. As the number of jellyfishes
+and maximum tries increase, the success rate also increases. For simpler
+problems (6 dimensions), small parameter values (20 jellyfishes and a
+maximum of 50 tries) can achieve a high success rate. However, for
+higher-dimensional problems (8, 10, and 12 dimensions), larger parameter
+values (i.e.~100 jellyfishes and a maximum of 100 tries) are necessary.
+Increasing both parameters enhances the performance of JSO, but it also
+extends the computational time required for the optimisation, which can
+be computationally intensive when evaluating the index function (such as
+scagnostic indexes) multiple times across numerous iterations.
 
@@
 
-\caption{\label{fig-proj}Projections found by the JSO and SBO at each
+\caption{\label{fig-proj}Projections found by the JSO and CRS at each
 10th quantile across 50 simulations. The projection pursuit problem is
@@
 12-dimensional spaces. The JSO uses 100 jellyfishes and a maximum number
-of tries of 100. The SBO uses a maximum of 1000 tries in each step of
+of tries of 100. The CRS uses a maximum of 1000 tries in each step of
 random sampling step before the algorithm terminates. In the 6-D data
-space, JSO always finds a clear pipe shape while the SBO also finds the
+space, JSO always finds a clear pipe shape while the CRS also finds the
 pipe shape but with a wide rim. At higher data dimensions, JSO finds a
 higher index value and a clearer pipe shape across all the quantiles
-than the SBO.}
+than the CRS}
 
@@
 
-Table~\ref{tbl-smoothness-squintability} presents the parameters
-estimated from the Gaussian process (variance, range, smooth, and
-nugget) and the scaled logistic function (\(\theta_1\) to \(\theta_4\))
-for each case. The column ``smooth'' is used as the smoothness measure
-and the column ``squint'' is calculated as equation
-\ref{eq-squintability-parametric} as squintability. The low squint value
-of MIC and TIC index is due to its convex shape of the index value
-against the projection distance, as shown in
-Figure~\ref{fig-idx-proj-dist}. Comparing to other concave shapes, the
-maximum first derivative happens at a smaller projection distance
-(\(\theta_2\)), requiring the optimiser to get closer to the optimal to
-see a significant change in the index value, hence a small squintability
-measure.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{optim_files/figure-pdf/fig-idx-proj-dist-1.pdf}
-
-}
-
-\caption{\label{fig-idx-proj-dist}Index values versus projection
-distance for the 12 pipe/sine-wave finding problem, after the binning
-procedure for calculating the squintability measure. The index values,
-averaged at bin width of 0.005, are scaled from 0 to 1 for comparison
-(holes, TIC, and stringy). The \texttt{MIC} and \texttt{TIC} index
-curves are convex while others are concave. The stringy curve shows an
-instantaneous jump to the optimum when aproaching the best basis.}
-
-\end{figure}%
+The optimisation properties, including smoothness and squintability,
+offer numerical metrics to characterise the complexity of projection
+pursuit optimisation problems. Table~\ref{tbl-smoothness-squintability}
+presents the parameters estimated from the Gaussian process (variance,
+range, smooth, and nugget) and the scaled logistic function
+(\(\theta_1\) to \(\theta_4\)) for calculating both metrics for each
+case considered in the simulation. The column ``smooth'' is used as the
+smoothness metric and the column ``squint'' is calculated as equation
+\ref{eq-squintability-parametric} as the squintability metric.
+Table~\ref{tbl-mod-output} presents the results of fitting a generalised
+linear model with a binomial family and a logit link function to the
+data in Table~\ref{tbl-mod-data}, where all the three components
+(success rate, JS hyper-parameters, and optimisation properties) are
+combined. The model suggests that JSO success rate is positively
+associated with the two hyper-parameters, as well as with the
+optimisation properties: smoothness and squintability. Specifically,
+using 10 more jellyfishes and 10 more tries increases the odd ratio of
+success by 24.11\% and 11.93\%, respectively. However, being flagged
+with long runtime and an increase of data dimension reduce the success
+rate by 41.72\% and 53.36\%, respectively. The variable
+\texttt{squintability} and \texttt{dimension} are significant,
+suggesting their importance relative to JSO hyper-parameters in the
+optimisation success. In defining a projection pursuit problem, factors
+such as the shape-to-find, the index function used, and the data
+dimension, determine the optimisation properties like smoothness and
+squintability. These metrics can then be compared across different
+problems to understand their relative complexity. The results suggest
+that squintability has a more significant influence than smoothness on
+the success rate of JSO. Once the characteristics of the projection
+pursuit problem are fully understood, increasing JSO hyper-parameters
+can enhance the search effectiveness, however, it is also important to
+consider the resulting increase in computational complexity.
 
@@
 \toprule
-Index & D & Smoothness & Squintability & Num. of jellyfish & Max. Num. of tries & Prop. of success & Time\\
+Index & D & Smoothness & Squintability & Num. of jellyfish & Max. Num. of tries & Success rate & Time\\
 \midrule
@@
 
-Table~\ref{tbl-mod-data} combines all the variables calculated above
-(JSO success rate, hyper-parameters, and optimisation properties) into
-one table. Table~\ref{tbl-mod-output} presents the model outputs. The
-model suggests that JSO success rate is positively associated with the
-two hyper-parameters, as well as with the optimisation properties --
-smoothness and squintability. Specifically, using 10 more jellyfishes
-and 10 more tries will increase the odd ratio of success by 24.11\% and
-11.93\%, respectively. However, being flagged with long runtime and an
-increase of data dimension will reduce the success rate by 41.72\% and
-53.36\%, respectively. The variable \texttt{squintability} and
-\texttt{dimension} are significant, suggesting their importance relative
-to JSO hyper-parameters in the optimisation success. Users should
-consider increasing the number of jellyfish and the number of tries to
-improve the success rate, until JSO exceeds a 30-second runtime.
-
 \section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
@@
 
+Main findings:
+
+\begin{itemize}
+\item
+  JSO tends to obtian a clearer projection and higher index value when
+  searching higher dimension spaces than CRS (and other current
+  optimisers) - Figure 4
+\item
+  Increasing JS hyper-parameter enhances its performance, but the
+  computational time required for the optimisation can quickly become
+  intensive, especially when evaluating the index function (such as
+  scagnostic indexes) multiple times across numerous iterations. -
+  Figure 5
+\item
+  Smoothness and squintability can be calculated and compared across
+  different problems to understand their relative complexity. Once a
+  projection pursuit problem is fully characterised (by the
+  shape-to-find, the index function used, and the data dimension),
+  increasing JSO hyper-parameters can enhance the search effectiveness,
+  however, the resulting increase in computational complexity will also
+  need to be considered. - modelling
+\end{itemize}
+
 \renewcommand\refname{References}

### 2024-06-30 optim.tex
compiled files

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, breakable, opacityback=0, toptitle=1mm, leftrule=.75mm, titlerule=0mm, coltitle=black, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, rightrule=.15mm, arc=.35mm, toprule=.15mm, colback=white, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, colback=white, toptitle=1mm, opacityback=0, colbacktitle=quarto-callout-note-color!10!white, breakable, bottomtitle=1mm, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, rightrule=.15mm, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, opacitybacktitle=0.6, toprule=.15mm, leftrule=.75mm, left=2mm]
 

### 2024-06-30 optim.qmd
remove space after each five rows in tables

@@
       "variance", "range", "smooth", "nugget",
-      "$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", "squint"),
+      "$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", "squint"), linesep = "",
     booktabs = T, escape = FALSE, row.names = TRUE) |> 
@@
                       "Success rate", "Time"), 
+        linesep = "", 
         booktabs = T) |> 
@@
                   "long time", "number of jellyfish", "maximum number of tries")) |>
-  kable(digits = 3) |> 
+  kable(digits = 3, linesep = "") |> 
   column_spec(c(1,5), border_right = TRUE) |> 

### 2024-06-30 optim.tex
remove space after each five rows in tables

@@
 
-\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, colback=white, toptitle=1mm, opacityback=0, colbacktitle=quarto-callout-note-color!10!white, breakable, bottomtitle=1mm, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, rightrule=.15mm, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, opacitybacktitle=0.6, toprule=.15mm, leftrule=.75mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, left=2mm, opacityback=0, toprule=.15mm, opacitybacktitle=0.6, coltitle=black, breakable, colframe=quarto-callout-note-color-frame, colback=white, arc=.35mm, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, titlerule=0mm, rightrule=.15mm, leftrule=.75mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white]
 
@@
 5 & sine & MIC & 6 & 0.016 & 0.092 & 2.457 & 0.083 & 0.894 & 0.571 & 1.623 & -0.024 & 1.257\\
-\addlinespace
 6 & sine & MIC & 8 & 0.016 & 0.081 & 2.636 & 0.084 & 0.932 & 0.328 & 1.314 & -0.030 & 0.772\\
@@
 10 & sine & loess2d & 6 & 0.079 & 0.336 & 1.988 & 0.309 & 1.016 & 1.039 & 2.648 & 0.080 & 2.756\\
-\addlinespace
 11 & sine & splines2d & 6 & 0.042 & 0.242 & 2.537 & 0.102 & 1.014 & 1.051 & 2.730 & -0.009 & 3.118\\

### 2024-06-30 optim.qmd
lots of small typos fixed and to the best of our knowledge statement removed

--- 2024-06-29 optim.qmd
@@
 It is of interest to assess how quickly and closely the optimiser reaches a gobal optima, for various PP indexes that may have differing complexities.
-To observe the performance of JSO with different types of PP indexes, metrics are introduced to capture specific properties of the index including squintability and smoothness as introduced in @laa_using_2020.
-Metrics for squintability and smoothness are precisely defined here, something that is new for the field.
+To observe the performance of JSO with different types of PP indexes, metrics are introduced to capture specific properties of the index including squintability (based on @barnett1981interpreting's squint angle) and smoothness as introduced in @laa_using_2020.
+Here we mathematically define metrics for squintability and smoothness, which is new for the field.
 A series of simulation experiments using specially structured datasets and various PP indexes are conducted to assess the behaviour of JSO under different settings and its sensitivity to hyper-parameter choices.
@@
 
-An initial investigation of PP indexes, and the potential for
-scagnostics is described in @laa_using_2020. To be useful here an
-optimiser needs to be able to handle index functions which are possibly
-not very smooth. In addition, because data structures might be
-relatively fine, the optimiser needs to be able to find maxima that
-occur with a small squint angle, that can only be seen from very close
-by. One last aspect that is useful is for an optimiser to return local
-maxima in addition to global because data can contain many different and
-interesting features.
+An initial investigation of PP indexes, and the potential for scagnostics is described in @laa_using_2020. To be useful here an optimiser needs to be able to handle index functions which are possibly
+not very smooth.
+In addition, because data structures might be relatively fine, the optimiser needs to be able to find maxima that occur with a small squint angle, that can only be seen from very close
+by. One last aspect that is useful is for an optimiser to return local maxima in addition to global because data can contain many different and interesting features.
 
@@
 
-@laa_using_2020 has proposed five criteria for assessing projection
-pursuit indexes (smoothness, squintability, flexibility, rotation
-invariance, and speed). Since not all the properties affects the
-execution of the optimisation, we choose to focus here on the first two properties, 
-namely *smoothness* (@sec-smoothness) and *squintability* (@sec-squintability), and propose two metrics to evaluate them.
+@laa_using_2020 has proposed five criteria for assessing projection pursuit indexes (smoothness, squintability, flexibility, rotation invariance, and speed). Since not all the properties affects the
+execution of the optimisation, we choose to focus here on the first two properties, namely *smoothness* (@sec-smoothness) and *squintability* (@sec-squintability), and propose two metrics to evaluate them.
 
@@
 expressed analytically when $\nu$ is a half-integer, the most popular
-values in the literature being $1/2$, $3/2$ and $5/2$ . The parameter
+values in the literature being $1/2$, $3/2$ and $5/2$. The parameter
 $\nu$, called *smoothness parameter*, controls the decay of the covariance
@@
 
-In this subsection, a metric to capture the notion of squintability is proposed, and two approaches to compute this metric numerically are detailed.
+Squintability has been a useful concept. Here, it is defined mathematically and two approaches to compute this metric numerically, are described.
 
@@
 
-It is expected that for a PP index with high squint angle, the optimization
-(@eq-optimization) should make substantial progress early on.
+It is expected that for a PP index with high squint angle, the optimization (@eq-optimization) should make substantial progress early on.
 Conversely, for a PP index with low squint angle, it might take a long
-while for the optimization to make substantial progress, as the
-candidate projections would need to be very close to the optimal one for
-the structure of the index function to be visible enough to be amenable
-to efficient optimization. This observation suggests that the extreme
-values of $f'$ (the ones for which $f''=0$, assuming that $f$ is twice
-differentiable), and the projection distances for which these values are attained,
-should play an important role in the mathematical definition of
-squintability. Noticing that $f$ is expected to be a decreasing
-function, we propose the following squintability metric:
+while for the optimization to make substantial progress, as the candidate projections would need to be very close to the optimal one for the structure of the index function to be visible enough to be amenable to efficient optimization.
+This observation suggests that the extreme values of $f'$ (the ones for which $f''=0$, assuming that $f$ is twice differentiable), and the projection distances for which these values are attained, should play an important role in the mathematical definition of squintability.
+Noticing that $f$ is expected to be a decreasing function, we propose the following squintability metric:
 
@@
 
-On the one hand, $-\min_{x}f'(x)$ represents the largest value of the
-gradient of $-f$. On the other hand, $\underset{x}{\arg\min}f'(x)$
-represents the projection distance at which this largest gradient is attained.
-Based on the above discussion, it is expected that these two values
-should be both high in the case of high squintability (fast increase in
-$f$ early on), and both low in the case of low squintability (any
-substantial increase in $f$ happens very late, close to the optimal
-angle). This suggest that their product (@eq-squintability) should
-provide a sensible measure of squintability. The multiplicative constant
-$4$, which can be deemed arbitrary, does not change the interpretation
-of the squintability metric $\varsigma$; it is here to adjust the range
-of values of $\varsigma$ and simplify the explicit formula for
-$\varsigma$ obtained later on.
+On the one hand, $-\min_{x}f'(x)$ represents the largest value of the gradient of $-f$.
+On the other hand, $\underset{x}{\arg\min}f'(x)$ represents the projection distance at which this largest gradient is attained.
+Based on the above discussion, it is expected that these two values should be both high in the case of high squintability (fast increase in $f$ early on), and both low in the case of low squintability (any substantial increase in $f$ happens very late, close to the optimal angle).
+This suggest that their product (@eq-squintability) should provide a sensible measure of squintability.
+The multiplicative constant $4$, which can be deemed arbitrary, does not change the interpretation of the squintability metric $\varsigma$; it is here to adjust the range of values of $\varsigma$ and simplify the explicit formula for $\varsigma$ obtained later on.
 
-To compute the squintability metric (@eq-squintability) in practice,
-several approaches are possible. The first one is to propose a
-parametric model for $f$, and use it to obtain an explicit formula for
-$\varsigma$. Numerical experiments suggest a scaled sigmoid shape as
+To compute the squintability metric (@eq-squintability) in practice, several approaches are possible.
+The first one is to propose a parametric model for $f$, and use it to obtain an explicit formula for $\varsigma$.
+Numerical experiments suggest a scaled sigmoid shape as
 described below. Define
@@
 
-which is a decreasing logistic function depending on two parameters
-$\theta_2$ and $\theta_3$ , such that $\ell(\theta_{2})=\frac{1}{2}$.
-Then, define
+which is a decreasing logistic function depending on two parameters $\theta_2$ and $\theta_3$, such that $\ell(\theta_{2})=\frac{1}{2}$. Then, define
 
@@
 
-which depends on three additional parameters, $\theta_1$, $\theta_2$ ,
-and $x_{\max}$, such that $f(0)=\theta_1$ and $f(x_{\max})=\theta_4$.
-Under the parametric model (@eq-parametric), the squintability metric
-(@eq-squintability) can be shown to be equal to
+which depends on three additional parameters, $\theta_1$, $\theta_2$, and $x_{\max}$, such that $f(0)=\theta_1$ and $f(x_{\max})=\theta_4$.
+Under the parametric model (@eq-parametric), the squintability metric (@eq-squintability) can be shown to be equal to
 
@@
 
-In practice, the parameters of this model (@eq-parametric) can be
-estimated numerically, for example by non-linear least squares, and then
+In practice, the parameters of this model (@eq-parametric) can be estimated numerically, for example by non-linear least squares, and then
 used to evaluate $\varsigma$ as in equation
@@
 
-Alternatively, one can estimate (@eq-squintability) in a nonparametric
-way, for example by fitting $f$ using kernel regression, then
-numerically estimate the angle at which $-f'$ attains its highest value.
+Alternatively, one can estimate (@eq-squintability) in a nonparametric way, for example by fitting $f$ using kernel regression, then numerically estimate the angle at which $-f'$ attains its highest value.
 
@@
 
-To the best of our knowledge, this is the first attempt to measure mathematically the
-notion of squintability.
+<!--To the best of our knowledge, -->
 
@@
   knitr::kable(
-    digits = 3, format = "latex", 
+    digits = 2, format = "latex", 
     col.names = c(
@@
   column_spec(c(4, 8, 13), border_right = TRUE) |> 
-  kable_styling(font_size = 7) 
+  kable_styling(font_size = 10) 
 ```
@@
 sim_df |> 
-  dplyr::select(index, d, smoothness, squintability, n_jellies, max_tries, 
+  dplyr::select(index, d, smoothness, 
+    squintability, n_jellies, max_tries, 
     -I_max, P_J, time) |> 
+  mutate(time = as.numeric(time)) |>
   arrange(index, d, n_jellies, max_tries) |> 
   head(7) |> 
-  kable(digits = 3,
-        col.names = c("Index", "D", "Smoothness", "Squintability", 
-                      "Num. of jellyfish", "Max. Num. of tries", 
-                      "Success rate", "Time"), 
+  kable(digits = 2, format = "latex", 
+        col.names = c("Index", "D", 
+          "Smoothness", "Squintability", 
+          "N. jellyfish", 
+          "Max. tries", 
+          "Success rate", "Time (sec)"),
+        linesep = "",
         booktabs = T) |> 
-  kable_styling(latex_options = "hold_position", font_size = 7)
+  kable_styling(#latex_options = "hold_position",
+    font_size = 10)
 ```
@@
 broom::tidy(mod1) |> 
-  mutate(term = c("Intercept", "Smoothness", "Squintability", "dimension (d)", 
-                  "long time", "number of jellyfish", "maximum number of tries")) |>
-  kable(digits = 3) |> 
+  mutate(term = c("Intercept", "Smoothness",
+   "Squintability", "Dimension (d)", 
+   "Long time", "N. jellyfish", 
+   "Max. tries")) |>
+  kable(digits = 2, format = "latex", 
+    linesep = "", booktabs = T) |> 
   column_spec(c(1,5), border_right = TRUE) |> 
   column_spec(1, border_left = TRUE) |> 
-  kable_styling(font_size = 7) 
+  kable_styling(font_size = 10) 
 ```

### 2024-06-30 optim.tex
lots of small typos fixed and to the best of our knowledge statement removed

@@
 metrics are introduced to capture specific properties of the index
-including squintability and smoothness as introduced in
-\citet{laa_using_2020}. Metrics for squintability and smoothness are
-precisely defined here, something that is new for the field. A series of
-simulation experiments using specially structured datasets and various
-PP indexes are conducted to assess the behaviour of JSO under different
-settings and its sensitivity to hyper-parameter choices. The
-relationship between the JSO performance, hyper-parameter tuning and
-properties of PP indexes is analysed to provide guidance for
-practitioners using projection pursuit, and developers of new indexes.
+including squintability (based on \citet{barnett1981interpreting}'s
+squint angle) and smoothness as introduced in \citet{laa_using_2020}.
+Here we mathematically define metrics for squintability and smoothness,
+which is new for the field. A series of simulation experiments using
+specially structured datasets and various PP indexes are conducted to
+assess the behaviour of JSO under different settings and its sensitivity
+to hyper-parameter choices. The relationship between the JSO
+performance, hyper-parameter tuning and properties of PP indexes is
+analysed to provide guidance for practitioners using projection pursuit,
+and developers of new indexes.
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, left=2mm, opacityback=0, toprule=.15mm, opacitybacktitle=0.6, coltitle=black, breakable, colframe=quarto-callout-note-color-frame, colback=white, arc=.35mm, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, titlerule=0mm, rightrule=.15mm, leftrule=.75mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white]
+\begin{tcolorbox}[enhanced jigsaw, left=2mm, colback=white, opacityback=0, leftrule=.75mm, bottomtitle=1mm, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm, coltitle=black, breakable, bottomrule=.15mm, titlerule=0mm, toprule=.15mm, opacitybacktitle=0.6]
 
@@
 half-integer, the most popular values in the literature being \(1/2\),
-\(3/2\) and \(5/2\) . The parameter \(\nu\), called \emph{smoothness
+\(3/2\) and \(5/2\). The parameter \(\nu\), called \emph{smoothness
 parameter}, controls the decay of the covariance function. As such, it
@@
 
-In this subsection, a metric to capture the notion of squintability is
-proposed, and two approaches to compute this metric numerically are
-detailed.
+Squintability has been a useful concept. Here, it is defined
+mathematically and two approaches to compute this metric numerically,
+are described.
 
@@
 which is a decreasing logistic function depending on two parameters
-\(\theta_2\) and \(\theta_3\) , such that
+\(\theta_2\) and \(\theta_3\), such that
 \(\ell(\theta_{2})=\frac{1}{2}\). Then, define
@@
 
-which depends on three additional parameters, \(\theta_1\), \(\theta_2\)
-, and \(x_{\max}\), such that \(f(0)=\theta_1\) and
+which depends on three additional parameters, \(\theta_1\),
+\(\theta_2\), and \(x_{\max}\), such that \(f(0)=\theta_1\) and
 \(f(x_{\max})=\theta_4\). Under the parametric model
@@
 value.
-
-To the best of our knowledge, this is the first attempt to measure
-mathematically the notion of squintability.
 
@@
 
-\centering\begingroup\fontsize{7}{9}\selectfont
+\centering\begingroup\fontsize{10}{12}\selectfont
 
@@
 \midrule
-1 & pipe & holes & 6 & 0.002 & 0.371 & 2.355 & 0.221 & 1.015 & 0.860 & 3.368 & 0.018 & 3.053\\
-2 & pipe & holes & 8 & 0.000 & 0.178 & 2.189 & 0.820 & 1.014 & 0.869 & 3.264 & 0.029 & 2.960\\
-3 & pipe & holes & 10 & 0.000 & 0.107 & 2.192 & 3.027 & 1.016 & 0.885 & 3.151 & 0.022 & 2.948\\
-4 & pipe & holes & 12 & 0.000 & 0.148 & 2.295 & 1.581 & 1.011 & 0.878 & 3.345 & 0.004 & 3.117\\
-5 & sine & MIC & 6 & 0.016 & 0.092 & 2.457 & 0.083 & 0.894 & 0.571 & 1.623 & -0.024 & 1.257\\
-6 & sine & MIC & 8 & 0.016 & 0.081 & 2.636 & 0.084 & 0.932 & 0.328 & 1.314 & -0.030 & 0.772\\
-7 & sine & TIC & 6 & 0.124 & 0.110 & 2.444 & 0.086 & 0.951 & 0.536 & 1.719 & -0.027 & 1.322\\
-8 & sine & TIC & 8 & 0.125 & 0.099 & 2.475 & 0.085 & 0.949 & 0.564 & 1.723 & -0.028 & 1.369\\
-9 & sine & dcor2d & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019 & 2.947\\
-10 & sine & loess2d & 6 & 0.079 & 0.336 & 1.988 & 0.309 & 1.016 & 1.039 & 2.648 & 0.080 & 2.756\\
-11 & sine & splines2d & 6 & 0.042 & 0.242 & 2.537 & 0.102 & 1.014 & 1.051 & 2.730 & -0.009 & 3.118\\
-12 & sine & stringy & 6 & 0.000 & 0.007 & 1.536 & 38.166 & 1.045 & 0.011 & 254.734 & 0.053 & 2.957\\
+1 & pipe & holes & 6 & 0.00 & 0.37 & 2.36 & 0.22 & 1.02 & 0.86 & 3.37 & 0.02 & 3.05\\
+2 & pipe & holes & 8 & 0.00 & 0.18 & 2.19 & 0.82 & 1.01 & 0.87 & 3.26 & 0.03 & 2.96\\
+3 & pipe & holes & 10 & 0.00 & 0.11 & 2.19 & 3.03 & 1.02 & 0.88 & 3.15 & 0.02 & 2.95\\
+4 & pipe & holes & 12 & 0.00 & 0.15 & 2.29 & 1.58 & 1.01 & 0.88 & 3.34 & 0.00 & 3.12\\
+5 & sine & MIC & 6 & 0.02 & 0.09 & 2.46 & 0.08 & 0.89 & 0.57 & 1.62 & -0.02 & 1.26\\
+6 & sine & MIC & 8 & 0.02 & 0.08 & 2.64 & 0.08 & 0.93 & 0.33 & 1.31 & -0.03 & 0.77\\
+7 & sine & TIC & 6 & 0.12 & 0.11 & 2.44 & 0.09 & 0.95 & 0.54 & 1.72 & -0.03 & 1.32\\
+8 & sine & TIC & 8 & 0.12 & 0.10 & 2.47 & 0.09 & 0.95 & 0.56 & 1.72 & -0.03 & 1.37\\
+9 & sine & dcor2d & 6 & 0.03 & 0.17 & 2.66 & 0.11 & 0.95 & 1.04 & 2.74 & -0.02 & 2.95\\
+10 & sine & loess2d & 6 & 0.08 & 0.34 & 1.99 & 0.31 & 1.02 & 1.04 & 2.65 & 0.08 & 2.76\\
+11 & sine & splines2d & 6 & 0.04 & 0.24 & 2.54 & 0.10 & 1.01 & 1.05 & 2.73 & -0.01 & 3.12\\
+12 & sine & stringy & 6 & 0.00 & 0.01 & 1.54 & 38.17 & 1.05 & 0.01 & 254.73 & 0.05 & 2.96\\
 \bottomrule
@@
 
-\begingroup\fontsize{7}{9}\selectfont
-
-\begin{longtable}[t]{lrrrrrrl}
+\begin{table}
 
@@
 
-\tabularnewline
-
+\centering{
+
+\centering\begingroup\fontsize{10}{12}\selectfont
+
+\begin{tabular}{lrrrrrrr}
 \toprule
-Index & D & Smoothness & Squintability & Num. of jellyfish & Max. Num. of tries & Success rate & Time\\
+Index & D & Smoothness & Squintability & N. jellyfish & Max. tries & Success rate & Time (sec)\\
 \midrule
-MIC & 6 & 2.457 & 1.257 & 20 & 50 & 0.12 & 2.479 secs\\
-MIC & 6 & 2.457 & 1.257 & 20 & 100 & 0.24 & 8.950 secs\\
-MIC & 6 & 2.457 & 1.257 & 50 & 50 & 0.52 & 5.651 secs\\
-MIC & 6 & 2.457 & 1.257 & 50 & 100 & 0.64 & 13.223 secs\\
-MIC & 6 & 2.457 & 1.257 & 100 & 50 & 0.76 & 19.453 secs\\
-\addlinespace
-MIC & 8 & 2.636 & 0.772 & 20 & 50 & 0.08 & 2.566 secs\\
-MIC & 8 & 2.636 & 0.772 & 20 & 100 & 0.08 & 4.960 secs\\
+MIC & 6 & 2.46 & 1.26 & 20 & 50 & 0.12 & 2.48\\
+MIC & 6 & 2.46 & 1.26 & 20 & 100 & 0.24 & 8.95\\
+MIC & 6 & 2.46 & 1.26 & 50 & 50 & 0.52 & 5.65\\
+MIC & 6 & 2.46 & 1.26 & 50 & 100 & 0.64 & 13.22\\
+MIC & 6 & 2.46 & 1.26 & 100 & 50 & 0.76 & 19.45\\
+MIC & 8 & 2.64 & 0.77 & 20 & 50 & 0.08 & 2.57\\
+MIC & 8 & 2.64 & 0.77 & 20 & 100 & 0.08 & 4.96\\
 \bottomrule
-
-\end{longtable}
-
+\end{tabular}
 \endgroup{}
 
-\begingroup\fontsize{7}{9}\selectfont
-
-\begin{longtable}[t]{|>{}lrrr>{}r|}
+}
+
+\end{table}%
+
+\begin{table}
 
@@
 
-\tabularnewline
-
+\centering{
+
+\centering\begingroup\fontsize{10}{12}\selectfont
+
+\begin{tabular}{|>{}lrrr>{}r|}
 \toprule
@@
 \midrule
-Intercept & -4.523 & 5.332 & -0.848 & 0.396\\
-Smoothness & 1.195 & 1.915 & 0.624 & 0.533\\
-Squintability & 2.060 & 0.684 & 3.011 & 0.003\\
-dimension (d) & -0.628 & 0.255 & -2.462 & 0.014\\
-long time & -0.874 & 1.286 & -0.680 & 0.497\\
-\addlinespace
-number of jellyfish & 0.216 & 0.127 & 1.701 & 0.089\\
-maximum number of tries & 0.113 & 0.150 & 0.752 & 0.452\\
+Intercept & -4.52 & 5.33 & -0.85 & 0.40\\
+Smoothness & 1.19 & 1.92 & 0.62 & 0.53\\
+Squintability & 2.06 & 0.68 & 3.01 & 0.00\\
+Dimension (d) & -0.63 & 0.26 & -2.46 & 0.01\\
+Long time & -0.87 & 1.29 & -0.68 & 0.50\\
+N. jellyfish & 0.22 & 0.13 & 1.70 & 0.09\\
+Max. tries & 0.11 & 0.15 & 0.75 & 0.45\\
 \bottomrule
-
-\end{longtable}
-
+\end{tabular}
 \endgroup{}
+
+}
+
+\end{table}%
 

### 2024-06-30 optim.qmd
conclusions drafted

@@
 Here we mathematically define metrics for squintability and smoothness, which is new for the field.
-A series of simulation experiments using specially structured datasets and various PP indexes are conducted to assess the behaviour of JSO under different settings and its sensitivity to hyper-parameter choices.
-The relationship between the JSO performance, hyper-parameter tuning and properties of PP
-indexes is analysed to provide guidance for practitioners using projection pursuit, and developers of new indexes.
+A series of simulation experiments using specially structured datasets and various PP indexes are conducted to assess the behaviour of JSO under different settings and its sensitivity to hyper-parameter choices (number of jellies and maximum number of tries).
+The relationship between the JSO performance, hyper-parameter tuning and properties of PP indexes is analysed to provide guidance for practitioners using projection pursuit, and developers of new indexes.
 
@@
 the index function, as done at the initialization stage of the jellyfish search optimization. There exist several R packages, such as GpGp [@guinness2021gpgp] or ExaGeoStatR [@abdulah2023large], to
-fit the hyperparameters of a GP covariance function on data. In this
-project, we make use of the GpGp package.
+fit the hyperparameters of a GP covariance function on data. In this project, we make use of the GpGp package.
 
@@
 
-JSO provides a new swarm-based search strategy for optimising
-projection pursuit problems. This section describes the setups for two simulations
-to demonstrate JSO's performance. The first simulation compares JSO's 
-performance with an exist optimiser, Creeping Random Search (CRS), and explores how JSO behaves
-with different hyper-parameters combinations. The second simulation
-studies the effect of optimisation properties (smoothness and squintability),
-along with JSO hyper-parameters, on the success rate of the
-optimisation.
+The JSO performance is compared with the existing Creeping Random Search (CRS) currently used in guided tour, and explores how JSO behaves with different hyper-parameters combinations.
+A second simulation
+studies the effect of index properties (smoothness and squintability), along with JSO hyper-parameters, on the success rate of the JSO performance.
 
@@
 In this simulation, 50 jellyfishes (`n_jellies`) and
-a maximum of 50 tries (`max_tries`) are used. The full data contains 50
-simulations, indexed by the variable `sim`, with the associated `seed`s. 
-Recorded in columns `basis`, `index_val`, `tries`, `loop` and
-`time`, are the sampled projection basis, index value,
-iteration ID, jellyfish ID, and computational time
-respectively.
+a maximum of 50 tries (`max_tries`) are used.
+The full data contains 50 simulations, indexed by the variable `sim`, with the associated `seed`s.
+Recorded in columns `basis`, `index_val`, `tries`, `loop` and `time`, are the sampled projection basis, index value, iteration ID, jellyfish ID, and computational time, respectively.
 
@@
 
-In the second experiment, fifty simulations are conducted at each
-hyper-parameter combination below to analyse its effects on the JSO success
-rate: number of jellyfish: 20, 50, and 100, and  maximum number of 
-tries: 50 and 100). The success 
-rate is defined as the proportion of
-simulations that achieves a final index value within 0.05 of the best
-index value found among all 50 simulations. @fig-success-rate illustrates 
-the calculation based on the projection found by the above simulation. Here
-the optimal index value is
-`r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`.
-Out of the 50 simulations,
-`r sim_sine_6d_spline_best |> filter(index_val >= aa - 0.05) |> nrow() -> bb; bb`
-achieved an index value within \[`r aa - 0.05`, `r aa`\], resulting
-in a success rate of
-`r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.
+In the second experiment, fifty simulations are conducted at each hyper-parameter combination below to analyse its effects on the JSO success rate: number of jellyfish: 20, 50, and 100, and  maximum number of  tries: 50 and 100).
+The success rate is defined as the proportion of simulations that achieves a final index value within 0.05 of the best index value found among all 50 simulations.
+@fig-success-rate illustrates
+the calculation based on the projection found by the above simulation. Here, the optimal index value is
+`r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`. Out of the 50 simulations,
+`r sim_sine_6d_spline_best |> filter(index_val >= aa - 0.05) |> nrow() -> bb; bb` achieved an index value within \[`r aa - 0.05`, `r aa`\], resulting in a success rate of `r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.
 
@@
 
-## Factors affecting JSO success rate: optimisation properties and jellyfish hyper-parameters {#sec-app-2}
+## Factors affecting JSO success rate: index properties and jellyfish hyper-parameters {#sec-app-2}
 
-In addition to the pipe-finding problem, a new shape, sine wave, is
-investigated in 6D and 8D spaces with six indexes considered:
-`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`. Combining
-with two JSO hyper-parameters, a total of `r nrow(sim_df)` cases
-is produced, comprising of `r nrow(filter(sim_df, index == "holes"))`
-pipe-finding cases and `r nrow(filter(sim_df, index != "holes"))`
-sine-wave finding cases. For each case, JSO is run fifty times and the
-success rate is calculated.
+In addition to the pipe-finding problem, a new shape, sine wave, is investigated in 6D and 8D spaces with six indexes considered:
+`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`.
+Combining with two JSO hyper-parameters, a total of `r nrow(sim_df)` cases is produced, comprising of `r nrow(filter(sim_df, index == "holes"))` pipe-finding cases and `r nrow(filter(sim_df, index != "holes"))` sine-wave finding cases.
+For each case, JSO is run fifty times and the success rate is calculated.
 
-Smoothness and squintability are computed for each case, following the
-procedures outlined in @sec-smoothness and @sec-squintability and illustrated in @fig-smoothness and  @fig-squintability. For smoothness, three hundred random bases are simulated. Index
-values and projection distance (to the optimal basis) are calculated for
-each random basis before fitting into a Gaussian process model to obtain 
-the smoothness measure. For
-squintability, fifty random bases are sampled and interpolated to the
-optimal basis with a step size of 0.005. Index values and projection 
-distances are calculated for these interpolated bases and the index values 
-are averaged with a bin width of 0.005. The squintability
-measure is then calculated from fitting a 4-parameter scaled logistic
-function of index value against projection distance, estimated by
-non-linear least squares.
+Smoothness and squintability are computed for each case, following the procedures outlined in @sec-smoothness and @sec-squintability and illustrated in @fig-smoothness and  @fig-squintability.
+For smoothness, three hundred random bases are simulated.
+Index values and projection distance (to the optimal basis) are calculated for each random basis before fitting into a Gaussian process model to obtain
+the smoothness measure. For squintability, fifty random bases are sampled and interpolated to the optimal basis with a step size of 0.005. Index values and projection 
+distances are calculated for these interpolated bases and the index values are averaged with a bin width of 0.005.
+The squintability measure is then calculated from fitting a 4-parameter scaled logistic
+function of index value against projection distance, estimated by non-linear least squares.
 
-Combining with the success rate calculated for each case, the dataset is fitted using 
-a generalised linear model with a binomial family and a logit link
-function with three pre-processing steps are applied: 1) scaling
-the JSO hyper-parameters by a factor of 10 for interpretation, 2)
-creating a new binary variable `long_time` to indicate cases with an
-average run time over 30 seconds, and 3) encoding the success rate for
-the stringy case as 0, because upon visual inspection of the final
-projection, none of the 50 simulations identified the sine-wave shape.
+Combining with the success rate calculated for each case, the dataset is fitted using  a generalised linear model with a binomial family and a logit link
+function with three pre-processing steps are applied: 
+1) scaling the JSO hyper-parameters by a factor of 10 for interpretation, 2) creating a new binary variable `long_time` to indicate cases with an average run time over 30 seconds, and 3) encoding the success rate for
+the stringy case as 0, because upon visual inspection of the final projection, none of the 50 simulations identified the sine-wave shape.
 
@@
 The results from the two simulations described in @sec-sim-deets are analysed based on the final projections across two optimisers (JSO and CRS) and the success rate across JSO hyper-parameters.
-Smoothness and squintability are calculated across a collection of pipe-finding and sine-wave finding problems to construct the relationship between success rate, JSO hyper-parameters, and optimisation properties.
+Smoothness and squintability are calculated across a collection of pipe-finding and sine-wave finding problems to construct the relationship between success rate, JSO hyper-parameters, and index properties.
 
@@
 
-The final projections found by the two optimisers (JSO and CRS) are presented in 
-@fig-proj, broken down by 10th quantile, faceted by the data dimension.
+The final projections found by the two optimisers (JSO and CRS) are presented in @fig-proj, broken down by 10th quantile, faceted by the data dimension.
 The JSO uses 100 jellyfishes and a maximum number of tries of 100.
@@
 The success rate calculated for each hyper-parameter combination is presented in @fig-proportion.
-As the number of jellyfishes and maximum tries increase, the success rate also increases. For simpler problems (6 dimensions), small parameter values (20 jellyfishes and a maximum
-of 50 tries) can achieve a high success rate.
+As the number of jellyfishes and maximum tries increase, the success rate also increases.
+For simpler problems (6 dimensions), small parameter values (20 jellyfishes and a maximum of 50 tries) can achieve a high success rate.
 However, for higher-dimensional problems (8, 10, and 12 dimensions), larger parameter values (i.e.
@@
 
-## Factors affecting JSO success rate: optimisation properties and jellyfish hyper-parameters {#sec-app-2}
+## Factors affecting JSO success rate: index properties and jellyfish hyper-parameters {#sec-app-2}
 
@@
 
-The optimisation properties, including smoothness and squintability,
-offer numerical metrics to characterise the complexity of
-projection pursuit optimisation problems. @tbl-smoothness-squintability presents 
-the parameters estimated from the
-Gaussian process (variance, range, smooth, and nugget) and the scaled
+The index properties, including smoothness and squintability, offer numerical metrics to characterise the complexity of projection pursuit optimisation problems.
+@tbl-smoothness-squintability presents
+the parameters estimated from the Gaussian process (variance, range, smooth, and nugget) and the scaled
 logistic function ($\theta_1$ to $\theta_4$) for calculating both metrics for each case considered in the simulation. The column
@@
 calculated as equation @eq-squintability-parametric as the squintability metric.  @tbl-mod-output presents the results of fitting a generalised linear model with a binomial family and a logit link
-function to the data in @tbl-mod-data, where all the three components (success rate, JS hyper-parameters, and optimisation properties) are combined.
-The model suggests that JSO success rate is positively
-associated with the two hyper-parameters, as well as with the
-optimisation properties: smoothness and squintability. Specifically,
-using 10 more jellyfishes and 10 more tries increases the odd ratio of
-success by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`%
+function to the data in @tbl-mod-data, where all the three components (success rate, JS hyper-parameters, and index properties) are combined.
+The model suggests that JSO success rate is positively associated with the two hyper-parameters, as well as with the index properties: smoothness and squintability.
+Specifically,
+using 10 more jellyfishes and 10 more tries increases the odd ratio of success by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`%
 and `r round(100 * (exp(broom::tidy(mod1)$estimate[7])-1), 2)`%,
-respectively. However, being flagged with long runtime and an increase
-of data dimension reduce the success rate by
-`r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and
-`r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively.
-The variable `squintability` and `dimension` are significant, suggesting
-their importance relative to JSO hyper-parameters in the optimisation
-success.
-In defining a projection pursuit problem, factors such as the shape-to-find, the index function used, and the data dimension, determine the optimisation properties like smoothness and squintability.
+respectively.
+However, being flagged with long runtime and an increase of data dimension reduce the success rate by `r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and `r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively.
+The variable `squintability` and `dimension` are significant, suggesting their importance relative to JSO hyper-parameters in the optimisation success.
+In defining a projection pursuit problem, factors such as the shape-to-find, the index function used, and the data dimension, determine the properties like smoothness and squintability.
 These metrics can then be compared across different problems to understand their relative complexity.
@@
 #| label: tbl-mod-output
-#| tbl-cap: Model estimates of proportion of jellyfish success on optimisation properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function.
+#| tbl-cap: Model estimates of proportion of jellyfish success on index properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function.
 The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with JSO success rate while data dimension and being flagged as long runtime are negatively associated with the success rate.
@@
 
+This paper has presented new metrics to mathematically define desirable features of PP indexes, squintability and smoothness, and used these to assess the performance of the new jellyfish search optimiser.
+The metrics will be generally useful for characterising PP indexes, and halp with developing nex indexes.
+
+In the comparison of the JSO against the currently used CRS, as expected the JSO vastly out-performs CRS, and provides a high probability of finding the global optima.
+The JSO obtains the maximum more cleanly, with a slightly higher index value, and plot of the projected data showing the structure more clearly.
+
+The JSO performance is affected by the hyper-parameters, with a higher chance of reaching the global optima when more jellies are used and the maximum number of tries is increased.
+However, it comes at a computational cost, as expected.
+The performance declines if the projection dimension increases and if the PP index has low squintability.
+The higher the squintability the better chance the JSO can find the optima.
+However, interestingly smoothness does not affect the JSO performance.
+
+The new JSO is integrated with the current implementation of the projection pursuit guided tour in the `tourr` package, and can be further examined using PP optimisation diagnostics in the `ferrn` package.
+Using the JSO is a little different than the current CRS, because it runs many paths.
+The recommended approach is to conduct the optimisation off-line, and then use the guided tour to follow selected jellies.
+The tools for this are available in the `tourr` package, too.
+
+<!--
 - Summarise the result of the simulation: JSO immense improvement on current methods, ...
@@
 - modelling
+-->
 

### 2024-06-30 optim.tex
conclusions drafted

@@
 assess the behaviour of JSO under different settings and its sensitivity
-to hyper-parameter choices. The relationship between the JSO
-performance, hyper-parameter tuning and properties of PP indexes is
-analysed to provide guidance for practitioners using projection pursuit,
-and developers of new indexes.
+to hyper-parameter choices (number of jellies and maximum number of
+tries). The relationship between the JSO performance, hyper-parameter
+tuning and properties of PP indexes is analysed to provide guidance for
+practitioners using projection pursuit, and developers of new indexes.
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, left=2mm, colback=white, opacityback=0, leftrule=.75mm, bottomtitle=1mm, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm, coltitle=black, breakable, bottomrule=.15mm, titlerule=0mm, toprule=.15mm, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, bottomtitle=1mm, bottomrule=.15mm, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, leftrule=.75mm, coltitle=black, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, opacityback=0, colback=white, left=2mm, toprule=.15mm, arc=.35mm, colframe=quarto-callout-note-color-frame]
 
@@
 
-JSO provides a new swarm-based search strategy for optimising projection
-pursuit problems. This section describes the setups for two simulations
-to demonstrate JSO's performance. The first simulation compares JSO's
-performance with an exist optimiser, Creeping Random Search (CRS), and
-explores how JSO behaves with different hyper-parameters combinations.
-The second simulation studies the effect of optimisation properties
-(smoothness and squintability), along with JSO hyper-parameters, on the
-success rate of the optimisation.
+The JSO performance is compared with the existing Creeping Random Search
+(CRS) currently used in guided tour, and explores how JSO behaves with
+different hyper-parameters combinations. A second simulation studies the
+effect of index properties (smoothness and squintability), along with
+JSO hyper-parameters, on the success rate of the JSO performance.
 
@@
 projection basis, index value, iteration ID, jellyfish ID, and
-computational time respectively.
+computational time, respectively.
 
@@
 Figure~\ref{fig-success-rate} illustrates the calculation based on the
-projection found by the above simulation. Here the optimal index value
+projection found by the above simulation. Here, the optimal index value
 is 0.999. Out of the 50 simulations, 44 achieved an index value within
@@
 
-\subsection{Factors affecting JSO success rate: optimisation properties
-and jellyfish hyper-parameters}\label{sec-app-2}
+\subsection{Factors affecting JSO success rate: index properties and
+jellyfish hyper-parameters}\label{sec-app-2}
 
@@
 collection of pipe-finding and sine-wave finding problems to construct
-the relationship between success rate, JSO hyper-parameters, and
-optimisation properties.
+the relationship between success rate, JSO hyper-parameters, and index
+properties.
 
@@
 
-\subsection{Factors affecting JSO success rate: optimisation properties
-and jellyfish hyper-parameters}\label{sec-app-2}
-
-The optimisation properties, including smoothness and squintability,
-offer numerical metrics to characterise the complexity of projection
-pursuit optimisation problems. Table~\ref{tbl-smoothness-squintability}
-presents the parameters estimated from the Gaussian process (variance,
-range, smooth, and nugget) and the scaled logistic function
-(\(\theta_1\) to \(\theta_4\)) for calculating both metrics for each
-case considered in the simulation. The column ``smooth'' is used as the
-smoothness metric and the column ``squint'' is calculated as equation
+\subsection{Factors affecting JSO success rate: index properties and
+jellyfish hyper-parameters}\label{sec-app-2}
+
+The index properties, including smoothness and squintability, offer
+numerical metrics to characterise the complexity of projection pursuit
+optimisation problems. Table~\ref{tbl-smoothness-squintability} presents
+the parameters estimated from the Gaussian process (variance, range,
+smooth, and nugget) and the scaled logistic function (\(\theta_1\) to
+\(\theta_4\)) for calculating both metrics for each case considered in
+the simulation. The column ``smooth'' is used as the smoothness metric
+and the column ``squint'' is calculated as equation
 \ref{eq-squintability-parametric} as the squintability metric.
@@
 data in Table~\ref{tbl-mod-data}, where all the three components
-(success rate, JS hyper-parameters, and optimisation properties) are
-combined. The model suggests that JSO success rate is positively
-associated with the two hyper-parameters, as well as with the
-optimisation properties: smoothness and squintability. Specifically,
-using 10 more jellyfishes and 10 more tries increases the odd ratio of
-success by 24.11\% and 11.93\%, respectively. However, being flagged
-with long runtime and an increase of data dimension reduce the success
-rate by 41.72\% and 53.36\%, respectively. The variable
-\texttt{squintability} and \texttt{dimension} are significant,
-suggesting their importance relative to JSO hyper-parameters in the
-optimisation success. In defining a projection pursuit problem, factors
-such as the shape-to-find, the index function used, and the data
-dimension, determine the optimisation properties like smoothness and
-squintability. These metrics can then be compared across different
-problems to understand their relative complexity. The results suggest
-that squintability has a more significant influence than smoothness on
-the success rate of JSO. Once the characteristics of the projection
-pursuit problem are fully understood, increasing JSO hyper-parameters
-can enhance the search effectiveness, however, it is also important to
-consider the resulting increase in computational complexity.
+(success rate, JS hyper-parameters, and index properties) are combined.
+The model suggests that JSO success rate is positively associated with
+the two hyper-parameters, as well as with the index properties:
+smoothness and squintability. Specifically, using 10 more jellyfishes
+and 10 more tries increases the odd ratio of success by 24.11\% and
+11.93\%, respectively. However, being flagged with long runtime and an
+increase of data dimension reduce the success rate by 41.72\% and
+53.36\%, respectively. The variable \texttt{squintability} and
+\texttt{dimension} are significant, suggesting their importance relative
+to JSO hyper-parameters in the optimisation success. In defining a
+projection pursuit problem, factors such as the shape-to-find, the index
+function used, and the data dimension, determine the properties like
+smoothness and squintability. These metrics can then be compared across
+different problems to understand their relative complexity. The results
+suggest that squintability has a more significant influence than
+smoothness on the success rate of JSO. Once the characteristics of the
+projection pursuit problem are fully understood, increasing JSO
+hyper-parameters can enhance the search effectiveness, however, it is
+also important to consider the resulting increase in computational
+complexity.
 
@@
 \caption{\label{tbl-mod-output}Model estimates of proportion of
-jellyfish success on optimisation properties and jellyfish
-hyper-parameters from the generalised linear model with a binomial
-family and a logit link function. The variable smoothness,
-squintability, number of jellyfish and maximum number of tries are
-positively associated with JSO success rate while data dimension and
-being flagged as long runtime are negatively associated with the success
-rate. The variable squintability and dimension are significant,
-suggesting their importance relative to jellyfish hyper-parameters in
-the optimisation success.}
+jellyfish success on index properties and jellyfish hyper-parameters
+from the generalised linear model with a binomial family and a logit
+link function. The variable smoothness, squintability, number of
+jellyfish and maximum number of tries are positively associated with JSO
+success rate while data dimension and being flagged as long runtime are
+negatively associated with the success rate. The variable squintability
+and dimension are significant, suggesting their importance relative to
+jellyfish hyper-parameters in the optimisation success.}
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Summarise the result of the simulation: JSO immense improvement on
-  current methods, \ldots{}
-\item
-  JSO for a guided tour, what changes are needed? Do we follow one
-  tentacle, do we perform off-line and then pull tentacles out to watch,
-  how does it integrate with ferrn
-\end{itemize}
-
-Main findings:
-
-\begin{itemize}
-\item
-  JSO tends to obtian a clearer projection and higher index value when
-  searching higher dimension spaces than CRS (and other current
-  optimisers) - Figure 4
-\item
-  Increasing JS hyper-parameter enhances its performance, but the
-  computational time required for the optimisation can quickly become
-  intensive, especially when evaluating the index function (such as
-  scagnostic indexes) multiple times across numerous iterations. -
-  Figure 5
-\item
-  Smoothness and squintability can be calculated and compared across
-  different problems to understand their relative complexity. Once a
-  projection pursuit problem is fully characterised (by the
-  shape-to-find, the index function used, and the data dimension),
-  increasing JSO hyper-parameters can enhance the search effectiveness,
-  however, the resulting increase in computational complexity will also
-  need to be considered. - modelling
-\end{itemize}
+This paper has presented new metrics to mathematically define desirable
+features of PP indexes, squintability and smoothness, and used these to
+assess the performance of the new jellyfish search optimiser. The
+metrics will be generally useful for characterising PP indexes, and halp
+with developing nex indexes.
+
+In the comparison of the JSO against the currently used CRS, as expected
+the JSO vastly out-performs CRS, and provides a high probability of
+finding the global optima. The JSO obtains the maximum more cleanly,
+with a slightly higher index value, and plot of the projected data
+showing the structure more clearly.
+
+The JSO performance is affected by the hyper-parameters, with a higher
+chance of reaching the global optima when more jellies are used and the
+maximum number of tries is increased. However, it comes at a
+computational cost, as expected. The performance declines if the
+projection dimension increases and if the PP index has low
+squintability. The higher the squintability the better chance the JSO
+can find the optima. However, interestingly smoothness does not affect
+the JSO performance.
+
+The new JSO is integrated with the current implementation of the
+projection pursuit guided tour in the \texttt{tourr} package, and can be
+further examined using PP optimisation diagnostics in the \texttt{ferrn}
+package. Using the JSO is a little different than the current CRS,
+because it runs many paths. The recommended approach is to conduct the
+optimisation off-line, and then use the guided tour to follow selected
+jellies. The tools for this are available in the \texttt{tourr} package,
+too.
 

### 2024-06-30 optim.qmd
update illustration, small edits

@@
 Additionally, this work proposes novel methods to quantify two properties of the PP index -- smoothness and squintability -- that capture the complexities inherent in PP optimisation problems.
-These two metrics are evaluated with JSO hyper-parameters to determine their effect on JSO success rate.
-The JSO algorithm has been implemented in the `tourr` package, while functions to calculate  smoothness and squintability of indexes are available in the `ferrn` package.
+These two metrics are evaluated with JSO hyper-parameters to determine their effects on JSO success rate.
+The JSO algorithm has been implemented in the `tourr` package and functions to calculate  smoothness and squintability of indexes are available in the `ferrn` package.
 keywords: 
@@
 
-# Introduction \[Nicolas and Jessica\]
+# Introduction
 
@@
 @sec-theory details JSO and introduces metrics that measure different properties of PP indexes.
-@sec-sim-deets outlines the data structures, numerical summaries, and visualisations used in the simulation studies conducted in @sec-app to assess JSO's performance.
-@sec-sim-res presents two sets of simulation experiments, comparing JSO permformance improvements relative to an existing optimiser, and studies the impact of different PP index properties on the optimisation performance.
+@sec-sim-deets details two simulation experiments to assess JSO's performance: one comparing JSO's permformance improvements relative to an existing optimiser, Creeping Random Search (CRS), and the other studying the impact of different PP index properties on optimisation performance.
+@sec-sim-res presents the results.
 @sec-conclusion summarises the work and provides suggestions for future directions.
 
-# Projection pursuit, tours, index functions and optimisation \[Di and Sherry\] {#sec-background}
+# Projection pursuit, tours, index functions and optimisation {#sec-background}
 
@@
 
-# The jellyfish optimiser and properties of PP indexes \[Nicolas and Jessica\] {#sec-theory}
+# The jellyfish optimiser and properties of PP indexes {#sec-theory}
 
@@
 
-# Results \[Di and Sherry\] {#sec-sim-res}
+# Results {#sec-sim-res}
 
@@
 Smoothness and squintability are calculated across a collection of pipe-finding and sine-wave finding problems to construct the relationship between success rate, JSO hyper-parameters, and index properties.
-
-## Performance of JSO in pipe-finding problems {#sec-app-1}
 
@@
 
-## Factors affecting JSO success rate: index properties and jellyfish hyper-parameters {#sec-app-2}
-
 ```{r}
@@
 
-# Conclusion \[Di and Sherry\] {#sec-conclusion}
+# Conclusion {#sec-conclusion}
 
@@
 
+# Acknowledgement
+
+The article is created using Quarto [@Allaire_Quarto_2022] in R [@R].
+The source code for reproducing the work reported in this paper can be found at: <https://github.com/huizezhang-sherry/paper-optim>.
+Nicolas Langrené acknowledges the partial support of the Guangdong Provincial Key Laboratory IRADS (2022B1212010006, R0400001-22) and the UIC Start-up Research Fund UICR0700041-22.
+
 # References

### 2024-06-30 optim.tex
update illustration, small edits

@@
 These two metrics are evaluated with JSO hyper-parameters to determine
-their effect on JSO success rate. The JSO algorithm has been implemented
-in the \texttt{tourr} package, while functions to calculate smoothness
-and squintability of indexes are available in the \texttt{ferrn}
-package.
+their effects on JSO success rate. The JSO algorithm has been
+implemented in the \texttt{tourr} package and functions to calculate
+smoothness and squintability of indexes are available in the
+\texttt{ferrn} package.
 \end{abstract}
@@
 
-\section{Introduction {[}Nicolas and
-Jessica{]}}\label{introduction-nicolas-and-jessica}
+\section{Introduction}\label{introduction}
 
@@
 Section~\ref{sec-theory} details JSO and introduces metrics that measure
-different properties of PP indexes. Section~\ref{sec-sim-deets} outlines
-the data structures, numerical summaries, and visualisations used in the
-simulation studies conducted in \textbf{?@sec-app} to assess JSO's
-performance. Section~\ref{sec-sim-res} presents two sets of simulation
-experiments, comparing JSO permformance improvements relative to an
-existing optimiser, and studies the impact of different PP index
-properties on the optimisation performance. Section~\ref{sec-conclusion}
-summarises the work and provides suggestions for future directions.
-
-\section{Projection pursuit, tours, index functions and optimisation
-{[}Di and Sherry{]}}\label{sec-background}
+different properties of PP indexes. Section~\ref{sec-sim-deets} details
+two simulation experiments to assess JSO's performance: one comparing
+JSO's permformance improvements relative to an existing optimiser,
+Creeping Random Search (CRS), and the other studying the impact of
+different PP index properties on optimisation performance.
+Section~\ref{sec-sim-res} presents the results.
+Section~\ref{sec-conclusion} summarises the work and provides
+suggestions for future directions.
+
+\section{Projection pursuit, tours, index functions and
+optimisation}\label{sec-background}
 
@@
 
-\section{The jellyfish optimiser and properties of PP indexes {[}Nicolas
-and Jessica{]}}\label{sec-theory}
+\section{The jellyfish optimiser and properties of PP
+indexes}\label{sec-theory}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, bottomtitle=1mm, bottomrule=.15mm, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, leftrule=.75mm, coltitle=black, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, opacityback=0, colback=white, left=2mm, toprule=.15mm, arc=.35mm, colframe=quarto-callout-note-color-frame]
+\begin{tcolorbox}[enhanced jigsaw, titlerule=0mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, arc=.35mm, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, leftrule=.75mm, coltitle=black, bottomrule=.15mm, toprule=.15mm, opacityback=0, left=2mm, rightrule=.15mm, breakable]
 
@@
 
-\includegraphics[width=4.49in,height=\textheight]{figures/success-rate.png}
+\includegraphics[width=4.52in,height=\textheight]{figures/success-rate.png}
 
@@
 
-\section{Results {[}Di and Sherry{]}}\label{sec-sim-res}
+\section{Results}\label{sec-sim-res}
 
@@
 properties.
-
-\subsection{Performance of JSO in pipe-finding
-problems}\label{sec-app-1}
 
@@
 \end{figure}%
-
-\subsection{Factors affecting JSO success rate: index properties and
-jellyfish hyper-parameters}\label{sec-app-2}
 
@@
 
-\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
+\section{Conclusion}\label{sec-conclusion}
 
@@
 
+\section{Acknowledgement}\label{acknowledgement}
+
+The article is created using Quarto \citep{Allaire_Quarto_2022} in R
+\citep{R}. The source code for reproducing the work reported in this
+paper can be found at:
+\url{https://github.com/huizezhang-sherry/paper-optim}. Nicolas Langrené
+acknowledges the partial support of the Guangdong Provincial Key
+Laboratory IRADS (2022B1212010006, R0400001-22) and the UIC Start-up
+Research Fund UICR0700041-22.
+
 \renewcommand\refname{References}

### 2024-07-01 optim.qmd
section 4.1 revised

@@
 
-There are several PP index functions available: `holes()` and `cmass()`
-[@cook1993projection]; `lda_pp()` [@lee2005projection]; `pda_pp()`
-[@lee2010projection]; `dcor2d()` and `splines2d()` [@Grimm2016];
-`norm_bin()` and `norm_kol()` [@huber85]; `slice_index()`
-[@Laa:2020wkm]. Most are relatively simply defined, for any projection
-dimension, and implemented because they are relatively easy to optimise.
-A goal is to be able to incorporate more complex PP indexes, for example
-based on scagnostics (@scag, @WW08).
+There are several PP index functions available: `holes()` and `cmass()` [@cook1993projection]; `lda_pp()` [@lee2005projection]; `pda_pp()`
+[@lee2010projection]; `dcor2d()` and `splines2d()` [@Grimm2016]; `norm_bin()` and `norm_kol()` [@huber85]; `slice_index()` [@Laa:2020wkm].
+Most are relatively simply defined, for any projection dimension, and implemented because they are relatively easy to optimise.
+A goal is to be able to incorporate more complex PP indexes, for example based on scagnostics (@scag, @WW08).
 
@@
 
-The JSO performance is compared with the existing Creeping Random Search (CRS) currently used in guided tour, and explores how JSO behaves with different hyper-parameters combinations.
-A second simulation
-studies the effect of index properties (smoothness and squintability), along with JSO hyper-parameters, on the success rate of the JSO performance.
+The JSO performance is compared with the existing Creeping Random Search (CRS) [@RJ-2021-105;
+@laa_using_2020] currently used in guided tour, and explores how JSO behaves with different hyper-parameter (number of jellies, maximum tries) and data dimension combinations. A second simulation
+studies the effect of index properties (smoothness and squintability), along with JSO hyper-parameters, and data dimension, on the success rate of the JSO performance.
 
-## Performance of JSO in pipe-finding problems {#sec-app-1}
+## Performance of JSO relative to CRS {#sec-app-1}
 
-The performance of JSO is investigated in two folds: 1) compare it with a commonly used optimiser: the Csreeping Random Search (CRS) [@RJ-2021-105;
-@laa_using_2020], and 2) examine its success rate under two hyper-parameters (number of jellyfishes and maximum number of tries).
-The projection pursuit problem used is finding the pipe shape using the holes index, investigated by @laa_using_2020.
-Fifty simulations are conducted with both JSO and CRS, in four dimensions ($d = 6, 8, 10, 12$).
-JSO uses 100 jellyfishes and 100 maximum number of tries, while the CRS allows a maximum of 1000 samples at each iteration before the algorithm terminates.
-The results of the simulation are collected using the data structure proposed in @RJ-2021-105 is used here for assessing JSO.
-A sample for the sine-wave structure in 6D data using a spline index function (`splines2d`) is printed below:
+The performance of JSO is investigated in comparison with the existing optimiser, CRS, with respect to two hyper-parameters (number of jellyfishes and maximum number of tries).
+The performance is measured by success rate, defined as the proportion of simulations that achieves a final index value within 0.05 of the best index value found among all 50 simulations.
+This comparison is based on projection pursuit to find the pipe shape investigated by @laa_using_2020 using two simple and well-behaved indexes, the `holes` index, and the `splines2d` index.
+Fifty simulations are conducted with both JSO and CRS, in four different data dimensions ($d = 6, 8, 10, 12$).
+JSO uses 100 jellyfishes with 100 maximum number of tries, while the CRS allows a maximum of 1000 samples at each iteration before the algorithm terminates.
+The different numbers accounts for the multiple paths of JSO to enable fairer comparison with CRS.
+The results of the simulation are collected using the data structure proposed in @RJ-2021-105 is used here for assessing JSO, where the design parameters are stored along with index value, projection basis, random seed, and computation time.
+<!--A sample for the sine-wave structure in 6D data using a spline index function (`splines2d`) is printed below:
 
@@
 ```
+-->
 
-In this simulation, 50 jellyfishes (`n_jellies`) and
+In this simulation, 50 jellies (`n_jellies`) and
 a maximum of 50 tries (`max_tries`) are used.
 The full data contains 50 simulations, indexed by the variable `sim`, with the associated `seed`s.
-Recorded in columns `basis`, `index_val`, `tries`, `loop` and `time`, are the sampled projection basis, index value, iteration ID, jellyfish ID, and computational time, respectively.
-
+Recorded in columns `basis`, `index_val`, `tries`, `loop` and `time`, are the projection basis, index value, iteration ID, jellyfish ID, and computational time, respectively.
 <!-- The basis column records every visited basis before -->
@@
 
-In the second experiment, fifty simulations are conducted at each hyper-parameter combination below to analyse its effects on the JSO success rate: number of jellyfish: 20, 50, and 100, and  maximum number of  tries: 50 and 100).
-The success rate is defined as the proportion of simulations that achieves a final index value within 0.05 of the best index value found among all 50 simulations.
-@fig-success-rate illustrates
-the calculation based on the projection found by the above simulation. Here, the optimal index value is
-`r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`. Out of the 50 simulations,
+In the second experiment, 50 simulations are conducted at each hyper-parameter combination to analyse its effect on the JSO success rate: number of jellies (20, 50, and 100), and  maximum number of  tries (50 and 100).
+@fig-success-rate illustrates the calculation based on the projection found by the above simulation.
+Here, the optimal index value is `r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`.
+Out of the 50 simulations,
 `r sim_sine_6d_spline_best |> filter(index_val >= aa - 0.05) |> nrow() -> bb; bb` achieved an index value within \[`r aa - 0.05`, `r aa`\], resulting in a success rate of `r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.

### 2024-07-01 optim.tex
section 4.1 revised

@@
 
-\begin{tcolorbox}[enhanced jigsaw, titlerule=0mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, arc=.35mm, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, leftrule=.75mm, coltitle=black, bottomrule=.15mm, toprule=.15mm, opacityback=0, left=2mm, rightrule=.15mm, breakable]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, opacityback=0, breakable, toprule=.15mm, arc=.35mm, colback=white, coltitle=black, bottomrule=.15mm, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm, rightrule=.15mm, opacitybacktitle=0.6]
 
@@
 The JSO performance is compared with the existing Creeping Random Search
-(CRS) currently used in guided tour, and explores how JSO behaves with
-different hyper-parameters combinations. A second simulation studies the
-effect of index properties (smoothness and squintability), along with
-JSO hyper-parameters, on the success rate of the JSO performance.
-
-\subsection{Performance of JSO in pipe-finding
-problems}\label{sec-app-1}
-
-The performance of JSO is investigated in two folds: 1) compare it with
-a commonly used optimiser: the Csreeping Random Search (CRS)
-\citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
-under two hyper-parameters (number of jellyfishes and maximum number of
-tries). The projection pursuit problem used is finding the pipe shape
-using the holes index, investigated by \citet{laa_using_2020}. Fifty
-simulations are conducted with both JSO and CRS, in four dimensions
-(\(d = 6, 8, 10, 12\)). JSO uses 100 jellyfishes and 100 maximum number
-of tries, while the CRS allows a maximum of 1000 samples at each
-iteration before the algorithm terminates. The results of the simulation
-are collected using the data structure proposed in \citet{RJ-2021-105}
-is used here for assessing JSO. A sample for the sine-wave structure in
-6D data using a spline index function (\texttt{splines2d}) is printed
-below:
-
-\begin{verbatim}
-Rows: 10
-Columns: 12
-$ index     <chr> "splines2d", "splines2d", "splines2d", "splines2d", "splines~
-$ d         <dbl> 6, 6, 6, 6, 6, 6, 6, 6, 6, 6
-$ n_jellies <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50
-$ max_tries <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50
-$ sim       <int> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1
-$ seed      <int> 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462
-$ basis     <list> <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<m~
-$ index_val <dbl> 0.036664532, 0.005121912, 0.058683104, 0.056527991, 0.015734~
-$ method    <chr> "search_jellyfish", "search_jellyfish", "search_jellyfish", ~
-$ tries     <dbl> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1
-$ loop      <dbl> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10
-$ time      <drtn> 4.362133 mins, 4.362133 mins, 4.362133 mins, 4.362133 mins, ~
-\end{verbatim}
-
-In this simulation, 50 jellyfishes (\texttt{n\_jellies}) and a maximum
-of 50 tries (\texttt{max\_tries}) are used. The full data contains 50
+(CRS) \citep{RJ-2021-105, laa_using_2020} currently used in guided tour,
+and explores how JSO behaves with different hyper-parameter (number of
+jellies, maximum tries) and data dimension combinations. A second
+simulation studies the effect of index properties (smoothness and
+squintability), along with JSO hyper-parameters, and data dimension, on
+the success rate of the JSO performance.
+
+\subsection{Performance of JSO relative to CRS}\label{sec-app-1}
+
+The performance of JSO is investigated in comparison with the existing
+optimiser, CRS, with respect to two hyper-parameters (number of
+jellyfishes and maximum number of tries). The performance is measured by
+success rate, defined as the proportion of simulations that achieves a
+final index value within 0.05 of the best index value found among all 50
+simulations. This comparison is based on projection pursuit to find the
+pipe shape investigated by \citet{laa_using_2020} using two simple and
+well-behaved indexes, the \texttt{holes} index, and the
+\texttt{splines2d} index. Fifty simulations are conducted with both JSO
+and CRS, in four different data dimensions (\(d = 6, 8, 10, 12\)). JSO
+uses 100 jellyfishes with 100 maximum number of tries, while the CRS
+allows a maximum of 1000 samples at each iteration before the algorithm
+terminates. The different numbers accounts for the multiple paths of JSO
+to enable fairer comparison with CRS. The results of the simulation are
+collected using the data structure proposed in \citet{RJ-2021-105} is
+used here for assessing JSO, where the design parameters are stored
+along with index value, projection basis, random seed, and computation
+time.
+
+In this simulation, 50 jellies (\texttt{n\_jellies}) and a maximum of 50
+tries (\texttt{max\_tries}) are used. The full data contains 50
 simulations, indexed by the variable \texttt{sim}, with the associated
 \texttt{seed}s. Recorded in columns \texttt{basis}, \texttt{index\_val},
-\texttt{tries}, \texttt{loop} and \texttt{time}, are the sampled
-projection basis, index value, iteration ID, jellyfish ID, and
-computational time, respectively.
-
-In the second experiment, fifty simulations are conducted at each
-hyper-parameter combination below to analyse its effects on the JSO
-success rate: number of jellyfish: 20, 50, and 100, and maximum number
-of tries: 50 and 100). The success rate is defined as the proportion of
-simulations that achieves a final index value within 0.05 of the best
-index value found among all 50 simulations.
-Figure~\ref{fig-success-rate} illustrates the calculation based on the
-projection found by the above simulation. Here, the optimal index value
-is 0.999. Out of the 50 simulations, 44 achieved an index value within
-{[}0.949, 0.999{]}, resulting in a success rate of 0.88.
+\texttt{tries}, \texttt{loop} and \texttt{time}, are the projection
+basis, index value, iteration ID, jellyfish ID, and computational time,
+respectively.
+
+In the second experiment, 50 simulations are conducted at each
+hyper-parameter combination to analyse its effect on the JSO success
+rate: number of jellies (20, 50, and 100), and maximum number of tries
+(50 and 100). Figure~\ref{fig-success-rate} illustrates the calculation
+based on the projection found by the above simulation. Here, the optimal
+index value is 0.999. Out of the 50 simulations, 44 achieved an index
+value within {[}0.949, 0.999{]}, resulting in a success rate of 0.88.
 

### 2024-07-01 optim.qmd
revised section 4.2

@@
 
-In addition to the pipe-finding problem, a new shape, sine wave, is investigated in 6D and 8D spaces with six indexes considered:
-`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`.
-Combining with two JSO hyper-parameters, a total of `r nrow(sim_df)` cases is produced, comprising of `r nrow(filter(sim_df, index == "holes"))` pipe-finding cases and `r nrow(filter(sim_df, index != "holes"))` sine-wave finding cases.
-For each case, JSO is run fifty times and the success rate is calculated.
+This simulation uses two different data shapes, pipe and a sine wave, investigated in 6D and 8D spaces with six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`.
+The two JSO hyper-parameters are varied.
+A total of `r nrow(sim_df)` combinations result, comprising of `r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and `r nrow(filter(sim_df, index != "holes"))` on the sine-wave data.
+For each case, JSO is run 50 times to calculate the success rate.
 
-Smoothness and squintability are computed for each case, following the procedures outlined in @sec-smoothness and @sec-squintability and illustrated in @fig-smoothness and  @fig-squintability.
-For smoothness, three hundred random bases are simulated.
-Index values and projection distance (to the optimal basis) are calculated for each random basis before fitting into a Gaussian process model to obtain
-the smoothness measure. For squintability, fifty random bases are sampled and interpolated to the optimal basis with a step size of 0.005. Index values and projection 
-distances are calculated for these interpolated bases and the index values are averaged with a bin width of 0.005.
-The squintability measure is then calculated from fitting a 4-parameter scaled logistic
-function of index value against projection distance, estimated by non-linear least squares.
+Smoothness and squintability are computed for each run, following the procedures outlined in @sec-smoothness and @sec-squintability and as illustrated in @fig-smoothness and  @fig-squintability.
 
-Combining with the success rate calculated for each case, the dataset is fitted using  a generalised linear model with a binomial family and a logit link
-function with three pre-processing steps are applied: 
-1) scaling the JSO hyper-parameters by a factor of 10 for interpretation, 2) creating a new binary variable `long_time` to indicate cases with an average run time over 30 seconds, and 3) encoding the success rate for
-the stringy case as 0, because upon visual inspection of the final projection, none of the 50 simulations identified the sine-wave shape.
+To compute smoothness, 300 random bases are simulated.
+Index values and projection distance (to the optimal basis) are calculated for each random basis before fitting the Gaussian process model to obtain the smoothness measure for the index.
+
+To compute squintability, 50 random bases are sampled and interpolated to the optimal basis with a step size of 0.005.
+Index values and projection distances are calculated for these interpolated bases and the index values are averaged with a bin width of 0.005.
+The squintability measure is then calculated by fitting a four-parameter scaled logistic function of index value against projection distance, estimated by non-linear least squares.
+
+A generalised linear model is fitted to the resulting data, using a binomial family and a logit link function.
+The data is pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10 for interpretation, 2) creating a new binary variable `long_time` to indicate cases with an average run time over 30 seconds, and 3) re-coding the success rate for
+the `stringy` index as 0, because none of the 50 simulations correctly identified the sine-wave shape.
 
@@
 the parameters estimated from the Gaussian process (variance, range, smooth, and nugget) and the scaled
-logistic function ($\theta_1$ to $\theta_4$) for calculating both metrics for each case considered in the simulation. The column
-"smooth" is used as the smoothness metric and the column "squint" is
-calculated as equation @eq-squintability-parametric as the squintability metric.  @tbl-mod-output presents the results of fitting a generalised linear model with a binomial family and a logit link
-function to the data in @tbl-mod-data, where all the three components (success rate, JS hyper-parameters, and index properties) are combined.
+logistic function ($\theta_1$ to $\theta_4$) for calculating both metrics for each case considered in the simulation.
+The column "smooth" is used as the smoothness metric and the column "squint" is calculated as equation @eq-squintability-parametric as the squintability metric.
+@tbl-mod-output presents the results of fitting a generalised linear model with a binomial family and a logit link function to the data in @tbl-mod-data, where all the three components (success rate, JS hyper-parameters, and index properties) are combined.
 The model suggests that JSO success rate is positively associated with the two hyper-parameters, as well as with the index properties: smoothness and squintability.
@@
   kable(digits = 2, format = "latex", 
-        col.names = c("Index", "D", 
-          "Smoothness", "Squintability", 
-          "N. jellyfish", 
-          "Max. tries", 
-          "Success rate", "Time (sec)"),
+        col.names = c("index", "d", 
+          "smoothness", "squintability", 
+          "n. jellyfish", 
+          "max. tries", 
+          "success rate", "time (sec)"),
         linesep = "",

### 2024-07-01 optim.tex
revised section 4.2

@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, opacityback=0, breakable, toprule=.15mm, arc=.35mm, colback=white, coltitle=black, bottomrule=.15mm, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm, rightrule=.15mm, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, coltitle=black, colback=white, colbacktitle=quarto-callout-note-color!10!white, left=2mm, leftrule=.75mm, titlerule=0mm, rightrule=.15mm, toptitle=1mm, opacityback=0, arc=.35mm, bottomrule=.15mm, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, opacitybacktitle=0.6, breakable]
 
@@
 
-In addition to the pipe-finding problem, a new shape, sine wave, is
-investigated in 6D and 8D spaces with six indexes considered:
+This simulation uses two different data shapes, pipe and a sine wave,
+investigated in 6D and 8D spaces with six different PP indexes:
 \texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
-\texttt{spline}, and \texttt{stringy}. Combining with two JSO
-hyper-parameters, a total of 52 cases is produced, comprising of 24
-pipe-finding cases and 28 sine-wave finding cases. For each case, JSO is
-run fifty times and the success rate is calculated.
-
-Smoothness and squintability are computed for each case, following the
+\texttt{spline}, and \texttt{stringy}. The two JSO hyper-parameters are
+varied. A total of 52 combinations result, comprising of 24 computed on
+the pipe data and 28 on the sine-wave data. For each case, JSO is run 50
+times to calculate the success rate.
+
+Smoothness and squintability are computed for each run, following the
 procedures outlined in Section~\ref{sec-smoothness} and
-Section~\ref{sec-squintability} and illustrated in
-Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}. For
-smoothness, three hundred random bases are simulated. Index values and
+Section~\ref{sec-squintability} and as illustrated in
+Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}.
+
+To compute smoothness, 300 random bases are simulated. Index values and
 projection distance (to the optimal basis) are calculated for each
-random basis before fitting into a Gaussian process model to obtain the
-smoothness measure. For squintability, fifty random bases are sampled
-and interpolated to the optimal basis with a step size of 0.005. Index
-values and projection distances are calculated for these interpolated
-bases and the index values are averaged with a bin width of 0.005. The
-squintability measure is then calculated from fitting a 4-parameter
-scaled logistic function of index value against projection distance,
-estimated by non-linear least squares.
-
-Combining with the success rate calculated for each case, the dataset is
-fitted using a generalised linear model with a binomial family and a
-logit link function with three pre-processing steps are applied: 1)
-scaling the JSO hyper-parameters by a factor of 10 for interpretation,
-2) creating a new binary variable \texttt{long\_time} to indicate cases
-with an average run time over 30 seconds, and 3) encoding the success
-rate for the stringy case as 0, because upon visual inspection of the
-final projection, none of the 50 simulations identified the sine-wave
-shape.
+random basis before fitting the Gaussian process model to obtain the
+smoothness measure for the index.
+
+To compute squintability, 50 random bases are sampled and interpolated
+to the optimal basis with a step size of 0.005. Index values and
+projection distances are calculated for these interpolated bases and the
+index values are averaged with a bin width of 0.005. The squintability
+measure is then calculated by fitting a four-parameter scaled logistic
+function of index value against projection distance, estimated by
+non-linear least squares.
+
+A generalised linear model is fitted to the resulting data, using a
+binomial family and a logit link function. The data is pre-processed by
+1) scaling the JSO hyper-parameters by a factor of 10 for
+interpretation, 2) creating a new binary variable \texttt{long\_time} to
+indicate cases with an average run time over 30 seconds, and 3)
+re-coding the success rate for the \texttt{stringy} index as 0, because
+none of the 50 simulations correctly identified the sine-wave shape.
 
@@
 \toprule
-Index & D & Smoothness & Squintability & N. jellyfish & Max. tries & Success rate & Time (sec)\\
+index & d & smoothness & squintability & n. jellyfish & max. tries & success rate & time (sec)\\
 \midrule

### 2024-07-01 optim.qmd
update abstract

@@
   The projection pursuit (PP) guided tour interactively optimises a criteria function known as the PP index, to explore high-dimensional data by revealing interesting projections.
-The optimisation in PP can be non-trivial, involving non-smooth functions and optima with small a "squint angle", detectable only from close proximity.
+The optimisation in PP can be non-trivial, involving non-smooth functions and optima with a small "squint angle", detectable only from close proximity.
 To address these challenges, this study investigates the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
@@
 Additionally, this work proposes novel methods to quantify two properties of the PP index -- smoothness and squintability -- that capture the complexities inherent in PP optimisation problems.
-These two metrics are evaluated with JSO hyper-parameters to determine their effects on JSO success rate.
+These two metrics are evaluated along with JSO hyper-parameters to determine their effects on JSO success rate.
+Our numerical results confirm the positive impact of these metrics on the JSO success rate, with squintability being the most significant.
 The JSO algorithm has been implemented in the `tourr` package and functions to calculate smoothness and squintability of indexes are available in the `ferrn` package.
@@
 
-Projection Pursuit (PP) (@kr69, @FT74) is a dimension reduction technique aimed at identifying informative linear projections of data.
-The method involves optimising a objective function known as the PP index
-(e.g.
-@hall1989polynomial, @cook1993projection, @lee2010projection, @Loperfido2018, @Loperfido2020), which defines the criteria for what constitutes  interesting or informative projections.
-Given high-dimensional data $X \in \mathbf{R}^{n\times p}$ and an index function $f(\cdot)$, PP finds the orthonormal basis $A \in \mathbf{R}^{p \times d}$ that maximises the index value of the projection, $Y = XA$:
+Projection Pursuit (PP) (@kr69, @FT74) is a dimension reduction
+technique aimed at identifying informative linear projections of data.
+The method involves optimising a objective function known as the PP
+index (e.g. @hall1989polynomial, @cook1993projection,
+@lee2010projection, @Loperfido2018, @Loperfido2020), which defines the
+criteria for what constitutes interesting or informative projections.
+Given high-dimensional data $X \in \mathbf{R}^{n\times p}$ and an index
+function $f(\cdot)$, PP finds the orthonormal basis
+$A \in \mathbf{R}^{p \times d}$ that maximises the index value of the
+projection, $Y = XA$:
 
@@
 
-The index functions can be non-linear and non-convex, hence an effective and efficient optimisation procedure is essential to explore the data landscape and achieving a globally optimal viewpoint of the data.
+The index functions can be non-linear and non-convex, hence an effective
+and efficient optimisation procedure is essential to explore the data
+landscape and achieving a globally optimal viewpoint of the data.
 
-Significant efforts have been devoted to enhancing the functionality of PP, improving optimisation (e.g.
-@posse1995, @marie-sainte2010, @grochowski2011) and extending it's use to visually explore high-dimensional data.
-@cook1995grand introduced the PP guided tour, which monitors the optimisation visually to see the data projections leading in and out of the optima.
-An implementation is available in the R [@R] package `tourr` [@tourr].
-@RJ-2021-105 illustrated how to diagnose optimisation processes, particularly focusing on the guided tour, and revealed a need to for improved optimisation.
-While improving the quality of the optimisation solutions in the tour is essential, it is also important to be able to view the data projections as the optimisation progresses.
-Integrating the guided tour with a global optimisation algorithm that is efficient in finding the global optimal and enables viewing of the projected data during the exploration process is a goal.
+Significant efforts have been devoted to enhancing the functionality of
+PP, improving optimisation (e.g. @posse1995, @marie-sainte2010,
+@grochowski2011) and extending it's use to visually explore
+high-dimensional data. @cook1995grand introduced the PP guided tour,
+which monitors the optimisation visually to see the data projections
+leading in and out of the optima. An implementation is available in the
+R [@R] package `tourr` [@tourr]. @RJ-2021-105 illustrated how to
+diagnose optimisation processes, particularly focusing on the guided
+tour, and revealed a need to for improved optimisation. While improving
+the quality of the optimisation solutions in the tour is essential, it
+is also important to be able to view the data projections as the
+optimisation progresses. Integrating the guided tour with a global
+optimisation algorithm that is efficient in finding the global optimal
+and enables viewing of the projected data during the exploration process
+is a goal.
 
-Here, the potential for a Jellyfish Search Optimiser (JSO) to be integrated with the projection pursuit guided tour is explored.
-JSO [@chou_novel_2021] -[@rajwar_exhaustive_2023], inspired by the search behaviour of jellyfish in the ocean, is a swarm-based metaheuristic designed to solve global optimisation problems.
-Compared to traditional methods, JSO demonstrated stronger search ability and faster convergence, and requires fewer tuning parameters.
-These practical advantages make JSO a promising candidate for enhancing PP optimisation.
+Here, the potential for a Jellyfish Search Optimiser (JSO) to be
+integrated with the projection pursuit guided tour is explored. JSO
+[@chou_novel_2021] -[@rajwar_exhaustive_2023], inspired by the search
+behaviour of jellyfish in the ocean, is a swarm-based metaheuristic
+designed to solve global optimisation problems. Compared to traditional
+methods, JSO demonstrated stronger search ability and faster
+convergence, and requires fewer tuning parameters. These practical
+advantages make JSO a promising candidate for enhancing PP optimisation.
 <!--The study reported here explores JSO's potential in enhancing projection pursuit and examines the characteristics
@@
 
-The primary goal of the study reported here is to investigate the performance of JSO in PP optimisation for the guided tour.
-It is of interest to assess how quickly and closely the optimiser reaches a gobal optima, for various PP indexes that may have differing complexities.
-To observe the performance of JSO with different types of PP indexes, metrics are introduced to capture specific properties of the index including squintability (based on @barnett1981interpreting's squint angle) and smoothness as introduced in @laa_using_2020.
-Here we mathematically define metrics for squintability and smoothness, which is new for the field.
-A series of simulation experiments using specially structured datasets and various PP indexes are conducted to assess the behaviour of JSO under different settings and its sensitivity to hyper-parameter choices (number of jellies and maximum number of tries).
-The relationship between the JSO performance, hyper-parameter tuning and properties of PP indexes is analysed to provide guidance for practitioners using projection pursuit, and developers of new indexes.
+The primary goal of the study reported here is to investigate the
+performance of JSO in PP optimisation for the guided tour. It is of
+interest to assess how quickly and closely the optimiser reaches a gobal
+optima, for various PP indexes that may have differing complexities. To
+observe the performance of JSO with different types of PP indexes,
+metrics are introduced to capture specific properties of the index
+including squintability (based on @barnett1981interpreting's squint
+angle) and smoothness as introduced in @laa_using_2020. Here we
+mathematically define metrics for squintability and smoothness, which is
+new for the field. A series of simulation experiments using specially
+structured datasets and various PP indexes are conducted to assess the
+behaviour of JSO under different settings and its sensitivity to
+hyper-parameter choices (number of jellies and maximum number of tries).
+The relationship between the JSO performance, hyper-parameter tuning and
+properties of PP indexes is analysed to provide guidance for
+practitioners using projection pursuit, and developers of new indexes.
 
-The paper is structured as follows.
-@sec-background introduces the background of PP guided tour and reviews existing optimisers and index functions in the literature.
-@sec-theory details JSO and introduces metrics that measure different properties of PP indexes.
-@sec-sim-deets details two simulation experiments to assess JSO's performance: one comparing JSO's permformance improvements relative to an existing optimiser, Creeping Random Search (CRS), and the other studying the impact of different PP index properties on optimisation performance.
-@sec-sim-res presents the results.
-@sec-conclusion summarises the work and provides suggestions for future directions.
+The paper is structured as follows. @sec-background introduces the
+background of PP guided tour and reviews existing optimisers and index
+functions in the literature. @sec-theory details JSO and introduces
+metrics that measure different properties of PP indexes. @sec-sim-deets
+details two simulation experiments to assess JSO's performance: one
+comparing JSO's permformance improvements relative to an existing
+optimiser, Creeping Random Search (CRS), and the other studying the
+impact of different PP index properties on optimisation performance.
+@sec-sim-res presents the results. @sec-conclusion summarises the work
+and provides suggestions for future directions.
 
@@
 
-A tour on high-dimensional data is constructed by geodesically interpolating between pairs of planes.
-Any plane is described by an orthonormal basis, $A_t$, where $t$ represents time in the sequence.
-The term "geodesic" refers to maintaining the orthonormality constraint so that each view shown is correctly a projection of the data.
-The PP guided tour operates by geodesically interpolating to target planes (projections) which have high PP index values, as provided by the optimiser.
-The geodesic interpolation means that the viewer sees a continuous sequence of projections of the data, so they can watch patterns of interest forming as the function is optimised.
-There are five (unsophisticated) optimisation methods implemented in the `tourr` package:
+A tour on high-dimensional data is constructed by geodesically
+interpolating between pairs of planes. Any plane is described by an
+orthonormal basis, $A_t$, where $t$ represents time in the sequence. The
+term "geodesic" refers to maintaining the orthonormality constraint so
+that each view shown is correctly a projection of the data. The PP
+guided tour operates by geodesically interpolating to target planes
+(projections) which have high PP index values, as provided by the
+optimiser. The geodesic interpolation means that the viewer sees a
+continuous sequence of projections of the data, so they can watch
+patterns of interest forming as the function is optimised. There are
+five (unsophisticated) optimisation methods implemented in the `tourr`
+package:
 
--   `search_geodesic()`: provides a pseudo-derivative optimisation.
-It searches locally for the best direction, based on differencing the index values for very close projections.
-Then it follows the direction along the geodesic path between planes, stopping when the next index value fails to increase.
--   `search_better()`: is a brute-force optimisation searching randomly for projections with higher index values.
--   `search_better_random()`: is essentially simulated annealing [@Bertsimas93] where the search space is reduced as the optimisation progresses.
+-   `search_geodesic()`: provides a pseudo-derivative optimisation. It
+    searches locally for the best direction, based on differencing the
+    index values for very close projections. Then it follows the
+    direction along the geodesic path between planes, stopping when the
+    next index value fails to increase.
+-   `search_better()`: is a brute-force optimisation searching randomly
+    for projections with higher index values.
+-   `search_better_random()`: is essentially simulated annealing
+    [@Bertsimas93] where the search space is reduced as the optimisation
+    progresses.
 -   `search_posse()`: implements the algorithm described in @posse95.
--   `search_polish()`: is a very localised search, to take tiny steps to get closer to the local maximum.
+-   `search_polish()`: is a very localised search, to take tiny steps to
+    get closer to the local maximum.
 
-There are several PP index functions available: `holes()` and `cmass()` [@cook1993projection]; `lda_pp()` [@lee2005projection]; `pda_pp()`
-[@lee2010projection]; `dcor2d()` and `splines2d()` [@Grimm2016]; `norm_bin()` and `norm_kol()` [@huber85]; `slice_index()` [@Laa:2020wkm].
-Most are relatively simply defined, for any projection dimension, and implemented because they are relatively easy to optimise.
-A goal is to be able to incorporate more complex PP indexes, for example based on scagnostics (@scag, @WW08).
+There are several PP index functions available: `holes()` and `cmass()`
+[@cook1993projection]; `lda_pp()` [@lee2005projection]; `pda_pp()`
+[@lee2010projection]; `dcor2d()` and `splines2d()` [@Grimm2016];
+`norm_bin()` and `norm_kol()` [@huber85]; `slice_index()`
+[@Laa:2020wkm]. Most are relatively simply defined, for any projection
+dimension, and implemented because they are relatively easy to optimise.
+A goal is to be able to incorporate more complex PP indexes, for example
+based on scagnostics (@scag, @WW08).
 
-An initial investigation of PP indexes, and the potential for scagnostics is described in @laa_using_2020. To be useful here an optimiser needs to be able to handle index functions which are possibly
-not very smooth.
-In addition, because data structures might be relatively fine, the optimiser needs to be able to find maxima that occur with a small squint angle, that can only be seen from very close
-by. One last aspect that is useful is for an optimiser to return local maxima in addition to global because data can contain many different and interesting features.
+An initial investigation of PP indexes, and the potential for
+scagnostics is described in @laa_using_2020. To be useful here an
+optimiser needs to be able to handle index functions which are possibly
+not very smooth. In addition, because data structures might be
+relatively fine, the optimiser needs to be able to find maxima that
+occur with a small squint angle, that can only be seen from very close
+by. One last aspect that is useful is for an optimiser to return local
+maxima in addition to global because data can contain many different and
+interesting features.
 
@@
 
-@laa_using_2020 has proposed five criteria for assessing projection pursuit indexes (smoothness, squintability, flexibility, rotation invariance, and speed). Since not all the properties affects the
-execution of the optimisation, we choose to focus here on the first two properties, namely *smoothness* (@sec-smoothness) and *squintability* (@sec-squintability), and propose two metrics to evaluate them.
+@laa_using_2020 has proposed five criteria for assessing projection
+pursuit indexes (smoothness, squintability, flexibility, rotation
+invariance, and speed). Since not all the properties affects the
+execution of the optimisation, we choose to focus here on the first two
+properties, namely *smoothness* (@sec-smoothness) and *squintability*
+(@sec-squintability), and propose two metrics to evaluate them.
 
@@
 ```
-
 This subsection proposes a metric for the smoothness of a projection
@@
 
-If a PP index function is evaluated at some random points, as is done at the 
-initialization stage of the jellyfish search optimization, then 
-these random index values can be interpreted as a random field, 
-indexed by a space parameter, namely the random projection angle. 
-This analogy suggests to use this random training sample to fit a spatial model, 
-a simple one being a (spatial) Gaussian process.
+If a PP index function is evaluated at some random points, as is done at
+the initialization stage of the jellyfish search optimization, then
+these random index values can be interpreted as a random field, indexed
+by a space parameter, namely the random projection angle. This analogy
+suggests to use this random training sample to fit a spatial model, a
+simple one being a (spatial) Gaussian process.
 
-The distribution of a Gaussian process is fully determined by its mean function and
-covariance function. The smoothness property comes into play in the definition
-of the covariance function: if a PP index is very smooth, then two close
-projection angles should produce close index values (strong
-correlation); by contrast, if a PP index is not very smooth, then two close
-projection angles might give very different index values (fast decay of
-correlations with respect to distance between angles).
+The distribution of a Gaussian process is fully determined by its mean
+function and covariance function. The smoothness property comes into
+play in the definition of the covariance function: if a PP index is very
+smooth, then two close projection angles should produce close index
+values (strong correlation); by contrast, if a PP index is not very
+smooth, then two close projection angles might give very different index
+values (fast decay of correlations with respect to distance between
+angles).
 
@@
 values in the literature being $1/2$, $3/2$ and $5/2$. The parameter
-$\nu$, called *smoothness parameter*, controls the decay of the covariance
-function. As such, it is an appropriate measure of smoothness of a
-random field.
+$\nu$, called *smoothness parameter*, controls the decay of the
+covariance function. As such, it is an appropriate measure of smoothness
+of a random field.
 
-In our context, we suggest to use this parameter $\nu$ as a measure of the
-smoothness of the index function by fitting a Gaussian process prior
+In our context, we suggest to use this parameter $\nu$ as a measure of
+the smoothness of the index function by fitting a Gaussian process prior
 with Matérn covariance on a dataset generated by random evaluations of
-the index function, as done at the initialization stage of the jellyfish search optimization. There exist several R packages, such as GpGp [@guinness2021gpgp] or ExaGeoStatR [@abdulah2023large], to
-fit the hyperparameters of a GP covariance function on data. In this project, we make use of the GpGp package.
+the index function, as done at the initialization stage of the jellyfish
+search optimization. There exist several R packages, such as GpGp
+[@guinness2021gpgp] or ExaGeoStatR [@abdulah2023large], to fit the
+hyperparameters of a GP covariance function on data. In this project, we
+make use of the GpGp package.
 
-After fitting the smoothness parameter $\nu>0$, its value can be interpreted as follows: the higher $\nu$, the smoother the index function.
+After fitting the smoothness parameter $\nu>0$, its value can be
+interpreted as follows: the higher $\nu$, the smoother the index
+function.
 
@@
 
-Squintability has been a useful concept. Here, it is defined mathematically and two approaches to compute this metric numerically, are described.
+Squintability has been a useful concept. Here, it is defined
+mathematically and two approaches to compute this metric numerically,
+are described.
 
-In the existing literature, it is commonly understood that a large squint angle implies that the objective function value is close to optimal even when
-we are not very close to the perfect view to see the structure.
-A small squint angle means that the PP index value improves substantially only when we are very close to the perfect view.
-As such, low squintability implies rapid improvement in the index value when near the perfect view.
-For PP, a small squint angle is considered to be undesirable because it means that the optimiser needs to be very close to be able to "see" the optima.
-Thus, it could be difficult for the optimiser to find the optima.
-We now propose a mathematical formulation of this intuition.
+In the existing literature, it is commonly understood that a large
+squint angle implies that the objective function value is close to
+optimal even when we are not very close to the perfect view to see the
+structure. A small squint angle means that the PP index value improves
+substantially only when we are very close to the perfect view. As such,
+low squintability implies rapid improvement in the index value when near
+the perfect view. For PP, a small squint angle is considered to be
+undesirable because it means that the optimiser needs to be very close
+to be able to "see" the optima. Thus, it could be difficult for the
+optimiser to find the optima. We now propose a mathematical formulation
+of this intuition.
 
-Let $A \in \mathbf{R}^{p \times 2}$ be the 2D projection basis and $A^*$ the optimal basis achieving the maximum index value for a given projection pursuit problem defined by the shape to find, data dimension, and the index function $f(.)$.
-The projection distance $d(A, A^*)$, is defined as the Frobenius norm $\lVert .
-\rVert _F$, between the squared difference of these two bases:
+Let $A \in \mathbf{R}^{p \times 2}$ be the 2D projection basis and $A^*$
+the optimal basis achieving the maximum index value for a given
+projection pursuit problem defined by the shape to find, data dimension,
+and the index function $f(.)$. The projection distance $d(A, A^*)$, is
+defined as the Frobenius norm $\lVert . \rVert _F$, between the squared
+difference of these two bases:
 
@@
 
-The index value $f(XA)$ is defined as described in the introduction.
-We now focus our attention to how the index value $f(XA)$ changes with respect to the projection distance $d(A, A^*)$, over the course of the JSO.
+The index value $f(XA)$ is defined as described in the introduction. We
+now focus our attention to how the index value $f(XA)$ changes with
+respect to the projection distance $d(A, A^*)$, over the course of the
+JSO.
 
@@
 ```
-
-It is expected that for a PP index with high squint angle, the optimization (@eq-optimization) should make substantial progress early on.
-Conversely, for a PP index with low squint angle, it might take a long
-while for the optimization to make substantial progress, as the candidate projections would need to be very close to the optimal one for the structure of the index function to be visible enough to be amenable to efficient optimization.
-This observation suggests that the extreme values of $f'$ (the ones for which $f''=0$, assuming that $f$ is twice differentiable), and the projection distances for which these values are attained, should play an important role in the mathematical definition of squintability.
-Noticing that $f$ is expected to be a decreasing function, we propose the following squintability metric:
+It is expected that for a PP index with high squint angle, the
+optimization (@eq-optimization) should make substantial progress early
+on. Conversely, for a PP index with low squint angle, it might take a
+long while for the optimization to make substantial progress, as the
+candidate projections would need to be very close to the optimal one for
+the structure of the index function to be visible enough to be amenable
+to efficient optimization. This observation suggests that the extreme
+values of $f'$ (the ones for which $f''=0$, assuming that $f$ is twice
+differentiable), and the projection distances for which these values are
+attained, should play an important role in the mathematical definition
+of squintability. Noticing that $f$ is expected to be a decreasing
+function, we propose the following squintability metric:
 
@@
 
-On the one hand, $-\min_{x}f'(x)$ represents the largest value of the gradient of $-f$.
-On the other hand, $\underset{x}{\arg\min}f'(x)$ represents the projection distance at which this largest gradient is attained.
-Based on the above discussion, it is expected that these two values should be both high in the case of high squintability (fast increase in $f$ early on), and both low in the case of low squintability (any substantial increase in $f$ happens very late, close to the optimal angle).
-This suggest that their product (@eq-squintability) should provide a sensible measure of squintability.
-The multiplicative constant $4$, which can be deemed arbitrary, does not change the interpretation of the squintability metric $\varsigma$; it is here to adjust the range of values of $\varsigma$ and simplify the explicit formula for $\varsigma$ obtained later on.
+On the one hand, $-\min_{x}f'(x)$ represents the largest value of the
+gradient of $-f$. On the other hand, $\underset{x}{\arg\min}f'(x)$
+represents the projection distance at which this largest gradient is
+attained. Based on the above discussion, it is expected that these two
+values should be both high in the case of high squintability (fast
+increase in $f$ early on), and both low in the case of low squintability
+(any substantial increase in $f$ happens very late, close to the optimal
+angle). This suggest that their product (@eq-squintability) should
+provide a sensible measure of squintability. The multiplicative constant
+$4$, which can be deemed arbitrary, does not change the interpretation
+of the squintability metric $\varsigma$; it is here to adjust the range
+of values of $\varsigma$ and simplify the explicit formula for
+$\varsigma$ obtained later on.
 
-To compute the squintability metric (@eq-squintability) in practice, several approaches are possible.
-The first one is to propose a parametric model for $f$, and use it to obtain an explicit formula for $\varsigma$.
-Numerical experiments suggest a scaled sigmoid shape as
+To compute the squintability metric (@eq-squintability) in practice,
+several approaches are possible. The first one is to propose a
+parametric model for $f$, and use it to obtain an explicit formula for
+$\varsigma$. Numerical experiments suggest a scaled sigmoid shape as
 described below. Define
@@
 
-which is a decreasing logistic function depending on two parameters $\theta_2$ and $\theta_3$, such that $\ell(\theta_{2})=\frac{1}{2}$. Then, define
+which is a decreasing logistic function depending on two parameters
+$\theta_2$ and $\theta_3$, such that $\ell(\theta_{2})=\frac{1}{2}$.
+Then, define
 
@@
 
-which depends on three additional parameters, $\theta_1$, $\theta_2$, and $x_{\max}$, such that $f(0)=\theta_1$ and $f(x_{\max})=\theta_4$.
-Under the parametric model (@eq-parametric), the squintability metric (@eq-squintability) can be shown to be equal to
+which depends on three additional parameters, $\theta_1$, $\theta_2$,
+and $x_{\max}$, such that $f(0)=\theta_1$ and $f(x_{\max})=\theta_4$.
+Under the parametric model (@eq-parametric), the squintability metric
+(@eq-squintability) can be shown to be equal to
 
@@
 
-In practice, the parameters of this model (@eq-parametric) can be estimated numerically, for example by non-linear least squares, and then
+In practice, the parameters of this model (@eq-parametric) can be
+estimated numerically, for example by non-linear least squares, and then
 used to evaluate $\varsigma$ as in equation
@@
 
-Alternatively, one can estimate (@eq-squintability) in a nonparametric way, for example by fitting $f$ using kernel regression, then numerically estimate the angle at which $-f'$ attains its highest value.
+Alternatively, one can estimate (@eq-squintability) in a nonparametric
+way, for example by fitting $f$ using kernel regression, then
+numerically estimate the angle at which $-f'$ attains its highest value.
 
@@
 <!-- The speed of optimizing an index function can be measured empirically by -->
+
 <!-- recording the duration of the optimisation. Alternatively, one can also -->
+
 <!-- gauge the speed in terms of computational complexity of evaluating the -->
+
 <!-- index function (for instance, in big O notation, with respect to the -->
+
 <!-- sample size) of computing the index function. -->
@@
 
-The JSO performance is compared with the existing Creeping Random Search (CRS) [@RJ-2021-105;
-@laa_using_2020] currently used in guided tour, and explores how JSO behaves with different hyper-parameter (number of jellies, maximum tries) and data dimension combinations. A second simulation
-studies the effect of index properties (smoothness and squintability), along with JSO hyper-parameters, and data dimension, on the success rate of the JSO performance.
+The JSO performance is compared with the existing Creeping Random Search
+(CRS) [@RJ-2021-105; @laa_using_2020] currently used in guided tour, and
+explores how JSO behaves with different hyper-parameter (number of
+jellies, maximum tries) and data dimension combinations. A second
+simulation studies the effect of index properties (smoothness and
+squintability), along with JSO hyper-parameters, and data dimension, on
+the success rate of the JSO performance.
 
@@
 
-The performance of JSO is investigated in comparison with the existing optimiser, CRS, with respect to two hyper-parameters (number of jellyfishes and maximum number of tries).
-The performance is measured by success rate, defined as the proportion of simulations that achieves a final index value within 0.05 of the best index value found among all 50 simulations.
-This comparison is based on projection pursuit to find the pipe shape investigated by @laa_using_2020 using two simple and well-behaved indexes, the `holes` index, and the `splines2d` index.
-Fifty simulations are conducted with both JSO and CRS, in four different data dimensions ($d = 6, 8, 10, 12$).
-JSO uses 100 jellyfishes with 100 maximum number of tries, while the CRS allows a maximum of 1000 samples at each iteration before the algorithm terminates.
-The different numbers accounts for the multiple paths of JSO to enable fairer comparison with CRS.
-The results of the simulation are collected using the data structure proposed in @RJ-2021-105 is used here for assessing JSO, where the design parameters are stored along with index value, projection basis, random seed, and computation time.
+The performance of JSO is investigated in comparison with the existing
+optimiser, CRS, with respect to two hyper-parameters (number of
+jellyfishes and maximum number of tries). The performance is measured by
+success rate, defined as the proportion of simulations that achieves a
+final index value within 0.05 of the best index value found among all 50
+simulations. This comparison is based on projection pursuit to find the
+pipe shape investigated by @laa_using_2020 using two simple and
+well-behaved indexes, the `holes` index, and the `splines2d` index.
+Fifty simulations are conducted with both JSO and CRS, in four different
+data dimensions ($d = 6, 8, 10, 12$). JSO uses 100 jellyfishes with 100
+maximum number of tries, while the CRS allows a maximum of 1000 samples
+at each iteration before the algorithm terminates. The different numbers
+accounts for the multiple paths of JSO to enable fairer comparison with
+CRS. The results of the simulation are collected using the data
+structure proposed in @RJ-2021-105 is used here for assessing JSO, where
+the design parameters are stored along with index value, projection
+basis, random seed, and computation time.
 <!--A sample for the sine-wave structure in 6D data using a spline index function (`splines2d`) is printed below:
@@
 
-In this simulation, 50 jellies (`n_jellies`) and
-a maximum of 50 tries (`max_tries`) are used.
-The full data contains 50 simulations, indexed by the variable `sim`, with the associated `seed`s.
-Recorded in columns `basis`, `index_val`, `tries`, `loop` and `time`, are the projection basis, index value, iteration ID, jellyfish ID, and computational time, respectively.
+In this simulation, 50 jellies (`n_jellies`) and a maximum of 50 tries
+(`max_tries`) are used. The full data contains 50 simulations, indexed
+by the variable `sim`, with the associated `seed`s. Recorded in columns
+`basis`, `index_val`, `tries`, `loop` and `time`, are the projection
+basis, index value, iteration ID, jellyfish ID, and computational time,
+respectively.
 <!-- The basis column records every visited basis before -->
@@
 
-In the second experiment, 50 simulations are conducted at each hyper-parameter combination to analyse its effect on the JSO success rate: number of jellies (20, 50, and 100), and  maximum number of  tries (50 and 100).
-@fig-success-rate illustrates the calculation based on the projection found by the above simulation.
-Here, the optimal index value is `r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`.
+In the second experiment, 50 simulations are conducted at each
+hyper-parameter combination to analyse its effect on the JSO success
+rate: number of jellies (20, 50, and 100), and maximum number of tries
+(50 and 100). @fig-success-rate illustrates the calculation based on the
+projection found by the above simulation. Here, the optimal index value
+is
+`r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`.
 Out of the 50 simulations,
-`r sim_sine_6d_spline_best |> filter(index_val >= aa - 0.05) |> nrow() -> bb; bb` achieved an index value within \[`r aa - 0.05`, `r aa`\], resulting in a success rate of `r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.
+`r sim_sine_6d_spline_best |> filter(index_val >= aa - 0.05) |> nrow() -> bb; bb`
+achieved an index value within \[`r aa - 0.05`, `r aa`\], resulting in a
+success rate of
+`r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.
 
 <!-- @fig-projection shows the best projection found in each simulation, -->
+
 <!-- sorted by the index value. Most of the projections show a clear -->
+
 <!-- sine-wave structure, while the last six show only a straight line -->
+
 <!-- with clusters at the end. When local optima exist in the projection -->
+
 <!-- pursuit problem, plotting the projections found in each simulation -->
+
 <!-- or by each jellyfish can help identify these local optima. -->
@@
 
-This simulation uses two different data shapes, pipe and a sine wave, investigated in 6D and 8D spaces with six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`.
-The two JSO hyper-parameters are varied.
-A total of `r nrow(sim_df)` combinations result, comprising of `r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and `r nrow(filter(sim_df, index != "holes"))` on the sine-wave data.
-For each case, JSO is run 50 times to calculate the success rate.
+This simulation uses two different data shapes, pipe and a sine wave,
+investigated in 6D and 8D spaces with six different PP indexes:
+`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`. The two
+JSO hyper-parameters are varied. A total of `r nrow(sim_df)`
+combinations result, comprising of
+`r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and
+`r nrow(filter(sim_df, index != "holes"))` on the sine-wave data. For
+each case, JSO is run 50 times to calculate the success rate.
 
-Smoothness and squintability are computed for each run, following the procedures outlined in @sec-smoothness and @sec-squintability and as illustrated in @fig-smoothness and  @fig-squintability.
+Smoothness and squintability are computed for each run, following the
+procedures outlined in @sec-smoothness and @sec-squintability and as
+illustrated in @fig-smoothness and @fig-squintability.
 
-To compute smoothness, 300 random bases are simulated.
-Index values and projection distance (to the optimal basis) are calculated for each random basis before fitting the Gaussian process model to obtain the smoothness measure for the index.
+To compute smoothness, 300 random bases are simulated. Index values and
+projection distance (to the optimal basis) are calculated for each
+random basis before fitting the Gaussian process model to obtain the
+smoothness measure for the index.
 
-To compute squintability, 50 random bases are sampled and interpolated to the optimal basis with a step size of 0.005.
-Index values and projection distances are calculated for these interpolated bases and the index values are averaged with a bin width of 0.005.
-The squintability measure is then calculated by fitting a four-parameter scaled logistic function of index value against projection distance, estimated by non-linear least squares.
+To compute squintability, 50 random bases are sampled and interpolated
+to the optimal basis with a step size of 0.005. Index values and
+projection distances are calculated for these interpolated bases and the
+index values are averaged with a bin width of 0.005. The squintability
+measure is then calculated by fitting a four-parameter scaled logistic
+function of index value against projection distance, estimated by
+non-linear least squares.
 
-A generalised linear model is fitted to the resulting data, using a binomial family and a logit link function.
-The data is pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10 for interpretation, 2) creating a new binary variable `long_time` to indicate cases with an average run time over 30 seconds, and 3) re-coding the success rate for
-the `stringy` index as 0, because none of the 50 simulations correctly identified the sine-wave shape.
+A generalised linear model is fitted to the resulting data, using a
+binomial family and a logit link function. The data is pre-processed
+by 1) scaling the JSO hyper-parameters by a factor of 10 for
+interpretation, 2) creating a new binary variable `long_time` to
+indicate cases with an average run time over 30 seconds, and 3)
+re-coding the success rate for the `stringy` index as 0, because none of
+the 50 simulations correctly identified the sine-wave shape.
 
 <!-- **Projection distance**: The projection distance between projection -->
+
 <!-- bases and the optimal best basis (or the empirical best basis if -->
+
 <!-- the optimal is unknown) measures how close a visited basis is to  -->
+
 <!-- the best basis and it is calculated as the Frobenius norm in Equation -->
+
 <!-- @eq-proj-dist. Plotting projection distances against index values for the  -->
+
 <!-- bases visited reveals how the index value changes as the basis -->
+
 <!-- approaches the optimal basis, helping to calculate the squintability -->
+
 <!-- measure. -->
@@
 
-The results from the two simulations described in @sec-sim-deets are analysed based on the final projections across two optimisers (JSO and CRS) and the success rate across JSO hyper-parameters.
-Smoothness and squintability are calculated across a collection of pipe-finding and sine-wave finding problems to construct the relationship between success rate, JSO hyper-parameters, and index properties.
+The results from the two simulations described in @sec-sim-deets are
+analysed based on the final projections across two optimisers (JSO and
+CRS) and the success rate across JSO hyper-parameters. Smoothness and
+squintability are calculated across a collection of pipe-finding and
+sine-wave finding problems to construct the relationship between success
+rate, JSO hyper-parameters, and index properties.
 
-The final projections found by the two optimisers (JSO and CRS) are presented in @fig-proj, broken down by 10th quantile, faceted by the data dimension.
-The JSO uses 100 jellyfishes and a maximum number of tries of 100.
-The CRS uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates.
-In the 6-dimensional data scenario, JSO consistently identifies a clear pipe shape.
-The CRS also finds the pipe shape but with a wide rim, suggesting a further polish search may be required.
-With increasing dimensions, JSO may not always identify the pipe shape due to random sampling, but it still presents a pipe shape in over 50% of cases.
-When compared to the CRS, JSO reaches higher index values and clearer pipe shapes across all quantiles.
-This suggests JSO can be more effective when searching higher dimension spaces, which may be preferred by users.
+The final projections found by the two optimisers (JSO and CRS) are
+presented in @fig-proj, broken down by 10th quantile, faceted by the
+data dimension. The JSO uses 100 jellyfishes and a maximum number of
+tries of 100. The CRS uses a maximum of 1000 tries in each step of
+random sampling step before the algorithm terminates. In the
+6-dimensional data scenario, JSO consistently identifies a clear pipe
+shape. The CRS also finds the pipe shape but with a wide rim, suggesting
+a further polish search may be required. With increasing dimensions, JSO
+may not always identify the pipe shape due to random sampling, but it
+still presents a pipe shape in over 50% of cases. When compared to the
+CRS, JSO reaches higher index values and clearer pipe shapes across all
+quantiles. This suggests JSO can be more effective when searching higher
+dimension spaces, which may be preferred by users.
 
-The success rate calculated for each hyper-parameter combination is presented in @fig-proportion.
-As the number of jellyfishes and maximum tries increase, the success rate also increases.
-For simpler problems (6 dimensions), small parameter values (20 jellyfishes and a maximum of 50 tries) can achieve a high success rate.
-However, for higher-dimensional problems (8, 10, and 12 dimensions), larger parameter values (i.e.
-100 jellyfishes and a maximum of 100 tries) are necessary.
-Increasing both parameters enhances the performance of JSO, but it also extends the computational time required for the optimisation, which can be computationally intensive when evaluating the index function (such as scagnostic indexes) multiple times across numerous iterations.
+The success rate calculated for each hyper-parameter combination is
+presented in @fig-proportion. As the number of jellyfishes and maximum
+tries increase, the success rate also increases. For simpler problems (6
+dimensions), small parameter values (20 jellyfishes and a maximum of 50
+tries) can achieve a high success rate. However, for higher-dimensional
+problems (8, 10, and 12 dimensions), larger parameter values (i.e. 100
+jellyfishes and a maximum of 100 tries) are necessary. Increasing both
+parameters enhances the performance of JSO, but it also extends the
+computational time required for the optimisation, which can be
+computationally intensive when evaluating the index function (such as
+scagnostic indexes) multiple times across numerous iterations.
 
@@
 <!-- The low -->
+
 <!-- squint value of MIC and TIC index is due to its convex shape of the -->
+
 <!-- index value against the projection distance, as shown in -->
+
 <!-- @fig-idx-proj-dist. Comparing to other concave shapes, the maximum first -->
+
 <!-- derivative happens at a smaller projection distance ($\theta_2$), -->
+
 <!-- requiring the optimiser to get closer to the optimal to see a -->
+
 <!-- significant change in the index value, hence a small squintability -->
+
 <!-- measure.  -->
 
-The index properties, including smoothness and squintability, offer numerical metrics to characterise the complexity of projection pursuit optimisation problems.
-@tbl-smoothness-squintability presents
-the parameters estimated from the Gaussian process (variance, range, smooth, and nugget) and the scaled
-logistic function ($\theta_1$ to $\theta_4$) for calculating both metrics for each case considered in the simulation.
-The column "smooth" is used as the smoothness metric and the column "squint" is calculated as equation @eq-squintability-parametric as the squintability metric.
-@tbl-mod-output presents the results of fitting a generalised linear model with a binomial family and a logit link function to the data in @tbl-mod-data, where all the three components (success rate, JS hyper-parameters, and index properties) are combined.
-The model suggests that JSO success rate is positively associated with the two hyper-parameters, as well as with the index properties: smoothness and squintability.
-Specifically,
-using 10 more jellyfishes and 10 more tries increases the odd ratio of success by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`%
-and `r round(100 * (exp(broom::tidy(mod1)$estimate[7])-1), 2)`%,
-respectively.
-However, being flagged with long runtime and an increase of data dimension reduce the success rate by `r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and `r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively.
-The variable `squintability` and `dimension` are significant, suggesting their importance relative to JSO hyper-parameters in the optimisation success.
-In defining a projection pursuit problem, factors such as the shape-to-find, the index function used, and the data dimension, determine the properties like smoothness and squintability.
-These metrics can then be compared across different problems to understand their relative complexity.
-The results suggest that squintability has a more significant influence than smoothness on the success rate of JSO.
-Once the characteristics of the projection pursuit problem are fully understood, increasing JSO hyper-parameters can enhance the search effectiveness, however, it is also important to consider the resulting increase in computational complexity.
+The index properties, including smoothness and squintability, offer
+numerical metrics to characterise the complexity of projection pursuit
+optimisation problems. @tbl-smoothness-squintability presents the
+parameters estimated from the Gaussian process (variance, range, smooth,
+and nugget) and the scaled logistic function ($\theta_1$ to $\theta_4$)
+for calculating both metrics for each case considered in the simulation.
+The column "smooth" is used as the smoothness metric and the column
+"squint" is calculated as equation @eq-squintability-parametric as the
+squintability metric. @tbl-mod-output presents the results of fitting a
+generalised linear model with a binomial family and a logit link
+function to the data in @tbl-mod-data, where all the three components
+(success rate, JS hyper-parameters, and index properties) are combined.
+The model suggests that JSO success rate is positively associated with
+the two hyper-parameters, as well as with the index properties:
+smoothness and squintability. Specifically, using 10 more jellyfishes
+and 10 more tries increases the odd ratio of success by
+`r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`% and
+`r round(100 * (exp(broom::tidy(mod1)$estimate[7])-1), 2)`%,
+respectively. However, being flagged with long runtime and an increase
+of data dimension reduce the success rate by
+`r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and
+`r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively.
+The variable `squintability` and `dimension` are significant, suggesting
+their importance relative to JSO hyper-parameters in the optimisation
+success. In defining a projection pursuit problem, factors such as the
+shape-to-find, the index function used, and the data dimension,
+determine the properties like smoothness and squintability. These
+metrics can then be compared across different problems to understand
+their relative complexity. The results suggest that squintability has a
+more significant influence than smoothness on the success rate of JSO.
+Once the characteristics of the projection pursuit problem are fully
+understood, increasing JSO hyper-parameters can enhance the search
+effectiveness, however, it is also important to consider the resulting
+increase in computational complexity.
 
@@
 
-This paper has presented new metrics to mathematically define desirable features of PP indexes, squintability and smoothness, and used these to assess the performance of the new jellyfish search optimiser.
-The metrics will be generally useful for characterising PP indexes, and halp with developing nex indexes.
+This paper has presented new metrics to mathematically define desirable
+features of PP indexes, squintability and smoothness, and used these to
+assess the performance of the new jellyfish search optimiser. The
+metrics will be generally useful for characterising PP indexes, and halp
+with developing nex indexes.
 
-In the comparison of the JSO against the currently used CRS, as expected the JSO vastly out-performs CRS, and provides a high probability of finding the global optima.
-The JSO obtains the maximum more cleanly, with a slightly higher index value, and plot of the projected data showing the structure more clearly.
+In the comparison of the JSO against the currently used CRS, as expected
+the JSO vastly out-performs CRS, and provides a high probability of
+finding the global optima. The JSO obtains the maximum more cleanly,
+with a slightly higher index value, and plot of the projected data
+showing the structure more clearly.
 
-The JSO performance is affected by the hyper-parameters, with a higher chance of reaching the global optima when more jellies are used and the maximum number of tries is increased.
-However, it comes at a computational cost, as expected.
-The performance declines if the projection dimension increases and if the PP index has low squintability.
-The higher the squintability the better chance the JSO can find the optima.
-However, interestingly smoothness does not affect the JSO performance.
+The JSO performance is affected by the hyper-parameters, with a higher
+chance of reaching the global optima when more jellies are used and the
+maximum number of tries is increased. However, it comes at a
+computational cost, as expected. The performance declines if the
+projection dimension increases and if the PP index has low
+squintability. The higher the squintability the better chance the JSO
+can find the optima. However, interestingly smoothness does not affect
+the JSO performance.
 
-The new JSO is integrated with the current implementation of the projection pursuit guided tour in the `tourr` package, and can be further examined using PP optimisation diagnostics in the `ferrn` package.
-Using the JSO is a little different than the current CRS, because it runs many paths.
-The recommended approach is to conduct the optimisation off-line, and then use the guided tour to follow selected jellies.
-The tools for this are available in the `tourr` package, too.
+The new JSO is integrated with the current implementation of the
+projection pursuit guided tour in the `tourr` package, and can be
+further examined using PP optimisation diagnostics in the `ferrn`
+package. Using the JSO is a little different than the current CRS,
+because it runs many paths. The recommended approach is to conduct the
+optimisation off-line, and then use the guided tour to follow selected
+jellies. The tools for this are available in the `tourr` package, too.
 
+```{=html}
 <!--
@@
 -->
-
+```
 # Acknowledgement
@@
 The article is created using Quarto [@Allaire_Quarto_2022] in R [@R].
-The source code for reproducing the work reported in this paper can be found at: <https://github.com/huizezhang-sherry/paper-optim>.
-Nicolas Langrené acknowledges the partial support of the Guangdong Provincial Key Laboratory IRADS (2022B1212010006, R0400001-22) and the UIC Start-up Research Fund UICR0700041-22.
+The source code for reproducing the work reported in this paper can be
+found at: <https://github.com/huizezhang-sherry/paper-optim>. Nicolas
+Langrené acknowledges the partial support of the Guangdong Provincial
+Key Laboratory IRADS (2022B1212010006, R0400001-22) and the UIC Start-up
+Research Fund UICR0700041-22.
 

### 2024-07-01 optim.tex
update abstract

@@
 \affiliation[3]{organization={Guangdong Provincial Key Laboratory of
-Interdisciplinary Research and Application for Data Science, BNU-HKBU
-United International College, Department of Mathematical
+Interdisciplinary Research and Application for Data Science,
+BNU-HKBU~United~International~College, Department of Mathematical
 Sciences},city={Zhuhai},country={China},countrysep={,},postcode={519087},postcodesep={}}
@@
 data by revealing interesting projections. The optimisation in PP can be
-non-trivial, involving non-smooth functions and optima with small a
+non-trivial, involving non-smooth functions and optima with a small
 ``squint angle'', detectable only from close proximity. To address these
@@
 existing optimisers. Additionally, this work proposes novel methods to
-quantify two properties of the PP index -- smoothness and squintability
--- that capture the complexities inherent in PP optimisation problems.
-These two metrics are evaluated with JSO hyper-parameters to determine
-their effects on JSO success rate. The JSO algorithm has been
-implemented in the \texttt{tourr} package and functions to calculate
-smoothness and squintability of indexes are available in the
-\texttt{ferrn} package.
+quantify two properties of the PP index --~smoothness and
+squintability~-- that capture the complexities inherent in PP
+optimisation problems. These two metrics are evaluated along with JSO
+hyper-parameters to determine their effects on JSO success rate. Our
+numerical results confirm the positive impact of these metrics on the
+JSO success rate, with squintability being the most significant. The JSO
+algorithm has been implemented in the \texttt{tourr} package and
+functions to calculate smoothness and squintability of indexes are
+available in the \texttt{ferrn} package.
 \end{abstract}
@@
 \end{frontmatter}
-
-\section{Introduction}\label{introduction}
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, boxrule=0pt, enhanced, sharp corners, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
+\hypertarget{sec-background}{%
 \section{Projection pursuit, tours, index functions and
-optimisation}\label{sec-background}
+optimisation}\label{sec-background}}
 
@@
 
+\hypertarget{sec-theory}{%
 \section{The jellyfish optimiser and properties of PP
-indexes}\label{sec-theory}
+indexes}\label{sec-theory}}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, coltitle=black, colback=white, colbacktitle=quarto-callout-note-color!10!white, left=2mm, leftrule=.75mm, titlerule=0mm, rightrule=.15mm, toptitle=1mm, opacityback=0, arc=.35mm, bottomrule=.15mm, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, opacitybacktitle=0.6, breakable]
+\begin{tcolorbox}[enhanced jigsaw, coltitle=black, toprule=.15mm, colback=white, breakable, colbacktitle=quarto-callout-note-color!10!white, bottomrule=.15mm, arc=.35mm, leftrule=.75mm, opacityback=0, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, toptitle=1mm, titlerule=0mm, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, opacitybacktitle=0.6]
 
@@
 
-\begin{equation}\phantomsection\label{eq-proj-dist}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .}\end{equation}
+\begin{equation}\protect\hypertarget{eq-proj-dist}{}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .}\label{eq-proj-dist}\end{equation}
 
@@
 
-\begin{equation}\phantomsection\label{eq-squintability}{
+\begin{equation}\protect\hypertarget{eq-squintability}{}{
 \varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ .
-}\end{equation}
+}\label{eq-squintability}\end{equation}
 
@@
 
-\begin{equation}\phantomsection\label{eq-squintability-parametric}{
+\begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
 \varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
-}\end{equation}
+}\label{eq-squintability-parametric}\end{equation}
 
@@
 
-\section{Simulation details}\label{sec-sim-deets}
+\hypertarget{sec-sim-deets}{%
+\section{Simulation details}\label{sec-sim-deets}}
 
@@
 
-\subsection{Performance of JSO relative to CRS}\label{sec-app-1}
+\hypertarget{sec-app-1}{%
+\subsection{Performance of JSO relative to CRS}\label{sec-app-1}}
 
@@
 
-\centering{
-
-\includegraphics[width=4.52in,height=\textheight]{figures/success-rate.png}
+{\centering \includegraphics[width=4.52in,height=\textheight]{figures/success-rate.png}
 
@@
 
-\end{figure}%
-
+\end{figure}
+
+\hypertarget{sec-app-2}{%
 \subsection{Factors affecting JSO success rate: index properties and
-jellyfish hyper-parameters}\label{sec-app-2}
+jellyfish hyper-parameters}\label{sec-app-2}}
 
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{figures/smoothness.png}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/smoothness.png}
 
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{figures/squintability.png}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/squintability.png}
 
@@
 
-\end{figure}%
-
-\section{Results}\label{sec-sim-res}
+\end{figure}
+
+\hypertarget{sec-sim-res}{%
+\section{Results}\label{sec-sim-res}}
 
@@
 \endgroup{}
-
-}
-
-\end{table}%
-
+\end{table}
+
+\hypertarget{tbl-mod-data}{}
 \begin{table}
-
-\caption{\label{tbl-mod-data}The first 7 rows of the datasets processed
-for modelling.}
-
-\centering{
+\caption{\label{tbl-mod-data}The first 7 rows of the datasets processed for modelling. }\tabularnewline
 
@@
 \endgroup{}
-
-}
-
-\end{table}%
-
+\end{table}
+
+\hypertarget{tbl-mod-output}{}
 \begin{table}
-
-\caption{\label{tbl-mod-output}Model estimates of proportion of
-jellyfish success on index properties and jellyfish hyper-parameters
-from the generalised linear model with a binomial family and a logit
-link function. The variable smoothness, squintability, number of
-jellyfish and maximum number of tries are positively associated with JSO
-success rate while data dimension and being flagged as long runtime are
-negatively associated with the success rate. The variable squintability
-and dimension are significant, suggesting their importance relative to
-jellyfish hyper-parameters in the optimisation success.}
-
-\centering{
+\caption{\label{tbl-mod-output}Model estimates of proportion of jellyfish success on index properties
+and jellyfish hyper-parameters from the generalised linear model with a
+binomial family and a logit link function. The variable smoothness,
+squintability, number of jellyfish and maximum number of tries are
+positively associated with JSO success rate while data dimension and
+being flagged as long runtime are negatively associated with the success
+rate. The variable squintability and dimension are significant,
+suggesting their importance relative to jellyfish hyper-parameters in
+the optimisation success. }\tabularnewline
 
@@
 \endgroup{}
-
-}
-
-\end{table}%
-
-\section{Conclusion}\label{sec-conclusion}
+\end{table}
+
+\hypertarget{sec-conclusion}{%
+\section{Conclusion}\label{sec-conclusion}}
 
@@
 
-\section{Acknowledgement}\label{acknowledgement}
+\hypertarget{acknowledgement}{%
+\section{Acknowledgement}\label{acknowledgement}}
 

### 2024-07-01 optim.qmd
small editsg

@@
 PP, improving optimisation (e.g. @posse1995, @marie-sainte2010,
-@grochowski2011) and extending it's use to visually explore
+@grochowski2011) and extending its use to visually explore
 high-dimensional data. @cook1995grand introduced the PP guided tour,
@@
 diagnose optimisation processes, particularly focusing on the guided
-tour, and revealed a need to for improved optimisation. While improving
+tour, and revealed a need for improved optimisation. While improving
 the quality of the optimisation solutions in the tour is essential, it
@@
 behaviour of JSO under different settings and its sensitivity to
-hyper-parameter choices (number of jellies and maximum number of tries).
+hyper-parameter choices (number of jellyfishes and maximum number of tries).
 The relationship between the JSO performance, hyper-parameter tuning and
@@
 details two simulation experiments to assess JSO's performance: one
-comparing JSO's permformance improvements relative to an existing
+comparing JSO's performance improvements relative to an existing
 optimiser, Creeping Random Search (CRS), and the other studying the
@@
 the specific implementation details can vary depending on the version of
-the algorithm or its application, we focus on two main parameters that
+the algorithm or its application, the focus is on two main parameters that
 are most relevant to our application: the number of jellyfish and the
@@
 invariance, and speed). Since not all the properties affects the
-execution of the optimisation, we choose to focus here on the first two
+execution of the optimisation, the focus here is on the first two
 properties, namely *smoothness* (@sec-smoothness) and *squintability*
@@
 
-In our context, we suggest to use this parameter $\nu$ as a measure of
+In this context, the parameter $\nu$ is suggested as a measure of
 the smoothness of the index function by fitting a Gaussian process prior
@@
 squint angle implies that the objective function value is close to
-optimal even when we are not very close to the perfect view to see the
-structure. A small squint angle means that the PP index value improves
-substantially only when we are very close to the perfect view. As such,
+optimal even when the perfect view to see the structure is far away.
+A small squint angle means that the PP index value improves
+substantially only when the perfect view is close by. As such,
 low squintability implies rapid improvement in the index value when near
@@
 to be able to "see" the optima. Thus, it could be difficult for the
-optimiser to find the optima. We now propose a mathematical formulation
-of this intuition.
+optimiser to find the optima. The mathematical formulation
+of this intuition is proposed below:
 
@@
 of squintability. Noticing that $f$ is expected to be a decreasing
-function, we propose the following squintability metric:
+function, the squintability metric is proposed as follows:
 
@@
 explores how JSO behaves with different hyper-parameter (number of
-jellies, maximum tries) and data dimension combinations. A second
+jellyfishes and maximum tries) and data dimension combinations. A second
 simulation studies the effect of index properties (smoothness and
@@
 
-In this simulation, 50 jellies (`n_jellies`) and a maximum of 50 tries
+In this simulation, 50 jellyfishes (`n_jellies`) and a maximum of 50 tries
 (`max_tries`) are used. The full data contains 50 simulations, indexed
@@
 hyper-parameter combination to analyse its effect on the JSO success
-rate: number of jellies (20, 50, and 100), and maximum number of tries
+rate: number of jellyfishes (20, 50, and 100), and maximum number of tries
 (50 and 100). @fig-success-rate illustrates the calculation based on the
@@
 assess the performance of the new jellyfish search optimiser. The
-metrics will be generally useful for characterising PP indexes, and halp
-with developing nex indexes.
+metrics will be generally useful for characterising PP indexes, and help
+with developing new indexes.
 
@@
 The JSO performance is affected by the hyper-parameters, with a higher
-chance of reaching the global optima when more jellies are used and the
+chance of reaching the global optima when more jellyfishes are used and the
 maximum number of tries is increased. However, it comes at a
@@
 optimisation off-line, and then use the guided tour to follow selected
-jellies. The tools for this are available in the `tourr` package, too.
+jellyfish. The tools for this are available in the `tourr` package, too.
 

### 2024-07-01 optim.tex
small editsg

--- 2024-07-01 optim.tex
@@
 PP, improving optimisation (e.g. \citet{posse1995},
-\citet{marie-sainte2010}, \citet{grochowski2011}) and extending it's use
+\citet{marie-sainte2010}, \citet{grochowski2011}) and extending its use
 to visually explore high-dimensional data. \citet{cook1995grand}
@@
 optimisation processes, particularly focusing on the guided tour, and
-revealed a need to for improved optimisation. While improving the
-quality of the optimisation solutions in the tour is essential, it is
-also important to be able to view the data projections as the
-optimisation progresses. Integrating the guided tour with a global
-optimisation algorithm that is efficient in finding the global optimal
-and enables viewing of the projected data during the exploration process
-is a goal.
+revealed a need for improved optimisation. While improving the quality
+of the optimisation solutions in the tour is essential, it is also
+important to be able to view the data projections as the optimisation
+progresses. Integrating the guided tour with a global optimisation
+algorithm that is efficient in finding the global optimal and enables
+viewing of the projected data during the exploration process is a goal.
 
@@
 assess the behaviour of JSO under different settings and its sensitivity
-to hyper-parameter choices (number of jellies and maximum number of
+to hyper-parameter choices (number of jellyfishes and maximum number of
 tries). The relationship between the JSO performance, hyper-parameter
@@
 two simulation experiments to assess JSO's performance: one comparing
-JSO's permformance improvements relative to an existing optimiser,
+JSO's performance improvements relative to an existing optimiser,
 Creeping Random Search (CRS), and the other studying the impact of
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, coltitle=black, colback=white, colbacktitle=quarto-callout-note-color!10!white, left=2mm, leftrule=.75mm, titlerule=0mm, rightrule=.15mm, toptitle=1mm, opacityback=0, arc=.35mm, bottomrule=.15mm, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, opacitybacktitle=0.6, breakable]
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, rightrule=.15mm, breakable, opacityback=0, toptitle=1mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, bottomrule=.15mm, arc=.35mm, toprule=.15mm, bottomtitle=1mm, titlerule=0mm, opacitybacktitle=0.6, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, left=2mm]
 
@@
 the specific implementation details can vary depending on the version of
-the algorithm or its application, we focus on two main parameters that
-are most relevant to our application: the number of jellyfish and the
-maximum number of trials.
+the algorithm or its application, the focus is on two main parameters
+that are most relevant to our application: the number of jellyfish and
+the maximum number of trials.
 
@@
 rotation invariance, and speed). Since not all the properties affects
-the execution of the optimisation, we choose to focus here on the first
-two properties, namely \emph{smoothness} (Section~\ref{sec-smoothness})
-and \emph{squintability} (Section~\ref{sec-squintability}), and propose
-two metrics to evaluate them.
+the execution of the optimisation, the focus here is on the first two
+properties, namely \emph{smoothness} (Section~\ref{sec-smoothness}) and
+\emph{squintability} (Section~\ref{sec-squintability}), and propose two
+metrics to evaluate them.
 
@@
 
-In our context, we suggest to use this parameter \(\nu\) as a measure of
-the smoothness of the index function by fitting a Gaussian process prior
+In this context, the parameter \(\nu\) is suggested as a measure of the
+smoothness of the index function by fitting a Gaussian process prior
 with Matérn covariance on a dataset generated by random evaluations of
@@
 squint angle implies that the objective function value is close to
-optimal even when we are not very close to the perfect view to see the
-structure. A small squint angle means that the PP index value improves
-substantially only when we are very close to the perfect view. As such,
-low squintability implies rapid improvement in the index value when near
-the perfect view. For PP, a small squint angle is considered to be
-undesirable because it means that the optimiser needs to be very close
-to be able to ``see'' the optima. Thus, it could be difficult for the
-optimiser to find the optima. We now propose a mathematical formulation
-of this intuition.
+optimal even when the perfect view to see the structure is far away. A
+small squint angle means that the PP index value improves substantially
+only when the perfect view is close by. As such, low squintability
+implies rapid improvement in the index value when near the perfect view.
+For PP, a small squint angle is considered to be undesirable because it
+means that the optimiser needs to be very close to be able to ``see''
+the optima. Thus, it could be difficult for the optimiser to find the
+optima. The mathematical formulation of this intuition is proposed
+below:
 
@@
 mathematical definition of squintability. Noticing that \(f\) is
-expected to be a decreasing function, we propose the following
-squintability metric:
+expected to be a decreasing function, the squintability metric is
+proposed as follows:
 
@@
 and explores how JSO behaves with different hyper-parameter (number of
-jellies, maximum tries) and data dimension combinations. A second
+jellyfishes and maximum tries) and data dimension combinations. A second
 simulation studies the effect of index properties (smoothness and
@@
 
-In this simulation, 50 jellies (\texttt{n\_jellies}) and a maximum of 50
-tries (\texttt{max\_tries}) are used. The full data contains 50
+In this simulation, 50 jellyfishes (\texttt{n\_jellies}) and a maximum
+of 50 tries (\texttt{max\_tries}) are used. The full data contains 50
 simulations, indexed by the variable \texttt{sim}, with the associated
@@
 hyper-parameter combination to analyse its effect on the JSO success
-rate: number of jellies (20, 50, and 100), and maximum number of tries
-(50 and 100). Figure~\ref{fig-success-rate} illustrates the calculation
-based on the projection found by the above simulation. Here, the optimal
-index value is 0.999. Out of the 50 simulations, 44 achieved an index
-value within {[}0.949, 0.999{]}, resulting in a success rate of 0.88.
+rate: number of jellyfishes (20, 50, and 100), and maximum number of
+tries (50 and 100). Figure~\ref{fig-success-rate} illustrates the
+calculation based on the projection found by the above simulation. Here,
+the optimal index value is 0.999. Out of the 50 simulations, 44 achieved
+an index value within {[}0.949, 0.999{]}, resulting in a success rate of
+0.88.
 
@@
 The JSO performance is affected by the hyper-parameters, with a higher
-chance of reaching the global optima when more jellies are used and the
-maximum number of tries is increased. However, it comes at a
+chance of reaching the global optima when more jellyfishes are used and
+the maximum number of tries is increased. However, it comes at a
 computational cost, as expected. The performance declines if the
@@
 optimisation off-line, and then use the guided tour to follow selected
-jellies. The tools for this are available in the \texttt{tourr} package,
-too.
+jellyfish. The tools for this are available in the \texttt{tourr}
+package, too.
 

### 2024-07-01 optim.tex
update illustration

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, rightrule=.15mm, breakable, opacityback=0, toptitle=1mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, bottomrule=.15mm, arc=.35mm, toprule=.15mm, bottomtitle=1mm, titlerule=0mm, opacitybacktitle=0.6, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, arc=.35mm, breakable, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, colback=white, opacityback=0, left=2mm, toprule=.15mm, opacitybacktitle=0.6, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm]
 
@@
 because it runs many paths. The recommended approach is to conduct the
-optimisation off-line, and then use the guided tour to follow selected
-jellyfish. The tools for this are available in the \texttt{tourr}
-package, too.
+optimisation off-line, extract the bases of a selected jellyfish, and
+then use the planned tour to follow selected jellyfish. The tools for
+this are available in the \texttt{tourr} package, too.
 

### 2024-07-15 jso.qmd
proofread

--- 2024-07-01 optim.qmd
@@
 Our numerical results confirm the positive impact of these metrics on the JSO success rate, with squintability being the most significant.
-The JSO algorithm has been implemented in the `tourr` package and functions to calculate smoothness and squintability of indexes are available in the `ferrn` package.
+The JSO algorithm has been implemented in the `tourr` package and functions to calculate smoothness and squintability are available in the `ferrn` package.
 keywords: 
@@
 technique aimed at identifying informative linear projections of data.
-The method involves optimising a objective function known as the PP
+The method involves optimising an objective function known as the PP
 index (e.g. @hall1989polynomial, @cook1993projection,
@@
 and efficient optimisation procedure is essential to explore the data
-landscape and achieving a globally optimal viewpoint of the data.
-
-Significant efforts have been devoted to enhancing the functionality of
-PP, improving optimisation (e.g. @posse1995, @marie-sainte2010,
-@grochowski2011) and extending its use to visually explore
-high-dimensional data. @cook1995grand introduced the PP guided tour,
-which monitors the optimisation visually to see the data projections
-leading in and out of the optima. An implementation is available in the
-R [@R] package `tourr` [@tourr]. @RJ-2021-105 illustrated how to
-diagnose optimisation processes, particularly focusing on the guided
-tour, and revealed a need for improved optimisation. While improving
-the quality of the optimisation solutions in the tour is essential, it
-is also important to be able to view the data projections as the
-optimisation progresses. Integrating the guided tour with a global
-optimisation algorithm that is efficient in finding the global optimal
-and enables viewing of the projected data during the exploration process
-is a goal.
+landscape and achieve a globally optimal viewpoint of the data.
+
+Optimisation of PP is typically investigated in the literature when new
+indexes are proposed [@posse1995; @marie-sainte2010; @grochowski2011] or
+when visualisation method are used to track the optimisation process.
+@cook1995grand introduced the PP guided tour, which monitors the
+optimisation visually to see the data projections leading in and out of
+the optima. An implementation is available in the `tourr` package
+[@tourr] in R [@R]. @RJ-2021-105 illustrated how to diagnose
+optimisation processes, particularly focusing on the guided tour, and
+revealed a need for improved optimisation. While improving the quality
+of the optimisation solutions in the tour is essential, it is also
+important to be able to view the data projections as the optimisation
+progresses. Integrating the guided tour with a global optimisation
+algorithm that is efficient in finding the global optimal and enables
+viewing of the projected data during the exploration process is a goal.
 
 Here, the potential for a Jellyfish Search Optimiser (JSO) to be
-integrated with the projection pursuit guided tour is explored. JSO
-[@chou_novel_2021] -[@rajwar_exhaustive_2023], inspired by the search
+integrated with the projection pursuit guided tour is explored. JSO,
+[@chou_novel_2021] - [@rajwar_exhaustive_2023], inspired by the search
 behaviour of jellyfish in the ocean, is a swarm-based metaheuristic
 designed to solve global optimisation problems. Compared to traditional
-methods, JSO demonstrated stronger search ability and faster
+methods, JSO has demonstrated stronger search ability and faster
 convergence, and requires fewer tuning parameters. These practical
@@
 mathematically define metrics for squintability and smoothness, which is
-new for the field. A series of simulation experiments using specially
-structured datasets and various PP indexes are conducted to assess the
-behaviour of JSO under different settings and its sensitivity to
-hyper-parameter choices (number of jellyfishes and maximum number of tries).
-The relationship between the JSO performance, hyper-parameter tuning and
-properties of PP indexes is analysed to provide guidance for
-practitioners using projection pursuit, and developers of new indexes.
+new for the field. A series of simulation experiments using various
+datasets and PP indexes are conducted to assess JSO's behaviour and its
+sensitivity to hyper-parameter choices (number of jellyfish and maximum
+number of tries). The relationship between the JSO performance,
+hyper-parameter choices and properties of PP indexes (smoothness and
+squintability) is analysed to provide guidance on selecting optimisers
+for practitioners using projection pursuit. Additionally, it aims to
+guide the design of new indexes that facilitate easy optimization for PP
+researchers.
 
@@
 functions in the literature. @sec-theory details JSO and introduces
-metrics that measure different properties of PP indexes. @sec-sim-deets
-details two simulation experiments to assess JSO's performance: one
-comparing JSO's performance improvements relative to an existing
-optimiser, Creeping Random Search (CRS), and the other studying the
-impact of different PP index properties on optimisation performance.
-@sec-sim-res presents the results. @sec-conclusion summarises the work
-and provides suggestions for future directions.
+metrics that measure different properties of PP indexes, smoothness and
+squintability. @sec-sim-deets details two simulation experiments to
+assess JSO's performance: one comparing JSO's performance improvements
+relative to an existing optimiser, Creeping Random Search (CRS), and the
+other studying the impact of different PP index properties on
+optimisation performance. @sec-sim-res presents the results.
+@sec-conclusion summarises the work and provides suggestions for future
+directions.
 
@@
     next index value fails to increase.
--   `search_better()`: is a brute-force optimisation searching randomly
-    for projections with higher index values.
+-   `search_better()`: also known as Creeping Random Search (CRS), is a
+    brute-force optimisation searching randomly for projections with
+    higher index values.
 -   `search_better_random()`: is essentially simulated annealing
@@
 dimension, and implemented because they are relatively easy to optimise.
-A goal is to be able to incorporate more complex PP indexes, for example
-based on scagnostics (@scag, @WW08).
+A goal is to be able to incorporate more complex PP indexes, for
+example, based on scagnostics (@scag, @WW08).
 
@@
 respectively. In the context of optimization, these movements are
-abstracted to explore the search space in a way that balances
-exploration (searching new areas) and exploitation (focusing on
-promising areas). The algorithm aims to find the optimal solution by
-adapting the behaviour of jellyfish to navigate towards the best
-solution over iterations [@chou_novel_2021].
+abstracted to explore the search space, aiming to balance exploration
+(searching new areas) and exploitation (focusing on promising areas).
+The algorithm aims to find the optimal solution by adapting the
+behaviour of jellyfish to navigate towards the best solution over
+iterations [@chou_novel_2021].
 
 To solve the optimisation problem embedded in the PP guided tour, a
-starting projection, an index function and the maximum number of
-iterations are provided as input. Then, the current projection is
-evaluated by the index function. The projection is then moved in a
-direction determined by a random factor, influenced by how far along we
-are in the optimization process. Occasionally, completely new directions
-may be taken like a jellyfish might with ocean currents. A new
-projection is accepted if it is an improvement compared to the current
-one, rejected otherwise. This process continues and iteratively improves
-the projection, until the pre-specified maximum number of trials is
-reached.
+starting projection, an index function, the number of jellyfish, and the
+maximum number of trials (tries) are provided as input. Then, the
+current projection is evaluated by the index function. The projection is
+then moved in a direction determined by a random factor, influenced by
+how far along we are in the optimisation process. Occasionally,
+completely new directions may be taken like a jellyfish might with ocean
+currents. A new projection is accepted if it is an improvement compared
+to the current one, rejected otherwise. This process continues and
+iteratively improves the projection, until the pre-specified maximum
+number of trials is reached.
 
@@
 
-**Output**: `optimized_projection`
+**Output**: `optimised_projection`
 
@@
 value from `current_projections`, and `current_idx` as the array of
-index function values for each projection in `current_projections`
-
-**for** each trial_id in 1 to max_tries **do**
-
-> Calculate $c_t$ based on the current_idx and max_trial
+index values for each projection in `current_projections`
+
+**for** each `trial_id` in 1 to `max_tries` **do**
+
+> Calculate the time control value, $c_t$, based on `current_idx` and
+> `max_trial`
 
@@
 >
-> > Define trend based on the current_best and current_projections
+> > Define trend based on the `current_best` and `current_projections`
 >
@@
 > >
-> > > For each projection, compare with a random jelly and adjust
+> > > For each projection, compare with a random jellyfish and adjust
 > > > towards or away from it (active)
@@
 > > Determine the projection with the best index value as the new
-> > current_best
-
-> **if** trial_id $\geq$ max_trial, print the last best projection
+> > `current_best`
+
+> **if** `trial_id` $\ge$ `max_trial`, print the last best projection
 > **exit**
 
-**return** the set of projections with the updated current_best as the
-optimized_projection
+**return** the set of projections with the updated `current_best` as the
+`optimised_projection`
 :::
@@
 the specific implementation details can vary depending on the version of
-the algorithm or its application, the focus is on two main parameters that
-are most relevant to our application: the number of jellyfish and the
-maximum number of trials.
+the algorithm or its application, the focus is on two main parameters
+that are most relevant to our application: the number of jellyfish and
+the maximum number of tries.
 
@@
 pursuit indexes (smoothness, squintability, flexibility, rotation
-invariance, and speed). Since not all the properties affects the
-execution of the optimisation, the focus here is on the first two
-properties, namely *smoothness* (@sec-smoothness) and *squintability*
-(@sec-squintability), and propose two metrics to evaluate them.
+invariance, and speed). Since not all index properties affect the
+optimisation process, the focus here is on the first two properties,
+*smoothness* (@sec-smoothness) and *squintability* (@sec-squintability),
+for which metrics are proposed to quantify them.
 
@@
 
-```{=html}
-<!--
-An intuitive way to measure smoothness of a function would be to count
-how many continuous derivatives exist. 
-To help define smoothness in our
-context, we can make use of the Sobolev spaces: functions $f$ in the
-Sobolev space $W^{p,\infty}$ have all derivatives of order less than $p$
-continuous. Smoothness would then be the highest $p$ such that the index
-function belongs to $W^{p,\infty}$. We can relax this and consider
-$W^{p,q}$ Sobolev spaces.
-
-Consider the following definition of partial derivative. Let $U$ be an
-open subset of $\mathbf{R}^n$ and $f: U \rightarrow \mathbf{R}$. The
-partial derivative of function $f$ at the point
-$\mathbf{a} = (a_1, \ldots, a_n) \in U$ with respective to the $i$-th
-variable $x_i$ is defined as
-
-$$
-\frac{\delta}{\delta x_i} f(\mathbf{a}) =  \lim_{h \rightarrow 0} \frac{f(a_1, \ldots, a_{i-1}, a_i +h, a_{i+1}, \ldots, a_n) - f(a_1 , \ldots, a_i, \ldots, a_n)}{h} = \lim_{h\rightarrow 0} \frac{f(\mathbf{a} + h\mathbf{e_i}) - f(\mathbf{a})}{h}.
-$$ where $\mathbf{e_i}$ is the unit vector of the $i$-th variable $x_i$.
-
-If the derivative is not well-defined, we propose to approximate the
-derivative using the following expression:
-
-$$\frac{1}{n_i}\sum_{i} \frac{\|f(\mathbf{a} + h_i\mathbf{e_i}) - f(\mathbf{a})\|^p}{h_i}$$
-
-where $p \in [0,1]$. The choice of $p$ reflects the penalty behaviour.
-For the application in this paper, we choose $p = 1$.
-
--   $h_i$ is an neighbourhood area around $a$. We can insert different
-    values of $h$ to approximate the above quantity and observe how this
-    quantity changes.
--->
-```
 This subsection proposes a metric for the smoothness of a projection
@@
 If a PP index function is evaluated at some random points, as is done at
-the initialization stage of the jellyfish search optimization, then
-these random index values can be interpreted as a random field, indexed
-by a space parameter, namely the random projection angle. This analogy
-suggests to use this random training sample to fit a spatial model, a
-simple one being a (spatial) Gaussian process.
+the initialization stage of JSO, then these random index values can be
+interpreted as a random field, indexed by a space parameter, namely the
+random projection angle. This analogy suggests to use this random
+training sample to fit a spatial model, a simple one being a (spatial)
+Gaussian process.
 
 The distribution of a Gaussian process is fully determined by its mean
-function and covariance function. The smoothness property comes into
-play in the definition of the covariance function: if a PP index is very
-smooth, then two close projection angles should produce close index
-values (strong correlation); by contrast, if a PP index is not very
-smooth, then two close projection angles might give very different index
-values (fast decay of correlations with respect to distance between
-angles).
+and covariance function. The smoothness property comes into play in the
+definition of the covariance function: if a PP index is very smooth,
+then two close projection angles should produce close index values
+(strong correlation); by contrast, if a PP index is not very smooth,
+then two close projection angles might give very different index values
+(fast decay of correlations with respect to distance between angles).
 
@@
 expressed analytically when $\nu$ is a half-integer, the most popular
-values in the literature being $1/2$, $3/2$ and $5/2$. The parameter
-$\nu$, called *smoothness parameter*, controls the decay of the
-covariance function. As such, it is an appropriate measure of smoothness
-of a random field.
-
-In this context, the parameter $\nu$ is suggested as a measure of
-the smoothness of the index function by fitting a Gaussian process prior
+values in the literature being $\frac{1}{2}$, $\frac{3}{2}$ and
+$\frac{5}{2}$. The parameter $\nu$, called *smoothness parameter*,
+controls the decay of the covariance function. As such, it is an
+appropriate measure of smoothness of a random field.
+
+In this context, the parameter $\nu$ is suggested as a measure of the
+smoothness of the index function by fitting a Gaussian process prior
 with Matérn covariance on a dataset generated by random evaluations of
 the index function, as done at the initialization stage of the jellyfish
-search optimization. There exist several R packages, such as GpGp
-[@guinness2021gpgp] or ExaGeoStatR [@abdulah2023large], to fit the
-hyperparameters of a GP covariance function on data. In this project, we
-make use of the GpGp package.
+search optimization. There exist several R packages, such as `GpGp`
+[@guinness2021gpgp] or `ExaGeoStatR` [@abdulah2023large], to fit the
+hyperparameters of a GP covariance function on data. In this project,
+the `GpGp` package is used.
 
@@
 
-In the existing literature, it is commonly understood that a large
-squint angle implies that the objective function value is close to
-optimal even when the perfect view to see the structure is far away.
-A small squint angle means that the PP index value improves
-substantially only when the perfect view is close by. As such,
-low squintability implies rapid improvement in the index value when near
-the perfect view. For PP, a small squint angle is considered to be
-undesirable because it means that the optimiser needs to be very close
-to be able to "see" the optima. Thus, it could be difficult for the
-optimiser to find the optima. The mathematical formulation
-of this intuition is proposed below:
-
-Let $A \in \mathbf{R}^{p \times 2}$ be the 2D projection basis and $A^*$
-the optimal basis achieving the maximum index value for a given
+From @barnett1981interpreting and @laa_using_2020, a large squint angle
+implies that the objective function value is close to optimal even when
+the perfect view to see the structure is far away. A small squint angle
+means that the PP index value improves substantially only when the
+perfect view is close by. As such, low squintability implies rapid
+improvement in the index value when near the perfect view. For PP, a
+small squint angle is considered to be undesirable because it means that
+the optimiser needs to be very close to be able to "see" the optima.
+Thus, it could be difficult for the optimiser to find the optima. The
+mathematical formulation of this intuition is proposed below:
+
+Let $A \in \mathbf{R}^{p \times d}$ be the d-D projection basis and
+$A^*$ be the optimal basis achieving the maximum index value for a given
 projection pursuit problem defined by the shape to find, data dimension,
@@
 
-The index value $f(XA)$ is defined as described in the introduction. We
-now focus our attention to how the index value $f(XA)$ changes with
+We now focus our attention to how the index value $f(XA)$ changes with
 respect to the projection distance $d(A, A^*)$, over the course of the
@@
 
-```{=html}
-<!--
-\[We generate random points that is beyond 1.5 projection distance and
-interpolate. Then we fit a kernel or use nonlinear least squares.\]
--->
-```
 It is expected that for a PP index with high squint angle, the
@@
 differentiable), and the projection distances for which these values are
-attained, should play an important role in the mathematical definition
-of squintability. Noticing that $f$ is expected to be a decreasing
-function, the squintability metric is proposed as follows:
+attained, are crucial in the mathematical definition of squintability.
+Noticing that $f$ is expected to be a decreasing function, the
+squintability metric is proposed as follows:
 
@@
 
-On the one hand, $-\min_{x}f'(x)$ represents the largest value of the
-gradient of $-f$. On the other hand, $\underset{x}{\arg\min}f'(x)$
-represents the projection distance at which this largest gradient is
-attained. Based on the above discussion, it is expected that these two
+$-\min_{x}f'(x)$ represents the largest value of the gradient of $-f$
+and $\underset{x}{\arg\min}f'(x)$ represents the projection distance at
+which this largest gradient is attained. It is expected that these two
 values should be both high in the case of high squintability (fast
@@
 
-The JSO performance is compared with the existing Creeping Random Search
-(CRS) [@RJ-2021-105; @laa_using_2020] currently used in guided tour, and
-explores how JSO behaves with different hyper-parameter (number of
-jellyfishes and maximum tries) and data dimension combinations. A second
-simulation studies the effect of index properties (smoothness and
-squintability), along with JSO hyper-parameters, and data dimension, on
-the success rate of the JSO performance.
+The JSO performance is compared with an existing optimiser, Creeping
+Random Search (CRS) [@RJ-2021-105; @laa_using_2020] used in the PP
+guided tour to explore JSO's behaviour under different hyper-parameter
+and data dimension combinations. The second simulation studies the
+effect of index properties (smoothness and squintability), along with
+JSO hyper-parameters, and data dimension, on the success rate of the JSO
+performance. This section describes the simulation details, with the
+results deferred to Section @sec-sim-res.
 
@@
 
-The performance of JSO is investigated in comparison with the existing
-optimiser, CRS, with respect to two hyper-parameters (number of
-jellyfishes and maximum number of tries). The performance is measured by
-success rate, defined as the proportion of simulations that achieves a
-final index value within 0.05 of the best index value found among all 50
-simulations. This comparison is based on projection pursuit to find the
-pipe shape investigated by @laa_using_2020 using two simple and
-well-behaved indexes, the `holes` index, and the `splines2d` index.
+he performance of JSO is investigated both in comparison to the existing
+optimizer, CRS, and across various hyper-parameter values. The
+performance is measured by the success rate, defined as the proportion
+of simulations that achieves a final index value within 0.05 of the best
+index value found among all 50 simulations (see @fig-success-rate for an
+illustration). This comparison is based on projection pursuit to find
+the pipe shape investigated by @laa_using_2020 using the `holes` index.
+
 Fifty simulations are conducted with both JSO and CRS, in four different
-data dimensions ($d = 6, 8, 10, 12$). JSO uses 100 jellyfishes with 100
-maximum number of tries, while the CRS allows a maximum of 1000 samples
-at each iteration before the algorithm terminates. The different numbers
-accounts for the multiple paths of JSO to enable fairer comparison with
+data dimensions ($d = 6, 8, 10, 12$). JSO uses 100 jellyfishes with a
+maximum of 100 tries, while the CRS allows a maximum of 1000 samples at
+each iteration before the algorithm terminates. The different numbers
+account for the multiple paths of JSO to enable fairer comparison with
 CRS. The results of the simulation are collected using the data
-structure proposed in @RJ-2021-105 is used here for assessing JSO, where
-the design parameters are stored along with index value, projection
-basis, random seed, and computation time.
-<!--A sample for the sine-wave structure in 6D data using a spline index function (`splines2d`) is printed below:
-
-```{r}
-dplyr::glimpse(sim_sine_6d_spline_head |> select(-id, -info)) 
-```
--->
-
-In this simulation, 50 jellyfishes (`n_jellies`) and a maximum of 50 tries
-(`max_tries`) are used. The full data contains 50 simulations, indexed
-by the variable `sim`, with the associated `seed`s. Recorded in columns
-`basis`, `index_val`, `tries`, `loop` and `time`, are the projection
-basis, index value, iteration ID, jellyfish ID, and computational time,
-respectively.
-<!-- The basis column records every visited basis before -->
-<!-- comparing with the current (last) basis. If a visited basis has a -->
-<!-- smaller index value than the current (last) one, JSO will keep the -->
-<!-- current basis for the next iteration in the search but still records the -->
-<!-- visited inferior basis.  -->
-
-In the second experiment, 50 simulations are conducted at each
-hyper-parameter combination to analyse its effect on the JSO success
-rate: number of jellyfishes (20, 50, and 100), and maximum number of tries
-(50 and 100). @fig-success-rate illustrates the calculation based on the
-projection found by the above simulation. Here, the optimal index value
-is
-`r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`.
-Out of the 50 simulations,
-`r sim_sine_6d_spline_best |> filter(index_val >= aa - 0.05) |> nrow() -> bb; bb`
-achieved an index value within \[`r aa - 0.05`, `r aa`\], resulting in a
-success rate of
-`r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.
-
-<!-- @fig-projection shows the best projection found in each simulation, -->
-
-<!-- sorted by the index value. Most of the projections show a clear -->
-
-<!-- sine-wave structure, while the last six show only a straight line -->
-
-<!-- with clusters at the end. When local optima exist in the projection -->
-
-<!-- pursuit problem, plotting the projections found in each simulation -->
-
-<!-- or by each jellyfish can help identify these local optima. -->
+structure proposed in @RJ-2021-105 for assessing JSO, where the design
+parameters are stored along with index value, projection basis, random
+seed, and computation time.
+
+Fifty additional simulations are conducted for each hyper-parameter
+combination to analyze how they affect the JSO success rate. This
+includes variations in the number of jellyfish (20, 50, and 100) and the
+maximum number of tries (50 and 100).
 
@@
 #| label: fig-success-rate
-#| fig.cap: "Illustration of success rate calculation: The best projections identified by the JSO in 50 simulations are sorted by their index values.
-The highest index value found across all simulation is 0.999.
-Out of the 50 simulations, 44 achieved an index value within 0.05 of the best, resulting in a success rate of 0.88 (44/50)."
-# proj_dt <- map_dfr(sim_sine_6d_spline_best |> pull(basis), 
-#               ~as_tibble(as.matrix(sine1000) %*% .x), .id = "sim") |> 
+#| fig.cap: "Illustration of success rate calculation: Final projections based on projection pursuit to find the pipe shape in 8D data using the holes index, optimised by CRS, in 50 simulations.
+The 50 final projections are sorted by their index values.
+The highest index value found across all simulations is 0.969.
+Out of the 50 simulations, 43 achieved an index value within 0.05 of the best, resulting in a success rate of 0.86 (43/50)."
+# pipe_jellyfish_12d <- pipe_better |> 
+#   filter(dim == 2) |> 
+#   mutate(dim = as.numeric(dim) * 8, sim = as.numeric(sim)) 
+# proj_dt <- map_dfr(pipe_jellyfish_12d |> pull(basis),
+#               ~as_tibble(pipe1000_8d %*% .x), .id = "sim") |>
 #   mutate(sim = as.numeric(sim)) |>
-#   left_join(sim_sine_6d_spline_best |> select(sim, index_val), by = "sim")
+#   left_join(pipe_jellyfish_12d |> select(sim, index_val), by = "sim")
 # 
@@
 # proj_dt |>
-#   #filter(index_val %in% c(idx_val_sorted[1:12], idx_val_sorted[39:50])) |>
 #   ggplot(aes(x = V1, y = V2)) +
@@
 
-This simulation uses two different data shapes, pipe and a sine wave,
-investigated in 6D and 8D spaces with six different PP indexes:
-`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`. The two
-JSO hyper-parameters are varied. A total of `r nrow(sim_df)`
-combinations result, comprising of
+To assess JSO's performance across various scenarios, two different data
+shapes, pipe and a sine wave, are investigated in 6D and 8D spaces using
+six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`,
+and `stringy`, with varied JSO hyper-parameters. A total of
+`r nrow(sim_df)` combinations result, comprising of
 `r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and
-`r nrow(filter(sim_df, index != "holes"))` on the sine-wave data. For
-each case, JSO is run 50 times to calculate the success rate.
-
-Smoothness and squintability are computed for each run, following the
-procedures outlined in @sec-smoothness and @sec-squintability and as
-illustrated in @fig-smoothness and @fig-squintability.
+`r nrow(filter(sim_df, index != "holes"))` on the sine-wave data. Again,
+JSO is run 50 times to calculate the success rate for each projection
+pursuit.
+
+Smoothness and squintability are computed following the procedures
+outlined in @sec-smoothness and @sec-squintability and as illustrated in
+@fig-smoothness and @fig-squintability.
 
@@
 projection distances are calculated for these interpolated bases and the
-index values are averaged with a bin width of 0.005. The squintability
-measure is then calculated by fitting a four-parameter scaled logistic
-function of index value against projection distance, estimated by
-non-linear least squares.
-
-A generalised linear model is fitted to the resulting data, using a
-binomial family and a logit link function. The data is pre-processed
-by 1) scaling the JSO hyper-parameters by a factor of 10 for
-interpretation, 2) creating a new binary variable `long_time` to
-indicate cases with an average run time over 30 seconds, and 3)
-re-coding the success rate for the `stringy` index as 0, because none of
-the 50 simulations correctly identified the sine-wave shape.
+index values are averaged with a bin width of 0.005. A four-parameter
+scaled logistic function is fitted to the index values against
+projection distances, estimated by non-linear least squares. The
+squintability measure is then calculated as
+@eq-squintability-parametric.
+
+To construct a relationship among success rate, index properties
+(smoothness and squintability), and jellyfish hyper-parameters, a
+generalised linear model is fitted using a binomial family and a logit
+link function. The data is pre-processed by 1) scaling the JSO
+hyper-parameters by a factor of 10 for interpretation, 2) creating a new
+binary variable `long_time` to indicate cases with an average run time
+over 30 seconds, and 3) re-coding the success rate for the `stringy`
+index as 0, because none of the 50 simulations correctly identified the
+sine-wave shape.
 
@@
 #| fig.cap: "Illustration of steps to calculate smoothness.
-For a given projection pursuit problem defined by the shape to find, data dimension and the index function, sample random bases given the orthonormality contraint.
-Calculate the projection distance and the index value for each random basis.
-Fit a Gaussian process model to the index values against the projection distances to obtain the smoothness measure.
+For a given projection pursuit problem defined by the shape to find, data dimension and the index function, 1) sample random bases given the orthonormality contraint, 2) calculate the projection distance and the index value for each random basis, and 3) fit a Gaussian process model of index values against projection distances to obtain the smoothness measure.
 "
@@
 #| fig.cap: "Illustration of steps to calculate squintability.
-For a given projection pursuit problem defined by the shape to find, data dimension and the index function, sample random bases given the orthonormality and projection distance contraint.
-Interpolate the sampled bases to the optimal basis and calculate the projection distance and the index value for each interpolated basis.
-Bin the index value by the projection distance, then claculate the average index value for each bin.
-Fit the scaled sigmoid function from Equation @eq-squintability to the binned index values against the projection distance using non-linear least square.
-Use Equation @eq-parametric to calculate the squintability measure from the estimated $\\theta_1$ to $\\theta_4$.
-"
+For a given projection pursuit problem defined by the shape to find, data dimension and the index function, 1) sample random bases given the orthonormality and projection distance contraint, 2) interpolate the sampled bases to the optimal basis and calculate the projection distance and the index value for each interpolated basis.
+3) bin the index values by projection distances to obtain the average index value for each bin, 4) fit the scaled sigmoid function in Equation @eq-squintability to the binned index values against projection distances using non-linear least square, 5) calculate the squintability measure using Equation @eq-squintability-parametric with parameters estimated from the model."
 # p1 <- sim_sine_6d_spline_projdist |> 
@@
 
-The results from the two simulations described in @sec-sim-deets are
+The results from the first simulation described in @sec-sim-deets are
 analysed based on the final projections across two optimisers (JSO and
-CRS) and the success rate across JSO hyper-parameters. Smoothness and
-squintability are calculated across a collection of pipe-finding and
-sine-wave finding problems to construct the relationship between success
-rate, JSO hyper-parameters, and index properties.
+CRS) and the success rate across JSO hyper-parameters. In the second
+simulation, smoothness and squintability are calculated across a
+collection of pipe-finding and sine-wave finding problems to construct
+the relationship between success rate, JSO hyper-parameters, and index
+properties.
 
@@
 presented in @fig-proj, broken down by 10th quantile, faceted by the
-data dimension. The JSO uses 100 jellyfishes and a maximum number of
-tries of 100. The CRS uses a maximum of 1000 tries in each step of
-random sampling step before the algorithm terminates. In the
-6-dimensional data scenario, JSO consistently identifies a clear pipe
-shape. The CRS also finds the pipe shape but with a wide rim, suggesting
-a further polish search may be required. With increasing dimensions, JSO
-may not always identify the pipe shape due to random sampling, but it
-still presents a pipe shape in over 50% of cases. When compared to the
-CRS, JSO reaches higher index values and clearer pipe shapes across all
-quantiles. This suggests JSO can be more effective when searching higher
-dimension spaces, which may be preferred by users.
-
-The success rate calculated for each hyper-parameter combination is
-presented in @fig-proportion. As the number of jellyfishes and maximum
-tries increase, the success rate also increases. For simpler problems (6
-dimensions), small parameter values (20 jellyfishes and a maximum of 50
-tries) can achieve a high success rate. However, for higher-dimensional
-problems (8, 10, and 12 dimensions), larger parameter values (i.e. 100
-jellyfishes and a maximum of 100 tries) are necessary. Increasing both
+data dimensions. In the 6-dimensional data scenario, JSO consistently
+identifies a clear pipe shape. The CRS also finds the pipe shape but
+with a wide rim, suggesting a further polish search may be required.
+With increasing dimensions, JSO may not always identify the pipe shape
+due to random sampling, but it still finds the pipe shape in over 50% of
+cases. Compared to CRS, JSO achieves higher index values and clearer
+pipe shapes across all quantiles in data of 8, 10, 12 dimensions,
+suggesting its advantage in exploring high-dimensional spaces.
+
+The success rate calculated at each hyper-parameter combination (number
+of jellyfish and the maximum number of tries) is presented in
+@fig-proportion. As the number of jellyfish and maximum tries increase,
+the success rate also increases. For simpler problems (6 dimensions),
+small parameter values (20 jellyfishes and a maximum of 50 tries) can
+already achieve a high success rate. However, larger parameter values
+(i.e. 100 jellyfishes and a maximum of 100 tries) are necessary for
+higher-dimensional problems (8, 10, and 12 dimensions). Increasing both
 parameters enhances the performance of JSO, but it also extends the
@@
 
-<!-- The low -->
-
-<!-- squint value of MIC and TIC index is due to its convex shape of the -->
-
-<!-- index value against the projection distance, as shown in -->
-
-<!-- @fig-idx-proj-dist. Comparing to other concave shapes, the maximum first -->
-
-<!-- derivative happens at a smaller projection distance ($\theta_2$), -->
-
-<!-- requiring the optimiser to get closer to the optimal to see a -->
-
-<!-- significant change in the index value, hence a small squintability -->
-
-<!-- measure.  -->
-
 The index properties, including smoothness and squintability, offer
@@
 optimisation problems. @tbl-smoothness-squintability presents the
-parameters estimated from the Gaussian process (variance, range, smooth,
-and nugget) and the scaled logistic function ($\theta_1$ to $\theta_4$)
-for calculating both metrics for each case considered in the simulation.
-The column "smooth" is used as the smoothness metric and the column
-"squint" is calculated as equation @eq-squintability-parametric as the
-squintability metric. @tbl-mod-output presents the results of fitting a
-generalised linear model with a binomial family and a logit link
-function to the data in @tbl-mod-data, where all the three components
-(success rate, JS hyper-parameters, and index properties) are combined.
-The model suggests that JSO success rate is positively associated with
-the two hyper-parameters, as well as with the index properties:
-smoothness and squintability. Specifically, using 10 more jellyfishes
-and 10 more tries increases the odd ratio of success by
+parameters for calculating both metrics estimated from the Gaussian
+process (variance, range, smooth, and nugget) and the scaled logistic
+function ($\theta_1$ to $\theta_4$) for each case considered in the
+simulation. The column "smooth" is used as the smoothness metric and the
+column "squint" is calculated as equation @eq-squintability-parametric
+as the squintability metric. @tbl-mod-output presents the results of
+fitting a generalised linear model with a binomial family and a logit
+link function to a sample of data in @tbl-mod-data, where all the three
+components (success rate, JS hyper-parameters, and index properties) are
+combined. The model suggests that JSO success rate is positively
+associated with the two hyper-parameters, as well as with the index
+properties: smoothness and squintability. Specifically, using 10 more
+jellyfish and 10 more tries increases the odd ratio of success by
 `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`% and
@@
 shape-to-find, the index function used, and the data dimension,
-determine the properties like smoothness and squintability. These
+determine the properties such as smoothness and squintability. These
 metrics can then be compared across different problems to understand
-their relative complexity. The results suggest that squintability has a
+their relative complexites. The results suggest that squintability has a
 more significant influence than smoothness on the success rate of JSO.
@@
 The JSO performance is affected by the hyper-parameters, with a higher
-chance of reaching the global optima when more jellyfishes are used and the
-maximum number of tries is increased. However, it comes at a
+chance of reaching the global optima when more jellyfish are used and
+the maximum number of tries is increased. However, it comes at a
 computational cost, as expected. The performance declines if the
@@
 because it runs many paths. The recommended approach is to conduct the
-optimisation off-line, and then use the guided tour to follow selected
-jellyfish. The tools for this are available in the `tourr` package, too.
+optimisation off-line, extract the bases of a selected jellyfish, and
+then use the planned tour to follow selected jellyfish. The tools for
+this are available in the `tourr` package, too.
 
@@
 The source code for reproducing the work reported in this paper can be
-found at: <https://github.com/huizezhang-sherry/paper-optim>. Nicolas
+found at: <https://github.com/huizezhang-sherry/paper-jso>. Nicolas
 Langrené acknowledges the partial support of the Guangdong Provincial

### 2024-07-15 jso.tex
proofread

--- 2024-07-01 optim.tex
@@
 algorithm has been implemented in the \texttt{tourr} package and
-functions to calculate smoothness and squintability of indexes are
-available in the \texttt{ferrn} package.
+functions to calculate smoothness and squintability are available in the
+\texttt{ferrn} package.
 \end{abstract}
@@
 reduction technique aimed at identifying informative linear projections
-of data. The method involves optimising a objective function known as
+of data. The method involves optimising an objective function known as
 the PP index (e.g. \citet{hall1989polynomial},
@@
 and efficient optimisation procedure is essential to explore the data
-landscape and achieving a globally optimal viewpoint of the data.
-
-Significant efforts have been devoted to enhancing the functionality of
-PP, improving optimisation (e.g. \citet{posse1995},
-\citet{marie-sainte2010}, \citet{grochowski2011}) and extending its use
-to visually explore high-dimensional data. \citet{cook1995grand}
-introduced the PP guided tour, which monitors the optimisation visually
-to see the data projections leading in and out of the optima. An
-implementation is available in the R \citep{R} package \texttt{tourr}
-\citep{tourr}. \citet{RJ-2021-105} illustrated how to diagnose
-optimisation processes, particularly focusing on the guided tour, and
-revealed a need for improved optimisation. While improving the quality
-of the optimisation solutions in the tour is essential, it is also
-important to be able to view the data projections as the optimisation
-progresses. Integrating the guided tour with a global optimisation
-algorithm that is efficient in finding the global optimal and enables
-viewing of the projected data during the exploration process is a goal.
+landscape and achieve a globally optimal viewpoint of the data.
+
+Optimisation of PP is typically investigated in the literature when new
+indexes are proposed \citep{posse1995, marie-sainte2010, grochowski2011}
+or when visualisation method are used to track the optimisation process.
+\citet{cook1995grand} introduced the PP guided tour, which monitors the
+optimisation visually to see the data projections leading in and out of
+the optima. An implementation is available in the \texttt{tourr} package
+\citep{tourr} in R \citep{R}. \citet{RJ-2021-105} illustrated how to
+diagnose optimisation processes, particularly focusing on the guided
+tour, and revealed a need for improved optimisation. While improving the
+quality of the optimisation solutions in the tour is essential, it is
+also important to be able to view the data projections as the
+optimisation progresses. Integrating the guided tour with a global
+optimisation algorithm that is efficient in finding the global optimal
+and enables viewing of the projected data during the exploration process
+is a goal.
 
 Here, the potential for a Jellyfish Search Optimiser (JSO) to be
-integrated with the projection pursuit guided tour is explored. JSO
-\citep{chou_novel_2021} -\citep{rajwar_exhaustive_2023}, inspired by the
-search behaviour of jellyfish in the ocean, is a swarm-based
+integrated with the projection pursuit guided tour is explored. JSO,
+\citep{chou_novel_2021} - \citep{rajwar_exhaustive_2023}, inspired by
+the search behaviour of jellyfish in the ocean, is a swarm-based
 metaheuristic designed to solve global optimisation problems. Compared
-to traditional methods, JSO demonstrated stronger search ability and
+to traditional methods, JSO has demonstrated stronger search ability and
 faster convergence, and requires fewer tuning parameters. These
@@
 which is new for the field. A series of simulation experiments using
-specially structured datasets and various PP indexes are conducted to
-assess the behaviour of JSO under different settings and its sensitivity
-to hyper-parameter choices (number of jellyfishes and maximum number of
-tries). The relationship between the JSO performance, hyper-parameter
-tuning and properties of PP indexes is analysed to provide guidance for
-practitioners using projection pursuit, and developers of new indexes.
+various datasets and PP indexes are conducted to assess JSO's behaviour
+and its sensitivity to hyper-parameter choices (number of jellyfish and
+maximum number of tries). The relationship between the JSO performance,
+hyper-parameter choices and properties of PP indexes (smoothness and
+squintability) is analysed to provide guidance on selecting optimisers
+for practitioners using projection pursuit. Additionally, it aims to
+guide the design of new indexes that facilitate easy optimization for PP
+researchers.
 
@@
 Section~\ref{sec-theory} details JSO and introduces metrics that measure
-different properties of PP indexes. Section~\ref{sec-sim-deets} details
-two simulation experiments to assess JSO's performance: one comparing
-JSO's performance improvements relative to an existing optimiser,
-Creeping Random Search (CRS), and the other studying the impact of
-different PP index properties on optimisation performance.
-Section~\ref{sec-sim-res} presents the results.
+different properties of PP indexes, smoothness and squintability.
+Section~\ref{sec-sim-deets} details two simulation experiments to assess
+JSO's performance: one comparing JSO's performance improvements relative
+to an existing optimiser, Creeping Random Search (CRS), and the other
+studying the impact of different PP index properties on optimisation
+performance. Section~\ref{sec-sim-res} presents the results.
 Section~\ref{sec-conclusion} summarises the work and provides
@@
 \item
-  \texttt{search\_better()}: is a brute-force optimisation searching
-  randomly for projections with higher index values.
+  \texttt{search\_better()}: also known as Creeping Random Search (CRS),
+  is a brute-force optimisation searching randomly for projections with
+  higher index values.
 \item
@@
 relatively easy to optimise. A goal is to be able to incorporate more
-complex PP indexes, for example based on scagnostics (\citet{scag},
+complex PP indexes, for example, based on scagnostics (\citet{scag},
 \citet{WW08}).
@@
 respectively. In the context of optimization, these movements are
-abstracted to explore the search space in a way that balances
-exploration (searching new areas) and exploitation (focusing on
-promising areas). The algorithm aims to find the optimal solution by
-adapting the behaviour of jellyfish to navigate towards the best
-solution over iterations \citep{chou_novel_2021}.
+abstracted to explore the search space, aiming to balance exploration
+(searching new areas) and exploitation (focusing on promising areas).
+The algorithm aims to find the optimal solution by adapting the
+behaviour of jellyfish to navigate towards the best solution over
+iterations \citep{chou_novel_2021}.
 
 To solve the optimisation problem embedded in the PP guided tour, a
-starting projection, an index function and the maximum number of
-iterations are provided as input. Then, the current projection is
-evaluated by the index function. The projection is then moved in a
-direction determined by a random factor, influenced by how far along we
-are in the optimization process. Occasionally, completely new directions
-may be taken like a jellyfish might with ocean currents. A new
-projection is accepted if it is an improvement compared to the current
-one, rejected otherwise. This process continues and iteratively improves
-the projection, until the pre-specified maximum number of trials is
-reached.
-
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, arc=.35mm, breakable, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, colback=white, opacityback=0, left=2mm, toprule=.15mm, opacitybacktitle=0.6, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm]
+starting projection, an index function, the number of jellyfish, and the
+maximum number of trials (tries) are provided as input. Then, the
+current projection is evaluated by the index function. The projection is
+then moved in a direction determined by a random factor, influenced by
+how far along we are in the optimisation process. Occasionally,
+completely new directions may be taken like a jellyfish might with ocean
+currents. A new projection is accepted if it is an improvement compared
+to the current one, rejected otherwise. This process continues and
+iteratively improves the projection, until the pre-specified maximum
+number of trials is reached.
+
+\begin{tcolorbox}[enhanced jigsaw, breakable, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, opacityback=0, arc=.35mm, toptitle=1mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, left=2mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, coltitle=black, rightrule=.15mm, bottomrule=.15mm, colback=white]
 
@@
 
-\textbf{Output}: \texttt{optimized\_projection}
+\textbf{Output}: \texttt{optimised\_projection}
 
@@
 best index value from \texttt{current\_projections}, and
-\texttt{current\_idx} as the array of index function values for each
-projection in \texttt{current\_projections}
-
-\textbf{for} each trial\_id in 1 to max\_tries \textbf{do}
+\texttt{current\_idx} as the array of index values for each projection
+in \texttt{current\_projections}
+
+\textbf{for} each \texttt{trial\_id} in 1 to \texttt{max\_tries}
+\textbf{do}
 
 \begin{quote}
-Calculate \(c_t\) based on the current\_idx and max\_trial
+Calculate the time control value, \(c_t\), based on
+\texttt{current\_idx} and \texttt{max\_trial}
 \end{quote}
@@
 \begin{quote}
-Define trend based on the current\_best and current\_projections
+Define trend based on the \texttt{current\_best} and
+\texttt{current\_projections}
 \end{quote}
@@
 \begin{quote}
-For each projection, compare with a random jelly and adjust towards or
-away from it (active)
+For each projection, compare with a random jellyfish and adjust towards
+or away from it (active)
 \end{quote}
@@
 Determine the projection with the best index value as the new
-current\_best
+\texttt{current\_best}
 \end{quote}
@@
 \begin{quote}
-\textbf{if} trial\_id \(\geq\) max\_trial, print the last best
-projection \textbf{exit}
+\textbf{if} \texttt{trial\_id} \(\ge\) \texttt{max\_trial}, print the
+last best projection \textbf{exit}
 \end{quote}
 
-\textbf{return} the set of projections with the updated current\_best as
-the optimized\_projection
+\textbf{return} the set of projections with the updated
+\texttt{current\_best} as the \texttt{optimised\_projection}
 
@@
 that are most relevant to our application: the number of jellyfish and
-the maximum number of trials.
+the maximum number of tries.
 
@@
 projection pursuit indexes (smoothness, squintability, flexibility,
-rotation invariance, and speed). Since not all the properties affects
-the execution of the optimisation, the focus here is on the first two
-properties, namely \emph{smoothness} (Section~\ref{sec-smoothness}) and
-\emph{squintability} (Section~\ref{sec-squintability}), and propose two
-metrics to evaluate them.
+rotation invariance, and speed). Since not all index properties affect
+the optimisation process, the focus here is on the first two properties,
+\emph{smoothness} (Section~\ref{sec-smoothness}) and
+\emph{squintability} (Section~\ref{sec-squintability}), for which
+metrics are proposed to quantify them.
 
@@
 covariance function can be expressed analytically when \(\nu\) is a
-half-integer, the most popular values in the literature being \(1/2\),
-\(3/2\) and \(5/2\). The parameter \(\nu\), called \emph{smoothness
-parameter}, controls the decay of the covariance function. As such, it
-is an appropriate measure of smoothness of a random field.
+half-integer, the most popular values in the literature being
+\(\frac{1}{2}\), \(\frac{3}{2}\) and \(\frac{5}{2}\). The parameter
+\(\nu\), called \emph{smoothness parameter}, controls the decay of the
+covariance function. As such, it is an appropriate measure of smoothness
+of a random field.
 
@@
 the index function, as done at the initialization stage of the jellyfish
-search optimization. There exist several R packages, such as GpGp
-\citep{guinness2021gpgp} or ExaGeoStatR \citep{abdulah2023large}, to fit
-the hyperparameters of a GP covariance function on data. In this
-project, we make use of the GpGp package.
+search optimization. There exist several R packages, such as
+\texttt{GpGp} \citep{guinness2021gpgp} or \texttt{ExaGeoStatR}
+\citep{abdulah2023large}, to fit the hyperparameters of a GP covariance
+function on data. In this project, the \texttt{GpGp} package is used.
 
@@
 
-In the existing literature, it is commonly understood that a large
+From \citet{barnett1981interpreting} and \citet{laa_using_2020}, a large
 squint angle implies that the objective function value is close to
@@
 
-Let \(A \in \mathbf{R}^{p \times 2}\) be the 2D projection basis and
-\(A^*\) the optimal basis achieving the maximum index value for a given
-projection pursuit problem defined by the shape to find, data dimension,
-and the index function \(f(.)\). The projection distance \(d(A, A^*)\),
-is defined as the Frobenius norm \(\lVert . \rVert _F\), between the
-squared difference of these two bases:
+Let \(A \in \mathbf{R}^{p \times d}\) be the d-D projection basis and
+\(A^*\) be the optimal basis achieving the maximum index value for a
+given projection pursuit problem defined by the shape to find, data
+dimension, and the index function \(f(.)\). The projection distance
+\(d(A, A^*)\), is defined as the Frobenius norm \(\lVert . \rVert _F\),
+between the squared difference of these two bases:
 
@@
 
-The index value \(f(XA)\) is defined as described in the introduction.
 We now focus our attention to how the index value \(f(XA)\) changes with
@@
 \(f\) is twice differentiable), and the projection distances for which
-these values are attained, should play an important role in the
-mathematical definition of squintability. Noticing that \(f\) is
-expected to be a decreasing function, the squintability metric is
-proposed as follows:
+these values are attained, are crucial in the mathematical definition of
+squintability. Noticing that \(f\) is expected to be a decreasing
+function, the squintability metric is proposed as follows:
 
@@
 
-On the one hand, \(-\min_{x}f'(x)\) represents the largest value of the
-gradient of \(-f\). On the other hand, \(\underset{x}{\arg\min}f'(x)\)
-represents the projection distance at which this largest gradient is
-attained. Based on the above discussion, it is expected that these two
-values should be both high in the case of high squintability (fast
-increase in \(f\) early on), and both low in the case of low
+\(-\min_{x}f'(x)\) represents the largest value of the gradient of
+\(-f\) and \(\underset{x}{\arg\min}f'(x)\) represents the projection
+distance at which this largest gradient is attained. It is expected that
+these two values should be both high in the case of high squintability
+(fast increase in \(f\) early on), and both low in the case of low
 squintability (any substantial increase in \(f\) happens very late,
@@
 
-The JSO performance is compared with the existing Creeping Random Search
-(CRS) \citep{RJ-2021-105, laa_using_2020} currently used in guided tour,
-and explores how JSO behaves with different hyper-parameter (number of
-jellyfishes and maximum tries) and data dimension combinations. A second
-simulation studies the effect of index properties (smoothness and
-squintability), along with JSO hyper-parameters, and data dimension, on
-the success rate of the JSO performance.
+The JSO performance is compared with an existing optimiser, Creeping
+Random Search (CRS) \citep{RJ-2021-105, laa_using_2020} used in the PP
+guided tour to explore JSO's behaviour under different hyper-parameter
+and data dimension combinations. The second simulation studies the
+effect of index properties (smoothness and squintability), along with
+JSO hyper-parameters, and data dimension, on the success rate of the JSO
+performance. This section describes the simulation details, with the
+results deferred to Section Section~\ref{sec-sim-res}.
 
@@
 
-The performance of JSO is investigated in comparison with the existing
-optimiser, CRS, with respect to two hyper-parameters (number of
-jellyfishes and maximum number of tries). The performance is measured by
-success rate, defined as the proportion of simulations that achieves a
-final index value within 0.05 of the best index value found among all 50
-simulations. This comparison is based on projection pursuit to find the
-pipe shape investigated by \citet{laa_using_2020} using two simple and
-well-behaved indexes, the \texttt{holes} index, and the
-\texttt{splines2d} index. Fifty simulations are conducted with both JSO
-and CRS, in four different data dimensions (\(d = 6, 8, 10, 12\)). JSO
-uses 100 jellyfishes with 100 maximum number of tries, while the CRS
-allows a maximum of 1000 samples at each iteration before the algorithm
-terminates. The different numbers accounts for the multiple paths of JSO
-to enable fairer comparison with CRS. The results of the simulation are
-collected using the data structure proposed in \citet{RJ-2021-105} is
-used here for assessing JSO, where the design parameters are stored
-along with index value, projection basis, random seed, and computation
-time.
-
-In this simulation, 50 jellyfishes (\texttt{n\_jellies}) and a maximum
-of 50 tries (\texttt{max\_tries}) are used. The full data contains 50
-simulations, indexed by the variable \texttt{sim}, with the associated
-\texttt{seed}s. Recorded in columns \texttt{basis}, \texttt{index\_val},
-\texttt{tries}, \texttt{loop} and \texttt{time}, are the projection
-basis, index value, iteration ID, jellyfish ID, and computational time,
-respectively.
-
-In the second experiment, 50 simulations are conducted at each
-hyper-parameter combination to analyse its effect on the JSO success
-rate: number of jellyfishes (20, 50, and 100), and maximum number of
-tries (50 and 100). Figure~\ref{fig-success-rate} illustrates the
-calculation based on the projection found by the above simulation. Here,
-the optimal index value is 0.999. Out of the 50 simulations, 44 achieved
-an index value within {[}0.949, 0.999{]}, resulting in a success rate of
-0.88.
+he performance of JSO is investigated both in comparison to the existing
+optimizer, CRS, and across various hyper-parameter values. The
+performance is measured by the success rate, defined as the proportion
+of simulations that achieves a final index value within 0.05 of the best
+index value found among all 50 simulations (see
+Figure~\ref{fig-success-rate} for an illustration). This comparison is
+based on projection pursuit to find the pipe shape investigated by
+\citet{laa_using_2020} using the \texttt{holes} index.
+
+Fifty simulations are conducted with both JSO and CRS, in four different
+data dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100 jellyfishes with a
+maximum of 100 tries, while the CRS allows a maximum of 1000 samples at
+each iteration before the algorithm terminates. The different numbers
+account for the multiple paths of JSO to enable fairer comparison with
+CRS. The results of the simulation are collected using the data
+structure proposed in \citet{RJ-2021-105} for assessing JSO, where the
+design parameters are stored along with index value, projection basis,
+random seed, and computation time.
+
+Fifty additional simulations are conducted for each hyper-parameter
+combination to analyze how they affect the JSO success rate. This
+includes variations in the number of jellyfish (20, 50, and 100) and the
+maximum number of tries (50 and 100).
 
@@
 \caption{\label{fig-success-rate}Illustration of success rate
-calculation: The best projections identified by the JSO in 50
-simulations are sorted by their index values. The highest index value
-found across all simulation is 0.999. Out of the 50 simulations, 44
-achieved an index value within 0.05 of the best, resulting in a success
-rate of 0.88 (44/50).}
+calculation: Final projections based on projection pursuit to find the
+pipe shape in 8D data using the holes index, optimised by CRS, in 50
+simulations. The 50 final projections are sorted by their index values.
+The highest index value found across all simulations is 0.969. Out of
+the 50 simulations, 43 achieved an index value within 0.05 of the best,
+resulting in a success rate of 0.86 (43/50).}
 
@@
 
-This simulation uses two different data shapes, pipe and a sine wave,
-investigated in 6D and 8D spaces with six different PP indexes:
-\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
-\texttt{spline}, and \texttt{stringy}. The two JSO hyper-parameters are
-varied. A total of 52 combinations result, comprising of 24 computed on
-the pipe data and 28 on the sine-wave data. For each case, JSO is run 50
-times to calculate the success rate.
-
-Smoothness and squintability are computed for each run, following the
-procedures outlined in Section~\ref{sec-smoothness} and
+To assess JSO's performance across various scenarios, two different data
+shapes, pipe and a sine wave, are investigated in 6D and 8D spaces using
+six different PP indexes: \texttt{dcor2d\_2}, \texttt{loess2d},
+\texttt{MIC}, \texttt{TIC}, \texttt{spline}, and \texttt{stringy}, with
+varied JSO hyper-parameters. A total of 52 combinations result,
+comprising of 24 computed on the pipe data and 28 on the sine-wave data.
+Again, JSO is run 50 times to calculate the success rate for each
+projection pursuit.
+
+Smoothness and squintability are computed following the procedures
+outlined in Section~\ref{sec-smoothness} and
 Section~\ref{sec-squintability} and as illustrated in
@@
 projection distances are calculated for these interpolated bases and the
-index values are averaged with a bin width of 0.005. The squintability
-measure is then calculated by fitting a four-parameter scaled logistic
-function of index value against projection distance, estimated by
-non-linear least squares.
-
-A generalised linear model is fitted to the resulting data, using a
-binomial family and a logit link function. The data is pre-processed by
-1) scaling the JSO hyper-parameters by a factor of 10 for
-interpretation, 2) creating a new binary variable \texttt{long\_time} to
-indicate cases with an average run time over 30 seconds, and 3)
-re-coding the success rate for the \texttt{stringy} index as 0, because
-none of the 50 simulations correctly identified the sine-wave shape.
+index values are averaged with a bin width of 0.005. A four-parameter
+scaled logistic function is fitted to the index values against
+projection distances, estimated by non-linear least squares. The
+squintability measure is then calculated as
+\ref{eq-squintability-parametric}.
+
+To construct a relationship among success rate, index properties
+(smoothness and squintability), and jellyfish hyper-parameters, a
+generalised linear model is fitted using a binomial family and a logit
+link function. The data is pre-processed by 1) scaling the JSO
+hyper-parameters by a factor of 10 for interpretation, 2) creating a new
+binary variable \texttt{long\_time} to indicate cases with an average
+run time over 30 seconds, and 3) re-coding the success rate for the
+\texttt{stringy} index as 0, because none of the 50 simulations
+correctly identified the sine-wave shape.
 
@@
 smoothness. For a given projection pursuit problem defined by the shape
-to find, data dimension and the index function, sample random bases
-given the orthonormality contraint. Calculate the projection distance
-and the index value for each random basis. Fit a Gaussian process model
-to the index values against the projection distances to obtain the
+to find, data dimension and the index function, 1) sample random bases
+given the orthonormality contraint, 2) calculate the projection distance
+and the index value for each random basis, and 3) fit a Gaussian process
+model of index values against projection distances to obtain the
 smoothness measure.}
@@
 squintability. For a given projection pursuit problem defined by the
-shape to find, data dimension and the index function, sample random
-bases given the orthonormality and projection distance contraint.
-Interpolate the sampled bases to the optimal basis and calculate the
-projection distance and the index value for each interpolated basis. Bin
-the index value by the projection distance, then claculate the average
-index value for each bin. Fit the scaled sigmoid function from Equation
-\ref{eq-squintability} to the binned index values against the projection
-distance using non-linear least square. Use Equation \ref{eq-parametric}
-to calculate the squintability measure from the estimated \(\theta_1\)
-to \(\theta_4\).}
+shape to find, data dimension and the index function, 1) sample random
+bases given the orthonormality and projection distance contraint, 2)
+interpolate the sampled bases to the optimal basis and calculate the
+projection distance and the index value for each interpolated basis. 3)
+bin the index values by projection distances to obtain the average index
+value for each bin, 4) fit the scaled sigmoid function in Equation
+\ref{eq-squintability} to the binned index values against projection
+distances using non-linear least square, 5) calculate the squintability
+measure using Equation \ref{eq-squintability-parametric} with parameters
+estimated from the model.}
 
@@
 
-The results from the two simulations described in
+The results from the first simulation described in
 Section~\ref{sec-sim-deets} are analysed based on the final projections
 across two optimisers (JSO and CRS) and the success rate across JSO
-hyper-parameters. Smoothness and squintability are calculated across a
-collection of pipe-finding and sine-wave finding problems to construct
-the relationship between success rate, JSO hyper-parameters, and index
-properties.
+hyper-parameters. In the second simulation, smoothness and squintability
+are calculated across a collection of pipe-finding and sine-wave finding
+problems to construct the relationship between success rate, JSO
+hyper-parameters, and index properties.
 
@@
 presented in Figure~\ref{fig-proj}, broken down by 10th quantile,
-faceted by the data dimension. The JSO uses 100 jellyfishes and a
-maximum number of tries of 100. The CRS uses a maximum of 1000 tries in
-each step of random sampling step before the algorithm terminates. In
-the 6-dimensional data scenario, JSO consistently identifies a clear
-pipe shape. The CRS also finds the pipe shape but with a wide rim,
-suggesting a further polish search may be required. With increasing
-dimensions, JSO may not always identify the pipe shape due to random
-sampling, but it still presents a pipe shape in over 50\% of cases. When
-compared to the CRS, JSO reaches higher index values and clearer pipe
-shapes across all quantiles. This suggests JSO can be more effective
-when searching higher dimension spaces, which may be preferred by users.
-
-The success rate calculated for each hyper-parameter combination is
-presented in Figure~\ref{fig-proportion}. As the number of jellyfishes
-and maximum tries increase, the success rate also increases. For simpler
-problems (6 dimensions), small parameter values (20 jellyfishes and a
-maximum of 50 tries) can achieve a high success rate. However, for
-higher-dimensional problems (8, 10, and 12 dimensions), larger parameter
-values (i.e.~100 jellyfishes and a maximum of 100 tries) are necessary.
+faceted by the data dimensions. In the 6-dimensional data scenario, JSO
+consistently identifies a clear pipe shape. The CRS also finds the pipe
+shape but with a wide rim, suggesting a further polish search may be
+required. With increasing dimensions, JSO may not always identify the
+pipe shape due to random sampling, but it still finds the pipe shape in
+over 50\% of cases. Compared to CRS, JSO achieves higher index values
+and clearer pipe shapes across all quantiles in data of 8, 10, 12
+dimensions, suggesting its advantage in exploring high-dimensional
+spaces.
+
+The success rate calculated at each hyper-parameter combination (number
+of jellyfish and the maximum number of tries) is presented in
+Figure~\ref{fig-proportion}. As the number of jellyfish and maximum
+tries increase, the success rate also increases. For simpler problems (6
+dimensions), small parameter values (20 jellyfishes and a maximum of 50
+tries) can already achieve a high success rate. However, larger
+parameter values (i.e.~100 jellyfishes and a maximum of 100 tries) are
+necessary for higher-dimensional problems (8, 10, and 12 dimensions).
 Increasing both parameters enhances the performance of JSO, but it also
@@
 
-\includegraphics{optim_files/figure-pdf/fig-proj-1.pdf}
+\includegraphics{jso_files/figure-pdf/fig-proj-1.pdf}
 
@@
 
-\includegraphics{optim_files/figure-pdf/fig-proportion-1.pdf}
+\includegraphics{jso_files/figure-pdf/fig-proportion-1.pdf}
 
@@
 optimisation problems. Table~\ref{tbl-smoothness-squintability} presents
-the parameters estimated from the Gaussian process (variance, range,
-smooth, and nugget) and the scaled logistic function (\(\theta_1\) to
-\(\theta_4\)) for calculating both metrics for each case considered in
-the simulation. The column ``smooth'' is used as the smoothness metric
-and the column ``squint'' is calculated as equation
+the parameters for calculating both metrics estimated from the Gaussian
+process (variance, range, smooth, and nugget) and the scaled logistic
+function (\(\theta_1\) to \(\theta_4\)) for each case considered in the
+simulation. The column ``smooth'' is used as the smoothness metric and
+the column ``squint'' is calculated as equation
 \ref{eq-squintability-parametric} as the squintability metric.
 Table~\ref{tbl-mod-output} presents the results of fitting a generalised
-linear model with a binomial family and a logit link function to the
-data in Table~\ref{tbl-mod-data}, where all the three components
-(success rate, JS hyper-parameters, and index properties) are combined.
-The model suggests that JSO success rate is positively associated with
-the two hyper-parameters, as well as with the index properties:
-smoothness and squintability. Specifically, using 10 more jellyfishes
-and 10 more tries increases the odd ratio of success by 24.11\% and
-11.93\%, respectively. However, being flagged with long runtime and an
-increase of data dimension reduce the success rate by 41.72\% and
-53.36\%, respectively. The variable \texttt{squintability} and
-\texttt{dimension} are significant, suggesting their importance relative
-to JSO hyper-parameters in the optimisation success. In defining a
-projection pursuit problem, factors such as the shape-to-find, the index
-function used, and the data dimension, determine the properties like
-smoothness and squintability. These metrics can then be compared across
-different problems to understand their relative complexity. The results
-suggest that squintability has a more significant influence than
-smoothness on the success rate of JSO. Once the characteristics of the
-projection pursuit problem are fully understood, increasing JSO
-hyper-parameters can enhance the search effectiveness, however, it is
-also important to consider the resulting increase in computational
-complexity.
+linear model with a binomial family and a logit link function to a
+sample of data in Table~\ref{tbl-mod-data}, where all the three
+components (success rate, JS hyper-parameters, and index properties) are
+combined. The model suggests that JSO success rate is positively
+associated with the two hyper-parameters, as well as with the index
+properties: smoothness and squintability. Specifically, using 10 more
+jellyfish and 10 more tries increases the odd ratio of success by
+24.11\% and 11.93\%, respectively. However, being flagged with long
+runtime and an increase of data dimension reduce the success rate by
+41.72\% and 53.36\%, respectively. The variable \texttt{squintability}
+and \texttt{dimension} are significant, suggesting their importance
+relative to JSO hyper-parameters in the optimisation success. In
+defining a projection pursuit problem, factors such as the
+shape-to-find, the index function used, and the data dimension,
+determine the properties such as smoothness and squintability. These
+metrics can then be compared across different problems to understand
+their relative complexites. The results suggest that squintability has a
+more significant influence than smoothness on the success rate of JSO.
+Once the characteristics of the projection pursuit problem are fully
+understood, increasing JSO hyper-parameters can enhance the search
+effectiveness, however, it is also important to consider the resulting
+increase in computational complexity.
 
@@
 The JSO performance is affected by the hyper-parameters, with a higher
-chance of reaching the global optima when more jellyfishes are used and
+chance of reaching the global optima when more jellyfish are used and
 the maximum number of tries is increased. However, it comes at a
@@
 paper can be found at:
-\url{https://github.com/huizezhang-sherry/paper-optim}. Nicolas Langrené
+\url{https://github.com/huizezhang-sherry/paper-jso}. Nicolas Langrené
 acknowledges the partial support of the Guangdong Provincial Key

### 2024-07-17 jso.tex
fixed a typo in bibliography

@@
 
-\begin{tcolorbox}[enhanced jigsaw, breakable, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, opacityback=0, arc=.35mm, toptitle=1mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, left=2mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, coltitle=black, rightrule=.15mm, bottomrule=.15mm, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, toptitle=1mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, opacityback=0, opacitybacktitle=0.6, coltitle=black, leftrule=.75mm, colback=white, arc=.35mm, bottomtitle=1mm, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm]
 

### 2024-07-21 jso.qmd
latex theorem
Added latex machinery for creating theorems, propositions, definitions, etc.

@@
    - \usepackage{algorithm}
+   - \usepackage{amsmath}
+   - \usepackage{amsthm}
+   - \usepackage{amssymb}
+   - \theoremstyle{plain}
+   - \newtheorem{assumption}{\protect\assumptionname}
+   - \newtheorem{claim}{\protect\claimname}
+   - \newtheorem{condition}{\protect\conditionname}
+   - \newtheorem{conjecture}{\protect\conjecturename}
+   - \newtheorem{cor}{\protect\corollaryname}
+   - \newtheorem{defn}{\protect\definitionname}
+   - \newtheorem{example}{\protect\examplename}
+   - \newtheorem{lem}{\protect\lemmaname}
+   - \newtheorem{notation}{\protect\notationname}
+   - \newtheorem{problem}{\protect\problemname}
+   - \newtheorem{prop}{\protect\propositionname}
+   - \newtheorem{rem}{\protect\remarkname}
+   - \newtheorem{thm}{\protect\theoremname}
+   - \providecommand{\assumptionname}{Assumption}
+   - \providecommand{\claimname}{Claim}
+   - \providecommand{\conditionname}{Condition}
+   - \providecommand{\conjecturename}{Conjecture}
+   - \providecommand{\corollaryname}{Corollary}
+   - \providecommand{\definitionname}{Definition}
+   - \providecommand{\examplename}{Example}
+   - \providecommand{\lemmaname}{Lemma}
+   - \providecommand{\notationname}{Notation}
+   - \providecommand{\problemname}{Problem}
+   - \providecommand{\propositionname}{Proposition}
+   - \providecommand{\remarkname}{Remark}
+   - \providecommand{\theoremname}{Theorem}
 author:
@@
 
+<!--  
+\begin{assumption}\label{assum:my_assumption}
+This is an assumption.
+\end{assumption}
+Assumption \ref{assum:my_assumption} is convenient.
+
+\begin{claim}\label{claim:my_claim}
+This is a claim.
+\end{claim}
+Claim \ref{claim:my_claim} is suspicious.
+
+\begin{condition}\label{cond:my_condition}
+This is a condition.
+\end{condition}
+Condition \ref{cond:my_condition} is necessary.
+
+\begin{conjecture}\label{conj:my_conjecture}
+This is a conjecture.
+\end{conjecture}
+Conjecture \ref{conj:my_conjecture} is intriguing.
+
+\begin{cor}\label{cor:my_corollary}
+This is a corollary.
+\end{cor}
+Corollary \ref{cor:my_corollary} is trivial.
+
+\begin{defn}\label{def:my_definition}
+This is a definition.
+\end{defn}
+Definition \ref{def:my_definition} is clear.
+
+\begin{example}\label{ex:my_example}
+This is an example.
+\end{example}
+Example \ref{ex:my_example} is enlightening.
+
+\begin{lem}\label{lem:my_lemma}
+This is a lemma.
+\end{lem}
+Lemma \ref{lem:my_lemma} is helpful.
+
+\begin{notation}\label{not:my_notation}
+This is a notation.
+\end{notation}
+Notation \ref{not:my_notation} is cumbersome.
+
+\begin{problem}\label{prob:my_problem}
+This is a problem.
+\end{problem}
+Problem \ref{prob:my_problem} is difficult.
+
+\begin{proof}\label{proof:my_proof}
+This is a proof.
+\end{proof}
+Proof \ref{proof:my_proof} is compelling.
+
+\begin{prop}\label{prop:my_prop}
+This is a proposition.
+\end{prop}
+Proposition \ref{prop:my_prop} is useful.
+
+\begin{rem}\label{rem:my_remark}
+This is a remark.
+\end{rem}
+Remark \ref{rem:my_remark} is interesting.
+
+\begin{thm}\label{thm:my_theorem}
+This is a theorem.
+\end{thm}
+Theorem \ref{thm:my_theorem} is important.
+-->
+
 Projection Pursuit (PP) (@kr69, @FT74) is a dimension reduction

### 2024-07-21 jso.tex
latex theorem
Added latex machinery for creating theorems, propositions, definitions, etc.

@@
 \usepackage{algorithm}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{amssymb}
+\theoremstyle{plain}
+\newtheorem{assumption}{\protect\assumptionname}
+\newtheorem{claim}{\protect\claimname}
+\newtheorem{condition}{\protect\conditionname}
+\newtheorem{conjecture}{\protect\conjecturename}
+\newtheorem{cor}{\protect\corollaryname}
+\newtheorem{defn}{\protect\definitionname}
+\newtheorem{example}{\protect\examplename}
+\newtheorem{lem}{\protect\lemmaname}
+\newtheorem{notation}{\protect\notationname}
+\newtheorem{problem}{\protect\problemname}
+\newtheorem{prop}{\protect\propositionname}
+\newtheorem{rem}{\protect\remarkname}
+\newtheorem{thm}{\protect\theoremname}
+\providecommand{\assumptionname}{Assumption}
+\providecommand{\claimname}{Claim}
+\providecommand{\conditionname}{Condition}
+\providecommand{\conjecturename}{Conjecture}
+\providecommand{\corollaryname}{Corollary}
+\providecommand{\definitionname}{Definition}
+\providecommand{\examplename}{Example}
+\providecommand{\lemmaname}{Lemma}
+\providecommand{\notationname}{Notation}
+\providecommand{\problemname}{Problem}
+\providecommand{\propositionname}{Proposition}
+\providecommand{\remarkname}{Remark}
+\providecommand{\theoremname}{Theorem}
 \makeatletter
@@
 \end{frontmatter}
-
-\section{Introduction}\label{introduction}
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, sharp corners, enhanced, boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, toptitle=1mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, opacityback=0, opacitybacktitle=0.6, coltitle=black, leftrule=.75mm, colback=white, arc=.35mm, bottomtitle=1mm, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, opacityback=0, left=2mm, toptitle=1mm, rightrule=.15mm, arc=.35mm, titlerule=0mm, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, bottomrule=.15mm, breakable, leftrule=.75mm, bottomtitle=1mm, opacitybacktitle=0.6, colback=white]
 
@@
 
-\centering{
-
-\includegraphics{jso_files/figure-pdf/fig-proj-1.pdf}
+{\centering \includegraphics{jso_files/figure-pdf/fig-proj-1.pdf}
 
@@
 
-\centering{
-
-\includegraphics{jso_files/figure-pdf/fig-proportion-1.pdf}
+{\centering \includegraphics{jso_files/figure-pdf/fig-proportion-1.pdf}
 

### 2024-07-24 jso.qmd
add definition for proj distance and squintability

@@
 
+\begin{defn}\label{def:proj-dist}
+Let $A \in \mathbf{R}^{p \times d}$ be the d-dimensional projection basis and
+$A^*$ be the optimal basis that achieves the maximum index value for a given
+PP problem defined by the shape to find, data dimension,
+and the index function. The projection distance $d(A, A^*)$ between $A$ 
+and $A^*$ is defined as 
+$d(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\  \rVert _F$
+where $\lVert . \rVert _F$ denotes the Frobenius norm, given by
+$\lVert M \rVert _F = \sqrt{\sum_{ij} M_{ij}^2}$. 
+\end{defn}
+
+We now focus our attention to how the index value $f(XA)$ changes with
+respect to the projection distance $d(A, A^*)$, over the course of the
+JSO.
+
+\begin{defn}\label{def:squintability}
+Let $g$ be the function that maps the projection distance, $d = d(A, A^*)$, 
+to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$.
+The squintability of $f(\cdot)$, denoted $\varsigma(f)$, is defined as 
+$$\varsigma(f) = c \times \max_{d} \Delta g(d) \times \arg \max_{d}  \Delta g(d)$$ 
+where $c$ is a constant scaling factor, $\max_A \Delta g(x)$ represents the 
+maximum change in the index value along the curve $g$, and $\arg \max_{d} \Delta g(x)$ represents the projection distance at which this largest gradient is attained.
+
+\end{defn}
+Definition \ref{def:squintability} is ...
+
 From @barnett1981interpreting and @laa_using_2020, a large squint angle
@@
 mathematical formulation of this intuition is proposed below:
-
-Let $A \in \mathbf{R}^{p \times d}$ be the d-D projection basis and
-$A^*$ be the optimal basis achieving the maximum index value for a given
-projection pursuit problem defined by the shape to find, data dimension,
-and the index function $f(.)$. The projection distance $d(A, A^*)$, is
-defined as the Frobenius norm $\lVert . \rVert _F$, between the squared
-difference of these two bases:
-
-$$d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .$$ {#eq-proj-dist}
-
-We now focus our attention to how the index value $f(XA)$ changes with
-respect to the projection distance $d(A, A^*)$, over the course of the
-JSO.
 

### 2024-07-24 jso.tex
add definition for proj distance and squintability

--- 2024-07-17 jso.tex
@@
 
-\begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, toptitle=1mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, opacityback=0, opacitybacktitle=0.6, coltitle=black, leftrule=.75mm, colback=white, arc=.35mm, bottomtitle=1mm, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, bottomrule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, colback=white, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, titlerule=0mm, arc=.35mm, opacityback=0, leftrule=.75mm, toprule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, left=2mm]
 
@@
 are described.
+
+\begin{defn}\label{def:proj-dist}
+Let $A \in \mathbf{R}^{p \times d}$ be the d-dimensional projection basis and
+$A^*$ be the optimal basis that achieves the maximum index value for a given
+PP problem defined by the shape to find, data dimension,
+and the index function. The projection distance $d(A, A^*)$ between $A$ 
+and $A^*$ is defined as 
+$d(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\  \rVert _F$
+where $\lVert . \rVert _F$ denotes the Frobenius norm, given by
+$\lVert M \rVert _F = \sqrt{\sum_{ij} M_{ij}^2}$. 
+\end{defn}
+
+We now focus our attention to how the index value \(f(XA)\) changes with
+respect to the projection distance \(d(A, A^*)\), over the course of the
+JSO.
+
+\begin{defn}\label{def:squintability}
+Let $g$ be the function that maps the projection distance, $d = d(A, A^*)$, 
+to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$.
+The squintability of $f(\cdot)$, denoted $\varsigma(f)$, is defined as 
+$$\varsigma(f) = c \times \max_{d} \Delta g(d) \times \arg \max_{d}  \Delta g(d)$$ 
+where $c$ is a constant scaling factor, $\max_A \Delta g(x)$ represents the 
+maximum change in the index value along the curve $g$, and $\arg \max_{d} \Delta g(x)$ represents the projection distance at which this largest gradient is attained.
+
+\end{defn}
+
+Definition \ref{def:squintability} is \ldots{}
 
@@
 below:
-
-Let \(A \in \mathbf{R}^{p \times d}\) be the d-D projection basis and
-\(A^*\) be the optimal basis achieving the maximum index value for a
-given projection pursuit problem defined by the shape to find, data
-dimension, and the index function \(f(.)\). The projection distance
-\(d(A, A^*)\), is defined as the Frobenius norm \(\lVert . \rVert _F\),
-between the squared difference of these two bases:
-
-\begin{equation}\phantomsection\label{eq-proj-dist}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .}\end{equation}
-
-We now focus our attention to how the index value \(f(XA)\) changes with
-respect to the projection distance \(d(A, A^*)\), over the course of the
-JSO.
 

### 2024-07-25 jso.qmd
smoothness metric
Updated smoothness metrics subsection, with definitions and more details.

@@
 
+```{=html}
 <!--  
@@
 -->
-
+```
 Projection Pursuit (PP) (@kr69, @FT74) is a dimension reduction
@@
 
-If a PP index function is evaluated at some random points, as is done at
-the initialization stage of JSO, then these random index values can be
+An classical way to describe the smoothness of a function is to identify
+how many continuous derivatives of the function exist. This can be
+characterized by Sobolev spaces [@adams2003sobolev].
+
+```{=tex}
+\begin{defn}\label{def:sobolev_space}
+The Sobolev space $W^{k,p}(\mathbb{R})$ for $1\leq p\leq \infty$ is the set of all functions $f$ in $L^p(\mathbb{R})$ for which all weak derivatives $f^{(\ell)}$ of order $\ell\leq k$ exist and have a finite $L^p$ norm.
+\end{defn}
+```
+The Sobolev index $k$ in Definition \ref{def:sobolev_space} can be used
+to characterize the smoothness of a function: if $f\in W^{k,p}$, then
+the higher $k$, the smoother $f$. While this Sobolev index $k$ is a
+useful measure of smoothness, it can be difficult to compute or even estimate
+in practice.
+
+To obtain a computable estimator of the smoothness of the index function
+$f$, we propose an approach based on random fields. If a PP index
+function $f$ is evaluated at some random bases, as is done at the
+initialization stage of JSO, then these random index values can be
 interpreted as a random field, indexed by a space parameter, namely the
-random projection angle. This analogy suggests to use this random
-training sample to fit a spatial model, a simple one being a (spatial)
-Gaussian process.
+random projection basis. This analogy suggests to use this random
+training sample to fit a spatial model. We propose to use a Gaussian
+process equipped with a Matérn covariance function, due to the
+connections between this model and Sobolev spaces, see for example
+@porcu2024matern.
 
@@
 definition of the covariance function: if a PP index is very smooth,
-then two close projection angles should produce close index values
+then two close projection bases should produce close index values
 (strong correlation); by contrast, if a PP index is not very smooth,
-then two close projection angles might give very different index values
-(fast decay of correlations with respect to distance between angles).
-
-Popular covariance functions are parametric positive semi-definite
-functions, some of which have a parameter to capture the smoothness of
-the Gaussian field. In particular, consider the Matérn class of
-covariance functions, defined by
-
-$$
-K(u):=\frac{(\sqrt{2\nu}u)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}(\sqrt{2\nu}u)\ ,
-$$
-
-where $\nu>0$ is the smoothness parameter and where $\mathcal{K}_\nu$ is
-the modified Bessel function. The Matérn covariance function can be
-expressed analytically when $\nu$ is a half-integer, the most popular
-values in the literature being $\frac{1}{2}$, $\frac{3}{2}$ and
-$\frac{5}{2}$. The parameter $\nu$, called *smoothness parameter*,
-controls the decay of the covariance function. As such, it is an
-appropriate measure of smoothness of a random field.
-
-In this context, the parameter $\nu$ is suggested as a measure of the
-smoothness of the index function by fitting a Gaussian process prior
-with Matérn covariance on a dataset generated by random evaluations of
-the index function, as done at the initialization stage of the jellyfish
-search optimization. There exist several R packages, such as `GpGp`
-[@guinness2021gpgp] or `ExaGeoStatR` [@abdulah2023large], to fit the
-hyperparameters of a GP covariance function on data. In this project,
-the `GpGp` package is used.
-
-After fitting the smoothness parameter $\nu>0$, its value can be
-interpreted as follows: the higher $\nu$, the smoother the index
-function.
+then two close projection bases might give very different index values
+(fast decay of correlations with respect to distance between bases). Popular covariance functions are parametric positive semi-definite
+functions.  In particular, the Matérn class of
+covariance functions has a dedicated parameter to capture the smoothness of
+the Gaussian field.
+
+```{=tex}
+\begin{defn}
+The Matérn covariance function $K$ is defined by
+\begin{equation}
+K(u)=K_{\nu,\eta,\ell}(u):=\eta^2\frac{\left(\sqrt{2\nu}\frac{u}{\ell}\right)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}\left(\sqrt{2\nu}\frac{u}{\ell}\right)\ , u\geq0\label{eq:matern}
+\end{equation}
+where $\nu>0$ is the smoothness parameter, $\eta$ is the outputscale, $\ell$ is the lengthscale, and $\mathcal{K}_\nu$ is
+the modified Bessel function \citep[10.25]{DLMF}. A multivariate extension $K(u)$, $u\in\mathbb{R}^d$ can be obtained by products of univariate covariance functions \eqref{eq:matern}.
+\end{defn}
+```
+The Matérn covariance function can be expressed analytically when $\nu$
+is a half-integer, the most popular values in the literature being
+$\frac{1}{2}$, $\frac{3}{2}$ and $\frac{5}{2}$ [@rasmussen2006gaussian]. The parameter $\nu$,
+called *smoothness parameter*, controls the decay of the covariance
+function. As such, it is an appropriate measure of smoothness of a
+random field. For example, @karvonen2023asymptotic showed that if a
+function $f$ has a Sobolev index of $k$, then the smoothness parameter
+estimate $\nu$ in \eqref{eq:matern} cannot be asymptotically less than
+$k$. See the survey @porcu2024matern for additional results on the connection between the Matérn model and Sobolev spaces. An interesting result is that the asymptotic case
+$\nu\rightarrow\infty$ coincides with the Gaussian kernel:
+$K_\infty(u)=\exp(-u^{2}/2)$.
+
+In view of these results, the parameter $\nu$ is suggested as a measure
+of the smoothness of the PP index function by fitting a Gaussian process
+prior with Matérn covariance on a dataset generated by random
+evaluations of the index function, as done at the initialization stage
+of the jellyfish search optimization. There exist several R packages,
+such as `GpGp` [@guinness2021gpgp] or `ExaGeoStatR` [@abdulah2023large],
+to fit the hyperparameters of a GP covariance function on data, which is
+usually done by maximum likelihood estimation. In this project, the
+`GpGp` package is used.
+
+```{=tex}
+\begin{defn}
+Let $\mathbf{A}=[A_1, \ldots, A_N] \in (\mathbf{R}^{p \times d})^N$ be d-dimensional projection bases, let $\mathbf{y}=[f(XA_1),\ldots,f(XA_N)]$ be the corresponding PP index values, and let $\mathbf{K}=[K_\theta(A_{i},A_{j})]_{1\leq i,j\leq N}\in\mathbb{R}^{N\times N}$ be the Matérn covariance matrix evaluated at the input bases, where the vector $\theta$ contains all the parameters of the multivariate Matérn covariance function $K$ (smoothness, outputscale, lengthscales).
+The log-likelihood of the parameters $\theta$ is defined by
+\begin{equation}
+\mathcal{L}(\theta)=\log p(\mathbf{y}\left|\mathbf{A},\theta\right.)=-\frac{1}{2}\mathbf{y}^{\top}(\mathbf{K}+\sigma^{2}\mathbf{I})^{-1}\mathbf{y}-\frac{1}{2}\mathrm{\log}(\det(\mathbf{K}+\sigma^{2}\mathbf{I}))-\frac{N}{2}\log(2\pi)\,.\label{eq:gp_log_likelihood}
+\end{equation}
+where the nugget parameter $\sigma$ is the standard deviation of the intrinsic noise of the Gaussian process.
+The optimal parameters (including smoothness) are obtained by maximum log-likelihood
+\begin{equation}
+\theta^* = \underset{\theta}{\max}\mathcal{L}(\theta)
+\end{equation}
+The resulting optimal smoothness parameter $\nu$ is chosen as our smoothness metric.
+\end{defn}
+```
+
+The value of the optimal smoothness parameter $\nu>0$ can be
+naturally interpreted as follows: the higher $\nu$, the smoother the
+index function.
 
@@
 
+```{=tex}
 \begin{defn}\label{def:proj-dist}
@@
 \end{defn}
-
+```
 We now focus our attention to how the index value $f(XA)$ changes with
@@
 
+```{=tex}
 \begin{defn}\label{def:squintability}
@@
 \end{defn}
+```
 Definition \ref{def:squintability} is ...

### 2024-07-25 jso.tex
smoothness metric
Updated smoothness metrics subsection, with definitions and more details.

--- 2024-07-21 jso.tex
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, sharp corners, enhanced, boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, opacityback=0, left=2mm, toptitle=1mm, rightrule=.15mm, arc=.35mm, titlerule=0mm, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, bottomrule=.15mm, breakable, leftrule=.75mm, bottomtitle=1mm, opacitybacktitle=0.6, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, opacityback=0, coltitle=black, arc=.35mm, toprule=.15mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, breakable, colframe=quarto-callout-note-color-frame, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, bottomrule=.15mm, titlerule=0mm, rightrule=.15mm, toptitle=1mm, left=2mm]
 
@@
 
-If a PP index function is evaluated at some random points, as is done at
-the initialization stage of JSO, then these random index values can be
+An classical way to describe the smoothness of a function is to identify
+how many continuous derivatives of the function exist. This can be
+characterized by Sobolev spaces \citep{adams2003sobolev}.
+
+\begin{defn}\label{def:sobolev_space}
+The Sobolev space $W^{k,p}(\mathbb{R})$ for $1\leq p\leq \infty$ is the set of all functions $f$ in $L^p(\mathbb{R})$ for which all weak derivatives $f^{(\ell)}$ of order $\ell\leq k$ exist and have a finite $L^p$ norm.
+\end{defn}
+
+The Sobolev index \(k\) in Definition \ref{def:sobolev_space} can be
+used to characterize the smoothness of a function: if \(f\in W^{k,p}\),
+then the higher \(k\), the smoother \(f\). While this Sobolev index
+\(k\) is a useful measure of smoothness, it can be difficult to compute
+or even estimate in practice.
+
+To obtain a computable estimator of the smoothness of the index function
+\(f\), we propose an approach based on random fields. If a PP index
+function \(f\) is evaluated at some random bases, as is done at the
+initialization stage of JSO, then these random index values can be
 interpreted as a random field, indexed by a space parameter, namely the
-random projection angle. This analogy suggests to use this random
-training sample to fit a spatial model, a simple one being a (spatial)
-Gaussian process.
+random projection basis. This analogy suggests to use this random
+training sample to fit a spatial model. We propose to use a Gaussian
+process equipped with a Matérn covariance function, due to the
+connections between this model and Sobolev spaces, see for example
+\citet{porcu2024matern}.
 
@@
 definition of the covariance function: if a PP index is very smooth,
-then two close projection angles should produce close index values
+then two close projection bases should produce close index values
 (strong correlation); by contrast, if a PP index is not very smooth,
-then two close projection angles might give very different index values
-(fast decay of correlations with respect to distance between angles).
-
+then two close projection bases might give very different index values
+(fast decay of correlations with respect to distance between bases).
 Popular covariance functions are parametric positive semi-definite
-functions, some of which have a parameter to capture the smoothness of
-the Gaussian field. In particular, consider the Matérn class of
-covariance functions, defined by
-
-\[
-K(u):=\frac{(\sqrt{2\nu}u)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}(\sqrt{2\nu}u)\ ,
-\]
-
-where \(\nu>0\) is the smoothness parameter and where
-\(\mathcal{K}_\nu\) is the modified Bessel function. The Matérn
-covariance function can be expressed analytically when \(\nu\) is a
-half-integer, the most popular values in the literature being
-\(\frac{1}{2}\), \(\frac{3}{2}\) and \(\frac{5}{2}\). The parameter
-\(\nu\), called \emph{smoothness parameter}, controls the decay of the
-covariance function. As such, it is an appropriate measure of smoothness
-of a random field.
-
-In this context, the parameter \(\nu\) is suggested as a measure of the
-smoothness of the index function by fitting a Gaussian process prior
-with Matérn covariance on a dataset generated by random evaluations of
-the index function, as done at the initialization stage of the jellyfish
-search optimization. There exist several R packages, such as
-\texttt{GpGp} \citep{guinness2021gpgp} or \texttt{ExaGeoStatR}
+functions. In particular, the Matérn class of covariance functions has a
+dedicated parameter to capture the smoothness of the Gaussian field.
+
+\begin{defn}
+The Matérn covariance function $K$ is defined by
+\begin{equation}
+K(u)=K_{\nu,\eta,\ell}(u):=\eta^2\frac{\left(\sqrt{2\nu}\frac{u}{\ell}\right)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}\left(\sqrt{2\nu}\frac{u}{\ell}\right)\ , u\geq0\label{eq:matern}
+\end{equation}
+where $\nu>0$ is the smoothness parameter, $\eta$ is the outputscale, $\ell$ is the lengthscale, and $\mathcal{K}_\nu$ is
+the modified Bessel function \citep[10.25]{DLMF}. A multivariate extension $K(u)$, $u\in\mathbb{R}^d$ can be obtained by products of univariate covariance functions \eqref{eq:matern}.
+\end{defn}
+
+The Matérn covariance function can be expressed analytically when
+\(\nu\) is a half-integer, the most popular values in the literature
+being \(\frac{1}{2}\), \(\frac{3}{2}\) and \(\frac{5}{2}\)
+\citep{rasmussen2006gaussian}. The parameter \(\nu\), called
+\emph{smoothness parameter}, controls the decay of the covariance
+function. As such, it is an appropriate measure of smoothness of a
+random field. For example, \citet{karvonen2023asymptotic} showed that if
+a function \(f\) has a Sobolev index of \(k\), then the smoothness
+parameter estimate \(\nu\) in \eqref{eq:matern} cannot be asymptotically
+less than \(k\). See the survey \citet{porcu2024matern} for additional
+results on the connection between the Matérn model and Sobolev spaces.
+An interesting result is that the asymptotic case
+\(\nu\rightarrow\infty\) coincides with the Gaussian kernel:
+\(K_\infty(u)=\exp(-u^{2}/2)\).
+
+In view of these results, the parameter \(\nu\) is suggested as a
+measure of the smoothness of the PP index function by fitting a Gaussian
+process prior with Matérn covariance on a dataset generated by random
+evaluations of the index function, as done at the initialization stage
+of the jellyfish search optimization. There exist several R packages,
+such as \texttt{GpGp} \citep{guinness2021gpgp} or \texttt{ExaGeoStatR}
 \citep{abdulah2023large}, to fit the hyperparameters of a GP covariance
-function on data. In this project, the \texttt{GpGp} package is used.
-
-After fitting the smoothness parameter \(\nu>0\), its value can be
+function on data, which is usually done by maximum likelihood
+estimation. In this project, the \texttt{GpGp} package is used.
+
+\begin{defn}
+Let $\mathbf{A}=[A_1, \ldots, A_N] \in (\mathbf{R}^{p \times d})^N$ be d-dimensional projection bases, let $\mathbf{y}=[f(XA_1),\ldots,f(XA_N)]$ be the corresponding PP index values, and let $\mathbf{K}=[K_\theta(A_{i},A_{j})]_{1\leq i,j\leq N}\in\mathbb{R}^{N\times N}$ be the Matérn covariance matrix evaluated at the input bases, where the vector $\theta$ contains all the parameters of the multivariate Matérn covariance function $K$ (smoothness, outputscale, lengthscales).
+The log-likelihood of the parameters $\theta$ is defined by
+\begin{equation}
+\mathcal{L}(\theta)=\log p(\mathbf{y}\left|\mathbf{A},\theta\right.)=-\frac{1}{2}\mathbf{y}^{\top}(\mathbf{K}+\sigma^{2}\mathbf{I})^{-1}\mathbf{y}-\frac{1}{2}\mathrm{\log}(\det(\mathbf{K}+\sigma^{2}\mathbf{I}))-\frac{N}{2}\log(2\pi)\,.\label{eq:gp_log_likelihood}
+\end{equation}
+where the nugget parameter $\sigma$ is the standard deviation of the intrinsic noise of the Gaussian process.
+The optimal parameters (including smoothness) are obtained by maximum log-likelihood
+\begin{equation}
+\theta^* = \underset{\theta}{\max}\mathcal{L}(\theta)
+\end{equation}
+The resulting optimal smoothness parameter $\nu$ is chosen as our smoothness metric.
+\end{defn}
+
+The value of the optimal smoothness parameter \(\nu>0\) can be naturally
 interpreted as follows: the higher \(\nu\), the smoother the index
@@
 below:
-
-Let \(A \in \mathbf{R}^{p \times d}\) be the d-D projection basis and
-\(A^*\) be the optimal basis achieving the maximum index value for a
-given projection pursuit problem defined by the shape to find, data
-dimension, and the index function \(f(.)\). The projection distance
-\(d(A, A^*)\), is defined as the Frobenius norm \(\lVert . \rVert _F\),
-between the squared difference of these two bases:
-
-\begin{equation}\protect\hypertarget{eq-proj-dist}{}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .}\label{eq-proj-dist}\end{equation}
-
-We now focus our attention to how the index value \(f(XA)\) changes with
-respect to the projection distance \(d(A, A^*)\), over the course of the
-JSO.
 

### 2024-07-28 jso.qmd
consolidate the definition in PP and squintability

@@
   eq-prefix: ""
+editor_options: 
+  chunk_output_type: console
 ---
@@
 criteria for what constitutes interesting or informative projections.
-Given high-dimensional data $X \in \mathbf{R}^{n\times p}$ and an index
-function $f(\cdot)$, PP finds the orthonormal basis
-$A \in \mathbf{R}^{p \times d}$ that maximises the index value of the
-projection, $Y = XA$:
+Let $X \in \mathbb{R}^{n\times p}$ be the data matrix, $A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix, where $A$ belongs to the Stiefel manifold $\mathcal{A} = V_d(\mathbb{R}^p)$.
+The projection $Y = XA$ is a linear transformation that maps data from a $p$-dimensional space into a $d$-dimensional space.
+The index function $f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ is a scalar function that measures an interesting aspect of the projected data, such as deviation from normality, presence of clusters, non-linear structure, or other features of interest.
+For a fixed data, PP finds the orthonormal basis $A$ that maximises the index value of the projection, $Y = XA$:
 
 $$
-\underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
+\underset{A \in \mathcal{A}}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I_d
 $$ {#eq-optimization}
@@
 ```{=tex}
-\begin{defn}\label{def:proj-dist}
+\begin{defn}[projection distance]\label{def:proj-dist}
 Let $A \in \mathbf{R}^{p \times d}$ be the d-dimensional projection basis and
@@
 
+<!-- Squintability is defined as how rapidly and significantly the index value, $f(XA)$ response to the decrease of projection distance to optimal, $d(A, A^*)$. -->
+
 ```{=tex}
-\begin{defn}\label{def:squintability}
-Let $g$ be the function that maps the projection distance, $d = d(A, A^*)$, 
-to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$.
-The squintability of $f(\cdot)$, denoted $\varsigma(f)$, is defined as 
-$$\varsigma(f) = c \times \max_{d} \Delta g(d) \times \arg \max_{d}  \Delta g(d)$$ 
-where $c$ is a constant scaling factor, $\max_A \Delta g(x)$ represents the 
-maximum change in the index value along the curve $g$, and $\arg \max_{d} \Delta g(x)$ represents the projection distance at which this largest gradient is attained.
+\begin{defn}[squintability]\label{def:squintability}
+Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance, $d = d(A, A^*)$, 
+to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$. The squintability of the index function $f(\cdot)$, is defined as 
+$$\varsigma(f) = -c \times \max_{d} g'(d) \times \arg \max_{d} g'(d)$$ 
+where $c$ is a constant scaling factor, $-\max_d g'(d)$ represents the 
+largest gradient of $-g$ and $\arg \max_{d} g'(d)$ represents the projection distance at which this largest gradient is attained.
 
@@
 ```
-Definition \ref{def:squintability} is ...
+
+<!-- Definition \ref{def:squintability} is ... -->
+
+\begin{prop}\label{prop:my_prop}
+This is a proposition.
+\end{prop}
+Proposition \ref{prop:my_prop} is useful.
 
@@
   geom_text(data = aaa |> select(-V1, -V2) |> unique(), 
-            aes(label = round(index_val,3)), x =0, y = 3.7, 
+            aes(label = round(index_val,3)), x = 0, y = 3.7, 
             size = 3) + 
@@
     levels = paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th")), 
-    d= 4 + 2 * as.numeric(dim)) |> 
+    d = 4 + 2 * as.numeric(dim)) |> 
   unnest(proj)
@@
   geom_text(data = bbb |> select(-V1, -V2) |> unique(), 
-            aes(label = round(index_val,3)), x =0, y = 3.7, 
+            aes(label = round(index_val,3)), x = 0, y = 3.7, 
             size = 3) + 

### 2024-07-28 jso.tex
consolidate the definition in PP and squintability

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced]}{\end{tcolorbox}}\fi
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+
+\section{Introduction}\label{introduction}
 
@@
 criteria for what constitutes interesting or informative projections.
-Given high-dimensional data \(X \in \mathbf{R}^{n\times p}\) and an
-index function \(f(\cdot)\), PP finds the orthonormal basis
-\(A \in \mathbf{R}^{p \times d}\) that maximises the index value of the
+Let \(X \in \mathbb{R}^{n\times p}\) be the data matrix,
+\(A \in\mathbb{R}^{p \times d}\) be an orthonormal matrix, where \(A\)
+belongs to the Stiefel manifold \(\mathcal{A} = V_d(\mathbb{R}^p)\). The
+projection \(Y = XA\) is a linear transformation that maps data from a
+\(p\)-dimensional space into a \(d\)-dimensional space. The index
+function \(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar
+function that measures an interesting aspect of the projected data, such
+as deviation from normality, presence of clusters, non-linear structure,
+or other features of interest. For a fixed data, PP finds the
+orthonormal basis \(A\) that maximises the index value of the
 projection, \(Y = XA\):
 
-\begin{equation}\protect\hypertarget{eq-optimization}{}{
-\underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
-}\label{eq-optimization}\end{equation}
+\begin{equation}\phantomsection\label{eq-optimization}{
+\underset{A \in \mathcal{A}}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I_d
+}\end{equation}
 
@@
 
-\hypertarget{sec-background}{%
 \section{Projection pursuit, tours, index functions and
-optimisation}\label{sec-background}}
+optimisation}\label{sec-background}
 
@@
 
-\hypertarget{sec-theory}{%
 \section{The jellyfish optimiser and properties of PP
-indexes}\label{sec-theory}}
+indexes}\label{sec-theory}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, opacityback=0, coltitle=black, arc=.35mm, toprule=.15mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, breakable, colframe=quarto-callout-note-color-frame, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, bottomrule=.15mm, titlerule=0mm, rightrule=.15mm, toptitle=1mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, toprule=.15mm, bottomtitle=1mm, toptitle=1mm, arc=.35mm, coltitle=black, left=2mm, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, rightrule=.15mm, bottomrule=.15mm, opacitybacktitle=0.6, colback=white, opacityback=0]
 
@@
 
-\begin{defn}\label{def:proj-dist}
+\begin{defn}[projection distance]\label{def:proj-dist}
 Let $A \in \mathbf{R}^{p \times d}$ be the d-dimensional projection basis and
@@
 
-\begin{defn}\label{def:squintability}
-Let $g$ be the function that maps the projection distance, $d = d(A, A^*)$, 
-to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$.
-The squintability of $f(\cdot)$, denoted $\varsigma(f)$, is defined as 
-$$\varsigma(f) = c \times \max_{d} \Delta g(d) \times \arg \max_{d}  \Delta g(d)$$ 
-where $c$ is a constant scaling factor, $\max_A \Delta g(x)$ represents the 
-maximum change in the index value along the curve $g$, and $\arg \max_{d} \Delta g(x)$ represents the projection distance at which this largest gradient is attained.
+\begin{defn}[squintability]\label{def:squintability}
+Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance, $d = d(A, A^*)$, 
+to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$. The squintability of the index function $f(\cdot)$, is defined as 
+$$\varsigma(f) = -c \times \max_{d} g'(d) \times \arg \max_{d} g'(d)$$ 
+where $c$ is a constant scaling factor, $-\max_d g'(d)$ represents the 
+largest gradient of $-g$ and $\arg \max_{d} g'(d)$ represents the projection distance at which this largest gradient is attained.
 
@@
 
-Definition \ref{def:squintability} is \ldots{}
+\begin{prop}\label{prop:my_prop}
+This is a proposition.
+\end{prop}
+
+Proposition \ref{prop:my_prop} is useful.
 
@@
 
-\hypertarget{sec-sim-deets}{%
-\section{Simulation details}\label{sec-sim-deets}}
+\section{Simulation details}\label{sec-sim-deets}
 
@@
 
-\hypertarget{sec-app-1}{%
-\subsection{Performance of JSO relative to CRS}\label{sec-app-1}}
+\subsection{Performance of JSO relative to CRS}\label{sec-app-1}
 
@@
 
-{\centering \includegraphics[width=4.52in,height=\textheight]{figures/success-rate.png}
+\centering{
+
+\includegraphics[width=4.52in,height=\textheight]{figures/success-rate.png}
 
@@
 
-\end{figure}
-
-\hypertarget{sec-app-2}{%
+\end{figure}%
+
 \subsection{Factors affecting JSO success rate: index properties and
-jellyfish hyper-parameters}\label{sec-app-2}}
+jellyfish hyper-parameters}\label{sec-app-2}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/smoothness.png}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{figures/smoothness.png}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/squintability.png}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{figures/squintability.png}
 
@@
 
-\end{figure}
-
-\hypertarget{sec-sim-res}{%
-\section{Results}\label{sec-sim-res}}
+\end{figure}%
+
+\section{Results}\label{sec-sim-res}
 
@@
 
-{\centering \includegraphics{jso_files/figure-pdf/fig-proj-1.pdf}
+\centering{
+
+\includegraphics{jso_files/figure-pdf/fig-proj-1.pdf}
 
@@
 
-{\centering \includegraphics{jso_files/figure-pdf/fig-proportion-1.pdf}
+\centering{
+
+\includegraphics{jso_files/figure-pdf/fig-proportion-1.pdf}
 
@@
 \endgroup{}
-\end{table}
-
-\hypertarget{tbl-mod-data}{}
+
+}
+
+\end{table}%
+
 \begin{table}
-\caption{\label{tbl-mod-data}The first 7 rows of the datasets processed for modelling. }\tabularnewline
+
+\caption{\label{tbl-mod-data}The first 7 rows of the datasets processed
+for modelling.}
+
+\centering{
 
@@
 \endgroup{}
-\end{table}
-
-\hypertarget{tbl-mod-output}{}
+
+}
+
+\end{table}%
+
 \begin{table}
-\caption{\label{tbl-mod-output}Model estimates of proportion of jellyfish success on index properties
-and jellyfish hyper-parameters from the generalised linear model with a
-binomial family and a logit link function. The variable smoothness,
-squintability, number of jellyfish and maximum number of tries are
-positively associated with JSO success rate while data dimension and
-being flagged as long runtime are negatively associated with the success
-rate. The variable squintability and dimension are significant,
-suggesting their importance relative to jellyfish hyper-parameters in
-the optimisation success. }\tabularnewline
+
+\caption{\label{tbl-mod-output}Model estimates of proportion of
+jellyfish success on index properties and jellyfish hyper-parameters
+from the generalised linear model with a binomial family and a logit
+link function. The variable smoothness, squintability, number of
+jellyfish and maximum number of tries are positively associated with JSO
+success rate while data dimension and being flagged as long runtime are
+negatively associated with the success rate. The variable squintability
+and dimension are significant, suggesting their importance relative to
+jellyfish hyper-parameters in the optimisation success.}
+
+\centering{
 
@@
 \endgroup{}
-\end{table}
-
-\hypertarget{sec-conclusion}{%
-\section{Conclusion}\label{sec-conclusion}}
+
+}
+
+\end{table}%
+
+\section{Conclusion}\label{sec-conclusion}
 
@@
 
-\hypertarget{acknowledgement}{%
-\section{Acknowledgement}\label{acknowledgement}}
+\section{Acknowledgement}\label{acknowledgement}
 

### 2024-07-29 jso.qmd
add definition for squint angle

@@
 \begin{defn}[projection distance]\label{def:proj-dist}
-Let $A \in \mathbf{R}^{p \times d}$ be the d-dimensional projection basis and
-$A^*$ be the optimal basis that achieves the maximum index value for a given
-PP problem defined by the shape to find, data dimension,
-and the index function. The projection distance $d(A, A^*)$ between $A$ 
+Recall that $A \in \mathbf{R}^{p \times d}$ is a $d$-dimensional orthonormal matrix.
+Let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.
+The projection distance $d(A, A^*)$ between $A$
 and $A^*$ is defined as 
@@
 ```
+
 We now focus our attention to how the index value $f(XA)$ changes with
@@
 
-<!-- Squintability is defined as how rapidly and significantly the index value, $f(XA)$ response to the decrease of projection distance to optimal, $d(A, A^*)$. -->
+```{=tex}
+\begin{defn}[squint angle]\label{def:squint-angle}
+Let $A$ and $B$ be two $d$-dimensional orthonormal matrices in $\mathbb{R}^p$.
+The squint angle $\theta$ between the subspace spanned by $A$ and $B$ is defined as the smallest principal angle between these subspaces: $\theta = \min_{i \in \{1, \cdots, d\}} \arccos(\tau_i)$, where $\tau_i$ are the singular values of the matrix $M = A^T B$ obtained from its singular value decomposition.
+
+\end{defn}
+```
 
@@
 \begin{defn}[squintability]\label{def:squintability}
-Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance, $d = d(A, A^*)$, 
-to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$. The squintability of the index function $f(\cdot)$, is defined as 
-$$\varsigma(f) = -c \times \max_{d} g'(d) \times \arg \max_{d} g'(d)$$ 
+Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance $d(A, A^*)$ to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$.
+For brevity, denote $g(d)$ as $g(d(A, A^*))$.
+The squintability of an index function $f(XA)$ is defined as
+
+$$\varsigma(f) = -c \times \max_{d} g'(d) \times \arg \max_{d} g'(d) $$
+
 where $c$ is a constant scaling factor, $-\max_d g'(d)$ represents the 

### 2024-07-29 jso.tex
add definition for squint angle

@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, toprule=.15mm, bottomtitle=1mm, toptitle=1mm, arc=.35mm, coltitle=black, left=2mm, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, rightrule=.15mm, bottomrule=.15mm, opacitybacktitle=0.6, colback=white, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, opacityback=0, titlerule=0mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, opacitybacktitle=0.6, rightrule=.15mm, coltitle=black, breakable, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, arc=.35mm, bottomrule=.15mm, leftrule=.75mm, toprule=.15mm]
 
@@
 
+\begin{defn}[squint angle]\label{def:squint-angle}
+Let $A$ and $B$ be two $d$-dimensional orthonormal matrices in $\mathbb{R}^p$.
+The squint angle $\theta$ between the subspace spanned by $A$ and $B$ is defined as the smallest principal angle between these subspaces: $\theta = \min_{i \in \{1, \cdots, d\}} \arccos(\tau_i)$, where $\tau_i$ are the singular values of the matrix $M = A^T B$ obtained from its singular value decomposition.
+
+\end{defn}
+
 \begin{defn}[squintability]\label{def:squintability}
-Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance, $d = d(A, A^*)$, 
-to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$. The squintability of the index function $f(\cdot)$, is defined as 
-$$\varsigma(f) = -c \times \max_{d} g'(d) \times \arg \max_{d} g'(d)$$ 
+Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance $d(A, A^*)$ to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$.
+For brevity, denote $g(d)$ as $g(d(A, A^*))$.
+The squintability of an index function $f(XA)$ is defined as
+
+$$\varsigma(f) = -c \times \max_{d} g'(d) \times \arg \max_{d} g'(d) $$
+
 where $c$ is a constant scaling factor, $-\max_d g'(d)$ represents the 
@@
 these values are attained, are crucial in the mathematical definition of
-squintability. Noticing that \(f\) is expected to be a decreasing
-function, the squintability metric is proposed as follows:
+squintability.
+
+Noticing that \(f\) is expected to be a decreasing function, the
+squintability metric is proposed as follows:
 
@@
 \(-f\) and \(\underset{x}{\arg\min}f'(x)\) represents the projection
-distance at which this largest gradient is attained. It is expected that
-these two values should be both high in the case of high squintability
-(fast increase in \(f\) early on), and both low in the case of low
-squintability (any substantial increase in \(f\) happens very late,
-close to the optimal angle). This suggest that their product
+distance at which this largest gradient is attained.
+
+It is expected that these two values should be both high in the case of
+high squintability (fast increase in \(f\) early on), and both low in
+the case of low squintability (any substantial increase in \(f\) happens
+very late, close to the optimal angle). This suggest that their product
 (\ref{eq-squintability}) should provide a sensible measure of
-squintability. The multiplicative constant \(4\), which can be deemed
-arbitrary, does not change the interpretation of the squintability
-metric \(\varsigma\); it is here to adjust the range of values of
-\(\varsigma\) and simplify the explicit formula for \(\varsigma\)
-obtained later on.
+squintability.
+
+The multiplicative constant \(4\), which can be deemed arbitrary, does
+not change the interpretation of the squintability metric \(\varsigma\);
+it is here to adjust the range of values of \(\varsigma\) and simplify
+the explicit formula for \(\varsigma\) obtained later on.
 

### 2024-08-02 jso.qmd
a proposition

@@
 
-Squintability has been a useful concept. Here, it is defined
-mathematically and two approaches to compute this metric numerically,
+Here the formal definition of projection distance and squint angle are given, before the definition of squintability. Two approaches to compute this metric numerically,
 are described.
@@
 
-We now focus our attention to how the index value $f(XA)$ changes with
-respect to the projection distance $d(A, A^*)$, over the course of the
-JSO.
-
 ```{=tex}
@@
 ```
+
+Squintability can be defined as how the index value $f(XA)$ changes with
+respect to the projection distance $d(A, A^*)$, over the course of the
+JSO:
 
@@
 
-$$\varsigma(f) = -c \times \max_{d} g'(d) \times \arg \max_{d} g'(d) $$
+\begin{equation}
+\varsigma(f) = -c \times \max_{d} g'(d) \times \arg \max_{d} g'(d)
+\label{eq-squintability}
+\end{equation}
 
@@
 
-<!-- Definition \ref{def:squintability} is ... -->
-
-\begin{prop}\label{prop:my_prop}
-This is a proposition.
+It is expected that these two
+values should be both high in the case of high squintability (fast
+increase in $f$ early on), and both low in the case of low squintability
+(any substantial increase in $f$ happens very late, close to the optimal
+angle). This suggests that their product (@eq-squintability) should
+provide a sensible measure of squintability. The multiplicative constant
+$4$, which can be deemed arbitrary, does not change the interpretation
+of the squintability metric $\varsigma$; it is here to adjust the range
+of values of $\varsigma$ and simplify the explicit formula for
+$\varsigma$ obtained later on.
+
+From the definition, the following proposition can be derived:
+
+\begin{prop}\label{prop:convex-concave}
+Let $g_1(d)$ be a convex function and $g_2(d)$ be a concave function, where $d \in [0, D]$ and $g_1(d), g_2(d) \in [0, 1]$.
+Assume both functions achieve the same maximum gradient, i.e., $\max g'_1(d_1) = \max g'_2(d_2) = -g_{max}$, it follows that $d_1 < d_2$ and $\varsigma(f_1) < \varsigma(f_2)$.
 \end{prop}
-Proposition \ref{prop:my_prop} is useful.
 
@@
 attained, are crucial in the mathematical definition of squintability.
-Noticing that $f$ is expected to be a decreasing function, the
-squintability metric is proposed as follows:
-
-$$
-\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ .
-$$ {#eq-squintability}
-
-$-\min_{x}f'(x)$ represents the largest value of the gradient of $-f$
-and $\underset{x}{\arg\min}f'(x)$ represents the projection distance at
-which this largest gradient is attained. It is expected that these two
-values should be both high in the case of high squintability (fast
-increase in $f$ early on), and both low in the case of low squintability
-(any substantial increase in $f$ happens very late, close to the optimal
-angle). This suggest that their product (@eq-squintability) should
-provide a sensible measure of squintability. The multiplicative constant
-$4$, which can be deemed arbitrary, does not change the interpretation
-of the squintability metric $\varsigma$; it is here to adjust the range
-of values of $\varsigma$ and simplify the explicit formula for
-$\varsigma$ obtained later on.
+
+<!-- Noticing that $f$ is expected to be a decreasing function, the -->
+<!-- squintability metric is proposed as follows: -->
+
+<!-- $$ -->
+<!-- \varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ . -->
+<!-- $$ {#eq-squintability} -->
+
+<!-- $-\min_{x}f'(x)$ represents the largest value of the gradient of $-f$ -->
+<!-- and $\underset{x}{\arg\min}f'(x)$ represents the projection distance at -->
+<!-- which this largest gradient is attained.  -->
 

### 2024-08-02 jso.tex
a proposition

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, opacityback=0, titlerule=0mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, opacitybacktitle=0.6, rightrule=.15mm, coltitle=black, breakable, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, arc=.35mm, bottomrule=.15mm, leftrule=.75mm, toprule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, breakable, arc=.35mm, colframe=quarto-callout-note-color-frame, colback=white, rightrule=.15mm, coltitle=black, toprule=.15mm, toptitle=1mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacitybacktitle=0.6, bottomtitle=1mm, bottomrule=.15mm, left=2mm]
 
@@
 
-Squintability has been a useful concept. Here, it is defined
-mathematically and two approaches to compute this metric numerically,
-are described.
+Here the formal definition of projection distance and squint angle are
+given, before the definition of squintability. Two approaches to compute
+this metric numerically, are described.
 
@@
 
-We now focus our attention to how the index value \(f(XA)\) changes with
-respect to the projection distance \(d(A, A^*)\), over the course of the
-JSO.
-
 \begin{defn}[squint angle]\label{def:squint-angle}
@@
 \end{defn}
+
+Squintability can be defined as how the index value \(f(XA)\) changes
+with respect to the projection distance \(d(A, A^*)\), over the course
+of the JSO:
 
@@
 
-\begin{prop}\label{prop:my_prop}
-This is a proposition.
+It is expected that these two values should be both high in the case of
+high squintability (fast increase in \(f\) early on), and both low in
+the case of low squintability (any substantial increase in \(f\) happens
+very late, close to the optimal angle). This suggests that their product
+(\textbf{?@eq-squintability}) should provide a sensible measure of
+squintability. The multiplicative constant \(4\), which can be deemed
+arbitrary, does not change the interpretation of the squintability
+metric \(\varsigma\); it is here to adjust the range of values of
+\(\varsigma\) and simplify the explicit formula for \(\varsigma\)
+obtained later on.
+
+From the definition, the following proposition can be derived:
+
+\begin{prop}\label{prop:convex-concave}
+Let $g_1(d)$ be a convex function and $g_2(d)$ be a concave function, where $d \in [0, D]$ and $g_1(d), g_2(d) \in [0, 1]$.
+Assume both functions achieve the same maximum gradient, i.e., $\max g'_1(d_1) = \max g'_2(d_2) = -g_{max}$, it follows that $d_1 < d_2$ and $\varsigma(f_1) < \varsigma(f_2)$.
 \end{prop}
-
-Proposition \ref{prop:my_prop} is useful.
 
@@
 
-Noticing that \(f\) is expected to be a decreasing function, the
-squintability metric is proposed as follows:
-
-\begin{equation}\phantomsection\label{eq-squintability}{
-\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ .
-}\end{equation}
-
-\(-\min_{x}f'(x)\) represents the largest value of the gradient of
-\(-f\) and \(\underset{x}{\arg\min}f'(x)\) represents the projection
-distance at which this largest gradient is attained.
-
-It is expected that these two values should be both high in the case of
-high squintability (fast increase in \(f\) early on), and both low in
-the case of low squintability (any substantial increase in \(f\) happens
-very late, close to the optimal angle). This suggest that their product
-(\ref{eq-squintability}) should provide a sensible measure of
-squintability.
-
-The multiplicative constant \(4\), which can be deemed arbitrary, does
-not change the interpretation of the squintability metric \(\varsigma\);
-it is here to adjust the range of values of \(\varsigma\) and simplify
-the explicit formula for \(\varsigma\) obtained later on.
-
-To compute the squintability metric (\ref{eq-squintability}) in
+To compute the squintability metric (\textbf{?@eq-squintability}) in
 practice, several approaches are possible. The first one is to propose a
@@
 \(f(x_{\max})=\theta_4\). Under the parametric model
-(\ref{eq-parametric}), the squintability metric (\ref{eq-squintability})
-can be shown to be equal to
+(\ref{eq-parametric}), the squintability metric
+(\textbf{?@eq-squintability}) can be shown to be equal to
 
@@
 
-Alternatively, one can estimate (\ref{eq-squintability}) in a
+Alternatively, one can estimate (\textbf{?@eq-squintability}) in a
 nonparametric way, for example by fitting \(f\) using kernel regression,
@@
 value for each bin, 4) fit the scaled sigmoid function in Equation
-\ref{eq-squintability} to the binned index values against projection
-distances using non-linear least square, 5) calculate the squintability
-measure using Equation \ref{eq-squintability-parametric} with parameters
-estimated from the model.}
+\textbf{?@eq-squintability} to the binned index values against
+projection distances using non-linear least square, 5) calculate the
+squintability measure using Equation \ref{eq-squintability-parametric}
+with parameters estimated from the model.}
 

### 2024-08-10 jso.qmd
Matérn simulation figures
added Matérn simulation figures in R and R^2 (Figure 1 and Figure 2)

@@
 
-An classical way to describe the smoothness of a function is to identify
+A classical way to describe the smoothness of a function is to identify
 how many continuous derivatives of the function exist. This can be
@@
 function. As such, it is an appropriate measure of smoothness of a
-random field. For example, @karvonen2023asymptotic showed that if a
+random field, as shown by the simulations on @fig-matern-1d and @fig-matern-2d. For example, @karvonen2023asymptotic showed that if a
 function $f$ has a Sobolev index of $k$, then the smoothness parameter
@@
 $K_\infty(u)=\exp(-u^{2}/2)$.
+
+```{r}
+#| label: fig-matern-1d
+#| out.width: 100%
+#| fig.cap: |
+#|   Five random simulations from a Gaussian Process defined on $\mathbb{R}$ with zero mean and Matérn-$\nu$ covariance function, with $\nu=1$ (left), $\nu=2$ (middle), and $\nu=4$ (right), showing that higher values of $\nu$ produce smoother curves.
+knitr::include_graphics(here::here("figures", "matern_simulation_1d.png"))
+```
+
+```{r}
+#| label: fig-matern-2d
+#| out.width: 100%
+#| fig.cap: |
+#|   One random simulation from a Gaussian Process defined on $\mathbb{R}^2$ with zero mean and Matérn-$\nu$ covariance function, with $\nu=1$ (left), $\nu=2$ (middle), and $\nu=4$ (right), showing that higher values of $\nu$ produce smoother surfaces.
+knitr::include_graphics(here::here("figures", "matern_simulation_2d.png"))
+```
 
@@
 (any substantial increase in $f$ happens very late, close to the optimal
-angle). This suggests that their product (@eq-squintability) should
+angle). This suggests that their product \eqref{eq-squintability} should
 provide a sensible measure of squintability. The multiplicative constant
@@
 
-To compute the squintability metric (@eq-squintability) in practice,
+To compute the squintability metric \eqref{eq-squintability} in practice,
 several approaches are possible. The first one is to propose a
@@
 Under the parametric model (@eq-parametric), the squintability metric
-(@eq-squintability) can be shown to be equal to
+\eqref{eq-squintability} can be shown to be equal to
 
@@
 
-Alternatively, one can estimate (@eq-squintability) in a nonparametric
+Alternatively, one can estimate \eqref{eq-squintability} in a nonparametric
 way, for example by fitting $f$ using kernel regression, then
@@
 squintability measure is then calculated as
-@eq-squintability-parametric.
+(@eq-squintability-parametric).
 
@@
 
-<!-- the best basis and it is calculated as the Frobenius norm in Equation -->
-
-<!-- @eq-proj-dist. Plotting projection distances against index values for the  -->
+<!-- the best basis and it is calculated as the Frobenius norm in equation -->
+
+<!-- (@eq-proj-dist). Plotting projection distances against index values for the  -->
 
@@
 #| out.width: 100%
-#| fig.cap: "Illustration of steps to calculate squintability.
+#| fig.cap: |
+#|   Illustration of steps to calculate squintability.
 For a given projection pursuit problem defined by the shape to find, data dimension and the index function, 1) sample random bases given the orthonormality and projection distance contraint, 2) interpolate the sampled bases to the optimal basis and calculate the projection distance and the index value for each interpolated basis.
-3) bin the index values by projection distances to obtain the average index value for each bin, 4) fit the scaled sigmoid function in Equation @eq-squintability to the binned index values against projection distances using non-linear least square, 5) calculate the squintability measure using Equation @eq-squintability-parametric with parameters estimated from the model."
+3) bin the index values by projection distances to obtain the average index value for each bin, 4) fit the scaled sigmoid function in equation \eqref{eq-squintability} to the binned index values against projection distances using non-linear least square, 5) calculate the squintability measure using equation (@eq-squintability-parametric) with parameters estimated from the model.
 # p1 <- sim_sine_6d_spline_projdist |> 
@@
 simulation. The column "smooth" is used as the smoothness metric and the
-column "squint" is calculated as equation @eq-squintability-parametric
+column "squint" is calculated as equation (@eq-squintability-parametric)
 as the squintability metric. @tbl-mod-output presents the results of

### 2024-08-10 jso.tex
Matérn simulation figures
added Matérn simulation figures in R and R^2 (Figure 1 and Figure 2)

@@
 \end{frontmatter}
-
-\section{Introduction}\label{introduction}
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-\begin{equation}\phantomsection\label{eq-optimization}{
+\begin{equation}\protect\hypertarget{eq-optimization}{}{
 \underset{A \in \mathcal{A}}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I_d
-}\end{equation}
+}\label{eq-optimization}\end{equation}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, breakable, arc=.35mm, colframe=quarto-callout-note-color-frame, colback=white, rightrule=.15mm, coltitle=black, toprule=.15mm, toptitle=1mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacitybacktitle=0.6, bottomtitle=1mm, bottomrule=.15mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, left=2mm, toptitle=1mm, opacityback=0, opacitybacktitle=0.6, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, leftrule=.75mm, colback=white, rightrule=.15mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, coltitle=black, bottomtitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm]
 
@@
 function. As such, it is an appropriate measure of smoothness of a
-random field. For example, \citet{karvonen2023asymptotic} showed that if
-a function \(f\) has a Sobolev index of \(k\), then the smoothness
-parameter estimate \(\nu\) in \eqref{eq:matern} cannot be asymptotically
-less than \(k\). See the survey \citet{porcu2024matern} for additional
-results on the connection between the Matérn model and Sobolev spaces.
-An interesting result is that the asymptotic case
-\(\nu\rightarrow\infty\) coincides with the Gaussian kernel:
-\(K_\infty(u)=\exp(-u^{2}/2)\).
+random field, as shown by the simulations on Figure~\ref{fig-matern-1d}
+and Figure~\ref{fig-matern-2d}. For example,
+\citet{karvonen2023asymptotic} showed that if a function \(f\) has a
+Sobolev index of \(k\), then the smoothness parameter estimate \(\nu\)
+in \eqref{eq:matern} cannot be asymptotically less than \(k\). See the
+survey \citet{porcu2024matern} for additional results on the connection
+between the Matérn model and Sobolev spaces. An interesting result is
+that the asymptotic case \(\nu\rightarrow\infty\) coincides with the
+Gaussian kernel: \(K_\infty(u)=\exp(-u^{2}/2)\).
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_1d.png}
+
+}
+
+\caption{\label{fig-matern-1d}Five random simulations from a Gaussian
+Process defined on \(\mathbb{R}\) with zero mean and Matérn-\(\nu\)
+covariance function, with \(\nu=1\) (left), \(\nu=2\) (middle), and
+\(\nu=4\) (right), showing that higher values of \(\nu\) produce
+smoother curves.}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_2d.png}
+
+}
+
+\caption{\label{fig-matern-2d}One random simulation from a Gaussian
+Process defined on \(\mathbb{R}^2\) with zero mean and Matérn-\(\nu\)
+covariance function, with \(\nu=1\) (left), \(\nu=2\) (middle), and
+\(\nu=4\) (right), showing that higher values of \(\nu\) produce
+smoother surfaces.}
+
+\end{figure}
 
@@
 very late, close to the optimal angle). This suggests that their product
-(\textbf{?@eq-squintability}) should provide a sensible measure of
+\eqref{eq-squintability} should provide a sensible measure of
 squintability. The multiplicative constant \(4\), which can be deemed
@@
 
-To compute the squintability metric (\textbf{?@eq-squintability}) in
+To compute the squintability metric \eqref{eq-squintability} in
 practice, several approaches are possible. The first one is to propose a
@@
 \(f(x_{\max})=\theta_4\). Under the parametric model
-(\ref{eq-parametric}), the squintability metric
-(\textbf{?@eq-squintability}) can be shown to be equal to
-
-\begin{equation}\phantomsection\label{eq-squintability-parametric}{
+(\ref{eq-parametric}), the squintability metric \eqref{eq-squintability}
+can be shown to be equal to
+
+\begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
 \varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
-}\end{equation}
+}\label{eq-squintability-parametric}\end{equation}
 
@@
 
-Alternatively, one can estimate (\textbf{?@eq-squintability}) in a
+Alternatively, one can estimate \eqref{eq-squintability} in a
 nonparametric way, for example by fitting \(f\) using kernel regression,
@@
 squintability measure is then calculated as
-\ref{eq-squintability-parametric}.
+(\ref{eq-squintability-parametric}).
 
@@
 bin the index values by projection distances to obtain the average index
-value for each bin, 4) fit the scaled sigmoid function in Equation
-\textbf{?@eq-squintability} to the binned index values against
-projection distances using non-linear least square, 5) calculate the
-squintability measure using Equation \ref{eq-squintability-parametric}
-with parameters estimated from the model.}
-
-\end{figure}%
-
-\section{Results}\label{sec-sim-res}
+value for each bin, 4) fit the scaled sigmoid function in equation
+\eqref{eq-squintability} to the binned index values against projection
+distances using non-linear least square, 5) calculate the squintability
+measure using equation (\ref{eq-squintability-parametric}) with
+parameters estimated from the model.}
+
+\end{figure}
+
+\hypertarget{sec-sim-res}{%
+\section{Results}\label{sec-sim-res}}
 
@@
 the column ``squint'' is calculated as equation
-\ref{eq-squintability-parametric} as the squintability metric.
+(\ref{eq-squintability-parametric}) as the squintability metric.
 Table~\ref{tbl-mod-output} presents the results of fitting a generalised

### 2024-08-15 jso.qmd
added on prop 1

@@
 \begin{prop}\label{prop:convex-concave}
-Let $g_1(d)$ be a convex function and $g_2(d)$ be a concave function, where $d \in [0, D]$ and $g_1(d), g_2(d) \in [0, 1]$.
-Assume both functions achieve the same maximum gradient, i.e., $\max g'_1(d_1) = \max g'_2(d_2) = -g_{max}$, it follows that $d_1 < d_2$ and $\varsigma(f_1) < \varsigma(f_2)$.
+Let $g_1(d)$ be a convex decreasing function and $g_2(d)$ be a concave decreasing function defined on [0, 1] where $d \in [0, D]$.
+Let $d_1 := \underset{d}{\arg \max} |g_1'(d)|$ and $d_2 := \underset{d}{\arg \max} |g_2'(d)|$.
+
+$$\varsigma(g_1) < \varsigma(g_2)$$
+
+For a convex function, $g_1'(d) < 0 \; \forall d$ and is non-decreasing. For a concave function, $g_2'(d) < 0 \; \forall d$ and is non-increasing. As such, it follows that $d_1 < d_2$.
+
+Suppose both functions achieve the same maximum gradient $-g_{\max}$, i.e., $\max g'_1(d_1) = \max g'_2(d_2) = -g_{\max}$,
+
+$$
+\varsigma(g_1(d_1)) = -c  \max g'_1(d_1)  d_1 = -c  (-g_{\max}) d_1 \leq -c  (-g_{\max}) d_2 = -c \max g'_2(d_2)  d_2 = \varsigma(g_2(d_2))
+$$
+
 \end{prop}

### 2024-08-15 jso.tex
added on prop 1

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, left=2mm, toptitle=1mm, opacityback=0, opacitybacktitle=0.6, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, leftrule=.75mm, colback=white, rightrule=.15mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, coltitle=black, bottomtitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, toprule=.15mm, coltitle=black, colframe=quarto-callout-note-color-frame, breakable, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, bottomtitle=1mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, rightrule=.15mm, bottomrule=.15mm, left=2mm, leftrule=.75mm, opacityback=0, colback=white]
 

### 2024-08-16 jso.qmd
change to JCGS templateg

@@
 format:
-  elsevier-pdf:
+  tandf-pdf:
     keep-tex: true
-    journal:
-      name: Journal of Multivariate Analysis
-      formatting: preprint
-      model: 3p
-      cite-style: numbername
+    linestretch: 2
+    fontsize: 12pt
 editor: 
@@
 where $\nu>0$ is the smoothness parameter, $\eta$ is the outputscale, $\ell$ is the lengthscale, and $\mathcal{K}_\nu$ is
-the modified Bessel function \citep[10.25]{DLMF}. A multivariate extension $K(u)$, $u\in\mathbb{R}^d$ can be obtained by products of univariate covariance functions \eqref{eq:matern}.
+the modified Bessel function @DLMF. A multivariate extension $K(u)$, $u\in\mathbb{R}^d$ can be obtained by products of univariate covariance functions \eqref{eq:matern}.
 \end{defn}

### 2024-08-16 jso.tex
change to JCGS templateg

--- 2024-08-02 jso.tex
@@
 \documentclass[
-  number,
-  preprint,
-  3p]{elsarticle}
+  12pt,
+]{interact}
 
 \usepackage{amsmath,amssymb}
+\usepackage{setspace}
 \usepackage{iftex}
@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 
@@
 \makeatother
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
+  \fi
+  % set entry spacing
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 \usepackage{xcolor}
+\usepackage{orcidlink}
 \usepackage{algorithm}
@@
 \makeatother
-\journal{Journal of Multivariate Analysis}
+
 \ifLuaTeX
@@
 \fi
-\usepackage[]{natbib}
-\bibliographystyle{elsarticle-num-names}
 \usepackage{bookmark}
@@
 
-\setlength{\parindent}{6pt}
-\begin{document}
-
-\begin{frontmatter}
 \title{Studying the Performance of the Jellyfish Search Optimiser for
 the Application of Projection Pursuit}
-\author[1]{H. Sherry Zhang%
-\corref{cor1}%
-}
- \ead{huize.zhang@austin.utexas.edu} 
-\author[2]{Dianne Cook%
-%
-}
- \ead{dicook@monash.edu} 
-\author[3]{Nicolas Langrené%
-%
-}
- \ead{nicolaslangrene@uic.edu.cn} 
-\author[2]{Jessica Wai Yin Leung%
-%
-}
- \ead{Jessica.Leung@monash.edu}
-
-\affiliation[1]{organization={University of Texas at Austin, Department
-of Statistics and Data Sciences},city={Austin},country={United
-States},countrysep={,},postcode={78751},postcodesep={}}
-\affiliation[2]{organization={Monash University, Department of
-Econometrics and Business
-Statistics},city={Melbourne},country={Australia},countrysep={,},postcode={3800},postcodesep={}}
-\affiliation[3]{organization={Guangdong Provincial Key Laboratory of
+\author{H. Sherry Zhang$\textsuperscript{1}$, Dianne
+Cook$\textsuperscript{2}$, Nicolas
+Langrené$\textsuperscript{3}$, Jessica Wai Yin
+Leung$\textsuperscript{2}$}
+
+\thanks{CONTACT: H. Sherry
+Zhang. Email: \href{mailto:huize.zhang@austin.utexas.edu}{\nolinkurl{huize.zhang@austin.utexas.edu}}. Dianne
+Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}. Nicolas
+Langrené. Email: \href{mailto:nicolaslangrene@uic.edu.cn}{\nolinkurl{nicolaslangrene@uic.edu.cn}}. Jessica
+Wai Yin
+Leung. Email: \href{mailto:Jessica.Leung@monash.edu}{\nolinkurl{Jessica.Leung@monash.edu}}. }
+\begin{document}
+\captionsetup{labelsep=space}
+\maketitle
+\textsuperscript{1} Department of Statistics and Data
+Sciences, University of Texas at Austin, Austin, United
+States\\ \textsuperscript{2} Department of Econometrics and Business
+Statistics, Monash
+University, Melbourne, Australia\\ \textsuperscript{3} Department of
+Mathematical Sciences, Guangdong Provincial Key Laboratory of
 Interdisciplinary Research and Application for Data Science,
-BNU-HKBU~United~International~College, Department of Mathematical
-Sciences},city={Zhuhai},country={China},countrysep={,},postcode={519087},postcodesep={}}
-
-\cortext[cor1]{Corresponding author}
-
+BNU-HKBU~United~International~College, Zhuhai, China
 \begin{abstract}
@@
 \end{abstract}
-
-\begin{keyword}
-    projection pursuit \sep jellyfish search optimiser
-(JSO) \sep optimisation \sep grand tour \sep high-dimensional data \sep 
+\begin{keywords}
+\def\sep{;\ }
+projection pursuit\sep jellyfish search optimiser
+(JSO)\sep optimisation\sep grand tour\sep high-dimensional data\sep 
 exploratory data analysis
-\end{keyword}
-\end{frontmatter}
-
+\end{keywords}
+
+\setstretch{2}
 \section{Introduction}\label{introduction}
 
-Projection Pursuit (PP) (\citet{kr69}, \citet{FT74}) is a dimension
-reduction technique aimed at identifying informative linear projections
-of data. The method involves optimising an objective function known as
-the PP index (e.g. \citet{hall1989polynomial},
-\citet{cook1993projection}, \citet{lee2010projection},
-\citet{Loperfido2018}, \citet{Loperfido2020}), which defines the
-criteria for what constitutes interesting or informative projections.
-Let \(X \in \mathbb{R}^{n\times p}\) be the data matrix,
-\(A \in\mathbb{R}^{p \times d}\) be an orthonormal matrix, where \(A\)
-belongs to the Stiefel manifold \(\mathcal{A} = V_d(\mathbb{R}^p)\). The
-projection \(Y = XA\) is a linear transformation that maps data from a
-\(p\)-dimensional space into a \(d\)-dimensional space. The index
-function \(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar
-function that measures an interesting aspect of the projected data, such
-as deviation from normality, presence of clusters, non-linear structure,
-or other features of interest. For a fixed data, PP finds the
-orthonormal basis \(A\) that maximises the index value of the
-projection, \(Y = XA\):
+Projection Pursuit (PP) (Kruskal (1969), Friedman and Tukey (1974)) is a
+dimension reduction technique aimed at identifying informative linear
+projections of data. The method involves optimising an objective
+function known as the PP index (e.g. Hall (1989), D. Cook, Buja, and
+Cabrera (1993), Lee and Cook (2010), Loperfido (2018), Loperfido
+(2020)), which defines the criteria for what constitutes interesting or
+informative projections. Let \(X \in \mathbb{R}^{n\times p}\) be the
+data matrix, \(A \in\mathbb{R}^{p \times d}\) be an orthonormal matrix,
+where \(A\) belongs to the Stiefel manifold
+\(\mathcal{A} = V_d(\mathbb{R}^p)\). The projection \(Y = XA\) is a
+linear transformation that maps data from a \(p\)-dimensional space into
+a \(d\)-dimensional space. The index function
+\(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar function
+that measures an interesting aspect of the projected data, such as
+deviation from normality, presence of clusters, non-linear structure, or
+other features of interest. For a fixed data, PP finds the orthonormal
+basis \(A\) that maximises the index value of the projection,
+\(Y = XA\):
 
@@
 Optimisation of PP is typically investigated in the literature when new
-indexes are proposed \citep{posse1995, marie-sainte2010, grochowski2011}
-or when visualisation method are used to track the optimisation process.
-\citet{cook1995grand} introduced the PP guided tour, which monitors the
-optimisation visually to see the data projections leading in and out of
-the optima. An implementation is available in the \texttt{tourr} package
-\citep{tourr} in R \citep{R}. \citet{RJ-2021-105} illustrated how to
-diagnose optimisation processes, particularly focusing on the guided
-tour, and revealed a need for improved optimisation. While improving the
-quality of the optimisation solutions in the tour is essential, it is
-also important to be able to view the data projections as the
-optimisation progresses. Integrating the guided tour with a global
-optimisation algorithm that is efficient in finding the global optimal
-and enables viewing of the projected data during the exploration process
-is a goal.
+indexes are proposed (Posse 1995a; Marie-Sainte, Berro, and Ruiz-Gazen
+2010; Grochowski and Duch 2011) or when visualisation method are used to
+track the optimisation process. Dianne Cook et al. (1995) introduced the
+PP guided tour, which monitors the optimisation visually to see the data
+projections leading in and out of the optima. An implementation is
+available in the \texttt{tourr} package (Wickham et al. 2011) in R (R
+Core Team 2023). Zhang et al. (2021) illustrated how to diagnose
+optimisation processes, particularly focusing on the guided tour, and
+revealed a need for improved optimisation. While improving the quality
+of the optimisation solutions in the tour is essential, it is also
+important to be able to view the data projections as the optimisation
+progresses. Integrating the guided tour with a global optimisation
+algorithm that is efficient in finding the global optimal and enables
+viewing of the projected data during the exploration process is a goal.
 
@@
 integrated with the projection pursuit guided tour is explored. JSO,
-\citep{chou_novel_2021} - \citep{rajwar_exhaustive_2023}, inspired by
-the search behaviour of jellyfish in the ocean, is a swarm-based
+(Chou and Truong 2021) - (Rajwar, Deep, and Das 2023), inspired by the
+search behaviour of jellyfish in the ocean, is a swarm-based
 metaheuristic designed to solve global optimisation problems. Compared
@@
 metrics are introduced to capture specific properties of the index
-including squintability (based on \citet{barnett1981interpreting}'s
-squint angle) and smoothness as introduced in \citet{laa_using_2020}.
-Here we mathematically define metrics for squintability and smoothness,
-which is new for the field. A series of simulation experiments using
-various datasets and PP indexes are conducted to assess JSO's behaviour
-and its sensitivity to hyper-parameter choices (number of jellyfish and
-maximum number of tries). The relationship between the JSO performance,
+including squintability (based on Tukey and Tukey (1981)'s squint angle)
+and smoothness as introduced in Laa and Cook (2020). Here we
+mathematically define metrics for squintability and smoothness, which is
+new for the field. A series of simulation experiments using various
+datasets and PP indexes are conducted to assess JSO's behaviour and its
+sensitivity to hyper-parameter choices (number of jellyfish and maximum
+number of tries). The relationship between the JSO performance,
 hyper-parameter choices and properties of PP indexes (smoothness and
@@
   \texttt{search\_better\_random()}: is essentially simulated annealing
-  \citep{Bertsimas93} where the search space is reduced as the
-  optimisation progresses.
+  (Bertsimas and Tsitsiklis 1993) where the search space is reduced as
+  the optimisation progresses.
 \item
-  \texttt{search\_posse()}: implements the algorithm described in
-  \citet{posse95}.
+  \texttt{search\_posse()}: implements the algorithm described in Posse
+  (1995b).
 \item
@@
 There are several PP index functions available: \texttt{holes()} and
-\texttt{cmass()} \citep{cook1993projection}; \texttt{lda\_pp()}
-\citep{lee2005projection}; \texttt{pda\_pp()} \citep{lee2010projection};
-\texttt{dcor2d()} and \texttt{splines2d()} \citep{Grimm2016};
-\texttt{norm\_bin()} and \texttt{norm\_kol()} \citep{huber85};
-\texttt{slice\_index()} \citep{Laa:2020wkm}. Most are relatively simply
-defined, for any projection dimension, and implemented because they are
-relatively easy to optimise. A goal is to be able to incorporate more
-complex PP indexes, for example, based on scagnostics (\citet{scag},
-\citet{WW08}).
+\texttt{cmass()} (D. Cook, Buja, and Cabrera 1993); \texttt{lda\_pp()}
+(Lee et al. 2005); \texttt{pda\_pp()} (Lee and Cook 2010);
+\texttt{dcor2d()} and \texttt{splines2d()} (Grimm 2016);
+\texttt{norm\_bin()} and \texttt{norm\_kol()} (Huber 1985);
+\texttt{slice\_index()} (Ursula Laa and Valencia 2022). Most are
+relatively simply defined, for any projection dimension, and implemented
+because they are relatively easy to optimise. A goal is to be able to
+incorporate more complex PP indexes, for example, based on scagnostics
+(L. Wilkinson, Anand, and Grossman (2005), Leland Wilkinson and Wills
+(2008)).
 
 An initial investigation of PP indexes, and the potential for
-scagnostics is described in \citet{laa_using_2020}. To be useful here an
+scagnostics is described in Laa and Cook (2020). To be useful here an
 optimiser needs to be able to handle index functions which are possibly
@@
 behaviour of jellyfish to navigate towards the best solution over
-iterations \citep{chou_novel_2021}.
+iterations (Chou and Truong 2021).
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, breakable, arc=.35mm, colframe=quarto-callout-note-color-frame, colback=white, rightrule=.15mm, coltitle=black, toprule=.15mm, toptitle=1mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacitybacktitle=0.6, bottomtitle=1mm, bottomrule=.15mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, opacitybacktitle=0.6, colback=white, left=2mm, toprule=.15mm, bottomrule=.15mm, opacityback=0, toptitle=1mm, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm, titlerule=0mm, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, leftrule=.75mm]
 
@@
 
-\citet{laa_using_2020} has proposed five criteria for assessing
-projection pursuit indexes (smoothness, squintability, flexibility,
-rotation invariance, and speed). Since not all index properties affect
-the optimisation process, the focus here is on the first two properties,
+Laa and Cook (2020) has proposed five criteria for assessing projection
+pursuit indexes (smoothness, squintability, flexibility, rotation
+invariance, and speed). Since not all index properties affect the
+optimisation process, the focus here is on the first two properties,
 \emph{smoothness} (Section~\ref{sec-smoothness}) and
@@
 
-An classical way to describe the smoothness of a function is to identify
+A classical way to describe the smoothness of a function is to identify
 how many continuous derivatives of the function exist. This can be
-characterized by Sobolev spaces \citep{adams2003sobolev}.
+characterized by Sobolev spaces (Adams and Fournier 2003).
 
@@
 process equipped with a Matérn covariance function, due to the
-connections between this model and Sobolev spaces, see for example
-\citet{porcu2024matern}.
+connections between this model and Sobolev spaces, see for example Porcu
+et al. (2024).
 
@@
 \(\nu\) is a half-integer, the most popular values in the literature
-being \(\frac{1}{2}\), \(\frac{3}{2}\) and \(\frac{5}{2}\)
-\citep{rasmussen2006gaussian}. The parameter \(\nu\), called
-\emph{smoothness parameter}, controls the decay of the covariance
-function. As such, it is an appropriate measure of smoothness of a
-random field. For example, \citet{karvonen2023asymptotic} showed that if
+being \(\frac{1}{2}\), \(\frac{3}{2}\) and \(\frac{5}{2}\) (Rasmussen
+and Williams 2006). The parameter \(\nu\), called \emph{smoothness
+parameter}, controls the decay of the covariance function. As such, it
+is an appropriate measure of smoothness of a random field, as shown by
+the simulations on Figure~\ref{fig-matern-1d} and
+Figure~\ref{fig-matern-2d}. For example, Karvonen (2023) showed that if
 a function \(f\) has a Sobolev index of \(k\), then the smoothness
 parameter estimate \(\nu\) in \eqref{eq:matern} cannot be asymptotically
-less than \(k\). See the survey \citet{porcu2024matern} for additional
+less than \(k\). See the survey Porcu et al. (2024) for additional
 results on the connection between the Matérn model and Sobolev spaces.
@@
 \(K_\infty(u)=\exp(-u^{2}/2)\).
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_1d.png}
+
+}
+
+\caption{\label{fig-matern-1d}Five random simulations from a Gaussian
+Process defined on \(\mathbb{R}\) with zero mean and Matérn-\(\nu\)
+covariance function, with \(\nu=1\) (left), \(\nu=2\) (middle), and
+\(\nu=4\) (right), showing that higher values of \(\nu\) produce
+smoother curves.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_2d.png}
+
+}
+
+\caption{\label{fig-matern-2d}One random simulation from a Gaussian
+Process defined on \(\mathbb{R}^2\) with zero mean and Matérn-\(\nu\)
+covariance function, with \(\nu=1\) (left), \(\nu=2\) (middle), and
+\(\nu=4\) (right), showing that higher values of \(\nu\) produce
+smoother surfaces.}
+
+\end{figure}%
 
@@
 of the jellyfish search optimization. There exist several R packages,
-such as \texttt{GpGp} \citep{guinness2021gpgp} or \texttt{ExaGeoStatR}
-\citep{abdulah2023large}, to fit the hyperparameters of a GP covariance
-function on data, which is usually done by maximum likelihood
-estimation. In this project, the \texttt{GpGp} package is used.
+such as \texttt{GpGp} (Guinness, Katzfuss, and Fahmy 2021) or
+\texttt{ExaGeoStatR} (Abdulah et al. 2023), to fit the hyperparameters
+of a GP covariance function on data, which is usually done by maximum
+likelihood estimation. In this project, the \texttt{GpGp} package is
+used.
 
@@
 \begin{prop}\label{prop:convex-concave}
-Let $g_1(d)$ be a convex function and $g_2(d)$ be a concave function, where $d \in [0, D]$ and $g_1(d), g_2(d) \in [0, 1]$.
-Assume both functions achieve the same maximum gradient, i.e., $\max g'_1(d_1) = \max g'_2(d_2) = -g_{max}$, it follows that $d_1 < d_2$ and $\varsigma(f_1) < \varsigma(f_2)$.
+Let $g_1(d)$ be a convex decreasing function and $g_2(d)$ be a concave decreasing function defined on [0, 1] where $d \in [0, D]$.
+Let $d_1 := \underset{d}{\arg \max} |g_1'(d)|$ and $d_2 := \underset{d}{\arg \max} |g_2'(d)|$.
+
+$$\varsigma(g_1) < \varsigma(g_2)$$
+
+For a convex function, $g_1'(d) < 0 \; \forall d$ and is non-decreasing. For a concave function, $g_2'(d) < 0 \; \forall d$ and is non-increasing. As such, it follows that $d_1 < d_2$.
+
+Suppose both functions achieve the same maximum gradient $-g_{\max}$, i.e., $\max g'_1(d_1) = \max g'_2(d_2) = -g_{\max}$,
+
+$$
+\varsigma(g_1(d_1)) = -c  \max g'_1(d_1)  d_1 = -c  (-g_{\max}) d_1 \leq -c  (-g_{\max}) d_2 = -c \max g'_2(d_2)  d_2 = \varsigma(g_2(d_2))
+$$
+
 \end{prop}
 
-From \citet{barnett1981interpreting} and \citet{laa_using_2020}, a large
-squint angle implies that the objective function value is close to
-optimal even when the perfect view to see the structure is far away. A
-small squint angle means that the PP index value improves substantially
-only when the perfect view is close by. As such, low squintability
-implies rapid improvement in the index value when near the perfect view.
-For PP, a small squint angle is considered to be undesirable because it
-means that the optimiser needs to be very close to be able to ``see''
-the optima. Thus, it could be difficult for the optimiser to find the
-optima. The mathematical formulation of this intuition is proposed
-below:
+From Tukey and Tukey (1981) and Laa and Cook (2020), a large squint
+angle implies that the objective function value is close to optimal even
+when the perfect view to see the structure is far away. A small squint
+angle means that the PP index value improves substantially only when the
+perfect view is close by. As such, low squintability implies rapid
+improvement in the index value when near the perfect view. For PP, a
+small squint angle is considered to be undesirable because it means that
+the optimiser needs to be very close to be able to ``see'' the optima.
+Thus, it could be difficult for the optimiser to find the optima. The
+mathematical formulation of this intuition is proposed below:
 
@@
 \(f(x_{\max})=\theta_4\). Under the parametric model
-(\ref{eq-parametric}), the squintability metric
-(\textbf{?@eq-squintability}) can be shown to be equal to
+(\ref{eq-parametric}), the squintability metric \eqref{eq-squintability}
+can be shown to be equal to
 
@@
 The JSO performance is compared with an existing optimiser, Creeping
-Random Search (CRS) \citep{RJ-2021-105, laa_using_2020} used in the PP
-guided tour to explore JSO's behaviour under different hyper-parameter
-and data dimension combinations. The second simulation studies the
-effect of index properties (smoothness and squintability), along with
-JSO hyper-parameters, and data dimension, on the success rate of the JSO
-performance. This section describes the simulation details, with the
-results deferred to Section Section~\ref{sec-sim-res}.
+Random Search (CRS) (Zhang et al. 2021; Laa and Cook 2020) used in the
+PP guided tour to explore JSO's behaviour under different
+hyper-parameter and data dimension combinations. The second simulation
+studies the effect of index properties (smoothness and squintability),
+along with JSO hyper-parameters, and data dimension, on the success rate
+of the JSO performance. This section describes the simulation details,
+with the results deferred to Section Section~\ref{sec-sim-res}.
 
@@
 Figure~\ref{fig-success-rate} for an illustration). This comparison is
-based on projection pursuit to find the pipe shape investigated by
-\citet{laa_using_2020} using the \texttt{holes} index.
+based on projection pursuit to find the pipe shape investigated by Laa
+and Cook (2020) using the \texttt{holes} index.
 
@@
 CRS. The results of the simulation are collected using the data
-structure proposed in \citet{RJ-2021-105} for assessing JSO, where the
+structure proposed in Zhang et al. (2021) for assessing JSO, where the
 design parameters are stored along with index value, projection basis,
@@
 bin the index values by projection distances to obtain the average index
-value for each bin, 4) fit the scaled sigmoid function in Equation
-\textbf{?@eq-squintability} to the binned index values against
-projection distances using non-linear least square, 5) calculate the
-squintability measure using Equation \ref{eq-squintability-parametric}
-with parameters estimated from the model.}
+value for each bin, 4) fit the scaled sigmoid function in equation
+\eqref{eq-squintability} to the binned index values against projection
+distances using non-linear least square, 5) calculate the squintability
+measure using equation (\ref{eq-squintability-parametric}) with
+parameters estimated from the model.}
 
@@
 
-The article is created using Quarto \citep{Allaire_Quarto_2022} in R
-\citep{R}. The source code for reproducing the work reported in this
+The article is created using Quarto (Allaire et al. 2022) in R (R Core
+Team 2023). The source code for reproducing the work reported in this
 paper can be found at:
@@
 
-\renewcommand\refname{References}
-  \bibliography{bibliography.bib}
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
+\phantomsection\label{refs}
+\begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-abdulah2023large}
+Abdulah, Sameh, Yuxiao Li, Jian Cao, Hatem Ltaief, David Keyes, Marc
+Genton, and Ying Sun. 2023. {``Large-Scale Environmental Data Science
+with {ExaGeoStatR}.''} \emph{Environmetrics} 34 (1): e2770.
+https://doi.org/\url{https://doi.org/10.1002/env.2770}.
+
+\bibitem[\citeproctext]{ref-adams2003sobolev}
+Adams, Robert, and John Fournier. 2003. \emph{Sobolev Spaces}. Vol. 140.
+Pure and Applied Mathematics. Elsevier.
+
+\bibitem[\citeproctext]{ref-Allaire_Quarto_2022}
+Allaire, J. J., Charles Teague, Carlos Scheidegger, Yihui Xie, and
+Christophe Dervieux. 2022. \emph{{Quarto}} (version 1.2).
+\url{https://doi.org/10.5281/zenodo.5960048}.
+
+\bibitem[\citeproctext]{ref-Bertsimas93}
+Bertsimas, Dimitris, and John Tsitsiklis. 1993. {``{Simulated
+Annealing}.''} \emph{Statistical Science} 8 (1): 10--15.
+\url{https://doi.org/10.1214/ss/1177011077}.
+
+\bibitem[\citeproctext]{ref-chou_novel_2021}
+Chou, Jui-Sheng, and Dinh-Nhat Truong. 2021. {``A Novel Metaheuristic
+Optimizer Inspired by Behavior of Jellyfish in Ocean.''} \emph{Applied
+Mathematics and Computation} 389 (January): 125535.
+\url{https://doi.org/10.1016/j.amc.2020.125535}.
+
+\bibitem[\citeproctext]{ref-cook1993projection}
+Cook, D., A. Buja, and J. Cabrera. 1993. {``Projection Pursuit Indexes
+Based on Orthonormal Function Expansions.''} \emph{Journal of
+Computational and Graphical Statistics} 2 (3): 225--50.
+\url{https://doi.org/10.2307/1390644}.
+
+\bibitem[\citeproctext]{ref-cook1995grand}
+Cook, Dianne, Andreas Buja, Javier Cabrera, and Catherine Hurley. 1995.
+{``Grand Tour and Projection Pursuit.''} \emph{Journal of Computational
+and Graphical Statistics} 4 (3): 155--72.
+\url{https://doi.org/10.1080/10618600.1995.10474674}.
+
+\bibitem[\citeproctext]{ref-FT74}
+Friedman, J. H., and J. W. Tukey. 1974. {``{A} {P}rojection {P}ursuit
+{A}lgorithm for {E}xploratory {D}ata {A}nalysis.''} \emph{IEEE
+Transactions on Computing C} 23: 881--89.
+
+\bibitem[\citeproctext]{ref-Grimm2016}
+Grimm, Katrin. 2016. {``Kennzahlenbasierte Grafikauswahl.''} Doctoral
+thesis, Universit{ä}t Augsburg.
+
+\bibitem[\citeproctext]{ref-grochowski2011}
+Grochowski, Marek, and Włodzisław Duch. 2011. {``Fast Projection Pursuit
+Based on Quality of Projected Clusters.''} In \emph{Adaptive and Natural
+Computing Algorithms}, edited by Andrej Dobnikar, Uroš Lotrič, and
+Branko Šter, 89--97. Berlin, Heidelberg: Springer Berlin Heidelberg.
+
+\bibitem[\citeproctext]{ref-guinness2021gpgp}
+Guinness, Joseph, Matthias Katzfuss, and Youssef Fahmy. 2021. {``{GpGp}:
+Fast {G}aussian Process Computation Using {V}ecchia's Approximation.''}
+R package.
+
+\bibitem[\citeproctext]{ref-hall1989polynomial}
+Hall, Peter. 1989. {``On Polynomial-Based Projection Indices for
+Exploratory Projection Pursuit.''} \emph{The Annals of Statistics} 17
+(2): 589--605. \url{https://doi.org/10.1214/aos/1176347127}.
+
+\bibitem[\citeproctext]{ref-huber85}
+Huber, Peter J. 1985. {``Projection Pursuit.''} \emph{Ann. Statist.} 13
+(2): 435--75. \url{https://doi.org/10.1214/aos/1176349519}.
+
+\bibitem[\citeproctext]{ref-karvonen2023asymptotic}
+Karvonen, Toni. 2023. {``Asymptotic Bounds for Smoothness Parameter
+Estimates in {G}aussian Process Interpolation.''} \emph{SIAM/ASA Journal
+on Uncertainty Quantification} 11 (4): 1225--57.
+
+\bibitem[\citeproctext]{ref-kr69}
+Kruskal, J. B. 1969. {``Toward a Practical Method Which Helps Uncover
+the Structure of a Set of Observations by Finding the Line
+Transformation Which Optimizes a New {`Index of Condensation'}.''} In
+\emph{Statistical Computation}, edited by R. C. Milton and J. A. Nelder,
+427--40. New York: Academic Press.
+
+\bibitem[\citeproctext]{ref-laa_using_2020}
+Laa, Ursula, and Dianne Cook. 2020. {``Using Tours to Visually
+Investigate Properties of New Projection Pursuit Indexes with
+Application to Problems in Physics.''} \emph{Computational Statistics}
+35 (3): 1171--1205. \url{https://doi.org/10.1007/s00180-020-00954-8}.
+
+\bibitem[\citeproctext]{ref-lee2010projection}
+Lee, Eun-Kyung, and Dianne Cook. 2010. {``A Projection Pursuit Index for
+Large p Small n Data.''} \emph{Statistics and Computing} 20 (3):
+381--92. \url{https://doi.org/10.1007/s11222-009-9131-1}.
+
+\bibitem[\citeproctext]{ref-lee2005projection}
+Lee, Eun-Kyung, Dianne Cook, Sigbert Klinke, and Thomas Lumley. 2005.
+{``Projection Pursuit for Exploratory Supervised Classification.''}
+\emph{Journal of Computational and Graphical Statistics} 14 (4):
+831--46. \url{https://doi.org/10.1198/106186005X77702}.
+
+\bibitem[\citeproctext]{ref-Loperfido2018}
+Loperfido, Nicola. 2018. {``Skewness-Based Projection Pursuit: A
+Computational Approach.''} \emph{Computational Statistics and Data
+Analysis} 120 (C): 42--57.
+https://doi.org/\url{https://doi.org/10.1016/j.csda.2017.11.001}.
+
+\bibitem[\citeproctext]{ref-Loperfido2020}
+---------. 2020. {``Kurtosis-Based Projection Pursuit for Outlier
+Detection in Financial Time Series.''} \emph{The European Journal of
+Finance} 26 (2-3): 142--64.
+https://doi.org/\url{https://doi.org/10.1080/1351847X.2019.1647864}.
+
+\bibitem[\citeproctext]{ref-marie-sainte2010}
+Marie-Sainte, Souad Larabi, Alain Berro, and Anne Ruiz-Gazen. 2010.
+{``An Efficient Optimization Method for Revealing Local Optima of
+Projection Pursuit Indices.''} In \emph{Swarm Intelligence}, edited by
+Marco Dorigo, Mauro Birattari, Gianni A. Di Caro, René Doursat, Andries
+P Engelbrecht, Dario Floreano, Luca Maria Gambardella, et al., 60--71.
+Berlin, Heidelberg: Springer Berlin Heidelberg.
+
+\bibitem[\citeproctext]{ref-porcu2024matern}
+Porcu, Emilio, Moreno Bevilacqua, Robert Schaback, and Chris Oates.
+2024. {``The {M}atérn Model: A Journey Through Statistics, Numerical
+Analysis and Machine Learning.''} \emph{Statistical Science} 39 (3):
+469--92.
+
+\bibitem[\citeproctext]{ref-posse95}
+Posse, Christian. 1995b. {``Projection Pursuit Exploratory Data
+Analysis.''} \emph{Computational Statistics and Data Analysis} 20 (6):
+669--87.
+https://doi.org/\url{https://doi.org/10.1016/0167-9473(95)00002-8}.
+
+\bibitem[\citeproctext]{ref-posse1995}
+---------. 1995a. {``Projection Pursuit Exploratory Data Analysis.''}
+\emph{Computational Statistics \& Data Analysis} 20 (6): 669--87.
+https://doi.org/\url{https://doi.org/10.1016/0167-9473(95)00002-8}.
+
+\bibitem[\citeproctext]{ref-R}
+R Core Team. 2023. \emph{R: A Language and Environment for Statistical
+Computing}. Vienna, Austria: R Foundation for Statistical Computing.
+\url{https://www.R-project.org/}.
+
+\bibitem[\citeproctext]{ref-rajwar_exhaustive_2023}
+Rajwar, Kanchan, Kusum Deep, and Swagatam Das. 2023. {``An Exhaustive
+Review of the Metaheuristic Algorithms for Search and Optimization:
+Taxonomy, Applications, and Open Challenges.''} \emph{Artificial
+Intelligence Review}, 1--71.
+\url{https://doi.org/10.1007/s10462-023-10470-y}.
+
+\bibitem[\citeproctext]{ref-rasmussen2006gaussian}
+Rasmussen, Carl Edward, and Christopher K. I. Williams. 2006.
+\emph{Gaussian Processes for Machine Learning}. The MIT Press.
+
+\bibitem[\citeproctext]{ref-barnett1981interpreting}
+Tukey, P. A., and J. W. Tukey. 1981. \emph{Graphical Display of Data in
+Three and Higher Dimensions}. Wiley Series in Probability and
+Mathematical Statistics: Applied Probability and Statistics. Wiley.
+\url{https://books.google.com.au/books?id=WBzvAAAAMAAJ}.
+
+\bibitem[\citeproctext]{ref-Laa:2020wkm}
+Ursula Laa, Andreas Buja, Dianne Cook, and German Valencia. 2022.
+{``Hole or Grain? A Section Pursuit Index for Finding Hidden Structure
+in Multiple Dimensions.''} \emph{Journal of Computational and Graphical
+Statistics} 31 (3): 739--52.
+\url{https://doi.org/10.1080/10618600.2022.2035230}.
+
+\bibitem[\citeproctext]{ref-tourr}
+Wickham, H., D. Cook, H. Hofmann, and A. Buja. 2011. {``{tourr}: An {R}
+Package for Exploring Multivariate Data with Projections.''}
+\emph{Journal of Statistical Software} 40 (2): 1--18.
+\url{http://doi.org/10.18637/jss.v040.i02}.
+
+\bibitem[\citeproctext]{ref-scag}
+Wilkinson, L., A. Anand, and R. Grossman. 2005. {``Graph-Theoretic
+Scagnostics.''} In \emph{IEEE Symposium on Information Visualization,
+2005. INFOVIS 2005.}, 157--64.
+\url{https://doi.org/10.1109/INFVIS.2005.1532142}.
+
+\bibitem[\citeproctext]{ref-WW08}
+Wilkinson, Leland, and Graham Wills. 2008. {``Scagnostics
+Distributions.''} \emph{Journal of Computational and Graphical
+Statistics} 17 (2): 473--91.
+\url{https://doi.org/10.1198/106186008X320465}.
+
+\bibitem[\citeproctext]{ref-RJ-2021-105}
+Zhang, H. Sherry, Dianne Cook, Ursula Laa, Nicolas Langrené, and
+Patricia Menéndez. 2021. {``Visual Diagnostics for Constrained
+Optimisation with Application to Guided Tours.''} \emph{The R Journal}
+13: 624--41. \url{https://doi.org/10.32614/RJ-2021-105}.
+
+\end{CSLReferences}
 

### 2024-08-20 jso.qmd
add the discussion section

@@
     keep-tex: true
+    fontsize: 12pt
     linestretch: 2
-    fontsize: 12pt
 editor: 
@@
 their importance relative to JSO hyper-parameters in the optimisation
-success. In defining a projection pursuit problem, factors such as the
-shape-to-find, the index function used, and the data dimension,
-determine the properties such as smoothness and squintability. These
-metrics can then be compared across different problems to understand
-their relative complexites. The results suggest that squintability has a
-more significant influence than smoothness on the success rate of JSO.
-Once the characteristics of the projection pursuit problem are fully
-understood, increasing JSO hyper-parameters can enhance the search
-effectiveness, however, it is also important to consider the resulting
-increase in computational complexity.
+success.
 
@@
 
+# Discussion {#sec-discussion}
+
+The two properties, smoothness and squintability, can be calculated once factors such as the target shape, the chosen index function, and the data dimension are defined in a projection pursuit problem.
+In practice, these metrics can be used to characterise the complexity of the optimisation task and inform the optimiser selection.
+For example, a large smoothness value suggests a smooth index function and both gradient-based and random sampling optimisers tend to perform well.
+A large squintability value indicates a distinct difference in index values between optimal regions and others.
+When one jellyfish enters the optimal region, the high index value it generates will be clearly distinguishable from spurious values due to this pronounced difference and lead other jellyfish to move toward the optimal region.
+In contrast, a smaller squintability value results in a less distinct separation between desirable and undesirable regions, making it more difficult for other jellyfish to accurately navigate toward the target area.
+When comparing between multiple optimisers, the smoothness and squintability metrics proposed in this paper are relative and comparison can only be made with common parameters, i.e.
+the same seed for generating random bases and the same binning width for squintability.
+Furthermore, the range of the index values should be standardised to the [0, 1] interval as demonstrated for the `holes` and `TIC` indexes.
+
+As a population-based metaheuristic method, JSO requires evaluating multiple jellyfish in each iteration, which increased its computational complexity compared to single-candidate random sampling methods.
+When choosing optimisers, this higher computational costs needs to be considered.
+Tuning JSO hyperparameters, such as the number of jellyfish and the maximum number of attempts, can enhance the likelihood of finding the optimal region.
+However, in some cases, increasing these parameters may not be effective if the index surface is complex or has low squintability.
+Moreover, the increased computational cost associated with higher parameter can make JSO impractical.
+
 # Conclusion {#sec-conclusion}

### 2024-08-20 jso.tex
add the discussion section

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, opacitybacktitle=0.6, colback=white, left=2mm, toprule=.15mm, bottomrule=.15mm, opacityback=0, toptitle=1mm, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm, titlerule=0mm, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, leftrule=.75mm]
+\begin{tcolorbox}[enhanced jigsaw, breakable, toptitle=1mm, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, opacityback=0, rightrule=.15mm, bottomrule=.15mm, bottomtitle=1mm, leftrule=.75mm, titlerule=0mm, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, colframe=quarto-callout-note-color-frame, opacitybacktitle=0.6]
 
@@
 and \texttt{dimension} are significant, suggesting their importance
-relative to JSO hyper-parameters in the optimisation success. In
-defining a projection pursuit problem, factors such as the
-shape-to-find, the index function used, and the data dimension,
-determine the properties such as smoothness and squintability. These
-metrics can then be compared across different problems to understand
-their relative complexites. The results suggest that squintability has a
-more significant influence than smoothness on the success rate of JSO.
-Once the characteristics of the projection pursuit problem are fully
-understood, increasing JSO hyper-parameters can enhance the search
-effectiveness, however, it is also important to consider the resulting
-increase in computational complexity.
+relative to JSO hyper-parameters in the optimisation success.
 
@@
 
+\section{Discussion}\label{sec-discussion}
+
+The two properties, smoothness and squintability, can be calculated once
+factors such as the target shape, the chosen index function, and the
+data dimension are defined in a projection pursuit problem. In practice,
+these metrics can be used to characterise the complexity of the
+optimisation task and inform the optimiser selection. For example, a
+large smoothness value suggests a smooth index function and both
+gradient-based and random sampling optimisers tend to perform well. A
+large squintability value indicates a distinct difference in index
+values between optimal regions and others. When one jellyfish enters the
+optimal region, the high index value it generates will be clearly
+distinguishable from spurious values due to this pronounced difference
+and lead other jellyfish to move toward the optimal region. In contrast,
+a smaller squintability value results in a less distinct separation
+between desirable and undesirable regions, making it more difficult for
+other jellyfish to accurately navigate toward the target area. When
+comparing between multiple optimisers, the smoothness and squintability
+metrics proposed in this paper are relative and comparison can only be
+made with common parameters, i.e.~the same seed for generating random
+bases and the same binning width for squintability. Furthermore, the
+range of the index values should be standardised to the {[}0, 1{]}
+interval as demonstrated for the \texttt{holes} and \texttt{TIC}
+indexes.
+
+As a population-based metaheuristic method, JSO requires evaluating
+multiple jellyfish in each iteration, which increased its computational
+complexity compared to single-candidate random sampling methods. When
+choosing optimisers, this higher computational costs needs to be
+considered. Tuning JSO hyperparameters, such as the number of jellyfish
+and the maximum number of attempts, can enhance the likelihood of
+finding the optimal region. However, in some cases, increasing these
+parameters may not be effective if the index surface is complex or has
+low squintability. Moreover, the increased computational cost associated
+with higher parameter can make JSO impractical.
+
 \section{Conclusion}\label{sec-conclusion}

### 2024-08-21 jso.qmd
update on sec6

@@
 \begin{prop}\label{prop:convex-concave}
-Let $g_1(d)$ be a convex decreasing function and $g_2(d)$ be a concave decreasing function defined on [0, 1] where $d \in [0, D]$.
+Let $g_1(d)$ be a convex decreasing function and $g_2(d)$ be a concave decreasing function both defined on [0, 1] where $d \in [0, D]$.
 Let $d_1 := \underset{d}{\arg \max} |g_1'(d)|$ and $d_2 := \underset{d}{\arg \max} |g_2'(d)|$.
@@
 
-The two properties, smoothness and squintability, can be calculated once factors such as the target shape, the chosen index function, and the data dimension are defined in a projection pursuit problem.
-In practice, these metrics can be used to characterise the complexity of the optimisation task and inform the optimiser selection.
-For example, a large smoothness value suggests a smooth index function and both gradient-based and random sampling optimisers tend to perform well.
-A large squintability value indicates a distinct difference in index values between optimal regions and others.
+This section discusses the insights in implementation yielded from the previous sections to guide the usage of projection pursuit guided tour amongst practitioners.
+Smoothness and squintability in a projection pursuit problem can now be measured given the target shape, the chosen index function, and the data dimension.
+In practice, these metrics can be used to characterise the difficulty of the optimisation task and inform the choice of optimiser.
+For instance, a large squintability value indicates a distinct difference in index values between optimal regions and others.
 When one jellyfish enters the optimal region, the high index value it generates will be clearly distinguishable from spurious values due to this pronounced difference and lead other jellyfish to move toward the optimal region.
 In contrast, a smaller squintability value results in a less distinct separation between desirable and undesirable regions, making it more difficult for other jellyfish to accurately navigate toward the target area.
-When comparing between multiple optimisers, the smoothness and squintability metrics proposed in this paper are relative and comparison can only be made with common parameters, i.e.
+In addition, a larger smoothness value implies a smoother index function, which suggests the use of both gradient-based and random sampling optimisers.
+
+This study enables comparison of new indices with existing indices with the proposed smoothness and squintability measures, as well as benchmarking with existing well-known data sets.
+Note that when comparing between multiple optimisers, the smoothness and squintability metrics proposed in this paper are relative.
+Comparison should be made with common parameters, i.e.
 the same seed for generating random bases and the same binning width for squintability.
@@
 As a population-based metaheuristic method, JSO requires evaluating multiple jellyfish in each iteration, which increased its computational complexity compared to single-candidate random sampling methods.
-When choosing optimisers, this higher computational costs needs to be considered.
-Tuning JSO hyperparameters, such as the number of jellyfish and the maximum number of attempts, can enhance the likelihood of finding the optimal region.
+This additional computational costs should be taken into consideration when choosing optimisers.
+Tuning JSO hyperparameters, such as the number of jellyfish and the maximum number of attempts, can enhance the likelihood of finding the optimal region at the expense of computational costs.
 However, in some cases, increasing these parameters may not be effective if the index surface is complex or has low squintability.
-Moreover, the increased computational cost associated with higher parameter can make JSO impractical.
+
+The new JSO is integrated with the current implementation of the projection pursuit guided tour in the `tourr` package, and can be further examined using PP optimisation diagnostics in the `ferrn` package.
+The recommended approach is to conduct the optimisation off-line, extract the bases of a selected jellyfish, and then use the planned tour to follow selected jellyfish.
+These tools are also available in the `tourr` package.
 
@@
 
-The new JSO is integrated with the current implementation of the
-projection pursuit guided tour in the `tourr` package, and can be
-further examined using PP optimisation diagnostics in the `ferrn`
-package. Using the JSO is a little different than the current CRS,
-because it runs many paths. The recommended approach is to conduct the
-optimisation off-line, extract the bases of a selected jellyfish, and
-then use the planned tour to follow selected jellyfish. The tools for
-this are available in the `tourr` package, too.
-
 ```{=html}

### 2024-08-21 jso.tex
update on sec6

@@
 % Make \paragraph and \subparagraph free-standing
-\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}{
-    \@ifstar
-      \xxxParagraphStar
-      \xxxParagraphNoStar
-  }
-  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
-  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
-\makeatother
 
@@
 \makeatother
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
   \fi
   % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 \makeatletter
-\makeatother
-\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatother
-
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
 \ifLuaTeX
@@
 \fi
-\usepackage{bookmark}
-
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \end{keywords}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, breakable, toptitle=1mm, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, opacityback=0, rightrule=.15mm, bottomrule=.15mm, bottomtitle=1mm, leftrule=.75mm, titlerule=0mm, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, colframe=quarto-callout-note-color-frame, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, breakable, rightrule=.15mm, leftrule=.75mm, arc=.35mm, titlerule=0mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, coltitle=black, colframe=quarto-callout-note-color-frame, toptitle=1mm, bottomrule=.15mm, toprule=.15mm, left=2mm]
 
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_1d.png}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_1d.png}
 
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_2d.png}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_2d.png}
 
@@
 \endgroup{}
-
-}
-
-\end{table}%
-
-\section{Discussion}\label{sec-discussion}
-
-The two properties, smoothness and squintability, can be calculated once
-factors such as the target shape, the chosen index function, and the
-data dimension are defined in a projection pursuit problem. In practice,
-these metrics can be used to characterise the complexity of the
-optimisation task and inform the optimiser selection. For example, a
-large smoothness value suggests a smooth index function and both
-gradient-based and random sampling optimisers tend to perform well. A
-large squintability value indicates a distinct difference in index
-values between optimal regions and others. When one jellyfish enters the
-optimal region, the high index value it generates will be clearly
-distinguishable from spurious values due to this pronounced difference
-and lead other jellyfish to move toward the optimal region. In contrast,
-a smaller squintability value results in a less distinct separation
-between desirable and undesirable regions, making it more difficult for
-other jellyfish to accurately navigate toward the target area. When
+\end{table}
+
+\hypertarget{sec-discussion}{%
+\section{Discussion}\label{sec-discussion}}
+
+This section discusses the insights in implementation yielded from the
+previous sections to guide the usage of projection pursuit guided tour
+amongst practitioners. Smoothness and squintability in a projection
+pursuit problem can now be measured given the target shape, the chosen
+index function, and the data dimension. In practice, these metrics can
+be used to characterise the difficulty of the optimisation task and
+inform the choice of optimiser. For instance, a large squintability
+value indicates a distinct difference in index values between optimal
+regions and others. When one jellyfish enters the optimal region, the
+high index value it generates will be clearly distinguishable from
+spurious values due to this pronounced difference and lead other
+jellyfish to move toward the optimal region. In contrast, a smaller
+squintability value results in a less distinct separation between
+desirable and undesirable regions, making it more difficult for other
+jellyfish to accurately navigate toward the target area. In addition, a
+larger smoothness value implies a smoother index function, which
+suggests the use of both gradient-based and random sampling optimisers.
+
+This study enables comparison of new indices with existing indices with
+the proposed smoothness and squintability measures, as well as
+benchmarking with existing well-known data sets. Note that when
 comparing between multiple optimisers, the smoothness and squintability
-metrics proposed in this paper are relative and comparison can only be
-made with common parameters, i.e.~the same seed for generating random
-bases and the same binning width for squintability. Furthermore, the
-range of the index values should be standardised to the {[}0, 1{]}
-interval as demonstrated for the \texttt{holes} and \texttt{TIC}
-indexes.
+metrics proposed in this paper are relative. Comparison should be made
+with common parameters, i.e.~the same seed for generating random bases
+and the same binning width for squintability. Furthermore, the range of
+the index values should be standardised to the {[}0, 1{]} interval as
+demonstrated for the \texttt{holes} and \texttt{TIC} indexes.
 
@@
 multiple jellyfish in each iteration, which increased its computational
-complexity compared to single-candidate random sampling methods. When
-choosing optimisers, this higher computational costs needs to be
-considered. Tuning JSO hyperparameters, such as the number of jellyfish
-and the maximum number of attempts, can enhance the likelihood of
-finding the optimal region. However, in some cases, increasing these
-parameters may not be effective if the index surface is complex or has
-low squintability. Moreover, the increased computational cost associated
-with higher parameter can make JSO impractical.
-
-\section{Conclusion}\label{sec-conclusion}
+complexity compared to single-candidate random sampling methods. This
+additional computational costs should be taken into consideration when
+choosing optimisers. Tuning JSO hyperparameters, such as the number of
+jellyfish and the maximum number of attempts, can enhance the likelihood
+of finding the optimal region at the expense of computational costs.
+However, in some cases, increasing these parameters may not be effective
+if the index surface is complex or has low squintability.
+
+The new JSO is integrated with the current implementation of the
+projection pursuit guided tour in the \texttt{tourr} package, and can be
+further examined using PP optimisation diagnostics in the \texttt{ferrn}
+package. The recommended approach is to conduct the optimisation
+off-line, extract the bases of a selected jellyfish, and then use the
+planned tour to follow selected jellyfish. These tools are also
+available in the \texttt{tourr} package.
+
+\hypertarget{sec-conclusion}{%
+\section{Conclusion}\label{sec-conclusion}}
 
@@
 
-The new JSO is integrated with the current implementation of the
-projection pursuit guided tour in the \texttt{tourr} package, and can be
-further examined using PP optimisation diagnostics in the \texttt{ferrn}
-package. Using the JSO is a little different than the current CRS,
-because it runs many paths. The recommended approach is to conduct the
-optimisation off-line, extract the bases of a selected jellyfish, and
-then use the planned tour to follow selected jellyfish. The tools for
-this are available in the \texttt{tourr} package, too.
-
-\section{Acknowledgement}\label{acknowledgement}
+\hypertarget{acknowledgement}{%
+\section{Acknowledgement}\label{acknowledgement}}
 
@@
 
-\section*{References}\label{references}
+\hypertarget{references}{%
+\section*{References}\label{references}}
 \addcontentsline{toc}{section}{References}
 
-\phantomsection\label{refs}
+\hypertarget{refs}{}
 \begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-abdulah2023large}
+\leavevmode\vadjust pre{\hypertarget{ref-abdulah2023large}{}}%
 Abdulah, Sameh, Yuxiao Li, Jian Cao, Hatem Ltaief, David Keyes, Marc
@@
 
-\bibitem[\citeproctext]{ref-adams2003sobolev}
+\leavevmode\vadjust pre{\hypertarget{ref-adams2003sobolev}{}}%
 Adams, Robert, and John Fournier. 2003. \emph{Sobolev Spaces}. Vol. 140.
@@
 
-\bibitem[\citeproctext]{ref-Allaire_Quarto_2022}
+\leavevmode\vadjust pre{\hypertarget{ref-Allaire_Quarto_2022}{}}%
 Allaire, J. J., Charles Teague, Carlos Scheidegger, Yihui Xie, and
@@
 
-\bibitem[\citeproctext]{ref-Bertsimas93}
+\leavevmode\vadjust pre{\hypertarget{ref-Bertsimas93}{}}%
 Bertsimas, Dimitris, and John Tsitsiklis. 1993. {``{Simulated
@@
 
-\bibitem[\citeproctext]{ref-chou_novel_2021}
+\leavevmode\vadjust pre{\hypertarget{ref-chou_novel_2021}{}}%
 Chou, Jui-Sheng, and Dinh-Nhat Truong. 2021. {``A Novel Metaheuristic
@@
 
-\bibitem[\citeproctext]{ref-cook1993projection}
+\leavevmode\vadjust pre{\hypertarget{ref-cook1993projection}{}}%
 Cook, D., A. Buja, and J. Cabrera. 1993. {``Projection Pursuit Indexes
@@
 
-\bibitem[\citeproctext]{ref-cook1995grand}
+\leavevmode\vadjust pre{\hypertarget{ref-cook1995grand}{}}%
 Cook, Dianne, Andreas Buja, Javier Cabrera, and Catherine Hurley. 1995.
@@
 
-\bibitem[\citeproctext]{ref-FT74}
+\leavevmode\vadjust pre{\hypertarget{ref-FT74}{}}%
 Friedman, J. H., and J. W. Tukey. 1974. {``{A} {P}rojection {P}ursuit
@@
 
-\bibitem[\citeproctext]{ref-Grimm2016}
+\leavevmode\vadjust pre{\hypertarget{ref-Grimm2016}{}}%
 Grimm, Katrin. 2016. {``Kennzahlenbasierte Grafikauswahl.''} Doctoral
@@
 
-\bibitem[\citeproctext]{ref-grochowski2011}
+\leavevmode\vadjust pre{\hypertarget{ref-grochowski2011}{}}%
 Grochowski, Marek, and Włodzisław Duch. 2011. {``Fast Projection Pursuit
@@
 
-\bibitem[\citeproctext]{ref-guinness2021gpgp}
+\leavevmode\vadjust pre{\hypertarget{ref-guinness2021gpgp}{}}%
 Guinness, Joseph, Matthias Katzfuss, and Youssef Fahmy. 2021. {``{GpGp}:
@@
 
-\bibitem[\citeproctext]{ref-hall1989polynomial}
+\leavevmode\vadjust pre{\hypertarget{ref-hall1989polynomial}{}}%
 Hall, Peter. 1989. {``On Polynomial-Based Projection Indices for
@@
 
-\bibitem[\citeproctext]{ref-huber85}
+\leavevmode\vadjust pre{\hypertarget{ref-huber85}{}}%
 Huber, Peter J. 1985. {``Projection Pursuit.''} \emph{Ann. Statist.} 13
@@
 
-\bibitem[\citeproctext]{ref-karvonen2023asymptotic}
+\leavevmode\vadjust pre{\hypertarget{ref-karvonen2023asymptotic}{}}%
 Karvonen, Toni. 2023. {``Asymptotic Bounds for Smoothness Parameter
@@
 
-\bibitem[\citeproctext]{ref-kr69}
+\leavevmode\vadjust pre{\hypertarget{ref-kr69}{}}%
 Kruskal, J. B. 1969. {``Toward a Practical Method Which Helps Uncover
@@
 
-\bibitem[\citeproctext]{ref-laa_using_2020}
+\leavevmode\vadjust pre{\hypertarget{ref-laa_using_2020}{}}%
 Laa, Ursula, and Dianne Cook. 2020. {``Using Tours to Visually
@@
 
-\bibitem[\citeproctext]{ref-lee2010projection}
+\leavevmode\vadjust pre{\hypertarget{ref-lee2010projection}{}}%
 Lee, Eun-Kyung, and Dianne Cook. 2010. {``A Projection Pursuit Index for
@@
 
-\bibitem[\citeproctext]{ref-lee2005projection}
+\leavevmode\vadjust pre{\hypertarget{ref-lee2005projection}{}}%
 Lee, Eun-Kyung, Dianne Cook, Sigbert Klinke, and Thomas Lumley. 2005.
@@
 
-\bibitem[\citeproctext]{ref-Loperfido2018}
+\leavevmode\vadjust pre{\hypertarget{ref-Loperfido2018}{}}%
 Loperfido, Nicola. 2018. {``Skewness-Based Projection Pursuit: A
@@
 
-\bibitem[\citeproctext]{ref-Loperfido2020}
+\leavevmode\vadjust pre{\hypertarget{ref-Loperfido2020}{}}%
 ---------. 2020. {``Kurtosis-Based Projection Pursuit for Outlier
@@
 
-\bibitem[\citeproctext]{ref-marie-sainte2010}
+\leavevmode\vadjust pre{\hypertarget{ref-marie-sainte2010}{}}%
 Marie-Sainte, Souad Larabi, Alain Berro, and Anne Ruiz-Gazen. 2010.
@@
 
-\bibitem[\citeproctext]{ref-porcu2024matern}
+\leavevmode\vadjust pre{\hypertarget{ref-porcu2024matern}{}}%
 Porcu, Emilio, Moreno Bevilacqua, Robert Schaback, and Chris Oates.
@@
 
-\bibitem[\citeproctext]{ref-posse95}
+\leavevmode\vadjust pre{\hypertarget{ref-posse95}{}}%
 Posse, Christian. 1995b. {``Projection Pursuit Exploratory Data
@@
 
-\bibitem[\citeproctext]{ref-posse1995}
+\leavevmode\vadjust pre{\hypertarget{ref-posse1995}{}}%
 ---------. 1995a. {``Projection Pursuit Exploratory Data Analysis.''}
@@
 
-\bibitem[\citeproctext]{ref-R}
+\leavevmode\vadjust pre{\hypertarget{ref-R}{}}%
 R Core Team. 2023. \emph{R: A Language and Environment for Statistical
@@
 
-\bibitem[\citeproctext]{ref-rajwar_exhaustive_2023}
+\leavevmode\vadjust pre{\hypertarget{ref-rajwar_exhaustive_2023}{}}%
 Rajwar, Kanchan, Kusum Deep, and Swagatam Das. 2023. {``An Exhaustive
@@
 
-\bibitem[\citeproctext]{ref-rasmussen2006gaussian}
+\leavevmode\vadjust pre{\hypertarget{ref-rasmussen2006gaussian}{}}%
 Rasmussen, Carl Edward, and Christopher K. I. Williams. 2006.
@@
 
-\bibitem[\citeproctext]{ref-barnett1981interpreting}
+\leavevmode\vadjust pre{\hypertarget{ref-barnett1981interpreting}{}}%
 Tukey, P. A., and J. W. Tukey. 1981. \emph{Graphical Display of Data in
@@
 
-\bibitem[\citeproctext]{ref-Laa:2020wkm}
+\leavevmode\vadjust pre{\hypertarget{ref-Laa:2020wkm}{}}%
 Ursula Laa, Andreas Buja, Dianne Cook, and German Valencia. 2022.
@@
 
-\bibitem[\citeproctext]{ref-tourr}
+\leavevmode\vadjust pre{\hypertarget{ref-tourr}{}}%
 Wickham, H., D. Cook, H. Hofmann, and A. Buja. 2011. {``{tourr}: An {R}
@@
 
-\bibitem[\citeproctext]{ref-scag}
+\leavevmode\vadjust pre{\hypertarget{ref-scag}{}}%
 Wilkinson, L., A. Anand, and R. Grossman. 2005. {``Graph-Theoretic
@@
 
-\bibitem[\citeproctext]{ref-WW08}
+\leavevmode\vadjust pre{\hypertarget{ref-WW08}{}}%
 Wilkinson, Leland, and Graham Wills. 2008. {``Scagnostics
@@
 
-\bibitem[\citeproctext]{ref-RJ-2021-105}
+\leavevmode\vadjust pre{\hypertarget{ref-RJ-2021-105}{}}%
 Zhang, H. Sherry, Dianne Cook, Ursula Laa, Nicolas Langrené, and

### 2024-08-23 jso.tex
adding sec6 to intro paragraph

@@
 \end{keywords}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, frame hidden, sharp corners, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 performance. Section~\ref{sec-sim-res} presents the results.
-Section~\ref{sec-conclusion} summarises the work and provides
-suggestions for future directions.
+Section~\ref{sec-discussion} discusses implementation details and
+insights for practitioners. Section~\ref{sec-conclusion} summarises the
+work and provides suggestions for future directions.
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, breakable, rightrule=.15mm, leftrule=.75mm, arc=.35mm, titlerule=0mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, coltitle=black, colframe=quarto-callout-note-color-frame, toptitle=1mm, bottomrule=.15mm, toprule=.15mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, breakable, colframe=quarto-callout-note-color-frame, arc=.35mm, titlerule=0mm, left=2mm, bottomrule=.15mm, toprule=.15mm, coltitle=black, bottomtitle=1mm, opacitybacktitle=0.6, colback=white, toptitle=1mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white]
 

### 2024-08-23 jso.tex
typo

@@
 \end{keywords}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, frame hidden, sharp corners, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, breakable, colframe=quarto-callout-note-color-frame, arc=.35mm, titlerule=0mm, left=2mm, bottomrule=.15mm, toprule=.15mm, coltitle=black, bottomtitle=1mm, opacitybacktitle=0.6, colback=white, toptitle=1mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white]
+\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, toptitle=1mm, bottomtitle=1mm, coltitle=black, opacityback=0, left=2mm, bottomrule=.15mm, leftrule=.75mm, rightrule=.15mm, opacitybacktitle=0.6, breakable, colback=white, colbacktitle=quarto-callout-note-color!10!white]
 
@@
 of the JSO performance. This section describes the simulation details,
-with the results deferred to Section Section~\ref{sec-sim-res}.
+with the results deferred to Section~\ref{sec-sim-res}.
 

### 2024-08-31 jso.qmd
adding sec6 to intro paragraph
typo
revised abstract

@@
   The projection pursuit (PP) guided tour interactively optimises a criteria function known as the PP index, to explore high-dimensional data by revealing interesting projections.
-The optimisation in PP can be non-trivial, involving non-smooth functions and optima with a small "squint angle", detectable only from close proximity.
+Optimisation of some PP indexes can be non-trivial, if they are non-smooth functions, or the optima has a small "squint angle", detectable only from close proximity.
 To address these challenges, this study investigates the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
 The performance of JSO for visualising data is evaluated across various hyper-parameter settings and compared with existing optimisers.
-Additionally, this work proposes novel methods to quantify two properties of the PP index -- smoothness and squintability -- that capture the complexities inherent in PP optimisation problems.
-These two metrics are evaluated along with JSO hyper-parameters to determine their effects on JSO success rate.
-Our numerical results confirm the positive impact of these metrics on the JSO success rate, with squintability being the most significant.
-The JSO algorithm has been implemented in the `tourr` package and functions to calculate smoothness and squintability are available in the `ferrn` package.
+Additionally, methods for calculating the smoothness and squintability properties of the PP index are proposed.
+There are used to assess the optimiser performance in the presence of PP index complexities.
+A simulation study illustrates the use of these performance metrics to compare the JSO with existing optimisation methods available for the guided tour.
+The JSO algorithm has been implemented in the R package, `tourr`, and functions to calculate smoothness and squintability are available in the `ferrn` package.
 keywords: 
@@
 other studying the impact of different PP index properties on
-optimisation performance. @sec-sim-res presents the results.
+optimisation performance. @sec-sim-res presents the results. @sec-discussion discusses implementation details and insights for practitioners. 
 @sec-conclusion summarises the work and provides suggestions for future
@@
 performance. This section describes the simulation details, with the
-results deferred to Section @sec-sim-res.
+results deferred to @sec-sim-res.
 

### 2024-08-31 jso.tex
revised abstract

@@
 \makeatother
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
   \fi
   % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 \fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 criteria function known as the PP index, to explore high-dimensional
-data by revealing interesting projections. The optimisation in PP can be
-non-trivial, involving non-smooth functions and optima with a small
-``squint angle'', detectable only from close proximity. To address these
-challenges, this study investigates the performance of a recently
-introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for
-optimising PP indexes. The performance of JSO for visualising data is
-evaluated across various hyper-parameter settings and compared with
-existing optimisers. Additionally, this work proposes novel methods to
-quantify two properties of the PP index --~smoothness and
-squintability~-- that capture the complexities inherent in PP
-optimisation problems. These two metrics are evaluated along with JSO
-hyper-parameters to determine their effects on JSO success rate. Our
-numerical results confirm the positive impact of these metrics on the
-JSO success rate, with squintability being the most significant. The JSO
-algorithm has been implemented in the \texttt{tourr} package and
-functions to calculate smoothness and squintability are available in the
+data by revealing interesting projections. Optimisation of some PP
+indexes can be non-trivial, if they are non-smooth functions, or the
+optima has a small ``squint angle'', detectable only from close
+proximity. To address these challenges, this study investigates the
+performance of a recently introduced swarm-based algorithm, Jellyfish
+Search Optimiser (JSO), for optimising PP indexes. The performance of
+JSO for visualising data is evaluated across various hyper-parameter
+settings and compared with existing optimisers. Additionally, methods
+for calculating the smoothness and squintability properties of the PP
+index are proposed. There are used to assess the optimiser performance
+in the presence of PP index complexities. A simulation study illustrates
+the use of these performance metrics to compare the JSO with existing
+optimisation methods available for the guided tour. The JSO algorithm
+has been implemented in the R package, \texttt{tourr}, and functions to
+calculate smoothness and squintability are available in the
 \texttt{ferrn} package.
@@
 \end{keywords}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-optimization}{}{
+\begin{equation}\phantomsection\label{eq-optimization}{
 \underset{A \in \mathcal{A}}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I_d
-}\label{eq-optimization}\end{equation}
+}\end{equation}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, toptitle=1mm, bottomtitle=1mm, coltitle=black, opacityback=0, left=2mm, bottomrule=.15mm, leftrule=.75mm, rightrule=.15mm, opacitybacktitle=0.6, breakable, colback=white, colbacktitle=quarto-callout-note-color!10!white]
+\begin{tcolorbox}[enhanced jigsaw, coltitle=black, breakable, arc=.35mm, opacityback=0, toprule=.15mm, opacitybacktitle=0.6, left=2mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, colback=white, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, bottomtitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, bottomrule=.15mm]
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_1d.png}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_1d.png}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_2d.png}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_2d.png}
 
@@
 \endgroup{}
-\end{table}
-
-\hypertarget{sec-discussion}{%
-\section{Discussion}\label{sec-discussion}}
+
+}
+
+\end{table}%
+
+\section{Discussion}\label{sec-discussion}
 
@@
 
-\hypertarget{sec-conclusion}{%
-\section{Conclusion}\label{sec-conclusion}}
+\section{Conclusion}\label{sec-conclusion}
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
 
-\hypertarget{refs}{}
+\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-abdulah2023large}{}}%
+\bibitem[\citeproctext]{ref-abdulah2023large}
 Abdulah, Sameh, Yuxiao Li, Jian Cao, Hatem Ltaief, David Keyes, Marc
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-adams2003sobolev}{}}%
+\bibitem[\citeproctext]{ref-adams2003sobolev}
 Adams, Robert, and John Fournier. 2003. \emph{Sobolev Spaces}. Vol. 140.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Allaire_Quarto_2022}{}}%
+\bibitem[\citeproctext]{ref-Allaire_Quarto_2022}
 Allaire, J. J., Charles Teague, Carlos Scheidegger, Yihui Xie, and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Bertsimas93}{}}%
+\bibitem[\citeproctext]{ref-Bertsimas93}
 Bertsimas, Dimitris, and John Tsitsiklis. 1993. {``{Simulated
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-chou_novel_2021}{}}%
+\bibitem[\citeproctext]{ref-chou_novel_2021}
 Chou, Jui-Sheng, and Dinh-Nhat Truong. 2021. {``A Novel Metaheuristic
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cook1993projection}{}}%
+\bibitem[\citeproctext]{ref-cook1993projection}
 Cook, D., A. Buja, and J. Cabrera. 1993. {``Projection Pursuit Indexes
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cook1995grand}{}}%
+\bibitem[\citeproctext]{ref-cook1995grand}
 Cook, Dianne, Andreas Buja, Javier Cabrera, and Catherine Hurley. 1995.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-FT74}{}}%
+\bibitem[\citeproctext]{ref-FT74}
 Friedman, J. H., and J. W. Tukey. 1974. {``{A} {P}rojection {P}ursuit
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Grimm2016}{}}%
+\bibitem[\citeproctext]{ref-Grimm2016}
 Grimm, Katrin. 2016. {``Kennzahlenbasierte Grafikauswahl.''} Doctoral
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-grochowski2011}{}}%
+\bibitem[\citeproctext]{ref-grochowski2011}
 Grochowski, Marek, and Włodzisław Duch. 2011. {``Fast Projection Pursuit
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-guinness2021gpgp}{}}%
+\bibitem[\citeproctext]{ref-guinness2021gpgp}
 Guinness, Joseph, Matthias Katzfuss, and Youssef Fahmy. 2021. {``{GpGp}:
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hall1989polynomial}{}}%
+\bibitem[\citeproctext]{ref-hall1989polynomial}
 Hall, Peter. 1989. {``On Polynomial-Based Projection Indices for
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-huber85}{}}%
+\bibitem[\citeproctext]{ref-huber85}
 Huber, Peter J. 1985. {``Projection Pursuit.''} \emph{Ann. Statist.} 13
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-karvonen2023asymptotic}{}}%
+\bibitem[\citeproctext]{ref-karvonen2023asymptotic}
 Karvonen, Toni. 2023. {``Asymptotic Bounds for Smoothness Parameter
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kr69}{}}%
+\bibitem[\citeproctext]{ref-kr69}
 Kruskal, J. B. 1969. {``Toward a Practical Method Which Helps Uncover
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-laa_using_2020}{}}%
+\bibitem[\citeproctext]{ref-laa_using_2020}
 Laa, Ursula, and Dianne Cook. 2020. {``Using Tours to Visually
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lee2010projection}{}}%
+\bibitem[\citeproctext]{ref-lee2010projection}
 Lee, Eun-Kyung, and Dianne Cook. 2010. {``A Projection Pursuit Index for
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lee2005projection}{}}%
+\bibitem[\citeproctext]{ref-lee2005projection}
 Lee, Eun-Kyung, Dianne Cook, Sigbert Klinke, and Thomas Lumley. 2005.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Loperfido2018}{}}%
+\bibitem[\citeproctext]{ref-Loperfido2018}
 Loperfido, Nicola. 2018. {``Skewness-Based Projection Pursuit: A
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Loperfido2020}{}}%
+\bibitem[\citeproctext]{ref-Loperfido2020}
 ---------. 2020. {``Kurtosis-Based Projection Pursuit for Outlier
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-marie-sainte2010}{}}%
+\bibitem[\citeproctext]{ref-marie-sainte2010}
 Marie-Sainte, Souad Larabi, Alain Berro, and Anne Ruiz-Gazen. 2010.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-porcu2024matern}{}}%
+\bibitem[\citeproctext]{ref-porcu2024matern}
 Porcu, Emilio, Moreno Bevilacqua, Robert Schaback, and Chris Oates.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-posse95}{}}%
+\bibitem[\citeproctext]{ref-posse95}
 Posse, Christian. 1995b. {``Projection Pursuit Exploratory Data
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-posse1995}{}}%
+\bibitem[\citeproctext]{ref-posse1995}
 ---------. 1995a. {``Projection Pursuit Exploratory Data Analysis.''}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-R}{}}%
+\bibitem[\citeproctext]{ref-R}
 R Core Team. 2023. \emph{R: A Language and Environment for Statistical
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-rajwar_exhaustive_2023}{}}%
+\bibitem[\citeproctext]{ref-rajwar_exhaustive_2023}
 Rajwar, Kanchan, Kusum Deep, and Swagatam Das. 2023. {``An Exhaustive
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-rasmussen2006gaussian}{}}%
+\bibitem[\citeproctext]{ref-rasmussen2006gaussian}
 Rasmussen, Carl Edward, and Christopher K. I. Williams. 2006.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-barnett1981interpreting}{}}%
+\bibitem[\citeproctext]{ref-barnett1981interpreting}
 Tukey, P. A., and J. W. Tukey. 1981. \emph{Graphical Display of Data in
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Laa:2020wkm}{}}%
+\bibitem[\citeproctext]{ref-Laa:2020wkm}
 Ursula Laa, Andreas Buja, Dianne Cook, and German Valencia. 2022.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-tourr}{}}%
+\bibitem[\citeproctext]{ref-tourr}
 Wickham, H., D. Cook, H. Hofmann, and A. Buja. 2011. {``{tourr}: An {R}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-scag}{}}%
+\bibitem[\citeproctext]{ref-scag}
 Wilkinson, L., A. Anand, and R. Grossman. 2005. {``Graph-Theoretic
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-WW08}{}}%
+\bibitem[\citeproctext]{ref-WW08}
 Wilkinson, Leland, and Graham Wills. 2008. {``Scagnostics
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-RJ-2021-105}{}}%
+\bibitem[\citeproctext]{ref-RJ-2021-105}
 Zhang, H. Sherry, Dianne Cook, Ursula Laa, Nicolas Langrené, and

### 2024-09-01 jso.qmd
revised the introduction

@@
 title: Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit
-header-includes:
-   - \usepackage{algorithm}
-   - \usepackage{amsmath}
-   - \usepackage{amsthm}
-   - \usepackage{amssymb}
-   - \theoremstyle{plain}
-   - \newtheorem{assumption}{\protect\assumptionname}
-   - \newtheorem{claim}{\protect\claimname}
-   - \newtheorem{condition}{\protect\conditionname}
-   - \newtheorem{conjecture}{\protect\conjecturename}
-   - \newtheorem{cor}{\protect\corollaryname}
-   - \newtheorem{defn}{\protect\definitionname}
-   - \newtheorem{example}{\protect\examplename}
-   - \newtheorem{lem}{\protect\lemmaname}
-   - \newtheorem{notation}{\protect\notationname}
-   - \newtheorem{problem}{\protect\problemname}
-   - \newtheorem{prop}{\protect\propositionname}
-   - \newtheorem{rem}{\protect\remarkname}
-   - \newtheorem{thm}{\protect\theoremname}
-   - \providecommand{\assumptionname}{Assumption}
-   - \providecommand{\claimname}{Claim}
-   - \providecommand{\conditionname}{Condition}
-   - \providecommand{\conjecturename}{Conjecture}
-   - \providecommand{\corollaryname}{Corollary}
-   - \providecommand{\definitionname}{Definition}
-   - \providecommand{\examplename}{Example}
-   - \providecommand{\lemmaname}{Lemma}
-   - \providecommand{\notationname}{Notation}
-   - \providecommand{\problemname}{Problem}
-   - \providecommand{\propositionname}{Proposition}
-   - \providecommand{\remarkname}{Remark}
-   - \providecommand{\theoremname}{Theorem}
 author:
@@
   chunk_output_type: console
+header-includes:
+   - \usepackage{algorithm}
+   - \usepackage{amsmath}
+   - \usepackage{amsthm}
+   - \usepackage{amssymb}
+   - \theoremstyle{plain}
+   - \newtheorem{assumption}{\protect\assumptionname}
+   - \newtheorem{claim}{\protect\claimname}
+   - \newtheorem{condition}{\protect\conditionname}
+   - \newtheorem{conjecture}{\protect\conjecturename}
+   - \newtheorem{cor}{\protect\corollaryname}
+   - \newtheorem{defn}{\protect\definitionname}
+   - \newtheorem{example}{\protect\examplename}
+   - \newtheorem{lem}{\protect\lemmaname}
+   - \newtheorem{notation}{\protect\notationname}
+   - \newtheorem{problem}{\protect\problemname}
+   - \newtheorem{prop}{\protect\propositionname}
+   - \newtheorem{rem}{\protect\remarkname}
+   - \newtheorem{thm}{\protect\theoremname}
+   - \providecommand{\assumptionname}{Assumption}
+   - \providecommand{\claimname}{Claim}
+   - \providecommand{\conditionname}{Condition}
+   - \providecommand{\conjecturename}{Conjecture}
+   - \providecommand{\corollaryname}{Corollary}
+   - \providecommand{\definitionname}{Definition}
+   - \providecommand{\examplename}{Example}
+   - \providecommand{\lemmaname}{Lemma}
+   - \providecommand{\notationname}{Notation}
+   - \providecommand{\problemname}{Problem}
+   - \providecommand{\propositionname}{Proposition}
+   - \providecommand{\remarkname}{Remark}
+   - \providecommand{\theoremname}{Theorem}
 ---
@@
 library(kableExtra)
+library(tourr)
+library(PPtreeViz)
 library(ferrn)
+library(colorspace)
 load(here::here("data/sim_pipe_run_best.rda"))
@@
 
+<!--  
 ```{=html}
-<!--  
 \begin{assumption}\label{assum:my_assumption}
@@
 Theorem \ref{thm:my_theorem} is important.
+```
 -->
-```
-Projection Pursuit (PP) (@kr69, @FT74) is a dimension reduction
-technique aimed at identifying informative linear projections of data.
-The method involves optimising an objective function known as the PP
-index (e.g. @hall1989polynomial, @cook1993projection,
-@lee2010projection, @Loperfido2018, @Loperfido2020), which defines the
-criteria for what constitutes interesting or informative projections.
+
+Projection Pursuit (PP) (@kr69, @FT74, @huber85) is a dimension reduction technique aimed at identifying informative linear projections of data.
+This is useful for exploring high-dimensional data, and creating plots of the data for publication that reveal the main features.
+The method involves optimising an objective function known as the PP index (e.g.
+@hall1989polynomial, @cook1993projection,
+@lee2010projection, @Loperfido2018, @Loperfido2020), which defines the criteria for what constitutes interesting or informative projections.
 Let $X \in \mathbb{R}^{n\times p}$ be the data matrix, $A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix, where $A$ belongs to the Stiefel manifold $\mathcal{A} = V_d(\mathbb{R}^p)$.
-The projection $Y = XA$ is a linear transformation that maps data from a $p$-dimensional space into a $d$-dimensional space.
+(It is interesting to note that when using PP visually, one cares less about $A$ than the plane described by $A$, because the orientation in the plane is irrelevant.
+The the space of planes belongs to a Grassmann manifold.) The projection $Y = XA$ is a linear transformation that maps data from a $p$-dimensional space into a $d$-dimensional space.
 The index function $f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ is a scalar function that measures an interesting aspect of the projected data, such as deviation from normality, presence of clusters, non-linear structure, or other features of interest.
@@
 
-The index functions can be non-linear and non-convex, hence an effective
-and efficient optimisation procedure is essential to explore the data
-landscape and achieve a globally optimal viewpoint of the data.
-
-Optimisation of PP is typically investigated in the literature when new
-indexes are proposed [@posse1995; @marie-sainte2010; @grochowski2011] or
-when visualisation method are used to track the optimisation process.
-@cook1995grand introduced the PP guided tour, which monitors the
-optimisation visually to see the data projections leading in and out of
-the optima. An implementation is available in the `tourr` package
-[@tourr] in R [@R]. @RJ-2021-105 illustrated how to diagnose
-optimisation processes, particularly focusing on the guided tour, and
-revealed a need for improved optimisation. While improving the quality
-of the optimisation solutions in the tour is essential, it is also
-important to be able to view the data projections as the optimisation
-progresses. Integrating the guided tour with a global optimisation
-algorithm that is efficient in finding the global optimal and enables
-viewing of the projected data during the exploration process is a goal.
+The index functions vary in structure, partly depending on the data that is being projected.
+@fig-example-functions shows two examples.
+Huber plots [@huberplot] of 2D data sets are in (a) and (c).
+The PP index values for all 1D projections of the 2D data are shown in polar coordinates, revealing the nature of the function.
+The dashed circle is a baseline set at the average value, and the straight line marks the optimal projection.
+Plots (b) and (d) show the respective best projections of the data as histograms.
+Indexes like the holes, central mass and skewness [@cook1993projection] are generally smooth for most data sets.
+Many indexes noisy and non-convex, requiring an effective and efficient optimisation procedure to explore the data landscape and achieve a globally optimal viewpoint of the data.
+The skewness index computed for trimodal data, in (a), is smooth with a large squint angle but has three modes, and thus is not convex.
+The binned normality index computed on the famous RANDU data, in (c), is noisier and has a very small squint angle.
+The discreteness cannot be seen unless the optimiser is very close to the optimal projection.
+
+```{r}
+#| label: huber-plot-function
+#| echo: false
+huber_plot <- function (origdata2D, origclass=NULL, 
+  PPmethod = "LDA", weight = TRUE, r = 1, lambda = 0.5, 
+  opt.proj = TRUE, UserDefFtn = NULL, ...) 
+{
+    index <- NULL
+    best.proj <- NULL
+    best.index <- 0
+    origdata2D <- as.matrix(origdata2D)
+    for (i in 0:360) {
+        theta <- pi/180 * i
+        proj.data <- matrix(cos(theta) * origdata2D[, 1] + sin(theta) * 
+            origdata2D[, 2])
+        proj <- matrix(c(cos(theta), sin(theta)), ncol = 1)
+        if (PPmethod == "LDA") {
+            newindex <- PPtreeViz::LDAindex(origclass, origdata2D, proj = proj, 
+                weight = weight)
+        }
+        else if (PPmethod == "PDA") {
+            newindex <- PPtreeViz::PDAindex(origclass, origdata2D, proj, 
+                weight = weight, lambda = lambda)
+        }
+        else if (PPmethod == "Lr") {
+            newindex <- PPtreeViz::Lrindex(origclass, origdata2D, proj, 
+                weight = weight, r = r)
+        }
+        else if (PPmethod == "GINI") {
+            newindex <- PPtreeViz::GINIindex1D(origclass, origdata2D, proj)
+        }
+        else if (PPmethod == "ENTROPY") {
+            newindex <- PPtreeViz::ENTROPYindex1D(origclass, origdata2D, 
+                proj)
+        }
+        else if (PPmethod == "holes") {
+          newindex <- tourr::holes()(proj.data)
+        }
+        else if (PPmethod == "skewness") {
+          newindex <- tourr::skewness()(proj.data)
+        }
+        else if (PPmethod == "norm_bin") {
+          newindex <- tourr::norm_bin(nrow(proj.data))(proj.data)
+        }
+        else if (PPmethod == "UserDef") {
+            newindex <- UserDefFtn(proj.data, ...)
+        }
+        index <- c(index, newindex)
+    }
+    sel.index <- which(index[1:360] > signif(max(index), 6) - 
+        1e-06)
+    theta.best.all <- pi/180 * (sel.index - 1)
+    theta.best <- theta.best.all[1]
+    proj.data.best <- matrix(cos(theta.best) * origdata2D[, 1] + 
+        sin(theta.best) * origdata2D[, 2])
+    index.best <- max(index)
+    range <- round(max(index) - min(index), 5)
+    if (range == 0) {
+      PPindex <- rep(4, length(index))
+    }
+    else {
+      PPindex <- (index - min(index))/range * 2 + 3
+    }
+    data.circle <- NULL
+    data.index <- NULL
+    for (i in 1:361) {
+        theta <- pi/180 * (i - 1)
+        data.index <- rbind(data.index, c(PPindex[i] * cos(theta), 
+            PPindex[i] * sin(theta)))
+        data.circle <- rbind(data.circle, c(4 * cos(theta), 4 * 
+            sin(theta)))
+    }
+    maxdiff <- max(c(diff(range(origdata2D[, 1])), diff(range(origdata2D[, 
+        2]))))
+    orig.scaled <- apply(origdata2D, 2, function(x) (x - mean(x))/maxdiff * 
+        3.5)
+    data.cX <- data.circle[, 1]
+    data.cY <- data.circle[, 2]
+    data.X <- data.index[, 1]
+    data.Y <- data.index[, 2]
+    plot.data <- data.frame(data.cX, data.cY, data.X, data.Y)
+    x <- orig.scaled[, 1]
+    y <- orig.scaled[, 2]
+    if (is.null(origclass)) 
+      group <- rep(1, nrow(origdata2D))
+    else
+      group <- origclass
+    point.data <- data.frame(x, y, group)
+    min.X <- min(unlist(plot.data))
+    max.X <- max(unlist(plot.data))
+    if (is.null(origclass)) {
+      P1 <- ggplot(data = plot.data, aes(x = data.X, y = data.Y)) + 
+        geom_path() + geom_path(aes(x = data.cX, y = data.cY), 
+        linetype = "dashed") + geom_point(data = point.data, 
+        aes(x = x, y = y)) + scale_x_continuous(breaks = NULL) + 
+        scale_y_continuous(breaks = NULL) + xlab("") + ylab("") + 
+        coord_fixed() + theme_bw() + theme(panel.border = element_blank())
+    } else {
+      P1 <- ggplot(data = plot.data, aes(x = data.X, y = data.Y)) + 
+        geom_path() + geom_path(aes(x = data.cX, y = data.cY), 
+        linetype = "dashed") + geom_point(data = point.data, 
+        aes(x = x, y = y, color = group, shape = group)) +
+        scale_x_continuous(breaks = NULL) + 
+        scale_y_continuous(breaks = NULL) + xlab("") + ylab("") + 
+        coord_fixed() + theme_bw() + theme(panel.border = element_blank())
+    }
+    if (opt.proj) {
+        P1 <- P1 + geom_abline(intercept = 0, slope = sin(theta.best)/cos(theta.best))
+        if (length(theta.best.all) > 1) 
+            for (i in 2:length(theta.best.all)) P1 <- P1 + geom_abline(intercept = 0, 
+                slope = sin(theta.best.all[i])/cos(theta.best.all[i]), 
+                linetype = "dashed")
+    }
+    best.proj.data <- proj.data.best
+    #group <- origclass
+    if (is.null(origclass)) {
+      hist.data <- data.frame(best.proj.data)
+      P2 <- ggplot(data = hist.data, aes(x = best.proj.data)) + 
+        geom_histogram()
+    } else {
+      hist.data <- data.frame(best.proj.data, group)
+      P2 <- ggplot(data = hist.data, aes(x = best.proj.data, 
+                                         group = group)) + 
+        geom_histogram(aes(fill = group), position = "stack")
+    }
+    #gridExtra::grid.arrange(P1, P2, nrow = 1)
+    return(list(plot=P1, proj_data=best.proj.data))
+}
+
+```
+
+```{r}
+#| label: fig-example-functions
+#| echo: false
+#| fig-width: 9
+#| fig-height: 6
+#| out-width: 80%
+#| fig-cap: "Examples of PP indexes with large (top row) and small (bottom row) squint angles, shown with a Huber plot, and histogram of the projected data corresponding to the optimal projection.
+Huber plot shows PP index values for all 1D projections in polar coordinates."
+data(randu)
+randu_std <- as.data.frame(apply(randu, 2, function(x) (x-mean(x))/sd(x)))
+randu_std$yz <- sqrt(35)/6*randu_std$y-randu_std$z/6
+
+#proj <- animate_xy(flea[,1:6], guided_tour(lda_pp(flea$species)))
+best_proj <- matrix(c(-0.44157005,  0.80959831,
+                      0.14622089, -0.08228193,
+                      -0.08454314, -0.23234287,
+                      0.70367852,  0.28185021,
+                      0.27876825,  0.44819691,
+                      0.45123453,  0.05896646), ncol=2, byrow=T)
+flea2D <- as.matrix(flea[,1:6]) %*% best_proj
+colnames(flea2D) <- c("X1", "X2")
+flea2D <- as.data.frame(flea2D)
+
+# Make both plots
+large_sqa <- huber_plot(flea2D, PPmethod="skewness")
+small_sqa <- huber_plot(randu_std[c(1,4)], PPmethod="norm_bin")
+
+sqa1 <- large_sqa$plot + ggtitle("(a) Skewness index")
+lsqa_d <- tibble(d = large_sqa$proj_data[,1])
+sqa2 <- ggplot(lsqa_d, aes(x=d)) + 
+  geom_histogram(binwidth=0.25, colour="white") +
+  xlab("") + ylab("") +
+  ggtitle("(b) Optimally projected data") +
+  theme_bw() +
+  theme(axis.text.y = element_blank())
+sqa3 <- small_sqa$plot + ggtitle("(c) Binned normality index")
+ssqa_d <- tibble(d = small_sqa$proj_data[,1])
+sqa4 = ggplot(ssqa_d, aes(x=d)) + 
+  geom_histogram(breaks = seq(-2.2, 2.4, 0.12)) +
+  xlab("") + ylab("") +
+  ggtitle("(d) Optimally projected data") +
+  theme_bw() +
+  theme(axis.text.y = element_blank())
+
+sqa1 + sqa2 + sqa3 + sqa4 + plot_layout(ncol=2, widths=c(2,2))
+```
+
+Optimisation of PP is often discussed when new indexes are proposed [@posse1995; @marie-sainte2010; @grochowski2011].
+@cook1995grand tied the optimisation more closely to the index, when they introduced the PP guided tour, which monitors the optimisation visually so that the user can see the projected data leading in and out of the optima.
+An implementation is available in the `tourr` package [@tourr] in R [@R].
+@RJ-2021-105 illustrated how to diagnose optimisation processes, particularly focusing on the guided tour, and revealed a need for improved optimisation.
+While improving the quality of the optimisation solutions in the tour is essential, it is also important to be able to view the data projections as the optimisation progresses.
+Integrating the guided tour with a global optimisation algorithm that is efficient in finding the global optimal and enables viewing of the projected data during the exploration process is a goal.
 

### 2024-09-01 jso.tex
revised the introduction

@@
 
-Projection Pursuit (PP) (Kruskal (1969), Friedman and Tukey (1974)) is a
-dimension reduction technique aimed at identifying informative linear
-projections of data. The method involves optimising an objective
-function known as the PP index (e.g. Hall (1989), D. Cook, Buja, and
-Cabrera (1993), Lee and Cook (2010), Loperfido (2018), Loperfido
+Projection Pursuit (PP) (Kruskal (1969), Friedman and Tukey (1974),
+Huber (1985)) is a dimension reduction technique aimed at identifying
+informative linear projections of data. This is useful for exploring
+high-dimensional data, and creating plots of the data for publication
+that reveal the main features. The method involves optimising an
+objective function known as the PP index (e.g. Hall (1989), Cook, Buja,
+and Cabrera (1993), Lee and Cook (2010), Loperfido (2018), Loperfido
 (2020)), which defines the criteria for what constitutes interesting or
@@
 where \(A\) belongs to the Stiefel manifold
-\(\mathcal{A} = V_d(\mathbb{R}^p)\). The projection \(Y = XA\) is a
-linear transformation that maps data from a \(p\)-dimensional space into
-a \(d\)-dimensional space. The index function
-\(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar function
-that measures an interesting aspect of the projected data, such as
-deviation from normality, presence of clusters, non-linear structure, or
-other features of interest. For a fixed data, PP finds the orthonormal
-basis \(A\) that maximises the index value of the projection,
-\(Y = XA\):
+\(\mathcal{A} = V_d(\mathbb{R}^p)\). (It is interesting to note that
+when using PP visually, one cares less about \(A\) than the plane
+described by \(A\), because the orientation in the plane is irrelevant.
+The the space of planes belongs to a Grassmann manifold.) The projection
+\(Y = XA\) is a linear transformation that maps data from a
+\(p\)-dimensional space into a \(d\)-dimensional space. The index
+function \(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar
+function that measures an interesting aspect of the projected data, such
+as deviation from normality, presence of clusters, non-linear structure,
+or other features of interest. For a fixed data, PP finds the
+orthonormal basis \(A\) that maximises the index value of the
+projection, \(Y = XA\):
 
@@
 
-The index functions can be non-linear and non-convex, hence an effective
-and efficient optimisation procedure is essential to explore the data
-landscape and achieve a globally optimal viewpoint of the data.
-
-Optimisation of PP is typically investigated in the literature when new
-indexes are proposed (Posse 1995a; Marie-Sainte, Berro, and Ruiz-Gazen
-2010; Grochowski and Duch 2011) or when visualisation method are used to
-track the optimisation process. Dianne Cook et al. (1995) introduced the
-PP guided tour, which monitors the optimisation visually to see the data
-projections leading in and out of the optima. An implementation is
-available in the \texttt{tourr} package (Wickham et al. 2011) in R (R
-Core Team 2023). Zhang et al. (2021) illustrated how to diagnose
-optimisation processes, particularly focusing on the guided tour, and
-revealed a need for improved optimisation. While improving the quality
-of the optimisation solutions in the tour is essential, it is also
-important to be able to view the data projections as the optimisation
-progresses. Integrating the guided tour with a global optimisation
-algorithm that is efficient in finding the global optimal and enables
-viewing of the projected data during the exploration process is a goal.
+The index functions vary in structure, partly depending on the data that
+is being projected. Figure~\ref{fig-example-functions} shows two
+examples. Huber plots (Huber 1990) of 2D data sets are in (a) and (c).
+The PP index values for all 1D projections of the 2D data are shown in
+polar coordinates, revealing the nature of the function. The dashed
+circle is a baseline set at the average value, and the straight line
+marks the optimal projection. Plots (b) and (d) show the respective best
+projections of the data as histograms. Indexes like the holes, central
+mass and skewness (Cook, Buja, and Cabrera 1993) are generally smooth
+for most data sets. Many indexes noisy and non-convex, requiring an
+effective and efficient optimisation procedure to explore the data
+landscape and achieve a globally optimal viewpoint of the data. The
+skewness index computed for trimodal data, in (a), is smooth with a
+large squint angle but has three modes, and thus is not convex. The
+binned normality index computed on the famous RANDU data, in (c), is
+noisier and has a very small squint angle. The discreteness cannot be
+seen unless the optimiser is very close to the optimal projection.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=0.8\textwidth,height=\textheight]{jso_files/figure-pdf/fig-example-functions-1.pdf}
+
+}
+
+\caption{\label{fig-example-functions}Examples of PP indexes with large
+(top row) and small (bottom row) squint angles, shown with a Huber plot,
+and histogram of the projected data corresponding to the optimal
+projection. Huber plot shows PP index values for all 1D projections in
+polar coordinates.}
+
+\end{figure}%
+
+Optimisation of PP is often discussed when new indexes are proposed
+(Posse 1995a; Marie-Sainte, Berro, and Ruiz-Gazen 2010; Grochowski and
+Duch 2011). Cook et al. (1995) tied the optimisation more closely to the
+index, when they introduced the PP guided tour, which monitors the
+optimisation visually so that the user can see the projected data
+leading in and out of the optima. An implementation is available in the
+\texttt{tourr} package (Wickham et al. 2011) in R (R Core Team 2023).
+Zhang et al. (2021) illustrated how to diagnose optimisation processes,
+particularly focusing on the guided tour, and revealed a need for
+improved optimisation. While improving the quality of the optimisation
+solutions in the tour is essential, it is also important to be able to
+view the data projections as the optimisation progresses. Integrating
+the guided tour with a global optimisation algorithm that is efficient
+in finding the global optimal and enables viewing of the projected data
+during the exploration process is a goal.
 
@@
 There are several PP index functions available: \texttt{holes()} and
-\texttt{cmass()} (D. Cook, Buja, and Cabrera 1993); \texttt{lda\_pp()}
-(Lee et al. 2005); \texttt{pda\_pp()} (Lee and Cook 2010);
-\texttt{dcor2d()} and \texttt{splines2d()} (Grimm 2016);
-\texttt{norm\_bin()} and \texttt{norm\_kol()} (Huber 1985);
-\texttt{slice\_index()} (Ursula Laa and Valencia 2022). Most are
-relatively simply defined, for any projection dimension, and implemented
-because they are relatively easy to optimise. A goal is to be able to
-incorporate more complex PP indexes, for example, based on scagnostics
-(L. Wilkinson, Anand, and Grossman (2005), Leland Wilkinson and Wills
-(2008)).
+\texttt{cmass()} (Cook, Buja, and Cabrera 1993); \texttt{lda\_pp()} (Lee
+et al. 2005); \texttt{pda\_pp()} (Lee and Cook 2010); \texttt{dcor2d()}
+and \texttt{splines2d()} (Grimm 2016); \texttt{norm\_bin()} and
+\texttt{norm\_kol()} (Huber 1985); \texttt{slice\_index()} (Ursula Laa
+and Valencia 2022). Most are relatively simply defined, for any
+projection dimension, and implemented because they are relatively easy
+to optimise. A goal is to be able to incorporate more complex PP
+indexes, for example, based on scagnostics (L. Wilkinson, Anand, and
+Grossman (2005), Leland Wilkinson and Wills (2008)).
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, coltitle=black, breakable, arc=.35mm, opacityback=0, toprule=.15mm, opacitybacktitle=0.6, left=2mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, colback=white, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, bottomtitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, bottomrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, toptitle=1mm, rightrule=.15mm, colback=white, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, arc=.35mm, bottomrule=.15mm, toprule=.15mm, titlerule=0mm, bottomtitle=1mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm, breakable, opacitybacktitle=0.6]
 
@@
 \bibitem[\citeproctext]{ref-cook1995grand}
-Cook, Dianne, Andreas Buja, Javier Cabrera, and Catherine Hurley. 1995.
-{``Grand Tour and Projection Pursuit.''} \emph{Journal of Computational
-and Graphical Statistics} 4 (3): 155--72.
+Cook, D., A. Buja, J. Cabrera, and C. Hurley. 1995. {``Grand Tour and
+Projection Pursuit.''} \emph{Journal of Computational and Graphical
+Statistics} 4 (3): 155--72.
 \url{https://doi.org/10.1080/10618600.1995.10474674}.
@@
 (2): 435--75. \url{https://doi.org/10.1214/aos/1176349519}.
+
+\bibitem[\citeproctext]{ref-huberplot}
+---------. 1990. {``Data Analysis and Projection Pursuit.''} Technical
+Report PJH-90-1. Dept. of Mathematics, Massachusetts Institute of
+Technology.
 

### 2024-09-02 jso.qmd
done revising introduction

@@
 
-Projection Pursuit (PP) (@kr69, @FT74, @huber85) is a dimension reduction technique aimed at identifying informative linear projections of data.
-This is useful for exploring high-dimensional data, and creating plots of the data for publication that reveal the main features.
+Projection pursuit (PP) (@kr69, @FT74, @huber85) is a dimension reduction technique aimed at identifying informative linear projections of data.
+This is useful for exploring high-dimensional data, and creating plots of the data that reveal the main features to use for publication.
 The method involves optimising an objective function known as the PP index (e.g.
@@
 Let $X \in \mathbb{R}^{n\times p}$ be the data matrix, $A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix, where $A$ belongs to the Stiefel manifold $\mathcal{A} = V_d(\mathbb{R}^p)$.
-(It is interesting to note that when using PP visually, one cares less about $A$ than the plane described by $A$, because the orientation in the plane is irrelevant.
-The the space of planes belongs to a Grassmann manifold.) The projection $Y = XA$ is a linear transformation that maps data from a $p$-dimensional space into a $d$-dimensional space.
+The projection $Y = XA$ is a linear transformation that maps data from a $p$-dimensional space into a $d$-dimensional space.
 The index function $f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ is a scalar function that measures an interesting aspect of the projected data, such as deviation from normality, presence of clusters, non-linear structure, or other features of interest.
-For a fixed data, PP finds the orthonormal basis $A$ that maximises the index value of the projection, $Y = XA$:
+For a fixed sample of data, PP finds the orthonormal basis $A$ that maximises the index value of the projection, $Y = XA$:
 
@@
 
-The index functions vary in structure, partly depending on the data that is being projected.
+It is interesting to note that when using PP visually, one cares less about $A$ than the plane described by $A$, because the orientation in the plane is irrelevant.
+The space of planes belongs to a Grassmann manifold.
+This is usually how the projection pursuit guided tour (PPGT) [@cook1995grand] operates, when using geodesic interpolation between starting and target planes.
+It interpolates plane to plane, removing irrelevant within plane spin, and is agnostic to the basis ($A$) used to define the plane.
+Thus, indexes which are used for the PPGT should be rotationally invariant.
+
+Index functions are quite varied in form, partially depending on the data that is being projected.
 @fig-example-functions shows two examples.
-Huber plots [@huberplot] of 2D data sets are in (a) and (c).
-The PP index values for all 1D projections of the 2D data are shown in polar coordinates, revealing the nature of the function.
+Huber plots [@huberplot] of 2D data sets are in (a) and (c), showing the PP index values for all 1D projections of the 2D data in polar coordinates, which reveals the form of these functions.
 The dashed circle is a baseline set at the average value, and the straight line marks the optimal projection.
 Plots (b) and (d) show the respective best projections of the data as histograms.
-Indexes like the holes, central mass and skewness [@cook1993projection] are generally smooth for most data sets.
+Indexes like the holes, central mass and skewness [@cook1993projection] are generally smooth for most data sets, but capture only large patterns.
 Many indexes noisy and non-convex, requiring an effective and efficient optimisation procedure to explore the data landscape and achieve a globally optimal viewpoint of the data.
 The skewness index computed for trimodal data, in (a), is smooth with a large squint angle but has three modes, and thus is not convex.
-The binned normality index computed on the famous RANDU data, in (c), is noisier and has a very small squint angle.
+The binned normality index (a simple version of a non-normality index as described in @huber85) computed on the famous RANDU data, in (c), is noisier and has a very small squint angle.
 The discreteness cannot be seen unless the optimiser is very close to the optimal projection.
@@
 #| fig-cap: "Examples of PP indexes with large (top row) and small (bottom row) squint angles, shown with a Huber plot, and histogram of the projected data corresponding to the optimal projection.
-Huber plot shows PP index values for all 1D projections in polar coordinates."
+A Huber plot shows the PP index values for all 1D data projections in polar coordinates."
 data(randu)
@@
 Optimisation of PP is often discussed when new indexes are proposed [@posse1995; @marie-sainte2010; @grochowski2011].
-@cook1995grand tied the optimisation more closely to the index, when they introduced the PP guided tour, which monitors the optimisation visually so that the user can see the projected data leading in and out of the optima.
+@cook1995grand tied the optimisation more closely to the index, when they introduced the PPGT, which monitors the optimisation visually so that the user can see the projected data leading in and out of the optima.
 An implementation is available in the `tourr` package [@tourr] in R [@R].
@@
 
-Here, the potential for a Jellyfish Search Optimiser (JSO) to be
-integrated with the projection pursuit guided tour is explored. JSO,
-[@chou_novel_2021] - [@rajwar_exhaustive_2023], inspired by the search
-behaviour of jellyfish in the ocean, is a swarm-based metaheuristic
-designed to solve global optimisation problems. Compared to traditional
-methods, JSO has demonstrated stronger search ability and faster
-convergence, and requires fewer tuning parameters. These practical
-advantages make JSO a promising candidate for enhancing PP optimisation.
+Here, the potential for a Jellyfish Search Optimiser (JSO) (see @chou_novel_2021, and @rajwar_exhaustive_2023) for the PPGT is explored.
+JSO, inspired by the search behaviour of jellyfish in the ocean, is a swarm-based metaheuristic designed to solve global optimisation problems.
+Compared to traditional methods, JSO has demonstrated stronger search ability and faster
+convergence, and requires fewer tuning parameters. These practicalities
+make JSO a promising candidate for enhancing PP optimisation.
 <!--The study reported here explores JSO's potential in enhancing projection pursuit and examines the characteristics
@@
 including squintability (based on @barnett1981interpreting's squint
-angle) and smoothness as introduced in @laa_using_2020. Here we
-mathematically define metrics for squintability and smoothness, which is
-new for the field. A series of simulation experiments using various
-datasets and PP indexes are conducted to assess JSO's behaviour and its
-sensitivity to hyper-parameter choices (number of jellyfish and maximum
-number of tries). The relationship between the JSO performance,
-hyper-parameter choices and properties of PP indexes (smoothness and
-squintability) is analysed to provide guidance on selecting optimisers
-for practitioners using projection pursuit. Additionally, it aims to
-guide the design of new indexes that facilitate easy optimization for PP
-researchers.
+angle) and smoothness.
+Here, we mathematically define metrics for squintability and smoothness, which is a new contribution for PP research.
+A series of simulation experiments using various datasets and PP indexes are conducted to assess JSO's behaviour and its sensitivity to hyper-parameter choices (number of jellyfish and maximum number of tries).
+The relationship between the JSO performance, hyper-parameter choices and properties of PP indexes (smoothness and squintability) is analysed to provide guidance on selecting optimisers for practitioners using projection pursuit.
+Additionally, this work should guide the design of new PP indexes and facilitate better optimization for PP.
 
 The paper is structured as follows. @sec-background introduces the
-background of PP guided tour and reviews existing optimisers and index
-functions in the literature. @sec-theory details JSO and introduces
+background of the PPGT, reviews existing optimisers and index
+functions in the literature. @sec-theory describes JSO and introduces the 
 metrics that measure different properties of PP indexes, smoothness and
-squintability. @sec-sim-deets details two simulation experiments to
-assess JSO's performance: one comparing JSO's performance improvements
-relative to an existing optimiser, Creeping Random Search (CRS), and the
-other studying the impact of different PP index properties on
-optimisation performance. @sec-sim-res presents the results. @sec-discussion discusses implementation details and insights for practitioners. 
+squintability.
+@sec-sim-deets summarises the results of two simulation experiments done to assess JSO's performance: one comparing JSO's performance improvements relative to an existing optimiser, Creeping Random Search (CRS), and the other studying the impact of different PP index properties on optimisation performance.
+@sec-sim-res presents the results.
+@sec-discussion discusses implementation details and insights for practitioners.
 @sec-conclusion summarises the work and provides suggestions for future

### 2024-09-02 jso.tex
done revising introduction

@@
 
-Projection Pursuit (PP) (Kruskal (1969), Friedman and Tukey (1974),
+Projection pursuit (PP) (Kruskal (1969), Friedman and Tukey (1974),
 Huber (1985)) is a dimension reduction technique aimed at identifying
 informative linear projections of data. This is useful for exploring
-high-dimensional data, and creating plots of the data for publication
-that reveal the main features. The method involves optimising an
+high-dimensional data, and creating plots of the data that reveal the
+main features to use for publication. The method involves optimising an
 objective function known as the PP index (e.g. Hall (1989), Cook, Buja,
@@
 where \(A\) belongs to the Stiefel manifold
-\(\mathcal{A} = V_d(\mathbb{R}^p)\). (It is interesting to note that
-when using PP visually, one cares less about \(A\) than the plane
-described by \(A\), because the orientation in the plane is irrelevant.
-The the space of planes belongs to a Grassmann manifold.) The projection
-\(Y = XA\) is a linear transformation that maps data from a
-\(p\)-dimensional space into a \(d\)-dimensional space. The index
-function \(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar
-function that measures an interesting aspect of the projected data, such
-as deviation from normality, presence of clusters, non-linear structure,
-or other features of interest. For a fixed data, PP finds the
+\(\mathcal{A} = V_d(\mathbb{R}^p)\). The projection \(Y = XA\) is a
+linear transformation that maps data from a \(p\)-dimensional space into
+a \(d\)-dimensional space. The index function
+\(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar function
+that measures an interesting aspect of the projected data, such as
+deviation from normality, presence of clusters, non-linear structure, or
+other features of interest. For a fixed sample of data, PP finds the
 orthonormal basis \(A\) that maximises the index value of the
@@
 
-The index functions vary in structure, partly depending on the data that
-is being projected. Figure~\ref{fig-example-functions} shows two
-examples. Huber plots (Huber 1990) of 2D data sets are in (a) and (c).
-The PP index values for all 1D projections of the 2D data are shown in
-polar coordinates, revealing the nature of the function. The dashed
-circle is a baseline set at the average value, and the straight line
-marks the optimal projection. Plots (b) and (d) show the respective best
-projections of the data as histograms. Indexes like the holes, central
-mass and skewness (Cook, Buja, and Cabrera 1993) are generally smooth
-for most data sets. Many indexes noisy and non-convex, requiring an
-effective and efficient optimisation procedure to explore the data
-landscape and achieve a globally optimal viewpoint of the data. The
-skewness index computed for trimodal data, in (a), is smooth with a
-large squint angle but has three modes, and thus is not convex. The
-binned normality index computed on the famous RANDU data, in (c), is
-noisier and has a very small squint angle. The discreteness cannot be
-seen unless the optimiser is very close to the optimal projection.
+It is interesting to note that when using PP visually, one cares less
+about \(A\) than the plane described by \(A\), because the orientation
+in the plane is irrelevant. The space of planes belongs to a Grassmann
+manifold. This is usually how the projection pursuit guided tour (PPGT)
+(Cook et al. 1995) operates, when using geodesic interpolation between
+starting and target planes. It interpolates plane to plane, removing
+irrelevant within plane spin, and is agnostic to the basis (\(A\)) used
+to define the plane. Thus, indexes which are used for the PPGT should be
+rotationally invariant.
+
+Index functions are quite varied in form, partially depending on the
+data that is being projected. Figure~\ref{fig-example-functions} shows
+two examples. Huber plots (Huber 1990) of 2D data sets are in (a) and
+(c), showing the PP index values for all 1D projections of the 2D data
+in polar coordinates, which reveals the form of these functions. The
+dashed circle is a baseline set at the average value, and the straight
+line marks the optimal projection. Plots (b) and (d) show the respective
+best projections of the data as histograms. Indexes like the holes,
+central mass and skewness (Cook, Buja, and Cabrera 1993) are generally
+smooth for most data sets, but capture only large patterns. Many indexes
+noisy and non-convex, requiring an effective and efficient optimisation
+procedure to explore the data landscape and achieve a globally optimal
+viewpoint of the data. The skewness index computed for trimodal data, in
+(a), is smooth with a large squint angle but has three modes, and thus
+is not convex. The binned normality index (a simple version of a
+non-normality index as described in Huber (1985)) computed on the famous
+RANDU data, in (c), is noisier and has a very small squint angle. The
+discreteness cannot be seen unless the optimiser is very close to the
+optimal projection.
 
@@
 and histogram of the projected data corresponding to the optimal
-projection. Huber plot shows PP index values for all 1D projections in
-polar coordinates.}
+projection. A Huber plot shows the PP index values for all 1D data
+projections in polar coordinates.}
 
@@
 Duch 2011). Cook et al. (1995) tied the optimisation more closely to the
-index, when they introduced the PP guided tour, which monitors the
-optimisation visually so that the user can see the projected data
-leading in and out of the optima. An implementation is available in the
-\texttt{tourr} package (Wickham et al. 2011) in R (R Core Team 2023).
-Zhang et al. (2021) illustrated how to diagnose optimisation processes,
-particularly focusing on the guided tour, and revealed a need for
-improved optimisation. While improving the quality of the optimisation
-solutions in the tour is essential, it is also important to be able to
-view the data projections as the optimisation progresses. Integrating
-the guided tour with a global optimisation algorithm that is efficient
-in finding the global optimal and enables viewing of the projected data
-during the exploration process is a goal.
-
-Here, the potential for a Jellyfish Search Optimiser (JSO) to be
-integrated with the projection pursuit guided tour is explored. JSO,
-(Chou and Truong 2021) - (Rajwar, Deep, and Das 2023), inspired by the
-search behaviour of jellyfish in the ocean, is a swarm-based
-metaheuristic designed to solve global optimisation problems. Compared
-to traditional methods, JSO has demonstrated stronger search ability and
-faster convergence, and requires fewer tuning parameters. These
-practical advantages make JSO a promising candidate for enhancing PP
-optimisation.
+index, when they introduced the PPGT, which monitors the optimisation
+visually so that the user can see the projected data leading in and out
+of the optima. An implementation is available in the \texttt{tourr}
+package (Wickham et al. 2011) in R (R Core Team 2023). Zhang et al.
+(2021) illustrated how to diagnose optimisation processes, particularly
+focusing on the guided tour, and revealed a need for improved
+optimisation. While improving the quality of the optimisation solutions
+in the tour is essential, it is also important to be able to view the
+data projections as the optimisation progresses. Integrating the guided
+tour with a global optimisation algorithm that is efficient in finding
+the global optimal and enables viewing of the projected data during the
+exploration process is a goal.
+
+Here, the potential for a Jellyfish Search Optimiser (JSO) (see Chou and
+Truong (2021), and Rajwar, Deep, and Das (2023)) for the PPGT is
+explored. JSO, inspired by the search behaviour of jellyfish in the
+ocean, is a swarm-based metaheuristic designed to solve global
+optimisation problems. Compared to traditional methods, JSO has
+demonstrated stronger search ability and faster convergence, and
+requires fewer tuning parameters. These practicalities make JSO a
+promising candidate for enhancing PP optimisation.
 
@@
 including squintability (based on Tukey and Tukey (1981)'s squint angle)
-and smoothness as introduced in Laa and Cook (2020). Here we
-mathematically define metrics for squintability and smoothness, which is
-new for the field. A series of simulation experiments using various
-datasets and PP indexes are conducted to assess JSO's behaviour and its
-sensitivity to hyper-parameter choices (number of jellyfish and maximum
-number of tries). The relationship between the JSO performance,
-hyper-parameter choices and properties of PP indexes (smoothness and
-squintability) is analysed to provide guidance on selecting optimisers
-for practitioners using projection pursuit. Additionally, it aims to
-guide the design of new indexes that facilitate easy optimization for PP
-researchers.
+and smoothness. Here, we mathematically define metrics for squintability
+and smoothness, which is a new contribution for PP research. A series of
+simulation experiments using various datasets and PP indexes are
+conducted to assess JSO's behaviour and its sensitivity to
+hyper-parameter choices (number of jellyfish and maximum number of
+tries). The relationship between the JSO performance, hyper-parameter
+choices and properties of PP indexes (smoothness and squintability) is
+analysed to provide guidance on selecting optimisers for practitioners
+using projection pursuit. Additionally, this work should guide the
+design of new PP indexes and facilitate better optimization for PP.
 
 The paper is structured as follows. Section~\ref{sec-background}
-introduces the background of PP guided tour and reviews existing
-optimisers and index functions in the literature.
-Section~\ref{sec-theory} details JSO and introduces metrics that measure
-different properties of PP indexes, smoothness and squintability.
-Section~\ref{sec-sim-deets} details two simulation experiments to assess
+introduces the background of the PPGT, reviews existing optimisers and
+index functions in the literature. Section~\ref{sec-theory} describes
+JSO and introduces the metrics that measure different properties of PP
+indexes, smoothness and squintability. Section~\ref{sec-sim-deets}
+summarises the results of two simulation experiments done to assess
 JSO's performance: one comparing JSO's performance improvements relative
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, toptitle=1mm, rightrule=.15mm, colback=white, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, arc=.35mm, bottomrule=.15mm, toprule=.15mm, titlerule=0mm, bottomtitle=1mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm, breakable, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, left=2mm, rightrule=.15mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, opacitybacktitle=0.6, colback=white, colframe=quarto-callout-note-color-frame, toprule=.15mm, toptitle=1mm, arc=.35mm, leftrule=.75mm, bottomtitle=1mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black]
 

### 2024-09-04 jso.qmd
Update jso.qmd
working on Practicalities section

@@
 Additionally, methods for calculating the smoothness and squintability properties of the PP index are proposed.
-There are used to assess the optimiser performance in the presence of PP index complexities.
+They are used to assess the optimiser performance in the presence of PP index complexities.
 A simulation study illustrates the use of these performance metrics to compare the JSO with existing optimisation methods available for the guided tour.
@@
 
-# Simulation details {#sec-sim-deets}
+# Simulation study {#sec-sim-deets}
 
@@
 
-# Discussion {#sec-discussion}
-
+# Practicalities {#sec-discussion}
+
+## Computing the metrics for your new index
+
+## Using the JSO in a PPGT
+
+```{r}
+#| eval: false
+#| echo: false
+jso_optima <- save_history(flea[,1:6], 
+                           guided_tour(holes()), 
+                           search_f = search_jellyfish)
+```
+<!--
 This section discusses the insights in implementation yielded from the previous sections to guide the usage of projection pursuit guided tour amongst practitioners.
@@
 These tools are also available in the `tourr` package.
+-->
 
@@
 ```
+
 # Acknowledgement

### 2024-09-04 jso.tex
working on Practicalities section

@@
 for calculating the smoothness and squintability properties of the PP
-index are proposed. There are used to assess the optimiser performance
-in the presence of PP index complexities. A simulation study illustrates
+index are proposed. They are used to assess the optimiser performance in
+the presence of PP index complexities. A simulation study illustrates
 the use of these performance metrics to compare the JSO with existing
@@
 
-\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, left=2mm, rightrule=.15mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, opacitybacktitle=0.6, colback=white, colframe=quarto-callout-note-color-frame, toprule=.15mm, toptitle=1mm, arc=.35mm, leftrule=.75mm, bottomtitle=1mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, coltitle=black, bottomrule=.15mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, leftrule=.75mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, left=2mm, opacitybacktitle=0.6, opacityback=0, bottomtitle=1mm, arc=.35mm, colback=white]
 
@@
 
-\section{Simulation details}\label{sec-sim-deets}
+\section{Simulation study}\label{sec-sim-deets}
 
@@
 
-\section{Discussion}\label{sec-discussion}
-
-This section discusses the insights in implementation yielded from the
-previous sections to guide the usage of projection pursuit guided tour
-amongst practitioners. Smoothness and squintability in a projection
-pursuit problem can now be measured given the target shape, the chosen
-index function, and the data dimension. In practice, these metrics can
-be used to characterise the difficulty of the optimisation task and
-inform the choice of optimiser. For instance, a large squintability
-value indicates a distinct difference in index values between optimal
-regions and others. When one jellyfish enters the optimal region, the
-high index value it generates will be clearly distinguishable from
-spurious values due to this pronounced difference and lead other
-jellyfish to move toward the optimal region. In contrast, a smaller
-squintability value results in a less distinct separation between
-desirable and undesirable regions, making it more difficult for other
-jellyfish to accurately navigate toward the target area. In addition, a
-larger smoothness value implies a smoother index function, which
-suggests the use of both gradient-based and random sampling optimisers.
-
-This study enables comparison of new indices with existing indices with
-the proposed smoothness and squintability measures, as well as
-benchmarking with existing well-known data sets. Note that when
-comparing between multiple optimisers, the smoothness and squintability
-metrics proposed in this paper are relative. Comparison should be made
-with common parameters, i.e.~the same seed for generating random bases
-and the same binning width for squintability. Furthermore, the range of
-the index values should be standardised to the {[}0, 1{]} interval as
-demonstrated for the \texttt{holes} and \texttt{TIC} indexes.
-
-As a population-based metaheuristic method, JSO requires evaluating
-multiple jellyfish in each iteration, which increased its computational
-complexity compared to single-candidate random sampling methods. This
-additional computational costs should be taken into consideration when
-choosing optimisers. Tuning JSO hyperparameters, such as the number of
-jellyfish and the maximum number of attempts, can enhance the likelihood
-of finding the optimal region at the expense of computational costs.
-However, in some cases, increasing these parameters may not be effective
-if the index surface is complex or has low squintability.
-
-The new JSO is integrated with the current implementation of the
-projection pursuit guided tour in the \texttt{tourr} package, and can be
-further examined using PP optimisation diagnostics in the \texttt{ferrn}
-package. The recommended approach is to conduct the optimisation
-off-line, extract the bases of a selected jellyfish, and then use the
-planned tour to follow selected jellyfish. These tools are also
-available in the \texttt{tourr} package.
+\section{Practicalities}\label{sec-discussion}
+
+\subsection{Computing the metrics for your new
+index}\label{computing-the-metrics-for-your-new-index}
+
+\subsection{Using the JSO in a PPGT}\label{using-the-jso-in-a-ppgt}
 

### 2024-09-04 jso.qmd
add practical guide for computing smoothness and squintability

@@
 
+The `ferrn` package provides functionality for computing the smoothness and squintability metrics.
+Both cases require sampling random bases using the `sample_bases()` function, followed by  calculating the specific metric with either `calc_smoothness()` or `calc_squintability()`.
+
+### Smoothness
+
+To sample bases for calculating smoothness, the required arguments include `idx` (the index function used), `data` (the dataset), `n_basis` (the number of random bases to sample), and `best` (the best basis to compute projection distance).
+The following code samples 300 bases for using the `holes` index to find the pipe shape in the 6D `sine1000` dataset:
+
+```{r}
+#| echo: true
+library(GpGp)
+library(fields)
+library(tourr)
+library(ferrn)
+basis_smoothness <- sample_bases(
+  idx = "holes", data = sine1000, n_basis = 300, 
+  best = matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1), nrow = 6))
+basis_smoothness |> head(3)
+```
+
+The output of `sample_bases()` is a data frame with four columns: `id` (the basis ID), `basis` (the basis matrix sampled), `dist` (the projection distance to the best basis), and `index` (the index value).
+To speed up the computation of index values,  parallellasation is available through the `parallel = TRUE` argument.
+
+The `calc_smoothness()` function takes the output from `sample_bases()` and fits a Gaussian process model to the index values against the projection distances to compute the smoothness metric.
+The starting parameter for the fit can be specified using the `start_params` argument, and additional Gaussian process arguments can be provided through `other_gp_params`.
+The following code calculates the smoothness metric from the previously sampled bases:
+
+```{r}
+#| echo: true
+(sm_res <- calc_smoothness(basis_smoothness))
+```
+
+The smoothness metric is the smoothness parameter from the Gussian process and all details of the fit can be accessed using `attr(sm_res, "fit_res")`.
+
+### Squintability
+
+Sampling bases for calculating squintability includes an additional step of interpolating between the sampled bases and the best basis.
+This is specified through setting the `step_size` and `min_proj_dist` arguments in the `sample_bases()` function to non-NA numerical values.
+Since the projection distance typically ranges from 0 to 2, it is recommended to set `step_size` to 0.1 or lower, and `min_proj_dist` to be at least 0.5 to ensure a meaningful interpolation length.
+
+The `calc_squintability()` function computes squintability using two methods: 1) parametrically, by fitting a scaled sigmoid function through non-linear least square (`method = "nls"`), and 2) non-parametrically, using kernel smoothing (`method = "ks"`).
+A `bin_width` argument is required to average the index values over the projection distance before the fit.
+The following code samples 100 bases for calculating squintability in the same scenario:
+
+```{r}
+#| echo: true
+basis_squint <- sample_bases(
+   idx = "holes", data = sine1000, n_basis = 100, 
+  best = matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1), nrow = 6),
+  step_size = 0.1, min_proj_dist = 1.5)
+calc_squintability(basis_squint, method = "nls", bin_width = 0.01)
+```
+
+The output provides the estimated parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$) and the squintability metric for the parametric case.
+For the non-parametric case, it shows the (`max_d`), the corresponding projection distance (`max_dist`), and teh squintability metric as their products.
+
 ## Using the JSO in a PPGT
@@
 ```
+
 <!--

### 2024-09-04 jso.tex
add practical guide for computing smoothness and squintability

@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
+
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{241,243,245}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
+\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
 
@@
 \makeatother
+
 \ifLuaTeX
@@
 
-\begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, coltitle=black, bottomrule=.15mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, leftrule=.75mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, left=2mm, opacitybacktitle=0.6, opacityback=0, bottomtitle=1mm, arc=.35mm, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, coltitle=black, rightrule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, breakable, opacitybacktitle=0.6, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, opacityback=0, bottomtitle=1mm, titlerule=0mm, toptitle=1mm, bottomrule=.15mm, toprule=.15mm]
 
@@
 
+The \texttt{ferrn} package provides functionality for computing the
+smoothness and squintability metrics. Both cases require sampling random
+bases using the \texttt{sample\_bases()} function, followed by
+calculating the specific metric with either \texttt{calc\_smoothness()}
+or \texttt{calc\_squintability()}.
+
+\subsubsection{Smoothness}\label{smoothness}
+
+To sample bases for calculating smoothness, the required arguments
+include \texttt{idx} (the index function used), \texttt{data} (the
+dataset), \texttt{n\_basis} (the number of random bases to sample), and
+\texttt{best} (the best basis to compute projection distance). The
+following code samples 300 bases for using the \texttt{holes} index to
+find the pipe shape in the 6D \texttt{sine1000} dataset:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(GpGp)}
+\FunctionTok{library}\NormalTok{(fields)}
+\FunctionTok{library}\NormalTok{(tourr)}
+\FunctionTok{library}\NormalTok{(ferrn)}
+\NormalTok{basis\_smoothness }\OtherTok{\textless{}{-}} \FunctionTok{sample\_bases}\NormalTok{(}
+  \AttributeTok{idx =} \StringTok{"holes"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sine1000, }\AttributeTok{n\_basis =} \DecValTok{300}\NormalTok{, }
+  \AttributeTok{best =} \FunctionTok{matrix}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{nrow =} \DecValTok{6}\NormalTok{))}
+\NormalTok{basis\_smoothness }\SpecialCharTok{|\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{3}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# PP index:                holes
+# No. of bases:            3
+# interpolation step size: NA
+# Min. proj. dist.:        NA
+     id basis          dist index
+  <int> <list>        <dbl> <dbl>
+1     1 <dbl [6 x 2]>  1.15 0.676
+2     2 <dbl [6 x 2]>  1.77 0.790
+3     3 <dbl [6 x 2]>  1.79 0.778
+\end{verbatim}
+
+The output of \texttt{sample\_bases()} is a data frame with four
+columns: \texttt{id} (the basis ID), \texttt{basis} (the basis matrix
+sampled), \texttt{dist} (the projection distance to the best basis), and
+\texttt{index} (the index value). To speed up the computation of index
+values, parallellasation is available through the
+\texttt{parallel\ =\ TRUE} argument.
+
+The \texttt{calc\_smoothness()} function takes the output from
+\texttt{sample\_bases()} and fits a Gaussian process model to the index
+values against the projection distances to compute the smoothness
+metric. The starting parameter for the fit can be specified using the
+\texttt{start\_params} argument, and additional Gaussian process
+arguments can be provided through \texttt{other\_gp\_params}. The
+following code calculates the smoothness metric from the previously
+sampled bases:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{(sm\_res }\OtherTok{\textless{}{-}} \FunctionTok{calc\_smoothness}\NormalTok{(basis\_smoothness))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# PP index:     holes
+# No. of bases: 300 [6 x 2]
+    variance range smoothness nugget convergence
+       <dbl> <dbl>      <dbl>  <dbl> <lgl>      
+1 0.00000672  18.1       1.03  1138. TRUE       
+\end{verbatim}
+
+The smoothness metric is the smoothness parameter from the Gussian
+process and all details of the fit can be accessed using
+\texttt{attr(sm\_res,\ "fit\_res")}.
+
+\subsubsection{Squintability}\label{squintability}
+
+Sampling bases for calculating squintability includes an additional step
+of interpolating between the sampled bases and the best basis. This is
+specified through setting the \texttt{step\_size} and
+\texttt{min\_proj\_dist} arguments in the \texttt{sample\_bases()}
+function to non-NA numerical values. Since the projection distance
+typically ranges from 0 to 2, it is recommended to set
+\texttt{step\_size} to 0.1 or lower, and \texttt{min\_proj\_dist} to be
+at least 0.5 to ensure a meaningful interpolation length.
+
+The \texttt{calc\_squintability()} function computes squintability using
+two methods: 1) parametrically, by fitting a scaled sigmoid function
+through non-linear least square (\texttt{method\ =\ "nls"}), and 2)
+non-parametrically, using kernel smoothing (\texttt{method\ =\ "ks"}). A
+\texttt{bin\_width} argument is required to average the index values
+over the projection distance before the fit. The following code samples
+100 bases for calculating squintability in the same scenario:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{basis\_squint }\OtherTok{\textless{}{-}} \FunctionTok{sample\_bases}\NormalTok{(}
+   \AttributeTok{idx =} \StringTok{"holes"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sine1000, }\AttributeTok{n\_basis =} \DecValTok{100}\NormalTok{, }
+  \AttributeTok{best =} \FunctionTok{matrix}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{nrow =} \DecValTok{6}\NormalTok{),}
+  \AttributeTok{step\_size =} \FloatTok{0.1}\NormalTok{, }\AttributeTok{min\_proj\_dist =} \FloatTok{1.5}\NormalTok{)}
+\FunctionTok{calc\_squintability}\NormalTok{(basis\_squint, }\AttributeTok{method =} \StringTok{"nls"}\NormalTok{, }\AttributeTok{bin\_width =} \FloatTok{0.01}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# PP index:     holes
+# No. of bases: 100 -> 1853
+# method:       nls
+  theta1 theta2 theta3 theta4 squint
+   <dbl>  <dbl>  <dbl>  <dbl>  <dbl>
+1  0.274  0.960   2.21  0.612  0.812
+\end{verbatim}
+
+The output provides the estimated parameters for the scaled sigmoid
+function (\(\theta_1\) to \(\theta_4\)) and the squintability metric for
+the parametric case. For the non-parametric case, it shows the
+(\texttt{max\_d}), the corresponding projection distance
+(\texttt{max\_dist}), and teh squintability metric as their products.
+
 \subsection{Using the JSO in a PPGT}\label{using-the-jso-in-a-ppgt}

### 2024-09-04 jso.qmd
added opening paragraph for Practicalities

@@
 
+The simulation study illustrates how these metrics can be used to assess a PP index, and the optimisation process.
+In general, we would expect that the JSO is going to outperform most other optimisers.
+From an exploratory perspective, it take s shift in the GTPP practice because one no longer has a single path to follow.
+Here we describe how to adjust to using the JSO optimiser for exploring high-dimensional data, and how to use the new metrics for evaluating a new index.
+
 ## Computing the metrics for your new index

### 2024-09-04 jso.tex
added opening paragraph for Practicalities

@@
 \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
 \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
@@
 \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
 \newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
@@
 \makeatother
-
 \ifLuaTeX
@@
 
-\begin{tcolorbox}[enhanced jigsaw, coltitle=black, rightrule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, breakable, opacitybacktitle=0.6, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, opacityback=0, bottomtitle=1mm, titlerule=0mm, toptitle=1mm, bottomrule=.15mm, toprule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, coltitle=black, breakable, colback=white, colframe=quarto-callout-note-color-frame, rightrule=.15mm, arc=.35mm, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, left=2mm, bottomrule=.15mm, opacitybacktitle=0.6, toprule=.15mm, toptitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0]
 
@@
 
+The simulation study illustrates how these metrics can be used to assess
+a PP index, and the optimisation process. In general, we would expect
+that the JSO is going to outperform most other optimisers. From an
+exploratory perspective, it take s shift in the GTPP practice because
+one no longer has a single path to follow. Here we describe how to
+adjust to using the JSO optimiser for exploring high-dimensional data,
+and how to use the new metrics for evaluating a new index.
+
 \subsection{Computing the metrics for your new

### 2024-09-04 jso.qmd
working on JSO in PPGT

--- 2024-09-04 jso.qmd
@@
 
+(not done yet)
+
+The JSO can be used by specifying the argument `search_f = search_jellyfish` in the `guided_tour()` function in the `tourr` package.
+In the example below, the `lda_pp` index is used to cluster the penguins data set.
+When binding animation `animate_xy()` to an object (`res<-`), it will save the visited bases of each jellyfish as a data object.
+The bases from each jellyfish can be extracted adn supplied as a planned tour to visualise its path.
+Figure @fig-jso-penguins shows the tour path of one particular jellyfish in the penguins data set.
+The jellyfish successfully finds the structure of the data set, separating the three species of penguins.
+
+```{r}
+#| echo: true
+#| eval: false
+# run the projection pursuit with JSO 
+res <- animate_xy(penguins_df, 
+                  guided_tour(lda_pp(cl = penguins_cl$species),
+                              search_f = search_jellyfish, 
+                              n_jellies = 20, max.tries = 50))
+# extract the bases from the first jellyfish
+bases <- res |> filter(loop == 1) |> pull(basis) |> check_dup(0.1)
+# visualise its path with a planned tour
+animate_xy(data = penguins_df, tour_path = planned_tour(bases), 
+           col = penguins_cl$species)
+```
+
+```{r}
+#| echo: false
+#| eval: false
+library(dplyr)
+library(palmerpenguins)
+library(ggplot2)
+library(tourr)
+check_dup <- function(bases, min_dist) {
+  res <- list()
+  res[[1]] <- bases[[1]]
+  i <- 1; j <- 2
+  while(j <= length(bases)){
+    if (proj_dist(bases[[i]], bases[[j]]) >= min_dist) {
+      res <- c(res, list(bases[[j]]))
+      i <- i + 1; j <- j + 1
+    } else{
+      j <- j + 1
+    }
+  }
+  return(res)
+}
+
+penguins_cl <- palmerpenguins::penguins |> filter(!is.na(bill_length_mm)) |>
+  rename(bl = bill_length_mm, bd = bill_depth_mm,
+         fl = flipper_length_mm, bm = body_mass_g)
+penguins_df <- penguins_cl |> select(bl:bm) |> rescale()
+res <- animate_xy(penguins_df, guided_tour(lda_pp(cl = penguins_cl$species), search_f = search_jellyfish, n_jellies = 20, max.tries = 50))
+bases <- res |> filter(loop == 1) |> pull(basis) |> check_dup(0.1)
+animate_xy(data = penguins_df, tour_path = planned_tour(bases), col = penguins_cl$species)
+render_gif(penguins_df, planned_tour(bases), display_xy(col = penguins_cl$species), 
+           here::here("figures/tour/penguins.gif"))
+render( 
+  penguins_df,
+  tour_path = planned_tour(bases),
+  dev = "png",
+  display = display_xy(col = penguins_cl$species),
+  frames = 100,
+  here::here("figures","tour", "penguins%03d.png")
+)
+```
+
+```{r}
+#| label: fig-jso-penguins
+#| fig-cap: "Visualising the tour path of one particular jellyfish in the penguins data set.
+The jellyfish is optimised by the JSO with 20 jellyfishes and a maximum of 50 tries.
+The tour path is planned off-line and then followed in the guided tour.
+The jellyfish successfully finds the structure of the data set, separating the three species of penguins."
+frames <- c("001", "010", "019", "028", "036", "048")
+ani <- paste0(here::here("figures/"), "tour/", "penguins", frames, ".png")
+rl <- lapply(ani, png::readPNG)
+gl <-  lapply(rl, grid::rasterGrob)
+wrap_plots(gl)
+```
+
 ```{r}

### 2024-09-04 jso.tex
working on JSO in PPGT

--- 2024-09-04 jso.tex
@@
 
-\begin{tcolorbox}[enhanced jigsaw, coltitle=black, rightrule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, breakable, opacitybacktitle=0.6, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, opacityback=0, bottomtitle=1mm, titlerule=0mm, toptitle=1mm, bottomrule=.15mm, toprule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, bottomrule=.15mm, arc=.35mm, breakable, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, bottomtitle=1mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, coltitle=black, opacitybacktitle=0.6, left=2mm, opacityback=0, titlerule=0mm]
 
@@
 
+(not done yet)
+
+The JSO can be used by specifying the argument
+\texttt{search\_f\ =\ search\_jellyfish} in the \texttt{guided\_tour()}
+function in the \texttt{tourr} package. In the example below, the
+\texttt{lda\_pp} index is used to cluster the penguins data set. When
+binding animation \texttt{animate\_xy()} to an object
+(\texttt{res\textless{}-}), it will save the visited bases of each
+jellyfish as a data object. The bases from each jellyfish can be
+extracted adn supplied as a planned tour to visualise its path. Figure
+Figure~\ref{fig-jso-penguins} shows the tour path of one particular
+jellyfish in the penguins data set. The jellyfish successfully finds the
+structure of the data set, separating the three species of penguins.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# run the projection pursuit with JSO }
+\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{animate\_xy}\NormalTok{(penguins\_df, }
+                  \FunctionTok{guided\_tour}\NormalTok{(}\FunctionTok{lda\_pp}\NormalTok{(}\AttributeTok{cl =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species),}
+                              \AttributeTok{search\_f =}\NormalTok{ search\_jellyfish, }
+                              \AttributeTok{n\_jellies =} \DecValTok{20}\NormalTok{, }\AttributeTok{max.tries =} \DecValTok{50}\NormalTok{))}
+\CommentTok{\# extract the bases from the first jellyfish}
+\NormalTok{bases }\OtherTok{\textless{}{-}}\NormalTok{ res }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(loop }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{pull}\NormalTok{(basis) }\SpecialCharTok{|\textgreater{}} \FunctionTok{check\_dup}\NormalTok{(}\FloatTok{0.1}\NormalTok{)}
+\CommentTok{\# visualise its path with a planned tour}
+\FunctionTok{animate\_xy}\NormalTok{(}\AttributeTok{data =}\NormalTok{ penguins\_df, }\AttributeTok{tour\_path =} \FunctionTok{planned\_tour}\NormalTok{(bases), }
+           \AttributeTok{col =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{jso_files/figure-pdf/fig-jso-penguins-1.pdf}
+
+}
+
+\caption{\label{fig-jso-penguins}Visualising the tour path of one
+particular jellyfish in the penguins data set. The jellyfish is
+optimised by the JSO with 20 jellyfishes and a maximum of 50 tries. The
+tour path is planned off-line and then followed in the guided tour. The
+jellyfish successfully finds the structure of the data set, separating
+the three species of penguins.}
+
+\end{figure}%
+
 \section{Conclusion}\label{sec-conclusion}

### 2024-09-04 jso.tex
knitr

@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, bottomrule=.15mm, arc=.35mm, breakable, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, bottomtitle=1mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, coltitle=black, opacitybacktitle=0.6, left=2mm, opacityback=0, titlerule=0mm]
+\begin{tcolorbox}[enhanced jigsaw, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, opacitybacktitle=0.6, arc=.35mm, colback=white, bottomtitle=1mm, opacityback=0, breakable, bottomrule=.15mm, rightrule=.15mm, left=2mm, toprule=.15mm, toptitle=1mm, titlerule=0mm, coltitle=black, leftrule=.75mm]
 

### 2024-09-06 jso.qmd
fixed DLMF reference
Workaround for including the DLMF reference inside a LaTeX block.

@@
    - \providecommand{\theoremname}{Theorem}
+nocite: |
+   @DLMF
 ---
@@
 where $\nu>0$ is the smoothness parameter, $\eta$ is the outputscale, $\ell$ is the lengthscale, and $\mathcal{K}_\nu$ is
-the modified Bessel function @DLMF. A multivariate extension $K(u)$, $u\in\mathbb{R}^d$ can be obtained by products of univariate covariance functions \eqref{eq:matern}.
+the modified Bessel function [DLMF 10.25]. A multivariate extension $K(u)$, $u\in\mathbb{R}^d$ can be obtained by products of univariate covariance functions \eqref{eq:matern}.
 \end{defn}

### 2024-09-06 jso.tex
fixed DLMF reference
Workaround for including the DLMF reference inside a LaTeX block.

@@
 \end{keywords}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, sharp corners, interior hidden, frame hidden, breakable]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-\centering{
-
-\includegraphics[width=0.8\textwidth,height=\textheight]{jso_files/figure-pdf/fig-example-functions-1.pdf}
+{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{jso_files/figure-pdf/fig-example-functions-1.pdf}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, opacitybacktitle=0.6, arc=.35mm, colback=white, bottomtitle=1mm, opacityback=0, breakable, bottomrule=.15mm, rightrule=.15mm, left=2mm, toprule=.15mm, toptitle=1mm, titlerule=0mm, coltitle=black, leftrule=.75mm]
+\begin{tcolorbox}[enhanced jigsaw, breakable, rightrule=.15mm, leftrule=.75mm, opacitybacktitle=0.6, arc=.35mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, toprule=.15mm, bottomrule=.15mm, coltitle=black, titlerule=0mm, bottomtitle=1mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, opacityback=0]
 
@@
 
-\section{Simulation study}\label{sec-sim-deets}
+\hypertarget{sec-sim-deets}{%
+\section{Simulation study}\label{sec-sim-deets}}
 
@@
 \endgroup{}
-
-}
-
-\end{table}%
-
-\section{Practicalities}\label{sec-discussion}
+\end{table}
+
+\hypertarget{sec-discussion}{%
+\section{Practicalities}\label{sec-discussion}}
 
@@
 
+\hypertarget{computing-the-metrics-for-your-new-index}{%
 \subsection{Computing the metrics for your new
-index}\label{computing-the-metrics-for-your-new-index}
+index}\label{computing-the-metrics-for-your-new-index}}
 
@@
 
-\subsubsection{Smoothness}\label{smoothness}
+\hypertarget{smoothness}{%
+\subsubsection{Smoothness}\label{smoothness}}
 
@@
 
-\subsubsection{Squintability}\label{squintability}
+\hypertarget{squintability}{%
+\subsubsection{Squintability}\label{squintability}}
 
@@
 
-\subsection{Using the JSO in a PPGT}\label{using-the-jso-in-a-ppgt}
+\hypertarget{using-the-jso-in-a-ppgt}{%
+\subsection{Using the JSO in a PPGT}\label{using-the-jso-in-a-ppgt}}
 
@@
 
-\centering{
-
-\includegraphics{jso_files/figure-pdf/fig-jso-penguins-1.pdf}
+{\centering \includegraphics{jso_files/figure-pdf/fig-jso-penguins-1.pdf}
 
@@
 
-\end{figure}%
-
-\section{Conclusion}\label{sec-conclusion}
+\end{figure}
+
+\hypertarget{sec-conclusion}{%
+\section{Conclusion}\label{sec-conclusion}}
 
@@
 
-\bibitem[\citeproctext]{ref-cook1995grand}
+\leavevmode\vadjust pre{\hypertarget{ref-cook1995grand}{}}%
 Cook, D., A. Buja, J. Cabrera, and C. Hurley. 1995. {``Grand Tour and
@@
 
-\bibitem[\citeproctext]{ref-FT74}
+\leavevmode\vadjust pre{\hypertarget{ref-DLMF}{}}%
+DLMF. 2024. {``{NIST Digital Library of Mathematical Functions}.''}
+\url{https://dlmf.nist.gov/10.25}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-FT74}{}}%
 Friedman, J. H., and J. W. Tukey. 1974. {``{A} {P}rojection {P}ursuit
@@
 
-\bibitem[\citeproctext]{ref-huberplot}
+\leavevmode\vadjust pre{\hypertarget{ref-huberplot}{}}%
 ---------. 1990. {``Data Analysis and Projection Pursuit.''} Technical

### 2024-09-08 jso.qmd
practicalities section

--- 2024-09-04 jso.qmd
@@
 
-<!--  
-```{=html}
-\begin{assumption}\label{assum:my_assumption}
-This is an assumption.
-\end{assumption}
-Assumption \ref{assum:my_assumption} is convenient.
-
-\begin{claim}\label{claim:my_claim}
-This is a claim.
-\end{claim}
-Claim \ref{claim:my_claim} is suspicious.
-
-\begin{condition}\label{cond:my_condition}
-This is a condition.
-\end{condition}
-Condition \ref{cond:my_condition} is necessary.
-
-\begin{conjecture}\label{conj:my_conjecture}
-This is a conjecture.
-\end{conjecture}
-Conjecture \ref{conj:my_conjecture} is intriguing.
-
-\begin{cor}\label{cor:my_corollary}
-This is a corollary.
-\end{cor}
-Corollary \ref{cor:my_corollary} is trivial.
-
-\begin{defn}\label{def:my_definition}
-This is a definition.
-\end{defn}
-Definition \ref{def:my_definition} is clear.
-
-\begin{example}\label{ex:my_example}
-This is an example.
-\end{example}
-Example \ref{ex:my_example} is enlightening.
-
-\begin{lem}\label{lem:my_lemma}
-This is a lemma.
-\end{lem}
-Lemma \ref{lem:my_lemma} is helpful.
-
-\begin{notation}\label{not:my_notation}
-This is a notation.
-\end{notation}
-Notation \ref{not:my_notation} is cumbersome.
-
-\begin{problem}\label{prob:my_problem}
-This is a problem.
-\end{problem}
-Problem \ref{prob:my_problem} is difficult.
-
-\begin{proof}\label{proof:my_proof}
-This is a proof.
-\end{proof}
-Proof \ref{proof:my_proof} is compelling.
-
-\begin{prop}\label{prop:my_prop}
-This is a proposition.
-\end{prop}
-Proposition \ref{prop:my_prop} is useful.
-
-\begin{rem}\label{rem:my_remark}
-This is a remark.
-\end{rem}
-Remark \ref{rem:my_remark} is interesting.
-
-\begin{thm}\label{thm:my_theorem}
-This is a theorem.
-\end{thm}
-Theorem \ref{thm:my_theorem} is important.
-```
--->
-
 Projection pursuit (PP) (@kr69, @FT74, @huber85) is a dimension reduction technique aimed at identifying informative linear projections of data.
@@
 
-## Computing the metrics for your new index
-
-The `ferrn` package provides functionality for computing the smoothness and squintability metrics.
-Both cases require sampling random bases using the `sample_bases()` function, followed by  calculating the specific metric with either `calc_smoothness()` or `calc_squintability()`.
-
-### Smoothness
-
-To sample bases for calculating smoothness, the required arguments include `idx` (the index function used), `data` (the dataset), `n_basis` (the number of random bases to sample), and `best` (the best basis to compute projection distance).
-The following code samples 300 bases for using the `holes` index to find the pipe shape in the 6D `sine1000` dataset:
-
-```{r}
-#| echo: true
-library(GpGp)
-library(fields)
-library(tourr)
-library(ferrn)
-basis_smoothness <- sample_bases(
-  idx = "holes", data = sine1000, n_basis = 300, 
-  best = matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1), nrow = 6))
-basis_smoothness |> head(3)
-```
-
-The output of `sample_bases()` is a data frame with four columns: `id` (the basis ID), `basis` (the basis matrix sampled), `dist` (the projection distance to the best basis), and `index` (the index value).
-To speed up the computation of index values,  parallellasation is available through the `parallel = TRUE` argument.
-
-The `calc_smoothness()` function takes the output from `sample_bases()` and fits a Gaussian process model to the index values against the projection distances to compute the smoothness metric.
-The starting parameter for the fit can be specified using the `start_params` argument, and additional Gaussian process arguments can be provided through `other_gp_params`.
-The following code calculates the smoothness metric from the previously sampled bases:
-
-```{r}
-#| echo: true
-(sm_res <- calc_smoothness(basis_smoothness))
-```
-
-The smoothness metric is the smoothness parameter from the Gussian process and all details of the fit can be accessed using `attr(sm_res, "fit_res")`.
-
-### Squintability
-
-Sampling bases for calculating squintability includes an additional step of interpolating between the sampled bases and the best basis.
-This is specified through setting the `step_size` and `min_proj_dist` arguments in the `sample_bases()` function to non-NA numerical values.
-Since the projection distance typically ranges from 0 to 2, it is recommended to set `step_size` to 0.1 or lower, and `min_proj_dist` to be at least 0.5 to ensure a meaningful interpolation length.
-
-The `calc_squintability()` function computes squintability using two methods: 1) parametrically, by fitting a scaled sigmoid function through non-linear least square (`method = "nls"`), and 2) non-parametrically, using kernel smoothing (`method = "ks"`).
-A `bin_width` argument is required to average the index values over the projection distance before the fit.
-The following code samples 100 bases for calculating squintability in the same scenario:
-
-```{r}
-#| echo: true
-basis_squint <- sample_bases(
-   idx = "holes", data = sine1000, n_basis = 100, 
-  best = matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1), nrow = 6),
-  step_size = 0.1, min_proj_dist = 1.5)
-calc_squintability(basis_squint, method = "nls", bin_width = 0.01)
-```
-
-The output provides the estimated parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$) and the squintability metric for the parametric case.
-For the non-parametric case, it shows the (`max_d`), the corresponding projection distance (`max_dist`), and teh squintability metric as their products.
-
 ## Using the JSO in a PPGT
 
-(not done yet)
-
-The JSO can be used by specifying the argument `search_f = search_jellyfish` in the `guided_tour()` function in the `tourr` package.
-In the example below, the `lda_pp` index is used to cluster the penguins data set.
-When binding animation `animate_xy()` to an object (`res<-`), it will save the visited bases of each jellyfish as a data object.
-The bases from each jellyfish can be extracted adn supplied as a planned tour to visualise its path.
-Figure @fig-jso-penguins shows the tour path of one particular jellyfish in the penguins data set.
-The jellyfish successfully finds the structure of the data set, separating the three species of penguins.
+To use JSO for PPGT in the `tourr` package, specify `search_f = search_jellyfish` in the guided tour.
+Since multiple tour paths are explored by the jellyfish, the animation function `animate_*()` will not render the tour path for JSO.
+To visualise path visited by each individual jellyfish, assign the animation to an object (`res <- animate_xy(...)`).
+This will save the bases visited by the optimiser, along with relevant metadata, as a tibble data object (see @RJ-2021-105 for more details on the data object).
+The bases visited for individual jellyfish can then be extracted and viewed as a planned tour.
+The following code uses JSO, with 20 jellyfish and a maximum of 50 tries, on the penguins data using the `lda_pp` index.
 
@@
 #| eval: false
+# clean up the penguins data
+penguins_cl <- palmerpenguins::penguins |> 
+  filter(!is.na(bill_length_mm)) |>
+  rename(bl = bill_length_mm, bd = bill_depth_mm,
+         fl = flipper_length_mm, bm = body_mass_g)
+penguins_df <- penguins_cl |> select(bl:bm) |> rescale()
 # run the projection pursuit with JSO 
@@
                               n_jellies = 20, max.tries = 50))
+```
+
+The bases visited by jellyfish No.
+1 is extracted with `filter(loop = 1)` (the variable `loop` in the data object is the jellyfish ID) and the `check_dup()` function removes the bases in the iteration without improvements.
+The bases are then supplied as a `planned_tour()` in `animate_xy()` to visualise the tour path of jellyfish No.1.
+
+```{r}
+#| echo: true
+#| eval: false
 # extract the bases from the first jellyfish
@@
 ```
+
+@fig-jso-penguins shows six frames from the tour jellyfish No.1 animation.
+The Gentoo species (red) was mixed with Chinstrap (yellow) initially and gradually separated out as flipper length (`fl`) gaining more weights.
 
@@
 #| label: fig-jso-penguins
-#| fig-cap: "Visualising the tour path of one particular jellyfish in the penguins data set.
-The jellyfish is optimised by the JSO with 20 jellyfishes and a maximum of 50 tries.
-The tour path is planned off-line and then followed in the guided tour.
+#| fig-cap: "Visualising the tour path of a single jellyfish (No.1) in the penguins data set.
+The projection pursuit is optimised by the JSO with 20 jellyfishes and a maximum of 50 tries.
+The bases visited by jellyfish No.1 is extracted and constructed as a planned tour.
 The jellyfish successfully finds the structure of the data set, separating the three species of penguins."
@@
 
-```{r}
-#| eval: false
-#| echo: false
-jso_optima <- save_history(flea[,1:6], 
-                           guided_tour(holes()), 
-                           search_f = search_jellyfish)
-```
+## Computing the metrics for your new index
+
+The `ferrn` package provides functionality for computing the smoothness and squintability metrics.
+Both cases require sampling random bases using the `sample_bases()` function, followed by  calculating the specific metric with either `calc_smoothness()` or `calc_squintability()`.
+
+### Smoothness
+
+To sample bases for calculating smoothness, the required arguments include `idx` (the index function used), `data` (the dataset), `n_basis` (the number of random bases to sample), and `best` (the best basis to compute projection distance).
+The following code samples 300 bases for using the `holes` index to find the pipe shape in the 6D `sine1000` dataset:
+
+```{r}
+#| echo: true
+library(GpGp)
+library(fields)
+library(tourr)
+library(ferrn)
+basis_smoothness <- sample_bases(
+  idx = "holes", data = sine1000, n_basis = 300, 
+  best = matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1), nrow = 6))
+basis_smoothness |> head(3)
+```
+
+The output of `sample_bases()` is a data frame with four columns: `id` (the basis ID), `basis` (the basis matrix sampled), `dist` (the projection distance to the best basis), and `index` (the index value).
+To speed up the computation of index values,  parallellasation is available through the `parallel = TRUE` argument.
+
+The `calc_smoothness()` function takes the output from `sample_bases()` and fits a Gaussian process model to the index values against the projection distances to compute the smoothness metric.
+The starting parameter for the fit can be specified using the `start_params` argument, and additional Gaussian process arguments can be provided through `other_gp_params`.
+The following code calculates the smoothness metric from the previously sampled bases:
+
+```{r}
+#| echo: true
+(sm_res <- calc_smoothness(basis_smoothness))
+```
+
+The smoothness metric is the smoothness parameter from the Gussian process and all details of the fit can be accessed using `attr(sm_res, "fit_res")`.
+
+### Squintability
+
+Sampling bases for calculating squintability includes an additional step of interpolating between the sampled bases and the best basis.
+This is specified through setting the `step_size` and `min_proj_dist` arguments in the `sample_bases()` function to non-NA numerical values.
+Since the projection distance typically ranges from 0 to 2, it is recommended to set `step_size` to 0.1 or lower, and `min_proj_dist` to be at least 0.5 to ensure a meaningful interpolation length.
+
+The `calc_squintability()` function computes squintability using two methods: 1) parametrically, by fitting a scaled sigmoid function through non-linear least square (`method = "nls"`), and 2) non-parametrically, using kernel smoothing (`method = "ks"`).
+A `bin_width` argument is required to average the index values over the projection distance before the fit.
+The following code samples 100 bases for calculating squintability in the same scenario:
+
+```{r}
+#| echo: true
+basis_squint <- sample_bases(
+   idx = "holes", data = sine1000, n_basis = 100, 
+  best = matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1), nrow = 6),
+  step_size = 0.1, min_proj_dist = 1.5)
+calc_squintability(basis_squint, method = "nls", bin_width = 0.01)
+```
+
+The output provides the estimated parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$) and the squintability metric for the parametric case.
+For the non-parametric case, it shows the (`max_d`), the corresponding projection distance (`max_dist`), and teh squintability metric as their products.
 

### 2024-09-08 jso.tex
practicalities section

--- 2024-09-04 jso.tex
@@
 
-\begin{tcolorbox}[enhanced jigsaw, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, opacitybacktitle=0.6, arc=.35mm, colback=white, bottomtitle=1mm, opacityback=0, breakable, bottomrule=.15mm, rightrule=.15mm, left=2mm, toprule=.15mm, toptitle=1mm, titlerule=0mm, coltitle=black, leftrule=.75mm]
+\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, colback=white, breakable, colbacktitle=quarto-callout-note-color!10!white, bottomrule=.15mm, coltitle=black, opacityback=0, titlerule=0mm, leftrule=.75mm, bottomtitle=1mm, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, toprule=.15mm, left=2mm]
 
@@
 
+\subsection{Using the JSO in a PPGT}\label{using-the-jso-in-a-ppgt}
+
+To use JSO for PPGT in the \texttt{tourr} package, specify
+\texttt{search\_f\ =\ search\_jellyfish} in the guided tour. Since
+multiple tour paths are explored by the jellyfish, the animation
+function \texttt{animate\_*()} will not render the tour path for JSO. To
+visualise path visited by each individual jellyfish, assign the
+animation to an object (\texttt{res\ \textless{}-\ animate\_xy(...)}).
+This will save the bases visited by the optimiser, along with relevant
+metadata, as a tibble data object (see Zhang et al. (2021) for more
+details on the data object). The bases visited for individual jellyfish
+can then be extracted and viewed as a planned tour. The following code
+uses JSO, with 20 jellyfish and a maximum of 50 tries, on the penguins
+data using the \texttt{lda\_pp} index.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# clean up the penguins data}
+\NormalTok{penguins\_cl }\OtherTok{\textless{}{-}}\NormalTok{ palmerpenguins}\SpecialCharTok{::}\NormalTok{penguins }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(bill\_length\_mm)) }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{rename}\NormalTok{(}\AttributeTok{bl =}\NormalTok{ bill\_length\_mm, }\AttributeTok{bd =}\NormalTok{ bill\_depth\_mm,}
+         \AttributeTok{fl =}\NormalTok{ flipper\_length\_mm, }\AttributeTok{bm =}\NormalTok{ body\_mass\_g)}
+\NormalTok{penguins\_df }\OtherTok{\textless{}{-}}\NormalTok{ penguins\_cl }\SpecialCharTok{|\textgreater{}} \FunctionTok{select}\NormalTok{(bl}\SpecialCharTok{:}\NormalTok{bm) }\SpecialCharTok{|\textgreater{}} \FunctionTok{rescale}\NormalTok{()}
+\CommentTok{\# run the projection pursuit with JSO }
+\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{animate\_xy}\NormalTok{(penguins\_df, }
+                  \FunctionTok{guided\_tour}\NormalTok{(}\FunctionTok{lda\_pp}\NormalTok{(}\AttributeTok{cl =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species),}
+                              \AttributeTok{search\_f =}\NormalTok{ search\_jellyfish, }
+                              \AttributeTok{n\_jellies =} \DecValTok{20}\NormalTok{, }\AttributeTok{max.tries =} \DecValTok{50}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+The bases visited by jellyfish No.~1 is extracted with
+\texttt{filter(loop\ =\ 1)} (the variable \texttt{loop} in the data
+object is the jellyfish ID) and the \texttt{check\_dup()} function
+removes the bases in the iteration without improvements. The bases are
+then supplied as a \texttt{planned\_tour()} in \texttt{animate\_xy()} to
+visualise the tour path of jellyfish No.1.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# extract the bases from the first jellyfish}
+\NormalTok{bases }\OtherTok{\textless{}{-}}\NormalTok{ res }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(loop }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{pull}\NormalTok{(basis) }\SpecialCharTok{|\textgreater{}} \FunctionTok{check\_dup}\NormalTok{(}\FloatTok{0.1}\NormalTok{)}
+\CommentTok{\# visualise its path with a planned tour}
+\FunctionTok{animate\_xy}\NormalTok{(}\AttributeTok{data =}\NormalTok{ penguins\_df, }\AttributeTok{tour\_path =} \FunctionTok{planned\_tour}\NormalTok{(bases), }
+           \AttributeTok{col =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species)}
+\end{Highlighting}
+\end{Shaded}
+
+Figure~\ref{fig-jso-penguins} shows six frames from the tour jellyfish
+No.1 animation. The Gentoo species (red) was mixed with Chinstrap
+(yellow) initially and gradually separated out as flipper length
+(\texttt{fl}) gaining more weights.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{jso_files/figure-pdf/fig-jso-penguins-1.pdf}
+
+}
+
+\caption{\label{fig-jso-penguins}Visualising the tour path of a single
+jellyfish (No.1) in the penguins data set. The projection pursuit is
+optimised by the JSO with 20 jellyfishes and a maximum of 50 tries. The
+bases visited by jellyfish No.1 is extracted and constructed as a
+planned tour. The jellyfish successfully finds the structure of the data
+set, separating the three species of penguins.}
+
+\end{figure}%
+
 \subsection{Computing the metrics for your new
@@
 
-\subsection{Using the JSO in a PPGT}\label{using-the-jso-in-a-ppgt}
-
-(not done yet)
-
-The JSO can be used by specifying the argument
-\texttt{search\_f\ =\ search\_jellyfish} in the \texttt{guided\_tour()}
-function in the \texttt{tourr} package. In the example below, the
-\texttt{lda\_pp} index is used to cluster the penguins data set. When
-binding animation \texttt{animate\_xy()} to an object
-(\texttt{res\textless{}-}), it will save the visited bases of each
-jellyfish as a data object. The bases from each jellyfish can be
-extracted adn supplied as a planned tour to visualise its path. Figure
-Figure~\ref{fig-jso-penguins} shows the tour path of one particular
-jellyfish in the penguins data set. The jellyfish successfully finds the
-structure of the data set, separating the three species of penguins.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# run the projection pursuit with JSO }
-\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{animate\_xy}\NormalTok{(penguins\_df, }
-                  \FunctionTok{guided\_tour}\NormalTok{(}\FunctionTok{lda\_pp}\NormalTok{(}\AttributeTok{cl =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species),}
-                              \AttributeTok{search\_f =}\NormalTok{ search\_jellyfish, }
-                              \AttributeTok{n\_jellies =} \DecValTok{20}\NormalTok{, }\AttributeTok{max.tries =} \DecValTok{50}\NormalTok{))}
-\CommentTok{\# extract the bases from the first jellyfish}
-\NormalTok{bases }\OtherTok{\textless{}{-}}\NormalTok{ res }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(loop }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{pull}\NormalTok{(basis) }\SpecialCharTok{|\textgreater{}} \FunctionTok{check\_dup}\NormalTok{(}\FloatTok{0.1}\NormalTok{)}
-\CommentTok{\# visualise its path with a planned tour}
-\FunctionTok{animate\_xy}\NormalTok{(}\AttributeTok{data =}\NormalTok{ penguins\_df, }\AttributeTok{tour\_path =} \FunctionTok{planned\_tour}\NormalTok{(bases), }
-           \AttributeTok{col =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{jso_files/figure-pdf/fig-jso-penguins-1.pdf}
-
-}
-
-\caption{\label{fig-jso-penguins}Visualising the tour path of one
-particular jellyfish in the penguins data set. The jellyfish is
-optimised by the JSO with 20 jellyfishes and a maximum of 50 tries. The
-tour path is planned off-line and then followed in the guided tour. The
-jellyfish successfully finds the structure of the data set, separating
-the three species of penguins.}
-
-\end{figure}%
-
 \section{Conclusion}\label{sec-conclusion}

### 2024-09-09 jso.qmd
Update jso.qmd
edits
typos and commented out the last sentence from one of the squintability paragraph

@@
 \begin{equation}
-\mathcal{L}(\theta)=\log p(\mathbf{y}\left|\mathbf{A},\theta\right.)=-\frac{1}{2}\mathbf{y}^{\top}(\mathbf{K}+\sigma^{2}\mathbf{I})^{-1}\mathbf{y}-\frac{1}{2}\mathrm{\log}(\det(\mathbf{K}+\sigma^{2}\mathbf{I}))-\frac{N}{2}\log(2\pi)\,.\label{eq:gp_log_likelihood}
+\mathcal{L}(\theta)=\log p(\mathbf{y}\left|\mathbf{A},\theta\right.)=-\frac{1}{2}\mathbf{y}^{\top}(\mathbf{K}+\sigma^{2}\mathbf{I})^{-1}\mathbf{y}-\frac{1}{2}\mathrm{\log}(\det(\mathbf{K}+\sigma^{2}\mathbf{I}))-\frac{N}{2}\log(2\pi)\, \label{eq:gp_log_likelihood}
 \end{equation}
@@
 
-It is expected that these two
-values should be both high in the case of high squintability (fast
+It is expected that these two values should be both high in the case of high squintability (fast
 increase in $f$ early on), and both low in the case of low squintability
@@
 the optimiser needs to be very close to be able to "see" the optima.
-Thus, it could be difficult for the optimiser to find the optima. The
-mathematical formulation of this intuition is proposed below:
+Thus, it could be difficult for the optimiser to find the optima. <!--The
+mathematical formulation of this intuition is proposed below:-->
 
@@
 
-he performance of JSO is investigated both in comparison to the existing
+The performance of JSO is investigated both in comparison to the existing
 optimizer, CRS, and across various hyper-parameter values. The
@@
 In general, we would expect that the JSO is going to outperform most other optimisers.
-From an exploratory perspective, it take s shift in the GTPP practice because one no longer has a single path to follow.
+From an exploratory perspective, it take a shift in the GTPP practice because one no longer has a single path to follow.
 Here we describe how to adjust to using the JSO optimiser for exploring high-dimensional data, and how to use the new metrics for evaluating a new index.

### 2024-09-09 jso.tex
edits
typos and commented out the last sentence from one of the squintability paragraph

--- 2024-09-06 jso.tex
@@
 \end{keywords}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, sharp corners, interior hidden, frame hidden, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, boxrule=0pt, breakable, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, breakable, rightrule=.15mm, leftrule=.75mm, opacitybacktitle=0.6, arc=.35mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, toprule=.15mm, bottomrule=.15mm, coltitle=black, titlerule=0mm, bottomtitle=1mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, opacitybacktitle=0.6, left=2mm, colframe=quarto-callout-note-color-frame, breakable, toprule=.15mm, opacityback=0, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, rightrule=.15mm, colback=white, toptitle=1mm, leftrule=.75mm]
 
@@
 the optimiser needs to be very close to be able to ``see'' the optima.
-Thus, it could be difficult for the optimiser to find the optima. The
-mathematical formulation of this intuition is proposed below:
+Thus, it could be difficult for the optimiser to find the optima.
 
@@
 
-he performance of JSO is investigated both in comparison to the existing
-optimizer, CRS, and across various hyper-parameter values. The
+The performance of JSO is investigated both in comparison to the
+existing optimizer, CRS, and across various hyper-parameter values. The
 performance is measured by the success rate, defined as the proportion
@@
 
+\hypertarget{using-the-jso-in-a-ppgt}{%
+\subsection{Using the JSO in a PPGT}\label{using-the-jso-in-a-ppgt}}
+
+To use JSO for PPGT in the \texttt{tourr} package, specify
+\texttt{search\_f\ =\ search\_jellyfish} in the guided tour. Since
+multiple tour paths are explored by the jellyfish, the animation
+function \texttt{animate\_*()} will not render the tour path for JSO. To
+visualise path visited by each individual jellyfish, assign the
+animation to an object (\texttt{res\ \textless{}-\ animate\_xy(...)}).
+This will save the bases visited by the optimiser, along with relevant
+metadata, as a tibble data object (see Zhang et al. (2021) for more
+details on the data object). The bases visited for individual jellyfish
+can then be extracted and viewed as a planned tour. The following code
+uses JSO, with 20 jellyfish and a maximum of 50 tries, on the penguins
+data using the \texttt{lda\_pp} index.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# clean up the penguins data}
+\NormalTok{penguins\_cl }\OtherTok{\textless{}{-}}\NormalTok{ palmerpenguins}\SpecialCharTok{::}\NormalTok{penguins }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(bill\_length\_mm)) }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{rename}\NormalTok{(}\AttributeTok{bl =}\NormalTok{ bill\_length\_mm, }\AttributeTok{bd =}\NormalTok{ bill\_depth\_mm,}
+         \AttributeTok{fl =}\NormalTok{ flipper\_length\_mm, }\AttributeTok{bm =}\NormalTok{ body\_mass\_g)}
+\NormalTok{penguins\_df }\OtherTok{\textless{}{-}}\NormalTok{ penguins\_cl }\SpecialCharTok{|\textgreater{}} \FunctionTok{select}\NormalTok{(bl}\SpecialCharTok{:}\NormalTok{bm) }\SpecialCharTok{|\textgreater{}} \FunctionTok{rescale}\NormalTok{()}
+\CommentTok{\# run the projection pursuit with JSO }
+\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{animate\_xy}\NormalTok{(penguins\_df, }
+                  \FunctionTok{guided\_tour}\NormalTok{(}\FunctionTok{lda\_pp}\NormalTok{(}\AttributeTok{cl =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species),}
+                              \AttributeTok{search\_f =}\NormalTok{ search\_jellyfish, }
+                              \AttributeTok{n\_jellies =} \DecValTok{20}\NormalTok{, }\AttributeTok{max.tries =} \DecValTok{50}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+The bases visited by jellyfish No.~1 is extracted with
+\texttt{filter(loop\ =\ 1)} (the variable \texttt{loop} in the data
+object is the jellyfish ID) and the \texttt{check\_dup()} function
+removes the bases in the iteration without improvements. The bases are
+then supplied as a \texttt{planned\_tour()} in \texttt{animate\_xy()} to
+visualise the tour path of jellyfish No.1.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# extract the bases from the first jellyfish}
+\NormalTok{bases }\OtherTok{\textless{}{-}}\NormalTok{ res }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(loop }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{pull}\NormalTok{(basis) }\SpecialCharTok{|\textgreater{}} \FunctionTok{check\_dup}\NormalTok{(}\FloatTok{0.1}\NormalTok{)}
+\CommentTok{\# visualise its path with a planned tour}
+\FunctionTok{animate\_xy}\NormalTok{(}\AttributeTok{data =}\NormalTok{ penguins\_df, }\AttributeTok{tour\_path =} \FunctionTok{planned\_tour}\NormalTok{(bases), }
+           \AttributeTok{col =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species)}
+\end{Highlighting}
+\end{Shaded}
+
+Figure~\ref{fig-jso-penguins} shows six frames from the tour jellyfish
+No.1 animation. The Gentoo species (red) was mixed with Chinstrap
+(yellow) initially and gradually separated out as flipper length
+(\texttt{fl}) gaining more weights.
+
+\begin{figure}
+
+{\centering \includegraphics{jso_files/figure-pdf/fig-jso-penguins-1.pdf}
+
+}
+
+\caption{\label{fig-jso-penguins}Visualising the tour path of a single
+jellyfish (No.1) in the penguins data set. The projection pursuit is
+optimised by the JSO with 20 jellyfishes and a maximum of 50 tries. The
+bases visited by jellyfish No.1 is extracted and constructed as a
+planned tour. The jellyfish successfully finds the structure of the data
+set, separating the three species of penguins.}
+
+\end{figure}
+
 \hypertarget{computing-the-metrics-for-your-new-index}{%
@@
    <dbl>  <dbl>  <dbl>  <dbl>  <dbl>
-1  0.274  0.960   2.21  0.612  0.812
+1  0.274  0.960   2.21  0.612  0.203
 \end{verbatim}
@@
 (\texttt{max\_dist}), and teh squintability metric as their products.
-
-\hypertarget{using-the-jso-in-a-ppgt}{%
-\subsection{Using the JSO in a PPGT}\label{using-the-jso-in-a-ppgt}}
-
-(not done yet)
-
-The JSO can be used by specifying the argument
-\texttt{search\_f\ =\ search\_jellyfish} in the \texttt{guided\_tour()}
-function in the \texttt{tourr} package. In the example below, the
-\texttt{lda\_pp} index is used to cluster the penguins data set. When
-binding animation \texttt{animate\_xy()} to an object
-(\texttt{res\textless{}-}), it will save the visited bases of each
-jellyfish as a data object. The bases from each jellyfish can be
-extracted adn supplied as a planned tour to visualise its path. Figure
-Figure~\ref{fig-jso-penguins} shows the tour path of one particular
-jellyfish in the penguins data set. The jellyfish successfully finds the
-structure of the data set, separating the three species of penguins.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# run the projection pursuit with JSO }
-\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{animate\_xy}\NormalTok{(penguins\_df, }
-                  \FunctionTok{guided\_tour}\NormalTok{(}\FunctionTok{lda\_pp}\NormalTok{(}\AttributeTok{cl =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species),}
-                              \AttributeTok{search\_f =}\NormalTok{ search\_jellyfish, }
-                              \AttributeTok{n\_jellies =} \DecValTok{20}\NormalTok{, }\AttributeTok{max.tries =} \DecValTok{50}\NormalTok{))}
-\CommentTok{\# extract the bases from the first jellyfish}
-\NormalTok{bases }\OtherTok{\textless{}{-}}\NormalTok{ res }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(loop }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{pull}\NormalTok{(basis) }\SpecialCharTok{|\textgreater{}} \FunctionTok{check\_dup}\NormalTok{(}\FloatTok{0.1}\NormalTok{)}
-\CommentTok{\# visualise its path with a planned tour}
-\FunctionTok{animate\_xy}\NormalTok{(}\AttributeTok{data =}\NormalTok{ penguins\_df, }\AttributeTok{tour\_path =} \FunctionTok{planned\_tour}\NormalTok{(bases), }
-           \AttributeTok{col =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-
-{\centering \includegraphics{jso_files/figure-pdf/fig-jso-penguins-1.pdf}
-
-}
-
-\caption{\label{fig-jso-penguins}Visualising the tour path of one
-particular jellyfish in the penguins data set. The jellyfish is
-optimised by the JSO with 20 jellyfishes and a maximum of 50 tries. The
-tour path is planned off-line and then followed in the guided tour. The
-jellyfish successfully finds the structure of the data set, separating
-the three species of penguins.}
-
-\end{figure}
 

### 2024-09-11 jso.qmd
finished revising the Practicalities section

@@
 ---
-title: Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit
+title: New Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices
 author:
@@
 background of the PPGT, reviews existing optimisers and index
-functions in the literature. @sec-theory describes JSO and introduces the 
+functions in the literature. @sec-PP-properties introduces the 
 metrics that measure different properties of PP indexes, smoothness and
-squintability.
+squintability, and @sec-JSO describes the new JSO to be used for the PPGT.
 @sec-sim-deets summarises the results of two simulation experiments done to assess JSO's performance: one comparing JSO's performance improvements relative to an existing optimiser, Creeping Random Search (CRS), and the other studying the impact of different PP index properties on optimisation performance.
@@
 patterns of interest forming as the function is optimised. There are
-five (unsophisticated) optimisation methods implemented in the `tourr`
+five optimisation methods implemented in the `tourr`
 package:
 
+- a pseudo-derivative, that searches locally for the best direction, based on differencing the index values for very close projections.
+- a brute-force optimisation (CRS). 
+- a modified brute force algorithm described in @posse95. 
+- an essentially simulated annealing [@Bertsimas93] where the search space is reduced as the optimisation.
+- a very localised search, to take tiny steps to get closer to the local maximum.
+
+<!-- 
 -   `search_geodesic()`: provides a pseudo-derivative optimisation. It
@@
     get closer to the local maximum.
-
-There are several PP index functions available: `holes()` and `cmass()`
-[@cook1993projection]; `lda_pp()` [@lee2005projection]; `pda_pp()`
-[@lee2010projection]; `dcor2d()` and `splines2d()` [@Grimm2016];
-`norm_bin()` and `norm_kol()` [@huber85]; `slice_index()`
-[@Laa:2020wkm]. Most are relatively simply defined, for any projection
-dimension, and implemented because they are relatively easy to optimise.
-A goal is to be able to incorporate more complex PP indexes, for
-example, based on scagnostics (@scag, @WW08).
+-->
+
+There are numerous PP index functions available: introduced in @cook1993projection, @lee2005projection, @lee2010projection, @Grimm2016, @huber85, @Laa:2020wkm.
+Most are relatively simply defined, for any projection dimension, and implemented because they are relatively easy to optimise.
+A goal is to develop PP indexes based on scagnostics (@scag, @WW08), but the blockage is optimisation that these tend to be noisy, with potentially small squint angles.
 
@@
 
-# The jellyfish optimiser and properties of PP indexes {#sec-theory}
-
-JSO mimics the natural movements of jellyfish, which include passive and
-active motions driven by ocean currents and their swimming patterns,
+# Properties of PP indexes {#sec-PP-properties}
+
+@laa_using_2020 has proposed five criteria for assessing projection
+pursuit indexes (smoothness, squintability, flexibility, rotation
+invariance, and speed). Since not all index properties affect the
+optimisation process, the focus here is on the first two properties,
+*smoothness* (@sec-smoothness) and *squintability* (@sec-squintability),
+for which metrics are proposed to quantify them.
+
+## Smoothness {#sec-smoothness}
+
+This subsection proposes a metric for the smoothness of a projection
+pursuit index.
+
+A classical way to describe the smoothness of a function is to identify
+how many continuous derivatives of the function exist. This can be
+characterized by Sobolev spaces [@adams2003sobolev].
+
+```{=tex}
+\begin{defn}\label{def:sobolev_space}
+The Sobolev space $W^{k,p}(\mathbb{R})$ for $1\leq p\leq \infty$ is the set of all functions $f$ in $L^p(\mathbb{R})$ for which all weak derivatives $f^{(\ell)}$ of order $\ell\leq k$ exist and have a finite $L^p$ norm.
+\end{defn}
+```
+The Sobolev index $k$ in Definition \ref{def:sobolev_space} can be used
+to characterize the smoothness of a function: if $f\in W^{k,p}$, then
+the higher $k$, the smoother $f$. While this Sobolev index $k$ is a
+useful measure of smoothness, it can be difficult to compute or even estimate
+in practice.
+
+To obtain a computable estimator of the smoothness of the index function
+$f$, we propose an approach based on random fields. If a PP index
+function $f$ is evaluated at some random bases, as is done at the
+initialization stage of JSO, then these random index values can be
+interpreted as a random field, indexed by a space parameter, namely the
+random projection basis. This analogy suggests to use this random
+training sample to fit a spatial model. We propose to use a Gaussian
+process equipped with a Matérn covariance function, due to the
+connections between this model and Sobolev spaces, see for example
+@porcu2024matern.
+
+The distribution of a Gaussian process is fully determined by its mean
+and covariance function. The smoothness property comes into play in the
+definition of the covariance function: if a PP index is very smooth,
+then two close projection bases should produce close index values
+(strong correlation); by contrast, if a PP index is not very smooth,
+then two close projection bases might give very different index values
+(fast decay of correlations with respect to distance between bases). Popular covariance functions are parametric positive semi-definite
+functions.  In particular, the Matérn class of
+covariance functions has a dedicated parameter to capture the smoothness of
+the Gaussian field.
+
+```{=tex}
+\begin{defn}
+The Matérn covariance function $K$ is defined by
+\begin{equation}
+K(u)=K_{\nu,\eta,\ell}(u):=\eta^2\frac{\left(\sqrt{2\nu}\frac{u}{\ell}\right)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}\left(\sqrt{2\nu}\frac{u}{\ell}\right)\ , u\geq0\label{eq:matern}
+\end{equation}
+where $\nu>0$ is the smoothness parameter, $\eta$ is the outputscale, $\ell$ is the lengthscale, and $\mathcal{K}_\nu$ is
+the modified Bessel function [DLMF 10.25]. A multivariate extension $K(u)$, $u\in\mathbb{R}^d$ can be obtained by products of univariate covariance functions \eqref{eq:matern}.
+\end{defn}
+```
+The Matérn covariance function can be expressed analytically when $\nu$
+is a half-integer, the most popular values in the literature being
+$\frac{1}{2}$, $\frac{3}{2}$ and $\frac{5}{2}$ [@rasmussen2006gaussian]. The parameter $\nu$,
+called *smoothness parameter*, controls the decay of the covariance
+function. As such, it is an appropriate measure of smoothness of a
+random field, as shown by the simulations on @fig-matern-1d and @fig-matern-2d. For example, @karvonen2023asymptotic showed that if a
+function $f$ has a Sobolev index of $k$, then the smoothness parameter
+estimate $\nu$ in \eqref{eq:matern} cannot be asymptotically less than
+$k$. See the survey @porcu2024matern for additional results on the connection between the Matérn model and Sobolev spaces. An interesting result is that the asymptotic case
+$\nu\rightarrow\infty$ coincides with the Gaussian kernel:
+$K_\infty(u)=\exp(-u^{2}/2)$.
+
+```{r}
+#| label: fig-matern-1d
+#| out.width: 100%
+#| fig.cap: |
+#|   Five random simulations from a Gaussian Process defined on $\mathbb{R}$ with zero mean and Matérn-$\nu$ covariance function, with $\nu=1$ (left), $\nu=2$ (middle), and $\nu=4$ (right), showing that higher values of $\nu$ produce smoother curves.
+knitr::include_graphics(here::here("figures", "matern_simulation_1d.png"))
+```
+
+```{r}
+#| label: fig-matern-2d
+#| out.width: 100%
+#| fig.cap: |
+#|   One random simulation from a Gaussian Process defined on $\mathbb{R}^2$ with zero mean and Matérn-$\nu$ covariance function, with $\nu=1$ (left), $\nu=2$ (middle), and $\nu=4$ (right), showing that higher values of $\nu$ produce smoother surfaces.
+knitr::include_graphics(here::here("figures", "matern_simulation_2d.png"))
+```
+
+In view of these results, the parameter $\nu$ is suggested as a measure
+of the smoothness of the PP index function by fitting a Gaussian process
+prior with Matérn covariance on a dataset generated by random
+evaluations of the index function, as done at the initialization stage
+of the jellyfish search optimization. There exist several R packages,
+such as `GpGp` [@guinness2021gpgp] or `ExaGeoStatR` [@abdulah2023large],
+to fit the hyperparameters of a GP covariance function on data, which is
+usually done by maximum likelihood estimation. In this project, the
+`GpGp` package is used.
+
+```{=tex}
+\begin{defn}
+Let $\mathbf{A}=[A_1, \ldots, A_N] \in (\mathbf{R}^{p \times d})^N$ be d-dimensional projection bases, let $\mathbf{y}=[f(XA_1),\ldots,f(XA_N)]$ be the corresponding PP index values, and let $\mathbf{K}=[K_\theta(A_{i},A_{j})]_{1\leq i,j\leq N}\in\mathbb{R}^{N\times N}$ be the Matérn covariance matrix evaluated at the input bases, where the vector $\theta$ contains all the parameters of the multivariate Matérn covariance function $K$ (smoothness, outputscale, lengthscales).
+The log-likelihood of the parameters $\theta$ is defined by
+\begin{equation}
+\mathcal{L}(\theta)=\log p(\mathbf{y}\left|\mathbf{A},\theta\right.)=-\frac{1}{2}\mathbf{y}^{\top}(\mathbf{K}+\sigma^{2}\mathbf{I})^{-1}\mathbf{y}-\frac{1}{2}\mathrm{\log}(\det(\mathbf{K}+\sigma^{2}\mathbf{I}))-\frac{N}{2}\log(2\pi)\, \label{eq:gp_log_likelihood}
+\end{equation}
+where the nugget parameter $\sigma$ is the standard deviation of the intrinsic noise of the Gaussian process.
+The optimal parameters (including smoothness) are obtained by maximum log-likelihood
+\begin{equation}
+\theta^* = \underset{\theta}{\max}\mathcal{L}(\theta)
+\end{equation}
+The resulting optimal smoothness parameter $\nu$ is chosen as our smoothness metric.
+\end{defn}
+```
+
+The value of the optimal smoothness parameter $\nu>0$ can be
+naturally interpreted as follows: the higher $\nu$, the smoother the
+index function.
+
+## Squintability {#sec-squintability}
+
+Here the formal definition of projection distance and squint angle are given, before the definition of squintability. Two approaches to compute this metric numerically,
+are described.
+
+```{=tex}
+\begin{defn}[projection distance]\label{def:proj-dist}
+Recall that $A \in \mathbf{R}^{p \times d}$ is a $d$-dimensional orthonormal matrix.
+Let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.
+The projection distance $d(A, A^*)$ between $A$
+and $A^*$ is defined as 
+$d(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\  \rVert _F$
+where $\lVert . \rVert _F$ denotes the Frobenius norm, given by
+$\lVert M \rVert _F = \sqrt{\sum_{ij} M_{ij}^2}$. 
+\end{defn}
+```
+
+```{=tex}
+\begin{defn}[squint angle]\label{def:squint-angle}
+Let $A$ and $B$ be two $d$-dimensional orthonormal matrices in $\mathbb{R}^p$.
+The squint angle $\theta$ between the subspace spanned by $A$ and $B$ is defined as the smallest principal angle between these subspaces: $\theta = \min_{i \in \{1, \cdots, d\}} \arccos(\tau_i)$, where $\tau_i$ are the singular values of the matrix $M = A^T B$ obtained from its singular value decomposition.
+
+\end{defn}
+```
+
+Squintability can be defined as how the index value $f(XA)$ changes with
+respect to the projection distance $d(A, A^*)$, over the course of the
+JSO:
+
+```{=tex}
+\begin{defn}[squintability]\label{def:squintability}
+Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance $d(A, A^*)$ to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$.
+For brevity, denote $g(d)$ as $g(d(A, A^*))$.
+The squintability of an index function $f(XA)$ is defined as
+
+\begin{equation}
+\varsigma(f) = -c \times \max_{d} g'(d) \times \arg \max_{d} g'(d)
+\label{eq-squintability}
+\end{equation}
+
+where $c$ is a constant scaling factor, $-\max_d g'(d)$ represents the 
+largest gradient of $-g$ and $\arg \max_{d} g'(d)$ represents the projection distance at which this largest gradient is attained.
+
+\end{defn}
+```
+
+It is expected that these two values should be both high in the case of high squintability (fast
+increase in $f$ early on), and both low in the case of low squintability
+(any substantial increase in $f$ happens very late, close to the optimal
+angle). This suggests that their product \eqref{eq-squintability} should
+provide a sensible measure of squintability. The multiplicative constant
+$4$, which can be deemed arbitrary, does not change the interpretation
+of the squintability metric $\varsigma$; it is here to adjust the range
+of values of $\varsigma$ and simplify the explicit formula for
+$\varsigma$ obtained later on.
+
+From the definition, the following proposition can be derived:
+
+\begin{prop}\label{prop:convex-concave}
+Let $g_1(d)$ be a convex decreasing function and $g_2(d)$ be a concave decreasing function both defined on [0, 1] where $d \in [0, D]$.
+Let $d_1 := \underset{d}{\arg \max} |g_1'(d)|$ and $d_2 := \underset{d}{\arg \max} |g_2'(d)|$.
+
+$$\varsigma(g_1) < \varsigma(g_2)$$
+
+For a convex function, $g_1'(d) < 0 \; \forall d$ and is non-decreasing. For a concave function, $g_2'(d) < 0 \; \forall d$ and is non-increasing. As such, it follows that $d_1 < d_2$.
+
+Suppose both functions achieve the same maximum gradient $-g_{\max}$, i.e., $\max g'_1(d_1) = \max g'_2(d_2) = -g_{\max}$,
+
+$$
+\varsigma(g_1(d_1)) = -c  \max g'_1(d_1)  d_1 = -c  (-g_{\max}) d_1 \leq -c  (-g_{\max}) d_2 = -c \max g'_2(d_2)  d_2 = \varsigma(g_2(d_2))
+$$
+
+\end{prop}
+
+From @barnett1981interpreting and @laa_using_2020, a large squint angle
+implies that the objective function value is close to optimal even when
+the perfect view to see the structure is far away. A small squint angle
+means that the PP index value improves substantially only when the
+perfect view is close by. As such, low squintability implies rapid
+improvement in the index value when near the perfect view. For PP, a
+small squint angle is considered to be undesirable because it means that
+the optimiser needs to be very close to be able to "see" the optima.
+Thus, it could be difficult for the optimiser to find the optima. <!--The
+mathematical formulation of this intuition is proposed below:-->
+
+It is expected that for a PP index with high squint angle, the
+optimization (@eq-optimization) should make substantial progress early
+on. Conversely, for a PP index with low squint angle, it might take a
+long while for the optimization to make substantial progress, as the
+candidate projections would need to be very close to the optimal one for
+the structure of the index function to be visible enough to be amenable
+to efficient optimization. This observation suggests that the extreme
+values of $f'$ (the ones for which $f''=0$, assuming that $f$ is twice
+differentiable), and the projection distances for which these values are
+attained, are crucial in the mathematical definition of squintability.
+
+<!-- Noticing that $f$ is expected to be a decreasing function, the -->
+<!-- squintability metric is proposed as follows: -->
+
+<!-- $$ -->
+<!-- \varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ . -->
+<!-- $$ {#eq-squintability} -->
+
+<!-- $-\min_{x}f'(x)$ represents the largest value of the gradient of $-f$ -->
+<!-- and $\underset{x}{\arg\min}f'(x)$ represents the projection distance at -->
+<!-- which this largest gradient is attained.  -->
+
+To compute the squintability metric \eqref{eq-squintability} in practice,
+several approaches are possible. The first one is to propose a
+parametric model for $f$, and use it to obtain an explicit formula for
+$\varsigma$. Numerical experiments suggest a scaled sigmoid shape as
+described below. Define
+
+$$
+\ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
+$$ {#eq-logistic}
+
+which is a decreasing logistic function depending on two parameters
+$\theta_2$ and $\theta_3$, such that $\ell(\theta_{2})=\frac{1}{2}$.
+Then, define
+
+$$
+f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
+$$ {#eq-parametric}
+
+which depends on three additional parameters, $\theta_1$, $\theta_2$,
+and $x_{\max}$, such that $f(0)=\theta_1$ and $f(x_{\max})=\theta_4$.
+Under the parametric model (@eq-parametric), the squintability metric
+\eqref{eq-squintability} can be shown to be equal to
+
+$$
+\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
+$$ {#eq-squintability-parametric}
+
+In practice, the parameters of this model (@eq-parametric) can be
+estimated numerically, for example by non-linear least squares, and then
+used to evaluate $\varsigma$ as in equation
+(@eq-squintability-parametric).
+
+Alternatively, one can estimate \eqref{eq-squintability} in a nonparametric
+way, for example by fitting $f$ using kernel regression, then
+numerically estimate the angle at which $-f'$ attains its highest value.
+
+<!-- \textcolor{orange}{See whether further details are needed for the nonparametric approach.} -->
+
+<!--To the best of our knowledge, -->
+
+<!-- ## Speed -->
+
+<!-- The speed of optimizing an index function can be measured empirically by -->
+
+<!-- recording the duration of the optimisation. Alternatively, one can also -->
+
+<!-- gauge the speed in terms of computational complexity of evaluating the -->
+
+<!-- index function (for instance, in big O notation, with respect to the -->
+
+<!-- sample size) of computing the index function. -->
+
+# The jellyfish optimiser {#sec-JSO}
+
+JSO mimics the natural movements of jellyfish, which include passive and active motions driven by ocean currents and their swimming patterns,
 respectively. In the context of optimization, these movements are
@@
 the maximum number of tries.
-
-@laa_using_2020 has proposed five criteria for assessing projection
-pursuit indexes (smoothness, squintability, flexibility, rotation
-invariance, and speed). Since not all index properties affect the
-optimisation process, the focus here is on the first two properties,
-*smoothness* (@sec-smoothness) and *squintability* (@sec-squintability),
-for which metrics are proposed to quantify them.
-
-## Smoothness {#sec-smoothness}
-
-This subsection proposes a metric for the smoothness of a projection
-pursuit index.
-
-A classical way to describe the smoothness of a function is to identify
-how many continuous derivatives of the function exist. This can be
-characterized by Sobolev spaces [@adams2003sobolev].
-
-```{=tex}
-\begin{defn}\label{def:sobolev_space}
-The Sobolev space $W^{k,p}(\mathbb{R})$ for $1\leq p\leq \infty$ is the set of all functions $f$ in $L^p(\mathbb{R})$ for which all weak derivatives $f^{(\ell)}$ of order $\ell\leq k$ exist and have a finite $L^p$ norm.
-\end{defn}
-```
-The Sobolev index $k$ in Definition \ref{def:sobolev_space} can be used
-to characterize the smoothness of a function: if $f\in W^{k,p}$, then
-the higher $k$, the smoother $f$. While this Sobolev index $k$ is a
-useful measure of smoothness, it can be difficult to compute or even estimate
-in practice.
-
-To obtain a computable estimator of the smoothness of the index function
-$f$, we propose an approach based on random fields. If a PP index
-function $f$ is evaluated at some random bases, as is done at the
-initialization stage of JSO, then these random index values can be
-interpreted as a random field, indexed by a space parameter, namely the
-random projection basis. This analogy suggests to use this random
-training sample to fit a spatial model. We propose to use a Gaussian
-process equipped with a Matérn covariance function, due to the
-connections between this model and Sobolev spaces, see for example
-@porcu2024matern.
-
-The distribution of a Gaussian process is fully determined by its mean
-and covariance function. The smoothness property comes into play in the
-definition of the covariance function: if a PP index is very smooth,
-then two close projection bases should produce close index values
-(strong correlation); by contrast, if a PP index is not very smooth,
-then two close projection bases might give very different index values
-(fast decay of correlations with respect to distance between bases). Popular covariance functions are parametric positive semi-definite
-functions.  In particular, the Matérn class of
-covariance functions has a dedicated parameter to capture the smoothness of
-the Gaussian field.
-
-```{=tex}
-\begin{defn}
-The Matérn covariance function $K$ is defined by
-\begin{equation}
-K(u)=K_{\nu,\eta,\ell}(u):=\eta^2\frac{\left(\sqrt{2\nu}\frac{u}{\ell}\right)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}\left(\sqrt{2\nu}\frac{u}{\ell}\right)\ , u\geq0\label{eq:matern}
-\end{equation}
-where $\nu>0$ is the smoothness parameter, $\eta$ is the outputscale, $\ell$ is the lengthscale, and $\mathcal{K}_\nu$ is
-the modified Bessel function [DLMF 10.25]. A multivariate extension $K(u)$, $u\in\mathbb{R}^d$ can be obtained by products of univariate covariance functions \eqref{eq:matern}.
-\end{defn}
-```
-The Matérn covariance function can be expressed analytically when $\nu$
-is a half-integer, the most popular values in the literature being
-$\frac{1}{2}$, $\frac{3}{2}$ and $\frac{5}{2}$ [@rasmussen2006gaussian]. The parameter $\nu$,
-called *smoothness parameter*, controls the decay of the covariance
-function. As such, it is an appropriate measure of smoothness of a
-random field, as shown by the simulations on @fig-matern-1d and @fig-matern-2d. For example, @karvonen2023asymptotic showed that if a
-function $f$ has a Sobolev index of $k$, then the smoothness parameter
-estimate $\nu$ in \eqref{eq:matern} cannot be asymptotically less than
-$k$. See the survey @porcu2024matern for additional results on the connection between the Matérn model and Sobolev spaces. An interesting result is that the asymptotic case
-$\nu\rightarrow\infty$ coincides with the Gaussian kernel:
-$K_\infty(u)=\exp(-u^{2}/2)$.
-
-```{r}
-#| label: fig-matern-1d
-#| out.width: 100%
-#| fig.cap: |
-#|   Five random simulations from a Gaussian Process defined on $\mathbb{R}$ with zero mean and Matérn-$\nu$ covariance function, with $\nu=1$ (left), $\nu=2$ (middle), and $\nu=4$ (right), showing that higher values of $\nu$ produce smoother curves.
-knitr::include_graphics(here::here("figures", "matern_simulation_1d.png"))
-```
-
-```{r}
-#| label: fig-matern-2d
-#| out.width: 100%
-#| fig.cap: |
-#|   One random simulation from a Gaussian Process defined on $\mathbb{R}^2$ with zero mean and Matérn-$\nu$ covariance function, with $\nu=1$ (left), $\nu=2$ (middle), and $\nu=4$ (right), showing that higher values of $\nu$ produce smoother surfaces.
-knitr::include_graphics(here::here("figures", "matern_simulation_2d.png"))
-```
-
-In view of these results, the parameter $\nu$ is suggested as a measure
-of the smoothness of the PP index function by fitting a Gaussian process
-prior with Matérn covariance on a dataset generated by random
-evaluations of the index function, as done at the initialization stage
-of the jellyfish search optimization. There exist several R packages,
-such as `GpGp` [@guinness2021gpgp] or `ExaGeoStatR` [@abdulah2023large],
-to fit the hyperparameters of a GP covariance function on data, which is
-usually done by maximum likelihood estimation. In this project, the
-`GpGp` package is used.
-
-```{=tex}
-\begin{defn}
-Let $\mathbf{A}=[A_1, \ldots, A_N] \in (\mathbf{R}^{p \times d})^N$ be d-dimensional projection bases, let $\mathbf{y}=[f(XA_1),\ldots,f(XA_N)]$ be the corresponding PP index values, and let $\mathbf{K}=[K_\theta(A_{i},A_{j})]_{1\leq i,j\leq N}\in\mathbb{R}^{N\times N}$ be the Matérn covariance matrix evaluated at the input bases, where the vector $\theta$ contains all the parameters of the multivariate Matérn covariance function $K$ (smoothness, outputscale, lengthscales).
-The log-likelihood of the parameters $\theta$ is defined by
-\begin{equation}
-\mathcal{L}(\theta)=\log p(\mathbf{y}\left|\mathbf{A},\theta\right.)=-\frac{1}{2}\mathbf{y}^{\top}(\mathbf{K}+\sigma^{2}\mathbf{I})^{-1}\mathbf{y}-\frac{1}{2}\mathrm{\log}(\det(\mathbf{K}+\sigma^{2}\mathbf{I}))-\frac{N}{2}\log(2\pi)\, \label{eq:gp_log_likelihood}
-\end{equation}
-where the nugget parameter $\sigma$ is the standard deviation of the intrinsic noise of the Gaussian process.
-The optimal parameters (including smoothness) are obtained by maximum log-likelihood
-\begin{equation}
-\theta^* = \underset{\theta}{\max}\mathcal{L}(\theta)
-\end{equation}
-The resulting optimal smoothness parameter $\nu$ is chosen as our smoothness metric.
-\end{defn}
-```
-
-The value of the optimal smoothness parameter $\nu>0$ can be
-naturally interpreted as follows: the higher $\nu$, the smoother the
-index function.
-
-## Squintability {#sec-squintability}
-
-Here the formal definition of projection distance and squint angle are given, before the definition of squintability. Two approaches to compute this metric numerically,
-are described.
-
-```{=tex}
-\begin{defn}[projection distance]\label{def:proj-dist}
-Recall that $A \in \mathbf{R}^{p \times d}$ is a $d$-dimensional orthonormal matrix.
-Let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.
-The projection distance $d(A, A^*)$ between $A$
-and $A^*$ is defined as 
-$d(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\  \rVert _F$
-where $\lVert . \rVert _F$ denotes the Frobenius norm, given by
-$\lVert M \rVert _F = \sqrt{\sum_{ij} M_{ij}^2}$. 
-\end{defn}
-```
-
-```{=tex}
-\begin{defn}[squint angle]\label{def:squint-angle}
-Let $A$ and $B$ be two $d$-dimensional orthonormal matrices in $\mathbb{R}^p$.
-The squint angle $\theta$ between the subspace spanned by $A$ and $B$ is defined as the smallest principal angle between these subspaces: $\theta = \min_{i \in \{1, \cdots, d\}} \arccos(\tau_i)$, where $\tau_i$ are the singular values of the matrix $M = A^T B$ obtained from its singular value decomposition.
-
-\end{defn}
-```
-
-Squintability can be defined as how the index value $f(XA)$ changes with
-respect to the projection distance $d(A, A^*)$, over the course of the
-JSO:
-
-```{=tex}
-\begin{defn}[squintability]\label{def:squintability}
-Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance $d(A, A^*)$ to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$.
-For brevity, denote $g(d)$ as $g(d(A, A^*))$.
-The squintability of an index function $f(XA)$ is defined as
-
-\begin{equation}
-\varsigma(f) = -c \times \max_{d} g'(d) \times \arg \max_{d} g'(d)
-\label{eq-squintability}
-\end{equation}
-
-where $c$ is a constant scaling factor, $-\max_d g'(d)$ represents the 
-largest gradient of $-g$ and $\arg \max_{d} g'(d)$ represents the projection distance at which this largest gradient is attained.
-
-\end{defn}
-```
-
-It is expected that these two values should be both high in the case of high squintability (fast
-increase in $f$ early on), and both low in the case of low squintability
-(any substantial increase in $f$ happens very late, close to the optimal
-angle). This suggests that their product \eqref{eq-squintability} should
-provide a sensible measure of squintability. The multiplicative constant
-$4$, which can be deemed arbitrary, does not change the interpretation
-of the squintability metric $\varsigma$; it is here to adjust the range
-of values of $\varsigma$ and simplify the explicit formula for
-$\varsigma$ obtained later on.
-
-From the definition, the following proposition can be derived:
-
-\begin{prop}\label{prop:convex-concave}
-Let $g_1(d)$ be a convex decreasing function and $g_2(d)$ be a concave decreasing function both defined on [0, 1] where $d \in [0, D]$.
-Let $d_1 := \underset{d}{\arg \max} |g_1'(d)|$ and $d_2 := \underset{d}{\arg \max} |g_2'(d)|$.
-
-$$\varsigma(g_1) < \varsigma(g_2)$$
-
-For a convex function, $g_1'(d) < 0 \; \forall d$ and is non-decreasing. For a concave function, $g_2'(d) < 0 \; \forall d$ and is non-increasing. As such, it follows that $d_1 < d_2$.
-
-Suppose both functions achieve the same maximum gradient $-g_{\max}$, i.e., $\max g'_1(d_1) = \max g'_2(d_2) = -g_{\max}$,
-
-$$
-\varsigma(g_1(d_1)) = -c  \max g'_1(d_1)  d_1 = -c  (-g_{\max}) d_1 \leq -c  (-g_{\max}) d_2 = -c \max g'_2(d_2)  d_2 = \varsigma(g_2(d_2))
-$$
-
-\end{prop}
-
-From @barnett1981interpreting and @laa_using_2020, a large squint angle
-implies that the objective function value is close to optimal even when
-the perfect view to see the structure is far away. A small squint angle
-means that the PP index value improves substantially only when the
-perfect view is close by. As such, low squintability implies rapid
-improvement in the index value when near the perfect view. For PP, a
-small squint angle is considered to be undesirable because it means that
-the optimiser needs to be very close to be able to "see" the optima.
-Thus, it could be difficult for the optimiser to find the optima. <!--The
-mathematical formulation of this intuition is proposed below:-->
-
-It is expected that for a PP index with high squint angle, the
-optimization (@eq-optimization) should make substantial progress early
-on. Conversely, for a PP index with low squint angle, it might take a
-long while for the optimization to make substantial progress, as the
-candidate projections would need to be very close to the optimal one for
-the structure of the index function to be visible enough to be amenable
-to efficient optimization. This observation suggests that the extreme
-values of $f'$ (the ones for which $f''=0$, assuming that $f$ is twice
-differentiable), and the projection distances for which these values are
-attained, are crucial in the mathematical definition of squintability.
-
-<!-- Noticing that $f$ is expected to be a decreasing function, the -->
-<!-- squintability metric is proposed as follows: -->
-
-<!-- $$ -->
-<!-- \varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ . -->
-<!-- $$ {#eq-squintability} -->
-
-<!-- $-\min_{x}f'(x)$ represents the largest value of the gradient of $-f$ -->
-<!-- and $\underset{x}{\arg\min}f'(x)$ represents the projection distance at -->
-<!-- which this largest gradient is attained.  -->
-
-To compute the squintability metric \eqref{eq-squintability} in practice,
-several approaches are possible. The first one is to propose a
-parametric model for $f$, and use it to obtain an explicit formula for
-$\varsigma$. Numerical experiments suggest a scaled sigmoid shape as
-described below. Define
-
-$$
-\ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
-$$ {#eq-logistic}
-
-which is a decreasing logistic function depending on two parameters
-$\theta_2$ and $\theta_3$, such that $\ell(\theta_{2})=\frac{1}{2}$.
-Then, define
-
-$$
-f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
-$$ {#eq-parametric}
-
-which depends on three additional parameters, $\theta_1$, $\theta_2$,
-and $x_{\max}$, such that $f(0)=\theta_1$ and $f(x_{\max})=\theta_4$.
-Under the parametric model (@eq-parametric), the squintability metric
-\eqref{eq-squintability} can be shown to be equal to
-
-$$
-\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
-$$ {#eq-squintability-parametric}
-
-In practice, the parameters of this model (@eq-parametric) can be
-estimated numerically, for example by non-linear least squares, and then
-used to evaluate $\varsigma$ as in equation
-(@eq-squintability-parametric).
-
-Alternatively, one can estimate \eqref{eq-squintability} in a nonparametric
-way, for example by fitting $f$ using kernel regression, then
-numerically estimate the angle at which $-f'$ attains its highest value.
-
-<!-- \textcolor{orange}{See whether further details are needed for the nonparametric approach.} -->
-
-<!--To the best of our knowledge, -->
-
-<!-- ## Speed -->
-
-<!-- The speed of optimizing an index function can be measured empirically by -->
-
-<!-- recording the duration of the optimisation. Alternatively, one can also -->
-
-<!-- gauge the speed in terms of computational complexity of evaluating the -->
-
-<!-- index function (for instance, in big O notation, with respect to the -->
-
-<!-- sample size) of computing the index function. -->
 
@@
 fitting a generalised linear model with a binomial family and a logit
-link function to a sample of data in @tbl-mod-data, where all the three
+link function.<!--, where all the three
 components (success rate, JS hyper-parameters, and index properties) are
-combined. The model suggests that JSO success rate is positively
+combined.--> The model suggests that JSO success rate is positively
 associated with the two hyper-parameters, as well as with the index
@@
 
+<!-- 
 ```{r}
 #| label: tbl-mod-data
-#| tbl-cap: The first 7 rows of the datasets processed for modelling. 
+#| tbl-cap: Data constructed to store simulation results (the first 7 rows) includes name of index, dimension, index metrics, JSO parameters, and performance metrics. 
 sim_df |> 
@@
 ```
+-->
 
@@
 
-The simulation study illustrates how these metrics can be used to assess a PP index, and the optimisation process.
-In general, we would expect that the JSO is going to outperform most other optimisers.
-From an exploratory perspective, it take a shift in the GTPP practice because one no longer has a single path to follow.
-Here we describe how to adjust to using the JSO optimiser for exploring high-dimensional data, and how to use the new metrics for evaluating a new index.
+The simulation study demonstrated the that these metrics can be used to assess a PP index, and compare optimisers.
+Here we explain how they can be used for new index and optimiser development.
+Using the JSO optimiser for GTPP requires a change in user behaviour.
+For all the existing optimisers a single optimisation path is followed.
+The JSO has multiple optimisation paths, and needs additional tools to incorporate into the PPGT, as explained here.
 
@@
 To use JSO for PPGT in the `tourr` package, specify `search_f = search_jellyfish` in the guided tour.
-Since multiple tour paths are explored by the jellyfish, the animation function `animate_*()` will not render the tour path for JSO.
+Since multiple tour paths are generated by the jellyfish, the animation function `animate_*()` will not render the tour path for JSO.
 To visualise path visited by each individual jellyfish, assign the animation to an object (`res <- animate_xy(...)`).
 This will save the bases visited by the optimiser, along with relevant metadata, as a tibble data object (see @RJ-2021-105 for more details on the data object).
-The bases visited for individual jellyfish can then be extracted and viewed as a planned tour.
-The following code uses JSO, with 20 jellyfish and a maximum of 50 tries, on the penguins data using the `lda_pp` index.
-
-```{r}
-#| echo: true
+The bases visited for individual jellyfish can then be extracted and viewed using the `planned_tour()` function.
+<!-- The following code uses JSO, with 20 jellyfish and a maximum of 50 tries, on the penguins data using the `lda_pp` index.-->
+
+```{r}
+#| echo: false
 #| eval: false
@@
 ```{r}
-#| echo: true
+#| echo: false
 #| eval: false
@@
 
-@fig-jso-penguins shows six frames from the tour jellyfish No.1 animation.
-The Gentoo species (red) was mixed with Chinstrap (yellow) initially and gradually separated out as flipper length (`fl`) gaining more weights.
+<!-- @fig-jso-penguins shows six frames from the tour jellyfish No.1 animation.
+The Gentoo species (red) was mixed with Chinstrap (yellow) initially and gradually separated out as flipper length (`fl`) gaining more weights.-->
 
@@
 
+<!-- 
 ```{r}
@@
 
-The `ferrn` package provides functionality for computing the smoothness and squintability metrics.
+The `ferrn` package [@RJ-2021-105] provides functionality for computing the smoothness and squintability metrics.
 Both cases require sampling random bases using the `sample_bases()` function, followed by  calculating the specific metric with either `calc_smoothness()` or `calc_squintability()`.
@@
 To sample bases for calculating smoothness, the required arguments include `idx` (the index function used), `data` (the dataset), `n_basis` (the number of random bases to sample), and `best` (the best basis to compute projection distance).
-The following code samples 300 bases for using the `holes` index to find the pipe shape in the 6D `sine1000` dataset:
-
-```{r}
-#| echo: true
+<!--The following code samples 300 bases for using the `holes` index to find the pipe shape in the 6D `sine1000` dataset: -->
+
+```{r}
+#| echo: false
+#| eval: false
 library(GpGp)
@@
 The starting parameter for the fit can be specified using the `start_params` argument, and additional Gaussian process arguments can be provided through `other_gp_params`.
-The following code calculates the smoothness metric from the previously sampled bases:
-
-```{r}
-#| echo: true
+<!--The following code calculates the smoothness metric from the previously sampled bases:-->
+
+```{r}
+#| echo: false
+#| eval: false
 (sm_res <- calc_smoothness(basis_smoothness))
@@
 A `bin_width` argument is required to average the index values over the projection distance before the fit.
-The following code samples 100 bases for calculating squintability in the same scenario:
-
-```{r}
-#| echo: true
+<!-- The following code samples 100 bases for calculating squintability in the same scenario:-->
+
+```{r}
+#| echo: false
+#| eval: false
 basis_squint <- sample_bases(
@@
 The source code for reproducing the work reported in this paper can be
-found at: <https://github.com/huizezhang-sherry/paper-jso>. Nicolas
+found at: [https://github.com/huizezhang-sherry/paper-jso](https://github.com/huizezhang-sherry/paper-jso). Nicolas
 Langrené acknowledges the partial support of the Guangdong Provincial

### 2024-09-11 jso.tex
finished revising the Practicalities section

--- 2024-09-08 jso.tex
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
-\makeatother
-
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{241,243,245}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
 
@@
 \hypersetup{
-  pdftitle={Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit},
+  pdftitle={New Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices},
   pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
@@
 
-\title{Studying the Performance of the Jellyfish Search Optimiser for
-the Application of Projection Pursuit}
+\title{New Metrics for Assessing Projection Pursuit Indexes, and Guiding
+Optimisation Choices}
 \author{H. Sherry Zhang$\textsuperscript{1}$, Dianne
@@
 introduces the background of the PPGT, reviews existing optimisers and
-index functions in the literature. Section~\ref{sec-theory} describes
-JSO and introduces the metrics that measure different properties of PP
-indexes, smoothness and squintability. Section~\ref{sec-sim-deets}
-summarises the results of two simulation experiments done to assess
-JSO's performance: one comparing JSO's performance improvements relative
-to an existing optimiser, Creeping Random Search (CRS), and the other
-studying the impact of different PP index properties on optimisation
-performance. Section~\ref{sec-sim-res} presents the results.
+index functions in the literature. Section~\ref{sec-PP-properties}
+introduces the metrics that measure different properties of PP indexes,
+smoothness and squintability, and Section~\ref{sec-JSO} describes the
+new JSO to be used for the PPGT. Section~\ref{sec-sim-deets} summarises
+the results of two simulation experiments done to assess JSO's
+performance: one comparing JSO's performance improvements relative to an
+existing optimiser, Creeping Random Search (CRS), and the other studying
+the impact of different PP index properties on optimisation performance.
+Section~\ref{sec-sim-res} presents the results.
 Section~\ref{sec-discussion} discusses implementation details and
@@
 patterns of interest forming as the function is optimised. There are
-five (unsophisticated) optimisation methods implemented in the
-\texttt{tourr} package:
+five optimisation methods implemented in the \texttt{tourr} package:
 
@@
 \item
-  \texttt{search\_geodesic()}: provides a pseudo-derivative
-  optimisation. It searches locally for the best direction, based on
-  differencing the index values for very close projections. Then it
-  follows the direction along the geodesic path between planes, stopping
-  when the next index value fails to increase.
+  a pseudo-derivative, that searches locally for the best direction,
+  based on differencing the index values for very close projections.
 \item
-  \texttt{search\_better()}: also known as Creeping Random Search (CRS),
-  is a brute-force optimisation searching randomly for projections with
-  higher index values.
+  a brute-force optimisation (CRS).
 \item
-  \texttt{search\_better\_random()}: is essentially simulated annealing
-  (Bertsimas and Tsitsiklis 1993) where the search space is reduced as
-  the optimisation progresses.
+  a modified brute force algorithm described in Posse (1995b).
 \item
-  \texttt{search\_posse()}: implements the algorithm described in Posse
-  (1995b).
+  an essentially simulated annealing (Bertsimas and Tsitsiklis 1993)
+  where the search space is reduced as the optimisation.
 \item
-  \texttt{search\_polish()}: is a very localised search, to take tiny
-  steps to get closer to the local maximum.
+  a very localised search, to take tiny steps to get closer to the local
+  maximum.
 \end{itemize}
 
-There are several PP index functions available: \texttt{holes()} and
-\texttt{cmass()} (Cook, Buja, and Cabrera 1993); \texttt{lda\_pp()} (Lee
-et al. 2005); \texttt{pda\_pp()} (Lee and Cook 2010); \texttt{dcor2d()}
-and \texttt{splines2d()} (Grimm 2016); \texttt{norm\_bin()} and
-\texttt{norm\_kol()} (Huber 1985); \texttt{slice\_index()} (Ursula Laa
-and Valencia 2022). Most are relatively simply defined, for any
-projection dimension, and implemented because they are relatively easy
-to optimise. A goal is to be able to incorporate more complex PP
-indexes, for example, based on scagnostics (L. Wilkinson, Anand, and
-Grossman (2005), Leland Wilkinson and Wills (2008)).
+There are numerous PP index functions available: introduced in Cook,
+Buja, and Cabrera (1993), Lee et al. (2005), Lee and Cook (2010), Grimm
+(2016), Huber (1985), Ursula Laa and Valencia (2022). Most are
+relatively simply defined, for any projection dimension, and implemented
+because they are relatively easy to optimise. A goal is to develop PP
+indexes based on scagnostics (L. Wilkinson, Anand, and Grossman (2005),
+Leland Wilkinson and Wills (2008)), but the blockage is optimisation
+that these tend to be noisy, with potentially small squint angles.
 
@@
 
-\section{The jellyfish optimiser and properties of PP
-indexes}\label{sec-theory}
-
-JSO mimics the natural movements of jellyfish, which include passive and
-active motions driven by ocean currents and their swimming patterns,
-respectively. In the context of optimization, these movements are
-abstracted to explore the search space, aiming to balance exploration
-(searching new areas) and exploitation (focusing on promising areas).
-The algorithm aims to find the optimal solution by adapting the
-behaviour of jellyfish to navigate towards the best solution over
-iterations (Chou and Truong 2021).
-
-To solve the optimisation problem embedded in the PP guided tour, a
-starting projection, an index function, the number of jellyfish, and the
-maximum number of trials (tries) are provided as input. Then, the
-current projection is evaluated by the index function. The projection is
-then moved in a direction determined by a random factor, influenced by
-how far along we are in the optimisation process. Occasionally,
-completely new directions may be taken like a jellyfish might with ocean
-currents. A new projection is accepted if it is an improvement compared
-to the current one, rejected otherwise. This process continues and
-iteratively improves the projection, until the pre-specified maximum
-number of trials is reached.
-
-\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, colback=white, breakable, colbacktitle=quarto-callout-note-color!10!white, bottomrule=.15mm, coltitle=black, opacityback=0, titlerule=0mm, leftrule=.75mm, bottomtitle=1mm, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, toprule=.15mm, left=2mm]
-
-\textbf{Input}: \texttt{current\_projections}, \texttt{index\_function},
-\texttt{trial\_id}, \texttt{max\_trial}
-
-\textbf{Output}: \texttt{optimised\_projection}
-
-\textbf{Initialize} \texttt{current\_best} as the projection with the
-best index value from \texttt{current\_projections}, and
-\texttt{current\_idx} as the array of index values for each projection
-in \texttt{current\_projections}
-
-\textbf{for} each \texttt{trial\_id} in 1 to \texttt{max\_tries}
-\textbf{do}
-
-\begin{quote}
-Calculate the time control value, \(c_t\), based on
-\texttt{current\_idx} and \texttt{max\_trial}
-\end{quote}
-
-\begin{quote}
-\textbf{if} \(c_t\) is greater than or equal to \(0.5\) \textbf{then}
-
-\begin{quote}
-Define trend based on the \texttt{current\_best} and
-\texttt{current\_projections}
-\end{quote}
-
-\begin{quote}
-Update each projection towards the trend using a random factor and
-orthonormalisation
-\end{quote}
-
-\textbf{else}
-
-\begin{quote}
-\textbf{if} a random number is greater than \(1 - c_t\) \textbf{then}
-
-\begin{quote}
-Slightly adjust each projection with a small random factor (passive)
-\end{quote}
-
-\textbf{else}
-
-\begin{quote}
-For each projection, compare with a random jellyfish and adjust towards
-or away from it (active)
-\end{quote}
-\end{quote}
-
-Update the orientation of each projection to maintain consistency
-
-Evaluate the new projections using the index function
-\end{quote}
-
-\begin{quote}
-\textbf{if} any new projection is worse than the current, revert to the
-\texttt{current\_projections} for that case
-
-\begin{quote}
-Determine the projection with the best index value as the new
-\texttt{current\_best}
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\textbf{if} \texttt{trial\_id} \(\ge\) \texttt{max\_trial}, print the
-last best projection \textbf{exit}
-\end{quote}
-
-\textbf{return} the set of projections with the updated
-\texttt{current\_best} as the \texttt{optimised\_projection}
-
-\end{tcolorbox}
-
-The JSO implementation involves several key parameters that control its
-search process in optimization problems. These parameters are designed
-to guide the exploration and exploitation phases of the algorithm. While
-the specific implementation details can vary depending on the version of
-the algorithm or its application, the focus is on two main parameters
-that are most relevant to our application: the number of jellyfish and
-the maximum number of tries.
+\section{Properties of PP indexes}\label{sec-PP-properties}
 
@@
 
+\section{The jellyfish optimiser}\label{sec-JSO}
+
+JSO mimics the natural movements of jellyfish, which include passive and
+active motions driven by ocean currents and their swimming patterns,
+respectively. In the context of optimization, these movements are
+abstracted to explore the search space, aiming to balance exploration
+(searching new areas) and exploitation (focusing on promising areas).
+The algorithm aims to find the optimal solution by adapting the
+behaviour of jellyfish to navigate towards the best solution over
+iterations (Chou and Truong 2021).
+
+To solve the optimisation problem embedded in the PP guided tour, a
+starting projection, an index function, the number of jellyfish, and the
+maximum number of trials (tries) are provided as input. Then, the
+current projection is evaluated by the index function. The projection is
+then moved in a direction determined by a random factor, influenced by
+how far along we are in the optimisation process. Occasionally,
+completely new directions may be taken like a jellyfish might with ocean
+currents. A new projection is accepted if it is an improvement compared
+to the current one, rejected otherwise. This process continues and
+iteratively improves the projection, until the pre-specified maximum
+number of trials is reached.
+
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, colback=white, rightrule=.15mm, opacitybacktitle=0.6, leftrule=.75mm, coltitle=black, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, opacityback=0, left=2mm, bottomtitle=1mm, titlerule=0mm, arc=.35mm, toprule=.15mm, bottomrule=.15mm]
+
+\textbf{Input}: \texttt{current\_projections}, \texttt{index\_function},
+\texttt{trial\_id}, \texttt{max\_trial}
+
+\textbf{Output}: \texttt{optimised\_projection}
+
+\textbf{Initialize} \texttt{current\_best} as the projection with the
+best index value from \texttt{current\_projections}, and
+\texttt{current\_idx} as the array of index values for each projection
+in \texttt{current\_projections}
+
+\textbf{for} each \texttt{trial\_id} in 1 to \texttt{max\_tries}
+\textbf{do}
+
+\begin{quote}
+Calculate the time control value, \(c_t\), based on
+\texttt{current\_idx} and \texttt{max\_trial}
+\end{quote}
+
+\begin{quote}
+\textbf{if} \(c_t\) is greater than or equal to \(0.5\) \textbf{then}
+
+\begin{quote}
+Define trend based on the \texttt{current\_best} and
+\texttt{current\_projections}
+\end{quote}
+
+\begin{quote}
+Update each projection towards the trend using a random factor and
+orthonormalisation
+\end{quote}
+
+\textbf{else}
+
+\begin{quote}
+\textbf{if} a random number is greater than \(1 - c_t\) \textbf{then}
+
+\begin{quote}
+Slightly adjust each projection with a small random factor (passive)
+\end{quote}
+
+\textbf{else}
+
+\begin{quote}
+For each projection, compare with a random jellyfish and adjust towards
+or away from it (active)
+\end{quote}
+\end{quote}
+
+Update the orientation of each projection to maintain consistency
+
+Evaluate the new projections using the index function
+\end{quote}
+
+\begin{quote}
+\textbf{if} any new projection is worse than the current, revert to the
+\texttt{current\_projections} for that case
+
+\begin{quote}
+Determine the projection with the best index value as the new
+\texttt{current\_best}
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\textbf{if} \texttt{trial\_id} \(\ge\) \texttt{max\_trial}, print the
+last best projection \textbf{exit}
+\end{quote}
+
+\textbf{return} the set of projections with the updated
+\texttt{current\_best} as the \texttt{optimised\_projection}
+
+\end{tcolorbox}
+
+The JSO implementation involves several key parameters that control its
+search process in optimization problems. These parameters are designed
+to guide the exploration and exploitation phases of the algorithm. While
+the specific implementation details can vary depending on the version of
+the algorithm or its application, the focus is on two main parameters
+that are most relevant to our application: the number of jellyfish and
+the maximum number of tries.
+
 \section{Simulation study}\label{sec-sim-deets}
@@
 Table~\ref{tbl-mod-output} presents the results of fitting a generalised
-linear model with a binomial family and a logit link function to a
-sample of data in Table~\ref{tbl-mod-data}, where all the three
-components (success rate, JS hyper-parameters, and index properties) are
-combined. The model suggests that JSO success rate is positively
-associated with the two hyper-parameters, as well as with the index
-properties: smoothness and squintability. Specifically, using 10 more
-jellyfish and 10 more tries increases the odd ratio of success by
-24.11\% and 11.93\%, respectively. However, being flagged with long
-runtime and an increase of data dimension reduce the success rate by
-41.72\% and 53.36\%, respectively. The variable \texttt{squintability}
-and \texttt{dimension} are significant, suggesting their importance
-relative to JSO hyper-parameters in the optimisation success.
+linear model with a binomial family and a logit link function. The model
+suggests that JSO success rate is positively associated with the two
+hyper-parameters, as well as with the index properties: smoothness and
+squintability. Specifically, using 10 more jellyfish and 10 more tries
+increases the odd ratio of success by 24.11\% and 11.93\%, respectively.
+However, being flagged with long runtime and an increase of data
+dimension reduce the success rate by 41.72\% and 53.36\%, respectively.
+The variable \texttt{squintability} and \texttt{dimension} are
+significant, suggesting their importance relative to JSO
+hyper-parameters in the optimisation success.
 
@@
 
-\caption{\label{tbl-mod-data}The first 7 rows of the datasets processed
-for modelling.}
-
-\centering{
-
-\centering\begingroup\fontsize{10}{12}\selectfont
-
-\begin{tabular}{lrrrrrrr}
-\toprule
-index & d & smoothness & squintability & n. jellyfish & max. tries & success rate & time (sec)\\
-\midrule
-MIC & 6 & 2.46 & 1.26 & 20 & 50 & 0.12 & 2.48\\
-MIC & 6 & 2.46 & 1.26 & 20 & 100 & 0.24 & 8.95\\
-MIC & 6 & 2.46 & 1.26 & 50 & 50 & 0.52 & 5.65\\
-MIC & 6 & 2.46 & 1.26 & 50 & 100 & 0.64 & 13.22\\
-MIC & 6 & 2.46 & 1.26 & 100 & 50 & 0.76 & 19.45\\
-MIC & 8 & 2.64 & 0.77 & 20 & 50 & 0.08 & 2.57\\
-MIC & 8 & 2.64 & 0.77 & 20 & 100 & 0.08 & 4.96\\
-\bottomrule
-\end{tabular}
-\endgroup{}
-
-}
-
-\end{table}%
-
-\begin{table}
-
 \caption{\label{tbl-mod-output}Model estimates of proportion of
@@
 
-The simulation study illustrates how these metrics can be used to assess
-a PP index, and the optimisation process. In general, we would expect
-that the JSO is going to outperform most other optimisers. From an
-exploratory perspective, it take s shift in the GTPP practice because
-one no longer has a single path to follow. Here we describe how to
-adjust to using the JSO optimiser for exploring high-dimensional data,
-and how to use the new metrics for evaluating a new index.
+The simulation study demonstrated the that these metrics can be used to
+assess a PP index, and compare optimisers. Here we explain how they can
+be used for new index and optimiser development. Using the JSO optimiser
+for GTPP requires a change in user behaviour. For all the existing
+optimisers a single optimisation path is followed. The JSO has multiple
+optimisation paths, and needs additional tools to incorporate into the
+PPGT, as explained here.
 
@@
 \texttt{search\_f\ =\ search\_jellyfish} in the guided tour. Since
-multiple tour paths are explored by the jellyfish, the animation
+multiple tour paths are generated by the jellyfish, the animation
 function \texttt{animate\_*()} will not render the tour path for JSO. To
@@
 details on the data object). The bases visited for individual jellyfish
-can then be extracted and viewed as a planned tour. The following code
-uses JSO, with 20 jellyfish and a maximum of 50 tries, on the penguins
-data using the \texttt{lda\_pp} index.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# clean up the penguins data}
-\NormalTok{penguins\_cl }\OtherTok{\textless{}{-}}\NormalTok{ palmerpenguins}\SpecialCharTok{::}\NormalTok{penguins }\SpecialCharTok{|\textgreater{}} 
-  \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(bill\_length\_mm)) }\SpecialCharTok{|\textgreater{}}
-  \FunctionTok{rename}\NormalTok{(}\AttributeTok{bl =}\NormalTok{ bill\_length\_mm, }\AttributeTok{bd =}\NormalTok{ bill\_depth\_mm,}
-         \AttributeTok{fl =}\NormalTok{ flipper\_length\_mm, }\AttributeTok{bm =}\NormalTok{ body\_mass\_g)}
-\NormalTok{penguins\_df }\OtherTok{\textless{}{-}}\NormalTok{ penguins\_cl }\SpecialCharTok{|\textgreater{}} \FunctionTok{select}\NormalTok{(bl}\SpecialCharTok{:}\NormalTok{bm) }\SpecialCharTok{|\textgreater{}} \FunctionTok{rescale}\NormalTok{()}
-\CommentTok{\# run the projection pursuit with JSO }
-\NormalTok{res }\OtherTok{\textless{}{-}} \FunctionTok{animate\_xy}\NormalTok{(penguins\_df, }
-                  \FunctionTok{guided\_tour}\NormalTok{(}\FunctionTok{lda\_pp}\NormalTok{(}\AttributeTok{cl =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species),}
-                              \AttributeTok{search\_f =}\NormalTok{ search\_jellyfish, }
-                              \AttributeTok{n\_jellies =} \DecValTok{20}\NormalTok{, }\AttributeTok{max.tries =} \DecValTok{50}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
+can then be extracted and viewed using the \texttt{planned\_tour()}
+function.
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# extract the bases from the first jellyfish}
-\NormalTok{bases }\OtherTok{\textless{}{-}}\NormalTok{ res }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(loop }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \FunctionTok{pull}\NormalTok{(basis) }\SpecialCharTok{|\textgreater{}} \FunctionTok{check\_dup}\NormalTok{(}\FloatTok{0.1}\NormalTok{)}
-\CommentTok{\# visualise its path with a planned tour}
-\FunctionTok{animate\_xy}\NormalTok{(}\AttributeTok{data =}\NormalTok{ penguins\_df, }\AttributeTok{tour\_path =} \FunctionTok{planned\_tour}\NormalTok{(bases), }
-           \AttributeTok{col =}\NormalTok{ penguins\_cl}\SpecialCharTok{$}\NormalTok{species)}
-\end{Highlighting}
-\end{Shaded}
-
-Figure~\ref{fig-jso-penguins} shows six frames from the tour jellyfish
-No.1 animation. The Gentoo species (red) was mixed with Chinstrap
-(yellow) initially and gradually separated out as flipper length
-(\texttt{fl}) gaining more weights.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{jso_files/figure-pdf/fig-jso-penguins-1.pdf}
-
-}
-
-\caption{\label{fig-jso-penguins}Visualising the tour path of a single
-jellyfish (No.1) in the penguins data set. The projection pursuit is
-optimised by the JSO with 20 jellyfishes and a maximum of 50 tries. The
-bases visited by jellyfish No.1 is extracted and constructed as a
-planned tour. The jellyfish successfully finds the structure of the data
-set, separating the three species of penguins.}
-
-\end{figure}%
-
 \subsection{Computing the metrics for your new
@@
 
-The \texttt{ferrn} package provides functionality for computing the
-smoothness and squintability metrics. Both cases require sampling random
-bases using the \texttt{sample\_bases()} function, followed by
-calculating the specific metric with either \texttt{calc\_smoothness()}
-or \texttt{calc\_squintability()}.
+The \texttt{ferrn} package (Zhang et al. 2021) provides functionality
+for computing the smoothness and squintability metrics. Both cases
+require sampling random bases using the \texttt{sample\_bases()}
+function, followed by calculating the specific metric with either
+\texttt{calc\_smoothness()} or \texttt{calc\_squintability()}.
 
@@
 dataset), \texttt{n\_basis} (the number of random bases to sample), and
-\texttt{best} (the best basis to compute projection distance). The
-following code samples 300 bases for using the \texttt{holes} index to
-find the pipe shape in the 6D \texttt{sine1000} dataset:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{library}\NormalTok{(GpGp)}
-\FunctionTok{library}\NormalTok{(fields)}
-\FunctionTok{library}\NormalTok{(tourr)}
-\FunctionTok{library}\NormalTok{(ferrn)}
-\NormalTok{basis\_smoothness }\OtherTok{\textless{}{-}} \FunctionTok{sample\_bases}\NormalTok{(}
-  \AttributeTok{idx =} \StringTok{"holes"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sine1000, }\AttributeTok{n\_basis =} \DecValTok{300}\NormalTok{, }
-  \AttributeTok{best =} \FunctionTok{matrix}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{nrow =} \DecValTok{6}\NormalTok{))}
-\NormalTok{basis\_smoothness }\SpecialCharTok{|\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{3}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# PP index:                holes
-# No. of bases:            3
-# interpolation step size: NA
-# Min. proj. dist.:        NA
-     id basis          dist index
-  <int> <list>        <dbl> <dbl>
-1     1 <dbl [6 x 2]>  1.15 0.676
-2     2 <dbl [6 x 2]>  1.77 0.790
-3     3 <dbl [6 x 2]>  1.79 0.778
-\end{verbatim}
+\texttt{best} (the best basis to compute projection distance).
 
@@
 \texttt{start\_params} argument, and additional Gaussian process
-arguments can be provided through \texttt{other\_gp\_params}. The
-following code calculates the smoothness metric from the previously
-sampled bases:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{(sm\_res }\OtherTok{\textless{}{-}} \FunctionTok{calc\_smoothness}\NormalTok{(basis\_smoothness))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# PP index:     holes
-# No. of bases: 300 [6 x 2]
-    variance range smoothness nugget convergence
-       <dbl> <dbl>      <dbl>  <dbl> <lgl>      
-1 0.00000672  18.1       1.03  1138. TRUE       
-\end{verbatim}
+arguments can be provided through \texttt{other\_gp\_params}.
 
@@
 \texttt{bin\_width} argument is required to average the index values
-over the projection distance before the fit. The following code samples
-100 bases for calculating squintability in the same scenario:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{basis\_squint }\OtherTok{\textless{}{-}} \FunctionTok{sample\_bases}\NormalTok{(}
-   \AttributeTok{idx =} \StringTok{"holes"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sine1000, }\AttributeTok{n\_basis =} \DecValTok{100}\NormalTok{, }
-  \AttributeTok{best =} \FunctionTok{matrix}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{nrow =} \DecValTok{6}\NormalTok{),}
-  \AttributeTok{step\_size =} \FloatTok{0.1}\NormalTok{, }\AttributeTok{min\_proj\_dist =} \FloatTok{1.5}\NormalTok{)}
-\FunctionTok{calc\_squintability}\NormalTok{(basis\_squint, }\AttributeTok{method =} \StringTok{"nls"}\NormalTok{, }\AttributeTok{bin\_width =} \FloatTok{0.01}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# PP index:     holes
-# No. of bases: 100 -> 1853
-# method:       nls
-  theta1 theta2 theta3 theta4 squint
-   <dbl>  <dbl>  <dbl>  <dbl>  <dbl>
-1  0.274  0.960   2.21  0.612  0.812
-\end{verbatim}
+over the projection distance before the fit.
 
@@
 
+\bibitem[\citeproctext]{ref-DLMF}
+DLMF. 2024. {``{NIST Digital Library of Mathematical Functions}.''}
+\url{https://dlmf.nist.gov/10.25}.
+
 \bibitem[\citeproctext]{ref-FT74}

### 2024-09-11 jso.qmd
Table 1 math notations
GP parameters' names replaced by their mathematical notations in Table 1

@@
 #| label: tbl-smoothness-squintability
-#| tbl-cap: Parameters estimated from the Gaussian process (including variance, range, smoothness, and nugget) and scaled logistic function ($\theta_1$ to $\theta_4$) for the pipe-finding and sine-wave finding problems.
-The column "smooth" and "squint" represent the smoothness and squintability measures.
+#| tbl-cap: Parameters estimated from the Gaussian process (outputscale $\eta$, lengthscale $\ell$, smoothness $\nu$, and nugget $\sigma$) and scaled logistic function ($\theta_1$ to $\theta_4$ and $\varsigma$) for the pipe-finding and sine-wave finding problems.
+The columns $\nu$ and $\varsigma$ represent the smoothness and squintability measures respectively.
 dt <- tibble(shape = c(rep("pipe", 4), rep("sine", 8)), smoothness) |>
@@
       "shape", "index", "d",
-      "variance", "range", "smooth", "nugget",
-      "$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", "squint"), linesep = "",
+      "$\\eta$", "$\\ell$", "$\\nu$", "$\\sigma$",
+      "$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", "$\\varsigma$"), linesep = "",
     booktabs = T, escape = FALSE, row.names = TRUE) |> 

### 2024-09-11 jso.tex
Table 1 math notations
GP parameters' names replaced by their mathematical notations in Table 1

@@
 \end{keywords}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, breakable, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-\section{Properties of PP indexes}\label{sec-PP-properties}
+\hypertarget{sec-PP-properties}{%
+\section{Properties of PP indexes}\label{sec-PP-properties}}
 
@@
 
-\section{The jellyfish optimiser}\label{sec-JSO}
+\hypertarget{sec-JSO}{%
+\section{The jellyfish optimiser}\label{sec-JSO}}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, colback=white, rightrule=.15mm, opacitybacktitle=0.6, leftrule=.75mm, coltitle=black, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, opacityback=0, left=2mm, bottomtitle=1mm, titlerule=0mm, arc=.35mm, toprule=.15mm, bottomrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, breakable, colframe=quarto-callout-note-color-frame, toptitle=1mm, titlerule=0mm, opacitybacktitle=0.6, leftrule=.75mm, toprule=.15mm, colback=white, bottomrule=.15mm, bottomtitle=1mm, coltitle=black, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, arc=.35mm, left=2mm]
 
@@
 
+\hypertarget{tbl-smoothness-squintability}{}
 \begin{table}
-
-\caption{\label{tbl-smoothness-squintability}Parameters estimated from
-the Gaussian process (including variance, range, smoothness, and nugget)
-and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
-pipe-finding and sine-wave finding problems. The column ``smooth'' and
-``squint'' represent the smoothness and squintability measures.}
-
-\centering{
+\caption{\label{tbl-smoothness-squintability}Parameters estimated from the Gaussian process (outputscale \(\eta\),
+lengthscale \(\ell\), smoothness \(\nu\), and nugget \(\sigma\)) and
+scaled logistic function (\(\theta_1\) to \(\theta_4\) and
+\(\varsigma\)) for the pipe-finding and sine-wave finding problems. The
+columns \(\nu\) and \(\varsigma\) represent the smoothness and
+squintability measures respectively. }\tabularnewline
 
@@
 \toprule
-  & shape & index & d & variance & range & smooth & nugget & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & squint\\
+  & shape & index & d & $\eta$ & $\ell$ & $\nu$ & $\sigma$ & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & $\varsigma$\\
 \midrule
@@
 
-\section{Conclusion}\label{sec-conclusion}
+\hypertarget{sec-conclusion}{%
+\section{Conclusion}\label{sec-conclusion}}
 
@@
 
-\bibitem[\citeproctext]{ref-DLMF}
+\leavevmode\vadjust pre{\hypertarget{ref-DLMF}{}}%
 DLMF. 2024. {``{NIST Digital Library of Mathematical Functions}.''}

### 2024-09-11 jso.qmd
add code as a separate file and clean up supplementary materials

@@
 The bases visited for individual jellyfish can then be extracted and viewed using the `planned_tour()` function.
-<!-- The following code uses JSO, with 20 jellyfish and a maximum of 50 tries, on the penguins data using the `lda_pp` index.-->
-
-```{r}
-#| echo: false
-#| eval: false
-# clean up the penguins data
-penguins_cl <- palmerpenguins::penguins |> 
-  filter(!is.na(bill_length_mm)) |>
-  rename(bl = bill_length_mm, bd = bill_depth_mm,
-         fl = flipper_length_mm, bm = body_mass_g)
-penguins_df <- penguins_cl |> select(bl:bm) |> rescale()
-# run the projection pursuit with JSO 
-res <- animate_xy(penguins_df, 
-                  guided_tour(lda_pp(cl = penguins_cl$species),
-                              search_f = search_jellyfish, 
-                              n_jellies = 20, max.tries = 50))
-```
 
@@
 
-```{r}
-#| echo: false
-#| eval: false
-# extract the bases from the first jellyfish
-bases <- res |> filter(loop == 1) |> pull(basis) |> check_dup(0.1)
-# visualise its path with a planned tour
-animate_xy(data = penguins_df, tour_path = planned_tour(bases), 
-           col = penguins_cl$species)
-```
-
-<!-- @fig-jso-penguins shows six frames from the tour jellyfish No.1 animation.
-The Gentoo species (red) was mixed with Chinstrap (yellow) initially and gradually separated out as flipper length (`fl`) gaining more weights.-->
-
-```{r}
-#| echo: false
-#| eval: false
-library(dplyr)
-library(palmerpenguins)
-library(ggplot2)
-library(tourr)
-check_dup <- function(bases, min_dist) {
-  res <- list()
-  res[[1]] <- bases[[1]]
-  i <- 1; j <- 2
-  while(j <= length(bases)){
-    if (proj_dist(bases[[i]], bases[[j]]) >= min_dist) {
-      res <- c(res, list(bases[[j]]))
-      i <- i + 1; j <- j + 1
-    } else{
-      j <- j + 1
-    }
-  }
-  return(res)
-}
-
-penguins_cl <- palmerpenguins::penguins |> filter(!is.na(bill_length_mm)) |>
-  rename(bl = bill_length_mm, bd = bill_depth_mm,
-         fl = flipper_length_mm, bm = body_mass_g)
-penguins_df <- penguins_cl |> select(bl:bm) |> rescale()
-res <- animate_xy(penguins_df, guided_tour(lda_pp(cl = penguins_cl$species), search_f = search_jellyfish, n_jellies = 20, max.tries = 50))
-bases <- res |> filter(loop == 1) |> pull(basis) |> check_dup(0.1)
-animate_xy(data = penguins_df, tour_path = planned_tour(bases), col = penguins_cl$species)
-render_gif(penguins_df, planned_tour(bases), display_xy(col = penguins_cl$species), 
-           here::here("figures/tour/penguins.gif"))
-render( 
-  penguins_df,
-  tour_path = planned_tour(bases),
-  dev = "png",
-  display = display_xy(col = penguins_cl$species),
-  frames = 100,
-  here::here("figures","tour", "penguins%03d.png")
-)
-```
-
-<!-- 
-```{r}
-#| label: fig-jso-penguins
-#| fig-cap: "Visualising the tour path of a single jellyfish (No.1) in the penguins data set.
-The projection pursuit is optimised by the JSO with 20 jellyfishes and a maximum of 50 tries.
-The bases visited by jellyfish No.1 is extracted and constructed as a planned tour.
-The jellyfish successfully finds the structure of the data set, separating the three species of penguins."
-frames <- c("001", "010", "019", "028", "036", "048")
-ani <- paste0(here::here("figures/"), "tour/", "penguins", frames, ".png")
-rl <- lapply(ani, png::readPNG)
-gl <-  lapply(rl, grid::rasterGrob)
-wrap_plots(gl)
-```
--->
-
 ## Computing the metrics for your new index
@@
 To sample bases for calculating smoothness, the required arguments include `idx` (the index function used), `data` (the dataset), `n_basis` (the number of random bases to sample), and `best` (the best basis to compute projection distance).
-<!--The following code samples 300 bases for using the `holes` index to find the pipe shape in the 6D `sine1000` dataset: -->
-
-```{r}
-#| echo: false
-#| eval: false
-library(GpGp)
-library(fields)
-library(tourr)
-library(ferrn)
-basis_smoothness <- sample_bases(
-  idx = "holes", data = sine1000, n_basis = 300, 
-  best = matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1), nrow = 6))
-basis_smoothness |> head(3)
-```
-
 The output of `sample_bases()` is a data frame with four columns: `id` (the basis ID), `basis` (the basis matrix sampled), `dist` (the projection distance to the best basis), and `index` (the index value).
@@
 The starting parameter for the fit can be specified using the `start_params` argument, and additional Gaussian process arguments can be provided through `other_gp_params`.
-<!--The following code calculates the smoothness metric from the previously sampled bases:-->
-
-```{r}
-#| echo: false
-#| eval: false
-(sm_res <- calc_smoothness(basis_smoothness))
-```
-
 The smoothness metric is the smoothness parameter from the Gussian process and all details of the fit can be accessed using `attr(sm_res, "fit_res")`.
@@
 A `bin_width` argument is required to average the index values over the projection distance before the fit.
-<!-- The following code samples 100 bases for calculating squintability in the same scenario:-->
-
-```{r}
-#| echo: false
-#| eval: false
-basis_squint <- sample_bases(
-   idx = "holes", data = sine1000, n_basis = 100, 
-  best = matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1), nrow = 6),
-  step_size = 0.1, min_proj_dist = 1.5)
-calc_squintability(basis_squint, method = "nls", bin_width = 0.01)
-```
-
 The output provides the estimated parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$) and the squintability metric for the parametric case.
@@
 The source code for reproducing the work reported in this paper can be
-found at: [https://github.com/huizezhang-sherry/paper-jso](https://github.com/huizezhang-sherry/paper-jso). Nicolas
+found at: <https://github.com/huizezhang-sherry/paper-jso>. The simulation data produced in Section 5 can be found at <https://figshare.com/articles/dataset/Simulated_raw_data/26039506>. Nicolas
 Langrené acknowledges the partial support of the Guangdong Provincial
@@
 
+# Supplementary materials {-}
+
+The supplementary materials include: 1) description of all the index functions used in the simulations and 2) the script that complements the Practicalities section for using JSO in a PPGT and calculating smoothness and squintability.
+
 # References

### 2024-09-11 jso.tex
add code as a separate file and clean up supplementary materials

--- 2024-09-11 jso.tex
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, colback=white, rightrule=.15mm, opacitybacktitle=0.6, leftrule=.75mm, coltitle=black, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, opacityback=0, left=2mm, bottomtitle=1mm, titlerule=0mm, arc=.35mm, toprule=.15mm, bottomrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, colback=white, opacitybacktitle=0.6, left=2mm, titlerule=0mm, toprule=.15mm, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, colframe=quarto-callout-note-color-frame, opacityback=0, breakable, toptitle=1mm, rightrule=.15mm, bottomrule=.15mm]
 
@@
 \caption{\label{tbl-smoothness-squintability}Parameters estimated from
-the Gaussian process (including variance, range, smoothness, and nugget)
-and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
-pipe-finding and sine-wave finding problems. The column ``smooth'' and
-``squint'' represent the smoothness and squintability measures.}
+the Gaussian process (outputscale \(\eta\), lengthscale \(\ell\),
+smoothness \(\nu\), and nugget \(\sigma\)) and scaled logistic function
+(\(\theta_1\) to \(\theta_4\) and \(\varsigma\)) for the pipe-finding
+and sine-wave finding problems. The columns \(\nu\) and \(\varsigma\)
+represent the smoothness and squintability measures respectively.}
 
@@
 dataset), \texttt{n\_basis} (the number of random bases to sample), and
-\texttt{best} (the best basis to compute projection distance).
-
-The output of \texttt{sample\_bases()} is a data frame with four
-columns: \texttt{id} (the basis ID), \texttt{basis} (the basis matrix
-sampled), \texttt{dist} (the projection distance to the best basis), and
+\texttt{best} (the best basis to compute projection distance). The
+output of \texttt{sample\_bases()} is a data frame with four columns:
+\texttt{id} (the basis ID), \texttt{basis} (the basis matrix sampled),
+\texttt{dist} (the projection distance to the best basis), and
 \texttt{index} (the index value). To speed up the computation of index
@@
 \texttt{start\_params} argument, and additional Gaussian process
-arguments can be provided through \texttt{other\_gp\_params}.
-
-The smoothness metric is the smoothness parameter from the Gussian
-process and all details of the fit can be accessed using
+arguments can be provided through \texttt{other\_gp\_params}. The
+smoothness metric is the smoothness parameter from the Gussian process
+and all details of the fit can be accessed using
 \texttt{attr(sm\_res,\ "fit\_res")}.
@@
 \texttt{bin\_width} argument is required to average the index values
-over the projection distance before the fit.
-
-The output provides the estimated parameters for the scaled sigmoid
-function (\(\theta_1\) to \(\theta_4\)) and the squintability metric for
-the parametric case. For the non-parametric case, it shows the
-(\texttt{max\_d}), the corresponding projection distance
-(\texttt{max\_dist}), and teh squintability metric as their products.
+over the projection distance before the fit. The output provides the
+estimated parameters for the scaled sigmoid function (\(\theta_1\) to
+\(\theta_4\)) and the squintability metric for the parametric case. For
+the non-parametric case, it shows the (\texttt{max\_d}), the
+corresponding projection distance (\texttt{max\_dist}), and teh
+squintability metric as their products.
 
@@
 paper can be found at:
-\url{https://github.com/huizezhang-sherry/paper-jso}. Nicolas Langrené
-acknowledges the partial support of the Guangdong Provincial Key
-Laboratory IRADS (2022B1212010006, R0400001-22) and the UIC Start-up
-Research Fund UICR0700041-22.
+\url{https://github.com/huizezhang-sherry/paper-jso}. The simulation
+data produced as described in Section 5 in can be found at
+\url{https://figshare.com/articles/dataset/Simulated_raw_data/26039506}.
+Nicolas Langrené acknowledges the partial support of the Guangdong
+Provincial Key Laboratory IRADS (2022B1212010006, R0400001-22) and the
+UIC Start-up Research Fund UICR0700041-22.
+
+\section*{Supplementary materials}\label{supplementary-materials}
+\addcontentsline{toc}{section}{Supplementary materials}
+
+The supplementary materials include: 1) description of all the index
+functions used in the simulations and 2) the script that complements the
+Practicalities section for using JSO in a PPGT and calculating
+smoothness and squintability.
 

### 2024-09-11 supplementary/index-def.qmd
add code as a separate file and clean up supplementary materials

--- 2024-06-23 appendix/appendix.qmd
@@
 ---
-title: 'Appendix to "Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit"'
-author:
-  - name: H. Sherry Zhang
-    email: huize.zhang@austin.utexas.edu
-    affiliations: 
-        - id: 1
-          name: University of Texas at Austin
-          department: Department of Statistics and Data Sciences
-          city: Austin
-          country: United States
-          postal-code: 78751
-    attributes:
-        corresponding: true
-  - name: Dianne Cook
-    email: dicook@monash.edu
-    affiliations:
-        - id: 2
-          name: Monash University
-          department: Department of Econometrics and Business Statistics
-          city: Melbourne
-          country: Australia
-          postal-code: 3800
-  - name: Nicolas  Langrené
-    email: nicolaslangrene@uic.edu.cn
-    affiliations:
-        - id: 3
-          name: Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College
-          department: Department of Mathematical Sciences
-          city: Zhuhai
-          country: China
-          postal-code: 519087
-  - name: Jessica Wai Yin Leung
-    email: Jessica.Leung@monash.edu
-    affiliations:
-        - id: 2
-          name: Monash University
-          department: Department of Econometrics and Business Statistics
-          city: Melbourne
-          country: Australia
-          postal-code: 3800
+title: 'Appendix to " New Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices"'
+author: H. Sherry Zhang, Dianne Cook, Nicolas Langrené, Jessica Wai Yin Leung
 date: last-modified
-abstract: |
-  Projection pursuit (PP) is a dimension reduction technique that identifies low-dimensional projections in high-dimensional data by optimising a criteria function known as the PP index.
-The optimisation in PP can be non-smooth and requires identifying optima with a small "squint angle", detectable only from close proximity.
-To address these challenges, this study investigates the performance of a recent swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
-The performance of JSO is evaluated across various hyper-parameter settings and compared with existing optimisers.
-Additionally, this work proposes novel methods to quantify two properties of the PP index -- smoothness and squintability -- that capture the complexities inherent in PP optimisation problems.
-These two metrics are evaluated with JSO hyper-parameters to determine their effect on JSO success rate.
-The JSO algorithm has been implemented in the `tourr` package, while calculations for smoothness and squintability are available in the `ferrn` package.
-format:
-  elsevier-pdf:
-    journal:
-      name: Journal of Multivariate Analysis
-      formatting: preprint
-      model: 3p
-      cite-style: numbername
+format: pdf
 bibliography: ../bibliography.bib
@@
 
-   -  column sum, row sum, and overall sum: $$a_{i \cdot} = \sum_{l = 1}^n a_{il}, \quad a_{\cdot j} = \sum_{k = 1}^n a_{kj}, \quad a_{\cdot \cdot} = \sum_{k,l = 1}^n a_{kl} \quad b_{i \cdot} = \sum_{l = 1}^n b_{il}, \quad b_{\cdot j} = \sum_{k = 1}^n b_{kj}, \quad b_{\cdot \cdot} = \sum_{k,l = 1}^n b_{kl}$$
+   -  column sum, row sum, and overall sum: $$a_{i \cdot} = \sum_{l = 1}^n a_{il}, \quad a_{\cdot j} = \sum_{k = 1}^n a_{kj}, \quad a_{\cdot \cdot} = \sum_{k,l = 1}^n a_{kl} $$
 
-  -   distance covariance (and variance defined similarly): $$\text{dCov}(a_{ij}, b_{ij}) = \frac{1}{n(n-3)} \sum_{i \neq j} a_{ij} b_{ij} -\frac{2}{n (n-2)(n-3)}\sum_{i = 1}^n a_{i \cdot} b_{i \cdot} + \frac{1}{n(n - 1)(n-2)(n-3)} a_{\cdot \cdot} b_{\cdot \cdot}$$
+   $$\quad b_{i \cdot} = \sum_{l = 1}^n b_{il}, \quad b_{\cdot j} = \sum_{k = 1}^n b_{kj}, \quad b_{\cdot \cdot} = \sum_{k,l = 1}^n b_{kl}$$
+
+  -   distance covariance (and variance defined similarly): $$\text{dCov}(a_{ij}, b_{ij}) = \frac{1}{n(n-3)} \sum_{i \neq j} a_{ij} b_{ij}  -\frac{2 \sum_{i = 1}^n a_{i \cdot} b_{i \cdot}}{n (n-2)(n-3)} + \frac{a_{\cdot \cdot} b_{\cdot \cdot}}{n(n - 1)(n-2)(n-3)}$$
 
@@
 $$I_{stringy}(A) = \frac{\text{number of vertices with 2 edges}}{\text{number of total vertices with more than one edge}} $$
+
+# Reference

### 2024-09-11 supplementary/code-sec-practicalities.R

@@ -1,0 +1,41 @@
+library(palmerpenguins)
+library(tidyverse)
+library(tourr)
+library(ferrn)
+library(GpGp)
+library(fields)
+################################################################################
+# Using the JSO in a PPGT
+# clean up the penguins data
+penguins_cl <- palmerpenguins::penguins |>
+  filter(!is.na(bill_length_mm)) |>
+  rename(bl = bill_length_mm, bd = bill_depth_mm,
+         fl = flipper_length_mm, bm = body_mass_g)
+penguins_df <- penguins_cl |> select(bl:bm) |> rescale()
+# run the projection pursuit with JSO
+res <- animate_xy(penguins_df,
+                  guided_tour(lda_pp(cl = penguins_cl$species),
+                              search_f = search_jellyfish,
+                              n_jellies = 20, max.tries = 50))
+
+# extract the bases from the first jellyfish
+bases <- res |> filter(loop == 1) |> pull(basis) |> check_dup(0.1)
+# visualise its path with a planned tour
+animate_xy(data = penguins_df, tour_path = planned_tour(bases),
+           col = penguins_cl$species)
+
+################################################################################
+# calculate smoothness and squintability
+# smoothness
+basis_smoothness <- sample_bases(
+  idx = "holes", data = sine1000, n_basis = 300,
+  best = matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1), nrow = 6))
+basis_smoothness |> head(3)
+calc_smoothness(basis_smoothness)
+
+# squintability
+basis_squint <- sample_bases(
+  idx = "holes", data = sine1000, n_basis = 100,
+  best = matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1), nrow = 6),
+  step_size = 0.1, min_proj_dist = 1.5)
+calc_squintability(basis_squint, method = "nls", bin_width = 0.01)

### 2024-09-11 jso.qmd
small changes to table 1

@@
 
-<!-- **Projection distance**: The projection distance between projection -->
-
-<!-- bases and the optimal best basis (or the empirical best basis if -->
-
-<!-- the optimal is unknown) measures how close a visited basis is to  -->
-
-<!-- the best basis and it is calculated as the Frobenius norm in equation -->
-
-<!-- (@eq-proj-dist). Plotting projection distances against index values for the  -->
-
-<!-- bases visited reveals how the index value changes as the basis -->
-
-<!-- approaches the optimal basis, helping to calculate the squintability -->
-
-<!-- measure. -->
-
 ```{r}
@@
 fitting a generalised linear model with a binomial family and a logit
-link function.<!--, where all the three
-components (success rate, JS hyper-parameters, and index properties) are
-combined.--> The model suggests that JSO success rate is positively
+link function. The model suggests that JSO success rate is positively
 associated with the two hyper-parameters, as well as with the index
@@
   rename(d = n, smooth = smoothness) |> 
-  mutate(index = ifelse(index == "dcor2d_2", "dcor2d", index)) 
+  mutate(index = ifelse(index == "dcor2d_2", "dcor", index),
+         index = ifelse(index == "loess2d", "loess", index),
+         index = ifelse(index == "splines2d", "splines", index),
+         variance = sqrt(variance)) 
 dt |> 
@@
       "$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", "$\\varsigma$"), linesep = "",
-    booktabs = T, escape = FALSE, row.names = TRUE) |> 
+    booktabs = TRUE, escape = FALSE, row.names = TRUE, align = "c") |> 
   column_spec(1, border_left = TRUE) |> 
@@
 
-<!-- 
-```{r}
-#| label: tbl-mod-data
-#| tbl-cap: Data constructed to store simulation results (the first 7 rows) includes name of index, dimension, index metrics, JSO parameters, and performance metrics. 
-sim_df |> 
-  dplyr::select(index, d, smoothness, 
-    squintability, n_jellies, max_tries, 
-    -I_max, P_J, time) |> 
-  mutate(time = as.numeric(time)) |>
-  arrange(index, d, n_jellies, max_tries) |> 
-  head(7) |> 
-  kable(digits = 2, format = "latex", 
-        col.names = c("index", "d", 
-          "smoothness", "squintability", 
-          "n. jellyfish", 
-          "max. tries", 
-          "success rate", "time (sec)"),
-        linesep = "",
-        booktabs = T) |> 
-  kable_styling(#latex_options = "hold_position",
-    font_size = 10)
-```
--->
-
 ```{r}
 #| label: tbl-mod-output
+#| results: "asis"
 #| tbl-cap: Model estimates of proportion of jellyfish success on index properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function.

### 2024-09-11 jso.tex
small changes to table 1

@@
 
-\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, colback=white, opacitybacktitle=0.6, left=2mm, titlerule=0mm, toprule=.15mm, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, colframe=quarto-callout-note-color-frame, opacityback=0, breakable, toptitle=1mm, rightrule=.15mm, bottomrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, rightrule=.15mm, bottomrule=.15mm, bottomtitle=1mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacitybacktitle=0.6, breakable, arc=.35mm, titlerule=0mm, toprule=.15mm, opacityback=0, left=2mm, colframe=quarto-callout-note-color-frame, colback=white]
 
@@
 
-\begin{tabular}{|>{}lll>{}r|rrr>{}r|rrrr>{}r|}
+\begin{tabular}{|>{}lcc>{}c|ccc>{}c|cccc>{}c|}
 \toprule
@@
 \midrule
-1 & pipe & holes & 6 & 0.00 & 0.37 & 2.36 & 0.22 & 1.02 & 0.86 & 3.37 & 0.02 & 3.05\\
-2 & pipe & holes & 8 & 0.00 & 0.18 & 2.19 & 0.82 & 1.01 & 0.87 & 3.26 & 0.03 & 2.96\\
-3 & pipe & holes & 10 & 0.00 & 0.11 & 2.19 & 3.03 & 1.02 & 0.88 & 3.15 & 0.02 & 2.95\\
-4 & pipe & holes & 12 & 0.00 & 0.15 & 2.29 & 1.58 & 1.01 & 0.88 & 3.34 & 0.00 & 3.12\\
-5 & sine & MIC & 6 & 0.02 & 0.09 & 2.46 & 0.08 & 0.89 & 0.57 & 1.62 & -0.02 & 1.26\\
-6 & sine & MIC & 8 & 0.02 & 0.08 & 2.64 & 0.08 & 0.93 & 0.33 & 1.31 & -0.03 & 0.77\\
-7 & sine & TIC & 6 & 0.12 & 0.11 & 2.44 & 0.09 & 0.95 & 0.54 & 1.72 & -0.03 & 1.32\\
-8 & sine & TIC & 8 & 0.12 & 0.10 & 2.47 & 0.09 & 0.95 & 0.56 & 1.72 & -0.03 & 1.37\\
-9 & sine & dcor2d & 6 & 0.03 & 0.17 & 2.66 & 0.11 & 0.95 & 1.04 & 2.74 & -0.02 & 2.95\\
-10 & sine & loess2d & 6 & 0.08 & 0.34 & 1.99 & 0.31 & 1.02 & 1.04 & 2.65 & 0.08 & 2.76\\
-11 & sine & splines2d & 6 & 0.04 & 0.24 & 2.54 & 0.10 & 1.01 & 1.05 & 2.73 & -0.01 & 3.12\\
+1 & pipe & holes & 6 & 0.04 & 0.37 & 2.36 & 0.22 & 1.02 & 0.86 & 3.37 & 0.02 & 3.05\\
+2 & pipe & holes & 8 & 0.01 & 0.18 & 2.19 & 0.82 & 1.01 & 0.87 & 3.26 & 0.03 & 2.96\\
+3 & pipe & holes & 10 & 0.01 & 0.11 & 2.19 & 3.03 & 1.02 & 0.88 & 3.15 & 0.02 & 2.95\\
+4 & pipe & holes & 12 & 0.01 & 0.15 & 2.29 & 1.58 & 1.01 & 0.88 & 3.34 & 0.00 & 3.12\\
+5 & sine & MIC & 6 & 0.13 & 0.09 & 2.46 & 0.08 & 0.89 & 0.57 & 1.62 & -0.02 & 1.26\\
+6 & sine & MIC & 8 & 0.13 & 0.08 & 2.64 & 0.08 & 0.93 & 0.33 & 1.31 & -0.03 & 0.77\\
+7 & sine & TIC & 6 & 0.35 & 0.11 & 2.44 & 0.09 & 0.95 & 0.54 & 1.72 & -0.03 & 1.32\\
+8 & sine & TIC & 8 & 0.35 & 0.10 & 2.47 & 0.09 & 0.95 & 0.56 & 1.72 & -0.03 & 1.37\\
+9 & sine & dcor & 6 & 0.18 & 0.17 & 2.66 & 0.11 & 0.95 & 1.04 & 2.74 & -0.02 & 2.95\\
+10 & sine & loess & 6 & 0.28 & 0.34 & 1.99 & 0.31 & 1.02 & 1.04 & 2.65 & 0.08 & 2.76\\
+11 & sine & splines & 6 & 0.20 & 0.24 & 2.54 & 0.10 & 1.01 & 1.05 & 2.73 & -0.01 & 3.12\\
 12 & sine & stringy & 6 & 0.00 & 0.01 & 1.54 & 38.17 & 1.05 & 0.01 & 254.73 & 0.05 & 2.96\\
@@
 \url{https://github.com/huizezhang-sherry/paper-jso}. The simulation
-data produced as described in Section 5 in can be found at
+data produced in Section 5 can be found at
 \url{https://figshare.com/articles/dataset/Simulated_raw_data/26039506}.

### 2024-09-12 jso.qmd
add uncertainty accessment for figure 8

@@
 of jellyfish and the maximum number of tries) is presented in
-@fig-proportion. As the number of jellyfish and maximum tries increase,
-the success rate also increases. For simpler problems (6 dimensions),
+@fig-proportion.
+The uncertainty is quantified through 500 bootstrap replicates for each case.
+As the number of jellyfish and maximum tries increase, the success rate also increases.
+For simpler problems (6 dimensions),
 small parameter values (20 jellyfishes and a maximum of 50 tries) can
@@
 #| label = fig-proportion,
-#| fig.width = 6,
-#| fig.height = 2.5,
+#| fig.width = 8,
+#| fig.height = 3.5,
 #| fig.align = "center",
-#| fig.cap = "Proportion of simulations reaches near-optimal index values in the pipe-finding problem using the holes index.
+#| fig.cap = "The proportion of simulations that reach near-optimal index values in the pipe-finding problem using the holes index.
 The proportion is calculated based on the number of simulations, out of 50, that achieve an index value within 0.05 of the best-performing simulation.
+To quantify uncertainty, Bootstrap samples with 500 are generated.
+The thin lines represent the proportion for each of the 500 bootstrap replicates, while the thicker lines represent the mean of these bootstrap samples, connnected by lines.
 As the dimensionality increases, the proportion of simulations reaching the optimal index value increases."
@@
 
-pipe_res |> 
-  rename(`max # of tries` = max_tries) |>
-  ggplot(aes(x = n_jellies, y = proportion, 
-             group = interaction(d, `max # of tries`),
-             color = d)) + 
-  geom_line() + 
+res_df <- sim_pipe_run_best |> 
+  left_join(pipe_sim_best |> ungroup() |> mutate(case_id = row_number()),
+            by = c("n_jellies", "max_tries", "d")) |> 
+  ungroup() |> 
+  group_split(case_id)
+
+resampling <- function(dt, size = size, replace = TRUE){
+  row_id <- sample(1:50, size = size, replace = replace)
+
+  dt |> filter(row_number() %in% row_id) |> 
+    mutate(diff = abs(I_max - I_best) < 0.05) |> 
+    summarise(proportion = sum(diff) / n()) |> 
+    pull(proportion)
+
+}
+
+res_interval_df <- tibble(df = res_df) |> 
+  crossing(rep = 1:500) |> 
+  rowwise() |> 
+  mutate(proportion = resampling(df, size = 50))
+
+res_quantile <- res_interval_df |> 
+  unnest(df) |> 
+  group_by(n_jellies, max_tries, case_id, d) |> 
+  select(-I_max, -id) |> 
+  distinct() |> 
+  mutate(d = as.factor(d),
+         n_jellies = ifelse(d == 10, n_jellies - 1, n_jellies),
+         n_jellies = ifelse(d == 12, n_jellies + 1, n_jellies)) |> 
+  rename(`max # of tries` = max_tries)
+
+res_quantile |> 
+  ungroup() |> 
+  ggplot(aes(x = n_jellies - 0.5, y = proportion, 
+             group = d, color = d)) + 
+  geom_segment(aes(xend = n_jellies + 0.5), size = 0.05) + 
+  geom_segment(aes(x = n_jellies - 1, xend = n_jellies + 1), 
+               data = res_quantile |> summarise(proportion = mean(proportion)), 
+               size = 2) + 
+  geom_line(aes(x = n_jellies), 
+               data = res_quantile |> summarise(proportion = mean(proportion)), 
+            alpha = 0.3) + 
   scale_color_brewer(palette = "Dark2", name = "dimension") +
+  scale_x_continuous(breaks = c(20, 50, 100), labels = c("20", "50", "100")) +
   facet_wrap(vars(`max # of tries`), 
@@
   theme_bw() +
+  theme(panel.grid.minor = element_blank()) + 
   xlab("Number of jellyfish") + 
   ylab("Proportion of success")
-
 ```

### 2024-09-12 jso.tex
add uncertainty accessment for figure 8

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, leftrule=.75mm, rightrule=.15mm, bottomrule=.15mm, bottomtitle=1mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacitybacktitle=0.6, breakable, arc=.35mm, titlerule=0mm, toprule=.15mm, opacityback=0, left=2mm, colframe=quarto-callout-note-color-frame, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, opacityback=0, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, left=2mm, arc=.35mm, breakable, bottomrule=.15mm, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, opacitybacktitle=0.6, toprule=.15mm, titlerule=0mm, leftrule=.75mm]
 
@@
 of jellyfish and the maximum number of tries) is presented in
-Figure~\ref{fig-proportion}. As the number of jellyfish and maximum
-tries increase, the success rate also increases. For simpler problems (6
-dimensions), small parameter values (20 jellyfishes and a maximum of 50
-tries) can already achieve a high success rate. However, larger
-parameter values (i.e.~100 jellyfishes and a maximum of 100 tries) are
-necessary for higher-dimensional problems (8, 10, and 12 dimensions).
-Increasing both parameters enhances the performance of JSO, but it also
-extends the computational time required for the optimisation, which can
-be computationally intensive when evaluating the index function (such as
-scagnostic indexes) multiple times across numerous iterations.
+Figure~\ref{fig-proportion}. The uncertainty is quantified through 500
+bootstrap replicates for each case. As the number of jellyfish and
+maximum tries increase, the success rate also increases. For simpler
+problems (6 dimensions), small parameter values (20 jellyfishes and a
+maximum of 50 tries) can already achieve a high success rate. However,
+larger parameter values (i.e.~100 jellyfishes and a maximum of 100
+tries) are necessary for higher-dimensional problems (8, 10, and 12
+dimensions). Increasing both parameters enhances the performance of JSO,
+but it also extends the computational time required for the
+optimisation, which can be computationally intensive when evaluating the
+index function (such as scagnostic indexes) multiple times across
+numerous iterations.
 
@@
 
-\caption{\label{fig-proportion}Proportion of simulations reaches
+\caption{\label{fig-proportion}The proportion of simulations that reach
 near-optimal index values in the pipe-finding problem using the holes
@@
 out of 50, that achieve an index value within 0.05 of the
-best-performing simulation. As the dimensionality increases, the
-proportion of simulations reaching the optimal index value increases.}
+best-performing simulation. To quantify uncertainty, Bootstrap samples
+with 500 are generated. The thin lines represent the proportion for each
+of the 500 bootstrap replicates, while the thicker lines represent the
+mean of these bootstrap samples, connnected by lines. As the
+dimensionality increases, the proportion of simulations reaching the
+optimal index value increases.}
 

### 2024-09-12 jso.qmd
use the ferrn pkg to make the huber plot

@@
 ```{r}
-#| label: huber-plot-function
-#| echo: false
-huber_plot <- function (origdata2D, origclass=NULL, 
-  PPmethod = "LDA", weight = TRUE, r = 1, lambda = 0.5, 
-  opt.proj = TRUE, UserDefFtn = NULL, ...) 
-{
-    index <- NULL
-    best.proj <- NULL
-    best.index <- 0
-    origdata2D <- as.matrix(origdata2D)
-    for (i in 0:360) {
-        theta <- pi/180 * i
-        proj.data <- matrix(cos(theta) * origdata2D[, 1] + sin(theta) * 
-            origdata2D[, 2])
-        proj <- matrix(c(cos(theta), sin(theta)), ncol = 1)
-        if (PPmethod == "LDA") {
-            newindex <- PPtreeViz::LDAindex(origclass, origdata2D, proj = proj, 
-                weight = weight)
-        }
-        else if (PPmethod == "PDA") {
-            newindex <- PPtreeViz::PDAindex(origclass, origdata2D, proj, 
-                weight = weight, lambda = lambda)
-        }
-        else if (PPmethod == "Lr") {
-            newindex <- PPtreeViz::Lrindex(origclass, origdata2D, proj, 
-                weight = weight, r = r)
-        }
-        else if (PPmethod == "GINI") {
-            newindex <- PPtreeViz::GINIindex1D(origclass, origdata2D, proj)
-        }
-        else if (PPmethod == "ENTROPY") {
-            newindex <- PPtreeViz::ENTROPYindex1D(origclass, origdata2D, 
-                proj)
-        }
-        else if (PPmethod == "holes") {
-          newindex <- tourr::holes()(proj.data)
-        }
-        else if (PPmethod == "skewness") {
-          newindex <- tourr::skewness()(proj.data)
-        }
-        else if (PPmethod == "norm_bin") {
-          newindex <- tourr::norm_bin(nrow(proj.data))(proj.data)
-        }
-        else if (PPmethod == "UserDef") {
-            newindex <- UserDefFtn(proj.data, ...)
-        }
-        index <- c(index, newindex)
-    }
-    sel.index <- which(index[1:360] > signif(max(index), 6) - 
-        1e-06)
-    theta.best.all <- pi/180 * (sel.index - 1)
-    theta.best <- theta.best.all[1]
-    proj.data.best <- matrix(cos(theta.best) * origdata2D[, 1] + 
-        sin(theta.best) * origdata2D[, 2])
-    index.best <- max(index)
-    range <- round(max(index) - min(index), 5)
-    if (range == 0) {
-      PPindex <- rep(4, length(index))
-    }
-    else {
-      PPindex <- (index - min(index))/range * 2 + 3
-    }
-    data.circle <- NULL
-    data.index <- NULL
-    for (i in 1:361) {
-        theta <- pi/180 * (i - 1)
-        data.index <- rbind(data.index, c(PPindex[i] * cos(theta), 
-            PPindex[i] * sin(theta)))
-        data.circle <- rbind(data.circle, c(4 * cos(theta), 4 * 
-            sin(theta)))
-    }
-    maxdiff <- max(c(diff(range(origdata2D[, 1])), diff(range(origdata2D[, 
-        2]))))
-    orig.scaled <- apply(origdata2D, 2, function(x) (x - mean(x))/maxdiff * 
-        3.5)
-    data.cX <- data.circle[, 1]
-    data.cY <- data.circle[, 2]
-    data.X <- data.index[, 1]
-    data.Y <- data.index[, 2]
-    plot.data <- data.frame(data.cX, data.cY, data.X, data.Y)
-    x <- orig.scaled[, 1]
-    y <- orig.scaled[, 2]
-    if (is.null(origclass)) 
-      group <- rep(1, nrow(origdata2D))
-    else
-      group <- origclass
-    point.data <- data.frame(x, y, group)
-    min.X <- min(unlist(plot.data))
-    max.X <- max(unlist(plot.data))
-    if (is.null(origclass)) {
-      P1 <- ggplot(data = plot.data, aes(x = data.X, y = data.Y)) + 
-        geom_path() + geom_path(aes(x = data.cX, y = data.cY), 
-        linetype = "dashed") + geom_point(data = point.data, 
-        aes(x = x, y = y)) + scale_x_continuous(breaks = NULL) + 
-        scale_y_continuous(breaks = NULL) + xlab("") + ylab("") + 
-        coord_fixed() + theme_bw() + theme(panel.border = element_blank())
-    } else {
-      P1 <- ggplot(data = plot.data, aes(x = data.X, y = data.Y)) + 
-        geom_path() + geom_path(aes(x = data.cX, y = data.cY), 
-        linetype = "dashed") + geom_point(data = point.data, 
-        aes(x = x, y = y, color = group, shape = group)) +
-        scale_x_continuous(breaks = NULL) + 
-        scale_y_continuous(breaks = NULL) + xlab("") + ylab("") + 
-        coord_fixed() + theme_bw() + theme(panel.border = element_blank())
-    }
-    if (opt.proj) {
-        P1 <- P1 + geom_abline(intercept = 0, slope = sin(theta.best)/cos(theta.best))
-        if (length(theta.best.all) > 1) 
-            for (i in 2:length(theta.best.all)) P1 <- P1 + geom_abline(intercept = 0, 
-                slope = sin(theta.best.all[i])/cos(theta.best.all[i]), 
-                linetype = "dashed")
-    }
-    best.proj.data <- proj.data.best
-    #group <- origclass
-    if (is.null(origclass)) {
-      hist.data <- data.frame(best.proj.data)
-      P2 <- ggplot(data = hist.data, aes(x = best.proj.data)) + 
-        geom_histogram()
-    } else {
-      hist.data <- data.frame(best.proj.data, group)
-      P2 <- ggplot(data = hist.data, aes(x = best.proj.data, 
-                                         group = group)) + 
-        geom_histogram(aes(fill = group), position = "stack")
-    }
-    #gridExtra::grid.arrange(P1, P2, nrow = 1)
-    return(list(plot=P1, proj_data=best.proj.data))
-}
-
-```
-
-```{r}
 #| label: fig-example-functions
@@
 A Huber plot shows the PP index values for all 1D data projections in polar coordinates."
-data(randu)
-randu_std <- as.data.frame(apply(randu, 2, function(x) (x-mean(x))/sd(x)))
-randu_std$yz <- sqrt(35)/6*randu_std$y-randu_std$z/6
-
 #proj <- animate_xy(flea[,1:6], guided_tour(lda_pp(flea$species)))
@@
                       0.45123453,  0.05896646), ncol=2, byrow=T)
+
 flea2D <- as.matrix(flea[,1:6]) %*% best_proj
@@
 flea2D <- as.data.frame(flea2D)
-
-# Make both plots
-large_sqa <- huber_plot(flea2D, PPmethod="skewness")
-small_sqa <- huber_plot(randu_std[c(1,4)], PPmethod="norm_bin")
-
-sqa1 <- large_sqa$plot + ggtitle("(a) Skewness index")
-lsqa_d <- tibble(d = large_sqa$proj_data[,1])
-sqa2 <- ggplot(lsqa_d, aes(x=d)) + 
+flea_huber <- prep_huber(flea2D, index = skewness())
+sqa1 <- ggplot() +
+  geom_huber(data = flea_huber$idx_df, aes(x = x, y = y)) +
+  geom_point(data = flea2D, aes(x = X1, y = X2)) +
+  geom_abline(slope = flea_huber$slope, intercept = 0) + 
+  coord_fixed() +
+  theme_huber() +
+  ggtitle("(a) Skewness index")
+
+sqa2 <- ggplot(flea_huber$proj_df, aes(x=x)) + 
   geom_histogram(binwidth=0.25, colour="white") +
@@
   theme(axis.text.y = element_blank())
-sqa3 <- small_sqa$plot + ggtitle("(c) Binned normality index")
-ssqa_d <- tibble(d = small_sqa$proj_data[,1])
-sqa4 = ggplot(ssqa_d, aes(x=d)) + 
+
+data(randu)
+randu_std <- as.data.frame(apply(randu, 2, function(x) (x-mean(x))/sd(x)))
+randu_std$yz <- sqrt(35)/6*randu_std$y-randu_std$z/6
+randu_df <- randu_std[c(1,4)]
+randu_huber <- prep_huber(randu_df, index = norm_bin(nr = nrow(randu_df)))
+
+sqa3 <- ggplot() +
+  geom_huber(data = randu_huber$idx_df, aes(x = x, y = y)) +
+  geom_point(data = randu_df, aes(x = x, y = yz)) +
+  geom_abline(slope = randu_huber$slope, intercept = 0) +
+  theme_huber() +
+  coord_fixed() + 
+  ggtitle("(c) Binned normality index")
+
+sqa4 <- ggplot(randu_huber$proj_df, aes(x = x)) +
   geom_histogram(breaks = seq(-2.2, 2.4, 0.12)) +
   xlab("") + ylab("") +
-  ggtitle("(d) Optimally projected data") +
   theme_bw() +
-  theme(axis.text.y = element_blank())
+  theme(axis.text.y = element_blank()) + 
+  ggtitle("(d) Optimally projected data")
 
 sqa1 + sqa2 + sqa3 + sqa4 + plot_layout(ncol=2, widths=c(2,2))
+
 ```
@@
 make JSO a promising candidate for enhancing PP optimisation.
-<!--The study reported here explores JSO's potential in enhancing projection pursuit and examines the characteristics
-and advantages of such implementation.-->
 
@@
 # Projection pursuit, tours, index functions and optimisation {#sec-background}
-
-<!-- Need to assume that the special issue will have the general introduction to PP, so the focus here will be on background needed for this paper, tours, guidance using PP, optimisation, visualisation, ...
--->
 

### 2024-09-12 jso.tex
use the ferrn pkg to make the huber plot

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, opacityback=0, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, left=2mm, arc=.35mm, breakable, bottomrule=.15mm, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, opacitybacktitle=0.6, toprule=.15mm, titlerule=0mm, leftrule=.75mm]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, opacityback=0, left=2mm, toprule=.15mm, breakable, leftrule=.75mm, toptitle=1mm, bottomtitle=1mm, titlerule=0mm, coltitle=black, arc=.35mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, bottomrule=.15mm]
 

### 2024-09-12 jso.qmd
use stargazer for model output

@@
    - \usepackage{algorithm}
+   - \usepackage{float} 
    - \usepackage{amsmath}
@@
          long_time = ifelse(time > 20, 1, 0),
-         P_J = ifelse(index == "stringy", 0, P_J))
-
-mod1 <- glm(P_J ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
+         P_J = ifelse(index == "stringy", 0, P_J)) |> 
+  rename(proportion = P_J)
+
+mod1 <- glm(proportion ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
 ```
@@
 ```{r}
-#| label: tbl-mod-output
 #| results: "asis"
-#| tbl-cap: Model estimates of proportion of jellyfish success on index properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function.
-The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with JSO success rate while data dimension and being flagged as long runtime are negatively associated with the success rate.
+stargazer::stargazer(mod1, header = FALSE, label = "tbl-mod-output",
+                     table.placement = "H", 
+                     title = "Model estimates of proportion of jellyfish success on index properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function.The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with JSO success rate while data dimension and being flagged as long runtime are negatively associated with the success rate.
 The variable squintability and dimension are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.
-broom::tidy(mod1) |> 
-  mutate(term = c("Intercept", "Smoothness",
-   "Squintability", "Dimension (d)", 
-   "Long time", "N. jellyfish", 
-   "Max. tries")) |>
-  kable(digits = 2, format = "latex", 
-    linesep = "", booktabs = T) |> 
-  column_spec(c(1,5), border_right = TRUE) |> 
-  column_spec(1, border_left = TRUE) |> 
-  kable_styling(font_size = 10) 
+")
 ```
@@
 For the non-parametric case, it shows the (`max_d`), the corresponding projection distance (`max_dist`), and teh squintability metric as their products.
-
-<!--
-This section discusses the insights in implementation yielded from the previous sections to guide the usage of projection pursuit guided tour amongst practitioners.
-Smoothness and squintability in a projection pursuit problem can now be measured given the target shape, the chosen index function, and the data dimension.
-In practice, these metrics can be used to characterise the difficulty of the optimisation task and inform the choice of optimiser.
-For instance, a large squintability value indicates a distinct difference in index values between optimal regions and others.
-When one jellyfish enters the optimal region, the high index value it generates will be clearly distinguishable from spurious values due to this pronounced difference and lead other jellyfish to move toward the optimal region.
-In contrast, a smaller squintability value results in a less distinct separation between desirable and undesirable regions, making it more difficult for other jellyfish to accurately navigate toward the target area.
-In addition, a larger smoothness value implies a smoother index function, which suggests the use of both gradient-based and random sampling optimisers.
-
-This study enables comparison of new indices with existing indices with the proposed smoothness and squintability measures, as well as benchmarking with existing well-known data sets.
-Note that when comparing between multiple optimisers, the smoothness and squintability metrics proposed in this paper are relative.
-Comparison should be made with common parameters, i.e.
-the same seed for generating random bases and the same binning width for squintability.
-Furthermore, the range of the index values should be standardised to the [0, 1] interval as demonstrated for the `holes` and `TIC` indexes.
-
-As a population-based metaheuristic method, JSO requires evaluating multiple jellyfish in each iteration, which increased its computational complexity compared to single-candidate random sampling methods.
-This additional computational costs should be taken into consideration when choosing optimisers.
-Tuning JSO hyperparameters, such as the number of jellyfish and the maximum number of attempts, can enhance the likelihood of finding the optimal region at the expense of computational costs.
-However, in some cases, increasing these parameters may not be effective if the index surface is complex or has low squintability.
-
-The new JSO is integrated with the current implementation of the projection pursuit guided tour in the `tourr` package, and can be further examined using PP optimisation diagnostics in the `ferrn` package.
-The recommended approach is to conduct the optimisation off-line, extract the bases of a selected jellyfish, and then use the planned tour to follow selected jellyfish.
-These tools are also available in the `tourr` package.
--->
 

### 2024-09-12 jso.tex
use stargazer for model output

@@
 \usepackage{algorithm}
+\usepackage{float}
 \usepackage{amsmath}
@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, opacityback=0, left=2mm, toprule=.15mm, breakable, leftrule=.75mm, toptitle=1mm, bottomtitle=1mm, titlerule=0mm, coltitle=black, arc=.35mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, bottomrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, colback=white, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, rightrule=.15mm, breakable, left=2mm, coltitle=black, bottomrule=.15mm, bottomtitle=1mm, arc=.35mm, opacityback=0, opacitybacktitle=0.6]
 
@@
 
-\caption{\label{tbl-mod-output}Model estimates of proportion of
-jellyfish success on index properties and jellyfish hyper-parameters
-from the generalised linear model with a binomial family and a logit
-link function. The variable smoothness, squintability, number of
-jellyfish and maximum number of tries are positively associated with JSO
-success rate while data dimension and being flagged as long runtime are
-negatively associated with the success rate. The variable squintability
-and dimension are significant, suggesting their importance relative to
-jellyfish hyper-parameters in the optimisation success.}
+\caption{\label{tbl-mod-output}Model estimates of proportion of jellyfish success on index properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function.The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with JSO success rate while data dimension and being flagged as long runtime are negatively associated with the success rate.
+The variable squintability and dimension are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.
+}
 
@@
 
-\centering\begingroup\fontsize{10}{12}\selectfont
-
-\begin{tabular}{|>{}lrrr>{}r|}
-\toprule
-term & estimate & std.error & statistic & p.value\\
-\midrule
-Intercept & -4.52 & 5.33 & -0.85 & 0.40\\
-Smoothness & 1.19 & 1.92 & 0.62 & 0.53\\
-Squintability & 2.06 & 0.68 & 3.01 & 0.00\\
-Dimension (d) & -0.63 & 0.26 & -2.46 & 0.01\\
-Long time & -0.87 & 1.29 & -0.68 & 0.50\\
-N. jellyfish & 0.22 & 0.13 & 1.70 & 0.09\\
-Max. tries & 0.11 & 0.15 & 0.75 & 0.45\\
-\bottomrule
+[H] \centering
+
+\begin{tabular}{@{\extracolsep{5pt}}lc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-2} 
+\\[-1.8ex] & proportion \\ 
+\hline \\[-1.8ex] 
+ smoothness & 1.195 \\ 
+  & (1.915) \\ 
+  & \\ 
+ squintability & 2.060$^{***}$ \\ 
+  & (0.684) \\ 
+  & \\ 
+ d & $-$0.628$^{**}$ \\ 
+  & (0.255) \\ 
+  & \\ 
+ long\_time & $-$0.874 \\ 
+  & (1.286) \\ 
+  & \\ 
+ n\_jellies & 0.216$^{*}$ \\ 
+  & (0.127) \\ 
+  & \\ 
+ max\_tries & 0.113 \\ 
+  & (0.150) \\ 
+  & \\ 
+ Constant & $-$4.523 \\ 
+  & (5.332) \\ 
+  & \\ 
+\hline \\[-1.8ex] 
+Observations & 52 \\ 
+Log Likelihood & $-$18.224 \\ 
+Akaike Inf. Crit. & 50.447 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 \end{tabular}
-\endgroup{}
 

### 2024-09-12 jso.qmd
use better covariate namesg

@@
 #| results: "asis"
-stargazer::stargazer(mod1, header = FALSE, label = "tbl-mod-output",
+stargazer::stargazer(
+  mod1, header = FALSE, label = "tbl-mod-output",
   table.placement = "H", 
+  covariate.labels = c("Smoothness", "Squintability", "Dimension", "Long runtime", "Number of jellyfish", "Maximum number of tries"),
   title = "Model estimates of proportion of jellyfish success on index properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function.The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with JSO success rate while data dimension and being flagged as long runtime are negatively associated with the success rate.

### 2024-09-12 jso.tex
use better covariate namesg

@@
 
-\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, colback=white, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, rightrule=.15mm, breakable, left=2mm, coltitle=black, bottomrule=.15mm, bottomtitle=1mm, arc=.35mm, opacityback=0, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, bottomtitle=1mm, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, breakable, toptitle=1mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm, left=2mm, bottomrule=.15mm, rightrule=.15mm, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, titlerule=0mm, opacityback=0]
 
@@
 
-[H] \centering
+\centering
 
@@
 \hline \\[-1.8ex]
- smoothness & 1.195 \\ 
+ Smoothness & 1.195 \\
   & (1.915) \\
   & \\
- squintability & 2.060$^{***}$ \\ 
+ Squintability & 2.060$^{***}$ \\
   & (0.684) \\
   & \\
- d & $-$0.628$^{**}$ \\ 
+ Dimension & $-$0.628$^{**}$ \\
   & (0.255) \\
   & \\
- long\_time & $-$0.874 \\ 
+ Long runtime & $-$0.874 \\
   & (1.286) \\
   & \\
- n\_jellies & 0.216$^{*}$ \\ 
+ Number of jellyfish & 0.216$^{*}$ \\
   & (0.127) \\
   & \\
- max\_tries & 0.113 \\ 
+ Maximum number of tries & 0.113 \\
   & (0.150) \\

### 2024-09-12 jso.qmd
small edits on the Practicalities section

@@
 outlined in @sec-smoothness and @sec-squintability and as illustrated in
-@fig-smoothness and @fig-squintability.
-
-To compute smoothness, 300 random bases are simulated. Index values and
+@fig-smoothness and @fig-squintability. To compute smoothness, 300 random
+bases are simulated. Index values and
 projection distance (to the optimal basis) are calculated for each
@@
 
-The bases visited by jellyfish No.
-1 is extracted with `filter(loop = 1)` (the variable `loop` in the data object is the jellyfish ID) and the `check_dup()` function removes the bases in the iteration without improvements.
-The bases are then supplied as a `planned_tour()` in `animate_xy()` to visualise the tour path of jellyfish No.1.
-
 ## Computing the metrics for your new index
@@
 
-To sample bases for calculating smoothness, the required arguments include `idx` (the index function used), `data` (the dataset), `n_basis` (the number of random bases to sample), and `best` (the best basis to compute projection distance).
-The output of `sample_bases()` is a data frame with four columns: `id` (the basis ID), `basis` (the basis matrix sampled), `dist` (the projection distance to the best basis), and `index` (the index value).
-To speed up the computation of index values,  parallellasation is available through the `parallel = TRUE` argument.
+To sample bases for calculating smoothness, four arguments are required: 1) the index function, the dataset, the number of random bases to sample, and the best basis to compute projection distance.
+The output of `sample_bases()` is a data frame containing columns/ list-columns of sampled basis matrix, calculated projection distance and index value.
+Parallellasation is available through the `parallel = TRUE` argument to speed up the computation.
 
 The `calc_smoothness()` function takes the output from `sample_bases()` and fits a Gaussian process model to the index values against the projection distances to compute the smoothness metric.
-The starting parameter for the fit can be specified using the `start_params` argument, and additional Gaussian process arguments can be provided through `other_gp_params`.
-The smoothness metric is the smoothness parameter from the Gussian process and all details of the fit can be accessed using `attr(sm_res, "fit_res")`.
+Starting parameters and additional Gaussian process arguments can be specified and details of the fit can be accessed through the `fit_res` attribute of the `calc_smoothness()` output.
 
@@
 The output provides the estimated parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$) and the squintability metric for the parametric case.
-For the non-parametric case, it shows the (`max_d`), the corresponding projection distance (`max_dist`), and teh squintability metric as their products.
+For the non-parametric case, it shows the maximum gradient attained(`max_d`), the corresponding projection distance (`max_dist`), and the squintability metric as their products.
 

### 2024-09-12 jso.tex
small edits on the Practicalities section

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, bottomtitle=1mm, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, breakable, toptitle=1mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm, left=2mm, bottomrule=.15mm, rightrule=.15mm, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, titlerule=0mm, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, breakable, toprule=.15mm, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, arc=.35mm, titlerule=0mm, coltitle=black, bottomrule=.15mm, opacityback=0, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, left=2mm, colback=white, leftrule=.75mm, toptitle=1mm]
 
@@
 Section~\ref{sec-squintability} and as illustrated in
-Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}.
-
-To compute smoothness, 300 random bases are simulated. Index values and
+Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}. To
+compute smoothness, 300 random bases are simulated. Index values and
 projection distance (to the optimal basis) are calculated for each
@@
 
-\centering
+[H] \centering
 
@@
 
-The bases visited by jellyfish No.~1 is extracted with
-\texttt{filter(loop\ =\ 1)} (the variable \texttt{loop} in the data
-object is the jellyfish ID) and the \texttt{check\_dup()} function
-removes the bases in the iteration without improvements. The bases are
-then supplied as a \texttt{planned\_tour()} in \texttt{animate\_xy()} to
-visualise the tour path of jellyfish No.1.
-
 \subsection{Computing the metrics for your new
@@
 
-To sample bases for calculating smoothness, the required arguments
-include \texttt{idx} (the index function used), \texttt{data} (the
-dataset), \texttt{n\_basis} (the number of random bases to sample), and
-\texttt{best} (the best basis to compute projection distance). The
-output of \texttt{sample\_bases()} is a data frame with four columns:
-\texttt{id} (the basis ID), \texttt{basis} (the basis matrix sampled),
-\texttt{dist} (the projection distance to the best basis), and
-\texttt{index} (the index value). To speed up the computation of index
-values, parallellasation is available through the
-\texttt{parallel\ =\ TRUE} argument.
+To sample bases for calculating smoothness, four arguments are required:
+1) the index function, the dataset, the number of random bases to
+sample, and the best basis to compute projection distance. The output of
+\texttt{sample\_bases()} is a data frame containing columns/
+list-columns of sampled basis matrix, calculated projection distance and
+index value. Parallellasation is available through the
+\texttt{parallel\ =\ TRUE} argument to speed up the computation.
 
@@
 values against the projection distances to compute the smoothness
-metric. The starting parameter for the fit can be specified using the
-\texttt{start\_params} argument, and additional Gaussian process
-arguments can be provided through \texttt{other\_gp\_params}. The
-smoothness metric is the smoothness parameter from the Gussian process
-and all details of the fit can be accessed using
-\texttt{attr(sm\_res,\ "fit\_res")}.
+metric. Starting parameters and additional Gaussian process arguments
+can be specified and details of the fit can be accessed through the
+\texttt{fit\_res} attribute of the \texttt{calc\_smoothness()} output.
 
@@
 \(\theta_4\)) and the squintability metric for the parametric case. For
-the non-parametric case, it shows the (\texttt{max\_d}), the
-corresponding projection distance (\texttt{max\_dist}), and teh
-squintability metric as their products.
+the non-parametric case, it shows the maximum gradient
+attained(\texttt{max\_d}), the corresponding projection distance
+(\texttt{max\_dist}), and the squintability metric as their products.
 

### 2024-09-12 jso.qmd
fix stargazer table renderg

@@
 #| results: "asis"
+#| label: tbl-mod-output
 stargazer::stargazer(
-  mod1, header = FALSE, label = "tbl-mod-output",
+  mod1, header = FALSE, 
   table.placement = "H", 
+  float = FALSE,
   covariate.labels = c("Smoothness", "Squintability", "Dimension", "Long runtime", "Number of jellyfish", "Maximum number of tries"),

### 2024-09-12 jso.tex
fix stargazer table renderg

@@
 
-\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, breakable, toprule=.15mm, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, arc=.35mm, titlerule=0mm, coltitle=black, bottomrule=.15mm, opacityback=0, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, left=2mm, colback=white, leftrule=.75mm, toptitle=1mm]
+\begin{tcolorbox}[enhanced jigsaw, breakable, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, leftrule=.75mm, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, coltitle=black, colback=white, bottomrule=.15mm, bottomtitle=1mm, toptitle=1mm, left=2mm, titlerule=0mm, arc=.35mm, toprule=.15mm, opacitybacktitle=0.6]
 
@@
 
-\caption{\label{tbl-mod-output}Model estimates of proportion of jellyfish success on index properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function.The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with JSO success rate while data dimension and being flagged as long runtime are negatively associated with the success rate.
-The variable squintability and dimension are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.
-}
+\caption{\label{tbl-mod-output}}
 
 \centering{
-
-[H] \centering
 

### 2024-09-13 jso.qmd
modified results section and added papers code to supplementary

--- 2024-09-12 jso.qmd
@@
 
-# Simulation study {#sec-sim-deets}
+# Assessing optimisers: a simulation study {#sec-sim-deets}
 
@@
 
-The results from the first simulation described in @sec-sim-deets are
-analysed based on the final projections across two optimisers (JSO and
-CRS) and the success rate across JSO hyper-parameters. In the second
-simulation, smoothness and squintability are calculated across a
-collection of pipe-finding and sine-wave finding problems to construct
-the relationship between success rate, JSO hyper-parameters, and index
-properties.
-
+## Performance of JSO relative to CRS
+
+Performance is assessed based on the final projections across two optimisers (JSO and CRS) and the success rate across JSO hyper-parameters.
 The final projections found by the two optimisers (JSO and CRS) are
@@
 
-The success rate calculated at each hyper-parameter combination (number
-of jellyfish and the maximum number of tries) is presented in
-@fig-proportion.
+## Factors affecting JSO success rate: index properties and jellyfish hyper-parameters
+
+Smoothness and squintability are calculated across a collection of pipe-finding and sine-wave finding problems to construct the relationship between success rate, JSO hyper-parameters, and index properties.
+The success rate calculated at each hyper-parameter combination (number of jellyfish and the maximum number of tries) is presented in @fig-proportion.
 The uncertainty is quantified through 500 bootstrap replicates for each case.
 As the number of jellyfish and maximum tries increase, the success rate also increases.
-For simpler problems (6 dimensions),
-small parameter values (20 jellyfishes and a maximum of 50 tries) can
-already achieve a high success rate. However, larger parameter values
+For simpler problems (6 dimensions), small parameter values (20 jellyfishes and a maximum of 50 tries) can already achieve a high success rate.
+However, larger parameter values
 (i.e. 100 jellyfishes and a maximum of 100 tries) are necessary for
@@
 
-The index properties, including smoothness and squintability, offer
-numerical metrics to characterise the complexity of projection pursuit
-optimisation problems. @tbl-smoothness-squintability presents the
-parameters for calculating both metrics estimated from the Gaussian
-process (variance, range, smooth, and nugget) and the scaled logistic
-function ($\theta_1$ to $\theta_4$) for each case considered in the
-simulation. The column "smooth" is used as the smoothness metric and the
+The index properties, smoothness and squintability, numerically characterise the complexity of projection pursuit index, and hence potential optimisation problems.
+@tbl-smoothness-squintability presents the parameters for calculating both metrics estimated from the Gaussian process (variance, range, smooth, and nugget) and the scaled logistic function ($\theta_1$ to $\theta_4$) for each case considered in the simulation.
+The column "smooth" is used as the smoothness metric and the
 column "squint" is calculated as equation (@eq-squintability-parametric)
-as the squintability metric. @tbl-mod-output presents the results of
-fitting a generalised linear model with a binomial family and a logit
-link function. The model suggests that JSO success rate is positively
-associated with the two hyper-parameters, as well as with the index
-properties: smoothness and squintability. Specifically, using 10 more
-jellyfish and 10 more tries increases the odd ratio of success by
-`r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`% and
-`r round(100 * (exp(broom::tidy(mod1)$estimate[7])-1), 2)`%,
-respectively. However, being flagged with long runtime and an increase
-of data dimension reduce the success rate by
-`r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and
-`r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively.
-The variable `squintability` and `dimension` are significant, suggesting
-their importance relative to JSO hyper-parameters in the optimisation
-success.
+as the squintability metric.
+
+@tbl-mod-output presents the results from fitting a logistic regression model using proportion of success as the response and smoothness, squintability, dimension, running time, number of jellyfish and maximum number of tries, as predictors.
+Long runtime is a binary variable, where greater than 20 seconds is labelled 1.
+The fit suggests that JSO success rate is only affected by squintability, dimension and number of jellies.
+Interestingly, the smoothness of the index is not a problem for the JSO - this is promising!
+Running the optimisation longer and allowing more tries also does not affect success rate.
+An increase in squintability of one unit increases the success rate by `r round(exp(broom::tidy(mod1)$estimate[3]) * 100, 0)`%.
+As dimension increases by one the success rate halves.
+Increasing the number of jellies by 10, increases the succes rate by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 0)`%.
 
@@
 ```{r}
-#| results: "asis"
-stargazer::stargazer(
-  mod1, header = FALSE, label = "tbl-mod-output",
-  table.placement = "H", 
-  covariate.labels = c("Smoothness", "Squintability", "Dimension", "Long runtime", "Number of jellyfish", "Maximum number of tries"),
-  title = "Model estimates of proportion of jellyfish success on index properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function.The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with JSO success rate while data dimension and being flagged as long runtime are negatively associated with the success rate.
-The variable squintability and dimension are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.
-")
+#| label: tbl-mod-output
+#| tbl-cap: "Jellyfish success rate relative to index properties and jellyfish hyper-parameters.
+This is the summary from a logistic regression fit to smoothness, squintability, dimension, running time, number of jellyfish and maximum number of tries.
+Interestingly, squintability and dimension strongly affect jellyfish optimisation success.
+The number of jellies marginally affects success, but index smoothness, running longer and increasing the number of tries do not."
+#| fig-pos: t
+library(gtsummary)
+set_gtsummary_theme(theme_gtsummary_journal("jama"))
+mod1 |> tbl_regression(exponentiate=TRUE, 
+                       pvalue_fun = label_style_pvalue(digits = 3),
+                       label = list(smoothness = "Smoothness", 
+                                    squintability = "Squintability",
+                                    d = "Dimension", 
+                                    long_time = "Long runtime", 
+                                    n_jellies = "Number of jellyfish", 
+                                    max_tries = "Maximum number of tries")) |>
+  bold_p(t = 0.1)
 ```
@@
 
-The supplementary materials include: 1) description of all the index functions used in the simulations and 2) the script that complements the Practicalities section for using JSO in a PPGT and calculating smoothness and squintability.
+The supplementary materials available at <https://github.com/huizezhang-sherry/paper-jso> include: 1) details of the indexes used in the simulation study, 2) the script to get started with using JSO in a PPGT and calculating smoothness and squintability, as explained in @sec-discussion, and the full code to reproduce the plots and summaries in this paper.
 

### 2024-09-13 jso.tex
modified results section and added papers code to supplementary

@@
 \usepackage{xcolor}
+\usepackage{caption}
 \usepackage{orcidlink}
@@
 
-\begin{tcolorbox}[enhanced jigsaw, breakable, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, leftrule=.75mm, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, coltitle=black, colback=white, bottomrule=.15mm, bottomtitle=1mm, toptitle=1mm, left=2mm, titlerule=0mm, arc=.35mm, toprule=.15mm, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, opacitybacktitle=0.6, arc=.35mm, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, coltitle=black, bottomrule=.15mm, left=2mm, breakable, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, bottomtitle=1mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm]
 
@@
 
-\section{Simulation study}\label{sec-sim-deets}
+\section{Assessing optimisers: a simulation study}\label{sec-sim-deets}
 
@@
 
-The results from the first simulation described in
-Section~\ref{sec-sim-deets} are analysed based on the final projections
-across two optimisers (JSO and CRS) and the success rate across JSO
-hyper-parameters. In the second simulation, smoothness and squintability
-are calculated across a collection of pipe-finding and sine-wave finding
-problems to construct the relationship between success rate, JSO
-hyper-parameters, and index properties.
-
-The final projections found by the two optimisers (JSO and CRS) are
-presented in Figure~\ref{fig-proj}, broken down by 10th quantile,
-faceted by the data dimensions. In the 6-dimensional data scenario, JSO
-consistently identifies a clear pipe shape. The CRS also finds the pipe
-shape but with a wide rim, suggesting a further polish search may be
-required. With increasing dimensions, JSO may not always identify the
-pipe shape due to random sampling, but it still finds the pipe shape in
-over 50\% of cases. Compared to CRS, JSO achieves higher index values
-and clearer pipe shapes across all quantiles in data of 8, 10, 12
-dimensions, suggesting its advantage in exploring high-dimensional
-spaces.
-
-The success rate calculated at each hyper-parameter combination (number
-of jellyfish and the maximum number of tries) is presented in
-Figure~\ref{fig-proportion}. The uncertainty is quantified through 500
-bootstrap replicates for each case. As the number of jellyfish and
-maximum tries increase, the success rate also increases. For simpler
-problems (6 dimensions), small parameter values (20 jellyfishes and a
-maximum of 50 tries) can already achieve a high success rate. However,
-larger parameter values (i.e.~100 jellyfishes and a maximum of 100
-tries) are necessary for higher-dimensional problems (8, 10, and 12
-dimensions). Increasing both parameters enhances the performance of JSO,
-but it also extends the computational time required for the
-optimisation, which can be computationally intensive when evaluating the
-index function (such as scagnostic indexes) multiple times across
-numerous iterations.
+\subsection{Performance of JSO relative to
+CRS}\label{performance-of-jso-relative-to-crs}
+
+Performance is assessed based on the final projections across two
+optimisers (JSO and CRS) and the success rate across JSO
+hyper-parameters. The final projections found by the two optimisers (JSO
+and CRS) are presented in Figure~\ref{fig-proj}, broken down by 10th
+quantile, faceted by the data dimensions. In the 6-dimensional data
+scenario, JSO consistently identifies a clear pipe shape. The CRS also
+finds the pipe shape but with a wide rim, suggesting a further polish
+search may be required. With increasing dimensions, JSO may not always
+identify the pipe shape due to random sampling, but it still finds the
+pipe shape in over 50\% of cases. Compared to CRS, JSO achieves higher
+index values and clearer pipe shapes across all quantiles in data of 8,
+10, 12 dimensions, suggesting its advantage in exploring
+high-dimensional spaces.
+
+\subsection{Factors affecting JSO success rate: index properties and
+jellyfish
+hyper-parameters}\label{factors-affecting-jso-success-rate-index-properties-and-jellyfish-hyper-parameters}
+
+Smoothness and squintability are calculated across a collection of
+pipe-finding and sine-wave finding problems to construct the
+relationship between success rate, JSO hyper-parameters, and index
+properties. The success rate calculated at each hyper-parameter
+combination (number of jellyfish and the maximum number of tries) is
+presented in Figure~\ref{fig-proportion}. The uncertainty is quantified
+through 500 bootstrap replicates for each case. As the number of
+jellyfish and maximum tries increase, the success rate also increases.
+For simpler problems (6 dimensions), small parameter values (20
+jellyfishes and a maximum of 50 tries) can already achieve a high
+success rate. However, larger parameter values (i.e.~100 jellyfishes and
+a maximum of 100 tries) are necessary for higher-dimensional problems
+(8, 10, and 12 dimensions). Increasing both parameters enhances the
+performance of JSO, but it also extends the computational time required
+for the optimisation, which can be computationally intensive when
+evaluating the index function (such as scagnostic indexes) multiple
+times across numerous iterations.
 
@@
 
-The index properties, including smoothness and squintability, offer
-numerical metrics to characterise the complexity of projection pursuit
-optimisation problems. Table~\ref{tbl-smoothness-squintability} presents
-the parameters for calculating both metrics estimated from the Gaussian
-process (variance, range, smooth, and nugget) and the scaled logistic
-function (\(\theta_1\) to \(\theta_4\)) for each case considered in the
+The index properties, smoothness and squintability, numerically
+characterise the complexity of projection pursuit index, and hence
+potential optimisation problems.
+Table~\ref{tbl-smoothness-squintability} presents the parameters for
+calculating both metrics estimated from the Gaussian process (variance,
+range, smooth, and nugget) and the scaled logistic function
+(\(\theta_1\) to \(\theta_4\)) for each case considered in the
 simulation. The column ``smooth'' is used as the smoothness metric and
@@
 (\ref{eq-squintability-parametric}) as the squintability metric.
-Table~\ref{tbl-mod-output} presents the results of fitting a generalised
-linear model with a binomial family and a logit link function. The model
-suggests that JSO success rate is positively associated with the two
-hyper-parameters, as well as with the index properties: smoothness and
-squintability. Specifically, using 10 more jellyfish and 10 more tries
-increases the odd ratio of success by 24.11\% and 11.93\%, respectively.
-However, being flagged with long runtime and an increase of data
-dimension reduce the success rate by 41.72\% and 53.36\%, respectively.
-The variable \texttt{squintability} and \texttt{dimension} are
-significant, suggesting their importance relative to JSO
-hyper-parameters in the optimisation success.
+
+Table~\ref{tbl-mod-output} presents the results from fitting a logistic
+regression model using proportion of success as the response and
+smoothness, squintability, dimension, running time, number of jellyfish
+and maximum number of tries, as predictors. Long runtime is a binary
+variable, where greater than 20 seconds is labelled 1. The fit suggests
+that JSO success rate is only affected by squintability, dimension and
+number of jellies. Interestingly, the smoothness of the index is not a
+problem for the JSO - this is promising! Running the optimisation longer
+and allowing more tries also does not affect success rate. An increase
+in squintability of one unit increases the success rate by 785\%. As
+dimension increases by one the success rate halves. Increasing the
+number of jellies by 10, increases the succes rate by 24\%.
 
@@
 
-\begin{table}
-
-\caption{\label{tbl-mod-output}}
-
-\centering{
-
-\begin{tabular}{@{\extracolsep{5pt}}lc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-2} 
-\\[-1.8ex] & proportion \\ 
-\hline \\[-1.8ex] 
- Smoothness & 1.195 \\ 
-  & (1.915) \\ 
-  & \\ 
- Squintability & 2.060$^{***}$ \\ 
-  & (0.684) \\ 
-  & \\ 
- Dimension & $-$0.628$^{**}$ \\ 
-  & (0.255) \\ 
-  & \\ 
- Long runtime & $-$0.874 \\ 
-  & (1.286) \\ 
-  & \\ 
- Number of jellyfish & 0.216$^{*}$ \\ 
-  & (0.127) \\ 
-  & \\ 
- Maximum number of tries & 0.113 \\ 
-  & (0.150) \\ 
-  & \\ 
- Constant & $-$4.523 \\ 
-  & (5.332) \\ 
-  & \\ 
-\hline \\[-1.8ex] 
-Observations & 52 \\ 
-Log Likelihood & $-$18.224 \\ 
-Akaike Inf. Crit. & 50.447 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular}
-
-}
-
-\end{table}%
+\setlength{\LTpost}{0mm}
+
+\begin{longtable}{lcc}
+
+\caption{\label{tbl-mod-output}Jellyfish success rate relative to index
+properties and jellyfish hyper-parameters. This is the summary from a
+logistic regression fit to smoothness, squintability, dimension, running
+time, number of jellyfish and maximum number of tries. Interestingly,
+squintability and dimension strongly affect jellyfish optimisation
+success. The number of jellies marginally affects success, but index
+smoothness, running longer and increasing the number of tries do not.}
+
+\tabularnewline
+
+\toprule
+\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
+\midrule\addlinespace[2.5pt]
+Smoothness & 3.30 (0.08 to 274) & 0.533 \\ 
+Squintability & 7.85 (2.40 to 37.7) & \textbf{0.003} \\ 
+Dimension & 0.53 (0.30 to 0.83) & \textbf{0.014} \\ 
+Long runtime & 0.42 (0.03 to 5.44) & 0.497 \\ 
+Number of jellyfish & 1.24 (0.98 to 1.63) & \textbf{0.089} \\ 
+Maximum number of tries & 1.12 (0.84 to 1.52) & 0.452 \\ 
+\bottomrule
+
+\end{longtable}
+
+\begin{minipage}{\linewidth}
+\textsuperscript{\textit{1}}OR = Odds Ratio, CI = Confidence Interval\\
+\end{minipage}
 
@@
 
-The supplementary materials include: 1) description of all the index
-functions used in the simulations and 2) the script that complements the
-Practicalities section for using JSO in a PPGT and calculating
-smoothness and squintability.
+The supplementary materials available at
+\url{https://github.com/huizezhang-sherry/paper-jso} include: 1) details
+of the indexes used in the simulation study, 2) the script to get
+started with using JSO in a PPGT and calculating smoothness and
+squintability, as explained in Section~\ref{sec-discussion}, and the
+full code to reproduce the plots and summaries in this paper.
 

### 2024-09-13 supplementary/jso.R

@@ -1,0 +1,361 @@
+## ----setup, echo = FALSE, message=FALSE, warning=FALSE-----------------------------
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+library(tidyverse)
+library(ggplot2)
+library(patchwork)
+library(ggh4x)
+library(broom)
+library(kableExtra)
+library(tourr)
+library(PPtreeViz)
+library(ferrn)
+library(colorspace)
+load(here::here("data/sim_pipe_run_best.rda"))
+load(here::here("data/sim_sine_6d_spline_head.rda"))
+load(here::here("data/sim_sine_6d_spline_best.rda"))
+load(here::here("data/sim_sine_6d_spline_projdist.rda"))
+load(here::here("data/pipe_better.rda"))
+load(here::here("data/pipe_jellyfish.rda"))
+load(here::here("data/smoothness.rda"))
+load(here::here("data/squintability.rda"))
+load(here::here("data/sq_basis_dist_idx.rda"))
+load(here::here("data/sim_df.rda"))
+
+
+## ----------------------------------------------------------------------------------
+#| label: fig-example-functions
+#| echo: false
+#| fig-width: 9
+#| fig-height: 6
+#| out-width: 80%
+#| fig-cap: "Examples of PP indexes with large (top row) and small (bottom row) squint angles, shown with a Huber plot, and histogram of the projected data corresponding to the optimal projection. A Huber plot shows the PP index values for all 1D data projections in polar coordinates."
+#proj <- animate_xy(flea[,1:6], guided_tour(lda_pp(flea$species)))
+best_proj <- matrix(c(-0.44157005,  0.80959831,
+                      0.14622089, -0.08228193,
+                      -0.08454314, -0.23234287,
+                      0.70367852,  0.28185021,
+                      0.27876825,  0.44819691,
+                      0.45123453,  0.05896646), ncol=2, byrow=T)
+
+flea2D <- as.matrix(flea[,1:6]) %*% best_proj
+colnames(flea2D) <- c("X1", "X2")
+flea2D <- as.data.frame(flea2D)
+flea_huber <- prep_huber(flea2D, index = skewness())
+sqa1 <- ggplot() +
+  geom_huber(data = flea_huber$idx_df, aes(x = x, y = y)) +
+  geom_point(data = flea2D, aes(x = X1, y = X2)) +
+  geom_abline(slope = flea_huber$slope, intercept = 0) + 
+  coord_fixed() +
+  theme_huber() +
[…]

### 2024-09-15 jso.qmd
small fix in supplementary

@@
 
-The supplementary materials available at <https://github.com/huizezhang-sherry/paper-jso> include: 1) details of the indexes used in the simulation study, 2) the script to get started with using JSO in a PPGT and calculating smoothness and squintability, as explained in @sec-discussion, and the full code to reproduce the plots and summaries in this paper.
+The supplementary materials available at <https://github.com/huizezhang-sherry/paper-jso> include: 1) details of the indexes used in the simulation study, 2) the script to get started with using JSO in a PPGT and calculating smoothness and squintability, as explained in @sec-discussion, and (3) the full code to reproduce the plots and summaries in this paper.
 

### 2024-09-15 jso.tex
small fix in supplementary

@@
 
-\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, opacitybacktitle=0.6, arc=.35mm, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, coltitle=black, bottomrule=.15mm, left=2mm, breakable, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, bottomtitle=1mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, opacityback=0, toprule=.15mm, coltitle=black, opacitybacktitle=0.6, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, breakable, colback=white, toptitle=1mm, titlerule=0mm, left=2mm, arc=.35mm, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}]
 
@@
 started with using JSO in a PPGT and calculating smoothness and
-squintability, as explained in Section~\ref{sec-discussion}, and the
+squintability, as explained in Section~\ref{sec-discussion}, and (3) the
 full code to reproduce the plots and summaries in this paper.

### 2024-09-16 jso.qmd
Typos

@@
 @tbl-smoothness-squintability presents the parameters for calculating both metrics estimated from the Gaussian process (variance, range, smooth, and nugget) and the scaled logistic function ($\theta_1$ to $\theta_4$) for each case considered in the simulation.
-The column "smooth" is used as the smoothness metric and the
-column "squint" is calculated as equation (@eq-squintability-parametric)
+The column $\nu$ is used as the smoothness metric and the
+column $\varsigma$ is calculated as equation (@eq-squintability-parametric)
 as the squintability metric.
@@
 
-The simulation study demonstrated the that these metrics can be used to assess a PP index, and compare optimisers.
+The simulation study demonstrated that these metrics can be used to assess a PP index, and compare optimisers.
 Here we explain how they can be used for new index and optimiser development.
-Using the JSO optimiser for GTPP requires a change in user behaviour.
+Using the JSO optimiser for PPGT requires a change in user behaviour.
 For all the existing optimisers a single optimisation path is followed.

### 2024-09-16 jso.tex
Typos

@@
 \usepackage{caption}
+\usepackage{anyfontsize}
 \usepackage{orcidlink}
@@
 \end{keywords}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, boxrule=0pt, sharp corners, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, opacityback=0, toprule=.15mm, coltitle=black, opacitybacktitle=0.6, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, breakable, colback=white, toptitle=1mm, titlerule=0mm, left=2mm, arc=.35mm, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, colframe=quarto-callout-note-color-frame, colback=white, toptitle=1mm, breakable, coltitle=black, opacitybacktitle=0.6, arc=.35mm, toprule=.15mm, left=2mm, bottomtitle=1mm, leftrule=.75mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}]
 
@@
 
-\section{Assessing optimisers: a simulation study}\label{sec-sim-deets}
+\hypertarget{sec-sim-deets}{%
+\section{Assessing optimisers: a simulation study}\label{sec-sim-deets}}
 
@@
 
-\end{figure}%
-
-\section{Results}\label{sec-sim-res}
-
+\end{figure}
+
+\hypertarget{sec-sim-res}{%
+\section{Results}\label{sec-sim-res}}
+
+\hypertarget{performance-of-jso-relative-to-crs}{%
 \subsection{Performance of JSO relative to
-CRS}\label{performance-of-jso-relative-to-crs}
+CRS}\label{performance-of-jso-relative-to-crs}}
 
@@
 
+\hypertarget{factors-affecting-jso-success-rate-index-properties-and-jellyfish-hyper-parameters}{%
 \subsection{Factors affecting JSO success rate: index properties and
 jellyfish
-hyper-parameters}\label{factors-affecting-jso-success-rate-index-properties-and-jellyfish-hyper-parameters}
+hyper-parameters}\label{factors-affecting-jso-success-rate-index-properties-and-jellyfish-hyper-parameters}}
 
@@
 (\(\theta_1\) to \(\theta_4\)) for each case considered in the
-simulation. The column ``smooth'' is used as the smoothness metric and
-the column ``squint'' is calculated as equation
+simulation. The column \(\nu\) is used as the smoothness metric and the
+column \(\varsigma\) is calculated as equation
 (\ref{eq-squintability-parametric}) as the squintability metric.
@@
 
+\hypertarget{tbl-smoothness-squintability}{}
 \begin{table}
-
-\caption{\label{tbl-smoothness-squintability}Parameters estimated from
-the Gaussian process (outputscale \(\eta\), lengthscale \(\ell\),
-smoothness \(\nu\), and nugget \(\sigma\)) and scaled logistic function
-(\(\theta_1\) to \(\theta_4\) and \(\varsigma\)) for the pipe-finding
-and sine-wave finding problems. The columns \(\nu\) and \(\varsigma\)
-represent the smoothness and squintability measures respectively.}
-
-\centering{
+\caption{\label{tbl-smoothness-squintability}Parameters estimated from the Gaussian process (outputscale \(\eta\),
+lengthscale \(\ell\), smoothness \(\nu\), and nugget \(\sigma\)) and
+scaled logistic function (\(\theta_1\) to \(\theta_4\) and
+\(\varsigma\)) for the pipe-finding and sine-wave finding problems. The
+columns \(\nu\) and \(\varsigma\) represent the smoothness and
+squintability measures respectively. }\tabularnewline
 
@@
 \endgroup{}
-
-}
-
-\end{table}%
-
+\end{table}
+
+\hypertarget{tbl-mod-output}{}
+\begingroup
+\fontsize{12.0pt}{14.4pt}\selectfont
 \setlength{\LTpost}{0mm}
-
 \begin{longtable}{lcc}
-
-\caption{\label{tbl-mod-output}Jellyfish success rate relative to index
-properties and jellyfish hyper-parameters. This is the summary from a
-logistic regression fit to smoothness, squintability, dimension, running
-time, number of jellyfish and maximum number of tries. Interestingly,
-squintability and dimension strongly affect jellyfish optimisation
-success. The number of jellies marginally affects success, but index
-smoothness, running longer and increasing the number of tries do not.}
-
-\tabularnewline
+\caption{\label{tbl-mod-output}Jellyfish success rate relative to index properties and jellyfish
+hyper-parameters. This is the summary from a logistic regression fit to
+smoothness, squintability, dimension, running time, number of jellyfish
+and maximum number of tries. Interestingly, squintability and dimension
+strongly affect jellyfish optimisation success. The number of jellies
+marginally affects success, but index smoothness, running longer and
+increasing the number of tries do not. }\tabularnewline
 
@@
 Smoothness & 3.30 (0.08 to 274) & 0.533 \\ 
-Squintability & 7.85 (2.40 to 37.7) & \textbf{0.003} \\ 
-Dimension & 0.53 (0.30 to 0.83) & \textbf{0.014} \\ 
+Squintability & 7.85 (2.40 to 37.7) & {\bfseries 0.003} \\ 
+Dimension & 0.53 (0.30 to 0.83) & {\bfseries 0.014} \\ 
 Long runtime & 0.42 (0.03 to 5.44) & 0.497 \\ 
-Number of jellyfish & 1.24 (0.98 to 1.63) & \textbf{0.089} \\ 
+Number of jellyfish & 1.24 (0.98 to 1.63) & {\bfseries 0.089} \\ 
 Maximum number of tries & 1.12 (0.84 to 1.52) & 0.452 \\ 
 \bottomrule
-
 \end{longtable}
-
 \begin{minipage}{\linewidth}
@@
 \end{minipage}
-
-\section{Practicalities}\label{sec-discussion}
-
-The simulation study demonstrated the that these metrics can be used to
+\endgroup
+
+\hypertarget{sec-discussion}{%
+\section{Practicalities}\label{sec-discussion}}
+
+The simulation study demonstrated that these metrics can be used to
 assess a PP index, and compare optimisers. Here we explain how they can
 be used for new index and optimiser development. Using the JSO optimiser
-for GTPP requires a change in user behaviour. For all the existing
+for PPGT requires a change in user behaviour. For all the existing
 optimisers a single optimisation path is followed. The JSO has multiple
@@
 
-\section*{Supplementary materials}\label{supplementary-materials}
+\hypertarget{supplementary-materials}{%
+\section*{Supplementary materials}\label{supplementary-materials}}
 \addcontentsline{toc}{section}{Supplementary materials}

### 2024-09-16 jso.qmd
minor typos

@@
 
-To sample bases for calculating smoothness, four arguments are required: 1) the index function, the dataset, the number of random bases to sample, and the best basis to compute projection distance.
+To sample bases for calculating smoothness, four arguments are required: the index function, the dataset, the number of random bases to sample, and the best basis to compute projection distance.
 The output of `sample_bases()` is a data frame containing columns/ list-columns of sampled basis matrix, calculated projection distance and index value.
@@
 The output provides the estimated parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$) and the squintability metric for the parametric case.
-For the non-parametric case, it shows the maximum gradient attained(`max_d`), the corresponding projection distance (`max_dist`), and the squintability metric as their products.
+For the non-parametric case, it shows the maximum gradient attained (`max_d`), the corresponding projection distance (`max_dist`), and the squintability metric as their products.
 

### 2024-09-16 jso.tex
minor typos

@@
 \end{keywords}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, boxrule=0pt, sharp corners, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, enhanced, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, colframe=quarto-callout-note-color-frame, colback=white, toptitle=1mm, breakable, coltitle=black, opacitybacktitle=0.6, arc=.35mm, toprule=.15mm, left=2mm, bottomtitle=1mm, leftrule=.75mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}]
+\begin{tcolorbox}[enhanced jigsaw, breakable, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, opacityback=0, opacitybacktitle=0.6, colback=white, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, arc=.35mm, toprule=.15mm, rightrule=.15mm, left=2mm, bottomtitle=1mm]
 
@@
 To sample bases for calculating smoothness, four arguments are required:
-1) the index function, the dataset, the number of random bases to
-sample, and the best basis to compute projection distance. The output of
+the index function, the dataset, the number of random bases to sample,
+and the best basis to compute projection distance. The output of
 \texttt{sample\_bases()} is a data frame containing columns/
@@
 \(\theta_4\)) and the squintability metric for the parametric case. For
-the non-parametric case, it shows the maximum gradient
-attained(\texttt{max\_d}), the corresponding projection distance
+the non-parametric case, it shows the maximum gradient attained
+(\texttt{max\_d}), the corresponding projection distance
 (\texttt{max\_dist}), and the squintability metric as their products.

### 2024-09-16 jso.qmd
update biblio
doi and other bibliography fixes

@@
 abstract: |
-  The projection pursuit (PP) guided tour interactively optimises a criteria function known as the PP index, to explore high-dimensional data by revealing interesting projections.
+  The projection pursuit (PP) guided tour interactively optimises a criterion function known as the PP index, to explore high-dimensional data by revealing interesting projections.
 Optimisation of some PP indexes can be non-trivial, if they are non-smooth functions, or the optima has a small "squint angle", detectable only from close proximity.
@@
 @hall1989polynomial, @cook1993projection,
-@lee2010projection, @Loperfido2018, @Loperfido2020), which defines the criteria for what constitutes interesting or informative projections.
+@lee2010projection, @Loperfido2018, @Loperfido2020), which defines the criterion for what constitutes interesting or informative projections.
 Let $X \in \mathbb{R}^{n\times p}$ be the data matrix, $A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix, where $A$ belongs to the Stiefel manifold $\mathcal{A} = V_d(\mathbb{R}^p)$.
@@
 
-Optimisation of PP is often discussed when new indexes are proposed [@posse1995; @marie-sainte2010; @grochowski2011].
+Optimisation of PP is often discussed when new indexes are proposed [@posse95; @marie-sainte2010; @grochowski2011].
 @cook1995grand tied the optimisation more closely to the index, when they introduced the PPGT, which monitors the optimisation visually so that the user can see the projected data leading in and out of the optima.

### 2024-09-16 jso.tex
update biblio
doi and other bibliography fixes

@@
 The projection pursuit (PP) guided tour interactively optimises a
-criteria function known as the PP index, to explore high-dimensional
+criterion function known as the PP index, to explore high-dimensional
 data by revealing interesting projections. Optimisation of some PP
@@
 \end{keywords}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, enhanced, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, frame hidden, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 and Cabrera (1993), Lee and Cook (2010), Loperfido (2018), Loperfido
-(2020)), which defines the criteria for what constitutes interesting or
+(2020)), which defines the criterion for what constitutes interesting or
 informative projections. Let \(X \in \mathbb{R}^{n\times p}\) be the
@@
 Optimisation of PP is often discussed when new indexes are proposed
-(Posse 1995a; Marie-Sainte, Berro, and Ruiz-Gazen 2010; Grochowski and
+(Posse 1995; Marie-Sainte, Berro, and Ruiz-Gazen 2010; Grochowski and
 Duch 2011). Cook et al. (1995) tied the optimisation more closely to the
@@
 \item
-  a modified brute force algorithm described in Posse (1995b).
+  a modified brute force algorithm described in Posse (1995).
 \item
@@
 
-\begin{tcolorbox}[enhanced jigsaw, breakable, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, opacityback=0, opacitybacktitle=0.6, colback=white, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, arc=.35mm, toprule=.15mm, rightrule=.15mm, left=2mm, bottomtitle=1mm]
+\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, breakable, toprule=.15mm, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, colback=white, bottomtitle=1mm, opacityback=0, rightrule=.15mm, toptitle=1mm, coltitle=black]
 
@@
 with {ExaGeoStatR}.''} \emph{Environmetrics} 34 (1): e2770.
-https://doi.org/\url{https://doi.org/10.1002/env.2770}.
+\url{https://doi.org/10.1002/env.2770}.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-FT74}{}}%
-Friedman, J. H., and J. W. Tukey. 1974. {``{A} {P}rojection {P}ursuit
-{A}lgorithm for {E}xploratory {D}ata {A}nalysis.''} \emph{IEEE
-Transactions on Computing C} 23: 881--89.
+Friedman, J. H., and J. W. Tukey. 1974. {``A Projection Pursuit
+Algorithm for Exploratory Data Analysis.''} \emph{IEEE Transactions on
+Computers} C-23 (9): 881--90.
+\url{https://doi.org/10.1109/T-C.1974.224051}.
 
@@
 Branko Šter, 89--97. Berlin, Heidelberg: Springer Berlin Heidelberg.
+\url{https://doi.org/10.1007/978-3-642-20267-4_10}.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-huber85}{}}%
-Huber, Peter J. 1985. {``Projection Pursuit.''} \emph{Ann. Statist.} 13
-(2): 435--75. \url{https://doi.org/10.1214/aos/1176349519}.
+Huber, Peter J. 1985. {``Projection Pursuit.''} \emph{The Annals of
+Statistics} 13 (2): 435--75.
+\url{https://doi.org/10.1214/aos/1176349519}.
 
@@
 on Uncertainty Quantification} 11 (4): 1225--57.
+\url{https://doi.org/10.1137/22M149288X}.
 
@@
 427--40. New York: Academic Press.
+\url{https://doi.org/10.1016/B978-0-12-498150-8.50024-0}.
 
@@
 Lee, Eun-Kyung, and Dianne Cook. 2010. {``A Projection Pursuit Index for
-Large p Small n Data.''} \emph{Statistics and Computing} 20 (3):
-381--92. \url{https://doi.org/10.1007/s11222-009-9131-1}.
+Large {\(p\)} Small {\(n\)} Data.''} \emph{Statistics and Computing} 20
+(3): 381--92. \url{https://doi.org/10.1007/s11222-009-9131-1}.
 
@@
 Analysis} 120 (C): 42--57.
-https://doi.org/\url{https://doi.org/10.1016/j.csda.2017.11.001}.
+\url{https://doi.org/10.1016/j.csda.2017.11.001}.
 
@@
 Finance} 26 (2-3): 142--64.
-https://doi.org/\url{https://doi.org/10.1080/1351847X.2019.1647864}.
+\url{https://doi.org/10.1080/1351847X.2019.1647864}.
 
@@
 Berlin, Heidelberg: Springer Berlin Heidelberg.
+\url{https://doi.org/10.1007/978-3-642-15461-4_6}.
 
@@
 Analysis and Machine Learning.''} \emph{Statistical Science} 39 (3):
-469--92.
+469--92. \url{https://doi.org/10.1214/24-STS923}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-posse95}{}}%
-Posse, Christian. 1995b. {``Projection Pursuit Exploratory Data
+Posse, Christian. 1995. {``Projection Pursuit Exploratory Data
 Analysis.''} \emph{Computational Statistics and Data Analysis} 20 (6):
-669--87.
-https://doi.org/\url{https://doi.org/10.1016/0167-9473(95)00002-8}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-posse1995}{}}%
----------. 1995a. {``Projection Pursuit Exploratory Data Analysis.''}
-\emph{Computational Statistics \& Data Analysis} 20 (6): 669--87.
-https://doi.org/\url{https://doi.org/10.1016/0167-9473(95)00002-8}.
+669--87. \url{https://doi.org/10.1016/0167-9473(95)00002-8}.
 
@@
 \emph{Journal of Statistical Software} 40 (2): 1--18.
-\url{http://doi.org/10.18637/jss.v040.i02}.
+\url{https://doi.org/10.18637/jss.v040.i02}.
 

### 2024-09-16 jso.qmd
minor corrections

@@
   The projection pursuit (PP) guided tour interactively optimises a criterion function known as the PP index, to explore high-dimensional data by revealing interesting projections.
-Optimisation of some PP indexes can be non-trivial, if they are non-smooth functions, or the optima has a small "squint angle", detectable only from close proximity.
+Optimisation of some PP indexes can be non-trivial, if they are non-smooth functions, or the optimum has a small "squint angle", detectable only from close proximity.
 To address these challenges, this study investigates the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
@@
 Indexes like the holes, central mass and skewness [@cook1993projection] are generally smooth for most data sets, but capture only large patterns.
-Many indexes noisy and non-convex, requiring an effective and efficient optimisation procedure to explore the data landscape and achieve a globally optimal viewpoint of the data.
+Many indexes are noisy and non-convex, requiring an effective and efficient optimisation procedure to explore the data landscape and achieve a globally optimal viewpoint of the data.
 The skewness index computed for trimodal data, in (a), is smooth with a large squint angle but has three modes, and thus is not convex.
@@
 Optimisation of PP is often discussed when new indexes are proposed [@posse95; @marie-sainte2010; @grochowski2011].
-@cook1995grand tied the optimisation more closely to the index, when they introduced the PPGT, which monitors the optimisation visually so that the user can see the projected data leading in and out of the optima.
+@cook1995grand tied the optimisation more closely to the index, when they introduced the PPGT, which monitors the optimisation visually so that the user can see the projected data leading in and out of the optimum.
 An implementation is available in the `tourr` package [@tourr] in R [@R].
@@
 performance of JSO in PP optimisation for the guided tour. It is of
-interest to assess how quickly and closely the optimiser reaches a gobal
-optima, for various PP indexes that may have differing complexities. To
+interest to assess how quickly and closely the optimiser reaches a global
+optimum, for various PP indexes that may have differing complexities. To
 observe the performance of JSO with different types of PP indexes,
@@
 - a modified brute force algorithm described in @posse95. 
-- an essentially simulated annealing [@Bertsimas93] where the search space is reduced as the optimisation.
+- an essentially simulated annealing [@Bertsimas93] where the search space is reduced during the optimisation.
 - a very localised search, to take tiny steps to get closer to the local maximum.
@@
 
-There are numerous PP index functions available: introduced in @cook1993projection, @lee2005projection, @lee2010projection, @Grimm2016, @huber85, @Laa:2020wkm.
+There are numerous PP index functions available: introduced in @huber85, @cook1993projection, @lee2005projection, @lee2010projection, @Grimm2016, @Laa:2020wkm.
 Most are relatively simply defined, for any projection dimension, and implemented because they are relatively easy to optimise.
-A goal is to develop PP indexes based on scagnostics (@scag, @WW08), but the blockage is optimisation that these tend to be noisy, with potentially small squint angles.
+A goal is to develop PP indexes based on scagnostics (@scag, @WW08), but the blockage is their optimisation as these tend to be noisy, with potentially small squint angles.
 
@@
 scagnostics is described in @laa_using_2020. To be useful here an
-optimiser needs to be able to handle index functions which are possibly
+optimiser needs to be able to handle index functions that are possibly
 not very smooth. In addition, because data structures might be
@@
 by. One last aspect that is useful is for an optimiser to return local
-maxima in addition to global because data can contain many different and
+maxima in addition to the global one because data can contain many different and
 interesting features.
@@
 \begin{equation}
-K(u)=K_{\nu,\eta,\ell}(u):=\eta^2\frac{\left(\sqrt{2\nu}\frac{u}{\ell}\right)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}\left(\sqrt{2\nu}\frac{u}{\ell}\right)\ , u\geq0\label{eq:matern}
+K(u)=K_{\nu,\eta,\ell}(u):=\eta^2\frac{\left(\sqrt{2\nu}\frac{\left\Vert u\right\Vert}{\ell}\right)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}\left(\sqrt{2\nu}\frac{\left\Vert u\right\Vert}{\ell}\right)\ ,\label{eq:matern}
 \end{equation}
-where $\nu>0$ is the smoothness parameter, $\eta$ is the outputscale, $\ell$ is the lengthscale, and $\mathcal{K}_\nu$ is
-the modified Bessel function [DLMF 10.25]. A multivariate extension $K(u)$, $u\in\mathbb{R}^d$ can be obtained by products of univariate covariance functions \eqref{eq:matern}.
+where $\left\Vert u\right\Vert$ is the Euclidean norm of $u\in\mathbb{R}^{p{\times}d}$, $\nu>0$ is the smoothness parameter, $\eta$ is the outputscale, $\ell$ is the lengthscale, and $\mathcal{K}_\nu$ is
+the modified Bessel function [DLMF 10.25].
 \end{defn}
@@
 $\nu\rightarrow\infty$ coincides with the Gaussian kernel:
-$K_\infty(u)=\exp(-u^{2}/2)$.
+$K_\infty(u)=\exp(-{\left\Vert u\right\Vert}^{2}/2)$.
 
@@
 such as `GpGp` [@guinness2021gpgp] or `ExaGeoStatR` [@abdulah2023large],
-to fit the hyperparameters of a GP covariance function on data, which is
+to fit the hyper-parameters of a GP covariance function on data, which is
 usually done by maximum likelihood estimation. In this project, the
@@
 \begin{defn}
-Let $\mathbf{A}=[A_1, \ldots, A_N] \in (\mathbf{R}^{p \times d})^N$ be d-dimensional projection bases, let $\mathbf{y}=[f(XA_1),\ldots,f(XA_N)]$ be the corresponding PP index values, and let $\mathbf{K}=[K_\theta(A_{i},A_{j})]_{1\leq i,j\leq N}\in\mathbb{R}^{N\times N}$ be the Matérn covariance matrix evaluated at the input bases, where the vector $\theta$ contains all the parameters of the multivariate Matérn covariance function $K$ (smoothness, outputscale, lengthscales).
+Let $\mathbf{A}=[A_1, \ldots, A_N] \in (\mathbf{R}^{p \times d})^N$ be d-dimensional projection bases, let $\mathbf{y}=[f(XA_1),\ldots,f(XA_N)]$ be the corresponding PP index values, and let $\mathbf{K}=[K_\theta(A_{i}-A_{j})]_{1\leq i,j\leq N}\in\mathbb{R}^{N\times N}$ be the Matérn covariance matrix evaluated at the input bases, where the vector $\theta$ contains all the parameters of the multivariate Matérn covariance function $K$ (smoothness, outputscale, lengthscales).
 The log-likelihood of the parameters $\theta$ is defined by
@@
 
-Here the formal definition of projection distance and squint angle are given, before the definition of squintability. Two approaches to compute this metric numerically,
-are described.
+Here the formal definitions of projection distance and squint angle are given, before the definition of squintability. Two approaches to compute this metric numerically are described.
 
@@
 and $A^*$ is defined as 
-$d(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\  \rVert _F$
+$d(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\,\rVert _F$
 where $\lVert . \rVert _F$ denotes the Frobenius norm, given by
@@
 \begin{equation}
-\varsigma(f) = -c \times \max_{d} g'(d) \times \arg \max_{d} g'(d)
+\varsigma(f) = -4 \times \max_{d} g'(d) \times \arg \max_{d} g'(d)
 \label{eq-squintability}
@@
 
-where $c$ is a constant scaling factor, $-\max_d g'(d)$ represents the 
+where $4$ is a constant scaling factor, $-\max_d g'(d)$ represents the 
 largest gradient of $-g$ and $\arg \max_{d} g'(d)$ represents the projection distance at which this largest gradient is attained.
@@
 It is expected that these two values should be both high in the case of high squintability (fast
-increase in $f$ early on), and both low in the case of low squintability
-(any substantial increase in $f$ happens very late, close to the optimal
+increase in $g$ early on), and both low in the case of low squintability
+(any substantial increase in $g$ happens very late, close to the optimal
 angle). This suggests that their product \eqref{eq-squintability} should
@@
 $$
-\varsigma(g_1(d_1)) = -c  \max g'_1(d_1)  d_1 = -c  (-g_{\max}) d_1 \leq -c  (-g_{\max}) d_2 = -c \max g'_2(d_2)  d_2 = \varsigma(g_2(d_2))
+\varsigma(g_1(d_1)) = -4  \max g'_1(d_1)  d_1 = -4  (-g_{\max}) d_1 \leq -4  (-g_{\max}) d_2 = -4 \max g'_2(d_2)  d_2 = \varsigma(g_2(d_2))
 $$
@@
 small squint angle is considered to be undesirable because it means that
-the optimiser needs to be very close to be able to "see" the optima.
-Thus, it could be difficult for the optimiser to find the optima. <!--The
+the optimiser needs to be very close to be able to "see" the optimum.
+Thus, it could be difficult for the optimiser to find the optimum. <!--The
 mathematical formulation of this intuition is proposed below:-->
@@
 to efficient optimization. This observation suggests that the extreme
-values of $f'$ (the ones for which $f''=0$, assuming that $f$ is twice
+values of $g'$ (the ones for which $g''=0$, assuming that $g$ is twice
 differentiable), and the projection distances for which these values are
@@
 several approaches are possible. The first one is to propose a
-parametric model for $f$, and use it to obtain an explicit formula for
+parametric model for $g$, and use it to obtain an explicit formula for
 $\varsigma$. Numerical experiments suggest a scaled sigmoid shape as
@@
 $$
-f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
+g(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
 $$ {#eq-parametric}
@@
 which depends on three additional parameters, $\theta_1$, $\theta_2$,
-and $x_{\max}$, such that $f(0)=\theta_1$ and $f(x_{\max})=\theta_4$.
+and $x_{\max}$, such that $g(0)=\theta_1$ and $g(x_{\max})=\theta_4$.
 Under the parametric model (@eq-parametric), the squintability metric
@@
 Alternatively, one can estimate \eqref{eq-squintability} in a nonparametric
-way, for example by fitting $f$ using kernel regression, then
-numerically estimate the angle at which $-f'$ attains its highest value.
+way, for example by fitting $g$ using kernel regression, then
+numerically estimate the angle at which $-g'$ attains its highest value.
 
@@
 In the comparison of the JSO against the currently used CRS, as expected
-the JSO vastly out-performs CRS, and provides a high probability of
-finding the global optima. The JSO obtains the maximum more cleanly,
+the JSO vastly outperforms CRS, and provides a high probability of
+finding the global optimum. The JSO obtains the maximum more cleanly,
 with a slightly higher index value, and plot of the projected data
@@
 The JSO performance is affected by the hyper-parameters, with a higher
-chance of reaching the global optima when more jellyfish are used and
+chance of reaching the global optimum when more jellyfish are used and
 the maximum number of tries is increased. However, it comes at a
@@
 squintability. The higher the squintability the better chance the JSO
-can find the optima. However, interestingly smoothness does not affect
+can find the optimum. However, interestingly smoothness does not affect
 the JSO performance.

### 2024-09-16 jso.tex
minor corrections

@@
 indexes can be non-trivial, if they are non-smooth functions, or the
-optima has a small ``squint angle'', detectable only from close
+optimum has a small ``squint angle'', detectable only from close
 proximity. To address these challenges, this study investigates the
@@
 \end{keywords}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, frame hidden, enhanced, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, boxrule=0pt, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
 
@@
 smooth for most data sets, but capture only large patterns. Many indexes
-noisy and non-convex, requiring an effective and efficient optimisation
-procedure to explore the data landscape and achieve a globally optimal
-viewpoint of the data. The skewness index computed for trimodal data, in
-(a), is smooth with a large squint angle but has three modes, and thus
-is not convex. The binned normality index (a simple version of a
-non-normality index as described in Huber (1985)) computed on the famous
-RANDU data, in (c), is noisier and has a very small squint angle. The
-discreteness cannot be seen unless the optimiser is very close to the
-optimal projection.
+are noisy and non-convex, requiring an effective and efficient
+optimisation procedure to explore the data landscape and achieve a
+globally optimal viewpoint of the data. The skewness index computed for
+trimodal data, in (a), is smooth with a large squint angle but has three
+modes, and thus is not convex. The binned normality index (a simple
+version of a non-normality index as described in Huber (1985)) computed
+on the famous RANDU data, in (c), is noisier and has a very small squint
+angle. The discreteness cannot be seen unless the optimiser is very
+close to the optimal projection.
 
@@
 visually so that the user can see the projected data leading in and out
-of the optima. An implementation is available in the \texttt{tourr}
+of the optimum. An implementation is available in the \texttt{tourr}
 package (Wickham et al. 2011) in R (R Core Team 2023). Zhang et al.
@@
 performance of JSO in PP optimisation for the guided tour. It is of
-interest to assess how quickly and closely the optimiser reaches a gobal
-optima, for various PP indexes that may have differing complexities. To
-observe the performance of JSO with different types of PP indexes,
-metrics are introduced to capture specific properties of the index
-including squintability (based on Tukey and Tukey (1981)'s squint angle)
-and smoothness. Here, we mathematically define metrics for squintability
-and smoothness, which is a new contribution for PP research. A series of
-simulation experiments using various datasets and PP indexes are
-conducted to assess JSO's behaviour and its sensitivity to
-hyper-parameter choices (number of jellyfish and maximum number of
+interest to assess how quickly and closely the optimiser reaches a
+global optimum, for various PP indexes that may have differing
+complexities. To observe the performance of JSO with different types of
+PP indexes, metrics are introduced to capture specific properties of the
+index including squintability (based on Tukey and Tukey (1981)'s squint
+angle) and smoothness. Here, we mathematically define metrics for
+squintability and smoothness, which is a new contribution for PP
+research. A series of simulation experiments using various datasets and
+PP indexes are conducted to assess JSO's behaviour and its sensitivity
+to hyper-parameter choices (number of jellyfish and maximum number of
 tries). The relationship between the JSO performance, hyper-parameter
@@
   an essentially simulated annealing (Bertsimas and Tsitsiklis 1993)
-  where the search space is reduced as the optimisation.
+  where the search space is reduced during the optimisation.
 \item
@@
 
-There are numerous PP index functions available: introduced in Cook,
-Buja, and Cabrera (1993), Lee et al. (2005), Lee and Cook (2010), Grimm
-(2016), Huber (1985), Ursula Laa and Valencia (2022). Most are
+There are numerous PP index functions available: introduced in Huber
+(1985), Cook, Buja, and Cabrera (1993), Lee et al. (2005), Lee and Cook
+(2010), Grimm (2016), Ursula Laa and Valencia (2022). Most are
 relatively simply defined, for any projection dimension, and implemented
@@
 indexes based on scagnostics (L. Wilkinson, Anand, and Grossman (2005),
-Leland Wilkinson and Wills (2008)), but the blockage is optimisation
-that these tend to be noisy, with potentially small squint angles.
+Leland Wilkinson and Wills (2008)), but the blockage is their
+optimisation as these tend to be noisy, with potentially small squint
+angles.
 
@@
 scagnostics is described in Laa and Cook (2020). To be useful here an
-optimiser needs to be able to handle index functions which are possibly
+optimiser needs to be able to handle index functions that are possibly
 not very smooth. In addition, because data structures might be
@@
 by. One last aspect that is useful is for an optimiser to return local
-maxima in addition to global because data can contain many different and
-interesting features.
+maxima in addition to the global one because data can contain many
+different and interesting features.
 
@@
 \(\nu\rightarrow\infty\) coincides with the Gaussian kernel:
-\(K_\infty(u)=\exp(-u^{2}/2)\).
+\(K_\infty(u)=\exp(-{\left\Vert u\right\Vert}^{2}/2)\).
 
@@
 such as \texttt{GpGp} (Guinness, Katzfuss, and Fahmy 2021) or
-\texttt{ExaGeoStatR} (Abdulah et al. 2023), to fit the hyperparameters
+\texttt{ExaGeoStatR} (Abdulah et al. 2023), to fit the hyper-parameters
 of a GP covariance function on data, which is usually done by maximum
@@
 
-Here the formal definition of projection distance and squint angle are
+Here the formal definitions of projection distance and squint angle are
 given, before the definition of squintability. Two approaches to compute
-this metric numerically, are described.
+this metric numerically are described.
 
@@
 It is expected that these two values should be both high in the case of
-high squintability (fast increase in \(f\) early on), and both low in
-the case of low squintability (any substantial increase in \(f\) happens
+high squintability (fast increase in \(g\) early on), and both low in
+the case of low squintability (any substantial increase in \(g\) happens
 very late, close to the optimal angle). This suggests that their product
@@
 small squint angle is considered to be undesirable because it means that
-the optimiser needs to be very close to be able to ``see'' the optima.
-Thus, it could be difficult for the optimiser to find the optima.
+the optimiser needs to be very close to be able to ``see'' the optimum.
+Thus, it could be difficult for the optimiser to find the optimum.
 
@@
 amenable to efficient optimization. This observation suggests that the
-extreme values of \(f'\) (the ones for which \(f''=0\), assuming that
-\(f\) is twice differentiable), and the projection distances for which
+extreme values of \(g'\) (the ones for which \(g''=0\), assuming that
+\(g\) is twice differentiable), and the projection distances for which
 these values are attained, are crucial in the mathematical definition of
@@
 practice, several approaches are possible. The first one is to propose a
-parametric model for \(f\), and use it to obtain an explicit formula for
+parametric model for \(g\), and use it to obtain an explicit formula for
 \(\varsigma\). Numerical experiments suggest a scaled sigmoid shape as
@@
 \begin{equation}\protect\hypertarget{eq-parametric}{}{
-f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
+g(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
 }\label{eq-parametric}\end{equation}
@@
 which depends on three additional parameters, \(\theta_1\),
-\(\theta_2\), and \(x_{\max}\), such that \(f(0)=\theta_1\) and
-\(f(x_{\max})=\theta_4\). Under the parametric model
+\(\theta_2\), and \(x_{\max}\), such that \(g(0)=\theta_1\) and
+\(g(x_{\max})=\theta_4\). Under the parametric model
 (\ref{eq-parametric}), the squintability metric \eqref{eq-squintability}
@@
 Alternatively, one can estimate \eqref{eq-squintability} in a
-nonparametric way, for example by fitting \(f\) using kernel regression,
-then numerically estimate the angle at which \(-f'\) attains its highest
+nonparametric way, for example by fitting \(g\) using kernel regression,
+then numerically estimate the angle at which \(-g'\) attains its highest
 value.
@@
 
-\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, breakable, toprule=.15mm, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, colback=white, bottomtitle=1mm, opacityback=0, rightrule=.15mm, toptitle=1mm, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, left=2mm, titlerule=0mm, rightrule=.15mm, opacityback=0, bottomrule=.15mm, leftrule=.75mm, colback=white, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, toptitle=1mm, arc=.35mm, opacitybacktitle=0.6, breakable, toprule=.15mm, coltitle=black]
 

### 2024-09-17 jso.qmd
squintability
fixed Definition 6, Proposition 1, and its proof.

@@
 \begin{equation}
-\varsigma(f) = -4 \times \max_{d} g'(d) \times \arg \max_{d} g'(d)
+\varsigma(f) = -4 \times \min_{d} g'(d) \times \arg \min_{d} g'(d)
 \label{eq-squintability}
@@
 
-where $4$ is a constant scaling factor, $-\max_d g'(d)$ represents the 
-largest gradient of $-g$ and $\arg \max_{d} g'(d)$ represents the projection distance at which this largest gradient is attained.
+where $4$ is a constant scaling factor, $-\min_d g'(d)$ represents the 
+largest gradient of $-g$ and $\arg \min_{d} g'(d)$ represents the projection distance at which this largest gradient is attained.
 
@@
 \begin{prop}\label{prop:convex-concave}
-Let $g_1(d)$ be a convex decreasing function and $g_2(d)$ be a concave decreasing function both defined on [0, 1] where $d \in [0, D]$.
-Let $d_1 := \underset{d}{\arg \max} |g_1'(d)|$ and $d_2 := \underset{d}{\arg \max} |g_2'(d)|$.
-
+Let $g_1$ be a convex, strictly decreasing function and $g_2$ be a concave, strictly decreasing function, both defined on $[0, D]$ where $D>0$, and taking values in $[0, 1]$. Then
 $$\varsigma(g_1) < \varsigma(g_2)$$
-
-For a convex function, $g_1'(d) < 0 \; \forall d$ and is non-decreasing. For a concave function, $g_2'(d) < 0 \; \forall d$ and is non-increasing. As such, it follows that $d_1 < d_2$.
-
-Suppose both functions achieve the same maximum gradient $-g_{\max}$, i.e., $\max g'_1(d_1) = \max g'_2(d_2) = -g_{\max}$,
-
-$$
-\varsigma(g_1(d_1)) = -4  \max g'_1(d_1)  d_1 = -4  (-g_{\max}) d_1 \leq -4  (-g_{\max}) d_2 = -4 \max g'_2(d_2)  d_2 = \varsigma(g_2(d_2))
-$$
-
 \end{prop}
+
+\begin{proof}
+Let $d_1 := \underset{d}{\arg \min }\,g_1'(d)$ and $d_2 := \underset{d}{\arg \min }\,g_2'(d)$.
+Since $g_1$ is convex, $g_1''(d) >= 0 \; \forall d \in [0, D]$, meaning that $g_1'$ is an increasing function, taking its minimum for $d_1=0$.
+As a result, $\varsigma(g_1)=-4 \times g'(0) \times 0 =0$.
+Since $g_2$ is concave, $g_2''(d) <= 0 \; \forall d \in [0, D]$, meaning that $g_2'$ is a decreasing function, taking its minimum for $d_2=D>0$.
+Since $g_2$ is strictly decreasing, $g_2'(d) < 0 \; \forall d \in [0, D]$.
+In particular $g_2'(D)<0$.
+As a result $\varsigma(g_2)=-4 \times g'(D) \times D >0$.
+This proves that $\varsigma(g_1) < \varsigma(g_2)$.
+\end{proof}
 

### 2024-09-17 jso.tex
squintability
fixed Definition 6, Proposition 1, and its proof.

@@
 \end{keywords}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, boxrule=0pt, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, left=2mm, titlerule=0mm, rightrule=.15mm, opacityback=0, bottomrule=.15mm, leftrule=.75mm, colback=white, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, toptitle=1mm, arc=.35mm, opacitybacktitle=0.6, breakable, toprule=.15mm, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, opacityback=0, breakable, coltitle=black, colframe=quarto-callout-note-color-frame, left=2mm, toprule=.15mm, bottomrule=.15mm, toptitle=1mm, titlerule=0mm, arc=.35mm, opacitybacktitle=0.6, colback=white]
 

### 2024-09-17 jso.qmd
minor corrections

@@
 largest gradient of $-g$ and $\arg \min_{d} g'(d)$ represents the projection distance at which this largest gradient is attained.
+If the set $\mathcal{D}:=\arg \min_{d} g'(d)$ contains more than one value $d$, the largest value $\max(\mathcal{D})$ is used in equation \eqref{eq-squintability}.
 
@@
 to efficient optimization. This observation suggests that the extreme
-values of $g'$ (the ones for which $g''=0$, assuming that $g$ is twice
-differentiable), and the projection distances for which these values are
+values of $g'$ and the projection distances for which these values are
 attained, are crucial in the mathematical definition of squintability.
-
-<!-- Noticing that $f$ is expected to be a decreasing function, the -->
-<!-- squintability metric is proposed as follows: -->
-
-<!-- $$ -->
-<!-- \varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ . -->
-<!-- $$ {#eq-squintability} -->
-
-<!-- $-\min_{x}f'(x)$ represents the largest value of the gradient of $-f$ -->
-<!-- and $\underset{x}{\arg\min}f'(x)$ represents the projection distance at -->
-<!-- which this largest gradient is attained.  -->
 
@@
 described below. Define
-
 $$
@@
 Then, define
-
 $$
@@
 \eqref{eq-squintability} can be shown to be equal to
-
 $$
@@
 projection distances are calculated for these interpolated bases and the
-index values are averaged with a bin width of 0.005. A four-parameter
-scaled logistic function is fitted to the index values against
+index values are averaged with a bin width of 0.005. The four-parameter
+scaled logistic function (@eq-parametric) is fitted to the index values against
 projection distances, estimated by non-linear least squares. The
@@
 The proportion is calculated based on the number of simulations, out of 50, that achieve an index value within 0.05 of the best-performing simulation.
-To quantify uncertainty, Bootstrap samples with 500 are generated.
-The thin lines represent the proportion for each of the 500 bootstrap replicates, while the thicker lines represent the mean of these bootstrap samples, connnected by lines.
-As the dimensionality increases, the proportion of simulations reaching the optimal index value increases."
+To quantify uncertainty, 500 bootstrap samples are generated.
+The thin segments represent the proportion for each of the 500 bootstrap replicates, while the thicker segments represent the mean of these bootstrap samples, connnected by lines.
+As the dimensionality increases, the proportion of simulations reaching the optimal index value decreases."
 pipe_sim_best <- sim_pipe_run_best |> 
@@
 
-@tbl-mod-output presents the results from fitting a logistic regression model using proportion of success as the response and smoothness, squintability, dimension, running time, number of jellyfish and maximum number of tries, as predictors.
+@tbl-mod-output presents the results from fitting a logistic regression model using the proportion of success as the response and smoothness, squintability, dimension, running time, number of jellyfish and maximum number of tries, as predictors.
 Long runtime is a binary variable, where greater than 20 seconds is labelled 1.

### 2024-09-17 jso.tex
minor corrections

@@
 \end{keywords}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 amenable to efficient optimization. This observation suggests that the
-extreme values of \(g'\) (the ones for which \(g''=0\), assuming that
-\(g\) is twice differentiable), and the projection distances for which
-these values are attained, are crucial in the mathematical definition of
+extreme values of \(g'\) and the projection distances for which these
+values are attained, are crucial in the mathematical definition of
 squintability.
@@
 described below. Define
-
 \begin{equation}\protect\hypertarget{eq-logistic}{}{
@@
 \(\ell(\theta_{2})=\frac{1}{2}\). Then, define
-
 \begin{equation}\protect\hypertarget{eq-parametric}{}{
@@
 can be shown to be equal to
-
 \begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, opacityback=0, breakable, coltitle=black, colframe=quarto-callout-note-color-frame, left=2mm, toprule=.15mm, bottomrule=.15mm, toptitle=1mm, titlerule=0mm, arc=.35mm, opacitybacktitle=0.6, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, breakable, bottomtitle=1mm, titlerule=0mm, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, colback=white, left=2mm, opacityback=0, coltitle=black, opacitybacktitle=0.6, arc=.35mm, rightrule=.15mm]
 
@@
 projection distances are calculated for these interpolated bases and the
-index values are averaged with a bin width of 0.005. A four-parameter
-scaled logistic function is fitted to the index values against
-projection distances, estimated by non-linear least squares. The
-squintability measure is then calculated as
+index values are averaged with a bin width of 0.005. The four-parameter
+scaled logistic function (\ref{eq-parametric}) is fitted to the index
+values against projection distances, estimated by non-linear least
+squares. The squintability measure is then calculated as
 (\ref{eq-squintability-parametric}).
@@
 out of 50, that achieve an index value within 0.05 of the
-best-performing simulation. To quantify uncertainty, Bootstrap samples
-with 500 are generated. The thin lines represent the proportion for each
-of the 500 bootstrap replicates, while the thicker lines represent the
-mean of these bootstrap samples, connnected by lines. As the
-dimensionality increases, the proportion of simulations reaching the
-optimal index value increases.}
+best-performing simulation. To quantify uncertainty, 500 bootstrap
+samples are generated. The thin segments represent the proportion for
+each of the 500 bootstrap replicates, while the thicker segments
+represent the mean of these bootstrap samples, connnected by lines. As
+the dimensionality increases, the proportion of simulations reaching the
+optimal index value decreases.}
 
@@
 Table~\ref{tbl-mod-output} presents the results from fitting a logistic
-regression model using proportion of success as the response and
+regression model using the proportion of success as the response and
 smoothness, squintability, dimension, running time, number of jellyfish

### 2024-09-18 jso.qmd
rerun the smoothness metric after ferrn update

@@
 Fifty simulations are conducted with both JSO and CRS, in four different
-data dimensions ($d = 6, 8, 10, 12$). JSO uses 100 jellyfishes with a
+data dimensions ($d = 6, 8, 10, 12$). JSO uses 100 jellyfish with a
 maximum of 100 tries, while the CRS allows a maximum of 1000 samples at
@@
 binary variable `long_time` to indicate cases with an average run time
-over 30 seconds, and 3) re-coding the success rate for the `stringy`
+over 20 seconds, and 3) re-coding the success rate for the `stringy`
 index as 0, because none of the 50 simulations correctly identified the
@@
 As the number of jellyfish and maximum tries increase, the success rate also increases.
-For simpler problems (6 dimensions), small parameter values (20 jellyfishes and a maximum of 50 tries) can already achieve a high success rate.
+For simpler problems (6 dimensions), small parameter values (20 jellyfish and a maximum of 50 tries) can already achieve a high success rate.
 However, larger parameter values
-(i.e. 100 jellyfishes and a maximum of 100 tries) are necessary for
+(i.e. 100 jellyfish and a maximum of 100 tries) are necessary for
 higher-dimensional problems (8, 10, and 12 dimensions). Increasing both
@@
 The projection pursuit problem is to find the pipe shape using the holes index in the 6, 8, 10, and 12-dimensional spaces.
-The JSO uses 100 jellyfishes and a maximum number of tries of 100.
+The JSO uses 100 jellyfish and a maximum number of tries of 100.
 The CRS uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates.

### 2024-09-18 jso.tex
rerun the smoothness metric after ferrn update

@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
+
 \ifLuaTeX
@@
 \end{keywords}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 
-{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{jso_files/figure-pdf/fig-example-functions-1.pdf}
+\centering{
+
+\includegraphics[width=0.8\textwidth,height=\textheight]{jso_files/figure-pdf/fig-example-functions-1.pdf}
 
@@
 
-\hypertarget{sec-PP-properties}{%
-\section{Properties of PP indexes}\label{sec-PP-properties}}
+\section{Properties of PP indexes}\label{sec-PP-properties}
 
@@
 described below. Define
-\begin{equation}\protect\hypertarget{eq-logistic}{}{
+\begin{equation}\phantomsection\label{eq-logistic}{
 \ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
-}\label{eq-logistic}\end{equation}
+}\end{equation}
 
@@
 \(\ell(\theta_{2})=\frac{1}{2}\). Then, define
-\begin{equation}\protect\hypertarget{eq-parametric}{}{
+\begin{equation}\phantomsection\label{eq-parametric}{
 g(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
-}\label{eq-parametric}\end{equation}
+}\end{equation}
 
@@
 can be shown to be equal to
-\begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
+\begin{equation}\phantomsection\label{eq-squintability-parametric}{
 \varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
-}\label{eq-squintability-parametric}\end{equation}
+}\end{equation}
 
@@
 
-\hypertarget{sec-JSO}{%
-\section{The jellyfish optimiser}\label{sec-JSO}}
+\section{The jellyfish optimiser}\label{sec-JSO}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, breakable, bottomtitle=1mm, titlerule=0mm, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, colback=white, left=2mm, opacityback=0, coltitle=black, opacitybacktitle=0.6, arc=.35mm, rightrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, rightrule=.15mm, arc=.35mm, colback=white, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, leftrule=.75mm, breakable, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, opacityback=0, left=2mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, coltitle=black]
 
@@
 
-\hypertarget{sec-sim-deets}{%
-\section{Assessing optimisers: a simulation study}\label{sec-sim-deets}}
+\section{Assessing optimisers: a simulation study}\label{sec-sim-deets}
 
@@
 Fifty simulations are conducted with both JSO and CRS, in four different
-data dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100 jellyfishes with a
+data dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100 jellyfish with a
 maximum of 100 tries, while the CRS allows a maximum of 1000 samples at
@@
 binary variable \texttt{long\_time} to indicate cases with an average
-run time over 30 seconds, and 3) re-coding the success rate for the
+run time over 20 seconds, and 3) re-coding the success rate for the
 \texttt{stringy} index as 0, because none of the 50 simulations
@@
 
-\end{figure}
-
-\hypertarget{sec-sim-res}{%
-\section{Results}\label{sec-sim-res}}
-
-\hypertarget{performance-of-jso-relative-to-crs}{%
+\end{figure}%
+
+\section{Results}\label{sec-sim-res}
+
 \subsection{Performance of JSO relative to
-CRS}\label{performance-of-jso-relative-to-crs}}
+CRS}\label{performance-of-jso-relative-to-crs}
 
@@
 
-\hypertarget{factors-affecting-jso-success-rate-index-properties-and-jellyfish-hyper-parameters}{%
 \subsection{Factors affecting JSO success rate: index properties and
 jellyfish
-hyper-parameters}\label{factors-affecting-jso-success-rate-index-properties-and-jellyfish-hyper-parameters}}
+hyper-parameters}\label{factors-affecting-jso-success-rate-index-properties-and-jellyfish-hyper-parameters}
 
@@
 For simpler problems (6 dimensions), small parameter values (20
-jellyfishes and a maximum of 50 tries) can already achieve a high
-success rate. However, larger parameter values (i.e.~100 jellyfishes and
-a maximum of 100 tries) are necessary for higher-dimensional problems
-(8, 10, and 12 dimensions). Increasing both parameters enhances the
-performance of JSO, but it also extends the computational time required
-for the optimisation, which can be computationally intensive when
-evaluating the index function (such as scagnostic indexes) multiple
-times across numerous iterations.
+jellyfish and a maximum of 50 tries) can already achieve a high success
+rate. However, larger parameter values (i.e.~100 jellyfish and a maximum
+of 100 tries) are necessary for higher-dimensional problems (8, 10, and
+12 dimensions). Increasing both parameters enhances the performance of
+JSO, but it also extends the computational time required for the
+optimisation, which can be computationally intensive when evaluating the
+index function (such as scagnostic indexes) multiple times across
+numerous iterations.
 
@@
 to find the pipe shape using the holes index in the 6, 8, 10, and
-12-dimensional spaces. The JSO uses 100 jellyfishes and a maximum number
+12-dimensional spaces. The JSO uses 100 jellyfish and a maximum number
 of tries of 100. The CRS uses a maximum of 1000 tries in each step of
@@
 and allowing more tries also does not affect success rate. An increase
-in squintability of one unit increases the success rate by 785\%. As
+in squintability of one unit increases the success rate by 1235\%. As
 dimension increases by one the success rate halves. Increasing the
-number of jellies by 10, increases the succes rate by 24\%.
-
-\hypertarget{tbl-smoothness-squintability}{}
+number of jellies by 10, increases the succes rate by 27\%.
+
 \begin{table}
-\caption{\label{tbl-smoothness-squintability}Parameters estimated from the Gaussian process (outputscale \(\eta\),
-lengthscale \(\ell\), smoothness \(\nu\), and nugget \(\sigma\)) and
-scaled logistic function (\(\theta_1\) to \(\theta_4\) and
-\(\varsigma\)) for the pipe-finding and sine-wave finding problems. The
-columns \(\nu\) and \(\varsigma\) represent the smoothness and
-squintability measures respectively. }\tabularnewline
+
+\caption{\label{tbl-smoothness-squintability}Parameters estimated from
+the Gaussian process (outputscale \(\eta\), lengthscale \(\ell\),
+smoothness \(\nu\), and nugget \(\sigma\)) and scaled logistic function
+(\(\theta_1\) to \(\theta_4\) and \(\varsigma\)) for the pipe-finding
+and sine-wave finding problems. The columns \(\nu\) and \(\varsigma\)
+represent the smoothness and squintability measures respectively.}
+
+\centering{
 
@@
 \midrule
-1 & pipe & holes & 6 & 0.04 & 0.37 & 2.36 & 0.22 & 1.02 & 0.86 & 3.37 & 0.02 & 3.05\\
-2 & pipe & holes & 8 & 0.01 & 0.18 & 2.19 & 0.82 & 1.01 & 0.87 & 3.26 & 0.03 & 2.96\\
-3 & pipe & holes & 10 & 0.01 & 0.11 & 2.19 & 3.03 & 1.02 & 0.88 & 3.15 & 0.02 & 2.95\\
-4 & pipe & holes & 12 & 0.01 & 0.15 & 2.29 & 1.58 & 1.01 & 0.88 & 3.34 & 0.00 & 3.12\\
-5 & sine & MIC & 6 & 0.13 & 0.09 & 2.46 & 0.08 & 0.89 & 0.57 & 1.62 & -0.02 & 1.26\\
-6 & sine & MIC & 8 & 0.13 & 0.08 & 2.64 & 0.08 & 0.93 & 0.33 & 1.31 & -0.03 & 0.77\\
-7 & sine & TIC & 6 & 0.35 & 0.11 & 2.44 & 0.09 & 0.95 & 0.54 & 1.72 & -0.03 & 1.32\\
-8 & sine & TIC & 8 & 0.35 & 0.10 & 2.47 & 0.09 & 0.95 & 0.56 & 1.72 & -0.03 & 1.37\\
-9 & sine & dcor & 6 & 0.18 & 0.17 & 2.66 & 0.11 & 0.95 & 1.04 & 2.74 & -0.02 & 2.95\\
-10 & sine & loess & 6 & 0.28 & 0.34 & 1.99 & 0.31 & 1.02 & 1.04 & 2.65 & 0.08 & 2.76\\
-11 & sine & splines & 6 & 0.20 & 0.24 & 2.54 & 0.10 & 1.01 & 1.05 & 2.73 & -0.01 & 3.12\\
-12 & sine & stringy & 6 & 0.00 & 0.01 & 1.54 & 38.17 & 1.05 & 0.01 & 254.73 & 0.05 & 2.96\\
+1 & pipe & holes & 6 & 0.03 & 0.39 & 3.08 & 0.0 & 1.02 & 0.86 & 3.37 & 0.02 & 3.05\\
+2 & pipe & holes & 8 & 0.02 & 0.32 & 2.86 & 0.0 & 1.01 & 0.87 & 3.26 & 0.03 & 2.96\\
+3 & pipe & holes & 10 & 0.02 & 0.36 & 2.80 & 0.0 & 1.02 & 0.88 & 3.15 & 0.02 & 2.95\\
+4 & pipe & holes & 12 & 0.02 & 0.36 & 2.82 & 0.0 & 1.01 & 0.88 & 3.34 & 0.00 & 3.12\\
+5 & sine & MIC & 6 & 0.06 & 0.29 & 3.15 & 0.0 & 0.89 & 0.57 & 1.62 & -0.02 & 1.26\\
+6 & sine & MIC & 8 & 0.06 & 0.33 & 3.09 & 0.0 & 0.93 & 0.33 & 1.31 & -0.03 & 0.77\\
+7 & sine & TIC & 6 & 0.17 & 0.34 & 3.06 & 0.0 & 0.95 & 0.54 & 1.72 & -0.03 & 1.32\\
+8 & sine & TIC & 8 & 0.16 & 0.31 & 3.17 & 0.0 & 0.95 & 0.56 & 1.72 & -0.03 & 1.37\\
+9 & sine & dcor & 6 & 0.09 & 0.31 & 3.06 & 0.0 & 0.95 & 1.04 & 2.74 & -0.02 & 2.95\\
+10 & sine & loess & 6 & 0.24 & 0.45 & 3.53 & 0.0 & 1.02 & 1.04 & 2.65 & 0.08 & 2.76\\
+11 & sine & splines & 6 & 0.10 & 0.32 & 3.11 & 0.0 & 1.01 & 1.05 & 2.73 & -0.01 & 3.12\\
+12 & sine & stringy & 6 & 0.00 & 1128.28 & 1.03 & 134.8 & 1.05 & 0.01 & 254.73 & 0.05 & 2.96\\
 \bottomrule
@@
 \endgroup{}
-\end{table}
-
-\hypertarget{tbl-mod-output}{}
+
+}
+
+\end{table}%
+
 \begingroup
@@
 \setlength{\LTpost}{0mm}
+
 \begin{longtable}{lcc}
-\caption{\label{tbl-mod-output}Jellyfish success rate relative to index properties and jellyfish
-hyper-parameters. This is the summary from a logistic regression fit to
-smoothness, squintability, dimension, running time, number of jellyfish
-and maximum number of tries. Interestingly, squintability and dimension
-strongly affect jellyfish optimisation success. The number of jellies
-marginally affects success, but index smoothness, running longer and
-increasing the number of tries do not. }\tabularnewline
+
+\caption{\label{tbl-mod-output}Jellyfish success rate relative to index
+properties and jellyfish hyper-parameters. This is the summary from a
+logistic regression fit to smoothness, squintability, dimension, running
+time, number of jellyfish and maximum number of tries. Interestingly,
+squintability and dimension strongly affect jellyfish optimisation
+success. The number of jellies marginally affects success, but index
+smoothness, running longer and increasing the number of tries do not.}
+
+\tabularnewline
 
@@
 \midrule\addlinespace[2.5pt]
-Smoothness & 3.30 (0.08 to 274) & 0.533 \\ 
-Squintability & 7.85 (2.40 to 37.7) & {\bfseries 0.003} \\ 
-Dimension & 0.53 (0.30 to 0.83) & {\bfseries 0.014} \\ 
-Long runtime & 0.42 (0.03 to 5.44) & 0.497 \\ 
-Number of jellyfish & 1.24 (0.98 to 1.63) & {\bfseries 0.089} \\ 
-Maximum number of tries & 1.12 (0.84 to 1.52) & 0.452 \\ 
+Smoothness & 89.1 (1.23 to 4,865,159) & 0.289 \\ 
+Squintability & 12.4 (3.08 to 125) & {\bfseries 0.006} \\ 
+Dimension & 0.57 (0.30 to 0.96) & {\bfseries 0.048} \\ 
+Long runtime & 0.17 (0.00 to 4.96) & 0.357 \\ 
+Number of jellyfish & 1.27 (0.99 to 1.68) & {\bfseries 0.069} \\ 
+Maximum number of tries & 1.10 (0.82 to 1.52) & 0.524 \\ 
 \bottomrule
+
 \end{longtable}
+
 \begin{minipage}{\linewidth}
@@
 
-\hypertarget{sec-discussion}{%
-\section{Practicalities}\label{sec-discussion}}
+\section{Practicalities}\label{sec-discussion}
 
@@
 
-\hypertarget{using-the-jso-in-a-ppgt}{%
-\subsection{Using the JSO in a PPGT}\label{using-the-jso-in-a-ppgt}}
+\subsection{Using the JSO in a PPGT}\label{using-the-jso-in-a-ppgt}
 
@@
 
-\hypertarget{computing-the-metrics-for-your-new-index}{%
 \subsection{Computing the metrics for your new
-index}\label{computing-the-metrics-for-your-new-index}}
+index}\label{computing-the-metrics-for-your-new-index}
 
@@
 
-\hypertarget{smoothness}{%
-\subsubsection{Smoothness}\label{smoothness}}
+\subsubsection{Smoothness}\label{smoothness}
 
@@
 
-\hypertarget{squintability}{%
-\subsubsection{Squintability}\label{squintability}}
+\subsubsection{Squintability}\label{squintability}
 
@@
 
-\hypertarget{supplementary-materials}{%
-\section*{Supplementary materials}\label{supplementary-materials}}
+\section*{Supplementary materials}\label{supplementary-materials}
 \addcontentsline{toc}{section}{Supplementary materials}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cook1995grand}{}}%
+\bibitem[\citeproctext]{ref-cook1995grand}
 Cook, D., A. Buja, J. Cabrera, and C. Hurley. 1995. {``Grand Tour and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-DLMF}{}}%
+\bibitem[\citeproctext]{ref-DLMF}
 DLMF. 2024. {``{NIST Digital Library of Mathematical Functions}.''}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-FT74}{}}%
+\bibitem[\citeproctext]{ref-FT74}
 Friedman, J. H., and J. W. Tukey. 1974. {``A Projection Pursuit
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-huber85}{}}%
+\bibitem[\citeproctext]{ref-huber85}
 Huber, Peter J. 1985. {``Projection Pursuit.''} \emph{The Annals of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-huberplot}{}}%
+\bibitem[\citeproctext]{ref-huberplot}
 ---------. 1990. {``Data Analysis and Projection Pursuit.''} Technical
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-posse95}{}}%
+\bibitem[\citeproctext]{ref-posse95}
 Posse, Christian. 1995. {``Projection Pursuit Exploratory Data

### 2024-09-18 code/07-calc-smoothness-squintability.R

@@ -18,7 +18,7 @@ smoothness_holes <- tibble::tibble(
               matrix(c(rep(0, 20), 1, 0, 0, 1), nrow = 12, byrow = TRUE))) |>
   dplyr::mutate(basis_df = purrr::pmap(
     list(data, best),
-    function(data, best, n){sample_bases("holes", data = data, best = best)}))
+    function(data, best, n){sample_bases("holes", data = data, best = best, n_basis = 500)}))
 
 holes_tidy <- smoothness_holes |>
   rowwise() |>
@@ -27,12 +27,12 @@ holes_tidy <- smoothness_holes |>
 
 idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "stringy", "splines2d")
 smoothness_sine <- tibble::tibble(
-  n = 6, index = idx_names[6], data = list(sine1000),
+  n = 6, index = idx_names, data = list(sine1000),
   best = list(matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE))) |>
   dplyr::bind_rows(sine_8d_tbl("MIC")) |>
   dplyr::bind_rows(sine_8d_tbl("TIC")) |>
   rowwise() |>
-  dplyr::mutate(basis_df = list(sample_bases(index)))
+  dplyr::mutate(basis_df = list(sample_bases(index, n_basis = 500)))
 
 sine_tidy <- smoothness_sine |>
   rowwise() |>

### 2024-09-18 jso.qmd
update diagram

@@
 outlined in @sec-smoothness and @sec-squintability and as illustrated in
-@fig-smoothness and @fig-squintability. To compute smoothness, 300 random
-bases are simulated. Index values and
-projection distance (to the optimal basis) are calculated for each
+@fig-smoothness and @fig-squintability. To compute smoothness, 500 random
+bases are simulated. Index values are calculated for each
 random basis before fitting the Gaussian process model to obtain the
@@
 #| fig.cap: "Illustration of steps to calculate smoothness.
-For a given projection pursuit problem defined by the shape to find, data dimension and the index function, 1) sample random bases given the orthonormality contraint, 2) calculate the projection distance and the index value for each random basis, and 3) fit a Gaussian process model of index values against projection distances to obtain the smoothness measure.
+For a given projection pursuit problem defined by the shape to find, data dimension and the index function, 1) sample random bases given the orthonormality contraint and calculate the index value associated with each random basis, and 2) fit a Gaussian process model of index values against projection distances to obtain the smoothness measure.
 "

### 2024-09-18 jso.tex
update diagram

@@
 
-\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, rightrule=.15mm, arc=.35mm, colback=white, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, leftrule=.75mm, breakable, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, opacityback=0, left=2mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, bottomtitle=1mm, opacityback=0, left=2mm, toptitle=1mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, coltitle=black, titlerule=0mm, breakable, arc=.35mm, colback=white, bottomrule=.15mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white]
 
@@
 Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}. To
-compute smoothness, 300 random bases are simulated. Index values and
-projection distance (to the optimal basis) are calculated for each
-random basis before fitting the Gaussian process model to obtain the
-smoothness measure for the index.
+compute smoothness, 500 random bases are simulated. Index values are
+calculated for each random basis before fitting the Gaussian process
+model to obtain the smoothness measure for the index.
 
@@
 to find, data dimension and the index function, 1) sample random bases
-given the orthonormality contraint, 2) calculate the projection distance
-and the index value for each random basis, and 3) fit a Gaussian process
-model of index values against projection distances to obtain the
-smoothness measure.}
+given the orthonormality contraint and calculate the index value
+associated with each random basis, and 2) fit a Gaussian process model
+of index values against projection distances to obtain the smoothness
+measure.}
 

### 2024-09-18 jso.qmd
use r to denote distance since d is used for dimension

@@
 Let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.
-The projection distance $d(A, A^*)$ between $A$
+The projection distance $r(A, A^*)$ between $A$
 and $A^*$ is defined as 
-$d(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\,\rVert _F$
+$r(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\,\rVert _F$
 where $\lVert . \rVert _F$ denotes the Frobenius norm, given by
@@
 Squintability can be defined as how the index value $f(XA)$ changes with
-respect to the projection distance $d(A, A^*)$, over the course of the
+respect to the projection distance $r(A, A^*)$, over the course of the
 JSO:
@@
 \begin{defn}[squintability]\label{def:squintability}
-Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance $d(A, A^*)$ to the index value $f(XA)$, such that $g(d(A, A^*)) = f(XA)$.
-For brevity, denote $g(d)$ as $g(d(A, A^*))$.
+Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance $r(A, A^*)$ to the index value $f(XA)$, such that $g(r(A, A^*)) = f(XA)$.
+For brevity, denote $g(r)$ as $g(r(A, A^*))$.
 The squintability of an index function $f(XA)$ is defined as
@@
 \begin{equation}
-\varsigma(f) = -4 \times \min_{d} g'(d) \times \arg \min_{d} g'(d)
+\varsigma(f) = -4 \times \min_{r} g'(r) \times \arg \min_{r} g'(r)
 \label{eq-squintability}
@@
 
-where $4$ is a constant scaling factor, $-\min_d g'(d)$ represents the 
-largest gradient of $-g$ and $\arg \min_{d} g'(d)$ represents the projection distance at which this largest gradient is attained.
-If the set $\mathcal{D}:=\arg \min_{d} g'(d)$ contains more than one value $d$, the largest value $\max(\mathcal{D})$ is used in equation \eqref{eq-squintability}.
+where $4$ is a constant scaling factor, $-\min_r g'(r)$ represents the 
+largest gradient of $-g$ and $\arg \min_{r} g'(r)$ represents the projection distance at which this largest gradient is attained.
+If the set $\mathcal{D}:=\arg \min_{r} g'(r)$ contains more than one value $r$, the largest value $\max(\mathcal{D})$ is used in equation \eqref{eq-squintability}.
 
@@
 \begin{proof}
-Let $d_1 := \underset{d}{\arg \min }\,g_1'(d)$ and $d_2 := \underset{d}{\arg \min }\,g_2'(d)$.
-Since $g_1$ is convex, $g_1''(d) >= 0 \; \forall d \in [0, D]$, meaning that $g_1'$ is an increasing function, taking its minimum for $d_1=0$.
+Let $r_1 := \underset{r}{\arg \min }\,g_1'(r)$ and $r_2 := \underset{r}{\arg \min }\,g_2'(r)$.
+Since $g_1$ is convex, $g_1''(r) >= 0 \; \forall r \in [0, D]$, meaning that $g_1'$ is an increasing function, taking its minimum for $r_1=0$.
 As a result, $\varsigma(g_1)=-4 \times g'(0) \times 0 =0$.
-Since $g_2$ is concave, $g_2''(d) <= 0 \; \forall d \in [0, D]$, meaning that $g_2'$ is a decreasing function, taking its minimum for $d_2=D>0$.
-Since $g_2$ is strictly decreasing, $g_2'(d) < 0 \; \forall d \in [0, D]$.
+Since $g_2$ is concave, $g_2''(r) <= 0 \; \forall r \in [0, D]$, meaning that $g_2'$ is a decreasing function, taking its minimum for $r_2=D>0$.
+Since $g_2$ is strictly decreasing, $g_2'(r) < 0 \; \forall r \in [0, D]$.
 In particular $g_2'(D)<0$.

### 2024-09-18 jso.tex
use r to denote distance since d is used for dimension

@@
 Squintability can be defined as how the index value \(f(XA)\) changes
-with respect to the projection distance \(d(A, A^*)\), over the course
+with respect to the projection distance \(r(A, A^*)\), over the course
 of the JSO:
@@
 
-\begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, bottomtitle=1mm, opacityback=0, left=2mm, toptitle=1mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, coltitle=black, titlerule=0mm, breakable, arc=.35mm, colback=white, bottomrule=.15mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, colback=white, bottomtitle=1mm, breakable, colframe=quarto-callout-note-color-frame, left=2mm, toptitle=1mm, arc=.35mm, opacityback=0, leftrule=.75mm, rightrule=.15mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, coltitle=black, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white]
 

### 2024-09-23 jso.qmd
change \mathbf{R} to \mathbb{R}

@@
 \begin{defn}
-Let $\mathbf{A}=[A_1, \ldots, A_N] \in (\mathbf{R}^{p \times d})^N$ be d-dimensional projection bases, let $\mathbf{y}=[f(XA_1),\ldots,f(XA_N)]$ be the corresponding PP index values, and let $\mathbf{K}=[K_\theta(A_{i}-A_{j})]_{1\leq i,j\leq N}\in\mathbb{R}^{N\times N}$ be the Matérn covariance matrix evaluated at the input bases, where the vector $\theta$ contains all the parameters of the multivariate Matérn covariance function $K$ (smoothness, outputscale, lengthscales).
+Let $\mathbf{A}=[A_1, \ldots, A_N] \in (\mathbb{R}^{p \times d})^N$ be d-dimensional projection bases, let $\mathbf{y}=[f(XA_1),\ldots,f(XA_N)]$ be the corresponding PP index values, and let $\mathbf{K}=[K_\theta(A_{i}-A_{j})]_{1\leq i,j\leq N}\in\mathbb{R}^{N\times N}$ be the Matérn covariance matrix evaluated at the input bases, where the vector $\theta$ contains all the parameters of the multivariate Matérn covariance function $K$ (smoothness, outputscale, lengthscales).
 The log-likelihood of the parameters $\theta$ is defined by
@@
 \begin{defn}[projection distance]\label{def:proj-dist}
-Recall that $A \in \mathbf{R}^{p \times d}$ is a $d$-dimensional orthonormal matrix.
+Recall that $A \in \mathbb{R}^{p \times d}$ is a $d$-dimensional orthonormal matrix.
 Let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.

### 2024-09-23 jso.tex
change \mathbf{R} to \mathbb{R}

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, colback=white, bottomtitle=1mm, breakable, colframe=quarto-callout-note-color-frame, left=2mm, toptitle=1mm, arc=.35mm, opacityback=0, leftrule=.75mm, rightrule=.15mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, coltitle=black, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, opacitybacktitle=0.6, breakable, coltitle=black, titlerule=0mm, bottomrule=.15mm, arc=.35mm, rightrule=.15mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, colback=white, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, left=2mm, toptitle=1mm]
 

### 2024-09-23 jso.qmd
inequality signs

@@
 Let $r_1 := \underset{r}{\arg \min }\,g_1'(r)$ and $r_2 := \underset{r}{\arg \min }\,g_2'(r)$.
-Since $g_1$ is convex, $g_1''(r) >= 0 \; \forall r \in [0, D]$, meaning that $g_1'$ is an increasing function, taking its minimum for $r_1=0$.
+Since $g_1$ is convex, $g_1''(r) \geq 0 \; \forall r \in [0, D]$, meaning that $g_1'$ is an increasing function, taking its minimum for $r_1=0$.
 As a result, $\varsigma(g_1)=-4 \times g'(0) \times 0 =0$.
-Since $g_2$ is concave, $g_2''(r) <= 0 \; \forall r \in [0, D]$, meaning that $g_2'$ is a decreasing function, taking its minimum for $r_2=D>0$.
+Since $g_2$ is concave, $g_2''(r) \leq 0 \; \forall r \in [0, D]$, meaning that $g_2'$ is a decreasing function, taking its minimum for $r_2=D>0$.
 Since $g_2$ is strictly decreasing, $g_2'(r) < 0 \; \forall r \in [0, D]$.

### 2024-09-23 jso.tex
inequality signs

@@
 \end{keywords}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, boxrule=0pt, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 described below. Define
-\begin{equation}\phantomsection\label{eq-logistic}{
+\begin{equation}\protect\hypertarget{eq-logistic}{}{
 \ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
-}\end{equation}
+}\label{eq-logistic}\end{equation}
 
@@
 \(\ell(\theta_{2})=\frac{1}{2}\). Then, define
-\begin{equation}\phantomsection\label{eq-parametric}{
+\begin{equation}\protect\hypertarget{eq-parametric}{}{
 g(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
-}\end{equation}
+}\label{eq-parametric}\end{equation}
 
@@
 can be shown to be equal to
-\begin{equation}\phantomsection\label{eq-squintability-parametric}{
+\begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
 \varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
-}\end{equation}
+}\label{eq-squintability-parametric}\end{equation}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, opacitybacktitle=0.6, breakable, coltitle=black, titlerule=0mm, bottomrule=.15mm, arc=.35mm, rightrule=.15mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, colback=white, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, left=2mm, toptitle=1mm]
+\begin{tcolorbox}[enhanced jigsaw, opacityback=0, left=2mm, bottomtitle=1mm, breakable, bottomrule=.15mm, titlerule=0mm, opacitybacktitle=0.6, toprule=.15mm, colback=white, toptitle=1mm, leftrule=.75mm, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, colframe=quarto-callout-note-color-frame, coltitle=black, colbacktitle=quarto-callout-note-color!10!white]
 
@@
 \endgroup{}
-
-}
-
-\end{table}%
-
+\end{table}
+
+\hypertarget{tbl-mod-output}{}
 \begingroup
@@
 \setlength{\LTpost}{0mm}
-
 \begin{longtable}{lcc}
-
-\caption{\label{tbl-mod-output}Jellyfish success rate relative to index
-properties and jellyfish hyper-parameters. This is the summary from a
-logistic regression fit to smoothness, squintability, dimension, running
-time, number of jellyfish and maximum number of tries. Interestingly,
-squintability and dimension strongly affect jellyfish optimisation
-success. The number of jellies marginally affects success, but index
-smoothness, running longer and increasing the number of tries do not.}
-
-\tabularnewline
+\caption{\label{tbl-mod-output}Jellyfish success rate relative to index properties and jellyfish
+hyper-parameters. This is the summary from a logistic regression fit to
+smoothness, squintability, dimension, running time, number of jellyfish
+and maximum number of tries. Interestingly, squintability and dimension
+strongly affect jellyfish optimisation success. The number of jellies
+marginally affects success, but index smoothness, running longer and
+increasing the number of tries do not. }\tabularnewline
 
@@
 \bottomrule
-
 \end{longtable}
-
 \begin{minipage}{\linewidth}
@@
 
-\section{Practicalities}\label{sec-discussion}
+\hypertarget{sec-discussion}{%
+\section{Practicalities}\label{sec-discussion}}
 
@@
 
-\bibitem[\citeproctext]{ref-FT74}
+\leavevmode\vadjust pre{\hypertarget{ref-FT74}{}}%
 Friedman, J. H., and J. W. Tukey. 1974. {``A Projection Pursuit
@@
 
-\bibitem[\citeproctext]{ref-huber85}
+\leavevmode\vadjust pre{\hypertarget{ref-huber85}{}}%
 Huber, Peter J. 1985. {``Projection Pursuit.''} \emph{The Annals of
@@
 
-\bibitem[\citeproctext]{ref-posse95}
+\leavevmode\vadjust pre{\hypertarget{ref-posse95}{}}%
 Posse, Christian. 1995. {``Projection Pursuit Exploratory Data

### 2024-09-24 jso.qmd
change D -> R in proposition 1

--- 2024-09-23 jso.qmd
@@
 \begin{prop}\label{prop:convex-concave}
-Let $g_1$ be a convex, strictly decreasing function and $g_2$ be a concave, strictly decreasing function, both defined on $[0, D]$ where $D>0$, and taking values in $[0, 1]$. Then
+Let $g_1$ be a convex, strictly decreasing function and $g_2$ be a concave, strictly decreasing function, both defined on $[0, R]$ where $R>0$, and taking values in $[0, 1]$. Then
 $$\varsigma(g_1) < \varsigma(g_2)$$
@@
 Let $r_1 := \underset{r}{\arg \min }\,g_1'(r)$ and $r_2 := \underset{r}{\arg \min }\,g_2'(r)$.
-Since $g_1$ is convex, $g_1''(r) >= 0 \; \forall r \in [0, D]$, meaning that $g_1'$ is an increasing function, taking its minimum for $r_1=0$.
+Since $g_1$ is convex, $g_1''(r) \geq 0 \; \forall r \in [0, R]$, meaning that $g_1'$ is an increasing function, taking its minimum for $r_1=0$.
 As a result, $\varsigma(g_1)=-4 \times g'(0) \times 0 =0$.
-Since $g_2$ is concave, $g_2''(r) <= 0 \; \forall r \in [0, D]$, meaning that $g_2'$ is a decreasing function, taking its minimum for $r_2=D>0$.
-Since $g_2$ is strictly decreasing, $g_2'(r) < 0 \; \forall r \in [0, D]$.
-In particular $g_2'(D)<0$.
-As a result $\varsigma(g_2)=-4 \times g'(D) \times D >0$.
+Since $g_2$ is concave, $g_2''(r) \leq 0 \; \forall r \in [0, R]$, meaning that $g_2'$ is a decreasing function, taking its minimum for $r_2=D>0$.
+Since $g_2$ is strictly decreasing, $g_2'(r) < 0 \; \forall r \in [0, R]$.
+In particular $g_2'(R)<0$.
+As a result $\varsigma(g_2)=-4 \times g'(R) \times R >0$.
 This proves that $\varsigma(g_1) < \varsigma(g_2)$.

### 2024-09-24 jso.tex
change D -> R in proposition 1

--- 2024-09-23 jso.tex
@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, opacitybacktitle=0.6, breakable, coltitle=black, titlerule=0mm, bottomrule=.15mm, arc=.35mm, rightrule=.15mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, colback=white, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, left=2mm, toptitle=1mm]
+\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, breakable, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, toprule=.15mm, left=2mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, colback=white, leftrule=.75mm, toptitle=1mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, coltitle=black]
 

### 2024-10-07 jso.qmd
line wrap with visual mode

@@
 
-Projection pursuit (PP) (@kr69, @FT74, @huber85) is a dimension reduction technique aimed at identifying informative linear projections of data.
-This is useful for exploring high-dimensional data, and creating plots of the data that reveal the main features to use for publication.
-The method involves optimising an objective function known as the PP index (e.g.
-@hall1989polynomial, @cook1993projection,
-@lee2010projection, @Loperfido2018, @Loperfido2020), which defines the criterion for what constitutes interesting or informative projections.
-Let $X \in \mathbb{R}^{n\times p}$ be the data matrix, $A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix, where $A$ belongs to the Stiefel manifold $\mathcal{A} = V_d(\mathbb{R}^p)$.
-The projection $Y = XA$ is a linear transformation that maps data from a $p$-dimensional space into a $d$-dimensional space.
-The index function $f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ is a scalar function that measures an interesting aspect of the projected data, such as deviation from normality, presence of clusters, non-linear structure, or other features of interest.
-For a fixed sample of data, PP finds the orthonormal basis $A$ that maximises the index value of the projection, $Y = XA$:
+Projection pursuit (PP) (@kr69, @FT74, @huber85) is a dimension
+reduction technique aimed at identifying informative linear projections
+of data. This is useful for exploring high-dimensional data, and
+creating plots of the data that reveal the main features to use for
+publication. The method involves optimising an objective function known
+as the PP index (e.g. @hall1989polynomial, @cook1993projection,
+@lee2010projection, @Loperfido2018, @Loperfido2020), which defines the
+criterion for what constitutes interesting or informative projections.
+Let $X \in \mathbb{R}^{n\times p}$ be the data matrix,
+$A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix, where $A$
+belongs to the Stiefel manifold $\mathcal{A} = V_d(\mathbb{R}^p)$. The
+projection $Y = XA$ is a linear transformation that maps data from a
+$p$-dimensional space into a $d$-dimensional space. The index function
+$f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ is a scalar function
+that measures an interesting aspect of the projected data, such as
+deviation from normality, presence of clusters, or non-linear structure.
+For a fixed sample of data, PP finds the orthonormal basis $A$ that
+maximises the index value of the projection, $Y = XA$:
 
@@
 
-It is interesting to note that when using PP visually, one cares less about $A$ than the plane described by $A$, because the orientation in the plane is irrelevant.
-The space of planes belongs to a Grassmann manifold.
-This is usually how the projection pursuit guided tour (PPGT) [@cook1995grand] operates, when using geodesic interpolation between starting and target planes.
-It interpolates plane to plane, removing irrelevant within plane spin, and is agnostic to the basis ($A$) used to define the plane.
-Thus, indexes which are used for the PPGT should be rotationally invariant.
-
-Index functions are quite varied in form, partially depending on the data that is being projected.
-@fig-example-functions shows two examples.
-Huber plots [@huberplot] of 2D data sets are in (a) and (c), showing the PP index values for all 1D projections of the 2D data in polar coordinates, which reveals the form of these functions.
-The dashed circle is a baseline set at the average value, and the straight line marks the optimal projection.
-Plots (b) and (d) show the respective best projections of the data as histograms.
-Indexes like the holes, central mass and skewness [@cook1993projection] are generally smooth for most data sets, but capture only large patterns.
-Many indexes are noisy and non-convex, requiring an effective and efficient optimisation procedure to explore the data landscape and achieve a globally optimal viewpoint of the data.
-The skewness index computed for trimodal data, in (a), is smooth with a large squint angle but has three modes, and thus is not convex.
-The binned normality index (a simple version of a non-normality index as described in @huber85) computed on the famous RANDU data, in (c), is noisier and has a very small squint angle.
-The discreteness cannot be seen unless the optimiser is very close to the optimal projection.
+It is interesting to note that when using PP visually, one cares less
+about $A$ than the plane described by $A$, because the orientation in
+the plane is irrelevant. The space of planes belongs to a Grassmann
+manifold. This is usually how the projection pursuit guided tour (PPGT)
+[@cook1995grand] operates, when using geodesic interpolation between
+starting and target planes. It interpolates plane to plane, removing
+irrelevant within plane spin, and is agnostic to the basis ($A$) used to
+define the plane. Thus, indexes which are used for the PPGT should be
+rotationally invariant.
+
+Index functions are quite varied in form, partially depending on the
+data that is being projected. @fig-example-functions shows two examples.
+Huber plots [@huberplot] of 2D data sets are in (a) and (c), showing the
+PP index values for all 1D projections of the 2D data in polar
+coordinates, which reveals the form of these functions. The dashed
+circle is a baseline set at the average value, and the straight line
+marks the optimal projection. Plots (b) and (d) show the respective best
+projections of the data as histograms. Indexes like the holes, central
+mass and skewness [@cook1993projection] are generally smooth for most
+data sets, but capture only large patterns. Many indexes are noisy and
+non-convex, requiring an effective and efficient optimisation procedure
+to explore the data landscape and achieve a globally optimal viewpoint
+of the data. The skewness index computed for trimodal data, in (a), is
+smooth with a large squint angle but has three modes, and thus is not
+convex. The binned normality index (a simple version of a non-normality
+index as described in @huber85) computed on the famous RANDU data, in
+(c), is noisier and has a very small squint angle. The discreteness
+cannot be seen unless the optimiser is very close to the optimal
+projection.
 
@@
 
-Optimisation of PP is often discussed when new indexes are proposed [@posse95; @marie-sainte2010; @grochowski2011].
-@cook1995grand tied the optimisation more closely to the index, when they introduced the PPGT, which monitors the optimisation visually so that the user can see the projected data leading in and out of the optimum.
-An implementation is available in the `tourr` package [@tourr] in R [@R].
-@RJ-2021-105 illustrated how to diagnose optimisation processes, particularly focusing on the guided tour, and revealed a need for improved optimisation.
-While improving the quality of the optimisation solutions in the tour is essential, it is also important to be able to view the data projections as the optimisation progresses.
-Integrating the guided tour with a global optimisation algorithm that is efficient in finding the global optimal and enables viewing of the projected data during the exploration process is a goal.
-
-Here, the potential for a Jellyfish Search Optimiser (JSO) (see @chou_novel_2021, and @rajwar_exhaustive_2023) for the PPGT is explored.
-JSO, inspired by the search behaviour of jellyfish in the ocean, is a swarm-based metaheuristic designed to solve global optimisation problems.
-Compared to traditional methods, JSO has demonstrated stronger search ability and faster
-convergence, and requires fewer tuning parameters. These practicalities
-make JSO a promising candidate for enhancing PP optimisation.
+Optimisation of PP is often discussed when new indexes are proposed
+[@posse95; @marie-sainte2010; @grochowski2011]. @cook1995grand tied the
+optimisation more closely to the index, when they introduced the PPGT,
+which monitors the optimisation visually so that the user can see the
+projected data leading in and out of the optimum. An implementation is
+available in the `tourr` package [@tourr] in R [@R]. @RJ-2021-105
+illustrated how to diagnose optimisation processes, particularly
+focusing on the guided tour, and revealed a need for improved
+optimisation. While improving the quality of the optimisation solutions
+in the tour is essential, it is also important to be able to view the
+data projections as the optimisation progresses. Integrating the guided
+tour with a global optimisation algorithm that is efficient in finding
+the global optimal and enables viewing of the projected data during the
+exploration process is a goal.
+
+Here, the potential for a Jellyfish Search Optimiser (JSO) (see
+@chou_novel_2021, and @rajwar_exhaustive_2023) for the PPGT is explored.
+JSO, inspired by the search behaviour of jellyfish in the ocean, is a
+swarm-based metaheuristic designed to solve global optimisation
+problems. Compared to traditional methods, JSO has demonstrated stronger
+search ability and faster convergence, and requires fewer tuning
+parameters. These practicalities make JSO a promising candidate for
+enhancing PP optimisation.
 
@@
 performance of JSO in PP optimisation for the guided tour. It is of
-interest to assess how quickly and closely the optimiser reaches a global
-optimum, for various PP indexes that may have differing complexities. To
-observe the performance of JSO with different types of PP indexes,
-metrics are introduced to capture specific properties of the index
-including squintability (based on @barnett1981interpreting's squint
-angle) and smoothness.
-Here, we mathematically define metrics for squintability and smoothness, which is a new contribution for PP research.
-A series of simulation experiments using various datasets and PP indexes are conducted to assess JSO's behaviour and its sensitivity to hyper-parameter choices (number of jellyfish and maximum number of tries).
-The relationship between the JSO performance, hyper-parameter choices and properties of PP indexes (smoothness and squintability) is analysed to provide guidance on selecting optimisers for practitioners using projection pursuit.
-Additionally, this work should guide the design of new PP indexes and facilitate better optimization for PP.
+interest to assess how quickly and closely the optimiser reaches a
+global optimum, for various PP indexes that may have differing
+complexities. To observe the performance of JSO with different types of
+PP indexes, metrics are introduced to capture specific properties of the
+index including squintability (based on @barnett1981interpreting's
+squint angle) and smoothness. Here, we mathematically define metrics for
+squintability and smoothness, which is a new contribution for PP
+research. A series of simulation experiments using various datasets and
+PP indexes are conducted to assess JSO's behaviour and its sensitivity
+to hyper-parameter choices (number of jellyfish and maximum number of
+tries). The relationship between the JSO performance, hyper-parameter
+choices and properties of PP indexes (smoothness and squintability) is
+analysed to provide guidance on selecting optimisers for practitioners
+using projection pursuit. Additionally, this work should guide the
+design of new PP indexes and facilitate better optimization for PP.
 
 The paper is structured as follows. @sec-background introduces the
-background of the PPGT, reviews existing optimisers and index
-functions in the literature. @sec-PP-properties introduces the 
-metrics that measure different properties of PP indexes, smoothness and
-squintability, and @sec-JSO describes the new JSO to be used for the PPGT.
-@sec-sim-deets summarises the results of two simulation experiments done to assess JSO's performance: one comparing JSO's performance improvements relative to an existing optimiser, Creeping Random Search (CRS), and the other studying the impact of different PP index properties on optimisation performance.
-@sec-sim-res presents the results.
-@sec-discussion discusses implementation details and insights for practitioners.
-@sec-conclusion summarises the work and provides suggestions for future
-directions.
+background of the PPGT, reviews existing optimisers and index functions
+in the literature. @sec-PP-properties introduces the metrics that
+measure different properties of PP indexes, smoothness and
+squintability, and @sec-JSO describes the new JSO to be used for the
+PPGT. @sec-sim-deets summarises the results of two simulation
+experiments done to assess JSO's performance: one comparing JSO's
+performance improvements relative to an existing optimiser, Creeping
+Random Search (CRS), and the other studying the impact of different PP
+index properties on optimisation performance. @sec-sim-res presents the
+results. @sec-discussion discusses implementation details and insights
+for practitioners. @sec-conclusion summarises the work and provides
+suggestions for future directions.
 
@@
 patterns of interest forming as the function is optimised. There are
-five optimisation methods implemented in the `tourr`
-package:
-
-- a pseudo-derivative, that searches locally for the best direction, based on differencing the index values for very close projections.
+five optimisation methods implemented in the `tourr` package:
+
+-   a pseudo-derivative, that searches locally for the best direction,
+    based on differencing the index values for very close projections.
 -   a brute-force optimisation (CRS).
 -   a modified brute force algorithm described in @posse95.
-- an essentially simulated annealing [@Bertsimas93] where the search space is reduced during the optimisation.
-- a very localised search, to take tiny steps to get closer to the local maximum.
+-   an essentially simulated annealing [@Bertsimas93] where the search
+    space is reduced during the optimisation.
+-   a very localised search, to take tiny steps to get closer to the
+    local maximum.
 
@@
 
-There are numerous PP index functions available: introduced in @huber85, @cook1993projection, @lee2005projection, @lee2010projection, @Grimm2016, @Laa:2020wkm.
-Most are relatively simply defined, for any projection dimension, and implemented because they are relatively easy to optimise.
-A goal is to develop PP indexes based on scagnostics (@scag, @WW08), but the blockage is their optimisation as these tend to be noisy, with potentially small squint angles.
+There are numerous PP index functions available: introduced in @huber85,
+@cook1993projection, @lee2005projection, @lee2010projection, @Grimm2016,
+@Laa:2020wkm. Most are relatively simply defined, for any projection
+dimension, and implemented because they are relatively easy to optimise.
+A goal is to develop PP indexes based on scagnostics (@scag, @WW08), but
+the blockage is their optimisation as these tend to be noisy, with
+potentially small squint angles.
 
@@
 by. One last aspect that is useful is for an optimiser to return local
-maxima in addition to the global one because data can contain many different and
-interesting features.
+maxima in addition to the global one because data can contain many
+different and interesting features.
 
@@
 
-```{=tex}
 \begin{defn}\label{def:sobolev_space}
@@
 \end{defn}
-```
+
 The Sobolev index $k$ in Definition \ref{def:sobolev_space} can be used
@@
 the higher $k$, the smoother $f$. While this Sobolev index $k$ is a
-useful measure of smoothness, it can be difficult to compute or even estimate
-in practice.
+useful measure of smoothness, it can be difficult to compute or even
+estimate in practice.
 
@@
 then two close projection bases might give very different index values
-(fast decay of correlations with respect to distance between bases). Popular covariance functions are parametric positive semi-definite
-functions.  In particular, the Matérn class of
-covariance functions has a dedicated parameter to capture the smoothness of
-the Gaussian field.
-
-```{=tex}
+(fast decay of correlations with respect to distance between bases).
+Popular covariance functions are parametric positive semi-definite
+functions. In particular, the Matérn class of covariance functions has a
+dedicated parameter to capture the smoothness of the Gaussian field.
+
 \begin{defn}
@@
 \end{defn}
-```
+
 The Matérn covariance function can be expressed analytically when $\nu$
 is a half-integer, the most popular values in the literature being
-$\frac{1}{2}$, $\frac{3}{2}$ and $\frac{5}{2}$ [@rasmussen2006gaussian]. The parameter $\nu$,
-called *smoothness parameter*, controls the decay of the covariance
-function. As such, it is an appropriate measure of smoothness of a
-random field, as shown by the simulations on @fig-matern-1d and @fig-matern-2d. For example, @karvonen2023asymptotic showed that if a
-function $f$ has a Sobolev index of $k$, then the smoothness parameter
-estimate $\nu$ in \eqref{eq:matern} cannot be asymptotically less than
-$k$. See the survey @porcu2024matern for additional results on the connection between the Matérn model and Sobolev spaces. An interesting result is that the asymptotic case
+$\frac{1}{2}$, $\frac{3}{2}$ and $\frac{5}{2}$ [@rasmussen2006gaussian].
+The parameter $\nu$, called *smoothness parameter*, controls the decay
+of the covariance function. As such, it is an appropriate measure of
+smoothness of a random field, as shown by the simulations on
+@fig-matern-1d and @fig-matern-2d. For example, @karvonen2023asymptotic
+showed that if a function $f$ has a Sobolev index of $k$, then the
+smoothness parameter estimate $\nu$ in \eqref{eq:matern} cannot be
+asymptotically less than $k$. See the survey @porcu2024matern for
+additional results on the connection between the Matérn model and
+Sobolev spaces. An interesting result is that the asymptotic case
 $\nu\rightarrow\infty$ coincides with the Gaussian kernel:
@@
 such as `GpGp` [@guinness2021gpgp] or `ExaGeoStatR` [@abdulah2023large],
-to fit the hyper-parameters of a GP covariance function on data, which is
-usually done by maximum likelihood estimation. In this project, the
+to fit the hyper-parameters of a GP covariance function on data, which
+is usually done by maximum likelihood estimation. In this project, the
 `GpGp` package is used.
 
-```{=tex}
 \begin{defn}
@@
 \end{defn}
-```
-
-The value of the optimal smoothness parameter $\nu>0$ can be
-naturally interpreted as follows: the higher $\nu$, the smoother the
-index function.
+
+The value of the optimal smoothness parameter $\nu>0$ can be naturally
+interpreted as follows: the higher $\nu$, the smoother the index
+function.
 
@@
 
-Here the formal definitions of projection distance and squint angle are given, before the definition of squintability. Two approaches to compute this metric numerically are described.
-
-```{=tex}
+Here the formal definitions of projection distance and squint angle are
+given, before the definition of squintability. Two approaches to compute
+this metric numerically are described.
+
 \begin{defn}[projection distance]\label{def:proj-dist}
@@
 \end{defn}
-```
-
-```{=tex}
 \begin{defn}[squint angle]\label{def:squint-angle}
@@
 \end{defn}
-```
 
@@
 
-```{=tex}
 \begin{defn}[squintability]\label{def:squintability}
@@
 \end{defn}
-```
-
-It is expected that these two values should be both high in the case of high squintability (fast
-increase in $g$ early on), and both low in the case of low squintability
-(any substantial increase in $g$ happens very late, close to the optimal
-angle). This suggests that their product \eqref{eq-squintability} should
-provide a sensible measure of squintability. The multiplicative constant
-$4$, which can be deemed arbitrary, does not change the interpretation
-of the squintability metric $\varsigma$; it is here to adjust the range
-of values of $\varsigma$ and simplify the explicit formula for
-$\varsigma$ obtained later on.
+
+It is expected that these two values should be both high in the case of
+high squintability (fast increase in $g$ early on), and both low in the
+case of low squintability (any substantial increase in $g$ happens very
+late, close to the optimal angle). This suggests that their product
+\eqref{eq-squintability} should provide a sensible measure of
+squintability. The multiplicative constant $4$, which can be deemed
+arbitrary, does not change the interpretation of the squintability
+metric $\varsigma$; it is here to adjust the range of values of
+$\varsigma$ and simplify the explicit formula for $\varsigma$ obtained
+later on.
 
@@
 \end{prop}
-
 \begin{proof}
@@
 the optimiser needs to be very close to be able to "see" the optimum.
-Thus, it could be difficult for the optimiser to find the optimum. <!--The
+Thus, it could be difficult for the optimiser to find the optimum.
+<!--The
 mathematical formulation of this intuition is proposed below:-->
@@
 
-To compute the squintability metric \eqref{eq-squintability} in practice,
-several approaches are possible. The first one is to propose a
+To compute the squintability metric \eqref{eq-squintability} in
+practice, several approaches are possible. The first one is to propose a
 parametric model for $g$, and use it to obtain an explicit formula for
 $\varsigma$. Numerical experiments suggest a scaled sigmoid shape as
-described below. Define
-$$
+described below. Define $$
 \ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
@@
 $\theta_2$ and $\theta_3$, such that $\ell(\theta_{2})=\frac{1}{2}$.
-Then, define
-$$
+Then, define $$
 g(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
@@
 Under the parametric model (@eq-parametric), the squintability metric
-\eqref{eq-squintability} can be shown to be equal to
-$$
+\eqref{eq-squintability} can be shown to be equal to $$
 \varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
@@
 
-Alternatively, one can estimate \eqref{eq-squintability} in a nonparametric
-way, for example by fitting $g$ using kernel regression, then
-numerically estimate the angle at which $-g'$ attains its highest value.
+Alternatively, one can estimate \eqref{eq-squintability} in a
+nonparametric way, for example by fitting $g$ using kernel regression,
+then numerically estimate the angle at which $-g'$ attains its highest
+value.
 
@@
 
-JSO mimics the natural movements of jellyfish, which include passive and active motions driven by ocean currents and their swimming patterns,
+JSO mimics the natural movements of jellyfish, which include passive and
+active motions driven by ocean currents and their swimming patterns,
 respectively. In the context of optimization, these movements are
@@
 
-The performance of JSO is investigated both in comparison to the existing
-optimizer, CRS, and across various hyper-parameter values. The
+The performance of JSO is investigated both in comparison to the
+existing optimizer, CRS, and across various hyper-parameter values. The
 performance is measured by the success rate, defined as the proportion
@@
 outlined in @sec-smoothness and @sec-squintability and as illustrated in
-@fig-smoothness and @fig-squintability. To compute smoothness, 500 random
-bases are simulated. Index values are calculated for each
-random basis before fitting the Gaussian process model to obtain the
-smoothness measure for the index.
+@fig-smoothness and @fig-squintability. To compute smoothness, 500
+random bases are simulated. Index values are calculated for each random
+basis before fitting the Gaussian process model to obtain the smoothness
+measure for the index.
 
@@
 index values are averaged with a bin width of 0.005. The four-parameter
-scaled logistic function (@eq-parametric) is fitted to the index values against
-projection distances, estimated by non-linear least squares. The
+scaled logistic function (@eq-parametric) is fitted to the index values
+against projection distances, estimated by non-linear least squares. The
 squintability measure is then calculated as
@@
 
-Performance is assessed based on the final projections across two optimisers (JSO and CRS) and the success rate across JSO hyper-parameters.
-The final projections found by the two optimisers (JSO and CRS) are
-presented in @fig-proj, broken down by 10th quantile, faceted by the
-data dimensions. In the 6-dimensional data scenario, JSO consistently
-identifies a clear pipe shape. The CRS also finds the pipe shape but
-with a wide rim, suggesting a further polish search may be required.
-With increasing dimensions, JSO may not always identify the pipe shape
-due to random sampling, but it still finds the pipe shape in over 50% of
-cases. Compared to CRS, JSO achieves higher index values and clearer
-pipe shapes across all quantiles in data of 8, 10, 12 dimensions,
-suggesting its advantage in exploring high-dimensional spaces.
+Performance is assessed based on the final projections across two
+optimisers (JSO and CRS) and the success rate across JSO
+hyper-parameters. The final projections found by the two optimisers (JSO
+and CRS) are presented in @fig-proj, broken down by 10th quantile,
+faceted by the data dimensions. In the 6-dimensional data scenario, JSO
+consistently identifies a clear pipe shape. The CRS also finds the pipe
+shape but with a wide rim, suggesting a further polish search may be
+required. With increasing dimensions, JSO may not always identify the
+pipe shape due to random sampling, but it still finds the pipe shape in
+over 50% of cases. Compared to CRS, JSO achieves higher index values and
+clearer pipe shapes across all quantiles in data of 8, 10, 12
+dimensions, suggesting its advantage in exploring high-dimensional
+spaces.
 
@@
 
-Smoothness and squintability are calculated across a collection of pipe-finding and sine-wave finding problems to construct the relationship between success rate, JSO hyper-parameters, and index properties.
-The success rate calculated at each hyper-parameter combination (number of jellyfish and the maximum number of tries) is presented in @fig-proportion.
-The uncertainty is quantified through 500 bootstrap replicates for each case.
-As the number of jellyfish and maximum tries increase, the success rate also increases.
-For simpler problems (6 dimensions), small parameter values (20 jellyfish and a maximum of 50 tries) can already achieve a high success rate.
-However, larger parameter values
-(i.e. 100 jellyfish and a maximum of 100 tries) are necessary for
-higher-dimensional problems (8, 10, and 12 dimensions). Increasing both
-parameters enhances the performance of JSO, but it also extends the
-computational time required for the optimisation, which can be
-computationally intensive when evaluating the index function (such as
-scagnostic indexes) multiple times across numerous iterations.
+Smoothness and squintability are calculated across a collection of
+pipe-finding and sine-wave finding problems to construct the
+relationship between success rate, JSO hyper-parameters, and index
+properties. The success rate calculated at each hyper-parameter
+combination (number of jellyfish and the maximum number of tries) is
+presented in @fig-proportion. The uncertainty is quantified through 500
+bootstrap replicates for each case. As the number of jellyfish and
+maximum tries increase, the success rate also increases. For simpler
+problems (6 dimensions), small parameter values (20 jellyfish and a
+maximum of 50 tries) can already achieve a high success rate. However,
+larger parameter values (i.e. 100 jellyfish and a maximum of 100 tries)
+are necessary for higher-dimensional problems (8, 10, and 12
+dimensions). Increasing both parameters enhances the performance of JSO,
+but it also extends the computational time required for the
+optimisation, which can be computationally intensive when evaluating the
+index function (such as scagnostic indexes) multiple times across
+numerous iterations.
 
@@
 
-The index properties, smoothness and squintability, numerically characterise the complexity of projection pursuit index, and hence potential optimisation problems.
-@tbl-smoothness-squintability presents the parameters for calculating both metrics estimated from the Gaussian process (variance, range, smooth, and nugget) and the scaled logistic function ($\theta_1$ to $\theta_4$) for each case considered in the simulation.
-The column $\nu$ is used as the smoothness metric and the
-column $\varsigma$ is calculated as equation (@eq-squintability-parametric)
-as the squintability metric.
-
-@tbl-mod-output presents the results from fitting a logistic regression model using the proportion of success as the response and smoothness, squintability, dimension, running time, number of jellyfish and maximum number of tries, as predictors.
-Long runtime is a binary variable, where greater than 20 seconds is labelled 1.
-The fit suggests that JSO success rate is only affected by squintability, dimension and number of jellies.
-Interestingly, the smoothness of the index is not a problem for the JSO - this is promising!
-Running the optimisation longer and allowing more tries also does not affect success rate.
-An increase in squintability of one unit increases the success rate by `r round(exp(broom::tidy(mod1)$estimate[3]) * 100, 0)`%.
-As dimension increases by one the success rate halves.
-Increasing the number of jellies by 10, increases the succes rate by `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 0)`%.
+The index properties, smoothness and squintability, numerically
+characterise the complexity of projection pursuit index, and hence
+potential optimisation problems. @tbl-smoothness-squintability presents
+the parameters for calculating both metrics estimated from the Gaussian
+process (variance, range, smooth, and nugget) and the scaled logistic
+function ($\theta_1$ to $\theta_4$) for each case considered in the
+simulation. The column $\nu$ is used as the smoothness metric and the
+column $\varsigma$ is calculated as equation
+(@eq-squintability-parametric) as the squintability metric.
+
+@tbl-mod-output presents the results from fitting a logistic regression
+model using the proportion of success as the response and smoothness,
+squintability, dimension, running time, number of jellyfish and maximum
+number of tries, as predictors. Long runtime is a binary variable, where
+greater than 20 seconds is labelled 1. The fit suggests that JSO success
+rate is only affected by squintability, dimension and number of jellies.
+Interestingly, the smoothness of the index is not a problem for the
+JSO - this is promising! Running the optimisation longer and allowing
+more tries also does not affect success rate. An increase in
+squintability of one unit increases the success rate by
+`r round(exp(broom::tidy(mod1)$estimate[3]) * 100, 0)`%. As dimension
+increases by one the success rate halves. Increasing the number of
+jellies by 10, increases the succes rate by
+`r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 0)`%.
 
@@
 
-The simulation study demonstrated that these metrics can be used to assess a PP index, and compare optimisers.
-Here we explain how they can be used for new index and optimiser development.
-Using the JSO optimiser for PPGT requires a change in user behaviour.
-For all the existing optimisers a single optimisation path is followed.
-The JSO has multiple optimisation paths, and needs additional tools to incorporate into the PPGT, as explained here.
+The simulation study demonstrated that these metrics can be used to
+assess a PP index, and compare optimisers. Here we explain how they can
+be used for new index and optimiser development. Using the JSO optimiser
+for PPGT requires a change in user behaviour. For all the existing
+optimisers a single optimisation path is followed. The JSO has multiple
+optimisation paths, and needs additional tools to incorporate into the
+PPGT, as explained here.
 
@@
 
-To use JSO for PPGT in the `tourr` package, specify `search_f = search_jellyfish` in the guided tour.
-Since multiple tour paths are generated by the jellyfish, the animation function `animate_*()` will not render the tour path for JSO.
-To visualise path visited by each individual jellyfish, assign the animation to an object (`res <- animate_xy(...)`).
-This will save the bases visited by the optimiser, along with relevant metadata, as a tibble data object (see @RJ-2021-105 for more details on the data object).
-The bases visited for individual jellyfish can then be extracted and viewed using the `planned_tour()` function.
+To use JSO for PPGT in the `tourr` package, specify
+`search_f = search_jellyfish` in the guided tour. Since multiple tour
+paths are generated by the jellyfish, the animation function
+`animate_*()` will not render the tour path for JSO. To visualise path
+visited by each individual jellyfish, assign the animation to an object
+(`res <- animate_xy(...)`). This will save the bases visited by the
+optimiser, along with relevant metadata, as a tibble data object (see
+@RJ-2021-105 for more details on the data object). The bases visited for
+individual jellyfish can then be extracted and viewed using the
+`planned_tour()` function.
 
@@
 
-The `ferrn` package [@RJ-2021-105] provides functionality for computing the smoothness and squintability metrics.
-Both cases require sampling random bases using the `sample_bases()` function, followed by  calculating the specific metric with either `calc_smoothness()` or `calc_squintability()`.
+The `ferrn` package [@RJ-2021-105] provides functionality for computing
+the smoothness and squintability metrics. Both cases require sampling
+random bases using the `sample_bases()` function, followed by
+calculating the specific metric with either `calc_smoothness()` or
+`calc_squintability()`.
 
@@
 
-To sample bases for calculating smoothness, four arguments are required: the index function, the dataset, the number of random bases to sample, and the best basis to compute projection distance.
-The output of `sample_bases()` is a data frame containing columns/ list-columns of sampled basis matrix, calculated projection distance and index value.
-Parallellasation is available through the `parallel = TRUE` argument to speed up the computation.
-
-The `calc_smoothness()` function takes the output from `sample_bases()` and fits a Gaussian process model to the index values against the projection distances to compute the smoothness metric.
-Starting parameters and additional Gaussian process arguments can be specified and details of the fit can be accessed through the `fit_res` attribute of the `calc_smoothness()` output.
+To sample bases for calculating smoothness, four arguments are required:
+the index function, the dataset, the number of random bases to sample,
+and the best basis to compute projection distance. The output of
+`sample_bases()` is a data frame containing columns/ list-columns of
+sampled basis matrix, calculated projection distance and index value.
+Parallellasation is available through the `parallel = TRUE` argument to
+speed up the computation.
+
+The `calc_smoothness()` function takes the output from `sample_bases()`
+and fits a Gaussian process model to the index values against the
+projection distances to compute the smoothness metric. Starting
+parameters and additional Gaussian process arguments can be specified
+and details of the fit can be accessed through the `fit_res` attribute
+of the `calc_smoothness()` output.
 
@@
 
-Sampling bases for calculating squintability includes an additional step of interpolating between the sampled bases and the best basis.
-This is specified through setting the `step_size` and `min_proj_dist` arguments in the `sample_bases()` function to non-NA numerical values.
-Since the projection distance typically ranges from 0 to 2, it is recommended to set `step_size` to 0.1 or lower, and `min_proj_dist` to be at least 0.5 to ensure a meaningful interpolation length.
-
-The `calc_squintability()` function computes squintability using two methods: 1) parametrically, by fitting a scaled sigmoid function through non-linear least square (`method = "nls"`), and 2) non-parametrically, using kernel smoothing (`method = "ks"`).
-A `bin_width` argument is required to average the index values over the projection distance before the fit.
-The output provides the estimated parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$) and the squintability metric for the parametric case.
-For the non-parametric case, it shows the maximum gradient attained (`max_d`), the corresponding projection distance (`max_dist`), and the squintability metric as their products.
+Sampling bases for calculating squintability includes an additional step
+of interpolating between the sampled bases and the best basis. This is
+specified through setting the `step_size` and `min_proj_dist` arguments
+in the `sample_bases()` function to non-NA numerical values. Since the
+projection distance typically ranges from 0 to 2, it is recommended to
+set `step_size` to 0.1 or lower, and `min_proj_dist` to be at least 0.5
+to ensure a meaningful interpolation length.
+
+The `calc_squintability()` function computes squintability using two
+methods: 1) parametrically, by fitting a scaled sigmoid function through
+non-linear least square (`method = "nls"`), and 2) non-parametrically,
+using kernel smoothing (`method = "ks"`). A `bin_width` argument is
+required to average the index values over the projection distance before
+the fit. The output provides the estimated parameters for the scaled
+sigmoid function ($\theta_1$ to $\theta_4$) and the squintability metric
+for the parametric case. For the non-parametric case, it shows the
+maximum gradient attained (`max_d`), the corresponding projection
+distance (`max_dist`), and the squintability metric as their products.
 
@@
 
-```{=html}
 <!--
@@
 -->
-```
 
@@
 The source code for reproducing the work reported in this paper can be
-found at: <https://github.com/huizezhang-sherry/paper-jso>. The simulation data produced in Section 5 can be found at <https://figshare.com/articles/dataset/Simulated_raw_data/26039506>. Nicolas
-Langrené acknowledges the partial support of the Guangdong Provincial
-Key Laboratory IRADS (2022B1212010006, R0400001-22) and the UIC Start-up
-Research Fund UICR0700041-22.
-
-# Supplementary materials {-}
-
-The supplementary materials available at <https://github.com/huizezhang-sherry/paper-jso> include: 1) details of the indexes used in the simulation study, 2) the script to get started with using JSO in a PPGT and calculating smoothness and squintability, as explained in @sec-discussion, and (3) the full code to reproduce the plots and summaries in this paper.
+found at: <https://github.com/huizezhang-sherry/paper-jso>. The
+simulation data produced in Section 5 can be found at
+<https://figshare.com/articles/dataset/Simulated_raw_data/26039506>.
+Nicolas Langrené acknowledges the partial support of the Guangdong
+Provincial Key Laboratory IRADS (2022B1212010006, R0400001-22) and the
+UIC Start-up Research Fund UICR0700041-22.
+
+# Supplementary materials {.unnumbered}
+
+The supplementary materials available at
+<https://github.com/huizezhang-sherry/paper-jso> include: 1) details of
+the indexes used in the simulation study, 2) the script to get started
+with using JSO in a PPGT and calculating smoothness and squintability,
+as explained in @sec-discussion, and (3) the full code to reproduce the
+plots and summaries in this paper.
 

### 2024-10-14 jso.qmd
go though the paper before submission

@@
    - \theoremstyle{plain}
-   - \newtheorem{assumption}{\protect\assumptionname}
-   - \newtheorem{claim}{\protect\claimname}
-   - \newtheorem{condition}{\protect\conditionname}
-   - \newtheorem{conjecture}{\protect\conjecturename}
-   - \newtheorem{cor}{\protect\corollaryname}
    - \newtheorem{defn}{\protect\definitionname}
-   - \newtheorem{example}{\protect\examplename}
-   - \newtheorem{lem}{\protect\lemmaname}
-   - \newtheorem{notation}{\protect\notationname}
-   - \newtheorem{problem}{\protect\problemname}
    - \newtheorem{prop}{\protect\propositionname}
-   - \newtheorem{rem}{\protect\remarkname}
-   - \newtheorem{thm}{\protect\theoremname}
-   - \providecommand{\assumptionname}{Assumption}
-   - \providecommand{\claimname}{Claim}
-   - \providecommand{\conditionname}{Condition}
-   - \providecommand{\conjecturename}{Conjecture}
-   - \providecommand{\corollaryname}{Corollary}
    - \providecommand{\definitionname}{Definition}
-   - \providecommand{\examplename}{Example}
-   - \providecommand{\lemmaname}{Lemma}
-   - \providecommand{\notationname}{Notation}
-   - \providecommand{\problemname}{Problem}
    - \providecommand{\propositionname}{Proposition}
-   - \providecommand{\remarkname}{Remark}
-   - \providecommand{\theoremname}{Theorem}
 nocite: |
@@
 in the literature. @sec-PP-properties introduces the metrics that
-measure different properties of PP indexes, smoothness and
-squintability, and @sec-JSO describes the new JSO to be used for the
-PPGT. @sec-sim-deets summarises the results of two simulation
-experiments done to assess JSO's performance: one comparing JSO's
-performance improvements relative to an existing optimiser, Creeping
-Random Search (CRS), and the other studying the impact of different PP
-index properties on optimisation performance. @sec-sim-res presents the
-results. @sec-discussion discusses implementation details and insights
-for practitioners. @sec-conclusion summarises the work and provides
-suggestions for future directions.
+measure two properties of PP indexes, smoothness and squintability, and
+@sec-JSO describes the new JSO to be used for the PPGT. @sec-sim-deets
+outlines two simulation experiments to assess JSO's performance: one
+comparing JSO's performance improvements relative to an existing
+optimiser, Creeping Random Search (CRS), and the other studying the
+impact of PP index properties on optimisation performance, and
+@sec-sim-res presents the results. @sec-discussion discusses the
+implementation of JSO in the `tourr` package and the PP property
+calculation in the `ferrn` package. @sec-conclusion summarises the work
+and provides suggestions for future directions.
 
@@
 
-\begin{defn}\label{def:sobolev_space}
+\begin{defn}[sobolev space]\label{def:sobolev_space}
 The Sobolev space $W^{k,p}(\mathbb{R})$ for $1\leq p\leq \infty$ is the set of all functions $f$ in $L^p(\mathbb{R})$ for which all weak derivatives $f^{(\ell)}$ of order $\ell\leq k$ exist and have a finite $L^p$ norm.
@@
 
-\begin{defn}
+\begin{defn}[Matérn covariance function]\label{def:matern}
 The Matérn covariance function $K$ is defined by
@@
 
-Here the formal definitions of projection distance and squint angle are
-given, before the definition of squintability. Two approaches to compute
-this metric numerically are described.
+This section defines the projection distance and introduces the
+squintability metric, followed by a description of two numerical
+approaches for computing squintability.
 
 \begin{defn}[projection distance]\label{def:proj-dist}
-Recall that $A \in \mathbb{R}^{p \times d}$ is a $d$-dimensional orthonormal matrix.
-Let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.
-The projection distance $r(A, A^*)$ between $A$
-and $A^*$ is defined as 
+Let $A \in \mathbb{R}^{p \times d}$ is a $d$-dimensional orthonormal matrix, and let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.
+The projection distance between $A$
+and $A^*$, $r(A, A^*)$, is defined by
 $r(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\,\rVert _F$
@@
 \begin{defn}[squint angle]\label{def:squint-angle}
+
 Let $A$ and $B$ be two $d$-dimensional orthonormal matrices in $\mathbb{R}^p$.
 The squint angle $\theta$ between the subspace spanned by $A$ and $B$ is defined as the smallest principal angle between these subspaces: $\theta = \min_{i \in \{1, \cdots, d\}} \arccos(\tau_i)$, where $\tau_i$ are the singular values of the matrix $M = A^T B$ obtained from its singular value decomposition.
-
 \end{defn}
@@
 \begin{defn}[squintability]\label{def:squintability}
-Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance $r(A, A^*)$ to the index value $f(XA)$, such that $g(r(A, A^*)) = f(XA)$.
-For brevity, denote $g(r)$ as $g(r(A, A^*))$.
-The squintability of an index function $f(XA)$ is defined as
+Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance $r(A, A^*)$ to the index value $f(XA)$, such that $g(r) = g(r(A, A^*)) = f(XA)$.
+The squintability of an index function $f$ is defined by
 
@@
 
-<!-- \textcolor{orange}{See whether further details are needed for the nonparametric approach.} -->
-
-<!--To the best of our knowledge, -->
-
-<!-- ## Speed -->
-
-<!-- The speed of optimizing an index function can be measured empirically by -->
-
-<!-- recording the duration of the optimisation. Alternatively, one can also -->
-
-<!-- gauge the speed in terms of computational complexity of evaluating the -->
-
-<!-- index function (for instance, in big O notation, with respect to the -->
-
-<!-- sample size) of computing the index function. -->
-
 # The jellyfish optimiser {#sec-JSO}
 
-JSO mimics the natural movements of jellyfish, which include passive and
-active motions driven by ocean currents and their swimming patterns,
-respectively. In the context of optimization, these movements are
-abstracted to explore the search space, aiming to balance exploration
-(searching new areas) and exploitation (focusing on promising areas).
-The algorithm aims to find the optimal solution by adapting the
-behaviour of jellyfish to navigate towards the best solution over
-iterations [@chou_novel_2021].
+The Jellyfish Search Optimiser (JSO) mimics the natural movements of
+jellyfish, which include passive and active motions driven by ocean
+currents and their swimming patterns, respectively. In the context of
+optimization, these movements are abstracted to explore the search
+space, aiming to balance exploration (searching new areas) and
+exploitation (focusing on promising areas). The algorithm aims to find
+the optimal solution by adapting the behaviour of jellyfish to navigate
+towards the best solution over iterations [@chou_novel_2021].
 
@@
 
-# Assessing optimisers: a simulation study {#sec-sim-deets}
-
-The JSO performance is compared with an existing optimiser, Creeping
-Random Search (CRS) [@RJ-2021-105; @laa_using_2020] used in the PP
-guided tour to explore JSO's behaviour under different hyper-parameter
-and data dimension combinations. The second simulation studies the
-effect of index properties (smoothness and squintability), along with
-JSO hyper-parameters, and data dimension, on the success rate of the JSO
-performance. This section describes the simulation details, with the
-results deferred to @sec-sim-res.
+# Assessing optimisers: two simulation studies {#sec-sim-deets}
+
+This section describes the details of two simulation studies: one
+compares the JSO performance with an existing optimiser, Creeping Random
+Search (CRS) [@RJ-2021-105; @laa_using_2020] to explore JSO's behaviour
+under different data dimensions. The second simulation first examines
+the JSO performance under different hyper-parameters combinations
+(number of jellyfish and the maximum number of tries), and then studies
+the effect of index properties (smoothness and squintability), along
+with JSO hyper-parameters, and data dimension, on the JSO performance
+through a logistic regression.
+
+The performance of JSO is measured by the success rate, defined as the
+proportion of simulations that achieves a final index value within 0.05
+of the best index value found among all 50 simulations (see
+@fig-success-rate for an illustration). This comparison is based on
+projection pursuit to find the pipe shape investigated by
+@laa_using_2020 using the `holes` index.
 
@@
 
-The performance of JSO is investigated both in comparison to the
-existing optimizer, CRS, and across various hyper-parameter values. The
-performance is measured by the success rate, defined as the proportion
-of simulations that achieves a final index value within 0.05 of the best
-index value found among all 50 simulations (see @fig-success-rate for an
-illustration). This comparison is based on projection pursuit to find
-the pipe shape investigated by @laa_using_2020 using the `holes` index.
-
-Fifty simulations are conducted with both JSO and CRS, in four different
-data dimensions ($d = 6, 8, 10, 12$). JSO uses 100 jellyfish with a
-maximum of 100 tries, while the CRS allows a maximum of 1000 samples at
-each iteration before the algorithm terminates. The different numbers
-account for the multiple paths of JSO to enable fairer comparison with
-CRS. The results of the simulation are collected using the data
-structure proposed in @RJ-2021-105 for assessing JSO, where the design
-parameters are stored along with index value, projection basis, random
-seed, and computation time.
-
-Fifty additional simulations are conducted for each hyper-parameter
-combination to analyze how they affect the JSO success rate. This
-includes variations in the number of jellyfish (20, 50, and 100) and the
-maximum number of tries (50 and 100).
+The JSO performance is investigated both in comparison to the existing
+optimizer, CRS, in a pipe-finding problem using the holes index in the
+6, 8, 10, and 12-dimensional space. Fifty simulations are conducted in
+four data dimensions ($d = 6, 8, 10, 12$) with each optimiser. JSO uses
+100 jellyfish with a maximum of 100 tries, while the CRS allows a
+maximum of 1000 samples at each iteration before the algorithm
+terminates. The different numbers account for the multiple paths of JSO
+to enable fairer comparison with CRS. The results of the simulation are
+collected using the data structure proposed in @RJ-2021-105 for
+assessing JSO, where the design parameters are stored along with index
+value, projection basis, random seed, and computation time.
 
@@
 
-## Factors affecting JSO success rate: index properties and jellyfish hyper-parameters {#sec-app-2}
-
-To assess JSO's performance across various scenarios, two different data
-shapes, pipe and a sine wave, are investigated in 6D and 8D spaces using
-six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`,
-and `stringy`, with varied JSO hyper-parameters. A total of
-`r nrow(sim_df)` combinations result, comprising of
+## Factors affecting JSO success rate: JSO hyper-parameters and index properties  {#sec-app-2}
+
+To examine JSO's performance across various hyper-parameter
+combinations, the pipe-finding problem using the holes index is repeated
+for fifty times in each hyper-parameter combination: 20, 50, and 100
+jellyfish and a maximum of 50 and 100 attempts. The success rate is
+calculated for each combination.
+
+To assess JSO's performance across various PP problems, two different
+data shapes, pipe and a sine wave, are investigated in 6D and 8D spaces
+using six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`,
+`spline`, and `stringy`, under varied JSO hyper-parameters. This results
+in a total of `r nrow(sim_df)` scenarios, comprising of
 `r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and
 `r nrow(filter(sim_df, index != "holes"))` on the sine-wave data. Again,
-JSO is run 50 times to calculate the success rate for each projection
-pursuit.
+JSO is run 50 times for each scenario to calculate the success rate for
+each PP problem.
 
@@
 
-To construct a relationship among success rate, index properties
-(smoothness and squintability), and jellyfish hyper-parameters, a
-generalised linear model is fitted using a binomial family and a logit
-link function. The data is pre-processed by 1) scaling the JSO
-hyper-parameters by a factor of 10 for interpretation, 2) creating a new
-binary variable `long_time` to indicate cases with an average run time
-over 20 seconds, and 3) re-coding the success rate for the `stringy`
-index as 0, because none of the 50 simulations correctly identified the
-sine-wave shape.
+To construct a relationship among success rate, smoothness,
+squintability, and JSO hyper-parameters, a generalised linear model is
+fitted using a binomial family and a logit link function. The data is
+pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
+for interpretation, 2) creating a new binary variable `long_time` to
+indicate cases with an average run time over 20 seconds, and 3)
+re-coding the success rate for the `stringy` index as 0, because none of
+the 50 simulations correctly identified the sine-wave shape.
 
@@
 #| out.width: 100%
-#| fig.cap: "Illustration of steps to calculate smoothness.
-For a given projection pursuit problem defined by the shape to find, data dimension and the index function, 1) sample random bases given the orthonormality contraint and calculate the index value associated with each random basis, and 2) fit a Gaussian process model of index values against projection distances to obtain the smoothness measure.
-"
+#| fig.cap: "Steps for calculating smoothness in a projection pursuit problem.
+Given the target shape, data dimension and the index function: (1) sample random bases given the orthonormality contraint and calculate their corresponding index values, and (2) fit a Gaussian process model of index values against the sampled bases to obtain the smoothness measure."
 # basis_smoothness <- sample_bases(idx = "splines2d", n_basis = 1000)
@@
 #| fig.cap: |
-#|   Illustration of steps to calculate squintability.
-For a given projection pursuit problem defined by the shape to find, data dimension and the index function, 1) sample random bases given the orthonormality and projection distance contraint, 2) interpolate the sampled bases to the optimal basis and calculate the projection distance and the index value for each interpolated basis.
-3) bin the index values by projection distances to obtain the average index value for each bin, 4) fit the scaled sigmoid function in equation \eqref{eq-squintability} to the binned index values against projection distances using non-linear least square, 5) calculate the squintability measure using equation (@eq-squintability-parametric) with parameters estimated from the model.
+#|   Steps for calculating squintability in a projection pursuit problem.
+Given the target shape, data dimension and the index function: (1) sample random bases given the orthonormality and projection distance contraint, (2) interpolate the sampled bases to the optimal basis to calculate the projection distance and the index value.
+(3) bin the index values by projection distances to obtain the average index value for each bin, (4) fit the scaled sigmoid function in equation \eqref{eq-squintability} to the binned index values against projection distances using non-linear least square, 5) calculate the squintability measure using equation (@eq-squintability-parametric) with parameters estimated from the model.
 # p1 <- sim_sine_6d_spline_projdist |> 
@@
 
-Performance is assessed based on the final projections across two
-optimisers (JSO and CRS) and the success rate across JSO
-hyper-parameters. The final projections found by the two optimisers (JSO
-and CRS) are presented in @fig-proj, broken down by 10th quantile,
+Performance of JSO is assessed based on the final projections across the
+two optimisers: JSO and CRS. The final projections found by the two
+optimisers are presented in @fig-proj, broken down by 10th quantile,
 faceted by the data dimensions. In the 6-dimensional data scenario, JSO
@@
 
-## Factors affecting JSO success rate: index properties and jellyfish hyper-parameters
+## Factors affecting JSO success rate: JSO hyper-parameters and index properties
+
+The effect of JSO hyper–parameters (number of jellyfish and the maximum
+number of tries) on the success rate is presented in @fig-proportion.
+The uncertainty is quantified through 500 bootstrap replicates for each
+case. As the number of jellyfish and maximum tries increase, the success
+rate also increases. For simpler problems (6 dimensions), small
+parameter values (20 jellyfish and a maximum of 50 tries) can already
+achieve a high success rate. However, larger parameter values (i.e. 100
+jellyfish and a maximum of 100 tries) are necessary for
+higher-dimensional problems (8, 10, and 12 dimensions). Increasing both
+parameters enhances the performance of JSO, but it also extends the
+computational time required for the optimisation, which can be
+computationally intensive when evaluating the index function (such as
+scagnostic indexes) multiple times across numerous iterations.
 
@@
 relationship between success rate, JSO hyper-parameters, and index
-properties. The success rate calculated at each hyper-parameter
-combination (number of jellyfish and the maximum number of tries) is
-presented in @fig-proportion. The uncertainty is quantified through 500
-bootstrap replicates for each case. As the number of jellyfish and
-maximum tries increase, the success rate also increases. For simpler
-problems (6 dimensions), small parameter values (20 jellyfish and a
-maximum of 50 tries) can already achieve a high success rate. However,
-larger parameter values (i.e. 100 jellyfish and a maximum of 100 tries)
-are necessary for higher-dimensional problems (8, 10, and 12
-dimensions). Increasing both parameters enhances the performance of JSO,
-but it also extends the computational time required for the
-optimisation, which can be computationally intensive when evaluating the
-index function (such as scagnostic indexes) multiple times across
-numerous iterations.
+properties. @tbl-smoothness-squintability presents the parameters
+estimated from the Gaussian process (outputscale $\eta$, lengthscale
+$\ell$, smoothness $\nu$, and nugget $\sigma$) and from the scaled
+logistic function ($\theta_1$ to $\theta_4$) for calculating smoothness
+and squintability in each PP problem considered. The column $\nu$ is
+used as the smoothness metric and the column $\varsigma$ is calculated
+as equation (@eq-squintability-parametric) as the squintability metric.
 
@@
 ```
-
-The index properties, smoothness and squintability, numerically
-characterise the complexity of projection pursuit index, and hence
-potential optimisation problems. @tbl-smoothness-squintability presents
-the parameters for calculating both metrics estimated from the Gaussian
-process (variance, range, smooth, and nugget) and the scaled logistic
-function ($\theta_1$ to $\theta_4$) for each case considered in the
-simulation. The column $\nu$ is used as the smoothness metric and the
-column $\varsigma$ is calculated as equation
-(@eq-squintability-parametric) as the squintability metric.
-
-@tbl-mod-output presents the results from fitting a logistic regression
-model using the proportion of success as the response and smoothness,
-squintability, dimension, running time, number of jellyfish and maximum
-number of tries, as predictors. Long runtime is a binary variable, where
-greater than 20 seconds is labelled 1. The fit suggests that JSO success
-rate is only affected by squintability, dimension and number of jellies.
-Interestingly, the smoothness of the index is not a problem for the
-JSO - this is promising! Running the optimisation longer and allowing
-more tries also does not affect success rate. An increase in
-squintability of one unit increases the success rate by
-`r round(exp(broom::tidy(mod1)$estimate[3]) * 100, 0)`%. As dimension
-increases by one the success rate halves. Increasing the number of
-jellies by 10, increases the succes rate by
-`r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 0)`%.
 
@@
 
+@tbl-mod-output presents the results from fitting a logistic regression
+model using the proportion of success as the response and smoothness,
+squintability, dimension, running time, number of jellyfish and maximum
+number of tries, as predictors. Long runtime is a binary variable, where
+greater than 20 seconds is labelled 1. The fit suggests that JSO success
+rate is only affected by squintability, dimension and number of jellies.
+Interestingly, the smoothness of the index is not a problem for the
+JSO - this is promising! Running the optimisation longer and allowing
+more tries also does not affect success rate. An increase in
+squintability of one unit increases the success rate by
+`r round(exp(broom::tidy(mod1)$estimate[3]) * 100, 0)`%. As dimension
+increases by one the success rate halves. Increasing the number of
+jellies by 10, increases the succes rate by
+`r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 0)`%.
+
 # Practicalities {#sec-discussion}
 
-The simulation study demonstrated that these metrics can be used to
-assess a PP index, and compare optimisers. Here we explain how they can
-be used for new index and optimiser development. Using the JSO optimiser
-for PPGT requires a change in user behaviour. For all the existing
-optimisers a single optimisation path is followed. The JSO has multiple
-optimisation paths, and needs additional tools to incorporate into the
-PPGT, as explained here.
+The simulation studies have compared the JSO with CRS and shown how
+smoothness and squintability affect the JSO success rate. Here we
+explain how they can be used for new index and optimiser development.
+Using the JSO optimiser for PPGT requires a change in user behaviour.
+For all the existing optimisers a single optimisation path is followed.
+The JSO has multiple optimisation paths, and needs additional tools to
+incorporate into the PPGT, as explained here.
 
@@
 To use JSO for PPGT in the `tourr` package, specify
-`search_f = search_jellyfish` in the guided tour. Since multiple tour
-paths are generated by the jellyfish, the animation function
-`animate_*()` will not render the tour path for JSO. To visualise path
-visited by each individual jellyfish, assign the animation to an object
-(`res <- animate_xy(...)`). This will save the bases visited by the
-optimiser, along with relevant metadata, as a tibble data object (see
-@RJ-2021-105 for more details on the data object). The bases visited for
-individual jellyfish can then be extracted and viewed using the
-`planned_tour()` function.
-
-## Computing the metrics for your new index
+`search_f = search_jellyfish` in the guided tour. Unlike existing
+optimisers, the animation function `animate_*()` won't directly render
+the tour path for JSO due to the generation of multiple tour paths. To
+visualise the path visited by each individual jellyfish, assign the
+animation to an object (`res <- animate_xy(...)`). This will save the
+bases visited by JSO, along with relevant metadata, as a tibble data
+object (see @RJ-2021-105 for more details on the data object). The bases
+visited for individual jellyfish can then be extracted and viewed using
+the `planned_tour()` function.
+
+## Computing the index properties for your new index
 
 The `ferrn` package [@RJ-2021-105] provides functionality for computing
-the smoothness and squintability metrics. Both cases require sampling
-random bases using the `sample_bases()` function, followed by
-calculating the specific metric with either `calc_smoothness()` or
-`calc_squintability()`.
+the smoothness and squintability metrics. Both metrics require sampling
+random bases using the `sample_bases()` function, followed by the metric
+calculation with `calc_smoothness()` or `calc_squintability()`.
 
@@
 
-To sample bases for calculating smoothness, four arguments are required:
-the index function, the dataset, the number of random bases to sample,
-and the best basis to compute projection distance. The output of
-`sample_bases()` is a data frame containing columns/ list-columns of
-sampled basis matrix, calculated projection distance and index value.
-Parallellasation is available through the `parallel = TRUE` argument to
-speed up the computation.
+To sample bases for calculating smoothness, the following arguments are
+required: the index function, the dataset, and the number of random
+bases to sample. The output of `sample_bases()` is a data frame with a
+list-column of sampled basis matrix and index value. Parallellasation is
+available to speed up the index value computation through the
+`parallel = TRUE` argument.
 
@@
 and fits a Gaussian process model to the index values against the
-projection distances to compute the smoothness metric. Starting
-parameters and additional Gaussian process arguments can be specified
-and details of the fit can be accessed through the `fit_res` attribute
-of the `calc_smoothness()` output.
+sampled basis, as the location, to obtain the smoothness metric.
+Starting parameters and additional Gaussian process arguments can be
+specified and details of the fit can be accessed through the `fit_res`
+attribute of the output.
 
@@
 
-Sampling bases for calculating squintability includes an additional step
-of interpolating between the sampled bases and the best basis. This is
-specified through setting the `step_size` and `min_proj_dist` arguments
-in the `sample_bases()` function to non-NA numerical values. Since the
-projection distance typically ranges from 0 to 2, it is recommended to
+Bases sampling for squintability includes an additional step of
+interpolating between the sampled bases and the best basis. This step is
+performed when the arguments `step_size` and `min_proj_dist` in the
+`sample_bases()` function are set to non-NA numerical values. Given the
+projection distance typically ranging from 0 to 2, it is recommended to
 set `step_size` to 0.1 or lower, and `min_proj_dist` to be at least 0.5
@@
 required to average the index values over the projection distance before
-the fit. The output provides the estimated parameters for the scaled
-sigmoid function ($\theta_1$ to $\theta_4$) and the squintability metric
-for the parametric case. For the non-parametric case, it shows the
+the fit. For the parametric case, the output provides the estimated
+parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$)
+and the calculated squintability metric as Equation
+@eq-squintability-parametric. For the non-parametric case, it shows the
 maximum gradient attained (`max_d`), the corresponding projection
@@
 
+One future direction of this work is to test the JSO, along with its
+variations and other swarm-based optimisers, on a broader range of
+indexes, for example, scagnostic indexes.
+
 <!--

### 2024-10-14 jso.tex
go though the paper before submission

@@
 \theoremstyle{plain}
-\newtheorem{assumption}{\protect\assumptionname}
-\newtheorem{claim}{\protect\claimname}
-\newtheorem{condition}{\protect\conditionname}
-\newtheorem{conjecture}{\protect\conjecturename}
-\newtheorem{cor}{\protect\corollaryname}
 \newtheorem{defn}{\protect\definitionname}
-\newtheorem{example}{\protect\examplename}
-\newtheorem{lem}{\protect\lemmaname}
-\newtheorem{notation}{\protect\notationname}
-\newtheorem{problem}{\protect\problemname}
 \newtheorem{prop}{\protect\propositionname}
-\newtheorem{rem}{\protect\remarkname}
-\newtheorem{thm}{\protect\theoremname}
-\providecommand{\assumptionname}{Assumption}
-\providecommand{\claimname}{Claim}
-\providecommand{\conditionname}{Condition}
-\providecommand{\conjecturename}{Conjecture}
-\providecommand{\corollaryname}{Corollary}
 \providecommand{\definitionname}{Definition}
-\providecommand{\examplename}{Example}
-\providecommand{\lemmaname}{Lemma}
-\providecommand{\notationname}{Notation}
-\providecommand{\problemname}{Problem}
 \providecommand{\propositionname}{Proposition}
-\providecommand{\remarkname}{Remark}
-\providecommand{\theoremname}{Theorem}
 \makeatletter
@@
 that measures an interesting aspect of the projected data, such as
-deviation from normality, presence of clusters, non-linear structure, or
-other features of interest. For a fixed sample of data, PP finds the
-orthonormal basis \(A\) that maximises the index value of the
-projection, \(Y = XA\):
+deviation from normality, presence of clusters, or non-linear structure.
+For a fixed sample of data, PP finds the orthonormal basis \(A\) that
+maximises the index value of the projection, \(Y = XA\):
 
@@
 index functions in the literature. Section~\ref{sec-PP-properties}
-introduces the metrics that measure different properties of PP indexes,
+introduces the metrics that measure two properties of PP indexes,
 smoothness and squintability, and Section~\ref{sec-JSO} describes the
-new JSO to be used for the PPGT. Section~\ref{sec-sim-deets} summarises
-the results of two simulation experiments done to assess JSO's
-performance: one comparing JSO's performance improvements relative to an
-existing optimiser, Creeping Random Search (CRS), and the other studying
-the impact of different PP index properties on optimisation performance.
+new JSO to be used for the PPGT. Section~\ref{sec-sim-deets} outlines
+two simulation experiments to assess JSO's performance: one comparing
+JSO's performance improvements relative to an existing optimiser,
+Creeping Random Search (CRS), and the other studying the impact of PP
+index properties on optimisation performance, and
 Section~\ref{sec-sim-res} presents the results.
-Section~\ref{sec-discussion} discusses implementation details and
-insights for practitioners. Section~\ref{sec-conclusion} summarises the
-work and provides suggestions for future directions.
+Section~\ref{sec-discussion} discusses the implementation of JSO in the
+\texttt{tourr} package and the PP property calculation in the
+\texttt{ferrn} package. Section~\ref{sec-conclusion} summarises the work
+and provides suggestions for future directions.
 
@@
 \end{defn}
-
 \begin{defn}[squint angle]\label{def:squint-angle}
+
 Let $A$ and $B$ be two $d$-dimensional orthonormal matrices in $\mathbb{R}^p$.
 The squint angle $\theta$ between the subspace spanned by $A$ and $B$ is defined as the smallest principal angle between these subspaces: $\theta = \min_{i \in \{1, \cdots, d\}} \arccos(\tau_i)$, where $\tau_i$ are the singular values of the matrix $M = A^T B$ obtained from its singular value decomposition.
-
 \end{defn}
@@
 
-JSO mimics the natural movements of jellyfish, which include passive and
-active motions driven by ocean currents and their swimming patterns,
-respectively. In the context of optimization, these movements are
-abstracted to explore the search space, aiming to balance exploration
-(searching new areas) and exploitation (focusing on promising areas).
-The algorithm aims to find the optimal solution by adapting the
-behaviour of jellyfish to navigate towards the best solution over
-iterations (Chou and Truong 2021).
+The Jellyfish Search Optimiser (JSO) mimics the natural movements of
+jellyfish, which include passive and active motions driven by ocean
+currents and their swimming patterns, respectively. In the context of
+optimization, these movements are abstracted to explore the search
+space, aiming to balance exploration (searching new areas) and
+exploitation (focusing on promising areas). The algorithm aims to find
+the optimal solution by adapting the behaviour of jellyfish to navigate
+towards the best solution over iterations (Chou and Truong 2021).
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, breakable, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, toprule=.15mm, left=2mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, colback=white, leftrule=.75mm, toptitle=1mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, bottomrule=.15mm, toptitle=1mm, titlerule=0mm, opacityback=0, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, coltitle=black, toprule=.15mm, leftrule=.75mm, left=2mm, opacitybacktitle=0.6, rightrule=.15mm]
 
@@
 
-\section{Assessing optimisers: a simulation study}\label{sec-sim-deets}
-
-The JSO performance is compared with an existing optimiser, Creeping
-Random Search (CRS) (Zhang et al. 2021; Laa and Cook 2020) used in the
-PP guided tour to explore JSO's behaviour under different
-hyper-parameter and data dimension combinations. The second simulation
-studies the effect of index properties (smoothness and squintability),
-along with JSO hyper-parameters, and data dimension, on the success rate
-of the JSO performance. This section describes the simulation details,
-with the results deferred to Section~\ref{sec-sim-res}.
-
-\subsection{Performance of JSO relative to CRS}\label{sec-app-1}
-
-The performance of JSO is investigated both in comparison to the
-existing optimizer, CRS, and across various hyper-parameter values. The
-performance is measured by the success rate, defined as the proportion
-of simulations that achieves a final index value within 0.05 of the best
-index value found among all 50 simulations (see
+\section{Assessing optimisers: two simulation
+studies}\label{sec-sim-deets}
+
+This section describes the details of two simulation studies: one
+compares the JSO performance with an existing optimiser, Creeping Random
+Search (CRS) (Zhang et al. 2021; Laa and Cook 2020) to explore JSO's
+behaviour under different data dimensions. The second simulation first
+examines the performance JSO under different JSO hyper-parameters
+(number of jellyfish and the maximum number of tries), and then studies
+the effect of index properties (smoothness and squintability), along
+with JSO hyper-parameters, and data dimension, on the performance of JSO
+through a logistic regression.
+
+The performance of JSO is measured by the success rate, defined as the
+proportion of simulations that achieves a final index value within 0.05
+of the best index value found among all 50 simulations (see
 Figure~\ref{fig-success-rate} for an illustration). This comparison is
@@
 
-Fifty simulations are conducted with both JSO and CRS, in four different
-data dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100 jellyfish with a
-maximum of 100 tries, while the CRS allows a maximum of 1000 samples at
-each iteration before the algorithm terminates. The different numbers
-account for the multiple paths of JSO to enable fairer comparison with
-CRS. The results of the simulation are collected using the data
-structure proposed in Zhang et al. (2021) for assessing JSO, where the
-design parameters are stored along with index value, projection basis,
-random seed, and computation time.
-
-Fifty additional simulations are conducted for each hyper-parameter
-combination to analyze how they affect the JSO success rate. This
-includes variations in the number of jellyfish (20, 50, and 100) and the
-maximum number of tries (50 and 100).
+\subsection{Performance of JSO relative to CRS}\label{sec-app-1}
+
+The JSO performance is investigated both in comparison to the existing
+optimizer, CRS, in a pipe-finding problem using the holes index in the
+6, 8, 10, and 12-dimensional space. Fifty simulations are conducted in
+four data dimensions (\(d = 6, 8, 10, 12\)) with each optimiser. JSO
+uses 100 jellyfish with a maximum of 100 tries, while the CRS allows a
+maximum of 1000 samples at each iteration before the algorithm
+terminates. The different numbers account for the multiple paths of JSO
+to enable fairer comparison with CRS. The results of the simulation are
+collected using the data structure proposed in Zhang et al. (2021) for
+assessing JSO, where the design parameters are stored along with index
+value, projection basis, random seed, and computation time.
 
@@
 
-\subsection{Factors affecting JSO success rate: index properties and
-jellyfish hyper-parameters}\label{sec-app-2}
-
-To assess JSO's performance across various scenarios, two different data
-shapes, pipe and a sine wave, are investigated in 6D and 8D spaces using
-six different PP indexes: \texttt{dcor2d\_2}, \texttt{loess2d},
-\texttt{MIC}, \texttt{TIC}, \texttt{spline}, and \texttt{stringy}, with
-varied JSO hyper-parameters. A total of 52 combinations result,
+\subsection{Factors affecting JSO success rate: JSO hyper-parameters and
+index properties}\label{sec-app-2}
+
+To examine JSO's performance across various hyper-parameter
+combinations, the pipe-finding problem using the holes index is repeated
+for fifty times in each hyper-parameter combination: 20, 50, and 100
+jellyfish and a maximum of 50 and 100 attempts. The success rate is
+calculated for each combination.
+
+To assess JSO's performance across various PP problems, two different
+data shapes, pipe and a sine wave, are investigated in 6D and 8D spaces
+using six different PP indexes: \texttt{dcor2d\_2}, \texttt{loess2d},
+\texttt{MIC}, \texttt{TIC}, \texttt{spline}, and \texttt{stringy}, under
+varied JSO hyper-parameters. This results in a total of 52 scenarios,
 comprising of 24 computed on the pipe data and 28 on the sine-wave data.
-Again, JSO is run 50 times to calculate the success rate for each
-projection pursuit.
+Again, JSO is run 50 times for each scenario to calculate the success
+rate for each PP problem.
 
@@
 
-To construct a relationship among success rate, index properties
-(smoothness and squintability), and jellyfish hyper-parameters, a
-generalised linear model is fitted using a binomial family and a logit
-link function. The data is pre-processed by 1) scaling the JSO
-hyper-parameters by a factor of 10 for interpretation, 2) creating a new
-binary variable \texttt{long\_time} to indicate cases with an average
-run time over 20 seconds, and 3) re-coding the success rate for the
-\texttt{stringy} index as 0, because none of the 50 simulations
-correctly identified the sine-wave shape.
+To construct a relationship among success rate, smoothness,
+squintability, and JSO hyper-parameters, a generalised linear model is
+fitted using a binomial family and a logit link function. The data is
+pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
+for interpretation, 2) creating a new binary variable
+\texttt{long\_time} to indicate cases with an average run time over 20
+seconds, and 3) re-coding the success rate for the \texttt{stringy}
+index as 0, because none of the 50 simulations correctly identified the
+sine-wave shape.
 
@@
 
-\caption{\label{fig-smoothness}Illustration of steps to calculate
-smoothness. For a given projection pursuit problem defined by the shape
-to find, data dimension and the index function, 1) sample random bases
-given the orthonormality contraint and calculate the index value
-associated with each random basis, and 2) fit a Gaussian process model
-of index values against projection distances to obtain the smoothness
-measure.}
+\caption{\label{fig-smoothness}Steps for calculating smoothness in a
+projection pursuit problem. Given the target shape, data dimension and
+the index function: (1) sample random bases given the orthonormality
+contraint and calculate their corresponding index values, and (2) fit a
+Gaussian process model of index values against the sampled bases to
+obtain the smoothness measure.}
 
@@
 
-\caption{\label{fig-squintability}Illustration of steps to calculate
-squintability. For a given projection pursuit problem defined by the
-shape to find, data dimension and the index function, 1) sample random
-bases given the orthonormality and projection distance contraint, 2)
-interpolate the sampled bases to the optimal basis and calculate the
-projection distance and the index value for each interpolated basis. 3)
-bin the index values by projection distances to obtain the average index
-value for each bin, 4) fit the scaled sigmoid function in equation
-\eqref{eq-squintability} to the binned index values against projection
-distances using non-linear least square, 5) calculate the squintability
-measure using equation (\ref{eq-squintability-parametric}) with
-parameters estimated from the model.}
+\caption{\label{fig-squintability}Steps for calculating squintability in
+a projection pursuit problem. Given the target shape, data dimension and
+the index function: (1) sample random bases given the orthonormality and
+projection distance contraint, (2) interpolate the sampled bases to the
+optimal basis to calculate the projection distance and the index value.
+(3) bin the index values by projection distances to obtain the average
+index value for each bin, (4) fit the scaled sigmoid function in
+equation \eqref{eq-squintability} to the binned index values against
+projection distances using non-linear least square, 5) calculate the
+squintability measure using equation (\ref{eq-squintability-parametric})
+with parameters estimated from the model.}
 
@@
 
-Performance is assessed based on the final projections across two
-optimisers (JSO and CRS) and the success rate across JSO
-hyper-parameters. The final projections found by the two optimisers (JSO
-and CRS) are presented in Figure~\ref{fig-proj}, broken down by 10th
+Performance of JSO is assessed based on the final projections across the
+two optimisers: JSO and CRS. The final projections found by the two
+optimisers are presented in Figure~\ref{fig-proj}, broken down by 10th
 quantile, faceted by the data dimensions. In the 6-dimensional data
@@
 
-\subsection{Factors affecting JSO success rate: index properties and
-jellyfish
-hyper-parameters}\label{factors-affecting-jso-success-rate-index-properties-and-jellyfish-hyper-parameters}
+\subsection{Factors affecting JSO success rate: JSO hyper-parameters and
+index
+properties}\label{factors-affecting-jso-success-rate-jso-hyper-parameters-and-index-properties}
+
+The effect of JSO hyper--parameters (number of jellyfish and the maximum
+number of tries) on the success rate is presented in
+Figure~\ref{fig-proportion}. The uncertainty is quantified through 500
+bootstrap replicates for each case. As the number of jellyfish and
+maximum tries increase, the success rate also increases. For simpler
+problems (6 dimensions), small parameter values (20 jellyfish and a
+maximum of 50 tries) can already achieve a high success rate. However,
+larger parameter values (i.e.~100 jellyfish and a maximum of 100 tries)
+are necessary for higher-dimensional problems (8, 10, and 12
+dimensions). Increasing both parameters enhances the performance of JSO,
+but it also extends the computational time required for the
+optimisation, which can be computationally intensive when evaluating the
+index function (such as scagnostic indexes) multiple times across
+numerous iterations.
 
@@
 relationship between success rate, JSO hyper-parameters, and index
-properties. The success rate calculated at each hyper-parameter
-combination (number of jellyfish and the maximum number of tries) is
-presented in Figure~\ref{fig-proportion}. The uncertainty is quantified
-through 500 bootstrap replicates for each case. As the number of
-jellyfish and maximum tries increase, the success rate also increases.
-For simpler problems (6 dimensions), small parameter values (20
-jellyfish and a maximum of 50 tries) can already achieve a high success
-rate. However, larger parameter values (i.e.~100 jellyfish and a maximum
-of 100 tries) are necessary for higher-dimensional problems (8, 10, and
-12 dimensions). Increasing both parameters enhances the performance of
-JSO, but it also extends the computational time required for the
-optimisation, which can be computationally intensive when evaluating the
-index function (such as scagnostic indexes) multiple times across
-numerous iterations.
+properties. Table~\ref{tbl-smoothness-squintability} presents the
+parameters estimated from the Gaussian process (outputscale \(\eta\),
+lengthscale \(\ell\), smoothness \(\nu\), and nugget \(\sigma\)) and
+from the scaled logistic function (\(\theta_1\) to \(\theta_4\)) for
+calculating smoothness and squintability in each PP problem considered.
+The column \(\nu\) is used as the smoothness metric and the column
+\(\varsigma\) is calculated as equation
+(\ref{eq-squintability-parametric}) as the squintability metric.
 
@@
 \end{figure}%
-
-The index properties, smoothness and squintability, numerically
-characterise the complexity of projection pursuit index, and hence
-potential optimisation problems.
-Table~\ref{tbl-smoothness-squintability} presents the parameters for
-calculating both metrics estimated from the Gaussian process (variance,
-range, smooth, and nugget) and the scaled logistic function
-(\(\theta_1\) to \(\theta_4\)) for each case considered in the
-simulation. The column \(\nu\) is used as the smoothness metric and the
-column \(\varsigma\) is calculated as equation
-(\ref{eq-squintability-parametric}) as the squintability metric.
-
-Table~\ref{tbl-mod-output} presents the results from fitting a logistic
-regression model using the proportion of success as the response and
-smoothness, squintability, dimension, running time, number of jellyfish
-and maximum number of tries, as predictors. Long runtime is a binary
-variable, where greater than 20 seconds is labelled 1. The fit suggests
-that JSO success rate is only affected by squintability, dimension and
-number of jellies. Interestingly, the smoothness of the index is not a
-problem for the JSO - this is promising! Running the optimisation longer
-and allowing more tries also does not affect success rate. An increase
-in squintability of one unit increases the success rate by 1235\%. As
-dimension increases by one the success rate halves. Increasing the
-number of jellies by 10, increases the succes rate by 27\%.
 
@@
 
+Table~\ref{tbl-mod-output} presents the results from fitting a logistic
+regression model using the proportion of success as the response and
+smoothness, squintability, dimension, running time, number of jellyfish
+and maximum number of tries, as predictors. Long runtime is a binary
+variable, where greater than 20 seconds is labelled 1. The fit suggests
+that JSO success rate is only affected by squintability, dimension and
+number of jellies. Interestingly, the smoothness of the index is not a
+problem for the JSO - this is promising! Running the optimisation longer
+and allowing more tries also does not affect success rate. An increase
+in squintability of one unit increases the success rate by 1235\%. As
+dimension increases by one the success rate halves. Increasing the
+number of jellies by 10, increases the succes rate by 27\%.
+
 \section{Practicalities}\label{sec-discussion}
 
-The simulation study demonstrated that these metrics can be used to
-assess a PP index, and compare optimisers. Here we explain how they can
-be used for new index and optimiser development. Using the JSO optimiser
-for PPGT requires a change in user behaviour. For all the existing
-optimisers a single optimisation path is followed. The JSO has multiple
-optimisation paths, and needs additional tools to incorporate into the
-PPGT, as explained here.
+The simulation studies have compared the JSO with CRS and shown how
+smoothness and squintability affect the JSO success rate. Here we
+explain how they can be used for new index and optimiser development.
+Using the JSO optimiser for PPGT requires a change in user behaviour.
+For all the existing optimisers a single optimisation path is followed.
+The JSO has multiple optimisation paths, and needs additional tools to
+incorporate into the PPGT, as explained here.
 
@@
 To use JSO for PPGT in the \texttt{tourr} package, specify
-\texttt{search\_f\ =\ search\_jellyfish} in the guided tour. Since
-multiple tour paths are generated by the jellyfish, the animation
-function \texttt{animate\_*()} will not render the tour path for JSO. To
-visualise path visited by each individual jellyfish, assign the
-animation to an object (\texttt{res\ \textless{}-\ animate\_xy(...)}).
-This will save the bases visited by the optimiser, along with relevant
-metadata, as a tibble data object (see Zhang et al. (2021) for more
-details on the data object). The bases visited for individual jellyfish
-can then be extracted and viewed using the \texttt{planned\_tour()}
-function.
-
-\subsection{Computing the metrics for your new
-index}\label{computing-the-metrics-for-your-new-index}
+\texttt{search\_f\ =\ search\_jellyfish} in the guided tour. Unlike
+existing optimisers, the animation function \texttt{animate\_*()} won't
+directly render the tour path for JSO due to the generation of multiple
+tour paths. To visualise the path visited by each individual jellyfish,
+assign the animation to an object
+(\texttt{res\ \textless{}-\ animate\_xy(...)}). This will save the bases
+visited by JSO, along with relevant metadata, as a tibble data object
+(see Zhang et al. (2021) for more details on the data object). The bases
+visited for individual jellyfish can then be extracted and viewed using
+the \texttt{planned\_tour()} function.
+
+\subsection{Computing the index properties for your new
+index}\label{computing-the-index-properties-for-your-new-index}
 
 The \texttt{ferrn} package (Zhang et al. 2021) provides functionality
-for computing the smoothness and squintability metrics. Both cases
+for computing the smoothness and squintability metrics. Both metrics
 require sampling random bases using the \texttt{sample\_bases()}
-function, followed by calculating the specific metric with either
+function, followed by the metric calculation with
 \texttt{calc\_smoothness()} or \texttt{calc\_squintability()}.
@@
 
-To sample bases for calculating smoothness, four arguments are required:
-the index function, the dataset, the number of random bases to sample,
-and the best basis to compute projection distance. The output of
-\texttt{sample\_bases()} is a data frame containing columns/
-list-columns of sampled basis matrix, calculated projection distance and
-index value. Parallellasation is available through the
-\texttt{parallel\ =\ TRUE} argument to speed up the computation.
+To sample bases for calculating smoothness, the following arguments are
+required: the index function, the dataset, and the number of random
+bases to sample. The output of \texttt{sample\_bases()} is a data frame
+with a list-column of sampled basis matrix and index value.
+Parallellasation is available to speed up the index value computation
+through the \texttt{parallel\ =\ TRUE} argument.
 
@@
 \texttt{sample\_bases()} and fits a Gaussian process model to the index
-values against the projection distances to compute the smoothness
-metric. Starting parameters and additional Gaussian process arguments
-can be specified and details of the fit can be accessed through the
-\texttt{fit\_res} attribute of the \texttt{calc\_smoothness()} output.
+values against the sampled basis, as the location, to obtain the
+smoothness metric. Starting parameters and additional Gaussian process
+arguments can be specified and details of the fit can be accessed
+through the \texttt{fit\_res} attribute of the output.
 
@@
 
-Sampling bases for calculating squintability includes an additional step
-of interpolating between the sampled bases and the best basis. This is
-specified through setting the \texttt{step\_size} and
-\texttt{min\_proj\_dist} arguments in the \texttt{sample\_bases()}
-function to non-NA numerical values. Since the projection distance
-typically ranges from 0 to 2, it is recommended to set
-\texttt{step\_size} to 0.1 or lower, and \texttt{min\_proj\_dist} to be
-at least 0.5 to ensure a meaningful interpolation length.
+Bases sampling for squintability includes an additional step of
+interpolating between the sampled bases and the best basis. This step is
+performed when the arguments \texttt{step\_size} and
+\texttt{min\_proj\_dist} in the \texttt{sample\_bases()} function are
+set to non-NA numerical values. Given the projection distance typically
+ranging from 0 to 2, it is recommended to set \texttt{step\_size} to 0.1
+or lower, and \texttt{min\_proj\_dist} to be at least 0.5 to ensure a
+meaningful interpolation length.
 
@@
 \texttt{bin\_width} argument is required to average the index values
-over the projection distance before the fit. The output provides the
-estimated parameters for the scaled sigmoid function (\(\theta_1\) to
-\(\theta_4\)) and the squintability metric for the parametric case. For
-the non-parametric case, it shows the maximum gradient attained
+over the projection distance before the fit. For the parametric case,
+the output provides the estimated parameters for the scaled sigmoid
+function (\(\theta_1\) to \(\theta_4\)) and the calculated squintability
+metric as Equation \ref{eq-squintability-parametric}. For the
+non-parametric case, it shows the maximum gradient attained
 (\texttt{max\_d}), the corresponding projection distance
@@
 the JSO performance.
+
+One future direction of this work is to test the JSO, along with its
+variations and other swarm-based optimisers, on a broader range of
+indexes, for example, scagnostic indexes.
 

### 2024-10-14 jso.qmd
update smoothness and squintaiblity diagram

@@
 random bases are simulated. Index values are calculated for each random
-basis before fitting the Gaussian process model to obtain the smoothness
-measure for the index.
+basis, followed by fitting a Gaussian process model to obtain the
+smoothness measure for the index.
 
@@
 to the optimal basis with a step size of 0.005. Index values and
-projection distances are calculated for these interpolated bases and the
-index values are averaged with a bin width of 0.005. The four-parameter
-scaled logistic function (@eq-parametric) is fitted to the index values
-against projection distances, estimated by non-linear least squares. The
-squintability measure is then calculated as
-(@eq-squintability-parametric).
+projection distances are calculated for these interpolated bases. A
+binning procedure is applied to average the index values within each
+0.005 projection distance bin. The four-parameter scaled logistic
+function (@eq-parametric) is fitted to the index values against
+projection distances, estimated by non-linear least squares to obtain
+the squintability measure as Equation @eq-squintability-parametric.
 
@@
 #| fig.cap: "Steps for calculating smoothness in a projection pursuit problem.
-Given the target shape, data dimension and the index function: (1) sample random bases given the orthonormality contraint and calculate their corresponding index values, and (2) fit a Gaussian process model of index values against the sampled bases to obtain the smoothness measure."
+Given the target shape, data dimension and the index function: 1) sample random bases given the orthonormality contraint and calculate their corresponding index values, and 2) fit a Gaussian process model of index values against the sampled bases to obtain the smoothness measure."
 # basis_smoothness <- sample_bases(idx = "splines2d", n_basis = 1000)
@@
 #|   Steps for calculating squintability in a projection pursuit problem.
-Given the target shape, data dimension and the index function: (1) sample random bases given the orthonormality and projection distance contraint, (2) interpolate the sampled bases to the optimal basis to calculate the projection distance and the index value.
-(3) bin the index values by projection distances to obtain the average index value for each bin, (4) fit the scaled sigmoid function in equation \eqref{eq-squintability} to the binned index values against projection distances using non-linear least square, 5) calculate the squintability measure using equation (@eq-squintability-parametric) with parameters estimated from the model.
+Given the target shape, data dimension and the index function: 1) sample random bases given the orthonormality and projection distance contraint and calculates their corresponding index values, 2) interpolate the sampled bases to the optimal basis and calculate the projection distance and the index value.
+3) bin the index values by projection distances through averaging index value, 4) fit the scaled sigmoid function in equation \eqref{eq-squintability} to the binned index values against projection distances using non-linear least square to obtian the squintability measure using equation (@eq-squintability-parametric).
 # p1 <- sim_sine_6d_spline_projdist |> 

### 2024-10-14 jso.tex
update smoothness and squintaiblity diagram

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, bottomrule=.15mm, toptitle=1mm, titlerule=0mm, opacityback=0, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, coltitle=black, toprule=.15mm, leftrule=.75mm, left=2mm, opacitybacktitle=0.6, rightrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, opacityback=0, toprule=.15mm, colframe=quarto-callout-note-color-frame, breakable, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, opacitybacktitle=0.6, bottomrule=.15mm, titlerule=0mm, rightrule=.15mm, arc=.35mm, bottomtitle=1mm, left=2mm]
 
@@
 behaviour under different data dimensions. The second simulation first
-examines the performance JSO under different JSO hyper-parameters
-(number of jellyfish and the maximum number of tries), and then studies
-the effect of index properties (smoothness and squintability), along
-with JSO hyper-parameters, and data dimension, on the performance of JSO
-through a logistic regression.
+examines the JSO performance under different hyper-parameters
+combinations (number of jellyfish and the maximum number of tries), and
+then studies the effect of index properties (smoothness and
+squintability), along with JSO hyper-parameters, and data dimension, on
+the JSO performance through a logistic regression.
 
@@
 compute smoothness, 500 random bases are simulated. Index values are
-calculated for each random basis before fitting the Gaussian process
+calculated for each random basis, followed by fitting a Gaussian process
 model to obtain the smoothness measure for the index.
@@
 to the optimal basis with a step size of 0.005. Index values and
-projection distances are calculated for these interpolated bases and the
-index values are averaged with a bin width of 0.005. The four-parameter
-scaled logistic function (\ref{eq-parametric}) is fitted to the index
-values against projection distances, estimated by non-linear least
-squares. The squintability measure is then calculated as
-(\ref{eq-squintability-parametric}).
+projection distances are calculated for these interpolated bases. A
+binning procedure is applied to average the index values within each
+0.005 projection distance bin. The four-parameter scaled logistic
+function (\ref{eq-parametric}) is fitted to the index values against
+projection distances, estimated by non-linear least squares to obtain
+the squintability measure as Equation \ref{eq-squintability-parametric}.
 
@@
 projection pursuit problem. Given the target shape, data dimension and
-the index function: (1) sample random bases given the orthonormality
-contraint and calculate their corresponding index values, and (2) fit a
+the index function: 1) sample random bases given the orthonormality
+contraint and calculate their corresponding index values, and 2) fit a
 Gaussian process model of index values against the sampled bases to
@@
 a projection pursuit problem. Given the target shape, data dimension and
-the index function: (1) sample random bases given the orthonormality and
-projection distance contraint, (2) interpolate the sampled bases to the
-optimal basis to calculate the projection distance and the index value.
-(3) bin the index values by projection distances to obtain the average
-index value for each bin, (4) fit the scaled sigmoid function in
-equation \eqref{eq-squintability} to the binned index values against
-projection distances using non-linear least square, 5) calculate the
-squintability measure using equation (\ref{eq-squintability-parametric})
-with parameters estimated from the model.}
+the index function: 1) sample random bases given the orthonormality and
+projection distance contraint and calculates their corresponding index
+values, 2) interpolate the sampled bases to the optimal basis and
+calculate the projection distance and the index value. 3) bin the index
+values by projection distances through averaging index value, 4) fit the
+scaled sigmoid function in equation \eqref{eq-squintability} to the
+binned index values against projection distances using non-linear least
+square to obtian the squintability measure using equation
+(\ref{eq-squintability-parametric}).}
 

### 2024-10-14 jso.tex
Table 1 before Table 2

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacityback=0, toprule=.15mm, colframe=quarto-callout-note-color-frame, breakable, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, opacitybacktitle=0.6, bottomrule=.15mm, titlerule=0mm, rightrule=.15mm, arc=.35mm, bottomtitle=1mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, opacitybacktitle=0.6, toprule=.15mm, colback=white, opacityback=0, left=2mm, bottomtitle=1mm, breakable, arc=.35mm, bottomrule=.15mm, titlerule=0mm, leftrule=.75mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black]
 
@@
 
-\begingroup
-\fontsize{12.0pt}{14.4pt}\selectfont
-\setlength{\LTpost}{0mm}
-
-\begin{longtable}{lcc}
-
-\caption{\label{tbl-mod-output}Jellyfish success rate relative to index
-properties and jellyfish hyper-parameters. This is the summary from a
-logistic regression fit to smoothness, squintability, dimension, running
-time, number of jellyfish and maximum number of tries. Interestingly,
-squintability and dimension strongly affect jellyfish optimisation
-success. The number of jellies marginally affects success, but index
-smoothness, running longer and increasing the number of tries do not.}
-
-\tabularnewline
-
-\toprule
-\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
-\midrule\addlinespace[2.5pt]
-Smoothness & 89.1 (1.23 to 4,865,159) & 0.289 \\ 
-Squintability & 12.4 (3.08 to 125) & {\bfseries 0.006} \\ 
-Dimension & 0.57 (0.30 to 0.96) & {\bfseries 0.048} \\ 
-Long runtime & 0.17 (0.00 to 4.96) & 0.357 \\ 
-Number of jellyfish & 1.27 (0.99 to 1.68) & {\bfseries 0.069} \\ 
-Maximum number of tries & 1.10 (0.82 to 1.52) & 0.524 \\ 
-\bottomrule
-
-\end{longtable}
-
-\begin{minipage}{\linewidth}
-\textsuperscript{\textit{1}}OR = Odds Ratio, CI = Confidence Interval\\
-\end{minipage}
-\endgroup
-
 Table~\ref{tbl-mod-output} presents the results from fitting a logistic
@@
 
+\begingroup
+\fontsize{12.0pt}{14.4pt}\selectfont
+\setlength{\LTpost}{0mm}
+
+\begin{longtable}{lcc}
+
+\caption{\label{tbl-mod-output}Jellyfish success rate relative to index
+properties and jellyfish hyper-parameters. This is the summary from a
+logistic regression fit to smoothness, squintability, dimension, running
+time, number of jellyfish and maximum number of tries. Interestingly,
+squintability and dimension strongly affect jellyfish optimisation
+success. The number of jellies marginally affects success, but index
+smoothness, running longer and increasing the number of tries do not.}
+
+\tabularnewline
+
+\toprule
+\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
+\midrule\addlinespace[2.5pt]
+Smoothness & 89.1 (1.23 to 4,865,159) & 0.289 \\ 
+Squintability & 12.4 (3.08 to 125) & {\bfseries 0.006} \\ 
+Dimension & 0.57 (0.30 to 0.96) & {\bfseries 0.048} \\ 
+Long runtime & 0.17 (0.00 to 4.96) & 0.357 \\ 
+Number of jellyfish & 1.27 (0.99 to 1.68) & {\bfseries 0.069} \\ 
+Maximum number of tries & 1.10 (0.82 to 1.52) & 0.524 \\ 
+\bottomrule
+
+\end{longtable}
+
+\begin{minipage}{\linewidth}
+\textsuperscript{\textit{1}}OR = Odds Ratio, CI = Confidence Interval\\
+\end{minipage}
+\endgroup
+
 \section{Practicalities}\label{sec-discussion}

### 2024-10-14 jso.qmd
Table 1 before Table 2
fixed alignment of columns in table 1

@@
       "$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", "$\\varsigma$"), linesep = "",
-    booktabs = TRUE, escape = FALSE, row.names = TRUE, align = "c") |> 
-  column_spec(1, border_left = TRUE) |> 
-  column_spec(c(4, 8, 13), border_right = TRUE) |> 
+    booktabs = TRUE, escape = FALSE, row.names = FALSE, align = "|llr|rrrr|rrrrr") |> 
+  #column_spec(1, border_left = TRUE) |> 
+  #column_spec(c(4, 8, 13), border_right = TRUE) |> 
   kable_styling(font_size = 10) 
 ```
+
+@tbl-mod-output presents the results from fitting a logistic regression
+model using the proportion of success as the response and smoothness,
+squintability, dimension, running time, number of jellyfish and maximum
+number of tries, as predictors. Long runtime is a binary variable, where
+greater than 20 seconds is labelled 1. The fit suggests that JSO success
+rate is only affected by squintability, dimension and number of jellies.
+Interestingly, the smoothness of the index is not a problem for the
+JSO - this is promising! Running the optimisation longer and allowing
+more tries also does not affect success rate. An increase in
+squintability of one unit increases the success rate by
+`r round(exp(broom::tidy(mod1)$estimate[3]) * 100, 0)`%. As dimension
+increases by one the success rate halves. Increasing the number of
+jellies by 10, increases the succes rate by
+`r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 0)`%.
 
@@
 
-@tbl-mod-output presents the results from fitting a logistic regression
-model using the proportion of success as the response and smoothness,
-squintability, dimension, running time, number of jellyfish and maximum
-number of tries, as predictors. Long runtime is a binary variable, where
-greater than 20 seconds is labelled 1. The fit suggests that JSO success
-rate is only affected by squintability, dimension and number of jellies.
-Interestingly, the smoothness of the index is not a problem for the
-JSO - this is promising! Running the optimisation longer and allowing
-more tries also does not affect success rate. An increase in
-squintability of one unit increases the success rate by
-`r round(exp(broom::tidy(mod1)$estimate[3]) * 100, 0)`%. As dimension
-increases by one the success rate halves. Increasing the number of
-jellies by 10, increases the succes rate by
-`r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 0)`%.
-
 # Practicalities {#sec-discussion}

### 2024-10-14 jso.tex
fixed alignment of columns in table 1

@@
 \usepackage{caption}
-\usepackage{anyfontsize}
 \usepackage{orcidlink}
@@
 
-\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, opacitybacktitle=0.6, toprule=.15mm, colback=white, opacityback=0, left=2mm, bottomtitle=1mm, breakable, arc=.35mm, bottomrule=.15mm, titlerule=0mm, leftrule=.75mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, bottomrule=.15mm, coltitle=black, leftrule=.75mm, left=2mm, arc=.35mm, colframe=quarto-callout-note-color-frame, opacityback=0, toprule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, opacitybacktitle=0.6, toptitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white]
 
@@
 
-\begin{tabular}{|>{}lcc>{}c|ccc>{}c|cccc>{}c|}
+\begin{tabular}{|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr}
 \toprule
-  & shape & index & d & $\eta$ & $\ell$ & $\nu$ & $\sigma$ & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & $\varsigma$\\
+shape & index & d & $\eta$ & $\ell$ & $\nu$ & $\sigma$ & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & $\varsigma$\\
 \midrule
-1 & pipe & holes & 6 & 0.03 & 0.39 & 3.08 & 0.0 & 1.02 & 0.86 & 3.37 & 0.02 & 3.05\\
-2 & pipe & holes & 8 & 0.02 & 0.32 & 2.86 & 0.0 & 1.01 & 0.87 & 3.26 & 0.03 & 2.96\\
-3 & pipe & holes & 10 & 0.02 & 0.36 & 2.80 & 0.0 & 1.02 & 0.88 & 3.15 & 0.02 & 2.95\\
-4 & pipe & holes & 12 & 0.02 & 0.36 & 2.82 & 0.0 & 1.01 & 0.88 & 3.34 & 0.00 & 3.12\\
-5 & sine & MIC & 6 & 0.06 & 0.29 & 3.15 & 0.0 & 0.89 & 0.57 & 1.62 & -0.02 & 1.26\\
-6 & sine & MIC & 8 & 0.06 & 0.33 & 3.09 & 0.0 & 0.93 & 0.33 & 1.31 & -0.03 & 0.77\\
-7 & sine & TIC & 6 & 0.17 & 0.34 & 3.06 & 0.0 & 0.95 & 0.54 & 1.72 & -0.03 & 1.32\\
-8 & sine & TIC & 8 & 0.16 & 0.31 & 3.17 & 0.0 & 0.95 & 0.56 & 1.72 & -0.03 & 1.37\\
-9 & sine & dcor & 6 & 0.09 & 0.31 & 3.06 & 0.0 & 0.95 & 1.04 & 2.74 & -0.02 & 2.95\\
-10 & sine & loess & 6 & 0.24 & 0.45 & 3.53 & 0.0 & 1.02 & 1.04 & 2.65 & 0.08 & 2.76\\
-11 & sine & splines & 6 & 0.10 & 0.32 & 3.11 & 0.0 & 1.01 & 1.05 & 2.73 & -0.01 & 3.12\\
-12 & sine & stringy & 6 & 0.00 & 1128.28 & 1.03 & 134.8 & 1.05 & 0.01 & 254.73 & 0.05 & 2.96\\
+pipe & holes & 6 & 0.03 & 0.39 & 3.08 & 0.0 & 1.02 & 0.86 & 3.37 & 0.02 & 3.05\\
+pipe & holes & 8 & 0.02 & 0.32 & 2.86 & 0.0 & 1.01 & 0.87 & 3.26 & 0.03 & 2.96\\
+pipe & holes & 10 & 0.02 & 0.36 & 2.80 & 0.0 & 1.02 & 0.88 & 3.15 & 0.02 & 2.95\\
+pipe & holes & 12 & 0.02 & 0.36 & 2.82 & 0.0 & 1.01 & 0.88 & 3.34 & 0.00 & 3.12\\
+sine & MIC & 6 & 0.06 & 0.29 & 3.15 & 0.0 & 0.89 & 0.57 & 1.62 & -0.02 & 1.26\\
+sine & MIC & 8 & 0.06 & 0.33 & 3.09 & 0.0 & 0.93 & 0.33 & 1.31 & -0.03 & 0.77\\
+sine & TIC & 6 & 0.17 & 0.34 & 3.06 & 0.0 & 0.95 & 0.54 & 1.72 & -0.03 & 1.32\\
+sine & TIC & 8 & 0.16 & 0.31 & 3.17 & 0.0 & 0.95 & 0.56 & 1.72 & -0.03 & 1.37\\
+sine & dcor & 6 & 0.09 & 0.31 & 3.06 & 0.0 & 0.95 & 1.04 & 2.74 & -0.02 & 2.95\\
+sine & loess & 6 & 0.24 & 0.45 & 3.53 & 0.0 & 1.02 & 1.04 & 2.65 & 0.08 & 2.76\\
+sine & splines & 6 & 0.10 & 0.32 & 3.11 & 0.0 & 1.01 & 1.05 & 2.73 & -0.01 & 3.12\\
+sine & stringy & 6 & 0.00 & 1128.28 & 1.03 & 134.8 & 1.05 & 0.01 & 254.73 & 0.05 & 2.96\\
 \bottomrule
@@
 
-\begingroup
-\fontsize{12.0pt}{14.4pt}\selectfont
 \setlength{\LTpost}{0mm}
@@
 Smoothness & 89.1 (1.23 to 4,865,159) & 0.289 \\ 
-Squintability & 12.4 (3.08 to 125) & {\bfseries 0.006} \\ 
-Dimension & 0.57 (0.30 to 0.96) & {\bfseries 0.048} \\ 
+Squintability & 12.4 (3.08 to 125) & \textbf{0.006} \\ 
+Dimension & 0.57 (0.30 to 0.96) & \textbf{0.048} \\ 
 Long runtime & 0.17 (0.00 to 4.96) & 0.357 \\ 
-Number of jellyfish & 1.27 (0.99 to 1.68) & {\bfseries 0.069} \\ 
+Number of jellyfish & 1.27 (0.99 to 1.68) & \textbf{0.069} \\ 
 Maximum number of tries & 1.10 (0.82 to 1.52) & 0.524 \\ 
@@
 \end{minipage}
-\endgroup
 

### 2024-10-14 jso.qmd
chunks labelled and plot of data added, that needs index values added

@@
 
+```{r}
+#| label: fig-plot-data
+#| fig-width: 10
+#| fig-height: 5
+#| fig-cap: "The sine and pipe structures in the example data, along with index values for these projections."
+data("sine1000")
+sine1000 <- as_tibble(sine1000)
+dp1 <- ggplot(sine1000, aes(x=V5, y=V6)) + geom_point()
+data("pipe1000")
+pipe1000 <- as_tibble(pipe1000)
+dp2 <- ggplot(pipe1000, aes(x=V5, y=V6)) + geom_point()
+dp1 + dp2 + plot_layout(ncol=2)
+```
+
 Smoothness and squintability are computed following the procedures
@@
 ```{r}
-#| fig.height = 7.5,
-#| fig.width = 10,
-#| label = fig-proj,
-#| fig.align = "center",
-#| fig.cap = "Projections found by the JSO and CRS at each 10th quantile across 50 simulations.
+#| fig.height: 7.5
+#| fig.width: 10
+#| label: fig-proj
+#| fig.align: "center"
+#| fig.cap: "Projections found by the JSO and CRS at each 10th quantile across 50 simulations.
 The projection pursuit problem is to find the pipe shape using the holes index in the 6, 8, 10, and 12-dimensional spaces.
@@
 ```{r}
-#| label = fig-proportion,
-#| fig.width = 8,
-#| fig.height = 3.5,
-#| fig.align = "center",
-#| fig.cap = "The proportion of simulations that reach near-optimal index values in the pipe-finding problem using the holes index.
+#| label: fig-proportion
+#| fig.width: 8
+#| fig.height: 3.5
+#| fig.align: "center"
+#| fig.cap: "The proportion of simulations that reach near-optimal index values in the pipe-finding problem using the holes index.
 The proportion is calculated based on the number of simulations, out of 50, that achieve an index value within 0.05 of the best-performing simulation.
@@
 ```{r}
+#| label: fit-logistic
 sim_df2 <- sim_df |>   
@@
 
-```{r fig-idx-proj-dist, eval = FALSE}
-#| fig.width = 9,
-#| fig.height = 5,
-#| fig.cap = "Index values versus projection distance for the 12 pipe/sine-wave finding problem, after the binning procedure for calculating the squintability measure.
+```{r}
+#| label: fig-idx-proj-dist
+#| fig.width: 9
+#| fig.height: 5
+#| eval: false
+#| fig.cap: "Index values versus projection distance for the 12 pipe/sine-wave finding problem, after the binning procedure for calculating the squintability measure.
 The index values, averaged at bin width of 0.005, are scaled from 0 to 1 for comparison (holes, TIC, and stringy).

### 2024-10-14 jso.tex
chunks labelled and plot of data added, that needs index values added

@@
 
-\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, bottomrule=.15mm, coltitle=black, leftrule=.75mm, left=2mm, arc=.35mm, colframe=quarto-callout-note-color-frame, opacityback=0, toprule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, opacitybacktitle=0.6, toptitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, breakable, colback=white, opacityback=0, leftrule=.75mm, toprule=.15mm, rightrule=.15mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, coltitle=black, bottomrule=.15mm, left=2mm, opacitybacktitle=0.6]
 
@@
 rate for each PP problem.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{jso_files/figure-pdf/fig-plot-data-1.pdf}
+
+}
+
+\caption{\label{fig-plot-data}The sine and pipe structures in the
+example data, along with index values for these projections.}
+
+\end{figure}%
 

### 2024-11-18 jso.qmd
update the modelling result

@@
          max_tries = max_tries/10,
-         long_time = ifelse(time > 20, 1, 0),
-         P_J = ifelse(index == "stringy", 0, P_J)) |> 
+         long_time = ifelse(time > 20, 1, 0)
+         ) |> 
   rename(proportion = P_J)
 
-mod1 <- glm(proportion ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
+mod1 <- glm(proportion ~ smoothness + squintability + d +  n_jellies + max_tries, data = sim_df2, family = binomial)
 ```
@@
 The columns $\nu$ and $\varsigma$ represent the smoothness and squintability measures respectively.
-dt <- tibble(shape = c(rep("pipe", 4), rep("sine", 8)), smoothness) |>
+dt <- tibble(shape = c(rep("pipe", 4), rep("sine", 9)), smoothness) |>
   left_join(squintability) |> 
@@
 set_gtsummary_theme(theme_gtsummary_journal("jama"))
-mod1 |> tbl_regression(exponentiate=TRUE, 
+mod1 |> tbl_regression(exponentiate = TRUE, 
                        pvalue_fun = label_style_pvalue(digits = 3),

### 2024-11-18 jso.tex
update the modelling result

@@
 
-\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, breakable, colback=white, opacityback=0, leftrule=.75mm, toprule=.15mm, rightrule=.15mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, coltitle=black, bottomrule=.15mm, left=2mm, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, opacityback=0, breakable, toptitle=1mm, opacitybacktitle=0.6, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, bottomrule=.15mm, leftrule=.75mm, titlerule=0mm, arc=.35mm, rightrule=.15mm, left=2mm]
 
@@
 \texttt{MIC}, \texttt{TIC}, \texttt{spline}, and \texttt{stringy}, under
-varied JSO hyper-parameters. This results in a total of 52 scenarios,
-comprising of 24 computed on the pipe data and 28 on the sine-wave data.
+varied JSO hyper-parameters. This results in a total of 54 scenarios,
+comprising of 24 computed on the pipe data and 30 on the sine-wave data.
 Again, JSO is run 50 times for each scenario to calculate the success
@@
 \midrule
-pipe & holes & 6 & 0.03 & 0.39 & 3.08 & 0.0 & 1.02 & 0.86 & 3.37 & 0.02 & 3.05\\
-pipe & holes & 8 & 0.02 & 0.32 & 2.86 & 0.0 & 1.01 & 0.87 & 3.26 & 0.03 & 2.96\\
-pipe & holes & 10 & 0.02 & 0.36 & 2.80 & 0.0 & 1.02 & 0.88 & 3.15 & 0.02 & 2.95\\
-pipe & holes & 12 & 0.02 & 0.36 & 2.82 & 0.0 & 1.01 & 0.88 & 3.34 & 0.00 & 3.12\\
-sine & MIC & 6 & 0.06 & 0.29 & 3.15 & 0.0 & 0.89 & 0.57 & 1.62 & -0.02 & 1.26\\
-sine & MIC & 8 & 0.06 & 0.33 & 3.09 & 0.0 & 0.93 & 0.33 & 1.31 & -0.03 & 0.77\\
-sine & TIC & 6 & 0.17 & 0.34 & 3.06 & 0.0 & 0.95 & 0.54 & 1.72 & -0.03 & 1.32\\
-sine & TIC & 8 & 0.16 & 0.31 & 3.17 & 0.0 & 0.95 & 0.56 & 1.72 & -0.03 & 1.37\\
-sine & dcor & 6 & 0.09 & 0.31 & 3.06 & 0.0 & 0.95 & 1.04 & 2.74 & -0.02 & 2.95\\
-sine & loess & 6 & 0.24 & 0.45 & 3.53 & 0.0 & 1.02 & 1.04 & 2.65 & 0.08 & 2.76\\
-sine & splines & 6 & 0.10 & 0.32 & 3.11 & 0.0 & 1.01 & 1.05 & 2.73 & -0.01 & 3.12\\
-sine & stringy & 6 & 0.00 & 1128.28 & 1.03 & 134.8 & 1.05 & 0.01 & 254.73 & 0.05 & 2.96\\
+pipe & holes & 6 & 0.03 & 0.39 & 3.08 & 0.00 & 1.02 & 0.86 & 3.37 & 0.02 & 3.14\\
+pipe & holes & 8 & 0.02 & 0.32 & 2.86 & 0.00 & 1.01 & 0.87 & 3.26 & 0.03 & 3.04\\
+pipe & holes & 10 & 0.02 & 0.36 & 2.80 & 0.00 & 1.02 & 0.88 & 3.15 & 0.02 & 3.05\\
+pipe & holes & 12 & 0.02 & 0.36 & 2.82 & 0.00 & 1.01 & 0.88 & 3.34 & 0.00 & 3.20\\
+sine & MIC & 6 & 0.06 & 0.31 & 3.08 & 0.00 & 0.89 & 0.57 & 1.62 & -0.03 & 1.38\\
+sine & MIC & 8 & 0.03 & 0.32 & 2.75 & 0.00 & 0.93 & 0.33 & 1.31 & -0.04 & 0.83\\
+sine & TIC & 6 & 0.16 & 0.31 & 3.08 & 0.00 & 0.95 & 0.54 & 1.72 & -0.03 & 1.43\\
+sine & TIC & 8 & 0.07 & 0.30 & 2.72 & 0.00 & 0.95 & 0.56 & 1.72 & -0.03 & 1.48\\
+sine & dcor & 6 & 0.08 & 0.29 & 3.13 & 0.00 & 0.95 & 1.04 & 2.74 & -0.03 & 3.20\\
+sine & loess & 6 & 0.24 & 0.46 & 3.56 & 0.00 & 1.02 & 1.04 & 2.65 & 0.07 & 3.02\\
+sine & splines & 6 & 0.10 & 0.32 & 3.11 & 0.00 & 1.01 & 1.05 & 2.73 & -0.02 & 3.40\\
+sine & stringy2 & 6 & 0.01 & 0.09 & 0.71 & 0.97 & 0.65 & 0.05 & 7.78 & 0.06 & 0.39\\
+sine & skinny & 6 & 0.05 & 0.19 & 2.24 & 0.01 & 0.83 & 0.00 & 2.98 & 0.05 & 0.00\\
 \bottomrule
@@
 and allowing more tries also does not affect success rate. An increase
-in squintability of one unit increases the success rate by 1235\%. As
+in squintability of one unit increases the success rate by 483\%. As
 dimension increases by one the success rate halves. Increasing the
-number of jellies by 10, increases the succes rate by 27\%.
-
+number of jellies by 10, increases the succes rate by 8\%.
+
+\begingroup
+\fontsize{12.0pt}{14.4pt}\selectfont
 \setlength{\LTpost}{0mm}
@@
 \midrule\addlinespace[2.5pt]
-Smoothness & 89.1 (1.23 to 4,865,159) & 0.289 \\ 
-Squintability & 12.4 (3.08 to 125) & \textbf{0.006} \\ 
-Dimension & 0.57 (0.30 to 0.96) & \textbf{0.048} \\ 
-Long runtime & 0.17 (0.00 to 4.96) & 0.357 \\ 
-Number of jellyfish & 1.27 (0.99 to 1.68) & \textbf{0.069} \\ 
-Maximum number of tries & 1.10 (0.82 to 1.52) & 0.524 \\ 
+Smoothness & 0.64 (0.09 to 14.1) & 0.673 \\ 
+Squintability & 4.83 (1.99 to 14.2) & {\bfseries 0.001} \\ 
+Dimension & 0.57 (0.36 to 0.85) & {\bfseries 0.008} \\ 
+Number of jellyfish & 1.23 (0.98 to 1.59) & {\bfseries 0.086} \\ 
+Maximum number of tries & 1.08 (0.82 to 1.44) & 0.593 \\ 
 \bottomrule
@@
 \end{minipage}
+\endgroup
 

### 2024-11-18 code/07-calc-smoothness-squintability.R

@@ -1,10 +1,12 @@
 library(ferrn)
 library(tourr)
 library(tidyverse)
+library(cassowaryr)
+library(igraph)
 
 sine_8d_tbl <- function(index){
   tibble::tibble(
-    n = 8, index = index, data = list(sine1000_8d),
+    n = 8, index = index, data = list(as.matrix(sine1000_8d)),
     best = list(matrix(c(rep(0, 12), 1, 0, 0, 1), nrow = 8, byrow = TRUE)))
 }
 
@@ -25,14 +27,15 @@ holes_tidy <- smoothness_holes |>
   dplyr::mutate(smooth = calc_smoothness(basis_df)) |>
   unnest(smooth)
 
-idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "stringy", "splines2d")
+idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "stringy2", "splines2d", "skinny")
 smoothness_sine <- tibble::tibble(
   n = 6, index = idx_names, data = list(sine1000),
   best = list(matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE))) |>
   dplyr::bind_rows(sine_8d_tbl("MIC")) |>
   dplyr::bind_rows(sine_8d_tbl("TIC")) |>
   rowwise() |>
-  dplyr::mutate(basis_df = list(sample_bases(index, n_basis = 500)))
+  dplyr::mutate(basis_df = list(sample_bases(index, n_basis = 500, best = best,
+                                             data = as.matrix(data))))
 
 sine_tidy <- smoothness_sine |>
   rowwise() |>
@@ -41,7 +44,7 @@ sine_tidy <- smoothness_sine |>
 
 smoothness <- bind_rows(sine_tidy, holes_tidy) |>
   mutate(index = factor(index, levels = c("holes", "MIC", "TIC", "dcor2d_2",
-                                          "loess2d", "splines2d", "stringy"))) |>
+                                          "loess2d", "splines2d", "stringy2", "skinny"))) |>
   arrange(index) |>
   select(index, n, variance:nugget)
 save(smoothness, file = here::here("data", "smoothness.rda"))
@@ -64,7 +67,7 @@ sq_holes_basis_df <- tibble::tibble(
                    step_size = 0.005, best = best)}))
[…]

### 2024-11-18 code/08-process-data-for-modelling.R

@@ -28,6 +28,7 @@ load(here::here("data-raw/sim_sine_6d_loess2d.rda"))
 load(here::here("data-raw/sim_sine_68d_TICMIC.rda"))
 load(here::here("data-raw/sim_sine_6d_splines2d.rda"))
 load(here::here("data-raw/sim_sine_6d_stringy.rda"))
+load(here::here("data-raw/sim_sine_6d_skinny.rda"))
 
 sim_data_sine <- bind_rows(
   sim_sine_6d_dcor2d, sim_sine_6d_loess2d, sim_sine_68d_TICMIC,
@@ -48,11 +49,45 @@ sine_setup_summ <- sine_run_df |>
     time = mean(time)
   )
 
+#############################################################################
+#############################################################################
+# process stringy and skinny
+
+row1 <- sim_sine_6d_stringy |> mutate(index = "stringy2") |>
+  mutate(id2 = paste0(index, n_jellies, max_tries, sim, d)) |>
+  get_best(group = id2) |>
+  group_by(index, d, n_jellies, max_tries) |>
+  reframe(
+    I_max = max(index_val),
+    P_J = sum(index_val > 0.46)/n(),
+    time = mean(time)
+  )
+
+
+row2 <- sim_sine_6d_skinny |> mutate(index = "skinny") |>
+  mutate(id2 = paste0(index, n_jellies, max_tries, sim, d)) |>
+  get_best(group = id2) |>
+  group_by(index, d, n_jellies, max_tries) |>
+  reframe(
+    I_max = max(index_val),
+    P_J = sum(index_val > 0.67)/n(),
+    time = mean(time)
+  )
+
+row3 <- sim_sine_6d_splines2d_100 |> mutate(index = "splines2d") |>
+  mutate(id2 = paste0(index, n_jellies, max_tries, sim, d)) |>
+  get_best(group = id2) |>
+  group_by(index, d, n_jellies, max_tries) |>
+  reframe(
+    I_max = max(index_val),
+    P_J = sum(abs(I_max - index_val) <= 0.05)/n(),
+    time = mean(time)
+  )
 
 #############################################################################
 #############################################################################
 # combine sine and pipe success rate
-sim_summary <- pipe_setup_summ |>  bind_rows(sine_setup_summ)
+sim_summary <- sim_summary |> filter(index != "stringy") |> bind_rows(row1, row2, row3)
 save(sim_summary, file = here::here("data", "sim_summary.rda"))
 
[…]

### 2024-11-18 code/09-sim-sine-6d-stringy.R

@@ -11,9 +11,9 @@ colnames(sine1000) <- paste0("V", 1:6)
 # d = 6
 stringy2 <- function(){
   function(mat){
-    x <- scree(mat[,1], mat[,2])
+    x <- cassowaryr::scree(mat[,1], mat[,2])
     b <- cassowaryr:::gen_mst(x$del, x$weights)
-    diameter <- length(get_diameter(b))
+    diameter <- length(igraph::get_diameter(b))
     length <-  length(b) - 1
     diameter / length
   }

### 2025-01-25 code/10-sim-4d.R
4D simulation script

@@ -1,0 +1,176 @@
+library(tidyverse)
+library(spinebil)
+library(tourr)
+
+skinny <- function(){
+  function(mat){
+    cassowaryr::sc_skinny(mat[,1], mat[,2])
+  }
+}
+
+stringy2 <- function(){
+  function(mat){
+    x <- cassowaryr::scree(mat[,1], mat[,2])
+    b <- cassowaryr:::gen_mst(x$del, x$weights)
+    diameter <- length(igraph::get_diameter(b))
+    length <-  length(b) - 1
+    diameter / length
+  }
+}
+
+
+sim_pipe <- function(d = d, n_jellies = n_jellies, max.tries = max.tries,
+                data_seed = 123456, optim_seed = seed){
+  set.seed(data_seed)
+  pipe1000 <- pipeData(d, 1000) %>% scale() %>% as_tibble()
+  colnames(pipe1000) <- paste0("V", 1:d)
+
+  cat("n_jellies: ", n_jellies, "\n")
+  cat("max.tries: ", max.tries, "\n")
+  cat("d: ", d, "\n")
+  res <- map_dfr(optim_seed, function(seed){
+    t1 <- Sys.time()
+    set.seed(seed)
+    res <- animate_xy(pipe1000, guided_tour(
+      holes(), d = 2, n_jellies = n_jellies,
+      search_f =  search_jellyfish,
+      max.tries = max.tries
+    ), max_frames = max.tries) |> dplyr::select(-id)
+    t2 <- Sys.time()
+    res <- res |> mutate(time = t2 - t1)
+    return(res)
+  })
+
+  return(res)
+}
+
+t1 <- Sys.time()
+set.seed(123)
+seed <- sample(1000:10000, size = 50)
+sim_setup <- tidyr::crossing(n_jellies = c(20, 50, 100), max_tries = c(50, 100),
+                             d = 4, sim = 1:50) |>
+  mutate(seed = seed[sim], id = row_number())
+
+sim_res <- sim_setup |>
+  rowwise() |>
+  mutate(res = list(sim_pipe(d, n_jellies, max_tries, optim_seed = seed))) |>
+  ungroup() |>
+  unnest(res)
+t2 <- Sys.time()
+t2 - t1
+
+sim_pipe <- sim_res |> mutate(index = "holes") |> select(index, d, n_jellies, max_tries, sim:time)
+save(sim_pipe, file = "data-raw/sim_pipe_4d.rda")
+
+################################################################################
[…]

### 2025-01-25 code/07-calc-smoothness-squintability.R
include 4d data

@@ -11,10 +11,11 @@ sine_8d_tbl <- function(index){
 }
 
 smoothness_holes <- tibble::tibble(
-  n = c(6, 8, 10, 12),
-  index = rep("holes", 4),
-  data = list(pipe1000, pipe1000_8d, pipe1000_10d, pipe1000_12d),
-  best = list(matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE),
+  n = c(4, 6, 8, 10, 12),
+  index = rep("holes", 5),
+  data = list(pipe1000_4d, pipe1000_6d, pipe1000_8d, pipe1000_10d, pipe1000_12d),
+  best = list(matrix(c(rep(0, 4), 1, 0, 0, 1), nrow = 4, byrow = TRUE),
+              matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE),
               matrix(c(rep(0, 12), 1, 0, 0, 1), nrow = 8, byrow = TRUE),
               matrix(c(rep(0, 16), 1, 0, 0, 1), nrow = 10, byrow = TRUE),
               matrix(c(rep(0, 20), 1, 0, 0, 1), nrow = 12, byrow = TRUE))) |>
@@ -37,8 +38,15 @@ smoothness_sine <- tibble::tibble(
   dplyr::mutate(basis_df = list(sample_bases(index, n_basis = 500, best = best,
                                              data = as.matrix(data))))
 
-sine_tidy <- smoothness_sine |>
+idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "splines2d")
+smoothness_4d <- tibble::tibble(
+  n = 4, index = idx_names, data = list(sine1000_4d),
+  best = list(matrix(c(rep(0, 4), 1, 0, 0, 1), nrow = 4, byrow = TRUE))) |>
   rowwise() |>
+  dplyr::mutate(basis_df = list(sample_bases(index, n_basis = 500, best = best,
+                                             data = as.matrix(data))))
+
+sine_tidy <- bind_rows(smoothness_sine, smoothness_4d) |>
   dplyr::mutate(smooth = calc_smoothness(basis_df)) |>
   unnest(smooth)
 
@@ -54,10 +62,11 @@ save(smoothness, file = here::here("data", "smoothness.rda"))
 # squintability - sample bases
 # 2-3 mins
 sq_holes_basis_df <- tibble::tibble(
-  n = c(6, 8, 10, 12),
-  index = rep("holes", 4),
[…]

### 2025-01-25 code/08-process-data-for-modelling.R
update post-processing of simulation data

@@ -52,7 +52,6 @@ sine_setup_summ <- sine_run_df |>
 #############################################################################
 #############################################################################
 # process stringy and skinny
-
 row1 <- sim_sine_6d_stringy |> mutate(index = "stringy2") |>
   mutate(id2 = paste0(index, n_jellies, max_tries, sim, d)) |>
   get_best(group = id2) |>
@@ -86,8 +85,43 @@ row3 <- sim_sine_6d_splines2d_100 |> mutate(index = "splines2d") |>
 
 #############################################################################
 #############################################################################
+# process 4d data
+load(here::here("data-raw/sim_pipe_4d.rda"))
+load(here::here("data-raw/sim_sine_4d_dcor2d.rda"))
+load(here::here("data-raw/sim_sine_4d_loess2d.rda"))
+load(here::here("data-raw/sim_sine_4d_MICTIC.rda"))
+load(here::here("data-raw/sim_sine_4d_splines2d.rda"))
+sim_pipe_4d <- sim_pipe |>
+  mutate(id2 = paste0(index, n_jellies, max_tries, sim, d)) |>
+  get_best(group = id2) |>
+  group_by(index, d, n_jellies, max_tries) |>
+  reframe(
+    I_max = max(index_val),
+    P_J = sum(index_val > 0.46)/n(),
+    time = mean(time)
+  )
+
+sim_data_sine_4d <- bind_rows(
+  sim_sine_4d_dcor2d |> mutate(index = "dcor2d_2") , sim_sine_4d_loess2d, sim_sine_4d_MICTIC,
+  sim_sine_4d_splines2d)
+
+sine_run_df_4d <- sim_data_sine_4d |>
+  mutate(id2 = paste0(index, n_jellies, max_tries, sim, d)) |>
+  get_best(group = id2)
+
+sine_setup_summ_4d <- sine_run_df_4d |>
+  group_by(index, d, n_jellies, max_tries) |>
+  reframe(
+    I_max = max(index_val),
+    P_J = sum(abs(I_max - index_val) <= 0.05)/n(),
+    time = mean(time)
+  )
+
+#############################################################################
+#############################################################################
 # combine sine and pipe success rate
-sim_summary <- sim_summary |> filter(index != "stringy") |> bind_rows(row1, row2, row3)
+sim_summary <- sim_summary |> filter(index != "stringy") |>
+  bind_rows(row1, row2, row3, sim_pipe_4d, sine_setup_summ_4d) |> select(-time)
[…]

### 2025-01-27 code/04-sim-sine-68d-MICTIC.R
add 8d loess & splines2d, fix MIC TIC

@@ -59,9 +59,8 @@ set.seed(123)
 seed <- sample(1000: 10000, size = 50)
 sim_setup <- crossing(idx_f = c("MIC", "TIC"),
                       d = c(6, 8),
-                      n_jellies = c(20, 50, 100),
+                      n_jellies = c(20, 50),
                       max_tries = c(50, 100)) |>
-  filter(!(n_jellies == 100 & max_tries == 100)) |>
   crossing(sim = 1) |>
   mutate(seed = seed[sim], id = row_number())
 

### 2025-01-27 code/08-process-data-for-modelling.R

@@ -29,10 +29,13 @@ load(here::here("data-raw/sim_sine_68d_TICMIC.rda"))
 load(here::here("data-raw/sim_sine_6d_splines2d.rda"))
 load(here::here("data-raw/sim_sine_6d_stringy.rda"))
 load(here::here("data-raw/sim_sine_6d_skinny.rda"))
+load(here::here("data-raw/sim_sine_8d_splines2d.rda"))
+load(here::here("data-raw/sim_sine_8d_loess2d.rda"))
 
 sim_data_sine <- bind_rows(
   sim_sine_6d_dcor2d, sim_sine_6d_loess2d, sim_sine_68d_TICMIC,
-  sim_sine_6d_splines2d, sim_sine_6d_stringy)
+  sim_sine_6d_splines2d, sim_sine_6d_stringy,
+  sim_sine_8d_splines2d, sim_sine_8d_loess2d)
 
 sine_run_df <- sim_data_sine |>
   mutate(id2 = paste0(index, n_jellies, max_tries, sim, d)) |>
@@ -45,8 +48,7 @@ sine_setup_summ <- sine_run_df |>
   group_by(index, d, n_jellies, max_tries) |>
   reframe(
     I_max = max(index_val),
-    P_J = sum(abs(I_max - index_val) <= 0.05)/n(),
-    time = mean(time)
+    P_J = sum(abs(I_max - index_val) <= 0.05)/n()
   )
 
 #############################################################################
@@ -132,3 +134,6 @@ sim_df <- sim_summary |>
   left_join(smoothness |> select(n, index, smoothness) |> rename(d = n)) |>
   left_join(squintability |> select(index, n, squint) |> rename(d = n, squintability = squint))
 save(sim_df, file = here::here("data", "sim_df.rda"))
+
+
+sim_summary <- sim_summary |> bind_rows(sine_setup_summ)

### 2025-01-27 code/10-sim-4d.R

@@ -174,3 +174,51 @@ t2 - t1
 sim_sine_4d_dcor2d <- sim_res
 save(sim_sine_4d_dcor2d, file = "data-raw/sim_sine_4d_dcor2d.rda")
 
+################################################################################
+################################################################################
+# about an hour
+# 20/ 50 can't find
+# 30/ 50 can't find
+# 30/ 100 can't find
+set.seed(123)
+seed <- sample(1000: 10000, size = 50)
+sim_setup <- tidyr::crossing(n_jellies = 50, max_tries = 100, d = 8) |>
+  tidyr::crossing(sim = 1:50) |>
+  mutate(seed = seed[sim], id = row_number())
+
+t1 <- Sys.time()
+set.seed(123)
+sim_res <- sim_setup |>
+  mutate(index = "splines2d") |>
+  rowwise() |>
+  mutate(res = list(sim(index = index, d, n_jellies, max_tries, optim_seed = seed, sim = sim))) |>
+  ungroup() |>
+  unnest(res)
+t2 <- Sys.time()
+t2 - t1
+sim_sine_8d_splines2d <- sim_res
+save(sim_sine_8d_splines2d, file = "data-raw/sim_sine_8d_splines2d.rda")
+
+################################################################################
+################################################################################
+# about an hour
+set.seed(123)
+seed <- sample(1000: 10000, size = 50)
+sim_setup <- tidyr::crossing(n_jellies = c(30, 50), max_tries = c(50, 100), d = 8) |>
+  filter(!(n_jellies == 50 & max_tries == 100)) |>
+  tidyr::crossing(sim = 1:50) |>
+  mutate(seed = seed[sim], id = row_number())
+
+t1 <- Sys.time()
+set.seed(123)
+sim_res <- sim_setup |>
+  mutate(index = "loess2d") |>
+  rowwise() |>
+  mutate(res = list(sim(index = index, d, n_jellies, max_tries, optim_seed = seed, sim = sim))) |>
+  ungroup() |>
+  unnest(res)
+t2 <- Sys.time()
+t2 - t1
+sim_sine_8d_loess2d <- sim_res
+save(sim_sine_8d_loess2d, file = "data-raw/sim_sine_8d_loess2d.rda")
+

### 2025-01-28 code/07-calc-smoothness-squintability.R
add 4d stringy2, 8d loess2d and splines2d

@@ -34,11 +34,13 @@ smoothness_sine <- tibble::tibble(
   best = list(matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE))) |>
   dplyr::bind_rows(sine_8d_tbl("MIC")) |>
   dplyr::bind_rows(sine_8d_tbl("TIC")) |>
+  dplyr::bind_rows(sine_8d_tbl("loess2d")) |>
+  dplyr::bind_rows(sine_8d_tbl("splines2d")) |>
   rowwise() |>
   dplyr::mutate(basis_df = list(sample_bases(index, n_basis = 500, best = best,
                                              data = as.matrix(data))))
 
-idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "splines2d")
+idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "splines2d", "stringy2")
 smoothness_4d <- tibble::tibble(
   n = 4, index = idx_names, data = list(sine1000_4d),
   best = list(matrix(c(rep(0, 4), 1, 0, 0, 1), nrow = 4, byrow = TRUE))) |>
@@ -83,6 +85,8 @@ sq_sine_basis_df <- tibble::tibble(
   best = list(matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE))) |>
   dplyr::bind_rows(sine_8d_tbl("MIC")) |>
   dplyr::bind_rows(sine_8d_tbl("TIC")) |>
+  dplyr::bind_rows(sine_8d_tbl("loess2d")) |>
+  dplyr::bind_rows(sine_8d_tbl("splines2d")) |>
   dplyr::rowwise() |>
   dplyr::mutate(basis_df = list(sample_bases(
     idx = index, data = as.matrix(data), n_basis = 50, min_proj_dist = 1.5,
@@ -93,7 +97,7 @@ sq_sine_basis_df <- tibble::tibble(
 sq_sine_basis_df2 <- sq_sine_basis_df |> filter(index != "stringy") |> bind_rows(stringy)
 
 # about 5 mins
-idx_names <- idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "splines2d")
+idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "splines2d", "stringy2")
 sq_4d_basis_df <- tibble::tibble(
[…]

### 2025-01-28 jso.qmd
add 4d stringy and other updates

@@
 sim_df2 <- sim_df |>   
-  mutate(n_jellies = n_jellies/10,
-         max_tries = max_tries/10,
-         long_time = ifelse(time > 20, 1, 0)
-         ) |> 
+  mutate(n_jellies = n_jellies/10, max_tries = max_tries/10) |> 
   rename(proportion = P_J)
@@
 The columns $\nu$ and $\varsigma$ represent the smoothness and squintability measures respectively.
-dt <- tibble(shape = c(rep("pipe", 4), rep("sine", 9)), smoothness) |>
+dt <- tibble(shape = c(rep("pipe", 4), rep("sine", 18)), smoothness) |>
   left_join(squintability) |> 

### 2025-01-28 jso.tex
add 4d stringy and other updates

@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 \usepackage{bookmark}
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{jso_files/figure-pdf/fig-example-functions-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{jso_files/figure-pdf/fig-example-functions-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_1d.png}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/matern_simulation_1d.png}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_2d.png}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/matern_simulation_2d.png}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, opacityback=0, breakable, toptitle=1mm, opacitybacktitle=0.6, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, bottomrule=.15mm, leftrule=.75mm, titlerule=0mm, arc=.35mm, rightrule=.15mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, colback=white, bottomrule=.15mm, opacitybacktitle=0.6, coltitle=black, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, breakable, rightrule=.15mm, left=2mm, opacityback=0, leftrule=.75mm]
 
@@
 
-\includegraphics[width=4.52in,height=\textheight]{figures/success-rate.png}
+\includegraphics[width=4.52in,height=\textheight,keepaspectratio]{figures/success-rate.png}
 
@@
 \texttt{MIC}, \texttt{TIC}, \texttt{spline}, and \texttt{stringy}, under
-varied JSO hyper-parameters. This results in a total of 54 scenarios,
-comprising of 24 computed on the pipe data and 30 on the sine-wave data.
+varied JSO hyper-parameters. This results in a total of 76 scenarios,
+comprising of 30 computed on the pipe data and 46 on the sine-wave data.
 Again, JSO is run 50 times for each scenario to calculate the success
@@
 
-\includegraphics{jso_files/figure-pdf/fig-plot-data-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-plot-data-1.pdf}}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{figures/smoothness.png}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/smoothness.png}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{figures/squintability.png}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/squintability.png}
 
@@
 
-\includegraphics{jso_files/figure-pdf/fig-proj-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-proj-1.pdf}}
 
@@
 
-\includegraphics{jso_files/figure-pdf/fig-proportion-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-proportion-1.pdf}}
 
@@
 \midrule
-pipe & holes & 6 & 0.03 & 0.39 & 3.08 & 0.00 & 1.02 & 0.86 & 3.37 & 0.02 & 3.14\\
-pipe & holes & 8 & 0.02 & 0.32 & 2.86 & 0.00 & 1.01 & 0.87 & 3.26 & 0.03 & 3.04\\
-pipe & holes & 10 & 0.02 & 0.36 & 2.80 & 0.00 & 1.02 & 0.88 & 3.15 & 0.02 & 3.05\\
-pipe & holes & 12 & 0.02 & 0.36 & 2.82 & 0.00 & 1.01 & 0.88 & 3.34 & 0.00 & 3.20\\
-sine & MIC & 6 & 0.06 & 0.31 & 3.08 & 0.00 & 0.89 & 0.57 & 1.62 & -0.03 & 1.38\\
-sine & MIC & 8 & 0.03 & 0.32 & 2.75 & 0.00 & 0.93 & 0.33 & 1.31 & -0.04 & 0.83\\
-sine & TIC & 6 & 0.16 & 0.31 & 3.08 & 0.00 & 0.95 & 0.54 & 1.72 & -0.03 & 1.43\\
-sine & TIC & 8 & 0.07 & 0.30 & 2.72 & 0.00 & 0.95 & 0.56 & 1.72 & -0.03 & 1.48\\
-sine & dcor & 6 & 0.08 & 0.29 & 3.13 & 0.00 & 0.95 & 1.04 & 2.74 & -0.03 & 3.20\\
-sine & loess & 6 & 0.24 & 0.46 & 3.56 & 0.00 & 1.02 & 1.04 & 2.65 & 0.07 & 3.02\\
-sine & splines & 6 & 0.10 & 0.32 & 3.11 & 0.00 & 1.01 & 1.05 & 2.73 & -0.02 & 3.40\\
-sine & stringy2 & 6 & 0.01 & 0.09 & 0.71 & 0.97 & 0.65 & 0.05 & 7.78 & 0.06 & 0.39\\
-sine & skinny & 6 & 0.05 & 0.19 & 2.24 & 0.01 & 0.83 & 0.00 & 2.98 & 0.05 & 0.00\\
+pipe & holes & 4 & 0.03 & 0.36 & 2.99 & 0.00 & 1.00 & 0.78 & 4.50 & 0.03 & 0.31\\
+pipe & holes & 6 & 0.03 & 0.38 & 3.09 & 0.00 & 1.02 & 0.86 & 3.37 & 0.02 & 0.41\\
+pipe & holes & 8 & 0.02 & 0.33 & 2.90 & 0.00 & 1.01 & 0.87 & 3.26 & 0.03 & 0.41\\
+pipe & holes & 10 & 0.02 & 0.34 & 2.88 & 0.00 & 1.02 & 0.88 & 3.15 & 0.02 & 0.43\\
+sine & holes & 12 & 0.02 & 0.43 & 2.72 & 0.00 & 1.01 & 0.88 & 3.34 & 0.00 & 0.41\\
+sine & MIC & 4 & 0.14 & 0.27 & 3.56 & 0.00 & 0.94 & 0.15 & 1.14 & -0.04 & 0.39\\
+sine & MIC & 6 & 0.06 & 0.31 & 3.08 & 0.00 & 0.89 & 0.57 & 1.62 & -0.03 & 0.39\\
+sine & MIC & 8 & 0.03 & 0.32 & 2.75 & 0.00 & 0.93 & 0.33 & 1.31 & -0.04 & 0.38\\
+sine & TIC & 4 & 0.40 & 0.27 & 3.49 & 0.00 & 0.97 & 0.46 & 1.55 & -0.04 & 0.38\\
+sine & TIC & 6 & 0.16 & 0.31 & 3.08 & 0.00 & 0.95 & 0.54 & 1.72 & -0.03 & 0.37\\
+sine & TIC & 8 & 0.07 & 0.30 & 2.72 & 0.00 & 0.95 & 0.56 & 1.72 & -0.03 & 0.38\\
+sine & dcor & 4 & 0.22 & 0.30 & 3.63 & 0.00 & 0.96 & 1.07 & 2.67 & -0.03 & 0.56\\
+sine & dcor & 6 & 0.08 & 0.29 & 3.13 & 0.00 & 0.95 & 1.04 & 2.74 & -0.03 & 0.53\\
+sine & loess & 4 & 0.34 & 0.50 & 2.90 & 0.00 & 1.02 & 1.10 & 2.50 & 0.05 & 0.57\\
+sine & loess & 6 & 0.24 & 0.46 & 3.56 & 0.00 & 1.02 & 1.04 & 2.65 & 0.07 & 0.53\\
+sine & loess & 8 & 0.19 & 0.46 & 3.36 & 0.00 & 1.02 & 1.04 & 2.62 & 0.08 & 0.53\\
+sine & skinny & 6 & 0.05 & 0.19 & 2.24 & 0.01 & 0.83 & 0.00 & 2.98 & 0.05 & 0.10\\
+sine & splines & 4 & 0.24 & 0.32 & 3.55 & 0.00 & 1.01 & 1.09 & 2.65 & -0.03 & 0.57\\
+sine & splines & 6 & 0.10 & 0.32 & 3.11 & 0.00 & 1.01 & 1.05 & 2.73 & -0.02 & 0.54\\
+sine & splines & 8 & 0.05 & 0.29 & 2.74 & 0.00 & 1.02 & 1.06 & 2.68 & -0.02 & 0.54\\
+sine & stringy2 & 4 & 0.02 & 0.30 & 2.76 & 1.01 & 0.64 & 0.05 & 7.89 & 0.05 & 0.00\\
+sine & stringy2 & 6 & 0.02 & 0.07 & 1.51 & 0.05 & 0.65 & 0.05 & 7.78 & 0.06 & 0.00\\
 \bottomrule
@@
 and allowing more tries also does not affect success rate. An increase
-in squintability of one unit increases the success rate by 483\%. As
+in squintability of one unit increases the success rate by 124\%. As
 dimension increases by one the success rate halves. Increasing the
-number of jellies by 10, increases the succes rate by 8\%.
+number of jellies by 10, increases the succes rate by 6\%.
 
@@
 \midrule\addlinespace[2.5pt]
-Smoothness & 0.64 (0.09 to 14.1) & 0.673 \\ 
-Squintability & 4.83 (1.99 to 14.2) & {\bfseries 0.001} \\ 
-Dimension & 0.57 (0.36 to 0.85) & {\bfseries 0.008} \\ 
-Number of jellyfish & 1.23 (0.98 to 1.59) & {\bfseries 0.086} \\ 
-Maximum number of tries & 1.08 (0.82 to 1.44) & 0.593 \\ 
+Smoothness & 1.00 (0.81 to 1.20) & 0.961 \\ 
+Squintability & 1.24 (1.05 to 1.58) & {\bfseries 0.033} \\ 
+Dimension & 0.57 (0.33 to 0.87) & {\bfseries 0.022} \\ 
+Number of jellyfish & 1.32 (1.06 to 1.71) & {\bfseries 0.022} \\ 
+Maximum number of tries & 1.06 (0.85 to 1.35) & 0.603 \\ 
 \bottomrule

### 2025-01-28 code/07-calc-smoothness-squintability.R

@@ -28,6 +28,7 @@ holes_tidy <- smoothness_holes |>
   dplyr::mutate(smooth = calc_smoothness(basis_df)) |>
   unnest(smooth)
 
+##################################################
 idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "stringy2", "splines2d", "skinny")
 smoothness_sine <- tibble::tibble(
   n = 6, index = idx_names, data = list(sine1000),
@@ -141,11 +142,10 @@ save(squintability, file = here::here("data", "squintability.rda"))
 ############################################################################
 ############################################################################
 # check smoothness against squintability
-smoothness |>
-  left_join(squintability, by = c("index", "n")) |>
-  dplyr::select(index,n, smoothness, squint) |>
-  ggplot(aes(x = smoothness, y = squint, group = index)) +
-  geom_point(aes(color = as.factor(n))) +
+smoothness |> select(n, index, smoothness) |>
+  left_join(squintability |> select(n, index, squint) |> rename(squintability = squint)) |>
+  ggplot(aes(x = smoothness, y = squintability, group = index, color = as.factor(index))) +
+  geom_point() +
   ggrepel::geom_label_repel(aes(label = index), nudge_x = 0.01, nudge_y = 0.01) +
   theme(aspect.ratio = 1)
 

### 2025-01-28 code/08-process-data-for-modelling.R

@@ -27,14 +27,16 @@ load(here::here("data-raw/sim_sine_6d_dcor2d.rda"))
 load(here::here("data-raw/sim_sine_6d_loess2d.rda"))
 load(here::here("data-raw/sim_sine_68d_TICMIC.rda"))
 load(here::here("data-raw/sim_sine_6d_splines2d.rda"))
-load(here::here("data-raw/sim_sine_6d_stringy.rda"))
+#load(here::here("data-raw/sim_sine_6d_stringy.rda"))
 load(here::here("data-raw/sim_sine_6d_skinny.rda"))
 load(here::here("data-raw/sim_sine_8d_splines2d.rda"))
 load(here::here("data-raw/sim_sine_8d_loess2d.rda"))
+load(here::here("data-raw/sim_sine_4d_stringy.rda"))
 
 sim_data_sine <- bind_rows(
-  sim_sine_6d_dcor2d, sim_sine_6d_loess2d, sim_sine_68d_TICMIC,
-  sim_sine_6d_splines2d, sim_sine_6d_stringy,
+  sim_sine_6d_dcor2d, sim_sine_6d_loess2d |> mutate(index = "loess2d"),
+  sim_sine_68d_TICMIC,
+  sim_sine_6d_splines2d,
   sim_sine_8d_splines2d, sim_sine_8d_loess2d)
 
 sine_run_df <- sim_data_sine |>
@@ -60,8 +62,16 @@ row1 <- sim_sine_6d_stringy |> mutate(index = "stringy2") |>
   group_by(index, d, n_jellies, max_tries) |>
   reframe(
     I_max = max(index_val),
-    P_J = sum(index_val > 0.46)/n(),
-    time = mean(time)
+    P_J = sum(index_val > 0.46)/n()
+  )
+
+row4 <- sim_sine_4d_stringy |> mutate(index = "stringy2") |>
+  mutate(id2 = paste0(index, n_jellies, max_tries, sim, d)) |>
+  get_best(group = id2) |>
+  group_by(index, d, n_jellies, max_tries) |>
+  reframe(
+    I_max = max(index_val),
+    P_J = sum(index_val > 0.595)/n()
   )
 
 
@@ -71,8 +81,7 @@ row2 <- sim_sine_6d_skinny |> mutate(index = "skinny") |>
   group_by(index, d, n_jellies, max_tries) |>
   reframe(
     I_max = max(index_val),
[…]

### 2025-01-28 code/10-sim-4d.R

@@ -59,8 +59,8 @@ sim_res <- sim_setup |>
 t2 <- Sys.time()
 t2 - t1
 
-sim_pipe <- sim_res |> mutate(index = "holes") |> select(index, d, n_jellies, max_tries, sim:time)
-save(sim_pipe, file = "data-raw/sim_pipe_4d.rda")
+sim_pipe_4d <- sim_res |> mutate(index = "holes") |> select(index, d, n_jellies, max_tries, sim:time)
+save(sim_pipe_4d, file = "data-raw/sim_pipe_4d.rda")
 
 ################################################################################
 ################################################################################
@@ -222,3 +222,25 @@ t2 - t1
 sim_sine_8d_loess2d <- sim_res
 save(sim_sine_8d_loess2d, file = "data-raw/sim_sine_8d_loess2d.rda")
 
+################################################################################
+################################################################################
+# n_jellies = 30/50, max_tries = 50, sim = 1:2, 4.5/ 7.8 mins
+# n_jellies = 20, max_tries = 50, sim = 1:2, 2.8 mins, about 2.5 hrs for 50 simulations
+set.seed(123)
+seed <- sample(1000: 10000, size = 50)
+sim_setup <- tidyr::crossing(n_jellies = 20, max_tries = 50, d = 4) |>
+  tidyr::crossing(sim = 1:50) |>
+  mutate(seed = seed[sim], id = row_number())
+
+t1 <- Sys.time()
+set.seed(123)
+sim_res <- sim_setup |>
+  mutate(index = "stringy2") |>
+  rowwise() |>
+  mutate(res = list(sim(index = index, d, n_jellies, max_tries, optim_seed = seed, sim = sim))) |>
+  ungroup() |>
+  unnest(res)
+t2 <- Sys.time()
+t2 - t1
+sim_sine_4d_stringy <- sim_res
+save(sim_sine_4d_stringy, file = "data-raw/sim_sine_4d_stringy.rda")

### 2025-02-05 jso.qmd
squintability
Updated squintability section + minor corrections

@@
         - id: 3
-          name: Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College
+          name: Guangdong Provincial/Zhuhai Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College
           department: Department of Mathematical Sciences
@@
 
-Projection pursuit (PP) (@kr69, @FT74, @huber85) is a dimension
+Projection pursuit (PP) [@kr69; @FT74; @huber85] is a dimension
 reduction technique aimed at identifying informative linear projections
@@
 publication. The method involves optimising an objective function known
-as the PP index (e.g. @hall1989polynomial, @cook1993projection,
-@lee2010projection, @Loperfido2018, @Loperfido2020), which defines the
+as the PP index [e.g., @hall1989polynomial; @cook1993projection;
+@lee2010projection; @Loperfido2018; @Loperfido2020], which defines the
 criterion for what constitutes interesting or informative projections.
@@
 smooth with a large squint angle but has three modes, and thus is not
-convex. The binned normality index (a simple version of a non-normality
-index as described in @huber85) computed on the famous RANDU data, in
+convex. The binned normality index [a simple version of a non-normality
+index as described in @huber85] computed on the famous RANDU data, in
 (c), is noisier and has a very small squint angle. The discreteness
@@
 
-Here, the potential for a Jellyfish Search Optimiser (JSO) (see
-@chou_novel_2021, and @rajwar_exhaustive_2023) for the PPGT is explored.
+Here, the potential for a Jellyfish Search Optimiser (JSO) [see
+@chou_novel_2021, and @rajwar_exhaustive_2023] for the PPGT is explored.
 JSO, inspired by the search behaviour of jellyfish in the ocean, is a
@@
 PP indexes, metrics are introduced to capture specific properties of the
-index including squintability (based on @barnett1981interpreting's
-squint angle) and smoothness. Here, we mathematically define metrics for
+index including squintability [based on @barnett1981interpreting's
+squint angle] and smoothness. Here, we mathematically define metrics for
 squintability and smoothness, which is a new contribution for PP
@@
 
+```{=html}
 <!-- 
@@
 -->
-
+```
 There are numerous PP index functions available: introduced in @huber85,
@@
 dimension, and implemented because they are relatively easy to optimise.
-A goal is to develop PP indexes based on scagnostics (@scag, @WW08), but
+A goal is to develop PP indexes based on scagnostics [@scag, @WW08], but
 the blockage is their optimisation as these tend to be noisy, with
@@
 
-\begin{defn}[sobolev space]\label{def:sobolev_space}
+```{=tex}
+\begin{defn}[Sobolev space]\label{def:sobolev_space}
 The Sobolev space $W^{k,p}(\mathbb{R})$ for $1\leq p\leq \infty$ is the set of all functions $f$ in $L^p(\mathbb{R})$ for which all weak derivatives $f^{(\ell)}$ of order $\ell\leq k$ exist and have a finite $L^p$ norm.
 \end{defn}
-
+```
 The Sobolev index $k$ in Definition \ref{def:sobolev_space} can be used
@@
 
+```{=tex}
 \begin{defn}[Matérn covariance function]\label{def:matern}
@@
 \end{defn}
-
+```
 The Matérn covariance function can be expressed analytically when $\nu$
@@
 
+```{=tex}
 \begin{defn}
@@
 \end{defn}
-
+```
 The value of the optimal smoothness parameter $\nu>0$ can be naturally
@@
 
+```{=tex}
 \begin{defn}[projection distance]\label{def:proj-dist}
-Let $A \in \mathbb{R}^{p \times d}$ is a $d$-dimensional orthonormal matrix, and let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.
+Let $A \in \mathbb{R}^{p \times d}$ be a $d$-dimensional orthonormal matrix, and let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.
 The projection distance between $A$
@@
 \end{defn}
-
-Squintability can be defined as how the index value $f(XA)$ changes with
+```
+Squintability can be defined based on how the index value $f(XA)$ changes with
 respect to the projection distance $r(A, A^*)$, over the course of the
-JSO:
-
+JSO. Suppose that the optimization starts with an arbitrary candidate
+matrix $A_0$ at a distance $r_0:=r(A_0, A^*)$ from the optimal one. A possible way to define squintability is as follows:
+
+```{=tex}
 \begin{defn}[squintability]\label{def:squintability}
@@
 \begin{equation}
-\varsigma(f) = -4 \times \min_{r} g'(r) \times \arg \min_{r} g'(r)
+\varsigma(f) = \frac{g(r_{0}/2)-g(r_{0})}{g(0)-g(r_{0})} \in [0,1]
 \label{eq-squintability}
@@
 
-where $4$ is a constant scaling factor, $-\min_r g'(r)$ represents the 
-largest gradient of $-g$ and $\arg \min_{r} g'(r)$ represents the projection distance at which this largest gradient is attained.
-If the set $\mathcal{D}:=\arg \min_{r} g'(r)$ contains more than one value $r$, the largest value $\max(\mathcal{D})$ is used in equation \eqref{eq-squintability}.
-
 \end{defn}
-
-It is expected that these two values should be both high in the case of
-high squintability (fast increase in $g$ early on), and both low in the
+```
+
+Remark that the amount by which the index value can improve from the starting matrix $A_0$ to the optimal one $A^*$ is given by $g(0)-g(r_{0})$.
+As a result, equation \eqref{eq-squintability} represent the proportion of this maximum improvement which has been achieved by the time the distance $r_0$ to the optimal matrix has been reduced by half ($r_0/2$).
+
+It is expected that this value should be high in the case of
+high squintability (fast increase in $g$ early on), and low in the
 case of low squintability (any substantial increase in $g$ happens very
-late, close to the optimal angle). This suggests that their product
-\eqref{eq-squintability} should provide a sensible measure of
-squintability. The multiplicative constant $4$, which can be deemed
-arbitrary, does not change the interpretation of the squintability
-metric $\varsigma$; it is here to adjust the range of values of
-$\varsigma$ and simplify the explicit formula for $\varsigma$ obtained
-later on.
-
-From the definition, the following proposition can be derived:
-
-\begin{prop}\label{prop:convex-concave}
-Let $g_1$ be a convex, strictly decreasing function and $g_2$ be a concave, strictly decreasing function, both defined on $[0, R]$ where $R>0$, and taking values in $[0, 1]$. Then
-$$\varsigma(g_1) < \varsigma(g_2)$$
-\end{prop}
-\begin{proof}
-Let $r_1 := \underset{r}{\arg \min }\,g_1'(r)$ and $r_2 := \underset{r}{\arg \min }\,g_2'(r)$.
-Since $g_1$ is convex, $g_1''(r) \geq 0 \; \forall r \in [0, R]$, meaning that $g_1'$ is an increasing function, taking its minimum for $r_1=0$.
-As a result, $\varsigma(g_1)=-4 \times g'(0) \times 0 =0$.
-Since $g_2$ is concave, $g_2''(r) \leq 0 \; \forall r \in [0, R]$, meaning that $g_2'$ is a decreasing function, taking its minimum for $r_2=D>0$.
-Since $g_2$ is strictly decreasing, $g_2'(r) < 0 \; \forall r \in [0, R]$.
-In particular $g_2'(R)<0$.
-As a result $\varsigma(g_2)=-4 \times g'(R) \times R >0$.
-This proves that $\varsigma(g_1) < \varsigma(g_2)$.
-\end{proof}
+late, close to the optimal angle). This suggests that this half-point improvement metric should provide a sensible measure of
+squintability.
 
@@
 the structure of the index function to be visible enough to be amenable
-to efficient optimization. This observation suggests that the extreme
-values of $g'$ and the projection distances for which these values are
-attained, are crucial in the mathematical definition of squintability.
+to efficient optimization.
+This observation suggests that the half-point index value improvement percentage provides an appropriate mathematical definition of squintability, which matches the intuition behind this concept, while being amenable to numerical computation.
 
@@
 Then, define $$
-g(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
+g(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(r_0)}{\ell(0)-\ell(r_0)}+\theta_{4}\ ,
 $$ {#eq-parametric}
@@
 which depends on three additional parameters, $\theta_1$, $\theta_2$,
-and $x_{\max}$, such that $g(0)=\theta_1$ and $g(x_{\max})=\theta_4$.
+and $r_0$, such that $g(0)=\theta_1$ and $g(r_0)=\theta_4$.
 Under the parametric model (@eq-parametric), the squintability metric
 \eqref{eq-squintability} can be shown to be equal to $$
-\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
+\varsigma=\frac{g\left(r_{0}/2\right)-\theta_{4}}{\theta_{1}-\theta_{4}}=\frac{\ell(r_0/2)-\ell(r_0)}{\ell(0)-\ell(r_0)}\ .
 $$ {#eq-squintability-parametric}
@@
 nonparametric way, for example by fitting $g$ using kernel regression,
-then numerically estimate the angle at which $-g'$ attains its highest
-value.
+then numerically estimate $\varsigma$ from its definition \eqref{eq-squintability}.
 
@@
 projection distances, estimated by non-linear least squares to obtain
-the squintability measure as Equation @eq-squintability-parametric.
+the squintability measure as Equation (@eq-squintability-parametric).
 
@@
 
-The effect of JSO hyper–parameters (number of jellyfish and the maximum
+The effect of JSO hyper--parameters (number of jellyfish and the maximum
 number of tries) on the success rate is presented in @fig-proportion.
@@
 number of tries, as predictors. Long runtime is a binary variable, where
-greater than 20 seconds is labelled 1. The fit suggests that JSO success
+greater than 20 seconds is labelled 1. The fit suggests that the JSO success
 rate is only affected by squintability, dimension and number of jellies.
-Interestingly, the smoothness of the index is not a problem for the
-JSO - this is promising! Running the optimisation longer and allowing
-more tries also does not affect success rate. An increase in
-squintability of one unit increases the success rate by
-`r round(exp(broom::tidy(mod1)$estimate[3]) * 100, 0)`%. As dimension
+As expected, the JSO success rate is higher with higher squintability and/or more jellyfish, and lower when dimension is higher (higher-dimensional optimization is more difficult).
+Interestingly, the JSO is unaffected by the smoothness of the index function.
+This is consistent with the way random search algorithms jump from value to value without taking local regularity into account, as opposed to gradient-based optimizers for example.
+Allowing
+for more tries also does not affect success rate significantly. An increase in
+squintability of 10\% increases the success rate by
+`r round(exp(broom::tidy(mod1)$estimate[3]) * 10, 0)`%. As dimension
 increases by one the success rate halves. Increasing the number of
-jellies by 10, increases the succes rate by
+jellies by 10, increases the success rate by
 `r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 0)`%.
@@
 bases visited by JSO, along with relevant metadata, as a tibble data
-object (see @RJ-2021-105 for more details on the data object). The bases
+object [see @RJ-2021-105 for more details on the data object]. The bases
 visited for individual jellyfish can then be extracted and viewed using
@@
 and the calculated squintability metric as Equation
-@eq-squintability-parametric. For the non-parametric case, it shows the
-maximum gradient attained (`max_d`), the corresponding projection
+(@eq-squintability-parametric). For the non-parametric case, it shows
+the maximum gradient attained (`max_d`), the corresponding projection
 distance (`max_dist`), and the squintability metric as their products.
@@
 -->
-
+```
 # Acknowledgement
 
-The article is created using Quarto [@Allaire_Quarto_2022] in R [@R].
-The source code for reproducing the work reported in this paper can be
-found at: <https://github.com/huizezhang-sherry/paper-jso>. The
+The article has been created using Quarto [@Allaire_Quarto_2022] in R
+[@R]. The source code for reproducing the work reported in this paper
+can be found at: <https://github.com/huizezhang-sherry/paper-jso>. The
 simulation data produced in Section 5 can be found at
@@
 Nicolas Langrené acknowledges the partial support of the Guangdong
-Provincial Key Laboratory IRADS (2022B1212010006, R0400001-22) and the
-UIC Start-up Research Fund UICR0700041-22.
+Provincial/Zhuhai Key Laboratory IRADS (2022B1212010006) and the UIC
+Start-up Research Fund UICR0700041-22.
 

### 2025-02-05 jso.tex
squintability
Updated squintability section + minor corrections

@@
 \makeatletter
-\newsavebox\pandoc@box
-\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
-  \sbox\pandoc@box{#1}%
-  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
-  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
-  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
-  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
-  \else\usebox{\pandoc@box}%
-  \fi%
-}
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 % Set default figure placement to htbp
+\makeatletter
 \def\fps@figure{htbp}
-\makeatother
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
 \makeatother
@@
 \makeatother
-
-\usepackage{bookmark}
-
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 University, Melbourne, Australia\\ \textsuperscript{3} Department of
-Mathematical Sciences, Guangdong Provincial Key Laboratory of
+Mathematical Sciences, Guangdong Provincial/Zhuhai Key Laboratory of
 Interdisciplinary Research and Application for Data Science,
@@
 \end{keywords}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, frame hidden, sharp corners, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\section{Introduction}\label{introduction}
-
-Projection pursuit (PP) (Kruskal (1969), Friedman and Tukey (1974),
-Huber (1985)) is a dimension reduction technique aimed at identifying
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+Projection pursuit (PP) (Kruskal 1969; Friedman and Tukey 1974; Huber
+1985) is a dimension reduction technique aimed at identifying
 informative linear projections of data. This is useful for exploring
@@
 main features to use for publication. The method involves optimising an
-objective function known as the PP index (e.g. Hall (1989), Cook, Buja,
-and Cabrera (1993), Lee and Cook (2010), Loperfido (2018), Loperfido
-(2020)), which defines the criterion for what constitutes interesting or
-informative projections. Let \(X \in \mathbb{R}^{n\times p}\) be the
-data matrix, \(A \in\mathbb{R}^{p \times d}\) be an orthonormal matrix,
-where \(A\) belongs to the Stiefel manifold
-\(\mathcal{A} = V_d(\mathbb{R}^p)\). The projection \(Y = XA\) is a
-linear transformation that maps data from a \(p\)-dimensional space into
-a \(d\)-dimensional space. The index function
-\(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar function
-that measures an interesting aspect of the projected data, such as
-deviation from normality, presence of clusters, or non-linear structure.
-For a fixed sample of data, PP finds the orthonormal basis \(A\) that
-maximises the index value of the projection, \(Y = XA\):
-
-\begin{equation}\phantomsection\label{eq-optimization}{
+objective function known as the PP index (e.g., Hall 1989; Cook, Buja,
+and Cabrera 1993; Lee and Cook 2010; Loperfido 2018, 2020), which
+defines the criterion for what constitutes interesting or informative
+projections. Let \(X \in \mathbb{R}^{n\times p}\) be the data matrix,
+\(A \in\mathbb{R}^{p \times d}\) be an orthonormal matrix, where \(A\)
+belongs to the Stiefel manifold \(\mathcal{A} = V_d(\mathbb{R}^p)\). The
+projection \(Y = XA\) is a linear transformation that maps data from a
+\(p\)-dimensional space into a \(d\)-dimensional space. The index
+function \(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar
+function that measures an interesting aspect of the projected data, such
+as deviation from normality, presence of clusters, or non-linear
+structure. For a fixed sample of data, PP finds the orthonormal basis
+\(A\) that maximises the index value of the projection, \(Y = XA\):
+
+\begin{equation}\protect\hypertarget{eq-optimization}{}{
 \underset{A \in \mathcal{A}}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I_d
-}\end{equation}
+}\label{eq-optimization}\end{equation}
 
@@
 modes, and thus is not convex. The binned normality index (a simple
-version of a non-normality index as described in Huber (1985)) computed
-on the famous RANDU data, in (c), is noisier and has a very small squint
+version of a non-normality index as described in Huber 1985) computed on
+the famous RANDU data, in (c), is noisier and has a very small squint
 angle. The discreteness cannot be seen unless the optimiser is very
@@
 
-\centering{
-
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{jso_files/figure-pdf/fig-example-functions-1.pdf}
+{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{jso_files/figure-pdf/fig-example-functions-1.pdf}
 
@@
 
-Here, the potential for a Jellyfish Search Optimiser (JSO) (see Chou and
-Truong (2021), and Rajwar, Deep, and Das (2023)) for the PPGT is
-explored. JSO, inspired by the search behaviour of jellyfish in the
-ocean, is a swarm-based metaheuristic designed to solve global
-optimisation problems. Compared to traditional methods, JSO has
-demonstrated stronger search ability and faster convergence, and
-requires fewer tuning parameters. These practicalities make JSO a
-promising candidate for enhancing PP optimisation.
+Here, the potential for a Jellyfish Search Optimiser (JSO) Rajwar, Deep,
+and Das (2023) for the PPGT is explored. JSO, inspired by the search
+behaviour of jellyfish in the ocean, is a swarm-based metaheuristic
+designed to solve global optimisation problems. Compared to traditional
+methods, JSO has demonstrated stronger search ability and faster
+convergence, and requires fewer tuning parameters. These practicalities
+make JSO a promising candidate for enhancing PP optimisation.
 
@@
 PP indexes, metrics are introduced to capture specific properties of the
-index including squintability (based on Tukey and Tukey (1981)'s squint
+index including squintability (based on Tukey and Tukey 1981's squint
 angle) and smoothness. Here, we mathematically define metrics for
@@
 because they are relatively easy to optimise. A goal is to develop PP
-indexes based on scagnostics (L. Wilkinson, Anand, and Grossman (2005),
-Leland Wilkinson and Wills (2008)), but the blockage is their
-optimisation as these tend to be noisy, with potentially small squint
-angles.
+indexes based on scagnostics Leland Wilkinson and Wills (2008), but the
+blockage is their optimisation as these tend to be noisy, with
+potentially small squint angles.
 
@@
 
-\begin{defn}[sobolev space]\label{def:sobolev_space}
+\begin{defn}[Sobolev space]\label{def:sobolev_space}
 The Sobolev space $W^{k,p}(\mathbb{R})$ for $1\leq p\leq \infty$ is the set of all functions $f$ in $L^p(\mathbb{R})$ for which all weak derivatives $f^{(\ell)}$ of order $\ell\leq k$ exist and have a finite $L^p$ norm.
@@
 
-\centering{
-
-\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/matern_simulation_1d.png}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_1d.png}
 
@@
 
-\centering{
-
-\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/matern_simulation_2d.png}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_2d.png}
 
@@
 \begin{defn}[projection distance]\label{def:proj-dist}
-Let $A \in \mathbb{R}^{p \times d}$ is a $d$-dimensional orthonormal matrix, and let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.
+Let $A \in \mathbb{R}^{p \times d}$ be a $d$-dimensional orthonormal matrix, and let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.
 The projection distance between $A$
@@
 
-Squintability can be defined as how the index value \(f(XA)\) changes
-with respect to the projection distance \(r(A, A^*)\), over the course
-of the JSO:
+Squintability can be defined based on how the index value \(f(XA)\)
+changes with respect to the projection distance \(r(A, A^*)\), over the
+course of the JSO. Suppose that the optimization starts with an
+arbitrary candidate matrix \(A_0\) at a distance \(r_0:=r(A_0, A^*)\)
+from the optimal one. A possible way to define squintability is as
+follows:
 
@@
 
-where $4$ is a constant scaling factor, $-\min_r g'(r)$ represents the 
-largest gradient of $-g$ and $\arg \min_{r} g'(r)$ represents the projection distance at which this largest gradient is attained.
-If the set $\mathcal{D}:=\arg \min_{r} g'(r)$ contains more than one value $r$, the largest value $\max(\mathcal{D})$ is used in equation \eqref{eq-squintability}.
-
 \end{defn}
 
-It is expected that these two values should be both high in the case of
-high squintability (fast increase in \(g\) early on), and both low in
-the case of low squintability (any substantial increase in \(g\) happens
-very late, close to the optimal angle). This suggests that their product
-\eqref{eq-squintability} should provide a sensible measure of
-squintability. The multiplicative constant \(4\), which can be deemed
-arbitrary, does not change the interpretation of the squintability
-metric \(\varsigma\); it is here to adjust the range of values of
-\(\varsigma\) and simplify the explicit formula for \(\varsigma\)
-obtained later on.
-
-From the definition, the following proposition can be derived:
-
-\begin{prop}\label{prop:convex-concave}
-Let $g_1$ be a convex, strictly decreasing function and $g_2$ be a concave, strictly decreasing function, both defined on $[0, R]$ where $R>0$, and taking values in $[0, 1]$. Then
-$$\varsigma(g_1) < \varsigma(g_2)$$
-\end{prop}
-\begin{proof}
-Let $r_1 := \underset{r}{\arg \min }\,g_1'(r)$ and $r_2 := \underset{r}{\arg \min }\,g_2'(r)$.
-Since $g_1$ is convex, $g_1''(r) \geq 0 \; \forall r \in [0, R]$, meaning that $g_1'$ is an increasing function, taking its minimum for $r_1=0$.
-As a result, $\varsigma(g_1)=-4 \times g'(0) \times 0 =0$.
-Since $g_2$ is concave, $g_2''(r) \leq 0 \; \forall r \in [0, R]$, meaning that $g_2'$ is a decreasing function, taking its minimum for $r_2=D>0$.
-Since $g_2$ is strictly decreasing, $g_2'(r) < 0 \; \forall r \in [0, R]$.
-In particular $g_2'(R)<0$.
-As a result $\varsigma(g_2)=-4 \times g'(R) \times R >0$.
-This proves that $\varsigma(g_1) < \varsigma(g_2)$.
-\end{proof}
+Remark that the amount by which the index value can improve from the
+starting matrix \(A_0\) to the optimal one \(A^*\) is given by
+\(g(0)-g(r_{0})\). As a result, equation~\eqref{eq-squintability}
+represent the proportion of this maximum improvement which has been
+achieved by the time the distance \(r_0\) to the optimal matrix has been
+reduced by half (\(r_0/2\)).
+
+It is expected that this value should be high in the case of high
+squintability (fast increase in \(g\) early on), and low in the case of
+low squintability (any substantial increase in \(g\) happens very late,
+close to the optimal angle). This suggests that this half-point
+improvement metric should provide a sensible measure of squintability.
 
@@
 amenable to efficient optimization. This observation suggests that the
-extreme values of \(g'\) and the projection distances for which these
-values are attained, are crucial in the mathematical definition of
-squintability.
+half-point index value improvement percentage provides an appropriate
+mathematical definition of squintability, which matches the intuition
+behind this concept, while being amenable to numerical computation.
 
@@
 \(\ell(\theta_{2})=\frac{1}{2}\). Then, define
-\begin{equation}\phantomsection\label{eq-parametric}{
-g(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
-}\end{equation}
+\begin{equation}\protect\hypertarget{eq-parametric}{}{
+g(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(r_0)}{\ell(0)-\ell(r_0)}+\theta_{4}\ ,
+}\label{eq-parametric}\end{equation}
 
 which depends on three additional parameters, \(\theta_1\),
-\(\theta_2\), and \(x_{\max}\), such that \(g(0)=\theta_1\) and
-\(g(x_{\max})=\theta_4\). Under the parametric model
-(\ref{eq-parametric}), the squintability metric \eqref{eq-squintability}
-can be shown to be equal to
-\begin{equation}\phantomsection\label{eq-squintability-parametric}{
-\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
-}\end{equation}
+\(\theta_2\), and \(r_0\), such that \(g(0)=\theta_1\) and
+\(g(r_0)=\theta_4\). Under the parametric model (\ref{eq-parametric}),
+the squintability metric \eqref{eq-squintability} can be shown to be
+equal to
+\begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
+\varsigma=\frac{g\left(r_{0}/2\right)-\theta_{4}}{\theta_{1}-\theta_{4}}=\frac{\ell(r_0/2)-\ell(r_0)}{\ell(0)-\ell(r_0)}\ .
+}\label{eq-squintability-parametric}\end{equation}
 
@@
 nonparametric way, for example by fitting \(g\) using kernel regression,
-then numerically estimate the angle at which \(-g'\) attains its highest
-value.
-
-\section{The jellyfish optimiser}\label{sec-JSO}
+then numerically estimate \(\varsigma\) from its definition
+\eqref{eq-squintability}.
+
+\hypertarget{sec-JSO}{%
+\section{The jellyfish optimiser}\label{sec-JSO}}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, colback=white, bottomrule=.15mm, opacitybacktitle=0.6, coltitle=black, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, breakable, rightrule=.15mm, left=2mm, opacityback=0, leftrule=.75mm]
+\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, opacityback=0, colbacktitle=quarto-callout-note-color!10!white, opacitybacktitle=0.6, colback=white, breakable, left=2mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, bottomtitle=1mm, toprule=.15mm, arc=.35mm, rightrule=.15mm, bottomrule=.15mm, leftrule=.75mm]
 
@@
 
+\hypertarget{sec-sim-deets}{%
 \section{Assessing optimisers: two simulation
-studies}\label{sec-sim-deets}
+studies}\label{sec-sim-deets}}
 
@@
 
-\centering{
-
-\includegraphics[width=4.52in,height=\textheight,keepaspectratio]{figures/success-rate.png}
+{\centering \includegraphics[width=4.52in,height=\textheight]{figures/success-rate.png}
 
@@
 
-\end{figure}%
-
+\end{figure}
+
+\hypertarget{sec-app-2}{%
 \subsection{Factors affecting JSO success rate: JSO hyper-parameters and
-index properties}\label{sec-app-2}
+index properties}\label{sec-app-2}}
 
@@
 
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-plot-data-1.pdf}}
+{\centering \includegraphics{jso_files/figure-pdf/fig-plot-data-1.pdf}
 
@@
 projection distances, estimated by non-linear least squares to obtain
-the squintability measure as Equation \ref{eq-squintability-parametric}.
+the squintability measure as Equation
+(\ref{eq-squintability-parametric}).
 
@@
 
-\centering{
-
-\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/smoothness.png}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/smoothness.png}
 
@@
 
-\centering{
-
-\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/squintability.png}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/squintability.png}
 
@@
 
+\hypertarget{factors-affecting-jso-success-rate-jso-hyper-parameters-and-index-properties}{%
 \subsection{Factors affecting JSO success rate: JSO hyper-parameters and
 index
-properties}\label{factors-affecting-jso-success-rate-jso-hyper-parameters-and-index-properties}
+properties}\label{factors-affecting-jso-success-rate-jso-hyper-parameters-and-index-properties}}
 
@@
 
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-proj-1.pdf}}
+{\centering \includegraphics{jso_files/figure-pdf/fig-proj-1.pdf}
 
@@
 
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-proportion-1.pdf}}
+{\centering \includegraphics{jso_files/figure-pdf/fig-proportion-1.pdf}
 
@@
 
-\end{figure}%
-
+\end{figure}
+
+\hypertarget{tbl-smoothness-squintability}{}
 \begin{table}
-
-\caption{\label{tbl-smoothness-squintability}Parameters estimated from
-the Gaussian process (outputscale \(\eta\), lengthscale \(\ell\),
-smoothness \(\nu\), and nugget \(\sigma\)) and scaled logistic function
-(\(\theta_1\) to \(\theta_4\) and \(\varsigma\)) for the pipe-finding
-and sine-wave finding problems. The columns \(\nu\) and \(\varsigma\)
-represent the smoothness and squintability measures respectively.}
-
-\centering{
+\caption{\label{tbl-smoothness-squintability}Parameters estimated from the Gaussian process (outputscale \(\eta\),
+lengthscale \(\ell\), smoothness \(\nu\), and nugget \(\sigma\)) and
+scaled logistic function (\(\theta_1\) to \(\theta_4\) and
+\(\varsigma\)) for the pipe-finding and sine-wave finding problems. The
+columns \(\nu\) and \(\varsigma\) represent the smoothness and
+squintability measures respectively. }\tabularnewline
 
@@
 variable, where greater than 20 seconds is labelled 1. The fit suggests
-that JSO success rate is only affected by squintability, dimension and
-number of jellies. Interestingly, the smoothness of the index is not a
-problem for the JSO - this is promising! Running the optimisation longer
-and allowing more tries also does not affect success rate. An increase
-in squintability of one unit increases the success rate by 124\%. As
-dimension increases by one the success rate halves. Increasing the
-number of jellies by 10, increases the succes rate by 6\%.
-
+that the JSO success rate is only affected by squintability, dimension
+and number of jellies. As expected, the JSO success rate is higher with
+higher squintability and/or more jellyfish, and lower when dimension is
+higher (higher-dimensional optimization is more difficult).
+Interestingly, the JSO is unaffected by the smoothness of the index
+function. This is consistent with the way random search algorithms jump
+from value to value without taking local regularity into account, as
+opposed to gradient-based optimizers for example. Allowing for more
+tries also does not affect success rate significantly. An increase in
+squintability of 10\% increases the success rate by 12\%. As dimension
+increases by one the success rate halves. Increasing the number of
+jellies by 10, increases the success rate by 6\%.
+
+\hypertarget{tbl-mod-output}{}
 \begingroup
@@
 visited by JSO, along with relevant metadata, as a tibble data object
-(see Zhang et al. (2021) for more details on the data object). The bases
+(see Zhang et al. 2021 for more details on the data object). The bases
 visited for individual jellyfish can then be extracted and viewed using
@@
 
+\hypertarget{computing-the-index-properties-for-your-new-index}{%
 \subsection{Computing the index properties for your new
-index}\label{computing-the-index-properties-for-your-new-index}
+index}\label{computing-the-index-properties-for-your-new-index}}
 
@@
 function (\(\theta_1\) to \(\theta_4\)) and the calculated squintability
-metric as Equation \ref{eq-squintability-parametric}. For the
+metric as Equation (\ref{eq-squintability-parametric}). For the
 non-parametric case, it shows the maximum gradient attained
@@
 
-\section{Acknowledgement}\label{acknowledgement}
-
-The article is created using Quarto (Allaire et al. 2022) in R (R Core
-Team 2023). The source code for reproducing the work reported in this
-paper can be found at:
+\hypertarget{acknowledgement}{%
+\section{Acknowledgement}\label{acknowledgement}}
+
+The article has been created using Quarto (Allaire et al. 2022) in R (R
+Core Team 2023). The source code for reproducing the work reported in
+this paper can be found at:
 \url{https://github.com/huizezhang-sherry/paper-jso}. The simulation
@@
 Nicolas Langrené acknowledges the partial support of the Guangdong
-Provincial Key Laboratory IRADS (2022B1212010006, R0400001-22) and the
-UIC Start-up Research Fund UICR0700041-22.
-
-\section*{Supplementary materials}\label{supplementary-materials}
+Provincial/Zhuhai Key Laboratory IRADS (2022B1212010006) and the UIC
+Start-up Research Fund UICR0700041-22.
+
+\hypertarget{supplementary-materials}{%
+\section*{Supplementary materials}\label{supplementary-materials}}
 \addcontentsline{toc}{section}{Supplementary materials}

### 2025-02-06 supplementary/index-def.qmd
update appendix, more to come

@@
 ---
-title: 'Appendix to " New Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices"'
+title: 'Appendix to "New Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices"'
 author: H. Sherry Zhang, Dianne Cook, Nicolas Langrené, Jessica Wai Yin Leung
@@
 
-Given high-dimensional data $X \in \mathbf{R}^{n\times p}$ and the index function $f(\cdot)$, projection pursuit finds the orthonormal projection basis $A \in \mathbf{R}^{p \times d}$ by solving the following optimisation problem:
+```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+library(tidyverse)
+library(ferrn)
+library(tourr)
+library(tidygraph)
+library(ggraph)
+library(cassowaryr)
+library(patchwork)
+```
+
+Let $Y = XA$ denote the projection of a p-dimensional dataset, $X \in \mathbb{R}^{n\times p}$, onto d-dimensional space, $Y \in \mathbb{R}^{n \times d}$, where the projection is defined by an orthonormal matrix $A \in \mathbb{R}^{p \times d}$.
+Projection pursuit aims to find the matrix $A$ that maximizes an index function $f(\cdot)$, which measures interesting features of the projection, such as multi-modality, linear or non-linear relationship:
 
@@
 
-This appendix presents definition of all indexes $f(\cdot)$ used in the paper.
-These indexes are defined in 2D cases ($d = 2$), where $Y \in \mathbf{R}^{n \times 2} = (y_1, y_2)$ represents the projected data.
+This appendix defines all indexes $f(\cdot)$ used in the paper, namely holes, MIC, TIC, dcor, loess, skinny, splines, and stringy).
+Since all the indexes are defined for 2D projections ($d = 2$), the projected data can also be written as $Y = (y_1, y_2)$.
 
-## The `holes` index
+## The holes index
 
-The `holes` index [@cook1993projection] is a smooth index for detecting the presence of multi-modality in the projection. The index is defined as:
+The holes index [@cook1993projection] defines as follows detects the presence of multi-modality in the projection:
 
-$$I_{\text{holes}}(A) = \frac{1-\frac{1}{n} \sum^n_{i = 1} \exp(-\frac{1}{2y_iy_i^\prime})}{1-\exp(-\frac{d}{2})}$$
+$$I_{\text{holes}} = \frac{1-\frac{1}{n} \sum^n_{i = 1} \exp(-\frac{1}{2y_iy_i^\prime})}{1-\exp(-\frac{d}{2})}$$
 
-## The `dcor2d` index
+## The dcor index
 
-The `dcor2d` index [@Grimm2016] is a smooth index that measures the distance correlation between the two projection axes.
+The dcor index [@Grimm2016], sometimes called dcor2D, measures the **d**istance **cor**relation between the two projection axes $(y_1, y_2)$.
 The index uses pair-wise distance to define column sum, row sum, and overall sum and then calculates distance correlation from distance variance and covariance.
@@
 
-   -   distance correlation: $$I_{\text{dcor}}(A)  = \text{dCor} (a_{ij}, b_{ij}) = \frac{\text{dCov}(a_{ij}, b_{ij})}{\sqrt{\text{dVar}(a_{ij}) \text{dVar}(b_{ij})}}$$
+   -   distance correlation: $$I_{\text{dcor}} = \text{dCor} (a_{ij}, b_{ij}) = \frac{\text{dCov}(a_{ij}, b_{ij})}{\sqrt{\text{dVar}(a_{ij}) \text{dVar}(b_{ij})}}$$
 
@@
 
-$$I_{MIC}(A) = \max_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \max_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$
+$$I_{MIC} = \max_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \max_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$
 
@@
 
-$$I_{TIC}(A) = \sum_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \sum_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$
+$$I_{TIC} = \sum_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \sum_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$
 
@@
 
-# The `stringy` index
+# The skinny and stringy index
 
-The `stringy` [@scag; @WW08] index is a non-smooth index based on scagnostics.
-It measures the proportion of vertices with two edges in the minimum spanning tree (MST) of the projection to detect whether the projection forms a straight line.
-The index is calculated as
+Both skinny and stringy indexes are non-smooth indexes based on scagnostics [@WW08; @scag].
+Scagnostics are defined based on graph concepts, specifically, the alpha hull for the skinny index and the minimum spanning tree for the stringy index.
+The skinny index is defined based on the alpha hull ($\text{AH}$) as:
 
-$$I_{stringy}(A) = \frac{\text{number of vertices with 2 edges}}{\text{number of total vertices with more than one edge}} $$
+$$I_{\text{skinny}} = 1 - \frac{\sqrt{4 \pi \text{area(AH)}}}{\text{perimeter(AH)}}$$
+
+The stringy index is defined based on the minimal spanning tree ($T$).
+We adopt the definition in @scag based on the diameter and length of the tree.
+The diameter refers to the longest shortest path through $T$, and the length refers to the total number of edges in $T$.
+
+$$I_{\text{stringy}} = \frac{\text{diameter}(T)}{\text{length}(T)}$$
+
+An alternative definition of the stringy index is based on number of vertices, given by $\frac{|V^{(2)}|}{ |V| - |V^{(1)}|}$, where $|V^{(2)}|$ represents the number of degree-2 vertices and $|V| - |V^{(1)}|$ is the overall number of vertices minus the number of degree-1 vertices.
+However, we prefer the original definition based on the diameter and length of the minimal spanning tree, as it is more robust to noise, especially as the number of observations increases.
+In @fig-stringy, the stringy index defined based on the number of vertices (row 1) fails differentiate the non-linear sine wave from the non-pattern (left to right), while the diameter/length-based definition (row 2) does.
+
+```{r}
+count_vertices <- function(proj){
+  x <- scree(proj$V1, proj$V2)
+  a <- cassowaryr:::gen_mst(x$del, x$weights) 
+  vertices <- igraph::degree(a)
+  vertices |> table() |> as_tibble()
+}
+
+find_stringy <- function(proj_df,  longest_shortest = FALSE){
+  x <- scree(proj_df$V1, proj_df$V2)
+  b <- cassowaryr:::gen_mst(x$del, x$weights)
+
+  if (longest_shortest){
+    diameter <- length(igraph::get_diameter(b))
+    length <- length(b) 
+    tibble(diameter = diameter, length = length, stringy = round(diameter / length, 4))  
+  } else{
+    vertex_counts <- igraph::degree(b)
+    tibble(degree_one = sum(vertex_counts == 1), 
+           degree_two = sum(vertex_counts == 2), 
+           stringy = round(sum(vertex_counts == 2) / sum(vertex_counts > 1), 3))
+  }
+}
+
+plot_mst <- function(proj, size = 1, longest_shortest = FALSE){
+  x <- scree(proj$V1, proj$V2)
+  a <- cassowaryr:::gen_mst(x$del, x$weights)
+
+  edge_df <- as_tbl_graph(a) |> activate(edges) |> as_tibble()
+  graph_df <- tbl_graph(nodes = proj, edges = edge_df, directed = FALSE)
+
+  if (longest_shortest){
+    longest_path <- igraph::get_diameter(a)
+    df <- tibble(from = as.vector(longest_path),
+                 to = c(as.vector(longest_path)[-1], 0)) |> 
+      filter(to != 0) |> 
+      mutate(fromto = paste0(from, to), tofrom = paste0(to, from))
+    graph_df <- graph_df |> 
+      activate(edges) |> 
+      mutate(fromto = paste0(from, to)) |> 
+      mutate(main_path = ifelse(fromto %in% c(df$fromto, df$tofrom), TRUE, FALSE))
+
+  } else{
+    vertex_counts <- igraph::degree(a)
+    graph_df <- graph_df |> 
+      activate(nodes) |> 
+      mutate(size = as.factor(vertex_counts),
+             edge_col = ifelse(vertex_counts > 1, 3, 1),
+             edge_col = as.factor(ifelse(size == 2, 2, edge_col)))
+  }
+
+  res <- ggraph(graph_df, x = V1, y = V2) +
+    geom_node_point(size = size) + 
+    theme_bw() + 
+    theme(aspect.ratio = 1,
+          axis.ticks = ggplot2::element_blank(),
+          axis.text = ggplot2::element_blank(),
+          panel.grid = ggplot2::element_blank())
+
+  if (longest_shortest){
+    res <- res +
+      geom_edge_link(aes(color = main_path)) + 
+      scale_edge_color_brewer(palette = "Dark2", name= "Diameter")
+  } else{
+    res <- res + 
+      geom_edge_link(color = "black") + 
+      geom_node_point(aes(color = edge_col), size=size) + 
+      scale_color_brewer(palette = "Dark2", name = "Vertices")
+  }
+  return(res)
+
+}
+```
+
+```{r}
+#| label: fig-stringy
+#| fig.width: 9
+#| fig.height: 5 
+#| fig.cap: "Projections and the stringy index values calculated based on the number of vertices definition (row 1) and the diameter/length definition (row 2) for a sine wave pattern with 1000 observations.
+The color represents the number of vertices in row 1 and highlighs the longest minimal path (the points forming the diameter) in row 2.
+The noise around the sine wave creates local structures in the minimal spanning tree, resulting in a high number of degree-2 and higher vertices when calculated based on the number of vertices.
+This causes stringy index based on the number of vertices fails to differentiate the non-linear sine wave from the non-pattern, while the diameter/length definition, which captures the dominant main structure, is able to differentiate the sine wave form the non-pattern."
+
+mat0 <- matrix(c(           
+  0.42501543,  0.14630027,
+  0.39159945,  0.46379132,
+ -0.49791838,  0.45474161,
+ -0.09191581, -0.43749847,
+  0.63834704, -0.04640681,
+ -0.04640681,  0.60261547), nrow = 6, byrow = TRUE)
+mat1 <- matrix(c(
+  0.119619098, -0.12987621,
+  0.005504289, -0.03605812,
+  0.016274619,  0.03172812,
+ -0.020440544, -0.06033105,
+  0.992435619,  0.00705958,
+  0.007059580,  0.98850171), nrow = 6, byrow = TRUE)
+
+mat2 <- matrix(c(
+  0.026769505,  0.017326941,
+  0.031236712,  0.026519087,
+ -0.001670954, -0.014253910,
+  0.014841987,  0.056109302,
+  0.999041254, -0.001076086,
+ -0.001076086,  0.997819588), nrow = 6, byrow = TRUE)
+
+mat3 <- matrix(c(rep(0, 4), 0, 0.01, 0.01, 0, 0.99, 0, 0, 0.99), 
+               nrow = 6, byrow = TRUE) |> orthonormalise()
+
+df0 <- as_tibble(as.matrix(sine1000) %*% mat0) 
+df1 <- as_tibble(as.matrix(sine1000) %*% mat1) 
+df2 <- as_tibble(as.matrix(sine1000) %*% mat2) 
+df3 <- as_tibble(as.matrix(sine1000) %*% mat3) 
+p1 <- plot_mst(df0, size = 0.2) + 
+  labs(title = "Index = 0.77") +
+  theme(plot.title = ggtext::element_markdown())
+p2 <- plot_mst(df1, size = 0.2) + labs(title = "Index = 0.76")
+p3 <- plot_mst(df2, size = 0.2) + 
+  labs(title = "Index = 0.719") +
+  theme(plot.title = ggtext::element_markdown())
+p4 <- plot_mst(df3, size = 0.2) + 
+  labs(title = "Index = 0.771") +
+  theme(plot.title = ggtext::element_markdown())
+
+vert_df <- map_dfr(list(df0, df1, df2, df3), count_vertices, .id = "id") |> 
+  group_by(id) |> 
+  summarise(`degree 2 \n vertices` = n[2], `degree 2+ \n vertices` = sum(n[-1]),
+            stringy = round(n[2] / (sum(n) - n[1]), 3))
+
+p5 <- plot_mst(df0, size = 0.2, longest_shortest = TRUE) + labs(title = "Index = 0.176")
+p6 <- plot_mst(df1, size = 0.2, longest_shortest = TRUE) + labs(title = "Index = 0.284")
+p7 <- plot_mst(df2, size = 0.2, longest_shortest = TRUE) + labs(title = "Index = 0.474")
+p8 <- plot_mst(df3, size = 0.2, longest_shortest = TRUE) + labs(title = "Index = 0.705")
+
+stringy_df <- tibble(id = c(1,2,3, 4)) |> 
+  bind_cols(bind_rows(find_stringy(df0, longest_shortest = TRUE), 
+                      find_stringy(df1, longest_shortest = TRUE),
+                      find_stringy(df2, longest_shortest = TRUE),
+                      find_stringy(df3, longest_shortest = TRUE)))
+
+(p1 | p2 | p3 | p4 | gt::gt(vert_df)) /( p5 | p6 | p7 | p8 | gt::gt(stringy_df)) + 
+  plot_layout(guides = "collect")  & 
+  theme(legend.position='bottom', axis.text = element_blank(), axis.title = element_blank())
+
+```
 

### 2025-02-06 supplementary/index-def.qmd
update index definition

@@
 
-# The `MIC` and `TIC` index
+# The MIC and TIC index
 
-Both `MIC` and `TIC` are information-based indexes derived from mutual information.
-The indexes are defined based on a partition, $g(k, l)$, of the space $(y_1, y_2)$ into $k \times l$ rectangles, For example, $g(2, 3)$ divides the data space into 2 rectangles in the $y_2$ direction and 3 rectangles in the $y_1$ direction.
-Let $G$ denotes all the possible partition.
+Maximal Information Coefficient (MIC) and Total Information Coefficient (TIC) [@reshef2016measuring] are information-based indexes derived from mutual information.
+These indexes calculate the mutual information after partitioning the space $(y_1, y_2)$ into $k \times l$ rectangles, with a constraint on the grid size to limit complexity: $k \times l < B(n) = n^{\alpha}$, where $n$ is the number of observations in the data.
+A commonly used value for $\alpha$ in the previous studies is 0.3 to balance flexiblity and overfitting.
 
-MIC finds the **maximum** mutual information over G where $k$ and $l$ are bounded by the grid size: $k \times l < B(n) = n^{\alpha}$. $alpha = 0.3$ is used in our simulation.
+The MIC index finds the **maximum** mutual information over possible grids configeration that satisfy the constraint, and then normalized by the grid size: $\log(\min(k^*, l^*))$:
 
-$$I_{MIC} = \max_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \max_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$
+$$I_{MIC} = \frac{I(y_1, y_2 | k^*, l^*)}{\log(\min(k^*, l^*))}$$
 
-where $k^*$ and $l^*$ are the number of rectangles in the optimal partition $g$.
+where $(k^*, l^*)$ is the grid configuration that maximizes the mutual information $I(y_1, y_2 | k, l)$ under the constraint $k \times l < B(n) = n^{0.3}$.
 
-TIC calculates the **sum** of mutual information over G
+TIC calculates the **sum** of the normalized mutual information for all grid configurations under the size constraint:
 
-$$I_{TIC} = \sum_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \sum_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$
+$$I_{TIC} = \sum_{k \times l < n^{0.3}} \frac{I(y_1, y_2 | k, l)}{\log(\min(k, l))}$$
 
-# The `loess` and `splines` index
+# The loess and splines index
 
-The `loess` and `splines` indexes detect non-linear structure in the projection, as captured by their respective model.
-These indexes are computed by regressing both axes against each other, and find the maximum variance from the residual that can be explained by the model.
-Let $e^{\text{model}}_{y_1 \sim y_2}$ denote the residual from the loess/splines model $y_1 \sim y_2$, and $e^{\text{model}}_{y_2 \sim y_1}$ denote the residual from the loess/spliens model $y_2 \sim y_1$.
-The two indexes are calculated as
+The loess and splines indexes detect non-linear structure in the projection by capturing the variance explained by the loess/ spline model.
+To account for the orientation of the projection, the model is run for both $y_1 \sim y_2$ and $y_2 \sim y_1$ and the larger of the two is used as the index value.
+Let $e^{\text{model}}_{y_1 \sim y_2}$ and $e^{\text{model}}_{y_2 \sim y_1}$ denote the residual from the loess/splines model $y_1 \sim y_2$, and $y_2 \sim y_1$.
+The two indexes are calculated as:
 
- $$I_{\text{loess}}(A) = \max \left(1 - \frac{var(e^{\text{loess}}_{y_1 \sim y_2})}{var(y_1)}, 1 - \frac{var(e^{\text{loess}}_{y_2 \sim y_1})}{var(y_2)}\right)$$
+ $$I_{\text{loess}} = \max \left(1 - \frac{var(e^{\text{loess}}_{y_1 \sim y_2})}{var(y_1)}, 1 - \frac{var(e^{\text{loess}}_{y_2 \sim y_1})}{var(y_2)}\right)$$
 
-$$I_{\text{spline}}(A) = \max \left(1 - \frac{var(e^{\text{spline}}_{y_1 \sim y_2})}{var(y_1)}, 1 - \frac{var(e^{\text{spline}}_{y_2 \sim y_1})}{var(y_2)}\right)$$
+$$I_{\text{spline}} = \max \left(1 - \frac{var(e^{\text{spline}}_{y_1 \sim y_2})}{var(y_1)}, 1 - \frac{var(e^{\text{spline}}_{y_2 \sim y_1})}{var(y_2)}\right)$$
 

### 2025-02-09 jso.tex
update title

@@
 \hypersetup{
-  pdftitle={New Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices},
+  pdftitle={Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices},
   pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
@@
 
-\title{New Metrics for Assessing Projection Pursuit Indexes, and Guiding
-Optimisation Choices}
+\title{Squintability and Other Metrics for Assessing Projection Pursuit
+Indexes, and Guiding Optimisation Choices}
 \author{H. Sherry Zhang$\textsuperscript{1}$, Dianne
@@
 \end{keywords}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, frame hidden, sharp corners, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, boxrule=0pt, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, opacityback=0, colbacktitle=quarto-callout-note-color!10!white, opacitybacktitle=0.6, colback=white, breakable, left=2mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, bottomtitle=1mm, toprule=.15mm, arc=.35mm, rightrule=.15mm, bottomrule=.15mm, leftrule=.75mm]
+\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, titlerule=0mm, arc=.35mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, coltitle=black, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, rightrule=.15mm, colback=white, leftrule=.75mm, toprule=.15mm, left=2mm]
 

### 2025-02-11 jso.qmd
update title
edits on sec 5

@@
 ---
-title: New Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices
+title: Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices
 author:
@@
 
-This section describes the details of two simulation studies: one
+This section describes the details of two simulation studies. The first simulation
 compares the JSO performance with an existing optimiser, Creeping Random
 Search (CRS) [@RJ-2021-105; @laa_using_2020] to explore JSO's behaviour
-under different data dimensions. The second simulation first examines
-the JSO performance under different hyper-parameters combinations
-(number of jellyfish and the maximum number of tries), and then studies
-the effect of index properties (smoothness and squintability), along
-with JSO hyper-parameters, and data dimension, on the JSO performance
-through a logistic regression.
+under different data dimensions. The second simulation examines various factors affecting the performance of JSO.
+The JSO performance is evaluated under different hyper-parameters combinations
+(number of jellyfish and the maximum number of tries). The effect of index properties (smoothness and squintability) along
+with JSO hyper-parameters and data dimension variations is captured and studied using logistic regression.
 
@@
 
-The JSO performance is investigated both in comparison to the existing
-optimizer, CRS, in a pipe-finding problem using the holes index in the
-6, 8, 10, and 12-dimensional space. Fifty simulations are conducted in
-four data dimensions ($d = 6, 8, 10, 12$) with each optimiser. JSO uses
+The JSO performance is compared to the existing optimizer, CRS, in a pipe-finding problem using the holes index in different dimensional spaces.
+Fifty simulations are conducted in four data dimensions ($d = 6, 8, 10, 12$) with each optimiser.
+JSO uses
 100 jellyfish with a maximum of 100 tries, while the CRS allows a
@@
 
-To examine JSO's performance across various hyper-parameter
+To examine the performance of JSO across various hyper-parameter
 combinations, the pipe-finding problem using the holes index is repeated
@@
 
-To assess JSO's performance across various PP problems, two different
-data shapes, pipe and a sine wave, are investigated in 6D and 8D spaces
+In addition, to assess the performance of JSO across various PP problems, two different
+data shapes, pipe and a sine wave, are studied in 4D, 6D and 8D spaces
 using six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`,

### 2025-02-11 jso.qmd
update sec 5

@@
 In addition, to assess the performance of JSO across various PP problems, two different
-data shapes, pipe and a sine wave, are studied in 4D, 6D and 8D spaces
+data shapes, pipe and a sine wave, are considered in various dimensional spaces
 using six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`,
@@
 `r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and
-`r nrow(filter(sim_df, index != "holes"))` on the sine-wave data. Again,
-JSO is run 50 times for each scenario to calculate the success rate for
-each PP problem.
+`r nrow(filter(sim_df, index != "holes"))` on the sine-wave data.
+A list of combinations of the data shape, dimensions and PP indexes considered are provided in @tbl-smoothness-squintability.
+Again, JSO is run 50 times for each scenario to calculate the success rate for each hyper-parameter combination.
 

### 2025-02-11 jso.tex
update sec 5

@@
 \end{keywords}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, boxrule=0pt, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 \(\ell(\theta_{2})=\frac{1}{2}\). Then, define
-\begin{equation}\protect\hypertarget{eq-parametric}{}{
+\begin{equation}\phantomsection\label{eq-parametric}{
 g(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(r_0)}{\ell(0)-\ell(r_0)}+\theta_{4}\ ,
-}\label{eq-parametric}\end{equation}
+}\end{equation}
 
@@
 equal to
-\begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
+\begin{equation}\phantomsection\label{eq-squintability-parametric}{
 \varsigma=\frac{g\left(r_{0}/2\right)-\theta_{4}}{\theta_{1}-\theta_{4}}=\frac{\ell(r_0/2)-\ell(r_0)}{\ell(0)-\ell(r_0)}\ .
-}\label{eq-squintability-parametric}\end{equation}
+}\end{equation}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, titlerule=0mm, arc=.35mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, coltitle=black, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, rightrule=.15mm, colback=white, leftrule=.75mm, toprule=.15mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, breakable, opacitybacktitle=0.6, rightrule=.15mm, left=2mm, toprule=.15mm, titlerule=0mm, toptitle=1mm, bottomtitle=1mm, arc=.35mm, leftrule=.75mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, colback=white]
 
@@
 
-\hypertarget{sec-sim-deets}{%
 \section{Assessing optimisers: two simulation
-studies}\label{sec-sim-deets}}
-
-This section describes the details of two simulation studies: one
-compares the JSO performance with an existing optimiser, Creeping Random
-Search (CRS) (Zhang et al. 2021; Laa and Cook 2020) to explore JSO's
-behaviour under different data dimensions. The second simulation first
-examines the JSO performance under different hyper-parameters
-combinations (number of jellyfish and the maximum number of tries), and
-then studies the effect of index properties (smoothness and
-squintability), along with JSO hyper-parameters, and data dimension, on
-the JSO performance through a logistic regression.
+studies}\label{sec-sim-deets}
+
+This section describes the details of two simulation studies. The first
+simulation compares the JSO performance with an existing optimiser,
+Creeping Random Search (CRS) (Zhang et al. 2021; Laa and Cook 2020) to
+explore JSO's behaviour under different data dimensions. The second
+simulation examines various factors affecting the performance of JSO.
+The JSO performance is evaluated under different hyper-parameters
+combinations (number of jellyfish and the maximum number of tries). The
+effect of index properties (smoothness and squintability) along with JSO
+hyper-parameters and data dimension variations is captured and studied
+using logistic regression.
 
@@
 
-\hypertarget{sec-app-1}{%
-\subsection{Performance of JSO relative to CRS}\label{sec-app-1}}
-
-The JSO performance is investigated both in comparison to the existing
-optimizer, CRS, in a pipe-finding problem using the holes index in the
-6, 8, 10, and 12-dimensional space. Fifty simulations are conducted in
-four data dimensions (\(d = 6, 8, 10, 12\)) with each optimiser. JSO
-uses 100 jellyfish with a maximum of 100 tries, while the CRS allows a
-maximum of 1000 samples at each iteration before the algorithm
-terminates. The different numbers account for the multiple paths of JSO
-to enable fairer comparison with CRS. The results of the simulation are
-collected using the data structure proposed in Zhang et al. (2021) for
-assessing JSO, where the design parameters are stored along with index
-value, projection basis, random seed, and computation time.
+\subsection{Performance of JSO relative to CRS}\label{sec-app-1}
+
+The JSO performance is compared to the existing optimizer, CRS, in a
+pipe-finding problem using the holes index in different dimensional
+spaces. Fifty simulations are conducted in four data dimensions
+(\(d = 6, 8, 10, 12\)) with each optimiser. JSO uses 100 jellyfish with
+a maximum of 100 tries, while the CRS allows a maximum of 1000 samples
+at each iteration before the algorithm terminates. The different numbers
+account for the multiple paths of JSO to enable fairer comparison with
+CRS. The results of the simulation are collected using the data
+structure proposed in Zhang et al. (2021) for assessing JSO, where the
+design parameters are stored along with index value, projection basis,
+random seed, and computation time.
 
@@
 
-\end{figure}
-
-\hypertarget{sec-app-2}{%
+\end{figure}%
+
 \subsection{Factors affecting JSO success rate: JSO hyper-parameters and
-index properties}\label{sec-app-2}}
-
-To examine JSO's performance across various hyper-parameter
+index properties}\label{sec-app-2}
+
+To examine the performance of JSO across various hyper-parameter
 combinations, the pipe-finding problem using the holes index is repeated
@@
 
-To assess JSO's performance across various PP problems, two different
-data shapes, pipe and a sine wave, are investigated in 6D and 8D spaces
-using six different PP indexes: \texttt{dcor2d\_2}, \texttt{loess2d},
-\texttt{MIC}, \texttt{TIC}, \texttt{spline}, and \texttt{stringy}, under
-varied JSO hyper-parameters. This results in a total of 76 scenarios,
-comprising of 30 computed on the pipe data and 46 on the sine-wave data.
-Again, JSO is run 50 times for each scenario to calculate the success
-rate for each PP problem.
+In addition, to assess the performance of JSO across various PP
+problems, two different data shapes, pipe and a sine wave, are
+considered in various dimensional spaces using six different PP indexes:
+\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
+\texttt{spline}, and \texttt{stringy}, under varied JSO
+hyper-parameters. This results in a total of 76 scenarios, comprising of
+30 computed on the pipe data and 46 on the sine-wave data. A list of
+combinations of the data shape, dimensions and PP indexes considered are
+provided in Table~\ref{tbl-smoothness-squintability}. Again, JSO is run
+50 times for each scenario to calculate the success rate for each
+hyper-parameter combination.
 
@@
 
-{\centering \includegraphics{jso_files/figure-pdf/fig-plot-data-1.pdf}
+\centering{
+
+\includegraphics{jso_files/figure-pdf/fig-plot-data-1.pdf}
 
@@
 
-\hypertarget{factors-affecting-jso-success-rate-jso-hyper-parameters-and-index-properties}{%
 \subsection{Factors affecting JSO success rate: JSO hyper-parameters and
 index
-properties}\label{factors-affecting-jso-success-rate-jso-hyper-parameters-and-index-properties}}
+properties}\label{factors-affecting-jso-success-rate-jso-hyper-parameters-and-index-properties}
 
@@
 
-\end{figure}
-
-\hypertarget{tbl-smoothness-squintability}{}
+\end{figure}%
+
 \begin{table}
-\caption{\label{tbl-smoothness-squintability}Parameters estimated from the Gaussian process (outputscale \(\eta\),
-lengthscale \(\ell\), smoothness \(\nu\), and nugget \(\sigma\)) and
-scaled logistic function (\(\theta_1\) to \(\theta_4\) and
-\(\varsigma\)) for the pipe-finding and sine-wave finding problems. The
-columns \(\nu\) and \(\varsigma\) represent the smoothness and
-squintability measures respectively. }\tabularnewline
+
+\caption{\label{tbl-smoothness-squintability}Parameters estimated from
+the Gaussian process (outputscale \(\eta\), lengthscale \(\ell\),
+smoothness \(\nu\), and nugget \(\sigma\)) and scaled logistic function
+(\(\theta_1\) to \(\theta_4\) and \(\varsigma\)) for the pipe-finding
+and sine-wave finding problems. The columns \(\nu\) and \(\varsigma\)
+represent the smoothness and squintability measures respectively.}
+
+\centering{
 
@@
 
-\hypertarget{tbl-mod-output}{}
-\begingroup
+\begin{table}
+
+\caption{\label{tbl-mod-output}Jellyfish success rate relative to index
+properties and jellyfish hyper-parameters. This is the summary from a
+logistic regression fit to smoothness, squintability, dimension, running
+time, number of jellyfish and maximum number of tries. Interestingly,
+squintability and dimension strongly affect jellyfish optimisation
+success. The number of jellies marginally affects success, but index
+smoothness, running longer and increasing the number of tries do not.}
+
+\centering{
+
 \fontsize{12.0pt}{14.4pt}\selectfont
-\setlength{\LTpost}{0mm}
-\begin{longtable}{lcc}
-\caption{\label{tbl-mod-output}Jellyfish success rate relative to index properties and jellyfish
-hyper-parameters. This is the summary from a logistic regression fit to
-smoothness, squintability, dimension, running time, number of jellyfish
-and maximum number of tries. Interestingly, squintability and dimension
-strongly affect jellyfish optimisation success. The number of jellies
-marginally affects success, but index smoothness, running longer and
-increasing the number of tries do not. }\tabularnewline
-
+\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lcc}
 \toprule
@@
 \bottomrule
-\end{longtable}
+\end{tabular*}
 \begin{minipage}{\linewidth}
@@
 \end{minipage}
-\endgroup
-
-\hypertarget{sec-discussion}{%
-\section{Practicalities}\label{sec-discussion}}
+
+}
+
+\end{table}%
+
+\section{Practicalities}\label{sec-discussion}
 
@@
 
-\hypertarget{computing-the-index-properties-for-your-new-index}{%
 \subsection{Computing the index properties for your new
-index}\label{computing-the-index-properties-for-your-new-index}}
+index}\label{computing-the-index-properties-for-your-new-index}
 

### 2025-02-11 code/01-sim-pipe-681012d-holes.R
update sim_pipe_run_best to include 4D cases for Figure 9

@@ -48,7 +48,8 @@ save(sim_pipe, file = "data-raw/sim_pipe.rda")
 ################################################################################
 ################################################################################
 # additional smaller data
-sim_pipe_run_best <- sim_pipe |>
+load(here::here("data-raw/sim_pipe_4d.rda"))
+sim_pipe_run_best <- bind_rows(sim_pipe_4d, sim_pipe) |>
   group_by(id, n_jellies, max_tries, d) |>
   summarise(I_max = max(index_val))
 save(sim_pipe_run_best, file = "data/sim_pipe_run_best.rda")

### 2025-02-12 jso.qmd
edits in sec 6

@@
 case. As the number of jellyfish and maximum tries increase, the success
-rate also increases. For simpler problems (6 dimensions), small
-parameter values (20 jellyfish and a maximum of 50 tries) can already
-achieve a high success rate. However, larger parameter values (i.e. 100
-jellyfish and a maximum of 100 tries) are necessary for
-higher-dimensional problems (8, 10, and 12 dimensions). Increasing both
-parameters enhances the performance of JSO, but it also extends the
-computational time required for the optimisation, which can be
+rate also increases.
+For problems with relatively lower dimensional search spaces (4 and 6 dimensions), small parameter values (20 jellyfish and a maximum of 50 tries) are sufficient for space exploration and thus can already achieve a high success rate.
+Higher parameter values (i.e.
+100 jellyfish and a maximum of 100 tries) enhances exploration of the search space, which is particularly beneficial in higher-dimensional problems (8, 10, and 12 dimensions) where a more complex search is necessary.
+However, in lower-dimensional problems, the benefit may be less pronounced, as the space is alredy relatively easy to navigate.
+While increasing both parameters enhances the performance of JSO, it also extends the computational time required for the optimisation, which can be
 computationally intensive when evaluating the index function (such as
@@
 model using the proportion of success as the response and smoothness,
-squintability, dimension, running time, number of jellyfish and maximum
-number of tries, as predictors. Long runtime is a binary variable, where
-greater than 20 seconds is labelled 1. The fit suggests that the JSO success
+squintability, dimension, number of jellyfish and maximum
+number of tries, as predictors. The fit suggests that the JSO success
 rate is only affected by squintability, dimension and number of jellies.
@@
 Allowing
-for more tries also does not affect success rate significantly. An increase in
-squintability of 10\% increases the success rate by
-`r round(exp(broom::tidy(mod1)$estimate[3]) * 10, 0)`%. As dimension
-increases by one the success rate halves. Increasing the number of
+for more tries also does not affect success rate significantly. A unit increase in squintability increases the success rate by
+`r round(100 * (exp(broom::tidy(mod1)$estimate[3])-1), 0)`%. As dimension
+increases by one the success rate almost halves. Increasing the number of
 jellies by 10, increases the success rate by
-`r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 0)`%.
+`r round(100 * (exp(broom::tidy(mod1)$estimate[5])-1), 0)`%.
 
@@
 #| tbl-cap: "Jellyfish success rate relative to index properties and jellyfish hyper-parameters.
-This is the summary from a logistic regression fit to smoothness, squintability, dimension, running time, number of jellyfish and maximum number of tries.
+This is the summary from a logistic regression fit to smoothness, squintability, dimension, number of jellyfish and maximum number of tries.
 Interestingly, squintability and dimension strongly affect jellyfish optimisation success.
-The number of jellies marginally affects success, but index smoothness, running longer and increasing the number of tries do not."
+The number of jellies marginally affects success, but index smoothness, and increasing the number of tries do not."
 #| fig-pos: t
@@
                                     d = "Dimension", 
-                                    long_time = "Long runtime", 
+                                    #long_time = "Long runtime", 
                                     n_jellies = "Number of jellyfish", 

### 2025-02-12 jso.tex
edits in sec 6

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, breakable, opacitybacktitle=0.6, rightrule=.15mm, left=2mm, toprule=.15mm, titlerule=0mm, toptitle=1mm, bottomtitle=1mm, arc=.35mm, leftrule=.75mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, rightrule=.15mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, leftrule=.75mm, toptitle=1mm, colback=white, titlerule=0mm, breakable, opacityback=0, opacitybacktitle=0.6, bottomtitle=1mm, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, coltitle=black]
 
@@
 bootstrap replicates for each case. As the number of jellyfish and
-maximum tries increase, the success rate also increases. For simpler
-problems (6 dimensions), small parameter values (20 jellyfish and a
-maximum of 50 tries) can already achieve a high success rate. However,
-larger parameter values (i.e.~100 jellyfish and a maximum of 100 tries)
-are necessary for higher-dimensional problems (8, 10, and 12
-dimensions). Increasing both parameters enhances the performance of JSO,
-but it also extends the computational time required for the
-optimisation, which can be computationally intensive when evaluating the
-index function (such as scagnostic indexes) multiple times across
-numerous iterations.
+maximum tries increase, the success rate also increases. For problems
+with relatively lower dimensional search spaces (4 and 6 dimensions),
+small parameter values (20 jellyfish and a maximum of 50 tries) are
+sufficient for space exploration and thus can already achieve a high
+success rate. Higher parameter values (i.e.~100 jellyfish and a maximum
+of 100 tries) enhances exploration of the search space, which is
+particularly beneficial in higher-dimensional problems (8, 10, and 12
+dimensions) where a more complex search is necessary. However, in
+lower-dimensional problems, the benefit may be less pronounced, as the
+space is alredy relatively easy to navigate. While increasing both
+parameters enhances the performance of JSO, it also extends the
+computational time required for the optimisation, which can be
+computationally intensive when evaluating the index function (such as
+scagnostic indexes) multiple times across numerous iterations.
 
@@
 regression model using the proportion of success as the response and
-smoothness, squintability, dimension, running time, number of jellyfish
-and maximum number of tries, as predictors. Long runtime is a binary
-variable, where greater than 20 seconds is labelled 1. The fit suggests
-that the JSO success rate is only affected by squintability, dimension
-and number of jellies. As expected, the JSO success rate is higher with
-higher squintability and/or more jellyfish, and lower when dimension is
-higher (higher-dimensional optimization is more difficult).
-Interestingly, the JSO is unaffected by the smoothness of the index
-function. This is consistent with the way random search algorithms jump
-from value to value without taking local regularity into account, as
-opposed to gradient-based optimizers for example. Allowing for more
-tries also does not affect success rate significantly. An increase in
-squintability of 10\% increases the success rate by 12\%. As dimension
-increases by one the success rate halves. Increasing the number of
-jellies by 10, increases the success rate by 6\%.
+smoothness, squintability, dimension, number of jellyfish and maximum
+number of tries, as predictors. The fit suggests that the JSO success
+rate is only affected by squintability, dimension and number of jellies.
+As expected, the JSO success rate is higher with higher squintability
+and/or more jellyfish, and lower when dimension is higher
+(higher-dimensional optimization is more difficult). Interestingly, the
+JSO is unaffected by the smoothness of the index function. This is
+consistent with the way random search algorithms jump from value to
+value without taking local regularity into account, as opposed to
+gradient-based optimizers for example. Allowing for more tries also does
+not affect success rate significantly. A unit increase in squintability
+increases the success rate by 24\%. As dimension increases by one the
+success rate almost halves. Increasing the number of jellies by 10,
+increases the success rate by 32\%.
 
@@
 properties and jellyfish hyper-parameters. This is the summary from a
-logistic regression fit to smoothness, squintability, dimension, running
-time, number of jellyfish and maximum number of tries. Interestingly,
-squintability and dimension strongly affect jellyfish optimisation
-success. The number of jellies marginally affects success, but index
-smoothness, running longer and increasing the number of tries do not.}
+logistic regression fit to smoothness, squintability, dimension, number
+of jellyfish and maximum number of tries. Interestingly, squintability
+and dimension strongly affect jellyfish optimisation success. The number
+of jellies marginally affects success, but index smoothness, and
+increasing the number of tries do not.}
 

### 2025-02-17 jso.qmd
edited sec 6, moved Fig 5 to appendix

@@
 library(colorspace)
+library(broom)
 load(here::here("data/sim_pipe_run_best.rda"))
@@
 `r nrow(filter(sim_df, index != "holes"))` on the sine-wave data.
-A list of combinations of the data shape, dimensions and PP indexes considered are provided in @tbl-smoothness-squintability.
+A list of combinations of the data shape, dimensions and PP indexes that are considered in this study is provided in @tbl-smoothness-squintability.
 Again, JSO is run 50 times for each scenario to calculate the success rate for each hyper-parameter combination.
-
-```{r}
-#| label: fig-plot-data
-#| fig-width: 10
-#| fig-height: 5
-#| fig-cap: "The sine and pipe structures in the example data, along with index values for these projections."
-data("sine1000")
-sine1000 <- as_tibble(sine1000)
-dp1 <- ggplot(sine1000, aes(x=V5, y=V6)) + geom_point()
-data("pipe1000")
-pipe1000 <- as_tibble(pipe1000)
-dp2 <- ggplot(pipe1000, aes(x=V5, y=V6)) + geom_point()
-dp1 + dp2 + plot_layout(ncol=2)
-```
 
@@
 
-The effect of JSO hyper--parameters (number of jellyfish and the maximum
+The effect of JSO hyper-parameters (number of jellyfish and the maximum
 number of tries) on the success rate is presented in @fig-proportion.
@@
 100 jellyfish and a maximum of 100 tries) enhances exploration of the search space, which is particularly beneficial in higher-dimensional problems (8, 10, and 12 dimensions) where a more complex search is necessary.
-However, in lower-dimensional problems, the benefit may be less pronounced, as the space is alredy relatively easy to navigate.
+However, in lower-dimensional problems, the benefit may be less pronounced, as the space is already relatively easy to navigate.
 While increasing both parameters enhances the performance of JSO, it also extends the computational time required for the optimisation, which can be
@@
 increases by one the success rate almost halves. Increasing the number of
-jellies by 10, increases the success rate by
+jellies by 10 increases the success rate by
 `r round(100 * (exp(broom::tidy(mod1)$estimate[5])-1), 0)`%.
@@
 
+```{r}
+#| tbl-pos: H
+#| label: tbl-joint-output
+library(kableExtra)
+mod2 <- glm(proportion ~ squintability + d + n_jellies + max_tries, data = sim_df2, family = binomial)
+mod3 <- glm(proportion ~ smoothness + d + n_jellies + max_tries, data = sim_df2, family = binomial)
+summary_table <- bind_rows(
+  glance(mod2) %>% mutate(Model = "Squintability"),
+  glance(mod3) %>% mutate(Model = "Smoothness")
+) %>%
+  select(Model, null.deviance, df.null, logLik, df.residual, AIC, BIC) %>%
+  rename(
+    `Null Deviance` = null.deviance,
+    `DF Null` = df.null,
+    `Log-Likelihood` = logLik,
+    `DF Residual` = df.residual
+  ) %>%
+  mutate(across(where(is.numeric), ~ round(.x, 2)))
+summary_table %>%
+  kable(caption = "Model summaries for squintability and smoothness separately") %>%
+  kable_styling(font_size = 10) %>%
+  column_spec(1, bold = TRUE, width = "5em")
+```
+
+Squintability and smoothness are likely dependent, making it unnecessary to include both as independent factors in the model.
+Intuitively, a highly smooth search space exhibits gradual changes, making it easier to perceive the overall structure from a distance, thereby increasing the squint angle.
+@tbl-joint-output compare the model fits when squintability and smoothness are included separately.
+Given that the null deviance remains the same across models, the model with squintability yields a smaller residual deviance and lower AIC and BIC values, indicating a better fit.
+This suggests that squintability captures the relevant variation more effectively than smoothness.
+
+This study intends to serve as an illustration that the proposed measure effectively captures the key factors influencing the success rate of PP.
+The interactions between the two variables were not examined in this study due to the limited number of data points, nor was the linearity assumption of the relationship with success rate explicitly tested.
+
 # Practicalities {#sec-discussion}

### 2025-02-17 jso.tex
edited sec 6, moved Fig 5 to appendix

@@
 
-\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, rightrule=.15mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, leftrule=.75mm, toptitle=1mm, colback=white, titlerule=0mm, breakable, opacityback=0, opacitybacktitle=0.6, bottomtitle=1mm, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm, toprule=.15mm, colback=white, leftrule=.75mm, arc=.35mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, rightrule=.15mm, toptitle=1mm, bottomrule=.15mm, breakable, titlerule=0mm, colframe=quarto-callout-note-color-frame]
 
@@
 30 computed on the pipe data and 46 on the sine-wave data. A list of
-combinations of the data shape, dimensions and PP indexes considered are
-provided in Table~\ref{tbl-smoothness-squintability}. Again, JSO is run
-50 times for each scenario to calculate the success rate for each
-hyper-parameter combination.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{jso_files/figure-pdf/fig-plot-data-1.pdf}
-
-}
-
-\caption{\label{fig-plot-data}The sine and pipe structures in the
-example data, along with index values for these projections.}
-
-\end{figure}%
+combinations of the data shape, dimensions and PP indexes that are
+considered in this study is provided in
+Table~\ref{tbl-smoothness-squintability}. Again, JSO is run 50 times for
+each scenario to calculate the success rate for each hyper-parameter
+combination.
 
@@
 
-The effect of JSO hyper--parameters (number of jellyfish and the maximum
+The effect of JSO hyper-parameters (number of jellyfish and the maximum
 number of tries) on the success rate is presented in
@@
 lower-dimensional problems, the benefit may be less pronounced, as the
-space is alredy relatively easy to navigate. While increasing both
+space is already relatively easy to navigate. While increasing both
 parameters enhances the performance of JSO, it also extends the
@@
 increases the success rate by 24\%. As dimension increases by one the
-success rate almost halves. Increasing the number of jellies by 10,
+success rate almost halves. Increasing the number of jellies by 10
 increases the success rate by 32\%.
@@
 
+\begingroup\fontsize{10}{12}\selectfont
+
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{5em}rrrrrr}
+
+\caption{\label{tbl-joint-output}Model summaries for squintability and
+smoothness separately}
+
+\tabularnewline
+
+\\
+\toprule
+Model & Null Deviance & DF Null & Log-Likelihood & DF Residual & AIC & BIC\\
+\midrule
+\textbf{Squintability} & 49.29 & 75 & -26.44 & 71 & 62.87 & 74.53\\
+\textbf{Smoothness} & 49.29 & 75 & -33.23 & 71 & 76.47 & 88.12\\
+\bottomrule
+
+\end{longtable}
+
+\endgroup{}
+
+Squintability and smoothness are likely dependent, making it unnecessary
+to include both as independent factors in the model. Intuitively, a
+highly smooth search space exhibits gradual changes, making it easier to
+perceive the overall structure from a distance, thereby increasing the
+squint angle. Table~\ref{tbl-joint-output} compare the model fits when
+squintability and smoothness are included separately. Given that the
+null deviance remains the same across models, the model with
+squintability yields a smaller residual deviance and lower AIC and BIC
+values, indicating a better fit. This suggests that squintability
+captures the relevant variation more effectively than smoothness.
+
+This study intends to serve as an illustration that the proposed measure
+effectively captures the key factors influencing the success rate of PP.
+The interactions between the two variables were not examined in this
+study due to the limited number of data points, nor was the linearity
+assumption of the relationship with success rate explicitly tested.
+
 \section{Practicalities}\label{sec-discussion}

### 2025-02-17 supplementary/index-def.qmd
edited sec 6, moved Fig 5 to appendix

@@
 
+# The sine and pipe structures in the example data
+
+```{r}
+#| label: fig-plot-data
+#| fig-width: 10
+#| fig-height: 5
+#| fig-cap: "The sine and pipe structures in the example data, along with index values for these projections."
+data("sine1000")
+sine1000 <- as_tibble(sine1000)
+dp1 <- ggplot(sine1000, aes(x=V5, y=V6)) + geom_point()
+data("pipe1000")
+pipe1000 <- as_tibble(pipe1000)
+dp2 <- ggplot(pipe1000, aes(x=V5, y=V6)) + geom_point()
+dp1 + dp2 + plot_layout(ncol=2)
+```
+
 # Reference

### 2025-02-18 jso.qmd
edits in sec5

@@
 
-In addition, to assess the performance of JSO across various PP problems, two different
-data shapes, pipe and a sine wave, are considered in various dimensional spaces
-using six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`,
+In addition, to assess the performance of JSO across various PP problems, two different data shapes, pipe and a sine wave, are considered in various dimensional spaces using six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`,
 `spline`, and `stringy`, under varied JSO hyper-parameters. This results
@@
 `r nrow(filter(sim_df, index != "holes"))` on the sine-wave data.
+This study does not consider dimensions lower than four, as exploring the search space in such cases is relatively straightforward.
+Similarly, extremely high-dimensional cases with complex data structures, where no method can reliably and consistenly uncover the underlying structure due to the curse of dimensionality, are also excluded, as they fall outside the scope of this study.
 A list of combinations of the data shape, dimensions and PP indexes that are considered in this study is provided in @tbl-smoothness-squintability.
@@
 ```{r}
+#| tbl-pos: H
 #| label: tbl-mod-output
@@
 
-Squintability and smoothness are likely dependent, making it unnecessary to include both as independent factors in the model.
+The simulation result suggests that squintability and smoothness are likely dependent.
 Intuitively, a highly smooth search space exhibits gradual changes, making it easier to perceive the overall structure from a distance, thereby increasing the squint angle.

### 2025-02-18 jso.tex
edits in sec5

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm, toprule=.15mm, colback=white, leftrule=.75mm, arc=.35mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, rightrule=.15mm, toptitle=1mm, bottomrule=.15mm, breakable, titlerule=0mm, colframe=quarto-callout-note-color-frame]
+\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, titlerule=0mm, opacityback=0, coltitle=black, colback=white, colframe=quarto-callout-note-color-frame, opacitybacktitle=0.6, arc=.35mm, breakable, toptitle=1mm, left=2mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, bottomrule=.15mm, toprule=.15mm]
 
@@
 hyper-parameters. This results in a total of 76 scenarios, comprising of
-30 computed on the pipe data and 46 on the sine-wave data. A list of
-combinations of the data shape, dimensions and PP indexes that are
-considered in this study is provided in
-Table~\ref{tbl-smoothness-squintability}. Again, JSO is run 50 times for
-each scenario to calculate the success rate for each hyper-parameter
+30 computed on the pipe data and 46 on the sine-wave data. This study
+does not consider dimensions lower than four, as exploring the search
+space in such cases is relatively straightforward. Similarly, extremely
+high-dimensional cases with complex data structures, where no method can
+reliably and consistenly uncover the underlying structure due to the
+curse of dimensionality, are also excluded, as they fall outside the
+scope of this study. A list of combinations of the data shape,
+dimensions and PP indexes that are considered in this study is provided
+in Table~\ref{tbl-smoothness-squintability}. Again, JSO is run 50 times
+for each scenario to calculate the success rate for each hyper-parameter
 combination.
@@
 
-\begin{table}
+\begin{table}[H]
 
@@
 
-Squintability and smoothness are likely dependent, making it unnecessary
-to include both as independent factors in the model. Intuitively, a
-highly smooth search space exhibits gradual changes, making it easier to
-perceive the overall structure from a distance, thereby increasing the
-squint angle. Table~\ref{tbl-joint-output} compare the model fits when
-squintability and smoothness are included separately. Given that the
-null deviance remains the same across models, the model with
-squintability yields a smaller residual deviance and lower AIC and BIC
-values, indicating a better fit. This suggests that squintability
-captures the relevant variation more effectively than smoothness.
+The simulation result suggests that squintability and smoothness are
+likely dependent. Intuitively, a highly smooth search space exhibits
+gradual changes, making it easier to perceive the overall structure from
+a distance, thereby increasing the squint angle.
+Table~\ref{tbl-joint-output} compare the model fits when squintability
+and smoothness are included separately. Given that the null deviance
+remains the same across models, the model with squintability yields a
+smaller residual deviance and lower AIC and BIC values, indicating a
+better fit. This suggests that squintability captures the relevant
+variation more effectively than smoothness.
 

### 2025-02-18 jso.tex
update supplementary material

@@
 
-\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, titlerule=0mm, opacityback=0, coltitle=black, colback=white, colframe=quarto-callout-note-color-frame, opacitybacktitle=0.6, arc=.35mm, breakable, toptitle=1mm, left=2mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, bottomrule=.15mm, toprule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, toprule=.15mm, coltitle=black, rightrule=.15mm, colback=white, toptitle=1mm, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, leftrule=.75mm, breakable, opacitybacktitle=0.6, titlerule=0mm]
 
@@
 \texttt{spline}, and \texttt{stringy}, under varied JSO
-hyper-parameters. This results in a total of 76 scenarios, comprising of
-30 computed on the pipe data and 46 on the sine-wave data. This study
-does not consider dimensions lower than four, as exploring the search
-space in such cases is relatively straightforward. Similarly, extremely
-high-dimensional cases with complex data structures, where no method can
-reliably and consistenly uncover the underlying structure due to the
-curse of dimensionality, are also excluded, as they fall outside the
-scope of this study. A list of combinations of the data shape,
-dimensions and PP indexes that are considered in this study is provided
-in Table~\ref{tbl-smoothness-squintability}. Again, JSO is run 50 times
-for each scenario to calculate the success rate for each hyper-parameter
+hyper-parameters. An example of holes and pipes data with the syntax to
+generate them are provided in the supplementary material. This results
+in a total of 76 scenarios, comprising of 30 computed on the pipe data
+and 46 on the sine-wave data. This study does not consider dimensions
+lower than four, as exploring the search space in such cases is
+relatively straightforward. Similarly, extremely high-dimensional cases
+with complex data structures, where no method can reliably and
+consistenly uncover the underlying structure due to the curse of
+dimensionality, are also excluded, as they fall outside the scope of
+this study. A list of combinations of the data shape, dimensions and PP
+indexes that are considered in this study is provided in
+Table~\ref{tbl-smoothness-squintability}. Again, JSO is run 50 times for
+each scenario to calculate the success rate for each hyper-parameter
 combination.

### 2025-02-18 supplementary/index-def.qmd
update supplementary material

--- 2025-02-06 supplementary/index-def.qmd
@@
 library(patchwork)
+load(here::here("data/sim_pipe_run_best.rda"))
+load(here::here("data/sim_sine_6d_spline_head.rda"))
+load(here::here("data/sim_sine_6d_spline_best.rda"))
+load(here::here("data/sim_sine_6d_spline_projdist.rda"))
+load(here::here("data/pipe_better.rda"))
+load(here::here("data/pipe_jellyfish.rda"))
+load(here::here("data/smoothness.rda"))
+load(here::here("data/squintability.rda"))
+load(here::here("data/sq_basis_dist_idx.rda"))
+load(here::here("data/sim_df.rda"))
 ```
@@
 
+# The sine and pipe structures in the example data
+
+@fig-plot-data gives an example of the sine and pipe structures data, along with index values for these projections. The syntax to generate these figures are provided below.
+
+```{r, echo = TRUE}
+#| label: fig-plot-data
+#| fig-width: 10
+#| fig-height: 5
+#| fig-cap: "The sine and pipe structures in the example data, along with index values for these projections."
+data("sine1000")
+sine1000 <- as_tibble(sine1000)
+dp1 <- ggplot(sine1000, aes(x=V5, y=V6)) + geom_point()
+data("pipe1000")
+pipe1000 <- as_tibble(pipe1000)
+dp2 <- ggplot(pipe1000, aes(x=V5, y=V6)) + geom_point()
+dp1 + dp2 + plot_layout(ncol=2)
+```
+
 # Reference

### 2025-02-18 jso.qmd
update supplementary material
edit sec6

@@
 In addition, to assess the performance of JSO across various PP problems, two different data shapes, pipe and a sine wave, are considered in various dimensional spaces using six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`,
-`spline`, and `stringy`, under varied JSO hyper-parameters. This results
+`spline`, and `stringy`, under varied JSO hyper-parameters. An example of holes and pipes data with the syntax to generate them are provided in the supplementary material. This results
 in a total of `r nrow(sim_df)` scenarios, comprising of
@@
 pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
-for interpretation, 2) creating a new binary variable `long_time` to
-indicate cases with an average run time over 20 seconds, and 3)
+for interpretation, 2) <!-- creating a new binary variable `long_time` to
+indicate cases with an average run time over 20 seconds, and 3)-->
 re-coding the success rate for the `stringy` index as 0, because none of
@@
 This suggests that squintability captures the relevant variation more effectively than smoothness.
+While it may seem counterintuitive at first that smoothness is not a significant factor, this outcome is reasonable in this context, as the swarm-based method used in this study, JSO, is designed to perform well even in non-smooth search spaces by leveraging population-based exploration.
 

### 2025-02-18 jso.tex
edit sec6

@@
 
-\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, toprule=.15mm, coltitle=black, rightrule=.15mm, colback=white, toptitle=1mm, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, leftrule=.75mm, breakable, opacitybacktitle=0.6, titlerule=0mm]
+\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, left=2mm, toprule=.15mm, opacityback=0, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, opacitybacktitle=0.6, bottomrule=.15mm, toptitle=1mm, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, titlerule=0mm, coltitle=black, arc=.35mm, rightrule=.15mm]
 
@@
 pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
-for interpretation, 2) creating a new binary variable
-\texttt{long\_time} to indicate cases with an average run time over 20
-seconds, and 3) re-coding the success rate for the \texttt{stringy}
-index as 0, because none of the 50 simulations correctly identified the
-sine-wave shape.
+for interpretation, 2) re-coding the success rate for the
+\texttt{stringy} index as 0, because none of the 50 simulations
+correctly identified the sine-wave shape.
 
@@
 better fit. This suggests that squintability captures the relevant
-variation more effectively than smoothness.
+variation more effectively than smoothness. While it may seem
+counterintuitive at first that smoothness is not a significant factor,
+this outcome is reasonable in this context, as the swarm-based method
+used in this study, JSO, is designed to perform well even in non-smooth
+search spaces by leveraging population-based exploration.
 

### 2025-03-01 jso.qmd
shortened and tightened abstarct

@@
 Optimisation of some PP indexes can be non-trivial, if they are non-smooth functions, or the optimum has a small "squint angle", detectable only from close proximity.
-To address these challenges, this study investigates the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
-The performance of JSO for visualising data is evaluated across various hyper-parameter settings and compared with existing optimisers.
-Additionally, methods for calculating the smoothness and squintability properties of the PP index are proposed.
-They are used to assess the optimiser performance in the presence of PP index complexities.
-A simulation study illustrates the use of these performance metrics to compare the JSO with existing optimisation methods available for the guided tour.
-The JSO algorithm has been implemented in the R package, `tourr`, and functions to calculate smoothness and squintability are available in the `ferrn` package.
+Here, measures for calculating the smoothness and squintability properties of the PP index are defined.
+These are used to investigate the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
+The performance of JSO for visualising data is evaluated across various hyper-parameter settings and compared with existing optimisers for the guided tour.
+The JSO algorithm has been implemented in the R package, `tourr`, and functions to calculate smoothness and squintability measures are implemented in the `ferrn` package.
 keywords: 

### 2025-03-01 jso.tex
shortened and tightened abstarct

@@
 optimum has a small ``squint angle'', detectable only from close
-proximity. To address these challenges, this study investigates the
-performance of a recently introduced swarm-based algorithm, Jellyfish
-Search Optimiser (JSO), for optimising PP indexes. The performance of
-JSO for visualising data is evaluated across various hyper-parameter
-settings and compared with existing optimisers. Additionally, methods
-for calculating the smoothness and squintability properties of the PP
-index are proposed. They are used to assess the optimiser performance in
-the presence of PP index complexities. A simulation study illustrates
-the use of these performance metrics to compare the JSO with existing
-optimisation methods available for the guided tour. The JSO algorithm
-has been implemented in the R package, \texttt{tourr}, and functions to
-calculate smoothness and squintability are available in the
-\texttt{ferrn} package.
+proximity. Here, measures for calculating the smoothness and
+squintability properties of the PP index are defined. These are used to
+investigate the performance of a recently introduced swarm-based
+algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
+The performance of JSO for visualising data is evaluated across various
+hyper-parameter settings and compared with existing optimisers for the
+guided tour. The JSO algorithm has been implemented in the R package,
+\texttt{tourr}, and functions to calculate smoothness and squintability
+measures are implemented in the \texttt{ferrn} package.
 \end{abstract}
@@
 
-\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, left=2mm, toprule=.15mm, opacityback=0, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, opacitybacktitle=0.6, bottomrule=.15mm, toptitle=1mm, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, titlerule=0mm, coltitle=black, arc=.35mm, rightrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, left=2mm, toptitle=1mm, colback=white, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, opacityback=0, titlerule=0mm, bottomtitle=1mm, rightrule=.15mm, arc=.35mm, toprule=.15mm, opacitybacktitle=0.6, coltitle=black]
 
@@
 
-\begin{table}[H]
+\setlength{\LTpost}{0mm}
+
+\begin{longtable}{lcc}
 
@@
 
-\centering{
-
-\fontsize{12.0pt}{14.4pt}\selectfont
-\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lcc}
+\tabularnewline
+
 \toprule
@@
 Smoothness & 1.00 (0.81 to 1.20) & 0.961 \\ 
-Squintability & 1.24 (1.05 to 1.58) & {\bfseries 0.033} \\ 
-Dimension & 0.57 (0.33 to 0.87) & {\bfseries 0.022} \\ 
-Number of jellyfish & 1.32 (1.06 to 1.71) & {\bfseries 0.022} \\ 
+Squintability & 1.24 (1.05 to 1.58) & \textbf{0.033} \\ 
+Dimension & 0.57 (0.33 to 0.87) & \textbf{0.022} \\ 
+Number of jellyfish & 1.32 (1.06 to 1.71) & \textbf{0.022} \\ 
 Maximum number of tries & 1.06 (0.85 to 1.35) & 0.603 \\ 
 \bottomrule
-\end{tabular*}
+
+\end{longtable}
+
 \begin{minipage}{\linewidth}
@@
 \end{minipage}
-
-}
-
-\end{table}%
 

### 2025-03-01 jso.qmd
interaction model fitted

@@
 library(kableExtra)
-mod2 <- glm(proportion ~ squintability + d + n_jellies + max_tries, data = sim_df2, family = binomial)
-mod3 <- glm(proportion ~ smoothness + d + n_jellies + max_tries, data = sim_df2, family = binomial)
+mod2 <- glm(proportion ~ squintability + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
+mod3 <- glm(proportion ~ smoothness + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
+mod4 <- glm(proportion ~ squintability*d + n_jellies, data = sim_df2, family = quasibinomial)
+
 summary_table <- bind_rows(
-  glance(mod2) %>% mutate(Model = "Squintability"),
-  glance(mod3) %>% mutate(Model = "Smoothness")
+  glance(mod1) %>% mutate(Model = "Full"),
+  #glance(mod2) %>% mutate(Model = "Squintability"),
+  #glance(mod3) %>% mutate(Model = "Smoothness")
+  glance(mod4) %>% mutate(Model = "Interaction")
 ) %>%
-  select(Model, null.deviance, df.null, logLik, df.residual, AIC, BIC) %>%
+  select(Model, deviance, df.residual, null.deviance, df.null, #logLik, 
+         ) %>% #, AIC, BIC) %>%
   rename(
@@
     `DF Null` = df.null,
-    `Log-Likelihood` = logLik,
-    `DF Residual` = df.residual
+    `Deviance` = deviance,
+    `DF Residual` = df.residual#,
+    #`Log-Likelihood` = logLik,
   ) %>%
@@
 summary_table %>%
-  kable(caption = "Model summaries for squintability and smoothness separately") %>%
+  kable(caption = "Fit statistics for models: full, individual squintability and smoothness.
+The full model (and the model containing squintability but not smoothness) explains more than 50% of the variability in optimisation success.") %>%
   kable_styling(font_size = 10) %>%
-  column_spec(1, bold = TRUE, width = "5em")
+  row_spec(0, bold = TRUE) #, width = "5em")
 ```

### 2025-03-01 jso.tex
interaction model fitted

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, left=2mm, toptitle=1mm, colback=white, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, opacityback=0, titlerule=0mm, bottomtitle=1mm, rightrule=.15mm, arc=.35mm, toprule=.15mm, opacitybacktitle=0.6, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, colback=white, arc=.35mm, coltitle=black, leftrule=.75mm, opacityback=0, colframe=quarto-callout-note-color-frame, breakable, toptitle=1mm, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, left=2mm, bottomrule=.15mm, rightrule=.15mm]
 
@@
 
-\begin{longtable}[t]{>{\raggedright\arraybackslash}p{5em}rrrrrr}
-
-\caption{\label{tbl-joint-output}Model summaries for squintability and
-smoothness separately}
+\begin{longtable}[t]{lrrrr}
+
+\caption{\label{tbl-joint-output}Fit statistics for models: full,
+individual squintability and smoothness. The full model (and the model
+containing squintability but not smoothness) explains more than 50\% of
+the variability in optimisation success.}
 
@@
 \toprule
-Model & Null Deviance & DF Null & Log-Likelihood & DF Residual & AIC & BIC\\
+\textbf{Model} & \textbf{Deviance} & \textbf{DF Residual} & \textbf{Null Deviance} & \textbf{DF Null}\\
 \midrule
-\textbf{Squintability} & 49.29 & 75 & -26.44 & 71 & 62.87 & 74.53\\
-\textbf{Smoothness} & 49.29 & 75 & -33.23 & 71 & 76.47 & 88.12\\
+Full & 20.69 & 70 & 49.29 & 75\\
+Interaction & 18.00 & 71 & 49.29 & 75\\
 \bottomrule

### 2025-03-01 jso.qmd
model fits all in same chunk

--- 2025-03-01 jso.qmd
@@
 mod1 <- glm(proportion ~ smoothness + squintability + d + n_jellies + max_tries, data = sim_df2, family = binomial) 
+mod2 <- glm(proportion ~ squintability + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
+mod3 <- glm(proportion ~ smoothness + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
+mod4 <- glm(proportion ~ squintability*d + n_jellies, data = sim_df2, family = quasibinomial)
 ```
@@
 library(kableExtra)
-mod2 <- glm(proportion ~ squintability + d + n_jellies + max_tries, data = sim_df2, family = binomial)
-mod3 <- glm(proportion ~ smoothness + d + n_jellies + max_tries, data = sim_df2, family = binomial)
 summary_table <- bind_rows(
-  glance(mod2) %>% mutate(Model = "Squintability"),
-  glance(mod3) %>% mutate(Model = "Smoothness")
-) %>%
-  select(Model, null.deviance, df.null, logLik, df.residual, AIC, BIC) %>%
+  glance(mod1) %>% mutate(Model = "Full"),
+  #glance(mod2) %>% mutate(Model = "Squintability"),
+  #glance(mod3) %>% mutate(Model = "Smoothness")
+  glance(mod4) %>% mutate(Model = "Interaction")
+) |>
+  select(Model, deviance, df.residual, null.deviance, df.null, #logLik, 
+         ) |> #, AIC, BIC) %>%
   rename(
@@
     `DF Null` = df.null,
-    `Log-Likelihood` = logLik,
-    `DF Residual` = df.residual
-  ) %>%
+    `Deviance` = deviance,
+    `DF Residual` = df.residual#,
+    #`Log-Likelihood` = logLik,
+  ) |>
   mutate(across(where(is.numeric), ~ round(.x, 2)))
-summary_table %>%
-  kable(caption = "Model summaries for squintability and smoothness separately") %>%
-  kable_styling(font_size = 10) %>%
-  column_spec(1, bold = TRUE, width = "5em")
+summary_table |>
+  kable(caption = "Fit statistics for models: full, individual squintability and smoothness.
+The full model (and the model containing squintability but not smoothness) explains more than 50% of the variability in optimisation success.") |>
+  kable_styling(font_size = 10) |>
+  row_spec(0, bold = TRUE) #, width = "5em")
 ```

### 2025-03-01 jso.tex
model fits all in same chunk

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, colback=white, arc=.35mm, coltitle=black, leftrule=.75mm, opacityback=0, colframe=quarto-callout-note-color-frame, breakable, toptitle=1mm, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, left=2mm, bottomrule=.15mm, rightrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, toprule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, colframe=quarto-callout-note-color-frame, left=2mm, bottomtitle=1mm, bottomrule=.15mm, toptitle=1mm, opacityback=0, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacitybacktitle=0.6, leftrule=.75mm, arc=.35mm, coltitle=black]
 

### 2025-03-01 jso.qmd
caption on fig 1

--- 2025-03-01 jso.qmd
@@
 #| fig-cap: "Examples of PP indexes with large (top row) and small (bottom row) squint angles, shown with a Huber plot, and histogram of the projected data corresponding to the optimal projection.
-A Huber plot shows the PP index values for all 1D data projections in polar coordinates."
+A Huber plot shows the PP index values for all 1D data projections in polar coordinates.
+The skewness index on the trimodal data is also smoother than the binned normality index on RANDU."
 #proj <- animate_xy(flea[,1:6], guided_tour(lda_pp(flea$species)))
@@
   mutate(n_jellies = n_jellies/10, max_tries = max_tries/10) |> 
-  rename(proportion = P_J)
-
-mod1 <- glm(proportion ~ smoothness + squintability + d + n_jellies + max_tries, data = sim_df2, family = binomial) 
+  rename(proportion = P_J) #|>
+  #filter(d < 10)
+
+mod1 <- glm(proportion ~ smoothness + squintability + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial) 
+mod2 <- glm(proportion ~ squintability + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
+mod3 <- glm(proportion ~ smoothness + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
+mod4 <- glm(proportion ~ squintability*d + n_jellies, data = sim_df2, family = quasibinomial)
+mod5 <- glm(proportion ~ smoothness*d + squintability*d + n_jellies, data = sim_df2, family = quasibinomial)
 ```
@@
 library(kableExtra)
-mod2 <- glm(proportion ~ squintability + d + n_jellies + max_tries, data = sim_df2, family = binomial)
-mod3 <- glm(proportion ~ smoothness + d + n_jellies + max_tries, data = sim_df2, family = binomial)
-summary_table <- bind_rows(
-  glance(mod2) %>% mutate(Model = "Squintability"),
-  glance(mod3) %>% mutate(Model = "Smoothness")
-) %>%
-  select(Model, null.deviance, df.null, logLik, df.residual, AIC, BIC) %>%
+summary_table <- #bind_rows(
+  glance(mod1) |> #%>% mutate(Model = "Full"),
+  #glance(mod2) %>% mutate(Model = "Squintability"),
+  #glance(mod3) %>% mutate(Model = "Smoothness")
+#  glance(mod4) %>% mutate(Model = "Interaction")
+#) |>
+  select(#Model, 
+    deviance, df.residual, null.deviance, df.null, #logLik, 
+         ) |> #, AIC, BIC) %>%
   rename(
@@
     `DF Null` = df.null,
-    `Log-Likelihood` = logLik,
-    `DF Residual` = df.residual
-  ) %>%
+    `Deviance` = deviance,
+    `DF Residual` = df.residual#,
+    #`Log-Likelihood` = logLik,
+  ) |>
   mutate(across(where(is.numeric), ~ round(.x, 2)))
-summary_table %>%
-  kable(caption = "Model summaries for squintability and smoothness separately") %>%
-  kable_styling(font_size = 10) %>%
-  column_spec(1, bold = TRUE, width = "5em")
+summary_table |>
+  kable(caption = "Fit statistics for the model.") |>
+  kable_styling(font_size = 10) |>
+  row_spec(0, bold = TRUE) #, width = "5em")
 ```

### 2025-03-01 jso.tex
caption on fig 1

--- 2025-03-01 jso.tex
@@
 projection. A Huber plot shows the PP index values for all 1D data
-projections in polar coordinates.}
+projections in polar coordinates. The skewness index on the trimodal
+data is also smoother than the binned normality index on RANDU.}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, left=2mm, toptitle=1mm, colback=white, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, opacityback=0, titlerule=0mm, bottomtitle=1mm, rightrule=.15mm, arc=.35mm, toprule=.15mm, opacitybacktitle=0.6, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, breakable, left=2mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, opacityback=0, bottomrule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, colback=white, leftrule=.75mm, bottomtitle=1mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black]
 
@@
 \midrule\addlinespace[2.5pt]
-Smoothness & 1.00 (0.81 to 1.20) & 0.961 \\ 
-Squintability & 1.24 (1.05 to 1.58) & \textbf{0.033} \\ 
-Dimension & 0.57 (0.33 to 0.87) & \textbf{0.022} \\ 
-Number of jellyfish & 1.32 (1.06 to 1.71) & \textbf{0.022} \\ 
-Maximum number of tries & 1.06 (0.85 to 1.35) & 0.603 \\ 
+Smoothness & 1.00 (0.90 to 1.11) & 0.928 \\ 
+Squintability & 1.24 (1.13 to 1.39) & \textbf{<0.001} \\ 
+Dimension & 0.57 (0.44 to 0.73) & \textbf{<0.001} \\ 
+Number of jellyfish & 1.32 (1.17 to 1.50) & \textbf{<0.001} \\ 
+Maximum number of tries & 1.06 (0.94 to 1.21) & 0.333 \\ 
 \bottomrule
@@
 
-\begin{longtable}[t]{>{\raggedright\arraybackslash}p{5em}rrrrrr}
-
-\caption{\label{tbl-joint-output}Model summaries for squintability and
-smoothness separately}
+\begin{longtable}[t]{rrrr}
+
+\caption{\label{tbl-joint-output}Fit statistics for the model.}
 
@@
 \toprule
-Model & Null Deviance & DF Null & Log-Likelihood & DF Residual & AIC & BIC\\
+\textbf{Deviance} & \textbf{DF Residual} & \textbf{Null Deviance} & \textbf{DF Null}\\
 \midrule
-\textbf{Squintability} & 49.29 & 75 & -26.44 & 71 & 62.87 & 74.53\\
-\textbf{Smoothness} & 49.29 & 75 & -33.23 & 71 & 76.47 & 88.12\\
+20.69 & 70 & 49.29 & 75\\
 \bottomrule

### 2025-03-02 jso.qmd
revising section 6.1

@@
 
-# Assessing optimisers: two simulation studies {#sec-sim-deets}
+# Assessing the optimisers {#sec-sim-deets}
 
@@
 
+This section summarises the findings from the simulations that compare the JSO performance with the existing CRS optimiser, and the relationship between optimisation success and hyper-parameter choices and PP index properties.
+
+```{r}
+#| fig.height: 7.5
+#| fig.width: 10
+#| label: fig-proj
+#| fig.align: "center"
+#| fig.cap: "Visual comparison of JSO and CRS results, using optimal data projections obtained over 50 simulations of the pipe data.
+Rows correspond to data dimension.
+Columns correspond to quantiles of the index values, with 0 being minimum, 100 being maximum, and 50 the median.
+JSO achieves the better views of the pipe generally than CRS.
+As dimension increases both have more difficulty finding the pipe."
+
+aaa <- map_dfr(
+  list(pipe1000, pipe1000_8d, pipe1000_10d, pipe1000_12d), 
+  ~{
+    d_dim = dim(.x)[2]
+    pipe_jellyfish |> 
+      filter(d == d_dim) |> 
+      arrange(index_val) |>
+      filter(row_number() %in% c(1, seq(5, 50, 5))) |> 
+      mutate(id = factor(
+        paste0(c(0, seq(5, 50, 5)) / 50 * 100),
+        levels = c(0, seq(5, 50, 5) / 50 * 100))) |> 
+      compute_projection(data = .x, col = c("d", "index_val", "id"))
+  }
+)
+
+p1 <- aaa |> 
+  ggplot() + 
+  geom_point(aes(x = V1, y = V2), size = 0.05) + 
+  geom_text(data = aaa |> select(-V1, -V2) |> unique(), 
+            aes(label = sprintf("%.3f", index_val)),
+            x = 0, y = 3.7, 
+            size = 3) + 
+  facet_grid(d ~ id) + 
+  xlim(-4, 4) + ylim(-4, 4) +
+  theme_bw() +
+  theme(aspect.ratio = 1, axis.ticks = element_blank(),
+        axis.text = element_blank(), axis.title = element_blank(),
+        panel.grid = element_blank())
+
+bbb <- pipe_better |> 
+  group_by(dim) |> 
+  arrange(index_val) |> 
+  filter(row_number() %in% c(1, seq(5, 50, 5))) |>  
+  mutate(id = factor(
+    paste0(c(0, seq(5, 50, 5)) / 50 * 100),
+        levels = c(0, seq(5, 50, 5) / 50 * 100)), 
+    d = 4 + 2 * as.numeric(dim)) |> 
+  unnest(proj)
+
+p2 <- bbb |> 
+  ggplot() + 
+  geom_point(aes(x = V1, y = V2), size = 0.05) + 
+  geom_text(data = bbb |> select(-V1, -V2) |> unique(), 
+            aes(label = sprintf("%.3f", index_val)), x = 0, y = 3.7, 
+            size = 3) + 
+  facet_grid(d ~ id) + 
+  xlim(-4, 4) + ylim(-4, 4) +
+  theme_bw() +
+  theme(aspect.ratio = 1, axis.ticks = element_blank(),
+        axis.text = element_blank(), axis.title = element_blank(),
+        panel.grid = element_blank())
+
+(p1 + ggtitle("a. JSO"))/(p2 + ggtitle("b. CRS"))
+```
+
 ## Performance of JSO relative to CRS
 
-Performance of JSO is assessed based on the final projections across the
-two optimisers: JSO and CRS. The final projections found by the two
-optimisers are presented in @fig-proj, broken down by 10th quantile,
+The performance of JSO is assessed relative to the existing CRS optimiser based on the optimal projection obtained.
+The final projections for the pipe data sets found by the two optimisers are shown in @fig-proj, with rows corresponding to data dimension.
+The 50 simulations in each data dimension are sorted by index value, and the projections corresponding to each 10th quantile value are shown, with minimum at left and maximum at right.
+Index value is printed on the plot.
+The purpose is to summarise the views of the data resulting from different optimisations, and hence compare results between the two optimisers.
+Generally, the JSO does a more consistent job of finding the pipe structure clearly.
+As the data dimension increases, both optimisers struggle and less clearly capture the circle shape.
+
+<!-- broken down by 10th quantile,
 faceted by the data dimensions. In the 6-dimensional data scenario, JSO
@@
 dimensions, suggesting its advantage in exploring high-dimensional
-spaces.
-
-## Factors affecting JSO success rate: JSO hyper-parameters and index properties
+spaces.-->
+
+## Effect of hyper-parameters effect on JSO success rate
 
@@
 
+## Effect of index properties on JSO success rate
+
 Smoothness and squintability are calculated across a collection of
@@
 as equation (@eq-squintability-parametric) as the squintability metric.
-
-```{r}
-#| fig.height: 7.5
-#| fig.width: 10
-#| label: fig-proj
-#| fig.align: "center"
-#| fig.cap: "Projections found by the JSO and CRS at each 10th quantile across 50 simulations.
-The projection pursuit problem is to find the pipe shape using the holes index in the 6, 8, 10, and 12-dimensional spaces.
-The JSO uses 100 jellyfish and a maximum number of tries of 100.
-The CRS uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates.
-In the 6-D data space, JSO always finds a clear pipe shape while the CRS also finds the pipe shape but with a wide rim.
-At higher data dimensions, JSO finds a higher index value and a clearer pipe shape across all the quantiles than the CRS"
-
-aaa <- map_dfr(
-  list(pipe1000, pipe1000_8d, pipe1000_10d, pipe1000_12d), 
-  ~{
-    d_dim = dim(.x)[2]
-    pipe_jellyfish |> 
-      filter(d == d_dim) |> 
-      arrange(index_val) |>
-      filter(row_number() %in% c(1, seq(5, 50, 5))) |> 
-      mutate(id = factor(
-        paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th"),
-        levels = paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th"))) |> 
-      compute_projection(data = .x, col = c("d", "index_val", "id"))
-  }
-)
-
-p1 <- aaa |> 
-  ggplot() + 
-  geom_point(aes(x = V1, y = V2), size = 0.05) + 
-  geom_text(data = aaa |> select(-V1, -V2) |> unique(), 
-            aes(label = round(index_val,3)), x = 0, y = 3.7, 
-            size = 3) + 
-  facet_grid(d ~ id) + 
-  xlim(-4, 4) + ylim(-4, 4) +
-  theme_bw() +
-  theme(aspect.ratio = 1, axis.ticks = element_blank(),
-        axis.text = element_blank(), axis.title = element_blank(),
-        panel.grid = element_blank())
-
-bbb <- pipe_better |> 
-  group_by(dim) |> 
-  arrange(index_val) |> 
-  filter(row_number() %in% c(1, seq(5, 50, 5))) |>  
-  mutate(id = factor(
-    paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th"),
-    levels = paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th")), 
-    d = 4 + 2 * as.numeric(dim)) |> 
-  unnest(proj)
-
-p2 <- bbb |> 
-  ggplot() + 
-  geom_point(aes(x = V1, y = V2), size = 0.05) + 
-  geom_text(data = bbb |> select(-V1, -V2) |> unique(), 
-            aes(label = round(index_val,3)), x = 0, y = 3.7, 
-            size = 3) + 
-  facet_grid(d ~ id) + 
-  xlim(-4, 4) + ylim(-4, 4) +
-  theme_bw() +
-  theme(aspect.ratio = 1, axis.ticks = element_blank(),
-        axis.text = element_blank(), axis.title = element_blank(),
-        panel.grid = element_blank())
-
-(p1 + ggtitle("a. JSO"))/(p2 + ggtitle("b. CRS"))
-```
 

### 2025-03-02 jso.tex
revising section 6.1

@@
 
-\begin{tcolorbox}[enhanced jigsaw, breakable, left=2mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, opacityback=0, bottomrule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, colback=white, leftrule=.75mm, bottomtitle=1mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, left=2mm, colback=white, opacityback=0, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, breakable, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, bottomrule=.15mm, toprule=.15mm, rightrule=.15mm, titlerule=0mm]
 
@@
 
-\section{Assessing optimisers: two simulation
-studies}\label{sec-sim-deets}
+\section{Assessing the optimisers}\label{sec-sim-deets}
 
@@
 
+This section summarises the findings from the simulations that compare
+the JSO performance with the existing CRS optimiser, and the
+relationship between optimisation success and hyper-parameter choices
+and PP index properties.
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-proj-1.pdf}}
+
+}
+
+\caption{\label{fig-proj}Visual comparison of JSO and CRS results, using
+optimal data projections obtained over 50 simulations of the pipe data.
+Rows correspond to data dimension. Columns correspond to quantiles of
+the index values, with 0 being minimum, 100 being maximum, and 50 the
+median. JSO achieves the better views of the pipe generally than CRS. As
+dimension increases both have more difficulty finding the pipe.}
+
+\end{figure}%
+
 \subsection{Performance of JSO relative to
@@
 
-Performance of JSO is assessed based on the final projections across the
-two optimisers: JSO and CRS. The final projections found by the two
-optimisers are presented in Figure~\ref{fig-proj}, broken down by 10th
-quantile, faceted by the data dimensions. In the 6-dimensional data
-scenario, JSO consistently identifies a clear pipe shape. The CRS also
-finds the pipe shape but with a wide rim, suggesting a further polish
-search may be required. With increasing dimensions, JSO may not always
-identify the pipe shape due to random sampling, but it still finds the
-pipe shape in over 50\% of cases. Compared to CRS, JSO achieves higher
-index values and clearer pipe shapes across all quantiles in data of 8,
-10, 12 dimensions, suggesting its advantage in exploring
-high-dimensional spaces.
-
-\subsection{Factors affecting JSO success rate: JSO hyper-parameters and
-index
-properties}\label{factors-affecting-jso-success-rate-jso-hyper-parameters-and-index-properties}
+The performance of JSO is assessed relative to the existing CRS
+optimiser based on the optimal projection obtained. The final
+projections for the pipe data sets found by the two optimisers are shown
+in Figure~\ref{fig-proj}, with rows corresponding to data dimension. The
+50 simulations in each data dimension are sorted by index value, and the
+projections corresponding to each 10th quantile value are shown, with
+minimum at left and maximum at right. Index value is printed on the
+plot. The purpose is to summarise the views of the data resulting from
+different optimisations, and hence compare results between the two
+optimisers. Generally, the JSO does a more consistent job of finding the
+pipe structure clearly. As the data dimension increases, both optimisers
+struggle and less clearly capture the circle shape.
+
+\subsection{Effect of hyper-parameters effect on JSO success
+rate}\label{effect-of-hyper-parameters-effect-on-jso-success-rate}
 
@@
 
+\subsection{Effect of index properties on JSO success
+rate}\label{effect-of-index-properties-on-jso-success-rate}
+
 Smoothness and squintability are calculated across a collection of
@@
 (\ref{eq-squintability-parametric}) as the squintability metric.
-
-\begin{figure}
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-proj-1.pdf}}
-
-}
-
-\caption{\label{fig-proj}Projections found by the JSO and CRS at each
-10th quantile across 50 simulations. The projection pursuit problem is
-to find the pipe shape using the holes index in the 6, 8, 10, and
-12-dimensional spaces. The JSO uses 100 jellyfish and a maximum number
-of tries of 100. The CRS uses a maximum of 1000 tries in each step of
-random sampling step before the algorithm terminates. In the 6-D data
-space, JSO always finds a clear pipe shape while the CRS also finds the
-pipe shape but with a wide rim. At higher data dimensions, JSO finds a
-higher index value and a clearer pipe shape across all the quantiles
-than the CRS}
-
-\end{figure}%
 

### 2025-03-02 jso.qmd
add references to pkgs used in the work

@@
 library(tidyverse)
-library(ggplot2)
 library(patchwork)
@@
 library(tourr)
-library(PPtreeViz)
 library(ferrn)
-library(colorspace)
-library(broom)
 load(here::here("data/sim_pipe_run_best.rda"))
@@
 The columns $\nu$ and $\varsigma$ represent the smoothness and squintability measures respectively.
-dt <- tibble(shape = c(rep("pipe", 4), rep("sine", 18)), smoothness) |>
+dt <- tibble(shape = c(rep("pipe", 5), rep("sine", 17)), smoothness) |>
   left_join(squintability) |> 
@@
 
+The R packages used in this work include: `tidyverse` [@tidyverse]; 
+`patchwork` [@patchwork]; `ggh4x` [@ggh4x]; `broom` [@broom]; 
+`kableExtra` [@kabelextra]; `ferrn` [@RJ-2021-105]; and `cassowaryr` [@cassowaryr].
+
 # Supplementary materials {.unnumbered}

### 2025-03-02 jso.tex
add references to pkgs used in the work

@@
 of the optimum. An implementation is available in the \texttt{tourr}
-package (Wickham et al. 2011) in R (R Core Team 2023). Zhang et al.
+package (H. Wickham et al. 2011) in R (R Core Team 2023). Zhang et al.
 (2021) illustrated how to diagnose optimisation processes, particularly
@@
 
-\begin{tcolorbox}[enhanced jigsaw, left=2mm, colback=white, opacityback=0, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, breakable, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, bottomrule=.15mm, toprule=.15mm, rightrule=.15mm, titlerule=0mm]
+\begin{tcolorbox}[enhanced jigsaw, breakable, toprule=.15mm, titlerule=0mm, coltitle=black, left=2mm, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomtitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6]
 
@@
 pipe & holes & 10 & 0.02 & 0.34 & 2.88 & 0.00 & 1.02 & 0.88 & 3.15 & 0.02 & 0.43\\
-sine & holes & 12 & 0.02 & 0.43 & 2.72 & 0.00 & 1.01 & 0.88 & 3.34 & 0.00 & 0.41\\
+pipe & holes & 12 & 0.02 & 0.43 & 2.72 & 0.00 & 1.01 & 0.88 & 3.34 & 0.00 & 0.41\\
 sine & MIC & 4 & 0.14 & 0.27 & 3.56 & 0.00 & 0.94 & 0.15 & 1.14 & -0.04 & 0.39\\
@@
 
+\begingroup
+\fontsize{12.0pt}{14.4pt}\selectfont
 \setlength{\LTpost}{0mm}
@@
 Smoothness & 1.00 (0.90 to 1.11) & 0.928 \\ 
-Squintability & 1.24 (1.13 to 1.39) & \textbf{<0.001} \\ 
-Dimension & 0.57 (0.44 to 0.73) & \textbf{<0.001} \\ 
-Number of jellyfish & 1.32 (1.17 to 1.50) & \textbf{<0.001} \\ 
+Squintability & 1.24 (1.13 to 1.39) & {\bfseries <0.001} \\ 
+Dimension & 0.57 (0.44 to 0.73) & {\bfseries <0.001} \\ 
+Number of jellyfish & 1.32 (1.17 to 1.50) & {\bfseries <0.001} \\ 
 Maximum number of tries & 1.06 (0.94 to 1.21) & 0.333 \\ 
@@
 
+The R packages used in this work include: \texttt{tidyverse} (Hadley
+Wickham et al. 2019); \texttt{patchwork} (Pedersen 2024); \texttt{ggh4x}
+(van den Brand 2024); \texttt{broom} (Robinson, Hayes, and Couch 2024);
+\texttt{kableExtra} (Zhu 2024); \texttt{ferrn} (Zhang et al. 2021); and
+\texttt{cassowaryr} (Mason et al. 2022).
+
 \section*{Supplementary materials}\label{supplementary-materials}
@@
 
+\bibitem[\citeproctext]{ref-cassowaryr}
+Mason, Harriet, Stuart Lee, Ursula Laa, and Dianne Cook. 2022.
+\emph{Cassowaryr: Compute Scagnostics on Pairs of Numeric Variables in a
+Data Set}. \url{https://CRAN.R-project.org/package=cassowayr}.
+
+\bibitem[\citeproctext]{ref-patchwork}
+Pedersen, Thomas Lin. 2024. \emph{Patchwork: The Composer of Plots}.
+\url{https://CRAN.R-project.org/package=patchwork}.
+
 \bibitem[\citeproctext]{ref-porcu2024matern}
@@
 
+\bibitem[\citeproctext]{ref-broom}
+Robinson, David, Alex Hayes, and Simon Couch. 2024. \emph{Broom: Convert
+Statistical Objects into Tidy Tibbles}.
+\url{https://CRAN.R-project.org/package=broom}.
+
 \bibitem[\citeproctext]{ref-barnett1981interpreting}
@@
 
+\bibitem[\citeproctext]{ref-ggh4x}
+van den Brand, Teun. 2024. \emph{Ggh4x: Hacks for 'Ggplot2'}.
+\url{https://CRAN.R-project.org/package=ggh4x}.
+
+\bibitem[\citeproctext]{ref-tidyverse}
+Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
+D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
+{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
+4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.
+
 \bibitem[\citeproctext]{ref-tourr}
@@
 
+\bibitem[\citeproctext]{ref-kabelextra}
+Zhu, Hao. 2024. \emph{kableExtra: Construct Complex Table with 'Kable'
+and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
+
 \end{CSLReferences}

### 2025-03-02 jso.qmd
updating description of simulations

@@
 
-This section describes the details of two simulation studies. The first simulation
-compares the JSO performance with an existing optimiser, Creeping Random
-Search (CRS) [@RJ-2021-105; @laa_using_2020] to explore JSO's behaviour
-under different data dimensions. The second simulation examines various factors affecting the performance of JSO.
+This section explains the details of two simulation studies: (1) comparison between the existing Creeping Random Search (CRS) [@RJ-2021-105; @laa_using_2020] and JSO, and (2) examining the factors affecting the performance of JSO.
+
+The second simulation examines various factors affecting the performance of JSO.
 The JSO performance is evaluated under different hyper-parameters combinations
@@
 
-The performance of JSO is measured by the success rate, defined as the
-proportion of simulations that achieves a final index value within 0.05
-of the best index value found among all 50 simulations (see
-@fig-success-rate for an illustration). This comparison is based on
-projection pursuit to find the pipe shape investigated by
-@laa_using_2020 using the `holes` index.
-
 ## Performance of JSO relative to CRS {#sec-app-1}
 
-The JSO performance is compared to the existing optimizer, CRS, in a pipe-finding problem using the holes index in different dimensional spaces.
-Fifty simulations are conducted in four data dimensions ($d = 6, 8, 10, 12$) with each optimiser.
-JSO uses
-100 jellyfish with a maximum of 100 tries, while the CRS allows a
-maximum of 1000 samples at each iteration before the algorithm
-terminates. The different numbers account for the multiple paths of JSO
-to enable fairer comparison with CRS. The results of the simulation are
-collected using the data structure proposed in @RJ-2021-105 for
-assessing JSO, where the design parameters are stored along with index
-value, projection basis, random seed, and computation time.
+The CRS is the main optimisation routine currently used for the guided tour.
+Here the two optimisers are compared on the task of finding the pipe using the `holes` index in data with dimensions, $d = 6, 8, 10, 12$.
+Fifty simulated data sets are used for each dimension.
+JSO uses 100 jellyfish with a maximum of 100 tries, while the CRS allows a maximum of 1000 tries at each iteration before the algorithm terminates.
+These choices enable fair comparison between CRS and JSO, that conforms to how they are used in practice.
+
+The performance of the optmisers is measured by the success rate, which is defined as the proportion of simulations that achieves a final index value within 0.05
+of the best index value found among all 50 simulations.
+@fig-success-rate illustrates why the choice of 0.05 is reasonable: the pipe is not recognisable in the projected data.
+This is motivated by @laa_using_2020's approach to investigating PP indexes.
+
+The results of the simulation are collected using the data structure used in @RJ-2021-105 for assessing PP optimisers.
+The design parameters are stored along with index value, projection basis, random seed, and computation time.
 
@@
 #| label: fig-success-rate
-#| fig.cap: "Illustration of success rate calculation: Final projections based on projection pursuit to find the pipe shape in 8D data using the holes index, optimised by CRS, in 50 simulations.
-The 50 final projections are sorted by their index values.
-The highest index value found across all simulations is 0.969.
-Out of the 50 simulations, 43 achieved an index value within 0.05 of the best, resulting in a success rate of 0.86 (43/50)."
+#| fig.cap: "How success rate is calculated, illustrated using the optimal projections from 50 optimisations of 8D pipe data sorted by index value.
+The pipe shape is recognisable in the projection index values between 0.933-0.969.
+Of the 50 simulations, 43 achieved an index value within 0.05 of the best, resulting in a success rate of 0.86."
 # pipe_jellyfish_12d <- pipe_better |> 
@@
 #   geom_point(size = 0.5) +
-#   geom_text(aes(label = round(index_val, 3)), x = 0, y = 3.7,
+#   geom_text(aes(label = sprintf("%.3f", index_val)), x = 0, y = 3.7,
 #             size = 8) +
@@
 for fifty times in each hyper-parameter combination: 20, 50, and 100
-jellyfish and a maximum of 50 and 100 attempts. The success rate is
-calculated for each combination.
+jellyfish and a maximum of 50 and 100 attempts.
 
 In addition, to assess the performance of JSO across various PP problems, two different data shapes, pipe and a sine wave, are considered in various dimensional spaces using six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`,
-`spline`, and `stringy`, under varied JSO hyper-parameters. An example of holes and pipes data with the syntax to generate them are provided in the supplementary material. This results
-in a total of `r nrow(sim_df)` scenarios, comprising of
+`spline`, and `stringy`, under varied JSO hyper-parameters.
+An example of holes and pipes data with the syntax to generate them are provided in the supplementary material.
+This results in a total of `r nrow(sim_df)` scenarios, comprising of
 `r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and
@@
 
-To construct a relationship among success rate, smoothness,
-squintability, and JSO hyper-parameters, a generalised linear model is
-fitted using a binomial family and a logit link function. The data is
+A generalised linear model is fitted using a binomial family and a logit link function to assess the factors affecting success rate.  Predictors are smoothness,
+squintability, and JSO hyper-parameters. A success rate of 0 indicates the optimiser did not come sufficiently close to the optimal projection. <!-- The data is
 pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
-for interpretation, 2) <!-- creating a new binary variable `long_time` to
-indicate cases with an average run time over 20 seconds, and 3)-->
+for interpretation, 2) creating a new binary variable `long_time` to
+indicate cases with an average run time over 20 seconds, and 3)
 re-coding the success rate for the `stringy` index as 0, because none of
-the 50 simulations correctly identified the sine-wave shape.
+the 50 simulations correctly identified the sine-wave shape.-->
 
@@
 
-## Effect of index properties on JSO success rate
-
-Smoothness and squintability are calculated across a collection of
-pipe-finding and sine-wave finding problems to construct the
-relationship between success rate, JSO hyper-parameters, and index
-properties. @tbl-smoothness-squintability presents the parameters
-estimated from the Gaussian process (outputscale $\eta$, lengthscale
-$\ell$, smoothness $\nu$, and nugget $\sigma$) and from the scaled
-logistic function ($\theta_1$ to $\theta_4$) for calculating smoothness
-and squintability in each PP problem considered. The column $\nu$ is
-used as the smoothness metric and the column $\varsigma$ is calculated
-as equation (@eq-squintability-parametric) as the squintability metric.
-
 ```{r}
@@
 #| fig.align: "center"
-#| fig.cap: "The proportion of simulations that reach near-optimal index values in the pipe-finding problem using the holes index.
-The proportion is calculated based on the number of simulations, out of 50, that achieve an index value within 0.05 of the best-performing simulation.
-To quantify uncertainty, 500 bootstrap samples are generated.
-The thin segments represent the proportion for each of the 500 bootstrap replicates, while the thicker segments represent the mean of these bootstrap samples, connnected by lines.
-As the dimensionality increases, the proportion of simulations reaching the optimal index value decreases."
+#| fig.cap: "Summary of the relationship between JSO hyper-parameters on optimisation success, using the holes index for pipe data of dimensions 4-12.
+Bootstrap samples show the variability in success rate.
+Success rate mostly plateaus by 50 jellies.
+The JSO has some difficulty finding the pipe when dimension is higher than 8!
+Maximum number of tries has little effect."
 pipe_sim_best <- sim_pipe_run_best |> 
@@
 ```
+
+## Effect of index properties on JSO success rate
+
+Smoothness and squintability are calculated across a collection of
+pipe-finding and sine-wave finding problems to construct the
+relationship between success rate, JSO hyper-parameters, and index
+properties. @tbl-smoothness-squintability presents the parameters
+estimated from the Gaussian process (outputscale $\eta$, lengthscale
+$\ell$, smoothness $\nu$, and nugget $\sigma$) and from the scaled
+logistic function ($\theta_1$ to $\theta_4$) for calculating smoothness
+and squintability in each PP problem considered. The column $\nu$ is
+used as the smoothness metric and the column $\varsigma$ is calculated
+as equation (@eq-squintability-parametric) as the squintability metric.
 

### 2025-03-02 jso.tex
updating description of simulations

@@
 
-\begin{tcolorbox}[enhanced jigsaw, breakable, toprule=.15mm, titlerule=0mm, coltitle=black, left=2mm, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, rightrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toptitle=1mm, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, bottomtitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, opacitybacktitle=0.6, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, opacityback=0, breakable, bottomrule=.15mm, bottomtitle=1mm, titlerule=0mm, coltitle=black, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, toprule=.15mm, leftrule=.75mm, left=2mm]
 
@@
 
-This section describes the details of two simulation studies. The first
-simulation compares the JSO performance with an existing optimiser,
-Creeping Random Search (CRS) (Zhang et al. 2021; Laa and Cook 2020) to
-explore JSO's behaviour under different data dimensions. The second
-simulation examines various factors affecting the performance of JSO.
-The JSO performance is evaluated under different hyper-parameters
-combinations (number of jellyfish and the maximum number of tries). The
-effect of index properties (smoothness and squintability) along with JSO
-hyper-parameters and data dimension variations is captured and studied
-using logistic regression.
-
-The performance of JSO is measured by the success rate, defined as the
-proportion of simulations that achieves a final index value within 0.05
-of the best index value found among all 50 simulations (see
-Figure~\ref{fig-success-rate} for an illustration). This comparison is
-based on projection pursuit to find the pipe shape investigated by Laa
-and Cook (2020) using the \texttt{holes} index.
+This section explains the details of two simulation studies: (1)
+comparison between the existing Creeping Random Search (CRS) (Zhang et
+al. 2021; Laa and Cook 2020) and JSO, and (2) examining the factors
+affecting the performance of JSO.
+
+The second simulation examines various factors affecting the performance
+of JSO. The JSO performance is evaluated under different
+hyper-parameters combinations (number of jellyfish and the maximum
+number of tries). The effect of index properties (smoothness and
+squintability) along with JSO hyper-parameters and data dimension
+variations is captured and studied using logistic regression.
 
@@
 
-The JSO performance is compared to the existing optimizer, CRS, in a
-pipe-finding problem using the holes index in different dimensional
-spaces. Fifty simulations are conducted in four data dimensions
-(\(d = 6, 8, 10, 12\)) with each optimiser. JSO uses 100 jellyfish with
-a maximum of 100 tries, while the CRS allows a maximum of 1000 samples
-at each iteration before the algorithm terminates. The different numbers
-account for the multiple paths of JSO to enable fairer comparison with
-CRS. The results of the simulation are collected using the data
-structure proposed in Zhang et al. (2021) for assessing JSO, where the
-design parameters are stored along with index value, projection basis,
-random seed, and computation time.
+The CRS is the main optimisation routine currently used for the guided
+tour. Here the two optimisers are compared on the task of finding the
+pipe using the \texttt{holes} index in data with dimensions,
+\(d = 6, 8, 10, 12\). Fifty simulated data sets are used for each
+dimension. JSO uses 100 jellyfish with a maximum of 100 tries, while the
+CRS allows a maximum of 1000 tries at each iteration before the
+algorithm terminates. These choices enable fair comparison between CRS
+and JSO, that conforms to how they are used in practice.
+
+The performance of the optmisers is measured by the success rate, which
+is defined as the proportion of simulations that achieves a final index
+value within 0.05 of the best index value found among all 50
+simulations. Figure~\ref{fig-success-rate} illustrates why the choice of
+0.05 is reasonable: the pipe is not recognisable in the projected data.
+This is motivated by Laa and Cook (2020)'s approach to investigating PP
+indexes.
+
+The results of the simulation are collected using the data structure
+used in Zhang et al. (2021) for assessing PP optimisers. The design
+parameters are stored along with index value, projection basis, random
+seed, and computation time.
 
@@
 
-\caption{\label{fig-success-rate}Illustration of success rate
-calculation: Final projections based on projection pursuit to find the
-pipe shape in 8D data using the holes index, optimised by CRS, in 50
-simulations. The 50 final projections are sorted by their index values.
-The highest index value found across all simulations is 0.969. Out of
-the 50 simulations, 43 achieved an index value within 0.05 of the best,
-resulting in a success rate of 0.86 (43/50).}
+\caption{\label{fig-success-rate}How success rate is calculated,
+illustrated using the optimal projections from 50 optimisations of 8D
+pipe data sorted by index value. The pipe shape is recognisable in the
+projection index values between 0.933-0.969. Of the 50 simulations, 43
+achieved an index value within 0.05 of the best, resulting in a success
+rate of 0.86.}
 
@@
 for fifty times in each hyper-parameter combination: 20, 50, and 100
-jellyfish and a maximum of 50 and 100 attempts. The success rate is
-calculated for each combination.
+jellyfish and a maximum of 50 and 100 attempts.
 
@@
 
-To construct a relationship among success rate, smoothness,
-squintability, and JSO hyper-parameters, a generalised linear model is
-fitted using a binomial family and a logit link function. The data is
-pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
-for interpretation, 2) re-coding the success rate for the
-\texttt{stringy} index as 0, because none of the 50 simulations
-correctly identified the sine-wave shape.
+A generalised linear model is fitted using a binomial family and a logit
+link function to assess the factors affecting success rate. Predictors
+are smoothness, squintability, and JSO hyper-parameters. A success rate
+of 0 indicates the optimiser did not come sufficiently close to the
+optimal projection.
 
@@
 
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-proportion-1.pdf}}
+
+}
+
+\caption{\label{fig-proportion}Summary of the relationship between JSO
+hyper-parameters on optimisation success, using the holes index for pipe
+data of dimensions 4-12. Bootstrap samples show the variability in
+success rate. Success rate mostly plateaus by 50 jellies. The JSO has
+some difficulty finding the pipe when dimension is higher than 8!
+Maximum number of tries has little effect.}
+
+\end{figure}%
+
 \subsection{Effect of index properties on JSO success
@@
 (\ref{eq-squintability-parametric}) as the squintability metric.
-
-\begin{figure}
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-proportion-1.pdf}}
-
-}
-
-\caption{\label{fig-proportion}The proportion of simulations that reach
-near-optimal index values in the pipe-finding problem using the holes
-index. The proportion is calculated based on the number of simulations,
-out of 50, that achieve an index value within 0.05 of the
-best-performing simulation. To quantify uncertainty, 500 bootstrap
-samples are generated. The thin segments represent the proportion for
-each of the 500 bootstrap replicates, while the thicker segments
-represent the mean of these bootstrap samples, connnected by lines. As
-the dimensionality increases, the proportion of simulations reaching the
-optimal index value decreases.}
-
-\end{figure}%
 
@@
 Smoothness & 1.00 (0.90 to 1.11) & 0.928 \\ 
-Squintability & 1.24 (1.13 to 1.39) & {\bfseries <0.001} \\ 
-Dimension & 0.57 (0.44 to 0.73) & {\bfseries <0.001} \\ 
-Number of jellyfish & 1.32 (1.17 to 1.50) & {\bfseries <0.001} \\ 
+Squintability & 1.24 (1.13 to 1.39) & \textbf{<0.001} \\ 
+Dimension & 0.57 (0.44 to 0.73) & \textbf{<0.001} \\ 
+Number of jellyfish & 1.32 (1.17 to 1.50) & \textbf{<0.001} \\ 
 Maximum number of tries & 1.06 (0.94 to 1.21) & 0.333 \\ 

### 2025-03-02 code/00-sim-search-better.R
spinebil -> ferrn for pipe/sine data

@@ -19,10 +19,17 @@ historyarray2dt <- function(array, index_f){
 }
 
 sim <- function(d, optimiser = c("search_better_random", "search_better", "search_jellyfish"),
-                data_seed = 123456, optim_seed = seed){
-  set.seed(data_seed)
-  pipe1000 <- pipeData(d, 1000) %>% scale() %>% as_tibble()
-  colnames(pipe1000) <- paste0("V", 1:d)
+                optim_seed = seed){
+ if (d == 6){
+    pipe1000 <- ferrn::pipe1000_6d
+  } else if (d == 8){
+    pipe1000 <- ferrn::pipe1000_8d
+  } else if (d == 10){
+    pipe1000 <- ferrn::pipe1000_10d
+  } else if (d == 12){
+    pipe1000 <- ferrn::pipe1000_12d
+  }
+
 
   if (optimiser == "search_better_random"){
     res <- map_dfr(optim_seed, function(seed){
@@ -50,11 +57,9 @@ sim <- function(d, optimiser = c("search_better_random", "search_better", "searc
   return(res)
 }
 
-data <- map_dfr(c(6, 8, 10, 12), ~{
-  set.seed(123456)
-  tibble(x = list(pipeData(.x, 1000) %>% scale()))
-}) |>
-  mutate(dim = c(6, 8, 10, 12))
+data <- tibble(x = list(ferrn::pipe1000_4d, ferrn::pipe1000_6d,
+                        ferrn::pipe1000_10d, ferrn::pipe1000_12d),
+               dim = c(6, 8, 10, 12))
 # a <- map_dfr(c(6, 8, 10, 12),
 #              ~safe_sim(d = .x, optimiser = "search_better",
 #                        optim_seed = seed),

### 2025-03-02 code/01-sim-pipe-681012d-holes.R

@@ -1,12 +1,19 @@
 library(tidyverse)
-library(spinebil)
+library(ferrn)
 library(tourr)
 
 sim <- function(d = d, n_jellies = n_jellies, max.tries = max.tries,
-                data_seed = 123456, optim_seed = seed){
-  set.seed(data_seed)
-  pipe1000 <- pipeData(d, 1000) %>% scale() %>% as_tibble()
-  colnames(pipe1000) <- paste0("V", 1:d)
+                optim_seed = seed){
+  if (d == 6){
+    pipe1000 <- ferrn::pipe1000_6d
+  } else if (d == 8){
+    pipe1000 <- ferrn::pipe1000_8d
+  } else if (d == 10){
+    pipe1000 <- ferrn::pipe1000_10d
+  } else if (d == 12){
+    pipe1000 <- ferrn::pipe1000_12d
+  }
+
 
   cat("n_jellies: ", n_jellies, "\n")
   cat("max.tries: ", max.tries, "\n")
@@ -47,7 +54,7 @@ save(sim_pipe, file = "data-raw/sim_pipe.rda")
 
 ################################################################################
 ################################################################################
-# additional smaller data
+# 4D data is simulated in 10-sim-4d.R
 load(here::here("data-raw/sim_pipe_4d.rda"))
 sim_pipe_run_best <- bind_rows(sim_pipe_4d, sim_pipe) |>
   group_by(id, n_jellies, max_tries, d) |>

### 2025-03-02 code/02-sim-sine-6d-splines2d.R

@@ -1,14 +1,14 @@
 library(tidyverse)
-library(spinebil)
+library(ferrn)
 library(tourr)
+
 
 ################################################################################
 # d = 6
+# d = 4 and 8 data is simulated in 10-sim-4d.R
 sim <- function(d = d, n_jellies = n_jellies, max.tries = max.tries,
-                data_seed = 123456, optim_seed = seed, sim = sim){
-  set.seed(data_seed)
-  sine1000 <- sinData(d, 1000) %>% scale() %>% as_tibble()
-  colnames(sine1000) <- paste0("V", 1:d)
+                optim_seed = seed, sim = sim){
+  sine1000 <- ferrn::sine1000_6d
 
   cat("sim: ", sim, "\n")
   cat("n_jellies: ", n_jellies, "\n")
@@ -90,8 +90,7 @@ save(sim_sine_6d_spline_projdist, file = here::here("data/sim_sine_6d_spline_pro
 ################################################################################
 ################################################################################
 set.seed(123456)
-sine1000 <- spinebil::sinData(6, 1000) %>% scale() %>% as_tibble()
-colnames(sine1000) <- paste0("V", 1:6)
+sine1000 <- ferrn::sine1000_6d
 sum <- sim_sine_6d_spline |>
   group_by(id) |> filter(index_val == max(index_val)) |> filter(row_number() == 1)
 b <- sum |> pull(basis)

### 2025-03-02 code/03-sim-sine-6d-loess2d.R

@@ -1,14 +1,13 @@
 library(tidyverse)
-library(spinebil)
+library(ferrn)
 library(tourr)
 
 ################################################################################
 # d = 6
+# d = 4 and 8 data is simulated in 10-sim-4d.R
 sim <- function(d = d, n_jellies = n_jellies, max.tries = max.tries,
-                data_seed = 123456, optim_seed = seed, sim = sim){
-  set.seed(data_seed)
-  sine1000 <- sinData(d, 1000) %>% scale() %>% as_tibble()
-  colnames(sine1000) <- paste0("V", 1:d)
+                optim_seed = seed, sim = sim){
+  sine1000 <- ferrn::sine1000_6d
 
   cat("sim: ", sim, "\n")
   cat("n_jellies: ", n_jellies, "\n")

### 2025-03-02 code/04-sim-sine-68d-MICTIC.R

@@ -1,15 +1,18 @@
 library(tidyverse)
-library(spinebil)
+library(ferrn)
 library(tourr)
 library(minerva)
 
 ################################################################################
-# d = 6
+# d = 6 and 8
+# d = 4 data is simulated in 10-sim-4d.R
 sim <- function(d = d, idx_f = idx_f, n_jellies = n_jellies, max.tries = max.tries,
                 data_seed = 123456, optim_seed = seed, sim = sim){
-  set.seed(data_seed)
-  sine1000 <- sinData(d, 1000) %>% scale() %>% as_tibble()
-  colnames(sine1000) <- paste0("V", 1:d)
+
+  if (d == 6){
+    sine1000 <- ferrn::sine1000_6d
+  } else if (d == 8)
+    sine1000 <- ferrn::sine1000_8d
 
   cat("sim: ", sim, "\n")
   cat("n_jellies: ", n_jellies, "\n")

### 2025-03-02 code/05-sim-sine-6d-dcor2d.R

@@ -1,15 +1,12 @@
 library(tidyverse)
-library(spinebil)
+library(ferrn)
 library(tourr)
 
 ################################################################################
 # d = 6
-set.seed(123456)
-sine1000 <- sinData(6, 1000) %>% scale() %>% as_tibble()
-colnames(sine1000) <- paste0("V", 1:6)
-
+# d = 4 data is simulated in 10-sim-4d.R
 sim <- function(d = d, idx_f = idx_f, n_jellies = n_jellies, max.tries = max.tries,
-                data = sine1000 , optim_seed = seed, sim = sim){
+                data = ferrn::sine1000_6d, optim_seed = seed, sim = sim){
 
   cat("sim:", sim, "| n_jellies:", n_jellies, "| max.tries:", max.tries,
   "| idx_f:", idx_f, "| d:", d, "\n")
@@ -50,8 +47,7 @@ dcor2d_2 <- function() {
 # take less than 2 hrs
 set.seed(123)
 seed <- sample(1000: 10000, size = 50)
-sim_setup <- crossing(idx_f = c("dcor2d_2"),
-                      d = 6,
+sim_setup <- crossing(idx_f = c("dcor2d_2"), d = 6,
                       n_jellies = c(20, 50),
                       max_tries = c(50, 100)) |>
   filter(!(n_jellies == 50 & max_tries == 100)) |>

### 2025-03-02 code/07-calc-smoothness-squintability.R

@@ -1,8 +1,6 @@
 library(ferrn)
 library(tourr)
 library(tidyverse)
-library(cassowaryr)
-library(igraph)
 
 sine_8d_tbl <- function(index){
   tibble::tibble(

### 2025-03-02 code/08-process-data-for-modelling.R

@@ -1,4 +1,5 @@
 library(tidyverse)
+library(ferrn)
 
 # calculate success rate for pipe-finding problem
 load(file = "data-raw/sim_pipe.rda")

### 2025-03-02 code/09-sim-sine-6d-stringy.R

@@ -1,13 +1,8 @@
 library(cassowaryr)
 library(igraph)
 library(tidyverse)
-library(spinebil)
 library(tourr)
 
-set.seed(123456)
-sine1000 <- sinData(6, 1000) %>% scale() %>% as_tibble()
-colnames(sine1000) <- paste0("V", 1:6)
-################################################################################
 # d = 6
 stringy2 <- function(){
   function(mat){
@@ -20,10 +15,8 @@ stringy2 <- function(){
 }
 
 sim <- function(d = d, n_jellies = n_jellies, max.tries = max.tries,
-                data_seed = 123456, optim_seed = seed, sim = sim){
-  set.seed(data_seed)
-  sine1000 <- sinData(d, 1000) %>% scale() %>% as_tibble()
-  colnames(sine1000) <- paste0("V", 1:d)
+                optim_seed = seed, sim = sim){
+  sine1000 <- ferrn::sine1000_6d
 
   cat("sim: ", sim, "\n")
   cat("n_jellies: ", n_jellies, "\n")

### 2025-03-02 code/10-sim-4d.R

@@ -1,5 +1,5 @@
 library(tidyverse)
-library(spinebil)
+library(ferrn)
 library(tourr)
 
 skinny <- function(){
@@ -20,10 +20,8 @@ stringy2 <- function(){
 
 
 sim_pipe <- function(d = d, n_jellies = n_jellies, max.tries = max.tries,
-                data_seed = 123456, optim_seed = seed){
-  set.seed(data_seed)
-  pipe1000 <- pipeData(d, 1000) %>% scale() %>% as_tibble()
-  colnames(pipe1000) <- paste0("V", 1:d)
+                     optim_seed = seed){
+  pipe1000 <- ferrn::pipe1000_4d
 
   cat("n_jellies: ", n_jellies, "\n")
   cat("max.tries: ", max.tries, "\n")
@@ -65,10 +63,14 @@ save(sim_pipe_4d, file = "data-raw/sim_pipe_4d.rda")
 ################################################################################
 ################################################################################
 sim <- function(index, d = d, n_jellies = n_jellies, max.tries = max.tries,
-                data_seed = 123456, optim_seed = seed, sim = sim){
-  set.seed(data_seed)
-  sine1000 <- sinData(d, 1000) %>% scale() %>% as_tibble()
-  colnames(sine1000) <- paste0("V", 1:d)
+                optim_seed = seed, sim = sim){
+
+  if (d == 4){
+    sine1000 <- ferrn::sine1000_4d
+  } else if (d == 8){
+    sine1000 <- ferrn::sine1000_8d
+  }
+
 
   cat("sim: ", sim, "\n")
   cat("n_jellies: ", n_jellies, "\n")

### 2025-03-02 jso.tex
merged changes with Sherry

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, opacitybacktitle=0.6, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, opacityback=0, breakable, bottomrule=.15mm, bottomtitle=1mm, titlerule=0mm, coltitle=black, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, toprule=.15mm, leftrule=.75mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, colbacktitle=quarto-callout-note-color!10!white, opacitybacktitle=0.6, opacityback=0, bottomtitle=1mm, leftrule=.75mm, breakable, arc=.35mm, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, titlerule=0mm, toptitle=1mm, rightrule=.15mm, left=2mm, toprule=.15mm, bottomrule=.15mm]
 

### 2025-03-02 jso.qmd
trimmed down table 1

@@
 dt |> 
+  filter(shape == "sine", index %in% c("TIC", "splines", "skinny")) |>
+  select(shape, index, d, smooth, squint) |>
+  arrange(desc(squint)) |>
   knitr::kable(
-    digits = 2, format = "latex", 
+    digits = 4, format = "latex", 
     col.names = c(
       "shape", "index", "d",
-      "$\\eta$", "$\\ell$", "$\\nu$", "$\\sigma$",
-      "$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", "$\\varsigma$"), linesep = "",
-    booktabs = TRUE, escape = FALSE, row.names = FALSE, align = "|llr|rrrr|rrrrr") |> 
+      #"$\\eta$", "$\\ell$", 
+      "$\\nu$", #"$\\sigma$",
+      #"$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", 
+      "$\\varsigma$"), linesep = "",
+    booktabs = TRUE, escape = FALSE, row.names = FALSE, align = "|llr|rr") #align = "|llr|rrrr|rrrrr") |> 
   #column_spec(1, border_left = TRUE) |> 
   #column_spec(c(4, 8, 13), border_right = TRUE) |> 
-  kable_styling(font_size = 10) 
+  #kable_styling(font_size = 10) 
 ```

### 2025-03-02 jso.tex
trimmed down table 1

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, colbacktitle=quarto-callout-note-color!10!white, opacitybacktitle=0.6, opacityback=0, bottomtitle=1mm, leftrule=.75mm, breakable, arc=.35mm, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, titlerule=0mm, toptitle=1mm, rightrule=.15mm, left=2mm, toprule=.15mm, bottomrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, bottomrule=.15mm, toprule=.15mm, colback=white, colframe=quarto-callout-note-color-frame, leftrule=.75mm, coltitle=black, breakable, left=2mm, opacityback=0, bottomtitle=1mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm]
 
@@
 
-\centering\begingroup\fontsize{10}{12}\selectfont
-
-\begin{tabular}{|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr}
+\begin{tabular}{|llr|rr|llr|rr|llr|rr|llr|rr|llr|rr}
 \toprule
-shape & index & d & $\eta$ & $\ell$ & $\nu$ & $\sigma$ & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & $\varsigma$\\
+shape & index & d & $\nu$ & $\varsigma$\\
 \midrule
-pipe & holes & 4 & 0.03 & 0.36 & 2.99 & 0.00 & 1.00 & 0.78 & 4.50 & 0.03 & 0.31\\
-pipe & holes & 6 & 0.03 & 0.38 & 3.09 & 0.00 & 1.02 & 0.86 & 3.37 & 0.02 & 0.41\\
-pipe & holes & 8 & 0.02 & 0.33 & 2.90 & 0.00 & 1.01 & 0.87 & 3.26 & 0.03 & 0.41\\
-pipe & holes & 10 & 0.02 & 0.34 & 2.88 & 0.00 & 1.02 & 0.88 & 3.15 & 0.02 & 0.43\\
-pipe & holes & 12 & 0.02 & 0.43 & 2.72 & 0.00 & 1.01 & 0.88 & 3.34 & 0.00 & 0.41\\
-sine & MIC & 4 & 0.14 & 0.27 & 3.56 & 0.00 & 0.94 & 0.15 & 1.14 & -0.04 & 0.39\\
-sine & MIC & 6 & 0.06 & 0.31 & 3.08 & 0.00 & 0.89 & 0.57 & 1.62 & -0.03 & 0.39\\
-sine & MIC & 8 & 0.03 & 0.32 & 2.75 & 0.00 & 0.93 & 0.33 & 1.31 & -0.04 & 0.38\\
-sine & TIC & 4 & 0.40 & 0.27 & 3.49 & 0.00 & 0.97 & 0.46 & 1.55 & -0.04 & 0.38\\
-sine & TIC & 6 & 0.16 & 0.31 & 3.08 & 0.00 & 0.95 & 0.54 & 1.72 & -0.03 & 0.37\\
-sine & TIC & 8 & 0.07 & 0.30 & 2.72 & 0.00 & 0.95 & 0.56 & 1.72 & -0.03 & 0.38\\
-sine & dcor & 4 & 0.22 & 0.30 & 3.63 & 0.00 & 0.96 & 1.07 & 2.67 & -0.03 & 0.56\\
-sine & dcor & 6 & 0.08 & 0.29 & 3.13 & 0.00 & 0.95 & 1.04 & 2.74 & -0.03 & 0.53\\
-sine & loess & 4 & 0.34 & 0.50 & 2.90 & 0.00 & 1.02 & 1.10 & 2.50 & 0.05 & 0.57\\
-sine & loess & 6 & 0.24 & 0.46 & 3.56 & 0.00 & 1.02 & 1.04 & 2.65 & 0.07 & 0.53\\
-sine & loess & 8 & 0.19 & 0.46 & 3.36 & 0.00 & 1.02 & 1.04 & 2.62 & 0.08 & 0.53\\
-sine & skinny & 6 & 0.05 & 0.19 & 2.24 & 0.01 & 0.83 & 0.00 & 2.98 & 0.05 & 0.10\\
-sine & splines & 4 & 0.24 & 0.32 & 3.55 & 0.00 & 1.01 & 1.09 & 2.65 & -0.03 & 0.57\\
-sine & splines & 6 & 0.10 & 0.32 & 3.11 & 0.00 & 1.01 & 1.05 & 2.73 & -0.02 & 0.54\\
-sine & splines & 8 & 0.05 & 0.29 & 2.74 & 0.00 & 1.02 & 1.06 & 2.68 & -0.02 & 0.54\\
-sine & stringy2 & 4 & 0.02 & 0.30 & 2.76 & 1.01 & 0.64 & 0.05 & 7.89 & 0.05 & 0.00\\
-sine & stringy2 & 6 & 0.02 & 0.07 & 1.51 & 0.05 & 0.65 & 0.05 & 7.78 & 0.06 & 0.00\\
+sine & splines & 4 & 3.5489 & 0.5723\\
+sine & splines & 6 & 3.1149 & 0.5415\\
+sine & splines & 8 & 2.7400 & 0.5394\\
+sine & TIC & 4 & 3.4854 & 0.3839\\
+sine & TIC & 8 & 2.7223 & 0.3785\\
+sine & TIC & 6 & 3.0807 & 0.3747\\
+sine & skinny & 6 & 2.2352 & 0.0973\\
 \bottomrule
 \end{tabular}
-\endgroup{}
 

### 2025-03-02 supplementary/index-def.qmd
trimmed down table 1

@@
 
+# Gaussian process estimates for all PP indexes and data sets
+
+```{r}
+#| label: tbl-smoothness-squintability
+#| tbl-cap: Parameters estimated from the Gaussian process (outputscale $\eta$, lengthscale $\ell$, smoothness $\nu$, and nugget $\sigma$) and scaled logistic function ($\theta_1$ to $\theta_4$ and $\varsigma$) for the pipe-finding and sine-wave finding problems.
+The columns $\nu$ and $\varsigma$ represent the smoothness and squintability measures respectively.
+dt <- tibble(shape = c(rep("pipe", 5), rep("sine", 17)), smoothness) |>
+  left_join(squintability) |> 
+  rename(d = n, smooth = smoothness) |> 
+  mutate(index = ifelse(index == "dcor2d_2", "dcor", index),
+         index = ifelse(index == "loess2d", "loess", index),
+         index = ifelse(index == "splines2d", "splines", index),
+         variance = sqrt(variance)) 
+dt |> 
+  knitr::kable(
+    digits = 2, format = "latex", 
+    col.names = c(
+      "shape", "index", "d",
+      "$\\eta$", "$\\ell$", "$\\nu$", "$\\sigma$",
+      "$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", "$\\varsigma$"), linesep = "",
+    booktabs = TRUE, escape = FALSE, row.names = FALSE, align = "|llr|rrrr|rrrrr") |> 
+  #column_spec(1, border_left = TRUE) |> 
+  #column_spec(c(4, 8, 13), border_right = TRUE) |> 
+  kable_styling(font_size = 10) 
+```
+
 # Reference

### 2025-03-02 supplementary/supplementary.qmd
more clean up script and the save sample trace data

--- 2025-03-02 supplementary/index-def.qmd
@@
 library(patchwork)
+library(kableExtra)
 load(here::here("data/sim_pipe_run_best.rda"))

### 2025-03-02 code/07-calc-smoothness-squintability.R

@@ -79,7 +79,7 @@ save(sq_holes_basis_df, file = here::here("data/sq_holes_basis_df.rda"))
 
 idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "splines2d", "stringy2", "skinny")
 # about 40 mins
-sq_sine_basis_df <- tibble::tibble(
+sq_sine_68s_basis_df <- tibble::tibble(
   n = 6, index = idx_names, data = list(sine1000),
   best = list(matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE))) |>
   dplyr::bind_rows(sine_8d_tbl("MIC")) |>
@@ -106,7 +106,7 @@ sq_4d_basis_df <- tibble::tibble(
     step_size = 0.005, best = best, parallel = TRUE)
   ))
 
-sq_sine_basis_df <- bind_rows(sq_sine_basis_df, sq_4d_basis_df) |> arrange(index, n)
+sq_sine_basis_df <- bind_rows(sq_sine_68s_basis_df, sq_4d_basis_df) |> arrange(index, n)
 save(sq_sine_basis_df, file = here::here("data/sq_sine_basis_df.rda"))
 
 # calculate squintability
@@ -116,7 +116,7 @@ res_holes <- sq_holes_basis_df |>
     basis_df, method = "nls", bin_width = 0.005, scale = TRUE,
     other_params = list(start = c(theta1 = 1, theta2 = 1, theta3 = 3, theta4 = 0)))) |>
   unnest(res) |>
-  select(index, n, theta1: squint)
+  select(index, n, theta1:squint)
 
 param_tbl <- tibble(other_params =  c(
   rep(list(start = list(theta1 = 1, theta2 = 1, theta3 = 2, theta4 = 0)), 11),
@@ -150,24 +150,18 @@ smoothness |> select(n, index, smoothness) |>
 ############################################################################
 ############################################################################
 # TODO
-sq_basis_dist_idx <- bind_rows(sq_holes_basis_df, sq_sine_basis_df) |>
+sq_basis_df_6d_TIC_spline_skinny <- sq_sine_basis_df |>
+  filter(index %in% c("TIC", "splines2d", "skinny"),n  == 6) |>
   dplyr::select(basis_df, n, index) |>
   rename(index_name = index) |>
   unnest(basis_df) |>
-  mutate(dist = ceiling(dist / 0.005) * 0.005,
[…]

### 2025-03-02 jso.qmd
new trace plot on the TIC, skinny, and splines index

@@
 library(ferrn)
+library(gtsummary)
 load(here::here("data/sim_pipe_run_best.rda"))
@@
 load(here::here("data/squintability.rda"))
-load(here::here("data/sq_basis_dist_idx.rda"))
+load(here::here("data/sq_basis_df_6d_TIC_spline_skinny.rda"))
 load(here::here("data/sim_df.rda"))
@@
 #| label: fig-idx-proj-dist
-#| fig.width: 9
-#| fig.height: 5
-#| eval: false
-#| fig.cap: "Index values versus projection distance for the 12 pipe/sine-wave finding problem, after the binning procedure for calculating the squintability measure.
-The index values, averaged at bin width of 0.005, are scaled from 0 to 1 for comparison (holes, TIC, and stringy).
-The `MIC` and `TIC` index curves are convex while others are concave.
-The stringy curve shows an instantaneous jump to the optimum when aproaching the best basis."
-sq_basis_dist_idx |>
-  ggplot(aes(x = dist, y = index, group = index_name)) +
+#| fig.width: 8
+#| fig.height: 4
+#| fig.cap: "Five traces of index value against projection distance for each of the three indexes (TIC, spline, and skinny) in 6D sine data.
+The traces of the splines index are smooth, with a steep increaes before leveling off, suggesting the optimum can be detected from a far distance, hence a largest squintability value.
+In contrast, the skinny index (squintability of almost zeor) has a noisy trace that only stabilizes near the optimum, suggesting the optimizer needs get very close to detect it."
+sq_basis_df_6d_TIC_spline_skinny |>
+  ggplot(aes(x = dist, y = index, group = interaction(index_name, id), color = index_name)) +
   geom_line() + 
-  facet_nested_wrap(~index_name + n, nrow = 3, labeller = label_both) + 
-  xlab("projection distance") + ylab("index value") +
-  theme_bw()
+  scale_color_brewer(palette = "Dark2") +
+  xlab("Projection distance") + ylab("Index value") +
+  theme_bw() + 
+  theme(legend.position = "bottom")
 ```
@@
 #| fig-pos: t
-library(gtsummary)
 set_gtsummary_theme(theme_gtsummary_journal("jama"))
@@
 #| label: tbl-joint-output
-library(kableExtra)
 summary_table <- #bind_rows(

### 2025-03-02 jso.tex
new trace plot on the TIC, skinny, and splines index

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, bottomrule=.15mm, toprule=.15mm, colback=white, colframe=quarto-callout-note-color-frame, leftrule=.75mm, coltitle=black, breakable, left=2mm, opacityback=0, bottomtitle=1mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, coltitle=black, breakable, opacityback=0, rightrule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, toprule=.15mm, toptitle=1mm, bottomtitle=1mm, arc=.35mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, bottomrule=.15mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm]
 
@@
 
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-idx-proj-dist-1.pdf}}
+
+}
+
+\caption{\label{fig-idx-proj-dist}Five traces of index value against
+projection distance for each of the three indexes (TIC, spline, and
+skinny) in 6D sine data. The traces of the splines index are smooth,
+with a steep increaes before leveling off, suggesting the optimum can be
+detected from a far distance, hence a largest squintability value. In
+contrast, the skinny index (squintability of almost zeor) has a noisy
+trace that only stabilizes near the optimum, suggesting the optimizer
+needs get very close to detect it.}
+
+\end{figure}%
+
 \begin{table}

### 2025-03-03 jso.qmd
fix references

@@
 
-The R packages used in this work include: `tidyverse` [@tidyverse];
-`patchwork` [@patchwork]; `ggh4x` [@ggh4x]; `broom` [@broom]; 
-`kableExtra` [@kabelextra]; `ferrn` [@RJ-2021-105]; and `cassowaryr` [@cassowaryr].
+The R packages used in this work include: `tidyr` [@tidyr], `dplyr` [@dplyr], 
+`ggplot2` [@ggplot2], `knitr` [@knitr],`gtsummary` [@gtsummary],
+`patchwork` [@patchwork], `ggh4x` [@ggh4x], `broom` [@broom], 
+`kableExtra` [@kabelextra], `ferrn` [@RJ-2021-105], and `cassowaryr` [@cassowaryr].
 

### 2025-03-03 jso.tex
fix references

@@
 of the optimum. An implementation is available in the \texttt{tourr}
-package (H. Wickham et al. 2011) in R (R Core Team 2023). Zhang et al.
+package (Wickham et al. 2011) in R (R Core Team 2023). Zhang et al.
 (2021) illustrated how to diagnose optimisation processes, particularly
@@
 (1985), Cook, Buja, and Cabrera (1993), Lee et al. (2005), Lee and Cook
-(2010), Grimm (2016), Ursula Laa and Valencia (2022). Most are
-relatively simply defined, for any projection dimension, and implemented
-because they are relatively easy to optimise. A goal is to develop PP
-indexes based on scagnostics Leland Wilkinson and Wills (2008), but the
-blockage is their optimisation as these tend to be noisy, with
-potentially small squint angles.
+(2010), Grimm (2016), Laa et al. (2022). Most are relatively simply
+defined, for any projection dimension, and implemented because they are
+relatively easy to optimise. A goal is to develop PP indexes based on
+scagnostics Wilkinson and Wills (2008), but the blockage is their
+optimisation as these tend to be noisy, with potentially small squint
+angles.
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, coltitle=black, breakable, opacityback=0, rightrule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, toprule=.15mm, toptitle=1mm, bottomtitle=1mm, arc=.35mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, bottomrule=.15mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm]
+\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, colback=white, breakable, left=2mm, bottomtitle=1mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, rightrule=.15mm, toptitle=1mm, opacityback=0, bottomrule=.15mm, arc=.35mm, leftrule=.75mm]
 
@@
 
-The R packages used in this work include: \texttt{tidyverse} (Hadley
-Wickham et al. 2019); \texttt{patchwork} (Pedersen 2024); \texttt{ggh4x}
-(van den Brand 2024); \texttt{broom} (Robinson, Hayes, and Couch 2024);
-\texttt{kableExtra} (Zhu 2024); \texttt{ferrn} (Zhang et al. 2021); and
-\texttt{cassowaryr} (Mason et al. 2022).
+The R packages used in this work include: \texttt{tidyr} (Wickham,
+Vaughan, and Girlich 2024), \texttt{dplyr} (Wickham et al. 2023),
+\texttt{ggplot2} (Wickham 2016), \texttt{knitr} (Xie
+2014),\texttt{gtsummary} (Sjoberg et al. 2021), \texttt{patchwork}
+(Pedersen 2024), \texttt{ggh4x} (Van Den Brand 2024), \texttt{broom}
+(Robinson, Hayes, and Couch 2024), \texttt{kableExtra} (Zhu 2024),
+\texttt{ferrn} (Zhang et al. 2021), and \texttt{cassowaryr} (Mason et
+al. 2022).
 
@@
 \bibitem[\citeproctext]{ref-abdulah2023large}
-Abdulah, Sameh, Yuxiao Li, Jian Cao, Hatem Ltaief, David Keyes, Marc
-Genton, and Ying Sun. 2023. {``Large-Scale Environmental Data Science
-with {ExaGeoStatR}.''} \emph{Environmetrics} 34 (1): e2770.
+Abdulah, S., Y. Li, J. Cao, H. Ltaief, D. Keyes, M. Genton, and Y. Sun.
+2023. {``Large-Scale Environmental Data Science with {ExaGeoStatR}.''}
+\emph{Environmetrics} 34 (1): e2770.
 \url{https://doi.org/10.1002/env.2770}.
@@
 \bibitem[\citeproctext]{ref-adams2003sobolev}
-Adams, Robert, and John Fournier. 2003. \emph{Sobolev Spaces}. Vol. 140.
-Pure and Applied Mathematics. Elsevier.
+Adams, R., and J. Fournier. 2003. \emph{Sobolev Spaces}. Vol. 140. Pure
+and Applied Mathematics. Elsevier.
 
 \bibitem[\citeproctext]{ref-Allaire_Quarto_2022}
-Allaire, J. J., Charles Teague, Carlos Scheidegger, Yihui Xie, and
-Christophe Dervieux. 2022. \emph{{Quarto}} (version 1.2).
+Allaire, J. J., C. Teague, C. Scheidegger, Y. Xie, and C. Dervieux.
+2022. \emph{{Quarto}} (version 1.2).
 \url{https://doi.org/10.5281/zenodo.5960048}.
@@
 \bibitem[\citeproctext]{ref-Bertsimas93}
-Bertsimas, Dimitris, and John Tsitsiklis. 1993. {``{Simulated
-Annealing}.''} \emph{Statistical Science} 8 (1): 10--15.
+Bertsimas, D., and J. Tsitsiklis. 1993. {``Simulated Annealing.''}
+\emph{Statistical Science} 8 (1): 10--15.
 \url{https://doi.org/10.1214/ss/1177011077}.
@@
 \bibitem[\citeproctext]{ref-chou_novel_2021}
-Chou, Jui-Sheng, and Dinh-Nhat Truong. 2021. {``A Novel Metaheuristic
-Optimizer Inspired by Behavior of Jellyfish in Ocean.''} \emph{Applied
-Mathematics and Computation} 389 (January): 125535.
+Chou, J.-S., and D.-N. Truong. 2021. {``A Novel Metaheuristic Optimizer
+Inspired by Behavior of Jellyfish in Ocean.''} \emph{Applied Mathematics
+and Computation} 389 (January): 125535.
 \url{https://doi.org/10.1016/j.amc.2020.125535}.
@@
 \bibitem[\citeproctext]{ref-Grimm2016}
-Grimm, Katrin. 2016. {``Kennzahlenbasierte Grafikauswahl.''} Doctoral
-thesis, Universit{ä}t Augsburg.
+Grimm, K. 2016. {``Kennzahlenbasierte Grafikauswahl.''} Doctoral thesis,
+Universität Augsburg.
 
 \bibitem[\citeproctext]{ref-grochowski2011}
-Grochowski, Marek, and Włodzisław Duch. 2011. {``Fast Projection Pursuit
-Based on Quality of Projected Clusters.''} In \emph{Adaptive and Natural
-Computing Algorithms}, edited by Andrej Dobnikar, Uroš Lotrič, and
-Branko Šter, 89--97. Berlin, Heidelberg: Springer Berlin Heidelberg.
+Grochowski, M., and W. Duch. 2011. {``Fast Projection Pursuit Based on
+Quality of Projected Clusters.''} In \emph{Adaptive and Natural
+Computing Algorithms}, edited by A. Dobnikar, U. Lotrič, and B. Šter,
+89--97. Berlin, Heidelberg: Springer Berlin Heidelberg.
 \url{https://doi.org/10.1007/978-3-642-20267-4_10}.
@@
 \bibitem[\citeproctext]{ref-guinness2021gpgp}
-Guinness, Joseph, Matthias Katzfuss, and Youssef Fahmy. 2021. {``{GpGp}:
-Fast {G}aussian Process Computation Using {V}ecchia's Approximation.''}
-R package.
+Guinness, J., M. Katzfuss, and Y. Fahmy. 2021. {``{GpGp}: Fast
+{G}aussian Process Computation Using {V}ecchia's Approximation.''} R
+package. \url{https://cran.r-project.org/package=GpGp}.
 
 \bibitem[\citeproctext]{ref-hall1989polynomial}
-Hall, Peter. 1989. {``On Polynomial-Based Projection Indices for
-Exploratory Projection Pursuit.''} \emph{The Annals of Statistics} 17
-(2): 589--605. \url{https://doi.org/10.1214/aos/1176347127}.
+Hall, P. 1989. {``On Polynomial-Based Projection Indices for Exploratory
+Projection Pursuit.''} \emph{The Annals of Statistics} 17 (2): 589--605.
+\url{https://doi.org/10.1214/aos/1176347127}.
 
 \bibitem[\citeproctext]{ref-huber85}
-Huber, Peter J. 1985. {``Projection Pursuit.''} \emph{The Annals of
+Huber, P. J. 1985. {``Projection Pursuit.''} \emph{The Annals of
 Statistics} 13 (2): 435--75.
@@
 \bibitem[\citeproctext]{ref-karvonen2023asymptotic}
-Karvonen, Toni. 2023. {``Asymptotic Bounds for Smoothness Parameter
+Karvonen, T. 2023. {``Asymptotic Bounds for Smoothness Parameter
 Estimates in {G}aussian Process Interpolation.''} \emph{SIAM/ASA Journal
@@
 \bibitem[\citeproctext]{ref-laa_using_2020}
-Laa, Ursula, and Dianne Cook. 2020. {``Using Tours to Visually
-Investigate Properties of New Projection Pursuit Indexes with
-Application to Problems in Physics.''} \emph{Computational Statistics}
-35 (3): 1171--1205. \url{https://doi.org/10.1007/s00180-020-00954-8}.
+Laa, U., and D. Cook. 2020. {``Using Tours to Visually Investigate
+Properties of New Projection Pursuit Indexes with Application to
+Problems in Physics.''} \emph{Computational Statistics} 35 (3):
+1171--1205. \url{https://doi.org/10.1007/s00180-020-00954-8}.
+
+\bibitem[\citeproctext]{ref-Laa:2020wkm}
+Laa, U., D. Cook, A. Buja, and G. Valencia. 2022. {``Hole or Grain? A
+Section Pursuit Index for Finding Hidden Structure in Multiple
+Dimensions.''} \emph{Journal of Computational and Graphical Statistics}
+31 (3): 739--52. \url{https://doi.org/10.1080/10618600.2022.2035230}.
 
 \bibitem[\citeproctext]{ref-lee2010projection}
-Lee, Eun-Kyung, and Dianne Cook. 2010. {``A Projection Pursuit Index for
-Large {\(p\)} Small {\(n\)} Data.''} \emph{Statistics and Computing} 20
-(3): 381--92. \url{https://doi.org/10.1007/s11222-009-9131-1}.
+Lee, E.-K., and D. Cook. 2010. {``A Projection Pursuit Index for Large
+\(p\) Small \(n\) Data.''} \emph{Statistics and Computing} 20 (3):
+381--92. \url{https://doi.org/10.1007/s11222-009-9131-1}.
 
 \bibitem[\citeproctext]{ref-lee2005projection}
-Lee, Eun-Kyung, Dianne Cook, Sigbert Klinke, and Thomas Lumley. 2005.
-{``Projection Pursuit for Exploratory Supervised Classification.''}
-\emph{Journal of Computational and Graphical Statistics} 14 (4):
-831--46. \url{https://doi.org/10.1198/106186005X77702}.
+Lee, E.-K., D. Cook, S. Klinke, and T. Lumley. 2005. {``Projection
+Pursuit for Exploratory Supervised Classification.''} \emph{Journal of
+Computational and Graphical Statistics} 14 (4): 831--46.
+\url{https://doi.org/10.1198/106186005X77702}.
 
 \bibitem[\citeproctext]{ref-Loperfido2018}
-Loperfido, Nicola. 2018. {``Skewness-Based Projection Pursuit: A
+Loperfido, N. 2018. {``Skewness-Based Projection Pursuit: A
 Computational Approach.''} \emph{Computational Statistics and Data
@@
 \bibitem[\citeproctext]{ref-marie-sainte2010}
-Marie-Sainte, Souad Larabi, Alain Berro, and Anne Ruiz-Gazen. 2010.
-{``An Efficient Optimization Method for Revealing Local Optima of
-Projection Pursuit Indices.''} In \emph{Swarm Intelligence}, edited by
-Marco Dorigo, Mauro Birattari, Gianni A. Di Caro, René Doursat, Andries
-P Engelbrecht, Dario Floreano, Luca Maria Gambardella, et al., 60--71.
-Berlin, Heidelberg: Springer Berlin Heidelberg.
-\url{https://doi.org/10.1007/978-3-642-15461-4_6}.
+Marie-Sainte, S. L., A. Berro, and A. Ruiz-Gazen. 2010. {``An Efficient
+Optimization Method for Revealing Local Optima of Projection Pursuit
+Indices.''} In \emph{Swarm Intelligence}, edited by M. Dorigo, M.
+Birattari, G. A. Di Caro, R. Doursat, A. P. Engelbrecht, D. Floreano, L.
+M. Gambardella, et al., 60--71. Berlin, Heidelberg: Springer Berlin
+Heidelberg. \url{https://doi.org/10.1007/978-3-642-15461-4_6}.
 
 \bibitem[\citeproctext]{ref-cassowaryr}
-Mason, Harriet, Stuart Lee, Ursula Laa, and Dianne Cook. 2022.
-\emph{Cassowaryr: Compute Scagnostics on Pairs of Numeric Variables in a
-Data Set}. \url{https://CRAN.R-project.org/package=cassowayr}.
+Mason, H., S. Lee, U. Laa, and D. Cook. 2022. \emph{{c}assowaryr:
+Compute Scagnostics on Pairs of Numeric Variables in a Data Set}.
+\url{https://CRAN.R-project.org/package=cassowayr}.
 
 \bibitem[\citeproctext]{ref-patchwork}
-Pedersen, Thomas Lin. 2024. \emph{Patchwork: The Composer of Plots}.
+Pedersen, T. L. 2024. \emph{{p}atchwork: The Composer of Plots}.
 \url{https://CRAN.R-project.org/package=patchwork}.
@@
 \bibitem[\citeproctext]{ref-porcu2024matern}
-Porcu, Emilio, Moreno Bevilacqua, Robert Schaback, and Chris Oates.
-2024. {``The {M}atérn Model: A Journey Through Statistics, Numerical
-Analysis and Machine Learning.''} \emph{Statistical Science} 39 (3):
-469--92. \url{https://doi.org/10.1214/24-STS923}.
+Porcu, E., M. Bevilacqua, R. Schaback, and C. Oates. 2024. {``The
+{M}atérn Model: A Journey Through Statistics, Numerical Analysis and
+Machine Learning.''} \emph{Statistical Science} 39 (3): 469--92.
+\url{https://doi.org/10.1214/24-STS923}.
 
 \bibitem[\citeproctext]{ref-posse95}
-Posse, Christian. 1995. {``Projection Pursuit Exploratory Data
-Analysis.''} \emph{Computational Statistics and Data Analysis} 20 (6):
-669--87. \url{https://doi.org/10.1016/0167-9473(95)00002-8}.
+Posse, C. 1995. {``Projection Pursuit Exploratory Data Analysis.''}
+\emph{Computational Statistics and Data Analysis} 20 (6): 669--87.
+\url{https://doi.org/10.1016/0167-9473(95)00002-8}.
 
@@
 \bibitem[\citeproctext]{ref-rajwar_exhaustive_2023}
-Rajwar, Kanchan, Kusum Deep, and Swagatam Das. 2023. {``An Exhaustive
-Review of the Metaheuristic Algorithms for Search and Optimization:
-Taxonomy, Applications, and Open Challenges.''} \emph{Artificial
-Intelligence Review}, 1--71.
-\url{https://doi.org/10.1007/s10462-023-10470-y}.
+Rajwar, K., K. Deep, and S. Das. 2023. {``An Exhaustive Review of the
+Metaheuristic Algorithms for Search and Optimization: Taxonomy,
+Applications, and Open Challenges.''} \emph{Artificial Intelligence
+Review}, 1--71. \url{https://doi.org/10.1007/s10462-023-10470-y}.
 
 \bibitem[\citeproctext]{ref-rasmussen2006gaussian}
-Rasmussen, Carl Edward, and Christopher K. I. Williams. 2006.
-\emph{Gaussian Processes for Machine Learning}. The MIT Press.
+Rasmussen, C. E., and C. K. I. Williams. 2006. \emph{Gaussian Processes
+for Machine Learning}. The MIT Press.
 
 \bibitem[\citeproctext]{ref-broom}
-Robinson, David, Alex Hayes, and Simon Couch. 2024. \emph{Broom: Convert
+Robinson, D., A. Hayes, and S. Couch. 2024. \emph{{b}room: Convert
 Statistical Objects into Tidy Tibbles}.
 \url{https://CRAN.R-project.org/package=broom}.
+
+\bibitem[\citeproctext]{ref-gtsummary}
+Sjoberg, D. D., K. Whiting, M. Curry, J. A. Lavery, and J. Larmarange.
+2021. {``Reproducible Summary Tables with the {g}tsummary Package.''}
+\emph{{The R Journal}} 13: 570--80.
+\url{https://doi.org/10.32614/RJ-2021-053}.
 
@@
 
-\bibitem[\citeproctext]{ref-Laa:2020wkm}
-Ursula Laa, Andreas Buja, Dianne Cook, and German Valencia. 2022.
-{``Hole or Grain? A Section Pursuit Index for Finding Hidden Structure
-in Multiple Dimensions.''} \emph{Journal of Computational and Graphical
-Statistics} 31 (3): 739--52.
-\url{https://doi.org/10.1080/10618600.2022.2035230}.
-
 \bibitem[\citeproctext]{ref-ggh4x}
-van den Brand, Teun. 2024. \emph{Ggh4x: Hacks for 'Ggplot2'}.
+Van Den Brand, T. 2024. \emph{{g}gh4x: Hacks for {ggplot2}}.
 \url{https://CRAN.R-project.org/package=ggh4x}.
 
-\bibitem[\citeproctext]{ref-tidyverse}
-Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
-D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
-{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
-4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.
+\bibitem[\citeproctext]{ref-ggplot2}
+Wickham, H. 2016. \emph{{g}gplot2: Elegant Graphics for Data Analysis}.
+Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
 
 \bibitem[\citeproctext]{ref-tourr}
-Wickham, H., D. Cook, H. Hofmann, and A. Buja. 2011. {``{tourr}: An {R}
+Wickham, H., D. Cook, H. Hofmann, and A. Buja. 2011. {``{t}ourr: An r
 Package for Exploring Multivariate Data with Projections.''}
@@
 \url{https://doi.org/10.18637/jss.v040.i02}.
+
+\bibitem[\citeproctext]{ref-dplyr}
+Wickham, H., R. François, L. Henry, K. Müller, and D. Vaughan. 2023.
+\emph{{d}plyr: A Grammar of Data Manipulation}.
+\url{https://CRAN.R-project.org/package=dplyr}.
+
+\bibitem[\citeproctext]{ref-tidyr}
+Wickham, H., D. Vaughan, and M. Girlich. 2024. \emph{{t}idyr: Tidy Messy
+Data}. \url{https://CRAN.R-project.org/package=tidyr}.
 
@@
 \bibitem[\citeproctext]{ref-WW08}
-Wilkinson, Leland, and Graham Wills. 2008. {``Scagnostics
-Distributions.''} \emph{Journal of Computational and Graphical
-Statistics} 17 (2): 473--91.
-\url{https://doi.org/10.1198/106186008X320465}.
+Wilkinson, L., and G. Wills. 2008. {``Scagnostics Distributions.''}
+\emph{Journal of Computational and Graphical Statistics} 17 (2):
+473--91. \url{https://doi.org/10.1198/106186008X320465}.
+
+\bibitem[\citeproctext]{ref-knitr}
+Xie, Y. 2014. {``{k}nitr: A Comprehensive Tool for Reproducible Research
+in {R}.''} In \emph{Implementing Reproducible Computational Research},
+edited by V. Stodden, F. Leisch, and R. D. Peng. Chapman; Hall/CRC.
 
 \bibitem[\citeproctext]{ref-RJ-2021-105}
-Zhang, H. Sherry, Dianne Cook, Ursula Laa, Nicolas Langrené, and
-Patricia Menéndez. 2021. {``Visual Diagnostics for Constrained
-Optimisation with Application to Guided Tours.''} \emph{The R Journal}
-13: 624--41. \url{https://doi.org/10.32614/RJ-2021-105}.
+Zhang, H. S., D. Cook, U. Laa, N. Langrené, and P. Menéndez. 2021.
+{``Visual Diagnostics for Constrained Optimisation with Application to
+Guided Tours.''} \emph{The R Journal} 13: 624--41.
+\url{https://doi.org/10.32614/RJ-2021-105}.
 
 \bibitem[\citeproctext]{ref-kabelextra}
-Zhu, Hao. 2024. \emph{kableExtra: Construct Complex Table with 'Kable'
+Zhu, H. 2024. \emph{{k}ableExtra: Construct Complex Table with {kable}
 and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

### 2025-03-03 jso.qmd
done with proofing

@@
 
-The second simulation examines various factors affecting the performance of JSO.
-The JSO performance is evaluated under different hyper-parameters combinations
-(number of jellyfish and the maximum number of tries). The effect of index properties (smoothness and squintability) along
-with JSO hyper-parameters and data dimension variations is captured and studied using logistic regression.
-
 ## Performance of JSO relative to CRS {#sec-app-1}
@@
 
-To examine the performance of JSO across various hyper-parameter
-combinations, the pipe-finding problem using the holes index is repeated
-for fifty times in each hyper-parameter combination: 20, 50, and 100
-jellyfish and a maximum of 50 and 100 attempts.
-
-In addition, to assess the performance of JSO across various PP problems, two different data shapes, pipe and a sine wave, are considered in various dimensional spaces using six different PP indexes: `dcor2d_2`, `loess2d`, `MIC`, `TIC`,
-`spline`, and `stringy`, under varied JSO hyper-parameters.
-An example of holes and pipes data with the syntax to generate them are provided in the supplementary material.
-This results in a total of `r nrow(sim_df)` scenarios, comprising of
-`r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and
-`r nrow(filter(sim_df, index != "holes"))` on the sine-wave data.
-This study does not consider dimensions lower than four, as exploring the search space in such cases is relatively straightforward.
-Similarly, extremely high-dimensional cases with complex data structures, where no method can reliably and consistenly uncover the underlying structure due to the curse of dimensionality, are also excluded, as they fall outside the scope of this study.
-A list of combinations of the data shape, dimensions and PP indexes that are considered in this study is provided in @tbl-smoothness-squintability.
-Again, JSO is run 50 times for each scenario to calculate the success rate for each hyper-parameter combination.
-
-Smoothness and squintability are computed following the procedures
-outlined in @sec-smoothness and @sec-squintability and as illustrated in
-@fig-smoothness and @fig-squintability. To compute smoothness, 500
-random bases are simulated. Index values are calculated for each random
-basis, followed by fitting a Gaussian process model to obtain the
-smoothness measure for the index.
-
-To compute squintability, 50 random bases are sampled and interpolated
-to the optimal basis with a step size of 0.005. Index values and
-projection distances are calculated for these interpolated bases. A
-binning procedure is applied to average the index values within each
-0.005 projection distance bin. The four-parameter scaled logistic
-function (@eq-parametric) is fitted to the index values against
-projection distances, estimated by non-linear least squares to obtain
-the squintability measure as Equation (@eq-squintability-parametric).
-
-A generalised linear model is fitted using a binomial family and a logit link function to assess the factors affecting success rate.  Predictors are smoothness,
-squintability, and JSO hyper-parameters. A success rate of 0 indicates the optimiser did not come sufficiently close to the optimal projection. <!-- The data is
-pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
-for interpretation, 2) creating a new binary variable `long_time` to
-indicate cases with an average run time over 20 seconds, and 3)
-re-coding the success rate for the `stringy` index as 0, because none of
-the 50 simulations correctly identified the sine-wave shape.-->
+To examine the performance of JSO relative to hyper-parameter choices, the `holes` index is used to find the 2D pipe in data of dimensions 4-12.
+Additional variables are Gaussian noise.
+The levels of hyper-parameters are 20, 50, and 100
+jellyfish and a maximum of 50 and 100 attempts. Fifty samples of each data set are generated, and the success rate is calculated in the same manner as described in @sec-app-1.
+
+This set of simulations is expanded to assess the performance of JSO relative to the index properties, with a second data set type and additional PP indexes.
+The second data set has a sine wave in two dimensions and Gaussian noise in the remaining dimensions.
+Six additional PP indexes that should be able to extract the sine wave are used: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`, following @laa_using_2020.
+(The supplementary material includes definitions of these PP indexes, and descriptions of the pipe and sine wave data simulation.).
+The JSO hyper-parameters, number of jellies and maximum tries, are also included.
+This results in a total of `r nrow(sim_df)` scenarios, comprising of `r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and `r nrow(filter(sim_df, index != "holes"))` on the sine wave data.
+The choice of 4-12 dimensions covers the range of easy to relatively difficult for the optimisers to detect the structured projections.
+<!-- A list of combinations of the data shape, dimensions and PP indexes that are considered in this study is provided in @tbl-smoothness-squintability.--> The same approach is used, that 50 samples are generated and the success rate is calculated at each level of the simulation design.
+A success rate of 0 indicates the optimiser did not come sufficiently close to the optimal projection.
 
@@
 ```
+
+@fig-smoothness describes the procedure to compute smoothness.
+Index values are calculated for each of 500 simulated random bases.
+A Gaussian process model is fitted to the resulting data to obtain the smoothness measure for the index, as described in @sec-smoothness.
 
@@
 
+@fig-squintability describes the procedure to compute squintability. Index values and
+projection distances are computed for the projections along interpolated paths between 50 random bases and the optimal basis.
+An interpolation step size of 0.005 is used.
+The projection distances are binned to 0.005 intervals and index values are averaged in each bin.
+The four-parameter scaled logistic function (@eq-parametric) is fitted to the index values against projection distances, estimated by non-linear least squares to obtain the squintability measure as Equation (@eq-squintability-parametric).
+
+A generalised linear model is fitted using a quasibinomial family and a logit link function to assess the factors affecting success rate.  Predictors are smoothness,
+squintability, data dimension and JSO hyper-parameters. <!-- The data is
+pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
+for interpretation, 2) creating a new binary variable `long_time` to
+indicate cases with an average run time over 20 seconds, and 3)
+re-coding the success rate for the `stringy` index as 0, because none of
+the 50 simulations correctly identified the sine wave shape.-->
+
 # Results {#sec-sim-res}
@@
 The effect of JSO hyper-parameters (number of jellyfish and the maximum
-number of tries) on the success rate is presented in @fig-proportion.
+number of tries) on the success rate is summarised in @fig-proportion.
 The uncertainty is quantified through 500 bootstrap replicates for each
@@
 rate also increases.
-For problems with relatively lower dimensional search spaces (4 and 6 dimensions), small parameter values (20 jellyfish and a maximum of 50 tries) are sufficient for space exploration and thus can already achieve a high success rate.
-Higher parameter values (i.e.
-100 jellyfish and a maximum of 100 tries) enhances exploration of the search space, which is particularly beneficial in higher-dimensional problems (8, 10, and 12 dimensions) where a more complex search is necessary.
+For problems with relatively lower dimensional (4, 6) search spaces, small parameter values (20 jellyfish and a maximum of 50 tries) are sufficient to achieve a high success rate.
+Many more jellyfish (100 jellyfish and a maximum of 100 tries) are needed to obtain a mediocre success rate for dimensions 10 and 12.
+This suggests that even JSO will struggle with high dimensions, and require substantial more computational time to have an acceptable success rate, especially for difficult PP indexes based on scagnostics.
+<!-- dimensions) where a more complex search is necessary.
 However, in lower-dimensional problems, the benefit may be less pronounced, as the space is already relatively easy to navigate.
@@
 computationally intensive when evaluating the index function (such as
-scagnostic indexes) multiple times across numerous iterations.
+scagnostic indexes) multiple times across numerous iterations.-->
 
@@
 Smoothness and squintability are calculated across a collection of
-pipe-finding and sine-wave finding problems to construct the
+pipe and sine wave finding problems to construct the
 relationship between success rate, JSO hyper-parameters, and index
-properties. @tbl-smoothness-squintability presents the parameters
+properties. @tbl-smoothness-squintability presents values computed for a subset of PP indexes on the sine wave data for dimensions 4-8. (A full table of the parameters
 estimated from the Gaussian process (outputscale $\eta$, lengthscale
@@
 logistic function ($\theta_1$ to $\theta_4$) for calculating smoothness
-and squintability in each PP problem considered. The column $\nu$ is
+and squintability for all PP indexes and data combinations is included in the supplementary materials). The column $\nu$ is
 used as the smoothness metric and the column $\varsigma$ is calculated
@@
 
-```{r}
-#| label: fit-logistic
-sim_df2 <- sim_df |>   
-  mutate(n_jellies = n_jellies/10, max_tries = max_tries/10) |> 
-  rename(proportion = P_J) #|>
-  #filter(d < 10)
-
-mod1 <- glm(proportion ~ smoothness + squintability + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial) 
-mod2 <- glm(proportion ~ squintability + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
-mod3 <- glm(proportion ~ smoothness + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
-mod4 <- glm(proportion ~ squintability*d + n_jellies, data = sim_df2, family = quasibinomial)
-mod5 <- glm(proportion ~ smoothness*d + squintability*d + n_jellies, data = sim_df2, family = quasibinomial)
-```
-
-```{r}
-#| label: fig-idx-proj-dist
-#| fig.width: 8
-#| fig.height: 4
-#| fig.cap: "Five traces of index value against projection distance for each of the three indexes (TIC, spline, and skinny) in 6D sine data.
-The traces of the splines index are smooth, with a steep increaes before leveling off, suggesting the optimum can be detected from a far distance, hence a largest squintability value.
-In contrast, the skinny index (squintability of almost zeor) has a noisy trace that only stabilizes near the optimum, suggesting the optimizer needs get very close to detect it."
-sq_basis_df_6d_TIC_spline_skinny |>
-  ggplot(aes(x = dist, y = index, group = interaction(index_name, id), color = index_name)) +
-  geom_line() + 
-  scale_color_brewer(palette = "Dark2") +
-  xlab("Projection distance") + ylab("Index value") +
-  theme_bw() + 
-  theme(legend.position = "bottom")
-```
+@fig-idx-proj-dist shows a sample of traces from the indexes in @tbl-smoothness-squintability for the 6D sine wave data.
+The traces the PP index values computed for projections along an interpolated path between a noise projection and the optimal projection (sine wave).
+(Note index values were self-standardised to range between 0 and 1, for comparison.) The `splines` index is very smooth with a gradual increase in value between the noise projection and the optimal projection.
+The `TIC` index is slightly less smooth and slightly less gradual change in value.
+Least desirable is the `skinny` index which is noisier and the optimisation needs to be much closer to the optimal angle to have a higher value.
+This illustrates how the new measures for squintability and smoothness work to describe the PP indexes.
 
@@
 #| label: tbl-smoothness-squintability
-#| tbl-cap: Parameters estimated from the Gaussian process (outputscale $\eta$, lengthscale $\ell$, smoothness $\nu$, and nugget $\sigma$) and scaled logistic function ($\theta_1$ to $\theta_4$ and $\varsigma$) for the pipe-finding and sine-wave finding problems.
-The columns $\nu$ and $\varsigma$ represent the smoothness and squintability measures respectively.
+#| tbl-cap: The columns $\nu$ and $\varsigma$ show the smoothness and squintability measures for three PP indexes for the sine wave data.
+The splines index has the largest squint angle and is the smoothest.
+The skinny index has low smoothness and smallest squint angle.
+As dimension increases the squint angle gets smaller but smoothness doesn't change much.
 dt <- tibble(shape = c(rep("pipe", 5), rep("sine", 17)), smoothness) |>
@@
 
+```{r}
+#| label: fit-logistic
+sim_df2 <- sim_df |>   
+  mutate(n_jellies = n_jellies/10, max_tries = max_tries/10) |> 
+  rename(proportion = P_J) #|>
+  #filter(d < 10)
+
+mod1 <- glm(proportion ~ smoothness + squintability + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial) 
+mod2 <- glm(proportion ~ squintability + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
+mod3 <- glm(proportion ~ smoothness + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
+mod4 <- glm(proportion ~ squintability*d + n_jellies, data = sim_df2, family = quasibinomial)
+mod5 <- glm(proportion ~ smoothness*d + squintability*d + n_jellies, data = sim_df2, family = quasibinomial)
+```
+
+```{r}
+#| label: fig-idx-proj-dist
+#| fig.width: 8
+#| fig.height: 4
+#| fig.cap: "Five traces of index value against projection distance for each of the three indexes (TIC, spline, and skinny) on 6D sine wave data.
+The traces of the splines index are smooth, with a gradual change between noise and otimal projection.
+In contrast, the skinny index has a noisy trace that increases rapidly near the optimum.
+This illustrates the usefulness of the new measures for squintability and smoothness."
+sq_basis_df_6d_TIC_spline_skinny |>
+  ggplot(aes(x = dist, y = index, group = interaction(index_name, id), color = index_name)) +
+  geom_line() + 
+  scale_color_brewer(palette = "Dark2") +
+  xlab("Projection distance") + ylab("Index value") +
+  theme_bw() + 
+  theme(legend.position = "bottom")
+```
+
 @tbl-mod-output presents the results from fitting a logistic regression
@@
 jellies by 10 increases the success rate by
-`r round(100 * (exp(broom::tidy(mod1)$estimate[5])-1), 0)`%.
+`r round(100 * (exp(broom::tidy(mod1)$estimate[5])-1), 0)`%. @tbl-joint-output shows the fit statistics supporting that the model is reasonably strong.
+
+Note that the addition of interaction terms for squintability and data dimension, and smoothness and data dimension, improves the model a small amount, but does not change the interpretation above.
 
@@
 
-The simulation result suggests that squintability and smoothness are likely dependent.
+<!--The simulation result suggests that squintability and smoothness are likely dependent.
 Intuitively, a highly smooth search space exhibits gradual changes, making it easier to perceive the overall structure from a distance, thereby increasing the squint angle.
@@
 This suggests that squintability captures the relevant variation more effectively than smoothness.
-While it may seem counterintuitive at first that smoothness is not a significant factor, this outcome is reasonable in this context, as the swarm-based method used in this study, JSO, is designed to perform well even in non-smooth search spaces by leveraging population-based exploration.
-
-This study intends to serve as an illustration that the proposed measure effectively captures the key factors influencing the success rate of PP.
-The interactions between the two variables were not examined in this study due to the limited number of data points, nor was the linearity assumption of the relationship with success rate explicitly tested.
+While it may seem counterintuitive at first that smoothness is not a significant factor, this outcome is reasonable in this context, as the swarm-based method used in this study, JSO, is designed to perform well even in non-smooth search spaces by leveraging population-based exploration.-->
+
+This study intends to serve as an illustration that the proposed measures effectively captures the key factors influencing the success rate of PP.
+<!--The interactions between the two variables were not examined in this study due to the limited number of data points, nor was the linearity assumption of the relationship with success rate explicitly tested.-->
 

### 2025-03-03 jso.tex
done with proofing

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, colback=white, breakable, left=2mm, bottomtitle=1mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, rightrule=.15mm, toptitle=1mm, opacityback=0, bottomrule=.15mm, arc=.35mm, leftrule=.75mm]
+\begin{tcolorbox}[enhanced jigsaw, colback=white, opacitybacktitle=0.6, breakable, bottomrule=.15mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, opacityback=0, bottomtitle=1mm, coltitle=black, colframe=quarto-callout-note-color-frame, toptitle=1mm, arc=.35mm, toprule=.15mm, rightrule=.15mm, left=2mm]
 
@@
 affecting the performance of JSO.
-
-The second simulation examines various factors affecting the performance
-of JSO. The JSO performance is evaluated under different
-hyper-parameters combinations (number of jellyfish and the maximum
-number of tries). The effect of index properties (smoothness and
-squintability) along with JSO hyper-parameters and data dimension
-variations is captured and studied using logistic regression.
 
@@
 
-To examine the performance of JSO across various hyper-parameter
-combinations, the pipe-finding problem using the holes index is repeated
-for fifty times in each hyper-parameter combination: 20, 50, and 100
-jellyfish and a maximum of 50 and 100 attempts.
-
-In addition, to assess the performance of JSO across various PP
-problems, two different data shapes, pipe and a sine wave, are
-considered in various dimensional spaces using six different PP indexes:
-\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
-\texttt{spline}, and \texttt{stringy}, under varied JSO
-hyper-parameters. An example of holes and pipes data with the syntax to
-generate them are provided in the supplementary material. This results
-in a total of 76 scenarios, comprising of 30 computed on the pipe data
-and 46 on the sine-wave data. This study does not consider dimensions
-lower than four, as exploring the search space in such cases is
-relatively straightforward. Similarly, extremely high-dimensional cases
-with complex data structures, where no method can reliably and
-consistenly uncover the underlying structure due to the curse of
-dimensionality, are also excluded, as they fall outside the scope of
-this study. A list of combinations of the data shape, dimensions and PP
-indexes that are considered in this study is provided in
-Table~\ref{tbl-smoothness-squintability}. Again, JSO is run 50 times for
-each scenario to calculate the success rate for each hyper-parameter
-combination.
-
-Smoothness and squintability are computed following the procedures
-outlined in Section~\ref{sec-smoothness} and
-Section~\ref{sec-squintability} and as illustrated in
-Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}. To
-compute smoothness, 500 random bases are simulated. Index values are
-calculated for each random basis, followed by fitting a Gaussian process
-model to obtain the smoothness measure for the index.
-
-To compute squintability, 50 random bases are sampled and interpolated
-to the optimal basis with a step size of 0.005. Index values and
-projection distances are calculated for these interpolated bases. A
-binning procedure is applied to average the index values within each
-0.005 projection distance bin. The four-parameter scaled logistic
-function (\ref{eq-parametric}) is fitted to the index values against
-projection distances, estimated by non-linear least squares to obtain
-the squintability measure as Equation
-(\ref{eq-squintability-parametric}).
-
-A generalised linear model is fitted using a binomial family and a logit
-link function to assess the factors affecting success rate. Predictors
-are smoothness, squintability, and JSO hyper-parameters. A success rate
-of 0 indicates the optimiser did not come sufficiently close to the
-optimal projection.
+To examine the performance of JSO relative to hyper-parameter choices,
+the \texttt{holes} index is used to find the 2D pipe in data of
+dimensions 4-12. Additional variables are Gaussian noise. The levels of
+hyper-parameters are 20, 50, and 100 jellyfish and a maximum of 50 and
+100 attempts. Fifty samples of each data set are generated, and the
+success rate is calculated in the same manner as described in
+Section~\ref{sec-app-1}.
+
+This set of simulations is expanded to assess the performance of JSO
+relative to the index properties, with a second data set type and
+additional PP indexes. The second data set has a sine wave in two
+dimensions and Gaussian noise in the remaining dimensions. Six
+additional PP indexes that should be able to extract the sine wave are
+used: \texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
+\texttt{spline}, and \texttt{stringy}, following Laa and Cook (2020).
+(The supplementary material includes definitions of these PP indexes,
+and descriptions of the pipe and sine wave data simulation.). The JSO
+hyper-parameters, number of jellies and maximum tries, are also
+included. This results in a total of 76 scenarios, comprising of 30
+computed on the pipe data and 46 on the sine wave data. The choice of
+4-12 dimensions covers the range of easy to relatively difficult for the
+optimisers to detect the structured projections. The same approach is
+used, that 50 samples are generated and the success rate is calculated
+at each level of the simulation design. A success rate of 0 indicates
+the optimiser did not come sufficiently close to the optimal projection.
 
@@
 \end{figure}%
+
+Figure~\ref{fig-smoothness} describes the procedure to compute
+smoothness. Index values are calculated for each of 500 simulated random
+bases. A Gaussian process model is fitted to the resulting data to
+obtain the smoothness measure for the index, as described in
+Section~\ref{sec-smoothness}.
 
@@
 \end{figure}%
+
+Figure~\ref{fig-squintability} describes the procedure to compute
+squintability. Index values and projection distances are computed for
+the projections along interpolated paths between 50 random bases and the
+optimal basis. An interpolation step size of 0.005 is used. The
+projection distances are binned to 0.005 intervals and index values are
+averaged in each bin. The four-parameter scaled logistic function
+(\ref{eq-parametric}) is fitted to the index values against projection
+distances, estimated by non-linear least squares to obtain the
+squintability measure as Equation (\ref{eq-squintability-parametric}).
+
+A generalised linear model is fitted using a quasibinomial family and a
+logit link function to assess the factors affecting success rate.
+Predictors are smoothness, squintability, data dimension and JSO
+hyper-parameters.
 
@@
 The effect of JSO hyper-parameters (number of jellyfish and the maximum
-number of tries) on the success rate is presented in
+number of tries) on the success rate is summarised in
 Figure~\ref{fig-proportion}. The uncertainty is quantified through 500
@@
 maximum tries increase, the success rate also increases. For problems
-with relatively lower dimensional search spaces (4 and 6 dimensions),
-small parameter values (20 jellyfish and a maximum of 50 tries) are
-sufficient for space exploration and thus can already achieve a high
-success rate. Higher parameter values (i.e.~100 jellyfish and a maximum
-of 100 tries) enhances exploration of the search space, which is
-particularly beneficial in higher-dimensional problems (8, 10, and 12
-dimensions) where a more complex search is necessary. However, in
-lower-dimensional problems, the benefit may be less pronounced, as the
-space is already relatively easy to navigate. While increasing both
-parameters enhances the performance of JSO, it also extends the
-computational time required for the optimisation, which can be
-computationally intensive when evaluating the index function (such as
-scagnostic indexes) multiple times across numerous iterations.
+with relatively lower dimensional (4, 6) search spaces, small parameter
+values (20 jellyfish and a maximum of 50 tries) are sufficient to
+achieve a high success rate. Many more jellyfish (100 jellyfish and a
+maximum of 100 tries) are needed to obtain a mediocre success rate for
+dimensions 10 and 12. This suggests that even JSO will struggle with
+high dimensions, and require substantial more computational time to have
+an acceptable success rate, especially for difficult PP indexes based on
+scagnostics.
 
@@
 
-Smoothness and squintability are calculated across a collection of
-pipe-finding and sine-wave finding problems to construct the
-relationship between success rate, JSO hyper-parameters, and index
-properties. Table~\ref{tbl-smoothness-squintability} presents the
-parameters estimated from the Gaussian process (outputscale \(\eta\),
-lengthscale \(\ell\), smoothness \(\nu\), and nugget \(\sigma\)) and
-from the scaled logistic function (\(\theta_1\) to \(\theta_4\)) for
-calculating smoothness and squintability in each PP problem considered.
-The column \(\nu\) is used as the smoothness metric and the column
-\(\varsigma\) is calculated as equation
+Smoothness and squintability are calculated across a collection of pipe
+and sine wave finding problems to construct the relationship between
+success rate, JSO hyper-parameters, and index properties.
+Table~\ref{tbl-smoothness-squintability} presents values computed for a
+subset of PP indexes on the sine wave data for dimensions 4-8. (A full
+table of the parameters estimated from the Gaussian process (outputscale
+\(\eta\), lengthscale \(\ell\), smoothness \(\nu\), and nugget
+\(\sigma\)) and from the scaled logistic function (\(\theta_1\) to
+\(\theta_4\)) for calculating smoothness and squintability for all PP
+indexes and data combinations is included in the supplementary
+materials). The column \(\nu\) is used as the smoothness metric and the
+column \(\varsigma\) is calculated as equation
 (\ref{eq-squintability-parametric}) as the squintability metric.
 
-\begin{figure}
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-idx-proj-dist-1.pdf}}
-
-}
-
-\caption{\label{fig-idx-proj-dist}Five traces of index value against
-projection distance for each of the three indexes (TIC, spline, and
-skinny) in 6D sine data. The traces of the splines index are smooth,
-with a steep increaes before leveling off, suggesting the optimum can be
-detected from a far distance, hence a largest squintability value. In
-contrast, the skinny index (squintability of almost zeor) has a noisy
-trace that only stabilizes near the optimum, suggesting the optimizer
-needs get very close to detect it.}
-
-\end{figure}%
+Figure~\ref{fig-idx-proj-dist} shows a sample of traces from the indexes
+in Table~\ref{tbl-smoothness-squintability} for the 6D sine wave data.
+The traces the PP index values computed for projections along an
+interpolated path between a noise projection and the optimal projection
+(sine wave). (Note index values were self-standardised to range between
+0 and 1, for comparison.) The \texttt{splines} index is very smooth with
+a gradual increase in value between the noise projection and the optimal
+projection. The \texttt{TIC} index is slightly less smooth and slightly
+less gradual change in value. Least desirable is the \texttt{skinny}
+index which is noisier and the optimisation needs to be much closer to
+the optimal angle to have a higher value. This illustrates how the new
+measures for squintability and smoothness work to describe the PP
+indexes.
 
@@
 
-\caption{\label{tbl-smoothness-squintability}Parameters estimated from
-the Gaussian process (outputscale \(\eta\), lengthscale \(\ell\),
-smoothness \(\nu\), and nugget \(\sigma\)) and scaled logistic function
-(\(\theta_1\) to \(\theta_4\) and \(\varsigma\)) for the pipe-finding
-and sine-wave finding problems. The columns \(\nu\) and \(\varsigma\)
-represent the smoothness and squintability measures respectively.}
+\caption{\label{tbl-smoothness-squintability}The columns \(\nu\) and
+\(\varsigma\) show the smoothness and squintability measures for three
+PP indexes for the sine wave data. The splines index has the largest
+squint angle and is the smoothest. The skinny index has low smoothness
+and smallest squint angle. As dimension increases the squint angle gets
+smaller but smoothness doesn't change much.}
 
@@
 \end{table}%
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-idx-proj-dist-1.pdf}}
+
+}
+
+\caption{\label{fig-idx-proj-dist}Five traces of index value against
+projection distance for each of the three indexes (TIC, spline, and
+skinny) on 6D sine wave data. The traces of the splines index are
+smooth, with a gradual change between noise and otimal projection. In
+contrast, the skinny index has a noisy trace that increases rapidly near
+the optimum. This illustrates the usefulness of the new measures for
+squintability and smoothness.}
+
+\end{figure}%
 
@@
 success rate almost halves. Increasing the number of jellies by 10
-increases the success rate by 32\%.
-
-\begingroup
-\fontsize{12.0pt}{14.4pt}\selectfont
+increases the success rate by 32\%. Table~\ref{tbl-joint-output} shows
+the fit statistics supporting that the model is reasonably strong.
+
+Note that the addition of interaction terms for squintability and data
+dimension, and smoothness and data dimension, improves the model a small
+amount, but does not change the interpretation above.
+
 \setlength{\LTpost}{0mm}
@@
 
-The simulation result suggests that squintability and smoothness are
-likely dependent. Intuitively, a highly smooth search space exhibits
-gradual changes, making it easier to perceive the overall structure from
-a distance, thereby increasing the squint angle.
-Table~\ref{tbl-joint-output} compare the model fits when squintability
-and smoothness are included separately. Given that the null deviance
-remains the same across models, the model with squintability yields a
-smaller residual deviance and lower AIC and BIC values, indicating a
-better fit. This suggests that squintability captures the relevant
-variation more effectively than smoothness. While it may seem
-counterintuitive at first that smoothness is not a significant factor,
-this outcome is reasonable in this context, as the swarm-based method
-used in this study, JSO, is designed to perform well even in non-smooth
-search spaces by leveraging population-based exploration.
-
-This study intends to serve as an illustration that the proposed measure
-effectively captures the key factors influencing the success rate of PP.
-The interactions between the two variables were not examined in this
-study due to the limited number of data points, nor was the linearity
-assumption of the relationship with success rate explicitly tested.
+This study intends to serve as an illustration that the proposed
+measures effectively captures the key factors influencing the success
+rate of PP.
 

### 2025-03-03 jso.qmd
fix figure 4

@@
 #   geom_point(size = 0.5) +
-#   geom_text(aes(label = sprintf("%.3f", index_val)), x = 0, y = 3.7,
-#             size = 8) +
+#   geom_text(aes(label = sprintf("%.3f", index_val)), x = 0, y = 3.6,
+#             size = 5) +
 #   xlim(-4, 4) + ylim(-4, 4) +

### 2025-03-03 jso.tex
fix figure 4

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colback=white, opacitybacktitle=0.6, breakable, bottomrule=.15mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, opacityback=0, bottomtitle=1mm, coltitle=black, colframe=quarto-callout-note-color-frame, toptitle=1mm, arc=.35mm, toprule=.15mm, rightrule=.15mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, opacitybacktitle=0.6, arc=.35mm, left=2mm, opacityback=0, colframe=quarto-callout-note-color-frame, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, bottomtitle=1mm, bottomrule=.15mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, toprule=.15mm, colback=white, coltitle=black]
 
@@
 
-\includegraphics[width=4.52in,height=\textheight,keepaspectratio]{figures/success-rate.png}
+\includegraphics[width=5.41in,height=\textheight,keepaspectratio]{figures/success-rate.png}
 

### 2025-03-03 jso.qmd
all to US spelling

@@
 abstract: |
-  The projection pursuit (PP) guided tour interactively optimises a criterion function known as the PP index, to explore high-dimensional data by revealing interesting projections.
-Optimisation of some PP indexes can be non-trivial, if they are non-smooth functions, or the optimum has a small "squint angle", detectable only from close proximity.
+  The projection pursuit (PP) guided tour interactively optimizes a criterion function known as the PP index, to explore high-dimensional data by revealing interesting projections.
+Optimization of some PP indexes can be non-trivial, if they are non-smooth functions, or the optimum has a small "squint angle", detectable only from close proximity.
 Here, measures for calculating the smoothness and squintability properties of the PP index are defined.
-These are used to investigate the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
-The performance of JSO for visualising data is evaluated across various hyper-parameter settings and compared with existing optimisers for the guided tour.
+These are used to investigate the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimizer (JSO), for optimizing PP indexes.
+The performance of JSO for visualizing data is evaluated across various hyper-parameter settings and compared with existing optimisers for the guided tour.
 The JSO algorithm has been implemented in the R package, `tourr`, and functions to calculate smoothness and squintability measures are implemented in the `ferrn` package.
@@
 creating plots of the data that reveal the main features to use for
-publication. The method involves optimising an objective function known
+publication. The method involves optimizing an objective function known
 as the PP index [e.g., @hall1989polynomial; @cook1993projection;
@@
 For a fixed sample of data, PP finds the orthonormal basis $A$ that
-maximises the index value of the projection, $Y = XA$:
+maximizes the index value of the projection, $Y = XA$:
 
@@
 
-Optimisation of PP is often discussed when new indexes are proposed
+Optimization of PP is often discussed when new indexes are proposed
 [@posse95; @marie-sainte2010; @grochowski2011]. @cook1995grand tied the
@@
 
-Here, the potential for a Jellyfish Search Optimiser (JSO) [see
+Here, the potential for a Jellyfish Search Optimizer (JSO) [see
 @chou_novel_2021, and @rajwar_exhaustive_2023] for the PPGT is explored.
-JSO, inspired by the search behaviour of jellyfish in the ocean, is a
+JSO, inspired by the search behavior of jellyfish in the ocean, is a
 swarm-based metaheuristic designed to solve global optimisation
@@
 research. A series of simulation experiments using various datasets and
-PP indexes are conducted to assess JSO's behaviour and its sensitivity
+PP indexes are conducted to assess JSO's behavior and its sensitivity
 to hyper-parameter choices (number of jellyfish and maximum number of
@@
     space is reduced during the optimisation.
--   a very localised search, to take tiny steps to get closer to the
+-   a very localized search, to take tiny steps to get closer to the
     local maximum.
@@
 @Laa:2020wkm. Most are relatively simply defined, for any projection
-dimension, and implemented because they are relatively easy to optimise.
+dimension, and implemented because they are relatively easy to optimist.
 A goal is to develop PP indexes based on scagnostics [@scag, @WW08], but
@@
 
-The Jellyfish Search Optimiser (JSO) mimics the natural movements of
+The Jellyfish Search Optimizer (JSO) mimics the natural movements of
 jellyfish, which include passive and active motions driven by ocean
@@
 exploitation (focusing on promising areas). The algorithm aims to find
-the optimal solution by adapting the behaviour of jellyfish to navigate
+the optimal solution by adapting the behavior of jellyfish to navigate
 towards the best solution over iterations [@chou_novel_2021].
@@
 
-The performance of the optmisers is measured by the success rate, which is defined as the proportion of simulations that achieves a final index value within 0.05
+The performance of the opt misers is measured by the success rate, which is defined as the proportion of simulations that achieves a final index value within 0.05
 of the best index value found among all 50 simulations.
-@fig-success-rate illustrates why the choice of 0.05 is reasonable: the pipe is not recognisable in the projected data.
+@fig-success-rate illustrates why the choice of 0.05 is reasonable: the pipe is not recognizable in the projected data.
 This is motivated by @laa_using_2020's approach to investigating PP indexes.
@@
 
-A generalised linear model is fitted using a quasibinomial family and a logit link function to assess the factors affecting success rate.  Predictors are smoothness,
+A generalized linear model is fitted using a quasibinomial family and a logit link function to assess the factors affecting success rate.  Predictors are smoothness,
 squintability, data dimension and JSO hyper-parameters. <!-- The data is
@@
 Index value is printed on the plot.
-The purpose is to summarise the views of the data resulting from different optimisations, and hence compare results between the two optimisers.
+The purpose is to summarize the views of the data resulting from different optimization, and hence compare results between the two optimisers.
 Generally, the JSO does a more consistent job of finding the pipe structure clearly.
@@
 The effect of JSO hyper-parameters (number of jellyfish and the maximum
-number of tries) on the success rate is summarised in @fig-proportion.
+number of tries) on the success rate is summarized in @fig-proportion.
 The uncertainty is quantified through 500 bootstrap replicates for each
@@
 The traces the PP index values computed for projections along an interpolated path between a noise projection and the optimal projection (sine wave).
-(Note index values were self-standardised to range between 0 and 1, for comparison.) The `splines` index is very smooth with a gradual increase in value between the noise projection and the optimal projection.
+(Note index values were self-standardized to range between 0 and 1, for comparison.) The `splines` index is very smooth with a gradual increase in value between the noise projection and the optimal projection.
 The `TIC` index is slightly less smooth and slightly less gradual change in value.
@@
 explain how they can be used for new index and optimiser development.
-Using the JSO optimiser for PPGT requires a change in user behaviour.
+Using the JSO optimiser for PPGT requires a change in user behavior.
 For all the existing optimisers a single optimisation path is followed.
@@
 the tour path for JSO due to the generation of multiple tour paths. To
-visualise the path visited by each individual jellyfish, assign the
+visualize the path visited by each individual jellyfish, assign the
 animation to an object (`res <- animate_xy(...)`). This will save the
@@
 bases to sample. The output of `sample_bases()` is a data frame with a
-list-column of sampled basis matrix and index value. Parallellasation is
+list-column of sampled basis matrix and index value. Particularization is
 available to speed up the index value computation through the
@@
 assess the performance of the new jellyfish search optimiser. The
-metrics will be generally useful for characterising PP indexes, and help
+metrics will be generally useful for characterizing PP indexes, and help
 with developing new indexes.

### 2025-03-03 jso.tex
all to US spelling

@@
 \begin{abstract}
-The projection pursuit (PP) guided tour interactively optimises a
+The projection pursuit (PP) guided tour interactively optimizes a
 criterion function known as the PP index, to explore high-dimensional
-data by revealing interesting projections. Optimisation of some PP
+data by revealing interesting projections. Optimization of some PP
 indexes can be non-trivial, if they are non-smooth functions, or the
@@
 investigate the performance of a recently introduced swarm-based
-algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes.
-The performance of JSO for visualising data is evaluated across various
+algorithm, Jellyfish Search Optimizer (JSO), for optimizing PP indexes.
+The performance of JSO for visualizing data is evaluated across various
 hyper-parameter settings and compared with existing optimisers for the
@@
 high-dimensional data, and creating plots of the data that reveal the
-main features to use for publication. The method involves optimising an
+main features to use for publication. The method involves optimizing an
 objective function known as the PP index (e.g., Hall 1989; Cook, Buja,
@@
 structure. For a fixed sample of data, PP finds the orthonormal basis
-\(A\) that maximises the index value of the projection, \(Y = XA\):
+\(A\) that maximizes the index value of the projection, \(Y = XA\):
 
@@
 
-Optimisation of PP is often discussed when new indexes are proposed
+Optimization of PP is often discussed when new indexes are proposed
 (Posse 1995; Marie-Sainte, Berro, and Ruiz-Gazen 2010; Grochowski and
@@
 
-Here, the potential for a Jellyfish Search Optimiser (JSO) Rajwar, Deep,
+Here, the potential for a Jellyfish Search Optimizer (JSO) Rajwar, Deep,
 and Das (2023) for the PPGT is explored. JSO, inspired by the search
-behaviour of jellyfish in the ocean, is a swarm-based metaheuristic
+behavior of jellyfish in the ocean, is a swarm-based metaheuristic
 designed to solve global optimisation problems. Compared to traditional
@@
 research. A series of simulation experiments using various datasets and
-PP indexes are conducted to assess JSO's behaviour and its sensitivity
-to hyper-parameter choices (number of jellyfish and maximum number of
+PP indexes are conducted to assess JSO's behavior and its sensitivity to
+hyper-parameter choices (number of jellyfish and maximum number of
 tries). The relationship between the JSO performance, hyper-parameter
@@
 \item
-  a very localised search, to take tiny steps to get closer to the local
+  a very localized search, to take tiny steps to get closer to the local
   maximum.
@@
 defined, for any projection dimension, and implemented because they are
-relatively easy to optimise. A goal is to develop PP indexes based on
+relatively easy to optimist. A goal is to develop PP indexes based on
 scagnostics Wilkinson and Wills (2008), but the blockage is their
@@
 exploitation (focusing on promising areas). The algorithm aims to find
-the optimal solution by adapting the behaviour of jellyfish to navigate
+the optimal solution by adapting the behavior of jellyfish to navigate
 towards the best solution over iterations (Chou and Truong 2021).
@@
 
-\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, opacitybacktitle=0.6, arc=.35mm, left=2mm, opacityback=0, colframe=quarto-callout-note-color-frame, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, bottomtitle=1mm, bottomrule=.15mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, toprule=.15mm, colback=white, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, left=2mm, bottomrule=.15mm, colback=white, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, bottomtitle=1mm, opacityback=0, opacitybacktitle=0.6, coltitle=black, breakable, toptitle=1mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, leftrule=.75mm, rightrule=.15mm]
 
@@
 
-The performance of the optmisers is measured by the success rate, which
+The performance of the opt misers is measured by the success rate, which
 is defined as the proportion of simulations that achieves a final index
@@
 simulations. Figure~\ref{fig-success-rate} illustrates why the choice of
-0.05 is reasonable: the pipe is not recognisable in the projected data.
+0.05 is reasonable: the pipe is not recognizable in the projected data.
 This is motivated by Laa and Cook (2020)'s approach to investigating PP
@@
 
-A generalised linear model is fitted using a quasibinomial family and a
+A generalized linear model is fitted using a quasibinomial family and a
 logit link function to assess the factors affecting success rate.
@@
 minimum at left and maximum at right. Index value is printed on the
-plot. The purpose is to summarise the views of the data resulting from
-different optimisations, and hence compare results between the two
+plot. The purpose is to summarize the views of the data resulting from
+different optimization, and hence compare results between the two
 optimisers. Generally, the JSO does a more consistent job of finding the
@@
 The effect of JSO hyper-parameters (number of jellyfish and the maximum
-number of tries) on the success rate is summarised in
+number of tries) on the success rate is summarized in
 Figure~\ref{fig-proportion}. The uncertainty is quantified through 500
@@
 interpolated path between a noise projection and the optimal projection
-(sine wave). (Note index values were self-standardised to range between
+(sine wave). (Note index values were self-standardized to range between
 0 and 1, for comparison.) The \texttt{splines} index is very smooth with
@@
 explain how they can be used for new index and optimiser development.
-Using the JSO optimiser for PPGT requires a change in user behaviour.
-For all the existing optimisers a single optimisation path is followed.
-The JSO has multiple optimisation paths, and needs additional tools to
+Using the JSO optimiser for PPGT requires a change in user behavior. For
+all the existing optimisers a single optimisation path is followed. The
+JSO has multiple optimisation paths, and needs additional tools to
 incorporate into the PPGT, as explained here.
@@
 directly render the tour path for JSO due to the generation of multiple
-tour paths. To visualise the path visited by each individual jellyfish,
+tour paths. To visualize the path visited by each individual jellyfish,
 assign the animation to an object
@@
 with a list-column of sampled basis matrix and index value.
-Parallellasation is available to speed up the index value computation
+Particularization is available to speed up the index value computation
 through the \texttt{parallel\ =\ TRUE} argument.

### 2025-03-03 jso.qmd
proofread before Sec 3.2

@@
 ---
-title: Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices
+title: Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding optimization Choices
 author:
@@
 abstract: |
-  The projection pursuit (PP) guided tour interactively optimizes a criterion function known as the PP index, to explore high-dimensional data by revealing interesting projections.
-Optimization of some PP indexes can be non-trivial, if they are non-smooth functions, or the optimum has a small "squint angle", detectable only from close proximity.
+  The projection pursuit (PP) guided tour optimizes a criterion function, known as the PP index, to gradually reveal projections of interest from high-dimensional data through animation.
+Optimization of some PP indexes can be non-trivial, if they are non-smooth functions, or when the optimum has a small "squint angle", detectable only from close proximity.
 Here, measures for calculating the smoothness and squintability properties of the PP index are defined.
 These are used to investigate the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimizer (JSO), for optimizing PP indexes.
-The performance of JSO for visualizing data is evaluated across various hyper-parameter settings and compared with existing optimisers for the guided tour.
+The performance of JSO in detecting the target pattern (pipe shape) is compared with existing optimizers in PP.
+Additionally, JSO's performance on detecting the sine-wave shape is evaluated using different PP indexes (hence different smoothness and squintability) across various data dimensions (d = 4, 6, 8, 10, 12) and JSO hyper-parameters.
 The JSO algorithm has been implemented in the R package, `tourr`, and functions to calculate smoothness and squintability measures are implemented in the `ferrn` package.
 keywords: 
-  - projection pursuit
-  - jellyfish search optimiser (JSO)
-  - optimisation
-  - grand tour
-  - high-dimensional data
-  - exploratory data analysis
+  - Projection Pursuit Guided Tour (PPGT)
+  - Jellyfish Search Optimizer (JSO)
 date: last-modified
@@
 Let $X \in \mathbb{R}^{n\times p}$ be the data matrix,
-$A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix, where $A$
-belongs to the Stiefel manifold $\mathcal{A} = V_d(\mathbb{R}^p)$. The
-projection $Y = XA$ is a linear transformation that maps data from a
-$p$-dimensional space into a $d$-dimensional space. The index function
-$f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ is a scalar function
-that measures an interesting aspect of the projected data, such as
+$A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix in the Stiefel 
+manifold $\mathcal{A} = V_d(\mathbb{R}^p)$. The
+projection $Y = XA$ is a linear transformation of the $p$-dimensional data to
+a $d$-dimensional space. The index function
+$f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ defines a statistics to measure
+an pattern of interest in the projected data, such as
 deviation from normality, presence of clusters, or non-linear structure.
-For a fixed sample of data, PP finds the orthonormal basis $A$ that
-maximizes the index value of the projection, $Y = XA$:
+For a fixed sample of data, PP finds the orthonormal matrix $A$ (also called 
+*projection basis* in PP literature) that maximizes the index value of the 
+projection:
 
@@
 
+where $I_d$ is the identity matrix of dimension $d$.  
 It is interesting to note that when using PP visually, one cares less
@@
 data sets, but capture only large patterns. Many indexes are noisy and
-non-convex, requiring an effective and efficient optimisation procedure
+non-convex, requiring an effective and efficient optimization procedure
 to explore the data landscape and achieve a globally optimal viewpoint
@@
 (c), is noisier and has a very small squint angle. The discreteness
-cannot be seen unless the optimiser is very close to the optimal
+cannot be seen unless the optimizer is very close to the optimal
 projection.
@@
 [@posse95; @marie-sainte2010; @grochowski2011]. @cook1995grand tied the
-optimisation more closely to the index, when they introduced the PPGT,
-which monitors the optimisation visually so that the user can see the
+optimization more closely to the index, when they introduced the PPGT,
+which monitors the optimization visually so that the user can see the
 projected data leading in and out of the optimum. An implementation is
 available in the `tourr` package [@tourr] in R [@R]. @RJ-2021-105
-illustrated how to diagnose optimisation processes, particularly
+illustrated how to diagnose optimization processes, particularly
 focusing on the guided tour, and revealed a need for improved
-optimisation. While improving the quality of the optimisation solutions
+optimization. While improving the quality of the optimization solutions
 in the tour is essential, it is also important to be able to view the
-data projections as the optimisation progresses. Integrating the guided
-tour with a global optimisation algorithm that is efficient in finding
+data projections as the optimization progresses. Integrating the guided
+tour with a global optimization algorithm that is efficient in finding
 the global optimal and enables viewing of the projected data during the
@@
 
-Here, the potential for a Jellyfish Search Optimizer (JSO) [see
-@chou_novel_2021, and @rajwar_exhaustive_2023] for the PPGT is explored.
-JSO, inspired by the search behavior of jellyfish in the ocean, is a
-swarm-based metaheuristic designed to solve global optimisation
-problems. Compared to traditional methods, JSO has demonstrated stronger
+In this work, the potential for a Jellyfish Search Optimizer (JSO) [
+@chou_novel_2021; @rajwar_exhaustive_2023] for the PPGT is explored.
+The JSO, inspired by the search behavior of jellyfish in the ocean, is a
+swarm-based metaheuristic designed to solve global optimization
+problems. Compared to other metaheuristic methods, JSO has demonstrated stronger
 search ability and faster convergence, and requires fewer tuning
 parameters. These practicalities make JSO a promising candidate for
-enhancing PP optimisation.
+enhancing PP optimization.
 
 The primary goal of the study reported here is to investigate the
-performance of JSO in PP optimisation for the guided tour. It is of
-interest to assess how quickly and closely the optimiser reaches a
+performance of JSO in PP optimization for the guided tour. It is of
+interest to assess how quickly and closely the optimizer reaches a
 global optimum, for various PP indexes that may have differing
@@
 index including squintability [based on @barnett1981interpreting's
-squint angle] and smoothness. Here, we mathematically define metrics for
+squint angle] and smoothness. We mathematically define metrics for
 squintability and smoothness, which is a new contribution for PP
-research. A series of simulation experiments using various datasets and
-PP indexes are conducted to assess JSO's behavior and its sensitivity
-to hyper-parameter choices (number of jellyfish and maximum number of
-tries). The relationship between the JSO performance, hyper-parameter
-choices and properties of PP indexes (smoothness and squintability) is
-analysed to provide guidance on selecting optimisers for practitioners
-using projection pursuit. Additionally, this work should guide the
-design of new PP indexes and facilitate better optimization for PP.
+research. A series of simulation experiments are conducted using JSO to detect 
+different target patterns (pipe and sine-wave) with different PP indexes 
+(holes, MIC, TIC, dcor, loess, splines, skinny, and stringy), as well as JSO's 
+hyper-parameter choices (number of jellyfish and maximum number of
+tries). This work should facilitate better optimization for PP and
+guide the choice of optimizer when designing new PP indexes.
 
 The paper is structured as follows. @sec-background introduces the
-background of the PPGT, reviews existing optimisers and index functions
+background of the PPGT, reviews existing optimizers and index functions
 in the literature. @sec-PP-properties introduces the metrics that
-measure two properties of PP indexes, smoothness and squintability, and
-@sec-JSO describes the new JSO to be used for the PPGT. @sec-sim-deets
+measure two properties of PP indexes, smoothness and squintability. 
+@sec-JSO describes the new JSO for PP. @sec-sim-deets
 outlines two simulation experiments to assess JSO's performance: one
 comparing JSO's performance improvements relative to an existing
-optimiser, Creeping Random Search (CRS), and the other studying the
-impact of PP index properties on optimisation performance, and
+optimizer, Creeping Random Search (CRS), and the other studying the
+impact of PP index properties, controlling for data dimension and JSO 
+hyper-parameters, on optimization performance, and
 @sec-sim-res presents the results. @sec-discussion discusses the
 implementation of JSO in the `tourr` package and the PP property
-calculation in the `ferrn` package. @sec-conclusion summarises the work
+calculation in the `ferrn` package. @sec-conclusion summarizes the work
 and provides suggestions for future directions.
 
-# Projection pursuit, tours, index functions and optimisation {#sec-background}
+# Projection pursuit, tours, index functions and optimization {#sec-background}
 
@@
 (projections) which have high PP index values, as provided by the
-optimiser. The geodesic interpolation means that the viewer sees a
+optimizer. The geodesic interpolation means that the viewer sees a
 continuous sequence of projections of the data, so they can watch
-patterns of interest forming as the function is optimised. There are
-five optimisation methods implemented in the `tourr` package:
+patterns of interest forming as the function is optimized. There are
+five optimization methods implemented in the `tourr` package:
 
@@
     based on differencing the index values for very close projections.
--   a brute-force optimisation (CRS).
+-   a brute-force optimization (CRS).
 -   a modified brute force algorithm described in @posse95.
 -   an essentially simulated annealing [@Bertsimas93] where the search
-    space is reduced during the optimisation.
+    space is reduced during the optimization.
 -   a very localized search, to take tiny steps to get closer to the
@@
 <!-- 
--   `search_geodesic()`: provides a pseudo-derivative optimisation. It
+-   `search_geodesic()`: provides a pseudo-derivative optimization. It
     searches locally for the best direction, based on differencing the
@@
 -   `search_better()`: also known as Creeping Random Search (CRS), is a
-    brute-force optimisation searching randomly for projections with
+    brute-force optimization searching randomly for projections with
     higher index values.
 -   `search_better_random()`: is essentially simulated annealing
-    [@Bertsimas93] where the search space is reduced as the optimisation
+    [@Bertsimas93] where the search space is reduced as the optimization
     progresses.
 -   `search_posse()`: implements the algorithm described in @posse95.
--   `search_polish()`: is a very localised search, to take tiny steps to
+-   `search_polish()`: is a very localized search, to take tiny steps to
     get closer to the local maximum.
@@
 @Laa:2020wkm. Most are relatively simply defined, for any projection
-dimension, and implemented because they are relatively easy to optimist.
-A goal is to develop PP indexes based on scagnostics [@scag, @WW08], but
-the blockage is their optimisation as these tend to be noisy, with
+dimension, and implemented because they are relatively easy to optimize.
+A goal is to develop PP indexes based on scagnostics [@scag; @WW08], but
+the blockage is their optimization as these tend to be noisy, with
 potentially small squint angles.
@@
 scagnostics is described in @laa_using_2020. To be useful here an
-optimiser needs to be able to handle index functions that are possibly
-not very smooth. In addition, because data structures might be
-relatively fine, the optimiser needs to be able to find maxima that
+optimizer needs to be able to handle index functions that are possibly
+not very smooth. In addition, because the target structure in the data might be
+relatively fine, the optimizer needs to be able to find maxima that
 occur with a small squint angle, that can only be seen from very close
-by. One last aspect that is useful is for an optimiser to return local
+by. One last aspect that is useful is for an optimizer to return local
 maxima in addition to the global one because data can contain many
@@
 invariance, and speed). Since not all index properties affect the
-optimisation process, the focus here is on the first two properties,
+optimization process, the focus of this work is on the first two properties,
 *smoothness* (@sec-smoothness) and *squintability* (@sec-squintability),
@@
 ## Smoothness {#sec-smoothness}
-
-This subsection proposes a metric for the smoothness of a projection
-pursuit index.
 
@@
 ```
+
 The Sobolev index $k$ in Definition \ref{def:sobolev_space} can be used
@@
 useful measure of smoothness, it can be difficult to compute or even
-estimate in practice.
+estimate in practice. For example, the loess and splines indexes are not
+continuous due to the use of the $\max$ operator in their definitions. 
+Scagnostic indexes, such as skinny and stringy, are defined based on
+graph elements (area, diameter, perimeter, and length) of alph hull or 
+minimal spanning tree).
 
@@
 $f$, we propose an approach based on random fields. If a PP index
-function $f$ is evaluated at some random bases, as is done at the
-initialization stage of JSO, then these random index values can be
-interpreted as a random field, indexed by a space parameter, namely the
+function $f$ is evaluated at some random bases, then these random index values 
+can be interpreted as a random field, indexed by a space parameter, namely the
 random projection basis. This analogy suggests to use this random
@@
 prior with Matérn covariance on a dataset generated by random
-evaluations of the index function, as done at the initialization stage
-of the jellyfish search optimization. There exist several R packages,
+evaluations of the index function. There exist several R packages,
 such as `GpGp` [@guinness2021gpgp] or `ExaGeoStatR` [@abdulah2023large],
@@
 \begin{defn}
-Let $\mathbf{A}=[A_1, \ldots, A_N] \in (\mathbb{R}^{p \times d})^N$ be d-dimensional projection bases, let $\mathbf{y}=[f(XA_1),\ldots,f(XA_N)]$ be the corresponding PP index values, and let $\mathbf{K}=[K_\theta(A_{i}-A_{j})]_{1\leq i,j\leq N}\in\mathbb{R}^{N\times N}$ be the Matérn covariance matrix evaluated at the input bases, where the vector $\theta$ contains all the parameters of the multivariate Matérn covariance function $K$ (smoothness, outputscale, lengthscales).
+Let $\mathbf{A}=[A_1, \ldots, A_N] \in (\mathbb{R}^{p \times d})^N$ be d-dimensional projection bases, $\mathbf{y}=[f(XA_1),\ldots,f(XA_N)]$ be the corresponding PP index values, and $\mathbf{K}=[K_\theta(A_{i}-A_{j})]_{1\leq i,j\leq N}\in\mathbb{R}^{N\times N}$ be the Matérn covariance matrix evaluated at the input bases, where the vector $\theta$ contains all the parameters of the multivariate Matérn covariance function $K$ (smoothness, outputscale, lengthscales).
 The log-likelihood of the parameters $\theta$ is defined by
@@
 
-This section defines the projection distance and introduces the
-squintability metric, followed by a description of two numerical
-approaches for computing squintability.
+Squintability relates to the notion of squint angle and is defined based on how the index value $f(XA)$ changes with respect to the projection distance over the course of the optimization.
 
 ```{=tex}
+\begin{defn}[squint angle]\label{def:squint-angle}
+Let $A$ and $B$ be two $d$-dimensional orthonormal matrices in $\mathbb{R}^p$.
+The squint angle $\theta$ between the subspace spanned by $A$ and $B$ is defined as the smallest principal angle between these subspaces: $\theta = \min_{i \in \{1, \cdots, d\}} \arccos(\tau_i)$, where $\tau_i$ are the singular values of the matrix $M = A^T B$ obtained from its singular value decomposition.
+\end{defn}
+
 \begin{defn}[projection distance]\label{def:proj-dist}
@@
 \end{defn}
-\begin{defn}[squint angle]\label{def:squint-angle}
-
-Let $A$ and $B$ be two $d$-dimensional orthonormal matrices in $\mathbb{R}^p$.
-The squint angle $\theta$ between the subspace spanned by $A$ and $B$ is defined as the smallest principal angle between these subspaces: $\theta = \min_{i \in \{1, \cdots, d\}} \arccos(\tau_i)$, where $\tau_i$ are the singular values of the matrix $M = A^T B$ obtained from its singular value decomposition.
-\end{defn}
-```
-Squintability can be defined based on how the index value $f(XA)$ changes with
-respect to the projection distance $r(A, A^*)$, over the course of the
-JSO. Suppose that the optimization starts with an arbitrary candidate
-matrix $A_0$ at a distance $r_0:=r(A_0, A^*)$ from the optimal one. A possible way to define squintability is as follows:
+```
 
@@
 small squint angle is considered to be undesirable because it means that
-the optimiser needs to be very close to be able to "see" the optimum.
-Thus, it could be difficult for the optimiser to find the optimum.
+the optimizer needs to be very close to be able to "see" the optimum.
+Thus, it could be difficult for the optimizer to find the optimum.
 <!--The
@@
 
-# The jellyfish optimiser {#sec-JSO}
+# The jellyfish search optimizer {#sec-JSO}
 
@@
 
-To solve the optimisation problem embedded in the PP guided tour, a
+To solve the optimization problem embedded in the PP guided tour, a
 starting projection, an index function, the number of jellyfish, and the
@@
 then moved in a direction determined by a random factor, influenced by
-how far along we are in the optimisation process. Occasionally,
+how far along we are in the optimization process. Occasionally,
 completely new directions may be taken like a jellyfish might with ocean
@@
 
-**Output**: `optimised_projection`
+**Output**: `optimized_projection`
 
@@
 **return** the set of projections with the updated `current_best` as the
-`optimised_projection`
+`optimized_projection`
 :::
@@
 
-# Assessing the optimisers {#sec-sim-deets}
-
-This section explains the details of two simulation studies: (1) comparison between the existing Creeping Random Search (CRS) [@RJ-2021-105; @laa_using_2020] and JSO, and (2) examining the factors affecting the performance of JSO.
+# Assessing the optimizers {#sec-sim-deets}
+
+This section explains the details of two simulation studies: (1) comparison between the existing creeping random search [@RJ-2021-105; @laa_using_2020] and JSO, and (2) examining the factors affecting the performance of JSO.
 
@@
 
-The CRS is the main optimisation routine currently used for the guided tour.
-Here the two optimisers are compared on the task of finding the pipe using the `holes` index in data with dimensions, $d = 6, 8, 10, 12$.
+The CRS is the main optimization routine currently used for the guided tour.
+Here the two optimizers are compared on the task of finding the pipe using the `holes` index in data with dimensions, $d = 6, 8, 10, 12$.
 Fifty simulated data sets are used for each dimension.
@@
 
-The results of the simulation are collected using the data structure used in @RJ-2021-105 for assessing PP optimisers.
+The results of the simulation are collected using the data structure used in @RJ-2021-105 for assessing PP optimizers.
 The design parameters are stored along with index value, projection basis, random seed, and computation time.
@@
 #| label: fig-success-rate
-#| fig.cap: "How success rate is calculated, illustrated using the optimal projections from 50 optimisations of 8D pipe data sorted by index value.
+#| fig.cap: "How success rate is calculated, illustrated using the optimal projections from 50 optimizations of 8D pipe data sorted by index value.
 The pipe shape is recognisable in the projection index values between 0.933-0.969.
@@
 This results in a total of `r nrow(sim_df)` scenarios, comprising of `r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and `r nrow(filter(sim_df, index != "holes"))` on the sine wave data.
-The choice of 4-12 dimensions covers the range of easy to relatively difficult for the optimisers to detect the structured projections.
+The choice of 4-12 dimensions covers the range of easy to relatively difficult for the optimizers to detect the structured projections.
 <!-- A list of combinations of the data shape, dimensions and PP indexes that are considered in this study is provided in @tbl-smoothness-squintability.--> The same approach is used, that 50 samples are generated and the success rate is calculated at each level of the simulation design.
-A success rate of 0 indicates the optimiser did not come sufficiently close to the optimal projection.
+A success rate of 0 indicates the optimizer did not come sufficiently close to the optimal projection.
 
@@
 
-This section summarises the findings from the simulations that compare the JSO performance with the existing CRS optimiser, and the relationship between optimisation success and hyper-parameter choices and PP index properties.
+This section summarises the findings from the simulations that compare the JSO performance with the existing CRS optimizer, and the relationship between optimization success and hyper-parameter choices and PP index properties.
 
@@
 
-The performance of JSO is assessed relative to the existing CRS optimiser based on the optimal projection obtained.
-The final projections for the pipe data sets found by the two optimisers are shown in @fig-proj, with rows corresponding to data dimension.
+The performance of JSO is assessed relative to the existing CRS optimizer based on the optimal projection obtained.
+The final projections for the pipe data sets found by the two optimizers are shown in @fig-proj, with rows corresponding to data dimension.
 The 50 simulations in each data dimension are sorted by index value, and the projections corresponding to each 10th quantile value are shown, with minimum at left and maximum at right.
 Index value is printed on the plot.
-The purpose is to summarize the views of the data resulting from different optimization, and hence compare results between the two optimisers.
+The purpose is to summarize the views of the data resulting from different optimization, and hence compare results between the two optimizers.
 Generally, the JSO does a more consistent job of finding the pipe structure clearly.
-As the data dimension increases, both optimisers struggle and less clearly capture the circle shape.
+As the data dimension increases, both optimizers struggle and less clearly capture the circle shape.
 
@@
 However, in lower-dimensional problems, the benefit may be less pronounced, as the space is already relatively easy to navigate.
-While increasing both parameters enhances the performance of JSO, it also extends the computational time required for the optimisation, which can be
+While increasing both parameters enhances the performance of JSO, it also extends the computational time required for the optimization, which can be
 computationally intensive when evaluating the index function (such as
@@
 #| fig.align: "center"
-#| fig.cap: "Summary of the relationship between JSO hyper-parameters on optimisation success, using the holes index for pipe data of dimensions 4-12.
+#| fig.cap: "Summary of the relationship between JSO hyper-parameters on optimization success, using the holes index for pipe data of dimensions 4-12.
 Bootstrap samples show the variability in success rate.
@@
 The `TIC` index is slightly less smooth and slightly less gradual change in value.
-Least desirable is the `skinny` index which is noisier and the optimisation needs to be much closer to the optimal angle to have a higher value.
+Least desirable is the `skinny` index which is noisier and the optimization needs to be much closer to the optimal angle to have a higher value.
 This illustrates how the new measures for squintability and smoothness work to describe the PP indexes.
@@
 This is the summary from a logistic regression fit to smoothness, squintability, dimension, number of jellyfish and maximum number of tries.
-Interestingly, squintability and dimension strongly affect jellyfish optimisation success.
+Interestingly, squintability and dimension strongly affect jellyfish optimization success.
 The number of jellies marginally affects success, but index smoothness, and increasing the number of tries do not."
@@
 smoothness and squintability affect the JSO success rate. Here we
-explain how they can be used for new index and optimiser development.
-Using the JSO optimiser for PPGT requires a change in user behavior.
-For all the existing optimisers a single optimisation path is followed.
-The JSO has multiple optimisation paths, and needs additional tools to
+explain how they can be used for new index and optimizer development.
+Using the JSO optimizer for PPGT requires a change in user behavior.
+For all the existing optimizers a single optimization path is followed.
+The JSO has multiple optimization paths, and needs additional tools to
 incorporate into the PPGT, as explained here.
@@
 `search_f = search_jellyfish` in the guided tour. Unlike existing
-optimisers, the animation function `animate_*()` won't directly render
+optimizers, the animation function `animate_*()` won't directly render
 the tour path for JSO due to the generation of multiple tour paths. To
@@
 features of PP indexes, squintability and smoothness, and used these to
-assess the performance of the new jellyfish search optimiser. The
+assess the performance of the new jellyfish search optimizer. The
 metrics will be generally useful for characterizing PP indexes, and help
@@
 One future direction of this work is to test the JSO, along with its
-variations and other swarm-based optimisers, on a broader range of
+variations and other swarm-based optimizers, on a broader range of
 indexes, for example, scagnostic indexes.
@@
 
-  - JSO tends to obtian a clearer projection and higher index value when searching higher dimension spaces than CRS (and other current optimisers) - Figure 4
-
-  - Increasing JS hyper-parameter enhances its performance, but the computational time required for the optimisation can quickly become intensive, especially when evaluating the index function (such as scagnostic indexes) multiple times across numerous iterations.
+  - JSO tends to obtian a clearer projection and higher index value when searching higher dimension spaces than CRS (and other current optimizers) - Figure 4
+
+  - Increasing JS hyper-parameter enhances its performance, but the computational time required for the optimization can quickly become intensive, especially when evaluating the index function (such as scagnostic indexes) multiple times across numerous iterations.
 - Figure 5

### 2025-03-03 jso.tex
proofread before Sec 3.2

@@
 \hypersetup{
-  pdftitle={Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices},
+  pdftitle={Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding optimization Choices},
   pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
-  pdfkeywords={projection pursuit, jellyfish search optimiser
-(JSO), optimisation, grand tour, high-dimensional data, exploratory data
-analysis},
+  pdfkeywords={Projection Pursuit Guided Tour (PPGT), Jellyfish Search
+Optimizer (JSO)},
   colorlinks=true,
@@
 \title{Squintability and Other Metrics for Assessing Projection Pursuit
-Indexes, and Guiding Optimisation Choices}
+Indexes, and Guiding optimization Choices}
 \author{H. Sherry Zhang$\textsuperscript{1}$, Dianne
@@
 \begin{abstract}
-The projection pursuit (PP) guided tour interactively optimizes a
-criterion function known as the PP index, to explore high-dimensional
-data by revealing interesting projections. Optimization of some PP
-indexes can be non-trivial, if they are non-smooth functions, or the
+The projection pursuit (PP) guided tour optimizes a criterion function,
+known as the PP index, to gradually reveal projections of interest from
+high-dimensional data through animation. Optimization of some PP indexes
+can be non-trivial, if they are non-smooth functions, or when the
 optimum has a small ``squint angle'', detectable only from close
@@
 algorithm, Jellyfish Search Optimizer (JSO), for optimizing PP indexes.
-The performance of JSO for visualizing data is evaluated across various
-hyper-parameter settings and compared with existing optimisers for the
-guided tour. The JSO algorithm has been implemented in the R package,
-\texttt{tourr}, and functions to calculate smoothness and squintability
-measures are implemented in the \texttt{ferrn} package.
+The performance of JSO in detecting the target pattern (pipe shape) is
+compared with existing optimizers in PP. Additionally, JSO's performance
+on detecting the sine-wave shape is evaluated using different PP indexes
+(hence different smoothness and squintability) across various data
+dimensions (d = 4, 6, 8, 10, 12) and JSO hyper-parameters. The JSO
+algorithm has been implemented in the R package, \texttt{tourr}, and
+functions to calculate smoothness and squintability measures are
+implemented in the \texttt{ferrn} package.
 \end{abstract}
@@
 \def\sep{;\ }
-projection pursuit\sep jellyfish search optimiser
-(JSO)\sep optimisation\sep grand tour\sep high-dimensional data\sep 
-exploratory data analysis
+Projection Pursuit Guided Tour (PPGT)\sep 
+Jellyfish Search Optimizer (JSO)
 \end{keywords}
@@
 projections. Let \(X \in \mathbb{R}^{n\times p}\) be the data matrix,
-\(A \in\mathbb{R}^{p \times d}\) be an orthonormal matrix, where \(A\)
-belongs to the Stiefel manifold \(\mathcal{A} = V_d(\mathbb{R}^p)\). The
-projection \(Y = XA\) is a linear transformation that maps data from a
-\(p\)-dimensional space into a \(d\)-dimensional space. The index
-function \(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar
-function that measures an interesting aspect of the projected data, such
-as deviation from normality, presence of clusters, or non-linear
-structure. For a fixed sample of data, PP finds the orthonormal basis
-\(A\) that maximizes the index value of the projection, \(Y = XA\):
+\(A \in\mathbb{R}^{p \times d}\) be an orthonormal matrix in the Stiefel
+manifold \(\mathcal{A} = V_d(\mathbb{R}^p)\). The projection \(Y = XA\)
+is a linear transformation of the \(p\)-dimensional data to a
+\(d\)-dimensional space. The index function
+\(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) defines a statistics
+to measure an pattern of interest in the projected data, such as
+deviation from normality, presence of clusters, or non-linear structure.
+For a fixed sample of data, PP finds the orthonormal matrix \(A\) (also
+called \emph{projection basis} in PP literature) that maximizes the
+index value of the projection:
 
@@
 
+where \(I_d\) is the identity matrix of dimension \(d\).\\
 It is interesting to note that when using PP visually, one cares less
@@
 are noisy and non-convex, requiring an effective and efficient
-optimisation procedure to explore the data landscape and achieve a
+optimization procedure to explore the data landscape and achieve a
 globally optimal viewpoint of the data. The skewness index computed for
@@
 the famous RANDU data, in (c), is noisier and has a very small squint
-angle. The discreteness cannot be seen unless the optimiser is very
+angle. The discreteness cannot be seen unless the optimizer is very
 close to the optimal projection.
@@
 (Posse 1995; Marie-Sainte, Berro, and Ruiz-Gazen 2010; Grochowski and
-Duch 2011). Cook et al. (1995) tied the optimisation more closely to the
-index, when they introduced the PPGT, which monitors the optimisation
+Duch 2011). Cook et al. (1995) tied the optimization more closely to the
+index, when they introduced the PPGT, which monitors the optimization
 visually so that the user can see the projected data leading in and out
@@
 package (Wickham et al. 2011) in R (R Core Team 2023). Zhang et al.
-(2021) illustrated how to diagnose optimisation processes, particularly
+(2021) illustrated how to diagnose optimization processes, particularly
 focusing on the guided tour, and revealed a need for improved
-optimisation. While improving the quality of the optimisation solutions
+optimization. While improving the quality of the optimization solutions
 in the tour is essential, it is also important to be able to view the
-data projections as the optimisation progresses. Integrating the guided
-tour with a global optimisation algorithm that is efficient in finding
+data projections as the optimization progresses. Integrating the guided
+tour with a global optimization algorithm that is efficient in finding
 the global optimal and enables viewing of the projected data during the
@@
 
-Here, the potential for a Jellyfish Search Optimizer (JSO) Rajwar, Deep,
-and Das (2023) for the PPGT is explored. JSO, inspired by the search
-behavior of jellyfish in the ocean, is a swarm-based metaheuristic
-designed to solve global optimisation problems. Compared to traditional
-methods, JSO has demonstrated stronger search ability and faster
-convergence, and requires fewer tuning parameters. These practicalities
-make JSO a promising candidate for enhancing PP optimisation.
+In this work, the potential for a Jellyfish Search Optimizer (JSO) (Chou
+and Truong 2021; Rajwar, Deep, and Das 2023) for the PPGT is explored.
+The JSO, inspired by the search behavior of jellyfish in the ocean, is a
+swarm-based metaheuristic designed to solve global optimization
+problems. Compared to other metaheuristic methods, JSO has demonstrated
+stronger search ability and faster convergence, and requires fewer
+tuning parameters. These practicalities make JSO a promising candidate
+for enhancing PP optimization.
 
 The primary goal of the study reported here is to investigate the
-performance of JSO in PP optimisation for the guided tour. It is of
-interest to assess how quickly and closely the optimiser reaches a
+performance of JSO in PP optimization for the guided tour. It is of
+interest to assess how quickly and closely the optimizer reaches a
 global optimum, for various PP indexes that may have differing
@@
 index including squintability (based on Tukey and Tukey 1981's squint
-angle) and smoothness. Here, we mathematically define metrics for
+angle) and smoothness. We mathematically define metrics for
 squintability and smoothness, which is a new contribution for PP
-research. A series of simulation experiments using various datasets and
-PP indexes are conducted to assess JSO's behavior and its sensitivity to
-hyper-parameter choices (number of jellyfish and maximum number of
-tries). The relationship between the JSO performance, hyper-parameter
-choices and properties of PP indexes (smoothness and squintability) is
-analysed to provide guidance on selecting optimisers for practitioners
-using projection pursuit. Additionally, this work should guide the
-design of new PP indexes and facilitate better optimization for PP.
+research. A series of simulation experiments are conducted using JSO to
+detect different target patterns (pipe and sine-wave) with different PP
+indexes (holes, MIC, TIC, dcor, loess, splines, skinny, and stringy), as
+well as JSO's hyper-parameter choices (number of jellyfish and maximum
+number of tries). This work should facilitate better optimization for PP
+and guide the choice of optimizer when designing new PP indexes.
 
 The paper is structured as follows. Section~\ref{sec-background}
-introduces the background of the PPGT, reviews existing optimisers and
+introduces the background of the PPGT, reviews existing optimizers and
 index functions in the literature. Section~\ref{sec-PP-properties}
 introduces the metrics that measure two properties of PP indexes,
-smoothness and squintability, and Section~\ref{sec-JSO} describes the
-new JSO to be used for the PPGT. Section~\ref{sec-sim-deets} outlines
-two simulation experiments to assess JSO's performance: one comparing
-JSO's performance improvements relative to an existing optimiser,
-Creeping Random Search (CRS), and the other studying the impact of PP
-index properties on optimisation performance, and
-Section~\ref{sec-sim-res} presents the results.
+smoothness and squintability. Section~\ref{sec-JSO} describes the new
+JSO for PP. Section~\ref{sec-sim-deets} outlines two simulation
+experiments to assess JSO's performance: one comparing JSO's performance
+improvements relative to an existing optimizer, Creeping Random Search
+(CRS), and the other studying the impact of PP index properties,
+controlling for data dimension and JSO hyper-parameters, on optimization
+performance, and Section~\ref{sec-sim-res} presents the results.
 Section~\ref{sec-discussion} discusses the implementation of JSO in the
 \texttt{tourr} package and the PP property calculation in the
-\texttt{ferrn} package. Section~\ref{sec-conclusion} summarises the work
+\texttt{ferrn} package. Section~\ref{sec-conclusion} summarizes the work
 and provides suggestions for future directions.
@@
 \section{Projection pursuit, tours, index functions and
-optimisation}\label{sec-background}
+optimization}\label{sec-background}
 
@@
 target planes (projections) which have high PP index values, as provided
-by the optimiser. The geodesic interpolation means that the viewer sees
+by the optimizer. The geodesic interpolation means that the viewer sees
 a continuous sequence of projections of the data, so they can watch
-patterns of interest forming as the function is optimised. There are
-five optimisation methods implemented in the \texttt{tourr} package:
+patterns of interest forming as the function is optimized. There are
+five optimization methods implemented in the \texttt{tourr} package:
 
@@
 \item
-  a brute-force optimisation (CRS).
+  a brute-force optimization (CRS).
 \item
@@
   an essentially simulated annealing (Bertsimas and Tsitsiklis 1993)
-  where the search space is reduced during the optimisation.
+  where the search space is reduced during the optimization.
 \item
@@
 defined, for any projection dimension, and implemented because they are
-relatively easy to optimist. A goal is to develop PP indexes based on
-scagnostics Wilkinson and Wills (2008), but the blockage is their
-optimisation as these tend to be noisy, with potentially small squint
-angles.
+relatively easy to optimize. A goal is to develop PP indexes based on
+scagnostics (Wilkinson, Anand, and Grossman 2005; Wilkinson and Wills
+2008), but the blockage is their optimization as these tend to be noisy,
+with potentially small squint angles.
 
@@
 scagnostics is described in Laa and Cook (2020). To be useful here an
-optimiser needs to be able to handle index functions that are possibly
-not very smooth. In addition, because data structures might be
-relatively fine, the optimiser needs to be able to find maxima that
-occur with a small squint angle, that can only be seen from very close
-by. One last aspect that is useful is for an optimiser to return local
-maxima in addition to the global one because data can contain many
+optimizer needs to be able to handle index functions that are possibly
+not very smooth. In addition, because the target structure in the data
+might be relatively fine, the optimizer needs to be able to find maxima
+that occur with a small squint angle, that can only be seen from very
+close by. One last aspect that is useful is for an optimizer to return
+local maxima in addition to the global one because data can contain many
 different and interesting features.
@@
 invariance, and speed). Since not all index properties affect the
-optimisation process, the focus here is on the first two properties,
-\emph{smoothness} (Section~\ref{sec-smoothness}) and
+optimization process, the focus of this work is on the first two
+properties, \emph{smoothness} (Section~\ref{sec-smoothness}) and
 \emph{squintability} (Section~\ref{sec-squintability}), for which
@@
 \subsection{Smoothness}\label{sec-smoothness}
-
-This subsection proposes a metric for the smoothness of a projection
-pursuit index.
 
@@
 \(k\) is a useful measure of smoothness, it can be difficult to compute
-or even estimate in practice.
+or even estimate in practice. For example, the loess and splines indexes
+are not continuous due to the use of the \(\max\) operator in their
+definitions. Scagnostic indexes, such as skinny and stringy, are defined
+based on graph elements (area, diameter, perimeter, and length) of alph
+hull or minimal spanning tree).
 
@@
 \(f\), we propose an approach based on random fields. If a PP index
-function \(f\) is evaluated at some random bases, as is done at the
-initialization stage of JSO, then these random index values can be
-interpreted as a random field, indexed by a space parameter, namely the
-random projection basis. This analogy suggests to use this random
-training sample to fit a spatial model. We propose to use a Gaussian
-process equipped with a Matérn covariance function, due to the
-connections between this model and Sobolev spaces, see for example Porcu
-et al. (2024).
+function \(f\) is evaluated at some random bases, then these random
+index values can be interpreted as a random field, indexed by a space
+parameter, namely the random projection basis. This analogy suggests to
+use this random training sample to fit a spatial model. We propose to
+use a Gaussian process equipped with a Matérn covariance function, due
+to the connections between this model and Sobolev spaces, see for
+example Porcu et al. (2024).
 
@@
 process prior with Matérn covariance on a dataset generated by random
-evaluations of the index function, as done at the initialization stage
-of the jellyfish search optimization. There exist several R packages,
-such as \texttt{GpGp} (Guinness, Katzfuss, and Fahmy 2021) or
+evaluations of the index function. There exist several R packages, such
+as \texttt{GpGp} (Guinness, Katzfuss, and Fahmy 2021) or
 \texttt{ExaGeoStatR} (Abdulah et al. 2023), to fit the hyper-parameters
@@
 
-This section defines the projection distance and introduces the
-squintability metric, followed by a description of two numerical
-approaches for computing squintability.
+Squintability relates to the notion of squint angle and is defined based
+on how the index value \(f(XA)\) changes with respect to the projection
+distance over the course of the optimization.
+
+\begin{defn}[squint angle]\label{def:squint-angle}
+Let $A$ and $B$ be two $d$-dimensional orthonormal matrices in $\mathbb{R}^p$.
+The squint angle $\theta$ between the subspace spanned by $A$ and $B$ is defined as the smallest principal angle between these subspaces: $\theta = \min_{i \in \{1, \cdots, d\}} \arccos(\tau_i)$, where $\tau_i$ are the singular values of the matrix $M = A^T B$ obtained from its singular value decomposition.
+\end{defn}
 
@@
 \end{defn}
-\begin{defn}[squint angle]\label{def:squint-angle}
-
-Let $A$ and $B$ be two $d$-dimensional orthonormal matrices in $\mathbb{R}^p$.
-The squint angle $\theta$ between the subspace spanned by $A$ and $B$ is defined as the smallest principal angle between these subspaces: $\theta = \min_{i \in \{1, \cdots, d\}} \arccos(\tau_i)$, where $\tau_i$ are the singular values of the matrix $M = A^T B$ obtained from its singular value decomposition.
-\end{defn}
-
-Squintability can be defined based on how the index value \(f(XA)\)
-changes with respect to the projection distance \(r(A, A^*)\), over the
-course of the JSO. Suppose that the optimization starts with an
-arbitrary candidate matrix \(A_0\) at a distance \(r_0:=r(A_0, A^*)\)
-from the optimal one. A possible way to define squintability is as
-follows:
 
@@
 small squint angle is considered to be undesirable because it means that
-the optimiser needs to be very close to be able to ``see'' the optimum.
-Thus, it could be difficult for the optimiser to find the optimum.
+the optimizer needs to be very close to be able to ``see'' the optimum.
+Thus, it could be difficult for the optimizer to find the optimum.
 
@@
 
-\section{The jellyfish optimiser}\label{sec-JSO}
+\section{The jellyfish search optimizer}\label{sec-JSO}
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, left=2mm, bottomrule=.15mm, colback=white, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, bottomtitle=1mm, opacityback=0, opacitybacktitle=0.6, coltitle=black, breakable, toptitle=1mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, leftrule=.75mm, rightrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, titlerule=0mm, colframe=quarto-callout-note-color-frame, coltitle=black, left=2mm, colback=white, leftrule=.75mm, arc=.35mm, bottomtitle=1mm, bottomrule=.15mm, opacityback=0, toptitle=1mm, rightrule=.15mm, opacitybacktitle=0.6]
 
@@
 
-\textbf{Output}: \texttt{optimised\_projection}
+\textbf{Output}: \texttt{optimized\_projection}
 
@@
 \textbf{return} the set of projections with the updated
-\texttt{current\_best} as the \texttt{optimised\_projection}
+\texttt{current\_best} as the \texttt{optimized\_projection}
 
@@
 
-\section{Assessing the optimisers}\label{sec-sim-deets}
+\section{Assessing the optimizers}\label{sec-sim-deets}
 
 This section explains the details of two simulation studies: (1)
-comparison between the existing Creeping Random Search (CRS) (Zhang et
-al. 2021; Laa and Cook 2020) and JSO, and (2) examining the factors
+comparison between the existing creeping random search (Zhang et al.
+2021; Laa and Cook 2020) and JSO, and (2) examining the factors
 affecting the performance of JSO.
@@
 
-The CRS is the main optimisation routine currently used for the guided
-tour. Here the two optimisers are compared on the task of finding the
+The CRS is the main optimization routine currently used for the guided
+tour. Here the two optimizers are compared on the task of finding the
 pipe using the \texttt{holes} index in data with dimensions,
@@
 The results of the simulation are collected using the data structure
-used in Zhang et al. (2021) for assessing PP optimisers. The design
+used in Zhang et al. (2021) for assessing PP optimizers. The design
 parameters are stored along with index value, projection basis, random
@@
 \caption{\label{fig-success-rate}How success rate is calculated,
-illustrated using the optimal projections from 50 optimisations of 8D
+illustrated using the optimal projections from 50 optimizations of 8D
 pipe data sorted by index value. The pipe shape is recognisable in the
@@
 4-12 dimensions covers the range of easy to relatively difficult for the
-optimisers to detect the structured projections. The same approach is
+optimizers to detect the structured projections. The same approach is
 used, that 50 samples are generated and the success rate is calculated
 at each level of the simulation design. A success rate of 0 indicates
-the optimiser did not come sufficiently close to the optimal projection.
+the optimizer did not come sufficiently close to the optimal projection.
 
@@
 This section summarises the findings from the simulations that compare
-the JSO performance with the existing CRS optimiser, and the
-relationship between optimisation success and hyper-parameter choices
+the JSO performance with the existing CRS optimizer, and the
+relationship between optimization success and hyper-parameter choices
 and PP index properties.
@@
 The performance of JSO is assessed relative to the existing CRS
-optimiser based on the optimal projection obtained. The final
-projections for the pipe data sets found by the two optimisers are shown
+optimizer based on the optimal projection obtained. The final
+projections for the pipe data sets found by the two optimizers are shown
 in Figure~\ref{fig-proj}, with rows corresponding to data dimension. The
@@
 different optimization, and hence compare results between the two
-optimisers. Generally, the JSO does a more consistent job of finding the
-pipe structure clearly. As the data dimension increases, both optimisers
+optimizers. Generally, the JSO does a more consistent job of finding the
+pipe structure clearly. As the data dimension increases, both optimizers
 struggle and less clearly capture the circle shape.
@@
 \caption{\label{fig-proportion}Summary of the relationship between JSO
-hyper-parameters on optimisation success, using the holes index for pipe
+hyper-parameters on optimization success, using the holes index for pipe
 data of dimensions 4-12. Bootstrap samples show the variability in
@@
 less gradual change in value. Least desirable is the \texttt{skinny}
-index which is noisier and the optimisation needs to be much closer to
+index which is noisier and the optimization needs to be much closer to
 the optimal angle to have a higher value. This illustrates how the new
@@
 of jellyfish and maximum number of tries. Interestingly, squintability
-and dimension strongly affect jellyfish optimisation success. The number
+and dimension strongly affect jellyfish optimization success. The number
 of jellies marginally affects success, but index smoothness, and
@@
 smoothness and squintability affect the JSO success rate. Here we
-explain how they can be used for new index and optimiser development.
-Using the JSO optimiser for PPGT requires a change in user behavior. For
-all the existing optimisers a single optimisation path is followed. The
-JSO has multiple optimisation paths, and needs additional tools to
+explain how they can be used for new index and optimizer development.
+Using the JSO optimizer for PPGT requires a change in user behavior. For
+all the existing optimizers a single optimization path is followed. The
+JSO has multiple optimization paths, and needs additional tools to
 incorporate into the PPGT, as explained here.
@@
 \texttt{search\_f\ =\ search\_jellyfish} in the guided tour. Unlike
-existing optimisers, the animation function \texttt{animate\_*()} won't
+existing optimizers, the animation function \texttt{animate\_*()} won't
 directly render the tour path for JSO due to the generation of multiple

### 2025-03-03 jso.tex
typo in title fixed

@@
 \hypersetup{
-  pdftitle={Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding optimization Choices},
+  pdftitle={Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimization Choices},
   pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
@@
 \title{Squintability and Other Metrics for Assessing Projection Pursuit
-Indexes, and Guiding optimization Choices}
+Indexes, and Guiding Optimization Choices}
 \author{H. Sherry Zhang$\textsuperscript{1}$, Dianne
@@
 
-\begin{tcolorbox}[enhanced jigsaw, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, titlerule=0mm, colframe=quarto-callout-note-color-frame, coltitle=black, left=2mm, colback=white, leftrule=.75mm, arc=.35mm, bottomtitle=1mm, bottomrule=.15mm, opacityback=0, toptitle=1mm, rightrule=.15mm, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, bottomtitle=1mm, titlerule=0mm, leftrule=.75mm, opacitybacktitle=0.6, colback=white, breakable, toptitle=1mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, bottomrule=.15mm, opacityback=0, left=2mm, rightrule=.15mm, colframe=quarto-callout-note-color-frame]
 

### 2025-03-04 jso.qmd
typo in title fixed
Sec 3.2

@@
 $A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix in the Stiefel
-manifold $\mathcal{A} = V_d(\mathbb{R}^p)$. The
-projection $Y = XA$ is a linear transformation of the $p$-dimensional data to
-a $d$-dimensional space. The index function
-$f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ defines a statistics to measure
-an pattern of interest in the projected data, such as
-deviation from normality, presence of clusters, or non-linear structure.
-For a fixed sample of data, PP finds the orthonormal matrix $A$ (also called 
-*projection basis* in PP literature) that maximizes the index value of the 
-projection:
+manifold $\mathcal{A} = V_d(\mathbb{R}^p)$. The projection $Y = XA$ is a
+linear transformation of the $p$-dimensional data to a $d$-dimensional
+space. The index function
+$f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ defines a statistics to
+measure an pattern of interest in the projected data, such as deviation
+from normality, presence of clusters, or non-linear structure. For a
+fixed sample of data, PP finds the orthonormal matrix $A$ (also called
+*projection basis* in PP literature) that maximizes the index value of
+the projection:
 
@@
 
-where $I_d$ is the identity matrix of dimension $d$.  
+where $I_d$ is the identity matrix of dimension $d$.\
 It is interesting to note that when using PP visually, one cares less
@@
 
-In this work, the potential for a Jellyfish Search Optimizer (JSO) [
-@chou_novel_2021; @rajwar_exhaustive_2023] for the PPGT is explored.
+In this work, the potential for a Jellyfish Search Optimizer (JSO)
+[@chou_novel_2021; @rajwar_exhaustive_2023] for the PPGT is explored.
 The JSO, inspired by the search behavior of jellyfish in the ocean, is a
 swarm-based metaheuristic designed to solve global optimization
-problems. Compared to other metaheuristic methods, JSO has demonstrated stronger
-search ability and faster convergence, and requires fewer tuning
-parameters. These practicalities make JSO a promising candidate for
-enhancing PP optimization.
+problems. Compared to other metaheuristic methods, JSO has demonstrated
+stronger search ability and faster convergence, and requires fewer
+tuning parameters. These practicalities make JSO a promising candidate
+for enhancing PP optimization.
 
@@
 squintability and smoothness, which is a new contribution for PP
-research. A series of simulation experiments are conducted using JSO to detect 
-different target patterns (pipe and sine-wave) with different PP indexes 
-(holes, MIC, TIC, dcor, loess, splines, skinny, and stringy), as well as JSO's 
-hyper-parameter choices (number of jellyfish and maximum number of
-tries). This work should facilitate better optimization for PP and
-guide the choice of optimizer when designing new PP indexes.
+research. A series of simulation experiments are conducted using JSO to
+detect different target patterns (pipe and sine-wave) with different PP
+indexes (holes, MIC, TIC, dcor, loess, splines, skinny, and stringy), as
+well as JSO's hyper-parameter choices (number of jellyfish and maximum
+number of tries). This work should facilitate better optimization for PP
+and guide the choice of optimizer when designing new PP indexes.
 
@@
 measure two properties of PP indexes, smoothness and squintability.
-@sec-JSO describes the new JSO for PP. @sec-sim-deets
-outlines two simulation experiments to assess JSO's performance: one
-comparing JSO's performance improvements relative to an existing
-optimizer, Creeping Random Search (CRS), and the other studying the
-impact of PP index properties, controlling for data dimension and JSO 
-hyper-parameters, on optimization performance, and
-@sec-sim-res presents the results. @sec-discussion discusses the
-implementation of JSO in the `tourr` package and the PP property
-calculation in the `ferrn` package. @sec-conclusion summarizes the work
-and provides suggestions for future directions.
+@sec-JSO describes the new JSO for PP. @sec-sim-deets outlines two
+simulation experiments to assess JSO's performance: one comparing JSO's
+performance improvements relative to an existing optimizer, Creeping
+Random Search (CRS), and the other studying the impact of PP index
+properties, controlling for data dimension and JSO hyper-parameters, on
+optimization performance, and @sec-sim-res presents the results.
+@sec-discussion discusses the implementation of JSO in the `tourr`
+package and the PP property calculation in the `ferrn` package.
+@sec-conclusion summarizes the work and provides suggestions for future
+directions.
 
@@
 
-```{=html}
 <!-- 
@@
 -->
-```
+
 There are numerous PP index functions available: introduced in @huber85,
@@
 optimizer needs to be able to handle index functions that are possibly
-not very smooth. In addition, because the target structure in the data might be
-relatively fine, the optimizer needs to be able to find maxima that
-occur with a small squint angle, that can only be seen from very close
-by. One last aspect that is useful is for an optimizer to return local
-maxima in addition to the global one because data can contain many
+not very smooth. In addition, because the target structure in the data
+might be relatively fine, the optimizer needs to be able to find maxima
+that occur with a small squint angle, that can only be seen from very
+close by. One last aspect that is useful is for an optimizer to return
+local maxima in addition to the global one because data can contain many
 different and interesting features.
@@
 invariance, and speed). Since not all index properties affect the
-optimization process, the focus of this work is on the first two properties,
-*smoothness* (@sec-smoothness) and *squintability* (@sec-squintability),
-for which metrics are proposed to quantify them.
+optimization process, the focus of this work is on the first two
+properties, *smoothness* (@sec-smoothness) and *squintability*
+(@sec-squintability), for which metrics are proposed to quantify them.
 
@@
 
-```{=tex}
 \begin{defn}[Sobolev space]\label{def:sobolev_space}
@@
 $f$, we propose an approach based on random fields. If a PP index
-function $f$ is evaluated at some random bases, then these random index values 
-can be interpreted as a random field, indexed by a space parameter, namely the
-random projection basis. This analogy suggests to use this random
-training sample to fit a spatial model. We propose to use a Gaussian
-process equipped with a Matérn covariance function, due to the
-connections between this model and Sobolev spaces, see for example
-@porcu2024matern.
+function $f$ is evaluated at some random bases, then these random index
+values can be interpreted as a random field, indexed by a space
+parameter, namely the random projection basis. This analogy suggests to
+use this random training sample to fit a spatial model. We propose to
+use a Gaussian process equipped with a Matérn covariance function, due
+to the connections between this model and Sobolev spaces, see for
+example @porcu2024matern.
 
@@
 
-```{=tex}
 \begin{defn}[Matérn covariance function]\label{def:matern}
@@
 \end{defn}
-```
+
 The Matérn covariance function can be expressed analytically when $\nu$
@@
 prior with Matérn covariance on a dataset generated by random
-evaluations of the index function. There exist several R packages,
-such as `GpGp` [@guinness2021gpgp] or `ExaGeoStatR` [@abdulah2023large],
-to fit the hyper-parameters of a GP covariance function on data, which
-is usually done by maximum likelihood estimation. In this project, the
+evaluations of the index function. There exist several R packages, such
+as `GpGp` [@guinness2021gpgp] or `ExaGeoStatR` [@abdulah2023large], to
+fit the hyper-parameters of a GP covariance function on data, which is
+usually done by maximum likelihood estimation. In this project, the
 `GpGp` package is used.
 
-```{=tex}
 \begin{defn}
@@
 \end{defn}
-```
+
 The value of the optimal smoothness parameter $\nu>0$ can be naturally
@@
 
-Squintability relates to the notion of squint angle and is defined based on how the index value $f(XA)$ changes with respect to the projection distance over the course of the optimization.
-
-```{=tex}
+From @barnett1981interpreting and @laa_using_2020, a large squint angle
+implies that the objective function value is close to optimal even when
+the target view to see the structure is far away. A small squint angle
+means that the PP index value improves substantially only when the
+target view is close by. As such, low squintability implies rapid
+improvement in the index value when near the target view. For PP, a
+small squint angle is considered to be undesirable because it means that
+the optimizer needs to be very close to be able to "see" the optimum.
+Thus, it could be difficult for the optimizer to find the optimum.
+
 \begin{defn}[squint angle]\label{def:squint-angle}
@@
 \end{defn}
+
+It is expected that for a PP index with high squint angle, the
+optimization \eqref{eq-optimization} should make substantial progress
+early on. Conversely, for a PP index with low squint angle, it might
+take a long while for the optimization to make substantial progress, as
+the candidate projections would need to be very close to the optimal one
+for the structure of the index function to be visible enough to be
+amenable to efficient optimization. This observation suggests that the
+half-point index value improvement percentage provides an appropriate
+mathematical definition of squintability, which matches the intuition
+behind this concept, while being amenable to numerical computation.
 
@@
 \end{defn}
-```
-
-```{=tex}
 \begin{defn}[squintability]\label{def:squintability}
@@
 \end{defn}
-```
-
-Remark that the amount by which the index value can improve from the starting matrix $A_0$ to the optimal one $A^*$ is given by $g(0)-g(r_{0})$.
-As a result, equation \eqref{eq-squintability} represent the proportion of this maximum improvement which has been achieved by the time the distance $r_0$ to the optimal matrix has been reduced by half ($r_0/2$).
-
-It is expected that this value should be high in the case of
-high squintability (fast increase in $g$ early on), and low in the
-case of low squintability (any substantial increase in $g$ happens very
-late, close to the optimal angle). This suggests that this half-point improvement metric should provide a sensible measure of
-squintability.
-
-From @barnett1981interpreting and @laa_using_2020, a large squint angle
-implies that the objective function value is close to optimal even when
-the perfect view to see the structure is far away. A small squint angle
-means that the PP index value improves substantially only when the
-perfect view is close by. As such, low squintability implies rapid
-improvement in the index value when near the perfect view. For PP, a
-small squint angle is considered to be undesirable because it means that
-the optimizer needs to be very close to be able to "see" the optimum.
-Thus, it could be difficult for the optimizer to find the optimum.
-<!--The
-mathematical formulation of this intuition is proposed below:-->
-
-It is expected that for a PP index with high squint angle, the
-optimization (@eq-optimization) should make substantial progress early
-on. Conversely, for a PP index with low squint angle, it might take a
-long while for the optimization to make substantial progress, as the
-candidate projections would need to be very close to the optimal one for
-the structure of the index function to be visible enough to be amenable
-to efficient optimization.
-This observation suggests that the half-point index value improvement percentage provides an appropriate mathematical definition of squintability, which matches the intuition behind this concept, while being amenable to numerical computation.
+
+Remark that the amount by which the index value can improve from the
+starting matrix $A_0$ to the optimal one $A^*$ is given by
+$g(0)-g(r_{0})$. As a result, equation \eqref{eq-squintability}
+represent the proportion of this maximum improvement which has been
+achieved by the time the distance $r_0$ to the optimal matrix has been
+reduced by half ($r_0/2$).
 
@@
 which depends on three additional parameters, $\theta_1$, $\theta_2$,
-and $r_0$, such that $g(0)=\theta_1$ and $g(r_0)=\theta_4$.
-Under the parametric model (@eq-parametric), the squintability metric
+and $r_0$, such that $g(0)=\theta_1$ and $g(r_0)=\theta_4$. Under the
+parametric model (@eq-parametric), the squintability metric
 \eqref{eq-squintability} can be shown to be equal to $$
@@
 nonparametric way, for example by fitting $g$ using kernel regression,
-then numerically estimate $\varsigma$ from its definition \eqref{eq-squintability}.
+then numerically estimate $\varsigma$ from its definition
+\eqref{eq-squintability}.
 
@@
 
-To solve the optimization problem embedded in the PP guided tour, a
-starting projection, an index function, the number of jellyfish, and the
-maximum number of trials (tries) are provided as input. Then, the
-current projection is evaluated by the index function. The projection is
-then moved in a direction determined by a random factor, influenced by
-how far along we are in the optimization process. Occasionally,
-completely new directions may be taken like a jellyfish might with ocean
-currents. A new projection is accepted if it is an improvement compared
-to the current one, rejected otherwise. This process continues and
-iteratively improves the projection, until the pre-specified maximum
-number of trials is reached.
+To solve the optimization problem embedded in the PPGT, a starting
+projection, an index function, the number of jellyfish, and the maximum
+number of trials (tries) are provided as input. Then, the current
+projection is evaluated by the index function. The projection is then
+moved in a direction determined by a random factor, influenced by how
+far along we are in the optimization process. Occasionally, completely
+new directions may be taken like a jellyfish might with ocean currents.
+A new projection is accepted if it is an improvement compared to the
+current one, rejected otherwise. This process continues and iteratively
+improves the projection, until the pre-specified maximum number of
+trials is reached.
 
@@
 
-This section explains the details of two simulation studies: (1) comparison between the existing creeping random search [@RJ-2021-105; @laa_using_2020] and JSO, and (2) examining the factors affecting the performance of JSO.
+This section explains the details of two simulation studies: (1)
+comparison between the existing creeping random search [@RJ-2021-105;
+@laa_using_2020] and JSO, and (2) examining the factors affecting the
+performance of JSO.
 
@@
 
-The CRS is the main optimization routine currently used for the guided tour.
-Here the two optimizers are compared on the task of finding the pipe using the `holes` index in data with dimensions, $d = 6, 8, 10, 12$.
-Fifty simulated data sets are used for each dimension.
-JSO uses 100 jellyfish with a maximum of 100 tries, while the CRS allows a maximum of 1000 tries at each iteration before the algorithm terminates.
-These choices enable fair comparison between CRS and JSO, that conforms to how they are used in practice.
-
-The performance of the opt misers is measured by the success rate, which is defined as the proportion of simulations that achieves a final index value within 0.05
-of the best index value found among all 50 simulations.
-@fig-success-rate illustrates why the choice of 0.05 is reasonable: the pipe is not recognizable in the projected data.
-This is motivated by @laa_using_2020's approach to investigating PP indexes.
-
-The results of the simulation are collected using the data structure used in @RJ-2021-105 for assessing PP optimizers.
-The design parameters are stored along with index value, projection basis, random seed, and computation time.
+The CRS is the main optimization routine currently used for the guided
+tour. Here the two optimizers are compared on the task of finding the
+pipe using the `holes` index in data with dimensions,
+$d = 6, 8, 10, 12$. Fifty simulated data sets are used for each
+dimension. JSO uses 100 jellyfish with a maximum of 100 tries, while the
+CRS allows a maximum of 1000 tries at each iteration before the
+algorithm terminates. These choices enable fair comparison between CRS
+and JSO, that conforms to how they are used in practice.
+
+The performance of the opt misers is measured by the success rate, which
+is defined as the proportion of simulations that achieves a final index
+value within 0.05 of the best index value found among all 50
+simulations. @fig-success-rate illustrates why the choice of 0.05 is
+reasonable: the pipe is not recognizable in the projected data. This is
+motivated by @laa_using_2020's approach to investigating PP indexes.
+
+The results of the simulation are collected using the data structure
+used in @RJ-2021-105 for assessing PP optimizers. The design parameters
+are stored along with index value, projection basis, random seed, and
+computation time.
 
@@
 
-To examine the performance of JSO relative to hyper-parameter choices, the `holes` index is used to find the 2D pipe in data of dimensions 4-12.
-Additional variables are Gaussian noise.
-The levels of hyper-parameters are 20, 50, and 100
-jellyfish and a maximum of 50 and 100 attempts. Fifty samples of each data set are generated, and the success rate is calculated in the same manner as described in @sec-app-1.
-
-This set of simulations is expanded to assess the performance of JSO relative to the index properties, with a second data set type and additional PP indexes.
-The second data set has a sine wave in two dimensions and Gaussian noise in the remaining dimensions.
-Six additional PP indexes that should be able to extract the sine wave are used: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`, following @laa_using_2020.
-(The supplementary material includes definitions of these PP indexes, and descriptions of the pipe and sine wave data simulation.).
-The JSO hyper-parameters, number of jellies and maximum tries, are also included.
-This results in a total of `r nrow(sim_df)` scenarios, comprising of `r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and `r nrow(filter(sim_df, index != "holes"))` on the sine wave data.
-The choice of 4-12 dimensions covers the range of easy to relatively difficult for the optimizers to detect the structured projections.
-<!-- A list of combinations of the data shape, dimensions and PP indexes that are considered in this study is provided in @tbl-smoothness-squintability.--> The same approach is used, that 50 samples are generated and the success rate is calculated at each level of the simulation design.
-A success rate of 0 indicates the optimizer did not come sufficiently close to the optimal projection.
+To examine the performance of JSO relative to hyper-parameter choices,
+the `holes` index is used to find the 2D pipe in data of dimensions
+4-12. Additional variables are Gaussian noise. The levels of
+hyper-parameters are 20, 50, and 100 jellyfish and a maximum of 50 and
+100 attempts. Fifty samples of each data set are generated, and the
+success rate is calculated in the same manner as described in
+@sec-app-1.
+
+This set of simulations is expanded to assess the performance of JSO
+relative to the index properties, with a second data set type and
+additional PP indexes. The second data set has a sine wave in two
+dimensions and Gaussian noise in the remaining dimensions. Six
+additional PP indexes that should be able to extract the sine wave are
+used: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`,
+following @laa_using_2020. (The supplementary material includes
+definitions of these PP indexes, and descriptions of the pipe and sine
+wave data simulation.). The JSO hyper-parameters, number of jellies and
+maximum tries, are also included. This results in a total of
+`r nrow(sim_df)` scenarios, comprising of
+`r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and
+`r nrow(filter(sim_df, index != "holes"))` on the sine wave data. The
+choice of 4-12 dimensions covers the range of easy to relatively
+difficult for the optimizers to detect the structured projections.
+<!-- A list of combinations of the data shape, dimensions and PP indexes that are considered in this study is provided in @tbl-smoothness-squintability.-->
+The same approach is used, that 50 samples are generated and the success
+rate is calculated at each level of the simulation design. A success
+rate of 0 indicates the optimizer did not come sufficiently close to the
+optimal projection.
 
@@
 
-@fig-smoothness describes the procedure to compute smoothness.
-Index values are calculated for each of 500 simulated random bases.
-A Gaussian process model is fitted to the resulting data to obtain the smoothness measure for the index, as described in @sec-smoothness.
+@fig-smoothness describes the procedure to compute smoothness. Index
+values are calculated for each of 500 simulated random bases. A Gaussian
+process model is fitted to the resulting data to obtain the smoothness
+measure for the index, as described in @sec-smoothness.
 
@@
 
-@fig-squintability describes the procedure to compute squintability. Index values and
-projection distances are computed for the projections along interpolated paths between 50 random bases and the optimal basis.
-An interpolation step size of 0.005 is used.
-The projection distances are binned to 0.005 intervals and index values are averaged in each bin.
-The four-parameter scaled logistic function (@eq-parametric) is fitted to the index values against projection distances, estimated by non-linear least squares to obtain the squintability measure as Equation (@eq-squintability-parametric).
-
-A generalized linear model is fitted using a quasibinomial family and a logit link function to assess the factors affecting success rate.  Predictors are smoothness,
-squintability, data dimension and JSO hyper-parameters. <!-- The data is
+@fig-squintability describes the procedure to compute squintability.
+Index values and projection distances are computed for the projections
+along interpolated paths between 50 random bases and the optimal basis.
+An interpolation step size of 0.005 is used. The projection distances
+are binned to 0.005 intervals and index values are averaged in each bin.
+The four-parameter scaled logistic function (@eq-parametric) is fitted
+to the index values against projection distances, estimated by
+non-linear least squares to obtain the squintability measure as Equation
+(@eq-squintability-parametric).
+
+A generalized linear model is fitted using a quasibinomial family and a
+logit link function to assess the factors affecting success rate.
+Predictors are smoothness, squintability, data dimension and JSO
+hyper-parameters. <!-- The data is
 pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
@@
 
-This section summarises the findings from the simulations that compare the JSO performance with the existing CRS optimizer, and the relationship between optimization success and hyper-parameter choices and PP index properties.
+This section summarises the findings from the simulations that compare
+the JSO performance with the existing CRS optimizer, and the
+relationship between optimization success and hyper-parameter choices
+and PP index properties.
 
@@
 
-The performance of JSO is assessed relative to the existing CRS optimizer based on the optimal projection obtained.
-The final projections for the pipe data sets found by the two optimizers are shown in @fig-proj, with rows corresponding to data dimension.
-The 50 simulations in each data dimension are sorted by index value, and the projections corresponding to each 10th quantile value are shown, with minimum at left and maximum at right.
-Index value is printed on the plot.
-The purpose is to summarize the views of the data resulting from different optimization, and hence compare results between the two optimizers.
-Generally, the JSO does a more consistent job of finding the pipe structure clearly.
-As the data dimension increases, both optimizers struggle and less clearly capture the circle shape.
+The performance of JSO is assessed relative to the existing CRS
+optimizer based on the optimal projection obtained. The final
+projections for the pipe data sets found by the two optimizers are shown
+in @fig-proj, with rows corresponding to data dimension. The 50
+simulations in each data dimension are sorted by index value, and the
+projections corresponding to each 10th quantile value are shown, with
+minimum at left and maximum at right. Index value is printed on the
+plot. The purpose is to summarize the views of the data resulting from
+different optimization, and hence compare results between the two
+optimizers. Generally, the JSO does a more consistent job of finding the
+pipe structure clearly. As the data dimension increases, both optimizers
+struggle and less clearly capture the circle shape.
 
@@
 case. As the number of jellyfish and maximum tries increase, the success
-rate also increases.
-For problems with relatively lower dimensional (4, 6) search spaces, small parameter values (20 jellyfish and a maximum of 50 tries) are sufficient to achieve a high success rate.
-Many more jellyfish (100 jellyfish and a maximum of 100 tries) are needed to obtain a mediocre success rate for dimensions 10 and 12.
-This suggests that even JSO will struggle with high dimensions, and require substantial more computational time to have an acceptable success rate, especially for difficult PP indexes based on scagnostics.
+rate also increases. For problems with relatively lower dimensional (4,
+6) search spaces, small parameter values (20 jellyfish and a maximum of
+50 tries) are sufficient to achieve a high success rate. Many more
+jellyfish (100 jellyfish and a maximum of 100 tries) are needed to
+obtain a mediocre success rate for dimensions 10 and 12. This suggests
+that even JSO will struggle with high dimensions, and require
+substantial more computational time to have an acceptable success rate,
+especially for difficult PP indexes based on scagnostics.
 <!-- dimensions) where a more complex search is necessary.
@@
 
-Smoothness and squintability are calculated across a collection of
-pipe and sine wave finding problems to construct the
-relationship between success rate, JSO hyper-parameters, and index
-properties. @tbl-smoothness-squintability presents values computed for a subset of PP indexes on the sine wave data for dimensions 4-8. (A full table of the parameters
-estimated from the Gaussian process (outputscale $\eta$, lengthscale
-$\ell$, smoothness $\nu$, and nugget $\sigma$) and from the scaled
-logistic function ($\theta_1$ to $\theta_4$) for calculating smoothness
-and squintability for all PP indexes and data combinations is included in the supplementary materials). The column $\nu$ is
-used as the smoothness metric and the column $\varsigma$ is calculated
-as equation (@eq-squintability-parametric) as the squintability metric.
-
-@fig-idx-proj-dist shows a sample of traces from the indexes in @tbl-smoothness-squintability for the 6D sine wave data.
-The traces the PP index values computed for projections along an interpolated path between a noise projection and the optimal projection (sine wave).
-(Note index values were self-standardized to range between 0 and 1, for comparison.) The `splines` index is very smooth with a gradual increase in value between the noise projection and the optimal projection.
-The `TIC` index is slightly less smooth and slightly less gradual change in value.
-Least desirable is the `skinny` index which is noisier and the optimization needs to be much closer to the optimal angle to have a higher value.
-This illustrates how the new measures for squintability and smoothness work to describe the PP indexes.
+Smoothness and squintability are calculated across a collection of pipe
+and sine wave finding problems to construct the relationship between
+success rate, JSO hyper-parameters, and index properties.
+@tbl-smoothness-squintability presents values computed for a subset of
+PP indexes on the sine wave data for dimensions 4-8. (A full table of
+the parameters estimated from the Gaussian process (outputscale $\eta$,
+lengthscale $\ell$, smoothness $\nu$, and nugget $\sigma$) and from the
+scaled logistic function ($\theta_1$ to $\theta_4$) for calculating
+smoothness and squintability for all PP indexes and data combinations is
+included in the supplementary materials). The column $\nu$ is used as
+the smoothness metric and the column $\varsigma$ is calculated as
+equation (@eq-squintability-parametric) as the squintability metric.
+
+@fig-idx-proj-dist shows a sample of traces from the indexes in
+@tbl-smoothness-squintability for the 6D sine wave data. The traces the
+PP index values computed for projections along an interpolated path
+between a noise projection and the optimal projection (sine wave). (Note
+index values were self-standardized to range between 0 and 1, for
+comparison.) The `splines` index is very smooth with a gradual increase
+in value between the noise projection and the optimal projection. The
+`TIC` index is slightly less smooth and slightly less gradual change in
+value. Least desirable is the `skinny` index which is noisier and the
+optimization needs to be much closer to the optimal angle to have a
+higher value. This illustrates how the new measures for squintability
+and smoothness work to describe the PP indexes.
 
@@
 model using the proportion of success as the response and smoothness,
-squintability, dimension, number of jellyfish and maximum
-number of tries, as predictors. The fit suggests that the JSO success
-rate is only affected by squintability, dimension and number of jellies.
-As expected, the JSO success rate is higher with higher squintability and/or more jellyfish, and lower when dimension is higher (higher-dimensional optimization is more difficult).
-Interestingly, the JSO is unaffected by the smoothness of the index function.
-This is consistent with the way random search algorithms jump from value to value without taking local regularity into account, as opposed to gradient-based optimizers for example.
-Allowing
-for more tries also does not affect success rate significantly. A unit increase in squintability increases the success rate by
-`r round(100 * (exp(broom::tidy(mod1)$estimate[3])-1), 0)`%. As dimension
-increases by one the success rate almost halves. Increasing the number of
-jellies by 10 increases the success rate by
-`r round(100 * (exp(broom::tidy(mod1)$estimate[5])-1), 0)`%. @tbl-joint-output shows the fit statistics supporting that the model is reasonably strong.
-
-Note that the addition of interaction terms for squintability and data dimension, and smoothness and data dimension, improves the model a small amount, but does not change the interpretation above.
+squintability, dimension, number of jellyfish and maximum number of
+tries, as predictors. The fit suggests that the JSO success rate is only
+affected by squintability, dimension and number of jellies. As expected,
+the JSO success rate is higher with higher squintability and/or more
+jellyfish, and lower when dimension is higher (higher-dimensional
+optimization is more difficult). Interestingly, the JSO is unaffected by
+the smoothness of the index function. This is consistent with the way
+random search algorithms jump from value to value without taking local
+regularity into account, as opposed to gradient-based optimizers for
+example. Allowing for more tries also does not affect success rate
+significantly. A unit increase in squintability increases the success
+rate by `r round(100 * (exp(broom::tidy(mod1)$estimate[3])-1), 0)`%. As
+dimension increases by one the success rate almost halves. Increasing
+the number of jellies by 10 increases the success rate by
+`r round(100 * (exp(broom::tidy(mod1)$estimate[5])-1), 0)`%.
+@tbl-joint-output shows the fit statistics supporting that the model is
+reasonably strong.
+
+Note that the addition of interaction terms for squintability and data
+dimension, and smoothness and data dimension, improves the model a small
+amount, but does not change the interpretation above.
 
@@
 
-This study intends to serve as an illustration that the proposed measures effectively captures the key factors influencing the success rate of PP.
+This study intends to serve as an illustration that the proposed
+measures effectively captures the key factors influencing the success
+rate of PP.
 <!--The interactions between the two variables were not examined in this study due to the limited number of data points, nor was the linearity assumption of the relationship with success rate explicitly tested.-->
@@
 explain how they can be used for new index and optimizer development.
-Using the JSO optimizer for PPGT requires a change in user behavior.
-For all the existing optimizers a single optimization path is followed.
-The JSO has multiple optimization paths, and needs additional tools to
+Using the JSO optimizer for PPGT requires a change in user behavior. For
+all the existing optimizers a single optimization path is followed. The
+JSO has multiple optimization paths, and needs additional tools to
 incorporate into the PPGT, as explained here.
@@
 bases to sample. The output of `sample_bases()` is a data frame with a
-list-column of sampled basis matrix and index value. Particularization is
-available to speed up the index value computation through the
+list-column of sampled basis matrix and index value. Particularization
+is available to speed up the index value computation through the
 `parallel = TRUE` argument.
@@
 -->
-```
+
 # Acknowledgement
@@
 
-The R packages used in this work include: `tidyr` [@tidyr], `dplyr` [@dplyr], 
-`ggplot2` [@ggplot2], `knitr` [@knitr],`gtsummary` [@gtsummary],
-`patchwork` [@patchwork], `ggh4x` [@ggh4x], `broom` [@broom], 
-`kableExtra` [@kabelextra], `ferrn` [@RJ-2021-105], and `cassowaryr` [@cassowaryr].
+The R packages used in this work include: `tidyr` [@tidyr], `dplyr`
+[@dplyr], `ggplot2` [@ggplot2], `knitr` [@knitr],`gtsummary`
+[@gtsummary], `patchwork` [@patchwork], `ggh4x` [@ggh4x], `broom`
+[@broom], `kableExtra` [@kabelextra], `ferrn` [@RJ-2021-105], and
+`cassowaryr` [@cassowaryr].
 

### 2025-03-04 jso.tex
Sec 3.2

--- 2025-03-03 jso.tex
@@
 
-Squintability relates to the notion of squint angle and is defined based
-on how the index value \(f(XA)\) changes with respect to the projection
-distance over the course of the optimization.
+From Tukey and Tukey (1981) and Laa and Cook (2020), a large squint
+angle implies that the objective function value is close to optimal even
+when the target view to see the structure is far away. A small squint
+angle means that the PP index value improves substantially only when the
+target view is close by. As such, low squintability implies rapid
+improvement in the index value when near the target view. For PP, a
+small squint angle is considered to be undesirable because it means that
+the optimizer needs to be very close to be able to ``see'' the optimum.
+Thus, it could be difficult for the optimizer to find the optimum.
 
@@
 \end{defn}
+
+It is expected that for a PP index with high squint angle, the
+optimization (\eqref{eq-optimization}) should make substantial progress
+early on. Conversely, for a PP index with low squint angle, it might
+take a long while for the optimization to make substantial progress, as
+the candidate projections would need to be very close to the optimal one
+for the structure of the index function to be visible enough to be
+amenable to efficient optimization. This observation suggests that the
+half-point index value improvement percentage provides an appropriate
+mathematical definition of squintability, which matches the intuition
+behind this concept, while being amenable to numerical computation.
 
@@
 \end{defn}
-
 \begin{defn}[squintability]\label{def:squintability}
@@
 reduced by half (\(r_0/2\)).
-
-It is expected that this value should be high in the case of high
-squintability (fast increase in \(g\) early on), and low in the case of
-low squintability (any substantial increase in \(g\) happens very late,
-close to the optimal angle). This suggests that this half-point
-improvement metric should provide a sensible measure of squintability.
-
-From Tukey and Tukey (1981) and Laa and Cook (2020), a large squint
-angle implies that the objective function value is close to optimal even
-when the perfect view to see the structure is far away. A small squint
-angle means that the PP index value improves substantially only when the
-perfect view is close by. As such, low squintability implies rapid
-improvement in the index value when near the perfect view. For PP, a
-small squint angle is considered to be undesirable because it means that
-the optimizer needs to be very close to be able to ``see'' the optimum.
-Thus, it could be difficult for the optimizer to find the optimum.
-
-It is expected that for a PP index with high squint angle, the
-optimization (\ref{eq-optimization}) should make substantial progress
-early on. Conversely, for a PP index with low squint angle, it might
-take a long while for the optimization to make substantial progress, as
-the candidate projections would need to be very close to the optimal one
-for the structure of the index function to be visible enough to be
-amenable to efficient optimization. This observation suggests that the
-half-point index value improvement percentage provides an appropriate
-mathematical definition of squintability, which matches the intuition
-behind this concept, while being amenable to numerical computation.
 
@@
 
-\begin{tcolorbox}[enhanced jigsaw, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, titlerule=0mm, colframe=quarto-callout-note-color-frame, coltitle=black, left=2mm, colback=white, leftrule=.75mm, arc=.35mm, bottomtitle=1mm, bottomrule=.15mm, opacityback=0, toptitle=1mm, rightrule=.15mm, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, rightrule=.15mm, opacitybacktitle=0.6, left=2mm, leftrule=.75mm, opacityback=0, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, titlerule=0mm, bottomrule=.15mm, bottomtitle=1mm, toprule=.15mm, breakable, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}]
 

### 2025-03-04 jso.qmd
before Section 6

@@
 ---
-title: Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding optimization Choices
+title: Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimization Choices
 author:
@@
 
-To solve the optimization problem embedded in the PPGT, a starting
+To solve the optimization problem embedded in the PPGT with JSO, a starting
 projection, an index function, the number of jellyfish, and the maximum
@@
 
-> Calculate the time control value, $c_t$, based on `current_idx` and
+> Calculate the time control value, $c_t$, based on `trial_id` and
 > `max_trial`
@@
 > >
-> > > For each projection, compare with a random jellyfish and adjust
-> > > towards or away from it (active)
+> > > For each projection, compare with a random jellyfish in `current_projections`
+> > > and adjust towards or away from it based on their corresponding `current_idx` (active)
 >
@@
 
-> **if** `trial_id` $\ge$ `max_trial`, print the last best projection
 > **exit**
@@
 This section explains the details of two simulation studies: (1)
-comparison between the existing creeping random search [@RJ-2021-105;
-@laa_using_2020] and JSO, and (2) examining the factors affecting the
-performance of JSO.
+a comparison between the JSO and an existing optimizer, creeping random
+search [@RJ-2021-105; @laa_using_2020], and (2) a modelling study to examine 
+how factors (smoothness, squintability, data dimension, and JSO hyper-parameters)
+affect its success rate in detecting holes and sine-wave pattern across 
+a collection of index functions.
 
@@
 The CRS is the main optimization routine currently used for the guided
-tour. Here the two optimizers are compared on the task of finding the
-pipe using the `holes` index in data with dimensions,
-$d = 6, 8, 10, 12$. Fifty simulated data sets are used for each
-dimension. JSO uses 100 jellyfish with a maximum of 100 tries, while the
-CRS allows a maximum of 1000 tries at each iteration before the
-algorithm terminates. These choices enable fair comparison between CRS
-and JSO, that conforms to how they are used in practice.
-
-The performance of the opt misers is measured by the success rate, which
+tour. Here the two optimizers are compared on the task of finding the 2D
+pipe structure using the `holes` index in data at dimensions
+6, 8, 10, and 12. Additional variables are Gaussian noise. JSO uses 100 
+jellyfish with 100 iterations, while CRS allows up to 1000 trials per 
+iteration terminates if no better projection is found within these 
+trials. These choices enable fair comparison between CRS and JSO, 
+that conforms to how they are used in practice. Each task is 
+repeated 50 times to evaluate the performance between the two optimizers.
+
+The performance of the optimizers is measured by the success rate, which
 is defined as the proportion of simulations that achieves a final index
@@
 used in @RJ-2021-105 for assessing PP optimizers. The design parameters
-are stored along with index value, projection basis, random seed, and
-computation time.
+are stored along with index value, projection basis, and random seed used.
 
@@
 #| fig.cap: "How success rate is calculated, illustrated using the optimal projections from 50 optimizations of 8D pipe data sorted by index value.
-The pipe shape is recognisable in the projection index values between 0.933-0.969.
+The pipe shape is recognizable in the projection index values between 0.933-0.969.
 Of the 50 simulations, 43 achieved an index value within 0.05 of the best, resulting in a success rate of 0.86."
@@
 the `holes` index is used to find the 2D pipe in data of dimensions
-4-12. Additional variables are Gaussian noise. The levels of
-hyper-parameters are 20, 50, and 100 jellyfish and a maximum of 50 and
-100 attempts. Fifty samples of each data set are generated, and the
-success rate is calculated in the same manner as described in
-@sec-app-1.
+4, 6, 8, 10, and 12. The levels of hyper-parameters are 20, 50, and 
+100 jellyfish and a maximum of 50 and 100 attempts. Each task is repeated
+fifty times to calculate the success rate as described in @sec-app-1.
 
@@
 additional PP indexes. The second data set has a sine wave in two
-dimensions and Gaussian noise in the remaining dimensions. Six
+dimensions and Gaussian noise in the remaining dimensions. Seven
 additional PP indexes that should be able to extract the sine wave are
-used: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`,
+used: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `splines2d`, `skinny` and `stringy2`,
 following @laa_using_2020. (The supplementary material includes
@@
 <!-- A list of combinations of the data shape, dimensions and PP indexes that are considered in this study is provided in @tbl-smoothness-squintability.-->
-The same approach is used, that 50 samples are generated and the success
-rate is calculated at each level of the simulation design. A success
-rate of 0 indicates the optimizer did not come sufficiently close to the
-optimal projection.
+The same approach is used, that each experiment is repeated 50 times and success
+rate is calculated at each level of the simulation design.
 
@@
 #| out.width: 100%
-#| fig.cap: "Steps for calculating smoothness in a projection pursuit problem.
-Given the target shape, data dimension and the index function: 1) sample random bases given the orthonormality contraint and calculate their corresponding index values, and 2) fit a Gaussian process model of index values against the sampled bases to obtain the smoothness measure."
+#| fig.cap: "Steps for calculating smoothness in a projection pursuit problem: 1) sample random bases given the orthonormality contraint and calculate their corresponding index values, 2) fit a Gaussian process model of index values against the bases to obtain the smoothness measure, 3) take the smoothness parameter from the Gaussian process as the smoothness metric."
 # basis_smoothness <- sample_bases(idx = "splines2d", n_basis = 1000)
@@
 
-@fig-smoothness describes the procedure to compute smoothness. Index
-values are calculated for each of 500 simulated random bases. A Gaussian
-process model is fitted to the resulting data to obtain the smoothness
+Recall that smoothness and squintability are metrics used to characterize the index 
+function for PP tasks and are invariant to the optimizer and its 
+hyper-parameters. Thus simulations with different JSO hyper-parameters 
+for the same PP task share the same smoothness value. In total, smoothness and
+squintability metrics are calculated on the 22 PP tasks in our simulation.
+
+@fig-smoothness describes the procedure to compute smoothness. For each PP task,
+500 random bases are simulated and index values are calculated for each. A Gaussian
+process model is then fitted to the resulting data to obtain the smoothness
 measure for the index, as described in @sec-smoothness.
@@
 #| fig.cap: |
-#|   Steps for calculating squintability in a projection pursuit problem.
-Given the target shape, data dimension and the index function: 1) sample random bases given the orthonormality and projection distance contraint and calculates their corresponding index values, 2) interpolate the sampled bases to the optimal basis and calculate the projection distance and the index value.
-3) bin the index values by projection distances through averaging index value, 4) fit the scaled sigmoid function in equation \eqref{eq-squintability} to the binned index values against projection distances using non-linear least square to obtian the squintability measure using equation (@eq-squintability-parametric).
+#|   Steps for calculating squintability in a projection pursuit problem: 1) sample random bases given the orthonormality and projection distance contraint and calculates their corresponding index values, 2) interpolate the sampled bases to the optimal basis and calculate the projection distance and the index value.
+3) average the index values by projection distances at each 0.005 distance bin, 4) fit the scaled sigmoid function in Equation \eqref{eq-logistic} and \eqref{eq-parametric} , 5) calculate the squintability metric using Equation (@eq-squintability-parametric).
 # p1 <- sim_sine_6d_spline_projdist |> 
@@
 
-@fig-squintability describes the procedure to compute squintability.
-Index values and projection distances are computed for the projections
-along interpolated paths between 50 random bases and the optimal basis.
-An interpolation step size of 0.005 is used. The projection distances
-are binned to 0.005 intervals and index values are averaged in each bin.
-The four-parameter scaled logistic function (@eq-parametric) is fitted
-to the index values against projection distances, estimated by
-non-linear least squares to obtain the squintability measure as Equation
+@fig-squintability describes the procedure to compute squintability. For each 
+PP task, 50 random bases are generated and interpolated to the optimal basis 
+with a step size of 0.005. The index values are averaged over the distance 
+window of 0.005. Index values are calculated for all the interpolated bases and 
+averaged over the distance bin of 0.005. The logisitic function in
+Equation (@eq-logistic) and (@eq-parametric) is fitted with non-linear least squares 
+and calculate the squintability measure using Equation 
 (@eq-squintability-parametric).
@@
 A generalized linear model is fitted using a quasibinomial family and a
-logit link function to assess the factors affecting success rate.
-Predictors are smoothness, squintability, data dimension and JSO
+logit link function to assess the factors affecting the success rate.
+Predictors are smoothness, squintability, data dimension, and JSO
 hyper-parameters. <!-- The data is

### 2025-03-04 jso.tex
before Section 6

@@
 It is expected that for a PP index with high squint angle, the
-optimization (\eqref{eq-optimization}) should make substantial progress
+optimization \eqref{eq-optimization} should make substantial progress
 early on. Conversely, for a PP index with low squint angle, it might
@@
 
-To solve the optimization problem embedded in the PP guided tour, a
+To solve the optimization problem embedded in the PPGT with JSO, a
 starting projection, an index function, the number of jellyfish, and the
@@
 
-\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, rightrule=.15mm, opacitybacktitle=0.6, left=2mm, leftrule=.75mm, opacityback=0, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, coltitle=black, titlerule=0mm, bottomrule=.15mm, bottomtitle=1mm, toprule=.15mm, breakable, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}]
+\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, leftrule=.75mm, breakable, coltitle=black, colback=white, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, opacitybacktitle=0.6, opacityback=0, left=2mm, bottomtitle=1mm, rightrule=.15mm, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm]
 
@@
 \begin{quote}
-Calculate the time control value, \(c_t\), based on
-\texttt{current\_idx} and \texttt{max\_trial}
+Calculate the time control value, \(c_t\), based on \texttt{trial\_id}
+and \texttt{max\_trial}
 \end{quote}
@@
 \begin{quote}
-For each projection, compare with a random jellyfish and adjust towards
-or away from it (active)
+For each projection, compare with a random jellyfish in
+\texttt{current\_projections} and adjust towards or away from it based
+on their corresponding \texttt{current\_idx} (active)
 \end{quote}
@@
 \begin{quote}
-\textbf{if} \texttt{trial\_id} \(\ge\) \texttt{max\_trial}, print the
-last best projection \textbf{exit}
+\textbf{exit}
 \end{quote}
@@
 
-This section explains the details of two simulation studies: (1)
-comparison between the existing creeping random search (Zhang et al.
-2021; Laa and Cook 2020) and JSO, and (2) examining the factors
-affecting the performance of JSO.
+This section explains the details of two simulation studies: (1) a
+comparison between the JSO and an existing optimizer, creeping random
+search (Zhang et al. 2021; Laa and Cook 2020), and (2) a modelling study
+to examine how factors (smoothness, squintability, data dimension, and
+JSO hyper-parameters) affect its success rate in detecting holes and
+sine-wave pattern across a collection of index functions.
 
@@
 The CRS is the main optimization routine currently used for the guided
-tour. Here the two optimizers are compared on the task of finding the
-pipe using the \texttt{holes} index in data with dimensions,
-\(d = 6, 8, 10, 12\). Fifty simulated data sets are used for each
-dimension. JSO uses 100 jellyfish with a maximum of 100 tries, while the
-CRS allows a maximum of 1000 tries at each iteration before the
-algorithm terminates. These choices enable fair comparison between CRS
-and JSO, that conforms to how they are used in practice.
-
-The performance of the opt misers is measured by the success rate, which
+tour. Here the two optimizers are compared on the task of finding the 2D
+pipe structure using the \texttt{holes} index in data at dimensions 6,
+8, 10, and 12. Additional variables are Gaussian noise. JSO uses 100
+jellyfish with 100 iterations, while CRS allows up to 1000 trials per
+iteration terminates if no better projection is found within these
+trials. These choices enable fair comparison between CRS and JSO, that
+conforms to how they are used in practice. Each task is repeated 50
+times to evaluate the performance between the two optimizers.
+
+The performance of the optimizers is measured by the success rate, which
 is defined as the proportion of simulations that achieves a final index
@@
 used in Zhang et al. (2021) for assessing PP optimizers. The design
-parameters are stored along with index value, projection basis, random
-seed, and computation time.
+parameters are stored along with index value, projection basis, and
+random seed used.
 
@@
 illustrated using the optimal projections from 50 optimizations of 8D
-pipe data sorted by index value. The pipe shape is recognisable in the
+pipe data sorted by index value. The pipe shape is recognizable in the
 projection index values between 0.933-0.969. Of the 50 simulations, 43
@@
 the \texttt{holes} index is used to find the 2D pipe in data of
-dimensions 4-12. Additional variables are Gaussian noise. The levels of
-hyper-parameters are 20, 50, and 100 jellyfish and a maximum of 50 and
-100 attempts. Fifty samples of each data set are generated, and the
-success rate is calculated in the same manner as described in
+dimensions 4, 6, 8, 10, and 12. The levels of hyper-parameters are 20,
+50, and 100 jellyfish and a maximum of 50 and 100 attempts. Each task is
+repeated fifty times to calculate the success rate as described in
 Section~\ref{sec-app-1}.
@@
 additional PP indexes. The second data set has a sine wave in two
-dimensions and Gaussian noise in the remaining dimensions. Six
+dimensions and Gaussian noise in the remaining dimensions. Seven
 additional PP indexes that should be able to extract the sine wave are
 used: \texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
-\texttt{spline}, and \texttt{stringy}, following Laa and Cook (2020).
-(The supplementary material includes definitions of these PP indexes,
-and descriptions of the pipe and sine wave data simulation.). The JSO
-hyper-parameters, number of jellies and maximum tries, are also
-included. This results in a total of 76 scenarios, comprising of 30
-computed on the pipe data and 46 on the sine wave data. The choice of
-4-12 dimensions covers the range of easy to relatively difficult for the
-optimizers to detect the structured projections. The same approach is
-used, that 50 samples are generated and the success rate is calculated
-at each level of the simulation design. A success rate of 0 indicates
-the optimizer did not come sufficiently close to the optimal projection.
+\texttt{splines2d}, \texttt{skinny} and \texttt{stringy2}, following Laa
+and Cook (2020). (The supplementary material includes definitions of
+these PP indexes, and descriptions of the pipe and sine wave data
+simulation.). The JSO hyper-parameters, number of jellies and maximum
+tries, are also included. This results in a total of 76 scenarios,
+comprising of 30 computed on the pipe data and 46 on the sine wave data.
+The choice of 4-12 dimensions covers the range of easy to relatively
+difficult for the optimizers to detect the structured projections. The
+same approach is used, that each experiment is repeated 50 times and
+success rate is calculated at each level of the simulation design.
 
@@
 \caption{\label{fig-smoothness}Steps for calculating smoothness in a
-projection pursuit problem. Given the target shape, data dimension and
-the index function: 1) sample random bases given the orthonormality
-contraint and calculate their corresponding index values, and 2) fit a
-Gaussian process model of index values against the sampled bases to
-obtain the smoothness measure.}
+projection pursuit problem: 1) sample random bases given the
+orthonormality contraint and calculate their corresponding index values,
+2) fit a Gaussian process model of index values against the bases to
+obtain the smoothness measure, 3) take the smoothness parameter from the
+Gaussian process as the smoothness metric.}
 
@@
 
+Recall that smoothness and squintability are metrics used to
+characterize the index function for PP tasks and are invariant to the
+optimizer and its hyper-parameters. Thus simulations with different JSO
+hyper-parameters for the same PP task share the same smoothness value.
+In total, smoothness and squintability metrics are calculated on the 22
+PP tasks in our simulation.
+
 Figure~\ref{fig-smoothness} describes the procedure to compute
-smoothness. Index values are calculated for each of 500 simulated random
-bases. A Gaussian process model is fitted to the resulting data to
-obtain the smoothness measure for the index, as described in
-Section~\ref{sec-smoothness}.
+smoothness. For each PP task, 500 random bases are simulated and index
+values are calculated for each. A Gaussian process model is then fitted
+to the resulting data to obtain the smoothness measure for the index, as
+described in Section~\ref{sec-smoothness}.
 
@@
 \caption{\label{fig-squintability}Steps for calculating squintability in
-a projection pursuit problem. Given the target shape, data dimension and
-the index function: 1) sample random bases given the orthonormality and
-projection distance contraint and calculates their corresponding index
-values, 2) interpolate the sampled bases to the optimal basis and
-calculate the projection distance and the index value. 3) bin the index
-values by projection distances through averaging index value, 4) fit the
-scaled sigmoid function in equation \eqref{eq-squintability} to the
-binned index values against projection distances using non-linear least
-square to obtian the squintability measure using equation
+a projection pursuit problem: 1) sample random bases given the
+orthonormality and projection distance contraint and calculates their
+corresponding index values, 2) interpolate the sampled bases to the
+optimal basis and calculate the projection distance and the index value.
+3) average the index values by projection distances at each 0.005
+distance bin, 4) fit the scaled sigmoid function in Equation
+\eqref{eq-logistic} and \eqref{eq-parametric} , 5) calculate the
+squintability metric using Equation
 (\ref{eq-squintability-parametric}).}
@@
 Figure~\ref{fig-squintability} describes the procedure to compute
-squintability. Index values and projection distances are computed for
-the projections along interpolated paths between 50 random bases and the
-optimal basis. An interpolation step size of 0.005 is used. The
-projection distances are binned to 0.005 intervals and index values are
-averaged in each bin. The four-parameter scaled logistic function
-(\ref{eq-parametric}) is fitted to the index values against projection
-distances, estimated by non-linear least squares to obtain the
-squintability measure as Equation (\ref{eq-squintability-parametric}).
+squintability. For each PP task, 50 random bases are generated and
+interpolated to the optimal basis with a step size of 0.005. The index
+values are averaged over the distance window of 0.005. Index values are
+calculated for all the interpolated bases and averaged over the distance
+bin of 0.005. The logisitic function in Equation (\ref{eq-logistic}) and
+(\ref{eq-parametric}) is fitted with non-linear least squares and
+calculate the squintability measure using Equation
+(\ref{eq-squintability-parametric}).
 
 A generalized linear model is fitted using a quasibinomial family and a
-logit link function to assess the factors affecting success rate.
-Predictors are smoothness, squintability, data dimension and JSO
+logit link function to assess the factors affecting the success rate.
+Predictors are smoothness, squintability, data dimension, and JSO
 hyper-parameters.

### 2025-03-04 code/06-sim-sine-6d-skinny.R
include 4d skinny data

@@ -1,10 +1,7 @@
 library(tidyverse)
-library(spinebil)
+library(ferrn)
 library(tourr)
 
-set.seed(123456)
-sine1000 <- sinData(6, 1000) %>% scale() %>% as_tibble()
-colnames(sine1000) <- paste0("V", 1:6)
 ################################################################################
 # d = 6
 skinny <- function(){
@@ -14,10 +11,13 @@ skinny <- function(){
 }
 
 sim <- function(d = d, n_jellies = n_jellies, max.tries = max.tries,
-                data_seed = 123456, optim_seed = seed, sim = sim){
-  set.seed(data_seed)
-  sine1000 <- sinData(d, 1000) %>% scale() %>% as_tibble()
-  colnames(sine1000) <- paste0("V", 1:d)
+                optim_seed = seed, sim = sim){
+
+  if (d == 4){
+    sine1000 <- ferrn::sine1000_4d
+  } else if (d == 6){
+    sine1000 <- ferrn::sine1000_6d
+  }
 
   cat("sim: ", sim, "\n")
   cat("n_jellies: ", n_jellies, "\n")
@@ -41,10 +41,8 @@ sim <- function(d = d, n_jellies = n_jellies, max.tries = max.tries,
 
 set.seed(1234)
 seed <- sample(1000:10000, size = 50)
-sim_setup <- tidyr::crossing(n_jellies = 30,
-                      max_tries = 50,
-                      d = 6) |>
-  tidyr::crossing(sim = 1) |>
+sim_setup <- tidyr::crossing(n_jellies = 30, max_tries = 50, d = 6) |>
+  tidyr::crossing(sim = 1:50) |>
   mutate(seed = seed[sim], id = row_number())
 
 t1 <- Sys.time()
@@ -67,9 +65,37 @@ sim_sine_6d_skinny <- bind_rows(sim_sine_6d_skinny_0105, sim_sine_6d_skinny_0610
 save(sim_sine_6d_skinny, file = here::here("data/sim_sine_6d_skinny.rda"))
 
 ##################################################
+set.seed(1234)
+seed <- sample(1000:10000, size = 50)
+sim_setup <- tidyr::crossing(n_jellies = 30,
+                             max_tries = 50,
+                             d = 4) |>
+  tidyr::crossing(sim = 31:50) |>
[…]

### 2025-03-04 code/07-calc-smoothness-squintability.R

@@ -27,7 +27,7 @@ holes_tidy <- smoothness_holes |>
   unnest(smooth)
 
 ##################################################
-idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "stringy2", "splines2d", "skinny")
+idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "splines2d", "skinny", "stringy2")
 smoothness_sine <- tibble::tibble(
   n = 6, index = idx_names, data = list(sine1000),
   best = list(matrix(c(rep(0, 8), 1, 0, 0, 1), nrow = 6, byrow = TRUE))) |>
@@ -39,7 +39,7 @@ smoothness_sine <- tibble::tibble(
   dplyr::mutate(basis_df = list(sample_bases(index, n_basis = 500, best = best,
                                              data = as.matrix(data))))
 
-idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "splines2d", "stringy2")
+idx_names <- c("dcor2d_2", "loess2d", "MIC", "TIC", "splines2d", "skinny", "stringy2")
 smoothness_4d <- tibble::tibble(
   n = 4, index = idx_names, data = list(sine1000_4d),
   best = list(matrix(c(rep(0, 4), 1, 0, 0, 1), nrow = 4, byrow = TRUE))) |>
@@ -53,11 +53,10 @@ sine_tidy <- bind_rows(smoothness_sine, smoothness_4d) |>
 
 smoothness <- bind_rows(sine_tidy, holes_tidy) |>
   mutate(index = factor(index, levels = c("holes", "MIC", "TIC", "dcor2d_2",
-                                          "loess2d", "splines2d", "stringy2", "skinny"))) |>
+                                          "loess2d", "splines2d", "skinny", "stringy2"))) |>
   arrange(index) |>
   select(index, n, variance:nugget)
 save(smoothness, file = here::here("data", "smoothness.rda"))
-
 ################################################################################
 ################################################################################
 # squintability - sample bases
@@ -77,7 +76,7 @@ sq_holes_basis_df <- tibble::tibble(
                    step_size = 0.005, best = best)}))
[…]

### 2025-03-04 code/08-process-data-for-modelling.R

@@ -102,6 +102,7 @@ load(here::here("data-raw/sim_sine_4d_dcor2d.rda"))
 load(here::here("data-raw/sim_sine_4d_loess2d.rda"))
 load(here::here("data-raw/sim_sine_4d_MICTIC.rda"))
 load(here::here("data-raw/sim_sine_4d_splines2d.rda"))
+load(here::here("data-raw/sim_sine_4d_skinny.rda"))
 pipe_setup_summ_4d <- sim_pipe_4d |>
   mutate(id2 = paste0(index, n_jellies, max_tries, sim, d)) |>
   get_best(group = id2) |>
@@ -114,7 +115,7 @@ pipe_setup_summ_4d <- sim_pipe_4d |>
 
 sim_data_sine_4d <- bind_rows(
   sim_sine_4d_dcor2d |> mutate(index = "dcor2d_2") , sim_sine_4d_loess2d, sim_sine_4d_MICTIC,
-  sim_sine_4d_splines2d)
+  sim_sine_4d_splines2d, sim_sine_4d_skinny |> mutate(index = "skinny"))
 
 sine_run_df_4d <- sim_data_sine_4d |>
   mutate(id2 = paste0(index, n_jellies, max_tries, sim, d)) |>
@@ -127,6 +128,8 @@ sine_setup_summ_4d <- sine_run_df_4d |>
     P_J = sum(abs(I_max - index_val) <= 0.05)/n(),
     time = mean(time)
   )
+
+
 
 #############################################################################
 #############################################################################

### 2025-03-04 jso.qmd
proofread done

@@
 well as JSO's hyper-parameter choices (number of jellyfish and maximum
-number of tries). This work should facilitate better optimization for PP
+number of iterations). This work should facilitate better optimization for PP
 and guide the choice of optimizer when designing new PP indexes.
@@
 projection, an index function, the number of jellyfish, and the maximum
-number of trials (tries) are provided as input. Then, the current
+number of iterations are provided as input. Then, the current
 projection is evaluated by the index function. The projection is then
@@
 that are most relevant to our application: the number of jellyfish and
-the maximum number of tries.
+the maximum number of iterations
 
@@
 definitions of these PP indexes, and descriptions of the pipe and sine
-wave data simulation.). The JSO hyper-parameters, number of jellies and
-maximum tries, are also included. This results in a total of
+wave data simulation.). The JSO hyper-parameters, number of jellyfish and
+maximum iterations, are also included. This results in a total of
 `r nrow(sim_df)` scenarios, comprising of
@@
 for the same PP task share the same smoothness value. In total, smoothness and
-squintability metrics are calculated on the 22 PP tasks in our simulation.
+squintability metrics are calculated on the `r nrow(smoothness)` PP tasks in our simulation.
 
@@
 Predictors are smoothness, squintability, data dimension, and JSO
-hyper-parameters. <!-- The data is
-pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
-for interpretation, 2) creating a new binary variable `long_time` to
-indicate cases with an average run time over 20 seconds, and 3)
-re-coding the success rate for the `stringy` index as 0, because none of
-the 50 simulations correctly identified the sine wave shape.-->
+hyper-parameters. Due to the scale of the smoothness and squintability measure,
+the rank of the two metrics, rather than the nominal values, are used in the model.
 
@@
 Rows correspond to data dimension.
-Columns correspond to quantiles of the index values, with 0 being minimum, 100 being maximum, and 50 the median.
+Columns correspond to quantiles of the index values, with 0 being the minimum, 100 being the maximum, and 50 the median.
 JSO achieves the better views of the pipe generally than CRS.
@@
 
-The performance of JSO is assessed relative to the existing CRS
-optimizer based on the optimal projection obtained. The final
-projections for the pipe data sets found by the two optimizers are shown
-in @fig-proj, with rows corresponding to data dimension. The 50
-simulations in each data dimension are sorted by index value, and the
-projections corresponding to each 10th quantile value are shown, with
-minimum at left and maximum at right. Index value is printed on the
-plot. The purpose is to summarize the views of the data resulting from
+The optimal projection is compared between JSO and CRS for the pipe data at
+dimensions 6, 8, 10, and 12 in @fig-proj. The columns show the quantile of 
+index value in the 50 optimal projection, with 0 being the minimum, 100 being 
+the maximum, and 50 the median. The purpose is to summarize the views of the data resulting from
 different optimization, and hence compare results between the two
@@
 
-<!-- broken down by 10th quantile,
-faceted by the data dimensions. In the 6-dimensional data scenario, JSO
-consistently identifies a clear pipe shape. The CRS also finds the pipe
-shape but with a wide rim, suggesting a further polish search may be
-required. With increasing dimensions, JSO may not always identify the
-pipe shape due to random sampling, but it still finds the pipe shape in
-over 50% of cases. Compared to CRS, JSO achieves higher index values and
-clearer pipe shapes across all quantiles in data of 8, 10, 12
-dimensions, suggesting its advantage in exploring high-dimensional
-spaces.-->
-
 ## Effect of hyper-parameters effect on JSO success rate
@@
 The effect of JSO hyper-parameters (number of jellyfish and the maximum
-number of tries) on the success rate is summarized in @fig-proportion.
-The uncertainty is quantified through 500 bootstrap replicates for each
-case. As the number of jellyfish and maximum tries increase, the success
-rate also increases. For problems with relatively lower dimensional (4,
-6) search spaces, small parameter values (20 jellyfish and a maximum of
-50 tries) are sufficient to achieve a high success rate. Many more
-jellyfish (100 jellyfish and a maximum of 100 tries) are needed to
-obtain a mediocre success rate for dimensions 10 and 12. This suggests
+number of iteration) on the success rate is summarized in @fig-proportion.
+The uncertainty is quantified through 500 bootstrap replicates sampled from 
+the simulation for each case. As the number of jellyfish and maximum iteration allowed 
+increase, the success rate also increases. For problems in lower dimensional (4,
+6) search spaces, small parameter values (20 jellyfish and 50 iterations) 
+are sufficient to achieve a high success rate. Many more
+jellyfish (100 jellyfish and 100 iterations) are neede for finding the 
+pipe structure in d = 10, and 12. This suggests
 that even JSO will struggle with high dimensions, and require
@@
 especially for difficult PP indexes based on scagnostics.
-<!-- dimensions) where a more complex search is necessary.
-However, in lower-dimensional problems, the benefit may be less pronounced, as the space is already relatively easy to navigate.
-While increasing both parameters enhances the performance of JSO, it also extends the computational time required for the optimization, which can be
-computationally intensive when evaluating the index function (such as
-scagnostic indexes) multiple times across numerous iterations.-->
 
@@
 #| fig.align: "center"
-#| fig.cap: "Summary of the relationship between JSO hyper-parameters on optimization success, using the holes index for pipe data of dimensions 4-12.
+#| fig.cap: "Summary of the relationship between JSO hyper-parameters on optimization success rate, using the holes index for pipe data of dimensions 4-12.
 Bootstrap samples show the variability in success rate.
-Success rate mostly plateaus by 50 jellies.
-The JSO has some difficulty finding the pipe when dimension is higher than 8!
-Maximum number of tries has little effect."
+Success rate mostly plateaus by 50 jellyfish.
+The JSO has some difficulty finding the pipe when dimension is higher than 8 and maximum number of iteration has little effect."
 pipe_sim_best <- sim_pipe_run_best |> 
@@
          n_jellies = ifelse(d == 12, n_jellies + 1, n_jellies)) |> 
-  rename(`max # of tries` = max_tries)
+  rename(`max # of iteration` = max_tries)
 
@@
   scale_x_continuous(breaks = c(20, 50, 100), labels = c("20", "50", "100")) +
-  facet_wrap(vars(`max # of tries`), 
+  facet_wrap(vars(`max # of iteration`), 
              labeller = "label_both") + 
@@
 
-Smoothness and squintability are calculated across a collection of pipe
-and sine wave finding problems to construct the relationship between
-success rate, JSO hyper-parameters, and index properties.
 @tbl-smoothness-squintability presents values computed for a subset of
@@
 @fig-idx-proj-dist shows a sample of traces from the indexes in
-@tbl-smoothness-squintability for the 6D sine wave data. The traces the
+@tbl-smoothness-squintability for the 6D sine wave data. This traces the
 PP index values computed for projections along an interpolated path
-between a noise projection and the optimal projection (sine wave). (Note
+between a randomly generated noise projection and the optimal projection (sine wave). (Note
 index values were self-standardized to range between 0 and 1, for
@@
 As dimension increases the squint angle gets smaller but smoothness doesn't change much.
-dt <- tibble(shape = c(rep("pipe", 5), rep("sine", 17)), smoothness) |>
+dt <- tibble(shape = c(rep("pipe", 5), rep("sine", 18)), smoothness) |>
   left_join(squintability) |> 
   rename(d = n, smooth = smoothness) |> 
-  mutate(index = ifelse(index == "dcor2d_2", "dcor", index),
-         index = ifelse(index == "loess2d", "loess", index),
+  mutate(index = as.character(index),
          index = ifelse(index == "splines2d", "splines", index),
@@
 squintability, dimension, number of jellyfish and maximum number of
-tries, as predictors. The fit suggests that the JSO success rate is only
-affected by squintability, dimension and number of jellies. As expected,
+iteration, as predictors. The fit suggests that the JSO success rate is only
+affected by squintability, dimension and number of jellyfish. As expected,
 the JSO success rate is higher with higher squintability and/or more
@@
 regularity into account, as opposed to gradient-based optimizers for
-example. Allowing for more tries also does not affect success rate
+example. Allowing for more iterations also does not affect success rate
 significantly. A unit increase in squintability increases the success
@@
 dimension, and smoothness and data dimension, improves the model a small
-amount, but does not change the interpretation above.
+amount, but does not change the interpretation above. This study intends 
+to serve as an illustration that the proposed
+measures effectively captures the key factors influencing the success
+rate of PP.
 
@@
 #| tbl-cap: "Jellyfish success rate relative to index properties and jellyfish hyper-parameters.
-This is the summary from a logistic regression fit to smoothness, squintability, dimension, number of jellyfish and maximum number of tries.
+This is the summary from a logistic regression fit to smoothness, squintability, dimension, number of jellyfish and maximum number of iterations.
 Interestingly, squintability and dimension strongly affect jellyfish optimization success.
-The number of jellies marginally affects success, but index smoothness, and increasing the number of tries do not."
+The number of jellies marginally affects success, but index smoothness, and increasing the number of iterations do not."
 #| fig-pos: t
@@
                                     d = "Dimension", 
-                                    #long_time = "Long runtime", 
                                     n_jellies = "Number of jellyfish", 
-                                    max_tries = "Maximum number of tries")) |>
+                                    max_tries = "Maximum number of iterations")) |>
   bold_p(t = 0.1) 
@@
 
-<!--The simulation result suggests that squintability and smoothness are likely dependent.
-Intuitively, a highly smooth search space exhibits gradual changes, making it easier to perceive the overall structure from a distance, thereby increasing the squint angle.
-@tbl-joint-output compare the model fits when squintability and smoothness are included separately.
-Given that the null deviance remains the same across models, the model with squintability yields a smaller residual deviance and lower AIC and BIC values, indicating a better fit.
-This suggests that squintability captures the relevant variation more effectively than smoothness.
-While it may seem counterintuitive at first that smoothness is not a significant factor, this outcome is reasonable in this context, as the swarm-based method used in this study, JSO, is designed to perform well even in non-smooth search spaces by leveraging population-based exploration.-->
-
-This study intends to serve as an illustration that the proposed
-measures effectively captures the key factors influencing the success
-rate of PP.
-<!--The interactions between the two variables were not examined in this study due to the limited number of data points, nor was the linearity assumption of the relationship with success rate explicitly tested.-->
-
 # Practicalities {#sec-discussion}
 
-The simulation studies have compared the JSO with CRS and shown how
-smoothness and squintability affect the JSO success rate. Here we
-explain how they can be used for new index and optimizer development.
-Using the JSO optimizer for PPGT requires a change in user behavior. For
+Using the JSO optimizer for PPGT in the `tourr` package requires a change in user behavior. For
 all the existing optimizers a single optimization path is followed. The
@@
 To use JSO for PPGT in the `tourr` package, specify
-`search_f = search_jellyfish` in the guided tour. Unlike existing
-optimizers, the animation function `animate_*()` won't directly render
-the tour path for JSO due to the generation of multiple tour paths. To
+`search_f = search_jellyfish` in the guided tour.
+
+The animation function `animate_*()` won't directly render the animated 
+projection for JSO, as with other existing optimizers since multiple tour 
+paths will need to be generated. To
 visualize the path visited by each individual jellyfish, assign the
@@
 
-The `ferrn` package [@RJ-2021-105] provides functionality for computing
+The `ferrn` package [@RJ-2021-105] has now included functionality for computing
 the smoothness and squintability metrics. Both metrics require sampling
-random bases using the `sample_bases()` function, followed by the metric
+random bases using the `sample_bases()` function, followed by the
 calculation with `calc_smoothness()` or `calc_squintability()`.
@@
 
-To sample bases for calculating smoothness, the following arguments are
+To sample bases for calculating smoothness, the following inputs are
 required: the index function, the dataset, and the number of random
 bases to sample. The output of `sample_bases()` is a data frame with a
-list-column of sampled basis matrix and index value. Particularization
-is available to speed up the index value computation through the
+list-column of sampled bases and index values. Parallelization
+is available to speed up the index value calculation through the
 `parallel = TRUE` argument.
@@
 and fits a Gaussian process model to the index values against the
-sampled basis, as the location, to obtain the smoothness metric.
+sampled bases, as the location, to obtain the smoothness metric.
 Starting parameters and additional Gaussian process arguments can be
@@
 
-Bases sampling for squintability includes an additional step of
-interpolating between the sampled bases and the best basis. This step is
+Bases sampling for calculating squintability includes an additional step of
+interpolating between the sampled bases and the optimal basis. This step is
 performed when the arguments `step_size` and `min_proj_dist` in the
@@
 required to average the index values over the projection distance before
-the fit. For the parametric case, the output provides the estimated
-parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$)
+the fitting. For the parametric case, the output provides the estimated
+parameters for the logistic function ($\theta_1$ to $\theta_4$)
 and the calculated squintability metric as Equation
@@
 features of PP indexes, squintability and smoothness, and used these to
-assess the performance of the new jellyfish search optimizer. The
+assess the performance of the new JSO. The
 metrics will be generally useful for characterizing PP indexes, and help
@@
 chance of reaching the global optimum when more jellyfish are used and
-the maximum number of tries is increased. However, it comes at a
+the maximum number of iteration is increased. However, it comes at a
 computational cost, as expected. The performance declines if the
@@
 
-One future direction of this work is to test the JSO, along with its
-variations and other swarm-based optimizers, on a broader range of
-indexes, for example, scagnostic indexes.
+Future work can focus on developing more efficient optimization algorithm for
+noise indexes with low squintability metric, which resemble a needle-in-a-haystack 
+problem. Additionally, if the index function is differentiable, automatic 
+differentiation can be used to implement new gradient-based optimizers.
 

### 2025-03-04 jso.tex
proofread done

@@
 well as JSO's hyper-parameter choices (number of jellyfish and maximum
-number of tries). This work should facilitate better optimization for PP
-and guide the choice of optimizer when designing new PP indexes.
+number of iterations). This work should facilitate better optimization
+for PP and guide the choice of optimizer when designing new PP indexes.
 
@@
 starting projection, an index function, the number of jellyfish, and the
-maximum number of trials (tries) are provided as input. Then, the
-current projection is evaluated by the index function. The projection is
-then moved in a direction determined by a random factor, influenced by
-how far along we are in the optimization process. Occasionally,
-completely new directions may be taken like a jellyfish might with ocean
-currents. A new projection is accepted if it is an improvement compared
-to the current one, rejected otherwise. This process continues and
-iteratively improves the projection, until the pre-specified maximum
-number of trials is reached.
-
-\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, leftrule=.75mm, breakable, coltitle=black, colback=white, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, opacitybacktitle=0.6, opacityback=0, left=2mm, bottomtitle=1mm, rightrule=.15mm, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm]
+maximum number of iterations are provided as input. Then, the current
+projection is evaluated by the index function. The projection is then
+moved in a direction determined by a random factor, influenced by how
+far along we are in the optimization process. Occasionally, completely
+new directions may be taken like a jellyfish might with ocean currents.
+A new projection is accepted if it is an improvement compared to the
+current one, rejected otherwise. This process continues and iteratively
+improves the projection, until the pre-specified maximum number of
+trials is reached.
+
+\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, colback=white, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, breakable, toprule=.15mm, leftrule=.75mm, opacityback=0, opacitybacktitle=0.6, toptitle=1mm, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, left=2mm]
 
@@
 these PP indexes, and descriptions of the pipe and sine wave data
-simulation.). The JSO hyper-parameters, number of jellies and maximum
-tries, are also included. This results in a total of 76 scenarios,
-comprising of 30 computed on the pipe data and 46 on the sine wave data.
+simulation.). The JSO hyper-parameters, number of jellyfish and maximum
+iterations, are also included. This results in a total of 77 scenarios,
+comprising of 30 computed on the pipe data and 47 on the sine wave data.
 The choice of 4-12 dimensions covers the range of easy to relatively
@@
 hyper-parameters for the same PP task share the same smoothness value.
-In total, smoothness and squintability metrics are calculated on the 22
+In total, smoothness and squintability metrics are calculated on the 23
 PP tasks in our simulation.
@@
 Predictors are smoothness, squintability, data dimension, and JSO
-hyper-parameters.
+hyper-parameters. Due to the scale of the smoothness and squintability
+measure, the rank of the two metrics, rather than the nominal values,
+are used in the model.
 
@@
 Rows correspond to data dimension. Columns correspond to quantiles of
-the index values, with 0 being minimum, 100 being maximum, and 50 the
-median. JSO achieves the better views of the pipe generally than CRS. As
-dimension increases both have more difficulty finding the pipe.}
+the index values, with 0 being the minimum, 100 being the maximum, and
+50 the median. JSO achieves the better views of the pipe generally than
+CRS. As dimension increases both have more difficulty finding the pipe.}
 
@@
 
-The performance of JSO is assessed relative to the existing CRS
-optimizer based on the optimal projection obtained. The final
-projections for the pipe data sets found by the two optimizers are shown
-in Figure~\ref{fig-proj}, with rows corresponding to data dimension. The
-50 simulations in each data dimension are sorted by index value, and the
-projections corresponding to each 10th quantile value are shown, with
-minimum at left and maximum at right. Index value is printed on the
-plot. The purpose is to summarize the views of the data resulting from
-different optimization, and hence compare results between the two
-optimizers. Generally, the JSO does a more consistent job of finding the
-pipe structure clearly. As the data dimension increases, both optimizers
+The optimal projection is compared between JSO and CRS for the pipe data
+at dimensions 6, 8, 10, and 12 in Figure~\ref{fig-proj}. The columns
+show the quantile of index value in the 50 optimal projection, with 0
+being the minimum, 100 being the maximum, and 50 the median. The purpose
+is to summarize the views of the data resulting from different
+optimization, and hence compare results between the two optimizers.
+Generally, the JSO does a more consistent job of finding the pipe
+structure clearly. As the data dimension increases, both optimizers
 struggle and less clearly capture the circle shape.
@@
 The effect of JSO hyper-parameters (number of jellyfish and the maximum
-number of tries) on the success rate is summarized in
+number of iteration) on the success rate is summarized in
 Figure~\ref{fig-proportion}. The uncertainty is quantified through 500
-bootstrap replicates for each case. As the number of jellyfish and
-maximum tries increase, the success rate also increases. For problems
-with relatively lower dimensional (4, 6) search spaces, small parameter
-values (20 jellyfish and a maximum of 50 tries) are sufficient to
-achieve a high success rate. Many more jellyfish (100 jellyfish and a
-maximum of 100 tries) are needed to obtain a mediocre success rate for
-dimensions 10 and 12. This suggests that even JSO will struggle with
-high dimensions, and require substantial more computational time to have
-an acceptable success rate, especially for difficult PP indexes based on
+bootstrap replicates sampled from the simulation for each case. As the
+number of jellyfish and maximum iteration allowed increase, the success
+rate also increases. For problems in lower dimensional (4, 6) search
+spaces, small parameter values (20 jellyfish and 50 iterations) are
+sufficient to achieve a high success rate. Many more jellyfish (100
+jellyfish and 100 iterations) are neede for finding the pipe structure
+in d = 10, and 12. This suggests that even JSO will struggle with high
+dimensions, and require substantial more computational time to have an
+acceptable success rate, especially for difficult PP indexes based on
 scagnostics.
@@
 \caption{\label{fig-proportion}Summary of the relationship between JSO
-hyper-parameters on optimization success, using the holes index for pipe
-data of dimensions 4-12. Bootstrap samples show the variability in
-success rate. Success rate mostly plateaus by 50 jellies. The JSO has
-some difficulty finding the pipe when dimension is higher than 8!
-Maximum number of tries has little effect.}
+hyper-parameters on optimization success rate, using the holes index for
+pipe data of dimensions 4-12. Bootstrap samples show the variability in
+success rate. Success rate mostly plateaus by 50 jellyfish. The JSO has
+some difficulty finding the pipe when dimension is higher than 8 and
+maximum number of iteration has little effect.}
 
@@
 
-Smoothness and squintability are calculated across a collection of pipe
-and sine wave finding problems to construct the relationship between
-success rate, JSO hyper-parameters, and index properties.
 Table~\ref{tbl-smoothness-squintability} presents values computed for a
@@
 in Table~\ref{tbl-smoothness-squintability} for the 6D sine wave data.
-The traces the PP index values computed for projections along an
-interpolated path between a noise projection and the optimal projection
-(sine wave). (Note index values were self-standardized to range between
-0 and 1, for comparison.) The \texttt{splines} index is very smooth with
-a gradual increase in value between the noise projection and the optimal
-projection. The \texttt{TIC} index is slightly less smooth and slightly
-less gradual change in value. Least desirable is the \texttt{skinny}
-index which is noisier and the optimization needs to be much closer to
-the optimal angle to have a higher value. This illustrates how the new
-measures for squintability and smoothness work to describe the PP
-indexes.
+This traces the PP index values computed for projections along an
+interpolated path between a randomly generated noise projection and the
+optimal projection (sine wave). (Note index values were
+self-standardized to range between 0 and 1, for comparison.) The
+\texttt{splines} index is very smooth with a gradual increase in value
+between the noise projection and the optimal projection. The
+\texttt{TIC} index is slightly less smooth and slightly less gradual
+change in value. Least desirable is the \texttt{skinny} index which is
+noisier and the optimization needs to be much closer to the optimal
+angle to have a higher value. This illustrates how the new measures for
+squintability and smoothness work to describe the PP indexes.
 
@@
 sine & skinny & 6 & 2.2352 & 0.0973\\
+sine & skinny & 4 & 2.5213 & 0.0810\\
 \bottomrule
@@
 smoothness, squintability, dimension, number of jellyfish and maximum
-number of tries, as predictors. The fit suggests that the JSO success
-rate is only affected by squintability, dimension and number of jellies.
-As expected, the JSO success rate is higher with higher squintability
-and/or more jellyfish, and lower when dimension is higher
+number of iteration, as predictors. The fit suggests that the JSO
+success rate is only affected by squintability, dimension and number of
+jellyfish. As expected, the JSO success rate is higher with higher
+squintability and/or more jellyfish, and lower when dimension is higher
 (higher-dimensional optimization is more difficult). Interestingly, the
@@
 value without taking local regularity into account, as opposed to
-gradient-based optimizers for example. Allowing for more tries also does
-not affect success rate significantly. A unit increase in squintability
-increases the success rate by 24\%. As dimension increases by one the
-success rate almost halves. Increasing the number of jellies by 10
-increases the success rate by 32\%. Table~\ref{tbl-joint-output} shows
-the fit statistics supporting that the model is reasonably strong.
+gradient-based optimizers for example. Allowing for more iterations also
+does not affect success rate significantly. A unit increase in
+squintability increases the success rate by 23\%. As dimension increases
+by one the success rate almost halves. Increasing the number of jellies
+by 10 increases the success rate by 32\%. Table~\ref{tbl-joint-output}
+shows the fit statistics supporting that the model is reasonably strong.
 
@@
 dimension, and smoothness and data dimension, improves the model a small
-amount, but does not change the interpretation above.
+amount, but does not change the interpretation above. This study intends
+to serve as an illustration that the proposed measures effectively
+captures the key factors influencing the success rate of PP.
 
@@
 logistic regression fit to smoothness, squintability, dimension, number
-of jellyfish and maximum number of tries. Interestingly, squintability
-and dimension strongly affect jellyfish optimization success. The number
-of jellies marginally affects success, but index smoothness, and
-increasing the number of tries do not.}
+of jellyfish and maximum number of iterations. Interestingly,
+squintability and dimension strongly affect jellyfish optimization
+success. The number of jellies marginally affects success, but index
+smoothness, and increasing the number of iterations do not.}
 
@@
 \midrule\addlinespace[2.5pt]
-Smoothness & 1.00 (0.90 to 1.11) & 0.928 \\ 
-Squintability & 1.24 (1.13 to 1.39) & {\bfseries <0.001} \\ 
-Dimension & 0.57 (0.44 to 0.73) & {\bfseries <0.001} \\ 
+Smoothness & 1.01 (0.92 to 1.11) & 0.810 \\ 
+Squintability & 1.23 (1.12 to 1.38) & {\bfseries <0.001} \\ 
+Dimension & 0.58 (0.45 to 0.73) & {\bfseries <0.001} \\ 
 Number of jellyfish & 1.32 (1.17 to 1.50) & {\bfseries <0.001} \\ 
-Maximum number of tries & 1.06 (0.94 to 1.21) & 0.333 \\ 
+Maximum number of iterations & 1.06 (0.94 to 1.20) & 0.328 \\ 
 \bottomrule
@@
 \midrule
-20.69 & 70 & 49.29 & 75\\
+20.89 & 71 & 49.77 & 76\\
 \bottomrule
@@
 
-This study intends to serve as an illustration that the proposed
-measures effectively captures the key factors influencing the success
-rate of PP.
-
 \section{Practicalities}\label{sec-discussion}
 
-The simulation studies have compared the JSO with CRS and shown how
-smoothness and squintability affect the JSO success rate. Here we
-explain how they can be used for new index and optimizer development.
-Using the JSO optimizer for PPGT requires a change in user behavior. For
-all the existing optimizers a single optimization path is followed. The
-JSO has multiple optimization paths, and needs additional tools to
-incorporate into the PPGT, as explained here.
+Using the JSO optimizer for PPGT in the \texttt{tourr} package requires
+a change in user behavior. For all the existing optimizers a single
+optimization path is followed. The JSO has multiple optimization paths,
+and needs additional tools to incorporate into the PPGT, as explained
+here.
 
@@
 To use JSO for PPGT in the \texttt{tourr} package, specify
-\texttt{search\_f\ =\ search\_jellyfish} in the guided tour. Unlike
-existing optimizers, the animation function \texttt{animate\_*()} won't
-directly render the tour path for JSO due to the generation of multiple
-tour paths. To visualize the path visited by each individual jellyfish,
-assign the animation to an object
+\texttt{search\_f\ =\ search\_jellyfish} in the guided tour.
+
+The animation function \texttt{animate\_*()} won't directly render the
+animated projection for JSO, as with other existing optimizers since
+multiple tour paths will need to be generated. To visualize the path
+visited by each individual jellyfish, assign the animation to an object
 (\texttt{res\ \textless{}-\ animate\_xy(...)}). This will save the bases
@@
 
-The \texttt{ferrn} package (Zhang et al. 2021) provides functionality
-for computing the smoothness and squintability metrics. Both metrics
-require sampling random bases using the \texttt{sample\_bases()}
-function, followed by the metric calculation with
+The \texttt{ferrn} package (Zhang et al. 2021) has now included
+functionality for computing the smoothness and squintability metrics.
+Both metrics require sampling random bases using the
+\texttt{sample\_bases()} function, followed by the calculation with
 \texttt{calc\_smoothness()} or \texttt{calc\_squintability()}.
@@
 
-To sample bases for calculating smoothness, the following arguments are
+To sample bases for calculating smoothness, the following inputs are
 required: the index function, the dataset, and the number of random
 bases to sample. The output of \texttt{sample\_bases()} is a data frame
-with a list-column of sampled basis matrix and index value.
-Particularization is available to speed up the index value computation
-through the \texttt{parallel\ =\ TRUE} argument.
+with a list-column of sampled bases and index values. Parallelization is
+available to speed up the index value calculation through the
+\texttt{parallel\ =\ TRUE} argument.
 
@@
 \texttt{sample\_bases()} and fits a Gaussian process model to the index
-values against the sampled basis, as the location, to obtain the
+values against the sampled bases, as the location, to obtain the
 smoothness metric. Starting parameters and additional Gaussian process
@@
 
-Bases sampling for squintability includes an additional step of
-interpolating between the sampled bases and the best basis. This step is
-performed when the arguments \texttt{step\_size} and
+Bases sampling for calculating squintability includes an additional step
+of interpolating between the sampled bases and the optimal basis. This
+step is performed when the arguments \texttt{step\_size} and
 \texttt{min\_proj\_dist} in the \texttt{sample\_bases()} function are
@@
 \texttt{bin\_width} argument is required to average the index values
-over the projection distance before the fit. For the parametric case,
-the output provides the estimated parameters for the scaled sigmoid
+over the projection distance before the fitting. For the parametric
+case, the output provides the estimated parameters for the logistic
 function (\(\theta_1\) to \(\theta_4\)) and the calculated squintability
@@
 features of PP indexes, squintability and smoothness, and used these to
-assess the performance of the new jellyfish search optimizer. The
-metrics will be generally useful for characterizing PP indexes, and help
-with developing new indexes.
+assess the performance of the new JSO. The metrics will be generally
+useful for characterizing PP indexes, and help with developing new
+indexes.
 
@@
 
-One future direction of this work is to test the JSO, along with its
-variations and other swarm-based optimizers, on a broader range of
-indexes, for example, scagnostic indexes.
+Future work can focus on developing more efficient optimization
+algorithm for noise indexes with low squintability metric, which
+resemble a needle-in-a-haystack problem. Additionally, if the index
+function is differentiable, automatic differentiation can be used to
+implement new gradient-based optimizers.
 

### 2025-03-05 jso.qmd
small changes on tables

@@
 
-## Effect of index properties on JSO success rate
-
-@tbl-smoothness-squintability presents values computed for a subset of
-PP indexes on the sine wave data for dimensions 4-8. (A full table of
-the parameters estimated from the Gaussian process (outputscale $\eta$,
-lengthscale $\ell$, smoothness $\nu$, and nugget $\sigma$) and from the
-scaled logistic function ($\theta_1$ to $\theta_4$) for calculating
-smoothness and squintability for all PP indexes and data combinations is
-included in the supplementary materials). The column $\nu$ is used as
-the smoothness metric and the column $\varsigma$ is calculated as
-equation (@eq-squintability-parametric) as the squintability metric.
-
-@fig-idx-proj-dist shows a sample of traces from the indexes in
-@tbl-smoothness-squintability for the 6D sine wave data. This traces the
-PP index values computed for projections along an interpolated path
-between a randomly generated noise projection and the optimal projection (sine wave). (Note
-index values were self-standardized to range between 0 and 1, for
-comparison.) The `splines` index is very smooth with a gradual increase
-in value between the noise projection and the optimal projection. The
-`TIC` index is slightly less smooth and slightly less gradual change in
-value. Least desirable is the `skinny` index which is noisier and the
-optimization needs to be much closer to the optimal angle to have a
-higher value. This illustrates how the new measures for squintability
-and smoothness work to describe the PP indexes.
-
 ```{r}
@@
   filter(shape == "sine", index %in% c("TIC", "splines", "skinny")) |>
+  mutate(index = factor(index, levels = c("splines", "TIC", "skinny"))) |> 
   select(shape, index, d, smooth, squint) |>
-  arrange(desc(squint)) |>
+  arrange(index, d) |>
   knitr::kable(
     digits = 4, format = "latex", 
-    col.names = c(
-      "shape", "index", "d",
-      #"$\\eta$", "$\\ell$", 
-      "$\\nu$", #"$\\sigma$",
-      #"$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", 
-      "$\\varsigma$"), linesep = "",
-    booktabs = TRUE, escape = FALSE, row.names = FALSE, align = "|llr|rr") #align = "|llr|rrrr|rrrrr") |> 
+    col.names = c("shape", "index", "d", "$\\nu$", "$\\varsigma$"), linesep = c("", "", "\\hline"),
+    booktabs = TRUE, escape = FALSE, row.names = FALSE, align = "|llr|rr") |> 
   #column_spec(1, border_left = TRUE) |>
-  #column_spec(c(4, 8, 13), border_right = TRUE) |> 
+  column_spec(c(3,5), border_right = TRUE) 
+  #row_spec(c(3, 6), extra_latex_after = "\\cline{1-5}")
   #kable_styling(font_size = 10)   
@@
 
-@tbl-mod-output presents the results from fitting a logistic regression
-model using the proportion of success as the response and smoothness,
-squintability, dimension, number of jellyfish and maximum number of
-iteration, as predictors. The fit suggests that the JSO success rate is only
-affected by squintability, dimension and number of jellyfish. As expected,
-the JSO success rate is higher with higher squintability and/or more
-jellyfish, and lower when dimension is higher (higher-dimensional
-optimization is more difficult). Interestingly, the JSO is unaffected by
-the smoothness of the index function. This is consistent with the way
-random search algorithms jump from value to value without taking local
-regularity into account, as opposed to gradient-based optimizers for
-example. Allowing for more iterations also does not affect success rate
-significantly. A unit increase in squintability increases the success
-rate by `r round(100 * (exp(broom::tidy(mod1)$estimate[3])-1), 0)`%. As
-dimension increases by one the success rate almost halves. Increasing
-the number of jellies by 10 increases the success rate by
-`r round(100 * (exp(broom::tidy(mod1)$estimate[5])-1), 0)`%.
-@tbl-joint-output shows the fit statistics supporting that the model is
-reasonably strong.
-
-Note that the addition of interaction terms for squintability and data
-dimension, and smoothness and data dimension, improves the model a small
-amount, but does not change the interpretation above. This study intends 
-to serve as an illustration that the proposed
-measures effectively captures the key factors influencing the success
-rate of PP.
+## Effect of index properties on JSO success rate
 
@@
 
+@tbl-smoothness-squintability presents values computed for a subset of
+PP indexes on the sine wave data for dimensions 4-8. (A full table of
+the parameters estimated from the Gaussian process (outputscale $\eta$,
+lengthscale $\ell$, smoothness $\nu$, and nugget $\sigma$) and from the
+scaled logistic function ($\theta_1$ to $\theta_4$) for calculating
+smoothness and squintability for all PP indexes and data combinations is
+included in the supplementary materials). The column $\nu$ is used as
+the smoothness metric and the column $\varsigma$ is calculated as
+equation (@eq-squintability-parametric) as the squintability metric.
+
+@fig-idx-proj-dist shows a sample of traces from the indexes in
+@tbl-smoothness-squintability for the 6D sine wave data. This traces the
+PP index values computed for projections along an interpolated path
+between a randomly generated noise projection and the optimal projection (sine wave). (Note
+index values were self-standardized to range between 0 and 1, for
+comparison.) The `splines` index is very smooth with a gradual increase
+in value between the noise projection and the optimal projection. The
+`TIC` index is slightly less smooth and slightly less gradual change in
+value. Least desirable is the `skinny` index which is noisier and the
+optimization needs to be much closer to the optimal angle to have a
+higher value. This illustrates how the new measures for squintability
+and smoothness work to describe the PP indexes.
+@tbl-mod-output presents the results from fitting a logistic regression
+model using the proportion of success as the response and smoothness,
+squintability, dimension, number of jellyfish and maximum number of
+iteration, as predictors. The fit suggests that the JSO success rate is only
+affected by squintability, dimension and number of jellyfish. As expected,
+the JSO success rate is higher with higher squintability and/or more
+jellyfish, and lower when dimension is higher (higher-dimensional
+optimization is more difficult). Interestingly, the JSO is unaffected by
+the smoothness of the index function. This is consistent with the way
+random search algorithms jump from value to value without taking local
+regularity into account, as opposed to gradient-based optimizers for
+example. Allowing for more iterations also does not affect success rate
+significantly. A unit increase in squintability increases the success
+rate by `r round(100 * (exp(broom::tidy(mod1)$estimate[3])-1), 0)`%. As
+dimension increases by one the success rate almost halves. Increasing
+the number of jellies by 10 increases the success rate by
+`r round(100 * (exp(broom::tidy(mod1)$estimate[5])-1), 0)`%.
+@tbl-joint-output shows the fit statistics supporting that the model is
+reasonably strong.
+
+Note that the addition of interaction terms for squintability and data
+dimension, and smoothness and data dimension, improves the model a small
+amount, but does not change the interpretation above. This study intends 
+to serve as an illustration that the proposed
+measures effectively captures the key factors influencing the success
+rate of PP.
+
 # Practicalities {#sec-discussion}

### 2025-03-05 jso.tex
small changes on tables

@@
 
-\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, colback=white, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, breakable, toprule=.15mm, leftrule=.75mm, opacityback=0, opacitybacktitle=0.6, toptitle=1mm, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, opacityback=0, arc=.35mm, opacitybacktitle=0.6, left=2mm, colback=white, bottomtitle=1mm, titlerule=0mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, rightrule=.15mm, coltitle=black, toprule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm]
 
@@
 
+\begin{table}
+
+\caption{\label{tbl-smoothness-squintability}The columns \(\nu\) and
+\(\varsigma\) show the smoothness and squintability measures for three
+PP indexes for the sine wave data. The splines index has the largest
+squint angle and is the smoothest. The skinny index has low smoothness
+and smallest squint angle. As dimension increases the squint angle gets
+smaller but smoothness doesn't change much.}
+
+\centering{
+
+\begin{tabular}{|llr|rr|llr|rr|llr|rr|llr|rr|llr|rr}
+\toprule
+shape & index & d & $\nu$ & $\varsigma$\\
+\midrule
+sine & splines & 4 & 3.5489 & 0.5723\\
+sine & splines & 6 & 3.1149 & 0.5415\\
+sine & splines & 8 & 2.7400 & 0.5394\\
+\hline
+sine & TIC & 4 & 3.4854 & 0.3839\\
+sine & TIC & 6 & 3.0807 & 0.3747\\
+sine & TIC & 8 & 2.7223 & 0.3785\\
+\hline
+sine & skinny & 4 & 2.5213 & 0.0810\\
+sine & skinny & 6 & 2.2352 & 0.0973\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{table}%
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-idx-proj-dist-1.pdf}}
+
+}
+
+\caption{\label{fig-idx-proj-dist}Five traces of index value against
+projection distance for each of the three indexes (TIC, spline, and
+skinny) on 6D sine wave data. The traces of the splines index are
+smooth, with a gradual change between noise and otimal projection. In
+contrast, the skinny index has a noisy trace that increases rapidly near
+the optimum. This illustrates the usefulness of the new measures for
+squintability and smoothness.}
+
+\end{figure}%
+
 \subsection{Effect of index properties on JSO success
 rate}\label{effect-of-index-properties-on-jso-success-rate}
+
+\begingroup
+\fontsize{12.0pt}{14.4pt}\selectfont
+\setlength{\LTpost}{0mm}
+
+\begin{longtable}{lcc}
+
+\caption{\label{tbl-mod-output}Jellyfish success rate relative to index
+properties and jellyfish hyper-parameters. This is the summary from a
+logistic regression fit to smoothness, squintability, dimension, number
+of jellyfish and maximum number of iterations. Interestingly,
+squintability and dimension strongly affect jellyfish optimization
+success. The number of jellies marginally affects success, but index
+smoothness, and increasing the number of iterations do not.}
+
+\tabularnewline
+
+\toprule
+\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
+\midrule\addlinespace[2.5pt]
+Smoothness & 1.01 (0.92 to 1.11) & 0.810 \\ 
+Squintability & 1.23 (1.12 to 1.38) & {\bfseries <0.001} \\ 
+Dimension & 0.58 (0.45 to 0.73) & {\bfseries <0.001} \\ 
+Number of jellyfish & 1.32 (1.17 to 1.50) & {\bfseries <0.001} \\ 
+Maximum number of iterations & 1.06 (0.94 to 1.20) & 0.328 \\ 
+\bottomrule
+
+\end{longtable}
+
+\begin{minipage}{\linewidth}
+\textsuperscript{\textit{1}}OR = Odds Ratio, CI = Confidence Interval\\
+\end{minipage}
+\endgroup
+
+\begingroup\fontsize{10}{12}\selectfont
+
+\begin{longtable}[t]{rrrr}
+
+\caption{\label{tbl-joint-output}Fit statistics for the model.}
+
+\tabularnewline
+
+\\
+\toprule
+\textbf{Deviance} & \textbf{DF Residual} & \textbf{Null Deviance} & \textbf{DF Null}\\
+\midrule
+20.89 & 71 & 49.77 & 76\\
+\bottomrule
+
+\end{longtable}
+
+\endgroup{}
 
@@
 squintability and smoothness work to describe the PP indexes.
-
-\begin{table}
-
-\caption{\label{tbl-smoothness-squintability}The columns \(\nu\) and
-\(\varsigma\) show the smoothness and squintability measures for three
-PP indexes for the sine wave data. The splines index has the largest
-squint angle and is the smoothest. The skinny index has low smoothness
-and smallest squint angle. As dimension increases the squint angle gets
-smaller but smoothness doesn't change much.}
-
-\centering{
-
-\begin{tabular}{|llr|rr|llr|rr|llr|rr|llr|rr|llr|rr}
-\toprule
-shape & index & d & $\nu$ & $\varsigma$\\
-\midrule
-sine & splines & 4 & 3.5489 & 0.5723\\
-sine & splines & 6 & 3.1149 & 0.5415\\
-sine & splines & 8 & 2.7400 & 0.5394\\
-sine & TIC & 4 & 3.4854 & 0.3839\\
-sine & TIC & 8 & 2.7223 & 0.3785\\
-sine & TIC & 6 & 3.0807 & 0.3747\\
-sine & skinny & 6 & 2.2352 & 0.0973\\
-sine & skinny & 4 & 2.5213 & 0.0810\\
-\bottomrule
-\end{tabular}
-
-}
-
-\end{table}%
-
-\begin{figure}
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-idx-proj-dist-1.pdf}}
-
-}
-
-\caption{\label{fig-idx-proj-dist}Five traces of index value against
-projection distance for each of the three indexes (TIC, spline, and
-skinny) on 6D sine wave data. The traces of the splines index are
-smooth, with a gradual change between noise and otimal projection. In
-contrast, the skinny index has a noisy trace that increases rapidly near
-the optimum. This illustrates the usefulness of the new measures for
-squintability and smoothness.}
-
-\end{figure}%
-
 Table~\ref{tbl-mod-output} presents the results from fitting a logistic
@@
 
-\begingroup
-\fontsize{12.0pt}{14.4pt}\selectfont
-\setlength{\LTpost}{0mm}
-
-\begin{longtable}{lcc}
-
-\caption{\label{tbl-mod-output}Jellyfish success rate relative to index
-properties and jellyfish hyper-parameters. This is the summary from a
-logistic regression fit to smoothness, squintability, dimension, number
-of jellyfish and maximum number of iterations. Interestingly,
-squintability and dimension strongly affect jellyfish optimization
-success. The number of jellies marginally affects success, but index
-smoothness, and increasing the number of iterations do not.}
-
-\tabularnewline
-
-\toprule
-\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
-\midrule\addlinespace[2.5pt]
-Smoothness & 1.01 (0.92 to 1.11) & 0.810 \\ 
-Squintability & 1.23 (1.12 to 1.38) & {\bfseries <0.001} \\ 
-Dimension & 0.58 (0.45 to 0.73) & {\bfseries <0.001} \\ 
-Number of jellyfish & 1.32 (1.17 to 1.50) & {\bfseries <0.001} \\ 
-Maximum number of iterations & 1.06 (0.94 to 1.20) & 0.328 \\ 
-\bottomrule
-
-\end{longtable}
-
-\begin{minipage}{\linewidth}
-\textsuperscript{\textit{1}}OR = Odds Ratio, CI = Confidence Interval\\
-\end{minipage}
-\endgroup
-
-\begingroup\fontsize{10}{12}\selectfont
-
-\begin{longtable}[t]{rrrr}
-
-\caption{\label{tbl-joint-output}Fit statistics for the model.}
-
-\tabularnewline
-
-\\
-\toprule
-\textbf{Deviance} & \textbf{DF Residual} & \textbf{Null Deviance} & \textbf{DF Null}\\
-\midrule
-20.89 & 71 & 49.77 & 76\\
-\bottomrule
-
-\end{longtable}
-
-\endgroup{}
-
 \section{Practicalities}\label{sec-discussion}

### 2025-03-05 supplementary/supplementary.qmd
small changes on tables

@@
 
-# Gaussian process estimates for all PP indexes and data sets
+# Calculated smoothness and squintability metrics for all PP indexes and data sets
+
+Smoothness and squintability are metrics used to characterize the index 
+function for PP tasks and are invariant to the optimizer and its 
+hyper-parameters. @tbl-smoothness-squintability presents the Gaussian process
+parameters (outputscale $\eta$, lengthscale $\ell$, smoothness $\nu$, and 
+nugget $\sigma$) and scaled logistic function parameters ($\theta_1$ to 
+$\theta_4$ and $\varsigma$) for all 23 PP problems investigated in our work. 
+The columns $\nu$ and $\varsigma$ represent the smoothness and squintability 
+measures respectively.
 
@@
 The columns $\nu$ and $\varsigma$ represent the smoothness and squintability measures respectively.
-dt <- tibble(shape = c(rep("pipe", 5), rep("sine", 17)), smoothness) |>
+dt <- tibble(shape = c(rep("pipe", 5), rep("sine", 18)), smoothness) |>
   left_join(squintability) |> 
   rename(d = n, smooth = smoothness) |> 
-  mutate(index = ifelse(index == "dcor2d_2", "dcor", index),
+  mutate(index = as.character(index),
+         index = ifelse(index == "dcor2d_2", "dcor", index),
          index = ifelse(index == "loess2d", "loess", index),

### 2025-03-05 jso.qmd
added Gaussian table to list of items in supplementary and a few typos fixed

@@
 additional PP indexes that should be able to extract the sine wave are
-used: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `splines2d`, `skinny` and `stringy2`,
-following @laa_using_2020. (The supplementary material includes
+used (dcor, loess, MIC, TIC, splines, skinny and stringy2) following @laa_using_2020. 
+(The supplementary material includes
 definitions of these PP indexes, and descriptions of the pipe and sine
-wave data simulation.). The JSO hyper-parameters, number of jellyfish and
+wave data simulation.) The JSO hyper-parameters, number of jellyfish and
 maximum iterations, are also included. This results in a total of
@@
 index values were self-standardized to range between 0 and 1, for
-comparison.) The `splines` index is very smooth with a gradual increase
+comparison.) The splines index is very smooth with a gradual increase
 in value between the noise projection and the optimal projection. The
-`TIC` index is slightly less smooth and slightly less gradual change in
-value. Least desirable is the `skinny` index which is noisier and the
+TIC index is slightly less smooth and slightly less gradual change in
+value. Least desirable is the skinny index which is noisier and the
 optimization needs to be much closer to the optimal angle to have a
@@
 with using JSO in a PPGT and calculating smoothness and squintability,
-as explained in @sec-discussion, and (3) the full code to reproduce the
-plots and summaries in this paper.
+as explained in @sec-discussion, (3) the table of the Gaussian
+process parameters and logistic function parameters for all 23 PP problems investigated, along with (4) the full code to reproduce the plots and summaries in this paper.
 

### 2025-03-05 jso.tex
added Gaussian table to list of items in supplementary and a few typos fixed

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, opacityback=0, arc=.35mm, opacitybacktitle=0.6, left=2mm, colback=white, bottomtitle=1mm, titlerule=0mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomrule=.15mm, rightrule=.15mm, coltitle=black, toprule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, toptitle=1mm, opacityback=0, coltitle=black, titlerule=0mm, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, rightrule=.15mm, bottomrule=.15mm, toprule=.15mm, left=2mm, colback=white]
 
@@
 additional PP indexes that should be able to extract the sine wave are
-used: \texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
-\texttt{splines2d}, \texttt{skinny} and \texttt{stringy2}, following Laa
+used (dcor, loess, MIC, TIC, splines, skinny and stringy2) following Laa
 and Cook (2020). (The supplementary material includes definitions of
 these PP indexes, and descriptions of the pipe and sine wave data
-simulation.). The JSO hyper-parameters, number of jellyfish and maximum
+simulation.) The JSO hyper-parameters, number of jellyfish and maximum
 iterations, are also included. This results in a total of 77 scenarios,
@@
 Smoothness & 1.01 (0.92 to 1.11) & 0.810 \\ 
-Squintability & 1.23 (1.12 to 1.38) & {\bfseries <0.001} \\ 
-Dimension & 0.58 (0.45 to 0.73) & {\bfseries <0.001} \\ 
-Number of jellyfish & 1.32 (1.17 to 1.50) & {\bfseries <0.001} \\ 
+Squintability & 1.23 (1.12 to 1.38) & \textbf{<0.001} \\ 
+Dimension & 0.58 (0.45 to 0.73) & \textbf{<0.001} \\ 
+Number of jellyfish & 1.32 (1.17 to 1.50) & \textbf{<0.001} \\ 
 Maximum number of iterations & 1.06 (0.94 to 1.20) & 0.328 \\ 
@@
 optimal projection (sine wave). (Note index values were
-self-standardized to range between 0 and 1, for comparison.) The
-\texttt{splines} index is very smooth with a gradual increase in value
-between the noise projection and the optimal projection. The
-\texttt{TIC} index is slightly less smooth and slightly less gradual
-change in value. Least desirable is the \texttt{skinny} index which is
-noisier and the optimization needs to be much closer to the optimal
-angle to have a higher value. This illustrates how the new measures for
-squintability and smoothness work to describe the PP indexes.
-Table~\ref{tbl-mod-output} presents the results from fitting a logistic
-regression model using the proportion of success as the response and
-smoothness, squintability, dimension, number of jellyfish and maximum
-number of iteration, as predictors. The fit suggests that the JSO
-success rate is only affected by squintability, dimension and number of
-jellyfish. As expected, the JSO success rate is higher with higher
-squintability and/or more jellyfish, and lower when dimension is higher
-(higher-dimensional optimization is more difficult). Interestingly, the
-JSO is unaffected by the smoothness of the index function. This is
-consistent with the way random search algorithms jump from value to
-value without taking local regularity into account, as opposed to
-gradient-based optimizers for example. Allowing for more iterations also
-does not affect success rate significantly. A unit increase in
-squintability increases the success rate by 23\%. As dimension increases
-by one the success rate almost halves. Increasing the number of jellies
-by 10 increases the success rate by 32\%. Table~\ref{tbl-joint-output}
-shows the fit statistics supporting that the model is reasonably strong.
+self-standardized to range between 0 and 1, for comparison.) The splines
+index is very smooth with a gradual increase in value between the noise
+projection and the optimal projection. The TIC index is slightly less
+smooth and slightly less gradual change in value. Least desirable is the
+skinny index which is noisier and the optimization needs to be much
+closer to the optimal angle to have a higher value. This illustrates how
+the new measures for squintability and smoothness work to describe the
+PP indexes. Table~\ref{tbl-mod-output} presents the results from fitting
+a logistic regression model using the proportion of success as the
+response and smoothness, squintability, dimension, number of jellyfish
+and maximum number of iteration, as predictors. The fit suggests that
+the JSO success rate is only affected by squintability, dimension and
+number of jellyfish. As expected, the JSO success rate is higher with
+higher squintability and/or more jellyfish, and lower when dimension is
+higher (higher-dimensional optimization is more difficult).
+Interestingly, the JSO is unaffected by the smoothness of the index
+function. This is consistent with the way random search algorithms jump
+from value to value without taking local regularity into account, as
+opposed to gradient-based optimizers for example. Allowing for more
+iterations also does not affect success rate significantly. A unit
+increase in squintability increases the success rate by 23\%. As
+dimension increases by one the success rate almost halves. Increasing
+the number of jellies by 10 increases the success rate by 32\%.
+Table~\ref{tbl-joint-output} shows the fit statistics supporting that
+the model is reasonably strong.
 
@@
 started with using JSO in a PPGT and calculating smoothness and
-squintability, as explained in Section~\ref{sec-discussion}, and (3) the
-full code to reproduce the plots and summaries in this paper.
+squintability, as explained in Section~\ref{sec-discussion}, (3) the
+table of the Gaussian process parameters and logistic function
+parameters for all 23 PP problems investigated, along with (4) the full
+code to reproduce the plots and summaries in this paper.
 

### 2025-03-05 jso.qmd
reference in the conclusion

@@
 noise indexes with low squintability metric, which resemble a needle-in-a-haystack 
-problem. Additionally, if the index function is differentiable, automatic 
-differentiation can be used to implement new gradient-based optimizers.
+problem [@siemenn2023fast]. Additionally, if the index function is differentiable, automatic 
+differentiation can be used to implement new gradient-based optimizers [@prince2023understanding].
 

### 2025-03-05 jso.tex
reference in the conclusion

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, toptitle=1mm, opacityback=0, coltitle=black, titlerule=0mm, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, rightrule=.15mm, bottomrule=.15mm, toprule=.15mm, left=2mm, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, breakable, titlerule=0mm, coltitle=black, colframe=quarto-callout-note-color-frame, rightrule=.15mm, bottomtitle=1mm, colback=white, opacityback=0, leftrule=.75mm, opacitybacktitle=0.6, toprule=.15mm, left=2mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, bottomrule=.15mm]
 
@@
 Smoothness & 1.01 (0.92 to 1.11) & 0.810 \\ 
-Squintability & 1.23 (1.12 to 1.38) & \textbf{<0.001} \\ 
-Dimension & 0.58 (0.45 to 0.73) & \textbf{<0.001} \\ 
-Number of jellyfish & 1.32 (1.17 to 1.50) & \textbf{<0.001} \\ 
+Squintability & 1.23 (1.12 to 1.38) & {\bfseries <0.001} \\ 
+Dimension & 0.58 (0.45 to 0.73) & {\bfseries <0.001} \\ 
+Number of jellyfish & 1.32 (1.17 to 1.50) & {\bfseries <0.001} \\ 
 Maximum number of iterations & 1.06 (0.94 to 1.20) & 0.328 \\ 
@@
 algorithm for noise indexes with low squintability metric, which
-resemble a needle-in-a-haystack problem. Additionally, if the index
-function is differentiable, automatic differentiation can be used to
-implement new gradient-based optimizers.
+resemble a needle-in-a-haystack problem (Siemenn et al. 2023).
+Additionally, if the index function is differentiable, automatic
+differentiation can be used to implement new gradient-based optimizers
+(Prince 2023).
 
@@
 
+\bibitem[\citeproctext]{ref-prince2023understanding}
+Prince, Simon JD. 2023. \emph{Understanding Deep Learning}. MIT press.
+
 \bibitem[\citeproctext]{ref-R}
@@
 
+\bibitem[\citeproctext]{ref-siemenn2023fast}
+Siemenn, Alexander E, Zekun Ren, Qianxiao Li, and Tonio Buonassisi.
+2023. {``Fast Bayesian Optimization of Needle-in-a-Haystack Problems
+Using Zooming Memory-Based Initialization (ZoMBI).''} \emph{Npj
+Computational Materials} 9 (1): 79.
+
 \bibitem[\citeproctext]{ref-gtsummary}

### 2025-03-09 jso.qmd
minor edit

@@
   - name: Jessica Wai Yin Leung
-    email: Jessica.Leung@monash.edu
+    email: jessica.leung@monash.edu
     affiliations:
@@
 Additionally, JSO's performance on detecting the sine-wave shape is evaluated using different PP indexes (hence different smoothness and squintability) across various data dimensions (d = 4, 6, 8, 10, 12) and JSO hyper-parameters.
+We observe empirically that higher squintability improves the success rate of the PP index optimization, while smoothness has no significant effect.
 The JSO algorithm has been implemented in the R package, `tourr`, and functions to calculate smoothness and squintability measures are implemented in the `ferrn` package.
@@
 $f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ defines a statistics to
-measure an pattern of interest in the projected data, such as deviation
+measure a pattern of interest in the projected data, such as deviation
 from normality, presence of clusters, or non-linear structure. For a
@@
 starting and target planes. It interpolates plane to plane, removing
-irrelevant within plane spin, and is agnostic to the basis ($A$) used to
+irrelevant within-plane spin, and is agnostic to the basis ($A$) used to
 define the plane. Thus, indexes which are used for the PPGT should be
@@
 estimate in practice. For example, the loess and splines indexes are not
-continuous due to the use of the $\max$ operator in their definitions.
+differentiable due to the use of the $\max$ operator in their definitions.
 Scagnostic indexes, such as skinny and stringy, are defined based on
@@
 dimensions 6, 8, 10, and 12 in @fig-proj. The columns show the quantile of 
-index value in the 50 optimal projection, with 0 being the minimum, 100 being 
+index value in the 50 optimal projections, with 0 being the minimum, 100 being 
 the maximum, and 50 the median. The purpose is to summarize the views of the data resulting from
@@
 are sufficient to achieve a high success rate. Many more
-jellyfish (100 jellyfish and 100 iterations) are neede for finding the 
+jellyfish (100 jellyfish and 100 iterations) are needed for finding the 
 pipe structure in d = 10, and 12. This suggests
 that even JSO will struggle with high dimensions, and require
-substantial more computational time to have an acceptable success rate,
+substantially more computational time to have an acceptable success rate,
 especially for difficult PP indexes based on scagnostics.
@@
 methods: 1) parametrically, by fitting a scaled sigmoid function through
-non-linear least square (`method = "nls"`), and 2) non-parametrically,
+non-linear least squares (`method = "nls"`), and 2) non-parametrically,
 using kernel smoothing (`method = "ks"`). A `bin_width` argument is

### 2025-03-09 jso.tex
minor edit

@@
 Wai Yin
-Leung. Email: \href{mailto:Jessica.Leung@monash.edu}{\nolinkurl{Jessica.Leung@monash.edu}}. }
+Leung. Email: \href{mailto:jessica.leung@monash.edu}{\nolinkurl{jessica.leung@monash.edu}}. }
 \begin{document}
@@
 (hence different smoothness and squintability) across various data
-dimensions (d = 4, 6, 8, 10, 12) and JSO hyper-parameters. The JSO
-algorithm has been implemented in the R package, \texttt{tourr}, and
+dimensions (d = 4, 6, 8, 10, 12) and JSO hyper-parameters. We observe
+empirically that higher squintability improves the success rate of the
+PP index optimization, while smoothness has no significant effect. The
+JSO algorithm has been implemented in the R package, \texttt{tourr}, and
 functions to calculate smoothness and squintability measures are
@@
 \(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) defines a statistics
-to measure an pattern of interest in the projected data, such as
+to measure a pattern of interest in the projected data, such as
 deviation from normality, presence of clusters, or non-linear structure.
@@
 starting and target planes. It interpolates plane to plane, removing
-irrelevant within plane spin, and is agnostic to the basis (\(A\)) used
+irrelevant within-plane spin, and is agnostic to the basis (\(A\)) used
 to define the plane. Thus, indexes which are used for the PPGT should be
@@
 or even estimate in practice. For example, the loess and splines indexes
-are not continuous due to the use of the \(\max\) operator in their
+are not differentiable due to the use of the \(\max\) operator in their
 definitions. Scagnostic indexes, such as skinny and stringy, are defined
@@
 
-\begin{tcolorbox}[enhanced jigsaw, breakable, titlerule=0mm, coltitle=black, colframe=quarto-callout-note-color-frame, rightrule=.15mm, bottomtitle=1mm, colback=white, opacityback=0, leftrule=.75mm, opacitybacktitle=0.6, toprule=.15mm, left=2mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, bottomrule=.15mm]
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, breakable, rightrule=.15mm, toptitle=1mm, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, arc=.35mm, coltitle=black, colback=white, opacityback=0, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, bottomrule=.15mm, leftrule=.75mm, opacitybacktitle=0.6]
 
@@
 at dimensions 6, 8, 10, and 12 in Figure~\ref{fig-proj}. The columns
-show the quantile of index value in the 50 optimal projection, with 0
+show the quantile of index value in the 50 optimal projections, with 0
 being the minimum, 100 being the maximum, and 50 the median. The purpose
@@
 sufficient to achieve a high success rate. Many more jellyfish (100
-jellyfish and 100 iterations) are neede for finding the pipe structure
+jellyfish and 100 iterations) are needed for finding the pipe structure
 in d = 10, and 12. This suggests that even JSO will struggle with high
-dimensions, and require substantial more computational time to have an
+dimensions, and require substantially more computational time to have an
 acceptable success rate, especially for difficult PP indexes based on
@@
 two methods: 1) parametrically, by fitting a scaled sigmoid function
-through non-linear least square (\texttt{method\ =\ "nls"}), and 2)
+through non-linear least squares (\texttt{method\ =\ "nls"}), and 2)
 non-parametrically, using kernel smoothing (\texttt{method\ =\ "ks"}). A

### 2025-10-23 revision/response.qmd
revision dot 5 and part of 6

---
title: "Response to Reviewers JCGS-25-135"
subtitle: "Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimization Choices"
author: "H. Sherry Zhang, Dianne Cook, Nicolas  Langrené, Jessica Wai Yin Leung"
date: "2025-10-06"
format: pdf
---

We would like to thank the reviewers and editor for taking the time to review our paper and provide constructive comments. Below is a point-by-point explanation of how we have changed the paper. Reviewers' comments are in normal text and **\textcolor{purple}{our responses are in red}**.

# Reviewer 1

1. The paper introduces two interesting new developments to projection pursuit research: the quantification of index properties that enables better comparison, and a new optimization approach based on Jellyfish Search Optimizer. The results show the good performance of the new optimization strategy, and how it can be related to the new metrics introduced. While the overall presentation is clear, it feels like the authors are conflicted about what the main focus of the paper actually should be. The introduction mentions that the primary goal is to investigate the potential for the jellyfish search optimization, however the title does not even mention this, and much of the paper seems to focus on the new metrics introduced. Having a more well defined focus should help guide the flow of the paper better (currently it feels like jumping around between the two objectives).

**\textcolor{purple}{TODO: Di}**.

2. I also think additional clarifications are needed in the definition of squintability. It seems to me Definition 4 is simply defining an angle between planes (without reference to "squinting"), while in the following the definition of squintability is based on index values at given projection distances?

**\textcolor{purple}{The concept of squintability comes from squint angle, which is formally defined in Definition 4. We formally define squintability in Definition 6 as the change of index value when the projection distance is reduced by half. TODO}**.

3. In Definition 6 it is not clear why g should be strictly a decreasing function? Also it seems it is implicitly assumed that the index function will depend only on the projection distance, but not on the direction, and that in practice the index values across directions are averaged to find a single value for g at any given distance? In practice I would expect potentially large deviations from such an assumption, in particular since there may be local optima in certain directions.

**\textcolor{purple}{our responses are in red}**.

4. The definition in Eq. (5) depends on the selected distance r_0, but it is not explained how this is selected or derived.

**\textcolor{purple}{our responses are in red}**.

5. Some more practical aspects of how the new metrics are defined are given later in the paper (Fig. 5 and 6), in my opinion the illustrations in the Figures are not needed, but the practical details should be explained together with the definitions in section 3. For the smoothness metric the examples in Fig 2 and 3 are useful to get an intuition about the index, and something comparable would help with better understanding the squintability metric (probably some additional information on something like Fig 9 would help here).

**\textcolor{purple}{We decide to keep the two figures (Now Figure 6 and 7) in Section 5.2 since some audience may benefit from the illustration. To provide some intuition about the index for motivating squintability, we include a new figure, Figure 4, in Section 3.2 to illustrate the trace for different indexes.}**

6. The results in section 6 are very interesting. However, I was wondering about some details in 6.3: from the results shown it seems that the smoothness and squintability for the selected data examples are highly correlated, could this cause issues with the model fit shown in Table 2? It would be interesting if a similar result for CRS could be included, to see the effect of smoothness in that setting.

**\textcolor{purple}{The correlation between the two variables are 0.55, which should not affect the modelling. (TODO: Di - do you mind checking if we need more here?) We did try CRS for the same sine-wave optimisation reported in Section 6, but most runs yield a failed result with the optimum not found by CRS, hence not included.}**

7. I was also wondering how the dependence of the smoothness on d can be explained, since it is not intuitive that this should change so much? (Note also that the last sentence in the caption for Table 1 seems to mix up squintability and smoothness.)

**\textcolor{purple}{TODO The caption has been fixed.}**.

8. Finally, I appreciate the practical advice in section 7. Here, 7.1 should be somewhat extended in my opinion: does the planned tour then run through all the individual paths, should we follow the most successful jellyfish, or how would we pick?

**\textcolor{purple}{our responses are in red}**.

Finally I have also collected a few minor remarks:

  * After defining Eq. 4 the additional parameter should probably be theta_4 and not theta_2.  **\textcolor{purple}{our responses are in red}**.
  * In the algorithm pseudo-code it is not clear what c_t and the trend actually are.  **\textcolor{purple}{our responses are in red}**.
  * In Fig. 4 please specify what optimizer was used. From the last sentence of Section 5, it is not clear to me why the ranks need to be used, why not scale the metrics?  **\textcolor{purple}{our responses are in red}**.
  * In 6.2 it is not really clear what is being bootstrapped for the evaluation of the uncertainty. **\textcolor{purple}{our responses are in red}**.

### 2025-10-23 jso.qmd
revision dot 5 and part of 6

@@
 load(here::here("data/pipe_jellyfish.rda"))
+load(here::here("data/sq_sine_basis_df.rda"))
 load(here::here("data/smoothness.rda"))
@@
 small squint angle is considered to be undesirable because it means that
-the optimizer needs to be very close to be able to "see" the optimum.
-Thus, it could be difficult for the optimizer to find the optimum.
+the optimizer needs to be very close to be able to "see" the optimum, making it difficult for the optimizer to find the optimum.
 
@@
 for the structure of the index function to be visible enough to be
-amenable to efficient optimization. This observation suggests that the
-half-point index value improvement percentage provides an appropriate
+amenable to efficient optimization.
+This observation, shown in the simulated trace from different index functions in @fig-squintability-ill, suggests that the half-point index value improvement percentage provides an appropriate
 mathematical definition of squintability, which matches the intuition
 behind this concept, while being amenable to numerical computation.
+
+```{r}
+#| label: fig-squintability-ill
+#| fig.width: 8
+#| fig.height: 4
+#| fig.cap: "Simulated traces of index value as a function of projection distances for four PP index functions: MIC, splines2d, skinny, and stringy2.
+The index function MIC and splines2d make early progression towards the optimal index value, indicating a large squint angle, whereas the traces from skinny, and specially stringy2, show improvement only near the optimal projection, suggesting low squintability."
+sq_sine_basis_df |> 
+  filter(n == 4) |> 
+  unnest(basis_df, names_repair = "unique") |> 
+  filter(id <= 3) |> 
+  rename(index = index...2, value = index...8) |> 
+  filter(index %in% c("MIC", "skinny", "stringy2", "splines2d")) |> 
+  group_by(id, index, n) |> 
+  mutate(value = (value - min(value)) / (max(value) - min(value))) |>
+  ggplot(aes(x = dist, y = value, 
+             group = interaction(index, id, n), 
+             color = index)) +
+  geom_line() + 
+  scale_color_brewer(palette = "Dark2") +
+  xlab("Projection distance") + ylab("Index value") +
+  theme_bw() + 
+  theme(legend.position = "bottom")
+```
 
@@
 $g(0)-g(r_{0})$. As a result, equation \eqref{eq-squintability}
-represent the proportion of this maximum improvement which has been
+represents the proportion of this maximum improvement which has been
 achieved by the time the distance $r_0$ to the optimal matrix has been
@@
 The skinny index has low smoothness and smallest squint angle.
-As dimension increases the squint angle gets smaller but smoothness doesn't change much.
+As dimension increases the smoothness becomes smaller but squintability doesn't change much.
 dt <- tibble(shape = c(rep("pipe", 5), rep("sine", 18)), smoothness) |>
@@
 sim_df2 <- sim_df |>  
+  # the original smoothness and squintability score is in the data
+  # but we use their rank in the modelling.
+  select(-smoothness, -squintability) |> 
+  rename(smoothness = smoothness_rank,
+         squintability = squintability_rank) |>
   mutate(n_jellies = n_jellies/10, max_tries = max_tries/10) |> 
-  rename(proportion = P_J) #|>
-  #filter(d < 10)
+  rename(proportion = P_J)
 
@@
   geom_line() + 
-  scale_color_brewer(palette = "Dark2") +
+  scale_color_brewer(palette = "Dark2", name = "Index function") +
   xlab("Projection distance") + ylab("Index value") +

### 2025-10-23 jso.tex
revision dot 5 and part of 6

@@
+% Options for packages loaded elsewhere
 % Options for packages loaded elsewhere
@@
 ]{interact}
-
+\usepackage{xcolor}
 \usepackage{amsmath,amssymb}
-\usepackage{setspace}
+\setcounter{secnumdepth}{5}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
@@
 }{}
+\usepackage{setspace}
 \makeatletter
@@
 \makeatother
-\usepackage{xcolor}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
 % Make \paragraph and \subparagraph free-standing
@@
 
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{longtable,booktabs,array}
 \usepackage{calc} % for calculating minipage widths
@@
 \makeatother
+
 % definitions for citeproc citations
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \makeatother
-
 \usepackage{bookmark}
-
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
@@
 small squint angle is considered to be undesirable because it means that
-the optimizer needs to be very close to be able to ``see'' the optimum.
-Thus, it could be difficult for the optimizer to find the optimum.
+the optimizer needs to be very close to be able to ``see'' the optimum,
+making it difficult for the optimizer to find the optimum.
 
@@
 for the structure of the index function to be visible enough to be
-amenable to efficient optimization. This observation suggests that the
-half-point index value improvement percentage provides an appropriate
-mathematical definition of squintability, which matches the intuition
-behind this concept, while being amenable to numerical computation.
+amenable to efficient optimization. This observation, shown in the
+simulated paths from different index functions in
+Figure~\ref{fig-squintability-ill}, suggests that the half-point index
+value improvement percentage provides an appropriate mathematical
+definition of squintability, which matches the intuition behind this
+concept, while being amenable to numerical computation.
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-squintability-ill-1.pdf}}
+
+}
+
+\caption{\label{fig-squintability-ill}Simulated traces of index value as
+a function of projection distances for four PP index functions: MIC,
+splines2d, skinny, and stringy2. The index function MIC and splines2d
+make early progression towards the optimal index value, indicating a
+large squint angle, whereas the traces from skinny, and specially
+stringy2, show improvement only near the optimal projection, suggesting
+low squintability.}
+
+\end{figure}%
 
@@
 \(g(0)-g(r_{0})\). As a result, equation~\eqref{eq-squintability}
-represent the proportion of this maximum improvement which has been
+represents the proportion of this maximum improvement which has been
 achieved by the time the distance \(r_0\) to the optimal matrix has been
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, bottomtitle=1mm, breakable, rightrule=.15mm, toptitle=1mm, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, arc=.35mm, coltitle=black, colback=white, opacityback=0, left=2mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm, bottomrule=.15mm, leftrule=.75mm, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, breakable, opacityback=0, toprule=.15mm, toptitle=1mm, titlerule=0mm, left=2mm, leftrule=.75mm, colframe=quarto-callout-note-color-frame, coltitle=black, bottomrule=.15mm, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, colback=white, opacitybacktitle=0.6, bottomtitle=1mm]
 
@@
 
-\begingroup
-\fontsize{12.0pt}{14.4pt}\selectfont
-\setlength{\LTpost}{0mm}
-
-\begin{longtable}{lcc}
+\begin{table}[H]
 
@@
 
-\tabularnewline
-
+\centering{
+
+\fontsize{12.0pt}{14.4pt}\selectfont
+\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lcc}
 \toprule
-\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
+\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)} & \textbf{p-value} \\ 
 \midrule\addlinespace[2.5pt]
@@
 \bottomrule
-
-\end{longtable}
-
+\end{tabular*}
 \begin{minipage}{\linewidth}
-\textsuperscript{\textit{1}}OR = Odds Ratio, CI = Confidence Interval\\
+Abbreviations: CI = Confidence Interval, OR = Odds Ratio\\
 \end{minipage}
-\endgroup
+
+}
+
+\end{table}%
 

### 2025-10-23 code/08-process-data-for-modelling.R

@@ -145,8 +145,14 @@ save(sim_summary, file = here::here("data", "sim_summary.rda"))
 load(here::here("data", "squintability.rda"))
 load(here::here("data", "smoothness.rda"))
 sim_df <- sim_summary |>
-  left_join(smoothness |> mutate(smoothness = rank(smoothness)) |> select(n, index, smoothness) |> rename(d = n)) |>
-  left_join(squintability |> mutate(squint = rank(squint)) |> select(index, n, squint) |> rename(d = n, squintability = squint))
+  left_join(smoothness |>
+              mutate(smoothness_rank = rank(smoothness)) |>
+              select(n, index, smoothness, smoothness_rank) |>
+              rename(d = n)) |>
+  left_join(squintability |>
+              mutate(squint_rank = rank(squint)) |>
+              select(index, n, squint, squint_rank) |>
+              rename(d = n, squintability = squint, squintability_rank = squint_rank))
 save(sim_df, file = here::here("data", "sim_df.rda"))
 
 

### 2025-10-24 revision/response.qmd
more revision

@@
 
-**\textcolor{purple}{TODO The caption has been fixed.}**.
+**\textcolor{purple}{TODO The caption has been fixed.}**
 
@@
 
-**\textcolor{purple}{our responses are in red}**.
+**\textcolor{purple}{We make Section 7.1 clearer that the planned tour can be run to trace the optimisation of individual jellyfish path.
+Users may wish to follow the most successful jellyfish, a random selection, or interesting jellyfish that result in local optimum.
+DISCUSSION: should I include a small example with the code?}**
 
@@
 
-  * After defining Eq. 4 the additional parameter should probably be theta_4 and not theta_2.  **\textcolor{purple}{our responses are in red}**.
-  * In the algorithm pseudo-code it is not clear what c_t and the trend actually are.  **\textcolor{purple}{our responses are in red}**.
-  * In Fig.
-4 please specify what optimizer was used.
-From the last sentence of Section 5, it is not clear to me why the ranks need to be used, why not scale the metrics?
-**\textcolor{purple}{our responses are in red}**.
-  * In 6.2 it is not really clear what is being bootstrapped for the evaluation of the uncertainty. **\textcolor{purple}{our responses are in red}**.
+  * After defining Eq. 4 the additional parameter should probably be theta_4 and not theta_2.  **\textcolor{purple}{our responses are in red.}**
+  * In the algorithm pseudo-code it is not clear what c_t and the trend actually are.
+**\textcolor{purple}{In the text and the pseudo-code, we include information that $c_t$ is the time control function determine the exploration and exploitation phase of the algorithm.
+We also include information that the (ocean) trend is defined as the difference between the best jellyfish and the average of all current jellyfish.}**
+  * In Fig. 4 please specify what optimizer was used. **\textcolor{purple}{We have included the information in the caption that the CRS optimised is used for this illustration. }**
+  * From the last sentence of Section 5, it is not clear to me why the ranks need to be used, why not scale the metrics?  **\textcolor{purple}{our responses are in red}**.
+  * In 6.2 it is not really clear what is being bootstrapped for the evaluation of the uncertainty.
+**\textcolor{purple}{We have included further description in the paper to describe the bootstrap procedure as drawing 500 bootstrap samples with replacement from the original 50 simulations for each setting: an optimisation with high-dimensional data (d = 4, 6, 8, 10, or 12) with different number of jellyfish (20, 50, or 100) and maximum number of iteration (50 or 100)}**.

### 2025-10-24 jso.qmd
more revision

@@
 projection is evaluated by the index function. The projection is then
-moved in a direction determined by a random factor, influenced by how
-far along we are in the optimization process. Occasionally, completely
-new directions may be taken like a jellyfish might with ocean currents.
+moved in a direction determined by a random time control function, $c_t$, whose value decreases as 
+the number of trials increases (subject to randomness) to guide the exploration and exploitation phases 
+of the algorithm. When $c_t \ge 0.5$, new 
+directions will be taken like a jellyfish explores with the ocean trend, defined as 
+the difference of the current best jellyfish and average of all the current jellyfish.
+When $1- c_t$ is smaller than a randomly generated number in $[0, 1]$, jellyfish will move passively with small random perturbations, otherwise, it will move actively towards or away from another jellyfish based on their index values.
 A new projection is accepted if it is an improvement compared to the
@@
 >
-> > Define trend based on the `current_best` and `current_projections`
+> > Define the ocean current trend as the difference of the `current_best` and the average of the `current_projections`
 >
-> > Update each projection towards the trend using a random factor and
+> > Update each projection towards the ocean trend using a random factor and
 > > orthonormalisation
@@
 The JSO implementation involves several key parameters that control its
-search process in optimization problems. These parameters are designed
-to guide the exploration and exploitation phases of the algorithm. While
-the specific implementation details can vary depending on the version of
+search process in optimization problems. While the specific implementation
+details can vary depending on the version of
 the algorithm or its application, the focus is on two main parameters
@@
 #| label: fig-success-rate
-#| fig.cap: "How success rate is calculated, illustrated using the optimal projections from 50 optimizations of 8D pipe data sorted by index value.
+#| fig.cap: "How success rate is calculated, illustrated using the optimal projections from 50 optimizations of 8D pipe data, optimised by CRS, sorted by index value.
 The pipe shape is recognizable in the projection index values between 0.933-0.969.
@@
 number of iteration) on the success rate is summarized in @fig-proportion.
-The uncertainty is quantified through 500 bootstrap replicates sampled from 
-the simulation for each case. As the number of jellyfish and maximum iteration allowed 
+Bootstrap resampling is used to quantify the uncertainty of the success rate through 500 bootstrap samples with replacement from the original 50 simulation results for each task: optimizing a high-dimensional data (d = 4, 6, 8, 10, or 12) with different number of jellyfish ( 20, 50, or 100) and maximum number of iteration (50 or 100).
+As the number of jellyfish and maximum iteration allowed 
 increase, the success rate also increases. For problems in lower dimensional (4,
@@
 `search_f = search_jellyfish` in the guided tour. 
-
 The animation function `animate_*()` won't directly render the animated 
@@
 visualize the path visited by each individual jellyfish, assign the
-animation to an object (`res <- animate_xy(...)`). This will save the
+animation to an object (`res <- animate_xy(...)`). This will save all the
 bases visited by JSO, along with relevant metadata, as a tibble data
-object [see @RJ-2021-105 for more details on the data object]. The bases
-visited for individual jellyfish can then be extracted and viewed using
-the `planned_tour()` function.
+object [see @RJ-2021-105 for more details on the data object]. Users can 
+then view the path of individual jellyfish through extracting the relevant 
+bases recorded and construct it as a planned tour (using the `planned_tour()` function).
+Users may wish to view the 
+optimisation path of the most successful jellyfish, or a random selection of jellyfish, 
+or some specific runs that find local maximum.
 

### 2025-10-24 jso.tex
more revision

@@
 amenable to efficient optimization. This observation, shown in the
-simulated paths from different index functions in
+simulated trace from different index functions in
 Figure~\ref{fig-squintability-ill}, suggests that the half-point index
@@
 projection is evaluated by the index function. The projection is then
-moved in a direction determined by a random factor, influenced by how
-far along we are in the optimization process. Occasionally, completely
-new directions may be taken like a jellyfish might with ocean currents.
-A new projection is accepted if it is an improvement compared to the
-current one, rejected otherwise. This process continues and iteratively
-improves the projection, until the pre-specified maximum number of
-trials is reached.
-
-\begin{tcolorbox}[enhanced jigsaw, breakable, opacityback=0, toprule=.15mm, toptitle=1mm, titlerule=0mm, left=2mm, leftrule=.75mm, colframe=quarto-callout-note-color-frame, coltitle=black, bottomrule=.15mm, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, colback=white, opacitybacktitle=0.6, bottomtitle=1mm]
+moved in a direction determined by a random time control function,
+\(c_t\), whose value decreases as the number of trials increases
+(subject to randomness) to guide the exploration and exploitation phases
+of the algorithm. When \(c_t \ge 0.5\), new directions will be taken
+like a jellyfish explores with the ocean trend, defined as the
+difference of the current best jellyfish and average of all the current
+jellyfish. When \(1- c_t\) is smaller than a randomly generated number
+in \([0, 1]\), jellyfish will move passively with small random
+perturbations, otherwise, it will move actively towards or away from
+another jellyfish based on their index values. A new projection is
+accepted if it is an improvement compared to the current one, rejected
+otherwise. This process continues and iteratively improves the
+projection, until the pre-specified maximum number of trials is reached.
+
+\begin{tcolorbox}[enhanced jigsaw, titlerule=0mm, opacityback=0, rightrule=.15mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, breakable, colback=white, leftrule=.75mm, opacitybacktitle=0.6, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, arc=.35mm, coltitle=black, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm]
 
@@
 \begin{quote}
-Define trend based on the \texttt{current\_best} and
+Define the ocean current trend as the difference of the
+\texttt{current\_best} and the average of the
 \texttt{current\_projections}
@@
 \begin{quote}
-Update each projection towards the trend using a random factor and
+Update each projection towards the ocean trend using a random factor and
 orthonormalisation
@@
 The JSO implementation involves several key parameters that control its
-search process in optimization problems. These parameters are designed
-to guide the exploration and exploitation phases of the algorithm. While
-the specific implementation details can vary depending on the version of
-the algorithm or its application, the focus is on two main parameters
-that are most relevant to our application: the number of jellyfish and
-the maximum number of iterations
+search process in optimization problems. While the specific
+implementation details can vary depending on the version of the
+algorithm or its application, the focus is on two main parameters that
+are most relevant to our application: the number of jellyfish and the
+maximum number of iterations
 
@@
 illustrated using the optimal projections from 50 optimizations of 8D
-pipe data sorted by index value. The pipe shape is recognizable in the
-projection index values between 0.933-0.969. Of the 50 simulations, 43
-achieved an index value within 0.05 of the best, resulting in a success
-rate of 0.86.}
+pipe data, optimised by CRS, sorted by index value. The pipe shape is
+recognizable in the projection index values between 0.933-0.969. Of the
+50 simulations, 43 achieved an index value within 0.05 of the best,
+resulting in a success rate of 0.86.}
 
@@
 number of iteration) on the success rate is summarized in
-Figure~\ref{fig-proportion}. The uncertainty is quantified through 500
-bootstrap replicates sampled from the simulation for each case. As the
-number of jellyfish and maximum iteration allowed increase, the success
-rate also increases. For problems in lower dimensional (4, 6) search
-spaces, small parameter values (20 jellyfish and 50 iterations) are
-sufficient to achieve a high success rate. Many more jellyfish (100
-jellyfish and 100 iterations) are needed for finding the pipe structure
-in d = 10, and 12. This suggests that even JSO will struggle with high
-dimensions, and require substantially more computational time to have an
-acceptable success rate, especially for difficult PP indexes based on
-scagnostics.
+Figure~\ref{fig-proportion}. Bootstrap resampling is used to quantify
+the uncertainty of the success rate by drawing 500 bootstrap samples
+with replacement from the original 50 simulation results for each task:
+optimizing a high-dimensional data (d = 4, 6, 8, 10, or 12) with
+different number of jellyfish ( 20, 50, or 100) and maximum number of
+iteration (50 or 100). As the number of jellyfish and maximum iteration
+allowed increase, the success rate also increases. For problems in lower
+dimensional (4, 6) search spaces, small parameter values (20 jellyfish
+and 50 iterations) are sufficient to achieve a high success rate. Many
+more jellyfish (100 jellyfish and 100 iterations) are needed for finding
+the pipe structure in d = 10, and 12. This suggests that even JSO will
+struggle with high dimensions, and require substantially more
+computational time to have an acceptable success rate, especially for
+difficult PP indexes based on scagnostics.
 
@@
 squint angle and is the smoothest. The skinny index has low smoothness
-and smallest squint angle. As dimension increases the squint angle gets
-smaller but smoothness doesn't change much.}
+and smallest squint angle. As dimension increases the smoothness becomes
+smaller but squintability doesn't change much.}
 
@@
 To use JSO for PPGT in the \texttt{tourr} package, specify
-\texttt{search\_f\ =\ search\_jellyfish} in the guided tour.
-
-The animation function \texttt{animate\_*()} won't directly render the
+\texttt{search\_f\ =\ search\_jellyfish} in the guided tour. The
+animation function \texttt{animate\_*()} won't directly render the
 animated projection for JSO, as with other existing optimizers since
@@
 visited by each individual jellyfish, assign the animation to an object
-(\texttt{res\ \textless{}-\ animate\_xy(...)}). This will save the bases
-visited by JSO, along with relevant metadata, as a tibble data object
-(see Zhang et al. 2021 for more details on the data object). The bases
-visited for individual jellyfish can then be extracted and viewed using
-the \texttt{planned\_tour()} function.
+(\texttt{res\ \textless{}-\ animate\_xy(...)}). This will save all the
+bases visited by JSO, along with relevant metadata, as a tibble data
+object (see Zhang et al. 2021 for more details on the data object).
+Users can then view the path of individual jellyfish through extracting
+the relevant bases recorded and construct it as a planned tour (using
+the \texttt{planned\_tour()} function). Users may wish to view the
+optimisation path of the most successful jellyfish, or a random
+selection of jellyfish, or some specific runs that find local maximum.
 

### 2025-10-27 revision/response.qmd
more revision

@@
 
-**\textcolor{purple}{The concept of squintability comes from squint angle, which is formally defined in Definition 4.
-We formally define squintability in Definition 6 as the change of index value when the projection distance is reduced by half.
-TODO}**.
+**\textcolor{purple}{Squint angle is the core concept inspires the metric squintability, hence we provide a separate definition of it in Definition 4.
+For all the projection bases in the space, those near the optimal basis correspond to projections with interesting structure we're reesarching.
+While for different indexes, the size of the neighbourhood is different: some indexes may have projections resemble the target even when it is far away, which corresponds to a large squint angle and easy to optimize (since it is easy for the optimizer to sample into the neighbourhood).
+To quantify how close projections need be before structures of interest can be observed (the optimizer starts hill-climbing), we can track how the index value changes as we move from an initial projection basis to the optimal one (projection distance).
+This prompts the definition of projection distance in Definition 5 and squintability in Definition 6.
+We have revise the text to explain this connection more clearly in Section 3.2.}**
 
@@
 
-**\textcolor{purple}{TODO The caption has been fixed.}**
+**\textcolor{purple}{Table 1 is a subset of the PP index estimated and it happens that the three chosen ones all decrease as n goes larger.
+This is not necessary the case in the full table shown in the Appendix.
+The caption for Table 1 has been fixed.}**
 
@@
 
-  * After defining Eq. 4 the additional parameter should probably be theta_4 and not theta_2.  **\textcolor{purple}{our responses are in red.}**
+  * After defining Eq. 4 the additional parameter should probably be theta_4 and not theta_2.  **\textcolor{purple}{TODO: Nicolas}**
   * In the algorithm pseudo-code it is not clear what c_t and the trend actually are.
@@
   * In Fig. 4 please specify what optimizer was used. **\textcolor{purple}{We have included the information in the caption that the CRS optimised is used for this illustration. }**
-  * From the last sentence of Section 5, it is not clear to me why the ranks need to be used, why not scale the metrics?  **\textcolor{purple}{our responses are in red}**.
+  * From the last sentence of Section 5, it is not clear to me why the ranks need to be used, why not scale the metrics?
+**\textcolor{purple}{We include further explanation in the end of Section 5 on the choice of using rank for the modelling.
+The reason is because of the numerical instability due to the small squintability value associated with the string index (less than 0.002).
+Scaling is less ideal here because we define the two metrics to be between [0, 1 ] and scaling will distort the interpretation of the GLM coefficients.}**.
   * In 6.2 it is not really clear what is being bootstrapped for the evaluation of the uncertainty.

### 2025-10-27 jso.qmd
more revision

@@
 
-From @barnett1981interpreting and @laa_using_2020, a large squint angle
-implies that the objective function value is close to optimal even when
-the target view to see the structure is far away. A small squint angle
-means that the PP index value improves substantially only when the
-target view is close by. As such, low squintability implies rapid
-improvement in the index value when near the target view. For PP, a
-small squint angle is considered to be undesirable because it means that
-the optimizer needs to be very close to be able to "see" the optimum, making it difficult for the optimizer to find the optimum.
+The squintability metric is inspired by the concept of squint angle in Definition \ref{def:squint-angle}, originally introduced in @barnett1981interpreting and later discussed in @laa_using_2020.
+Consider the space of all projection bases: those near the optimal basis correspond to projections with interesting structure we're researching.
+The size of this neighborhood depends on the choice of the index function.
+For some indexes, projections resemble the target even when relatively far from the optimum, corresponding to a large squint angle, while for others, projections need to be close to the optimum to reveal structure we're interested, resulting in a small squint angle.
+For PP, a small squint angle is considered to be undesirable because it means that the optimizer needs to sample into a be very narrow region to be able to "see" the optimum, making it difficult for the optimizer to find the optimum.
 
@@
 
-It is expected that for a PP index with high squint angle, the
-optimization \eqref{eq-optimization} should make substantial progress
-early on. Conversely, for a PP index with low squint angle, it might
-take a long while for the optimization to make substantial progress, as
-the candidate projections would need to be very close to the optimal one
-for the structure of the index function to be visible enough to be
-amenable to efficient optimization.
-This observation, shown in the simulated trace from different index functions in @fig-squintability-ill, suggests that the half-point index value improvement percentage provides an appropriate
-mathematical definition of squintability, which matches the intuition
-behind this concept, while being amenable to numerical computation.
+To quantify how close projections need be before structures of interest can be observed (the optimizer starts hill-climbing), we can track how the index value changes as we move from an initial projection basis towards the optimal one, that is, how index value changes as the projection distance becomes smaller.
+
+\begin{defn}[projection distance]\label{def:proj-dist}
+Let $A \in \mathbb{R}^{p \times d}$ be a $d$-dimensional orthonormal matrix, and let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.
+The projection distance between $A$
+and $A^*$, $r(A, A^*)$, is defined by
+$r(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\,\rVert _F$
+where $\lVert . \rVert _F$ denotes the Frobenius norm, given by
+$\lVert M \rVert _F = \sqrt{\sum_{ij} M_{ij}^2}$. 
+\end{defn}
 
@@
 
-\begin{defn}[projection distance]\label{def:proj-dist}
-Let $A \in \mathbb{R}^{p \times d}$ be a $d$-dimensional orthonormal matrix, and let $A^*$ be the optimal matrix that achieves the maximum index value for a given data.
-The projection distance between $A$
-and $A^*$, $r(A, A^*)$, is defined by
-$r(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\,\rVert _F$
-where $\lVert . \rVert _F$ denotes the Frobenius norm, given by
-$\lVert M \rVert _F = \sqrt{\sum_{ij} M_{ij}^2}$. 
-\end{defn}
+@fig-squintability-ill illustrates some simulated traces for different PP index functions: MIC, splines2d, skinny, and stringy2.
+It is expected that for a PP index with high squint angle, the optimization \eqref{eq-optimization} should make substantial progress early on.
+Conversely, for a PP index with low squint angle, it might take a long while for the optimization to make substantial progress, as the candidate projections would need to be very close to the optimal one for the structure of the index function to be visible enough to be amenable to efficient optimization.
+This observation suggests that the half-point index value improvement percentage provides an appropriate
+mathematical definition of squintability, which matches the intuition
+behind this concept, while being amenable to numerical computation.
+
 \begin{defn}[squintability]\label{def:squintability}
@@
 Predictors are smoothness, squintability, data dimension, and JSO
-hyper-parameters. Due to the scale of the smoothness and squintability measure,
-the rank of the two metrics, rather than the nominal values, are used in the model.
+hyper-parameters.
+For numerical stability related to small squintability values associated with the stringy index (squintability less than 0.002), we decide to use the rank of the two metrics, rather than the nominal values, in the model.
 

### 2025-10-27 jso.tex
more revision

@@
 
-From Tukey and Tukey (1981) and Laa and Cook (2020), a large squint
-angle implies that the objective function value is close to optimal even
-when the target view to see the structure is far away. A small squint
-angle means that the PP index value improves substantially only when the
-target view is close by. As such, low squintability implies rapid
-improvement in the index value when near the target view. For PP, a
-small squint angle is considered to be undesirable because it means that
-the optimizer needs to be very close to be able to ``see'' the optimum,
-making it difficult for the optimizer to find the optimum.
+The squintability metric is inspired by the concept of squint angle in
+Definition \ref{def:squint-angle}, originally introduced in Tukey and
+Tukey (1981) and later discussed in Laa and Cook (2020). Consider the
+space of all projection bases: those near the optimal basis correspond
+to projections with interesting structure we're researching. The size of
+this neighborhood depends on the choice of the index function. For some
+indexes, projections resemble the target even when relatively far from
+the optimum, corresponding to a large squint angle, while for others,
+projections need to be close to the optimum to reveal structure we're
+interested, resulting in a small squint angle. For PP, a small squint
+angle is considered to be undesirable because it means that the
+optimizer needs to sample into a be very narrow region to be able to
+``see'' the optimum, making it difficult for the optimizer to find the
+optimum.
 
@@
 
-It is expected that for a PP index with high squint angle, the
-optimization \eqref{eq-optimization} should make substantial progress
-early on. Conversely, for a PP index with low squint angle, it might
-take a long while for the optimization to make substantial progress, as
-the candidate projections would need to be very close to the optimal one
-for the structure of the index function to be visible enough to be
-amenable to efficient optimization. This observation, shown in the
-simulated trace from different index functions in
-Figure~\ref{fig-squintability-ill}, suggests that the half-point index
-value improvement percentage provides an appropriate mathematical
-definition of squintability, which matches the intuition behind this
-concept, while being amenable to numerical computation.
-
-\begin{figure}
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-squintability-ill-1.pdf}}
-
-}
-
-\caption{\label{fig-squintability-ill}Simulated traces of index value as
-a function of projection distances for four PP index functions: MIC,
-splines2d, skinny, and stringy2. The index function MIC and splines2d
-make early progression towards the optimal index value, indicating a
-large squint angle, whereas the traces from skinny, and specially
-stringy2, show improvement only near the optimal projection, suggesting
-low squintability.}
-
-\end{figure}%
+To quantify how close projections need be before structures of interest
+can be observed (the optimizer starts hill-climbing), we can track how
+the index value changes as we move from an initial projection basis
+towards the optimal one, that is, how index value changes as the
+projection distance becomes smaller.
 
@@
 \end{defn}
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{jso_files/figure-pdf/fig-squintability-ill-1.pdf}}
+
+}
+
+\caption{\label{fig-squintability-ill}Simulated traces of index value as
+a function of projection distances for four PP index functions: MIC,
+splines2d, skinny, and stringy2. The index function MIC and splines2d
+make early progression towards the optimal index value, indicating a
+large squint angle, whereas the traces from skinny, and specially
+stringy2, show improvement only near the optimal projection, suggesting
+low squintability.}
+
+\end{figure}%
+
+Figure~\ref{fig-squintability-ill} illustrates some simulated traces for
+different PP index functions: MIC, splines2d, skinny, and stringy2. It
+is expected that for a PP index with high squint angle, the optimization
+\eqref{eq-optimization} should make substantial progress early on.
+Conversely, for a PP index with low squint angle, it might take a long
+while for the optimization to make substantial progress, as the
+candidate projections would need to be very close to the optimal one for
+the structure of the index function to be visible enough to be amenable
+to efficient optimization. This observation suggests that the half-point
+index value improvement percentage provides an appropriate mathematical
+definition of squintability, which matches the intuition behind this
+concept, while being amenable to numerical computation.
+
 \begin{defn}[squintability]\label{def:squintability}
@@
 
-\begin{tcolorbox}[enhanced jigsaw, titlerule=0mm, opacityback=0, rightrule=.15mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, breakable, colback=white, leftrule=.75mm, opacitybacktitle=0.6, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, arc=.35mm, coltitle=black, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm]
+\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, toprule=.15mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, breakable, colback=white, leftrule=.75mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, bottomrule=.15mm, coltitle=black, bottomtitle=1mm, opacitybacktitle=0.6, left=2mm, opacityback=0]
 
@@
 Predictors are smoothness, squintability, data dimension, and JSO
-hyper-parameters. Due to the scale of the smoothness and squintability
-measure, the rank of the two metrics, rather than the nominal values,
-are used in the model.
+hyper-parameters. For numerical stability related to small squintability
+values associated with the stringy index (squintability less than
+0.002), we decide to use the rank of the two metrics, rather than the
+nominal values, in the model.
 
@@
 Figure~\ref{fig-proportion}. Bootstrap resampling is used to quantify
-the uncertainty of the success rate by drawing 500 bootstrap samples
-with replacement from the original 50 simulation results for each task:
+the uncertainty of the success rate through 500 bootstrap samples with
+replacement from the original 50 simulation results for each task:
 optimizing a high-dimensional data (d = 4, 6, 8, 10, or 12) with

### 2025-10-28 revision/response.qmd
small changes

@@
 # Reviewer 1
+
+## Major comments
 
@@
 
-**\textcolor{purple}{our responses are in red}**.
+**\textcolor{purple}{TODO}**.
 
@@
 
-**\textcolor{purple}{our responses are in red}**.
+**\textcolor{purple}{TODO: Nicolas}**.
 
@@
 
-Finally I have also collected a few minor remarks:
+## Minor comments
 
-  * After defining Eq. 4 the additional parameter should probably be theta_4 and not theta_2.  **\textcolor{purple}{TODO: Nicolas}**
-  * In the algorithm pseudo-code it is not clear what c_t and the trend actually are.
+1. After defining Eq. 4 the additional parameter should probably be theta_4 and not theta_2.  **\textcolor{purple}{TODO: Nicolas}**
+
+2.
+In the algorithm pseudo-code it is not clear what c_t and the trend actually are.
 **\textcolor{purple}{In the text and the pseudo-code, we include information that $c_t$ is the time control function determine the exploration and exploitation phase of the algorithm.
 We also include information that the (ocean) trend is defined as the difference between the best jellyfish and the average of all current jellyfish.}**
-  * In Fig. 4 please specify what optimizer was used. **\textcolor{purple}{We have included the information in the caption that the CRS optimised is used for this illustration. }**
-  * From the last sentence of Section 5, it is not clear to me why the ranks need to be used, why not scale the metrics?
+
+3. In Fig. 4 please specify what optimizer was used. **\textcolor{purple}{We have included the information in the caption that the CRS optimised is used for this illustration. }**
+
+4.
+From the last sentence of Section 5, it is not clear to me why the ranks need to be used, why not scale the metrics?
 **\textcolor{purple}{We include further explanation in the end of Section 5 on the choice of using rank for the modelling.
@@
 Scaling is less ideal here because we define the two metrics to be between [0, 1 ] and scaling will distort the interpretation of the GLM coefficients.}**.
-  * In 6.2 it is not really clear what is being bootstrapped for the evaluation of the uncertainty.
+
+5.
+In 6.2 it is not really clear what is being bootstrapped for the evaluation of the uncertainty.
 **\textcolor{purple}{We have included further description in the paper to describe the bootstrap procedure as drawing 500 bootstrap samples with replacement from the original 50 simulations for each setting: an optimisation with high-dimensional data (d = 4, 6, 8, 10, or 12) with different number of jellyfish (20, 50, or 100) and maximum number of iteration (50 or 100)}**.

### 2025-11-07 revision/response.qmd
response

@@
 
-We would like to thank the reviewers and editor for taking the time to review our paper and provide constructive comments.
-Below is a point-by-point explanation of how we have changed the paper.
-Reviewers' comments are in normal text and **\textcolor{purple}{our responses are in red}**.
+We sincerely thank the reviewers and the editor for the time and care devoted to evaluating our manuscript and for providing thoughtful and constructive feedback.
+Below, we provide a point-by-point response detailing the revisions made.
+The reviewers’ comments appear in regular text, and \textcolor{purple}{our responses are in pink}.
 
@@
 
-**\textcolor{purple}{Squint angle is the core concept inspires the metric squintability, hence we provide a separate definition of it in Definition 4.
-For all the projection bases in the space, those near the optimal basis correspond to projections with interesting structure we're reesarching.
-While for different indexes, the size of the neighbourhood is different: some indexes may have projections resemble the target even when it is far away, which corresponds to a large squint angle and easy to optimize (since it is easy for the optimizer to sample into the neighbourhood).
-To quantify how close projections need be before structures of interest can be observed (the optimizer starts hill-climbing), we can track how the index value changes as we move from an initial projection basis to the optimal one (projection distance).
-This prompts the definition of projection distance in Definition 5 and squintability in Definition 6.
-We have revise the text to explain this connection more clearly in Section 3.2.}**
+\textcolor{purple}{We thank the reviewer for this helpful comment, which prompted us to clarify the relationship between the squint angle and the metric squintability.
+Squint angle is the fundamental geometric notion that motivates the metric squintability, and for this reason we introduce it separately in Definition 4.
+Intuitively, across all possible projection bases, those that lie close to the optimal basis yield projections that exhibit the structural features of interest.
+However, different indexes vary in how quickly such structure emerges: for some indexes, projections may resemble the target even when the basis is relatively far from optimal.
+This corresponds to a larger squint angle and a problem that is easier to optimize, as the optimizer can more readily sample points within the relevant neighbourhood.
+To quantify how close a projection must be before the structure of interest becomes detectable (i.e., before the optimizer begins effective hill-climbing), we track how the index value changes as we move from an initial basis toward the optimal one.
+This motivates the definition of projection distance (Definition 5) and squintability (Definition 6).
+We have revised Section 3.2 to articulate these connections more clearly.}
 

### 2025-11-07 revision/response.qmd
response

@@
 
-**\textcolor{purple}{We decide to keep the two figures (Now Figure 6 and 7) in Section 5.2 since some audience may benefit from the illustration.
-To provide some intuition about the index for motivating squintability, we include a new figure, Figure 4, in Section 3.2 to illustrate the trace for different indexes.}**
+\textcolor{purple}{We thank the reviewer for this insightful suggestion.
+In line with the reviewer’s recommendation, we have also strengthened Section 3.2 by adding a new figure (Figure 4) that provides intuition for the squintability metric.
+We have retained the two illustrative figures (now Figures 6 and 7) in Section 5.2, as we believe that some readers may still find the visualisation helpful when interpreting the practical behaviour of the metrics.}
 

### 2025-11-07 jso.qmd
response

@@
 
-To quantify how close projections need be before structures of interest can be observed (the optimizer starts hill-climbing), we can track how the index value changes as we move from an initial projection basis towards the optimal one, that is, how index value changes as the projection distance becomes smaller.
+To quantify how close projections need to be before structures of interest can be observed (the optimizer starts hill-climbing), we can track how the index value changes as we move from an initial projection basis towards the optimal one, that is, how index value changes as the projection distance becomes smaller.
 

### 2025-11-07 revision/response.qmd
response

@@
 
-**\textcolor{purple}{Table 1 is a subset of the PP index estimated and it happens that the three chosen ones all decrease as n goes larger.
+\textcolor{purple}{Table 1 is a subset of the PP index estimated and it happens that the three chosen ones all decrease as n goes larger.
 This is not necessary the case in the full table shown in the Appendix.
-The caption for Table 1 has been fixed.}**
+The caption for Table 1 has been fixed.}
 
@@
 
-**\textcolor{purple}{We make Section 7.1 clearer that the planned tour can be run to trace the optimisation of individual jellyfish path.
+\textcolor{purple}{We make Section 7.1 clearer that the planned tour can be run to trace the optimisation of individual jellyfish path.
 Users may wish to follow the most successful jellyfish, a random selection, or interesting jellyfish that result in local optimum.
-DISCUSSION: should I include a small example with the code?}**
+DISCUSSION: should I include a small example with the code?}
 
@@
 In the algorithm pseudo-code it is not clear what c_t and the trend actually are.
-**\textcolor{purple}{In the text and the pseudo-code, we include information that $c_t$ is the time control function determine the exploration and exploitation phase of the algorithm.
-We also include information that the (ocean) trend is defined as the difference between the best jellyfish and the average of all current jellyfish.}**
+\textcolor{purple}{In the text and the pseudo-code, we include information that $c_t$ is the time control function determine the exploration and exploitation phase of the algorithm.
+We also include information that the (ocean) trend is defined as the difference between the best jellyfish and the average of all current jellyfish.}
 
-3. In Fig. 4 please specify what optimizer was used. **\textcolor{purple}{We have included the information in the caption that the CRS optimised is used for this illustration. }**
+3. In Fig. 4 please specify what optimizer was used. \textcolor{purple}{We have included the information in the caption that the CRS optimised is used for this illustration. }
 
@@
 From the last sentence of Section 5, it is not clear to me why the ranks need to be used, why not scale the metrics?
-**\textcolor{purple}{We include further explanation in the end of Section 5 on the choice of using rank for the modelling.
+\textcolor{purple}{We include further explanation in the end of Section 5 on the choice of using rank for the modelling.
 The reason is because of the numerical instability due to the small squintability value associated with the string index (less than 0.002).
-Scaling is less ideal here because we define the two metrics to be between [0, 1 ] and scaling will distort the interpretation of the GLM coefficients.}**.
+Scaling is less ideal here because we define the two metrics to be between [0, 1 ] and scaling will distort the interpretation of the GLM coefficients.}.
 
@@
 In 6.2 it is not really clear what is being bootstrapped for the evaluation of the uncertainty.
-**\textcolor{purple}{We have included further description in the paper to describe the bootstrap procedure as drawing 500 bootstrap samples with replacement from the original 50 simulations for each setting: an optimisation with high-dimensional data (d = 4, 6, 8, 10, or 12) with different number of jellyfish (20, 50, or 100) and maximum number of iteration (50 or 100)}**.
+\textcolor{purple}{Thank you for the clarification.
+We have included further description in the paper to describe the bootstrap procedure as drawing 500 bootstrap samples with replacement from the original 50 simulations for each setting: an optimisation with high-dimensional data (d = 4, 6, 8, 10, or 12) with different number of jellyfish (20, 50, or 100) and maximum number of iteration (50 or 100)}.

### 2025-11-07 revision/response.qmd
change text color

@@
 Below, we provide a point-by-point response detailing the revisions made.
-The reviewers’ comments appear in regular text, and \textcolor{purple}{our responses are in pink}.
+The reviewers’ comments appear in regular text, and \textcolor{blue}{our responses are in blue}.
 
@@
 
-**\textcolor{purple}{TODO: Di}**.
+**\textcolor{blue}{TODO: Di}**.
 
@@
 
-\textcolor{purple}{We thank the reviewer for this helpful comment, which prompted us to clarify the relationship between the squint angle and the metric squintability.
+\textcolor{blue}{We thank the reviewer for this helpful comment, which prompted us to clarify the relationship between the squint angle and the metric squintability.
 Squint angle is the fundamental geometric notion that motivates the metric squintability, and for this reason we introduce it separately in Definition 4.
@@
 
-**\textcolor{purple}{TODO}**.
+**\textcolor{blue}{TODO}**.
 
@@
 
-**\textcolor{purple}{TODO: Nicolas}**.
+**\textcolor{blue}{TODO: Nicolas}**.
 
@@
 
-\textcolor{purple}{We thank the reviewer for this insightful suggestion.
+\textcolor{blue}{We thank the reviewer for this insightful suggestion.
 In line with the reviewer’s recommendation, we have also strengthened Section 3.2 by adding a new figure (Figure 4) that provides intuition for the squintability metric.
@@
 
-**\textcolor{purple}{The correlation between the two variables are 0.55, which should not affect the modelling.
+**\textcolor{blue}{The correlation between the two variables are 0.55, which should not affect the modelling.
 (TODO: Di - do you mind checking if we need more here?) We did try CRS for the same sine-wave optimisation reported in Section 6, but most runs yield a failed result with the optimum not found by CRS, hence not included.}**
@@
 
-\textcolor{purple}{Table 1 is a subset of the PP index estimated and it happens that the three chosen ones all decrease as n goes larger.
+\textcolor{blue}{Table 1 is a subset of the PP index estimated and it happens that the three chosen ones all decrease as n goes larger.
 This is not necessary the case in the full table shown in the Appendix.
@@
 
-\textcolor{purple}{We make Section 7.1 clearer that the planned tour can be run to trace the optimisation of individual jellyfish path.
+\textcolor{blue}{We make Section 7.1 clearer that the planned tour can be run to trace the optimisation of individual jellyfish path.
 Users may wish to follow the most successful jellyfish, a random selection, or interesting jellyfish that result in local optimum.
@@
 
-1. After defining Eq. 4 the additional parameter should probably be theta_4 and not theta_2.  **\textcolor{purple}{TODO: Nicolas}**
+1. After defining Eq. 4 the additional parameter should probably be theta_4 and not theta_2.  **\textcolor{blue}{TODO: Nicolas}**
 
@@
 In the algorithm pseudo-code it is not clear what c_t and the trend actually are.
-\textcolor{purple}{In the text and the pseudo-code, we include information that $c_t$ is the time control function determine the exploration and exploitation phase of the algorithm.
+\textcolor{blue}{In the text and the pseudo-code, we include information that $c_t$ is the time control function determine the exploration and exploitation phase of the algorithm.
 We also include information that the (ocean) trend is defined as the difference between the best jellyfish and the average of all current jellyfish.}
 
-3. In Fig. 4 please specify what optimizer was used. \textcolor{purple}{We have included the information in the caption that the CRS optimised is used for this illustration. }
+3. In Fig. 4 please specify what optimizer was used. \textcolor{blue}{We have included the information in the caption that the CRS optimised is used for this illustration. }
 
@@
 From the last sentence of Section 5, it is not clear to me why the ranks need to be used, why not scale the metrics?
-\textcolor{purple}{We include further explanation in the end of Section 5 on the choice of using rank for the modelling.
+\textcolor{blue}{We include further explanation in the end of Section 5 on the choice of using rank for the modelling.
 The reason is because of the numerical instability due to the small squintability value associated with the string index (less than 0.002).
@@
 In 6.2 it is not really clear what is being bootstrapped for the evaluation of the uncertainty.
-\textcolor{purple}{Thank you for the clarification.
+\textcolor{blue}{Thank you for the clarification.
 We have included further description in the paper to describe the bootstrap procedure as drawing 500 bootstrap samples with replacement from the original 50 simulations for each setting: an optimisation with high-dimensional data (d = 4, 6, 8, 10, or 12) with different number of jellyfish (20, 50, or 100) and maximum number of iteration (50 or 100)}.

### 2025-11-07 revision/response.qmd
minor polish

@@
 
-\textcolor{blue}{Table 1 is a subset of the PP index estimated and it happens that the three chosen ones all decrease as n goes larger.
+\textcolor{blue}{Table 1 is a subset of the PP index estimated and it happens that the three chosen ones all decrease as $n$ goes larger.
 This is not necessary the case in the full table shown in the Appendix.
@@
 \textcolor{blue}{Thank you for the clarification.
-We have included further description in the paper to describe the bootstrap procedure as drawing 500 bootstrap samples with replacement from the original 50 simulations for each setting: an optimisation with high-dimensional data (d = 4, 6, 8, 10, or 12) with different number of jellyfish (20, 50, or 100) and maximum number of iteration (50 or 100)}.
+We have included further description in the paper to describe the bootstrap procedure as drawing $500$ bootstrap samples with replacement from the original $50$ simulations for each setting: an optimisation with high-dimensional data ($d = 4, 6, 8, 10$, or $12$) with different number of jellyfish ($20, 50, or 100$) and maximum number of iteration ($50$ or $100$)}.

### 2025-11-07 revision/response.qmd
polish
dollar signs

@@
 
-1. After defining Eq. 4 the additional parameter should probably be theta_4 and not theta_2.  **\textcolor{blue}{TODO: Nicolas}**
+1. After defining Eq. 4 the additional parameter should probably be theta_4 and not theta_2.
 
-2.
-In the algorithm pseudo-code it is not clear what c_t and the trend actually are.
+**\textcolor{blue}{TODO: Nicolas}**
+
+2. In the algorithm pseudo-code it is not clear what c_t and the trend actually are.
+
 \textcolor{blue}{In the text and the pseudo-code, we include information that $c_t$ is the time control function determine the exploration and exploitation phase of the algorithm.
@@
 
-3. In Fig. 4 please specify what optimizer was used. \textcolor{blue}{We have included the information in the caption that the CRS optimised is used for this illustration. }
+3. In Fig. 4 please specify what optimizer was used.
 
-4.
-From the last sentence of Section 5, it is not clear to me why the ranks need to be used, why not scale the metrics?
+\textcolor{blue}{We have included the information in the caption that the CRS optimised is used for this illustration. }
+
+4. From the last sentence of Section 5, it is not clear to me why the ranks need to be used, why not scale the metrics?
+
 \textcolor{blue}{We include further explanation in the end of Section 5 on the choice of using rank for the modelling.
@@
 
-5.
-In 6.2 it is not really clear what is being bootstrapped for the evaluation of the uncertainty.
+5. In 6.2 it is not really clear what is being bootstrapped for the evaluation of the uncertainty.
+
 \textcolor{blue}{Thank you for the clarification.
-We have included further description in the paper to describe the bootstrap procedure as drawing $500$ bootstrap samples with replacement from the original $50$ simulations for each setting: an optimisation with high-dimensional data ($d = 4, 6, 8, 10$, or $12$) with different number of jellyfish ($20, 50, or 100$) and maximum number of iteration ($50$ or $100$)}.
+We have included further description in the paper to describe the bootstrap procedure as drawing $500$ bootstrap samples with replacement from the original $50$ simulations for each setting: an optimisation with high-dimensional data ($d = 4, 6, 8, 10$, or $12$) with different number of jellyfish ($20, 50$, or $100$) and maximum number of iteration ($50$ or $100$)}.

### 2025-11-17 jso.qmd
addressing title comment, and flow

@@
 library(broom)
+library(broom.helpers)
 library(kableExtra)
@@
 
-Optimization of PP is often discussed when new indexes are proposed
+Optimization of PP is sometimes discussed when new indexes are proposed
 [@posse95; @marie-sainte2010; @grochowski2011]. @cook1995grand tied the
-optimization more closely to the index, when they introduced the PPGT,
+optimization closely to the index with the introduction of the PPGT,
 which monitors the optimization visually so that the user can see the

### 2025-11-17 jso.tex
addressing title comment, and flow

@@
 
-Optimization of PP is often discussed when new indexes are proposed
+Optimization of PP is sometimes discussed when new indexes are proposed
 (Posse 1995; Marie-Sainte, Berro, and Ruiz-Gazen 2010; Grochowski and
-Duch 2011). Cook et al. (1995) tied the optimization more closely to the
-index, when they introduced the PPGT, which monitors the optimization
+Duch 2011). Cook et al. (1995) tied the optimization closely to the
+index with the introduction of the PPGT, which monitors the optimization
 visually so that the user can see the projected data leading in and out
@@
 
-To quantify how close projections need be before structures of interest
-can be observed (the optimizer starts hill-climbing), we can track how
-the index value changes as we move from an initial projection basis
-towards the optimal one, that is, how index value changes as the
+To quantify how close projections need to be before structures of
+interest can be observed (the optimizer starts hill-climbing), we can
+track how the index value changes as we move from an initial projection
+basis towards the optimal one, that is, how index value changes as the
 projection distance becomes smaller.
@@
 
-\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, toprule=.15mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, breakable, colback=white, leftrule=.75mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, bottomrule=.15mm, coltitle=black, bottomtitle=1mm, opacitybacktitle=0.6, left=2mm, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, toprule=.15mm, arc=.35mm, toptitle=1mm, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, left=2mm, breakable, colback=white, titlerule=0mm, rightrule=.15mm, bottomrule=.15mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white]
 

### 2025-11-17 revision/response.qmd
addressing title comment, and flow
mostly addressed the correlation between squintability and smoothness comment

@@
 
-**\textcolor{blue}{TODO: Di}**.
+\textcolor{blue}{Point taken. The title has been changed to better reflect the content. Attention has been paid to the flow of the paper in the revision.}
 
@@
 
-**\textcolor{blue}{The correlation between the two variables are 0.55, which should not affect the modelling.
-(TODO: Di - do you mind checking if we need more here?) We did try CRS for the same sine-wave optimisation reported in Section 6, but most runs yield a failed result with the optimum not found by CRS, hence not included.}**
+**\textcolor{blue}{We have revised the text to include a comment on the correlation and why CRS cannot be run for the sine-wave optimization.
+(TODO: Di - do you mind checking if we need more here?)}**
 

### 2025-11-17 jso.qmd
mostly addressed the correlation between squintability and smoothness comment

@@
 that are most relevant to our application: the number of jellyfish and
-the maximum number of iterations
+the maximum number of iterations.
 
@@
 Of the 50 simulations, 43 achieved an index value within 0.05 of the best, resulting in a success rate of 0.86."
+#| fig-pos: "!ht"
 # pipe_jellyfish_12d <- pipe_better |> 
@@
 #| fig.align: "center"
+#| fig-pos: "!ht"
 #| fig.cap: "Visual comparison of JSO and CRS results, using optimal data projections obtained over 50 simulations of the pipe data.
@@
 
-## Effect of index properties on JSO success rate
-
-```{r}
-#| tbl-pos: H
+```{r}
+#| tbl-pos: "!ht"
 #| label: tbl-mod-output
@@
 ```{r}
-#| tbl-pos: H
+#| tbl-pos: "!ht"
 #| label: tbl-joint-output
@@
 
+## Effect of index properties on JSO success rate
+
 @tbl-smoothness-squintability presents values computed for a subset of
@@
 and smoothness work to describe the PP indexes.
+
 @tbl-mod-output presents the results from fitting a logistic regression
@@
 
-Note that the addition of interaction terms for squintability and data
-dimension, and smoothness and data dimension, improves the model a small
-amount, but does not change the interpretation above. This study intends 
-to serve as an illustration that the proposed
-measures effectively captures the key factors influencing the success
+In relation to the model fit, it should be noted that there is moderate correlation between squintability and smoothness, but this does not substantially affect the fit.
+Fitting smoothness alone results in this variable being weakly significant for explaining success rate, but the effect is dominated by squintability.
+The addition of interaction terms for squintability and data
+dimension, and smoothness and data dimension, improves the model slightly, but not enough to be practically important or change the interpretation above.
+
+This results support the use of the proposed
+measures to effectively capture key factors affecting the success
 rate of PP.

### 2025-11-17 jso.tex
mostly addressed the correlation between squintability and smoothness comment

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, opacityback=0, colframe=quarto-callout-note-color-frame, leftrule=.75mm, toprule=.15mm, arc=.35mm, toptitle=1mm, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, left=2mm, breakable, colback=white, titlerule=0mm, rightrule=.15mm, bottomrule=.15mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white]
+\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, breakable, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, left=2mm, colback=white, bottomtitle=1mm, opacityback=0, toprule=.15mm, toptitle=1mm, titlerule=0mm, rightrule=.15mm, leftrule=.75mm, opacitybacktitle=0.6]
 
@@
 are most relevant to our application: the number of jellyfish and the
-maximum number of iterations
+maximum number of iterations.
 
@@
 
-\begin{figure}
+\begin{figure}[!ht]
 
@@
 
-\subsection{Effect of index properties on JSO success
-rate}\label{effect-of-index-properties-on-jso-success-rate}
-
-\begin{table}[H]
+\begin{table}[!ht]
 
@@
 
+\subsection{Effect of index properties on JSO success
+rate}\label{effect-of-index-properties-on-jso-success-rate}
+
 Table~\ref{tbl-smoothness-squintability} presents values computed for a
@@
 the new measures for squintability and smoothness work to describe the
-PP indexes. Table~\ref{tbl-mod-output} presents the results from fitting
-a logistic regression model using the proportion of success as the
-response and smoothness, squintability, dimension, number of jellyfish
-and maximum number of iteration, as predictors. The fit suggests that
-the JSO success rate is only affected by squintability, dimension and
-number of jellyfish. As expected, the JSO success rate is higher with
-higher squintability and/or more jellyfish, and lower when dimension is
-higher (higher-dimensional optimization is more difficult).
-Interestingly, the JSO is unaffected by the smoothness of the index
-function. This is consistent with the way random search algorithms jump
-from value to value without taking local regularity into account, as
-opposed to gradient-based optimizers for example. Allowing for more
-iterations also does not affect success rate significantly. A unit
-increase in squintability increases the success rate by 23\%. As
-dimension increases by one the success rate almost halves. Increasing
-the number of jellies by 10 increases the success rate by 32\%.
-Table~\ref{tbl-joint-output} shows the fit statistics supporting that
-the model is reasonably strong.
-
-Note that the addition of interaction terms for squintability and data
-dimension, and smoothness and data dimension, improves the model a small
-amount, but does not change the interpretation above. This study intends
-to serve as an illustration that the proposed measures effectively
-captures the key factors influencing the success rate of PP.
+PP indexes.
+
+Table~\ref{tbl-mod-output} presents the results from fitting a logistic
+regression model using the proportion of success as the response and
+smoothness, squintability, dimension, number of jellyfish and maximum
+number of iteration, as predictors. The fit suggests that the JSO
+success rate is only affected by squintability, dimension and number of
+jellyfish. As expected, the JSO success rate is higher with higher
+squintability and/or more jellyfish, and lower when dimension is higher
+(higher-dimensional optimization is more difficult). Interestingly, the
+JSO is unaffected by the smoothness of the index function. This is
+consistent with the way random search algorithms jump from value to
+value without taking local regularity into account, as opposed to
+gradient-based optimizers for example. Allowing for more iterations also
+does not affect success rate significantly. A unit increase in
+squintability increases the success rate by 23\%. As dimension increases
+by one the success rate almost halves. Increasing the number of jellies
+by 10 increases the success rate by 32\%. Table~\ref{tbl-joint-output}
+shows the fit statistics supporting that the model is reasonably strong.
+
+In relation to the model fit, it should be noted that there is moderate
+correlation between squintability and smoothness, but this does not
+substantially affect the fit. Fitting smoothness alone results in this
+variable being weakly significant for explaining success rate, but the
+effect is dominated by squintability. The addition of interaction terms
+for squintability and data dimension, and smoothness and data dimension,
+improves the model slightly, but not enough to be practically important
+or change the interpretation above.
+
+This results support the use of the proposed measures to effectively
+capture key factors affecting the success rate of PP.
 

### 2025-11-18 revision/response.qmd
finished revising the paper and response about correlation and CRS for the sine-wave

@@
 
-**\textcolor{blue}{We have revised the text to include a comment on the correlation and why CRS cannot be run for the sine-wave optimization.
-(TODO: Di - do you mind checking if we need more here?)}**
+\textcolor{blue}{We have revised the text to include a comment on the correlation.
+It is moderate and does not practically affect the fit.
+We have also noted that it is not feasible to run CRS for the sine-wave optimization because it is not an efficient optimizer and fails to converge too often.
+On small dimensionality, it can be seen that it is not affected by lack of smoothness of an index.
+This would be expected because it is not a derivative-based optimizer, and, similarly to JSO small local fluctuations won't affect it's search.}
 

### 2025-11-18 jso.qmd
finished revising the paper and response about correlation and CRS for the sine-wave

@@
 PP indexes on the sine wave data for dimensions 4-8. (A full table of
-the parameters estimated from the Gaussian process (outputscale $\eta$,
-lengthscale $\ell$, smoothness $\nu$, and nugget $\sigma$) and from the
-scaled logistic function ($\theta_1$ to $\theta_4$) for calculating
+the parameters estimated from the Gaussian process <!-- (outputscale $\eta$,
+lengthscale $\ell$, smoothness $\nu$, and nugget $\sigma$)--> and from the
+scaled logistic function <!--($\theta_1$ to $\theta_4$)--> for calculating
 smoothness and squintability for all PP indexes and data combinations is
@@
 dimension, and smoothness and data dimension, improves the model slightly, but not enough to be practically important or change the interpretation above.
-
-This results support the use of the proposed
+The CRS optimizer couldn't be used here because it is not efficient (@fig-proj) but we would expect that smoothness does not overly affect it either.
+Both JSO and CRS are expected to be robust to small local fluctuations.
+
+These results also support the use of the proposed
 measures to effectively capture key factors affecting the success

### 2025-11-18 jso.tex
finished revising the paper and response about correlation and CRS for the sine-wave

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, breakable, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, left=2mm, colback=white, bottomtitle=1mm, opacityback=0, toprule=.15mm, toptitle=1mm, titlerule=0mm, rightrule=.15mm, leftrule=.75mm, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, bottomtitle=1mm, leftrule=.75mm, titlerule=0mm, toprule=.15mm, left=2mm, colback=white, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, arc=.35mm, coltitle=black, toptitle=1mm, rightrule=.15mm, opacitybacktitle=0.6]
 
@@
 subset of PP indexes on the sine wave data for dimensions 4-8. (A full
-table of the parameters estimated from the Gaussian process (outputscale
-\(\eta\), lengthscale \(\ell\), smoothness \(\nu\), and nugget
-\(\sigma\)) and from the scaled logistic function (\(\theta_1\) to
-\(\theta_4\)) for calculating smoothness and squintability for all PP
-indexes and data combinations is included in the supplementary
-materials). The column \(\nu\) is used as the smoothness metric and the
-column \(\varsigma\) is calculated as equation
+table of the parameters estimated from the Gaussian process and from the
+scaled logistic function for calculating smoothness and squintability
+for all PP indexes and data combinations is included in the
+supplementary materials). The column \(\nu\) is used as the smoothness
+metric and the column \(\varsigma\) is calculated as equation
 (\ref{eq-squintability-parametric}) as the squintability metric.
@@
 improves the model slightly, but not enough to be practically important
-or change the interpretation above.
-
-This results support the use of the proposed measures to effectively
-capture key factors affecting the success rate of PP.
+or change the interpretation above. The CRS optimizer couldn't be used
+here because it is not efficient (Figure~\ref{fig-proj}) but we would
+expect that smoothness does not overly affect it either. Both JSO and
+CRS are expected to be robust to small local fluctuations.
+
+These results also support the use of the proposed measures to
+effectively capture key factors affecting the success rate of PP.
 

### 2025-11-20 revision/response.qmd
incorporate Nicolas's comments

@@
 author: "H. Sherry Zhang, Dianne Cook, Nicolas  Langrené, Jessica Wai Yin Leung"
-date: "2025-10-06"
+date: "2025-11-20"
 format: pdf
@@
 
-**\textcolor{blue}{TODO}**.
+\textcolor{blue}{In Definition 6, we assume g as a decreasing function and this assumption comes from the empirical evidence across all our examples.
+It is true that, by definition, our function g does not depend on directions, only on the distance to optimal.
+We cannot really avoid this averaging across directions in our definition, since eventually our goal is to summarize the whole optimization information into one single metrics, squintability.
+Some information is necessarily lost along the way, but the resulting metric still captures enough information to be able to predict the success rate of the optimization, as shown by the numerical results from Section 6.
+}
 
@@
 
-**\textcolor{blue}{TODO: Nicolas}**.
+\textcolor{blue}{We have now added included the definition of r\_0 right after Eq 5 to make it clear that r\_0 refers to the projection distance from the random start projection to the optimal.}
 
@@
 
-**\textcolor{blue}{TODO: Nicolas}**
+\textcolor{blue}{We thank the reviewer for pointing this out. We have corrected this in the revised manuscript.}
 

### 2025-11-20 jso.qmd
incorporate Nicolas's comments

@@
 flea2D <- as.data.frame(flea2D)
-flea_huber <- prep_huber(flea2D, index = skewness())
+
 sqa1 <- ggplot() + 
-  geom_huber(data = flea_huber$idx_df, aes(x = x, y = y)) +
-  geom_point(data = flea2D, aes(x = X1, y = X2)) +
-  geom_abline(slope = flea_huber$slope, intercept = 0) + 
+  geom_huber(data = flea2D, aes(x = X1, y = X2), 
+             index.fun = skewness()) + 
+  theme_huber() + 
   coord_fixed() + 
-  theme_huber() +
   ggtitle("(a) Skewness index")
 
-sqa2 <- ggplot(flea_huber$proj_df, aes(x=x)) + 
+flea_huber <- prep_huber_best_proj(flea2D, index_fun = skewness())
+sqa2 <- ggplot(flea_huber, aes(x=x)) + 
   geom_histogram(binwidth=0.25, colour="white") +
@@
 randu_df <- randu_std[c(1,4)]
-randu_huber <- prep_huber(randu_df, index = norm_bin(nr = nrow(randu_df)))
 
 sqa3 <- ggplot() + 
-  geom_huber(data = randu_huber$idx_df, aes(x = x, y = y)) +
-  geom_point(data = randu_df, aes(x = x, y = yz)) +
-  geom_abline(slope = randu_huber$slope, intercept = 0) +
+  geom_huber(data = randu_df, aes(x = x, y = yz), 
+             index.fun = norm_bin(nr = nrow(randu_df)))  + 
   theme_huber() + 
@@
 
-sqa4 <- ggplot(randu_huber$proj_df, aes(x = x)) +
+randu_huber <- prep_huber_best_proj(
+  randu_df, index_fun = norm_bin(nr = nrow(randu_df)))
+sqa4 <- ggplot(randu_huber, aes(x = x)) +
   geom_histogram(breaks = seq(-2.2, 2.4, 0.12)) +
@@
 where the nugget parameter $\sigma$ is the standard deviation of the intrinsic noise of the Gaussian process.
-The optimal parameters (including smoothness) are obtained by maximum log-likelihood
+The optimal parameters are obtained by maximum log-likelihood
 \begin{equation}
@@
 Conversely, for a PP index with low squint angle, it might take a long while for the optimization to make substantial progress, as the candidate projections would need to be very close to the optimal one for the structure of the index function to be visible enough to be amenable to efficient optimization.
-This observation suggests that the half-point index value improvement percentage provides an appropriate
+These empirical observations suggest that the half-point index value improvement percentage provides an appropriate
 mathematical definition of squintability, which matches the intuition
@@
 
-Remark that the amount by which the index value can improve from the
+where $r_0 = r(A_0, A^*)$ is the projection distance from the initial projection $A_0$ to the optimal one $A^*$. Remark that the amount by which the index value can improve from the
 starting matrix $A_0$ to the optimal one $A^*$ is given by
@@
 
-which depends on three additional parameters, $\theta_1$, $\theta_2$,
+which depends on three additional parameters, $\theta_1$, $\theta_4$,
 and $r_0$, such that $g(0)=\theta_1$ and $g(r_0)=\theta_4$. Under the

### 2025-11-20 jso.tex
incorporate Nicolas's comments

@@
 the structure of the index function to be visible enough to be amenable
-to efficient optimization. This observation suggests that the half-point
-index value improvement percentage provides an appropriate mathematical
-definition of squintability, which matches the intuition behind this
-concept, while being amenable to numerical computation.
+to efficient optimization. These empirical observations suggest that the
+half-point index value improvement percentage provides an appropriate
+mathematical definition of squintability, which matches the intuition
+behind this concept, while being amenable to numerical computation.
 
@@
 
-Remark that the amount by which the index value can improve from the
-starting matrix \(A_0\) to the optimal one \(A^*\) is given by
-\(g(0)-g(r_{0})\). As a result, equation~\eqref{eq-squintability}
-represents the proportion of this maximum improvement which has been
-achieved by the time the distance \(r_0\) to the optimal matrix has been
-reduced by half (\(r_0/2\)).
+where \(r_0 = r(A_0, A^*)\) is the projection distance from the initial
+projection \(A_0\) to the optimal one \(A^*\). Remark that the amount by
+which the index value can improve from the starting matrix \(A_0\) to
+the optimal one \(A^*\) is given by \(g(0)-g(r_{0})\). As a result,
+equation~\eqref{eq-squintability} represents the proportion of this
+maximum improvement which has been achieved by the time the distance
+\(r_0\) to the optimal matrix has been reduced by half (\(r_0/2\)).
 
@@
 which depends on three additional parameters, \(\theta_1\),
-\(\theta_2\), and \(r_0\), such that \(g(0)=\theta_1\) and
+\(\theta_4\), and \(r_0\), such that \(g(0)=\theta_1\) and
 \(g(r_0)=\theta_4\). Under the parametric model (\ref{eq-parametric}),
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, bottomtitle=1mm, leftrule=.75mm, titlerule=0mm, toprule=.15mm, left=2mm, colback=white, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, arc=.35mm, coltitle=black, toptitle=1mm, rightrule=.15mm, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, bottomtitle=1mm, colback=white, toprule=.15mm, left=2mm, opacityback=0, colframe=quarto-callout-note-color-frame, arc=.35mm, bottomrule=.15mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, breakable, coltitle=black, toptitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacitybacktitle=0.6]
 

### 2025-11-20 jso.qmd
change color palette

@@
   geom_line() + 
-  scale_color_brewer(palette = "Dark2") +
+  scale_color_manual(values = palette.colors()[1:4]) + 
   xlab("Projection distance") + ylab("Index value") +
@@
   geom_segment(aes(x = n_jellies - 1, xend = n_jellies + 1), 
-               data = res_quantile |> summarise(proportion = mean(proportion)), 
+               data = res_quantile |> 
+                 summarise(proportion = mean(proportion)), 
                size = 2) + 
   geom_line(aes(x = n_jellies), 
-               data = res_quantile |> summarise(proportion = mean(proportion)), 
+            data = res_quantile |> 
+              summarise(proportion = mean(proportion)), 
             alpha = 0.3) + 
-  scale_color_brewer(palette = "Dark2", name = "dimension") +
+  scale_color_manual(values = palette.colors()[1:5], name = "Dimension") + 
   scale_x_continuous(breaks = c(20, 50, 100), labels = c("20", "50", "100")) +
@@
   geom_line() + 
-  scale_color_brewer(palette = "Dark2", name = "Index function") +
+  scale_color_manual(values = palette.colors()[1:3],
+                     name = "Index function") + 
+  #scale_color_brewer(palette = "Dark2", name = "Index function") +
   xlab("Projection distance") + ylab("Index value") +
@@
 The R packages used in this work include: `tidyr` [@tidyr], `dplyr`
-[@dplyr], `ggplot2` [@ggplot2], `knitr` [@knitr],`gtsummary`
+[@dplyr], `ggplot2` [@ggplot2], `knitr` [@knitr], `gtsummary`
 [@gtsummary], `patchwork` [@patchwork], `ggh4x` [@ggh4x], `broom`

### 2025-11-20 jso.tex
change color palette

@@
 
-\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, bottomtitle=1mm, colback=white, toprule=.15mm, left=2mm, opacityback=0, colframe=quarto-callout-note-color-frame, arc=.35mm, bottomrule=.15mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, breakable, coltitle=black, toptitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, coltitle=black, titlerule=0mm, opacitybacktitle=0.6, left=2mm, toprule=.15mm, colback=white, leftrule=.75mm, arc=.35mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, rightrule=.15mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0]
 
@@
 Vaughan, and Girlich 2024), \texttt{dplyr} (Wickham et al. 2023),
-\texttt{ggplot2} (Wickham 2016), \texttt{knitr} (Xie
-2014),\texttt{gtsummary} (Sjoberg et al. 2021), \texttt{patchwork}
-(Pedersen 2024), \texttt{ggh4x} (Van Den Brand 2024), \texttt{broom}
-(Robinson, Hayes, and Couch 2024), \texttt{kableExtra} (Zhu 2024),
-\texttt{ferrn} (Zhang et al. 2021), and \texttt{cassowaryr} (Mason et
-al. 2022).
+\texttt{ggplot2} (Wickham 2016), \texttt{knitr} (Xie 2014),
+\texttt{gtsummary} (Sjoberg et al. 2021), \texttt{patchwork} (Pedersen
+2024), \texttt{ggh4x} (Van Den Brand 2024), \texttt{broom} (Robinson,
+Hayes, and Couch 2024), \texttt{kableExtra} (Zhu 2024), \texttt{ferrn}
+(Zhang et al. 2021), and \texttt{cassowaryr} (Mason et al. 2022).
 

### 2025-11-20 revision/response.qmd
remove discussion point

@@
 \textcolor{blue}{We make Section 7.1 clearer that the planned tour can be run to trace the optimisation of individual jellyfish path.
-Users may wish to follow the most successful jellyfish, a random selection, or interesting jellyfish that result in local optimum.
-DISCUSSION: should I include a small example with the code?}
+Users may wish to follow the most successful jellyfish, a random selection, or interesting jellyfish that result in local optimum.}
 

### 2025-11-20 jso.qmd
link to examples in search_jellyfish

@@
 then view the path of individual jellyfish through extracting the relevant 
-bases recorded and construct it as a planned tour (using the `planned_tour()` function).
+bases recorded and construct it as a planned tour using the `planned_tour()` function (see the examples via `?search_jellyfish`).
 Users may wish to view the 

### 2025-11-20 jso.tex
link to examples in search_jellyfish

@@
 
-\begin{tcolorbox}[enhanced jigsaw, coltitle=black, titlerule=0mm, opacitybacktitle=0.6, left=2mm, toprule=.15mm, colback=white, leftrule=.75mm, arc=.35mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, rightrule=.15mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, left=2mm, breakable, coltitle=black, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, colback=white, leftrule=.75mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, arc=.35mm, toptitle=1mm, titlerule=0mm, rightrule=.15mm, toprule=.15mm, opacitybacktitle=0.6]
 
@@
 Users can then view the path of individual jellyfish through extracting
-the relevant bases recorded and construct it as a planned tour (using
-the \texttt{planned\_tour()} function). Users may wish to view the
-optimisation path of the most successful jellyfish, or a random
-selection of jellyfish, or some specific runs that find local maximum.
+the relevant bases recorded and construct it as a planned tour using the
+\texttt{planned\_tour()} function (see the examples via
+\texttt{?search\_jellyfish}). Users may wish to view the optimisation
+path of the most successful jellyfish, or a random selection of
+jellyfish, or some specific runs that find local maximum.
 

### 2025-11-21 jso.qmd
update the color palette

@@
   group_by(id, index, n) |> 
-  mutate(value = (value - min(value)) / (max(value) - min(value))) |>
+  mutate(value = (value - min(value)) / (max(value) - min(value)),
+         index = fct_relevel(index, c("stringy2", "skinny", "MIC", "splines2d"))) |>
   ggplot(aes(x = dist, y = value, 
@@
   geom_line() + 
-  scale_color_manual(values = palette.colors()[1:4]) + 
+  colorspace::scale_color_discrete_sequential(palette = "RdPu", name = "Index function", 
+                                              c2 = 100, rev = FALSE) + 
   xlab("Projection distance") + ylab("Index value") +
@@
   ggplot(aes(x = n_jellies - 0.5, y = proportion, 
-             group = d, color = d)) + 
-  geom_segment(aes(xend = n_jellies + 0.5), size = 0.05) + 
-  geom_segment(aes(x = n_jellies - 1, xend = n_jellies + 1), 
-               data = res_quantile |> 
-                 summarise(proportion = mean(proportion)), 
+             group = d)) + 
+  geom_segment(aes(x = n_jellies - 1, xend = n_jellies + 1, color = d), 
+               data = res_quantile |> summarise(proportion = mean(proportion)), 
                size = 2) + 
-  geom_line(aes(x = n_jellies), 
-            data = res_quantile |> 
-              summarise(proportion = mean(proportion)), 
-            alpha = 0.3) + 
-  scale_color_manual(values = palette.colors()[1:5], name = "Dimension") + 
+  geom_segment(aes(xend = n_jellies + 0.5, color = d), size = 0.05) + 
+  geom_line(aes(x = n_jellies, color = d),  
+            data = res_quantile |> summarise(proportion = mean(proportion)), 
+            alpha = 0.5, show.legend = FALSE) +
+  scale_color_manual(values = colorspace::sequential_hcl(5, "RdPu", c2 = 200),
+                     name = "Dimension", labels = seq(4, 12, 2)) + 
+  # scale_linetype_manual(values = c(1, 5,2,4,3), guide = "legend", 
+  #                       name = "Dimension", labels = seq(4, 12, 2)) + 
   scale_x_continuous(breaks = c(20, 50, 100), labels = c("20", "50", "100")) +
-  facet_wrap(vars(`max # of iteration`), 
-             labeller = "label_both") + 
+  facet_wrap(vars(`max # of iteration`), labeller = "label_both") + 
   theme_bw() +
@@
 sq_basis_df_6d_TIC_spline_skinny |>
-  ggplot(aes(x = dist, y = index, group = interaction(index_name, id), color = index_name)) +
+  ggplot(aes(x = dist, y = index, group = interaction(index_name, id), 
+             color = fct_relevel(index_name, c("skinny", "TIC", "splines")))) +
   geom_line() + 
-  scale_color_manual(values = palette.colors()[1:3],
-                     name = "Index function") + 
-  #scale_color_brewer(palette = "Dark2", name = "Index function") +
+  colorspace::scale_color_discrete_sequential(palette = "RdPu", name = "Index function", 
+                                              c2 = 100, rev = FALSE) + 
   xlab("Projection distance") + ylab("Index value") +

### 2025-11-21 jso.tex
update the color palette

@@
 
-\begin{tcolorbox}[enhanced jigsaw, left=2mm, breakable, coltitle=black, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, colback=white, leftrule=.75mm, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, arc=.35mm, toptitle=1mm, titlerule=0mm, rightrule=.15mm, toprule=.15mm, opacitybacktitle=0.6]
+\begin{tcolorbox}[enhanced jigsaw, breakable, bottomtitle=1mm, opacitybacktitle=0.6, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, toprule=.15mm, left=2mm, bottomrule=.15mm, rightrule=.15mm, titlerule=0mm, coltitle=black, leftrule=.75mm, colframe=quarto-callout-note-color-frame, colback=white]
 

### 2025-11-21 code/11-sim-matern.py

@@ -21,7 +21,10 @@ nu_array = np.array([1.0,2.0,4.0])
 n_samples = 5
 x = np.linspace(0, 6, 2001)
 X = x.reshape(-1, 1)
-sns.set_palette("cividis")
+# RdPu but I change the end chroma to 200
+palette = ["#490062", "#AB1998", "#FF49B1", "#FF8A88", "#FFCB00"]
+sns.set_palette(palette)
+
 
 fig, axs = plt.subplots(ncols=np.size(nu_array), sharex=False, sharey=False, figsize=(12, 4))
 for (i,nu) in enumerate(nu_array):
@@ -30,7 +33,7 @@ for (i,nu) in enumerate(nu_array):
     gpr = GaussianProcessRegressor(kernel=kernel, random_state=100)
     y_samples = gpr.sample_y(X, n_samples)
     for n in np.arange(n_samples):
-        sns.lineplot(x=x,y=y_samples[:,n],ax=axs[i],linewidth=2,alpha=0.9)
+        sns.lineplot(x=x,y=y_samples[:,n],ax=axs[i],linewidth=1,alpha=0.9)
     axs[i].set_xlim([0,6])
     axs[i].set_ylim([-3,3])
     axs[i].set_xticks([0,1,2,3,4,5,6])
@@ -65,7 +68,7 @@ for (i,nu) in enumerate(nu_array):
     for n in np.arange(nx):
         zv[n,:] = np.squeeze(z_samples[n*nx:(n+1)*nx])
     
-    surf = ax.plot_surface(xv, yv, zv, cmap=cm.cividis,linewidth=0, antialiased=False)
+    surf = ax.plot_surface(xv, yv, zv, cmap=cm.RdPu,linewidth=0, antialiased=False)
     ax.zaxis.set_major_locator(LinearLocator(6))
     ax.zaxis.set_major_formatter('{x:.0f}')
     ax.set_xlim([0,5])

### 2025-11-21 jso.tex
small tweaks

@@
 
-\begin{tcolorbox}[enhanced jigsaw, breakable, bottomtitle=1mm, opacitybacktitle=0.6, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, toprule=.15mm, left=2mm, bottomrule=.15mm, rightrule=.15mm, titlerule=0mm, coltitle=black, leftrule=.75mm, colframe=quarto-callout-note-color-frame, colback=white]
+\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, coltitle=black, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, opacityback=0, toprule=.15mm, left=2mm, bottomtitle=1mm, rightrule=.15mm, bottomrule=.15mm, titlerule=0mm, colframe=quarto-callout-note-color-frame, colback=white, leftrule=.75mm, toptitle=1mm]
 

### 2025-11-21 code/11-sim-matern.py

@@ -21,8 +21,8 @@ nu_array = np.array([1.0,2.0,4.0])
 n_samples = 5
 x = np.linspace(0, 6, 2001)
 X = x.reshape(-1, 1)
-# RdPu but I change the end chroma to 200
-palette = ["#490062", "#AB1998", "#FF49B1", "#FF8A88", "#FFCB00"]
+# RdPu, but I change the end chroma to 100
+palette = ["#490062", "#963E87", "#E775A8", "#FFAEAD", "#FFE08C"]
 sns.set_palette(palette)
 
 

### 2025-11-22 code/11-sim-matern.py
update colormap Figure 3

@@ -12,7 +12,12 @@ from sklearn.gaussian_process.kernels import Matern
 
 
 
-
+# RdPu, but I change the end chroma to 100
+YlRdPu = ["#490062", "#963E87", "#E775A8", "#FFAEAD", "#FFE08C"]
+from matplotlib.colors import LinearSegmentedColormap
+clist = sns.color_palette(YlRdPu)
+my_cmap = LinearSegmentedColormap.from_list('my_cmap',clist,80)
+sns.set_palette(YlRdPu)
 
 
 
@@ -21,9 +26,7 @@ nu_array = np.array([1.0,2.0,4.0])
 n_samples = 5
 x = np.linspace(0, 6, 2001)
 X = x.reshape(-1, 1)
-# RdPu, but I change the end chroma to 100
-palette = ["#490062", "#963E87", "#E775A8", "#FFAEAD", "#FFE08C"]
-sns.set_palette(palette)
+
 
 
 fig, axs = plt.subplots(ncols=np.size(nu_array), sharex=False, sharey=False, figsize=(12, 4))
@@ -68,7 +71,7 @@ for (i,nu) in enumerate(nu_array):
     for n in np.arange(nx):
         zv[n,:] = np.squeeze(z_samples[n*nx:(n+1)*nx])
     
-    surf = ax.plot_surface(xv, yv, zv, cmap=cm.RdPu,linewidth=0, antialiased=False)
+    surf = ax.plot_surface(xv, yv, zv, cmap=my_cmap,linewidth=0, antialiased=False)
     ax.zaxis.set_major_locator(LinearLocator(6))
     ax.zaxis.set_major_formatter('{x:.0f}')
     ax.set_xlim([0,5])

### 2025-11-22 jso.tex
knit the pdf

@@
 
-\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, coltitle=black, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, opacityback=0, toprule=.15mm, left=2mm, bottomtitle=1mm, rightrule=.15mm, bottomrule=.15mm, titlerule=0mm, colframe=quarto-callout-note-color-frame, colback=white, leftrule=.75mm, toptitle=1mm]
+\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, toprule=.15mm, arc=.35mm, opacityback=0, opacitybacktitle=0.6, bottomrule=.15mm, breakable, left=2mm, coltitle=black, bottomtitle=1mm, toptitle=1mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, titlerule=0mm]
 

### 2025-11-23 revision/response.qmd
table 1 caption and response fixed

@@
 
-\textcolor{blue}{Table 1 is a subset of the PP index estimated and it happens that the three chosen ones all decrease as $n$ goes larger.
-This is not necessary the case in the full table shown in the Appendix.
-The caption for Table 1 has been fixed.}
+\textcolor{blue}{Table 1 is a subset of the estimated PP index, from the full table in the appendix.
+It happens that the three chosen ones all decrease as $n$ goes larger.
+This is not generally a pattern so this has been dropped from the caption.}
 

### 2025-11-23 jso.qmd
table 1 caption and response fixed

@@
 #| label: tbl-smoothness-squintability
-#| tbl-cap: The columns $\nu$ and $\varsigma$ show the smoothness and squintability measures for three PP indexes for the sine wave data.
+#| tbl-cap: The columns $\nu$ and $\varsigma$ show the smoothness and squintability measures for PP indexes for the sine wave data.
+Only three are shown for illustration, and a full table can be found in Supplementary materials.
 The splines index has the largest squint angle and is the smoothest.
 The skinny index has low smoothness and smallest squint angle.
-As dimension increases the smoothness becomes smaller but squintability doesn't change much.
 dt <- tibble(shape = c(rep("pipe", 5), rep("sine", 18)), smoothness) |>

### 2025-11-23 jso.tex
table 1 caption and response fixed

@@
 
-\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, toprule=.15mm, arc=.35mm, opacityback=0, opacitybacktitle=0.6, bottomrule=.15mm, breakable, left=2mm, coltitle=black, bottomtitle=1mm, toptitle=1mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, titlerule=0mm]
+\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, rightrule=.15mm, colback=white, toptitle=1mm, opacityback=0, arc=.35mm, left=2mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, bottomrule=.15mm, coltitle=black, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, toprule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white]
 
@@
 \caption{\label{tbl-smoothness-squintability}The columns \(\nu\) and
-\(\varsigma\) show the smoothness and squintability measures for three
-PP indexes for the sine wave data. The splines index has the largest
-squint angle and is the smoothest. The skinny index has low smoothness
-and smallest squint angle. As dimension increases the smoothness becomes
-smaller but squintability doesn't change much.}
+\(\varsigma\) show the smoothness and squintability measures for PP
+indexes for the sine wave data. Only three are shown for illustration,
+and a full table can be found in Supplementary materials. The splines
+index has the largest squint angle and is the smoothest. The skinny
+index has low smoothness and smallest squint angle.}
 

### 2025-11-23 revision/response.qmd
section 7.1 comment

@@
 
-\textcolor{blue}{We make Section 7.1 clearer that the planned tour can be run to trace the optimisation of individual jellyfish path.
-Users may wish to follow the most successful jellyfish, a random selection, or interesting jellyfish that result in local optimum.}
+\textcolor{blue}{In Section 7.1 some more details of using the planned tour with the jellyfish optimiser are given.
+More details on this are provided on the tourr package website, so that it can be kept updated if there are changes in the package.}
 

### 2025-11-23 jso.qmd
fig-alt AI generated added

@@
 The skewness index on the trimodal data is also smoother than the binned normality index on RANDU."
+#| fig-alt: "The figure contains four panels.
+Panels (a) and (c) are polar coordinate plots showing the values of projection pursuit (PP) indexes—skewness for trimodal data in (a) and binned normality for RANDU data in (c)—across all 1D projections of 2D datasets.
+The radius represents the index value for each projection angle, with a dashed circle indicating the average index and a straight line marking the optimal projection.
+Panel (a) displays a smooth but non-convex skewness index with three pronounced modes and a wide squint angle, while panel (c) shows a noisier, non-normality index with less smoothness and a narrow squint angle.
+Panels (b) and (d) are histograms representing data projected onto the optimal directions identified in (a) and (c), respectively; panel (b) reveals trimodality and panel (d) shows the discrete nature of the RANDU data, which is harder to discern except near the optimal projection."
 #proj <- animate_xy(flea[,1:6], guided_tour(lda_pp(flea$species)))

### 2025-11-23 jso.tex
fig-alt AI generated added

@@
 
-\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, rightrule=.15mm, colback=white, toptitle=1mm, opacityback=0, arc=.35mm, left=2mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, bottomrule=.15mm, coltitle=black, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, toprule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white]
+\begin{tcolorbox}[enhanced jigsaw, titlerule=0mm, opacityback=0, left=2mm, colback=white, leftrule=.75mm, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, breakable, rightrule=.15mm, arc=.35mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, coltitle=black]
 

### 2025-11-23 jso.qmd
all alt-text added

@@
 The index function MIC and splines2d make early progression towards the optimal index value, indicating a large squint angle, whereas the traces from skinny, and specially stringy2, show improvement only near the optimal projection, suggesting low squintability."
+#| fig-alt: "The figure presents four line graphs, each showing simulated optimization traces for different projection pursuit (PP) index functions: MIC, splines2d, skinny, and stringy2.
+The x-axis likely represents optimization iterations or search steps, while the y-axis shows the index value.
+Each trace illustrates how rapidly the index value improves during optimization.
+The MIC and stringy2 functions display steeper early increases (representing high squintability or high squint angle), indicating fast progress toward the optimal projection.
+In contrast, splines2d and skinny traces rise more slowly (lower squintability), reflecting that substantial improvement is only achieved when candidate projections get very close to optimal.
+The chart visually supports the concept of squintability as the percent improvement at the halfway point of the index value."
 sq_sine_basis_df |> 
@@
 Of the 50 simulations, 43 achieved an index value within 0.05 of the best, resulting in a success rate of 0.86."
+#| fig-alt: "The figure compares histograms or scatterplots of projected data from optimization simulations, illustrating results when the final projection index is within 0.05 of the optimal value.
+Axes likely represent the values along the projected data component.
+In these projections, the underlying 'pipe' structure is not visibly distinguishable, supporting the threshold choice for the success rate metric.
+The visual suggests that projections within this index margin do not reveal meaningful structure in the data, validating the 0.05 threshold as a criterion for optimization success."
 #| fig-pos: "!ht"
@@
 #| fig.cap: "Steps for calculating smoothness in a projection pursuit problem: 1) sample random bases given the orthonormality contraint and calculate their corresponding index values, 2) fit a Gaussian process model of index values against the bases to obtain the smoothness measure, 3) take the smoothness parameter from the Gaussian process as the smoothness metric."
+#| fig-alt: "The figure depicts a flowchart or schematic illustrating the smoothness computation process for projection pursuit (PP) indexes.
+It shows that for each PP task, 500 random basis directions are generated, index values are computed for these projections, and a Gaussian process model is fitted to the resulting data.
+The final output is a quantified smoothness measure of the index, capturing how smoothly the index value changes across different basis projections."
 # basis_smoothness <- sample_bases(idx = "splines2d", n_basis = 1000)
@@
 3) average the index values by projection distances at each 0.005 distance bin, 4) fit the scaled sigmoid function in Equation \eqref{eq-logistic} and \eqref{eq-parametric} , 5) calculate the squintability metric using Equation (@eq-squintability-parametric).
+#| fig-alt: "The diagram illustrates the procedure for calculating squintability of a projection pursuit (PP) index: 50 random initial bases are generated for each task, and linearly interpolated to the optimal basis in small increments of 0.005.
+For each interpolated basis, index values are computed and averaged within each 0.005 distance bin.
+These averaged values are then used to fit a logistic function via non-linear least squares, from which the squintability measure is derived, reflecting how quickly the index value improves as projections approach optimality."
 # p1 <- sim_sine_6d_spline_projdist |> 
@@
 As dimension increases both have more difficulty finding the pipe."
+#| fig-alt: "The figure presents a series of regression plots or bar charts depicting the relationships among squintability, smoothness, data dimension, and optimization success rate.
+The primary trend shows a moderate positive correlation between squintability and smoothness, but squintability is a much stronger predictor of success rate than smoothness.
+Addition of interaction terms with data dimension results in slight model improvement.
+Visual evidence suggests that higher squintability is strongly associated with higher success rates, while smoothness has only a minor influence.
+Differences between optimization algorithms, such as JSO and CRS, are noted, but optimizer efficiency is not strongly tied to smoothness."
 
@@
 The JSO has some difficulty finding the pipe when dimension is higher than 8 and maximum number of iteration has little effect."
+#| fig-alt: "A grouped bar chart shows the success rate of the JSO optimizer for projection pursuit tasks across varying data dimensions (d = 4, 6, 8, 10, and 12), number of jellyfish (20, 50, 100), and maximum iterations (50, 100).
+Each bar’s height indicates the mean success rate, with error bars representing uncertainty from bootstrap resampling.
+For low-dimensional tasks (d = 4, 6), high success rates are achieved with few jellyfish and iterations, while higher dimensions (d = 10, 12) require significantly more jellyfish and iterations to approach similar rates.
+The trend highlights that as dimensionality increases, optimizing becomes harder and more computational resources are necessary for acceptable performance."
 pipe_sim_best <- sim_pipe_run_best |> 
@@
 This illustrates the usefulness of the new measures for squintability and smoothness."
+#| fig-alt: "A line chart displays the self-standardized PP index values (ranging from 0 to 1 on the y-axis) plotted along the interpolation path from a random noise projection to the optimal projection (x-axis) for 6D sine wave data.
+Three index traces—splines, TIC, and skinny—are compared: the splines index line rises smoothly and steadily, indicating high smoothness and gradual improvement; the TIC index shows moderate smoothness with a slightly steeper increase; the skinny index trace is noticeably jagged, remaining low until near the optimal projection where it sharply increases.
+These patterns demonstrate the differences in squintability and smoothness among the PP indexes, with smoothness and gradual improvement aiding optimization."
 sq_basis_df_6d_TIC_spline_skinny |>

### 2025-11-23 jso.tex
all alt-text added

@@
 
-\begin{tcolorbox}[enhanced jigsaw, titlerule=0mm, opacityback=0, left=2mm, colback=white, leftrule=.75mm, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, breakable, rightrule=.15mm, arc=.35mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, coltitle=black]
+\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, toprule=.15mm, breakable, left=2mm, bottomtitle=1mm, coltitle=black, titlerule=0mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, bottomrule=.15mm, toptitle=1mm, arc=.35mm, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, opacityback=0]
 

### 2025-11-23 jso.qmd
use BNBU affiliation

@@
   - name: Nicolas  Langrené
-    email: nicolaslangrene@uic.edu.cn
+    email: nicolaslangrene@bnbu.edu.cn
     affiliations:
         - id: 3
-          name: Guangdong Provincial/Zhuhai Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College
+          name: Guangdong Provincial/Zhuhai Key Laboratory of Interdisciplinary Research and Application for Data Science, Beijing Normal-Hong Kong Baptist University
           department: Department of Mathematical Sciences
@@
 Nicolas Langrené acknowledges the partial support of the Guangdong
-Provincial/Zhuhai Key Laboratory IRADS (2022B1212010006) and the UIC
+Provincial/Zhuhai Key Laboratory IRADS (2022B1212010006) and the BNBU
 Start-up Research Fund UICR0700041-22.

### 2025-11-23 jso.qmd
update Def 6

@@
 \begin{defn}[squintability]\label{def:squintability}
-Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance $r(A, A^*)$ to the index value $f(XA)$, such that $g(r) = g(r(A, A^*)) = f(XA)$.
-The squintability of an index function $f$ is defined by
+Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that models the index value $f(XA)$ as a function of the projection distance $r(A, A^*)$, such that $g(r) = g(r(A, A^*)) \simeq f(XA)$.
+The squintability of an index function $f$ is then defined by
 

### 2025-11-23 jso.tex
knit

@@
 Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}. Nicolas
-Langrené. Email: \href{mailto:nicolaslangrene@uic.edu.cn}{\nolinkurl{nicolaslangrene@uic.edu.cn}}. Jessica
+Langrené. Email: \href{mailto:nicolaslangrene@bnbu.edu.cn}{\nolinkurl{nicolaslangrene@bnbu.edu.cn}}. Jessica
 Wai Yin
@@
 Mathematical Sciences, Guangdong Provincial/Zhuhai Key Laboratory of
-Interdisciplinary Research and Application for Data Science,
-BNU-HKBU~United~International~College, Zhuhai, China
+Interdisciplinary Research and Application for Data Science, Beijing
+Normal-Hong Kong Baptist University, Zhuhai, China
 \begin{abstract}
@@
 
-\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, toprule=.15mm, breakable, left=2mm, bottomtitle=1mm, coltitle=black, titlerule=0mm, leftrule=.75mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, bottomrule=.15mm, toptitle=1mm, arc=.35mm, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, opacityback=0]
+\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, toptitle=1mm, rightrule=.15mm, colback=white, coltitle=black, leftrule=.75mm, titlerule=0mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, bottomrule=.15mm, opacityback=0, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, toprule=.15mm, left=2mm]
 
